### 2021-09-21 chapter1/chapter.tex
thesisify from reverseME in Dropbox

%!TEX root = ../thesis.tex
\chapter{Introduction \label{ch:intro}}
A fluid queue is a two-dimensional stochastic process \(\{(X(t),\varphi(t))\}_{t\geq0}\). The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC). The level process, \(\{X(t)\}_{t\geq0}\), is a real-valued, continuous, and piecewise linear.

Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982} as a canonical application in this area), power systems \cite{hydro}, risk processes \cite{betal2005} and environmental modelling \cite{wurm2020}. Fluid queues are relatively well studied. Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \cite{anick1982,kk1995,beanetal2019}. %For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest. Anick Mitra Sondhi CITE, analysed fluid queues using a more direct differential equation-based method. Since Ramaswami's CITE initial work, there has been significant developments in the analysis of fluid queues CITE and related algorithms CITE.

More recently, Bean and O'Reilly \cite{bo2014} extended fluid queues to \emph{so-called} stochastic fluid-fluid queues. In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). The analysis, Bean and O'Reilly \cite{bo2014}, is in principal similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). For practical computation of the results \cite{bo2014} a matrix-discretisation of the infinitesimal generator of the fluid queue can be used. To this end, to date, two possible discretisation have been suggested. Taking a differential equations-based approach, Bean \emph{et al.}~\cite{beanetal2019} use the discontinuous Galerkin (DG) method to discretise this operator, while Bean and O'Reilly \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process. Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations. It turns out that the latter approach is a sub-class of the former; the QBD can be viewed as the simplest DG scheme where the operator is projected onto a basis of piecewise constant functions.

%A QBD can be viewed as a two-dimensional CTMC, \(\{(L(t),\varphi(t))\}_{t\geq0}\), where \(\{L(t)\}\) is the discrete level process, and \(\{\varphi(t)\}_{t\geq0}\) is the phase process. The level process \(\{L(t)\}\) is skip free, meaning that, given the process is at level \(L(t)=\ell\), is may only jump to \(\ell+1\) or \(\ell-1\) at jump epochs. The sojourn time of \(\{L(t)\}\) in a given level follows a phase-type distribution

In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \cite[Section 3.3]{koltai2011}, see also \cite{nodalDGBook} and references therein. One justification for the positivity preserving property is from the interpretation of the discretisation as a stochastic process thereby ensuring positivity. For higher order DG schemes there is no such interpretation. Moreover, higher-order DG approximation schemes may produce negative, or highly oscillatory solutions, particularly when discontinuities or steep gradients are present. Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \cite[Section 6.5]{nodalDGBook} and references therein). Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in theses regions. Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.

\subsection{Fluid queues}
A fluid queue is a two-dimensional stochastic process \(\{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer. The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we we assume to be \(\mathcal S=\{1,2,\dots,N\}\) without loss of generality, and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\). We assume that \(\bs T\) is \emph{conservative}. Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\).

Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{\i\in\calS\mid c_i<0\}\) and \(\calS_0 = \{i\in\calS\mid c_i=0\}\). We assume, without loss of generality, that the generator \(\bs T\) is partitioned into sub-matrices 
\[\bs T = \left[\begin{array}{ccc}\bs T_{++} & \bs T_{+-} & \bs T_{+0} \\ \bs T_{-+} & \bs T_{--} & \bs T_{-0} \\ \bs T_{0+} & \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
where \(\bs T_{mn} = [T_{ij}]_{i\in\mathcal S_m, j\in\mathcal S_n}\), \(m,n\in\{+,-,0\}\). Also define the diagonal matrices 
\begin{align*}
	\bs C &= \left[\begin{array}{ccc} \bs C_+ && \\ &\bs C_-& \\ && \bs 0\end{array}\right], && \bs C_+ = diag(c_i,i\in\calS_+), && \bs C_- = diag(|c_i|,i\in\calS_-),
\end{align*}
where \(diag(a_i,i\in\mathcal I)\) denotes a diagonal matrix with entries \(a_i\) down the diagonal.

When no boundary conditions are imposed, the level process is given by 
\[X(t) = X(0) + \int_{s=0}^t c_{\varphi(s)}\wrt s.\]
Sample paths of \(\{X(t)\}\) are continuous and piecewise linear, with \(\cfrac{\wrt }{\wrt t} X(t) = c_\varphi(t)\), when \(X(t)\) is differentiable. Given sample paths of \(\{\varphi(t)\}\), then \(\{X(t)\}\) is deterministic, and in this sense, \(\{\varphi(t)\}\) is the only stochastic element of the fluid queue.

Often, boundary conditions are imposed. Here, we consider a mixture of \emph{regulated} and \emph{reflecting} boundary conditions. Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition). At a lower boundary, if \(j\in\calS_0\cup\calS_-\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\), and the phase process continues to evolve according to the sub-generator 
\[\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
until such a time that \(\varphi(t)\) transitions to a phase \(k\in\calS_+\), at which time \(X(t)\) leaves the boundary. Similarly, at an upper boundary if \(j\in\calS_0\cup\calS_+\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\) and the phase process continues to evolve according to the sub-generator 
\[\left[\begin{array}{cc} \bs T_{++} & \bs T_{+0} \\ \bs T_{0+} & \bs T_{00}  \end{array}\right],\]
until such a time that \(\varphi(t)\) transitions to a phase \(k\in\calS_-\) at which time \(X(t)\) leaves the boundary. Without loss of generality, we assume the lower and upper boundaries (when present) are at \(x=0\) and \(x=M>0\), respectively.

In summary, the evolution of the level can be expressed as 
\[\cfrac{\wrt}{\wrt t} X(t) = \begin{cases} c_{\varphi(t)}, & \mbox{ if } X(t)>0, \\ \max\{0,c_{\varphi(t)}\}, & \mbox{ if } X(t)=0, \\ \min\{0,c_{\varphi(t)}\}, & \mbox{ if } X(t)=M.  \end{cases}\]

Let \(\bs f(x,t) = \vligne{f_i(x,t)}_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i)\), assuming it exists. When a differentiable density exists, the system of partial differential equation which describes the evolution of the densities \(\bs f(x,t)\) is 
\begin{equation}
	\cfrac{\partial}{\partial t} \bs f(x,t) = \bs f(x,t)\bs T - \cfrac{\partial}{\partial x}\bs f(x,t)\bs C.\label{eqn: pde}
\end{equation}
The initial condition is the initial distribution of the fluid queue, \(f_i(x,0)\). Often a differentiable density function does not exist and therefore the partial differential equation (\ref{eqn: pde}) is not well-defined. For example, for a fluid queue with a regulated boundary, if the initial distribution of the fluid queue is a point mass at any point \(x_0\geq 0\) and in phase \(i\in\calS_+\cup\calS_0\), then a density function \(f_i(x,t)\) will not exist for any finite \(t\). Specifically, a point mass will persist along the ray \(x_0+c_it\), \(t\geq 0\). In such situations, it is the \emph{weak solution} to (\ref{eqn: pde}) that we seek.

Boundary conditions may also be imposed on (\ref{eqn: pde}).

Discretisation methods approximate the operator on the right-hand side of (\ref{eqn: pde}) by a matrix, in our case, by the generator of a QBD-RAP.

### 2021-09-21 chapter2/chapter.tex
thesisify from reverseME in Dropbox

%!TEX root = ../thesis.tex
\chapter{A stochastic modelling approach to approximation\label{sec: construction and modelling}}
In this work we develop a new discretisation of a fluid queue using a \emph{quasi-birth-and-death process with rational arrival process components} (QBD-RAP), to be described in more detail later, and prove convergence of the approximation scheme. To improve upon the discretisation of Bean and O'Reilly \cite{bo2013}, we argue that the shift from QBD to QBD-RAP is necessary. The approximation can be made arbitrarily accurate. Due to its stochastic interpretation, the QBD-RAP discretisation ensures solutions are positive, even when discontinuities are present. Unlike slope limiters and post-hoc filtering, for the QBD-RAP the positivity preserving nature is a property of the discretised operator.

This paper details the construction of the approximation and proves its convergence. The QBD-RAP is constructed using \emph{matrix-exponential distributions}. The main result shows the convergence of the approximation scheme under the assumption that the variance of the matrix-exponential distribution(s) used in the construction tends to 0. Given that there is no known simple closed form for the transient distributions of a fluid queue, and that we make minimal assumptions about the matrix-exponential distributions used in the construction of the approximation, it is somewhat remarkable that we are able to establish a convergence result. The generality of the result with respect to the assumptions on the matrix-exponential distributions used in the construction of the approximation is necessitated by the fact that we use a class of \emph{concentrated matrix-exponential distributions} found numerically in \cite{hhat2020}, and for which there is relatively little known about their properties.

The structure of this paper is as follows. In Section~\ref{sec: prelim} we provide a brief background on matrix exponential random variables and the stochastic processes appearing in this work, namely fluid queues, quasi-birth-and-death (QBD) processes, rational arrival processes (RAPs), and quasi-birth-and-death processes with rational arrival process components (QBD-RAPs). Section~\ref{sec: construction and modelling} describes the modelling approach taken to develop the approximation. Section~\ref{sec: inspiration} motivates the idea using the QBD approximation of \cite{bo2014}. Sections \ref{sec: modelling} and \ref{sec: zero states} describe the modelling of certain events of the fluid queue with matrix exponential distributions to ultimately construct the behaviour of the QBD-RAP. Section~\ref{sec: dynamics} describes the dynamics of the resulting construction and constructs the level process of the QBD-RAP. Section~\ref{sec: boundary conditions} deals with modelling boundary behaviour. Section~\ref{sec: initial conditions} describes how to model initial conditions of the fluid queue. Section~\ref{sec: closing} introduces the concept of a \emph{closing operator} which maps states of the QBD-RAP to estimates of the distribution of the fluid queue. We then proceed to show convergence of the approximation scheme in Section~\ref{sec: conv}. First, in Section~\ref{sec: no change}, we derive expressions for certain Laplace transforms with respect to time of the QBD-RAP and fluid queue on the event that there is no change of level. Section~\ref{sec: no change convergence} provides error bounds between the Laplace transforms of the QBD-RAP and the fluid queue and also some domination conditions used so that we may apply the Dominated Convergence Theorem. Section~\ref{sec: local to global} uses the results of the preceding section to prove convergence Laplace transform with respect to time of the expected local time of the QBD-RAP to the Laplace transform with respect to time of the expected local time of the fluid queue. Technical proofs which do not offer a great deal of insight are reserved for appendices.

\section{Preliminaries}\label{sec: prelim}
This work approximates fluid queues using other stochastic processes, specifically quasi-birth-and-death processes with rational arrival process components (QBD-RAPs). The building blocks of QBD-RAPs are rational arrival processes (RAPs) and the building blocks of RAPs are matrix-exponential distributions. In this section, we introduce these concepts.

\subsection{Matrix-exponential distributions}
Here we recount some facts about matrix exponential distributions. See \cite{MEinAP} for a more detailed exposition. A random variable, \(Z\), is said to have a matrix-exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential. The density function of \(Z\) is given by \(f_Z(x) = \bs\alpha e^{\bs S x}\bs s\). The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) There is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility. These condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\). However, in general there is no way to determine whether, given a triplet \((\bs \alpha, \bs S, \bs s)\) whether it is a representation of a matrix-exponential distribution, or not. Nonetheless, some properties of a triplet \((\bs \alpha, \bs S, \bs s)\) are known, such as the following, which is used in the characterisation of QBD-RAPs. 
\begin{thm}[Theorem 4.1.3, \cite{MEinAP}]
	The density function of a matrix-exponential distribution with representation \((\bs\alpha, \bs S, \bs s)\) can be expressed in terms of real-valued constants as 
	\begin{align}
		\psi(x)&=\sum_{j=1}^{m_1} \sum_{k=1}^{p_j} c_{jk} \cfrac{x^{k-1}}{(k-1)!} e^{\mu_jx} + \sum_{j=1}^{m+2}\sum_{k=1}^{q_j} d_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\cos(\sigma_jx) + \sum_{j=1}^{m_2}\sum_{k=1}^{q_j}e_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\sin(\sigma_jx), \label{eqn: spec}
	\end{align}
	for integers \(m_1,\,m_2,\,p_j,\) and \(q_j\) and some real constants \(c_{jk},\,d_{jk},\,e_{jk},\,\mu_j,\,\eta_j,\) and \(\sigma_j\). Here \(\mu_j,\, j=1,\dots,m_1\) are the real eigenvalues of \(\bs S\), while \(\eta_j+i\sigma_j, \, \eta_j-i\sigma_j,\, j=1,\dots,m_2\) denote its complex eigenvalues, which come in conjugate pairs. Thus \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m+1}p_j + 2\sum_{j=1}^{m_2}q_j\). 
\end{thm}
\begin{thm}[Theorem 4.1.4, \cite{MEinAP}]\label{thm: 4.1.4}
	Consider the nonvanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}=0,\,d_{jk} =0,\) or \(e_{jk}=0\). Among the corresponding eigenvalues \(\lambda_j\), there is a real dominating eigenvalue \(\kappa\), say. That is, \(\kappa\) is real, \(\kappa \geq Re(\lambda_j)\) for all \(j\), and the multiplicity of \(\kappa\) is at least the multiplicity of every other eigenvalue with real part \(\kappa\).
\end{thm}
\begin{cor}[Corollary 4.1.5, \cite{MEinAP}]
If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix-exponential distribution, then \(\bs S\) has a real dominating eigenvalues.
\end{cor}
\begin{thm}[Theorem 4.1.6, \cite{MEinAP}]
	Let \(Z\) be a matrix-exponentially distributed random variable with density (\ref{eqn: spec}). Then the dominant real eigenvalue \(\kappa\) of Theorem~\ref{thm: 4.1.4} is strictly negative. 
\end{thm}
We define \(dev(\bs S)\) to be the real dominating eigenvalue of \(\bs S\), that is \(dev(\bs S)=\kappa\) in Theorem~\ref{thm: 4.1.4}.

The class of matrix-exponential distribution is characterised as the class of probability distributions which have a rational Laplace transform. That is, \(\displaystyle \int_{x=0}^\infty e^{-\lambda x}\bs \alpha e^{\bs S x}\bs s \wrt x\) is a ratio of two polynomial functions in \(\lambda\). Matrix exponential distributions are an extension of Phase-type distributions, where for the latter, \(\bs S\) must be a sub-generator matrix of a CTMC, \(\bs s = -\bs S\bs e\) where \(\bs e\) is a \(1\times p\) vector of ones, and \(\bs \alpha\) is a discrete probability distribution.

A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix-exponential distribution with this representation. The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\). Representations of matrix-exponential distributions are not unique \cite{MEinAP}. A representation is called \emph{minimal} when \(\bs S\) has the smallest possible dimension. Throughout this work, we assume that the representation of any matrix exponential distribution is minimal. Let \(\bs e_i\) be a vector with a 1 in the \(i\)th position and zeros elsewhere. We assume that \(\bs s = -\bs S\bs e\), and that \((\bs e_i,\bs S,\bs s)\) for \(i=1,\dots,p\) are representations of matrix exponential distributions. It is always possible to find such a representation \cite[Theorem 4.5.17, Corollary 4.5.18]{MEinAP}. As such, we abbreviate our notation \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to \(Z\sim ME(\bs \alpha, \bs S)\). Further, given \(\bs s=-\bs S\bs e\) then observe that  \(\displaystyle \int_{x=0}^\infty e^{\bs S x} \bs s\wrt x = (-\bs S)^{-1}\bs s= \bs e\).

For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal A\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a matrix exponential distribution. %By CITE it is known that \(\mathcal A\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).

%\subsection{Rational arrival processes}
%Let \(N\) be a simple point process, with event time \(Y_0=0<Y_1<Y_2<\cdot\). Let \(\{N(t)\}_{t\geq0}\) be the right-continuous counting process associated with \(N\); \(N(t)\) returns the number of events by time \(t\). Denote by \(f_{N,n}(y_1,\dots,y_n)\) the joint density of \(Y_1, Y_2-Y_1, \dots,Y_n-Y_{n-1}\), the first \(n\) inter-arrival times. For a matrix \(\bs B\) let \(dev(\bs B)\) denote the dominant eigenvalue of \(\bs B\) (the one with maximal real part). Theorem 1.1 of Asmussen and Bladt CITE states that the point process \(N\) is a RAP if there exists matrices \(\bs S\), \(\bs D\), a row vector \(\bs\alpha\), and a column vector \(\bs s\) such that \(dev(\bs S)<1\), \(dev(\bs S+\bs D)=0\), \((\bs S+\bs D)\bs e=0,\) and 
%\[f_{N,n}(y_1,\dots,y_n) = \bs \alpha e^{\bs Sy_1} \bs D e^{\bs Sy_1} \bs D \dots e^{\bs Sy_n} \bs s.\]
%Here \(\bs s\) can be taken to be \(\bs D e\). The \(n\)th inter-arrival, \(T_n-T_{n-1}\), has a matrix exponential distribution with density \(\bs \alpha(-\bs S\bs D )^{n-1}e^{\bs S x_n}\bs s\). Denote such a process \(N \sim RAP(\bs \alpha, \bs S, \bs D)\). 
%
%Asmussen and Bladt CITE, Corollary 2.2, show that associated with the RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\), 
%\[\bs A(t) = \cfrac{\bs \alpha\left( \prod_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs S\right)e^{\bs S t-Y_{N(t)}}}{\bs \alpha\left( \prod_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs S\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
%Thus, \(\{\bs A(t)\}\) is a piecewise-deterministic Markov process where, in between jumps \(\{\bs A(t)\}\) according to 
%\[\bs A(t) = \cfrac{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}}{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}\bs e},\]
%where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\). The process \(\{\bs A(t)\}\) jumps at event times of \(N\). At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\), and upon a jump at time \(t\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D/\bs A(t^-) \bs D\bs e\). 
%
%RAPs are an extension of Markovian arrival processes (MAPs) to include matrix-exponential inter-arrival times. For MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain. 
%
%Intuitively, \(\bs A(t)\) encodes all of the information about the jump times of the RAP up to time \(t\). Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\). Then \(N\mid \mathcal F_t \equiv N\mid \bs A(t) \sim ME(\bs A(t), \bs S, \bs S)\). In words, the future of the point process after time \(t\) given all of the information about the process up to and including time \(t\), is distributed as a RAP with initial vector \(\bs A(t)\).

\subsection{QBD-RAPs}
As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes. To define a QBD-RAP we first need a Batch (Marked) RAP.

Let \(\mathscr K\subset \mathbb Z\) be a set of \emph{marks}. Let \(N\) be a point process, and \(Y_0=0<Y_1<Y_2\cdots\) be event times of \(N\). Let \(\{N(t)\}\) be the counting process associated with \(N\) such that \(N(t)\) returns the number of events by time \(t\). Associated with the \(n\)th event is a mark \(M_n\). For \(i\in\mathscr K\), let \(N_i\) be simple point processes associated with events with marks of type \(i\) only, and let \(\{N_i(t)\}_{t\geq 0}\) be the associated counting processes of events of mark \(i\). For a matrix \(\bs B\) let \(dev(\bs B)\) denote the real part of the dominant eigenvalue of \(\bs B\) (the one with maximal real part). Denote by \(f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n)\) the joint density, probability mass function of the first \(n\) inter-arrival times, \(Y_1,Y_2-Y_1,\dots,Y_n-Y_{n-1}\), and the associated marks \(M_n\). From Bean and Nielsen \cite[Theorem 1]{bn2010} we have the following. 
\begin{thm}
A process \(N\) is a Marked RAP if there exist matrices \(\bs S\), \(\bs D_i,\,i\in\mathscr K\), and a row vector \(\bs \alpha\) such that \(dev(\bs S)<0\), \(dev(\bs S+\bs D)=0\), \((\bs S+\bs D)\bs e = 0\), \(\bs D = \sum_{i\in\mathscr K} \bs D_i\), and 
\begin{align}f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n) = \bs \alpha e^{\bs S y_1}\bs D_{m_1} e^{\bs S y_2} \bs D_{m_2}\dots e^{\bs S y_n} \bs D_{m_n}\bs e.\label{eqn: brap}\end{align}
Conversely, if a point process has the property (\ref{eqn: brap}) then it is a Marked RAP.
\end{thm}

Denote such a process \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\mathscr K).\)

Also from Bean and Nielsen \cite{bn2010}, associated with a Marked RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\)
\[\bs A(t) = \cfrac{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S (t-Y_{N(t)})}}{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
Thus, \(\{\bs A(t)\}\) is a piecewise-deterministic Markov process where, in between events \(\{\bs A(t)\}\) evolves deterministically according to 
\[\bs A(t) = \cfrac{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}}{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}\bs e},\]
where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\). The process \(\{\bs A(t)\}\) "jumps" at event times of \(N\) (the process may not always jump at these times, but typically the dynamics change discontinuously at this point). At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\). Upon an event the event is associated with mark \(i\) with probability \(\bs A(t) \bs D_i\bs e/\bs A(t) \bs D\bs e\). Upon on an event at time \(t\) with mark \(i\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D_i/\bs A(t^-) \bs D_i\bs e\).

Marked RAPs are an extension of Marked Markovian arrival processes to include matrix-exponential inter-arrival times. For Marked MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain.

Intuitively, \(\bs A(t)\) encodes all of the information about the event times of the Marked RAP and associated marks up to time \(t\) that is needed to determine the future behaviour of the point process. Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\). Then \(N\mid \mathcal F_t \equiv N\mid \bs A(t) \sim BRAP(\bs A(t), \bs S, \bs D_i,i\in\mathscr K)\). In words, the future of the point process after time \(t\) given all of the information about the process up to and including time \(t\), is distributed as a Marked RAP with initial vector \(\bs A(t)\).

Now consider a Marked RAP, \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\{-1,0,+1\}).\) The process \(\{(L(t),\bs A(t))\}_{t\geq0}\) formed by letting \(L(t) = N_{+1}(t) - N_{-1}(t)\) is a QBD-RAP.

\section{Inspiration and motivation}\label{sec: inspiration}
The inspiration for our approach stems from the QBD-based discretisation of Bean and O'Reilly \cite{bo2013}. A QBD can be seen as a two dimensional CTMC, \(\{(L(t),\phi(t))\}_{t\geq 0}\), where \(\{L(t)\}\) is the discrete level, and \(\{\phi(t)\}\) is the phase. Bean and O'Reilly \cite{bo2013} discretise the state space of the fluid into small intervals of width \(\Delta\); denote these by \(\mathcal D_k=[k\Delta,(k+1)\Delta]\). Their approximation captures the dynamics of the phase process exactly, but discretises the level process of the fluid queue. Their approximation supposes that when \(L(t)=k,\,\phi(t)=i\), then \(X(t)\approx k\Delta,\,\phi(t)=i\).

When in level \(k\) and phase \(i\), the approximation sees events at rate \(|T_{ii}| + |c_i|\Delta\). Upon an event, with probability \(T_{ij}/(|T_{ii}| + |c_i|\Delta)\) a change of phase from \(i\) to \(j\) is observed and the approximation remains in level \(k\); with probability \(|c_i|\Delta/(|T_{ii}| + |c_i|\Delta)\) a change of level occurs, to \(k+1\) if \(c_i>0\) and \(k-1\) if \(c_i<0\). The generator of their QBD approximation (for an unbounded fluid queue) is 
\[\bs B = \left[ \begin{array}{cccc} \ddots & \ddots & \ddots & \\ \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) & \\  & \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) \\ & \ddots & \ddots & \ddots \end{array} \right]\]
\begin{align}
\bs B_0 &= \bs T- \bs C\Delta, && \bs B_{-1} = \left[\begin{array}{ccc} \bs 0 & & \\ & \bs C_-\Delta & \\ & & \bs 0 \end{array}\right], && \bs B_{+1} = \left[\begin{array}{ccc} \bs C_+\Delta & & \\ & \bs 0 & \\ & & \bs 0\end{array}\right].
\end{align}
Bean and O'Reilly then take \(\Delta \to 0\) and show that the approximation converges weakly to the fluid queue. It seems that this is the best we can do if we want to keep the interpretation of the approximation as a QBD. We now elaborate on this point.

For a given \(\Delta\), the discretisation of Bean and O'Reilly supposes that, when the phase is \(i\) and on the event of no change of phase, the sojourn time in a given level has an exponential distribution with rate \(|c_i|\Delta\). However, the corresponding sojourn time of the fluid queue is deterministic: given \(X(t) = x \in\calD_k\), then \(\{X(t)\}\) will leave \(\calD_k\) in exactly \(((k+1)\Delta - x)/|c_i|\) units of time if \(c_i>0\), and \((x-k\Delta)/|c_i|\) units of time if \(c_i<0\), on the event that the phase does not change before this time. We can extend the QBD model of Bean and O'Reilly to model this determinism more accurately by supposing that the sojourn times in each interval have Erlang distributions (rather than exponential). However, we effectively realise the same QBD approximation as the original, but on a finer discretisation. For example, using Erlang distributions of order 2 we would get a generator of the QBD approximation
\[\bs B = \left[ \begin{array}{cccc} \ddots & \ddots & \ddots & \\ \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) & \\  & \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) \\ & \ddots & \ddots & \ddots \end{array} \right]\]
\begin{align}
\bs B_0 &= \bs T\otimes I_2 - \left[ \begin{array}{ccc} \left[\begin{array}{cc} -\Delta/2 & \Delta/2 \\ 0 & -\Delta/2 \end{array}\right] \otimes \bs C_+ & & \\ & \left[\begin{array}{cc} -\Delta/2 & 0 \\ \Delta/2 & -\Delta/2 \end{array}\right] \otimes C_- & \\ & & \bs 0 \end{array}\right], 
\end{align}
\begin{align}
\bs B_{-1} &= \left[\begin{array}{ccc} \bs 0 & & \\ & \bs C_-\otimes \left[\begin{array}{cc} 0 & \Delta/2 \\ 0 & 0 \end{array}\right] & \\ & & \bs 0 \end{array}\right],
&&\bs B_{+1} = \left[\begin{array}{ccc} \bs C_+\otimes \left[\begin{array}{cc} 0 & 0 \\ \Delta/2 & 0 \end{array}\right] & & \\ & \bs 0 & \\ & & \bs 0\end{array}\right].
\end{align}
But this is the same discretisation we would get if we were to use intervals of width \(\Delta/2\) in the QBD approximation of Bean and O'Reilly \cite{bo2013}, with the rows/columns reordered. Indeed, this appears to be the best we can do with a QBD-MAP approximation as the Erlang distribution is known to be the least variable Phase-type distribution for a given order \cite{as1987} and therefore gives the closest approximation to the deterministic behaviour we are trying to approximate. This necessitates the extension to QBD-RAPs, whereby we can find more concentrated distributions than any Phase-type, but retain (some of) the stochastic interpretation and matrix-analytic tools.

Here, the construction of the approximating QBD-RAP is developed intuitively, from a stochastic modelling perspective. The key observation is that, given \(\{\varphi(t)\}\) is constant, then \(\{X(t)\}\) moves deterministically. Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic. We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance). As in Bean and O'Reilly \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP. Unlike Bean and O'Reilly \cite{bo2013}, however, here we do not take \(\Delta \to 0\). Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small. This necessarily means that the order of the matrix exponential increases.

It turns out that the corresponding orbit process \(\{\bs A(t)\}\) can be seen to ``track", approximately, how far \(X(t)\) is from the left of a given interval when the phase is in \(\mathcal S_+\), or how far from the right of the interval \(X(t)\) is when the phase is in \(\mathcal S_-\). This leaves us with two problems. 1) on a transition from \(\mathcal S_+\) to \(\mathcal S_-\) or \(\mathcal S_-\) to \(\mathcal S_+\) how must \(\bs A(t)\) jump to retain this information about where \(X(t)\) is within a given interval. 2) how to use the orbit position at time \(t\), \(\bs A(t)\), to obtain an approximation for the density of \(X(t)\). Answers to both problems can be derived from the \emph{residual time} of a matrix exponential distribution. Let us now formalise these concepts some more.

\section{Time to exit an interval}\label{sec: modelling}
Consider partition of the state space of the level of a fluid queue, \([0,M]\) into \(K\) intervals of width \(\Delta = M/(K+1)\), specifically \(\mathcal D_{k,i} = [k\Delta,(k+1)\Delta)\) if \(i\in\mathcal S_+\) and \(\mathcal D_{k,i} = (k\Delta,(k+1)\Delta]\) if \(i\in\mathcal S_-\), \(k=0,\dots,K+1\). The distinction between \(\mathcal D_{k,i}\) for \(i\in\mathcal S_+\) and \(\calS_-\) is a technical one and will be discussed later. For now, one reason we might want to specify the discretisation in this way, is that it ensures the stochastic process 
\begin{align}
	\displaystyle \left\{\sum_{k\in\mathcal K}\sum_{i\in\calS} k1(X(t)\in\calD_{k,i})\right\}_{t\geq 0} \label{eqn: kfj}
\end{align}
is c\'adl\'ag. The expression (\ref{eqn: kfj}) is the discrete process of the fluid-queue which the level process of the QBD-RAP will approximate. Let \(y_k = k\Delta\), \(k=0,\dots,K+1\) and \(\calD_k = [y_k,y_{k+1}]\). We now look to model the behaviour of \(\{X(t)\}\) on an interval \(\calD_k\) over the time \([t,t+u]\) given information about the phase process over this time.

\paragraph{Modelling the residual time to exit on no change of phase} 
We first consider no change of phase in \([t,t+u]\). Suppose that \(\varphi(t)=i\in\mathcal S_+\) and \(X(t) = x\in\mathcal D_{k,i}\). Observe that on no change of phase, \(\varphi(t+s) = i\), \(s\in[0,u]\), and at time \(t+s\) the level of the fluid queue is given by \(X(t+s) = x+c_i s\), \(s\in[0,u]\). Further, for \(u>(y_{k+1}-x)/c_i\) the level process \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(y_{k+1} - x)/c_i\). Similarly, for \(i\in\mathcal S_-\), \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(x-y_k)/|c_i|\) on the event that there is no change of phase on \([t,t+u]\) for \(u>(x-y_k)/|c_i|\).

To approximate the deterministic behaviour observed above, consider a matrix-exponential distribution \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) with mean \(\Delta\) and low variance. As we shall see later, the assumption that \(Z\) has low variance will allow us to claim that \(Z\) can be used to approximate deterministic behaviour. Let \(Z_i\sim ME(\bs \alpha, \bs S_i, \bs s_i)\) where \(\bs S_i = |c_i|\bs S\), \(\bs s_i=-\bs S_i\bs e\), \(i\in\calS\). The random variable \(Z_i\) has mean \(\Delta/|c_i|\). Further, let \(R_i(u)\) be the residual time, \(R_i(u) = Z_i-u\mid \{Z_i-u>0\}\), \(i\in\calS\). Suppose first that \(X(t)=y_k\) and \(\varphi(t)=i\in\mathcal S_+\), so that \(X(t)\) will leave \(\mathcal D_{k,i}\) in exactly \(\Delta/|c_i|\) units of time on the event that the phase process remains in \(i\) for at least this amount of time.

After \(u\in[0,\Delta/|c_i|)\) amount of time has elapsed, on the event that \(\varphi(t+s)=i\), for all \(s\in[0,u)\), then \(X(t+u) = y_k + c_i u\). If the phase remains \(i\) for a further \(\Delta/|c_i|-u\) amount of time, then \(\{X(t)\}\) will leave \(\calD_{k,i}\) at exactly this time. Also at time \(t+u\), given \(Z_i>u\), the distribution of the residual time, \(R_i(u)\), has density 
\[f_{R_i(u)}(r) = \cfrac{\bs\alpha e^{\bs S_i(u+ r) } \bs s_i}{\bs\alpha e^{\bs S_iu } \bs e} = \bs A(t+u) e^{\bs S_i r}  \bs s_i.\]
Here, the event \(Z_i>u\) approximates the event \(X(t)\in\calD_{k,i}\) and we want the residual time, \(R_i(u)\), to approximate the time until \(\{X(t)\}\) leaves \(\calD_{k,i}\). That is, we want \(R_i(u)\) to approximate a deterministic random variable at \(\Delta/|c_i| - u\). To this end, we observe that, if the variance of \(Z_i\) is sufficiently low, then for \(\varepsilon>0\) and \(u<\Delta/|c_i|-\varepsilon\), then
\begin{align}
	\mathbb P(R_i(u) \in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon)) 
	&= \mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon)\mid Z_i > u) 
	\\&= \cfrac{\mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon))}{\mathbb P(Z_i > u)} 
	\\&= \cfrac{\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))}{\mathbb P(Z_i > u)} 
%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
	\\&\geq \cfrac{1-\cfrac{\var(Z)}{\varepsilon^2}}{1} 
\end{align}
since, by Chebyshev's inequality, \(\mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon)) \geq 1- \cfrac{\var(Z_i)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(X)^{1/3}\), then 
\[\mathbb P(R_i(u) \in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))\geq 1-\var(Z)^{1/3} \approx 1. \]
That is, when the variance of \(Z_i\) is low, the residual time \(R_i(u)\) will be concentrated around \(\Delta/|c_i| - u\), as required. Figure \ref{fig:residual distributions} gives an example of a density function for a \(Z_i\) with mean \(\Delta=1\) and \(c_i=1\), as well as the density function of \(R_i(0.3)\) and \(R_i(0.5)\) for comparison. Observe that the density of the residual life \(R_i(0.3)\) is concentrated around \(\Delta-0.3 = 0.7\) and, similarly the density of the residual life \(R_i(0.6)\) is concentrated around \(\Delta-0.6 = 0.4\).

\begin{figure}
\centering
\includegraphics[width=\textwidth]{chapter2/figs/ME_residual_life_density.pdf}
\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \cite{hhat2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.5}\) (blue). Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.5}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively. }
\label{fig:residual distributions}
\end{figure}

We can interpret the position of the orbit \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\). This gives a heuristic argument as to how we can model the sojourn times in a given interval \(\mathcal D_{k,i}\) on the event that the phase does not change.

We can apply analogous arguments to heuristically develop a model for the sojourn time of the fluid queue in an interval, \(\calD_{k,i}\), \(i\in\calS_-\), on the event that the phase does not change, given the fluid is \(X(t)=y_{k+1}\). In this case \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).

\paragraph{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\)} 
%Let \(Z_j\sim ME(\bs\alpha,\bs S_j)\) be a matrix exponential random variable with mean \(\Delta/|c_j|\) and let \(R_j(u)\) be the residual time \(R_j(u) = Z_j-u \mid\{Z_j-u>0\}\). %At time \(t+u\) the value of the orbit is \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\), which can be equivalently written as \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_ju/|c_i| } }{\bs\alpha e^{\bs S_ju/|c_i| } \bs e}\).
Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j\) occurs, given \(Z_j>|c_i|u/|c_j|\). The initial vector associated with the residual time \(R_j(|c_i|u/|c_j|)\) is \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\). This is the initial vector which is obtained if at time \(t\) the phase is \(\varphi(t)=j\) and the orbit is \(\bs A(t)=\alpha,\) so that the QBD-RAP has just entered level \(k\), and there is no change of phase by time \(|c_i|u/|c_j|\). Above, we argued that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) can be seen to correspond to the event that \(X(t+u)\) is \(|c_j|\left(|c_i|u/|c_j|\right) = |c_i|u\) units greater than \(y_k\). 
%We argued earlier that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the position of the orbit on the event that at time \(t\) there is a change of level so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=j\), followed by no change of phase by time \(t+|c_i|u/|c_j|\). The analogous event of the fluid queue is that at time \(t\) the phase is \(\varphi(t)=j\), the fluid level is \(X(t)=y_k\), so that the fluid level has just entered an interval \(\calD_k\) at time \(t\), and there is no change of phase by time \(t+|c_i|u/|c_j|\), in which case the fluid level will be \(X(t+|c_i|u/|c_j|)=y_\ell + |c_j|\left(|c_i|u/|c_j|\right)=y_\ell + |c_i|u\). Therefore, we claim that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the correct position of the orbit to approximate the fact that \(\{X(t)\}\) is \(|c_i|u\) units to the right of the point \(y_k\).

Notice that we can re-write the vector 
\(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) as \(\cfrac{\bs \alpha e^{\bs S_iu } }{\bs \alpha e^{\bs S_iu} \bs e }\). This is the orbit position on the event that at time \(t\) there is a change of level, so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=i\), followed by no change of phase by time \(t+u\). We have argued that this is the orbit position which approximates the fact that \(\{X(t)\}\) is \(|c_i|u\) units greater than \(y_k\).

Therefore, at time \(t+u\), if there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\) then the orbit position should not change as in both cases the orbit position \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) corresponds to an approximation to \(X(t)\) being \(|c_i|u\) units greater than \(y_k\).

%Therefore, suppose that at time \(t\), \(X(t)=y_k\), and that at time \(t+u\), \(u\in[0,\Delta/|c_i|)\), before \(\{X(t)\}\) exits the band \(\calD_{k}\), there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\). At time \(t+u\) the position of the fluid level is \(X(t+u) = y_k+c_iu\), which is \(|c_i|u\) units to the right of \(y_k\) and \(\Delta-c_iu\) to the left of \(y_{k+1}\), the right-hand boundary of \(\calD_{k,i}\). The residual time until the level of the fluid queue leaves \(\calD_{k,j}\), on the event that it remains in phase \(j\) until this time, is \((\Delta - |c_i|u)/|c_j|\).

Further to this argument, observe that the residual time \(R_j(|c_i|u/|c_j|)\) has density 
\[f_{R_j(|c_i|u/|c_j|)}(r) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}e^{\bs S_j r}\bs s_j,\]
%and corresponds to an orbit position of \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\). 
and 
\begin{align}
	&\mathbb P(R_j(|c_i|u/|c_j|) \in ((\Delta-|c_i|u)/|c_j|-\varepsilon, (\Delta-|c_i|u)/|c_j| + \varepsilon)) 
	\\&=\mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon)\mid Z_j > |c_i|u/|c_j|) 
	\\&= \cfrac{\mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon))}{\mathbb P(Z_j > |c_i|u/|c_j|)} 
%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
%	\\&\geq \cfrac{1-\cfrac{\var(Z_j)}{\varepsilon^2}}{1} 
	\\&\geq 1-\var(Z)^{1/3} \approx 1.
\end{align}
Hence, when the variance of \(Z\) is low, the residual time, \(R_j(|c_i|u/|c_j|)\), is concentrated around \((\Delta - |c_i|u)/|c_j|\) as required.

Therefore, in our model, upon the event of a change from \(i\in\calS_+\) to \(j\in\calS_+\) at time \(t+u\) the orbit should be \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\). That is, the initial condition for the next segment of the orbit process is just the current orbit position and the orbit should start to evolve according to the matrix \(\bs S_j\).

Analogous arguments suggest that the same applies for changes of phase from \(i\in\calS_-\) to \(j\in\calS_-\).

\paragraph{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\)} Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\). We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\). Meanwhile, the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units to the left of \(y_{k+1}\). We have also reasoned that for \(i\in\calS_-\) the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_{k+1}-x\), \(x\) units less than \(y_{k+1}\), and the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units greater than \(y_k\). Notice that in positive phases the orbit position describes the exact position of \(X(t)\) from the left of the interval but only approximates the position of \(X(t)\) from the right of the interval, and vice-versa for negative phases.

To construct the QBD-RAP we need to find a matrix \(\bs D\), say, to take the information about the position of \(X(t)\) contained in the orbit when in positive phases, and map it to information about the position of \(X(t)\) in negative phases. Then, when there is a change of phase at time \(t+u\) from \(i\in\calS_+\) to \(j\in\calS_-\), the orbit position will jump from \(\bs A(t+u^-)\) to \(\bs A(t+u)=\bs A(t+u^-)\bs D\). The matrix \(\bs D\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D, \bs S)\) is a representation of a matrix exponential distribution. The matrix \(\bs D\) should also have the property that \(\bs D \bs e = \bs e\) as this will mean that we can model the phase process of the fluid exactly.

To construct such a matrix \(\bs D\). If we were given \(R_i(u)=r\), then the appropriate orbit position upon a jump from phase \(i\in\calS_+\) to \(j\in\calS_-\) would be \(\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\) as this corresponds to the fluid level position \(y_{k+1}-|c_i|r\). However, at time \(t+u\), \(R_i(u)\) is a random variable about the future of the process and therefore not known, so instead, we take the expected initial vector 
\[\mathbb E \left[ \cfrac{\bs\alpha e^{ \bs S_iR_i(u) } }{\bs\alpha e^{\bs S_iR_i(u) } \bs e} \right] = \int_{r=0}^\infty \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} e^{\bs S_ir } \bs s_i \cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{r=0}^\infty e^{\bs S_ir } \bs s_i\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r.\]
After a change of variables \(x=|c_i|r\) and defining \(\bs D = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\), we get 
\[\cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{x=0}^\infty e^{\bs S x }|c_i| \bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x/|c_i| = \bs A(t+u)\int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x = \bs A(t+u) \bs D,\]
since at time \(t+u\), the orbit position is \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e}\). This suggests that upon a jump from \(i\in\calS_+\) to \(j\in\calS_-\), the orbit position jumps according to the matrix \(\bs D\).

Observe that \(\bs D\bs e = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\bs e =\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \wrt x=\bs e.\) This is an important observation as it implies that events corresponding to jumps from \(i\in\calS_+\) to \(j\in\calS_-\) occur with intensity \(\bs A(t)T_{ij}\bs D\bs e = T_{ij}\) which is constant with respect to orbit position and time and this implies they are exponential. Further, since \(\mathcal A\) is closed and convex \cite{MEinAP}, then \((\bs a \bs D,\bs S,\bs s)\) is a representation of a matrix exponential distribution for any \(\bs a\in\mathcal A\).

Note that the matrix \(\bs D\) is a modelling choice and other choices are possible. For example, \(\displaystyle \bs D = \int_{x=0}^{\Delta-\varepsilon}e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\). It may also be possible to construct other matrices \(\bs D\), perhaps via geometric arguments.

\paragraph{The Phase-type case}
If \(Z\sim ME(\bs \alpha, \bs S)\) is chosen to be a Phase-type distribution then there is another possibility for the matrix \(\bs D\). A Phase-type distribution is the distribution of time until absorption of a finite state absorbing Markov chain with transient states \(\{1,\dots,p\}.\) The sub-generator matrix describing the dynamics of the Markov chain on transient states is \(\bs S\), and \(\bs \alpha\) is an initial probability distribution over the transient states. Let \(\{J(t)\}_{t\in (-\infty,\infty)}\) be the Markov chain to which the Phase-type distribution corresponds.

Let \(\bs \Pi = diag\left(\bs \pi \right)\), where \(\vligne{\pi_k}_{k\in\{1,\dots,p\}}=:\bs \pi=\bs \alpha (-\bs{S})^{-1}/m\) and \(m=\bs \alpha (-\bs{S})^{-1}\bs e\). There is a time-reverse representation of a Phase-type distribution given by \((\widetilde{\bs \alpha}, \widetilde{\bs S}, \widetilde{\bs s})\), where \(\widetilde{\bs \alpha} = m\bs s' \bs \Pi \), \(\widetilde{\bs S} = \bs\Pi^{-1}\bs S'\bs \Pi\) and \(\widetilde{\bs s} = \bs \Pi^{-1}\bs \alpha'/m\) \cite[Page 91]{a2008}.

Rather than use the residual time \(R_i(u)\) to estimate how far \(X(t)\) is from the edge of an interval, we can use the information contained in \(J(t+u)\). Suppose there is a change of phase of the fluid queue from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\) at which time \(J(t+u)=k\). At a random time before \(t+u\) the Phase-type random variable was initialised. Let \(C_{t+u}\) be the \emph{age} of the Phase-type distribution at time \(t+u\), that is, the time since the Phase-type random variable was initialised. Suppose that \(\varphi(v)=i\), \(v\in [t+u-C_{t+u}, t+u)\), that is, there has been no change of phase of the fluid queue during the current life of the Phase-type random variable. Given \(J(t+u)=k\), we wish to construct an approximation to the distribution of \(C_{t+u}\).

According to Hautphenne \emph{et al.}~\cite{hmp2017}, the distribution of \(C_{t+u}\) depends on the method of observation. Here, the Phase-type distribution is observed after an exponential time with rate \(-T_{ii}\). Proposition 4.1, Hautphenne \emph{et al.}~\cite{hmp2017} states that the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) and the process is observed after an exponential time with rate \(-T_{ii}\) is 
\[\mathbb P(C_{t+u}\leq a )=1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}.\]
Let \(\widehat{ \bs S}_i(T_{ii}) = diag(\bs \nu)^{-1}\bs S_i'diag(\bs \nu)\), where \(\bs \nu = \bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}/(\bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e)\). Algebraic manipulations show
\begin{align}
	1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}
	= 1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)a} \bs e,\label{eqn: skgj}
\end{align}
which is of Phase-type. The conversion between the intensity at which events occurs when the fluid queue is in phase \(i\) compared to phase \(j\) is \(|c_j|/|c_i|\). %Taking the derivative of (\ref{eqn: skgj}) with respect to \(v\), and applying the chain rule, gives 
%\[\cfrac{\wrt }{\wrt u}\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)u} \bs e \cfrac{|c_j|}{|c_i|} = \bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)|c_j|v} \widehat{\bs s}_{ij}, \]
%where \(\widehat{\bs s}_{ij} = (\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\cfrac{|c_j|}{|c_i|} \bs e\). 
Therefore, when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), if the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the generator \(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\). The distribution of time until the next event in the QBD-RAP after time \(t+u\) is 
\[1-\bs e_k e^{(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I) + T_{jj}\bs I)a}\bs e.\]
While this does approximate the desired quantity, it is not quite satisfactory for the purpose of our approximation scheme due to dependence on the sample path of \(\{\varphi(t)\}\). Specifically, the evolution of the QBD-RAP from time \(t+u\) until the next event depends on the phase immediately before the change of phase at time \(t+u\), \(\varphi(t+u^-)=i\). This increases the size of the approximating QBD-RAP as we need a separate model for each \(\varphi(t+u^-)\in\calS_+\). Furthermore, we have not yet considered how to model further changes from \(\calS_-\) to \(\calS_+\) or beyond, which further complicates matters.

A solution is to suppose that, rather than observing \(C_{t+u}\) at an exponential time with rate \(-T_{ii}\), we instead observe \(\{J(t+u)\}\) uniformly randomly on the lifetime of length \(Z_i\). In this case, the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) is (Hautphenne \emph{et al.}~\cite[Lemma 3.1]{hmp2017})
\begin{align}\mathbb P(C_{t+u}\leq a) = 1-\cfrac{\bs \alpha e^{\bs S_iu}(-\bs S_i)^{-1}\bs e_k}{\bs \alpha (-\bs S_i)^{-1}\bs e_k} = 1- \bs e_k e^{\widetilde{ \bs S}_iu}\bs e.\label{eqn:ksda}\end{align}
Once again, noting that the conversion between the intensity at which events occurs when the fluid queue is in phase \(i\) compared to phase \(j\) is \(|c_j|/|c_i|\)%, and taking the derivative with respect to \(v\) of (\ref{eqn:ksda}) by applying the chain rule, gives 
%\begin{align}%
%	\cfrac{\wrt }{\wrt u} \bs e_k e^{\widetilde{ \bs S}_iu}\bs e|c_j|/|c_i|=\bs e_k e^{\widetilde{ \bs S}_iv|c_j|/|c_i|}\widetilde{ \bs S}_i\bs e|c_j|/|c_i| = \bs e_k e^{\widetilde{ \bs S}_jv}\widetilde{ \bs s}_j.\label{eqn:ksdadsva}
%\end{align}
then, upon a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the reverse-time generator \(\widetilde{ \bs S}_j\). The distribution of time until the next event in the QBD-RAP after time \(t+u\) is 
\[1-\bs e_k e^{(\widetilde{\bs S}_j + T_{jj}\bs I)a}\bs e.\] 
This suggests that, at a jump from \(\calS_+\) to \(\calS_-\) or from \(\calS_-\) to \(\calS_+\), the state of \(\{J(t+u)\}\) does not change, and begins to evolve according to the time-reverse generator at an appropriate speed.

With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of Bean and O'Reilly \cite{bo2014} with discretisation parameter \(\Delta/p\).

\paragraph{Upon exiting \(\calD_{k,i}\)} Suppose that upon exiting \(\calD_{k,i}\) at time \(t\) the phase is \(\varphi(t)=i\in\calS_+\). At this time \(X(t)=y_{k+1}\) which is the left-hand endpoint of \(\calD_{k+1,i}\). Hence, we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\). Similarly, upon exiting \(\calD_{k,i}\) at time \(t\) in phase \(\varphi(t)=i\in\calS_-\), then \(X(t)=y_k\), which is the right-hand endpoint of \(\calD_{k,i}\), and so we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).

\section{Computing \(\bs D\)}
In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \cite{hhat2020}. For this class of CMEs, we take the index \(p\) to be the order of the representation. Since the variance of \(Z^{(2p)}\) and \(Z^{(2p-1)}\) are relatively similar \cite{hhat2020}, we take \(p\) to be odd. For a given CME, with representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part.

We numerically evaluate \(\bs{D}\) where
\begin{align*}
	\bs{D} = \displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\cfrac{\bs \alpha e^{\bs{S}t}}{\bs \alpha e^{\bs{S}t} \bs e}\wrt t
\end{align*}
using a trapezoidal rule as follows.

For CMEs, the vector function \(\bs a(t) := \cfrac{\bs \alpha e^{\bs St}}{\bs \alpha e^{\bs S t} \bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\). Let \(\bs f(t) = e^{\bs S t}\bs s\). Then \(\bs f(t)e^{-\lambda t}\) where \(\lambda = \Re(\lambda_i)\) is the real part of the eigenvalues of \(\bs{S}\) (the all share the same real part), is also periodic with the same period. Hence we can simplify the integral to a finite one;
\begin{align}
	\nonumber\bs{D} &= \displaystyle \int_{t=0}^\infty \bs f(t)\cdot\bs a(t)\wrt t  %\displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\cfrac{\bs \alpha e^{\bs{S}t}}{\bs \alpha e^{\bs{S}t} \bs e}\wrt t 
	\\\nonumber&= \sum_{k=0}^\infty \int_{k\rho }^{(k+1)\rho } e^{\lambda t} e^{-\lambda t} \bs f(t)\cdot\bs a(t)\wrt t
	\\&= \sum_{k=0}^\infty \int_{0}^{\rho }  e^{\lambda(k\rho +t)}e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs a(k\rho +t)\wrt t. \label{eqn: sc6789gGHJ}
	\intertext{By periodicity, then \(e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs a(k\rho +t)=e^{ -\lambda t} \bs f(t)\cdot\bs a(t)\), hence (\ref{eqn: sc6789gGHJ}) is equal to}
	&\sum_{k=0}^\infty (e^{\lambda \rho })^k\int_{0}^{\rho } e^{\lambda t}e^{-\lambda t}\bs f(t)\cdot\bs a(t)\wrt t \nonumber 
	\\&= \cfrac{1}{1 - e^{\lambda \rho }}\int_{0}^{\rho } \bs f(t)\cdot\bs a(t)\wrt t,\label{eqn: hdsv45KJ}
\end{align}
where the sum converges as it is a geometric series and \(\lambda <0\), \(\rho >0\).

To approximate (\ref{eqn: hdsv45KJ}) numerically, we first partition \(\left[0,\rho \right]\) into \(N\) equal-width intervals \(\left[t_n,t_{n+1}\right)\), where \(t_n = (n-1)\rho /N\), \(n=1,2,...,N+1\). On \(\left[t_n,t_{n+1}\right)\) we approximate the orbit \(\bs a(t)\) by a constant \(\bs a(t) \approx \bs a_n := \cfrac{1}{2}\left( \bs a(t_n) + \bs a(t_{n+1}) \right), \, t \in \left[t_n,t_{n+1}\right)\). Substituting this approximation into the expression for \(\bs{D}\) gives 
\begin{align*}
	\bs{D} &\approx \cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\int_{t_n}^{t_{n+1}} \bs f(t)\cdot\bs a_n\wrt t
	\\&=\cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\left[e^{\bs{S}t_{n+1}}-e^{\bs{S}t_{n}}\right]\bs e \cdot\bs a_n.
\end{align*}
This approximation preserves the property that \(\bs{D}\bs e = \bs e\).

Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \cite{hhat2020}.

\section{The association of \(j\in\calS_0\) with \(\calS_+\) or \(\calS_-\)}\label{sec: zero states}
Let \(X(0)=y_\ell\) and consider the event where \(\left\{\varphi(t)\right\}\) transitions from \(j_0\to j_1\to j_2\) where \(j_0\in\calS_+\), \(j_1\in\calS_0\) and \(j_2\in\calS_-\), before there is a change of level, i.e.~
\[\varphi(t) = \begin{cases}
	j_0 & t\in\left[0,t_1\right), \\
	j_1 & t\in\left[t_1,t_2\right), \\
	j_2 & t\in\left[t_2,t_3\right),
\end{cases}\]
and \(X(t)\in\mathcal D_\ell\), \(t\in[0,t_3)\). 
On approximating this event, the initial orbit position is \(\bs A(0)=\bs \alpha\), and the condition \(X(t)\in\calD_\ell,\, t\in [0,t_3)\) is modelled by \(L(t)=\ell,\, t\in[0,t_3)\). The corresponding sample path of the orbit process on \(t\in[0,t_3)\), is 
\begin{align*}
\bs A(t) &= \begin{cases} 
	\cfrac{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t}}{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
	\cfrac{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t-t_1)}}{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t-t_1)}\bs e} & t\in\left[t_1,t_2\right), \\[10pt]
	\cfrac{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t_2-t_1)}\bs{D}({j_1},{j_2})e^{(\bs{S}_{j_2}+T_{{j_2}{j_2}}\bs{I})(t-t_2)}}{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t_2-t_1)}\bs{D}({j_1},{j_2})e^{(\bs{S}_{j_2}+T_{{j_2}{j_2}}\bs{I})(t-t_2)}\bs e} & t\in\left[t_2,t_3\right),
\end{cases}
%
\\&= \begin{cases} 
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})e^{\bs{S}_{j_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})e^{\bs{S}_{j_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right),
\end{cases}
\end{align*}
for some matrices \(\bs{D}({j_0},{j_1})\) and \(\bs{D}({j_2},{j_1})\). Notice that \(\left\{\bs A(t)\right\}\) is constant on \(t\in\left[t_1,t_2\right)\), and ultimately, \(\left\{\varphi(t)\right\}\) transitions from a positive phase to a negative phase. The matrix product \(\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})\) must capture that there is ultimately a change of phase from \(\calS_+\) to \(\calS_-\). Hence \(\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})\) should be equal to \(\bs{D}\). These types of sample paths are the reason we need to associate states \({j_1}\in\calS_0\) with either \(\calS_+\) or \(\calS_-\).

\subsubsection{Augmented state-space schemes}\label{subsec: augmented }
One way to approach this problem is to augment the state space of the phase process by duplicating \(\calS_0\) and associating one copy of \(\calS_0\) with \(\calS_+\) and one copy of \(\calS_0\) with \(\calS_-\). Let \(\left\{\varphi^*(t)\right\}\) be the augmented CTMC with state space \(\calS^*\) and generator \(\bs{T}^*\). Let \(\calS_+\) and \(\calS_-\) be as before and \(\calS_{m0} = \left\{(m,i)\mid i\in\calS_m\right\}\), \(m\in\left\{+,-\right\}\), then \(\calS^* = \calS_+\cup\calS_-\cup\calS_{+0}\cup\calS_{-0}\). 
The generator of \(\varphi^*(t)\) is
\[\bs{T}^* = \left[\begin{array}{cccc} \bs{T}_{++} & \bs{T}_{+-} & \bs{T}_{+0} & 0 \\ \bs{T}_{-+} & \bs{T}_{--} & 0 & \bs{T}_{-0} \\ \bs{T}_{0+} & \bs{T}_{0-} & \bs{T}_{00} & 0 \\  \bs{T}_{0+} & \bs{T}_{0-} & 0 & \bs{T}_{00} \end{array}\right]. \]
Also define a fluid level \(\left\{X^*(t)\right\}\) using \(\left\{\varphi^*(t)\right\}\), with rates \(c_i^*=c_i\) for \(i\in\calS_+\cup\calS_-\) and \(c_{(m,i)}^* = 0\) for \((m,i)\in\calS_{0+}\cup\calS_{0-}\). The process \(\left\{\varphi(t)\right\}\) is imbedded within \(\left\{\varphi^*(t)\right\}\) and is recovered by marginalising over \(\calS_{0+}\) and \(\calS_{0-}\). Given \(X^*(0)=X(0)\), the fluid levels \(X^*(t)\) and \(X(t)\) match exactly. Hence, if we approximate \(\left\{(X^*(t),\varphi^*(t))\right\}\), then we can recover an approximation to \(\left\{(X(t),\varphi(t))\right\}\). This construction removes the problem of having to choose how to associate states \(j\in\calS_0\) with either \(\calS_+\) or \(\calS_-\)

The generator for the QBD-RAP approximation to the augmented fluid process is 
\begin{align*}
\bs B &= \left[\begin{array}{cccccc}
 \ddots & \ddots & \ddots & & & \\
 & \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & & \\ 
 & &  \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & \\
 & & & \ddots & \ddots & \ddots 
\end{array}\right],
\intertext{where,}
\bs{B}_{0} &= \left[\begin{array}{cccc}
	\bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs{T}_{+0}\otimes \bs{I} & 0 \\
	\bs{T}_{-+}\otimes \bs{D} & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & 0 & \bs{T}_{-0}\otimes \bs{I} \\
	\bs{T}_{0+}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs{T}_{00}\otimes \bs{I} & 0 \\
	\bs{T}_{0+}\otimes \bs{D} & \bs{T}_{0-}\otimes \bs{I} & 0 & \bs{T}_{00}\otimes \bs{I} 
	\end{array}\right],
\end{align*}
\begin{align*}
\bs{B}_{-1} &= \left[\begin{array}{cccc}
	\bs 0 & & & \\
	& \bs{C}_-\otimes (\bs s\bs\alpha) & & \\
	& & \bs 0 & \\
	& & & \bs 0
	\end{array}\right],
%
&& \bs{B}_{+1} = \left[\begin{array}{cccc}
	\bs{C}_+\otimes (\bs s\bs\alpha) & & & \\
	& \bs 0 & & \\
	& & \bs 0 & \\
	& & & \bs 0
	\end{array}\right].
\end{align*}
With this construction, jumps according to the matrix \(\bs D\) occur only on transitions from \(\calS_m\to\calS_n\), or \(\calS_m\to\calS_{0m}\to\calS_{n}\), \(m,n\in\{+.-\}\), \(m\neq n\).

\subsubsection{Other approaches}
There are other possible approaches to model this behaviour which do not require duplicating states and may achieve a small computational saving, however they are less amenable to analysis and possibly lead to less accurate schemes. Instead of duplicating states in \(\calS_0\), we can choose to associate \(j_1\in\calS_0\) with either \(\calS_+\) or \(\calS_-\). If we associate \({j_1}\) with \(\calS_+\), this amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{I}\) and \(\bs{D}({j_1},{j_2})=\bs{D}\); if we associate \({j_1}\) with \(\calS_-\), this amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{D}\) and \(\bs{D}({j_1},{j_2})=\bs{I}\).

There are some consequences to this choice. Let \(k_2\in \calS_+\). Consider an event where the phase process of the fluid queue transitions from \({j_0}\to {j_1}\to k_2\) and there is no change of level. If \({j_1}\) is associated with \(\calS_+\), then \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{I}\) and the corresponding orbit process, given \(\bs A(0) = \bs \alpha\), is 
\begin{align*}
\bs A(t) &= \begin{cases} 
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
\end{cases}
\end{align*}
%This sample path of the orbit process captures the fact that the process initially spent \(t_1-t_0\) amount of time in \({j_0}\in\calS_+\), exactly. Hence, at time \(t_2\), the orbit process knows that it has spent exactly \(t_1-t_0\) amount of time moving at rate \(c_{j_0}\). 
Notice that there is no matrix \(\bs D\) in this expression.

Compare this to if \({j_1}\) is associated with \(\calS_-\). In this case \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{D}\) and the corresponding orbit process of the approximation, given \(\bs A(0) = \bs \alpha\) and there are no change of level in \(\left[t_1,t_3\right)\), is 
\begin{align*}
\bs A(t) &= \begin{cases} 
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
\end{cases}
\end{align*}
Ideally \(\bs D^2=\bs I\), however this is not the case here. Recall that a jump according to \(\bs{D}\) corresponds to approximating the residual life by an expectation. With this interpretation as an approximation, it suggests that we might want to minimise the number of jumps according to \(\bs D\) which occur. Therefore, for \(j_1\in \calS_0\), if transitions \(\calS_+\to j_1\to \calS_+\) occur with high probability compared to transition \(\calS_-\to j_1\to \calS_-\), then this suggests we might want to associate \(j_1\) with \(\calS_+\). Which association is chosen will depend on the parameters of the fluid queue and on which aspects of the model we wish to approximate. Although this advice is based on intuition only, numerical results in Section TBC suggest that it is reasonable. %Hence, at time \(t_2\) the process no longer knows that it has spent exactly \(t_1-t_0\) amount of time moving at rate \(c_{j_0}\); rather, it knows some approximation of this information in the form of an expectation.

%\section{Motivation and construction of approximations}
%
%We can approximate the generator of a fluid queue as a QBD, either by considering a uniformisation of the underlying phase process (as in \cite{bo2013}), or using a finite-volume scheme (see the DG paper, to be submitted). Let \(\boldsymbol M(t) = (X(t),\varphi(t))_{t\geq 0}\) be a fluid queue, with finite state space \(\mathcal S = \left\{1,2,...,N\right\}\), \(N\times N\) generator matrix \(\bs{T}\), and rates \(c_i\) (stored in the \(N\times N\) diagonal matrix \(\bs{C}\)). Also define the \(N\times N\) diagonal matrices \(\bs{C}_+ = \diag( c_i1(c_i>0),\,i\in\mathcal S)\), \(\bs{C}_- = \diag( c_i1(c_i<0),\,i\in\mathcal S)\), and define \(\calD_k = \left[k \Delta, (k+1)\Delta\right]\). 
%
%We approximate the fluid model by the QBD process \(\boldsymbol J(t) = (L(t),\varphi(t))\) with state space \(\mathbb Z\times \mathcal S\) and generator defined by the block matrices \(A_0 = \bs{T}- |\bs{C}|/\Delta\), \(A_1 = |\bs{C}_+|/\Delta\), \(A_{-1} = |\bs{C}_-|\Delta\), where \(|\cdot|\) takes the absolute values of the elements of the matrix. The phase variable of \(\bs J(t)\) and \(\bs M(t)\) match exactly. The process, \(L(t)\Delta\), approximates the fluid height \(X(t)\), and converges as \(\Delta \to 0\) \cite{bo2013}. We also have the approximations 
%\[\bb P(X(t)\in\calD_\ell,\varphi(t)=j\mid X(0)\in\calD_k,\varphi(0)=i)\approx \bb P(L(t) = \ell, \varphi(t) = i \mid L(0) = k, \varphi(0) = i).\] 
%
%These approximations are particularly nice as the ensure positivity and preservation of probability due to their stochastic interpretation. However, they are low order methods compared to more sophisticated numerical schemes such as DG. We wish to develop a numerical approximation method which is high order and has a probabilistic interpretation. 
%
%As motivation, consider again the QBD approximation, \(\bs J(t)\), to a fluid queue, \(\bs M(t)\), as described above. When the state is \((L(t),\varphi(t)) = (\ell, i)\), the process spends an exponentially distributed amount of time in this phase, with rate \(q_{i} = T_{ii}-|c_i|/\Delta\), after which there is a change of phase from phase \(i\) to phase \(j\) with probability \(T_{ij}/|q_i|\) while the level remains the same, otherwise there is a change of level: either \(L(t)\) increases from \(\ell\) to \(\ell +1\) if \(c_i>0\), or decreases from \(\ell\) to \(\ell-1\) if \(c_i<0\). 
%
%Rather than approximate the sojourn time of \((X(t),\varphi(t)) \in(\calD_k, i)\) by an exponential distribution, we could instead approximate this sojourn time by a phase-type distribution denoted \(\tau_i \sim PH(\bs \alpha, T_{ii}\bs{I}+\bs{S}_i)\) where \(-\bs\alpha \bs{S}^{-1}\bs 1 = \Delta\). That is, the mean of the Phase-type \(PH(\bs \alpha,\bs{S})\) is equal to the cell width, \(\Delta\). The distribution \(PH(\bs \alpha, T_{ii}\bs{I}+\bs{S}_i)\) can be viewed as \(min(U,V)\) where \(U\sim exp(T_{ii})\) and \( V\sim PH(\bs\alpha, \bs{S})\). Thus, expiry of the distribution \(\tau_i\) corresponds to either a change in phase if \(U\) is the minimum, or change in level if \(V\) is the minimum. The latter case we map to a change in level: \(L(t)\) moves from \(\ell\) to \(\ell+1\) if \(c_i>0\), and from \(\ell\) to \(\ell-1\) if \(c_i<0\).
%
%Now, we know that, given \(\varphi(t)=i\), and given that \(X(0)=y_k\), then \(X(t)\) has a degenerate distribution which is a point mass at \(t=\Delta\). Thus, we wan to choose \(PH(\bs\alpha, \bs{S})\) such that it approximates the \(\delta\) function as closely as possible. This suggests we choose \(PH(\bs\alpha, \bs{S})\) as an Erlang distribution with mean \(\Delta\) as the Erlang distribution has the lowest coefficient of variation of all PH distributions.
%
%Let \(\calS_+=\left\{i\in\calS\mid c_i>0\right\}\), \(\calS_-=\left\{i\in\calS\mid c_i<0\right\}\). Rather than have exactly the same representation \(PH(\bs\alpha, \bs{S})\) for each \(i\in\calS\), consider instead letting \(PH(\bs\alpha, \bs{S})\) for each \(i\in\calS_+\) and \(PH(\wt{\bs\alpha},\wt{\bs{S}})\) for each \(i \in \calS_-\), where \(PH(\wt{\bs\alpha},\wt{\bs S})\) is the time reverse representation. Assume without loss that the phases are ordered such that \(i<j\) for all \(i\in\calS_+,\,j\in\calS_-\). Construct an approximation to the distribution of \(X(t)\in\calD_k\) given \(X(0)=y_k,\,\varphi(0)=i,\, c_i>0\) as \(PH(\bs e_i \otimes \bs \alpha, \bs{Q})\) where 
%\[\bs{Q} = \left[\begin{array}{ccccc}
%	T_{11}\bs{I}+|c_1|\bs{S} & T_{12}\bs{I} & \hdots & T_{1,N-1}\bs{I} & T_{1,N}\bs{I} \\
%	T_{21}\bs{I} & T_{22}\bs{I}+|c_2|\bs{S} & \hdots & T_{2,N-1}\bs{I} & T_{2,N}\bs{I} \\
%	 \vdots & \vdots &  & \vdots & \vdots \\ 
%	 T_{N-1,1}\bs{I} & T_{N-1,2}\bs{I} & \hdots & T_{N-1,N-1}\bs{I}+|c_{N-1}|\wt{\bs S} & T_{N-1,N}\bs{I} \\ 
%	 T_{N,1}\bs{I} & T_{N,2}\bs{I}  & \hdots & T_{N,N-1}\bs{I} & T_{NN}\bs{I}+|c_N|\wt{\bs S} \\ 
%\end{array}\right].\]
%Thus, when in phases \(\varphi=i\) with \(c_i>0\), the Markov chain associated with \(\wt{\bs S}\) evolves \emph{forward} in time. When in phases \(\varphi=i\) with \(c_i<0\), the Markov chain associated with \(\wt{ \bs S}\) evolves \emph{backward} in time. When the process jumps from phase \(\varphi = i\) to phase \(\varphi = j\), the state of the Markov chain associated with \(\bs{S}\) (or \(\wt{\bs S}\)) is preserved. In this way the approximation can capture the \emph{memory} of the process through the state of \(\bs S/\wt{\bs S}\).
%
%Recall, that we claimed that we should choose \(PH(\bs\alpha, \bs{S})\) as an Erlang distribution in order to achieve the best approximation. So suppose we choose an \(Erlang(n,1/(n\Delta))\) representation. We can show that this is actually exactly the same approximation as if we had constructed the approximation via the uniformisation method with uniformisation parameter \(1/(n\Delta)\). 
%
%However, rather than use a \(PH(\bs\alpha, \bs{S})\) distribution to approximate the \(\delta\) function, what if we instead used a matrix exponential distribution. That is, choose \(ME(\bs\alpha, \bs{S})\) as a \emph{concentrated matrix exponential distribution}. If we could find a time-reverse version of \(ME(\bs\alpha, \bs{S})\), then we might be able to use the same structure as above to construct a high-order approximation to a fluid model with a stochastic interpretation (and therefore preserve positivity of the approximate solution). 
%
%\section{Time reverse ME/RAPs}
%\paragraph{Time-reverse PH renewal processes}
%Let \(\left\{X(t)\right\}_{t\in(-\infty,\infty)}\) be a CTMC with generator \(\bs{Q}=\left[q_{ij}\right]_{i,j\in\calS}\) and state space \(\calS\) and assume that \(\left\{X(t)\right\}\) is stationary with stationary distribution \(\bs\pi\). Let \(\wt X(t)=X(-t)\) be the time reverse of \(\left\{X(t)\right\}\). Recall that the generator of \(\wt X(t)\) is given by \(\wt{\bs Q}\), with entries \(\wt q_{ij} = \cfrac{\pi_j q_{ji}}{\pi_i}\). The processes \(X(t)\) and \(\wt X(t)\) share the same stationary distribution. Let \(P=\diag(\pi)\). 
%
%Consider now a PH renewal process defined by \(PH(\bs\alpha, \bs{S})\). Ignoring renewals jumps, the marginal phase process evolves according to the generator \(\bs{Q} = \bs{S}+\bs s\bs \alpha \) where \(\bs s = -\bs{S}\bs 1\). 
%
%Let \(\wt{ \bs S}=P^{-1}{\bs S}'P\), and \(\wt{\bs\alpha} = \bs s'P\mu\) where \(M'\) denotes the transpose of \(M\) and \(\mu = -\bs\alpha \bs{S}^{-1}\bs 1\). Then \(\wt{\bs Q} = \wt{\bs S}+\wt{ \bs s}\wt{\bs \alpha}\) where \(\wt{\bs s}=-\bs{S}\bs 1\), is the time-reverse representation of \(PH(\bs\alpha, \bs{S})\). That is, \(PH(\bs\alpha, \bs{S})\) and \(PH(\wt{\bs\alpha}, \wt{\bs S})\) are representations of the same distribution and the PH renewal process can be defined in terms of either representation. 
%
%\paragraph{Time-reverse (?) ME renewal processes} 
%Consider now a ME renewal process defined by \(ME(\bs\alpha, \bs{C})\) which we assume is minimal and of order \(p\). Can we come up with some time-reverse representation of this renewal process. To this end, we now detail some properties of ME renewal processes. 
%
%Let \(\bs{D} = \bs c\bs \alpha\) where \(\bs c = -\bs{C}\bs 1\). Asmussen and Bladt \cite{ab1999} gave such a process a stochastic interpretation (in fact, Asmussen and Bladt worked with rational arrival processes, but I think we need only consider ME renewal processes). Let \(V = span\left\{\alpha, \bs\alpha \bs{S},...,\bs\alpha \bs{S}^{p-1}\right\}\). There exists a vector-valued stochastic process \(\left\{\bs A(t)\right\}_{t}\) with \(\bs A(t)\in V\) for all \(t\) and \(\bs A(0)=\bs\alpha\). Suppose \(\{Y_k\}\), \(k=0,1,...\) are the renewal times, and \(Y_0=0\). Between renewal epochs \(\bs A(t)\) moves deterministically according to \(\bs A(t) = \cfrac{\bs A(Y_{N(t)}) e^{\bs{C}(t-Y_{N(t)})}}{\bs A(Y_{N(t)}) e^{\bs{C}(t-Y_{N(t)})}\bs 1}\). \(\bs A(t)\) jumps at rate \(\bs A(t)\bs{D}\bs 1\) at time \(t\), and upon a jump \(\bs A(Y_k) = \cfrac{\bs A(Y_k-) \bs{D}}{\bs A(Y_k-) \bs{D}\bs 1}=\bs \alpha\) where \(Y_k-\) is the left limit at \(Y_k\). Furthermore, \(\left\{\bs A(t)\right\}\) is a strong Markov process and \(E\left[\bs A(t+s) \mid \bs A(t)\right] = \bs A(t)e^{\bs{Q}t}\) where \(\bs{Q} = \bs{C}+\bs{D}\). We know \(\bs{Q}\bs 1=\bs 0\) and \(dev(\bs{Q})=0\), that is, the dominant eigenvalue of \(\bs{Q}\) is 0.
%
%Let \(\bs \pi\) be the left eigenvector of \(\bs{Q}\) corresponding to eigenvalue 0. Assume that \(\pi_i\neq0\), \(i=1,...,p\). If this is not the case then construct an invertible matrix \(\bs{V}\) such that \(\bs{V}\bs 1=\bs 1\), then consider the equivalent representation of the ME renewal process with \(ME(\bs\alpha \bs{V}^{-1}, \bs{V}\bs{C}\bs{V}^{-1})\), and pray that \(\pi_i\neq0\), \(i=1,...,p\) (can we show that there is always such a \(\bs{V}\)?) By choosing \(\bs{V}\) such that \(\bs{V}\bs 1=\bs 1\), then we maintain that the right eigenvector corresponding to \(\bs{V}\bs{C}\bs{V}^{-1}\) is still \(\bs 1\). 
%
%Let \(\wt{\bs C}=\bs{P}^{-1}\bs{C}'\bs{P}\), \(\wt{ \bs D} = \bs{P}^{-1}\bs{D}'\bs{P}\), and \(\wt{\bs\alpha} = \bs c'\bs{P}\mu\) where \(\mu = -\bs\alpha \bs{C}^{-1}\bs 1\). Is the ME renewal process defined by \(ME(\wt{\bs\alpha},\wt{ \bs C})\) in any meaningful way, the reverse of the ME renewal process defined by \(ME(\bs\alpha, \bs{C})\)?
%
%Let \(\tau_1,...,\tau_K\) be inter-arrival times for the renewal process defined by \(ME(\bs \alpha, \bs{C})\). Then joint density is of these arrivals is 
%\[\bs \alpha e^{\tau_1\bs{C}}\bs{D}e^{\tau_2\bs{C}}\bs{D}...e^{\tau_K\bs{C}}\bs c.\]
%Now let \(\wt \tau_1,...,\wt \tau_K\) be inter-arrival times for the renewal process defined by \(ME(\wt{\bs \alpha},\wt{\bs C})\).
%Then joint density is
%\begin{align*}
%	\wt{\bs \alpha} e^{\wt \tau_1\wt{\bs C}}\wt{ \bs D}e^{\wt \tau_2\wt{ \bs C} }\wt{ \bs D}...e^{\wt \tau_K\wt{ \bs C}}\wt{\bs c} 
%	&= \bs c'\bs{P}\mu e^{ \wt \tau_1 \bs{P}^{-1}\bs{C}'\bs{P}} \bs{P}^{-1}\bs{D}'\bs{P}e^{ \wt \tau_2 \bs{P}^{-1}\bs{C}'\bs{P}} \bs{P}^{-1}\bs{D}'\bs{P}...e^{ \wt \tau_K \bs{P}^{-1}\bs{C}'\bs{P}}{\bs{P}^{-1}\bs{C}'\bs{P}\bs 1}
%	%
%	\\&= \bs c'\bs{P}\mu \bs{P}^{-1} e^{ \wt \tau_1 \bs{C}'} \bs{P} \bs{P}^{-1}\bs{D}'\bs{P}\bs{P}^{-1}e^{ \wt \tau_2 \bs{C}'} \bs{P}\bs{P}^{-1}\bs{D}'\bs{P}...\bs{P}^{-1}e^{ \wt \tau_K \bs{C}'}\bs{P}{\bs{P}^{-1}\bs{C}'\bs{P}\bs 1}
%	%
%	\\&= \mu \bs c' e^{ \wt \tau_1 \bs{C}'} \bs{D}'e^{ \wt \tau_2 \bs{C}'} \bs{D}'...e^{ \wt \tau_K \bs{C}'}{\bs{C}'\bs \pi'}.
%\end{align*}
%Now use the fact that \(\bs \pi = -\bs \alpha \bs{C}^{-1} / \mu\) (shown in \cite{ab1999}) to get that the joint density is equal to 
%\[\bs c' e^{ \wt \tau_1 \bs{C}'} \bs{D}'e^{ \wt \tau_2 \bs{C}'} \bs{D}'...e^{ \wt \tau_K \bs{C}'}{\bs \alpha'} = \bs \alpha e^{\wt \tau_K\bs{C}}\bs{D}e^{\wt \tau_{K-1}\bs{C}}\bs{D}...e^{\wt \tau_1\bs{C}}\bs c,\]
%since transposing a scalar does noting. Therefore the joint density of \(\tau_1,...,\tau_K\) is the same as that of \(\wt \tau_K,...,\wt \tau_1\).
%
%\paragraph{Time-reverse using Baye's rule} 
%For Markov processes we find the time reverse probabilities using Bayes' rule, 
%\begin{align*}
%	&\wt p_{ij}(t) = P(\wt X(t)=j\mid \wt X(0)=i) = P( X(0)=j\mid X(t)=i) 
%	= \cfrac{P( X(t)=i\mid X(0)=j)P(X(0)=j)}{P(X(t)=i)} 
%	\\& = \cfrac{P( X(t)=i\mid X(0)=j)\pi_j}{\pi_i} = \cfrac{p_{ji}(t)\pi_j}{\pi_i}
%\end{align*}
%
%Do we want to do the same for \(\bs A(t)\)? 
%\begin{align*}
%	&P(\wt{\bs A}(t)=\bs j\mid \wt{\bs A}(0)=\bs i) = P( {\bs A}(0)=\bs j\mid {\bs A}(t)=\bs i) 
%	= \cfrac{P( {\bs A}(t)=\bs i\mid {\bs A}(0)=\bs j)P({\bs A}(0)=\bs j)}{P({\bs A}(t)=\bs i)}.
%\end{align*}
%But now the state space is not discrete, how do we do this for all \(\bs i,\,\bs j\)? I suspect we just need to do it for a linearly independent selection of \(\bs i\), \(\bs j\). What even are \(P( {\bs A}(t)=\bs i\mid {\bs A}(0)=\bs j)\), \(P({\bs A}(0)=\bs j)\) and \(P({\bs A}(t)=\bs i)\). Is \(P({\bs A}(t)=\bs i)\) time-dependent. 
%
%\paragraph{An adjoint process?}
%Asmussen and Bladt \cite{ab1999} showed that 
%\[\mathbb P (\theta_tN\in\cdot \mid \mathcal F_{\left[0,t\right]}) = \bs A(t)\bs v(\cdot),\]
%where \(\bs v\) is a columns vector of basis measures, and \(\theta_t\) is the shift operator, \(\theta_{t}N\left[s,s+u\right)=N\left[t+s,t+s+u\right)\).
%Given the result
%\begin{align*}
%	\wt{\bs \alpha} e^{\wt \tau_1\wt{\bs C}}\wt{ \bs D}e^{\wt \tau_2\wt{\bs C}}\wt{ \bs D}...e^{\wt \tau_K\wt{\bs C}}\wt{\bs c} 
%	&= 
%	\bs c' e^{ \wt \tau_1 \bs{C}'} \bs{D}'e^{ \wt \tau_2 \bs{C}'} \bs{D}'...e^{ \wt \tau_K \bs{C}'}{\bs \alpha'}
%\end{align*}
%is there some sort of \emph{adjoint} process \(\wt{\bs A}(t)\), a column vector, such that 
%\[\mathbb P(\theta_{-t} N \in \cdot \mid \mathcal F_{\left[t,\infty\right)}) = {\wt{\bs v}(\cdot)} \wt{\bs A}(t),\]
%where \(\theta_t\) is the shift operator \(\theta_{-t}N\left[t+s,t+s+u\right)=N\left[s,s+u\right)\), and \(\wt{\bs v}(\cdot)\) is a row-vector of basis measures. If so, do we have \(\mathbb E\left[\wt{\bs A}(t-s)|\mathcal F_{\left[t,\infty\right)}\right] = e^{\wt{\bs Q}s}\wt{\bs A}(t),\) and \(\wt{\bs A}(t) = \cfrac{e^{\wt{\bs Q}s}\wt{\bs A}(t)}{\bs e e^{\wt{ \bs Q}s}\wt{\bs A}(t)}\) on \(N\left[0,t\right]=0\)?
%
%\section{Phase-Type and ME renewal theory}
%Let \(\left\{N(t)\right\}\) be a stationary point process with \(ME(\bs \alpha, \bs{Q})\) inter-arrival times. Define \(\bs\pi=\bs\alpha (-\bs{Q}^{-1})/\mu\). Let \(C_t\) be the \emph{Age} of the process at time \(t\); that is \(C_t = t-Y_{N{t}}\) where \(Y_{N(t)}\) is the last event before time \(t\). Let \(R_t\) be the \emph{Residual} time; the time from \(t\) until we see the next arrival; \(R_t = Y_{N(t)+1}-t\). The spread \(S_t\) is time between arrivals if we view the process at time \(t\). \(S_t=C_t+R_t\). It is well-known that for any stationary renewal process \(C_t\) and \(R_t\) have the same distribution and that \(S_t\) has density \(\propto x\psi(x)\) where \(\psi(x)\) is the inter-arrival density. 
%
%For an ME renewal process we know that \(C_t\sim ME(\bs\pi, \bs{Q})\) and \(R_t\sim ME(\bs\pi, \bs{Q})\) \cite{ab1999}. Bladt and Nielsen use algebraic arguments to show that \(S_t\) has a matrix exponential distribution with representation 
%\[\left(\left[ \cfrac{\bs\alpha \bs{Q}^{-1}}{\bs\alpha \bs{Q}^{-1} \bs e}\,, \bs 0\right], \left[\begin{array}{cc}\bs{Q}&-\bs{Q}\\0&\bs{Q}\end{array}\right]\right).\]
%
%Furthermore, for when the ME renewal process is a PH renewal process, they use the stochastic interpretation of the PH distribution as time until absorption of a finite-state Markov chain to argue that \(S_t\) also is also a PH distribution. 
%
%We can give the ME representation of the spread a stochastic interpretation. Let us first consider the age, \(A\sim ME(\bs\pi,\bs{Q})\). Corresponding to the age is an orbit process; \(\bs A(s) = \bs\pi e^{\bs{Q}s}/(\bs\pi e^{\bs{Q}s}\bs e)\), which is entirely deterministic on \(s\in\left[0,A\right]\). At time \(Y_{N(t)}\) the age orbit is initialised. If we observe the process \(s\) amount of time later, then the age is \(A=Y_{N(t)}-t=s\) and the corresponding position of the orbit is \(\bs A(t) = \bs\pi e^{\bs{Q}s}\). On \(A=s\), the residual time has distribution \(ME\left(\cfrac{\bs\alpha e^{\bs{Q}s}}{\bs\alpha e^{\bs{Q}s}\bs e},\bs{Q}\right)\). That is, \(R_t\mid A=s \sim ME\left(\cfrac{\bs\alpha e^{\bs{Q}s}}{\bs\alpha e^{\bs{Q}s}\bs e},\bs{Q}\right)\). However, \(\bs A(s) \propto \bs\pi e^{\bs{Q}s} = \bs\alpha (-\bs{Q}^{-1}) e^{\bs{Q}s} = \bs\alpha e^{\bs{Q}s}(-\bs{Q}^{-1})\) and therefore \(\cfrac{\bs A(s)(-\bs{Q})}{\bs A(s)(-\bs{Q})\bs e} = \cfrac{\bs\alpha e^{\bs{Q}s}}{\bs\alpha e^{\bs{Q}s}\bs e}\). Thus we may think of the spread as the evolution of an orbit process. At time \(Y_{N(t)}\) initialise the age orbit process \(\bs A(0) = \bs \pi\) and evolve according to \(\bs A(s) =  \bs\pi e^{\bs{Q}s}/(\bs\pi e^{\bs{Q}s}\bs e)\). The age process expires with intensity \(\bs A(u) (-\bs{Q})\bs e\). Upon expiry we realise the age \(A=s\) and the orbit process jumps from \(\bs A(s-)\) to \(\cfrac{\bs A(s)(-\bs{Q})}{\bs A(s)(-\bs{Q})\bs e}\). The orbit process now continues to evolve along \(\bs A(u) = \cfrac{\bs A(s)(-\bs{Q})e^{\bs{Q}{(u-s)}}}{\bs A(s)(-\bs{Q})e^{\bs{Q}{(u-s)}}\bs e}\) until the residual time expires which occurs with intensity \(\bs A(s)(-\bs{Q})\bs e\). 
%
%An alternative representation of the spread is found by applying the similarity transform \(\bs{D} = \left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]\), where \(\Delta = \diag((-\bs{Q}^{-1})\bs e)\). The new representation is 
%\[ME(\bs \beta, U)\] 
%where 
%\[\bs \beta = \left[\begin{array}{cc}\bs \pi & 0\end{array}\right]\left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]=\left[\begin{array}{cc}\cfrac{\bs \alpha(-\bs{Q}^{-1})}{\bs \alpha(-\bs{Q}^{-1})\bs e} & 0\end{array}\right]\left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]=\left[\begin{array}{cc}\cfrac{\bs \alpha\Delta}{\bs \alpha(-\bs{Q}^{-1})\bs e} & 0\end{array}\right], \]
%and 
%\[U = \bs{D}^{-1}\bs{Q}\bs{D}=\left[\begin{array}{cc}-\Delta^{-1}\bs{Q}^{-1} & \\ & \bs{I}\end{array}\right] \left[\begin{array}{cc}\bs{Q}&-\bs{Q}\\0&\bs{Q}\end{array}\right] \left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]%=\left[\begin{array}{cc}\Delta^{-1}&\Delta^{-1}\\0&\bs{Q}\end{array}\right] \left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]
%=\left[\begin{array}{cc}\Delta^{-1}\bs{Q}\Delta&\Delta^{-1}\\0&\bs{Q}\end{array}\right] \]
%and the closing vector is unchanged \(\bs u = -U\bs e =  \left[\begin{array}{c} 0 \\ -\bs{Q}\bs e \end{array}\right]\). 
%
%Under this representation, the orbit process is initialised according to \(\bs A(0)=\cfrac{\bs \alpha\Delta}{\bs \alpha(-\bs{Q}^{-1})\bs e} \), evolves along \(\bs A(s)=\cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}}{{\bs \alpha\Delta}e^{\bs{Q}s}\bs e}\) and expires with intensity \(\bs A(s)\Delta^{-1}=\cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}}{{\bs \alpha\Delta}e^{\bs{Q}s}\bs e}\Delta^{-1}\bs e\). At expiry the age is realised, \(C_t=s\), and the orbit process jumps from \(\bs A(s-) = \cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}}{{\bs \alpha\Delta}e^{\bs{Q}s}\bs e}\) to \(\bs A(s) = \cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}\Delta^{-1}}{{\bs \alpha\Delta}e^{\bs{Q}s}\Delta^{-1}\bs e}\).
%
%When the ME inter-arrival times are also Phase-type, then this representation for the spread is also Phase-type.

%\section{SFM approximation}
%%\paragraph{SFMs} Let \(\left\{\varphi(t)\right\}_{t\geq 0}\) be an irreducible continuous-time Markov Chain (CTMC) with a finite state space \(\mathcal S = \left\{1,...,N\right\}\) and infinitesimal generator \(\bs{T} = \left[T_{ij}\right]_{i,j\in\mathcal S}\). Let \(\left\{(X(t),\varphi(t))\right\}_{t\geq0}\) be a Markovian stochastic fluid model (SFM) with phase variable \(\varphi(t)\) and level variable \(X(t)\in (-\infty,\infty)\). For each \(i\in\mathcal S\) there are rates \(c_i\in\mathbb R\) such that \(\cfrac{\wrt X(t)}{\wrt t} = c_i\) when \(\varphi(t)=i\). The variable \(\left\{X(t)\right\}\) is known as the buffer, or level, of the fluid process. 
%
%Partition the set of all phases, \(\mathcal S\), into \(\mathcal S_+ = \left\{i\in\mathcal S\mid c_i>0\right\},\, \mathcal S_- = \left\{i\in\mathcal S\mid c_i<0\right\}, \,\mathcal S_0 = \left\{i\in\mathcal S\mid c_i=0\right\},\). Use the same partition to partition the generator 
%\[\bs{T} = \left[\begin{array}{ccc} \bs{T}_{++} & \bs{T}_{+-} & \bs{T}_{+0} \\ \bs{T}_{-+} & \bs{T}_{--} & \bs{T}_{-0} \\ \bs{T}_{0+} & \bs{T}_{0-} & \bs{T}_{00} \end{array}\right].\]
%Let \(\bs{C}_+ = \diag(c_i,i\in\mathcal S_+)\), and \(\bs{C}_- = \diag(|c_i|,i\in\mathcal S_-)\).

%Let \(Z\sim ME_p(\bs \alpha, \bs{S}, \bs s)\) of dimension \(p\) with \(\bs\alpha\bs e = 1\), \(\bs s = -\bs{S}\bs e\) and mean \(\Delta\). Let \(\{\bs A(t)\} = \left\{\bs A(t)\right\}_{t\geq 0}\) be a \(N\times p\)-dimensional vector-valued stochastic process which is a piecewise-deterministic Markov process. The sample path behaviour of \(\{\bs A(t)\}\) is as follows. For some \(i\in \mathcal \bs{S}\), let \(\bs a_i\) be a vector such that \(\bs a_i\bs e = 1\) and \((\bs a_i,\bs{S},\bs s)\) is a representation of a ME distribution. Let 
%\[Q = \left[\begin{array}{ccc}  \bs{C}_+ \otimes \bs{S} &&\\ &   \bs{C}_- \otimes \bs{S} & \\ &&0\end{array}\right],\]
%and \(\bs q = -Q\bs e\). 
%Given \(\bs A_0 = \bs a = (\bs 0_{(i-1)p},\bs a_i,\bs 0_{(N-i-1)p})\), \(\{\bs A(t)\}\) evolves deterministically according to
%\[\bs A(t) = \cfrac{\bs a e^{Q t}}{\bs a e^{Q t} \bs e}\]

%Let \(Z\sim ME_p(\bs \alpha, \bs{S}, \bs s)\) of dimension \(p\) with \(\bs\alpha\bs e = 1\), \(\bs s = -\bs{S}\bs e\). For each \(i\in\mathcal S\), define \(\bs{S}_i = |c_i|\bs{S}\) and \(\bs s_i = |c_i|\bs s\). Let \(\{\bs A(t)\} = \left\{\bs A(t)\right\}_{t\geq 0}\) be a \(p\)-dimensional vector-valued stochastic process which is piecewise-deterministic and c\`adl\`ag. Additionally, let \(\left\{{\phi}(t)\right\}_{t\geq 0}\) be a real-valued stochastic process. Together, \(\left\{(\bs A(t),{\phi}(t))\right\}_{t\geq 0}\), form a Markov process, however, \(\{\bs A(t)\}\) alone is not Markovian. We will show later that \(\left\{{\phi}(t)\right\}\) is a CTMC the same as \(\left\{\varphi(t)\right\}\). The sample path behaviour of \(\{\bs A(t)\}\) and \({\phi}(t)\) is as follows. 
\section{The dynamics of the QBD-RAP approximation}\label{sec: dynamics}
We now have all the elements we need to describe the dynamics of the QBD-RAP approximation. Recall that a QBD-RAP is a stochastic process \(\{(L(t),\bs A(t))\}_{t\geq0}\) where \(L(t)\) is a discrete, level process and \(\bs A(t)\) is a piecewise deterministic orbit process. Since the phase dynamics are a CTMC, we choose to use an alternate notation, \(\{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process. We choose this representation to make explicit how the approximation captures the phase dynamics of the fluid queue. We will show later that \(\phi(t)\) captures the phase dynamics of the fluid queue exactly, provided the phase process is independent of the fluid level \(\{X(t)\}\). We use the level \(L(t)=\ell\) to approximate which band \(\calD_\ell\) that \(X(t)\) is in at time \(t\), and use the orbit \(\bs A(t)\) to obtain approximations about where \(X(t)\) is within the interval \(\calD_\ell\).

We now proceed to describe the evolution of the orbit and phase processes, before introducing the level variable later.

\paragraph{The orbit process and phase dynamics} On \({\phi}(t)=i\), between event epochs, the process \(\{\bs A(t)\}\) evolves deterministically according to the differential equation 
\begin{align}\label{eqn: At DE}
\cfrac{\wrt}{\wrt t}\bs A(t) = \bs A(t)(\bs{S}_i - \bs A(t) (\bs{S}_i + T_{ii}\bs{I}_p)\bs e)\bs{I}_p.
\end{align}
Let \(\bs a_0\) be a vector such that \(\bs a_0\bs e = 1\) and \((\bs a_0,\bs{S},\bs s)\) is a representation of a ME distribution. Given no events occur before time \(t+u\), \(\bs A(u) = \bs a_0\) and \({\phi}(u)=i\), the solution to (\ref{eqn: At DE}) states that \(\bs A({t+u})\) evolves deterministically according to 
\[\bs A({t+u})\mid \left\{\bs A(u)=\bs a_0\right\} =  \cfrac{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}}{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t} \bs e} = \cfrac{\bs a_0 e^{\bs{S}_i t}}{\bs a_0 e^{\bs{S}_i t} \bs e}.\]

At time \(t\), given \(\bs A(t)\), an event occurs at rate 
\[\bs A(t) (\bs{S}_i - T_{ii}\bs{I}_p)\bs e = \bs A(t)\bs s_i - T_{ii}.\]
More precisely, an event corresponding to a change in phase for \({\phi}(t)\) occurs at rate \(-\bs A(t) T_{ii}\bs e = -T_{ii}\) and an event corresponding to a \emph{change of level} occurs at rate \(-\bs A(t) \bs{S}_i\bs e = \bs A(t) \bs s_i\). Later, we will make clear why we say that the latter event corresponds to a change of level. Upon an event occurring at time \(t\), with probability \(\cfrac{-T_{ii}}{-T_{ii} + \bs A({t^-})\bs s_i}\) the event corresponds to a change of phase and with probability \(\cfrac{\bs A({t^-})\bs s_i}{-T_{ii} + \bs A({t^-})\bs s_i}\) the event corresponds a change of level.

Upon an event corresponding to a change of level occurring at time \(t\) the process \(\{\bs A(t)\}\) jumps to
\(\bs A(t) = \bs \alpha\).

Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), there are two possibilities; either \(sign(c_i)=sign(c_j)\), or \(sign(c_i)\neq sign(c_j)\). As discussed earlier, for states \(i\in\mathcal S_0\) we must specify some association with either \(\mathcal S_+\) or \(\mathcal S_-\). Here we choose the augmented state space approach and duplicate \(\calS_0\) and associate one copy with \(i\in\calS_+\) and one copy with \(\calS_-\); call these \(\calS_{+0}\) and \(\calS_{-0}\), respectively. We take \(sign(c_i)=+\) for \(i\in\calS_{+0}\) and \(sign(c_i)=-\) for \(i\in\calS_{-0}\).

Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)=sign(c_j)\), at the time of the event \(\bs A(t)\) is unchanged but begins to evolves according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), while \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).

Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)\neq sign(c_j)\), the process \(\{\bs A(t)\}\) jumps to \(\bs A(t) = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{\bs A({t^-}) T_{ij} \bs{D} \bs e} = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{T_{ij} } = \bs A({t^-}) \bs{D} \) and then proceeds to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), and \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).

The two scenarios, \(sign(c_i)\neq sign(c_j)\) and \(sign(c_i)= sign(c_j)\), can be written succinctly by stating that, at the time of an event corresponding to a change of phase from \(i\) to \(j\), \(\{\bs A(t)\}\) jumps to \(\bs A(t)\bs{D}^{1(sign(c_i)\neq sign(c_j))}\) and begins to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), meanwhile \(\left\{{\phi}(t)\right\}\) jumps from \(\phi(t^-)=i\) to \(\phi(t)=j\). Moreover, given a change of phase event occurs at time \(t\) with \(\phi(t^-)=i\) then, with probability \(\cfrac{\bs A(t^-)T_{ij}\bs{D}^{1(sign(c_i)\neq sign(c_j))}\bs e}{-\bs A(t^{-})T_{ii}\bs e} = \cfrac{T_{ij}}{-T_{ii}}\), the event corresponds to a change of phase from \(i\) to \(j\).

The following result states that \(\left\{{\phi}(t)\right\}\) has the same distribution as \(\left\{{\varphi}(t)\right\}\) when the fluid queue is unbounded, the latter being the phase process of the fluid queue. 
\begin{thm}\label{thm: 1}
	Let \(\Theta_i\) be the time at which the first jump of the phase process of the unbounded QBD-RAP, \(\left\{{\phi}(t)\right\}\), occurs given \(\phi(0)=i\). For any initial orbit \(\bs a\in\mathcal A\), then \(\Theta_i\) has an exponential distribution with rate parameter \(|T_{ii}|\). Furthermore, given \({\phi}(t)\) leaves state \(i\), it jumps to state \(j\) with probability \(T_{ij}/|T_{ii}|\). Hence \({\phi}(t)\) and \(\varphi(t)\) have the same probability law. 
\end{thm}
\begin{proof}
	Let \(\{\tau_n\}_{n\geq 0}\) with \(\tau_0=0\) and \(\tau_n\) the time of the \(n\)-th change of level of the QBD-RAP. Consider partitioning \(\left\{\Theta_i\leq t \right\}\) with respect to \(\left\{\tau_{n-1}< t \leq\tau_n\right\},\,n=1,2,\dots\). For \(n=1\) we can write
	\begin{align*}
		&\mathbb P(\Theta_i\leq  t, \tau_0< t\leq \tau_1 \mid \bs A(0) = \bs a) 
		= \bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}\bs e 
		\intertext{and since \(T_{ii}\bs{I}_p\) commutes with \(\bs{S}_i\) and \(e^{T_{ii}t}\) is a scalar, then this is equal to}
		&\bs a e^{\bs{S}_it}e^{T_{ii}\bs{I}_pt}\bs e  
		= \bs a e^{\bs{S}_it}\bs e e^{T_{ii}t} 
		= \mathbb P(\tau_0 < t\leq \tau_1)e^{T_{ii}t} .
	\end{align*}
	For \(n>1\), by partitioning on the times of the first \(n-1\) level changes, \(\tau_1,\dots,\tau_n\), we get 
	\begin{align*}
		&\mathbb P(\Theta_i\leq t, \tau_{n-1}< t \leq \tau_n \mid \bs A(0) = \bs a) 
		\\&=\displaystyle{\int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t} \bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t_1}\bs s_i
			\left(\prod_{k=2}^{n-1} \bs \alpha e^{(\bs{S}_i+T_{ii}\bs{I}_p)(t_k-t_{k-1})} \bs s_i\right) 
			\\&\qquad\displaystyle{\times\bs \alpha e^{(\bs{S}_i+T_{ii}\bs{I}_p)(t-t_{n-1})}\bs e}\wrt t_{n-1}\wrt t_{n-2}\hdots\wrt t_{1}
		\intertext{where \(t_k\) is the time of the \(k\) change of level. Since \(T_{ii}\bs{I}_p\) commutes with \(\bs{S}_i\), \(e^{T_{ii}t_k},\,k=1,...,n-1\) are scalars, and \(t_1+(t_2-t_1)+\hdots+(t_{n-1}-t_{n-2})+(t-t_{n-1})=t\), then this is equal to}
		& \displaystyle{\int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t} \bs a e^{\bs{S}_it_1}\bs s_i 
			\left(\prod_{k=2}^{n-1} \bs \alpha e^{\bs{S}_i(t_k-t_{k-1})} \bs s_i\right) \bs \alpha e^{\bs{S}_i(t-t_{n-1})}\bs e
			\times e^{T_{ii}t} \wrt t_{n-1}\wrt t_{n-2}\hdots\wrt t_{1}
		\\&= \mathbb P(\tau_{n-1}< t \leq \tau_n)\displaystyle{  e^{T_{ii}t} }.
	\end{align*}
	Hence, by the law of total probability, 
	\begin{align*}
		\mathbb P(\Theta_i\leq t) &= \sum_{n=1}^\infty \mathbb P(\Theta_i\leq t, \tau_{n-1} < t \leq \tau_n)
		\\&= \sum_{n=1}^\infty \mathbb P(\tau_{n-1} < t \leq \tau_n)\displaystyle{  e^{T_{ii}t} }
		\\&= e^{T_{ii}t},
	\end{align*}
	and therefore \(\Theta_i\) has an exponential distribution with rate \(|T_{ii}|\).

	Upon leaving state \(i\) at time \(t\), \({\phi}(t)\) transitions to state \(j\) with probability 
	\begin{align*}
            	\cfrac{\left(\cfrac{\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e}{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e + \bs A(t) \bs s_i}\right)}
            	{\left(\cfrac{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e}{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e + \bs A(t) \bs s_i}\right)}
            	&=\cfrac{\bs A(t) \bs eT_{ij}}
            	{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs eT_{ij}}
	= \cfrac{T_{ij}}
            	{-T_{ii}}.
	 \end{align*}
	 Therefore the process \(\left\{{\phi}(t)\right\}\) is has the same probability law as \(\left\{\varphi(t)\right\}\).
\end{proof}
\begin{rem}
The same result can be shown for a regulated boundary. For boundary conditions which interact with the phase dynamics, such as a reflecting boundary, the result does not hold. The cause is the fact that the phase dynamics are level dependent -- we see a forced change of phase upon a boundary being hit -- and the QBD-RAP can only approximate the level process of the fluid queue. However, until a boundary is hit (by either the fluid queue or QBD-RAP) then the phase processes match. We show later that, in the limit as the variance of the the matrix exponential distribution used in the construction of the QBD-RAP goes to zero, then the dynamics of the level process of the fluid queue, \(X(t)\), are captured by the QBD-RAP, and boundary behaviour which interacts with the phase dynamics can be captured too.
\end{rem}

Since the phase processes \(\left\{{\phi}(t)\right\}\) and \(\left\{\varphi(t)\right\}\) have the same law when boundaries are not present, henceforth, we shall assume \(\left\{{\phi}(t)\right\}\) and \(\left\{\varphi(t)\right\}\) are coupled when possible (they share the same sample path). Specifically, in Section~\ref{sec: no change}, we will analyse the QBD-RAP on the event that it remains in the same level, \(\ell\), say, and we compare this to the fluid queue on the event that the level remains in the band \(\calD_\ell\). No boundary behaviour is involved in this calculation, so we treat \(\left\{{\phi}(t)\right\}\) and \(\left\{\varphi(t)\right\}\) as coupled and use the latter notation. When we must distinguish the two processes, we use \(\phi(t)\) for the phase of the QBD-RAP and \(\varphi(t)\) for the phase of the fluid.

\paragraph{The level process}
To event epochs of \(\left\{(\bs A(t),{\varphi}(t))\right\}_{t\geq0}\) we associate marks \(\left\{-1,0,+1\right\}\) in the following way.
\begin{itemize}
	\item To events epochs corresponding to a change of phase of \( {\varphi}(t)\) we associate the mark \(0\).
	\item To event epochs at time \(t\) which correspond to a change in level and for which \( {\varphi}(t^-)=i\in\mathcal S_-\) we associate the mark \(-1\).
	\item To event epochs at time \(t\) which correspond to a change in level and for which \( {\varphi}(t^-)=i\in\mathcal S_+\) we associate the mark \(+1\).
\end{itemize}
Now define \(N_+(t)\) (\(N_-(t)\)) as the simple point process which counts the number of event epochs with marks \(+1\) (\(-1\)) which have occurred in the time up to and including time \(t\). The level process of the QBD-RAP is given by \( L(t) = N_+(t)-N_-(t)\). The process \(\left\{(L(t),\bs A(t),\varphi(t))\right\}_{t\geq 0}\) forms a QBD with RAP components.

One way to specify the QBD with RAP components, \(\left\{(L(t),\bs A(t),{\varphi}(t))\right\}\), is to describe its generator:
\begin{align*}
\bs{B} &= \left[\begin{array}{cccccc}
 \ddots & \ddots & \ddots & & & \\
 & \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & & \\ 
 & &  \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & \\
 & & & \ddots & \ddots & \ddots 
\end{array}\right],
\end{align*}
{where,}
\begin{align*}
\bs{B}_{0} &= \left[\begin{array}{cccc}
	\bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs{T}_{+0}\otimes \bs{I} & \bs 0 \\
	\bs{T}_{-+}\otimes \bs{D} & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & \bs 0 & \bs{T}_{-0}\otimes \bs{I} \\
	\bs{T}_{0+}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs{T}_{00}\otimes \bs{I} & \bs 0 \\
	\bs{T}_{0+}\otimes \bs{D} & \bs{T}_{0-}\otimes \bs{I} & \bs 0 & \bs{T}_{00}\otimes \bs{I}
	\end{array}\right],
\end{align*}
\begin{align*}
\bs{B}_{-1} &= \left[\begin{array}{cccc}
	\bs 0 & & & \\
	& \bs{C}_-\otimes (\bs s\bs\alpha) & & \\
	& & \bs 0 & \\ 
	& & & \bs 0
	\end{array}\right],
%
&& \bs{B}_{+1} = \left[\begin{array}{cccc}
	\bs{C}_+\otimes (\bs s\bs\alpha) & & & \\
	& \bs 0 & & \\
	& & \bs 0 & \\ 
	& & & \bs 0
	\end{array}\right].
\end{align*}
%The generator acts on infinite dimensional row-vectors 
%\[\bs v = \left( \begin{array}{ccccc} \dots & \bs v_{-1} & \bs v_0 & \bs v_{1} & \dots \end{array} \right),\]
%via right-multiplying the vector. The vector \(\bs v\) represents an initial distribution on the QBD-RAP. The vectors \(\bs v_k\), \(k\in\mathbb Z\) are 
%\[\bs v_k = \left( \begin{array}{ccc} \bs v_{k,1} & \dots & \bs v_{k,N} \end{array} \right),\]
%where each \(\bs v_{k,i}\) is such that \((\bs v_{k,i}, \bs{S}_i,\bs s_i)\) is a valid representation of an ME. The initial distribution of the phase and level is given by the product \(\bs v_{k,i}\bs e = \mathbb P( L(0)=k, \varphi(0)=i)\) and, given \(\left\{ L(0)=k, \varphi(0)=i\right\}\), the initial position of \(\left\{\bs A(t)\right\}\) is \(\cfrac{\bs v_{k,i}}{\bs v_{k,i}\bs e} \). 
%
%The product 
%\[\bs v(t) := \bs v e^{\bs{B}t} = \left( \begin{array}{ccccc} \dots & \bs v_{-1}(t) & \bs v_0(t) & \bs v_{1}(t) & \dots \end{array} \right),\]
%where 
%\[\bs v_k(t) = \left( \begin{array}{ccc} \bs v_{k,1}(t) & \dots & \bs v_{k,N}(t) \end{array} \right),\quad k\in\mathbb Z,\]
%gives the distribution at time \(t\): for \(k\in\mathbb Z\),  \(\mathbb P({L}(t) = k, \varphi(t)=i) = \bs v_{k,1}(t)\bs e\) and, given \(\left\{{L}(t) = k, \varphi(t)=i\right\}\), the position of \(\left\{\bs A(t)\right\}\) is \(\bs A(t) =\cfrac{\bs v_{k,i}(t)}{\bs v_{k,i}(t)\bs e}. \)

\section{Boundary conditions}\label{sec: boundary conditions}
In the study and practical application of fluid queues, regulated, reflecting, or a mixture of these boundary conditions may be imposed. We now present the intuition as to how one may include such boundary conditions in the QBD-RAP approximation scheme.

Without loss of generality, assume that there is a boundary for the fluid level at \(y_0 = 0\). This boundary can only be hit in phases \(i\in\mathcal S_-\). %For a regulated boundary, upon hitting the boundary in phase \(i\in\mathcal S_-\) the fluid level remains at \(0\), while the phase continues to evolve among states in the set \(\calS_-\cup\calS_0\). Upon the first transition of the phase process from \(\calS_-\cup \calS_+\) to \(\calS_+\) the fluid level leaves the boundary. For a reflecting boundary we suppose that, upon hitting the boundary in some phase \(i\in\mathcal S_-\) the phase of the process immediately jumps to \(j\in\mathcal S_+\) with probability \(p_{ij}\). 
Suppose that, upon hitting the boundary in phase \(i\in\calS_-\), the phase jumps from \(i\) to \(j\in\calS\) with probability \(p_{ij}\). If \(j\in\calS_-\cup\calS_0\) the process remains at the boundary and the phase process evolves among states in \(\calS_-\cup\calS_0\) until the first transition to a state in \(\calS_+\), at which point the level \(\{X(t)\}\) immediately leaves the boundary. If \(j\in\calS_+\) the process immediately leaves the boundary. 
We collect the probabilities \(p_{ij}\) into the matrices \(\bs P_{-+} = [p_{ij}]_{i\in\calS_-,j\in\calS_+},\) \(\bs P_{-0} = [p_{ij}]_{i\in\calS_-,j\in\calS_0},\) and \(\bs P_{--} = [p_{ij}]_{i\in\calS_-,j\in\calS_-}.\)

Adjacent to the boundary at \(y_0=0\) is the band \(\mathcal D_0=[0,\Delta]\) which corresponds to level \(0\) for the QBD-RAP. Modelling the behaviour of the fluid queue at boundaries can be broken down into three components. 1) Modelling the time and phase when the fluid level hits the boundary. 2) Modelling the phase whilst the fluid level remains at the boundary. 3) Modelling the fluid level and phase at the exit from the boundary.

We claim that the event \(\{L(t)=\ell,\phi(t)=i\}\) models the event \(\{X(t)\in\mathcal D_{\ell,i},\varphi(t)=i\}\), and further that instants \(u\) with \(L(u^-)\neq L(u), \phi(u)=i\) model the events \(\{X(u^-)\in \mathcal D_{\ell,i}, X(u)\notin\mathcal D_{\ell,i},\varphi(u)=i\}\). With this in mind, we suppose that when the QBD-RAP is in level \(0\) in phase \(i\in\mathcal S_-\) and there is a change of level, this corresponds approximating the event that the fluid level \(\{X(t)\}\) hits the boundary at \(0\). We refer to this as the QBD-RAP hitting the boundary also. We show later that, using matrix exponentials with sufficiently small variance in the construction of the QBD-RAP, then the distribution of time until the QBD-RAP first hits the boundary closely models the time until the fluid level hits the boundary closely. This addresses component 1).

%Assuming the fluid process \(\{X(t)\}\) is continuous then we may use the QBD-RAP to approximate the fluid queue hitting the boundary. For example, with a regulated lower boundary at \(x=0\), place \(x=0\) at the boundary of the first of the bands \(\mathcal D_0=[0,\Delta]\) which corresponds to level \(\ell =0\). When the QBD-RAP process, \(\{(L(t), \bs A(t), \varphi(t))\}\) exits level \(0\) in a phase \(j\in\mathcal S_-\), then we correspond this with the event that fluid queue is hits level \(x=0\) in phase \(j\). Corollary~\ref{cor: aln222} states that, for sufficiently large \(p\), the construction described here models distribution of time until \(\{X(t)\}\) leaves \(\mathcal D_0\) arbitrarily closely, and therefore we claim that the QBD-RAP approximation models hitting the boundary at \(x=0\) arbitrarily closely.

For regulated behaviour at a lower boundary at \(0\), given the fluid level reaches zero at time \(u\) in phase \(i\in\mathcal S_-\), then the phase transitions to some phase \(j\in\calS_-\cup\calS_0\) with probability \(p_{ij}\), and the distribution of the phase \(\varphi(t)\) until the fluid queue leaves the boundary for the first time after \(u\) is given by the elements of the vector
\[p_{ij} \bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right).\] 
The distribution for the time until \(\{X(t)\}\) leaves \(x=0\) for the first time after \(u\) is 
\[1-p_{ij}\bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right)\bs e.\] 
Therefore, in the case of a regulated boundary, we can capture the behaviour of the process at the boundary exactly. For a reflecting boundary there is no time spent at the boundary. This addresses component 2) of the modelling problem.

For both regulated boundary behaviour and reflecting boundary behaviour, given the QBD-RAP is in the correct phase at the instant upon which it hits the boundary, then upon exiting the boundary the phase can be captured exactly too. For a regulated boundary, given the fluid/QBD-RAP hits the boundary in phase \(i\in\mathcal S_-\) at time \(t\), it then jumps to phase \(j\in\calS_+\cup\calS_0\) and exits the boundary in phase \(k\in\mathcal S_+\) at time \((t-u)\) with density 
\[p_{ij}\bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right)\left[\begin{array}{c} \bs T_{-+} \\ \bs T_{0+}\end{array}\right]\bs e_k.\] 
For a reflecting boundary, given the boundary is hit in phase \(i\in\calS_-\), the phase upon leaving the boundary is \(j\in\calS_+\) with probability \(p_{ij}\). Upon leaving the boundary, the appropriate orbit position for the QBD-RAP is \(\bs \alpha\), as this corresponds to the fluid level \(X=0\).

To summarise, at time \(t\) the rate at which probability mass accumulates at the regulated boundary in phase \(j \in\calS_-\cup\calS_0\) upon hitting the boundary in phase \(i\in\calS_-\) is 
\[ c_ip_{ij} \bs A(t)\bs s.\]
The rate at which mass transitions from phase \(i\in\calS_-\) to \(j\in\calS_+\) upon hitting a boundary is 
\[ c_ip_{ij} \bs A(t)\bs s,\]
and upon this transition, the orbit jumps to \(\bs \alpha\). 
The rate at which mass leaves the regulated boundary from phase \(i\in\calS_-\) into phases \(j\in\calS_+\) is 
\[T_{ij},\]
and upon leaving the boundary the orbit is \(\bs\alpha\).

This information can be inscribed in the generator of the QBD-RAP. For example, using the augmented state-space scheme to account for phases \(\calS_0\), the generator of the QBD-RAP described above has boundary conditions included as 
\begin{align*}
&\left[\begin{array}{cc|cccc|cc}
	\bs T_{--} & \bs T_{0-} & \bs T_{-+} \otimes\bs \alpha & \bs 0 & \bs 0 & \bs 0 &&\\
	\bs T_{-0} & \bs T_{00} & \bs T_{0+} \otimes\bs \alpha & \bs 0 & \bs 0 & \bs 0 &&\\\hline
	\bs 0 & \bs 0 &\multicolumn{4}{c|}{\multirow{4}{*}{$\check{\bs B}_{0}$}} & \multirow{4}{*}{$\bs B_{+1}$}& \\
	(\bs C_-\bs P_{--})\otimes \bs s &(\bs C_-\bs P_{-0}) \otimes \bs s& \multicolumn{4}{c|}{} & &\\
	\bs 0 & \bs 0 & \multicolumn{4}{c|}{} &&\\
	\bs 0 & \bs 0 & \multicolumn{4}{c|}{} && \ddots\\ \hline
	& & \multicolumn{4}{c|}{\bs B_{-1}} & \bs B_0 &\\
	& & \multicolumn{3}{c}{}& \ddots &  & \ddots
	\end{array}\right],
\end{align*}
where 
\[\check{\bs B}_{0} = \left[\begin{array}{cccc}
	 \bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs{T}_{+0}\otimes \bs{I} &  \\
	 (\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha + \bs T_{-+}\otimes \bs D  & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} &  & \bs{T}_{-0}\otimes \bs{I} \\
	 \bs{T}_{0+}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs{T}_{00}\otimes \bs{I} &  \\
	 \bs{T}_{0+}\otimes \bs{D} & \bs{T}_{0-}\otimes \bs{I} &  & \bs{T}_{00}\otimes \bs{I} 
	\end{array}\right].\]
The top-left block represents the point masses due to the regulated boundary. The bottom-left block is the transition of density into point masses due to the regulated boundary. The top-right block is the transition of point masses into density as the process leaves the boundary. The term \((\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha\) in \(\check{\bs B}_{0}\) incorporates the reflecting boundary.

Upper boundaries can be included in an analogous manner.

The orbit process \(\{\bs A(t)\}\) is not required to model the behaviour at the regulated boundary. However, we suppose that \(\bs A(t)=1\) at the boundaries. This choice is arbitrary but allows us to generalise some notation later when we are describing the evolution of the QBD-RAP. Further, for boundaries at \(y_0=0\) and \(y_{K+1} = (K+1)\Delta\), we let the level process take the value \(L(t)=-1\) at the lower boundary, and \(L(t)=K+1\) at the upper boundary. Denote the set of levels (including boundary levels) by \(\mathcal K = \{-1,0,\dots,K,K+1\}\).

\begin{rem}
	It may be possible to extend the QBD-RAP approximation to model jumps into and out of the boundary, provided that the jumps into/out of the boundary happen at an intensity which can be described by a matrix exponential distribution.

	Furthermore, it may be possible to model more general boundary behaviours which can be approximated by a limit of matrix exponentials. For example, a boundary condition where the fluid queue spends a deterministic time at the boundary. 
\end{rem}

\section{Initial conditions}\label{sec: initial conditions}
%To this point, we have assumed that \((X(0),\varphi(0))=(y_\ell,i), \,i\in\mathcal S_+\) or \((X(0),\varphi(0))=(y_{\ell+1},i), \,i\in\mathcal S_-\), i.e., the initial position of the fluid is at the edge of the cell. In these cases, the appropriate initial orbit position is \(\bs \alpha\). We now generalise this so that we can approximate arbitrary initial conditions.

%To motivate the construction, first consider the initial condition for the QBD-RAP approximation, \((L(0),\bs A(0), \varphi(0))=(\ell,\bs \alpha,i)\), corresponding to the initial condition of the fluid queue \((X(0),\varphi(0))=(y_\ell,i),\) for some \(i\in\mathcal S_+\). After time \(t\), \(c_it\in[0,\Delta)\), and assuming that no events of the QBD-RAP have occurred by this time, the position of the QBD-RAP is \((L(t),\bs A(t),\varphi(t)) = (\ell, \bs A(t), i),\) where \(\bs A(t) = \cfrac{\bs \alpha e^{c_i\bs{S}t}}{\bs \alpha e^{c_i\bs{S}t}\bs e}\). The level of the fluid at the same time is \(X(t) = y_\ell + c_i t\). 
We argued earlier that we can think of the orbit \(\cfrac{\bs \alpha e^{\bs{S}|c_i|t}}{\bs \alpha e^{\bs{S}|c_i|t}\bs e}\) as corresponding to the fluid being a distance of \(|c_i|t\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\). With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is the set the initial orbit to \(\cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\). Similarly, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\cfrac{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}}{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}\bs e}.\)

We use the notation \(\bs a_{\ell,i}(x_0) = \cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\). Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \cfrac{\bs \alpha e^{\bs{S}(y_{\ell+1} - x_0)}}{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}\bs e}\).

More generally, given an initial measure \(\mu_i(\cdot):= \mathbb P(X(0)\in\cdot,\varphi(0)=i)\), an approximation to this initial condition is to set the orbit to \(\displaystyle\int_{x\in\calD_{\ell,i}}\bs a_{\ell,i} (x) \wrt \mu_i\), for \(i\in\mathcal S,\, \ell\in\mathcal K\setminus\{-1,K+1\}\).

%For technical reasons we must restrict the initial conditions so that there are no point masses at \(x_0=y_{\ell+1}\) when \(i\in\mathcal S_+\) or \(x_0=y_{\ell}\) when \(i\in\mathcal S_-\). In either of these cases, the initial orbit position would be \(\cfrac{\bs \alpha e^{\bs{S}\Delta}}{\bs \alpha e^{\bs{S}\Delta}\bs e}\) and the denominator is \(\bs \alpha e^{\bs{S}\Delta}\bs e\). To our knowledge, there are currently no results that ensure that \(\bs \alpha e^{\bs{S}\Delta}\bs e\) is not arbitrarily small for the class of matrix exponential distributions with which we work. For point masses at any other value of \(x_0\in\mathcal D_\ell, x_0\neq y_{\ell+1}\) and \(i\in\mathcal S_+\) the initial orbit position is \(\cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\) and the denominator is \(\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e\geq 1-\cfrac{\var\left(Z\right)}{(\Delta-(x_0-y_\ell))}\). Therefore, by choosing a matrix exponential \(Z\) with sufficiently small variance, then the denominator is arbitrarily close to 1. Similarly for initial point masses \(x_0\in\mathcal D_\ell\neq y_{\ell}\), \(i\in\mathcal S_-\). This is, however, no real practical loss as point masses at \(x_0=y_{\ell+1}\) when \(i\in\mathcal S_+\) (\(x_0=y_{\ell}\) when \(i\in\mathcal S_-\)) can, and should, have the initial conditions \(\bs v_{\ell+1,i}=\bs \alpha\) (respectively, \(\bs v_{\ell-1,i}\)), corresponding to the fluid starting at the left (right) of the adjacent interval which shares the boundary \(y_{\ell+1}\) (\(y_\ell\)).

\section{At time \(t\) -- closing operators}\label{sec: closing}
We now describe how the orbit position \(\bs A(t)\) can be used to approximate the density of the position of \(X(t)\) in a given level.

Suppose that the QBD-RAP is in level \(L(t)=\ell\), phase \(\phi(t)=j\in\calS_+\), and the orbit is \(\bs A(t)=\bs a\). If the QBD-RAP remains in phase \(j\), then QBD-RAP approximation will transition out of level \(\ell\) in the infinitesimal time interval \(t+\wrt u\) with density \(\bs a e^{|c_j|\bs{S}u}|c_j|\bs s\wrt u\). At the time of the change of level we estimate the position of \(X(t+u)\) by \(X(t+u)\approx y_{\ell+1}\). Tracing this back to time \(t\), we estimate the position of \(X(t)\) as \(X(t)\approx y_{\ell+1} - |c_j|u\). Reverse engineering this logic, the approximation to the density of \(X(t) \in\wrt x\) is \(\bs a e^{|c_j|\bs S(y_{\ell+1}-x)/|c_j|}|c_j|\bs s \wrt x/|c_j|\), since \(\wrt x = |c_j|\wrt u\) where \(\wrt x\) is an infinitesimal interval in space and \(\wrt u\) is an infinitesimal interval with respect to time.

Similarly, for \(j\in\mathcal S_-\), if the QBD-RAP remains in phase \(j\) then the QBD-RAP approximation will transition out of level \(\ell\) in the infinitesimal time interval \(t+\wrt u\) with density \(\bs a e^{|c_j|\bs{S}u}|c_j|\bs s\wrt u\). At the time of the transition of level, we estimate the position of \(X(t+u)\) by \(X(t+u)\approx y_{\ell}\). Tracing this back to time \(t\), we estimate the position of \(X(t)\) as \(X(t)\approx y_{\ell} + |c_j|u\). Reverse engineering this logic, the approximation to the density \(X(t) \in\wrt x\) is \(\bs a e^{|c_j|\bs S(x-y_{\ell})/|c_j|}|c_j|\bs s \wrt x/|c_j|\), since \(\wrt x = |c_j|\wrt u\).

For phases \(j\in\mathcal S_0\), if phase \(j\) is associated with \(\mathcal S_+\), we use the same approximation as if \(j\in \mathcal S_+\), and if phase \(j\in\mathcal S_0\) is associated with \(\mathcal S_-\), we use the same approximation as if \(j\in \mathcal S_-\).

This reasoning leads to an approximation of the distribution of the fluid at time \(t\), \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \), as
\begin{align}
		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x,\label{eqn: density approx aug}
		\intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
\end{align}
for \( j\in\mathcal S_-\cup\calS_{-0}\), where \(x\in\mathcal D_{\ell,j}\), \(x_0\in\mathcal D_{\ell_0,i}\).
%For example, for the augmented scheme in Section~\ref{subsec: augmented }, we approximate \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i)\) by 
%\begin{align}
%	\label{eqn: density approx aug}
%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_0, \varphi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x & j\in\mathcal S_+\cup \mathcal S_{+0},
%		\\\label{eqn: density approx aug2}
%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_0, \varphi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x & j\in\mathcal S_-\cup \mathcal S_{-0},
%\end{align}

While the estimates in the right-hand side of (\ref{eqn: density approx aug}) and (\ref{eqn: density approx aug2}) are appealing, they are, however, defective estimates -- they do not integrate to 1 for any finite-dimensional matrix exponential distribution. To see this, compute 
\begin{align*}
	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_+\cup \calS_{+0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}(y_{\ell+1}-x)} \bs s \wrt x
	\\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_+\cup \calS_{+0}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	\\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
	\\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
	\\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta} \bs e )
	\\&<1,
\end{align*}
since \(\bs a e^{\bs{S}u} \bs e>0\) for any \(u>0\). For an orbit position \(\bs A(t)=\bs a\), the amount of mass missing from each level and phase is 
\[\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}\Delta} \bs e \wrt x.\]

One way to rectify this is to renormalise the estimates in the right-hand side of (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}) with the missing mass, i.e.~the approximation to \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) is
\begin{align}
	\label{eqn: density approx 2}
		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(y_{\ell+1}-x)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x \intertext{for \(j\in\mathcal S_+\cup \calS_{+0}\), and}
		\label{eqn: density approx 22}
		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(x-y_\ell)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x ,
\end{align}
for \(j\in\mathcal S_-\cup \calS_{-0}.\) This renormalisation distributes any missing mass over the interval \(\calD_{\ell,j}\) proportional to the estimated density in the region.

More amenable to analysis, however, is to approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
\begin{align}
%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
	\label{eqn: density approx 3+}&
		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) u_+^\ell(\bs a,x) \wrt x ,
	\intertext{where}
		&u_+^\ell(\bs a,x) = \cfrac{\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right)}{1-\bs a e^{ \bs{S}2\Delta} \bs e},
	\intertext{for \(j\in\mathcal S_+\cup \calS_{+0}\), and }
%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
	\label{eqn: density approx 3-}& 
		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell_0,i}(x_0), \phi(0) = i) u_-^\ell(\bs a ,x) \wrt x ,
	\intertext{where}
		&u_-^\ell(\bs a,x) = \cfrac{\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right)}{1-\bs a e^{ \bs{S}2\Delta} \bs e},
\end{align}
for \(j\in\mathcal S_-\cup \calS_{-0}\).

Intuitively, we take the density function \(\cfrac{\bs ae^{\bs{S}x}\bs s}{1-\bs ae^{\bs{S}2\Delta}\bs e}\) on \(x \in [0,2\Delta)\), and `fold' it back on itself around \(\Delta\), to create a density function on \([0,\Delta)\), 
\[\cfrac{\bs ae^{\bs{S}x}\bs s +\bs ae^{\bs{S}(2\Delta-x)}\bs s}{1-\bs ae^{\bs{S}2\Delta}\bs e},\, x \in [0,\Delta).\]

The advantages of this approximation are;
\begin{itemize}
	\item the denominator \(1-\bs ae^{\bs{S}2\Delta}\bs s\) is provably close to \(1\) for concentrated matrix exponential distributions.
	\item in (\ref{eqn: density approx 2}) any of the once missing mass is spread across the interval \(\calD_{\ell,j}\) proportionally to the estimated density in the region, whereas the missing mass should be more concentrated around \(\Delta\), as this is the region around which the function \(\bs \alpha e^{\bs Sx}\bs s\) fails to be like the delta function most serverely. This is what (\ref{eqn: density approx 3+})-(\ref{eqn: density approx 3-}) do.
\end{itemize}
Meanwhile, (\ref{eqn: density approx 3+})-(\ref{eqn: density approx 3-}) maintains positivity and smoothness.

The difference between these approximation schemes is the way in which the vector \(\bs a\) is used. We generalise this idea with the concept of \emph{closing operators}. For \(x\in[0,\Delta)\), let \(V(x):\mathcal A \to \mathbb R,\) be a closing operator. For example, the closing operator in Equations (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}) is the operator \(V(x),\,x\in[0,\Delta)\) such that for any \(\bs a \in \mathcal A\), 
\begin{align}
	\bs a V(x) &= \bs a e^{\bs Sx}\bs s.\label{eqn: alal}
	%\\\bs a V_-(x-y_\ell) &= \bs a e^{\bs S(x-y_{\ell})}\bs s,\label{eqn:ndaenf}
\end{align}
Similarly, in (\ref{eqn: density approx 2}) and (\ref{eqn: density approx 22})
\begin{align}
	\bs a V(x) &=\bs a \bs e \cfrac{\bs a e^{ \bs{S}x} \bs s}{\bs a \bs e -\bs a e^{ \bs{S}\Delta} \bs e}, \label{eqn:epan}
%	\\\bs a V_-(x-y_\ell) &=  \cfrac{\bs a e^{ \bs{S}(x-y_\ell)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e},\label{eqn:epasan}
\end{align}
and in (\ref{eqn: density approx 3+}) and (\ref{eqn: density approx 3-}) 
\begin{align}
	\bs a V(x) &= \bs a \bs e\cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{\bs a\bs e-\bs a e^{ \bs{S}2\Delta} \bs e}.\label{eqn:546}
%	\\\bs a V_-(x-y_\ell) &=   \cfrac{\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right)}{1-\bs a e^{ \bs{S}2\Delta} \bs e}. \label{eqn: sadlgj}
\end{align}
The operator \(V(x)\) is a choice which forms part of the definition of the approximation scheme. As we shall see, given certain properties of \(V(x)\), we can prove that the approximation scheme converges, conserves probability, and ensures positivity. In the cases above, all the closing operators (\ref{eqn: alal})-(\ref{eqn:546}) lead to an approximation which ensures positivity due to their interpretation as probability densities. The operators (\ref{eqn:epan})-(\ref{eqn:546}) ensure the approximation scheme conserves probability by construction. We can prove that the operators (\ref{eqn: alal}) and (\ref{eqn:546}) lead to approximation schemes which converge. This is not to say that the scheme with \(V(x)\) as in (\ref{eqn:epan}) does not converge necessarily, just that we have not proven it.

### 2021-09-22 chapter1/chapter.tex
remove denominator in V(x) operator

@@
 
-Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982} as a canonical application in this area), power systems \cite{hydro}, risk processes \cite{betal2005} and environmental modelling \cite{wurm2020}.
+Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982} as a canonical application in this area), power systems \citep{hydro}, risk processes \citep{betal2005} and environmental modelling \citep{wurm2020}.
 Fluid queues are relatively well studied.
-Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \cite{anick1982,kk1995,beanetal2019}.
+Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods \citep{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \citep{anick1982,kk1995,beanetal2019}.
 %For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest.
@@
 
-More recently, Bean and O'Reilly \cite{bo2014} extended fluid queues to \emph{so-called} stochastic fluid-fluid queues.
+More recently, \cite{bo2014} extended fluid queues to \emph{so-called} stochastic fluid-fluid queues.
 In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
-The analysis, Bean and O'Reilly \cite{bo2014}, is in principal similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
-For practical computation of the results \cite{bo2014} a matrix-discretisation of the infinitesimal generator of the fluid queue can be used.
+The analysis,  \citep{bo2014}, is in principal similar to the matrix-analytic methods of \citep{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
+For practical computation of the results of \cite{bo2014}, a matrix-discretisation of the infinitesimal generator of the fluid queue can be used.
 To this end, to date, two possible discretisation have been suggested.
-Taking a differential equations-based approach, Bean \emph{et al.}~\cite{beanetal2019} use the discontinuous Galerkin (DG) method to discretise this operator, while Bean and O'Reilly \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process.
+Taking a differential equations-based approach, \cite{beanetal2019} use the discontinuous Galerkin (DG) method to discretise this operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process.
 Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations.
@@
 
-In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \cite[Section 3.3]{koltai2011}, see also \cite{nodalDGBook} and references therein.
+In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \citep[Section 3.3]{koltai2011}, see also \citep{nodalDGBook} and references therein.
 One justification for the positivity preserving property is from the interpretation of the discretisation as a stochastic process thereby ensuring positivity.
@@
 Moreover, higher-order DG approximation schemes may produce negative, or highly oscillatory solutions, particularly when discontinuities or steep gradients are present.
-Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \cite[Section 6.5]{nodalDGBook} and references therein).
+Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \citep[Section 6.5]{nodalDGBook} and references therein).
 Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in theses regions.

### 2021-09-22 chapter2/chapter.tex
remove denominator in V(x) operator

@@
 \chapter{A stochastic modelling approach to approximation\label{sec: construction and modelling}}
-In this work we develop a new discretisation of a fluid queue using a \emph{quasi-birth-and-death process with rational arrival process components} (QBD-RAP), to be described in more detail later, and prove convergence of the approximation scheme.
-To improve upon the discretisation of Bean and O'Reilly \cite{bo2013}, we argue that the shift from QBD to QBD-RAP is necessary.
+In this chapter we develop a new discretisation of a fluid queue using a \emph{quasi-birth-and-death process with rational arrival process components} (QBD-RAP), to be described in more detail later, and prove convergence of the approximation scheme.
+To improve upon the discretisation of \cite{bo2013}, we argue that the shift from QBD to QBD-RAP is necessary.
 The approximation can be made arbitrarily accurate.
@@
 
-This paper details the construction of the approximation and proves its convergence.
-The QBD-RAP is constructed using \emph{matrix-exponential distributions}.
-The main result shows the convergence of the approximation scheme under the assumption that the variance of the matrix-exponential distribution(s) used in the construction tends to 0.
-Given that there is no known simple closed form for the transient distributions of a fluid queue, and that we make minimal assumptions about the matrix-exponential distributions used in the construction of the approximation, it is somewhat remarkable that we are able to establish a convergence result.
-The generality of the result with respect to the assumptions on the matrix-exponential distributions used in the construction of the approximation is necessitated by the fact that we use a class of \emph{concentrated matrix-exponential distributions} found numerically in \cite{hhat2020}, and for which there is relatively little known about their properties.
-
-The structure of this paper is as follows.
+The structure of this chapter is as follows.
 In Section~\ref{sec: prelim} we provide a brief background on matrix exponential random variables and the stochastic processes appearing in this work, namely fluid queues, quasi-birth-and-death (QBD) processes, rational arrival processes (RAPs), and quasi-birth-and-death processes with rational arrival process components (QBD-RAPs).
-Section~\ref{sec: construction and modelling} describes the modelling approach taken to develop the approximation.
-Section~\ref{sec: inspiration} motivates the idea using the QBD approximation of \cite{bo2014}.
+Section~\ref{sec: inspiration} motivates the idea using the QBD approximation of \citep{bo2014}.
 Sections \ref{sec: modelling} and \ref{sec: zero states} describe the modelling of certain events of the fluid queue with matrix exponential distributions to ultimately construct the behaviour of the QBD-RAP.
@@
 Section~\ref{sec: closing} introduces the concept of a \emph{closing operator} which maps states of the QBD-RAP to estimates of the distribution of the fluid queue.
-We then proceed to show convergence of the approximation scheme in Section~\ref{sec: conv}.
-First, in Section~\ref{sec: no change}, we derive expressions for certain Laplace transforms with respect to time of the QBD-RAP and fluid queue on the event that there is no change of level.
-Section~\ref{sec: no change convergence} provides error bounds between the Laplace transforms of the QBD-RAP and the fluid queue and also some domination conditions used so that we may apply the Dominated Convergence Theorem.
-Section~\ref{sec: local to global} uses the results of the preceding section to prove convergence Laplace transform with respect to time of the expected local time of the QBD-RAP to the Laplace transform with respect to time of the expected local time of the fluid queue.
-Technical proofs which do not offer a great deal of insight are reserved for appendices.
 
@@
 Here we recount some facts about matrix exponential distributions.
-See \cite{MEinAP} for a more detailed exposition.
+See \citep{MEinAP} for a more detailed exposition.
 A random variable, \(Z\), is said to have a matrix-exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
@@
 The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\).
-Representations of matrix-exponential distributions are not unique \cite{MEinAP}.
+Representations of matrix-exponential distributions are not unique \citep{MEinAP}.
 A representation is called \emph{minimal} when \(\bs S\) has the smallest possible dimension.
@@
 
-Also from Bean and Nielsen \cite{bn2010}, associated with a Marked RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\)
+Also from \cite{bn2010}, associated with a Marked RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\)
 \[\bs A(t) = \cfrac{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S (t-Y_{N(t)})}}{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
@@
 \section{Inspiration and motivation}\label{sec: inspiration}
-The inspiration for our approach stems from the QBD-based discretisation of Bean and O'Reilly \cite{bo2013}.
+The inspiration for our approach stems from the QBD-based discretisation of \cite{bo2013}.
 A QBD can be seen as a two dimensional CTMC, \(\{(L(t),\phi(t))\}_{t\geq 0}\), where \(\{L(t)\}\) is the discrete level, and \(\{\phi(t)\}\) is the phase.
-Bean and O'Reilly \cite{bo2013} discretise the state space of the fluid into small intervals of width \(\Delta\); denote these by \(\mathcal D_k=[k\Delta,(k+1)\Delta]\).
+\cite{bo2013} discretise the state space of the fluid into small intervals of width \(\Delta\); denote these by \(\mathcal D_k=[k\Delta,(k+1)\Delta]\).
 Their approximation captures the dynamics of the phase process exactly, but discretises the level process of the fluid queue.
@@
 \end{align}
-But this is the same discretisation we would get if we were to use intervals of width \(\Delta/2\) in the QBD approximation of Bean and O'Reilly \cite{bo2013}, with the rows/columns reordered.
-Indeed, this appears to be the best we can do with a QBD-MAP approximation as the Erlang distribution is known to be the least variable Phase-type distribution for a given order \cite{as1987} and therefore gives the closest approximation to the deterministic behaviour we are trying to approximate.
+But this is the same discretisation we would get if we were to use intervals of width \(\Delta/2\) in the QBD approximation of \cite{bo2013}, with the rows/columns reordered.
+Indeed, this appears to be the best we can do with a QBD-MAP approximation as the Erlang distribution is known to be the least variable Phase-type distribution for a given order \citep{as1987} and therefore gives the closest approximation to the deterministic behaviour we are trying to approximate.
 This necessitates the extension to QBD-RAPs, whereby we can find more concentrated distributions than any Phase-type, but retain (some of) the stochastic interpretation and matrix-analytic tools.
@@
 We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance).
-As in Bean and O'Reilly \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP.
-Unlike Bean and O'Reilly \cite{bo2013}, however, here we do not take \(\Delta \to 0\).
+As in \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP.
+Unlike  \cite{bo2013}, however, here we do not take \(\Delta \to 0\).
 Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small.
@@
 \includegraphics[width=\textwidth]{chapter2/figs/ME_residual_life_density.pdf}
-\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \cite{hhat2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.5}\) (blue).
+\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hhat2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.5}\) (blue).
 Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.5}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively.
@@
 Observe that \(\bs D\bs e = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\bs e =\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \wrt x=\bs e.\) This is an important observation as it implies that events corresponding to jumps from \(i\in\calS_+\) to \(j\in\calS_-\) occur with intensity \(\bs A(t)T_{ij}\bs D\bs e = T_{ij}\) which is constant with respect to orbit position and time and this implies they are exponential.
-Further, since \(\mathcal A\) is closed and convex \cite{MEinAP}, then \((\bs a \bs D,\bs S,\bs s)\) is a representation of a matrix exponential distribution for any \(\bs a\in\mathcal A\).
+Further, since \(\mathcal A\) is closed and convex \citep{MEinAP}, then \((\bs a \bs D,\bs S,\bs s)\) is a representation of a matrix exponential distribution for any \(\bs a\in\mathcal A\).
 
@@
 
-According to Hautphenne \emph{et al.}~\cite{hmp2017}, the distribution of \(C_{t+u}\) depends on the method of observation.
+According to \cite{hmp2017}, the distribution of \(C_{t+u}\) depends on the method of observation.
 Here, the Phase-type distribution is observed after an exponential time with rate \(-T_{ii}\).
-Proposition 4.1, Hautphenne \emph{et al.}~\cite{hmp2017} states that the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) and the process is observed after an exponential time with rate \(-T_{ii}\) is
+Proposition 4.1, \cite{hmp2017} states that the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) and the process is observed after an exponential time with rate \(-T_{ii}\) is
 \[\mathbb P(C_{t+u}\leq a )=1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}.\]
@@
 
-With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of Bean and O'Reilly \cite{bo2014} with discretisation parameter \(\Delta/p\).
+With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2014} with discretisation parameter \(\Delta/p\).
 
@@
 \section{Computing \(\bs D\)}
-In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \cite{hhat2020}.
+In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hhat2020}.
 For this class of CMEs, we take the index \(p\) to be the order of the representation.
-Since the variance of \(Z^{(2p)}\) and \(Z^{(2p-1)}\) are relatively similar \cite{hhat2020}, we take \(p\) to be odd.
+Since the variance of \(Z^{(2p)}\) and \(Z^{(2p-1)}\) are relatively similar \citep{hhat2020}, we take \(p\) to be odd.
 For a given CME, with representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part.
@@
 
-Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \cite{hhat2020}.
+Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \citep{hhat2020}.
 
@@
 %
-%We can approximate the generator of a fluid queue as a QBD, either by considering a uniformisation of the underlying phase process (as in \cite{bo2013}), or using a finite-volume scheme (see the DG paper, to be submitted).
+%We can approximate the generator of a fluid queue as a QBD, either by considering a uniformisation of the underlying phase process (as in \citep{bo2013}), or using a finite-volume scheme (see the DG paper, to be submitted).
 Let \(\boldsymbol M(t) = (X(t),\varphi(t))_{t\geq 0}\) be a fluid queue, with finite state space \(\mathcal S = \left\{1,2,...,N\right\}\), \(N\times N\) generator matrix \(\bs{T}\), and rates \(c_i\) (stored in the \(N\times N\) diagonal matrix \(\bs{C}\)).
@@
 The phase variable of \(\bs J(t)\) and \(\bs M(t)\) match exactly.
-The process, \(L(t)\Delta\), approximates the fluid height \(X(t)\), and converges as \(\Delta \to 0\) \cite{bo2013}.
+The process, \(L(t)\Delta\), approximates the fluid height \(X(t)\), and converges as \(\Delta \to 0\) \citep{bo2013}.
 We also have the approximations
@@
 %Let \(\bs{D} = \bs c\bs \alpha\) where \(\bs c = -\bs{C}\bs 1\).
-Asmussen and Bladt \cite{ab1999} gave such a process a stochastic interpretation (in fact, Asmussen and Bladt worked with rational arrival processes, but I think we need only consider ME renewal processes).
+Asmussen and Bladt \citep{ab1999} gave such a process a stochastic interpretation (in fact, Asmussen and Bladt worked with rational arrival processes, but I think we need only consider ME renewal processes).
 Let \(V = span\left\{\alpha, \bs\alpha \bs{S},...,\bs\alpha \bs{S}^{p-1}\right\}\).
@@
 %\end{align*}
-%Now use the fact that \(\bs \pi = -\bs \alpha \bs{C}^{-1} / \mu\) (shown in \cite{ab1999}) to get that the joint density is equal to 
+%Now use the fact that \(\bs \pi = -\bs \alpha \bs{C}^{-1} / \mu\) (shown in \citep{ab1999}) to get that the joint density is equal to 
 %\[\bs c' e^{ \wt \tau_1 \bs{C}'} \bs{D}'e^{ \wt \tau_2 \bs{C}'} \bs{D}'...e^{ \wt \tau_K \bs{C}'}{\bs \alpha'} = \bs \alpha e^{\wt \tau_K\bs{C}}\bs{D}e^{\wt \tau_{K-1}\bs{C}}\bs{D}...e^{\wt \tau_1\bs{C}}\bs c,\]
@@
 %\paragraph{An adjoint process?}
-%Asmussen and Bladt \cite{ab1999} showed that 
+%Asmussen and Bladt \citep{ab1999} showed that 
 %\[\mathbb P (\theta_tN\in\cdot \mid \mathcal F_{\left[0,t\right]}) = \bs A(t)\bs v(\cdot),\]
@@
 %
-%For an ME renewal process we know that \(C_t\sim ME(\bs\pi, \bs{Q})\) and \(R_t\sim ME(\bs\pi, \bs{Q})\) \cite{ab1999}.
+%For an ME renewal process we know that \(C_t\sim ME(\bs\pi, \bs{Q})\) and \(R_t\sim ME(\bs\pi, \bs{Q})\) \citep{ab1999}.
 Bladt and Nielsen use algebraic arguments to show that \(S_t\) has a matrix exponential distribution with representation
@@
 
-One way to rectify this is to renormalise the estimates in the right-hand side of (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}) with the missing mass, i.e.~the approximation to \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) is
-\begin{align}
-	\label{eqn: density approx 2}
-		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(y_{\ell+1}-x)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x \intertext{for \(j\in\mathcal S_+\cup \calS_{+0}\), and}
-		\label{eqn: density approx 22}
-		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(x-y_\ell)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x ,
-\end{align}
-for \(j\in\mathcal S_-\cup \calS_{-0}.\) This renormalisation distributes any missing mass over the interval \(\calD_{\ell,j}\) proportional to the estimated density in the region.
-
-More amenable to analysis, however, is to approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
+%One way to rectify this is to renormalise the estimates in the right-hand side of (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}) with the missing mass, i.e.~the approximation to \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) is
+%\begin{align}
+%	\label{eqn: density approx 2}
+%		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(y_{\ell+1}-x)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x \intertext{for \(j\in\mathcal S_+\cup \calS_{+0}\), and}
+%		\label{eqn: density approx 22}
+%		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(x-y_\ell)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x ,
+%\end{align}
+%for \(j\in\mathcal S_-\cup \calS_{-0}.\) This renormalisation distributes any missing mass over the interval \(\calD_{\ell,j}\) proportional to the estimated density in the region.
+
+To partially rectify this we can instead approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
 \begin{align}
@@
 	\label{eqn: density approx 3+}&
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) u_+^\ell(\bs a,x) \wrt x ,
-	\intertext{where}
-		&u_+^\ell(\bs a,x) = \cfrac{\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right)}{1-\bs a e^{ \bs{S}2\Delta} \bs e},
-	\intertext{for \(j\in\mathcal S_+\cup \calS_{+0}\), and }
-%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
+		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_+^\ell(x) \wrt x ,
+\end{align}
+{where}
+\begin{align}
+		&\bs a u_+^\ell(x) = \bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right),
+\end{align}
+{for \(j\in\mathcal S_+\cup \calS_{+0}\), and }
+\begin{align}
 	\label{eqn: density approx 3-}& 
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell_0,i}(x_0), \phi(0) = i) u_-^\ell(\bs a ,x) \wrt x ,
-	\intertext{where}
-		&u_-^\ell(\bs a,x) = \cfrac{\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right)}{1-\bs a e^{ \bs{S}2\Delta} \bs e},
+		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell_0,i}(x_0), \phi(0) = i) \bs a u_-^\ell(x) \wrt x ,
+\end{align}
+{where}
+\begin{align}
+		&\bs a u_-^\ell(x) = \bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right),
 \end{align}
@@
 
-Intuitively, we take the density function \(\cfrac{\bs ae^{\bs{S}x}\bs s}{1-\bs ae^{\bs{S}2\Delta}\bs e}\) on \(x \in [0,2\Delta)\), and `fold' it back on itself around \(\Delta\), to create a density function on \([0,\Delta)\),
-\[\cfrac{\bs ae^{\bs{S}x}\bs s +\bs ae^{\bs{S}(2\Delta-x)}\bs s}{1-\bs ae^{\bs{S}2\Delta}\bs e},\, x \in [0,\Delta).\]
-
-The advantages of this approximation are;
-\begin{itemize}
-	\item the denominator \(1-\bs ae^{\bs{S}2\Delta}\bs s\) is provably close to \(1\) for concentrated matrix exponential distributions.
-	\item in (\ref{eqn: density approx 2}) any of the once missing mass is spread across the interval \(\calD_{\ell,j}\) proportionally to the estimated density in the region, whereas the missing mass should be more concentrated around \(\Delta\), as this is the region around which the function \(\bs \alpha e^{\bs Sx}\bs s\) fails to be like the delta function most serverely.
+Intuitively, we take the density function \(\bs ae^{\bs{S}x}\bs s\) on \(x \in [0,2\Delta)\), and `fold' it back on itself around \(\Delta\), to create a density function on \([0,\Delta)\), 
+\[\bs ae^{\bs{S}x}\bs s +\bs ae^{\bs{S}(2\Delta-x)}\bs s,\, x \in [0,\Delta).\]
+The missing mass is now proportional to 
+\[\bs a e^{\bs S 2\Delta}\bs e,\]
+which is less than the quantity \(\bs a e^{\bs S \Delta}\bs e\) from the approximation scheme (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}).
+
+%The advantages of this approximation are;
+%\begin{itemize}
+%	\item the denominator \(1-\bs ae^{\bs{S}2\Delta}\bs s\) is provably close to \(1\) for concentrated matrix exponential distributions.
+%	\item in (\ref{eqn: density approx 2}) any of the once missing mass is spread across the interval \(\calD_{\ell,j}\) proportionally to the estimated density in the region, whereas the missing mass should be more concentrated around \(\Delta\), as this is the region around which the function \(\bs \alpha e^{\bs Sx}\bs s\) fails to be like the delta function most serverely.
 This is what (\ref{eqn: density approx 3+})-(\ref{eqn: density approx 3-}) do.
-\end{itemize}
-Meanwhile, (\ref{eqn: density approx 3+})-(\ref{eqn: density approx 3-}) maintains positivity and smoothness.
-
-The difference between these approximation schemes is the way in which the vector \(\bs a\) is used.
-We generalise this idea with the concept of \emph{closing operators}.
-For \(x\in[0,\Delta)\), let \(V(x):\mathcal A \to \mathbb R,\) be a closing operator.
+%\end{itemize}
+%Meanwhile, (\ref{eqn: density approx 3+})-(\ref{eqn: density approx 3-}) maintains positivity and smoothness.
+
+The difference between the approximation schemes is the way in which the vector \(\bs a\) is used.
+We generalise this idea with the concept of \emph{closing operators} which is a linear operator \(V(x):\mathcal A \to \mathbb R,\) for each \(x\in[0,\Delta)\).
 For example, the closing operator in Equations (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}) is the operator \(V(x),\,x\in[0,\Delta)\) such that for any \(\bs a \in \mathcal A\),
@@
 \end{align}
-Similarly, in (\ref{eqn: density approx 2}) and (\ref{eqn: density approx 22})
-\begin{align}
-	\bs a V(x) &=\bs a \bs e \cfrac{\bs a e^{ \bs{S}x} \bs s}{\bs a \bs e -\bs a e^{ \bs{S}\Delta} \bs e}, \label{eqn:epan}
-%	\\\bs a V_-(x-y_\ell) &=  \cfrac{\bs a e^{ \bs{S}(x-y_\ell)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e},\label{eqn:epasan}
-\end{align}
-and in (\ref{eqn: density approx 3+}) and (\ref{eqn: density approx 3-}) 
-\begin{align}
-	\bs a V(x) &= \bs a \bs e\cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{\bs a\bs e-\bs a e^{ \bs{S}2\Delta} \bs e}.\label{eqn:546}
-%	\\\bs a V_-(x-y_\ell) &=   \cfrac{\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right)}{1-\bs a e^{ \bs{S}2\Delta} \bs e}. \label{eqn: sadlgj}
+Similarly, in (\ref{eqn: density approx 3+}) and (\ref{eqn: density approx 3-}) 
+\begin{align}
+	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right).\label{eqn:546}
 \end{align}
 The operator \(V(x)\) is a choice which forms part of the definition of the approximation scheme.
-As we shall see, given certain properties of \(V(x)\), we can prove that the approximation scheme converges, conserves probability, and ensures positivity.
-In the cases above, all the closing operators (\ref{eqn: alal})-(\ref{eqn:546}) lead to an approximation which ensures positivity due to their interpretation as probability densities.
-The operators (\ref{eqn:epan})-(\ref{eqn:546}) ensure the approximation scheme conserves probability by construction.
-We can prove that the operators (\ref{eqn: alal}) and (\ref{eqn:546}) lead to approximation schemes which converge.
-This is not to say that the scheme with \(V(x)\) as in (\ref{eqn:epan}) does not converge necessarily, just that we have not proven it.
+As we shall see, given certain properties of \(V(x)\), we can prove that the approximation scheme converges, and ensures positivity.
+In the cases above, all the closing operators (\ref{eqn: alal})-(\ref{eqn:546}) lead to an approximation which converges and, due to their interpretation as probability densities, ensure positivity.
+
+In (\ref{eqn: alal})-(\ref{eqn:546}), since \(V\) is linear, then we actually have 
+\begin{align}
+	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a V(x)
+	\\&= \mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i]V(x).
+\end{align}
+
+\begin{rem}
+	I considered taking the closing operators to be 
+	\begin{align}
+		\bs a V(x) &= \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e},\label{eqn: nonlin}
+	\end{align}
+	which also ensures that the solution always preserves mass (the result of integrating (\ref{eqn: nonlin}) over \([0,\Delta)\) is 1).
+However, (\ref{eqn: nonlin}) is not a linear operator.
+The importance of this fact is that for a linear operator,
+	\[\mathbb E[\bs A(t) V(x)] = \mathbb E[\bs A(t)]V(x),\]
+	thus, the computation of \(\mathbb E[\bs A(t) V(x)]\) can be achieved by first computing \(\mathbb E[\bs A(t)]\) and then applying \(V(x)\) to the result.
+In this sense, \(\mathbb E[\bs A(t)]\) contains all of the information about the history of the process up to time \(t\) needed to compute \(\mathbb E[\bs A(t) V(x)]\).
+This is much the same as in RAPs and QBD-RAPs where \(\mathbb E[\bs A(t)]\) is all that is required to compute probabilities about the future of the processes from time \(t\) onwards.
+
+	When \(V(x)\) is not linear then, in general, 
+	\[\mathbb E[\bs A(t) V(x)] \neq \mathbb E[\bs A(t)]V(x).\]
+	For example, with (\ref{eqn: nonlin}), 
+	\begin{align}
+		\mathbb E[\bs A(t) V(x)] = \int_{\bs a \in\mathcal A} \mathbb P(\bs A(t) \in \wrt \bs a) \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e}.\label{eqn:nonlin2}
+	\end{align}
+	While the calculation in (\ref{eqn:nonlin2}) is theoretically possible, practically it is not.
+The seemingly innocuous integral over \(\bs a \in\mathcal A\) actually amounts to an integral over all possible sample paths of the QBD-RAP (i.e.~all possible event times of the QBD-RAP before time \(t\)) and in general, there is no `nice' way to do this computation.
+
+	In the linear case, computation of
+	\[\mathbb E[\bs A(t)] = \int_{\bs a\in\mathcal A} \mathbb P(\bs A(t) \in \wrt \bs a) \bs a,\]
+	amounts to a matrix-exponential calculation. 
+\end{rem}
+
+In practice, to ensure that the approximation integrates to \(1\), we do re-normalise the estimates of the density. Specifically, we use the estimates
+	\begin{align}
+		&\cfrac{\mathbb E[\bs A(t) 1(L(t)=\ell,\phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i]V(x)}{\int_{x\in[0,\Delta)}\mathbb E[\bs A(t) 1(L(t)=\ell,\phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i]V(x)\wrt x}.
+	\end{align}

### 2021-09-22 chapter2/chapter.tex
formatting long equations, yet again

@@
 %!TEX root = ../thesis.tex
-\chapter{A stochastic modelling approach to approximation\label{sec: construction and modelling}}
-In this chapter we develop a new discretisation of a fluid queue using a \emph{quasi-birth-and-death process with rational arrival process components} (QBD-RAP), to be described in more detail later, and prove convergence of the approximation scheme.
+\chapter{A stochastic modelling approach to approximating fluid queues\label{sec: construction and modelling}}
+In this chapter we develop a new discretisation of a fluid queue using a \emph{quasi-birth-and-death process with rational arrival process components} (QBD-RAP).
+In the following chapter, Chapter \ref{sec: conv} we prove the scheme converges weakly.
 To improve upon the discretisation of \cite{bo2013}, we argue that the shift from QBD to QBD-RAP is necessary.
@@
 The structure of this chapter is as follows.
-In Section~\ref{sec: prelim} we provide a brief background on matrix exponential random variables and the stochastic processes appearing in this work, namely fluid queues, quasi-birth-and-death (QBD) processes, rational arrival processes (RAPs), and quasi-birth-and-death processes with rational arrival process components (QBD-RAPs).
-Section~\ref{sec: inspiration} motivates the idea using the QBD approximation of \citep{bo2014}.
-Sections \ref{sec: modelling} and \ref{sec: zero states} describe the modelling of certain events of the fluid queue with matrix exponential distributions to ultimately construct the behaviour of the QBD-RAP.
-Section~\ref{sec: dynamics} describes the dynamics of the resulting construction and constructs the level process of the QBD-RAP.
+In Section~\ref{sec: prelim} we provide a brief background on matrix exponential random variables and and quasi-birth-and-death processes with rational arrival process components (QBD-RAPs).
+Section~\ref{sec: inspiration} motivates the idea using the QBD approximation of \cite{bo2014}.
+Sections \ref{sec: modelling} and \ref{sec: zero states} describe the modelling of certain events of the fluid queue with matrix exponential distributions to ultimately construct the behaviour of the QBD-RAP on the event that the QBD-RAP remains in a given level.
+Section~\ref{sec: dynamics} describes the dynamics of the construction and also constructs the level process of the QBD-RAP.
 Section~\ref{sec: boundary conditions} deals with modelling boundary behaviour.
 Section~\ref{sec: initial conditions} describes how to model initial conditions of the fluid queue.
-Section~\ref{sec: closing} introduces the concept of a \emph{closing operator} which maps states of the QBD-RAP to estimates of the distribution of the fluid queue.
+Section~\ref{sec: closing} introduces the concept of a \emph{closing operator} which maps the state space of the QBD-RAP to estimates of the density of the fluid queue.
 
 \section{Preliminaries}\label{sec: prelim}
-This work approximates fluid queues using other stochastic processes, specifically quasi-birth-and-death processes with rational arrival process components (QBD-RAPs).
+%This work approximates fluid queues using other stochastic processes, specifically quasi-birth-and-death processes with rational arrival process components (QBD-RAPs).
 The building blocks of QBD-RAPs are rational arrival processes (RAPs) and the building blocks of RAPs are matrix-exponential distributions.
@@
 
-\section{Computing \(\bs D\)}
+\subsection{Computing \(\bs D\)}
 In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hhat2020}.
 For this class of CMEs, we take the index \(p\) to be the order of the representation.
-Since the variance of \(Z^{(2p)}\) and \(Z^{(2p-1)}\) are relatively similar \citep{hhat2020}, we take \(p\) to be odd.
-For a given CME, with representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part.
-
-We numerically evaluate \(\bs{D}\) where
+Moreover, we take \(p\) to be odd.
+The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hhat2020}.
+Hence, if we construct a QBD-RAP approximation of with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\).
+However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.
+
+For a given CME with odd order, \(p\), and representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part.
+
+We numerically evaluate the matrix \(\bs{D}\) where
 \begin{align*}
@@
 using a trapezoidal rule as follows.
-
 For CMEs, the vector function \(\bs a(t) := \cfrac{\bs \alpha e^{\bs St}}{\bs \alpha e^{\bs S t} \bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
@@
 	\intertext{By periodicity, then \(e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs a(k\rho +t)=e^{ -\lambda t} \bs f(t)\cdot\bs a(t)\), hence (\ref{eqn: sc6789gGHJ}) is equal to}
-	&\sum_{k=0}^\infty (e^{\lambda \rho })^k\int_{0}^{\rho } e^{\lambda t}e^{-\lambda t}\bs f(t)\cdot\bs a(t)\wrt t \nonumber 
-	\\&= \cfrac{1}{1 - e^{\lambda \rho }}\int_{0}^{\rho } \bs f(t)\cdot\bs a(t)\wrt t,\label{eqn: hdsv45KJ}
+	&\sum_{k=0}^\infty (e^{\lambda \rho })^k\int_{0}^{\rho } e^{\lambda t}e^{-\lambda t}\bs f(t)\cdot\bs a(t)\wrt t  
+	= \cfrac{1}{1 - e^{\lambda \rho }}\int_{0}^{\rho } \bs f(t)\cdot\bs a(t)\wrt t,\label{eqn: hdsv45KJ}
 \end{align}
@@
 
-To approximate (\ref{eqn: hdsv45KJ}) numerically, we first partition \(\left[0,\rho \right]\) into \(N\) equal-width intervals \(\left[t_n,t_{n+1}\right)\), where \(t_n = (n-1)\rho /N\), \(n=1,2,...,N+1\).
+To approximate (\ref{eqn: hdsv45KJ}) numerically, we first partition \(\left[0,\rho \right)\) into \(N\) equal-width intervals \(\left[t_n,t_{n+1}\right)\), where \(t_n = (n-1)\rho /N\), \(n=1,2,...,N+1\).
 On \(\left[t_n,t_{n+1}\right)\) we approximate the orbit \(\bs a(t)\) by a constant \(\bs a(t) \approx \bs a_n := \cfrac{1}{2}\left( \bs a(t_n) + \bs a(t_{n+1}) \right), \, t \in \left[t_n,t_{n+1}\right)\).
@@
 \begin{align}
-		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x,\label{eqn: density approx aug}
+		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
+		\\&\qquad\times\bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x,\label{eqn: density approx aug}
 		\intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
-		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
+		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
+		\\&\qquad\times\bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
 \end{align}
@@
 	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}(y_{\ell+1}-x)} \bs s \wrt x
-	\\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
-	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
+	\\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0,
+	\\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
 	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_+\cup \calS_{+0}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
@@
 \begin{align}
-	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a V(x)
+	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a V(x) \nonumber 
 	\\&= \mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i]V(x).

### 2021-09-22 chapter1/chapter.tex
Ch 2 modelling, Ch 3 analysis,Appendix A technical bounds, first drafts

@@
 
-\subsection{Fluid queues}
+\section{Fluid queues}
 A fluid queue is a two-dimensional stochastic process \(\{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer.
@@
 Discretisation methods approximate the operator on the right-hand side of (\ref{eqn: pde}) by a matrix, in our case, by the generator of a QBD-RAP.
+
+\section{Matrix-exponential distributions}
+Here we recount some facts about matrix exponential distributions.
+See \citep{MEinAP} for a more detailed exposition.
+A random variable, \(Z\), is said to have a matrix-exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
+The density function of \(Z\) is given by \(f_Z(x) = \bs\alpha e^{\bs S x}\bs s\).
+The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) There is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility.
+These condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\).
+However, in general there is no way to determine whether, given a triplet \((\bs \alpha, \bs S, \bs s)\) whether it is a representation of a matrix-exponential distribution, or not.
+Nonetheless, some properties of a triplet \((\bs \alpha, \bs S, \bs s)\) are known, such as the following, which is used in the characterisation of QBD-RAPs.
+\begin{thm}[Theorem 4.1.3, \cite{MEinAP}]
+	The density function of a matrix-exponential distribution with representation \((\bs\alpha, \bs S, \bs s)\) can be expressed in terms of real-valued constants as 
+	\begin{align}
+		\psi(x)&=\sum_{j=1}^{m_1} \sum_{k=1}^{p_j} c_{jk} \cfrac{x^{k-1}}{(k-1)!} e^{\mu_jx} + \sum_{j=1}^{m+2}\sum_{k=1}^{q_j} d_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\cos(\sigma_jx) \nonumber 
+		\\&\quad{}+ \sum_{j=1}^{m_2}\sum_{k=1}^{q_j}e_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\sin(\sigma_jx), \label{eqn: spec}
+	\end{align}
+	for integers \(m_1,\,m_2,\,p_j,\) and \(q_j\) and some real constants \(c_{jk},\,d_{jk},\,e_{jk},\,\mu_j,\,\eta_j,\) and \(\sigma_j\).
+Here \(\mu_j,\, j=1,\dots,m_1\) are the real eigenvalues of \(\bs S\), while \(\eta_j+i\sigma_j, \, \eta_j-i\sigma_j,\, j=1,\dots,m_2\) denote its complex eigenvalues, which come in conjugate pairs.
+Thus \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m+1}p_j + 2\sum_{j=1}^{m_2}q_j\).
+\end{thm}
+\begin{thm}[Theorem 4.1.4, \cite{MEinAP}]\label{thm: 4.1.4}
+	Consider the nonvanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}=0,\,d_{jk} =0,\) or \(e_{jk}=0\).
+Among the corresponding eigenvalues \(\lambda_j\), there is a real dominating eigenvalue \(\kappa\), say.
+That is, \(\kappa\) is real, \(\kappa \geq Re(\lambda_j)\) for all \(j\), and the multiplicity of \(\kappa\) is at least the multiplicity of every other eigenvalue with real part \(\kappa\).
+\end{thm}
+\begin{cor}[Corollary 4.1.5, \cite{MEinAP}]
+If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix-exponential distribution, then \(\bs S\) has a real dominating eigenvalues.
+\end{cor}
+\begin{thm}[Theorem 4.1.6, \cite{MEinAP}]
+	Let \(Z\) be a matrix-exponentially distributed random variable with density (\ref{eqn: spec}). Then the dominant real eigenvalue \(\kappa\) of Theorem~\ref{thm: 4.1.4} is strictly negative. 
+\end{thm}
+We define \(dev(\bs S)\) to be the real dominating eigenvalue of \(\bs S\), that is \(dev(\bs S)=\kappa\) in Theorem~\ref{thm: 4.1.4}.
+
+The class of matrix-exponential distribution is characterised as the class of probability distributions which have a rational Laplace transform.
+That is, \(\displaystyle \int_{x=0}^\infty e^{-\lambda x}\bs \alpha e^{\bs S x}\bs s \wrt x\) is a ratio of two polynomial functions in \(\lambda\).
+Matrix exponential distributions are an extension of Phase-type distributions, where for the latter, \(\bs S\) must be a sub-generator matrix of a CTMC, \(\bs s = -\bs S\bs e\) where \(\bs e\) is a \(1\times p\) vector of ones, and \(\bs \alpha\) is a discrete probability distribution.
+
+A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix-exponential distribution with this representation.
+The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\).
+Representations of matrix-exponential distributions are not unique \citep{MEinAP}.
+A representation is called \emph{minimal} when \(\bs S\) has the smallest possible dimension.
+Throughout this work, we assume that the representation of any matrix exponential distribution is minimal.
+Let \(\bs e_i\) be a vector with a 1 in the \(i\)th position and zeros elsewhere.
+We assume that \(\bs s = -\bs S\bs e\), and that \((\bs e_i,\bs S,\bs s)\) for \(i=1,\dots,p\) are representations of matrix exponential distributions.
+It is always possible to find such a representation \cite[Theorem 4.5.17, Corollary 4.5.18]{MEinAP}.
+As such, we abbreviate our notation \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to \(Z\sim ME(\bs \alpha, \bs S)\).
+Further, given \(\bs s=-\bs S\bs e\) then observe that  \(\displaystyle \int_{x=0}^\infty e^{\bs S x} \bs s\wrt x = (-\bs S)^{-1}\bs s= \bs e\).
+
+For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal A\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a matrix exponential distribution.
+%By CITE it is known that \(\mathcal A\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).
+
+%\section{Rational arrival processes}
+%Let \(N\) be a simple point process, with event time \(Y_0=0<Y_1<Y_2<\cdot\).
+Let \(\{N(t)\}_{t\geq0}\) be the right-continuous counting process associated with \(N\); \(N(t)\) returns the number of events by time \(t\).
+Denote by \(f_{N,n}(y_1,\dots,y_n)\) the joint density of \(Y_1, Y_2-Y_1, \dots,Y_n-Y_{n-1}\), the first \(n\) inter-arrival times.
+For a matrix \(\bs B\) let \(dev(\bs B)\) denote the dominant eigenvalue of \(\bs B\) (the one with maximal real part).
+Theorem 1.1 of Asmussen and Bladt CITE states that the point process \(N\) is a RAP if there exists matrices \(\bs S\), \(\bs D\), a row vector \(\bs\alpha\), and a column vector \(\bs s\) such that \(dev(\bs S)<1\), \(dev(\bs S+\bs D)=0\), \((\bs S+\bs D)\bs e=0,\) and
+%\[f_{N,n}(y_1,\dots,y_n) = \bs \alpha e^{\bs Sy_1} \bs D e^{\bs Sy_1} \bs D \dots e^{\bs Sy_n} \bs s.\]
+%Here \(\bs s\) can be taken to be \(\bs D e\).
+The \(n\)th inter-arrival, \(T_n-T_{n-1}\), has a matrix exponential distribution with density \(\bs \alpha(-\bs S\bs D )^{n-1}e^{\bs S x_n}\bs s\).
+Denote such a process \(N \sim RAP(\bs \alpha, \bs S, \bs D)\).
+%
+%Asmussen and Bladt CITE, Corollary 2.2, show that associated with the RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\), 
+%\[\bs A(t) = \cfrac{\bs \alpha\left( \prod_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs S\right)e^{\bs S t-Y_{N(t)}}}{\bs \alpha\left( \prod_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs S\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
+%Thus, \(\{\bs A(t)\}\) is a piecewise-deterministic Markov process where, in between jumps \(\{\bs A(t)\}\) according to 
+%\[\bs A(t) = \cfrac{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}}{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}\bs e},\]
+%where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
+The process \(\{\bs A(t)\}\) jumps at event times of \(N\).
+At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\), and upon a jump at time \(t\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D/\bs A(t^-) \bs D\bs e\).
+%
+%RAPs are an extension of Markovian arrival processes (MAPs) to include matrix-exponential inter-arrival times.
+For MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain.
+%
+%Intuitively, \(\bs A(t)\) encodes all of the information about the jump times of the RAP up to time \(t\).
+Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\).
+Then \(N\mid \mathcal F_t \equiv N\mid \bs A(t) \sim ME(\bs A(t), \bs S, \bs S)\).
+In words, the future of the point process after time \(t\) given all of the information about the process up to and including time \(t\), is distributed as a RAP with initial vector \(\bs A(t)\).
+
+\section{QBD-RAPs}
+As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes.
+To define a QBD-RAP we first need a Batch (Marked) RAP.
+
+Let \(\mathscr K\subset \mathbb Z\) be a set of \emph{marks}.
+Let \(N\) be a point process, and \(Y_0=0<Y_1<Y_2\cdots\) be event times of \(N\).
+Let \(\{N(t)\}\) be the counting process associated with \(N\) such that \(N(t)\) returns the number of events by time \(t\).
+Associated with the \(n\)th event is a mark \(M_n\).
+For \(i\in\mathscr K\), let \(N_i\) be simple point processes associated with events with marks of type \(i\) only, and let \(\{N_i(t)\}_{t\geq 0}\) be the associated counting processes of events of mark \(i\).
+For a matrix \(\bs B\) let \(dev(\bs B)\) denote the real part of the dominant eigenvalue of \(\bs B\) (the one with maximal real part).
+Denote by \(f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n)\) the joint density, probability mass function of the first \(n\) inter-arrival times, \(Y_1,Y_2-Y_1,\dots,Y_n-Y_{n-1}\), and the associated marks \(M_n\).
+From Bean and Nielsen \cite[Theorem 1]{bn2010} we have the following.
+\begin{thm}
+A process \(N\) is a Marked RAP if there exist matrices \(\bs S\), \(\bs D_i,\,i\in\mathscr K\), and a row vector \(\bs \alpha\) such that \(dev(\bs S)<0\), \(dev(\bs S+\bs D)=0\), \((\bs S+\bs D)\bs e = 0\), \(\bs D = \sum_{i\in\mathscr K} \bs D_i\), and
+\begin{align}f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n) = \bs \alpha e^{\bs S y_1}\bs D_{m_1} e^{\bs S y_2} \bs D_{m_2}\dots e^{\bs S y_n} \bs D_{m_n}\bs e.\label{eqn: brap}\end{align}
+Conversely, if a point process has the property (\ref{eqn: brap}) then it is a Marked RAP.
+\end{thm}
+
+Denote such a process \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\mathscr K).\)
+
+Also from \cite{bn2010}, associated with a Marked RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\)
+\[\bs A(t) = \cfrac{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S (t-Y_{N(t)})}}{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
+Thus, \(\{\bs A(t)\}\) is a piecewise-deterministic Markov process where, in between events \(\{\bs A(t)\}\) evolves deterministically according to 
+\[\bs A(t) = \cfrac{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}}{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}\bs e},\]
+where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
+The process \(\{\bs A(t)\}\) "jumps" at event times of \(N\) (the process may not always jump at these times, but typically the dynamics change discontinuously at this point).
+At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\).
+Upon an event the event is associated with mark \(i\) with probability \(\bs A(t) \bs D_i\bs e/\bs A(t) \bs D\bs e\).
+Upon on an event at time \(t\) with mark \(i\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D_i/\bs A(t^-) \bs D_i\bs e\).
+
+Marked RAPs are an extension of Marked Markovian arrival processes to include matrix-exponential inter-arrival times.
+For Marked MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain.
+
+Intuitively, \(\bs A(t)\) encodes all of the information about the event times of the Marked RAP and associated marks up to time \(t\) that is needed to determine the future behaviour of the point process.
+Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\).
+Then \(N\mid \mathcal F_t \equiv N\mid \bs A(t) \sim BRAP(\bs A(t), \bs S, \bs D_i,i\in\mathscr K)\).
+In words, the future of the point process after time \(t\) given all of the information about the process up to and including time \(t\), is distributed as a Marked RAP with initial vector \(\bs A(t)\).
+
+Now consider a Marked RAP, \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\{-1,0,+1\}).\) The process \(\{(L(t),\bs A(t))\}_{t\geq0}\) formed by letting \(L(t) = N_{+1}(t) - N_{-1}(t)\) is a QBD-RAP.

### 2021-09-22 chapter2/chapter.tex
Ch 2 modelling, Ch 3 analysis,Appendix A technical bounds, first drafts

@@
 The structure of this chapter is as follows.
-In Section~\ref{sec: prelim} we provide a brief background on matrix exponential random variables and and quasi-birth-and-death processes with rational arrival process components (QBD-RAPs).
 Section~\ref{sec: inspiration} motivates the idea using the QBD approximation of \cite{bo2014}.
@@
 Section~\ref{sec: closing} introduces the concept of a \emph{closing operator} which maps the state space of the QBD-RAP to estimates of the density of the fluid queue.
-
-\section{Preliminaries}\label{sec: prelim}
-%This work approximates fluid queues using other stochastic processes, specifically quasi-birth-and-death processes with rational arrival process components (QBD-RAPs).
-The building blocks of QBD-RAPs are rational arrival processes (RAPs) and the building blocks of RAPs are matrix-exponential distributions.
-In this section, we introduce these concepts.
-
-\subsection{Matrix-exponential distributions}
-Here we recount some facts about matrix exponential distributions.
-See \citep{MEinAP} for a more detailed exposition.
-A random variable, \(Z\), is said to have a matrix-exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
-The density function of \(Z\) is given by \(f_Z(x) = \bs\alpha e^{\bs S x}\bs s\).
-The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) There is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility.
-These condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\).
-However, in general there is no way to determine whether, given a triplet \((\bs \alpha, \bs S, \bs s)\) whether it is a representation of a matrix-exponential distribution, or not.
-Nonetheless, some properties of a triplet \((\bs \alpha, \bs S, \bs s)\) are known, such as the following, which is used in the characterisation of QBD-RAPs.
-\begin{thm}[Theorem 4.1.3, \cite{MEinAP}]
-	The density function of a matrix-exponential distribution with representation \((\bs\alpha, \bs S, \bs s)\) can be expressed in terms of real-valued constants as 
-	\begin{align}
-		\psi(x)&=\sum_{j=1}^{m_1} \sum_{k=1}^{p_j} c_{jk} \cfrac{x^{k-1}}{(k-1)!} e^{\mu_jx} + \sum_{j=1}^{m+2}\sum_{k=1}^{q_j} d_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\cos(\sigma_jx) + \sum_{j=1}^{m_2}\sum_{k=1}^{q_j}e_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\sin(\sigma_jx), \label{eqn: spec}
-	\end{align}
-	for integers \(m_1,\,m_2,\,p_j,\) and \(q_j\) and some real constants \(c_{jk},\,d_{jk},\,e_{jk},\,\mu_j,\,\eta_j,\) and \(\sigma_j\).
-Here \(\mu_j,\, j=1,\dots,m_1\) are the real eigenvalues of \(\bs S\), while \(\eta_j+i\sigma_j, \, \eta_j-i\sigma_j,\, j=1,\dots,m_2\) denote its complex eigenvalues, which come in conjugate pairs.
-Thus \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m+1}p_j + 2\sum_{j=1}^{m_2}q_j\).
-\end{thm}
-\begin{thm}[Theorem 4.1.4, \cite{MEinAP}]\label{thm: 4.1.4}
-	Consider the nonvanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}=0,\,d_{jk} =0,\) or \(e_{jk}=0\).
-Among the corresponding eigenvalues \(\lambda_j\), there is a real dominating eigenvalue \(\kappa\), say.
-That is, \(\kappa\) is real, \(\kappa \geq Re(\lambda_j)\) for all \(j\), and the multiplicity of \(\kappa\) is at least the multiplicity of every other eigenvalue with real part \(\kappa\).
-\end{thm}
-\begin{cor}[Corollary 4.1.5, \cite{MEinAP}]
-If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix-exponential distribution, then \(\bs S\) has a real dominating eigenvalues.
-\end{cor}
-\begin{thm}[Theorem 4.1.6, \cite{MEinAP}]
-	Let \(Z\) be a matrix-exponentially distributed random variable with density (\ref{eqn: spec}). Then the dominant real eigenvalue \(\kappa\) of Theorem~\ref{thm: 4.1.4} is strictly negative. 
-\end{thm}
-We define \(dev(\bs S)\) to be the real dominating eigenvalue of \(\bs S\), that is \(dev(\bs S)=\kappa\) in Theorem~\ref{thm: 4.1.4}.
-
-The class of matrix-exponential distribution is characterised as the class of probability distributions which have a rational Laplace transform.
-That is, \(\displaystyle \int_{x=0}^\infty e^{-\lambda x}\bs \alpha e^{\bs S x}\bs s \wrt x\) is a ratio of two polynomial functions in \(\lambda\).
-Matrix exponential distributions are an extension of Phase-type distributions, where for the latter, \(\bs S\) must be a sub-generator matrix of a CTMC, \(\bs s = -\bs S\bs e\) where \(\bs e\) is a \(1\times p\) vector of ones, and \(\bs \alpha\) is a discrete probability distribution.
-
-A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix-exponential distribution with this representation.
-The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\).
-Representations of matrix-exponential distributions are not unique \citep{MEinAP}.
-A representation is called \emph{minimal} when \(\bs S\) has the smallest possible dimension.
-Throughout this work, we assume that the representation of any matrix exponential distribution is minimal.
-Let \(\bs e_i\) be a vector with a 1 in the \(i\)th position and zeros elsewhere.
-We assume that \(\bs s = -\bs S\bs e\), and that \((\bs e_i,\bs S,\bs s)\) for \(i=1,\dots,p\) are representations of matrix exponential distributions.
-It is always possible to find such a representation \cite[Theorem 4.5.17, Corollary 4.5.18]{MEinAP}.
-As such, we abbreviate our notation \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to \(Z\sim ME(\bs \alpha, \bs S)\).
-Further, given \(\bs s=-\bs S\bs e\) then observe that  \(\displaystyle \int_{x=0}^\infty e^{\bs S x} \bs s\wrt x = (-\bs S)^{-1}\bs s= \bs e\).
-
-For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal A\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a matrix exponential distribution.
-%By CITE it is known that \(\mathcal A\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).
-
-%\subsection{Rational arrival processes}
-%Let \(N\) be a simple point process, with event time \(Y_0=0<Y_1<Y_2<\cdot\).
-Let \(\{N(t)\}_{t\geq0}\) be the right-continuous counting process associated with \(N\); \(N(t)\) returns the number of events by time \(t\).
-Denote by \(f_{N,n}(y_1,\dots,y_n)\) the joint density of \(Y_1, Y_2-Y_1, \dots,Y_n-Y_{n-1}\), the first \(n\) inter-arrival times.
-For a matrix \(\bs B\) let \(dev(\bs B)\) denote the dominant eigenvalue of \(\bs B\) (the one with maximal real part).
-Theorem 1.1 of Asmussen and Bladt CITE states that the point process \(N\) is a RAP if there exists matrices \(\bs S\), \(\bs D\), a row vector \(\bs\alpha\), and a column vector \(\bs s\) such that \(dev(\bs S)<1\), \(dev(\bs S+\bs D)=0\), \((\bs S+\bs D)\bs e=0,\) and
-%\[f_{N,n}(y_1,\dots,y_n) = \bs \alpha e^{\bs Sy_1} \bs D e^{\bs Sy_1} \bs D \dots e^{\bs Sy_n} \bs s.\]
-%Here \(\bs s\) can be taken to be \(\bs D e\).
-The \(n\)th inter-arrival, \(T_n-T_{n-1}\), has a matrix exponential distribution with density \(\bs \alpha(-\bs S\bs D )^{n-1}e^{\bs S x_n}\bs s\).
-Denote such a process \(N \sim RAP(\bs \alpha, \bs S, \bs D)\).
-%
-%Asmussen and Bladt CITE, Corollary 2.2, show that associated with the RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\), 
-%\[\bs A(t) = \cfrac{\bs \alpha\left( \prod_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs S\right)e^{\bs S t-Y_{N(t)}}}{\bs \alpha\left( \prod_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs S\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
-%Thus, \(\{\bs A(t)\}\) is a piecewise-deterministic Markov process where, in between jumps \(\{\bs A(t)\}\) according to 
-%\[\bs A(t) = \cfrac{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}}{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}\bs e},\]
-%where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
-The process \(\{\bs A(t)\}\) jumps at event times of \(N\).
-At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\), and upon a jump at time \(t\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D/\bs A(t^-) \bs D\bs e\).
-%
-%RAPs are an extension of Markovian arrival processes (MAPs) to include matrix-exponential inter-arrival times.
-For MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain.
-%
-%Intuitively, \(\bs A(t)\) encodes all of the information about the jump times of the RAP up to time \(t\).
-Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\).
-Then \(N\mid \mathcal F_t \equiv N\mid \bs A(t) \sim ME(\bs A(t), \bs S, \bs S)\).
-In words, the future of the point process after time \(t\) given all of the information about the process up to and including time \(t\), is distributed as a RAP with initial vector \(\bs A(t)\).
-
-\subsection{QBD-RAPs}
-As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes.
-To define a QBD-RAP we first need a Batch (Marked) RAP.
-
-Let \(\mathscr K\subset \mathbb Z\) be a set of \emph{marks}.
-Let \(N\) be a point process, and \(Y_0=0<Y_1<Y_2\cdots\) be event times of \(N\).
-Let \(\{N(t)\}\) be the counting process associated with \(N\) such that \(N(t)\) returns the number of events by time \(t\).
-Associated with the \(n\)th event is a mark \(M_n\).
-For \(i\in\mathscr K\), let \(N_i\) be simple point processes associated with events with marks of type \(i\) only, and let \(\{N_i(t)\}_{t\geq 0}\) be the associated counting processes of events of mark \(i\).
-For a matrix \(\bs B\) let \(dev(\bs B)\) denote the real part of the dominant eigenvalue of \(\bs B\) (the one with maximal real part).
-Denote by \(f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n)\) the joint density, probability mass function of the first \(n\) inter-arrival times, \(Y_1,Y_2-Y_1,\dots,Y_n-Y_{n-1}\), and the associated marks \(M_n\).
-From Bean and Nielsen \cite[Theorem 1]{bn2010} we have the following.
-\begin{thm}
-A process \(N\) is a Marked RAP if there exist matrices \(\bs S\), \(\bs D_i,\,i\in\mathscr K\), and a row vector \(\bs \alpha\) such that \(dev(\bs S)<0\), \(dev(\bs S+\bs D)=0\), \((\bs S+\bs D)\bs e = 0\), \(\bs D = \sum_{i\in\mathscr K} \bs D_i\), and
-\begin{align}f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n) = \bs \alpha e^{\bs S y_1}\bs D_{m_1} e^{\bs S y_2} \bs D_{m_2}\dots e^{\bs S y_n} \bs D_{m_n}\bs e.\label{eqn: brap}\end{align}
-Conversely, if a point process has the property (\ref{eqn: brap}) then it is a Marked RAP.
-\end{thm}
-
-Denote such a process \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\mathscr K).\)
-
-Also from \cite{bn2010}, associated with a Marked RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\)
-\[\bs A(t) = \cfrac{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S (t-Y_{N(t)})}}{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
-Thus, \(\{\bs A(t)\}\) is a piecewise-deterministic Markov process where, in between events \(\{\bs A(t)\}\) evolves deterministically according to 
-\[\bs A(t) = \cfrac{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}}{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}\bs e},\]
-where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
-The process \(\{\bs A(t)\}\) "jumps" at event times of \(N\) (the process may not always jump at these times, but typically the dynamics change discontinuously at this point).
-At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\).
-Upon an event the event is associated with mark \(i\) with probability \(\bs A(t) \bs D_i\bs e/\bs A(t) \bs D\bs e\).
-Upon on an event at time \(t\) with mark \(i\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D_i/\bs A(t^-) \bs D_i\bs e\).
-
-Marked RAPs are an extension of Marked Markovian arrival processes to include matrix-exponential inter-arrival times.
-For Marked MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain.
-
-Intuitively, \(\bs A(t)\) encodes all of the information about the event times of the Marked RAP and associated marks up to time \(t\) that is needed to determine the future behaviour of the point process.
-Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\).
-Then \(N\mid \mathcal F_t \equiv N\mid \bs A(t) \sim BRAP(\bs A(t), \bs S, \bs D_i,i\in\mathscr K)\).
-In words, the future of the point process after time \(t\) given all of the information about the process up to and including time \(t\), is distributed as a Marked RAP with initial vector \(\bs A(t)\).
-
-Now consider a Marked RAP, \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\{-1,0,+1\}).\) The process \(\{(L(t),\bs A(t))\}_{t\geq0}\) formed by letting \(L(t) = N_{+1}(t) - N_{-1}(t)\) is a QBD-RAP.
 
@@
 To this end, we observe that, if the variance of \(Z_i\) is sufficiently low, then for \(\varepsilon>0\) and \(u<\Delta/|c_i|-\varepsilon\), then
-\begin{align}
-	\mathbb P(R_i(u) \in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon)) 
-	&= \mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon)\mid Z_i > u) 
+\begin{align*}
+	&\mathbb P(R_i(u) \in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon)) 
+	\\&= \mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon)\mid Z_i > u) 
 	\\&= \cfrac{\mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon))}{\mathbb P(Z_i > u)} 
@@
 	\\&\geq \cfrac{1-\cfrac{\var(Z)}{\varepsilon^2}}{1} 
-\end{align}
+\end{align*}
 since, by Chebyshev's inequality, \(\mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon)) \geq 1- \cfrac{\var(Z_i)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(X)^{1/3}\), then 

### 2021-10-13 chapter1/chapter.tex
easy parts of the first round of revisions on Ch 1/2 from Nigel

@@
 In such situations, it is the \emph{weak solution} to (\ref{eqn: pde}) that we seek.
+A weak solution satisfies
+\begin{equation}
+	-\int_x\int_t \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= \int_{x}\int_t\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x + \int_x\int_t\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
+\end{equation}
+for every vector of test functions, \(\bs \psi(x,t)\), wich are smooth and have compact support.
 
@@
 
-\section{QBD-RAPs}
+\section{QBD-RAPs}\label{sec: qbd-rap}
 As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes.

### 2021-10-13 chapter2/chapter.tex
easy parts of the first round of revisions on Ch 1/2 from Nigel

@@
 To improve upon the discretisation of \cite{bo2013}, we argue that the shift from QBD to QBD-RAP is necessary.
-The approximation can be made arbitrarily accurate.
+The QBD construction of \cite{bo2013} uses Erlang distributions to approximate the deterministic time that the fluid queue spends in an interval on the event that the phase is constant -- it is well known that the Erlang distribution is the distribution with the smallest variance in the class of Phase-type distributions of a given order \cite{as1987}, and in this sense is the best approximation to the determinstic behaviour we want to capture.
+
+By choosing a matrix-exponential distribution with sufficiently small variance, the QBD-RAP approximation can be made arbitrarily accurate.
 Due to its stochastic interpretation, the QBD-RAP discretisation ensures solutions are positive, even when discontinuities are present.
@@
 But this is the same discretisation we would get if we were to use intervals of width \(\Delta/2\) in the QBD approximation of \cite{bo2013}, with the rows/columns reordered.
-Indeed, this appears to be the best we can do with a QBD-MAP approximation as the Erlang distribution is known to be the least variable Phase-type distribution for a given order \citep{as1987} and therefore gives the closest approximation to the deterministic behaviour we are trying to approximate.
-This necessitates the extension to QBD-RAPs, whereby we can find more concentrated distributions than any Phase-type, but retain (some of) the stochastic interpretation and matrix-analytic tools.
+Indeed, this appears to be the best we can do with a QBD approximation as the Erlang distribution is known to be the least variable Phase-type distribution for a given order \citep{as1987} and therefore gives the closest approximation to the deterministic behaviour we are trying to approximate.
+This necessitates the extension to QBD-RAPs, whereby we can find more concentrated distributions than any Phase-type and retain enough stochastic interpretation and matrix-analytic tools.
 
@@
 \section{Time to exit an interval}\label{sec: modelling}
-Consider partition of the state space of the level of a fluid queue, \([0,M]\) into \(K\) intervals of width \(\Delta = M/(K+1)\), specifically \(\mathcal D_{k,i} = [k\Delta,(k+1)\Delta)\) if \(i\in\mathcal S_+\) and \(\mathcal D_{k,i} = (k\Delta,(k+1)\Delta]\) if \(i\in\mathcal S_-\), \(k=0,\dots,K+1\).
+Consider the partition of the state space of the level of a fluid queue, \([0,M]\) into \(K\) intervals of width \(\Delta = M/(K+1)\), specifically \(\mathcal D_{k,i} = [k\Delta,(k+1)\Delta)\) if \(i\in\mathcal S_+\) and \(\mathcal D_{k,i} = (k\Delta,(k+1)\Delta]\) if \(i\in\mathcal S_-\), \(k=0,\dots,K+1\).
 The distinction between \(\mathcal D_{k,i}\) for \(i\in\mathcal S_+\) and \(\calS_-\) is a technical one and will be discussed later.
@@
 \end{align}
-is c\'adl\'ag.
+is right-continuous with left limits (c\'adl\'ag).
 The expression (\ref{eqn: kfj}) is the discrete process of the fluid-queue which the level process of the QBD-RAP will approximate.
@@
 The random variable \(Z_i\) has mean \(\Delta/|c_i|\).
-Further, let \(R_i(u)\) be the residual time, \(R_i(u) = Z_i-u\mid \{Z_i-u>0\}\), \(i\in\calS\).
+Further, let \(R_i(u)\) be the residual time, \(R_i(u) = (Z_i-u)1\{Z_i-u>0\}\), \(i\in\calS\).
 Suppose first that \(X(t)=y_k\) and \(\varphi(t)=i\in\mathcal S_+\), so that \(X(t)\) will leave \(\mathcal D_{k,i}\) in exactly \(\Delta/|c_i|\) units of time on the event that the phase process remains in \(i\) for at least this amount of time.
@@
 \[f_{R_i(u)}(r) = \cfrac{\bs\alpha e^{\bs S_i(u+ r) } \bs s_i}{\bs\alpha e^{\bs S_iu } \bs e} = \bs A(t+u) e^{\bs S_i r}  \bs s_i.\]
-Here, the event \(Z_i>u\) approximates the event \(X(t)\in\calD_{k,i}\) and we want the residual time, \(R_i(u)\), to approximate the time until \(\{X(t)\}\) leaves \(\calD_{k,i}\).
+Here, the event \(Z_i>u\) approximates the event \(X(t+u)\in\calD_{k,i}\) and we want the residual time, \(R_i(u)\), to approximate the time until \(\{X(t)\}\) leaves \(\calD_{k,i}\).
 That is, we want \(R_i(u)\) to approximate a deterministic random variable at \(\Delta/|c_i| - u\).
-To this end, we observe that, if the variance of \(Z_i\) is sufficiently low, then for \(\varepsilon>0\) and \(u<\Delta/|c_i|-\varepsilon\), then
-\begin{align*}
-	&\mathbb P(R_i(u) \in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon)) 
-	\\&= \mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon)\mid Z_i > u) 
-	\\&= \cfrac{\mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon))}{\mathbb P(Z_i > u)} 
-	\\&= \cfrac{\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))}{\mathbb P(Z_i > u)} 
+To this end, we observe that for \(\varepsilon>0\) and \(u<(\Delta-\varepsilon)/|c_i|\), then
+\begin{align*}
+	&\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|-u) ) 
+	\\&= \mathbb P(Z_i\in ((\Delta-\varepsilon)/|c_i|, (\Delta+\varepsilon)/|c_i|) ) 
+	\\&= \mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))
 %	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
 %	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
-	\\&\geq \cfrac{1-\cfrac{\var(Z)}{\varepsilon^2}}{1} 
-\end{align*}
-since, by Chebyshev's inequality, \(\mathbb P(Z_i\in (\Delta/|c_i|-\varepsilon, \Delta/|c_i| + \varepsilon)) \geq 1- \cfrac{\var(Z_i)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(X)^{1/3}\), then 
-\[\mathbb P(R_i(u) \in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))\geq 1-\var(Z)^{1/3} \approx 1. \]
-That is, when the variance of \(Z_i\) is low, the residual time \(R_i(u)\) will be concentrated around \(\Delta/|c_i| - u\), as required.
-Figure \ref{fig:residual distributions} gives an example of a density function for a \(Z_i\) with mean \(\Delta=1\) and \(c_i=1\), as well as the density function of \(R_i(0.3)\) and \(R_i(0.5)\) for comparison.
-Observe that the density of the residual life \(R_i(0.3)\) is concentrated around \(\Delta-0.3 = 0.7\) and, similarly the density of the residual life \(R_i(0.6)\) is concentrated around \(\Delta-0.6 = 0.4\).
+	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2}
+\end{align*}
+since, by Chebyshev's inequality, \(\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon)) \geq 1- \cfrac{\var(Z)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
+\[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
+when \(\var(Z)\) is small. 
+That is, when the variance of \(Z\) (equivalently \(Z_i\)) is low, the residual time \(R_i(u)\) will be concentrated around \(\Delta/|c_i| - u\), as required.
+Figure \ref{fig:residual distributions} gives an example of a density function for a \(Z_i\) with mean \(\Delta=1\) and \(c_i=1\), as well as the density function of \(R_i(0.3)\) conditional on \(Z_i>0.3\) and \(R_i(0.6)\) conditional on \(Z_i>0.6\), for comparison.
+Observe that the density of the residual life \(R_i(0.3)\) conditional on \(Z_i>0.3\) is concentrated around \(\Delta-0.3 = 0.7\) and, similarly the density of the residual life \(R_i(0.6)\) conditional in \(Z_i>0.6\) is concentrated around \(\Delta-0.6 = 0.4\).
 
@@
 \includegraphics[width=\textwidth]{chapter2/figs/ME_residual_life_density.pdf}
-\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hhat2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.5}\) (blue).
-Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.5}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively.
+\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hhat2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.6}\) (blue).
+Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.6}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively.
 }
@@
 %At time \(t+u\) the value of the orbit is \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\), which can be equivalently written as \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_ju/|c_i| } }{\bs\alpha e^{\bs S_ju/|c_i| } \bs e}\).
-Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j\) occurs, given \(Z_j>|c_i|u/|c_j|\).
+Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j-|c_i|u/|c_j|\) occurs, on the event \(Z_j>|c_i|u/|c_j|\).
 The initial vector associated with the residual time \(R_j(|c_i|u/|c_j|)\) is \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\).
-This is the initial vector which is obtained if at time \(t\) the phase is \(\varphi(t)=j\) and the orbit is \(\bs A(t)=\alpha,\) so that the QBD-RAP has just entered level \(k\), and there is no change of phase by time \(|c_i|u/|c_j|\).
+This is the initial vector which is obtained if at time \(t\) the phase is \(\varphi(t)=j\) and the orbit is \(\bs A(t)=\bs \alpha,\) so that the QBD-RAP has just entered level \(k\), and there is no change of phase by time \(|c_i|u/|c_j|\).
 Above, we argued that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) can be seen to correspond to the event that \(X(t+u)\) is \(|c_j|\left(|c_i|u/|c_j|\right) = |c_i|u\) units greater than \(y_k\).
@@
 \begin{align}
-	&\mathbb P(R_j(|c_i|u/|c_j|) \in ((\Delta-|c_i|u)/|c_j|-\varepsilon, (\Delta-|c_i|u)/|c_j| + \varepsilon)) 
-	\\&=\mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon)\mid Z_j > |c_i|u/|c_j|) 
-	\\&= \cfrac{\mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon))}{\mathbb P(Z_j > |c_i|u/|c_j|)} 
+	&\mathbb P(R_j(|c_i|u/|c_j|) \in ((\Delta-|c_i|u-\varepsilon)/|c_j|, (\Delta-|c_i|u + \varepsilon)/|c_j| )) \nonumber
+	\\&=\mathbb P(Z_j\in ((\Delta-\varepsilon)/|c_j|, (\Delta + \varepsilon)/|c_j|)) 
+	\\&=\mathbb P(Z\in (\Delta-\varepsilon ,  \Delta + \varepsilon )) 
+	% \\&=  \mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon)) 
 %	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
@@
 
-\paragraph{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\)} Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
+\paragraph{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\)} Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
 We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\).
-Meanwhile, the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units to the left of \(y_{k+1}\).
-We have also reasoned that for \(i\in\calS_-\) the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_{k+1}-x\), \(x\) units less than \(y_{k+1}\), and the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units greater than \(y_k\).
-Notice that in positive phases the orbit position describes the exact position of \(X(t)\) from the left of the interval but only approximates the position of \(X(t)\) from the right of the interval, and vice-versa for negative phases.
-
-To construct the QBD-RAP we need to find a matrix \(\bs D\), say, to take the information about the position of \(X(t)\) contained in the orbit when in positive phases, and map it to information about the position of \(X(t)\) in negative phases.
-Then, when there is a change of phase at time \(t+u\) from \(i\in\calS_+\) to \(j\in\calS_-\), the orbit position will jump from \(\bs A(t+u^-)\) to \(\bs A(t+u)=\bs A(t+u^-)\bs D\).
+Correspondingly, the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units less than \(y_{k+1}\).
+We have also reasoned that for \(i\in\calS_-\) the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_{k+1}-x\), that is \(x\) units less than \(y_{k+1}\), and the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units greater than \(y_k\).
+Notice that in positive phases the orbit position describes the exact position of \(X(t)\) from the left-hand boundary of the interval but only approximates the position of \(X(t)\) from the right-hand boundary of the interval, and vice-versa for negative phases.
+
+To construct the QBD-RAP we need to find a matrix \(\bs D\) such that when there is a change of phase at time \(t+u\) from \(i\in\calS_+\) to \(j\in\calS_-\), the orbit position will jump from \(\bs A(t+u^-)\) to \(\bs A(t+u)=\bs A(t+u^-)\bs D\).
+We want \(\bs A(t+u)\) to represent, in some way, a distance \(|c_i|u\) from \(y_k\) when viewed from a negative phase.
 The matrix \(\bs D\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D, \bs S)\) is a representation of a matrix exponential distribution.
-The matrix \(\bs D\) should also have the property that \(\bs D \bs e = \bs e\) as this will mean that we can model the phase process of the fluid exactly.
-
-To construct such a matrix \(\bs D\).
-If we were given \(R_i(u)=r\), then the appropriate orbit position upon a jump from phase \(i\in\calS_+\) to \(j\in\calS_-\) would be \(\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\) as this corresponds to the fluid level position \(y_{k+1}-|c_i|r\).
+We would also like the matrix \(\bs D\)to the property \(\bs D \bs e = \bs e\) as we can use this to show that phase process of the QBD-RAP and fluid queue have the same probability law (distribution).
+
+The matrix \(\bs D\) is a modelling choice.
+One way to construct such a matrix \(\bs D\) is as follows.
+Suppose we are given \(R_i(u)=r\), then the appropriate orbit position upon a jump from phase \(i\in\calS_+\) to \(j\in\calS_-\) would be \(\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\) as this corresponds to the fluid level position \(y_{k+1}-|c_i|r\).
 However, at time \(t+u\), \(R_i(u)\) is a random variable about the future of the process and therefore not known, so instead, we take the expected initial vector
@@
 
-Note that the matrix \(\bs D\) is a modelling choice and other choices are possible.
-For example, \(\displaystyle \bs D = \int_{x=0}^{\Delta-\varepsilon}e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\).
+Recall that we posed the choice of the matrix \(\bs D\) as a modelling choice.
+Other choices are possible, for example, \(\displaystyle \bs D = \int_{x=0}^{\Delta-\varepsilon}e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\).
 It may also be possible to construct other matrices \(\bs D\), perhaps via geometric arguments.
+
+\paragraph{Upon exiting \(\calD_{k,i}\)} Suppose that upon exiting \(\calD_{k,i}\) at time \(t\) the phase is \(\varphi(t)=i\in\calS_+\).
+At this time \(X(t)=y_{k+1}\) which is the left-hand endpoint of \(\calD_{k+1,i}\).
+Hence, we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
+Similarly, upon exiting \(\calD_{k,i}\) at time \(t\) in phase \(\varphi(t)=i\in\calS_-\), then \(X(t)=y_k\), which is the right-hand endpoint of \(\calD_{k,i}\), and so we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
 
@@
 %where \(\widehat{\bs s}_{ij} = (\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\cfrac{|c_j|}{|c_i|} \bs e\). 
-Therefore, when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), if the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the generator \(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\).
-The distribution of time until the next event in the QBD-RAP after time \(t+u\) is
+Therefore, when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), if the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the generator \(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\), then the distribution of time until the next event in the QBD-RAP after time \(t+u\) is
 \[1-\bs e_k e^{(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I) + T_{jj}\bs I)a}\bs e.\]
@@
 A solution is to suppose that, rather than observing \(C_{t+u}\) at an exponential time with rate \(-T_{ii}\), we instead observe \(\{J(t+u)\}\) uniformly randomly on the lifetime of length \(Z_i\).
-In this case, the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) is (Hautphenne \emph{et al.}~\cite[Lemma 3.1]{hmp2017})
+In this case, the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) is \cite[Lemma 3.1]{hmp2017}
 \begin{align}\mathbb P(C_{t+u}\leq a) = 1-\cfrac{\bs \alpha e^{\bs S_iu}(-\bs S_i)^{-1}\bs e_k}{\bs \alpha (-\bs S_i)^{-1}\bs e_k} = 1- \bs e_k e^{\widetilde{ \bs S}_iu}\bs e.\label{eqn:ksda}\end{align}
@@
 
-With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2014} with discretisation parameter \(\Delta/p\).
-
-\paragraph{Upon exiting \(\calD_{k,i}\)} Suppose that upon exiting \(\calD_{k,i}\) at time \(t\) the phase is \(\varphi(t)=i\in\calS_+\).
-At this time \(X(t)=y_{k+1}\) which is the left-hand endpoint of \(\calD_{k+1,i}\).
-Hence, we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
-Similarly, upon exiting \(\calD_{k,i}\) at time \(t\) in phase \(\varphi(t)=i\in\calS_-\), then \(X(t)=y_k\), which is the right-hand endpoint of \(\calD_{k,i}\), and so we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
+With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2013} with discretisation parameter \(\Delta/p\).
 
@@
 The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hhat2020}.
-Hence, if we construct a QBD-RAP approximation of with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\).
+Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\).
 However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.
@@
 Let \(\bs f(t) = e^{\bs S t}\bs s\).
-Then \(\bs f(t)e^{-\lambda t}\) where \(\lambda = \Re(\lambda_i)\) is the real part of the eigenvalues of \(\bs{S}\) (the all share the same real part), is also periodic with the same period.
+Then \(\bs f(t)e^{-\lambda t}\) where \(\lambda = \Re(\lambda_i)\) is the real part of the eigenvalues of \(\bs{S}\) (they all share the same real part), is also periodic with the same period.
 Hence we can simplify the integral to a finite one;
@@
 and \(X(t)\in\mathcal D_\ell\), \(t\in[0,t_3)\). 
-On approximating this event, the initial orbit position is \(\bs A(0)=\bs \alpha\), and the condition \(X(t)\in\calD_\ell,\, t\in [0,t_3)\) is modelled by \(L(t)=\ell,\, t\in[0,t_3)\).
-The corresponding sample path of the orbit process on \(t\in[0,t_3)\), is
+On approximating this event, the initial orbit position is \(\bs A(0)=\bs \alpha\). The corresponding sample path of the orbit process on \(t\in[0,t_3)\), is 
 \begin{align*}
@@
 for some matrices \(\bs{D}({j_0},{j_1})\) and \(\bs{D}({j_2},{j_1})\).
-Notice that \(\left\{\bs A(t)\right\}\) is constant on \(t\in\left[t_1,t_2\right)\), and ultimately, \(\left\{\varphi(t)\right\}\) transitions from a positive phase to a negative phase.
-The matrix product \(\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})\) must capture that there is ultimately a change of phase from \(\calS_+\) to \(\calS_-\).
+Notice that \(\left\{\bs A(t)\right\}\) is constant on \(t\in\left[t_1,t_2\right)\).
+The matrix product \(\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})\) is there to capture the change in direction due to the change from \(\calS_+\) to \(\calS_-\).
 Hence \(\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})\) should be equal to \(\bs{D}\).
@@
 We now have all the elements we need to describe the dynamics of the QBD-RAP approximation.
-Recall that a QBD-RAP is a stochastic process \(\{(L(t),\bs A(t))\}_{t\geq0}\) where \(L(t)\) is a discrete, level process and \(\bs A(t)\) is a piecewise deterministic orbit process.
-Since the phase dynamics are a CTMC, we choose to use an alternate notation, \(\{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process.
-We choose this representation to make explicit how the approximation captures the phase dynamics of the fluid queue.
+Since the phase dynamics are a CTMC, we choose to use an alternate notation to the standard presented in Section \ref{sec: qbd-rap}.
+Let \(\{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) be the QBD-RAP approximation of a fluid queue, where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process.
+We use this representation to make explicit how the approximation captures the phase dynamics of the fluid queue.
 We will show later that \(\phi(t)\) captures the phase dynamics of the fluid queue exactly, provided the phase process is independent of the fluid level \(\{X(t)\}\).
@@
 Let \(\bs a_0\) be a vector such that \(\bs a_0\bs e = 1\) and \((\bs a_0,\bs{S},\bs s)\) is a representation of a ME distribution.
-Given no events occur before time \(t+u\), \(\bs A(u) = \bs a_0\) and \({\phi}(u)=i\), the solution to (\ref{eqn: At DE}) states that \(\bs A({t+u})\) evolves deterministically according to
-\[\bs A({t+u})\mid \left\{\bs A(u)=\bs a_0\right\} =  \cfrac{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}}{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t} \bs e} = \cfrac{\bs a_0 e^{\bs{S}_i t}}{\bs a_0 e^{\bs{S}_i t} \bs e}.\]
-
-At time \(t\), given \(\bs A(t)\), an event occurs at rate 
+Given no events occur before time \(t+u\), \(\bs A(u) = \bs a_0\in\mathcal A\) and \({\phi}(u)=i\), the solution to (\ref{eqn: At DE}) states that \(\bs A({t+u})\) evolves deterministically according to
+\[\bs A({t+u}) =  \cfrac{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}}{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t} \bs e} = \cfrac{\bs a_0 e^{\bs{S}_i t}}{\bs a_0 e^{\bs{S}_i t} \bs e}.\]
+
+At time \(t\) an event occurs at rate 
 \[\bs A(t) (\bs{S}_i - T_{ii}\bs{I}_p)\bs e = \bs A(t)\bs s_i - T_{ii}.\]
@@
 
-Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)=sign(c_j)\), at the time of the event \(\bs A(t)\) is unchanged but begins to evolves according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), while \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
-
-Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)\neq sign(c_j)\), the process \(\{\bs A(t)\}\) jumps to \(\bs A(t) = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{\bs A({t^-}) T_{ij} \bs{D} \bs e} = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{T_{ij} } = \bs A({t^-}) \bs{D} \) and then proceeds to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), and \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
+Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)=sign(c_j)\), at the time of the event \(\bs A(t)\) is unchanged but immediately begins to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), while \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
+
+Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)\neq sign(c_j)\), the process \(\{\bs A(t)\}\) jumps to \(\bs A(t) = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{\bs A({t^-}) T_{ij} \bs{D} \bs e} = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{T_{ij} } = \bs A({t^-}) \bs{D} \) and then immediately proceeds to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), and \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
 
@@
 	Let \(\{\tau_n\}_{n\geq 0}\) with \(\tau_0=0\) and \(\tau_n\) the time of the \(n\)-th change of level of the QBD-RAP.
-Consider partitioning \(\left\{\Theta_i\leq t \right\}\) with respect to \(\left\{\tau_{n-1}< t \leq\tau_n\right\},\,n=1,2,\dots\).
+Consider partitioning \(\left\{\Theta_i> t \right\}\) with respect to \(\left\{\tau_{n-1}< t \leq\tau_n\right\},\,n=1,2,\dots\).
 For \(n=1\) we can write
 	\begin{align*}
-		&\mathbb P(\Theta_i\leq  t, \tau_0< t\leq \tau_1 \mid \bs A(0) = \bs a) 
+		&\mathbb P(\Theta_i>  t, \tau_0< t\leq \tau_1 \mid \bs A(0) = \bs a) 
 		= \bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}\bs e 
@@
 	\begin{align*}
-		&\mathbb P(\Theta_i\leq t, \tau_{n-1}< t \leq \tau_n \mid \bs A(0) = \bs a) 
+		&\mathbb P(\Theta_i> t, \tau_{n-1}< t \leq \tau_n \mid \bs A(0) = \bs a) 
+		\\&= \int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t \mathbb P(\Theta_i> t, t \leq \tau_n, \tau_{n-1}\in \wrt t_{n-1},\dots \tau_2\in\wrt t_2, \tau_1\in\wrt t_1 \mid \bs A(0) = \bs a)
 		\\&=\displaystyle{\int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t} \bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t_1}\bs s_i
 			\left(\prod_{k=2}^{n-1} \bs \alpha e^{(\bs{S}_i+T_{ii}\bs{I}_p)(t_k-t_{k-1})} \bs s_i\right) 
-			\\&\qquad\displaystyle{\times\bs \alpha e^{(\bs{S}_i+T_{ii}\bs{I}_p)(t-t_{n-1})}\bs e}\wrt t_{n-1}\wrt t_{n-2}\hdots\wrt t_{1}
-		\intertext{where \(t_k\) is the time of the \(k\) change of level.
-Since \(T_{ii}\bs{I}_p\) commutes with \(\bs{S}_i\), \(e^{T_{ii}t_k},\,k=1,...,n-1\) are scalars, and \(t_1+(t_2-t_1)+\hdots+(t_{n-1}-t_{n-2})+(t-t_{n-1})=t\), then this is equal to}
+			\\&\qquad\displaystyle{\times\bs \alpha e^{(\bs{S}_i+T_{ii}\bs{I}_p)(t-t_{n-1})}\bs e}\wrt t_{n-1}\wrt t_{n-2}\hdots\wrt t_{1}.
+		\intertext{Since \(T_{ii}\bs{I}_p\) commutes with \(\bs{S}_i\), \(e^{T_{ii}t_k},\,k=1,...,n-1\) are scalars, and \(t_1+(t_2-t_1)+\hdots+(t_{n-1}-t_{n-2})+(t-t_{n-1})=t\), then this is equal to}
 		& \displaystyle{\int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t} \bs a e^{\bs{S}_it_1}\bs s_i 
@@
 	\begin{align*}
-		\mathbb P(\Theta_i\leq t) &= \sum_{n=1}^\infty \mathbb P(\Theta_i\leq t, \tau_{n-1} < t \leq \tau_n)
+		\mathbb P(\Theta_i> t) &= \sum_{n=1}^\infty \mathbb P(\Theta_i> t, \tau_{n-1} < t \leq \tau_n)
 		\\&= \sum_{n=1}^\infty \mathbb P(\tau_{n-1} < t \leq \tau_n)\displaystyle{  e^{T_{ii}t} }
@@
 Without loss of generality, assume that there is a boundary for the fluid level at \(y_0 = 0\).
-This boundary can only be hit in phases \(i\in\mathcal S_-\).
+This boundary can only be hit from above in phases \(i\in\mathcal S_-\).
 %For a regulated boundary, upon hitting the boundary in phase \(i\in\mathcal S_-\) the fluid level remains at \(0\), while the phase continues to evolve among states in the set \(\calS_-\cup\calS_0\).
@@
 Adjacent to the boundary at \(y_0=0\) is the band \(\mathcal D_0=[0,\Delta]\) which corresponds to level \(0\) for the QBD-RAP.
+We denote the boundary of the QBD-RAP as level \(-1\), which corresponds to \(\{0\}\) for the fluid queue.
 Modelling the behaviour of the fluid queue at boundaries can be broken down into three components.
@@
 We claim that the event \(\{L(t)=\ell,\phi(t)=i\}\) models the event \(\{X(t)\in\mathcal D_{\ell,i},\varphi(t)=i\}\), and further that instants \(u\) with \(L(u^-)\neq L(u), \phi(u)=i\) model the events \(\{X(u^-)\in \mathcal D_{\ell,i}, X(u)\notin\mathcal D_{\ell,i},\varphi(u)=i\}\).
-With this in mind, we suppose that when the QBD-RAP is in level \(0\) in phase \(i\in\mathcal S_-\) and there is a change of level, this corresponds approximating the event that the fluid level \(\{X(t)\}\) hits the boundary at \(0\).
+With this in mind, we suppose that when the QBD-RAP is in level \(0\) in phase \(i\in\mathcal S_-\) and there is a change of level, this corresponds to approximating the event that the fluid level \(\{X(t)\}\) hits the boundary at \(0\).
 We refer to this as the QBD-RAP hitting the boundary also.
-We show later that, using matrix exponentials with sufficiently small variance in the construction of the QBD-RAP, then the distribution of time until the QBD-RAP first hits the boundary closely models the time until the fluid level hits the boundary closely.
+We show later that, using matrix exponentials with sufficiently small variance in the construction of the QBD-RAP, then the distribution of time until the QBD-RAP first hits the boundary closely models the time until the fluid level hits the boundary.
 This addresses component 1).
@@
 
-For regulated behaviour at a lower boundary at \(0\), given the fluid level reaches zero at time \(u\) in phase \(i\in\mathcal S_-\), then the phase transitions to some phase \(j\in\calS_-\cup\calS_0\) with probability \(p_{ij}\), and the distribution of the phase \(\varphi(t)\) until the fluid queue leaves the boundary for the first time after \(u\) is given by the elements of the vector
+For regulated behaviour at a lower boundary at \(0\), given the fluid level reaches zero at time \(u\) in phase \(i\in\mathcal S_-\), then the phase transitions to some phase \(j\in\calS_-\cup\calS_0\) with probability \(p_{ij}\), and the distribution of the phase \(\varphi(t)\) (on the event that the process has not left the boundary at or before time \(t\)) is given by the elements of the vector
 \[p_{ij} \bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right).\] 
@@
 
-To summarise, at time \(t\) the rate at which probability mass accumulates at the regulated boundary in phase \(j \in\calS_-\cup\calS_0\) upon hitting the boundary in phase \(i\in\calS_-\) is 
+To summarise, at time \(t\) the rate at which probability mass accumulates at the regulated boundary in phase \(j \in\calS_-\cup\calS_0\) and level \(-1\), upon hitting the boundary in phase \(i\in\calS_-\) and level \(0\), is
 \[ c_ip_{ij} \bs A(t)\bs s.\]
-The rate at which mass transitions from phase \(i\in\calS_-\) to \(j\in\calS_+\) upon hitting a boundary is 
+The rate at which density from phase \(i\in\calS_-\) and level \(0\) transitions to density in phase \(j\in\calS_+\) and level \(0\) upon hitting a boundary is 
 \[ c_ip_{ij} \bs A(t)\bs s,\]
 and upon this transition, the orbit jumps to \(\bs \alpha\). 
-The rate at which mass leaves the regulated boundary from phase \(i\in\calS_-\) into phases \(j\in\calS_+\) is 
+The rate at which mass leaves the regulated boundary from phase \(i\in\calS_-\cup\calS_0\) and level \(-1\) into phases \(j\in\calS_+\) and level \(0\) is 
 \[T_{ij},\]
@@
 The top-right block is the transition of point masses into density as the process leaves the boundary.
-The term \((\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha\) in \(\check{\bs B}_{0}\) incorporates the reflecting boundary.
+The term \((\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha\) in \(\check{\bs B}_{0}\) incorporates the reflecting boundary behaviour.
 
@@
 We argued earlier that we can think of the orbit \(\cfrac{\bs \alpha e^{\bs{S}|c_i|t}}{\bs \alpha e^{\bs{S}|c_i|t}\bs e}\) as corresponding to the fluid being a distance of \(|c_i|t\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\).
-With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is the set the initial orbit to \(\cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\).
+With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\).
 Similarly, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\cfrac{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}}{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}\bs e}.\)
@@
 \section{At time \(t\) -- closing operators}\label{sec: closing}
-We now describe how the orbit position \(\bs A(t)\) can be used to approximate the density of the position of \(X(t)\) in a given level.
+So far we have described how we can use a QBD-RAP to approximate a fluid queue. The level process \(\{L(t)\}\) of the QBD-RAP approximates the process 
+\begin{align}
+	\displaystyle \left\{\sum_{k\in\mathcal K}\sum_{i\in\calS} k1(X(t)\in\calD_{k,i})\right\}_{t\geq 0}. \label{eqn: kfafj}
+\end{align}
+This may be of interest in its own right. However, it is a somewhat granular approximation -- the level process tells us nothing about where in the intervals \(\calD_{k,i}\) the fluid queue might be.
+
+We now describe how the orbit position \(\bs A(t)\) can be used to approximate the density of the level of the fluid queue \(X(t)\).
 
 Suppose that the QBD-RAP is in level \(L(t)=\ell\), phase \(\phi(t)=j\in\calS_+\), and the orbit is \(\bs A(t)=\bs a\).
-If the QBD-RAP remains in phase \(j\), then QBD-RAP approximation will transition out of level \(\ell\) in the infinitesimal time interval \(t+\wrt u\) with density \(\bs a e^{|c_j|\bs{S}u}|c_j|\bs s\wrt u\).
+If the QBD-RAP remains in phase \(j\), the QBD-RAP approximation will transition out of level \(\ell\) in the infinitesimal time interval \(t+\wrt u\) with density \(\bs a e^{|c_j|\bs{S}u}|c_j|\bs s\wrt u\).
 At the time of the change of level we estimate the position of \(X(t+u)\) by \(X(t+u)\approx y_{\ell+1}\).
@@
 \begin{rem}
-	I considered taking the closing operators to be 
+	We considered taking the closing operators to be 
 	\begin{align}

### 2021-10-15 chapter2/chapter.tex
update hhat reference, add me references

@@
 Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic.
-We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance).
+We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \cite{ert2006,hstz2016,ert2006,hht2020,mt2021}.
 As in \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP.
@@
 \includegraphics[width=\textwidth]{chapter2/figs/ME_residual_life_density.pdf}
-\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hhat2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.6}\) (blue).
+\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hht2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.6}\) (blue).
 Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.6}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively.
@@
 \subsection{Computing \(\bs D\)}
-In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hhat2020}.
+In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hht2020}.
 For this class of CMEs, we take the index \(p\) to be the order of the representation.
 Moreover, we take \(p\) to be odd.
-The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hhat2020}.
+The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}.
 Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\).
@@
 
-Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \citep{hhat2020}.
+Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \citep{hht2020}.
 

### 2021-10-17 chapter2/chapter.tex
fix more of Nige's comments -- up to page 25. reorder PH-case section and rewrite. add requirements on D, most importanly linearity

@@
 We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \cite{ert2006,hstz2016,ert2006,hht2020,mt2021}.
-As in \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP.
+We construct a level process for the QBD-RAP and correspond it to a discretisation the level process of the fluid queue.
+This is similar to \cite{bo2013} where the level process of their QBD corresponds a discretisation of the fluid level.
 Unlike \cite{bo2013}, however, here we do not take \(\Delta \to 0\).
 Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small.
-This necessarily means that the order of the matrix exponential increases.
+% This necessarily means that the order of the matrix exponential increases.
 
@@
 
-\paragraph{Modelling the residual time to exit on no change of phase} 
+\subsection{Modelling the residual time to exit on no change of phase} 
 We first consider no change of phase in \([t,t+u]\).
@@
 The random variable \(Z_i\) has mean \(\Delta/|c_i|\).
-Further, let \(R_i(u)\) be the residual time, \(R_i(u) = (Z_i-u)1\{Z_i-u>0\}\), \(i\in\calS\).
-Suppose first that \(X(t)=y_k\) and \(\varphi(t)=i\in\mathcal S_+\), so that \(X(t)\) will leave \(\mathcal D_{k,i}\) in exactly \(\Delta/|c_i|\) units of time on the event that the phase process remains in \(i\) for at least this amount of time.
-
+
+Suppose first that \(X(t)=y_k\) and \(\varphi(t)=i\in\mathcal S_+\), so that \(X(t)\) will leave \(\mathcal D_{k,i}\) in exactly \(\Delta/|c_i|\) units of time on the event that the phase process of the QBD-RAP remains in \(i\) for at least this amount of time.
+A sensible approximation for this deterministic event is to choose the value of the orbit process at time \(t\) to be \(\bs \alpha\), \(\bs A(t)=\bs\alpha\).
+With this choice, the distribution of time until the next event of the QBD-RAP, on the event that the phase process remains constant will be the distribution of \(Z_i\).
+Observe,
+\begin{align*}
+	\mathbb P(Z_i \in ((\Delta-\varepsilon)/|c_i|, (\Delta+\varepsilon)/|c_i|) ) 
+	&= \mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))
+%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
+%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
+	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2},
+\end{align*}
+since, by Chebyshev's inequality, \(\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon)) \geq 1- \cfrac{\var(Z)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
+\[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
+when \(\var(Z)\) is small.
+Hence, when the variance of \(Z\) (equivalently \(Z_i\)) is low, then with high probability the time until the next event of the QBD-RAP, given the phase process remains constant will occur at approximately \(\Delta/|c_i|\).
+
+Now, let \(R_i(u)\) be the residual time, \(R_i(u) = (Z_i-u)1\{Z_i-u>0\}\), \(i\in\calS\).
 After \(u\in[0,\Delta/|c_i|)\) amount of time has elapsed, on the event that \(\varphi(t+s)=i\), for all \(s\in[0,u)\), then \(X(t+u) = y_k + c_i u\).
@@
 %	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
-	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2}
-\end{align*}
-since, by Chebyshev's inequality, \(\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon)) \geq 1- \cfrac{\var(Z)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
+	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2}.
+\end{align*}
+Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
 \[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
@@
 
-We can interpret the position of the orbit \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\).
+We interpret the position of the orbit \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\).
 This gives a heuristic argument as to how we can model the sojourn times in a given interval \(\mathcal D_{k,i}\) on the event that the phase does not change.
@@
 We can apply analogous arguments to heuristically develop a model for the sojourn time of the fluid queue in an interval, \(\calD_{k,i}\), \(i\in\calS_-\), on the event that the phase does not change, given the fluid is \(X(t)=y_{k+1}\).
-In this case \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).
-
-\paragraph{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\)} 
+For \(i\in\calS_-\), the orbit position \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).
+
+\subsection{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\)} 
 %Let \(Z_j\sim ME(\bs\alpha,\bs S_j)\) be a matrix exponential random variable with mean \(\Delta/|c_j|\) and let \(R_j(u)\) be the residual time \(R_j(u) = Z_j-u \mid\{Z_j-u>0\}\).
 %At time \(t+u\) the value of the orbit is \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\), which can be equivalently written as \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_ju/|c_i| } }{\bs\alpha e^{\bs S_ju/|c_i| } \bs e}\).
-Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j-|c_i|u/|c_j|\) occurs, on the event \(Z_j>|c_i|u/|c_j|\).
+
+Let \(E\) be the event that, at time \(t\) the orbit of the QBD-RAP is \(\bs A(t) = \bs \alpha\), the phase process is \(\varphi(t) = i\in\calS_+\), and there are no events of the QBD-RAP before time \(t+u\).
+On the event \(E\) the orbit position at time \(t+u\) will be
+\begin{align}
+	\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}.\label{eqn: orbit i}
+\end{align}
+Correspondingly, on the event that at time \(t\) the level process of the fluid queue is \(X(t)=y_{k}\), the phase is \(\varphi(t)=i\) and there are no change of phase by time \(t+u\), then \(X(t+u)=y_{k}+c_iu\).
+That is, \(X(t+u)\) is \(c_iu\) units from the left-hand edge of \(\calD_k\).
+
+If, at time \(t+u\) there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\), then we need to map the orbit position \(\bs A(t+u^-) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}\) to an orbit position which corresponds to the being \(c_iu\) units from the left-hand edge of \(\calD_k\) in phase \(j\).
+Call this mapping \(\bs D(i,j)(\cdot)\).
+As noted in Section \ref{sec: qbd-rap}, the mapping must be linear,
+\(\bs D(i,j)(\bs a) = \bs a \bs D(i,j),\, \bs a \in\mathcal A,\)
+for some matrix \(\bs D(i,j)\). The matrix \(\bs D(i,j)\) is a modelling choice.
+
+So that the process is a QBD-RAP, the matrix \(\bs D(i,j)\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D(i,j), \bs S)\) is a representation of a matrix exponential distribution.
+We would also like the matrix \(\bs D(i,j)\) to have the property \(\bs D(i,j) \bs e = \bs e\) as this will mean that the rate at which a change of phase from \(i\) to \(j\) of the QBD-RAP occurs proportional to
+\[\bs A(t+u) \bs D(i,j)\bs e = 1,\]
+which is constant and therefore the distribution of time until a change from phase \(i\) to \(j\) is exponential.
+We can use this to show, for certain models, the distribution of the phase process of the fluid queue and the distribution of the phase process of the QBD-RAP share the same distribution.
+
+An orbit position which corresponds to the level process of the QBD-RAP beign \(c_iu\) units to the right of \(y_k\) in phase \(j\in\calS_+\) is \(\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e}\).
+To see this, consider the event that at time \(t\) the orbit process of the QBD-RAP is \(\bs A(t) = \bs \alpha\), the phase process is \(\varphi(t) = j\in\calS_+\), and there are no events of the QBD-RAP before time \(t+|c_i|u/|c_j|\).
+On this event, the orbit position at time \(t+u\) will be \(\displaystyle \bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e}\).
+Correspondingly, on the event that at time \(t\) the level process of the fluid queue is \(X(t)=y_{k}\), the phase is \(\varphi(t)=j\) and there are no change of phase by time \(t+|c_i|u/|c_j|\), then \(X(t+u)=y_{k}+c_j|c_i|u/|c_j| = y_k+c_iu\).
+That is, \(X(t+u)\) is \(c_iu\) units from the left-hand edge of \(\calD_k\).
+
+Now, observe 
+\[\cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e} = \cfrac{\bs \alpha e^{\bs S_iu}}{\bs \alpha e^{\bs S_i u}\bs e},\] 
+which is exactly (\ref{eqn: orbit i}), the orbit position on the event \(E\). Hence, it is sensible to choose \(\bs D(i,j)=\bs I\).
+
+% Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j-|c_i|u/|c_j|\) occurs, on the event \(Z_j>|c_i|u/|c_j|\).
 The initial vector associated with the residual time \(R_j(|c_i|u/|c_j|)\) is \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\).
@@
 Above, we argued that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) can be seen to correspond to the event that \(X(t+u)\) is \(|c_j|\left(|c_i|u/|c_j|\right) = |c_i|u\) units greater than \(y_k\).
-%We argued earlier that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the position of the orbit on the event that at time \(t\) there is a change of level so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=j\), followed by no change of phase by time \(t+|c_i|u/|c_j|\).
+% %We argued earlier that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the position of the orbit on the event that at time \(t\) there is a change of level so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=j\), followed by no change of phase by time \(t+|c_i|u/|c_j|\).
 The analogous event of the fluid queue is that at time \(t\) the phase is \(\varphi(t)=j\), the fluid level is \(X(t)=y_k\), so that the fluid level has just entered an interval \(\calD_k\) at time \(t\), and there is no change of phase by time \(t+|c_i|u/|c_j|\), in which case the fluid level will be \(X(t+|c_i|u/|c_j|)=y_\ell + |c_j|\left(|c_i|u/|c_j|\right)=y_\ell + |c_i|u\).
@@
 
-Notice that we can re-write the vector 
-\(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) as \(\cfrac{\bs \alpha e^{\bs S_iu } }{\bs \alpha e^{\bs S_iu} \bs e }\).
+% Notice that we can re-write the vector 
+% \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) as \(\cfrac{\bs \alpha e^{\bs S_iu } }{\bs \alpha e^{\bs S_iu} \bs e }\).
 This is the orbit position on the event that at time \(t\) there is a change of level, so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=i\), followed by no change of phase by time \(t+u\).
@@
 
-Therefore, at time \(t+u\), if there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\) then the orbit position should not change as in both cases the orbit position \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) corresponds to an approximation to \(X(t)\) being \(|c_i|u\) units greater than \(y_k\).
+% Therefore, at time \(t+u\), if there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\) then the orbit position should not change as in both cases the orbit position \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) corresponds to an approximation to \(X(t)\) being \(|c_i|u\) units greater than \(y_k\).
 
@@
 
-Further to this argument, observe that the residual time \(R_j(|c_i|u/|c_j|)\) has density 
+Moreover, the residual time \(R_j(|c_i|u/|c_j|)\) has density 
 \[f_{R_j(|c_i|u/|c_j|)}(r) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}e^{\bs S_j r}\bs s_j,\]
@@
 \end{align}
-Hence, when the variance of \(Z\) is low, the residual time, \(R_j(|c_i|u/|c_j|)\), is concentrated around \((\Delta - |c_i|u)/|c_j|\) as required.
-
-Therefore, in our model, upon the event of a change from \(i\in\calS_+\) to \(j\in\calS_+\) at time \(t+u\) the orbit should be \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\).
+Hence, when the variance of \(Z\) is low, the residual time, \(R_j(|c_i|u/|c_j|)\), is concentrated around \((\Delta - |c_i|u)/|c_j|\), as required.
+
+% Therefore, in our model, upon the event of a change from \(i\in\calS_+\) to \(j\in\calS_+\) at time \(t+u\) the orbit should be \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\).
 That is, the initial condition for the next segment of the orbit process is just the current orbit position and the orbit should start to evolve according to the matrix \(\bs S_j\).
@@
 
-\paragraph{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\)} Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
+\subsection{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\)} 
+Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
 We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\).
-Correspondingly, the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units less than \(y_{k+1}\).
-We have also reasoned that for \(i\in\calS_-\) the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_{k+1}-x\), that is \(x\) units less than \(y_{k+1}\), and the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units greater than \(y_k\).
-Notice that in positive phases the orbit position describes the exact position of \(X(t)\) from the left-hand boundary of the interval but only approximates the position of \(X(t)\) from the right-hand boundary of the interval, and vice-versa for negative phases.
-
-To construct the QBD-RAP we need to find a matrix \(\bs D\) such that when there is a change of phase at time \(t+u\) from \(i\in\calS_+\) to \(j\in\calS_-\), the orbit position will jump from \(\bs A(t+u^-)\) to \(\bs A(t+u)=\bs A(t+u^-)\bs D\).
-We want \(\bs A(t+u)\) to represent, in some way, a distance \(|c_i|u\) from \(y_k\) when viewed from a negative phase.
-The matrix \(\bs D\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D, \bs S)\) is a representation of a matrix exponential distribution.
-We would also like the matrix \(\bs D\)to the property \(\bs D \bs e = \bs e\) as we can use this to show that phase process of the QBD-RAP and fluid queue have the same probability law (distribution).
-
-The matrix \(\bs D\) is a modelling choice.
-One way to construct such a matrix \(\bs D\) is as follows.
-Suppose we are given \(R_i(u)=r\), then the appropriate orbit position upon a jump from phase \(i\in\calS_+\) to \(j\in\calS_-\) would be \(\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\) as this corresponds to the fluid level position \(y_{k+1}-|c_i|r\).
-However, at time \(t+u\), \(R_i(u)\) is a random variable about the future of the process and therefore not known, so instead, we take the expected initial vector
-\[\mathbb E \left[ \cfrac{\bs\alpha e^{ \bs S_iR_i(u) } }{\bs\alpha e^{\bs S_iR_i(u) } \bs e} \right] = \int_{r=0}^\infty \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} e^{\bs S_ir } \bs s_i \cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{r=0}^\infty e^{\bs S_ir } \bs s_i\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r.\]
-After a change of variables \(x=|c_i|r\) and defining \(\bs D = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\), we get 
-\[\cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{x=0}^\infty e^{\bs S x }|c_i| \bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x/|c_i| = \bs A(t+u)\int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x = \bs A(t+u) \bs D,\]
-since at time \(t+u\), the orbit position is \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e}\).
-This suggests that upon a jump from \(i\in\calS_+\) to \(j\in\calS_-\), the orbit position jumps according to the matrix \(\bs D\).
-
-Observe that \(\bs D\bs e = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\bs e =\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \wrt x=\bs e.\) This is an important observation as it implies that events corresponding to jumps from \(i\in\calS_+\) to \(j\in\calS_-\) occur with intensity \(\bs A(t)T_{ij}\bs D\bs e = T_{ij}\) which is constant with respect to orbit position and time and this implies they are exponential.
-Further, since \(\mathcal A\) is closed and convex \citep{MEinAP}, then \((\bs a \bs D,\bs S,\bs s)\) is a representation of a matrix exponential distribution for any \(\bs a\in\mathcal A\).
-
-Recall that we posed the choice of the matrix \(\bs D\) as a modelling choice.
-Other choices are possible, for example, \(\displaystyle \bs D = \int_{x=0}^{\Delta-\varepsilon}e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\).
-It may also be possible to construct other matrices \(\bs D\), perhaps via geometric arguments.
-
-\paragraph{Upon exiting \(\calD_{k,i}\)} Suppose that upon exiting \(\calD_{k,i}\) at time \(t\) the phase is \(\varphi(t)=i\in\calS_+\).
-At this time \(X(t)=y_{k+1}\) which is the left-hand endpoint of \(\calD_{k+1,i}\).
-Hence, we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
-Similarly, upon exiting \(\calD_{k,i}\) at time \(t\) in phase \(\varphi(t)=i\in\calS_-\), then \(X(t)=y_k\), which is the right-hand endpoint of \(\calD_{k,i}\), and so we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
+
+As before, we need to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\), to
+\[\bs A(t+u) = \cfrac{\bs A(t+u^-)\bs D(i,j)}{\bs A(t+u^-)\bs D(i,j)\bs e}=\cfrac{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)}{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
+in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\) in phase \(j\).
+
+Again, the matrix \(\bs D(i,j)\) is a modelling choice. We first discuss how we might choose the matrix \(\bs D(i,j)\) for when the matrix-exponential \(Z\) is a phase-type distribution.
 
 \paragraph{The Phase-type case}
-If \(Z\sim ME(\bs \alpha, \bs S)\) is chosen to be a Phase-type distribution then there is another possibility for the matrix \(\bs D\).
-A Phase-type distribution is the distribution of time until absorption of a finite state absorbing Markov chain with transient states \(\{1,\dots,p\}.\) The sub-generator matrix describing the dynamics of the Markov chain on transient states is \(\bs S\), and \(\bs \alpha\) is an initial probability distribution over the transient states.
-Let \(\{J(t)\}_{t\in (-\infty,\infty)}\) be the Markov chain to which the Phase-type distribution corresponds.
-
+If \(Z\sim ME(\bs \alpha, \bs S)\) is chosen to be a phase-type distribution then \(Z\) has the interpretation as the distribution of time until absorption of a finite-state continuous-time Markov chain with transient states \(\{1,\dots,p\}\) and a single absrobing state.
+The sub-generator matrix describing the dynamics of the Markov chain on transient states is \(\bs S\), and \(\bs \alpha\) is an initial probability distribution over the transient states.
+Let \(\{J(t)\}\) be the Markov chain associated with the phase-type distribution.
+
+In the discussions above, we have relied on the relationship between the event \(E\) and the orbit position \(\displaystyle \bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\).
+The relationship allows us to associate this orbit position with the level of the fluid queue.
+For phase-type distributions the vector \(\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) is the vector of posterior probabilities \(\left[\mathbb P(J(u)=k \mid E)\right]_{k\in\{1,...,p\}}\).
+We can use this vector of posterior probabilities in the same way as we do for matrix-exponential distributions and QBD-RAPs.
+However, the orbit interpretation forgoes the presence of the phase process \(\{J(t)\}\).
+Here, we will use the phase process \(\{J(t)\}\) to derive a choice of \(\bs D(i,j)\).
+
+Recall the position of the orbit process can be used to determine the residual time.
+This idea can be replicated when we are given the phase of the phase-type rather than the value of the orbit process.
+Given the phase of the phase-type distribution is \(k\in\{1,...,p\}\), the distribution of the residual time is
+\[\mathbb P(R_i\leq r \mid \mbox{phase}=k)=1-\bs e_k e^{\bs S_i r}\bs e.\]
+Notice that this is independent of the time since the phase-type distribution was initialised.
+We call this time since the phase-type distribution was initialised the \emph{age}.
+As noted by \cite{hmp2017}, the distribution of the age given the phase of the phase-type is \(k\in\{1,...,p\}\) depends on the sampling scheme which determines the time of the observation.
+Here, on the event that at time \(t\) the phase of the QBD-RAP is \(i\), a change of phase occurs after an exponential amount of time with rate \(-T_{ii}\).
+Proposition 4.1, \cite{hmp2017} states that the distribution function of the age, given the phase of the phase-type is \(k\) and the process is observed after an exponential time with rate \(-T_{ii}\) is
+\[\mathbb P(\mbox{age}\leq u \mid \mbox{phase}=k )=1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)u}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}.\]
+Let \(\widehat{ \bs S}_i(T_{ii}) = diag(\bs \nu)^{-1}\bs S_i'diag(\bs \nu)\), where \(\bs \nu = \bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}/(\bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e)\).
+Algebraic manipulations show
+\begin{align}
+	1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)u}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}
+	= 1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)u} \bs e,\label{eqn: skgj}
+\end{align}
+which is of Phase-type.
+
+Recall that in the matrix-exponential case we associated the value of the orbit process \(\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) with being \(c_iu\) units to the right of \(y_{k}\).
+In contrast, for the phase-type case if we use the phase process \(\{J(t)\}\) to inform our modelling of the fluid queue (as opposed to the value of the orbit process), then we associate phase \(k\in\{1,...,p\}\) of \(\{J(t)\}\) with being a \emph{random} distance to the right of \(y_k\) where this distance has distribution
+\begin{align}
+	1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)u} \bs e.\label{eqn: skgj2}
+\end{align}
+Therefore, on the event \(E\), upon a change from of phase of the QBD-RAP from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), it seems reasonable to want the distribution of time until the next event of the QBD-RAP to be
+\begin{align}
+	1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)|c_j|u/|c_i| + T_{jj}Iu} \bs e.\label{eqn: skgj3}
+\end{align}
+The factor \(|c_j|u/|c_i|\) arises as a conversion between the speed at which the fluid level moves in phase \(j\) compared to phase \(i\).
+The rate \(T_{jj}\) in the exponent is the rate at which the next change of phase occurs.
+
+While this does achieve what we want, it is not quite satisfactory for the purpose of our approximation scheme due to dependence on the sample path of \(\{\varphi(t)\}\).
+Specifically, the evolution of the QBD-RAP from time \(t+u\) until the next event depends on the phase immediately before the change of phase at time \(t+u\), \(\varphi(t+u^-)=i\).
+This increases the size of the approximating QBD-RAP as we need a separate model for each \(\varphi(t+u^-)\in\calS_+\).
+Furthermore, we have not yet considered how to model any further changes from \(\calS_-\) to \(\calS_+\) or beyond, which further complicates matters.
+
+A solution is to suppose that, rather than observing the phase-type random variable at an exponential time with rate \(-T_{ii}\), we instead observe the process uniformly randomly on the lifetime of length \(Z_i\).
 Let \(\bs \Pi = diag\left(\bs \pi \right)\), where \(\vligne{\pi_k}_{k\in\{1,\dots,p\}}=:\bs \pi=\bs \alpha (-\bs{S})^{-1}/m\) and \(m=\bs \alpha (-\bs{S})^{-1}\bs e\).
 There is a time-reverse representation of a Phase-type distribution given by \((\widetilde{\bs \alpha}, \widetilde{\bs S}, \widetilde{\bs s})\), where \(\widetilde{\bs \alpha} = m\bs s' \bs \Pi \), \(\widetilde{\bs S} = \bs\Pi^{-1}\bs S'\bs \Pi\) and \(\widetilde{\bs s} = \bs \Pi^{-1}\bs \alpha'/m\) \cite[Page 91]{a2008}.
-
-Rather than use the residual time \(R_i(u)\) to estimate how far \(X(t)\) is from the edge of an interval, we can use the information contained in \(J(t+u)\).
+In the case where the phase is observed randomly on the lifetime of \(Z_i\), the distribution function of the age, given the phase is \(k\) is \cite[Lemma 3.1]{hmp2017}
+\begin{align}\mathbb P(\mbox{age}\leq u \mid \mbox{phase}=k) = 1-\cfrac{\bs \alpha e^{\bs S_iu}(-\bs S_i)^{-1}\bs e_k}{\bs \alpha (-\bs S_i)^{-1}\bs e_k} = 1- \bs e_k e^{\widetilde{ \bs S}_iu}\bs e.\label{eqn:ksda2}\end{align}
+With this interpretation, we may associate phase \(k\in\{1,...,p\}\) of \(\{J(t)\}\) with being a \emph{random} distance, with distribution (\ref{eqn:ksda2}), to the right of \(y_k\).
+
+Therefore, on the event \(E\) and on a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), a reasonable model for the time until the next event of the QBD-RAP has distribution 
+\begin{align}
+	1- \bs e_k e^{\widetilde{ \bs S}_iu|c_j|/|c_i| + T_{jj}\bs Iu}\bs e = 1- \bs e_k e^{\widetilde{ \bs S}_ju+ T_{jj}\bs Iu}\bs e.\label{eqn:ksda3}
+\end{align}
+This suggests that, at a jump from \(\calS_+\) to \(\calS_-\), the state of \(\{J(t)\}\) does not change, but begins to evolve according to the time-reverse generator at an appropriate speed.
+Since the time-reverse of \(\widetilde{\bs S}\) is \(\bs S\), then upon a jump back to \(\calS_+\) from \(\calS_-\), the phase of the phase-type random variable remains \(k\) but begins to evolve according to \(\calS\).
+This suggests that we use the representation \((\bs \alpha, \bs S)\) when in phases in \(\calS_+\), and use the time-reverse representation \((\widetilde{\bs \alpha}, \widetilde{\bs S})\) when in phases in \(\calS_-\).
+The matrices \(\bs D(i,j)=\bs I\) for all \(i,j\in\calS\).
+
+With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2013} with discretisation parameter \(\Delta/p\).
+
+% Let \(\bs \Pi = diag\left(\bs \pi \right)\), where \(\vligne{\pi_k}_{k\in\{1,\dots,p\}}=:\bs \pi=\bs \alpha (-\bs{S})^{-1}/m\) and \(m=\bs \alpha (-\bs{S})^{-1}\bs e\).
+There is a time-reverse representation of a Phase-type distribution given by \((\widetilde{\bs \alpha}, \widetilde{\bs S}, \widetilde{\bs s})\), where \(\widetilde{\bs \alpha} = m\bs s' \bs \Pi \), \(\widetilde{\bs S} = \bs\Pi^{-1}\bs S'\bs \Pi\) and \(\widetilde{\bs s} = \bs \Pi^{-1}\bs \alpha'/m\) \cite[Page 91]{a2008}.
+
+% Rather than use the residual time \(R_i(u)\) to estimate how far \(X(t)\) is from the edge of an interval, we can use the information contained in \(J(t+u)\).
 Suppose there is a change of phase of the fluid queue from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\) at which time \(J(t+u)=k\).
@@
 
-According to \cite{hmp2017}, the distribution of \(C_{t+u}\) depends on the method of observation.
+% According to \cite{hmp2017}, the distribution of \(C_{t+u}\) depends on the method of observation.
 Here, the Phase-type distribution is observed after an exponential time with rate \(-T_{ii}\).
 Proposition 4.1, \cite{hmp2017} states that the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) and the process is observed after an exponential time with rate \(-T_{ii}\) is
-\[\mathbb P(C_{t+u}\leq a )=1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}.\]
-Let \(\widehat{ \bs S}_i(T_{ii}) = diag(\bs \nu)^{-1}\bs S_i'diag(\bs \nu)\), where \(\bs \nu = \bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}/(\bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e)\).
+% \[\mathbb P(C_{t+u}\leq a )=1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}.\]
+% Let \(\widehat{ \bs S}_i(T_{ii}) = diag(\bs \nu)^{-1}\bs S_i'diag(\bs \nu)\), where \(\bs \nu = \bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}/(\bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e)\).
 Algebraic manipulations show
-\begin{align}
-	1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}
-	= 1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)a} \bs e,\label{eqn: skgj}
-\end{align}
-which is of Phase-type.
+% \begin{align}
+% 	1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}
+% 	= 1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)a} \bs e,\label{eqn: skgj}
+% \end{align}
+% which is of Phase-type.
 The conversion between the intensity at which events occurs when the fluid queue is in phase \(i\) compared to phase \(j\) is \(|c_j|/|c_i|\).
@@
 %where \(\widehat{\bs s}_{ij} = (\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\cfrac{|c_j|}{|c_i|} \bs e\). 
-Therefore, when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), if the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the generator \(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\), then the distribution of time until the next event in the QBD-RAP after time \(t+u\) is
-\[1-\bs e_k e^{(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I) + T_{jj}\bs I)a}\bs e.\]
-While this does approximate the desired quantity, it is not quite satisfactory for the purpose of our approximation scheme due to dependence on the sample path of \(\{\varphi(t)\}\).
+% Therefore, when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), if the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the generator \(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\), then the distribution of time until the next event in the QBD-RAP after time \(t+u\) is
+% \[1-\bs e_k e^{(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I) + T_{jj}\bs I)a}\bs e.\]
+% While this does approximate the desired quantity, it is not quite satisfactory for the purpose of our approximation scheme due to dependence on the sample path of \(\{\varphi(t)\}\).
 Specifically, the evolution of the QBD-RAP from time \(t+u\) until the next event depends on the phase immediately before the change of phase at time \(t+u\), \(\varphi(t+u^-)=i\).
@@
 
-A solution is to suppose that, rather than observing \(C_{t+u}\) at an exponential time with rate \(-T_{ii}\), we instead observe \(\{J(t+u)\}\) uniformly randomly on the lifetime of length \(Z_i\).
+% A solution is to suppose that, rather than observing \(C_{t+u}\) at an exponential time with rate \(-T_{ii}\), we instead observe \(\{J(t+u)\}\) uniformly randomly on the lifetime of length \(Z_i\).
 In this case, the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) is \cite[Lemma 3.1]{hmp2017}
-\begin{align}\mathbb P(C_{t+u}\leq a) = 1-\cfrac{\bs \alpha e^{\bs S_iu}(-\bs S_i)^{-1}\bs e_k}{\bs \alpha (-\bs S_i)^{-1}\bs e_k} = 1- \bs e_k e^{\widetilde{ \bs S}_iu}\bs e.\label{eqn:ksda}\end{align}
-Once again, noting that the conversion between the intensity at which events occurs when the fluid queue is in phase \(i\) compared to phase \(j\) is \(|c_j|/|c_i|\)%, and taking the derivative with respect to \(v\) of (\ref{eqn:ksda}) by applying the chain rule, gives
-%\begin{align}%
-%	\cfrac{\wrt }{\wrt u} \bs e_k e^{\widetilde{ \bs S}_iu}\bs e|c_j|/|c_i|=\bs e_k e^{\widetilde{ \bs S}_iv|c_j|/|c_i|}\widetilde{ \bs S}_i\bs e|c_j|/|c_i| = \bs e_k e^{\widetilde{ \bs S}_jv}\widetilde{ \bs s}_j.\label{eqn:ksdadsva}
-%\end{align}
-then, upon a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the reverse-time generator \(\widetilde{ \bs S}_j\).
+% \begin{align}\mathbb P(C_{t+u}\leq a) = 1-\cfrac{\bs \alpha e^{\bs S_iu}(-\bs S_i)^{-1}\bs e_k}{\bs \alpha (-\bs S_i)^{-1}\bs e_k} = 1- \bs e_k e^{\widetilde{ \bs S}_iu}\bs e.\label{eqn:ksda}\end{align}
+% Once again, noting that the conversion between the intensity at which events occurs when the fluid queue is in phase \(i\) compared to phase \(j\) is \(|c_j|/|c_i|\)%, and taking the derivative with respect to \(v\) of (\ref{eqn:ksda}) by applying the chain rule, gives
+% %\begin{align}%
+% %	\cfrac{\wrt }{\wrt u} \bs e_k e^{\widetilde{ \bs S}_iu}\bs e|c_j|/|c_i|=\bs e_k e^{\widetilde{ \bs S}_iv|c_j|/|c_i|}\widetilde{ \bs S}_i\bs e|c_j|/|c_i| = \bs e_k e^{\widetilde{ \bs S}_jv}\widetilde{ \bs s}_j.\label{eqn:ksdadsva}
+% %\end{align}
+% then, upon a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the reverse-time generator \(\widetilde{ \bs S}_j\).
 The distribution of time until the next event in the QBD-RAP after time \(t+u\) is
-\[1-\bs e_k e^{(\widetilde{\bs S}_j + T_{jj}\bs I)a}\bs e.\] 
-This suggests that, at a jump from \(\calS_+\) to \(\calS_-\) or from \(\calS_-\) to \(\calS_+\), the state of \(\{J(t+u)\}\) does not change, and begins to evolve according to the time-reverse generator at an appropriate speed.
-
-With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2013} with discretisation parameter \(\Delta/p\).
-
-\subsection{Computing \(\bs D\)}
+% \[1-\bs e_k e^{(\widetilde{\bs S}_j + T_{jj}\bs I)a}\bs e.\] 
+% This suggests that, at a jump from \(\calS_+\) to \(\calS_-\) or from \(\calS_-\) to \(\calS_+\), the state of \(\{J(t+u)\}\) does not change, and begins to evolve according to the time-reverse generator at an appropriate speed.
+
+% With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2013} with discretisation parameter \(\Delta/p\).
+
+\paragraph{The matrix-exponential case}
+% Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
+We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\).
+Correspondingly, the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units less than \(y_{k+1}\).
+We have also reasoned that for \(j\in\calS_-\) the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_{k+1}-x\), that is \(x\) units less than \(y_{k+1}\), and the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units greater than \(y_k\).
+Notice that in positive phases the orbit position describes the exact position of \(X(t)\) from the left-hand boundary of the interval but only approximates the position of \(X(t)\) from the right-hand boundary of the interval, and vice-versa for negative phases.
+
+% As before, we need to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\), \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) to 
+% \[\bs A(t+u) = \cfrac{\bs A(t+u^-)\bs D(i,j)}{\bs A(t+u^-)\bs D(i,j)\bs e}=\cfrac{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)}{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
+% in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\)or, equivalently, \(\Delta-|c_i|u\) from \(y_{k+1}\).
+
+% The matrix \(\bs D\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D, \bs S)\) is a representation of a matrix exponential distribution.
+We would also like the matrix \(\bs D\) to have the property \(\bs D \bs e = \bs e\) as this will mean that the rate at which a change of phase from \(i\) to \(j\) of the QBD-RAP occurs proportional to
+% \[\bs A(t+u) \bs D(i,j)\bs e = 1,\]
+% which is constant and therefore the distribution of time until a change from phase \(i\) to \(j\) is exponential.
+We can use this to show that for certain models, the distribution of the phase process of the fluid queue and the distribution of the phase process of the QBD-RAP share the same distribution.
+
+For matrix exponential distributions we cannot rely on the phase process \(\{J(t)\}\) as we did in the phase-type case.
+Recall that we want to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\), to
+\[\bs A(t+u) = \cfrac{\bs A(t+u^-)\bs D(i,j)}{\bs A(t+u^-)\bs D(i,j)\bs e}=\cfrac{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)}{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
+in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\) in phase \(j\).
+
+Given our interpretation of the orbit position, \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\), a solution would be find a mapping which takes \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) and maps it directly to \(\cfrac{\bs\alpha e^{\bs S(\Delta-x)} }{\bs\alpha e^{\bs S(\Delta-x) } \bs e}\).
+However, we have been unsuccessful in finding such a mapping.
+The main hurdle being that the map must be linear.
+Instead, we approximate as follows.
+
+Recall that we use \(R_i(u)\) to approximate the distance of the fluid level to the left of \(y_{k+1}\).
+Suppose we are given \(R_i(u)=r\) which corresponds, approximately, to the fluid level being \(|c_i|r\) units to the left of \(y_{k+1}\).
+The position of the orbit process which corresponds to a distance of \(|c_i|r\) to the left of \(y_{k+1}\) in phase \(j\) is \(\cfrac{\bs\alpha e^{ \bs S_j|c_i|r/|c_j| } }{\bs\alpha e^{\bs S_j |c_i|r/|c_j| } \bs e}=\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\).
+Hence, on \(R_i(u)=r\), the event \(E\) and on the event that a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) occurs at time \(t+u\), the orbit position should jump from \(\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) to \(\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\).
+However, at time \(t+u\), \(R_i(u)\) is a random variable about the future of the process and therefore not known, so instead, we take the expected initial vector
+\[\mathbb E \left[ \cfrac{\bs\alpha e^{ \bs S_iR_i(u) } }{\bs\alpha e^{\bs S_iR_i(u) } \bs e} \right] = \int_{r=0}^\infty \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} e^{\bs S_ir } \bs s_i \cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{r=0}^\infty e^{\bs S_ir } \bs s_i\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r.\]
+After a change of variables \(x=|c_i|r\) we get 
+\[\cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{x=0}^\infty e^{\bs S x }|c_i| \bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x/|c_i| = \bs A(t+u^-)\int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x,\]
+since at time \(t+u^-\), the orbit position is \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e}\). Thus, we find 
+\[\bs D(i,j)=\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x=:\bs D.\]
+
+Observe that \(\bs D\bs e = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\bs e =\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \wrt x=\bs e\).
+Further, since \(\mathcal A\) is closed and convex \citep{MEinAP}, then \((\bs a \bs D,\bs S,\bs s)\) is a representation of a matrix exponential distribution for any \(\bs a\in\mathcal A\).
+
+We pose the choice of the matrix \(\bs D\) as a modelling choice.
+Other choices are possible, for example, \(\displaystyle \bs D = \int_{x=0}^{\Delta-\varepsilon}e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\).
+It may also be possible to construct other matrices \(\bs D\), perhaps via geometric arguments.
+
+\paragraph{Computing \(\bs D\)}
 In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hht2020}.
@@
 The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}.
-Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\).
+Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform only marginally better than an approximation constructed with representations of order \(2p-1\).
 However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.
@@
 Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \citep{hht2020}.
+
+\subsection{Upon exiting \(\calD_{k,i}\)} Suppose that upon exiting \(\calD_{k,i}\) at time \(t\) the phase is \(\varphi(t)=i\in\calS_+\).
+At this time \(X(t)=y_{k+1}\) which is the left-hand endpoint of \(\calD_{k+1,i}\).
+Hence, we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
+Similarly, upon exiting \(\calD_{k,i}\) at time \(t\) in phase \(\varphi(t)=i\in\calS_-\), then \(X(t)=y_k\), which is the right-hand endpoint of \(\calD_{k,i}\), and so we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
 
@@
 
+% \paragraph{Other approaches}
+% There are other possible approaches to model this behaviour which do not require duplicating states and may achieve a small computational saving, however they are less amenable to analysis and possibly lead to less accurate schemes.
+Instead of duplicating states in \(\calS_0\), we can choose to associate \(j_1\in\calS_0\) with either \(\calS_+\) or \(\calS_-\).
+
+Associating \({j_1}\) with \(\calS_+\), amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{I}\) and \(\bs{D}({j_1},{j_2})=\bs{D}\); associating \({j_1}\) with \(\calS_-\), amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{D}\) and \(\bs{D}({j_1},{j_2})=\bs{I}\).
+
+There are some consequences to this choice.
+Let \(k_2\in \calS_+\).
+Consider an event where the phase process of the fluid queue transitions from \({j_0}\to {j_1}\to k_2\) and there is no change of level.
+If \({j_1}\) is associated with \(\calS_+\), then \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{I}\) and the corresponding orbit process, given \(\bs A(0) = \bs \alpha\), is
+\begin{align*}
+\bs A(t) &= \begin{cases} 
+	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
+	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
+	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
+\end{cases}
+\end{align*}
+%This sample path of the orbit process captures the fact that the process initially spent \(t_1-t_0\) amount of time in \({j_0}\in\calS_+\), exactly.
+Hence, at time \(t_2\), the orbit process knows that it has spent exactly \(t_1-t_0\) amount of time moving at rate \(c_{j_0}\).
+Notice that there is no matrix \(\bs D\) in this expression.
+
+Compare this to if \({j_1}\) is associated with \(\calS_-\).
+In this case \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{D}\) and the corresponding orbit process of the approximation, given \(\bs A(0) = \bs \alpha\) and there are no change of level in \(\left[t_1,t_3\right)\), is
+\begin{align*}
+\bs A(t) &= \begin{cases} 
+	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
+	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
+	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
+\end{cases}
+\end{align*}
+Ideally \(\bs D^2=\bs I\), however this is not the case here.
+Recall that a jump according to \(\bs{D}\) corresponds to approximating the residual life by an expectation.
+With this interpretation as an approximation, it suggests that we might want to minimise the number of jumps according to \(\bs D\) which occur.
+Therefore, for \(j_1\in \calS_0\), if transitions \(\calS_+\to j_1\to \calS_+\) occur with high probability compared to transition \(\calS_-\to j_1\to \calS_-\), then this suggests we might want to associate \(j_1\) with \(\calS_+\).
+Which association is chosen will depend on the parameters of the fluid queue and on which aspects of the model we wish to approximate.
+Although this advice is based on intuition only, numerical results in Section TBC suggest that it is reasonable.
+
+% It might be tempting to choose \(\bs D(j_0,j_1)=\bs D(j_1,j_2)=\bs D^{1/2}\) (provided that the square root matrix exists).
+With this choice \(\bs D(j_0,j_1)\bs D(j_1,j_2)=\bs D\) as desired.
+However, for sample paths which transition from \(j_0\to j_1\to k\), we have \(\bs D(j_0,j_1)\bs D(j_1,k) = \bs D^{1/2}\bs D(j_1,k)\).
+
 \subsubsection{Augmented state-space schemes}\label{subsec: augmented }
-One way to approach this problem is to augment the state space of the phase process by duplicating \(\calS_0\) and associating one copy of \(\calS_0\) with \(\calS_+\) and one copy of \(\calS_0\) with \(\calS_-\).
+Another way to approach the problem is to augment the state space of the phase process by duplicating \(\calS_0\) and associating one copy of \(\calS_0\) with \(\calS_+\) and one copy of \(\calS_0\) with \(\calS_-\).
 Let \(\left\{\varphi^*(t)\right\}\) be the augmented CTMC with state space \(\calS^*\) and generator \(\bs{T}^*\).
@@
 \]
-Also define a fluid level \(\left\{X^*(t)\right\}\) using \(\left\{\varphi^*(t)\right\}\), with rates \(c_i^*=c_i\) for \(i\in\calS_+\cup\calS_-\) and \(c_{(m,i)}^* = 0\) for \((m,i)\in\calS_{0+}\cup\calS_{0-}\).
-The process \(\left\{\varphi(t)\right\}\) is imbedded within \(\left\{\varphi^*(t)\right\}\) and is recovered by marginalising over \(\calS_{0+}\) and \(\calS_{0-}\).
-Given \(X^*(0)=X(0)\), the fluid levels \(X^*(t)\) and \(X(t)\) match exactly.
-Hence, if we approximate \(\left\{(X^*(t),\varphi^*(t))\right\}\), then we can recover an approximation to \(\left\{(X(t),\varphi(t))\right\}\).
+Also define a fluid level \(\left\{X^*(t)\right\}\) using \(\left\{\varphi^*(t)\right\}\), with rates \(c_i^*=c_i\) for \(i\in\calS_+\cup\calS_-\) and \(c_{(m,i)}^* = 0\) for \((m,i)\in\calS_{+0}\cup\calS_{-0}\).
+The process \(\left\{\varphi(t)\right\}\) is imbedded within \(\left\{\varphi^*(t)\right\}\) and is recovered by marginalising over \(\calS_{+0}\) and \(\calS_{-0}\).
+On the event \(X^*(0)=X(0)\), the fluid levels \(X^*(t)\) and \(X(t)\) match exactly.
+Hence, by approximating \(\left\{(X^*(t),\varphi^*(t))\right\}\), we can recover an approximation to \(\left\{(X(t),\varphi(t))\right\}\).
 This construction removes the problem of having to choose how to associate states \(j\in\calS_0\) with either \(\calS_+\) or \(\calS_-\)
 
-The generator for the QBD-RAP approximation to the augmented fluid process is 
+The generator for the QBD-RAP approximation to the augmented fluid process (ignoring boundaries) is 
 \begin{align*}
@@
 
-\subsubsection{Other approaches}
-There are other possible approaches to model this behaviour which do not require duplicating states and may achieve a small computational saving, however they are less amenable to analysis and possibly lead to less accurate schemes.
-Instead of duplicating states in \(\calS_0\), we can choose to associate \(j_1\in\calS_0\) with either \(\calS_+\) or \(\calS_-\).
-If we associate \({j_1}\) with \(\calS_+\), this amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{I}\) and \(\bs{D}({j_1},{j_2})=\bs{D}\); if we associate \({j_1}\) with \(\calS_-\), this amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{D}\) and \(\bs{D}({j_1},{j_2})=\bs{I}\).
-
-There are some consequences to this choice.
-Let \(k_2\in \calS_+\).
-Consider an event where the phase process of the fluid queue transitions from \({j_0}\to {j_1}\to k_2\) and there is no change of level.
-If \({j_1}\) is associated with \(\calS_+\), then \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{I}\) and the corresponding orbit process, given \(\bs A(0) = \bs \alpha\), is
-\begin{align*}
-\bs A(t) &= \begin{cases} 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
-\end{cases}
-\end{align*}
-%This sample path of the orbit process captures the fact that the process initially spent \(t_1-t_0\) amount of time in \({j_0}\in\calS_+\), exactly.
-Hence, at time \(t_2\), the orbit process knows that it has spent exactly \(t_1-t_0\) amount of time moving at rate \(c_{j_0}\).
-Notice that there is no matrix \(\bs D\) in this expression.
-
-Compare this to if \({j_1}\) is associated with \(\calS_-\).
-In this case \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{D}\) and the corresponding orbit process of the approximation, given \(\bs A(0) = \bs \alpha\) and there are no change of level in \(\left[t_1,t_3\right)\), is
-\begin{align*}
-\bs A(t) &= \begin{cases} 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
-\end{cases}
-\end{align*}
-Ideally \(\bs D^2=\bs I\), however this is not the case here.
-Recall that a jump according to \(\bs{D}\) corresponds to approximating the residual life by an expectation.
-With this interpretation as an approximation, it suggests that we might want to minimise the number of jumps according to \(\bs D\) which occur.
-Therefore, for \(j_1\in \calS_0\), if transitions \(\calS_+\to j_1\to \calS_+\) occur with high probability compared to transition \(\calS_-\to j_1\to \calS_-\), then this suggests we might want to associate \(j_1\) with \(\calS_+\).
-Which association is chosen will depend on the parameters of the fluid queue and on which aspects of the model we wish to approximate.
-Although this advice is based on intuition only, numerical results in Section TBC suggest that it is reasonable.
 %Hence, at time \(t_2\) the process no longer knows that it has spent exactly \(t_1-t_0\) amount of time moving at rate \(c_{j_0}\); rather, it knows some approximation of this information in the form of an expectation.
@@
 We will show later that \(\phi(t)\) captures the phase dynamics of the fluid queue exactly, provided the phase process is independent of the fluid level \(\{X(t)\}\).
-We use the level \(L(t)=\ell\) to approximate which band \(\calD_\ell\) that \(X(t)\) is in at time \(t\), and use the orbit \(\bs A(t)\) to obtain approximations about where \(X(t)\) is within the interval \(\calD_\ell\).
+The level \(L(t)=\ell\) approximates which band \(\calD_\ell\) that \(X(t)\) is in at time \(t\), and the orbit \(\bs A(t)\) can be use to obtain an approximation of where \(X(t)\) is within the interval \(\calD_\ell\).
 
@@
 \end{align}
-Let \(\bs a_0\) be a vector such that \(\bs a_0\bs e = 1\) and \((\bs a_0,\bs{S},\bs s)\) is a representation of a ME distribution.
-Given no events occur before time \(t+u\), \(\bs A(u) = \bs a_0\in\mathcal A\) and \({\phi}(u)=i\), the solution to (\ref{eqn: At DE}) states that \(\bs A({t+u})\) evolves deterministically according to
-\[\bs A({t+u}) =  \cfrac{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}}{\bs a_0 e^{(\bs{S}_i+T_{ii}\bs{I}_p)t} \bs e} = \cfrac{\bs a_0 e^{\bs{S}_i t}}{\bs a_0 e^{\bs{S}_i t} \bs e}.\]
+Let \(\bs a \in \mathcal A\) be an arbitrary vector in \(\mathcal A\).
+On the event that no events occur before time \(t+u\), \(\bs A(u) = \bs a\in\mathcal A\) and \({\phi}(u)=i\), the solution to (\ref{eqn: At DE}) states that \(\bs A({t+u})\) evolves deterministically according to
+\[\bs A({t+u}) =  \cfrac{\bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}}{\bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t} \bs e} = \cfrac{\bs a e^{\bs{S}_i t}}{\bs a e^{\bs{S}_i t} \bs e}.\]
 
@@
 The two scenarios, \(sign(c_i)\neq sign(c_j)\) and \(sign(c_i)= sign(c_j)\), can be written succinctly by stating that, at the time of an event corresponding to a change of phase from \(i\) to \(j\), \(\{\bs A(t)\}\) jumps to \(\bs A(t)\bs{D}^{1(sign(c_i)\neq sign(c_j))}\) and begins to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), meanwhile \(\left\{{\phi}(t)\right\}\) jumps from \(\phi(t^-)=i\) to \(\phi(t)=j\).
-Moreover, given a change of phase event occurs at time \(t\) with \(\phi(t^-)=i\) then, with probability \(\cfrac{\bs A(t^-)T_{ij}\bs{D}^{1(sign(c_i)\neq sign(c_j))}\bs e}{-\bs A(t^{-})T_{ii}\bs e} = \cfrac{T_{ij}}{-T_{ii}}\), the event corresponds to a change of phase from \(i\) to \(j\).
+% Moreover, given a change of phase event occurs at time \(t\) with \(\phi(t^-)=i\) then, with probability \(\cfrac{\bs A(t^-)T_{ij}\bs{D}^{1(sign(c_i)\neq sign(c_j))}\bs e}{-\bs A(t^{-})T_{ii}\bs e} = \cfrac{T_{ij}}{-T_{ii}}\), the event corresponds to a change of phase from \(i\) to \(j\).
 
@@
 %	\label{eqn: density approx aug}
-%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_0, \varphi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x & j\in\mathcal S_+\cup \mathcal S_{+0},
+%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x & j\in\mathcal S_+\cup \mathcal S_{+0},
 %		\\\label{eqn: density approx aug2}
-%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_0, \varphi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x & j\in\mathcal S_-\cup \mathcal S_{-0},
+%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x & j\in\mathcal S_-\cup \mathcal S_{-0},
 %\end{align}

### 2021-11-08 chapter2/chapter.tex
first steps of nigels changes to first part of ch 3. up to sec 3.3. about to change to matrix notation instead of partition into phases.

@@
 Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic.
-We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \cite{ert2006,hstz2016,ert2006,hht2020,mt2021}.
+We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \citep{ert2006,hstz2016,ert2006,hht2020,mt2021}.
 We construct a level process for the QBD-RAP and correspond it to a discretisation the level process of the fluid queue.
@@
 
-We interpret the position of the orbit \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\).
+For notational convenience, define the row-vector-valued function \(\bs k(t):\mathbb R\to \mathcal A\), 
+\begin{equation}
+	\bs k(t) = \cfrac{\bs \alpha e^{\bs S t}}{\bs \alpha e^{\bs S t}\bs e}, \label{eqn: orbit fun def}
+\end{equation}
+and the operator \(\bs K(t): \mathcal A\to \mathbb R\), \(t\geq 0\), 
+\begin{equation}
+	\bs a \bs K(t) = \cfrac{\bs a e^{\bs S t}}{\bs a e^{\bs S t}\bs e}, \label{eqn: operator fun def}
+\end{equation}
+for any \(a\in\mathcal A\).
+
+We interpret the position of the orbit \(\bs A(t+u) = \bs k(|c_i|u)\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\).
 This gives a heuristic argument as to how we can model the sojourn times in a given interval \(\mathcal D_{k,i}\) on the event that the phase does not change.
@@
 We can apply analogous arguments to heuristically develop a model for the sojourn time of the fluid queue in an interval, \(\calD_{k,i}\), \(i\in\calS_-\), on the event that the phase does not change, given the fluid is \(X(t)=y_{k+1}\).
-For \(i\in\calS_-\), the orbit position \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).
+For \(i\in\calS_-\), the orbit position \(\bs A(t+u) = \bs k(|c_i|u)\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).
 
@@
 \begin{align}
-	\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}.\label{eqn: orbit i}
+	\bs A(t+u) = \bs k(|c_i|u) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}.\label{eqn: orbit i}
 \end{align}
@@
 
-If, at time \(t+u\) there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\), then we need to map the orbit position \(\bs A(t+u^-) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}\) to an orbit position which corresponds to the being \(c_iu\) units from the left-hand edge of \(\calD_k\) in phase \(j\).
+If, at time \(t+u\) there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\), then we need to map the orbit position \(\bs A(t+u^-) = \bs k(|c_i|u)\) to an orbit position which corresponds to the being \(c_iu\) units from the left-hand edge of \(\calD_k\) in phase \(j\).
 Call this mapping \(\bs D(i,j)(\cdot)\).
@@
 
-An orbit position which corresponds to the level process of the QBD-RAP beign \(c_iu\) units to the right of \(y_k\) in phase \(j\in\calS_+\) is \(\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e}\).
+An orbit position which corresponds to the level process of the QBD-RAP beign \(c_iu\) units to the right of \(y_k\) in phase \(j\in\calS_+\) is 
+\[\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|)}}{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|) u}\bs e}.\] 
 To see this, consider the event that at time \(t\) the orbit process of the QBD-RAP is \(\bs A(t) = \bs \alpha\), the phase process is \(\varphi(t) = j\in\calS_+\), and there are no events of the QBD-RAP before time \(t+|c_i|u/|c_j|\).
-On this event, the orbit position at time \(t+u\) will be \(\displaystyle \bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e}\).
+On this event, the orbit position at time \(t+u\) will be
+\[\displaystyle \bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|) u}\bs e}.\] 
 Correspondingly, on the event that at time \(t\) the level process of the fluid queue is \(X(t)=y_{k}\), the phase is \(\varphi(t)=j\) and there are no change of phase by time \(t+|c_i|u/|c_j|\), then \(X(t+u)=y_{k}+c_j|c_i|u/|c_j| = y_k+c_iu\).
@@
 Now, observe 
-\[\cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j| u}\bs e} = \cfrac{\bs \alpha e^{\bs S_iu}}{\bs \alpha e^{\bs S_i u}\bs e},\] 
+\[\cfrac{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|)}}{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|) u}\bs e} = \cfrac{\bs \alpha e^{\bs S_iu}}{\bs \alpha e^{\bs S_i u}\bs e} = \bs k(|c_i|u),\] 
 which is exactly (\ref{eqn: orbit i}), the orbit position on the event \(E\). Hence, it is sensible to choose \(\bs D(i,j)=\bs I\).
@@
 Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
-We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\).
-
-As before, we need to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\), to
-\[\bs A(t+u) = \cfrac{\bs A(t+u^-)\bs D(i,j)}{\bs A(t+u^-)\bs D(i,j)\bs e}=\cfrac{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)}{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
+%We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\bs k(x)\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\).
+As before, we need to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \[\bs A(t+u^-) = \bs k(|c_i|u^-),\]
+to 
+\[\bs A(t+u) = \cfrac{\bs k(u^-)\bs D(i,j)}{\bs k(u^-)\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
 in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\) in phase \(j\).
@@
 
-In the discussions above, we have relied on the relationship between the event \(E\) and the orbit position \(\displaystyle \bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\).
+In the discussions above, we have relied on the relationship between the event \(E\) and the orbit position \(\displaystyle \bs k(|c_i|u)\).
 The relationship allows us to associate this orbit position with the level of the fluid queue.
-For phase-type distributions the vector \(\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) is the vector of posterior probabilities \(\left[\mathbb P(J(u)=k \mid E)\right]_{k\in\{1,...,p\}}\).
+For phase-type distributions the vector \(\bs k(|c_i|u)\) is the vector of posterior probabilities \(\left[\mathbb P(J(u)=k \mid E)\right]_{k\in\{1,...,p\}}\).
 We can use this vector of posterior probabilities in the same way as we do for matrix-exponential distributions and QBD-RAPs.
@@
 Notice that this is independent of the time since the phase-type distribution was initialised.
-We call this time since the phase-type distribution was initialised the \emph{age}.
-As noted by \cite{hmp2017}, the distribution of the age given the phase of the phase-type is \(k\in\{1,...,p\}\) depends on the sampling scheme which determines the time of the observation.
+We call the time since the phase-type distribution was initialised the \emph{age}.
+As noted by \cite{hmp2017}, the distribution of the age given the phase of the phase-type is \(k\in\{1,...,p\}\) depends on the sampling scheme which determines the observation time.
 Here, on the event that at time \(t\) the phase of the QBD-RAP is \(i\), a change of phase occurs after an exponential amount of time with rate \(-T_{ii}\).
@@
 
-Recall that in the matrix-exponential case we associated the value of the orbit process \(\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) with being \(c_iu\) units to the right of \(y_{k}\).
+Recall that in the matrix-exponential case we associated the value of the orbit process \(\bs k(|c_i|u)\) with being \(|c_i|u\) units to the right of \(y_{k}\).
 In contrast, for the phase-type case if we use the phase process \(\{J(t)\}\) to inform our modelling of the fluid queue (as opposed to the value of the orbit process), then we associate phase \(k\in\{1,...,p\}\) of \(\{J(t)\}\) with being a \emph{random} distance to the right of \(y_k\) where this distance has distribution
@@
 For matrix exponential distributions we cannot rely on the phase process \(\{J(t)\}\) as we did in the phase-type case.
-Recall that we want to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\), to
-\[\bs A(t+u) = \cfrac{\bs A(t+u^-)\bs D(i,j)}{\bs A(t+u^-)\bs D(i,j)\bs e}=\cfrac{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)}{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
+Recall that we want to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \(\bs A(t+u^-) = \bs k(|c_i|u)\), to
+\[\bs A(t+u) = \cfrac{\bs k(|c_i|u)\bs D(i,j)}{\bs k(|c_i|u)\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
 in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\) in phase \(j\).
 
-Given our interpretation of the orbit position, \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\), a solution would be find a mapping which takes \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) and maps it directly to \(\cfrac{\bs\alpha e^{\bs S(\Delta-x)} }{\bs\alpha e^{\bs S(\Delta-x) } \bs e}\).
-However, we have been unsuccessful in finding such a mapping.
+Given our interpretation of the orbit position, \(\bs k(x)\), a solution would be find a linear map which takes \(\bs k(x)\) and maps it directly to 
+\[\bs k(\Delta-x) = \cfrac{\bs\alpha e^{\bs S(\Delta-x)} }{\bs\alpha e^{\bs S(\Delta-x) } \bs e}.\] However, we have been unsuccessful in finding such a mapping.
 The main hurdle being that the map must be linear.
@@
 Suppose we are given \(R_i(u)=r\) which corresponds, approximately, to the fluid level being \(|c_i|r\) units to the left of \(y_{k+1}\).
-The position of the orbit process which corresponds to a distance of \(|c_i|r\) to the left of \(y_{k+1}\) in phase \(j\) is \(\cfrac{\bs\alpha e^{ \bs S_j|c_i|r/|c_j| } }{\bs\alpha e^{\bs S_j |c_i|r/|c_j| } \bs e}=\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\).
-Hence, on \(R_i(u)=r\), the event \(E\) and on the event that a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) occurs at time \(t+u\), the orbit position should jump from \(\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) to \(\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{\bs S_ir } \bs e}\).
+The position of the orbit process which corresponds to a distance of \(|c_i|r\) to the left of \(y_{k+1}\) in phase \(j\) is \(\bs k(|c_i|r).\)
+Hence, on \(R_i(u)=r\), the event \(E\) and on the event that a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) occurs at time \(t+u\), the orbit position should jump from \(\bs k(|c_i|u)\) to \(\bs k(|c_i|r)\).
 However, at time \(t+u\), \(R_i(u)\) is a random variable about the future of the process and therefore not known, so instead, we take the expected initial vector
-\[\mathbb E \left[ \cfrac{\bs\alpha e^{ \bs S_iR_i(u) } }{\bs\alpha e^{\bs S_iR_i(u) } \bs e} \right] = \int_{r=0}^\infty \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} e^{\bs S_ir } \bs s_i \cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{r=0}^\infty e^{\bs S_ir } \bs s_i\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r.\]
+\[\mathbb E \left[ \bs k(|c_i|R_i(u))\right] = \int_{r=0}^\infty \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} e^{\bs S_ir } \bs s_i \bs k(|c_i|r)\wrt r = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{r=0}^\infty e^{\bs S_ir } \bs s_i\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r.\]
 After a change of variables \(x=|c_i|r\) we get 
 \[\cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{x=0}^\infty e^{\bs S x }|c_i| \bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x/|c_i| = \bs A(t+u^-)\int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x,\]
-since at time \(t+u^-\), the orbit position is \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e}\). Thus, we find 
-\[\bs D(i,j)=\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x=:\bs D.\]
-
-Observe that \(\bs D\bs e = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\bs e =\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \wrt x=\bs e\).
+since at time \(t+u^-\), the orbit position is \(\bs A(t+u^-) = {\bs\alpha e^{\bs S_i u } }/{\bs\alpha e^{\bs S_iu } \bs e}\). Thus, we find 
+\[\bs D(i,j)=\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \bs k(x)\wrt x=:\bs D.\]
+
+Observe that \(\bs D\bs e = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \bs k(x)\wrt x\bs e =\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \wrt x=\bs e\), since \(\bs k(x)\bs e = 1\) for all \(x\geq 0\).
 Further, since \(\mathcal A\) is closed and convex \citep{MEinAP}, then \((\bs a \bs D,\bs S,\bs s)\) is a representation of a matrix exponential distribution for any \(\bs a\in\mathcal A\).
@@
 We pose the choice of the matrix \(\bs D\) as a modelling choice.
-Other choices are possible, for example, \(\displaystyle \bs D = \int_{x=0}^{\Delta-\varepsilon}e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x\).
+Other choices are possible, for example, \(\displaystyle \bs D = \int_{x=0}^{\Delta}e^{\bs S x }\bs s \bs k(x)\wrt x\) is another sensible choice.
 It may also be possible to construct other matrices \(\bs D\), perhaps via geometric arguments.
@@
 \begin{align*}
-	\bs{D} = \displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\cfrac{\bs \alpha e^{\bs{S}t}}{\bs \alpha e^{\bs{S}t} \bs e}\wrt t
+	\bs{D} = \displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\bs k(t)\wrt t
 \end{align*}
 using a trapezoidal rule as follows.
-For CMEs, the vector function \(\bs a(t) := \cfrac{\bs \alpha e^{\bs St}}{\bs \alpha e^{\bs S t} \bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
+For the CMEs considered here, the vector function \(\bs k(t)\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
 Let \(\bs f(t) = e^{\bs S t}\bs s\).
@@
 \begin{align}
-	\nonumber\bs{D} &= \displaystyle \int_{t=0}^\infty \bs f(t)\cdot\bs a(t)\wrt t  %\displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\cfrac{\bs \alpha e^{\bs{S}t}}{\bs \alpha e^{\bs{S}t} \bs e}\wrt t
-	\\\nonumber&= \sum_{k=0}^\infty \int_{k\rho }^{(k+1)\rho } e^{\lambda t} e^{-\lambda t} \bs f(t)\cdot\bs a(t)\wrt t
-	\\&= \sum_{k=0}^\infty \int_{0}^{\rho }  e^{\lambda(k\rho +t)}e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs a(k\rho +t)\wrt t. \label{eqn: sc6789gGHJ}
-	\intertext{By periodicity, then \(e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs a(k\rho +t)=e^{ -\lambda t} \bs f(t)\cdot\bs a(t)\), hence (\ref{eqn: sc6789gGHJ}) is equal to}
-	&\sum_{k=0}^\infty (e^{\lambda \rho })^k\int_{0}^{\rho } e^{\lambda t}e^{-\lambda t}\bs f(t)\cdot\bs a(t)\wrt t  
-	= \cfrac{1}{1 - e^{\lambda \rho }}\int_{0}^{\rho } \bs f(t)\cdot\bs a(t)\wrt t,\label{eqn: hdsv45KJ}
+	\nonumber\bs{D} &= \displaystyle \int_{t=0}^\infty \bs f(t)\cdot\bs k(t)\wrt t  %\displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\cfrac{\bs \alpha e^{\bs{S}t}}{\bs \alpha e^{\bs{S}t} \bs e}\wrt t
+	\\\nonumber&= \sum_{k=0}^\infty \int_{k\rho }^{(k+1)\rho } e^{\lambda t} e^{-\lambda t} \bs f(t)\cdot\bs k(t)\wrt t
+	\\&= \sum_{k=0}^\infty \int_{0}^{\rho }  e^{\lambda(k\rho +t)}e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs k(k\rho +t)\wrt t. \label{eqn: sc6789gGHJ}
+	\intertext{By periodicity, then \(e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs k(k\rho +t)=e^{ -\lambda t} \bs f(t)\cdot\bs k(t)\), hence (\ref{eqn: sc6789gGHJ}) is equal to}
+	&\sum_{k=0}^\infty (e^{\lambda \rho })^k\int_{0}^{\rho } e^{\lambda t}e^{-\lambda t}\bs f(t)\cdot\bs k(t)\wrt t  
+	= \cfrac{1}{1 - e^{\lambda \rho }}\int_{0}^{\rho } \bs f(t)\cdot\bs k(t)\wrt t,\label{eqn: hdsv45KJ}
 \end{align}
@@
 To approximate (\ref{eqn: hdsv45KJ}) numerically, we first partition \(\left[0,\rho \right)\) into \(N\) equal-width intervals \(\left[t_n,t_{n+1}\right)\), where \(t_n = (n-1)\rho /N\), \(n=1,2,...,N+1\).
-On \(\left[t_n,t_{n+1}\right)\) we approximate the orbit \(\bs a(t)\) by a constant \(\bs a(t) \approx \bs a_n := \cfrac{1}{2}\left( \bs a(t_n) + \bs a(t_{n+1}) \right), \, t \in \left[t_n,t_{n+1}\right)\).
+On \(\left[t_n,t_{n+1}\right)\) we approximate the orbit \(\bs k(t)\) by a constant \(\bs k(t) \approx \bs k_n := \cfrac{1}{2}\left( \bs k(t_n) + \bs k(t_{n+1}) \right), \, t \in \left[t_n,t_{n+1}\right)\).
 Substituting this approximation into the expression for \(\bs{D}\) gives
 \begin{align*}
-	\bs{D} &\approx \cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\int_{t_n}^{t_{n+1}} \bs f(t)\cdot\bs a_n\wrt t
-	\\&=\cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\left[e^{\bs{S}t_{n+1}}-e^{\bs{S}t_{n}}\right]\bs e \cdot\bs a_n.
+	\bs{D} &\approx \cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\int_{t_n}^{t_{n+1}} \bs f(t)\cdot\bs k_n\wrt t
+	\\&=\cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\left[e^{\bs{S}t_{n+1}}-e^{\bs{S}t_{n}}\right]\bs e \cdot\bs k_n.
 \end{align*}
@@
 
-Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \citep{hht2020}.
+Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs k(t)\) are provided in \citep{hht2020}.
 
@@
 Let \(\calS_+\) and \(\calS_-\) be as before and \(\calS_{m0} = \left\{(m,i)\mid i\in\calS_m\right\}\), \(m\in\left\{+,-\right\}\), then \(\calS^* = \calS_+\cup\calS_-\cup\calS_{+0}\cup\calS_{-0}\).
-The generator of \(\varphi^*(t)\) is
-\[\bs{T}^* = \left[\begin{array}{cccc} \bs{T}_{++} & \bs{T}_{+-} & \bs{T}_{+0} & 0 \\ \bs{T}_{-+} & \bs{T}_{--} & 0 & \bs{T}_{-0} \\ \bs{T}_{0+} & \bs{T}_{0-} & \bs{T}_{00} & 0 \\  \bs{T}_{0+} & \bs{T}_{0-} & 0 & \bs{T}_{00} \end{array}\right].
+The generator of \(\varphi^*(t)\) can be written as 
+\[\bs{T}^* = \left[\begin{array}{cccc} \bs{T}_{++} & \bs{T}_{+0} & \bs{T}_{+-} & 0 \\ \bs{T}_{0+} & \bs{T}_{00} & \bs{T}_{0-} & 0 \\ \bs{T}_{-+} & 0 & \bs{T}_{--} & \bs{T}_{-0} \\  \bs{T}_{0+} & 0 & \bs{T}_{0-} & \bs{T}_{00} \end{array}\right].
 \]
@@
 \bs{B}_{0} &= \left[\begin{array}{cccc}
-	\bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs{T}_{+0}\otimes \bs{I} & 0 \\
-	\bs{T}_{-+}\otimes \bs{D} & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & 0 & \bs{T}_{-0}\otimes \bs{I} \\
-	\bs{T}_{0+}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs{T}_{00}\otimes \bs{I} & 0 \\
-	\bs{T}_{0+}\otimes \bs{D} & \bs{T}_{0-}\otimes \bs{I} & 0 & \bs{T}_{00}\otimes \bs{I} 
+	\bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I}  & \bs{T}_{+0}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs 0 \\
+	\bs{T}_{0+}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs 0 \\
+	\bs{T}_{-+}\otimes \bs{D} & \bs 0 & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & \bs{T}_{-0}\otimes \bs{I} \\
+	\bs{T}_{0+}\otimes \bs{D} & \bs 0 & \bs{T}_{0-}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} 
 	\end{array}\right],
@@
 	\bs 0 & & & \\
-	& \bs{C}_-\otimes (\bs s\bs\alpha) & & \\
-	& & \bs 0 & \\
+	& \bs 0 & & \\
+	& & \bs{C}_-\otimes (\bs s\bs\alpha) & \\
 	& & & \bs 0
@@
 Later, we will make clear why we say that the latter event corresponds to a change of level.
-Upon an event occurring at time \(t\), with probability \(\cfrac{-T_{ii}}{-T_{ii} + \bs A({t^-})\bs s_i}\) the event corresponds to a change of phase and with probability \(\cfrac{\bs A({t^-})\bs s_i}{-T_{ii} + \bs A({t^-})\bs s_i}\) the event corresponds a change of level.
+Upon an event occurring at time \(t\), with probability \({-T_{ii}}/{\left(-T_{ii} + \bs A({t^-})\bs s_i\right)}\) the event corresponds to a change of phase and with probability \({\bs A({t^-})\bs s_i}/{\left(-T_{ii} + \bs A({t^-})\bs s_i\right)}\) the event corresponds a change of level.
 
@@
 
-Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)\neq sign(c_j)\), the process \(\{\bs A(t)\}\) jumps to \(\bs A(t) = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{\bs A({t^-}) T_{ij} \bs{D} \bs e} = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{T_{ij} } = \bs A({t^-}) \bs{D} \) and then immediately proceeds to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), and \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
+Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)\neq sign(c_j)\), the process \(\{\bs A(t)\}\) jumps to 
+\[\bs A(t) = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{\bs A({t^-}) T_{ij} \bs{D} \bs e} = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{T_{ij} } = \bs A({t^-}) \bs{D} \] 
+and then immediately proceeds to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), and \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
 
@@
 For boundary conditions which interact with the phase dynamics, such as a reflecting boundary, the result does not hold.
-The cause is the fact that the phase dynamics are level dependent -- we see a forced change of phase upon a boundary being hit -- and the QBD-RAP can only approximate the level process of the fluid queue.
+The cause is the fact that the phase dynamics are level dependent -- we may see a forced change of phase upon a boundary being hit -- and the QBD-RAP can only approximate the level process of the fluid queue.
 However, until a boundary is hit (by either the fluid queue or QBD-RAP) then the phase processes match.
@@
 \begin{align*}
-\bs{B}_{0} &= \left[\begin{array}{cccc}
-	\bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs{T}_{+0}\otimes \bs{I} & \bs 0 \\
-	\bs{T}_{-+}\otimes \bs{D} & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & \bs 0 & \bs{T}_{-0}\otimes \bs{I} \\
-	\bs{T}_{0+}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs{T}_{00}\otimes \bs{I} & \bs 0 \\
-	\bs{T}_{0+}\otimes \bs{D} & \bs{T}_{0-}\otimes \bs{I} & \bs 0 & \bs{T}_{00}\otimes \bs{I}
-	\end{array}\right],
+	\bs B &= \left[\begin{array}{ccc} \bs B_{++} & \bs B_{+-} \\ \bs B_{-+} & \bs B_{--}\end{array}\right]\\
 \end{align*}
@@
 		\bs 0 & & & \\
-	& \bs{C}_-\otimes (\bs s\bs\alpha) & & \\
-	& & \bs 0 & \\ 
+		& \bs 0 & & \\
+		& & \bs{C}_-\otimes (\bs s\bs\alpha) & \\ 
 		& & & \bs 0
@@
 		& & & \bs 0
-	\end{array}\right].
-\end{align*}
+		\end{array}\right],
+\end{align*}
+and
+\begin{align*}
+	\bs{B}_{++}  &= \left[\begin{array}{cc} \bs{C}_+\otimes \bs{S}  + \bs{T}_{++}\otimes \bs{I} & \bs T_{+0}\otimes \bs I \\ \bs T_{0+}\otimes \bs I & \bs T_{00}\otimes \bs I\end{array}\right] ,
+	%
+	\bs{B}_{+-}  = \left[\begin{array}{cc} \bs{T}_{+-}\otimes \bs{D}  & \bs 0 \\ \bs T_{0-}\otimes \bs D  & \bs 0 \end{array}\right],
+	%
+	\\ \bs{B}_{-+}  &=  \left[\begin{array}{cc} \bs{T}_{-+}\otimes \bs{D}  & \bs 0 \\ \bs T_{0+}\otimes \bs D  & \bs 0 \end{array}\right] ,
+	% 
+	\bs{B}_{--}  = \left[\begin{array}{cc} \bs{C}_-\otimes \bs{S}  - \bs{T}_{--}\otimes \bs{I} & \bs T_{-0}\otimes \bs I \\ \bs T_{0-}\otimes \bs I & \bs T_{00}\otimes \bs I\end{array}\right].
+\end{align*}
+
 %The generator acts on infinite dimensional row-vectors 
@@
 \section{Boundary conditions}\label{sec: boundary conditions}
-In the study and practical application of fluid queues, regulated, reflecting, or a mixture of these boundary conditions may be imposed.
+In the study and practical application of fluid queues, regulated, reflecting, sticky, or a mixture of these boundary conditions may be imposed.
 We now present the intuition as to how one may include such boundary conditions in the QBD-RAP approximation scheme.
@@
 Suppose that, upon hitting the boundary in phase \(i\in\calS_-\), the phase jumps from \(i\) to \(j\in\calS\) with probability \(p_{ij}\).
-If \(j\in\calS_-\cup\calS_0\) the process remains at the boundary and the phase process evolves among states in \(\calS_-\cup\calS_0\) until the first transition to a state in \(\calS_+\), at which point the level \(\{X(t)\}\) immediately leaves the boundary.
-If \(j\in\calS_+\) the process immediately leaves the boundary.
+If \(j\in\calS_-\cup\calS_0\) the process remains at the boundary and the phase process evolves among states in \(\calS_-\cup\calS_0\) until the first transition to a state in \(\calS_+\), at which point the level \(\{X(t)\}\) immediately leaves the boundary -- we will call this a \emph{sticky} boundary.
+If \(j\in\calS_+\) the process immediately leaves the boundary -- we will call this a refecting boundary.
 We collect the probabilities \(p_{ij}\) into the matrices \(\bs P_{-+} = [p_{ij}]_{i\in\calS_-,j\in\calS_+},\) \(\bs P_{-0} = [p_{ij}]_{i\in\calS_-,j\in\calS_0},\) and \(\bs P_{--} = [p_{ij}]_{i\in\calS_-,j\in\calS_-}.\)
@@
 
-For regulated behaviour at a lower boundary at \(0\), given the fluid level reaches zero at time \(u\) in phase \(i\in\mathcal S_-\), then the phase transitions to some phase \(j\in\calS_-\cup\calS_0\) with probability \(p_{ij}\), and the distribution of the phase \(\varphi(t)\) (on the event that the process has not left the boundary at or before time \(t\)) is given by the elements of the vector
+For a sticky lower boundary at \(0\), given the fluid level reaches zero at time \(u\) in phase \(i\in\mathcal S_-\), then the phase transitions to some phase \(j\in\calS_-\cup\calS_0\) with probability \(p_{ij}\), and the distribution of the phase \(\varphi(t)\) (on the event that the process has not left the boundary at or before time \(t\)) is given by the elements of the vector
 \[p_{ij} \bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right).\] 
@@
 \[1-p_{ij}\bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right)\bs e.\] 
-Therefore, in the case of a regulated boundary, we can capture the behaviour of the process at the boundary exactly.
+Therefore, in the case of a sticky boundary, we can capture the behaviour of the process at the boundary exactly.
 For a reflecting boundary there is no time spent at the boundary.
@@
 
-For both regulated boundary behaviour and reflecting boundary behaviour, given the QBD-RAP is in the correct phase at the instant upon which it hits the boundary, then upon exiting the boundary the phase can be captured exactly too.
-For a regulated boundary, given the fluid/QBD-RAP hits the boundary in phase \(i\in\mathcal S_-\) at time \(t\), it then jumps to phase \(j\in\calS_+\cup\calS_0\) and exits the boundary in phase \(k\in\mathcal S_+\) at time \((t-u)\) with density
+For both sticky and reflecting boundaries, given the QBD-RAP is in the correct phase at the instant upon which it hits the boundary, then upon exiting the boundary the phase can be captured exactly too.
+For a sticky boundary, given the fluid/QBD-RAP hits the boundary in phase \(i\in\mathcal S_-\) at time \(t\), it then jumps to phase \(j\in\calS_+\cup\calS_0\) and exits the boundary in phase \(k\in\mathcal S_+\) at time \((t-u)\) with density
 \[p_{ij}\bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right)\left[\begin{array}{c} \bs T_{-+} \\ \bs T_{0+}\end{array}\right]\bs e_k.\]
@@
 
-To summarise, at time \(t\) the rate at which probability mass accumulates at the regulated boundary in phase \(j \in\calS_-\cup\calS_0\) and level \(-1\), upon hitting the boundary in phase \(i\in\calS_-\) and level \(0\), is
+To summarise, at time \(t\) the rate at which probability mass accumulates at the sticky boundary in phase \(j \in\calS_-\cup\calS_0\) and level \(-1\), upon hitting the boundary in phase \(i\in\calS_-\) and level \(0\), is
 \[ c_ip_{ij} \bs A(t)\bs s.\]
@@
 and upon this transition, the orbit jumps to \(\bs \alpha\). 
-The rate at which mass leaves the regulated boundary from phase \(i\in\calS_-\cup\calS_0\) and level \(-1\) into phases \(j\in\calS_+\) and level \(0\) is 
+The rate at which mass leaves the sticky boundary from phase \(i\in\calS_-\cup\calS_0\) and level \(-1\) into phases \(j\in\calS_+\) and level \(0\) is 
 \[T_{ij},\]
@@
 	\bs 0 & \bs 0 &\multicolumn{4}{c|}{\multirow{4}{*}{$\check{\bs B}_{0}$}} & \multirow{4}{*}{$\bs B_{+1}$}& \\
-	(\bs C_-\bs P_{--})\otimes \bs s &(\bs C_-\bs P_{-0}) \otimes \bs s& \multicolumn{4}{c|}{} & &\\
-	\bs 0 & \bs 0 & \multicolumn{4}{c|}{} &&\\
+	\bs 0 & \bs 0 & \multicolumn{4}{c|}{} & &\\
+	(\bs C_-\bs P_{--})\otimes \bs s &(\bs C_-\bs P_{-0}) \otimes \bs s & \multicolumn{4}{c|}{} &&\\
 	\bs 0 & \bs 0 & \multicolumn{4}{c|}{} && \ddots\\ \hline
@@
 \[\check{\bs B}_{0} = \left[\begin{array}{cccc}
-	 \bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs{T}_{+0}\otimes \bs{I} &  \\
-	 (\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha + \bs T_{-+}\otimes \bs D  & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} &  & \bs{T}_{-0}\otimes \bs{I} \\
-	 \bs{T}_{0+}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs{T}_{00}\otimes \bs{I} &  \\
-	 \bs{T}_{0+}\otimes \bs{D} & \bs{T}_{0-}\otimes \bs{I} &  & \bs{T}_{00}\otimes \bs{I} 
+	 \bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+0}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \\
+	 \bs{T}_{0+}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \\
+	 (\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha + \bs T_{-+}\otimes \bs D  &  & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & \bs{T}_{-0}\otimes \bs{I} \\
+	 \bs{T}_{0+}\otimes \bs{D} & & \bs{T}_{0-}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} 
 	\end{array}\right].\]
-The top-left block represents the point masses due to the regulated boundary.
-The bottom-left block is the transition of density into point masses due to the regulated boundary.
+The top-left block represents the point masses due to the sticky boundary.
+The bottom-left block is the transition of density into point masses due to the sticky boundary.
 The top-right block is the transition of point masses into density as the process leaves the boundary.
@@
 
-The orbit process \(\{\bs A(t)\}\) is not required to model the behaviour at the regulated boundary.
+The orbit process \(\{\bs A(t)\}\) is not required to model the behaviour at the sticky boundary.
 However, we suppose that \(\bs A(t)=1\) at the boundaries.
@@
 The level of the fluid at the same time is \(X(t) = y_\ell + c_i t\).
-We argued earlier that we can think of the orbit \(\cfrac{\bs \alpha e^{\bs{S}|c_i|t}}{\bs \alpha e^{\bs{S}|c_i|t}\bs e}\) as corresponding to the fluid being a distance of \(|c_i|t\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\).
-With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\).
-Similarly, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\cfrac{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}}{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}\bs e}.\)
-
-We use the notation \(\bs a_{\ell,i}(x_0) = \cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\).
-Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \cfrac{\bs \alpha e^{\bs{S}(y_{\ell+1} - x_0)}}{\bs \alpha e^{\bs{S}(y_{\ell+1}-x_0)}\bs e}\).
+We argued earlier that we can think of the orbit \(\bs k(x)\) as corresponding to the fluid being a distance of \(x\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\).
+With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\bs k(x_0-y_{\ell})\).
+Similarly, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\bs k(y_{\ell+1}-x_0).\)
+
+We use the notation \(\bs a_{\ell,i}(x_0) = \bs k(x_0-y_\ell)\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\).
+Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \bs k(y_{\ell+1}-x_0)\).
 
@@
 \end{align}
-This may be of interest in its own right. However, it is a somewhat granular approximation -- the level process tells us nothing about where in the intervals \(\calD_{k,i}\) the fluid queue might be.
+This may be of interest in its own right.
+However, for some applications, this approximation may be too coarse -- the level process tells us nothing about where in the intervals \(\calD_{k,i}\) the fluid queue might be.
 
 We now describe how the orbit position \(\bs A(t)\) can be used to approximate the density of the level of the fluid queue \(X(t)\).
+This reconstruction is entirely post-hoc in the sense that it does not affect the dymaics of the QBD-RAP itself.
+For any time \(t\) we only take the information contained in \(\mathbb E[A(t)1(L(t)=\ell,\varphi(t)=i)]\) and rewrite it in a descriptive way.
 
@@
 \begin{align}
-		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
-		\\&\qquad\times\bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x,\label{eqn: density approx aug}
-		\intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
-		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
-		\\&\qquad\times\bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
-\end{align}
-for \( j\in\mathcal S_-\cup\calS_{-0}\), where \(x\in\mathcal D_{\ell,j}\), \(x_0\in\mathcal D_{\ell_0,i}\).
+		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i)  
+		\bs a e^{\bs{S}z} \bs s \wrt x,\label{eqn: density approx aug}
+		% \intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
+		% &\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
+		% \\&\qquad\times\bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
+\end{align}
+where \(x\in\mathcal D_{\ell,j}\), \(x_0\in\mathcal D_{\ell_0,i}\) and 
+\[
+	z = \begin{cases} 
+		x - y_\ell & i\in\calS_- \\
+		y_{\ell+1} - x & i\in\calS_+.
+	\end{cases}
+\]
 %For example, for the augmented scheme in Section~\ref{subsec: augmented }, we approximate \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i)\) by 
@@
 
-While the estimates in the right-hand side of (\ref{eqn: density approx aug}) and (\ref{eqn: density approx aug2}) are appealing, they are, however, defective estimates -- they do not integrate to 1 for any finite-dimensional matrix exponential distribution.
+While the estimate (\ref{eqn: density approx aug}) is appealing, it is, however, defective -- it does not integrate to 1 for any finite-dimensional matrix exponential distribution.
 To see this, compute
 \begin{align*}
-	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_+\cup \calS_{+0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
-	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}(y_{\ell+1}-x)} \bs s \wrt x
-	\\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0,
-	\\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
-	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_+\cup \calS_{+0}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
+	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
+	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}z} \bs s \wrt z
+	% \\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0,
+	% \\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
+	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
 	\\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
-	\\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
-	\\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta} \bs e )
 	\\&<1,
@@
 \[\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}\Delta} \bs e \wrt x.\]
+The problem is that the triple \((\bs a,\bs S, \bs s)\) defines a distribution on \(x\in[0,\infty)\), however, it only makes sense to take \(x\in[0,\Delta)\) in the approximation scheme.
 
@@
 	\label{eqn: density approx 3+}&
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_+^\ell(x) \wrt x ,
+		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_i^\ell(x) \wrt x ,
 \end{align}
@@
 \begin{align}
-		&\bs a u_+^\ell(x) = \bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right),
-\end{align}
-{for \(j\in\mathcal S_+\cup \calS_{+0}\), and }
-\begin{align}
-	\label{eqn: density approx 3-}& 
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell_0,i}(x_0), \phi(0) = i) \bs a u_-^\ell(x) \wrt x ,
-\end{align}
-{where}
-\begin{align}
-		&\bs a u_-^\ell(x) = \bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right),
-\end{align}
-for \(j\in\mathcal S_-\cup \calS_{-0}\).
-
+		\bs a u_i^\ell(x) = \begin{cases}
+			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right), & i\in\calS_+, \\ 
+			\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right), & i\in\calS_-. \\ 
+		\end{cases}\label{eqn: density approx 3+a}
+\end{align}
 Intuitively, we take the density function \(\bs ae^{\bs{S}x}\bs s\) on \(x \in [0,2\Delta)\), and `fold' it back on itself around \(\Delta\), to create a density function on \([0,\Delta)\), 
@@
 \[\bs a e^{\bs S 2\Delta}\bs e,\]
-which is less than the quantity \(\bs a e^{\bs S \Delta}\bs e\) from the approximation scheme (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}).
+which is less than the quantity \(\bs a e^{\bs S \Delta}\bs e\) from the approximation scheme (\ref{eqn: density approx aug}).
+There is nothing special about the choice of truncating the distribution at \(2\Delta\) in this construction, and we could choose any truncation value greater than \(\Delta\).
+
+A third option is to approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
+\begin{align}
+%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
+	\label{eqn: density approx 4+}&
+		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a v_i^\ell(x) \wrt x ,
+\end{align}
+{where}
+\begin{align}
+		\bs a v_i^\ell(x) = \begin{cases}
+			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \right)\left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_+, \\ 
+			\bs a \left(e^{ \bs{S}(x-y_\ell)}+e^{ \bs{S}(2\Delta-(x-y_\ell))}\right) \left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_-. \\ 
+		\end{cases}
+\end{align}
+Here, we take the density function \(\bs a e^{\bs S x}\bs s\), defined on \([0,\infty)\), and map it to a density function on \([0,\Delta)\) by \(\Delta - |(x \mod 2\Delta )-\Delta|\).
+The resulting density function is
+\begin{align*}
+	v(x) &= \sum_{m=0}^\infty \bs a \left(e^{\bs S (x+2\Delta m)} + e^{\bs S (2\Delta -x+2\Delta m)} \right)\bs s
+	%
+	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\sum_{m=0}^\infty e^{\bs S 2\Delta m} \bs s
+	%
+	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\left[I-e^{S2\Delta}\right]^{-1} \bs s.
+\end{align*}
 
@@
 We generalise this idea with the concept of \emph{closing operators} which is a linear operator \(V(x):\mathcal A \to \mathbb R,\) for each \(x\in[0,\Delta)\).
-For example, the closing operator in Equations (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}) is the operator \(V(x),\,x\in[0,\Delta)\) such that for any \(\bs a \in \mathcal A\),
+For example, the closing operator in Equations (\ref{eqn: density approx aug}) is the operator \(V(x),\,x\in[0,\Delta)\) such that for any \(\bs a \in \mathcal A\),
 \begin{align}
@@
 \end{align}
-Similarly, in (\ref{eqn: density approx 3+}) and (\ref{eqn: density approx 3-}) 
-\begin{align}
-	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right).\label{eqn:546}
+Similarly, in (\ref{eqn: density approx 3+})
+\begin{align}
+	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right),\label{eqn:546}
+\end{align}
+and in (\ref{eqn: density approx 4+})
+\begin{align}
+	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)}\left[I-e^{S2\Delta}\right]^{-1} \bs s\right).\label{eqn:5462}
 \end{align}
@@
 
-In (\ref{eqn: alal})-(\ref{eqn:546}), since \(V\) is linear, then we actually have 
-\begin{align}
-	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a V(x) \nonumber 
-	\\&= \mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i]V(x).
-\end{align}
-
-\begin{rem}
+% \begin{rem}
+% Since we require \(V\) to be linear, then we actually have 
+% \begin{align}
+% 	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a V(x) \nonumber 
+% 	\\&= \mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i]V(x).
+% \end{align}
+% \end{rem}
+
+% \begin{rem}
+	\paragraph{Comments on linearity and normalisation}
 	We considered taking the closing operators to be 
@@
 	amounts to a matrix-exponential calculation.
-\end{rem}
-
-In practice, to ensure that the approximation integrates to \(1\), we do re-normalise the estimates of the density. Specifically, we use the estimates
-\begin{align}
-	&\cfrac{\mathbb E[\bs A(t) 1(L(t)=\ell,\phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i]V(x)}{\int_{x\in[0,\Delta)}\mathbb E[\bs A(t) 1(L(t)=\ell,\phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i]V(x)\wrt x}.
-\end{align}
+
+	In practice we use 
+	\begin{align}
+		\cfrac{\mathbb E\left[\bs A(t)\right] \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\mathbb E\left[\bs A(t)\right] e^{\bs S 2\Delta}\bs e}
+	\end{align}
+	as a normalised version of (\ref{eqn:546}), and this seems to work well. However, this is not the same as using the closing operator (\ref{eqn: nonlin}).
+% \end{rem}
+
+% \begin{rem}
+% 	Consider approximating a function \(f(x)\) by a linear combination of the elements of \(e^{\bs S x}\bs s\), i.e.
+% 	\begin{align}
+% 		f(x) \approx \bs b e^{\bs S x} \bs s.
+% 	\end{align}
+% 	Further, suppose that we require the residual \(f(x) - \bs b e^{\bs S x} \bs s\) to be orthogonal to the space of functions spanned by the elements of 
+% 	\begin{align}
+% 		\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}.
+% 	\end{align}
+% 	This amounts to the requirement  
+% 	\begin{align}
+% 		\int_{0}^\infty \left(f(x) - \bs b e^{\bs S x} \bs s\right)\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}\wrt x = \bs 0%\int \bs e^{\bs Sx}\bs s \cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e} \bs c \wrt x
+% 	\end{align}
+% 	which implies
+% 	\begin{align}
+% 		\bs b \int_{0}^\infty e^{\bs S x} \bs s\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}\wrt x = \int_{0}^\infty f(x)\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}\wrt x
+% 	\end{align}
+% 	and therefore 
+% 	\begin{align}
+% 		\bs b = \int_{0}^\infty f(x)\bs k(x)\wrt x \bs D^{-1}.
+% 	\end{align}
+
+% 	This suggests that we may be able to reconstruct solutions as 
+% 	\[\mathbb E[\bs A(t)1(L(t)=\ell, \varphi(t)=j)]\bs D^{-1} e^{\bs S x}\bs s.\] 
+% 	Some basic numerical experiments suggest this might work and could be investigated as further research.
+Since this reconstruction does not have a stochastic interpretation, and does not enforce positivity, we do not persue it further here.
+% \end{rem}

### 2021-11-11 chapter2/chapter.tex
move aD initial condition to appendix, simplify presentation.

@@
 
-We use the notation \(\bs a_{\ell,i}(x_0) = \bs k(x_0-y_\ell)\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\).
-Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \bs k(y_{\ell+1}-x_0)\).
+We use the notation \(\bs a_{\ell,i}(x_0) = \bs a_{\ell,+}(x_0) = \bs k(x_0-y_\ell)\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\).
+Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \bs a_{\ell,-}(x_0) = \bs k(y_{\ell+1}-x_0)\).
 
 More generally, given an initial measure \(\mu_i(\cdot):= \mathbb P(X(0)\in\cdot,\varphi(0)=i)\), an approximation to this initial condition is to set the orbit to \(\displaystyle\int_{x\in\calD_{\ell,i}}\bs a_{\ell,i} (x) \wrt \mu_i\), for \(i\in\mathcal S,\, \ell\in\mathcal K\setminus\{-1,K+1\}\).
+
+\paragraph{Initial conditions for phases \(i\in\calS_0\)}
+While the augmented state-space scheme described above is a convergent one, the analysis of the QBD-RAP scheme is \emph{greatly} simplified if we allow initial conditions which allocate \(0\) probability to phases in \(i\in\calS_{+0}\cup\calS_{-0}\) only.
+To see why initial mass in \(\mathcal S_{+0}\cup\calS_{-0}\) might introduce further complexity, consider a sample path with \(\varphi(0)=k\in\calS_{+0}\) and \(\bs A(0)=\bs a\in\mathcal A\).
+On \(\varphi(u)\in\calS_{+0},\) for all \(u\in[0,t)\), then \(\bs A(u) = \bs a \) is constant.
+Upon the phase leaving \(\calS_{+0}\), at time \(t\), say, the phase will either jump to \(\calS_+\) or \(\calS_-\).
+If \(\varphi(t)\) jumps to \(i\in\calS_+\), the orbit process at time \(t\) will be \(\bs A(t)=\bs a \) and the process evolve from time \(t\) as if it has started in phase \(i\in\calS_+\) with initial orbit \(\bs a \) -- there are no major issues in this case.
+If, however, \(\varphi(t)\) jumps to \(i\in\calS_-\), then the orbit process will jump to \(\bs A(t) = \bs a \bs D\) at time \(t\).
+From time \(t\) onward the QBD-RAP process will evolve as if it has started in phase \(i\in\calS_-\) with initial orbit \(\bs a \bs D\).
+With the way we have structured the proof of convergence, we then have to complete two analyses of the QBD-RAP process -- one for the initial condition \(\bs a\) and another for the initial condition \(\bs a\bs D\).
+
+A solution to this problem is to augment a set of ephemeral states to the QBD-RAP which capture the initial sojourn in \(\calS_0\) only.
+Let us denote this set of ephemeral states \(\calS_0^*\) which has the same phase dynamics as \(\calS_0\).
+Suppose we wish to approximate a fluid queue with the initial condition \(\mu_k\) where \(\mu_k\) has mass 1.
+At time \(0\), we start the QBD-RAP in the ephemeral set of states \(\calS_0^*\) in phase \(k^*\in\calS_0^*\).
+While in \(\calS_0^*\), the phase process evolves according to the generator \(\bs T_{00}\).
+Upon exiting the ephemeral set \(\calS_0^*\) at time \(t\), say, the QBD-RAP process jumps to \(L(t)=\ell,\, \bs A(t)=\int_{\calD_{\ell,i}}\bs a_{\ell,i}(x)\wrt \mu_k,\,\varphi(t)=i\), where \(\ell\in\mathcal K\), \(i\in\calS_+\cup\calS_-\).
+
+Notice that we have not mentioned anything about the level or orbit process during the sojourn in \(\calS_0^*\).
+This is because all information about the level and phase during this time is contained within the initial condition \(\mu_k\), and the only thing that changes is the phase.
+
+To include the ephemeral states within the generator of the QBD-RAP we can do the following, 
+\begin{align*}
+	&\left[\begin{array}{c|cccc}
+		\bs T_{00} & \bs B_{-1}^* &\bs B_{0}^*&\bs B_{1}^* & \hdots \\\hline
+		&\check{\bs B}_{0} &  \check{\bs B}_{+1} &&\\
+		& \check{\bs B}_{-1}&\check{\bs B}_{0}&\bs B_{+1}& \\
+		&&\bs B_{-1} & \bs B_0&\ddots \\
+		&&&\ddots&\ddots 
+	\end{array}\right],
+\end{align*}
+where 
+\begin{align*}
+	\bs B_{-1}^* &= \vligne{\bs T_{0-}\mu_k(\{0\}) & \bs 0} & \bs B_{k}^* &= \vligne{\bs T_{0+}\otimes \bs a_k^+ & \bs 0 & \bs T_{0-}\otimes \bs a_k^- & \bs 0},
+	%
+	\\\check{\bs B}_{0} &= \left[\begin{array}{cc}\bs T_{--} & \bs T_{-0}\\\bs T_{0-} & \bs T_{00}\end{array}\right], 
+	& \check{\bs B}_{+1} &= \left[\begin{array}{cccc}\bs T_{-+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \\ \bs T_{0+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \end{array}\right],
+	%
+	\\\check{\bs B}_{-1} &= \left[\begin{array}{cc} \bs0&\bs0\\\bs0&\bs0\\(\bs C_-\bs P_{--})\otimes \bs s & (\bs C_-\bs P_{-0})\otimes \bs s \\\bs0&\bs0\end{array}\right], &&
+\end{align*}
+and 
+\begin{align*}
+	\bs a_\ell^r = \int_{\calD_{\ell,r}} \bs a_{\ell,r}(x)\wrt \mu_k,\, \ell\in\mathcal K\setminus\{-1,K+1\},\, r\in\{+,-\}. 
+\end{align*}
 

### 2021-11-15 chapter1/chapter.tex
update hhat reference, add me references
fix more of Nige's comments -- up to page 25. reorder PH-case section and rewrite. add requirements on D, most importanly linearity
nigel changes ch 3 next set 1, update QBDRAP and fluid queue to use vector notation, introduce notation for tau_n<t<=tau_n+1

@@
 \chapter{Introduction \label{ch:intro}} 
-A fluid queue is a two-dimensional stochastic process \(\{(X(t),\varphi(t))\}_{t\geq0}\).
+A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\).
 The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC).
@@
 \section{Fluid queues}
-A fluid queue is a two-dimensional stochastic process \(\{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer.
+A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer.
 The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we we assume to be \(\mathcal S=\{1,2,\dots,N\}\) without loss of generality, and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\).
@@
 Upon on an event at time \(t\) with mark \(i\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D_i/\bs A(t^-) \bs D_i\bs e\).
+It is important to note that the jumps of the orbit process are \emph{linear} transformations of the orbit process immediately before the time of the jump.
 

### 2021-11-15 chapter2/chapter.tex
nigel changes ch 3 next set 1, update QBDRAP and fluid queue to use vector notation, introduce notation for tau_n<t<=tau_n+1

@@
 Since the phase dynamics are a CTMC, we choose to use an alternate notation to the standard presented in Section \ref{sec: qbd-rap}.
-Let \(\{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) be the QBD-RAP approximation of a fluid queue, where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process.
+Let \(\{\bs Y(t)\}_{t\geq0} = \{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) be the QBD-RAP approximation of a fluid queue, where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process.
 We use this representation to make explicit how the approximation captures the phase dynamics of the fluid queue.
+Furthermore, define the notation \(\bs Y_d(t) = (L(t),\phi(t))\) for the discrete elements of \(\bs Y(t)\).
+
 We will show later that \(\phi(t)\) captures the phase dynamics of the fluid queue exactly, provided the phase process is independent of the fluid level \(\{X(t)\}\).
@@
 We argued earlier that we can think of the orbit \(\bs k(x)\) as corresponding to the fluid being a distance of \(x\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\).
-With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\bs k(x_0-y_{\ell})\).
-Similarly, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\bs k(y_{\ell+1}-x_0).\)
+With this interpretation, an approximation to the initial condition \(\bs X(0) = (X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\bs k(x_0-y_{\ell})\).
+Similarly, an approximation to the initial condition \(\bs X(0)=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\bs k(y_{\ell+1}-x_0).\)
 
@@
 A solution to this problem is to augment a set of ephemeral states to the QBD-RAP which capture the initial sojourn in \(\calS_0\) only.
-Let us denote this set of ephemeral states \(\calS_0^*\) which has the same phase dynamics as \(\calS_0\).
 Suppose we wish to approximate a fluid queue with the initial condition \(\mu_k\) where \(\mu_k\) has mass 1.
-At time \(0\), we start the QBD-RAP in the ephemeral set of states \(\calS_0^*\) in phase \(k^*\in\calS_0^*\).
-While in \(\calS_0^*\), the phase process evolves according to the generator \(\bs T_{00}\).
-Upon exiting the ephemeral set \(\calS_0^*\) at time \(t\), say, the QBD-RAP process jumps to \(L(t)=\ell,\, \bs A(t)=\int_{\calD_{\ell,i}}\bs a_{\ell,i}(x)\wrt \mu_k,\,\varphi(t)=i\), where \(\ell\in\mathcal K\), \(i\in\calS_+\cup\calS_-\).
-
-Notice that we have not mentioned anything about the level or orbit process during the sojourn in \(\calS_0^*\).
+Let us denote the set of ephemeral states by \(\calS_0^{*,k}\) and each state in \(\calS_0^{*,k}\) by \(i^*\).
+Each state \(i^*\in\calS_0^{*,k}\) is a copy of \(i\in\calS_0\), and, in \(\calS_0^{*,k}\) the phase process has the same phase dynamics as it does in \(\calS_0\).
+At time \(0\), we start the QBD-RAP in the ephemeral set of states \(\calS_0^{*,k}\) in phase \(k^*\in\calS_0^{*,k}\).
+While in \(\calS_0^{*,k}\), the phase process evolves according to the generator \(\bs T_{00}\).
+Upon exiting the ephemeral set \(\calS_0^{*,k}\) at time \(t\), say, the QBD-RAP process jumps to \(L(t)=\ell,\, \bs A(t)=\int_{\calD_{\ell,i}}\bs a_{\ell,i}(x)\wrt \mu_k,\,\varphi(t)=i\), where \(\ell\in\mathcal K\), \(i\in\calS_+\cup\calS_-\).
+
+Notice that we have not mentioned anything about the level or orbit process during the sojourn in \(\calS_0^{*,k}\).
 This is because all information about the level and phase during this time is contained within the initial condition \(\mu_k\), and the only thing that changes is the phase.
@@
 	&\left[\begin{array}{c|cccc}
-		\bs T_{00} & \bs B_{-1}^* &\bs B_{0}^*&\bs B_{1}^* & \hdots \\\hline
+		\bs T_{00} & \bs B_{-1}^{*,k} &\bs B_{0}^{*,k}&\bs B_{1}^{*,k} & \hdots \\\hline
 		&\check{\bs B}_{0} &  \check{\bs B}_{+1} &&\\
@@
 \begin{align*}
-	\bs B_{-1}^* &= \vligne{\bs T_{0-}\mu_k(\{0\}) & \bs 0} & \bs B_{k}^* &= \vligne{\bs T_{0+}\otimes \bs a_k^+ & \bs 0 & \bs T_{0-}\otimes \bs a_k^- & \bs 0},
+	\bs B_{-1}^{*,k} &= \vligne{\bs T_{0-}\mu_k(\{0\}) & \bs 0} & \bs B_{\ell}^{*,k} &= \vligne{\bs T_{0+}\otimes \bs a_\ell^+ & \bs 0 & \bs T_{0-}\otimes \bs a_\ell^- & \bs 0},
 	%
@@
 \end{align*}
+
+In general, we need to augment a set of states \(\calS_0^{*,k}\) to the QBD-RAP for each initial phase, \(k\in\calS_0\) such that \(\mu_k\) has positive mass.
 
@@
 
-This reasoning leads to an approximation of the distribution of the fluid at time \(t\), \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \), as
-\begin{align}
-		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i)  
-		\bs a e^{\bs{S}z} \bs s \wrt x,\label{eqn: density approx aug}
+This reasoning leads to an approximation of the distribution of the fluid at time \(t\), \(\mathbb P(\bs X(t)\in (\wrt x,j) \mid \bs X(0)=(x_0,i)) \), as
+\begin{align}
+		&\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a, j)\mid \bs Y(0)=(\ell_0, \bs a_{\ell_0,i}(x_0), i) )
+		\bs a e^{\bs{S}z} \bs s \wrt x,
+		%
+		% \\&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i)  
+		% \bs a e^{\bs{S}z} \bs s \wrt x,
+		\label{eqn: density approx aug}
 		% \intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
@@
 	z = \begin{cases} 
-		x - y_\ell & i\in\calS_- \\
+		x - y_\ell & i\in\calS_-, \\
 		y_{\ell+1} - x & i\in\calS_+.
@@
 \begin{align*}
-	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
-	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}z} \bs s \wrt z
+	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a, j)\mid \bs Y(0)=(\ell_0, \bs  a_{\ell,i}(x_0), i)) \bs a e^{ \bs{S}z} \bs s \wrt z
 	% \\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0,
 	% \\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
-	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
-	\\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
+	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a, j)\mid \bs Y(\ell_0, \bs  a_{\ell,i}(x_0), i)) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
 	\\&<1,
+	% &\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
+	% \\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}z} \bs s \wrt z
+	% % \\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0,
+	% % \\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
+	% \\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
+	% \\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
+	% \\&<1,
 \end{align*}
 since \(\bs a e^{\bs{S}u} \bs e>0\) for any \(u>0\). For an orbit position \(\bs A(t)=\bs a\), the amount of mass missing from each level and phase is 
-\[\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}\Delta} \bs e \wrt x.\]
+\[\mathbb P(\bs Y(t)\in (\ell, \wrt \bs a,j)\mid \bs Y(0)=(\ell_0,  \bs  a_{\ell,i}(x_0), i)) \bs a e^{\bs{S}\Delta} \bs e \wrt x.\]
 The problem is that the triple \((\bs a,\bs S, \bs s)\) defines a distribution on \(x\in[0,\infty)\), however, it only makes sense to take \(x\in[0,\Delta)\) in the approximation scheme.
@@
 
-To partially rectify this we can instead approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
+To partially rectify this we can instead approximate \(\mathbb P(\bs X(t)\in (\wrt x, j)\mid \bs X(0)=(x_0, i)) \) by
 \begin{align}
@@
 	\label{eqn: density approx 3+}&
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_i^\ell(x) \wrt x ,
+	\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in(\ell, \wrt \bs a, j)\mid \bs Y(0)=()\ell_0, \bs  a_{\ell,i}(x_0), i)) \bs a u_i^\ell(x) \wrt x ,
+	% \displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_i^\ell(x) \wrt x ,
 \end{align}
@@
 
-A third option is to approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
+A third option is to approximate \(\mathbb P(\bs X(t)\in (\wrt x, j)\mid \bs X(0)=(x_0, i)) \) by
 \begin{align}
@@
 	\label{eqn: density approx 4+}&
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a v_i^\ell(x) \wrt x ,
+		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(\bs Y(t) \in (\ell, \wrt \bs a,j)\mid \bs Y(0) = (\ell_0, \bs  a_{\ell,i}(x_0), i)) \bs a v_i^\ell(x) \wrt x ,
 \end{align}

### 2021-11-22 chapter2/chapter.tex
proof of properties for the closing operator for the normalised version

@@
 That is, the mean of the Phase-type \(PH(\bs \alpha,\bs{S})\) is equal to the cell width, \(\Delta\).
-The distribution \(PH(\bs \alpha, T_{ii}\bs{I}+\bs{S}_i)\) can be viewed as \(min(U,V)\) where \(U\sim exp(T_{ii})\) and \( V\sim PH(\bs\alpha, \bs{S})\).
-Thus, expiry of the distribution \(\tau_i\) corresponds to either a change in phase if \(U\) is the minimum, or change in level if \(V\) is the minimum.
+The distribution \(PH(\bs \alpha, T_{ii}\bs{I}+\bs{S}_i)\) can be viewed as \(min(U,{\bs v})\) where \(U\sim exp(T_{ii})\) and \( {\bs v}\sim PH(\bs\alpha, \bs{S})\).
+Thus, expiry of the distribution \(\tau_i\) corresponds to either a change in phase if \(U\) is the minimum, or change in level if \({\bs v}\) is the minimum.
 The latter case we map to a change in level: \(L(t)\) moves from \(\ell\) to \(\ell+1\) if \(c_i>0\), and from \(\ell\) to \(\ell-1\) if \(c_i<0\).
@@
 Asmussen and Bladt \citep{ab1999} gave such a process a stochastic interpretation (in fact, Asmussen and Bladt worked with rational arrival processes, but I think we need only consider ME renewal processes).
-Let \(V = span\left\{\alpha, \bs\alpha \bs{S},...,\bs\alpha \bs{S}^{p-1}\right\}\).
-There exists a vector-valued stochastic process \(\left\{\bs A(t)\right\}_{t}\) with \(\bs A(t)\in V\) for all \(t\) and \(\bs A(0)=\bs\alpha\).
+Let \({\bs v} = span\left\{\alpha, \bs\alpha \bs{S},...,\bs\alpha \bs{S}^{p-1}\right\}\).
+There exists a vector-valued stochastic process \(\left\{\bs A(t)\right\}_{t}\) with \(\bs A(t)\in {\bs v}\) for all \(t\) and \(\bs A(0)=\bs\alpha\).
 Suppose \(\{Y_k\}\), \(k=0,1,...\) are the renewal times, and \(Y_0=0\).
@@
 Assume that \(\pi_i\neq0\), \(i=1,...,p\).
-If this is not the case then construct an invertible matrix \(\bs{V}\) such that \(\bs{V}\bs 1=\bs 1\), then consider the equivalent representation of the ME renewal process with \(ME(\bs\alpha \bs{V}^{-1}, \bs{V}\bs{C}\bs{V}^{-1})\), and pray that \(\pi_i\neq0\), \(i=1,...,p\) (can we show that there is always such a \(\bs{V}\)?) By choosing \(\bs{V}\) such that \(\bs{V}\bs 1=\bs 1\), then we maintain that the right eigenvector corresponding to \(\bs{V}\bs{C}\bs{V}^{-1}\) is still \(\bs 1\).
+If this is not the case then construct an invertible matrix \(\bs{{\bs v}}\) such that \(\bs{{\bs v}}\bs 1=\bs 1\), then consider the equivalent representation of the ME renewal process with \(ME(\bs\alpha \bs{{\bs v}}^{-1}, \bs{{\bs v}}\bs{C}\bs{{\bs v}}^{-1})\), and pray that \(\pi_i\neq0\), \(i=1,...,p\) (can we show that there is always such a \(\bs{{\bs v}}\)?) By choosing \(\bs{{\bs v}}\) such that \(\bs{{\bs v}}\bs 1=\bs 1\), then we maintain that the right eigenvector corresponding to \(\bs{{\bs v}}\bs{C}\bs{{\bs v}}^{-1}\) is still \(\bs 1\).
 %
@@
 The difference between the approximation schemes is the way in which the vector \(\bs a\) is used.
-We generalise this idea with the concept of \emph{closing operators} which is a linear operator \(V(x):\mathcal A \to \mathbb R,\) for each \(x\in[0,\Delta)\).
-For example, the closing operator in Equations (\ref{eqn: density approx aug}) is the operator \(V(x),\,x\in[0,\Delta)\) such that for any \(\bs a \in \mathcal A\),
-\begin{align}
-	\bs a V(x) &= \bs a e^{\bs Sx}\bs s.\label{eqn: alal}
-	%\\\bs a V_-(x-y_\ell) &= \bs a e^{\bs S(x-y_{\ell})}\bs s,\label{eqn:ndaenf}
+We generalise this idea with the concept of \emph{closing operators} which is a linear operator \({\bs v}(x):\mathcal A \to \mathbb R,\) for each \(x\in[0,\Delta)\).
+For example, the closing operator in Equations (\ref{eqn: density approx aug}) is the operator \({\bs v}(x),\,x\in[0,\Delta)\) such that for any \(\bs a \in \mathcal A\),
+\begin{align}
+	\bs a {\bs v}(x) &= \bs a e^{\bs Sx}\bs s.\label{eqn: alal}
+	%\\\bs a {\bs v}_-(x-y_\ell) &= \bs a e^{\bs S(x-y_{\ell})}\bs s,\label{eqn:ndaenf}
 \end{align}
@@
 \begin{align}
-	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right),\label{eqn:546}
+	\bs a {\bs v}(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right),\label{eqn:546}
 \end{align}
@@
 \begin{align}
-	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)}\left[I-e^{S2\Delta}\right]^{-1} \bs s\right).\label{eqn:5462}
-\end{align}
-The operator \(V(x)\) is a choice which forms part of the definition of the approximation scheme.
-As we shall see, given certain properties of \(V(x)\), we can prove that the approximation scheme converges, and ensures positivity.
+	\bs a {\bs v}(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)}\right)\left[I-e^{S2\Delta}\right]^{-1} \bs s.\label{eqn:5462}
+\end{align}
+The operator \({\bs v}(x)\) is a choice which forms part of the definition of the approximation scheme.
+As we shall see, given certain properties of \({\bs v}(x)\), we can prove that the approximation scheme converges, and ensures positivity.
 In the cases above, all the closing operators (\ref{eqn: alal})-(\ref{eqn:546}) lead to an approximation which converges and, due to their interpretation as probability densities, ensure positivity.
@@
 % \begin{rem}
-% Since we require \(V\) to be linear, then we actually have 
+% Since we require \({\bs v}\) to be linear, then we actually have 
 % \begin{align}
-% 	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a V(x) \nonumber 
-% 	\\&= \mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i]V(x).
+% 	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a {\bs v}(x) \nonumber 
+% 	\\&= \mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i]{\bs v}(x).
 % \end{align}
@@
 	\begin{align}
-		\bs a V(x) &= \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e},\label{eqn: nonlin}
+		\bs a {\bs v}(x) &= \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e},\label{eqn: nonlin}
 	\end{align}
@@
 The importance of this fact is that for a linear operator,
-	\[\mathbb E[\bs A(t) V(x)] = \mathbb E[\bs A(t)]V(x),\]
-	thus, the computation of \(\mathbb E[\bs A(t) V(x)]\) can be achieved by first computing \(\mathbb E[\bs A(t)]\) and then applying \(V(x)\) to the result.
-In this sense, \(\mathbb E[\bs A(t)]\) contains all of the information about the history of the process up to time \(t\) needed to compute \(\mathbb E[\bs A(t) V(x)]\).
+	\[\mathbb E[\bs A(t) {\bs v}(x)] = \mathbb E[\bs A(t)]{\bs v}(x),\]
+	thus, the computation of \(\mathbb E[\bs A(t) {\bs v}(x)]\) can be achieved by first computing \(\mathbb E[\bs A(t)]\) and then applying \({\bs v}(x)\) to the result.
+In this sense, \(\mathbb E[\bs A(t)]\) contains all of the information about the history of the process up to time \(t\) needed to compute \(\mathbb E[\bs A(t) {\bs v}(x)]\).
 This is much the same as in RAPs and QBD-RAPs where \(\mathbb E[\bs A(t)]\) is all that is required to compute probabilities about the future of the processes from time \(t\) onwards.
 
-	When \(V(x)\) is not linear then, in general, 
-	\[\mathbb E[\bs A(t) V(x)] \neq \mathbb E[\bs A(t)]V(x).\]
+	When \({\bs v}(x)\) is not linear then, in general, 
+	\[\mathbb E[\bs A(t) {\bs v}(x)] \neq \mathbb E[\bs A(t)]{\bs v}(x).\]
 	For example, with (\ref{eqn: nonlin}), 
 	\begin{align}
-		\mathbb E[\bs A(t) V(x)] = \int_{\bs a \in\mathcal A} \mathbb P(\bs A(t) \in \wrt \bs a) \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e}.\label{eqn:nonlin2}
+		\mathbb E[\bs A(t) {\bs v}(x)] = \int_{\bs a \in\mathcal A} \mathbb P(\bs A(t) \in \wrt \bs a) \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e}.\label{eqn:nonlin2}
 	\end{align}
@@
 
-	In practice we use 
+	In practice we may use 
 	\begin{align}

### 2021-11-22 chapter1/chapter.tex
fix typo

@@
 
-Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{\i\in\calS\mid c_i<0\}\) and \(\calS_0 = \{i\in\calS\mid c_i=0\}\).
+Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{i\in\calS\mid c_i<0\}\) and \(\calS_0 = \{i\in\calS\mid c_i=0\}\).
 We assume, without loss of generality, that the generator \(\bs T\) is partitioned into sub-matrices

### 2021-11-24 appendix3/appendix.tex
story-ify appendix A

%!TEX root = ../thesis.tex
\chapter{Convergence without ephemeral phases\label{app:extend conv}}
For completeness, we include here, the results needed to prove the convergence of the QBD-RAP scheme without the need for the ephemeral initial states.

For \(\varphi(0)=k\in\mathcal S_{-0}\) (or \(\varphi(0)=k\in\mathcal S_{+0}\)) the added complexity comes from the fact that it is possible that, upon the phase process first leaving \(\mathcal S_{-0}\) (\(\mathcal S_{+0}\)) the phase transitions to a state in \(\mathcal S_+\) (\(\mathcal S_-\)). Since the orbit of the QBD-RAP is constant on \(\varphi(t)\in\mathcal S_{-0}\) (\(\varphi(t)\in\mathcal S_{+0}\)), then upon a first transition out of \(\mathcal S_{-0}\) (\(\mathcal S_{+0}\)) and into \(\mathcal S_+\) (\(\mathcal S_-\)) the orbit jumps to \(\bs   a_{\ell_0,i}^{(p)}(x_0)\bs D^{(p)}\). For \(k\in\mathcal S_{-0}\), \(m\geq 0\), the corresponding Laplace transform of the QBD-RAP is
\begin{align}
	&\widehat{f}_{m,-0,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\nonumber 
	\\&:=\int_{x_1=0}^\infty \dots \int_{x_{2m+1}=0}^\infty  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0+}\bs M^m_{++}(\lambda,x_1,\dots,x_{2m+1})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0) \bs D^{(p)} \bs N^{2m+1,(p)}(\lambda,x_1,\dots,x_{2m+1}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+1}\dots\wrt x_2 \wrt x_1  \nonumber
	\\& + \int_{x_1=0}^\infty \dots \int_{x_{2m+2}=0}^\infty  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0-}\bs M^{m+1}_{-+}(\lambda,x_1,\dots,x_{2m+2})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0)\bs N^{2m+2,(p)}(\lambda,x_1,\dots,x_{2m+2}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+2}\dots\wrt x_2 \wrt x_1.
	\label{eqn: k0jJ0}
\end{align}
% The second term in expression (\ref{eqn: k0jJ0}) is a linear combination of \(\widehat{f}_{m+1,-,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\). The first term is a linear combination of terms similar to \(\widehat{f}_{m,+,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\), except there is a different initial vector, \(\bs a_{\ell_0,i}^{(p)}(x_0)\bs D^{(p)}\).  
The Laplace transform of the fluid queue corresponding to (\ref{eqn: k0jJ0}) is 
\begin{align}
	\widehat \mu_{m,-0,+}^{\ell_0}(\lambda)(x,j;x_0,k) \nonumber 
	&:= \sum_{i\in\calS_+}\bs e_k\vligne{\lambda \bs I - \bs T_{00}}^{-1}\bs T_{0i}\widehat \mu_{m,+,+}^{\ell_0}(\lambda)(x,j;x_0,i) 
	\\&{}+ \sum_{i\in\calS_-}\bs e_k\vligne{\lambda \bs I - \bs T_{00}}^{-1}\bs T_{0i}\widehat \mu_{m+1,-,+}^{\ell_0}(\lambda)(x,j;x_0,i), \label{eqn: s0- actual}
\end{align}
\(m\geq 0\).

The second term of (\ref{eqn: k0jJ0}) is a linear combination of \(\widehat{f}_{m+1,-,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\) which converges to \(\widehat \mu_{m,-,+}^{\ell_0}(\lambda)(x,j;x_0,i)\), so there are no issues here. The first term of (\ref{eqn: k0jJ0}) creates significantly more work. Essentially, we need to repreat the results from Appendix~\ref{app:tand} but with the initial vector \(\bs a_{\ell_0,i}(x_0)\bs D\).

Analogously, for \(k\in\calS_{-0}\), \(m\geq 0\), we also have 
\begin{align*}
	\widehat{f}_{m,-0,-}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\nonumber 
	&:=\int_{x_1=0}^\infty \dots \int_{x_{2m+2}=0}^\infty  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0+} \bs M^{m+1}_{+-}(\lambda,x_1,\dots,x_{2m+2})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0) \bs D^{(p)} \bs N^{2m+2,(p)}(\lambda,x_1,\dots,x_{2m+2}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+2}\dots  \wrt x_1  \nonumber
	\\& + \int_{x_1=0}^\infty \dots \int_{x_{2m+1}=0}^\infty  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0-} \bs M^{m}_{--}(\lambda,x_1,\dots,x_{2m+1})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0)\bs N^{2m+1,(p)}(\lambda,x_1,\dots,x_{2m+1}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+1}\dots\wrt x_1.
\end{align*}
For \(k\in\calS_{+0}\), \(m\geq 0\), we have 
\begin{align*}
	\widehat{f}_{m,+0,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\nonumber 
	&:=\int_{x_1=0}^\infty \dots \int_{x_{2m+2}=0}^\infty \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0-} \bs M^{m+1}_{-+}(\lambda,x_1,\dots,x_{2m+2})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0) \bs D^{(p)} \bs N^{2m+2,(p)}(\lambda,x_1,\dots,x_{2m+2}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+2}\dots  \wrt x_1  \nonumber
	\\& + \int_{x_1=0}^\infty \dots \int_{x_{2m+1}=0}^\infty  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0+} \bs M^m_{++}(\lambda,x_1,\dots,x_{2m+1})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0)\bs N^{2m+1,(p)}(\lambda,x_1,\dots,x_{2m+1}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+1}\dots  \wrt x_1,
\end{align*}
and 
\begin{align*}
	\widehat{f}_{m,+0,-}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\nonumber 
	&:=\int_{x_1=0}^\infty \dots \int_{x_{2m+1}=0}^\infty  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0-} \bs M^m_{--}(\lambda,x_1,\dots,x_{2m+1})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0) \bs D^{(p)} \bs N^{2m+1,(p)}(\lambda,x_1,\dots,x_{2m+1}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+1}\dots  \wrt x_1  \nonumber
	\\& + \int_{x_1=0}^\infty \dots \int_{x_{2m+2}=0}^\infty  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0+} \bs M^{m+1}_{+-}(\lambda,x_1,\dots,x_{2m+2})\bs e_j \nonumber
	\\& \bs a_{\ell_0,k}^{(p)}(x_0)\bs N^{2m+2,(p)}(\lambda,x_1,\dots,x_{2m+2}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\wrt x_{2m+2}\dots  \wrt x_1.
\end{align*}
In general, for \(q\in \{+,-\}, \, q'\in\{+,-\}\), \(m\geq 0\),
\begin{align*}
	\widehat \mu_{m,q0,q'}^{\ell_0}(\lambda)(x,j;x_0,k) \nonumber 
	&:= \sum_{r\in\{+,-\}}\sum_{i\in\calS_r}\bs e_k\vligne{\lambda \bs I - \bs T_{00}}^{-1}\bs T_{0i}\widehat \mu_{m+1(r\neq q'),r,q'}^{\ell_0}(\lambda)(x,j;x_0,i).
\end{align*}

\begin{rem}
For technical reasons we should not have point masses at \(x_0\in\partial \calD_{\ell_0}\) when \(\varphi(0)\in\calS_{+0}\cup\calS_{-0}\). Intuitively, if \(\varphi(0)=k\in\calS_{+0}\) and \(x_0 = y_{\ell_0}\) then, if upon exiting \(\calS_{+0}\) the phase process transitions to \(\calS_-\) then the fluid queue will instantaneously leave the interval \(\calD_{\ell_0}\) upon this transition. On the same event, the orbit of the QBD-RAP will be \(\bs \alpha^{(p)} \bs D^{(p)}\) at the instant of the transition to \(\calS_-\). Roughly speaking \(\bs D^{(p)}\) maps \(\bs \alpha^{(p)}\) to approximately \(\cfrac{\bs \alpha^{(p)} e^{\bs S^{(p)}\Delta}}{\bs \alpha^{(p)} e^{\bs S^{(p)}\Delta}\bs e}\) (asymptotically). Our asymptotic arguments rely on Chebyshev's inequality, in the form of a bound in terms of the distance of the random variable \(Z^{(p)}\sim ME(\bs\alpha^{(p)},\bs S^{(p)})\) from its mean \(\Delta\). However, we cannot use such a technique to bound terms such as \(\cfrac{\bs \alpha^{(p)} e^{\bs S^{(p)}\Delta}}{\bs \alpha^{(p)} e^{\bs S^{(p)}\Delta}\bs e}e^{\bs S^{(p)}z}\bs s^{(p)}\). %Furthermore, for the QBD-RAP to capture the fact that the fluid queue instantaneously leaves \(\calD_{\ell_0}\) at the instant it the phase process transitions to \(\calS_-\), we need the intensity of a change of level of the QBD-RAP to become arbitrarily large in our limiting arguments. That is, we need \(\bs \alpha \bs D\bs s\) to become arbitrarily large. It is unknown, in general, whether that this is the case. %In these cases the initial condition \(\bs   a_{\ell_0,i}(x_0)\) has a non-negligible contribution from \(\bs \alpha e^{\bs S\Delta}/\bs \alpha e^{\bs S\Delta}\bs e\) and, to our knowledge, there is no way to ensure the denominator is bounded away from \(0\) when we construct limiting arguments. This issue is present when we try to do analysis on 
%\[\cfrac{\bs \alpha e^{\bs S\Delta}}{\bs \alpha e^{\bs S\Delta}\bs e}\bs D = \cfrac{\bs \alpha e^{\bs S\Delta}}{\bs \alpha e^{\bs S\Delta}\bs e} \int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e}\wrt u.\]

In practice, it may be possible to avoid this issue by choosing the intervals \(\{\mathcal D_\ell\}\) so that the boundaries do not align with any point masses. Another option is to append an ephemeral class of phases to the fluid queue as previously stated.

% Suppose we wish to have a point mass at \(x_0=y_{\ell}\), \(i\in\calS_{+0}\), and without loss of generality, assume the point mass has total mass 1. Let \(\mathcal S_{+0}^*\) with rates \(c_j=0\), \(j\in\mathcal S_{+0}^*\). These states represent a duplicated copy of \(\mathcal S_{+0}\). The initial orbit representing the point mass is \(\bs \alpha^{(p)}\), \(i\in\mathcal S_{+0}^*\). The dynamics of the QBD-RAP process regarding the ephemeral class \(\calS_{0+}^*\) is as follows. The phase process evolves according to the sub-generator matrix \(\bs T_{00}\) whilst it remains in \(\calS_{0+}^*\). Meanwhile, the orbit and level remain constant. When in phase \(k\in\calS_{0+}^*\), at rate \([\bs T_{0+}]_{kj}\) the process transitions out of \(\calS_{0+}^*\) and into phase \(j\in\mathcal S_+\). Similarly, when in phase \(k\in\calS_{0+}^*\), at rate \([\bs T_{0-}]_{kj}\) the process transitions out of \(\calS_{0+}^*\) and into phase \(j\in\mathcal S_-\). Upon a transition from \(\mathcal S_{0+}^*\) to \(j\in\mathcal S_{+}\) the orbit and level remain unchanged, and the process evolves as usual. Upon a transition from \(\mathcal S_{0+}^*\) to \(j\in\mathcal S_{-}\) the orbit remains unchanged, but the level jumps to \(\ell - 1\). Thus, the process evolves as if it has just entered level \(\ell-1\) in phase \(j\in\mathcal S_-\). 
\end{rem}

\begin{thm}\label{thm: a thm2!}
	As \(p\to \infty\) the following convergence statements hold.
	For \(r\in\{+,-\},\, q\in\{+0,-0\}\) and \(m=0\),  
	\begin{align}\int_{x\in\calD_{\ell_0}}\widehat f_{0,q,r}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\psi(x)\wrt x \to \int_{x\in\calD_{\ell_0}}\widehat \mu_{0,q,r}^{\ell_0}(\lambda)(x,j;x_0,k)\psi(x)\wrt x.\end{align}
	For \(r\in\{+,-\},\, q\in\{+0,-0\}\) and \(m\geq 1\), 
	\begin{align}\int_{x\in\calD_{\ell_0}}\widehat f_{m,q,r}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\psi(x)\wrt x \to \int_{x\in\calD_{\ell_0}}\widehat \mu_{m,q,r}^{\ell_0}(\lambda)(x,j;x_0,k)\psi(x)\wrt x.\label{eqn: thm 22}\end{align}
\end{thm}
\begin{proof}
	\textit{Cases \((q,r) \in \{(+0,-),(-0,+)\}\), and \(m=0\).} First, take \(q=-0\) and \(r=+\), then 
	\begin{align}
		&\int_{x\in\calD_k}\widehat{f}_{0,-0,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\psi(x)\wrt x \nonumber 
		\\&:=\int_{x_1=0}^\infty \int_{x\in\calD_k} \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0+} \bs M^0_{++}(\lambda,x_1)\bs e_j \nonumber
		\bs a_{\ell_0,k}^{(p)}(x_0) \bs D^{(p)} \bs N^{1,(p)}(\lambda,x_1) 
		\\& {\bs v}^{(p)}(y_{\ell_0+1}-x)\psi(x)\wrt x\wrt x_{1}  \nonumber
		\\& + \int_{x_1=0}^\infty\int_{x_2=0}^\infty \int_{x\in\calD_k}  \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0-} \bs M^1_{-+}(\lambda,x_1,x_{2})\bs e_j \nonumber
		\bs a_{\ell_0,k}^{(p)}(x_0)\bs N^{2,(p)}(\lambda,x_1,x_{2}) 
		\\& {\bs v}^{(p)}(y_{\ell_0+1}-x)\psi(x)\wrt x\wrt x_{2} \wrt x_1. \label{eqn: thm 1223}
	\end{align}
	The second term is a linear combination of \(\int_{x\in\calD_k}\widehat{f}_{1,-,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,i) \psi(x)\wrt x\) terms, each of which converge to \(\int_{x\in\calD_k}\widehat{\mu}_{1,-,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,i) \psi(x)\wrt x\), by the case we proved in Theorem~\ref{thm: a thm!}. As for the first term, it is a linear combination of terms
	\[\int_{x_1=0}^\infty \int_{x\in\calD_k} \bs e_i \bs H^{++}(\lambda,x_1)\bs e_j \nonumber
	\bs a_{\ell_0,k}^{(p)}(x_0) \bs D^{(p)} e^{\bs S^{(p)}x_1} 
	{\bs v}^{(p)}(y_{\ell_0+1}-x)\psi(x)\wrt x\wrt x_{1}.\]
	Lemma~\ref{lem: ppp} proves that such terms converge to \(\int_{x\in\calD_k}\widehat{\mu}_{0,+,+}^{\ell_0}(\lambda)(x,j;x_0,i) \psi(x)\wrt x\). Therefore (\ref{eqn: thm 1223}) is a finite linear combination of terms, each of which converge, hence (\ref{eqn: thm 1223}) converges and converges to 
	\begin{align}
		&\int_{x_1=0}^\infty \int_{x\in\calD_k} \bs e_k \sum_{i\in\calS_+} [\bs I - \bs T_{00}]^{-1}\bs T_{0i}\int_{x\in\calD_k}\widehat{\mu}_{0,+,+}^{\ell_0}(\lambda)(x,j;x_0,i) \psi(x)\wrt x \nonumber
		\\& + \int_{x_1=0}^\infty\int_{x_2=0}^\infty \int_{x\in\calD_k} \sum_{i\in\calS_-}\bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0i}\int_{x\in\calD_k}\widehat{\mu}_{1,-,+}^{\ell_0}(\lambda)(x,j;x_0,i) \psi(x)\wrt x, \label{eqn: thm 1223B}
	\end{align}
	which is \(\widehat{\mu}_{0,-0,+}^{\ell_0}(\lambda)(x,j;x_0,k)\). This proves the result for \(r=+\) and \(q=-0\). Analogous arguments prove the result for \(r=-\) and \(q=+0\).

	\textit{Cases \(r \in \{+,-\},\, q\in\{+0,-0\} \) and \(m\geq 1\).} First, take \(q=+0\) and \(r=+\), then 
	\begin{align}
		&\int_{x\in\calD_k}\widehat{f}_{m,-0,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\psi(x)\wrt x \nonumber 
		\\&:=\int_{x\in\calD_k}\int_{x_1=0}^\infty \dots \int_{x_{2m+1}=0}^\infty   \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0+}\bs M^m_{++}(\lambda,x_1,\dots,x_{2m+1})\bs e_j \nonumber
		\\& \bs a_{\ell_0,k}^{(p)}(x_0) \bs D^{(p)} \bs N^{2m+1,(p)}(\lambda,x_1,\dots,x_{2m+1}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\psi(x)\wrt x_{2m+1}\dots\wrt x_2 \wrt x_1  \wrt x\nonumber
		\\& + \int_{x\in\calD_k}\int_{x_1=0}^\infty \dots \int_{x_{2m+2}=0}^\infty   \bs e_k [\bs I - \bs T_{00}]^{-1}\bs T_{0-}\bs M^{m+1}_{-+}(\lambda,x_1,\dots,x_{2m+2})\bs e_j \nonumber
		\\& \bs a_{\ell_0,k}^{(p)}(x_0)\bs N^{2m+2,(p)}(\lambda,x_1,\dots,x_{2m+2}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\psi(x)\wrt x_{2m+2}\dots\wrt x_2 \wrt x_1 \wrt x.
	\end{align}
	The second term is a linear combination of \(\int_{x\in\calD_k}\widehat{f}_{m+1,-,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,i) \psi(x)\wrt x\) terms, each of which converge to \(\int_{x\in\calD_k}\widehat{\mu}_{m+1,-,+}^{\ell_0,(p)}(\lambda)(x,j;x_0,i) \psi(x)\wrt x\).
	As for the first term, it is a linear combination of terms 
	\begin{align*}
		&\int_{x\in\calD_k}\int_{x_1=0}^\infty \dots \int_{x_{2m+1}=0}^\infty \bs e_i \bs M^m_{++}(\lambda,x_1,\dots,x_{2m+1})\bs e_j \nonumber
		\\& \bs a_{\ell_0,k}^{(p)}(x_0)\bs D^{(p)}\bs N^{2m+1,(p)}(\lambda,x_1,\dots,x_{2m+1}) {\bs v}^{(p)}(y_{\ell_0+1}-x)\psi(x)\wrt x_{2m+1}\dots\wrt x_2 \wrt x_1 \wrt x
		%
		\\&=\int_{x\in\calD_k}\int_{x_1=0}^\infty \dots \int_{x_{2m+1}=0}^\infty \bs e_i \bs H^{+-}(\lambda,x_1)\prod_{r=1}^{m-1}\bs H^{-+}(\lambda,x_{2r}) \bs H^{+-}(\lambda,x_{2r+1}) \\&\bs H^{-+}(\lambda,x_{2m}) 
		\bs H^{++}(\lambda,x_{2m+1})\bs e_j \nonumber
		\bs a_{\ell_0,k}^{(p)}(x_0)\prod_{r=1}^{2m} e^{\bs{S}^{(p)}x_{r}}\bs{D}^{(p)} e^{\bs{S}^{(p)}x_{2m+1}}
		\\& {\bs v}^{(p)}(y_{\ell_0+1}-x)\psi(x)\wrt x_{2m+1}\dots\wrt x_2 \wrt x_1 \wrt x,
	\end{align*}
	which satisfies the assumptions of Lemma~\ref{lem: boobies2}. To see this take \(n=2m+1\), \(\bs G_1(x_1) = \bs e_i\bs H^{+-}(\lambda, x_1)\), \(\bs G_{2k}(x_{2k}) = \bs H^{-+}(\lambda, x_{2k})\), \(\bs G_{2k+1}(x_{2k}) = \bs H^{+-}(\lambda, x_{2k+1})\), \(k=1,\dots,m-1\) and \(\bs G_{2m}(x_{2m}) = \bs H^{-+}(x_{2m})\) and \(\bs G_{2m+1} = \bs H^{++}(\lambda,x_{2m+1})\bs e_j\) in Equation (\ref{eqn: KAFnnmna22G}). By the remarks following Lemma~\ref{lem: boobies2}, this gives the required convergence for this case. Analogous arguments prove the result for the remaining combinations of \((q,r)\). 
\end{proof}

\section{Technical results}

\begin{lem}\label{lem: ppp}
	Let \(\psi:\calD_{\ell_0}\to \mathbb R\) be bounded \(|\psi(x)|\leq F\) and Lipschitz continuous function and let \(\lambda >0\). For \(i\in\mathcal S_-,j\in\mathcal S_-\cup\calS_{-0}\), \(x_0\in(0,\Delta)\), 
	\begin{align}
		&\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\bs k^{(p)}(x_0) \bs D^{(p)} e^{\bs S x_1} {\bs v}^{(p)}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1  
		\to \int_{x=0}^{x_0} h_{ij}^{--}(\lambda,x_0-x)\psi(x)\wrt x, \label{eqn: LLLlllL}
	\end{align}
	as \(p\to\infty\). Similarly, for \(i\in\mathcal S_+,j\in\mathcal S_+\cup\calS_{+0}\)
	\begin{align}
		&\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\bs k^{(p)}(x_0) \bs D^{(p)} e^{\bs S^{(p)} x_1} {\bs v}^{(p)}(x) h_{ij}^{++}(\lambda, x_1) \psi(\Delta-x) \wrt x \wrt x_1  \nonumber 
		\\&\to\int_{x=\Delta-x_0}^{\Delta} h_{ij}^{++}(\lambda,x-x_0)\psi(x)\wrt x, \label{eqn: LLLlllL2222}
	\end{align}
\end{lem}
\begin{proof}
	Assume, without loss of generality \(\ell_0=0\) so \(\calD_{\ell_0}=[0,\Delta]\). Substituting the definition of \(\bs D\), then the left-hand side of (\ref{eqn: LLLlllL}) is
	\begin{align}
%		&\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\bs k^{(p)}(x_0) \bs D^{(p)} e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\nonumber
%		 - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber
		%
		&\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u=0}^\infty \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\nonumber
		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
		%
		\\&\leq \Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\nonumber
		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
		%
		\\&\qquad {}+\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\Bigg| \label{eqn: llllL}
	\end{align}
	Since \(|\psi(x)|\leq F\) and \(|h_{ij}^{--}(\lambda,x)|\leq G\), the third term in (\ref{eqn: llllL}) is less than or equal to 
	\begin{align}
		&\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) \wrt x \wrt x_1G F.\label{eqn: a big lola}
	\end{align} 
	Computing the integral over \(x_1\) in (\ref{eqn: a big lola}) gives
	\begin{align}
		& \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u (-\bs S)^{-1}{\bs v}(x) \wrt x G F\nonumber
		%
		\\&\leq  \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \wrt u \wrt x G_{\bs v} G F \nonumber
		%
		\\&= \int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s  \wrt u\Delta G_{\bs v} G F.\label{eqn: google the yes}
	\end{align}
	since, by property \ref{properties: 1}, \(\bs k(u)(-\bs S)^{-1}{\bs v}(x)\leq \bs k(u)\bs e G_{\bs v}=G_{\bs v}\). The bound (\ref{eqn: google the yes}) is equal to 
	\begin{align}
		\Delta \mathbb P(Z>x_0 + \Delta-\varepsilon) \wrt x G_{\bs v} G F
		\leq  \Delta \cfrac{\var(Z)}{(x_0-\varepsilon)^2} G_{\bs v} G F, \label{eqn: JjJ}
	\end{align}
	by Chebyshev's inequality.

	The first and second terms in (\ref{eqn: llllL}) are
	\begin{align}
		&\Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \wrt x_1 \wrt u \psi(x) \wrt x \nonumber
		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
		%
		\\&=\Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \left[ h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) + r_{\bs v}(u,x) \right] \psi(x) \wrt u \wrt x \nonumber
		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg|\nonumber
		\\&\leq \Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) \psi(x) \wrt u \wrt x \nonumber 
		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| + \Bigg| \int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s r_{\bs v}(u,x) \psi(x) \wrt u \wrt x \Bigg|
		\label{eqn: aKK}
	\end{align}
	where the first equality holds from Property \ref{properties: 2}. The last term in (\ref{eqn: aKK}) is less than or equal to 
	\begin{align}
		\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su} \int_{x=0}^{\Delta} \bs s \left|r_{\bs v}(u,x)\right|\wrt x \wrt u F 
		%
		&\leq \int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su} \bs sR_{{\bs v},2} \wrt u F \nonumber 
		%
		\\&\leq R_{{\bs v},2} F, \label{eqn: HhHj}
	\end{align}
	by Property \ref{properties: 2}. The first two terms in (\ref{eqn: aKK}) are 
	\begin{align}
		& \Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) \psi(x) \wrt u \wrt x \nonumber
		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber 
		\\&= \Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \wrt x \nonumber
		  \\&\qquad{}- \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber 
		\\&\leq \Bigg| \int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u 1(\Delta-x_0+\varepsilon\leq \Delta-x-\varepsilon) \wrt x\nonumber
		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber 
		\\&\qquad {}+\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\min(\Delta-x_0-\varepsilon,\Delta-x-\varepsilon)} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \wrt x\Bigg|\nonumber
		\\&\qquad {}+\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \nonumber 
		\\&\qquad\qquad \times1(\Delta-x_0+\varepsilon\leq \Delta - x -\varepsilon)\wrt x\Bigg|\nonumber
		\\&\qquad {}+\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \nonumber
		\\&\qquad\qquad \times 1(\Delta-x_0-\varepsilon\leq \Delta - x -\varepsilon < \Delta-x_0+\varepsilon)\wrt x\Bigg|   \label{eqn: LLkkK}
	\end{align}
	Since \(|\psi|\leq F\), the second and third terms in (\ref{eqn: LLkkK}) are less than or equal to 
	\begin{align}
		&\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\min(\Delta-x_0-\varepsilon,\Delta-x-\varepsilon)} \bs k(x_0)e^{\bs Su}\bs s  \wrt u \wrt xGF \nonumber
		\\&\qquad {}+\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u 1(\Delta-x_0+\varepsilon\leq \Delta - x -\varepsilon)\wrt x GF \nonumber
		\\&\leq \int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\Delta-x_0-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u \wrt xGF \nonumber
		\\&\qquad {}+\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\infty} \bs k(x_0)e^{\bs Su}\bs s  \wrt u 1(\Delta-x_0+\varepsilon\leq \Delta - x -\varepsilon)\wrt x GF \nonumber
		\\&\leq\Delta\int_{u=0}^{\Delta-x_0-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u \wrt xGF 
		+ \Delta \int_{u=\Delta-x_0+\varepsilon}^{\infty} \bs k(x_0)e^{\bs Su}\bs s  \wrt u GF \nonumber
		\\&\leq \Delta \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}GF. \label{eqn:aaAAAAd}
	\end{align}
	Since \(\displaystyle \int_{u=\Delta-x_0-\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \leq 1\), the last term in (\ref{eqn: LLkkK}) is less than or equal to 
	\begin{align}
		&\int_{x=0}^{\Delta-\varepsilon}GF \wrt u 
		 1(\Delta-x_0-\varepsilon\leq \Delta - x -\varepsilon < \Delta-x_0+\varepsilon)\wrt x =2\varepsilon GF. \label{eqn:Qwe}
	\end{align}
	Now consider the difference in the first absolute value in (\ref{eqn: LLkkK}),
	\begin{align}
		&\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u 1(x\leq x_0-2\varepsilon) \wrt x \nonumber
		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg|\nonumber
		%
		\\&\leq \Bigg|\int_{x=0}^{ x_0-2\varepsilon }\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u  \wrt x \nonumber
		\\&{}\qquad {} - \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| + \Bigg| \int_{x=x_0-2\varepsilon}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
		%
		\\&\leq  \int_{x=0}^{ x_0-2\varepsilon }\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \Bigg| h_{ij}^{--}(\lambda, \Delta - u - x) - h_{ij}^{--}(\lambda, x_0 - x) \Bigg| \wrt u \left|\psi(x)\right| \wrt x  \nonumber
		\\&{}\qquad {} + \Bigg| \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x) \wrt x \mathbb P(|Z-\Delta|>\varepsilon)\Bigg| \nonumber 
		\\&\qquad{} + \Bigg|\int_{x=x_0-2\varepsilon}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber
		%
		\\&\leq \int_{x=0}^{ x_0-2\varepsilon }\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s L\varepsilon  \wrt u \left|\psi(x)\right| \wrt x \nonumber
		+\Delta GF \cfrac{\var(Z)}{\varepsilon^2} + 2\varepsilon GF
		%
		\\&\leq \Delta L\varepsilon   F 
		+\Delta GF \cfrac{\var(Z)}{\varepsilon^2} + 2\varepsilon GF.\label{eqn: LLLaase}
	\end{align}
	In summary, we have shown 
	\begin{align}
		&\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\bs k^{(p)}(x_0) \bs D^{(p)} e^{\bs S^{(p)} x_1} {\bs v}^{(p)}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1 \nonumber
		\\&\qquad{}- \int_{x=0}^{x_0} h_{ij}^{--}(\lambda,x_0-x)\psi(x)\wrt x \Bigg| \nonumber 
		%
		\\& \leq  \Delta L\varepsilon^{(p)}   F 
		+\Delta GF \cfrac{\var(Z^{(p)})}{\left(\varepsilon^{(p)}\right)^2} + 4\varepsilon^{(p)} GF + \Delta \cfrac{\var(Z^{(p)})/\left(\varepsilon^{(p)}\right)^2}{1-\var(Z^{(p)})/\left(\varepsilon^{(p)}\right)^2}GF + R_{{\bs v},2}^{(p)} F \nonumber
		\\&\qquad{} +  \Delta \cfrac{\var(Z^{(p)})}{(x_0-\varepsilon^{(p)})^2} G_{\bs v} G F.
	\end{align}
	The result follows upon choosing \(\varepsilon^{(p)}=\var(Z^{(p)})^{1/3}\) and letting \(p\to\infty\). 
\end{proof}

\subsection{One integral}
\begin{lem} \label{lem:tttttt}
	Let \(g\) satisfy the Assumptions \ref{asu: g} and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\). Then
	\begin{align}
		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}4G + 3L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2}.
	\end{align}
\end{lem}
\begin{proof}
	First rewrite the left-hand side as 
	\begin{align}
%		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| 
		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}(g(x)-g(x_0))\bs s\wrt x \right|
		%
		&\leq \int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x.
	\end{align}
	Substituting in the expression for \(\bs D\) gives,
	\begin{align}
		&\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x  \nonumber 
%		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x 
		%
		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \nonumber
		\\&\quad {} + \int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \label{eqn: 2nd here}
	\end{align}
	Since \(g\) is bounded, the second term is less than or equal to 
	\begin{align}
		\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x 2G \nonumber
		&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u \bs e 2G \nonumber
		\\&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \wrt u2G \nonumber
		\\&= \cfrac{\mathbb P(Z\geq x_0+\Delta-\varepsilon)}{\mathbb P(Z>x_0)}2G \label{eqn" dgkjlwerhv}
	\end{align}
	For \(x_0\in(2\varepsilon,\Delta-\varepsilon),\) then (\ref{eqn" dgkjlwerhv}) is less than or equal to 
	\[\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}2G.\]

	As for the first term in (\ref{eqn: 2nd here}), it can be written as 
	\begin{align}
		&\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
		%
		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
		%
		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x. \label{eqn: 201}
	\end{align}
	Since \(g\) is bounded, then the last two terms in (\ref{eqn: 201}) are 
	\begin{align}
		&2G\Bigg(\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x \nonumber
		%
		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x\Bigg) \nonumber 
		%
		\\&=2G\Bigg(\bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e}  \bs e  \wrt u\nonumber
		%
		+ \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \bs e\wrt u \Bigg) \nonumber
		%
		\\&=2G \cfrac{\mathbb P(Z>x_0, Z\notin(\Delta-\varepsilon, \Delta+\varepsilon))}{\mathbb P(Z>x_0)}  \nonumber 
		%
		\\&\leq 2G\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2},
	\end{align}
	provided that \(x_0\in[0,\Delta-\varepsilon)\). 
	Exchanging the order of integration for the first term in (\ref{eqn: 201})
	\begin{align}
		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\int_{x=0}^\infty \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \wrt u \label{eqn: fdk13897}
	\end{align}
	from which we see that we can apply Corollary~\ref{cor: cond bnd} to the integral over \(x\), implying that (\ref{eqn: fdk13897}) is less than or equal to
	\begin{align}
		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(|g(\Delta-u)-g(x_0)|+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u. \label{eqn: sdkagh lkhvasfv}
	\end{align}
	Since \(g\) is Lipschitz, then (\ref{eqn: sdkagh lkhvasfv}) is less than or equal to 
	\begin{align}
		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u \leq L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon. \label{eqn: sdkagh lsfv}
	\end{align}
	Putting all the bounds together proves the result. 
\end{proof}

%\begin{lem}
%	Let \(f:[0,\Delta)\to \mathbb R\) be any bounded, Lipschitz continuous function with \(|\psi(x)|\leq F\). Then for \(u\leq \Delta-\varepsilon\), \(v\in[0,\Delta)\)
%	\begin{align}
%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[f(\bar X(t))1(\varphi(t)=j, L(t)=\ell_0, L(s)=\ell_0, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid L(0)=\ell_0, \nonumber
%		\\&\qquad{} \bs A(0)=\bs a_{\ell_0,i}(x_0), \varphi(0)=i]\wrt t \nonumber
%		\\&= \int_{t=0}^\infty e^{-\lambda t}\mathbb E[\psi(X(t))1(\varphi(t)=j,X(s)\in\calD_{\ell_0}, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid X(0)=x_0, \varphi(0)=i]\wrt t \nonumber
%		\\&\qquad {} + r_{11}^{(p)}
%	\end{align}
%	where 
%	\[|r_{11}^{(p)}| \leq F R_{{\bs v},2}+ GF\varepsilon.\]
%\end{lem}
%\begin{proof}
%	Assume, without loss of generality \(\ell_0=0\) so \(\calD_{\ell_0}=[0,\Delta]\). First observe that 
%	\begin{align}
%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[f(\bar X(t))1(\varphi(t)=j, L(t)=\ell_0, L(s)=\ell_0, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid L(0)=\ell_0, \nonumber
%		\\&\qquad{} \bs A(0)=\bs a_{\ell_0,i}(x_0), \varphi(0)=i]\wrt t \nonumber 
%		\\&= \int_{x_1=0}^\infty \int_{x=[0,\Delta)} \cfrac{\bs \alpha e^{\bs S(x_1+x_0+x)}\bs s}{\bs \alpha e^{\bs S x_0} \bs e}h_{ij}^{++}(\lambda, x_1)f(\Delta-x)\wrt x\wrt x_1
%	\end{align}
%	and 
%	\begin{align}
%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[\psi(X(t))1(\varphi(t)=j,X(s)\in\calD_{\ell_0}, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid X(0)=x_0, \varphi(0)=i]\wrt t \nonumber
%		\\&=  \int_{x=0}^{\Delta-x_0} h_{ij}^{++}(\lambda, x)\psi(X_0 + x)\wrt x
%	\end{align}
%	In Property \ref{properties: 2} take \(g(x) = h_{ij}^{--}(\lambda, x)\), which states 
%	\begin{align}
%		&\int_{x_1=0}^\infty \int_{x=0}^\Delta\cfrac{\bs \alpha e^{\bs S(x_1+x_0+x)}\bs s}{\bs \alpha e^{\bs S x_0} \bs e}h_{ij}^{++}(\lambda, x_1)f(\Delta-x)\wrt x\wrt x_1 \nonumber
%		\\& = \int_{x=0}^{\Delta}h_{ij}^{++}(\lambda,\Delta - x_0 - x)1(x+x_0\leq \Delta-\varepsilon)f(\Delta - x)\wrt x + \int_{x=0}^\Delta r_{\bs v}(x_0,x)\psi(x)\wrt x. 
%	\end{align}
%	The second term is less than or equal to 
%	\[F\int_{x=0}^\Delta r_{\bs v}(x_0,x)\wrt x = FR_{{\bs v},1}.\]
%	The first term is 
%	\[\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x - \int_{x=\Delta-x_0-\varepsilon}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x.\]
%	Now 
%	\[\int_{x=\Delta-x_0-\varepsilon}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x\leq GF\varepsilon\]
%	and 
%	\[\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x=\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,x)\psi(x_0+x)\wrt x.\]
%\end{proof}

%\begin{cor}\label{cor: Dcoajc2222}
%	Let \(\psi:\calD_{\ell_0}\to \mathbb R\) be bounded, \(|\psi(x)|\leq F\), and Lipschitz continuous. Then, for \(x_0\in(y_{\ell_0}, y_{\ell+1}),\) \(k\in\calS_{0+}\), \(i\in\calS_-\), \(j\in\calS_-\cup\calS_{-0}\), there exists \(r_{10}^{(p)}\to 0\) as \(p \to \infty\), 
%	\begin{align}
%		\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)}.\label{eqn:xvasfv}
%	\end{align}
%	Similarly, for \(k\in\calS_{0-}\), \(i\in\calS_+\), \(j\in\calS_+\cup\calS_{+0}\)
%	\begin{align}
%		\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,+0,+}^{\ell_0}(x,i,j;j,x_0)\psi(x)\wrt x  - \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,+0,+}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)}.\label{eqn:!124msfvcds}
%	\end{align}
%\end{cor}
%\begin{proof}[Proof of Corollary \ref{cor: Dcoajc}]
%	We show the result for (\ref{eqn:xs}) only, with the result for (\ref{eqn:!124mcds}) following analogously. 
%	
%	Observe that, for \(k\in\calS_{0+}\), \(i\in\calS_-\), \(j\in\calS_-\cup\calS_{-0}\), 
%	\begin{align}
%		&\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x \nonumber 
%		%
%		\\&{}= \vligne{\lambda \bs I - \bs T_{00}}^{-1}_{ki}\int_{x\in\calD_{\ell_0}}\int_{x_1=0}^\infty \bs k(x_0) \bs D e^{\bs S x_1} {\bs v}(x)h_{ij}^{--}(\lambda,x_1)\wrt x_1\psi(x)\wrt x,
%	\end{align}
%	and 
%	\begin{align}
%		\int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x \nonumber  = \vligne{\lambda \bs I - \bs T_{00}}^{-1}_{ki}\int_{x=0}^{x_0} h_{ij}^{--}(\lambda,x_0-x)\psi(x)\wrt x 
%	\end{align}
%	These terms appear in the proof of Corollary~\ref{lem: ppp} and using arguments applied there we can show
%	\begin{align}
%		\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x ,\label{eqn:xssgwg}
%	\end{align}
%	and therefore there exists a bound \(r_{10}^{(p)}\) such that 
%	\[\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)} ,\]
%	and \(r_{10}^{(p)}\to 0\) as \(p\to\infty\). 
%\end{proof}

%\begin{cor}\label{corL Tta} For \(i\in\mathcal S_-,j\in\mathcal S_-\cup\calS_{-0}\), \(x_0\in[0,\Delta)\), 
%	\begin{align}
%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t \nonumber 
%		%
%		\\& \to \int_{t=0}^\infty e^{-\lambda t}\mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) \wrt t \label{eqn: ffaA}
%	\end{align}
%	as \(p\to\infty\).
%\end{cor}
%\begin{proof}
%	From the convergence of Laplace transforms in Lemma~\ref{lem: ppp} and the Continuity Theorem \ref{thm: ext cont thm} then 
%	\begin{align}
%		&  \mathbb E[f(\bar X^{(p)}(t))1(\varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j) \mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i]  \nonumber 
%		\\& \to \mathbb E[\psi(X(t)) 1(\varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j) \mid X(0)=x_0, \varphi(0)=i],
%	\end{align}
%	as \(p\to\infty\), for every bounded, Lipschitz function \(f\). By the Portmanteau Theorem, then 
%	\begin{align}
%		& \mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i)   \nonumber 
%		%
%		\\& \to \mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) 
%	\end{align}
%	weakly as \(p\to\infty\). In addition 
%	\begin{align}
%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t\nonumber 
%		\\& \leq \int_{x=0}^\infty  h_{ij}^{--}(\lambda, x)\wrt x ,
%	\end{align}
%	which is bounded as a function of \(p\). Hence we can apply the Continuity Theorem \ref{thm: ext cont thm} again and claim that 
%	\begin{align}
%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t \nonumber 
%		%
%		\\& \to \int_{t=0}^\infty e^{-\lambda t}\mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) \wrt t
%	\end{align}
%	as \(p\to\infty,\) as required. 
%\end{proof}
%

\subsection{Many integrals}
The orbit process at the time at which \(\{\phi(t)\}\) first exits \(\calS_{+0}\) (\(\calS_{-0}\)) is \(\bs k(x_0) \bs D\), thus we can treat this case simply as a change in the initial condition of the QBD-RAP. We do so by showing that asymptotically, the  initial conditions \(\bs k(\Delta-x_0)\) and \(\bs k(x_0)\bs D\) produce results that are arbitrarily close to each other. We first show a Lipschitz-like condition in \(x_0\) for \(w_n(x_0,x)\).

\begin{cor}\label{cor: awrg}
	For \(x_0,x\in[0,\Delta)\), \(n\geq 2\), 
	\begin{align}
		&\left| w_n(x_0,x)-w_n(z_0,x)\right| 
		\leq 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + |x_0-z_0|G_n^*(G+L\Delta), \label{eqn: ssdmm}
	\end{align}
\end{cor}
\begin{proof}
	By adding and subtracting both \(\displaystyle \int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1\) and \(\displaystyle \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - z_0)g^*_{2,n}(u_1,x)\wrt u_1\), we can write the left-hand side of (\ref{eqn: ssdmm}) as 
	\begin{align}
\nonumber		&\Bigg| w_n(x_0,x)
		- \int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1
	%
		\\\nonumber&\qquad{}- w_n(z_0,x)
		{}+ \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - z_0)g^*_{2,n}(u_1,x)\wrt u_1
		%
		\\\nonumber&\qquad {} +\int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1
		%
		{} - \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - z_0)g^*_{2,n}(u_1,x)\wrt u_1\Bigg| \nonumber
		%
		%
		\intertext{which, by the triangle inequality, is less than or equal to}
		\nonumber& \Bigg| w_n(x_0,x)- \int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1\Bigg|
		%
		\\\nonumber&\qquad{}+ \Bigg|w_n(z_0,x)- \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - z_0)g^*_{2,n}(u_1,x)\wrt u_1\Bigg|
		%
		\\&\qquad{} +\Bigg|\int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1 - \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - z_0)g^*_{2,n}(u_1,x)\wrt u_1\Bigg|\label{eqn: kdfsdf}
		%
%		\\&{} \leq 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| 
%		\\&{}+\Bigg|\int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)w(\Delta-u_1)\wrt u_1 - \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - z_0)w(\Delta-u_1)\wrt u_1\Bigg|.
	\end{align}
	By Corollary~\ref{cor: a cor}, the first two terms of (\ref{eqn: kdfsdf}) are less than or equal to \(|r_5(n)| + |r_6(n)| + (n-1)|r_4(n)|\). 
	As for the last term, adding and subtracting \( \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1\) gives 
	\begin{align}
%		\nonumber &\Bigg|\int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1 - \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - z_0)g^*_{2,n}(u_1,x)\wrt u_1\Bigg|
%		%
		\nonumber &{} = \Bigg|\int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1 - \int_{u_1=0}^{\Delta-z_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1
		\\\nonumber &\qquad{} - \int_{u_1=0}^{\Delta-z_0}(g_1(\Delta - u_1 - z_0)-g_1(\Delta - u_1 - x_0))g^*_{2,n}(u_1,x)\wrt u_1\Bigg|
		%
		\\\nonumber& \leq  \Bigg|\int_{u_1=\Delta-z_0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)g^*_{2,n}(u_1,x)\wrt u_1\Bigg|
		\\\nonumber &\qquad{} + \int_{u_1=0}^{\Delta-z_0}|g_1(\Delta - u_1 - z_0)-g_1(\Delta - u_1 - x_0)|g^*_{2,n}(u_1,x)\wrt u_1
		%
		\\\label{eqn: shhhhhh}& \leq  GG^*_n|x_0-z_0|
		 + \int_{u_1=0}^{\Delta-z_0}L|x_0-z_0|G^*_n\wrt u_1
		 \end{align}
		{since \(g_1\) is Lipschitz by Assumption \ref{asu: lipschitz} and \(g_{2,n}^*\leq G_n^*\). Bounding the integral over \(u_1\) by \(\Delta\), then (\ref{eqn: shhhhhh}) is less than or equal to}
		\begin{align}
		&GG^*_n|x_0-z_0| + \Delta L|x_0-z_0|G^*_n.
	\end{align}
\end{proof}

\begin{cor}\label{cor: ahjg}
	Let \(g_1,g_2,\dots,\) be functions satisfying Assumptions \ref{asu: g} and let \({\bs v}(x)\), \(x\in[0,\Delta)\), be a closing operator with Properties \ref{properties: some props}. For \(x_0,x\in\mathcal [0,\Delta)\), \(n\geq 2\), 
	\begin{align}
		&\Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right]
            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \nonumber 
	%
		\\&\quad{}- w_n(\Delta - x_0,x) \Bigg| \nonumber
		\\&= r_8(n),
	\end{align}
	where 
	\begin{align*}
	|r_8(n)|&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G_n^{*}(G+L\Delta) \right) \\&\qquad{}+ 2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}.\end{align*}
\end{cor}
\begin{proof}
	Observe that 
	\begin{align}
		\nonumber&\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right]
            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) 
	%
		\\\nonumber&=\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs s\cfrac{\bs\alpha e^{\bs Sz_0}}{\bs\alpha e^{\bs Sz_0}\bs e} \wrt z_0 e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right]
            	\\\nonumber&\quad\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) 
	%
		\\&=\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0
	\end{align}

	Now
	\begin{align}
		\nonumber&\left|\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0 - w_n(\Delta - x_0,x) \right|
		\\\nonumber&= \left|\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs s(w_n(z_0,x) - w_n(\Delta - x_0,x)) \wrt z_0\right|
		%
		\\\nonumber&\leq \bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs s  \left|w_n(z_0,x) - w_n(\Delta - x_0,x)\right| \wrt z_0
		%
		\\\nonumber&= \bs k(x_0) \int_{z_0=0}^{\Delta-\varepsilon-x_0} e^{\bs Sz_0}\bs s  \left|w_n(z_0,x) - w_n(\Delta - x_0,x)\right| \wrt z_0
		\\\nonumber&\qquad{}+\bs k(x_0) \int_{z_0=\Delta+\varepsilon-x_0}^\infty e^{\bs Sz_0}\bs s  \left|w_n(z_0,x) - w_n(\Delta - x_0,x)\right| \wrt z_0
		\\\label{eqn: KASJF}&\qquad{}+\bs k(x_0) \int_{z_0=\Delta-\varepsilon-x_0}^{\Delta+\varepsilon-x_0} e^{\bs Sz_0}\bs s  \left|w_n(z_0,x) - w_n(\Delta - x_0,x)\right| \wrt z_0.
		%
		\end{align}
		By Corollary~\ref{cor: ksjkd}, for any \(x,y\in [0,\Delta)\), \(0\leq w_n(x,y)\leq \widehat G^{n-2}GG_{\bs v}\), so the sum of the first two terms is less than or equal to 
		\begin{align}
		&2\widehat G^{n-1}GG_{\bs v}\left(\int_{z_0=0}^{\Delta-\varepsilon-x_0}\bs k(x_0)e^{\bs S z_0} \bs s \wrt z_0+\int_{z_0=\Delta+\varepsilon-x_0}^\infty \bs k(x_0)e^{\bs S z_0} \bs s \wrt z_0\right)
		\\\nonumber &= 2\widehat G^{n-1}GG_{\bs v} \cfrac{\mathbb P(|Z-\Delta|> \varepsilon)}{\mathbb P(Z> x_0)}
		\\&\leq 2G^n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}
		\end{align}
		by Chebyshev's inequality. As for the last term in (\ref{eqn: KASJF}), we can use Corollary~\ref{cor: awrg} to bound the integrand so that the last term is less than or equal to 
		\begin{align}
		\nonumber&\bs k(x_0) \int_{z_0=\Delta-\varepsilon-x_0}^{\Delta+\varepsilon-x_0} e^{\bs Sz_0}\bs s \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G^{n-1}\Delta^{n-2}(G+L\Delta) \right) \wrt z_0
		%
		\\\nonumber&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G^{n-1}\Delta^{n-2}(G+L\Delta) \right)
		\\&\quad{}+2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2} ,
	\end{align}
	since \(\displaystyle \bs k(x_0) \int_{z_0=\Delta-\varepsilon-x_0}^{\Delta+\varepsilon-x_0} e^{\bs Sz_0}\bs s\wrt z_0\leq 1\). 
\end{proof}

\begin{cor} \label{cor: aaaaa}
	Let \(g_1,g_2,\dots,\) be functions satisfying Assumptions \ref{asu: g} and let \({\bs v}(x)\), \(x\in(0,\Delta)\), be a closing operator with Properties \ref{properties: some props}. For \(x_0,x\in(0,\Delta)\), \(n\geq 2\)
	\begin{align}
		&\Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right]
            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \nonumber 
	%
		\\&\qquad {}- \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
		\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
		%\\&{}\nonumber
            	%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
            	g_{n}(\Delta - x-u_{n-1})
	\\&\qquad{} 1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1} \Bigg| \nonumber
		\\&\leq |r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|,\label{eqn: KAFnn}
	\end{align}
	where 
	\begin{align}
		|r_8(n)|&\leq \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G^{n-1}\Delta^{n-2}(G+L\Delta) \right) \\&\quad{} 
	+2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}.
	\end{align}
\end{cor}
\begin{proof}
	Adding and subtracting \(w_n(\Delta-x_0,x)\) within the absolute value on the left-hand side of (\ref{eqn: KAFnn}) 
	\begin{align*}
		%&\Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
%            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
%		\bs D\right]
%            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \nonumber 
	%
%		\\&{}- g_{1,n}^*(\Delta - x_0,x) \Bigg| 
		&\Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right]
            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \nonumber 
	%
		\\\nonumber &{}\qquad - w_n(\Delta - x_0,x) + w_n(\Delta - x_0,x) - g_{1,n}^*(\Delta - x_0,x) \Bigg| 
		\\&\leq \Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right]
            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) ,\nonumber 
	\\&\qquad - w_n(\Delta - x_0,x)\Bigg| + \Bigg| w_n(\Delta - x_0,x) - g_{1,n}^*(\Delta - x_0,x) \Bigg| 
	\end{align*}
	where the first absolute value is less than or equal to \(|r_8(n)|\) by Corollary~\ref{cor: ahjg} and the second absolute value is less than or equal to \(|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\) by Corollary~\ref{cor: a cor}.
\end{proof}

\begin{cor}
	Let \(\psi\) be bounded and Lipschitz, let \(g_1,g_2,\dots,\) be functions satisfying Assumptions \ref{asu: g} and let \({\bs v}(x)\), \(x\in(0,\Delta)\), be a closing operator with Properties \ref{properties: some props}. For \(x_0,x\in(0,\Delta)\), \(n\geq 2\)
	\begin{align}
		&\Bigg| \int_{x\in[0,\Delta)} \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right] \nonumber 
            	\\&\qquad{}\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \psi(x)\wrt x \nonumber 
	% 
		\\&\qquad {}- \int_{x\in[0,\Delta)} \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
		\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
		%\\&{}\nonumber
            	%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
            	g_{n}(\Delta - x-u_{n-1})
	\\&\qquad{} \times1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\psi(x)\wrt x \Bigg| \nonumber
		\\&\leq \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta.\label{eqn: KAFnnmna2}
	\end{align}
\end{cor}
\begin{proof}
	The left-hand side of (\ref{eqn: KAFnnmna2}) is less than or equal to 
	\begin{align}
		& \int_{x\in[0,\Delta)} \Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
		\bs D\right] \nonumber 
            	\\&\qquad {}\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x)  \nonumber 
	%
		- \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
		\\&\qquad {}\times \left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
            	g_{n}(\Delta - x-u_{n-1})
	\\&\qquad{} 1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\Bigg| \left|\psi(x)\right|\wrt x . \label{eqn: lfj}
	\end{align}
	Now, apply Corollary~\ref{cor: aaaaa} to the first absolute value then (\ref{eqn: lfj}) is less than or equal to 
	\begin{align}
		&\int_{x\in[0,\Delta)} \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right) \left|\psi(x)\right|\wrt x  \nonumber 
		\\&\leq \int_{x\in[0,\Delta)} \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right) F\wrt x  \nonumber 
		\\&=  \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right) \Delta F
	\end{align}
\end{proof}

To summarise, the main results we need are 
\begin{align}
	&\Bigg| \int_{x=0}^\Delta w_n(x_0,x) \psi(x) \wrt x \nonumber 
%
	\\&{}- \int_{x=0}^\Delta \int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)
%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
	\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
	%\\&{}\nonumber
			%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
			g_{n}(\Delta - x-u_{n-1})
\\&\qquad{} 1(\Delta-x-u_{n-1}\geq0) \wrt u_{n-1}\psi(x) \wrt x \Bigg| \nonumber
	\\&\leq (|r_5(n)| + |r_6(n)| + (n-1)|r_4(n)|)\Delta F. \label{eqn: rhs g 4dvfklsmv2}
\end{align}
and 
\begin{align}
	&\Bigg| \int_{x\in[0,\Delta)} \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
			\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
	\bs D\right] \nonumber 
			\\&\qquad{}\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \psi(x)\wrt x \nonumber 
%
	\\&\qquad {}- \int_{x\in[0,\Delta)} \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
	\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
	%\\&{}\nonumber
			%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
			g_{n}(\Delta - x-u_{n-1})
\\&\qquad{} \times1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\psi(x)\wrt x \Bigg| \nonumber
	\\&\leq \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta.\label{eqn: KAFnnmna22}
\end{align}

We have assumed throughout the appendix that the functions \(g\) are scalar functions, however, we are ulitmately interested in expressions which contain matrix functions. We now extend the previous results to the matrix case.
\begin{lem}\label{lem: boobies2}
	Let \(\bs G_k(x)\), \(k\in\{1,2,...\}\), be matrix functions with dimensions \(N_k \times N_{k+1}\). Further, suppose \([\bs G_k(x)]_{ij}\), \(k\in\{1,2,...\}\) satisfy Assumptions \ref{asu: g}. Then, 
	\begin{align}
		&\Bigg| \int_{x\in[0,\Delta)} \int_{x_1=0}^\infty \bs G_1(x_1) \otimes \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
				\left[\prod_{k=2}^{n-1}\int_{x_k=0}^\infty \bs G_{k}(x_k) \otimes e^{\bs{S}x_k} \wrt x_k
		\bs D\right] \nonumber 
				\\&\qquad{}\times\int_{x_n=0}^\infty \bs G_{n}(x_n) \otimes e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \psi(x)\wrt x \nonumber 
	%
		\\&\qquad {}- \int_{x\in[0,\Delta)} \int_{u_1=0}^{x_0}\bs G_1(x_0 - u_1)
	%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
		\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} \bs {G}_{k}(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
		%\\&{}\nonumber
				%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
				\\&\qquad{} \bs G_{n}(\Delta - x-u_{n-1})
	 		\times1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\psi(x)\wrt x \Bigg| \nonumber
		\\&\leq \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta \prod_{k=2}^{n}N_{k}.\label{eqn: KAFnnmna22G}
	\end{align}
	Moreover, choosing \(\varepsilon=\var(Z)\), then, for fixed \(n\), the error term \(\left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta \prod_{k=2}^{n}N_{k}\) is \(\mathcal O(\var(Z)^{1/3})\). 
\end{lem}
\begin{proof}
	The proof is the same as the proof of Lemma~\ref{lem: boobies}.
\end{proof}
Lemma \ref{lem: boobies} effectively shows that, as \(p \to \infty\), then 
\begin{align}
	&\int_{x\in[0,\Delta)} \int_{x_1=0}^\infty \bs G_1(x_1) \otimes \bs k^{(p)} (x_0) \bs D^{(p)} e^{\bs{S}^{(p)}x_1}\wrt x_1\bs D^{(p)} 
			\left[\prod_{k=2}^{n-1}\int_{x_k=0}^\infty \bs G_{k}(x_k) \otimes e^{\bs{S}^{(p)}x_k} \wrt x_k
	\bs D^{(p)} \right] \nonumber 
			\\&\qquad{}\times\int_{x_n=0}^\infty \bs G_{n}(x_n) \otimes e^{\bs{S}^{(p)}x_n} \wrt x_n {\bs v}^{(p)}(x) \psi(x)\wrt x \nonumber 
%
	\\&\qquad {}\to \int_{x\in[0,\Delta)} \int_{u_1=0}^{x_0}\bs G_1(x_0 - u_1)
%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
	\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} \bs {G}_{k}(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
	%\\&{}\nonumber
			%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
			\\&\qquad{} \bs G_{n}(\Delta - x-u_{n-1})
		 \times1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\psi(x)\wrt x.  \nonumber
\end{align}

### 2021-11-24 appendix4/appendix.tex
story-ify appendix A

%!TEX root = ../thesis.tex
\chapter{Kronecker properties}\label{appendix: kronecker}
Here we detail some properties of Kronecker sum, products, and exponentials (see \citep{MEinAP}, Appendix A.4).

Let 
\[\bs A = \left[\begin{array}{ccc}a_{11} & \dots & a_{1m}\\\hdots & & \hdots \\ a_{n1}&\dots & a_{nm}\end{array}\right]
\qquad
\bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
be matrices. The operator \(\otimes\) is the Kronecker product of two matrices; 
\[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
which is an \(nn'\times mm'\) matrix.

Let \(\bs{C},\bs{D}\)  be matrices with dimensions \(m\times k\) and \(m'\times k'\). A property of the Kronecker Product is 
\begin{align}
	\left(\bs A\otimes \bs{B}\right)\left(\bs{C}\otimes \bs{D}\right) &= \bs A\bs{C}\otimes \bs{B}\bs{D}.\label{eqn:mpr}\tag{Mixed Product Rule}
\end{align}
\begin{proof}
	The proof follows from 
	\begin{align*}
		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j}\\\vdots\\c_{nj}\bs{D} \end{array}\right] 
		%
		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
	\end{align*}
\end{proof}

If \(\bs A\) and \(\bs{B}\) are invertible matrices, then 
\begin{align}\label{eqn:kron inverse}
	\left(\bs A\otimes \bs{B}\right)^{-1} = \bs A^{-1}\otimes \bs{B}^{-1}.
\end{align}

Let \(\bs A\) and \(\bs{B}\) be \(n\times n\) and \(m\times m\) matrices, respectively. The Kronecker sum of \(\bs A\) and \(\bs{B}\) is denoted by \(\oplus\) and defined as 
\[\bs A\oplus \bs{B} := \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]

A property of the Kronecker sum is 
\begin{align}\label{eqn:kron exp}
	e^{\bs A\oplus \bs{B}}= e^{\bs A}\otimes e^{\bs{B}}.
\end{align}
\begin{proof}
	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence 
	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\). The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\). The former follows from 
	\begin{align}
	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A\otimes \bs{I}_m\right) \nonumber
	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
	\end{align}

	Therefore 
	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
	and the result follows by the mixed product rule. 
\end{proof}

\begin{lem}\label{lem: lst mpr}
	Let \(\bs{T}\) and \(\bs{C}\) be \(n\times n\), square matrices with \(\bs{C}\) diagonal and invertible; let \(\bs{S}\) be a \(p\times p\) matrix. Further, suppose \(\left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]\) is invertible for \(\lambda>0\). Then
\begin{align}
	&\int_{t=0}^\infty e^{-\lambda t}  e^{{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t}} \wrt t 
	%
	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}  \label{eqn:lstsimplify}\end{align}
\end{lem}
\begin{proof}
	Computing the integral on the left-hand side and then factorising the result and using the \ref{eqn:mpr} multiple times gives
	\begin{align}
            	\int_{t=0}^\infty e^{-\lambda t} e^{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t} \wrt t\nonumber 
            	%
            	&= - \left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]^{-1}
		%
		\\&= -  \left[\bs{T}\otimes \bs{I} + \left(\bs{C}\otimes \bs{I}\right)\left(\bs{I}\otimes \bs{S}\right) - \lambda \bs{I}\right]^{-1}\nonumber
		%
		\\&= -  \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}\otimes \bs{I}\right)^{-1}\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}\otimes \bs{I}\right)^{-1}\lambda \bs{I}\right)\right]^{-1}. \label{eqn: ref this one 12}
		%
	\end{align}
	By Equation~(\ref{eqn:kron inverse}) and since \(\bs{C}\) is invertible, (\ref{eqn: ref this one 12}) is equal to
	\begin{align}
		& - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\otimes \bs{I}\right)\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\otimes \bs{I}\right)\lambda \bs{I}\right)\right]^{-1}. \label{eqn: ref this one 13}
		%
	\end{align}
	{Using the \ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
	\begin{align}
		&- \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\bs{T}\right)\otimes \bs{I} + \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\lambda \bs{I}\right)\otimes \bs{I}\right)\right]^{-1} \nonumber
		%
		\\&= - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right)\right]^{-1} \nonumber
		%
		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1} \nonumber
		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1},\label{eqn: is an integral}
	\end{align}
	by definition of the Kronecker sum.

	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore (\ref{eqn: is an integral}) is 
	\begin{align*}
		-\left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1}
		&= \int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x\right)\oplus \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1}.
	\end{align*}
	{Using the rule in Equation~(\ref{eqn:kron exp}) gives }
	\begin{align*}
		&\int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)x}\otimes e^{ \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1},
	\end{align*}
	which is the result.
\end{proof}

%\section{Matrix exponentials}\label{appendix: matrix exponentials}
%This appendix contains some technical results regarding the matrix exponential.

The exponential of a matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}B^n.\]

\cite{ln2015} show the following.
\begin{lem}\label{lem: sfkjgn}
	Let \(\bs B\) be the block-partitioned matrix
	\[\bs B = \left[\begin{array}{cc} \bs B_{11} & \bs B_{12} \\ \bs B_{21} & \bs B_{22} \end{array}\right]\]
	where \(\bs B_{11}\) and \(\bs B_{22}\) are matrices of order \(m_1\) and \(m_2\), respectively. Denote by \(\bs H_{11}(t)\) the top-left quadrant of order \(m_1\) of \(e^{\bs Bt}\):
	\[\bs H_{11}(t) = \vligne{\bs I_{m_1\times m_1} & \bs 0} e^{\bs Bt}\left[\begin{array}{c}\bs I_{m_1\times m_1} \\ \bs 0\end{array}\right].\]

	The matrix \(\bs H_{11}(t)\) is the solution of 
	\begin{align}\label{eqn: akg987LKJ}
		\bs H_{11}(t) = e^{\bs B_{11}t} + \int_{v=0}^t\int_{u=v}^t e^{\bs B_{11}(t-u)}\bs B_{12}e^{\bs B_{22}(u-v)}\bs B_{21}\bs H_{11}(v)\wrt u\wrt v.
	\end{align}
\end{lem}

Let \(\bs H_{12}(t)\) be the top-right quadrant of \(e^{\bs Bt}\) of size \(m_1\times m_2\), i.e.
\begin{align}
	\bs H_{12}(t) &= \vligne{  \bs I_{m_1\times m_1} & \bs 0}e^{\bs Bt} \left[\begin{array}{c} \bs 0 \\ \bs I_{m_2\times m_2}\end{array}\right]. \label{eqn: p03j}
\end{align}
Denote by \(\widehat{\bs H}_{11}(\lambda):=\displaystyle \int_{t=0}^\infty e^{-\lambda t}\bs H_{11}(t)\wrt t\) and by \(\widehat{\bs H}_{12}(\lambda):=\displaystyle \int_{t=0}^\infty e^{-\lambda t}\bs H_{12}(t)\wrt t\), the Laplace transforms of \(\bs H_{11}(t)\) and \(\bs H_{12}(t)\), respectively. Using Lemma~\ref{lem: sfkjgn} we can show the following result. 
\begin{lem}
	\begin{align}
		\widehat{\bs H}_{11}(\lambda) &= \int_{x=0}^\infty e^{\left(\bs B_{11} -\lambda \bs I_{m_1\times m_1} + \bs B_{12}(\lambda \bs I_{m_2\times m_2}- \bs B_{22})^{-1}\bs B_{21}\right)x} \wrt x,\label{eqn: 202}
		%
		\\\widehat{\bs H}_{12}(\lambda) &= \int_{x=0}^\infty e^{\left(\bs B_{11} -\lambda \bs I_{m_1\times m_1} + \bs B_{12}(\lambda \bs I_{m_2\times m_2}- \bs B_{22})^{-1}\bs B_{21}\right)x} \bs B_{12}(\lambda \bs I_{m_2\times m_2}-\bs B_{22})^{-1}\wrt x.\label{eqn: 203}
	\end{align}
\end{lem}
\begin{proof}
	First we show the result for \(\widehat{\bs H}_{11}(\lambda)\). Taking the Laplace transform of (\ref{eqn: akg987LKJ}) shows that \(\widehat{\bs H}_{11}(\lambda)\) is equal to 
	\begin{align}
		&\int_{t=0}^\infty \int_{v=0}^t\int_{u=v}^t e^{-\lambda (t-u)}e^{\bs B_{11}(t-u)}\bs B_{12}e^{-\lambda (u-v)}e^{\bs B_{22}(u-v)}\bs B_{21}e^{-\lambda v}\bs H_{11}(v)\wrt u\wrt v \nonumber 
		\\&\qquad{}+ (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1}\nonumber 
		%
		\\& = (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1} + (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1}\bs B_{12}(\lambda \bs I_{m_2\times m_2}-\bs B_{22})^{-1}\bs B_{21}\widehat{\bs H}_{11}(\lambda),
	\end{align}
	by the convolution theorem for Laplace transforms. This implies
	\begin{align*}
		&\left[\bs I_{m_1\times m_1} -  (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1}\bs B_{12}(\lambda \bs I_{m_2\times m_2}-\bs B_{22})^{-1}\bs B_{21}\right]\widehat{\bs H}_{11}(\lambda) 
		\\&= (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1},
	\end{align*}
	and therefore 
	\begin{align*}
		&\widehat{\bs H}_{11}(\lambda) 
		\\&= \left[\bs I_{m_1\times m_1} -  (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1}\bs B_{12}(\lambda \bs I_{m_2\times m_2}-\bs B_{22})^{-1}\bs B_{21}\right]^{-1} (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1}
		%
		\\&= \left[(\lambda \bs I_{m_1\times m_1} - \bs B_{11})\left(\bs I_{m_1\times m_1} -  (\lambda \bs I_{m_1\times m_1} - \bs B_{11})^{-1}\bs B_{12}(\lambda \bs I_{m_2\times m_2}-\bs B_{22})^{-1}\bs B_{21}\right)\right]^{-1} 
		%
		\\&= \left[\lambda \bs I_{m_1\times m_1} - \bs B_{11} - \bs B_{12}(\lambda \bs I_{m_2\times m_2}-\bs B_{22})^{-1}\bs B_{21}\right]^{-1}
		%
		\\&= \int_{t=0}^\infty e^{\left(\bs B_{11} - \lambda \bs I_{m_1\times m_1} + \bs B_{12}(\lambda \bs I_{m_2\times m_2}-\bs B_{22})^{-1}\bs B_{21}\right)t}\wrt t,
	\end{align*}
	which is (\ref{eqn: 202}).

	Now, to show (\ref{eqn: 203}), differentiate (\ref{eqn: p03j})
	\begin{align}
		\cfrac{\wrt}{\wrt t}\bs H_{12}(t) &= \vligne{  \bs I_{m_1\times m_1} & \bs 0}e^{\bs Bt} \left[\begin{array}{cc} \bs B_{11} & \bs B_{12} \\ \bs B_{21} & \bs B_{22} \end{array}\right] \left[\begin{array}{c} \bs 0 \\ \bs I_{m_2\times m_2}\end{array}\right] \nonumber
		%
		\\&= \vligne{  \bs I_{m_1\times m_1} & \bs 0}e^{\bs Bt} \left[\begin{array}{c} \bs B_{12} \\ \bs B_{22} \end{array}\right] \nonumber
		%
		\\&= \bs H_{11}(t) \bs B_{12} + \bs H_{12}(t)\bs B_{22}.
	\end{align}
	Now take the Laplace transform 
	\begin{align}
		\lambda \widehat{\bs H}_{12}(\lambda) - \bs H_{12}(0) = \widehat{\bs H}_{11}(\lambda) \bs B_{12} + \widehat{\bs H}_{12}(\lambda)\bs B_{22}.
	\end{align}
	Since \(\bs H_{12}(0)=\bs 0\) and after rearranging we get 
	\begin{align}
		\widehat{\bs H}_{12}(\lambda) = \widehat{\bs H}_{11}(\lambda) \bs B_{12} (\lambda \bs I_{m_2\times m_2} -\bs B_{22})^{-1},
	\end{align}
	which gives (\ref{eqn: 203}) upon substituting (\ref{eqn: 202}).
\end{proof}

\begin{cor}\label{cor: mpr B}
	For \(m\in\{+,-\}\) the top-left quadrant of size \(m_1\times m_1=|\calS_m|\cdot p \times |\calS_m|\cdot p\) of \(e^{\bs B_{mm}t}\), 
	\begin{align}
		&\vligne{\bs I & \bs 0}\int_{t=0}^\infty e^{-\lambda t} \exp\left\{\left[\begin{array}{cc} \bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S & \bs T_{m0}\otimes \bs I \\ \bs T_{0m} \otimes \bs I & \bs T_{00}\otimes \bs I \end{array}\right]t\right\} \wrt t\left[\begin{array}{c} \bs I \\ \bs 0 \end{array}\right], \nonumber
	\end{align}
	is given by 
	\begin{align}
		\int_{x=0}^\infty e^{\bs Q_{mm} (\lambda)x}\otimes  e^{\bs S x}\wrt x(\bs C_m^{-1}\otimes \bs I).\label{eqn: skagh87} 
	\end{align}
	For \(m\in\{+,-\}\) the top-right quadrant of size \(m_1\times m_2=|\calS_m|\cdot p\times |\calS_0|\cdot p\) of  \(e^{\bs B_{mm}t}\), 
	\begin{align}
		&\vligne{\bs I & \bs 0}\int_{t=0}^\infty e^{-\lambda t} \exp\left\{\left[\begin{array}{cc} \bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S & \bs T_{m0}\otimes \bs I \\ \bs T_{0m} \otimes \bs I & \bs T_{00}\otimes \bs I \end{array}\right]t\right\} \wrt t \left[\begin{array}{c}\bs 0 \\ \bs I\end{array}\right] ,\nonumber
	\end{align}
	is given by
	\begin{align}
		\int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x((\bs C_m^{-1}\bs T_{m0}(\lambda \bs I - \bs T_{00})^{-1})\otimes \bs I).\label{eqn: skagh873} 
	\end{align}
	Also, 
	\begin{align}
		&\vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t \bs{B}_{m{n}} %= \vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t  \left[\begin{array}{cc} \bs{T}_{mn}\otimes \bs{D} & \bs 0 \\ \bs T_{0n}\otimes \bs D & \bs 0 \end{array}\right]
	= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x\left(\vligne{\bs I_{pn} & \bs 0_{pn\times p|\calS_0|}}\right), \label{eqn: akgj987adKLDJ}
\end{align}
for \(m,n\in\{+,-\}\), \(m\neq n\).
\end{cor}
\begin{proof}
	From Lemma~\ref{lem: sfkjgn} the top-left quadrant of size \(m_1\times m_1=|\calS_m|\cdot p \times |\calS_m|\cdot p\) of the integral with respect to \(t\) on the left-hand side of (\ref{eqn: skagh87}) is 
	\begin{align}
		\int_{t=0}^\infty e^{\left(\bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S - \lambda \bs I+ (\bs T_{m0}\otimes \bs I)(\lambda \bs I-\bs T_{00}\otimes \bs I)^{-1}(\bs T_{0m}\otimes \bs I)\right)t}\wrt t. \label{eqn: akjf768}
	\end{align}
	By Lemma~\ref{lem: lst mpr}, (\ref{eqn: akjf768}) is equal to 
	\begin{align}
		&\int_{x=0}^\infty e^{\bs C_m^{-1}\left(\bs T_{mm} - \lambda \bs I+ \bs T_{m0} (\lambda \bs I-\bs T_{00})^{-1}\bs T_{0m} \right)x}\otimes e^{\bs Sx}\wrt x(\bs C_m\otimes \bs I)^{-1}\nonumber
		%
		\\&=\int_{x=0}^\infty e^{\bs Q_{mm} ( \lambda )x}\otimes e^{\bs Sx}\wrt x(\bs C_m\otimes \bs I)^{-1} , \label{eqn: akjf7623988}
	\end{align}
	from the definition of \(\bs Q_{mm}(\lambda)\). This proves (\ref{eqn: skagh87}).

	Now, from Lemma~\ref{lem: sfkjgn} the top-right quadrant of size \(m_1\times m_2=|\calS_m|\cdot p\times |\calS_0|\cdot p\) of the integral with respect to \(t\) on the left-hand side of (\ref{eqn: skagh873}) is 
	\begin{align}
		&\int_{t=0}^\infty e^{\left(\bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S - \lambda \bs I+ (\bs T_{m0}\otimes \bs I)(\lambda \bs I-\bs T_{00}\otimes \bs I)^{-1}(\bs T_{0m}\otimes \bs I)\right)t}(\bs T_{m0}\otimes \bs I)(\lambda \bs I - \bs T_{00}\otimes \bs I)^{-1} \nonumber 
		\\&\qquad\times (\bs T_{0m}\otimes \bs I)\wrt t. \label{eqn: akjf768f}
	\end{align}
	By Lemma~\ref{lem: lst mpr}, (\ref{eqn: akjf768f}) is equal to 
	\begin{align}
		\int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes e^{\bs Sx}\wrt x(\bs C_m\otimes \bs I)^{-1}(\bs T_{m0}\otimes \bs I)(\lambda \bs I - \bs T_{00}\otimes \bs I)^{-1}(\bs T_{0m}\otimes \bs I) . \label{eqn: akjf7623988g}
	\end{align}
	Now, 
	\begin{align*}
		\displaystyle (\lambda \bs I - \bs T_{00}\otimes \bs I)^{-1} &= \int_{u=0}^\infty e^{-(\lambda \bs I - \bs T_{00}\otimes \bs I)u}\wrt u \\&= \int_{u=0}^\infty e^{-\lambda u} e^{(\bs T_{00}\otimes \bs I)u}\wrt u \\&= \int_{u=0}^\infty e^{-\lambda u} e^{\bs T_{00}u}\otimes \bs I \wrt u,
	\end{align*} by (\ref{eqn:09ksdjgah}). Using this and the \ref{eqn:mpr} we can write  
	\begin{align}
		&(\bs C_m\otimes \bs I)^{-1}(\bs T_{m0}\otimes \bs I)(\lambda \bs I - \bs T_{00}\otimes \bs I)^{-1}(\bs T_{0m}\otimes \bs I) \nonumber
		\\&= (\bs C_m^{-1}\otimes \bs I)(\bs T_{m0}\otimes \bs I)\int_{u=0}^\infty e^{-\lambda u} e^{\bs T_{00}u}\otimes \bs I \wrt u(\bs T_{0m}\otimes \bs I) \nonumber
		\\&= \left(\bs C_m^{-1}\bs T_{m0}(\lambda \bs I-\bs T_{00}u)^{-1}\bs T_{0m}\right)\otimes \bs I).\label{eqn: q092}
	\end{align}
	Substituting (\ref{eqn: q092}) into (\ref{eqn: akjf7623988g}) completes the proof of (\ref{eqn: skagh873}).

Now, using (\ref{eqn: skagh87}) and (\ref{eqn: skagh873}) we can write 
\begin{align}
	\nonumber&\vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t \bs{B}_{m{n}} = \vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t  \left[\begin{array}{cc} \bs{T}_{mn}\otimes \bs{D} & \bs 0 \\ \bs T_{0n}\otimes \bs D & \bs 0 \end{array}\right]
	%
%	\\\nonumber& =\left[ \begin{array}{cc} \displaystyle \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x(\bs C_m^{-1} \otimes \bs I) & \displaystyle \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x((\bs C_m^{-1}\bs T_{m0}(\lambda \bs I - \bs T_{00})^{-1})\otimes \bs I)  \end{array}\right]
%	\\\nonumber&\quad\times\left[\begin{array}{cc} \bs{T}_{mn}\otimes \bs{D} & \bs 0 \\ \bs T_{0n}\otimes \bs D & \bs 0 \end{array}\right]
	%
	\\&= \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x (\bs C_m^{-1}(\bs{T}_{mn} + \bs T_{m0}(\lambda \bs I - \bs T_{00})^{-1}\bs T_{0n}) \vligne{\bs I_n & \bs 0_{n\times |\calS_0|}} \otimes \bs D) \nonumber
	%
	\\&= \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x \left( \left( \bs{Q}_{mn} (\lambda) \vligne{\bs I_n & \bs 0_{n\times |\calS_0|}} \right) \otimes \bs D\right)  \nonumber
	%
	\\&= \int_{x=0}^\infty \left(\bs H^{mn}(\lambda,x) \vligne{\bs I_n & \bs 0_{n\times |\calS_0|}}\right) \otimes  e^{\bs S x}\bs D\wrt x,\nonumber
	%
	\\&= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x,\vligne{\bs I_{pn} & \bs 0_{pn\times p|\calS_0|}} \label{eqn: akgj987ad}
\end{align}
for \(m,n\in\{+,-\}\), \(m\neq n\) which is (\ref{eqn: akgj987adKLDJ}), where the last line holds from the \ref{eqn:mpr}. 
\end{proof}

### 2021-11-25 appendix3/appendix.tex
begin rearranging appendix 3

@@
 \chapter{Convergence without ephemeral phases\label{app:extend conv}}
-For completeness, we include here, the results needed to prove the convergence of the QBD-RAP scheme without the need for the ephemeral initial states.
+For completeness, we include here proof of the convergence of the QBD-RAP scheme to the fluid queue without the need for the ephemeral initial states.
 
@@
 
+Theorem~\ref{thm: a thm2!} below is analogous to Theorem~\ref{thm: a thm!} and proves a certain convergence of the QBD-RAP scheme to the fluid queue in the case that \(\phi(0)\in\mathcal S_{+0}\cup\calS_{-0}\).
+Like the proof of Theorem~\ref{thm: a thm!}, the proof of Theorem~\ref{thm: a thm2!} relies on technical bounds which we establish with the remainder of this Appendix.
+
 \begin{thm}\label{thm: a thm2!}
-	As \(p\to \infty\) the following convergence statements hold.
-	For \(r\in\{+,-\},\, q\in\{+0,-0\}\) and \(m=0\),  
+	As \(p\to \infty\), for \(q\in\{+0,-0\},\, r\in\{+,-\}\) and \(m=0\),  
 	\begin{align}\int_{x\in\calD_{\ell_0}}\widehat f_{0,q,r}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\psi(x)\wrt x \to \int_{x\in\calD_{\ell_0}}\widehat \mu_{0,q,r}^{\ell_0}(\lambda)(x,j;x_0,k)\psi(x)\wrt x.\end{align}
-	For \(r\in\{+,-\},\, q\in\{+0,-0\}\) and \(m\geq 1\), 
+	For \(q\in\{+0,-0\}, r\in\{+,-\}\) and \(m\geq 1\), 
 	\begin{align}\int_{x\in\calD_{\ell_0}}\widehat f_{m,q,r}^{\ell_0,(p)}(\lambda)(x,j;x_0,k)\psi(x)\wrt x \to \int_{x\in\calD_{\ell_0}}\widehat \mu_{m,q,r}^{\ell_0}(\lambda)(x,j;x_0,k)\psi(x)\wrt x.\label{eqn: thm 22}\end{align}
@@
 \section{Technical results}
+As in Appendix~\ref{app:tand}, we treat the cases of \(m=0\) and \(m\geq 1\) seperately.
+We start with the case \(m=0\).
+The following result is analgous to Lemma~\ref{lem: Dcoajc}, though the proof is somewhat more tedious.
 
@@
 \begin{proof}
-	Assume, without loss of generality \(\ell_0=0\) so \(\calD_{\ell_0}=[0,\Delta]\). Substituting the definition of \(\bs D\), then the left-hand side of (\ref{eqn: LLLlllL}) is
+	Assume, for simplicity and without loss of generality, that \(\ell_0=0\) so \(\calD_{\ell_0}=[0,\Delta]\).
+Substituting the definition of \(\bs D\) and exchanging the order of integration, then the left-hand side of (\ref{eqn: LLLlllL}) is
 	\begin{align}
-%		&\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\bs k^{(p)}(x_0) \bs D^{(p)} e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\nonumber
-%		 - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber
-		%
-		&\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u=0}^\infty \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\nonumber
-		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
-		%
-		\\&\leq \Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\nonumber
-		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
-		%
-		\\&\qquad {}+\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1\Bigg| \label{eqn: llllL}
+		&\Bigg|\int_{u=0}^\infty \bs k(x_0)e^{\bs Su}\bs s \int_{x=0}^{\Delta}  \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x_1 \wrt x \wrt u\nonumber
+		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg|.\label{eqn: aksfKJJJJJ}
 	\end{align}
+	We wish to apply Property~\ref{properties: 2} to the integral over \(x_1\), however, to do so requires that \(u\leq \Delta-\varepsilon\).
+Breaking up the integral with respect to \(u\), then (\ref{eqn: aksfKJJJJJ}) is equal to
+	\begin{align}
+		&\Bigg|\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \int_{x=0}^{\Delta}  \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x_1 \wrt x \wrt u + d_1 \nonumber
+		\\&{}  - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg|
+		\\&\leq \Bigg|\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \int_{x=0}^{\Delta}  \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x_1 \wrt x \wrt u\nonumber
+		\\&{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| +|d_1| \label{eqnL MMMM}
+	\end{align}
+	where 
+	\[|d_1|=\left|\int_{u=\Delta-\varepsilon}^\infty \bs k(x_0)e^{\bs Su}\bs s \int_{x=0}^{\Delta}  \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x_1 \wrt x \wrt u\right|.\]
+	We show later that \(|d_1|\) can be made arbitrarily small by choosing \(Z\) with sufficiently small variance.
+For now, let us focus on the first absolute value in (\ref{eqnL MMMM}).
+By Property~\ref{properties: 2} the first absolute value in (\ref{eqnL MMMM}) is
+	\begin{align}
+		&\Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \left[ h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) + r_{\bs v}(u,x) \right] \psi(x) \wrt u \wrt x \nonumber
+		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg|\nonumber
+		\\&\leq \Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) \psi(x) \wrt u \wrt x \nonumber 
+		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| + |d_2|,
+		\label{eqn: aKK}
+	\end{align}
+	where 
+	\[|d_2| = \left| \int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s r_{\bs v}(u,x) \psi(x) \wrt u \wrt x \right|.\]
+	We show later that \(|d_2|\) can be made arbitrarily small by choosing \(Z\) with sufficiently small variance. The remaining term in (\ref{eqn: aKK}) can be written as 
+	\begin{align}
+		&\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \wrt x 
+		- \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg|.\label{eqn: FGHJKL}
+	\end{align}
+
+	Intuitively, when the variance of \(Z\) is low, we expect a significant constribution to the integral with respect to \(u\) in (\ref{eqn: FGHJKL}) to come from the portion of the integral over the interval \((\Delta-x_0-\varepsilon,\Delta-x_0+\varepsilon)\).
+Although, the integral with respect to \(u\) only contain this interval when \(x\) is sufficiently small.
+Breaking up the integral with respect to \(u\) in (\ref{eqn: FGHJKL}) into an integral over the interval \((\Delta-x_0-\varepsilon,\Delta-x_0+\varepsilon)\) and integrals over the rest, then applying the triangle inequality, (\ref{eqn: FGHJKL}) is less than or equal to
+	\begin{align}
+		&\Bigg| \int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u 1(x\leq x_0-2\varepsilon) \wrt x\nonumber
+		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| + |d_3|+|d_4|+|d_5|,\label{eqn: LLkkK} 
+	\intertext{where}
+		&|d_3|=\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\min(\Delta-x_0-\varepsilon,\Delta-x-\varepsilon)} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \wrt x\Bigg|,\nonumber
+		\\&|d_4|=\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \nonumber 
+		 1(x\leq x_0-2\varepsilon)\wrt x\Bigg|,\nonumber
+		\\&|d_5|=\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \times 1(x\in[x_0,x_0-2\varepsilon))\wrt x\Bigg|.
+\nonumber
+	\end{align}
+	We show later that \(|d_3|,\, |d_4|\) and \(|d_5|\) can be made arbitrarily small by choosing \(Z\) with sufficiently small variance.
+
+	In the first integral with respect to \(x\) in (\ref{eqn: LLkkK}), since \(x_0\in(2\varepsilon,\Delta-\varepsilon)\), then we can absorb the indicator function into the limits of the integral which results in
+	\begin{align}
+		\int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u 1(x\leq x_0-2\varepsilon) \wrt x.
+	\end{align}
+	With this, and breaking up the integral over \(h_{ij}^{--}\), we can write the first absolute value in (\ref{eqn: LLkkK}) as 
+	\begin{align}
+		\nonumber &\Bigg| \int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u  \wrt x 
+		\\\nonumber &\qquad{} - \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x  - \int_{x=x_0-2\varepsilon}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| 
+		\\\nonumber &\leq \Bigg| \int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u   \wrt x 
+		\\&\qquad{} - \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| + |d_6|,\label{eqn:PPPPPP}
+	\end{align}
+		where
+	\begin{align}
+		&|d_6|=\left|\int_{x=x_0-2\varepsilon}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \right|.\nonumber 
+	\end{align}
+	We show later that \(|d_6|\) can be made arbitrarily small by choosing \(Z\) with sufficiently small variance.
+
+	Now, since probability densities integrate to 1, then we can write 
+	\begin{align*}
+		\int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x
+		&=\int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\mathbb P(|Z-\Delta|\leq \varepsilon\mid Z>x_0) 
+		\\&{}+ \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\mathbb P(|Z-\Delta|> \varepsilon\mid Z>x_0)
+		\\&=\int_{x=0}^{x_0-2\varepsilon} \int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt u\wrt x
+		\\&{}+ \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\mathbb P(|Z-\Delta|> \varepsilon\mid Z>x_0).
+	\end{align*}
+	Therefore, the first absolute value in (\ref{eqn:PPPPPP}) can be written as 
+	\begin{align}
+		\nonumber &\Bigg| \int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u   \wrt x 
+		\\\nonumber &\qquad{} - \int_{x=0}^{x_0-2\varepsilon} \int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt u\wrt x
+		\\\nonumber &\qquad{}- \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\mathbb P(|Z-\Delta|> \varepsilon\mid Z>x_0)\Bigg|
+		\\\nonumber &=\Bigg| \int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \left[h_{ij}^{--}(\lambda, \Delta - u - x) -  h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\right]\wrt u\wrt x
+		\\\nonumber &\qquad{}- \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\mathbb P(|Z-\Delta|> \varepsilon\mid Z>x_0)\Bigg|
+		\\&\leq \Bigg| \int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \left[h_{ij}^{--}(\lambda, \Delta - u - x) -  h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\right]\wrt u\wrt x\Bigg| + |d_7| \nonumber
+		\\&\leq  \int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \left|h_{ij}^{--}(\lambda, \Delta - u - x) -  h_{ij}^{--}(\lambda, x_0 - x)\left|\right|\psi(x)\right|\wrt u\wrt x + |d_7|, \label{eqn: ajkhvakjhdannvnnvnvn}
+	\end{align}
+	where 
+	\[|d_7|=\left| \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\mathbb P(|Z-\Delta|> \varepsilon\mid Z>x_0)\right|.\]
+	We show later that \(|d_7|\) can be made arbitrarily small by choosing \(Z\) with sufficiently small variance.
+
+	What remains is to bound the terms \(|d_\ell|,\, \ell=1,\dots,7\).
+
+	Since \(|\psi|\leq F\) then 
+	\begin{align}
+		|d_1|&\leq \left|\int_{u=\Delta-\varepsilon}^\infty \bs k(x_0)e^{\bs Su}\bs s \int_{x=0}^{\Delta}  \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1)  \wrt x_1 \wrt x \wrt u\right|F.\label{eqn:  CVBNM}
+	\end{align}
+	From Property~\ref{properties: -2}, (\ref{eqn:  CVBNM}) is less than or equal to 
+	\begin{align}
+		\nonumber &\left|\int_{u=\Delta-\varepsilon}^\infty \bs k(x_0)e^{\bs Su}\bs s  \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs e} h_{ij}^{--}(\lambda, x_1) \wrt x_1 \wrt u\right|F
+		\\\nonumber &\leq\left|\int_{u=\Delta-\varepsilon}^\infty \bs k(x_0)e^{\bs Su}\bs s  \int_{x_1=0}^\infty \bs k(u) {\bs e} h_{ij}^{--}(\lambda, x_1) \wrt x_1 \wrt u\right|F
+		\\&=\left|\int_{u=\Delta-\varepsilon}^\infty \bs k(x_0)e^{\bs Su}\bs s  \int_{x_1=0}^\infty h_{ij}^{--}(\lambda, x_1) \wrt x_1 \wrt u\right|F\label{eqb:afsgh}
+	\end{align}
+	since \(\bs k(u)e^{\bs Sx_1}\bs e\) is decreasing in \(x_1\) and \(\bs k(u)\bs e=1\).
+Now, as \(h_{ij}^{--}(\lambda,x_1)\) is integrable with \(\int_{x_1=0}^\infty h_{ij}^{--}(\lambda, x_1)\wrt x_1\leq \widehat{ G}\), then (\ref{eqb:afsgh}) is less than or equal to
+	\begin{align}
+		\nonumber \left|\int_{u=\Delta-\varepsilon}^\infty \bs k(x_0)e^{\bs Su}\bs s \wrt u\right|\widehat GF
+		&=\mathbb P(Z>x_0+\Delta-\varepsilon\mid Z>x_0)\widehat G F
+		\\&\leq\cfrac{\var{(Z)}/\varepsilon^2}{1-\var{(Z)}/\varepsilon^2} \widehat G F,\nonumber 
+	\end{align}
+	by Chebyshev's inequality, since \(x_0\in(2\varepsilon,\Delta-\varepsilon)\).
+
+	Since \(h_{ij}^{--}\) is Lipschitz and \(|(\Delta-u-x)-(x_0-x)|\leq 2\varepsilon\) for all \(u\in(\Delta-x_0-\varepsilon,\Delta-x_0+\varepsilon)\), then the first absolute value of (\ref{eqn: ajkhvakjhdannvnnvnvn}) is less than or equal to
+	\begin{align}
+		&\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u 2L\varepsilon \int_{x=0}^{x_0-2\varepsilon} \left|\psi(x)\right|\wrt x. \label{eqn: the last one?}
+	\end{align}
+	Now, \(\displaystyle \int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u\leq 1\) as it is a probability, and \(\displaystyle \int_{x=0}^{x_0-2\varepsilon} \left|\psi(x)\right|\wrt x \leq F\Delta\) as \(|\psi|\leq F\) and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\).
+Therefore, (\ref{eqn: the last one?}) is less than or equal to \(2\varepsilon LF\Delta.\)
+
 	Since \(|\psi(x)|\leq F\) and \(|h_{ij}^{--}(\lambda,x)|\leq G\), the third term in (\ref{eqn: llllL}) is less than or equal to 
@@
 	The first and second terms in (\ref{eqn: llllL}) are
-	\begin{align}
-		&\Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \int_{x_1=0}^\infty \bs k(u)  e^{\bs S x_1} {\bs v}(x) h_{ij}^{--}(\lambda, x_1) \wrt x_1 \wrt u \psi(x) \wrt x \nonumber
-		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
-		%
-		\\&=\Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \left[ h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) + r_{\bs v}(u,x) \right] \psi(x) \wrt u \wrt x \nonumber
-		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg|\nonumber
-		\\&\leq \Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) \psi(x) \wrt u \wrt x \nonumber 
-		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| + \Bigg| \int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s r_{\bs v}(u,x) \psi(x) \wrt u \wrt x \Bigg|
-		\label{eqn: aKK}
-	\end{align}
+
 	where the first equality holds from Property \ref{properties: 2}. The last term in (\ref{eqn: aKK}) is less than or equal to 
@@
 	by Property \ref{properties: 2}. The first two terms in (\ref{eqn: aKK}) are
-	\begin{align}
-		& \Bigg|\int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x)1(u+x\leq \Delta-\varepsilon) \psi(x) \wrt u \wrt x \nonumber
-		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber 
-		\\&= \Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \wrt x \nonumber
-		  \\&\qquad{}- \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber 
-		\\&\leq \Bigg| \int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u 1(\Delta-x_0+\varepsilon\leq \Delta-x-\varepsilon) \wrt x\nonumber
-		\\&\qquad{} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber 
-		\\&\qquad {}+\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\min(\Delta-x_0-\varepsilon,\Delta-x-\varepsilon)} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \wrt x\Bigg|\nonumber
-		\\&\qquad {}+\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \nonumber 
-		\\&\qquad\qquad \times1(\Delta-x_0+\varepsilon\leq \Delta - x -\varepsilon)\wrt x\Bigg|\nonumber
-		\\&\qquad {}+\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u \nonumber
-		\\&\qquad\qquad \times 1(\Delta-x_0-\varepsilon\leq \Delta - x -\varepsilon < \Delta-x_0+\varepsilon)\wrt x\Bigg|   \label{eqn: LLkkK}
-	\end{align}
+
 	Since \(|\psi|\leq F\), the second and third terms in (\ref{eqn: LLkkK}) are less than or equal to 
@@
 \subsection{One integral}
-\begin{lem} \label{lem:tttttt}
-	Let \(g\) satisfy the Assumptions \ref{asu: g} and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\). Then
-	\begin{align}
-		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}4G + 3L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2}.
-	\end{align}
-\end{lem}
-\begin{proof}
-	First rewrite the left-hand side as 
-	\begin{align}
-%		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| 
-		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}(g(x)-g(x_0))\bs s\wrt x \right|
-		%
-		&\leq \int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x.
-	\end{align}
-	Substituting in the expression for \(\bs D\) gives,
-	\begin{align}
-		&\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x  \nonumber 
-%		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x 
-		%
-		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \nonumber
-		\\&\quad {} + \int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \label{eqn: 2nd here}
-	\end{align}
-	Since \(g\) is bounded, the second term is less than or equal to 
-	\begin{align}
-		\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x 2G \nonumber
-		&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u \bs e 2G \nonumber
-		\\&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \wrt u2G \nonumber
-		\\&= \cfrac{\mathbb P(Z\geq x_0+\Delta-\varepsilon)}{\mathbb P(Z>x_0)}2G \label{eqn" dgkjlwerhv}
-	\end{align}
-	For \(x_0\in(2\varepsilon,\Delta-\varepsilon),\) then (\ref{eqn" dgkjlwerhv}) is less than or equal to 
-	\[\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}2G.\]
+% \begin{lem} \label{lem:tttttt}
+% 	Let \(g\) satisfy the Assumptions \ref{asu: g} and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\). Then
+% 	\begin{align}
+% 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}4G + 3L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2}.
+% 	\end{align}
+% \end{lem}
+% \begin{proof}
+% 	First rewrite the left-hand side as 
+% 	\begin{align}
+% %		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| 
+% 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}(g(x)-g(x_0))\bs s\wrt x \right|
+% 		%
+% 		&\leq \int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x.
+% 	\end{align}
+% 	Substituting in the expression for \(\bs D\) gives,
+% 	\begin{align}
+% 		&\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x  \nonumber 
+% %		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x 
+% 		%
+% 		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \nonumber
+% 		\\&\quad {} + \int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \label{eqn: 2nd here}
+% 	\end{align}
+% 	Since \(g\) is bounded, the second term is less than or equal to 
+% 	\begin{align}
+% 		\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x 2G \nonumber
+% 		&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u \bs e 2G \nonumber
+% 		\\&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \wrt u2G \nonumber
+% 		\\&= \cfrac{\mathbb P(Z\geq x_0+\Delta-\varepsilon)}{\mathbb P(Z>x_0)}2G \label{eqn" dgkjlwerhv}
+% 	\end{align}
+% 	For \(x_0\in(2\varepsilon,\Delta-\varepsilon),\) then (\ref{eqn" dgkjlwerhv}) is less than or equal to 
+% 	\[\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}2G.\]
 
-	As for the first term in (\ref{eqn: 2nd here}), it can be written as 
-	\begin{align}
-		&\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
-		%
-		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
-		%
-		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x.
+% 	As for the first term in (\ref{eqn: 2nd here}), it can be written as 
+% 	\begin{align}
+% 		&\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
+% 		%
+% 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
+% 		%
+% 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x.
 \label{eqn: 201}
-	\end{align}
-	Since \(g\) is bounded, then the last two terms in (\ref{eqn: 201}) are 
-	\begin{align}
-		&2G\Bigg(\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x \nonumber
-		%
-		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x\Bigg) \nonumber
-		%
-		\\&=2G\Bigg(\bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e}  \bs e  \wrt u\nonumber
-		%
-		+ \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \bs e\wrt u \Bigg) \nonumber
-		%
-		\\&=2G \cfrac{\mathbb P(Z>x_0, Z\notin(\Delta-\varepsilon, \Delta+\varepsilon))}{\mathbb P(Z>x_0)}  \nonumber 
-		%
-		\\&\leq 2G\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2},
-	\end{align}
-	provided that \(x_0\in[0,\Delta-\varepsilon)\). 
-	Exchanging the order of integration for the first term in (\ref{eqn: 201})
-	\begin{align}
-		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\int_{x=0}^\infty \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \wrt u \label{eqn: fdk13897}
-	\end{align}
-	from which we see that we can apply Corollary~\ref{cor: cond bnd} to the integral over \(x\), implying that (\ref{eqn: fdk13897}) is less than or equal to
-	\begin{align}
-		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(|g(\Delta-u)-g(x_0)|+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u.
+% 	\end{align}
+% 	Since \(g\) is bounded, then the last two terms in (\ref{eqn: 201}) are 
+% 	\begin{align}
+% 		&2G\Bigg(\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x \nonumber
+% 		%
+% 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x\Bigg) \nonumber
+% 		%
+% 		\\&=2G\Bigg(\bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e}  \bs e  \wrt u\nonumber
+% 		%
+% 		+ \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \bs e\wrt u \Bigg) \nonumber
+% 		%
+% 		\\&=2G \cfrac{\mathbb P(Z>x_0, Z\notin(\Delta-\varepsilon, \Delta+\varepsilon))}{\mathbb P(Z>x_0)}  \nonumber 
+% 		%
+% 		\\&\leq 2G\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2},
+% 	\end{align}
+% 	provided that \(x_0\in[0,\Delta-\varepsilon)\). 
+% 	Exchanging the order of integration for the first term in (\ref{eqn: 201})
+% 	\begin{align}
+% 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\int_{x=0}^\infty \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \wrt u \label{eqn: fdk13897}
+% 	\end{align}
+% 	from which we see that we can apply Corollary~\ref{cor: cond bnd} to the integral over \(x\), implying that (\ref{eqn: fdk13897}) is less than or equal to
+% 	\begin{align}
+% 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(|g(\Delta-u)-g(x_0)|+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u.
 \label{eqn: sdkagh lkhvasfv}
-	\end{align}
-	Since \(g\) is Lipschitz, then (\ref{eqn: sdkagh lkhvasfv}) is less than or equal to 
-	\begin{align}
-		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u \leq L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon.
+% 	\end{align}
+% 	Since \(g\) is Lipschitz, then (\ref{eqn: sdkagh lkhvasfv}) is less than or equal to 
+% 	\begin{align}
+% 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u \leq L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon.
 \label{eqn: sdkagh lsfv}
-	\end{align}
-	Putting all the bounds together proves the result. 
-\end{proof}
+% 	\end{align}
+% 	Putting all the bounds together proves the result. 
+% \end{proof}
 

### 2021-11-26 appendix3/appendix.tex
update apendicies a,b,c

@@
 \chapter{Convergence without ephemeral phases\label{app:extend conv}}
-For completeness, we include here proof of the convergence of the QBD-RAP scheme to the fluid queue without the need for the ephemeral initial states.
+For completeness, we include here results needed to prove the convergence of the convergence of the QBD-RAP scheme to the fluid queue without the need for the ephemeral initial states.
+Note that we only need to prove that the convergence up to, and at, the time of the first change of level of the QBD-RAP.
 
@@
 \section{Technical results}
-As in Appendix~\ref{app:tand}, we treat the cases of \(m=0\) and \(m\geq 1\) seperately.
-We start with the case \(m=0\).
+As we did in Appendix~\ref{app:tand}, we treat the cases of no changes, and \(m\geq 1\) changes, of phase from \(\calS_+\) to \(\calS_-\) and \(\calS_-\) to \(\calS_+\) seperately.
+We start with the \emph{no changes} case.
 The following result is analgous to Lemma~\ref{lem: Dcoajc}, though the proof is somewhat more tedious.
@@
 
+	Since \(h_{ij}^{--}\) is Lipschitz and \(|(\Delta-u-x)-(x_0-x)|\leq 2\varepsilon\) for all \(u\in(\Delta-x_0-\varepsilon,\Delta-x_0+\varepsilon)\), then the first absolute value of (\ref{eqn: ajkhvakjhdannvnnvnvn}) is less than or equal to
+	\begin{align}
+		&\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u 2L\varepsilon \int_{x=0}^{x_0-2\varepsilon} \left|\psi(x)\right|\wrt x. \label{eqn: the last one?}
+	\end{align}
+	Now, \(\displaystyle \int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u\leq 1\) as it is a probability, and \(\displaystyle \int_{x=0}^{x_0-2\varepsilon} \left|\psi(x)\right|\wrt x \leq F\Delta\) as \(|\psi|\leq F\) and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\).
+Therefore, (\ref{eqn: the last one?}) is less than or equal to \(2\varepsilon LF\Delta.\)
+
 	What remains is to bound the terms \(|d_\ell|,\, \ell=1,\dots,7\).
 
+	%% d1 %%
 	Since \(|\psi|\leq F\) then 
@@
 
-	Since \(h_{ij}^{--}\) is Lipschitz and \(|(\Delta-u-x)-(x_0-x)|\leq 2\varepsilon\) for all \(u\in(\Delta-x_0-\varepsilon,\Delta-x_0+\varepsilon)\), then the first absolute value of (\ref{eqn: ajkhvakjhdannvnnvnvn}) is less than or equal to
+	%% d2 %% 
+	Since \(|\psi(x)|\leq F\), then
 	\begin{align}
-		&\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u 2L\varepsilon \int_{x=0}^{x_0-2\varepsilon} \left|\psi(x)\right|\wrt x. \label{eqn: the last one?}
+		|d_2| &\leq \left| \int_{x=0}^{\Delta}\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s r_{\bs v}(u,x) \wrt u \wrt x \right|F
+		%
+		\\&\leq \left| \int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u \right|R_{\bs v,2}F
+		%
+		\\&\leq R_{\bs v,2}F,
 	\end{align}
-	Now, \(\displaystyle \int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u\leq 1\) as it is a probability, and \(\displaystyle \int_{x=0}^{x_0-2\varepsilon} \left|\psi(x)\right|\wrt x \leq F\Delta\) as \(|\psi|\leq F\) and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\).
-Therefore, (\ref{eqn: the last one?}) is less than or equal to \(2\varepsilon LF\Delta.\)
+	where the first inequality holds from Property~\ref{properties: 2}, and the last inequality holds since \(\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u=\mathbb P(Z\in (x_0,x_0+\Delta-\varepsilon)\mid Z>x_0)\leq 1\).
 
-	Since \(|\psi(x)|\leq F\) and \(|h_{ij}^{--}(\lambda,x)|\leq G\), the third term in (\ref{eqn: llllL}) is less than or equal to 
+	%% d3 %%
+	Since \(|\psi(x)|\leq F\) and \(h_{ij}^{--}(\lambda,\Delta-u-x)\leq G\), then
 	\begin{align}
-		&\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u e^{\bs S x_1} {\bs v}(x) \wrt x \wrt x_1G F.\label{eqn: a big lola}
+		|d_3|&\leq \int_{u=0}^{\min(\Delta-x_0-\varepsilon,\Delta-x-\varepsilon)} \bs k(x_0)e^{\bs Su}\bs s \wrt u \Delta GF
+		\\ &\leq \mathbb P(Z\leq\Delta-\varepsilon \mid Z>x_0) \wrt u \Delta GF\nonumber
+		\\ &\leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2} \Delta GF,\nonumber
 	\end{align}
-	Computing the integral over \(x_1\) in (\ref{eqn: a big lola}) gives
+	where the last inequality holds from Chebyshev's inequality.
+
+	%% d4 %%
+	Similarly, since \(|\psi(x)|\leq F\) and \(h_{ij}^{--}(\lambda,\Delta-u-x)\leq G\), then
 	\begin{align} 
-		& \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \bs k(u) \wrt u (-\bs S)^{-1}{\bs v}(x) \wrt x G F\nonumber
-		%
-		\\&\leq  \int_{x=0}^{\Delta}\int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s \wrt u \wrt x G_{\bs v} G F \nonumber
-		%
-		\\&= \int_{u= \Delta-\varepsilon}^\infty  \bs k(x_0)e^{\bs Su}\bs s  \wrt u\Delta G_{\bs v} G F.\label{eqn: google the yes}
+		|d_4|&\leq \int_{x=0}^{x_0-2\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u \wrt x GF \nonumber
+			\\& = \int_{x=0}^{x_0-2\varepsilon}\mathbb P(Z\in[\Delta+\varepsilon, \Delta-x-\varepsilon + x_0]\mid Z>x_0) \wrt x GF \nonumber
+			\\& \leq \int_{x=0}^{x_0-2\varepsilon}\mathbb P(Z\in[\Delta+\varepsilon, \Delta-\varepsilon + x_0]\mid Z>x_0) \wrt x GF \nonumber
+			\\& \leq \mathbb P(Z\in[\Delta+\varepsilon, \Delta-\varepsilon + x_0]\mid Z>x_0)  \Delta GF \nonumber
+			\\& \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2} \Delta GF \nonumber
 	\end{align}
-	since, by property \ref{properties: 1}, \(\bs k(u)(-\bs S)^{-1}{\bs v}(x)\leq \bs k(u)\bs e G_{\bs v}=G_{\bs v}\). The bound (\ref{eqn: google the yes}) is equal to 
+
+	%% d5 %%
+	Since \(|\psi(x)|\leq F\) and \(h_{ij}^{--}(\lambda,\Delta-u-x)\leq G\), then
 	\begin{align} 
-		\Delta \mathbb P(Z>x_0 + \Delta-\varepsilon) \wrt x G_{\bs v} G F
-		\leq  \Delta \cfrac{\var(Z)}{(x_0-\varepsilon)^2} G_{\bs v} G F, \label{eqn: JjJ}
+		|d_5|&\leq \int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u 1(x\in[x_0,x_0-2\varepsilon))\wrt xGF
+		\\&\leq \int_{x=0}^{\Delta-\varepsilon} 1(x\in[x_0,x_0-2\varepsilon))\wrt xGF  \nonumber 
+		\\&= 2\varepsilon GF,  \nonumber 
 	\end{align}
-	by Chebyshev's inequality.
+	where the second inequality holds since \(\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \wrt u\leq 1\).
 
-	The first and second terms in (\ref{eqn: llllL}) are
+	%% d6 %% 
+	Since \(|\psi(x)|\leq F\) and \(h_{ij}^{--}(\lambda,\Delta-u-x)\leq G\), then
+	\begin{align}
+		&|d_6| \leq \int_{x=x_0-2\varepsilon}^{x_0} \wrt x GF\nonumber 
+		\\&= 2\varepsilon GF.
+	\end{align}
 
-	where the first equality holds from Property \ref{properties: 2}. The last term in (\ref{eqn: aKK}) is less than or equal to 
+	%% d7 %% 
+	Since \(|\psi(x)|\leq F\) and \(h_{ij}^{--}(\lambda,\Delta-u-x)\leq G\) then
 	\begin{align}
-		\int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su} \int_{x=0}^{\Delta} \bs s \left|r_{\bs v}(u,x)\right|\wrt x \wrt u F 
-		%
-		&\leq \int_{u=0}^{\Delta-\varepsilon} \bs k(x_0)e^{\bs Su} \bs sR_{{\bs v},2} \wrt u F \nonumber 
-		%
-		\\&\leq R_{{\bs v},2} F, \label{eqn: HhHj}
+		|d_7|&\leq \mathbb P(|Z-\Delta|> \varepsilon\mid Z>x_0)\Delta GF
+		\\&\leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}\Delta GF
 	\end{align}
-	by Property \ref{properties: 2}. The first two terms in (\ref{eqn: aKK}) are
 
-	Since \(|\psi|\leq F\), the second and third terms in (\ref{eqn: LLkkK}) are less than or equal to 
-	\begin{align}
-		&\int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\min(\Delta-x_0-\varepsilon,\Delta-x-\varepsilon)} \bs k(x_0)e^{\bs Su}\bs s  \wrt u \wrt xGF \nonumber
-		\\&\qquad {}+\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u 1(\Delta-x_0+\varepsilon\leq \Delta - x -\varepsilon)\wrt x GF \nonumber
-		\\&\leq \int_{x=0}^{\Delta-\varepsilon}\int_{u=0}^{\Delta-x_0-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u \wrt xGF \nonumber
-		\\&\qquad {}+\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0+\varepsilon}^{\infty} \bs k(x_0)e^{\bs Su}\bs s  \wrt u 1(\Delta-x_0+\varepsilon\leq \Delta - x -\varepsilon)\wrt x GF \nonumber
-		\\&\leq\Delta\int_{u=0}^{\Delta-x_0-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s  \wrt u \wrt xGF 
-		+ \Delta \int_{u=\Delta-x_0+\varepsilon}^{\infty} \bs k(x_0)e^{\bs Su}\bs s  \wrt u GF \nonumber
-		\\&\leq \Delta \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}GF. \label{eqn:aaAAAAd}
-	\end{align}
-	Since \(\displaystyle \int_{u=\Delta-x_0-\varepsilon}^{\Delta-x-\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \leq 1\), the last term in (\ref{eqn: LLkkK}) is less than or equal to 
-	\begin{align}
-		&\int_{x=0}^{\Delta-\varepsilon}GF \wrt u 
-		 1(\Delta-x_0-\varepsilon\leq \Delta - x -\varepsilon < \Delta-x_0+\varepsilon)\wrt x =2\varepsilon GF. \label{eqn:Qwe}
-	\end{align}
-	Now consider the difference in the first absolute value in (\ref{eqn: LLkkK}),
-	\begin{align}
-		&\Bigg|\int_{x=0}^{\Delta-\varepsilon}\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u 1(x\leq x_0-2\varepsilon) \wrt x \nonumber
-		\\&{}\qquad {} - \int_{x=0}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg|\nonumber
-		%
-		\\&\leq \Bigg|\int_{x=0}^{ x_0-2\varepsilon }\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s h_{ij}^{--}(\lambda, \Delta - u - x) \psi(x) \wrt u  \wrt x \nonumber
-		\\&{}\qquad {} - \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| + \Bigg| \int_{x=x_0-2\varepsilon}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x\Bigg| \nonumber
-		%
-		\\&\leq  \int_{x=0}^{ x_0-2\varepsilon }\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s \Bigg| h_{ij}^{--}(\lambda, \Delta - u - x) - h_{ij}^{--}(\lambda, x_0 - x) \Bigg| \wrt u \left|\psi(x)\right| \wrt x  \nonumber
-		\\&{}\qquad {} + \Bigg| \int_{x=0}^{x_0-2\varepsilon} h_{ij}^{--}(\lambda, x_0 - x)\psi(x) \wrt x \mathbb P(|Z-\Delta|>\varepsilon)\Bigg| \nonumber 
-		\\&\qquad{} + \Bigg|\int_{x=x_0-2\varepsilon}^{x_0} h_{ij}^{--}(\lambda, x_0 - x)\psi(x)\wrt x \Bigg| \nonumber
-		%
-		\\&\leq \int_{x=0}^{ x_0-2\varepsilon }\int_{u=\Delta-x_0-\varepsilon}^{\Delta-x_0+\varepsilon} \bs k(x_0)e^{\bs Su}\bs s L\varepsilon  \wrt u \left|\psi(x)\right| \wrt x \nonumber
-		+\Delta GF \cfrac{\var(Z)}{\varepsilon^2} + 2\varepsilon GF
-		%
-		\\&\leq \Delta L\varepsilon   F 
-		+\Delta GF \cfrac{\var(Z)}{\varepsilon^2} + 2\varepsilon GF.\label{eqn: LLLaase}
-	\end{align}
-	In summary, we have shown 
-	\begin{align}
-		&\Bigg|\int_{x_1=0}^\infty \int_{x=0}^{\Delta}\bs k^{(p)}(x_0) \bs D^{(p)} e^{\bs S^{(p)} x_1} {\bs v}^{(p)}(x) h_{ij}^{--}(\lambda, x_1) \psi(x) \wrt x \wrt x_1 \nonumber
-		\\&\qquad{}- \int_{x=0}^{x_0} h_{ij}^{--}(\lambda,x_0-x)\psi(x)\wrt x \Bigg| \nonumber 
-		%
-		\\& \leq  \Delta L\varepsilon^{(p)}   F 
-		+\Delta GF \cfrac{\var(Z^{(p)})}{\left(\varepsilon^{(p)}\right)^2} + 4\varepsilon^{(p)} GF + \Delta \cfrac{\var(Z^{(p)})/\left(\varepsilon^{(p)}\right)^2}{1-\var(Z^{(p)})/\left(\varepsilon^{(p)}\right)^2}GF + R_{{\bs v},2}^{(p)} F \nonumber
-		\\&\qquad{} +  \Delta \cfrac{\var(Z^{(p)})}{(x_0-\varepsilon^{(p)})^2} G_{\bs v} G F.
-	\end{align}
-	The result follows upon choosing \(\varepsilon^{(p)}=\var(Z^{(p)})^{1/3}\) and letting \(p\to\infty\). 
+	Convergence follows after setting \(\varepsilon^{(p)}=\var(Z^{(p)})^{1/3}\) and observing that all the bounds \(|d_1|,\dots,|d_7|\) tend to 0, as does the bound on (\ref{eqn: the last one?}), given by \(2\varepsilon^{(p)}LF\Delta\), as \(p\to\infty\).
 \end{proof}
 
-\subsection{One integral}
-% \begin{lem} \label{lem:tttttt}
-% 	Let \(g\) satisfy the Assumptions \ref{asu: g} and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\). Then
-% 	\begin{align}
-% 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}4G + 3L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2}.
-% 	\end{align}
-% \end{lem}
-% \begin{proof}
-% 	First rewrite the left-hand side as 
-% 	\begin{align}
-% %		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| 
-% 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}(g(x)-g(x_0))\bs s\wrt x \right|
-% 		%
-% 		&\leq \int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x.
-% 	\end{align}
-% 	Substituting in the expression for \(\bs D\) gives,
-% 	\begin{align}
-% 		&\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x  \nonumber 
-% %		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x 
-% 		%
-% 		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \nonumber
-% 		\\&\quad {} + \int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \label{eqn: 2nd here}
-% 	\end{align}
-% 	Since \(g\) is bounded, the second term is less than or equal to 
-% 	\begin{align}
-% 		\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x 2G \nonumber
-% 		&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u \bs e 2G \nonumber
-% 		\\&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \wrt u2G \nonumber
-% 		\\&= \cfrac{\mathbb P(Z\geq x_0+\Delta-\varepsilon)}{\mathbb P(Z>x_0)}2G \label{eqn" dgkjlwerhv}
-% 	\end{align}
-% 	For \(x_0\in(2\varepsilon,\Delta-\varepsilon),\) then (\ref{eqn" dgkjlwerhv}) is less than or equal to 
-% 	\[\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}2G.\]
+	Now we show bounds for certain Laplace transform expressions which arise when the QBD-RAP starts in phases in \(\calS_{+0}\cup\calS_{-0}\) and there is more than one change from \(\calS_+\) to \(\calS_-\) or \(\calS_-\) to \(\calS_+\) before the first change of level.
+These expressions have the form
+	\begin{align}
+		\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
+            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
+		\bs D\right]
+            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x). \label{eqn: c30}
+	\end{align}
+	Here, we ultimately wish to show that (\ref{eqn: c30}) converges to \(g_{1,n}^*(\Delta-x_0,x)\).
+We do not do this directly, instead, we show that (\ref{eqn: c30}) is `close' to \(w_n(\Delta-x_0,x)\), then rely on the results from Appendix~\ref{app:tand} to get the desired convergence.
 
-% 	As for the first term in (\ref{eqn: 2nd here}), it can be written as 
-% 	\begin{align}
-% 		&\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
-% 		%
-% 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
-% 		%
-% 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x.
-\label{eqn: 201}
-% 	\end{align}
-% 	Since \(g\) is bounded, then the last two terms in (\ref{eqn: 201}) are 
-% 	\begin{align}
-% 		&2G\Bigg(\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x \nonumber
-% 		%
-% 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x\Bigg) \nonumber
-% 		%
-% 		\\&=2G\Bigg(\bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e}  \bs e  \wrt u\nonumber
-% 		%
-% 		+ \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \bs e\wrt u \Bigg) \nonumber
-% 		%
-% 		\\&=2G \cfrac{\mathbb P(Z>x_0, Z\notin(\Delta-\varepsilon, \Delta+\varepsilon))}{\mathbb P(Z>x_0)}  \nonumber 
-% 		%
-% 		\\&\leq 2G\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2},
-% 	\end{align}
-% 	provided that \(x_0\in[0,\Delta-\varepsilon)\). 
-% 	Exchanging the order of integration for the first term in (\ref{eqn: 201})
-% 	\begin{align}
-% 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\int_{x=0}^\infty \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \wrt u \label{eqn: fdk13897}
-% 	\end{align}
-% 	from which we see that we can apply Corollary~\ref{cor: cond bnd} to the integral over \(x\), implying that (\ref{eqn: fdk13897}) is less than or equal to
-% 	\begin{align}
-% 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(|g(\Delta-u)-g(x_0)|+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u.
-\label{eqn: sdkagh lkhvasfv}
-% 	\end{align}
-% 	Since \(g\) is Lipschitz, then (\ref{eqn: sdkagh lkhvasfv}) is less than or equal to 
-% 	\begin{align}
-% 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon\right)\wrt u \leq L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2} + 2 L \varepsilon.
-\label{eqn: sdkagh lsfv}
-% 	\end{align}
-% 	Putting all the bounds together proves the result. 
-% \end{proof}
+	Observe that by substituting the first matrix \(\bs D\) in the expression above for its integral expression, then (\ref{eqn: c30}) is equal to 
+	\begin{align}
+		&\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs s\cfrac{\bs\alpha e^{\bs Sz_0}}{\bs\alpha e^{\bs Sz_0}\bs e} \wrt z_0 e^{\bs{S}x_1}\wrt x_1\bs D 
+            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
+		\bs D\right]
+            	\\\nonumber&\quad\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) 
+	%
+		\\&=\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0.\label{eqn: kd djv}
+	\end{align}
+	Intuitively, when the variance of \(Z\) is low, we expect that the integral in (\ref{eqn: kd djv}) above will be approximately equal to \(w_n(\Delta-x_0,x)\).
+Indeed, we proved in Lemma~\ref{lemma:bound} that this is the case for functions \(g\) satisfying the Assumptions~\ref{asu: g}.
+However, here we do not immediately have that \(w_n(x_0,x)\) is Lipschitz as we would need for it to satisfy Assumptions~\ref{asu: g}.
+Instead, we can show a Lipschitz-like condition in \(x_0\) for \(w_n(x_0,x)\), which suffices.
 
-%\begin{lem}
-%	Let \(f:[0,\Delta)\to \mathbb R\) be any bounded, Lipschitz continuous function with \(|\psi(x)|\leq F\). Then for \(u\leq \Delta-\varepsilon\), \(v\in[0,\Delta)\)
-%	\begin{align}
-%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[f(\bar X(t))1(\varphi(t)=j, L(t)=\ell_0, L(s)=\ell_0, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid L(0)=\ell_0, \nonumber
-%		\\&\qquad{} \bs A(0)=\bs a_{\ell_0,i}(x_0), \varphi(0)=i]\wrt t \nonumber
-%		\\&= \int_{t=0}^\infty e^{-\lambda t}\mathbb E[\psi(X(t))1(\varphi(t)=j,X(s)\in\calD_{\ell_0}, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid X(0)=x_0, \varphi(0)=i]\wrt t \nonumber
-%		\\&\qquad {} + r_{11}^{(p)}
-%	\end{align}
-%	where 
-%	\[|r_{11}^{(p)}| \leq F R_{{\bs v},2}+ GF\varepsilon.\]
-%\end{lem}
-%\begin{proof}
-%	Assume, without loss of generality \(\ell_0=0\) so \(\calD_{\ell_0}=[0,\Delta]\). First observe that 
-%	\begin{align}
-%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[f(\bar X(t))1(\varphi(t)=j, L(t)=\ell_0, L(s)=\ell_0, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid L(0)=\ell_0, \nonumber
-%		\\&\qquad{} \bs A(0)=\bs a_{\ell_0,i}(x_0), \varphi(0)=i]\wrt t \nonumber 
-%		\\&= \int_{x_1=0}^\infty \int_{x=[0,\Delta)} \cfrac{\bs \alpha e^{\bs S(x_1+x_0+x)}\bs s}{\bs \alpha e^{\bs S x_0} \bs e}h_{ij}^{++}(\lambda, x_1)f(\Delta-x)\wrt x\wrt x_1
-%	\end{align}
-%	and 
-%	\begin{align}
-%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[\psi(X(t))1(\varphi(t)=j,X(s)\in\calD_{\ell_0}, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid X(0)=x_0, \varphi(0)=i]\wrt t \nonumber
-%		\\&=  \int_{x=0}^{\Delta-x_0} h_{ij}^{++}(\lambda, x)\psi(X_0 + x)\wrt x
-%	\end{align}
-%	In Property \ref{properties: 2} take \(g(x) = h_{ij}^{--}(\lambda, x)\), which states 
-%	\begin{align}
-%		&\int_{x_1=0}^\infty \int_{x=0}^\Delta\cfrac{\bs \alpha e^{\bs S(x_1+x_0+x)}\bs s}{\bs \alpha e^{\bs S x_0} \bs e}h_{ij}^{++}(\lambda, x_1)f(\Delta-x)\wrt x\wrt x_1 \nonumber
-%		\\& = \int_{x=0}^{\Delta}h_{ij}^{++}(\lambda,\Delta - x_0 - x)1(x+x_0\leq \Delta-\varepsilon)f(\Delta - x)\wrt x + \int_{x=0}^\Delta r_{\bs v}(x_0,x)\psi(x)\wrt x. 
-%	\end{align}
-%	The second term is less than or equal to 
-%	\[F\int_{x=0}^\Delta r_{\bs v}(x_0,x)\wrt x = FR_{{\bs v},1}.\]
-%	The first term is 
-%	\[\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x - \int_{x=\Delta-x_0-\varepsilon}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x.\]
-%	Now 
-%	\[\int_{x=\Delta-x_0-\varepsilon}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x\leq GF\varepsilon\]
-%	and 
-%	\[\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x=\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,x)\psi(x_0+x)\wrt x.\]
-%\end{proof}
-
-%\begin{cor}\label{cor: Dcoajc2222}
-%	Let \(\psi:\calD_{\ell_0}\to \mathbb R\) be bounded, \(|\psi(x)|\leq F\), and Lipschitz continuous.
-Then, for \(x_0\in(y_{\ell_0}, y_{\ell+1}),\) \(k\in\calS_{0+}\), \(i\in\calS_-\), \(j\in\calS_-\cup\calS_{-0}\), there exists \(r_{10}^{(p)}\to 0\) as \(p \to \infty\),
-%	\begin{align}
-%		\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)}.\label{eqn:xvasfv}
-%	\end{align}
-%	Similarly, for \(k\in\calS_{0-}\), \(i\in\calS_+\), \(j\in\calS_+\cup\calS_{+0}\)
-%	\begin{align}
-%		\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,+0,+}^{\ell_0}(x,i,j;j,x_0)\psi(x)\wrt x  - \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,+0,+}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)}.\label{eqn:!124msfvcds}
-%	\end{align}
-%\end{cor}
-%\begin{proof}[Proof of Corollary \ref{cor: Dcoajc}]
-%	We show the result for (\ref{eqn:xs}) only, with the result for (\ref{eqn:!124mcds}) following analogously. 
+For later, observe that 
+\begin{align}
+	g^*_{2,n}(u_1,x) &= \int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1 \dots \nonumber 
+         	\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
+       	\\&\qquad{}g_{n}(\Delta - x-u_{n-1})1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1} \nonumber
  %
-%	Observe that, for \(k\in\calS_{0+}\), \(i\in\calS_-\), \(j\in\calS_-\cup\calS_{-0}\), 
-%	\begin{align}
-%		&\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x \nonumber 
-%		%
-%		\\&{}= \vligne{\lambda \bs I - \bs T_{00}}^{-1}_{ki}\int_{x\in\calD_{\ell_0}}\int_{x_1=0}^\infty \bs k(x_0) \bs D e^{\bs S x_1} {\bs v}(x)h_{ij}^{--}(\lambda,x_1)\wrt x_1\psi(x)\wrt x,
-%	\end{align}
-%	and 
-%	\begin{align}
-%		\int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x \nonumber  = \vligne{\lambda \bs I - \bs T_{00}}^{-1}_{ki}\int_{x=0}^{x_0} h_{ij}^{--}(\lambda,x_0-x)\psi(x)\wrt x
-%	\end{align}
-%	These terms appear in the proof of Corollary~\ref{lem: ppp} and using arguments applied there we can show
-%	\begin{align}
-%		\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x ,\label{eqn:xssgwg}
-%	\end{align}
-%	and therefore there exists a bound \(r_{10}^{(p)}\) such that 
-%	\[\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)} ,\]
-%	and \(r_{10}^{(p)}\to 0\) as \(p\to\infty\). 
-%\end{proof}
-
-%\begin{cor}\label{corL Tta} For \(i\in\mathcal S_-,j\in\mathcal S_-\cup\calS_{-0}\), \(x_0\in[0,\Delta)\), 
-%	\begin{align}
-%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t \nonumber
-%		%
-%		\\& \to \int_{t=0}^\infty e^{-\lambda t}\mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) \wrt t \label{eqn: ffaA}
-%	\end{align}
-%	as \(p\to\infty\).
-%\end{cor}
-%\begin{proof}
-%	From the convergence of Laplace transforms in Lemma~\ref{lem: ppp} and the Continuity Theorem \ref{thm: ext cont thm} then 
-%	\begin{align}
-%		&  \mathbb E[f(\bar X^{(p)}(t))1(\varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j) \mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i]  \nonumber 
-%		\\& \to \mathbb E[\psi(X(t)) 1(\varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j) \mid X(0)=x_0, \varphi(0)=i],
-%	\end{align}
-%	as \(p\to\infty\), for every bounded, Lipschitz function \(f\). By the Portmanteau Theorem, then 
-%	\begin{align}
-%		& \mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i)   \nonumber 
-%		%
-%		\\& \to \mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) 
-%	\end{align}
-%	weakly as \(p\to\infty\). In addition 
-%	\begin{align}
-%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t\nonumber
-%		\\& \leq \int_{x=0}^\infty  h_{ij}^{--}(\lambda, x)\wrt x ,
-%	\end{align}
-%	which is bounded as a function of \(p\). Hence we can apply the Continuity Theorem \ref{thm: ext cont thm} again and claim that 
-%	\begin{align}
-%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t \nonumber
-%		%
-%		\\& \to \int_{t=0}^\infty e^{-\lambda t}\mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) \wrt t
-%	\end{align}
-%	as \(p\to\infty,\) as required. 
-%\end{proof}
+ 	\\&\leq G^{n-1}\int_{u_2=0}^{\Delta-u_1}\wrt u_1 \dots\nonumber
+            	\int_{u_{n-1}=0}^{\Delta-u_{n-2}}  \wrt u_{n-1}
+		\\&\leq G^{n-1}\Delta^{n-2}:=G^*_n.
+\end{align}
+Also, 
+\begin{align}
+	g^*_{2,n}(u_1,x) &= \int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1 \dots \nonumber 
+         	\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
+       	\\&\qquad{}g_{n}(\Delta - x-u_{n-1})1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1} \nonumber
  %
-
-\subsection{Many integrals}
-The orbit process at the time at which \(\{\phi(t)\}\) first exits \(\calS_{+0}\) (\(\calS_{-0}\)) is \(\bs k(x_0) \bs D\), thus we can treat this case simply as a change in the initial condition of the QBD-RAP.
-We do so by showing that asymptotically, the  initial conditions \(\bs k(\Delta-x_0)\) and \(\bs k(x_0)\bs D\) produce results that are arbitrarily close to each other.
-We first show a Lipschitz-like condition in \(x_0\) for \(w_n(x_0,x)\).
+ 	\\&\leq G^{n-1}\int_{u_2=0}^{\Delta-u_1}\wrt u_1 \dots\nonumber
+            	\int_{u_{n-1}=0}^{\Delta-u_{n-2}}  \wrt u_{n-1}
+		\\&\leq G^{n-1}\Delta^{n-2}:=G^*_n.
+\end{align}
 
@@
 
+% \begin{cor}\label{cor: ksjkd}
+% 	Let \(g_1, g_2, \dots,\) be functions satisfying the Assumptions \ref{asu: g} and let \(\bs v(x)\) be a closing operator with the Properties \ref{properties: some props}, then,
+% 	\begin{align}
+% 		&\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) e^{\bs{S}x_1}\wrt x_1\bs D 
+%             	\left[\prod_{k=2}^{n-1}\int_{x_k=0}^\infty g_k(x_k) e^{\bs{S}x_k} \wrt x_k \bs D\right] \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n \bs v(x) \nonumber 
+% 		\\&\leq \cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e} \widehat{G}^{n-1}G(G_{\bs v}+\widehat G O(\var(Z)) \label{eqn :mmmm}
+% 	\end{align}
+% \end{cor}
+% \begin{proof}
+% 	By Corollary \ref{cor: amammme} 
+% 	\begin{align}
+% 		&\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) e^{\bs{S}x_1}\wrt x_1\bs D 
+%             	\left[\prod_{k=2}^{n-1}\int_{x_k=0}^\infty g_k(x_k) e^{\bs{S}x_k} \wrt x_k \bs D\right] \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n \bs v(x)\nonumber
+% 		\\&\leq \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) e^{\bs{S}x_1}\wrt x_1\bs D 
+% 		\left[\prod_{k=2}^{n-1}\int_{x_k=0}^\infty g_k(x_k) e^{\bs{S}x_k} \wrt x_k \bs D\right] \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n\bs w(x) \nonumber
+% 		\\&\quad{} + O(\var(Z)). \label{eqn: LLaaNNab}
+% 	\end{align}
+% 	The first term of (\ref{eqn: LLaaNNab}) can be seen to be equivalent to \(\mathcal J_{1,n+1}(x_0,x_1),\) with \(g_1(x_1)=1\), and the integrability condition on \(g_1\) is not required to prove the bound.
+% \end{proof}
+
+% \begin{cor}\label{cor: amammme}
+% 	Let \(g_1, g_2, \dots,\) be functions satisfying the Assumptions \ref{asu: g} and let \({\bs v}(x)\) be a closing operator with the Properties \ref{properties: some props}, then,
+% 	\begin{align}
+% 		&\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) e^{\bs{S}x_1}\wrt x_1\bs D 
+%             	\left[\prod_{k=2}^{n-1}\int_{x_k=0}^\infty g_k(x_k) e^{\bs{S}x_k} \wrt x_k \bs D\right] \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n \widetilde{\bs w}(x) \nonumber 
+% 		% \\&= \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) e^{\bs{S}x_1}\wrt x_1\bs D 
+% 		% \left[\prod_{k=2}^{n-1}\int_{x_k=0}^\infty g_k(x_k) e^{\bs{S}x_k} \wrt x_k \bs D\right] \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs w}(x) \nonumber
+% 		\\&= O(\var(Z)) \label{eqn :mmmm2}
+% 	\end{align}
+% \end{cor}
+% \begin{proof}
+% 	% First observe that 
+% 	% \begin{align*} 
+% 	% 	\int_{x_n=0}^\infty \bs\alpha e^{\bs S(u+x_n)}{\bs v}(x)\wrt x_n 
+% 	% 	&=\int_{x_n=0}^\infty \bs\alpha e^{\bs S(u+x_n)} (\bs w(x)+\widetilde{\bs w}(x))\wrt x_n
+% 	% 	\\&=\int_{x_n=0}^\infty \bs\alpha e^{\bs S(u+x_n)} \bs w(x)\wrt x_n + \bs\alpha e^{\bs S(u+x_n)} \widetilde{\bs w}(x))\wrt x_n
+% 	% \end{align*}
+% 	% and recall that Property \ref{properties: 0} states 
+% 	% \begin{align}
+% 	% 	\int_{x_n=0}^\infty \bs\alpha e^{\bs S(u+x_n)} \widetilde{\bs w}(x)\wrt x_n = \bs\alpha e^{\bs Su}(-\bs S)^{-1} \widetilde{\bs w}(x) = O(\var(Z)).
+% 	% \end{align}
+
+% 	Consider the left-hand side of (\ref{eqn :mmmm2}). Replacing the last \(\bs D\) matrix in (\ref{eqn :mmmm2}) by its integral definition, gives 
+% 	\begin{align*}
+% 		&\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) e^{\bs{S}x_1}\wrt x_1
+%             \left[\prod_{k=2}^{n-1}\bs D \int_{x_k=0}^\infty g_k(x_k) e^{\bs{S}x_k} \wrt x_k \right] \int_{u=0}^\infty e^{\bs S u}\bs s \cfrac{\bs \alpha e^{\bs S u}}{\bs \alpha e^{\bs S u} \bs e}\wrt u
+% 			\\&\qquad{}\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n \widetilde{\bs w}(x) 
+% 		\\&=\bs k(x_0) \int_{u=0}^\infty \mathcal I_{1,n}(u) \cfrac{\bs \alpha e^{\bs S u}}{\bs \alpha e^{\bs S u} \bs e}\wrt u \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n \widetilde{\bs w}(x) 
+% 		\\&\leq \cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e}G\widehat G^{n-1} \int_{u=0}^\infty \bs \alpha e^{\bs{S}u}\bs e \cfrac{\bs \alpha e^{\bs S u}}{\bs \alpha e^{\bs S u} \bs e}\wrt u \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n \widetilde{\bs w}(x)
+% 	\end{align*}
+% 	by Lemma \ref{lem: lh bnd}. Integrating over \(u\) gives 
+% 	\begin{align*}
+% 		&\cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e}G\widehat G^{n-1} \bs \alpha (-\bs S)^{-1} \int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n \widetilde{\bs w}(x) 
+% 		\\&\leq \cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e}G\widehat G^{n-1} \bs \alpha (-\bs S)^{-1} \int_{x_n=0}^\infty g_{n}(x_n) \wrt x_n \widetilde{\bs w}(x),
+% 	\end{align*}
+% 	by Property \ref{properties: -1}. Integrating over \(x_n\), gives
+% 	\begin{align*}
+% 		\cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e}G\widehat G^{n} \bs \alpha (-\bs S)^{-1} \widetilde{\bs w}(x) 
+% 		&=\cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e}G\widehat G^{n}O(\var(Z)),
+% 	\end{align*}
+% 	by Property \ref{properties: 0}.
+% \end{proof}
+
 \begin{cor}\label{cor: ahjg}
@@
 	\begin{align}
-		&\Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right]
-            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \nonumber 
-	%
-		\\&\quad{}- w_n(\Delta - x_0,x) \Bigg| \nonumber
-		\\&= r_8(n),
+		&\left|\int_{x=0}\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0 - w_n(\Delta - x_0,x) \right| \nonumber
+		= r_8(n),
 	\end{align}
@@
 	\begin{align*}
-	|r_8(n)|&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G_n^{*}(G+L\Delta) \right) \\&\qquad{}+ 2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}.\end{align*}
+		|r_8(n)|&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G_n^{*}(G+L\Delta) \right) \\&\qquad{}+ 2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}.
+	\end{align*}
 \end{cor}
 \begin{proof}
-	Observe that 
-	\begin{align}
-		\nonumber&\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right]
-            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) 
-	%
-		\\\nonumber&=\int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs s\cfrac{\bs\alpha e^{\bs Sz_0}}{\bs\alpha e^{\bs Sz_0}\bs e} \wrt z_0 e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right]
-            	\\\nonumber&\quad\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) 
-	%
-		\\&=\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0
-	\end{align}
-
 	Now
@@
 	\end{align}
-		By Corollary~\ref{cor: ksjkd}, for any \(x,y\in [0,\Delta)\), \(0\leq w_n(x,y)\leq \widehat G^{n-2}GG_{\bs v}\), so the sum of the first two terms is less than or equal to 
+
+		Using Equations~(\ref{eqn: in here}), (\ref{eqn: J bound}) and (\ref{eqn:FGHJSjjs sj}) we can claim
 		\begin{align}
-		&2\widehat G^{n-1}GG_{\bs v}\left(\int_{z_0=0}^{\Delta-\varepsilon-x_0}\bs k(x_0)e^{\bs S z_0} \bs s \wrt z_0+\int_{z_0=\Delta+\varepsilon-x_0}^\infty \bs k(x_0)e^{\bs S z_0} \bs s \wrt z_0\right)
-		\\\nonumber &= 2\widehat G^{n-1}GG_{\bs v} \cfrac{\mathbb P(|Z-\Delta|> \varepsilon)}{\mathbb P(Z> x_0)}
-		\\&\leq 2G^n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}
+			|w_n(x_0,x)| &\leq \cfrac{1}{\bs \alpha e^{\bs{S} x_0}\bs e}G^2 \widehat G^{n-2}  
+			\int_{u_k=0}^\infty \bs \alpha e^{\bs{S}u_k}\bs e \wrt u_k G_{\bs v} + \cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e}G\widehat G^{n}\widetilde G_{\bs v} \nonumber 
+			\\& = \cfrac{1}{\bs \alpha e^{\bs{S} x_0}\bs e}G^2
+			 \widehat G^{n-2} G_{\bs v} + \cfrac{1}{\bs \alpha e^{\bs{S}x_0}\bs e}G\widehat G^{n}\widetilde G_{\bs v} \nonumber 
+			 \\&=:W_n
+		\end{align}
+
+		Therefore, the sum of the first two terms in (\ref{eqn: KASJF}) is less than or equal to 
+		\begin{align}
+		&2W_n\left(\int_{z_0=0}^{\Delta-\varepsilon-x_0}\bs k(x_0)e^{\bs S z_0} \bs s \wrt z_0+\int_{z_0=\Delta+\varepsilon-x_0}^\infty \bs k(x_0)e^{\bs S z_0} \bs s \wrt z_0\right)\nonumber 
+		\\\nonumber &= 2W \cfrac{\mathbb P(|Z-\Delta|> \varepsilon)}{\mathbb P(Z> x_0)}
+		\\&\leq 2W_n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}
 		\end{align}
@@
 		\\\nonumber&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G^{n-1}\Delta^{n-2}(G+L\Delta) \right)
-		\\&\quad{}+2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2} ,
+		\\&\quad{}+2W_n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2} ,
 	\end{align}
@@
 	\begin{align}
-		&\Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right]
-            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \nonumber 
-	%
-		\\&\qquad {}- \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
-%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
-		\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
-		%\\&{}\nonumber
-            	%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
-            	g_{n}(\Delta - x-u_{n-1})
-	\\&\qquad{} 1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1} \Bigg| \nonumber
+		&\left|\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0- g_{1,n}^*(\Delta-x_0,x) \right| \nonumber
 		\\&\leq |r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|,\label{eqn: KAFnn}
@@
 %		\\&{}- g_{1,n}^*(\Delta - x_0,x) \Bigg| 
-		&\Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right]
-            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \nonumber 
+		&\left| \bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0 \nonumber 
 	%
-		\\\nonumber &{}\qquad - w_n(\Delta - x_0,x) + w_n(\Delta - x_0,x) - g_{1,n}^*(\Delta - x_0,x) \Bigg| 
-		\\&\leq \Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right]
-            	\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) ,\nonumber 
-	\\&\qquad - w_n(\Delta - x_0,x)\Bigg| + \Bigg| w_n(\Delta - x_0,x) - g_{1,n}^*(\Delta - x_0,x) \Bigg| 
+		 - w_n(\Delta - x_0,x) + w_n(\Delta - x_0,x) - g_{1,n}^*(\Delta - x_0,x) \right| 
+		\\&\leq \left| \bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0 - w_n(\Delta - x_0,x)\right| + \left| w_n(\Delta - x_0,x) - g_{1,n}^*(\Delta - x_0,x) \right| 
 	\end{align*}
@@
 	\begin{align}
-		&\Bigg| \int_{x\in[0,\Delta)} \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right] \nonumber 
-            	\\&\qquad{}\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \psi(x)\wrt x \nonumber 
+		&\left| \int_{x\in[0,\Delta)}\bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0 \psi(x)\wrt x  
 	% 
-		\\&\qquad {}- \int_{x\in[0,\Delta)} \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
-%		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
-		\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
-		%\\&{}\nonumber
-            	%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
-            	g_{n}(\Delta - x-u_{n-1})
-	\\&\qquad{} \times1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\psi(x)\wrt x \Bigg| \nonumber
+		- \int_{x\in[0,\Delta)} g_{1,n}^*(\Delta-x_0,x)\psi(x)\wrt x \right|\nonumber 
 		\\&\leq \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta.\label{eqn: KAFnnmna2}
@@
 	\begin{align}
-		& \int_{x\in[0,\Delta)} \Bigg| \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-            	\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-		\bs D\right] \nonumber 
-            	\\&\qquad {}\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x)  \nonumber 
+		& \int_{x\in[0,\Delta)} \left| \int_{x_1=0}^\infty \bs k(x_0) \int_{z_0=0}^\infty e^{\bs Sz_0}\bs sw_n(z_0,x)\wrt z_0 
 	%
-		- \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
-		\\&\qquad {}\times \left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
-            	g_{n}(\Delta - x-u_{n-1})
-	\\&\qquad{} 1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\Bigg| \left|\psi(x)\right|\wrt x . \label{eqn: lfj}
+		- g_{1,n}^*(\Delta-x_0,x)\right| \left|\psi(x)\right|\wrt x . \label{eqn: lfj}
 	\end{align}
-	Now, apply Corollary~\ref{cor: aaaaa} to the first absolute value then (\ref{eqn: lfj}) is less than or equal to 
-	\begin{align}
-		&\int_{x\in[0,\Delta)} \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right) \left|\psi(x)\right|\wrt x  \nonumber 
-		\\&\leq \int_{x\in[0,\Delta)} \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right) F\wrt x  \nonumber 
-		\\&=  \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right) \Delta F
-	\end{align}
+	Now, using \(|\psi(x)|\leq F\) and Corollary~\ref{cor: aaaaa} then (\ref{eqn: lfj}) is less than or equal to 
+		\[\left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right) \Delta F\]
 \end{proof}
-
-To summarise, the main results we need are 
-\begin{align}
-	&\Bigg| \int_{x=0}^\Delta w_n(x_0,x) \psi(x) \wrt x \nonumber 
 %
-	\\&{}- \int_{x=0}^\Delta \int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)
+%To summarise, the main results we need are 
+%\begin{align}
+%	&\Bigg| \int_{x=0}^\Delta w_n(x_0,x) \psi(x) \wrt x \nonumber 
+%
+%	\\&{}- \int_{x=0}^\Delta \int_{u_1=0}^{\Delta-x_0}g_1(\Delta - u_1 - x_0)
 %		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
-	\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
-	%\\&{}\nonumber
-			%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
-			g_{n}(\Delta - x-u_{n-1})
-\\&\qquad{} 1(\Delta-x-u_{n-1}\geq0) \wrt u_{n-1}\psi(x) \wrt x \Bigg| \nonumber
-	\\&\leq (|r_5(n)| + |r_6(n)| + (n-1)|r_4(n)|)\Delta F. \label{eqn: rhs g 4dvfklsmv2}
-\end{align}
-and 
-\begin{align}
-	&\Bigg| \int_{x\in[0,\Delta)} \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
-			\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
-	\bs D\right] \nonumber 
-			\\&\qquad{}\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \psi(x)\wrt x \nonumber 
+%	\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
+%	%\\&{}\nonumber
+%			%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
+%			g_{n}(\Delta - x-u_{n-1})
+%\\&\qquad{} 1(\Delta-x-u_{n-1}\geq0) \wrt u_{n-1}\psi(x) \wrt x \Bigg| \nonumber
+%	\\&\leq (|r_5(n)| + |r_6(n)| + (n-1)|r_4(n)|)\Delta F. \label{eqn: rhs g 4dvfklsmv2}
+%\end{align}
+%and 
+%\begin{align}
+%	&\Bigg| \int_{x\in[0,\Delta)} \int_{x_1=0}^\infty g_1(x_1) \bs k(x_0) \bs D e^{\bs{S}x_1}\wrt x_1\bs D 
+%			\left[\prod_{n=2}^{k-1}\int_{x_n=0}^\infty g_n(x_n) e^{\bs{S}x_n} \wrt x_n
+%	\bs D\right] \nonumber 
+%			\\&\qquad{}\times\int_{x_n=0}^\infty g_{n}(x_n) e^{\bs{S}x_n} \wrt x_n {\bs v}(x) \psi(x)\wrt x \nonumber 
 %
-	\\&\qquad {}- \int_{x\in[0,\Delta)} \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
+%	\\&\qquad {}- \int_{x\in[0,\Delta)} \int_{u_1=0}^{x_0}g_1(x_0 - u_1)
 %		\int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1  \nonumber 
-	\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
-	%\\&{}\nonumber
-			%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
-			g_{n}(\Delta - x-u_{n-1})
-\\&\qquad{} \times1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\psi(x)\wrt x \Bigg| \nonumber
-	\\&\leq \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta.\label{eqn: KAFnnmna22}
-\end{align}
-
+%	\left[\prod_{k=2}^{n-1} \int_{u_k=0}^{\Delta-u_{k-1}} g_k(\Delta-u_k-u_{k-1})\wrt u_{k-1}\right] \nonumber 
+%	%\\&{}\nonumber
+%			%\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
+%			g_{n}(\Delta - x-u_{n-1})
+%\\&\qquad{} \times1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1}\psi(x)\wrt x \Bigg| \nonumber
+%	\\&\leq \left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta.\label{eqn: KAFnnmna22}
+%\end{align}
+%
 We have assumed throughout the appendix that the functions \(g\) are scalar functions, however, we are ulitmately interested in expressions which contain matrix functions.
@@
 \end{align}
+
+Thus, we have estableshed a type of convergence of the QBD-RAP scheme to the fluid queue on the event that the phase is initially in \(\calS_{+0}\cup\calS_{-0}\), and before the first change of level.
+However, we're not quite done yet.
+The last thing we need to prove is convergence at the first change of level.
+Since the result in Corollary~\ref{cor: aaaaa} is pointwise in \(x\), then choosing the closing operator as \(e^{\bs Sx}\bs s\) and setting \(x=0\), then we get covergence at the first change of level, on the event that there is one or more phase transition from \(\calS_+\) to \(\calS_-\) or \(\calS_-\) to \(\calS_+\).
+The only thing that remains is to show convergence at the first change of level on the event that there is no phase transition from \(\calS_+\) to \(\calS_-\) or \(\calS_-\) to \(\calS_+\).
+
+ \begin{lem} \label{lem:tttttt}
+ 	Let \(g\) satisfy the Assumptions \ref{asu: g} and \(x_0\in(2\varepsilon,\Delta-\varepsilon)\). Then
+ 	\begin{align}
+ 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}4G + 3L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2}.
+ 	\end{align}
+ \end{lem}
+ \begin{proof}
+ 	First rewrite the left-hand side as 
+ 	\begin{align}
+ %		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| 
+ 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}(g(x)-g(x_0))\bs s\wrt x \right|
+ 		%
+ 		&\leq \int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x.
+ 	\end{align}
+ 	Substituting in the expression for \(\bs D\) gives,
+ 	\begin{align}
+ 		&\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x  \nonumber 
+ %		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x 
+ 		%
+ 		\\&= \int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \nonumber
+ 		\\&\quad {} + \int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \label{eqn: 2nd here}
+ 	\end{align}
+ 	Since \(g\) is bounded, the second term is less than or equal to 
+ 	\begin{align}
+ 		\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x 2G \nonumber
+ 		&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u \bs e 2G \nonumber
+ 		\\&= \bs k(x_0)\int_{u=\Delta-\varepsilon}^\infty e^{\bs Su}\bs s \wrt u2G \nonumber
+ 		\\&= \cfrac{\mathbb P(Z\geq x_0+\Delta-\varepsilon)}{\mathbb P(Z>x_0)}2G \label{eqn" dgkjlwerhv}
+ 	\end{align}
+ 	For \(x_0\in(2\varepsilon,\Delta-\varepsilon),\) then (\ref{eqn" dgkjlwerhv}) is less than or equal to 
+ 	\[\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}2G.\]
+
+ 	As for the first term in (\ref{eqn: 2nd here}), it can be written as 
+ 	\begin{align}
+ 		&\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
+ 		%
+ 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x \nonumber
+ 		%
+ 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}|g(x) - g(x_0)|\bs s\wrt x.
+\label{eqn: 201}
+ 	\end{align}
+ 	Since \(g\) is bounded, then the last two terms in (\ref{eqn: 201}) are 
+ 	\begin{align}
+ 		&2G\Bigg(\int_{x=0}^\infty \bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x \nonumber
+ 		%
+ 		\\&\qquad{}+\int_{x=0}^\infty \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \wrt u e^{\bs Sx}\bs s\wrt x\Bigg) \nonumber
+ 		%
+ 		\\&=2G\Bigg(\bs k(x_0)\int_{u=0}^{\Delta-x_0-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e}  \bs e  \wrt u\nonumber
+ 		%
+ 		+ \bs k(x_0)\int_{u=\Delta-x_0+\varepsilon}^{\Delta-\varepsilon} e^{\bs Su}\bs s \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} \bs e\wrt u \Bigg) \nonumber
+ 		%
+ 		\\&=2G \cfrac{\mathbb P(Z>x_0, Z\notin(\Delta-\varepsilon, \Delta+\varepsilon))}{\mathbb P(Z>x_0)}  \nonumber 
+ 		%
+ 		\\&\leq 2G\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2},
+ 	\end{align}
+ 	provided that \(x_0\in[0,\Delta-\varepsilon)\). 
+ 	Exchanging the order of integration for the first term in (\ref{eqn: 201})
+ 	\begin{align}
+ 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\int_{x=0}^\infty \cfrac{\bs \alpha e^{\bs Su}}{\bs \alpha e^{\bs Su}\bs e} e^{\bs Sx}|g(x)-g(x_0)|\bs s\wrt x \wrt u \label{eqn: fdk13897}
+ 	\end{align}
+ 	from which we see that we can apply Corollary~\ref{cor: cond bnd 2} with \(v=0\) to the integral over \(x\), implying that (\ref{eqn: fdk13897}) is less than or equal to
+ 	\begin{align}
+ 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(|g(\Delta-u)-g(x_0)|+r_3(u)\right)\wrt u. \label{eqn: sdkagh lkhvasfv}
+ 	\end{align}
+ 	Noting that \(\sup |r_3(u)|\leq|r_2|\) for \(u\leq \Delta-\varepsilon\), and since \(g\) is Lipschitz, then (\ref{eqn: sdkagh lkhvasfv}) is less than or equal to 
+ 	\begin{align}
+ 		\int_{u=\Delta-x_0-\varepsilon}^{u=\Delta-x_0+\varepsilon} \bs k(x_0) e^{\bs Su}\bs s\left(L\varepsilon+|r_2|\right)\wrt u \leq L\varepsilon+|r_2|. \label{eqn: sdkagh lsfv}
+ 	\end{align}
+ 	Putting all the bounds together proves the result. 
+ \end{proof}
+%
+%\begin{lem}
+%	Let \(f:[0,\Delta)\to \mathbb R\) be any bounded, Lipschitz continuous function with \(|\psi(x)|\leq F\). Then for \(u\leq \Delta-\varepsilon\), \(v\in[0,\Delta)\)
+%	\begin{align}
+%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[f(\bar X(t))1(\varphi(t)=j, L(t)=\ell_0, L(s)=\ell_0, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid L(0)=\ell_0, \nonumber
+%		\\&\qquad{} \bs A(0)=\bs a_{\ell_0,i}(x_0), \varphi(0)=i]\wrt t \nonumber
+%		\\&= \int_{t=0}^\infty e^{-\lambda t}\mathbb E[\psi(X(t))1(\varphi(t)=j,X(s)\in\calD_{\ell_0}, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid X(0)=x_0, \varphi(0)=i]\wrt t \nonumber
+%		\\&\qquad {} + r_{11}^{(p)}
+%	\end{align}
+%	where 
+%	\[|r_{11}^{(p)}| \leq F R_{{\bs v},2}+ GF\varepsilon.\]
+%\end{lem}
+%\begin{proof}
+%	Assume, without loss of generality \(\ell_0=0\) so \(\calD_{\ell_0}=[0,\Delta]\). First observe that 
+%	\begin{align}
+%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[f(\bar X(t))1(\varphi(t)=j, L(t)=\ell_0, L(s)=\ell_0, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid L(0)=\ell_0, \nonumber
+%		\\&\qquad{} \bs A(0)=\bs a_{\ell_0,i}(x_0), \varphi(0)=i]\wrt t \nonumber 
+%		\\&= \int_{x_1=0}^\infty \int_{x=[0,\Delta)} \cfrac{\bs \alpha e^{\bs S(x_1+x_0+x)}\bs s}{\bs \alpha e^{\bs S x_0} \bs e}h_{ij}^{++}(\lambda, x_1)f(\Delta-x)\wrt x\wrt x_1
+%	\end{align}
+%	and 
+%	\begin{align}
+%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E[\psi(X(t))1(\varphi(t)=j,X(s)\in\calD_{\ell_0}, \varphi(s)\in\calS_+\cup\calS_{+0},s\in[0,t])\mid X(0)=x_0, \varphi(0)=i]\wrt t \nonumber
+%		\\&=  \int_{x=0}^{\Delta-x_0} h_{ij}^{++}(\lambda, x)\psi(X_0 + x)\wrt x
+%	\end{align}
+%	In Property \ref{properties: 2} take \(g(x) = h_{ij}^{--}(\lambda, x)\), which states 
+%	\begin{align}
+%		&\int_{x_1=0}^\infty \int_{x=0}^\Delta\cfrac{\bs \alpha e^{\bs S(x_1+x_0+x)}\bs s}{\bs \alpha e^{\bs S x_0} \bs e}h_{ij}^{++}(\lambda, x_1)f(\Delta-x)\wrt x\wrt x_1 \nonumber
+%		\\& = \int_{x=0}^{\Delta}h_{ij}^{++}(\lambda,\Delta - x_0 - x)1(x+x_0\leq \Delta-\varepsilon)f(\Delta - x)\wrt x + \int_{x=0}^\Delta r_{\bs v}(x_0,x)\psi(x)\wrt x. 
+%	\end{align}
+%	The second term is less than or equal to 
+%	\[F\int_{x=0}^\Delta r_{\bs v}(x_0,x)\wrt x = FR_{{\bs v},1}.\]
+%	The first term is 
+%	\[\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x - \int_{x=\Delta-x_0-\varepsilon}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x.\]
+%	Now 
+%	\[\int_{x=\Delta-x_0-\varepsilon}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x\leq GF\varepsilon\]
+%	and 
+%	\[\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,\Delta - x_0 - x)f(\Delta - x)\wrt x=\int_{x=0}^{\Delta-x_0}h_{ij}^{++}(\lambda,x)\psi(x_0+x)\wrt x.\]
+%\end{proof}
+%
+%
+%\begin{cor}\label{cor: Dcoajc2222}
+%	Let \(\psi:\calD_{\ell_0}\to \mathbb R\) be bounded, \(|\psi(x)|\leq F\), and Lipschitz continuous.
+Then, for \(x_0\in(y_{\ell_0}, y_{\ell+1}),\) \(k\in\calS_{0+}\), \(i\in\calS_-\), \(j\in\calS_-\cup\calS_{-0}\), there exists \(r_{10}^{(p)}\to 0\) as \(p \to \infty\),
+%	\begin{align}
+%		\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)}.\label{eqn:xvasfv}
+%	\end{align}
+%	Similarly, for \(k\in\calS_{0-}\), \(i\in\calS_+\), \(j\in\calS_+\cup\calS_{+0}\)
+%	\begin{align}
+%		\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,+0,+}^{\ell_0}(x,i,j;j,x_0)\psi(x)\wrt x  - \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,+0,+}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)}.\label{eqn:!124msfvcds}
+%	\end{align}
+%\end{cor}
+%\begin{proof}[Proof of Corollary \ref{cor: Dcoajc}]
+%	We show the result for (\ref{eqn:xs}) only, with the result for (\ref{eqn:!124mcds}) following analogously. 
+%	
+%	Observe that, for \(k\in\calS_{0+}\), \(i\in\calS_-\), \(j\in\calS_-\cup\calS_{-0}\), 
+%	\begin{align}
+%		&\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x \nonumber 
+%		%
+%		\\&{}= \vligne{\lambda \bs I - \bs T_{00}}^{-1}_{ki}\int_{x\in\calD_{\ell_0}}\int_{x_1=0}^\infty \bs k(x_0) \bs D e^{\bs S x_1} {\bs v}(x)h_{ij}^{--}(\lambda,x_1)\wrt x_1\psi(x)\wrt x,
+%	\end{align}
+%	and 
+%	\begin{align}
+%		\int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x \nonumber  = \vligne{\lambda \bs I - \bs T_{00}}^{-1}_{ki}\int_{x=0}^{x_0} h_{ij}^{--}(\lambda,x_0-x)\psi(x)\wrt x
+%	\end{align}
+%	These terms appear in the proof of Corollary~\ref{lem: ppp} and using arguments applied there we can show
+%	\begin{align}
+%		\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x ,\label{eqn:xssgwg}
+%	\end{align}
+%	and therefore there exists a bound \(r_{10}^{(p)}\) such that 
+%	\[\left|\int_{x\in\calD_{\ell_0}} \widehat f_{0,-0,-}^{\ell_0,(p)}(x,i,j;j,x_0)\psi(x)\wrt x  \to \int_{x\in\calD_{\ell_0}} \widehat \mu_{0,-0,-}^{\ell_0}(x,i,j;k,x_0)\psi(x)\wrt x\right|\leq r_{10}^{(p)} ,\]
+%	and \(r_{10}^{(p)}\to 0\) as \(p\to\infty\). 
+%\end{proof}
+%
+%\begin{cor}\label{corL Tta} For \(i\in\mathcal S_-,j\in\mathcal S_-\cup\calS_{-0}\), \(x_0\in[0,\Delta)\), 
+%	\begin{align}
+%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t \nonumber
+%		%
+%		\\& \to \int_{t=0}^\infty e^{-\lambda t}\mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) \wrt t \label{eqn: ffaA}
+%	\end{align}
+%	as \(p\to\infty\).
+%\end{cor}
+%\begin{proof}
+%	From the convergence of Laplace transforms in Lemma~\ref{lem: ppp} and the Continuity Theorem \ref{thm: ext cont thm} then 
+%	\begin{align}
+%		&  \mathbb E[f(\bar X^{(p)}(t))1(\varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j) \mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i]  \nonumber 
+%		\\& \to \mathbb E[\psi(X(t)) 1(\varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j) \mid X(0)=x_0, \varphi(0)=i],
+%	\end{align}
+%	as \(p\to\infty\), for every bounded, Lipschitz function \(f\). By the Portmanteau Theorem, then 
+%	\begin{align}
+%		& \mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i)   \nonumber 
+%		%
+%		\\& \to \mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) 
+%	\end{align}
+%	weakly as \(p\to\infty\). In addition 
+%	\begin{align}
+%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t\nonumber
+%		\\& \leq \int_{x=0}^\infty  h_{ij}^{--}(\lambda, x)\wrt x ,
+%	\end{align}
+%	which is bounded as a function of \(p\). Hence we can apply the Continuity Theorem \ref{thm: ext cont thm} again and claim that 
+%	\begin{align}
+%		&\int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bar X^{(p)}(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid \bs A^{(p)}(0)=\bs k^{(p)}(x_0)\bs D^{(p)}, \varphi(0)=i) \wrt t \nonumber
+%		%
+%		\\& \to \int_{t=0}^\infty e^{-\lambda t}\mathbb P( X(t) \in \cdot, \varphi(s)\in\mathcal S_-\cup\mathcal S_{-0}, s\in[0,t], \varphi(t)=j\mid X(0)= x_0, \varphi(0)=i) \wrt t
+%	\end{align}
+%	as \(p\to\infty,\) as required. 
+%\end{proof}
+%
+%
+%
+%
+%
+%
+%

### 2021-11-28 appendix3/appendix.tex
finisehd rewriting appendices, but need to proof still

@@
 
-For \(\varphi(0)=k\in\mathcal S_{-0}\) (or \(\varphi(0)=k\in\mathcal S_{+0}\)) the added complexity comes from the fact that it is possible that, upon the phase process first leaving \(\mathcal S_{-0}\) (\(\mathcal S_{+0}\)) the phase transitions to a state in \(\mathcal S_+\) (\(\mathcal S_-\)).
+For \(\varphi(0)=k\in\mathcal S_{-0}\) (or \(\varphi(0)=k\in\mathcal S_{+0}\)) the added complexity comes from the fact that it is possible, upon the phase process first leaving \(\mathcal S_{-0}\) (\(\mathcal S_{+0}\)) the phase transitions to a state in \(\mathcal S_+\) (\(\mathcal S_-\)).
 Since the orbit of the QBD-RAP is constant on \(\varphi(t)\in\mathcal S_{-0}\) (\(\varphi(t)\in\mathcal S_{+0}\)), then upon a first transition out of \(\mathcal S_{-0}\) (\(\mathcal S_{+0}\)) and into \(\mathcal S_+\) (\(\mathcal S_-\)) the orbit jumps to \(\bs   a_{\ell_0,i}^{(p)}(x_0)\bs D^{(p)}\).
@@
 The first term of (\ref{eqn: k0jJ0}) creates significantly more work.
-Essentially, we need to repreat the results from Appendix~\ref{app:tand} but with the initial vector \(\bs a_{\ell_0,i}(x_0)\bs D\).
+Essentially, we need to derive more bounds, analogous to the results from Appendix~\ref{app:tand}, but with the initial vector \(\bs a_{\ell_0,i}(x_0)\bs D\).
 
@@
 \end{align}
-Also, 
-\begin{align}
-	g^*_{2,n}(u_1,x) &= \int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1 \dots \nonumber 
-         	\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
-       	\\&\qquad{}g_{n}(\Delta - x-u_{n-1})1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1} \nonumber
- %
- 	\\&\leq G^{n-1}\int_{u_2=0}^{\Delta-u_1}\wrt u_1 \dots\nonumber
-            	\int_{u_{n-1}=0}^{\Delta-u_{n-2}}  \wrt u_{n-1}
-		\\&\leq G^{n-1}\Delta^{n-2}:=G^*_n.
-\end{align}
+% Also, 
+% \begin{align}
+% 	g^*_{2,n}(u_1,x) &= \int_{u_2=0}^{\Delta-u_1}g_2(\Delta - u_2 - u_1)\wrt u_1 \dots \nonumber 
+%          	\int_{u_{n-1}=0}^{\Delta-u_{n-2}} g_{n-1}(\Delta - u_{n-1} - u_{n-2}) \wrt u_{n-2}
+%        	\\&\qquad{}g_{n}(\Delta - x-u_{n-1})1(\Delta-x-u_{n-1}\geq0)\wrt u_{n-1} \nonumber
+%  %
+%  	\\&\leq G^{n-1}\int_{u_2=0}^{\Delta-u_1}\wrt u_1 \dots\nonumber
+%             	\int_{u_{n-1}=0}^{\Delta-u_{n-2}}  \wrt u_{n-1}
+% 		\\&\leq G^{n-1}\Delta^{n-2}:=G^*_n.
+% \end{align}
 
@@
 	\begin{align*}
-		|r_8(n)|&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G_n^{*}(G+L\Delta) \right) \\&\qquad{}+ 2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}.
+		|r_8(n)|&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G_n^{*}(G+L\Delta) \right) \\&\qquad{}+ 2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/(\Delta-x_0)^2}.
 	\end{align*}
@@
 		\\\nonumber &= 2W \cfrac{\mathbb P(|Z-\Delta|> \varepsilon)}{\mathbb P(Z> x_0)}
-		\\&\leq 2W_n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}
+		\\&\leq 2W_n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/(\Delta-x_0)^2}
 		\end{align}
@@
 		\\\nonumber&\leq  \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G^{n-1}\Delta^{n-2}(G+L\Delta) \right)
-		\\&\quad{}+2W_n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2} ,
+		\\&\quad{}+2W_n\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/(\Delta-x_0)^2} ,
 	\end{align}
@@
 		|r_8(n)|&\leq \left( 2|r_5(n)| + 2|r_6(n)| + 2(n-1)|r_4(n)| + \varepsilon G^{n-1}\Delta^{n-2}(G+L\Delta) \right) \\&\quad{} 
-	+2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}.
+	+2\widehat G^{n-2}GG_{\bs v}\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/(\Delta-x_0)^2}.
 	\end{align}
@@
 %
-We have assumed throughout the appendix that the functions \(g\) are scalar functions, however, we are ulitmately interested in expressions which contain matrix functions.
 We now extend the previous results to the matrix case.
@@
 	\end{align}
-	Moreover, choosing \(\varepsilon=\var(Z)\), then, for fixed \(n\), the error term \(\left(|r_8(n)|+|r_5(n)|+|r_6(n)| + (n-1)|r_4(n)|\right)F\Delta \prod_{k=2}^{n}N_{k}\) is \(\mathcal O(\var(Z)^{1/3})\).
+	Moreover, choosing \(\varepsilon=\var(Z)\), then, for fixed \(n\), the bound is \(\mathcal O(\var(Z)^{1/3})\). 
 \end{lem}
@@
 
-Thus, we have estableshed a type of convergence of the QBD-RAP scheme to the fluid queue on the event that the phase is initially in \(\calS_{+0}\cup\calS_{-0}\), and before the first change of level.
-However, we're not quite done yet.
+Thus, we have established a type of convergence of the QBD-RAP scheme to the fluid queue on the event that the phase is initially in \(\calS_{+0}\cup\calS_{-0}\), and before the first change of level.
+However, we are not quite done yet.
 The last thing we need to prove is convergence at the first change of level.
@@
  	\begin{align}
- 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}4G + 3L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2}.
+ 		\left|\int_{x=0}^\infty \bs k(x_0)\bs De^{\bs Sx}g(x)\bs s\wrt x - g(x_0)\right| \leq \cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/(\Delta-x_0)^2}4G + 3L\varepsilon+6G\cfrac{\var(Z)}{\varepsilon^2}.
  	\end{align}
@@
  	For \(x_0\in(2\varepsilon,\Delta-\varepsilon),\) then (\ref{eqn" dgkjlwerhv}) is less than or equal to 
- 	\[\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2}2G.\]
+ 	\[\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/(\Delta-x_0)^2}2G.\]
 
@@
  		%
- 		\\&\leq 2G\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/\varepsilon^2},
+ 		\\&\leq 2G\cfrac{\var(Z)/\varepsilon^2}{1-\var(Z)/(\Delta-x_0)^2}.
  	\end{align}
- 	provided that \(x_0\in[0,\Delta-\varepsilon)\). 
  	Exchanging the order of integration for the first term in (\ref{eqn: 201})
@@
  \end{proof}
+
+ Lastly, the geometric domination results in Lemma~\ref{lem: gkjljklgagjklagsjlk} follow by the same arguments as in the proof of Lemma~\ref{lem: gkjljklgagjklagsjlk}, except with the initial vector \(\bs a_{\ell_0,i}(x_0)\) replaced by \(\bs a_{\ell_0,i}(x_0)\bs D\).
+Similarly, an anologue of the domination condition required in the proof of Lemma~\ref{cor: aln222} can be established by the same arguments except with the initial vector \(\bs a_{\ell_0,i}(x_0)\) replaced by \(\bs a_{\ell_0,i}(x_0)\bs D\).
 %

### 2021-11-29 chapter2/chapter.tex
first steps of numerics chapter

--- 2021-11-11 chapter2/chapter.tex
@@
 %!TEX root = ../thesis.tex
-\chapter{A stochastic modelling approach to approximating fluid queues\label{sec: construction and modelling}}
-In this chapter we develop a new discretisation of a fluid queue using a \emph{quasi-birth-and-death process with rational arrival process components} (QBD-RAP).
-In the following chapter, Chapter \ref{sec: conv} we prove the scheme converges weakly.
-To improve upon the discretisation of \cite{bo2013}, we argue that the shift from QBD to QBD-RAP is necessary.
-The QBD construction of \cite{bo2013} uses Erlang distributions to approximate the deterministic time that the fluid queue spends in an interval on the event that the phase is constant -- it is well known that the Erlang distribution is the distribution with the smallest variance in the class of Phase-type distributions of a given order \cite{as1987}, and in this sense is the best approximation to the determinstic behaviour we want to capture.
-
-By choosing a matrix-exponential distribution with sufficiently small variance, the QBD-RAP approximation can be made arbitrarily accurate.
-Due to its stochastic interpretation, the QBD-RAP discretisation ensures solutions are positive, even when discontinuities are present.
-Unlike slope limiters and post-hoc filtering, for the QBD-RAP the positivity preserving nature is a property of the discretised operator.
-
-The structure of this chapter is as follows.
-Section~\ref{sec: inspiration} motivates the idea using the QBD approximation of \cite{bo2014}.
-Sections \ref{sec: modelling} and \ref{sec: zero states} describe the modelling of certain events of the fluid queue with matrix exponential distributions to ultimately construct the behaviour of the QBD-RAP on the event that the QBD-RAP remains in a given level.
-Section~\ref{sec: dynamics} describes the dynamics of the construction and also constructs the level process of the QBD-RAP.
-Section~\ref{sec: boundary conditions} deals with modelling boundary behaviour.
-Section~\ref{sec: initial conditions} describes how to model initial conditions of the fluid queue.
-Section~\ref{sec: closing} introduces the concept of a \emph{closing operator} which maps the state space of the QBD-RAP to estimates of the density of the fluid queue.
-
-\section{Inspiration and motivation}\label{sec: inspiration}
-The inspiration for our approach stems from the QBD-based discretisation of \cite{bo2013}.
-A QBD can be seen as a two dimensional CTMC, \(\{(L(t),\phi(t))\}_{t\geq 0}\), where \(\{L(t)\}\) is the discrete level, and \(\{\phi(t)\}\) is the phase.
-\cite{bo2013} discretise the state space of the fluid into small intervals of width \(\Delta\); denote these by \(\mathcal D_k=[k\Delta,(k+1)\Delta]\).
-Their approximation captures the dynamics of the phase process exactly, but discretises the level process of the fluid queue.
-Their approximation supposes that when \(L(t)=k,\,\phi(t)=i\), then \(X(t)\approx k\Delta,\,\phi(t)=i\).
-
-When in level \(k\) and phase \(i\), the approximation sees events at rate \(|T_{ii}| + |c_i|\Delta\).
-Upon an event, with probability \(T_{ij}/(|T_{ii}| + |c_i|\Delta)\) a change of phase from \(i\) to \(j\) is observed and the approximation remains in level \(k\); with probability \(|c_i|\Delta/(|T_{ii}| + |c_i|\Delta)\) a change of level occurs, to \(k+1\) if \(c_i>0\) and \(k-1\) if \(c_i<0\).
-The generator of their QBD approximation (for an unbounded fluid queue) is
-\[\bs B = \left[ \begin{array}{cccc} \ddots & \ddots & \ddots & \\ \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) & \\  & \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) \\ & \ddots & \ddots & \ddots \end{array} \right]\]
-\begin{align}
-\bs B_0 &= \bs T- \bs C\Delta, && \bs B_{-1} = \left[\begin{array}{ccc} \bs 0 & & \\ & \bs C_-\Delta & \\ & & \bs 0 \end{array}\right], && \bs B_{+1} = \left[\begin{array}{ccc} \bs C_+\Delta & & \\ & \bs 0 & \\ & & \bs 0\end{array}\right].
-\end{align}
-Bean and O'Reilly then take \(\Delta \to 0\) and show that the approximation converges weakly to the fluid queue.
-It seems that this is the best we can do if we want to keep the interpretation of the approximation as a QBD.
-We now elaborate on this point.
-
-For a given \(\Delta\), the discretisation of Bean and O'Reilly supposes that, when the phase is \(i\) and on the event of no change of phase, the sojourn time in a given level has an exponential distribution with rate \(|c_i|\Delta\).
-However, the corresponding sojourn time of the fluid queue is deterministic: given \(X(t) = x \in\calD_k\), then \(\{X(t)\}\) will leave \(\calD_k\) in exactly \(((k+1)\Delta - x)/|c_i|\) units of time if \(c_i>0\), and \((x-k\Delta)/|c_i|\) units of time if \(c_i<0\), on the event that the phase does not change before this time.
-We can extend the QBD model of Bean and O'Reilly to model this determinism more accurately by supposing that the sojourn times in each interval have Erlang distributions (rather than exponential).
-However, we effectively realise the same QBD approximation as the original, but on a finer discretisation.
-For example, using Erlang distributions of order 2 we would get a generator of the QBD approximation
-\[\bs B = \left[ \begin{array}{cccc} \ddots & \ddots & \ddots & \\ \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) & \\  & \bs B_{-1}(\Delta) & \bs B_0(\Delta) & \bs B_{+1}(\Delta) \\ & \ddots & \ddots & \ddots \end{array} \right]\]
-\begin{align}
-\bs B_0 &= \bs T\otimes I_2 - \left[ \begin{array}{ccc} \left[\begin{array}{cc} -\Delta/2 & \Delta/2 \\ 0 & -\Delta/2 \end{array}\right] \otimes \bs C_+ & & \\ & \left[\begin{array}{cc} -\Delta/2 & 0 \\ \Delta/2 & -\Delta/2 \end{array}\right] \otimes C_- & \\ & & \bs 0 \end{array}\right],
-\end{align}
-\begin{align}
-\bs B_{-1} &= \left[\begin{array}{ccc} \bs 0 & & \\ & \bs C_-\otimes \left[\begin{array}{cc} 0 & \Delta/2 \\ 0 & 0 \end{array}\right] & \\ & & \bs 0 \end{array}\right],
-&&\bs B_{+1} = \left[\begin{array}{ccc} \bs C_+\otimes \left[\begin{array}{cc} 0 & 0 \\ \Delta/2 & 0 \end{array}\right] & & \\ & \bs 0 & \\ & & \bs 0\end{array}\right].
-\end{align}
-But this is the same discretisation we would get if we were to use intervals of width \(\Delta/2\) in the QBD approximation of \cite{bo2013}, with the rows/columns reordered.
-Indeed, this appears to be the best we can do with a QBD approximation as the Erlang distribution is known to be the least variable Phase-type distribution for a given order \citep{as1987} and therefore gives the closest approximation to the deterministic behaviour we are trying to approximate.
-This necessitates the extension to QBD-RAPs, whereby we can find more concentrated distributions than any Phase-type and retain enough stochastic interpretation and matrix-analytic tools.
-
-Here, the construction of the approximating QBD-RAP is developed intuitively, from a stochastic modelling perspective.
-The key observation is that, given \(\{\varphi(t)\}\) is constant, then \(\{X(t)\}\) moves deterministically.
-Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic.
-We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \citep{ert2006,hstz2016,ert2006,hht2020,mt2021}.
-We construct a level process for the QBD-RAP and correspond it to a discretisation the level process of the fluid queue.
-This is similar to \cite{bo2013} where the level process of their QBD corresponds a discretisation of the fluid level.
-Unlike \cite{bo2013}, however, here we do not take \(\Delta \to 0\).
-Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small.
-% This necessarily means that the order of the matrix exponential increases.
-
-It turns out that the corresponding orbit process \(\{\bs A(t)\}\) can be seen to ``track", approximately, how far \(X(t)\) is from the left of a given interval when the phase is in \(\mathcal S_+\), or how far from the right of the interval \(X(t)\) is when the phase is in \(\mathcal S_-\).
-This leaves us with two problems.
-1) on a transition from \(\mathcal S_+\) to \(\mathcal S_-\) or \(\mathcal S_-\) to \(\mathcal S_+\) how must \(\bs A(t)\) jump to retain this information about where \(X(t)\) is within a given interval.
-2) how to use the orbit position at time \(t\), \(\bs A(t)\), to obtain an approximation for the density of \(X(t)\).
-Answers to both problems can be derived from the \emph{residual time} of a matrix exponential distribution.
-Let us now formalise these concepts some more.
-
-\section{Time to exit an interval}\label{sec: modelling}
-Consider the partition of the state space of the level of a fluid queue, \([0,M]\) into \(K\) intervals of width \(\Delta = M/(K+1)\), specifically \(\mathcal D_{k,i} = [k\Delta,(k+1)\Delta)\) if \(i\in\mathcal S_+\) and \(\mathcal D_{k,i} = (k\Delta,(k+1)\Delta]\) if \(i\in\mathcal S_-\), \(k=0,\dots,K+1\).
-The distinction between \(\mathcal D_{k,i}\) for \(i\in\mathcal S_+\) and \(\calS_-\) is a technical one and will be discussed later.
-For now, one reason we might want to specify the discretisation in this way, is that it ensures the stochastic process
-\begin{align}
-	\displaystyle \left\{\sum_{k\in\mathcal K}\sum_{i\in\calS} k1(X(t)\in\calD_{k,i})\right\}_{t\geq 0} \label{eqn: kfj}
-\end{align}
-is right-continuous with left limits (c\'adl\'ag).
-The expression (\ref{eqn: kfj}) is the discrete process of the fluid-queue which the level process of the QBD-RAP will approximate.
-Let \(y_k = k\Delta\), \(k=0,\dots,K+1\) and \(\calD_k = [y_k,y_{k+1}]\).
-We now look to model the behaviour of \(\{X(t)\}\) on an interval \(\calD_k\) over the time \([t,t+u]\) given information about the phase process over this time.
-
-\subsection{Modelling the residual time to exit on no change of phase} 
-We first consider no change of phase in \([t,t+u]\).
-Suppose that \(\varphi(t)=i\in\mathcal S_+\) and \(X(t) = x\in\mathcal D_{k,i}\).
-Observe that on no change of phase, \(\varphi(t+s) = i\), \(s\in[0,u]\), and at time \(t+s\) the level of the fluid queue is given by \(X(t+s) = x+c_i s\), \(s\in[0,u]\).
-Further, for \(u>(y_{k+1}-x)/c_i\) the level process \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(y_{k+1} - x)/c_i\).
-Similarly, for \(i\in\mathcal S_-\), \(\{X(t)\}\) leaves the interval \(\mathcal D_{k,i}\) at exactly time \(t+(x-y_k)/|c_i|\) on the event that there is no change of phase on \([t,t+u]\) for \(u>(x-y_k)/|c_i|\).
-
-To approximate the deterministic behaviour observed above, consider a matrix-exponential distribution \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) with mean \(\Delta\) and low variance.
-As we shall see later, the assumption that \(Z\) has low variance will allow us to claim that \(Z\) can be used to approximate deterministic behaviour.
-Let \(Z_i\sim ME(\bs \alpha, \bs S_i, \bs s_i)\) where \(\bs S_i = |c_i|\bs S\), \(\bs s_i=-\bs S_i\bs e\), \(i\in\calS\).
-The random variable \(Z_i\) has mean \(\Delta/|c_i|\).
-
-Suppose first that \(X(t)=y_k\) and \(\varphi(t)=i\in\mathcal S_+\), so that \(X(t)\) will leave \(\mathcal D_{k,i}\) in exactly \(\Delta/|c_i|\) units of time on the event that the phase process of the QBD-RAP remains in \(i\) for at least this amount of time.
-A sensible approximation for this deterministic event is to choose the value of the orbit process at time \(t\) to be \(\bs \alpha\), \(\bs A(t)=\bs\alpha\).
-With this choice, the distribution of time until the next event of the QBD-RAP, on the event that the phase process remains constant will be the distribution of \(Z_i\).
-Observe,
-\begin{align*}
-	\mathbb P(Z_i \in ((\Delta-\varepsilon)/|c_i|, (\Delta+\varepsilon)/|c_i|) ) 
-	&= \mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))
-%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
-%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
-	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2},
-\end{align*}
-since, by Chebyshev's inequality, \(\mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon)) \geq 1- \cfrac{\var(Z)}{\varepsilon^2}\). Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
-\[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
-when \(\var(Z)\) is small.
-Hence, when the variance of \(Z\) (equivalently \(Z_i\)) is low, then with high probability the time until the next event of the QBD-RAP, given the phase process remains constant will occur at approximately \(\Delta/|c_i|\).
-
-Now, let \(R_i(u)\) be the residual time, \(R_i(u) = (Z_i-u)1\{Z_i-u>0\}\), \(i\in\calS\).
-After \(u\in[0,\Delta/|c_i|)\) amount of time has elapsed, on the event that \(\varphi(t+s)=i\), for all \(s\in[0,u)\), then \(X(t+u) = y_k + c_i u\).
-If the phase remains \(i\) for a further \(\Delta/|c_i|-u\) amount of time, then \(\{X(t)\}\) will leave \(\calD_{k,i}\) at exactly this time.
-Also at time \(t+u\), given \(Z_i>u\), the distribution of the residual time, \(R_i(u)\), has density
-\[f_{R_i(u)}(r) = \cfrac{\bs\alpha e^{\bs S_i(u+ r) } \bs s_i}{\bs\alpha e^{\bs S_iu } \bs e} = \bs A(t+u) e^{\bs S_i r}  \bs s_i.\]
-Here, the event \(Z_i>u\) approximates the event \(X(t+u)\in\calD_{k,i}\) and we want the residual time, \(R_i(u)\), to approximate the time until \(\{X(t)\}\) leaves \(\calD_{k,i}\).
-That is, we want \(R_i(u)\) to approximate a deterministic random variable at \(\Delta/|c_i| - u\).
-To this end, we observe that for \(\varepsilon>0\) and \(u<(\Delta-\varepsilon)/|c_i|\), then
-\begin{align*}
-	&\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|-u) ) 
-	\\&= \mathbb P(Z_i\in ((\Delta-\varepsilon)/|c_i|, (\Delta+\varepsilon)/|c_i|) ) 
-	\\&= \mathbb P(Z\in (\Delta-\varepsilon, \Delta + \varepsilon))
-%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
-%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
-	\\&\geq 1-\cfrac{\var(Z)}{\varepsilon^2}.
-\end{align*}
-Choosing \(\varepsilon = \var(Z)^{1/3}\), then 
-\[\mathbb P(R_i(u) \in ((\Delta-\varepsilon)/|c_i|-u, (\Delta+\varepsilon)/|c_i|) )\geq 1-\var(Z)^{1/3} \approx 1, \]
-when \(\var(Z)\) is small. 
-That is, when the variance of \(Z\) (equivalently \(Z_i\)) is low, the residual time \(R_i(u)\) will be concentrated around \(\Delta/|c_i| - u\), as required.
-Figure \ref{fig:residual distributions} gives an example of a density function for a \(Z_i\) with mean \(\Delta=1\) and \(c_i=1\), as well as the density function of \(R_i(0.3)\) conditional on \(Z_i>0.3\) and \(R_i(0.6)\) conditional on \(Z_i>0.6\), for comparison.
-Observe that the density of the residual life \(R_i(0.3)\) conditional on \(Z_i>0.3\) is concentrated around \(\Delta-0.3 = 0.7\) and, similarly the density of the residual life \(R_i(0.6)\) conditional in \(Z_i>0.6\) is concentrated around \(\Delta-0.6 = 0.4\).
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{chapter2/figs/ME_residual_life_density.pdf}
-\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hht2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.6}\) (blue).
-Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.6}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively.
-}
-\label{fig:residual distributions}
-\end{figure}
-
-For notational convenience, define the row-vector-valued function \(\bs k(t):\mathbb R\to \mathcal A\), 
-\begin{equation}
-	\bs k(t) = \cfrac{\bs \alpha e^{\bs S t}}{\bs \alpha e^{\bs S t}\bs e}, \label{eqn: orbit fun def}
-\end{equation}
-and the operator \(\bs K(t): \mathcal A\to \mathbb R\), \(t\geq 0\), 
-\begin{equation}
-	\bs a \bs K(t) = \cfrac{\bs a e^{\bs S t}}{\bs a e^{\bs S t}\bs e}, \label{eqn: operator fun def}
-\end{equation}
-for any \(a\in\mathcal A\).
-
-We interpret the position of the orbit \(\bs A(t+u) = \bs k(|c_i|u)\) as corresponding to the fact that \(X(t+u)\) is \(|c_i|u\) units from the left-hand boundary of the interval \(\calD_{k,i}\).
-This gives a heuristic argument as to how we can model the sojourn times in a given interval \(\mathcal D_{k,i}\) on the event that the phase does not change.
-
-We can apply analogous arguments to heuristically develop a model for the sojourn time of the fluid queue in an interval, \(\calD_{k,i}\), \(i\in\calS_-\), on the event that the phase does not change, given the fluid is \(X(t)=y_{k+1}\).
-For \(i\in\calS_-\), the orbit position \(\bs A(t+u) = \bs k(|c_i|u)\) is interpreted as corresponding to \(X(t)\) being \(|c_i|u\) units from the right-hand boundary of the interval \(\calD_{k,i}\), \(y_{k+1}\).
-
-\subsection{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\)} 
-%Let \(Z_j\sim ME(\bs\alpha,\bs S_j)\) be a matrix exponential random variable with mean \(\Delta/|c_j|\) and let \(R_j(u)\) be the residual time \(R_j(u) = Z_j-u \mid\{Z_j-u>0\}\).
-%At time \(t+u\) the value of the orbit is \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\), which can be equivalently written as \(\bs A(t+u) =  \cfrac{\bs\alpha e^{\bs S_ju/|c_i| } }{\bs\alpha e^{\bs S_ju/|c_i| } \bs e}\).
-
-Let \(E\) be the event that, at time \(t\) the orbit of the QBD-RAP is \(\bs A(t) = \bs \alpha\), the phase process is \(\varphi(t) = i\in\calS_+\), and there are no events of the QBD-RAP before time \(t+u\).
-On the event \(E\) the orbit position at time \(t+u\) will be
-\begin{align}
-	\bs A(t+u) = \bs k(|c_i|u) = \cfrac{\bs \alpha e^{\bs S_i u}}{\bs \alpha e^{\bs S_i u}\bs e}.\label{eqn: orbit i}
-\end{align}
-Correspondingly, on the event that at time \(t\) the level process of the fluid queue is \(X(t)=y_{k}\), the phase is \(\varphi(t)=i\) and there are no change of phase by time \(t+u\), then \(X(t+u)=y_{k}+c_iu\).
-That is, \(X(t+u)\) is \(c_iu\) units from the left-hand edge of \(\calD_k\).
-
-If, at time \(t+u\) there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\), then we need to map the orbit position \(\bs A(t+u^-) = \bs k(|c_i|u)\) to an orbit position which corresponds to the being \(c_iu\) units from the left-hand edge of \(\calD_k\) in phase \(j\).
-Call this mapping \(\bs D(i,j)(\cdot)\).
-As noted in Section \ref{sec: qbd-rap}, the mapping must be linear,
-\(\bs D(i,j)(\bs a) = \bs a \bs D(i,j),\, \bs a \in\mathcal A,\)
-for some matrix \(\bs D(i,j)\). The matrix \(\bs D(i,j)\) is a modelling choice.
-
-So that the process is a QBD-RAP, the matrix \(\bs D(i,j)\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D(i,j), \bs S)\) is a representation of a matrix exponential distribution.
-We would also like the matrix \(\bs D(i,j)\) to have the property \(\bs D(i,j) \bs e = \bs e\) as this will mean that the rate at which a change of phase from \(i\) to \(j\) of the QBD-RAP occurs proportional to
-\[\bs A(t+u) \bs D(i,j)\bs e = 1,\]
-which is constant and therefore the distribution of time until a change from phase \(i\) to \(j\) is exponential.
-We can use this to show, for certain models, the distribution of the phase process of the fluid queue and the distribution of the phase process of the QBD-RAP share the same distribution.
-
-An orbit position which corresponds to the level process of the QBD-RAP beign \(c_iu\) units to the right of \(y_k\) in phase \(j\in\calS_+\) is 
-\[\bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|)}}{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|) u}\bs e}.\] 
-To see this, consider the event that at time \(t\) the orbit process of the QBD-RAP is \(\bs A(t) = \bs \alpha\), the phase process is \(\varphi(t) = j\in\calS_+\), and there are no events of the QBD-RAP before time \(t+|c_i|u/|c_j|\).
-On this event, the orbit position at time \(t+u\) will be
-\[\displaystyle \bs A(t+u) = \cfrac{\bs \alpha e^{\bs S_j (|c_i|u)/|c_j|}}{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|) u}\bs e}.\] 
-Correspondingly, on the event that at time \(t\) the level process of the fluid queue is \(X(t)=y_{k}\), the phase is \(\varphi(t)=j\) and there are no change of phase by time \(t+|c_i|u/|c_j|\), then \(X(t+u)=y_{k}+c_j|c_i|u/|c_j| = y_k+c_iu\).
-That is, \(X(t+u)\) is \(c_iu\) units from the left-hand edge of \(\calD_k\).
-
-Now, observe 
-\[\cfrac{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|)}}{\bs \alpha e^{\bs S_j (|c_i|u/|c_j|) u}\bs e} = \cfrac{\bs \alpha e^{\bs S_iu}}{\bs \alpha e^{\bs S_i u}\bs e} = \bs k(|c_i|u),\] 
-which is exactly (\ref{eqn: orbit i}), the orbit position on the event \(E\). Hence, it is sensible to choose \(\bs D(i,j)=\bs I\).
-
-% Consider the residual time \(R_j(|c_i|u/|c_j|)\), which is the distribution of time until \(Z_j-|c_i|u/|c_j|\) occurs, on the event \(Z_j>|c_i|u/|c_j|\).
-The initial vector associated with the residual time \(R_j(|c_i|u/|c_j|)\) is \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\).
-This is the initial vector which is obtained if at time \(t\) the phase is \(\varphi(t)=j\) and the orbit is \(\bs A(t)=\bs \alpha,\) so that the QBD-RAP has just entered level \(k\), and there is no change of phase by time \(|c_i|u/|c_j|\).
-Above, we argued that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) can be seen to correspond to the event that \(X(t+u)\) is \(|c_j|\left(|c_i|u/|c_j|\right) = |c_i|u\) units greater than \(y_k\).
-% %We argued earlier that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the position of the orbit on the event that at time \(t\) there is a change of level so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=j\), followed by no change of phase by time \(t+|c_i|u/|c_j|\).
-The analogous event of the fluid queue is that at time \(t\) the phase is \(\varphi(t)=j\), the fluid level is \(X(t)=y_k\), so that the fluid level has just entered an interval \(\calD_k\) at time \(t\), and there is no change of phase by time \(t+|c_i|u/|c_j|\), in which case the fluid level will be \(X(t+|c_i|u/|c_j|)=y_\ell + |c_j|\left(|c_i|u/|c_j|\right)=y_\ell + |c_i|u\).
-Therefore, we claim that the vector \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) is the correct position of the orbit to approximate the fact that \(\{X(t)\}\) is \(|c_i|u\) units to the right of the point \(y_k\).
-
-% Notice that we can re-write the vector 
-% \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) as \(\cfrac{\bs \alpha e^{\bs S_iu } }{\bs \alpha e^{\bs S_iu} \bs e }\).
-This is the orbit position on the event that at time \(t\) there is a change of level, so that \(\bs A(t)=\bs \alpha\), and the phase is \(\varphi(t)=i\), followed by no change of phase by time \(t+u\).
-We have argued that this is the orbit position which approximates the fact that \(\{X(t)\}\) is \(|c_i|u\) units greater than \(y_k\).
-
-% Therefore, at time \(t+u\), if there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\) then the orbit position should not change as in both cases the orbit position \(\cfrac{\bs \alpha e^{\bs S_j |c_i|u/|c_j| } }{\bs \alpha e^{\bs S_j |c_i|u/|c_j|} \bs e }\) corresponds to an approximation to \(X(t)\) being \(|c_i|u\) units greater than \(y_k\).
-
-%Therefore, suppose that at time \(t\), \(X(t)=y_k\), and that at time \(t+u\), \(u\in[0,\Delta/|c_i|)\), before \(\{X(t)\}\) exits the band \(\calD_{k}\), there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_+\).
-At time \(t+u\) the position of the fluid level is \(X(t+u) = y_k+c_iu\), which is \(|c_i|u\) units to the right of \(y_k\) and \(\Delta-c_iu\) to the left of \(y_{k+1}\), the right-hand boundary of \(\calD_{k,i}\).
-The residual time until the level of the fluid queue leaves \(\calD_{k,j}\), on the event that it remains in phase \(j\) until this time, is \((\Delta - |c_i|u)/|c_j|\).
-
-Moreover, the residual time \(R_j(|c_i|u/|c_j|)\) has density 
-\[f_{R_j(|c_i|u/|c_j|)}(r) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}e^{\bs S_j r}\bs s_j,\]
-%and corresponds to an orbit position of \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\). 
-and 
-\begin{align}
-	&\mathbb P(R_j(|c_i|u/|c_j|) \in ((\Delta-|c_i|u-\varepsilon)/|c_j|, (\Delta-|c_i|u + \varepsilon)/|c_j| )) \nonumber
-	\\&=\mathbb P(Z_j\in ((\Delta-\varepsilon)/|c_j|, (\Delta + \varepsilon)/|c_j|)) 
-	\\&=\mathbb P(Z\in (\Delta-\varepsilon ,  \Delta + \varepsilon )) 
-	% \\&=  \mathbb P(Z_j\in (\Delta/|c_j|-\varepsilon, \Delta/|c_j| + \varepsilon)) 
-%	\\&= \cfrac{\mathbb P(Z_i-u\in (\Delta/|c_i|-u-\varepsilon, \Delta/|c_i|-u + \varepsilon))}{\mathbb P(Z_i-u > 0)} 
-%	\\&\geq \mathbb P(Z-u\in (\Delta-u-\varepsilon, \Delta-u + \varepsilon))
-%	\\&\geq \cfrac{1-\cfrac{\var(Z_j)}{\varepsilon^2}}{1} 
-	\\&\geq 1-\var(Z)^{1/3} \approx 1.
-\end{align}
-Hence, when the variance of \(Z\) is low, the residual time, \(R_j(|c_i|u/|c_j|)\), is concentrated around \((\Delta - |c_i|u)/|c_j|\), as required.
-
-% Therefore, in our model, upon the event of a change from \(i\in\calS_+\) to \(j\in\calS_+\) at time \(t+u\) the orbit should be \(\bs A(t+u) = \cfrac{\bs\alpha e^{\bs S_iu } }{\bs\alpha e^{\bs S_iu } \bs e}\).
-That is, the initial condition for the next segment of the orbit process is just the current orbit position and the orbit should start to evolve according to the matrix \(\bs S_j\).
-
-Analogous arguments suggest that the same applies for changes of phase from \(i\in\calS_-\) to \(j\in\calS_-\).
-
-\subsection{On a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\)} 
-Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
-%We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\bs k(x)\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\).
-As before, we need to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \[\bs A(t+u^-) = \bs k(|c_i|u^-),\]
-to 
-\[\bs A(t+u) = \cfrac{\bs k(u^-)\bs D(i,j)}{\bs k(u^-)\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
-in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\) in phase \(j\).
-
-Again, the matrix \(\bs D(i,j)\) is a modelling choice. We first discuss how we might choose the matrix \(\bs D(i,j)\) for when the matrix-exponential \(Z\) is a phase-type distribution.
-
-\paragraph{The Phase-type case}
-If \(Z\sim ME(\bs \alpha, \bs S)\) is chosen to be a phase-type distribution then \(Z\) has the interpretation as the distribution of time until absorption of a finite-state continuous-time Markov chain with transient states \(\{1,\dots,p\}\) and a single absrobing state.
-The sub-generator matrix describing the dynamics of the Markov chain on transient states is \(\bs S\), and \(\bs \alpha\) is an initial probability distribution over the transient states.
-Let \(\{J(t)\}\) be the Markov chain associated with the phase-type distribution.
-
-In the discussions above, we have relied on the relationship between the event \(E\) and the orbit position \(\displaystyle \bs k(|c_i|u)\).
-The relationship allows us to associate this orbit position with the level of the fluid queue.
-For phase-type distributions the vector \(\bs k(|c_i|u)\) is the vector of posterior probabilities \(\left[\mathbb P(J(u)=k \mid E)\right]_{k\in\{1,...,p\}}\).
-We can use this vector of posterior probabilities in the same way as we do for matrix-exponential distributions and QBD-RAPs.
-However, the orbit interpretation forgoes the presence of the phase process \(\{J(t)\}\).
-Here, we will use the phase process \(\{J(t)\}\) to derive a choice of \(\bs D(i,j)\).
-
-Recall the position of the orbit process can be used to determine the residual time.
-This idea can be replicated when we are given the phase of the phase-type rather than the value of the orbit process.
-Given the phase of the phase-type distribution is \(k\in\{1,...,p\}\), the distribution of the residual time is
-\[\mathbb P(R_i\leq r \mid \mbox{phase}=k)=1-\bs e_k e^{\bs S_i r}\bs e.\]
-Notice that this is independent of the time since the phase-type distribution was initialised.
-We call the time since the phase-type distribution was initialised the \emph{age}.
-As noted by \cite{hmp2017}, the distribution of the age given the phase of the phase-type is \(k\in\{1,...,p\}\) depends on the sampling scheme which determines the observation time.
-Here, on the event that at time \(t\) the phase of the QBD-RAP is \(i\), a change of phase occurs after an exponential amount of time with rate \(-T_{ii}\).
-Proposition 4.1, \cite{hmp2017} states that the distribution function of the age, given the phase of the phase-type is \(k\) and the process is observed after an exponential time with rate \(-T_{ii}\) is
-\[\mathbb P(\mbox{age}\leq u \mid \mbox{phase}=k )=1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)u}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}.\]
-Let \(\widehat{ \bs S}_i(T_{ii}) = diag(\bs \nu)^{-1}\bs S_i'diag(\bs \nu)\), where \(\bs \nu = \bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}/(\bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e)\).
-Algebraic manipulations show
-\begin{align}
-	1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)u}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}
-	= 1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)u} \bs e,\label{eqn: skgj}
-\end{align}
-which is of Phase-type.
-
-Recall that in the matrix-exponential case we associated the value of the orbit process \(\bs k(|c_i|u)\) with being \(|c_i|u\) units to the right of \(y_{k}\).
-In contrast, for the phase-type case if we use the phase process \(\{J(t)\}\) to inform our modelling of the fluid queue (as opposed to the value of the orbit process), then we associate phase \(k\in\{1,...,p\}\) of \(\{J(t)\}\) with being a \emph{random} distance to the right of \(y_k\) where this distance has distribution
-\begin{align}
-	1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)u} \bs e.\label{eqn: skgj2}
-\end{align}
-Therefore, on the event \(E\), upon a change from of phase of the QBD-RAP from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), it seems reasonable to want the distribution of time until the next event of the QBD-RAP to be
-\begin{align}
-	1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)|c_j|u/|c_i| + T_{jj}Iu} \bs e.\label{eqn: skgj3}
-\end{align}
-The factor \(|c_j|u/|c_i|\) arises as a conversion between the speed at which the fluid level moves in phase \(j\) compared to phase \(i\).
-The rate \(T_{jj}\) in the exponent is the rate at which the next change of phase occurs.
-
-While this does achieve what we want, it is not quite satisfactory for the purpose of our approximation scheme due to dependence on the sample path of \(\{\varphi(t)\}\).
-Specifically, the evolution of the QBD-RAP from time \(t+u\) until the next event depends on the phase immediately before the change of phase at time \(t+u\), \(\varphi(t+u^-)=i\).
-This increases the size of the approximating QBD-RAP as we need a separate model for each \(\varphi(t+u^-)\in\calS_+\).
-Furthermore, we have not yet considered how to model any further changes from \(\calS_-\) to \(\calS_+\) or beyond, which further complicates matters.
-
-A solution is to suppose that, rather than observing the phase-type random variable at an exponential time with rate \(-T_{ii}\), we instead observe the process uniformly randomly on the lifetime of length \(Z_i\).
-Let \(\bs \Pi = diag\left(\bs \pi \right)\), where \(\vligne{\pi_k}_{k\in\{1,\dots,p\}}=:\bs \pi=\bs \alpha (-\bs{S})^{-1}/m\) and \(m=\bs \alpha (-\bs{S})^{-1}\bs e\).
-There is a time-reverse representation of a Phase-type distribution given by \((\widetilde{\bs \alpha}, \widetilde{\bs S}, \widetilde{\bs s})\), where \(\widetilde{\bs \alpha} = m\bs s' \bs \Pi \), \(\widetilde{\bs S} = \bs\Pi^{-1}\bs S'\bs \Pi\) and \(\widetilde{\bs s} = \bs \Pi^{-1}\bs \alpha'/m\) \cite[Page 91]{a2008}.
-In the case where the phase is observed randomly on the lifetime of \(Z_i\), the distribution function of the age, given the phase is \(k\) is \cite[Lemma 3.1]{hmp2017}
-\begin{align}\mathbb P(\mbox{age}\leq u \mid \mbox{phase}=k) = 1-\cfrac{\bs \alpha e^{\bs S_iu}(-\bs S_i)^{-1}\bs e_k}{\bs \alpha (-\bs S_i)^{-1}\bs e_k} = 1- \bs e_k e^{\widetilde{ \bs S}_iu}\bs e.\label{eqn:ksda2}\end{align}
-With this interpretation, we may associate phase \(k\in\{1,...,p\}\) of \(\{J(t)\}\) with being a \emph{random} distance, with distribution (\ref{eqn:ksda2}), to the right of \(y_k\).
-
-Therefore, on the event \(E\) and on a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), a reasonable model for the time until the next event of the QBD-RAP has distribution 
-\begin{align}
-	1- \bs e_k e^{\widetilde{ \bs S}_iu|c_j|/|c_i| + T_{jj}\bs Iu}\bs e = 1- \bs e_k e^{\widetilde{ \bs S}_ju+ T_{jj}\bs Iu}\bs e.\label{eqn:ksda3}
-\end{align}
-This suggests that, at a jump from \(\calS_+\) to \(\calS_-\), the state of \(\{J(t)\}\) does not change, but begins to evolve according to the time-reverse generator at an appropriate speed.
-Since the time-reverse of \(\widetilde{\bs S}\) is \(\bs S\), then upon a jump back to \(\calS_+\) from \(\calS_-\), the phase of the phase-type random variable remains \(k\) but begins to evolve according to \(\calS\).
-This suggests that we use the representation \((\bs \alpha, \bs S)\) when in phases in \(\calS_+\), and use the time-reverse representation \((\widetilde{\bs \alpha}, \widetilde{\bs S})\) when in phases in \(\calS_-\).
-The matrices \(\bs D(i,j)=\bs I\) for all \(i,j\in\calS\).
-
-With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2013} with discretisation parameter \(\Delta/p\).
-
-% Let \(\bs \Pi = diag\left(\bs \pi \right)\), where \(\vligne{\pi_k}_{k\in\{1,\dots,p\}}=:\bs \pi=\bs \alpha (-\bs{S})^{-1}/m\) and \(m=\bs \alpha (-\bs{S})^{-1}\bs e\).
-There is a time-reverse representation of a Phase-type distribution given by \((\widetilde{\bs \alpha}, \widetilde{\bs S}, \widetilde{\bs s})\), where \(\widetilde{\bs \alpha} = m\bs s' \bs \Pi \), \(\widetilde{\bs S} = \bs\Pi^{-1}\bs S'\bs \Pi\) and \(\widetilde{\bs s} = \bs \Pi^{-1}\bs \alpha'/m\) \cite[Page 91]{a2008}.
-
-% Rather than use the residual time \(R_i(u)\) to estimate how far \(X(t)\) is from the edge of an interval, we can use the information contained in \(J(t+u)\).
-Suppose there is a change of phase of the fluid queue from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\) at which time \(J(t+u)=k\).
-At a random time before \(t+u\) the Phase-type random variable was initialised.
-Let \(C_{t+u}\) be the \emph{age} of the Phase-type distribution at time \(t+u\), that is, the time since the Phase-type random variable was initialised.
-Suppose that \(\varphi(v)=i\), \(v\in [t+u-C_{t+u}, t+u)\), that is, there has been no change of phase of the fluid queue during the current life of the Phase-type random variable.
-Given \(J(t+u)=k\), we wish to construct an approximation to the distribution of \(C_{t+u}\).
-
-% According to \cite{hmp2017}, the distribution of \(C_{t+u}\) depends on the method of observation.
-Here, the Phase-type distribution is observed after an exponential time with rate \(-T_{ii}\).
-Proposition 4.1, \cite{hmp2017} states that the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) and the process is observed after an exponential time with rate \(-T_{ii}\) is
-% \[\mathbb P(C_{t+u}\leq a )=1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}.\]
-% Let \(\widehat{ \bs S}_i(T_{ii}) = diag(\bs \nu)^{-1}\bs S_i'diag(\bs \nu)\), where \(\bs \nu = \bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}/(\bs\alpha(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e)\).
-Algebraic manipulations show
-% \begin{align}
-% 	1-\cfrac{\bs \alpha e^{(\bs S_i + T_{ii}\bs I)a}(-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}{\bs \alpha (-(\bs S_i+ T_{ii}\bs I))^{-1}\bs e_k}
-% 	= 1-\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)a} \bs e,\label{eqn: skgj}
-% \end{align}
-% which is of Phase-type.
-The conversion between the intensity at which events occurs when the fluid queue is in phase \(i\) compared to phase \(j\) is \(|c_j|/|c_i|\).
-%Taking the derivative of (\ref{eqn: skgj}) with respect to \(v\), and applying the chain rule, gives
-%\[\cfrac{\wrt }{\wrt u}\bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)u} \bs e \cfrac{|c_j|}{|c_i|} = \bs e_k' e^{(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)|c_j|v} \widehat{\bs s}_{ij}, \]
-%where \(\widehat{\bs s}_{ij} = (\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\cfrac{|c_j|}{|c_i|} \bs e\). 
-% Therefore, when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), if the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the generator \(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I)\), then the distribution of time until the next event in the QBD-RAP after time \(t+u\) is
-% \[1-\bs e_k e^{(|c_j|/|c_i|(\widehat{\bs S}_i(T_{ii}) + T_{ii}\bs I) + T_{jj}\bs I)a}\bs e.\]
-% While this does approximate the desired quantity, it is not quite satisfactory for the purpose of our approximation scheme due to dependence on the sample path of \(\{\varphi(t)\}\).
-Specifically, the evolution of the QBD-RAP from time \(t+u\) until the next event depends on the phase immediately before the change of phase at time \(t+u\), \(\varphi(t+u^-)=i\).
-This increases the size of the approximating QBD-RAP as we need a separate model for each \(\varphi(t+u^-)\in\calS_+\).
-Furthermore, we have not yet considered how to model further changes from \(\calS_-\) to \(\calS_+\) or beyond, which further complicates matters.
-
-% A solution is to suppose that, rather than observing \(C_{t+u}\) at an exponential time with rate \(-T_{ii}\), we instead observe \(\{J(t+u)\}\) uniformly randomly on the lifetime of length \(Z_i\).
-In this case, the cumulative distribution function (CDF) of \(C_{t+u}\), given \(J(t+u)=k\) is \cite[Lemma 3.1]{hmp2017}
-% \begin{align}\mathbb P(C_{t+u}\leq a) = 1-\cfrac{\bs \alpha e^{\bs S_iu}(-\bs S_i)^{-1}\bs e_k}{\bs \alpha (-\bs S_i)^{-1}\bs e_k} = 1- \bs e_k e^{\widetilde{ \bs S}_iu}\bs e.\label{eqn:ksda}\end{align}
-% Once again, noting that the conversion between the intensity at which events occurs when the fluid queue is in phase \(i\) compared to phase \(j\) is \(|c_j|/|c_i|\)%, and taking the derivative with respect to \(v\) of (\ref{eqn:ksda}) by applying the chain rule, gives
-% %\begin{align}%
-% %	\cfrac{\wrt }{\wrt u} \bs e_k e^{\widetilde{ \bs S}_iu}\bs e|c_j|/|c_i|=\bs e_k e^{\widetilde{ \bs S}_iv|c_j|/|c_i|}\widetilde{ \bs S}_i\bs e|c_j|/|c_i| = \bs e_k e^{\widetilde{ \bs S}_jv}\widetilde{ \bs s}_j.\label{eqn:ksdadsva}
-% %\end{align}
-% then, upon a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), the Phase-type process which approximates the fluid level starts in phase \(k\) and evolves according to the reverse-time generator \(\widetilde{ \bs S}_j\).
-The distribution of time until the next event in the QBD-RAP after time \(t+u\) is
-% \[1-\bs e_k e^{(\widetilde{\bs S}_j + T_{jj}\bs I)a}\bs e.\] 
-% This suggests that, at a jump from \(\calS_+\) to \(\calS_-\) or from \(\calS_-\) to \(\calS_+\), the state of \(\{J(t+u)\}\) does not change, and begins to evolve according to the time-reverse generator at an appropriate speed.
-
-% With this construction, and choosing \(Z\sim Erlang(p,\Delta/p)\), we recover the discretisation of \cite{bo2013} with discretisation parameter \(\Delta/p\).
-
-\paragraph{The matrix-exponential case}
-% Now suppose \(X(t)=y_k\), \(\varphi(t)=i\in\calS_+\), and the phase remains in state \(i\) until there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), \(u\in[0,\Delta/|c_i|)\).
-We have reasoned, intuitively, that, for \(i\in\calS_+\) and for \(x=|c_i|u\in[0,\Delta)\), the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_k+x\), that is \(x\) units greater than \(y_k\).
-Correspondingly, the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units less than \(y_{k+1}\).
-We have also reasoned that for \(j\in\calS_-\) the orbit position \(\cfrac{\bs\alpha e^{\bs Sx} }{\bs\alpha e^{\bs Sx } \bs e}\) corresponds to the fluid level position \(X(t) = y_{k+1}-x\), that is \(x\) units less than \(y_{k+1}\), and the residual time \(R_i(u)\) \emph{approximates} the fact that \(X(t)\) is \(\Delta-x\) units greater than \(y_k\).
-Notice that in positive phases the orbit position describes the exact position of \(X(t)\) from the left-hand boundary of the interval but only approximates the position of \(X(t)\) from the right-hand boundary of the interval, and vice-versa for negative phases.
-
-% As before, we need to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\), \(\bs A(t+u^-) = \cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\) to 
-% \[\bs A(t+u) = \cfrac{\bs A(t+u^-)\bs D(i,j)}{\bs A(t+u^-)\bs D(i,j)\bs e}=\cfrac{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)}{\cfrac{\bs\alpha e^{\bs S_iu} }{\bs\alpha e^{\bs S_iu } \bs e}\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
-% in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\)or, equivalently, \(\Delta-|c_i|u\) from \(y_{k+1}\).
-
-% The matrix \(\bs D\) must have the property that, for any \(\bs a\in\mathcal A\), \((\bs a\bs D, \bs S)\) is a representation of a matrix exponential distribution.
-We would also like the matrix \(\bs D\) to have the property \(\bs D \bs e = \bs e\) as this will mean that the rate at which a change of phase from \(i\) to \(j\) of the QBD-RAP occurs proportional to
-% \[\bs A(t+u) \bs D(i,j)\bs e = 1,\]
-% which is constant and therefore the distribution of time until a change from phase \(i\) to \(j\) is exponential.
-We can use this to show that for certain models, the distribution of the phase process of the fluid queue and the distribution of the phase process of the QBD-RAP share the same distribution.
-
-For matrix exponential distributions we cannot rely on the phase process \(\{J(t)\}\) as we did in the phase-type case.
-Recall that we want to find a matrix \(\bs D(i,j)\) to map the orbit position on the event \(E\) when there is a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) at time \(t+u\), from \(\bs A(t+u^-) = \bs k(|c_i|u)\), to
-\[\bs A(t+u) = \cfrac{\bs k(|c_i|u)\bs D(i,j)}{\bs k(|c_i|u)\bs D(i,j)\bs e}=\cfrac{\bs\alpha e^{\bs S_iu} \bs D(i,j)}{\bs\alpha e^{\bs S_iu}\bs D(i,j)\bs e}\]
-in such a way that the orbit position after the jump corresponds, in some way, to the fluid level being a distance of \(|c_i|u\) from \(y_k\) in phase \(j\).
-
-Given our interpretation of the orbit position, \(\bs k(x)\), a solution would be find a linear map which takes \(\bs k(x)\) and maps it directly to 
-\[\bs k(\Delta-x) = \cfrac{\bs\alpha e^{\bs S(\Delta-x)} }{\bs\alpha e^{\bs S(\Delta-x) } \bs e}.\] However, we have been unsuccessful in finding such a mapping.
-The main hurdle being that the map must be linear.
-Instead, we approximate as follows.
-
-Recall that we use \(R_i(u)\) to approximate the distance of the fluid level to the left of \(y_{k+1}\).
-Suppose we are given \(R_i(u)=r\) which corresponds, approximately, to the fluid level being \(|c_i|r\) units to the left of \(y_{k+1}\).
-The position of the orbit process which corresponds to a distance of \(|c_i|r\) to the left of \(y_{k+1}\) in phase \(j\) is \(\bs k(|c_i|r).\)
-Hence, on \(R_i(u)=r\), the event \(E\) and on the event that a change of phase from \(i\in\calS_+\) to \(j\in\calS_-\) occurs at time \(t+u\), the orbit position should jump from \(\bs k(|c_i|u)\) to \(\bs k(|c_i|r)\).
-However, at time \(t+u\), \(R_i(u)\) is a random variable about the future of the process and therefore not known, so instead, we take the expected initial vector
-\[\mathbb E \left[ \bs k(|c_i|R_i(u))\right] = \int_{r=0}^\infty \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} e^{\bs S_ir } \bs s_i \bs k(|c_i|r)\wrt r = \cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{r=0}^\infty e^{\bs S_ir } \bs s_i\cfrac{\bs\alpha e^{ \bs S_ir } }{\bs\alpha e^{ \bs S_ir } \bs e}\wrt r.\]
-After a change of variables \(x=|c_i|r\) we get 
-\[\cfrac{\bs\alpha e^{\bs S_i u } }{\bs\alpha e^{\bs S_iu } \bs e} \int_{x=0}^\infty e^{\bs S x }|c_i| \bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x/|c_i| = \bs A(t+u^-)\int_{x=0}^\infty e^{\bs S x }\bs s \cfrac{\bs\alpha e^{ \bs Sx } }{\bs\alpha e^{ \bs Sx } \bs e}\wrt x,\]
-since at time \(t+u^-\), the orbit position is \(\bs A(t+u^-) = {\bs\alpha e^{\bs S_i u } }/{\bs\alpha e^{\bs S_iu } \bs e}\). Thus, we find 
-\[\bs D(i,j)=\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \bs k(x)\wrt x=:\bs D.\]
-
-Observe that \(\bs D\bs e = \displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \bs k(x)\wrt x\bs e =\displaystyle \int_{x=0}^\infty e^{\bs S x }\bs s \wrt x=\bs e\), since \(\bs k(x)\bs e = 1\) for all \(x\geq 0\).
-Further, since \(\mathcal A\) is closed and convex \citep{MEinAP}, then \((\bs a \bs D,\bs S,\bs s)\) is a representation of a matrix exponential distribution for any \(\bs a\in\mathcal A\).
-
-We pose the choice of the matrix \(\bs D\) as a modelling choice.
-Other choices are possible, for example, \(\displaystyle \bs D = \int_{x=0}^{\Delta}e^{\bs S x }\bs s \bs k(x)\wrt x\) is another sensible choice.
-It may also be possible to construct other matrices \(\bs D\), perhaps via geometric arguments.
-
-\paragraph{Computing \(\bs D\)}
-In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hht2020}.
-For this class of CMEs, we take the index \(p\) to be the order of the representation.
-Moreover, we take \(p\) to be odd.
-The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}.
-Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform only marginally better than an approximation constructed with representations of order \(2p-1\).
-However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.
-
-For a given CME with odd order, \(p\), and representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part.
-
-We numerically evaluate the matrix \(\bs{D}\) where
-\begin{align*}
-	\bs{D} = \displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\bs k(t)\wrt t
-\end{align*}
-using a trapezoidal rule as follows.
-For the CMEs considered here, the vector function \(\bs k(t)\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
-Let \(\bs f(t) = e^{\bs S t}\bs s\).
-Then \(\bs f(t)e^{-\lambda t}\) where \(\lambda = \Re(\lambda_i)\) is the real part of the eigenvalues of \(\bs{S}\) (they all share the same real part), is also periodic with the same period.
-Hence we can simplify the integral to a finite one;
-\begin{align}
-	\nonumber\bs{D} &= \displaystyle \int_{t=0}^\infty \bs f(t)\cdot\bs k(t)\wrt t  %\displaystyle \int_{t=0}^\infty e^{\bs{S}t}\bs s\cdot\cfrac{\bs \alpha e^{\bs{S}t}}{\bs \alpha e^{\bs{S}t} \bs e}\wrt t
-	\\\nonumber&= \sum_{k=0}^\infty \int_{k\rho }^{(k+1)\rho } e^{\lambda t} e^{-\lambda t} \bs f(t)\cdot\bs k(t)\wrt t
-	\\&= \sum_{k=0}^\infty \int_{0}^{\rho }  e^{\lambda(k\rho +t)}e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs k(k\rho +t)\wrt t. \label{eqn: sc6789gGHJ}
-	\intertext{By periodicity, then \(e^{ -\lambda(k\rho  + t)} \bs f(k\rho +t)\cdot\bs k(k\rho +t)=e^{ -\lambda t} \bs f(t)\cdot\bs k(t)\), hence (\ref{eqn: sc6789gGHJ}) is equal to}
-	&\sum_{k=0}^\infty (e^{\lambda \rho })^k\int_{0}^{\rho } e^{\lambda t}e^{-\lambda t}\bs f(t)\cdot\bs k(t)\wrt t  
-	= \cfrac{1}{1 - e^{\lambda \rho }}\int_{0}^{\rho } \bs f(t)\cdot\bs k(t)\wrt t,\label{eqn: hdsv45KJ}
-\end{align}
-where the sum converges as it is a geometric series and \(\lambda <0\), \(\rho >0\).
-
-To approximate (\ref{eqn: hdsv45KJ}) numerically, we first partition \(\left[0,\rho \right)\) into \(N\) equal-width intervals \(\left[t_n,t_{n+1}\right)\), where \(t_n = (n-1)\rho /N\), \(n=1,2,...,N+1\).
-On \(\left[t_n,t_{n+1}\right)\) we approximate the orbit \(\bs k(t)\) by a constant \(\bs k(t) \approx \bs k_n := \cfrac{1}{2}\left( \bs k(t_n) + \bs k(t_{n+1}) \right), \, t \in \left[t_n,t_{n+1}\right)\).
-Substituting this approximation into the expression for \(\bs{D}\) gives
-\begin{align*}
-	\bs{D} &\approx \cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\int_{t_n}^{t_{n+1}} \bs f(t)\cdot\bs k_n\wrt t
-	\\&=\cfrac{1}{1 - e^{\lambda \rho }}\sum_{n=1}^N\left[e^{\bs{S}t_{n+1}}-e^{\bs{S}t_{n}}\right]\bs e \cdot\bs k_n.
-\end{align*}
-This approximation preserves the property that \(\bs{D}\bs e = \bs e\).
-
-Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs k(t)\) are provided in \citep{hht2020}.
-
-\subsection{Upon exiting \(\calD_{k,i}\)} Suppose that upon exiting \(\calD_{k,i}\) at time \(t\) the phase is \(\varphi(t)=i\in\calS_+\).
-At this time \(X(t)=y_{k+1}\) which is the left-hand endpoint of \(\calD_{k+1,i}\).
-Hence, we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
-Similarly, upon exiting \(\calD_{k,i}\) at time \(t\) in phase \(\varphi(t)=i\in\calS_-\), then \(X(t)=y_k\), which is the right-hand endpoint of \(\calD_{k,i}\), and so we restart the model of the sojourn time with the initial condition \(\bs A(t)=\bs \alpha\).
-
-\section{The association of \(j\in\calS_0\) with \(\calS_+\) or \(\calS_-\)}\label{sec: zero states}
-Let \(X(0)=y_\ell\) and consider the event where \(\left\{\varphi(t)\right\}\) transitions from \(j_0\to j_1\to j_2\) where \(j_0\in\calS_+\), \(j_1\in\calS_0\) and \(j_2\in\calS_-\), before there is a change of level, i.e.~
-\[\varphi(t) = \begin{cases}
-	j_0 & t\in\left[0,t_1\right), \\
-	j_1 & t\in\left[t_1,t_2\right), \\
-	j_2 & t\in\left[t_2,t_3\right),
-\end{cases}\]
-and \(X(t)\in\mathcal D_\ell\), \(t\in[0,t_3)\). 
-On approximating this event, the initial orbit position is \(\bs A(0)=\bs \alpha\). The corresponding sample path of the orbit process on \(t\in[0,t_3)\), is 
-\begin{align*}
-\bs A(t) &= \begin{cases} 
-	\cfrac{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t}}{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
-	\cfrac{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t-t_1)}}{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t-t_1)}\bs e} & t\in\left[t_1,t_2\right), \\[10pt]
-	\cfrac{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t_2-t_1)}\bs{D}({j_1},{j_2})e^{(\bs{S}_{j_2}+T_{{j_2}{j_2}}\bs{I})(t-t_2)}}{\bs \alpha e^{(\bs{S}_{j_0}+T_{{j_0}{j_0}}\bs{I})t_1}\bs{D}({j_0},{j_1})e^{T_{{j_1}{j_1}}(t_2-t_1)}\bs{D}({j_1},{j_2})e^{(\bs{S}_{j_2}+T_{{j_2}{j_2}}\bs{I})(t-t_2)}\bs e} & t\in\left[t_2,t_3\right),
-\end{cases}
-%
-\\&= \begin{cases} 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})e^{\bs{S}_{j_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})e^{\bs{S}_{j_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right),
-\end{cases}
-\end{align*}
-for some matrices \(\bs{D}({j_0},{j_1})\) and \(\bs{D}({j_2},{j_1})\).
-Notice that \(\left\{\bs A(t)\right\}\) is constant on \(t\in\left[t_1,t_2\right)\).
-The matrix product \(\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})\) is there to capture the change in direction due to the change from \(\calS_+\) to \(\calS_-\).
-Hence \(\bs{D}({j_0},{j_1})\bs{D}({j_1},{j_2})\) should be equal to \(\bs{D}\).
-These types of sample paths are the reason we need to associate states \({j_1}\in\calS_0\) with either \(\calS_+\) or \(\calS_-\).
-
-% \paragraph{Other approaches}
-% There are other possible approaches to model this behaviour which do not require duplicating states and may achieve a small computational saving, however they are less amenable to analysis and possibly lead to less accurate schemes.
-Instead of duplicating states in \(\calS_0\), we can choose to associate \(j_1\in\calS_0\) with either \(\calS_+\) or \(\calS_-\).
-
-Associating \({j_1}\) with \(\calS_+\), amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{I}\) and \(\bs{D}({j_1},{j_2})=\bs{D}\); associating \({j_1}\) with \(\calS_-\), amounts to choosing \(\bs{D}({j_0},{j_1})=\bs{D}\) and \(\bs{D}({j_1},{j_2})=\bs{I}\).
-
-There are some consequences to this choice.
-Let \(k_2\in \calS_+\).
-Consider an event where the phase process of the fluid queue transitions from \({j_0}\to {j_1}\to k_2\) and there is no change of level.
-If \({j_1}\) is associated with \(\calS_+\), then \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{I}\) and the corresponding orbit process, given \(\bs A(0) = \bs \alpha\), is
-\begin{align*}
-\bs A(t) &= \begin{cases} 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
-\end{cases}
-\end{align*}
-%This sample path of the orbit process captures the fact that the process initially spent \(t_1-t_0\) amount of time in \({j_0}\in\calS_+\), exactly.
-Hence, at time \(t_2\), the orbit process knows that it has spent exactly \(t_1-t_0\) amount of time moving at rate \(c_{j_0}\).
-Notice that there is no matrix \(\bs D\) in this expression.
-
-Compare this to if \({j_1}\) is associated with \(\calS_-\).
-In this case \(\bs{D}({j_0},{j_1})=\bs{D}({j_1},k_2)=\bs{D}\) and the corresponding orbit process of the approximation, given \(\bs A(0) = \bs \alpha\) and there are no change of level in \(\left[t_1,t_3\right)\), is
-\begin{align*}
-\bs A(t) &= \begin{cases} 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t}}{\bs \alpha e^{\bs{S}_{{j_0}}t}\bs e} & t\in\left[0,t_1\right), \\[10pt]
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs e} & t\in\left[t_1,t_2\right), \\[10pt] 
-	\cfrac{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}}{\bs \alpha e^{\bs{S}_{j_0}t_1}\bs{D}\bs{D}e^{\bs{S}_{k_2}(t-t_2)}\bs e} & t\in\left[t_2,t_3\right).
-\end{cases}
-\end{align*}
-Ideally \(\bs D^2=\bs I\), however this is not the case here.
-Recall that a jump according to \(\bs{D}\) corresponds to approximating the residual life by an expectation.
-With this interpretation as an approximation, it suggests that we might want to minimise the number of jumps according to \(\bs D\) which occur.
-Therefore, for \(j_1\in \calS_0\), if transitions \(\calS_+\to j_1\to \calS_+\) occur with high probability compared to transition \(\calS_-\to j_1\to \calS_-\), then this suggests we might want to associate \(j_1\) with \(\calS_+\).
-Which association is chosen will depend on the parameters of the fluid queue and on which aspects of the model we wish to approximate.
-Although this advice is based on intuition only, numerical results in Section TBC suggest that it is reasonable.
-
-% It might be tempting to choose \(\bs D(j_0,j_1)=\bs D(j_1,j_2)=\bs D^{1/2}\) (provided that the square root matrix exists).
-With this choice \(\bs D(j_0,j_1)\bs D(j_1,j_2)=\bs D\) as desired.
-However, for sample paths which transition from \(j_0\to j_1\to k\), we have \(\bs D(j_0,j_1)\bs D(j_1,k) = \bs D^{1/2}\bs D(j_1,k)\).
-
-\subsubsection{Augmented state-space schemes}\label{subsec: augmented }
-Another way to approach the problem is to augment the state space of the phase process by duplicating \(\calS_0\) and associating one copy of \(\calS_0\) with \(\calS_+\) and one copy of \(\calS_0\) with \(\calS_-\).
-Let \(\left\{\varphi^*(t)\right\}\) be the augmented CTMC with state space \(\calS^*\) and generator \(\bs{T}^*\).
-Let \(\calS_+\) and \(\calS_-\) be as before and \(\calS_{m0} = \left\{(m,i)\mid i\in\calS_m\right\}\), \(m\in\left\{+,-\right\}\), then \(\calS^* = \calS_+\cup\calS_-\cup\calS_{+0}\cup\calS_{-0}\).
-The generator of \(\varphi^*(t)\) can be written as 
-\[\bs{T}^* = \left[\begin{array}{cccc} \bs{T}_{++} & \bs{T}_{+0} & \bs{T}_{+-} & 0 \\ \bs{T}_{0+} & \bs{T}_{00} & \bs{T}_{0-} & 0 \\ \bs{T}_{-+} & 0 & \bs{T}_{--} & \bs{T}_{-0} \\  \bs{T}_{0+} & 0 & \bs{T}_{0-} & \bs{T}_{00} \end{array}\right].
-\]
-Also define a fluid level \(\left\{X^*(t)\right\}\) using \(\left\{\varphi^*(t)\right\}\), with rates \(c_i^*=c_i\) for \(i\in\calS_+\cup\calS_-\) and \(c_{(m,i)}^* = 0\) for \((m,i)\in\calS_{+0}\cup\calS_{-0}\).
-The process \(\left\{\varphi(t)\right\}\) is imbedded within \(\left\{\varphi^*(t)\right\}\) and is recovered by marginalising over \(\calS_{+0}\) and \(\calS_{-0}\).
-On the event \(X^*(0)=X(0)\), the fluid levels \(X^*(t)\) and \(X(t)\) match exactly.
-Hence, by approximating \(\left\{(X^*(t),\varphi^*(t))\right\}\), we can recover an approximation to \(\left\{(X(t),\varphi(t))\right\}\).
-This construction removes the problem of having to choose how to associate states \(j\in\calS_0\) with either \(\calS_+\) or \(\calS_-\)
-
-The generator for the QBD-RAP approximation to the augmented fluid process (ignoring boundaries) is 
-\begin{align*}
-\bs B &= \left[\begin{array}{cccccc}
- \ddots & \ddots & \ddots & & & \\
- & \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & & \\ 
- & &  \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & \\
- & & & \ddots & \ddots & \ddots 
-\end{array}\right],
-\intertext{where,}
-\bs{B}_{0} &= \left[\begin{array}{cccc}
-	\bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I}  & \bs{T}_{+0}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \bs 0 \\
-	\bs{T}_{0+}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \bs 0 \\
-	\bs{T}_{-+}\otimes \bs{D} & \bs 0 & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & \bs{T}_{-0}\otimes \bs{I} \\
-	\bs{T}_{0+}\otimes \bs{D} & \bs 0 & \bs{T}_{0-}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} 
-	\end{array}\right],
-\end{align*}
-\begin{align*}
-\bs{B}_{-1} &= \left[\begin{array}{cccc}
-	\bs 0 & & & \\
-	& \bs 0 & & \\
-	& & \bs{C}_-\otimes (\bs s\bs\alpha) & \\
-	& & & \bs 0
-	\end{array}\right],
-%
-&& \bs{B}_{+1} = \left[\begin{array}{cccc}
-	\bs{C}_+\otimes (\bs s\bs\alpha) & & & \\
-	& \bs 0 & & \\
-	& & \bs 0 & \\
-	& & & \bs 0
-	\end{array}\right].
-\end{align*}
-With this construction, jumps according to the matrix \(\bs D\) occur only on transitions from \(\calS_m\to\calS_n\), or \(\calS_m\to\calS_{0m}\to\calS_{n}\), \(m,n\in\{+.-\}\), \(m\neq n\).
-
-%Hence, at time \(t_2\) the process no longer knows that it has spent exactly \(t_1-t_0\) amount of time moving at rate \(c_{j_0}\); rather, it knows some approximation of this information in the form of an expectation.
-
-%\section{Motivation and construction of approximations}
-%
-%We can approximate the generator of a fluid queue as a QBD, either by considering a uniformisation of the underlying phase process (as in \citep{bo2013}), or using a finite-volume scheme (see the DG paper, to be submitted).
-Let \(\boldsymbol M(t) = (X(t),\varphi(t))_{t\geq 0}\) be a fluid queue, with finite state space \(\mathcal S = \left\{1,2,...,N\right\}\), \(N\times N\) generator matrix \(\bs{T}\), and rates \(c_i\) (stored in the \(N\times N\) diagonal matrix \(\bs{C}\)).
-Also define the \(N\times N\) diagonal matrices \(\bs{C}_+ = \diag( c_i1(c_i>0),\,i\in\mathcal S)\), \(\bs{C}_- = \diag( c_i1(c_i<0),\,i\in\mathcal S)\), and define \(\calD_k = \left[k \Delta, (k+1)\Delta\right]\).
-%
-%We approximate the fluid model by the QBD process \(\boldsymbol J(t) = (L(t),\varphi(t))\) with state space \(\mathbb Z\times \mathcal S\) and generator defined by the block matrices \(A_0 = \bs{T}- |\bs{C}|/\Delta\), \(A_1 = |\bs{C}_+|/\Delta\), \(A_{-1} = |\bs{C}_-|\Delta\), where \(|\cdot|\) takes the absolute values of the elements of the matrix.
-The phase variable of \(\bs J(t)\) and \(\bs M(t)\) match exactly.
-The process, \(L(t)\Delta\), approximates the fluid height \(X(t)\), and converges as \(\Delta \to 0\) \citep{bo2013}.
-We also have the approximations
-%\[\bb P(X(t)\in\calD_\ell,\varphi(t)=j\mid X(0)\in\calD_k,\varphi(0)=i)\approx \bb P(L(t) = \ell, \varphi(t) = i \mid L(0) = k, \varphi(0) = i).\] 
-%
-%These approximations are particularly nice as the ensure positivity and preservation of probability due to their stochastic interpretation.
-However, they are low order methods compared to more sophisticated numerical schemes such as DG.
-We wish to develop a numerical approximation method which is high order and has a probabilistic interpretation.
-%
-%As motivation, consider again the QBD approximation, \(\bs J(t)\), to a fluid queue, \(\bs M(t)\), as described above.
-When the state is \((L(t),\varphi(t)) = (\ell, i)\), the process spends an exponentially distributed amount of time in this phase, with rate \(q_{i} = T_{ii}-|c_i|/\Delta\), after which there is a change of phase from phase \(i\) to phase \(j\) with probability \(T_{ij}/|q_i|\) while the level remains the same, otherwise there is a change of level: either \(L(t)\) increases from \(\ell\) to \(\ell +1\) if \(c_i>0\), or decreases from \(\ell\) to \(\ell-1\) if \(c_i<0\).
-%
-%Rather than approximate the sojourn time of \((X(t),\varphi(t)) \in(\calD_k, i)\) by an exponential distribution, we could instead approximate this sojourn time by a phase-type distribution denoted \(\tau_i \sim PH(\bs \alpha, T_{ii}\bs{I}+\bs{S}_i)\) where \(-\bs\alpha \bs{S}^{-1}\bs 1 = \Delta\).
-That is, the mean of the Phase-type \(PH(\bs \alpha,\bs{S})\) is equal to the cell width, \(\Delta\).
-The distribution \(PH(\bs \alpha, T_{ii}\bs{I}+\bs{S}_i)\) can be viewed as \(min(U,V)\) where \(U\sim exp(T_{ii})\) and \( V\sim PH(\bs\alpha, \bs{S})\).
-Thus, expiry of the distribution \(\tau_i\) corresponds to either a change in phase if \(U\) is the minimum, or change in level if \(V\) is the minimum.
-The latter case we map to a change in level: \(L(t)\) moves from \(\ell\) to \(\ell+1\) if \(c_i>0\), and from \(\ell\) to \(\ell-1\) if \(c_i<0\).
-%
-%Now, we know that, given \(\varphi(t)=i\), and given that \(X(0)=y_k\), then \(X(t)\) has a degenerate distribution which is a point mass at \(t=\Delta\).
-Thus, we wan to choose \(PH(\bs\alpha, \bs{S})\) such that it approximates the \(\delta\) function as closely as possible.
-This suggests we choose \(PH(\bs\alpha, \bs{S})\) as an Erlang distribution with mean \(\Delta\) as the Erlang distribution has the lowest coefficient of variation of all PH distributions.
-%
-%Let \(\calS_+=\left\{i\in\calS\mid c_i>0\right\}\), \(\calS_-=\left\{i\in\calS\mid c_i<0\right\}\).
-Rather than have exactly the same representation \(PH(\bs\alpha, \bs{S})\) for each \(i\in\calS\), consider instead letting \(PH(\bs\alpha, \bs{S})\) for each \(i\in\calS_+\) and \(PH(\wt{\bs\alpha},\wt{\bs{S}})\) for each \(i \in \calS_-\), where \(PH(\wt{\bs\alpha},\wt{\bs S})\) is the time reverse representation.
-Assume without loss that the phases are ordered such that \(i<j\) for all \(i\in\calS_+,\,j\in\calS_-\).
-Construct an approximation to the distribution of \(X(t)\in\calD_k\) given \(X(0)=y_k,\,\varphi(0)=i,\, c_i>0\) as \(PH(\bs e_i \otimes \bs \alpha, \bs{Q})\) where
-%\[\bs{Q} = \left[\begin{array}{ccccc}
-%	T_{11}\bs{I}+|c_1|\bs{S} & T_{12}\bs{I} & \hdots & T_{1,N-1}\bs{I} & T_{1,N}\bs{I} \\
-%	T_{21}\bs{I} & T_{22}\bs{I}+|c_2|\bs{S} & \hdots & T_{2,N-1}\bs{I} & T_{2,N}\bs{I} \\
-%	 \vdots & \vdots &  & \vdots & \vdots \\ 
-%	 T_{N-1,1}\bs{I} & T_{N-1,2}\bs{I} & \hdots & T_{N-1,N-1}\bs{I}+|c_{N-1}|\wt{\bs S} & T_{N-1,N}\bs{I} \\ 
-%	 T_{N,1}\bs{I} & T_{N,2}\bs{I}  & \hdots & T_{N,N-1}\bs{I} & T_{NN}\bs{I}+|c_N|\wt{\bs S} \\ 
-%\end{array}\right].\]
-%Thus, when in phases \(\varphi=i\) with \(c_i>0\), the Markov chain associated with \(\wt{\bs S}\) evolves \emph{forward} in time.
-When in phases \(\varphi=i\) with \(c_i<0\), the Markov chain associated with \(\wt{ \bs S}\) evolves \emph{backward} in time.
-When the process jumps from phase \(\varphi = i\) to phase \(\varphi = j\), the state of the Markov chain associated with \(\bs{S}\) (or \(\wt{\bs S}\)) is preserved.
-In this way the approximation can capture the \emph{memory} of the process through the state of \(\bs S/\wt{\bs S}\).
-%
-%Recall, that we claimed that we should choose \(PH(\bs\alpha, \bs{S})\) as an Erlang distribution in order to achieve the best approximation.
-So suppose we choose an \(Erlang(n,1/(n\Delta))\) representation.
-We can show that this is actually exactly the same approximation as if we had constructed the approximation via the uniformisation method with uniformisation parameter \(1/(n\Delta)\).
-%
-%However, rather than use a \(PH(\bs\alpha, \bs{S})\) distribution to approximate the \(\delta\) function, what if we instead used a matrix exponential distribution.
-That is, choose \(ME(\bs\alpha, \bs{S})\) as a \emph{concentrated matrix exponential distribution}.
-If we could find a time-reverse version of \(ME(\bs\alpha, \bs{S})\), then we might be able to use the same structure as above to construct a high-order approximation to a fluid model with a stochastic interpretation (and therefore preserve positivity of the approximate solution).
-%
-%\section{Time reverse ME/RAPs}
-%\paragraph{Time-reverse PH renewal processes}
-%Let \(\left\{X(t)\right\}_{t\in(-\infty,\infty)}\) be a CTMC with generator \(\bs{Q}=\left[q_{ij}\right]_{i,j\in\calS}\) and state space \(\calS\) and assume that \(\left\{X(t)\right\}\) is stationary with stationary distribution \(\bs\pi\).
-Let \(\wt X(t)=X(-t)\) be the time reverse of \(\left\{X(t)\right\}\).
-Recall that the generator of \(\wt X(t)\) is given by \(\wt{\bs Q}\), with entries \(\wt q_{ij} = \cfrac{\pi_j q_{ji}}{\pi_i}\).
-The processes \(X(t)\) and \(\wt X(t)\) share the same stationary distribution.
-Let \(P=\diag(\pi)\).
-%
-%Consider now a PH renewal process defined by \(PH(\bs\alpha, \bs{S})\).
-Ignoring renewals jumps, the marginal phase process evolves according to the generator \(\bs{Q} = \bs{S}+\bs s\bs \alpha \) where \(\bs s = -\bs{S}\bs 1\).
-%
-%Let \(\wt{ \bs S}=P^{-1}{\bs S}'P\), and \(\wt{\bs\alpha} = \bs s'P\mu\) where \(M'\) denotes the transpose of \(M\) and \(\mu = -\bs\alpha \bs{S}^{-1}\bs 1\).
-Then \(\wt{\bs Q} = \wt{\bs S}+\wt{ \bs s}\wt{\bs \alpha}\) where \(\wt{\bs s}=-\bs{S}\bs 1\), is the time-reverse representation of \(PH(\bs\alpha, \bs{S})\).
-That is, \(PH(\bs\alpha, \bs{S})\) and \(PH(\wt{\bs\alpha}, \wt{\bs S})\) are representations of the same distribution and the PH renewal process can be defined in terms of either representation.
-%
-%\paragraph{Time-reverse (?) ME renewal processes} 
-%Consider now a ME renewal process defined by \(ME(\bs\alpha, \bs{C})\) which we assume is minimal and of order \(p\).
-Can we come up with some time-reverse representation of this renewal process.
-To this end, we now detail some properties of ME renewal processes.
-%
-%Let \(\bs{D} = \bs c\bs \alpha\) where \(\bs c = -\bs{C}\bs 1\).
-Asmussen and Bladt \citep{ab1999} gave such a process a stochastic interpretation (in fact, Asmussen and Bladt worked with rational arrival processes, but I think we need only consider ME renewal processes).
-Let \(V = span\left\{\alpha, \bs\alpha \bs{S},...,\bs\alpha \bs{S}^{p-1}\right\}\).
-There exists a vector-valued stochastic process \(\left\{\bs A(t)\right\}_{t}\) with \(\bs A(t)\in V\) for all \(t\) and \(\bs A(0)=\bs\alpha\).
-Suppose \(\{Y_k\}\), \(k=0,1,...\) are the renewal times, and \(Y_0=0\).
-Between renewal epochs \(\bs A(t)\) moves deterministically according to \(\bs A(t) = \cfrac{\bs A(Y_{N(t)}) e^{\bs{C}(t-Y_{N(t)})}}{\bs A(Y_{N(t)}) e^{\bs{C}(t-Y_{N(t)})}\bs 1}\).
-\(\bs A(t)\) jumps at rate \(\bs A(t)\bs{D}\bs 1\) at time \(t\), and upon a jump \(\bs A(Y_k) = \cfrac{\bs A(Y_k-) \bs{D}}{\bs A(Y_k-) \bs{D}\bs 1}=\bs \alpha\) where \(Y_k-\) is the left limit at \(Y_k\).
-Furthermore, \(\left\{\bs A(t)\right\}\) is a strong Markov process and \(E\left[\bs A(t+s) \mid \bs A(t)\right] = \bs A(t)e^{\bs{Q}t}\) where \(\bs{Q} = \bs{C}+\bs{D}\).
-We know \(\bs{Q}\bs 1=\bs 0\) and \(dev(\bs{Q})=0\), that is, the dominant eigenvalue of \(\bs{Q}\) is 0.
-%
-%Let \(\bs \pi\) be the left eigenvector of \(\bs{Q}\) corresponding to eigenvalue 0.
-Assume that \(\pi_i\neq0\), \(i=1,...,p\).
-If this is not the case then construct an invertible matrix \(\bs{V}\) such that \(\bs{V}\bs 1=\bs 1\), then consider the equivalent representation of the ME renewal process with \(ME(\bs\alpha \bs{V}^{-1}, \bs{V}\bs{C}\bs{V}^{-1})\), and pray that \(\pi_i\neq0\), \(i=1,...,p\) (can we show that there is always such a \(\bs{V}\)?) By choosing \(\bs{V}\) such that \(\bs{V}\bs 1=\bs 1\), then we maintain that the right eigenvector corresponding to \(\bs{V}\bs{C}\bs{V}^{-1}\) is still \(\bs 1\).
-%
-%Let \(\wt{\bs C}=\bs{P}^{-1}\bs{C}'\bs{P}\), \(\wt{ \bs D} = \bs{P}^{-1}\bs{D}'\bs{P}\), and \(\wt{\bs\alpha} = \bs c'\bs{P}\mu\) where \(\mu = -\bs\alpha \bs{C}^{-1}\bs 1\).
-Is the ME renewal process defined by \(ME(\wt{\bs\alpha},\wt{ \bs C})\) in any meaningful way, the reverse of the ME renewal process defined by \(ME(\bs\alpha, \bs{C})\)?
-%
-%Let \(\tau_1,...,\tau_K\) be inter-arrival times for the renewal process defined by \(ME(\bs \alpha, \bs{C})\). Then joint density is of these arrivals is 
-%\[\bs \alpha e^{\tau_1\bs{C}}\bs{D}e^{\tau_2\bs{C}}\bs{D}...e^{\tau_K\bs{C}}\bs c.\]
-%Now let \(\wt \tau_1,...,\wt \tau_K\) be inter-arrival times for the renewal process defined by \(ME(\wt{\bs \alpha},\wt{\bs C})\).
-%Then joint density is
-%\begin{align*}
-%	\wt{\bs \alpha} e^{\wt \tau_1\wt{\bs C}}\wt{ \bs D}e^{\wt \tau_2\wt{ \bs C} }\wt{ \bs D}...e^{\wt \tau_K\wt{ \bs C}}\wt{\bs c} 
-%	&= \bs c'\bs{P}\mu e^{ \wt \tau_1 \bs{P}^{-1}\bs{C}'\bs{P}} \bs{P}^{-1}\bs{D}'\bs{P}e^{ \wt \tau_2 \bs{P}^{-1}\bs{C}'\bs{P}} \bs{P}^{-1}\bs{D}'\bs{P}...e^{ \wt \tau_K \bs{P}^{-1}\bs{C}'\bs{P}}{\bs{P}^{-1}\bs{C}'\bs{P}\bs 1}
-%	%
-%	\\&= \bs c'\bs{P}\mu \bs{P}^{-1} e^{ \wt \tau_1 \bs{C}'} \bs{P} \bs{P}^{-1}\bs{D}'\bs{P}\bs{P}^{-1}e^{ \wt \tau_2 \bs{C}'} \bs{P}\bs{P}^{-1}\bs{D}'\bs{P}...\bs{P}^{-1}e^{ \wt \tau_K \bs{C}'}\bs{P}{\bs{P}^{-1}\bs{C}'\bs{P}\bs 1}
-%	%
-%	\\&= \mu \bs c' e^{ \wt \tau_1 \bs{C}'} \bs{D}'e^{ \wt \tau_2 \bs{C}'} \bs{D}'...e^{ \wt \tau_K \bs{C}'}{\bs{C}'\bs \pi'}.
-%\end{align*}
-%Now use the fact that \(\bs \pi = -\bs \alpha \bs{C}^{-1} / \mu\) (shown in \citep{ab1999}) to get that the joint density is equal to 
-%\[\bs c' e^{ \wt \tau_1 \bs{C}'} \bs{D}'e^{ \wt \tau_2 \bs{C}'} \bs{D}'...e^{ \wt \tau_K \bs{C}'}{\bs \alpha'} = \bs \alpha e^{\wt \tau_K\bs{C}}\bs{D}e^{\wt \tau_{K-1}\bs{C}}\bs{D}...e^{\wt \tau_1\bs{C}}\bs c,\]
-%since transposing a scalar does noting. Therefore the joint density of \(\tau_1,...,\tau_K\) is the same as that of \(\wt \tau_K,...,\wt \tau_1\).
-%
-%\paragraph{Time-reverse using Baye's rule} 
-%For Markov processes we find the time reverse probabilities using Bayes' rule, 
-%\begin{align*}
-%	&\wt p_{ij}(t) = P(\wt X(t)=j\mid \wt X(0)=i) = P( X(0)=j\mid X(t)=i) 
-%	= \cfrac{P( X(t)=i\mid X(0)=j)P(X(0)=j)}{P(X(t)=i)} 
-%	\\& = \cfrac{P( X(t)=i\mid X(0)=j)\pi_j}{\pi_i} = \cfrac{p_{ji}(t)\pi_j}{\pi_i}
-%\end{align*}
-%
-%Do we want to do the same for \(\bs A(t)\)? 
-%\begin{align*}
-%	&P(\wt{\bs A}(t)=\bs j\mid \wt{\bs A}(0)=\bs i) = P( {\bs A}(0)=\bs j\mid {\bs A}(t)=\bs i) 
-%	= \cfrac{P( {\bs A}(t)=\bs i\mid {\bs A}(0)=\bs j)P({\bs A}(0)=\bs j)}{P({\bs A}(t)=\bs i)}.
-%\end{align*}
-%But now the state space is not discrete, how do we do this for all \(\bs i,\,\bs j\)?
-I suspect we just need to do it for a linearly independent selection of \(\bs i\), \(\bs j\).
-What even are \(P( {\bs A}(t)=\bs i\mid {\bs A}(0)=\bs j)\), \(P({\bs A}(0)=\bs j)\) and \(P({\bs A}(t)=\bs i)\).
-Is \(P({\bs A}(t)=\bs i)\) time-dependent.
-%
-%\paragraph{An adjoint process?}
-%Asmussen and Bladt \citep{ab1999} showed that 
-%\[\mathbb P (\theta_tN\in\cdot \mid \mathcal F_{\left[0,t\right]}) = \bs A(t)\bs v(\cdot),\]
-%where \(\bs v\) is a columns vector of basis measures, and \(\theta_t\) is the shift operator, \(\theta_{t}N\left[s,s+u\right)=N\left[t+s,t+s+u\right)\).
-%Given the result
-%\begin{align*}
-%	\wt{\bs \alpha} e^{\wt \tau_1\wt{\bs C}}\wt{ \bs D}e^{\wt \tau_2\wt{\bs C}}\wt{ \bs D}...e^{\wt \tau_K\wt{\bs C}}\wt{\bs c} 
-%	&= 
-%	\bs c' e^{ \wt \tau_1 \bs{C}'} \bs{D}'e^{ \wt \tau_2 \bs{C}'} \bs{D}'...e^{ \wt \tau_K \bs{C}'}{\bs \alpha'}
-%\end{align*}
-%is there some sort of \emph{adjoint} process \(\wt{\bs A}(t)\), a column vector, such that 
-%\[\mathbb P(\theta_{-t} N \in \cdot \mid \mathcal F_{\left[t,\infty\right)}) = {\wt{\bs v}(\cdot)} \wt{\bs A}(t),\]
-%where \(\theta_t\) is the shift operator \(\theta_{-t}N\left[t+s,t+s+u\right)=N\left[s,s+u\right)\), and \(\wt{\bs v}(\cdot)\) is a row-vector of basis measures.
-If so, do we have \(\mathbb E\left[\wt{\bs A}(t-s)|\mathcal F_{\left[t,\infty\right)}\right] = e^{\wt{\bs Q}s}\wt{\bs A}(t),\) and \(\wt{\bs A}(t) = \cfrac{e^{\wt{\bs Q}s}\wt{\bs A}(t)}{\bs e e^{\wt{ \bs Q}s}\wt{\bs A}(t)}\) on \(N\left[0,t\right]=0\)?
-%
-%\section{Phase-Type and ME renewal theory}
-%Let \(\left\{N(t)\right\}\) be a stationary point process with \(ME(\bs \alpha, \bs{Q})\) inter-arrival times.
-Define \(\bs\pi=\bs\alpha (-\bs{Q}^{-1})/\mu\).
-Let \(C_t\) be the \emph{Age} of the process at time \(t\); that is \(C_t = t-Y_{N{t}}\) where \(Y_{N(t)}\) is the last event before time \(t\).
-Let \(R_t\) be the \emph{Residual} time; the time from \(t\) until we see the next arrival; \(R_t = Y_{N(t)+1}-t\).
-The spread \(S_t\) is time between arrivals if we view the process at time \(t\).
-\(S_t=C_t+R_t\).
-It is well-known that for any stationary renewal process \(C_t\) and \(R_t\) have the same distribution and that \(S_t\) has density \(\propto x\psi(x)\) where \(\psi(x)\) is the inter-arrival density.
-%
-%For an ME renewal process we know that \(C_t\sim ME(\bs\pi, \bs{Q})\) and \(R_t\sim ME(\bs\pi, \bs{Q})\) \citep{ab1999}.
-Bladt and Nielsen use algebraic arguments to show that \(S_t\) has a matrix exponential distribution with representation
-%\[\left(\left[ \cfrac{\bs\alpha \bs{Q}^{-1}}{\bs\alpha \bs{Q}^{-1} \bs e}\,, \bs 0\right], \left[\begin{array}{cc}\bs{Q}&-\bs{Q}\\0&\bs{Q}\end{array}\right]\right).\]
-%
-%Furthermore, for when the ME renewal process is a PH renewal process, they use the stochastic interpretation of the PH distribution as time until absorption of a finite-state Markov chain to argue that \(S_t\) also is also a PH distribution.
-%
-%We can give the ME representation of the spread a stochastic interpretation.
-Let us first consider the age, \(A\sim ME(\bs\pi,\bs{Q})\).
-Corresponding to the age is an orbit process; \(\bs A(s) = \bs\pi e^{\bs{Q}s}/(\bs\pi e^{\bs{Q}s}\bs e)\), which is entirely deterministic on \(s\in\left[0,A\right]\).
-At time \(Y_{N(t)}\) the age orbit is initialised.
-If we observe the process \(s\) amount of time later, then the age is \(A=Y_{N(t)}-t=s\) and the corresponding position of the orbit is \(\bs A(t) = \bs\pi e^{\bs{Q}s}\).
-On \(A=s\), the residual time has distribution \(ME\left(\cfrac{\bs\alpha e^{\bs{Q}s}}{\bs\alpha e^{\bs{Q}s}\bs e},\bs{Q}\right)\).
-That is, \(R_t\mid A=s \sim ME\left(\cfrac{\bs\alpha e^{\bs{Q}s}}{\bs\alpha e^{\bs{Q}s}\bs e},\bs{Q}\right)\).
-However, \(\bs A(s) \propto \bs\pi e^{\bs{Q}s} = \bs\alpha (-\bs{Q}^{-1}) e^{\bs{Q}s} = \bs\alpha e^{\bs{Q}s}(-\bs{Q}^{-1})\) and therefore \(\cfrac{\bs A(s)(-\bs{Q})}{\bs A(s)(-\bs{Q})\bs e} = \cfrac{\bs\alpha e^{\bs{Q}s}}{\bs\alpha e^{\bs{Q}s}\bs e}\).
-Thus we may think of the spread as the evolution of an orbit process.
-At time \(Y_{N(t)}\) initialise the age orbit process \(\bs A(0) = \bs \pi\) and evolve according to \(\bs A(s) =  \bs\pi e^{\bs{Q}s}/(\bs\pi e^{\bs{Q}s}\bs e)\).
-The age process expires with intensity \(\bs A(u) (-\bs{Q})\bs e\).
-Upon expiry we realise the age \(A=s\) and the orbit process jumps from \(\bs A(s-)\) to \(\cfrac{\bs A(s)(-\bs{Q})}{\bs A(s)(-\bs{Q})\bs e}\).
-The orbit process now continues to evolve along \(\bs A(u) = \cfrac{\bs A(s)(-\bs{Q})e^{\bs{Q}{(u-s)}}}{\bs A(s)(-\bs{Q})e^{\bs{Q}{(u-s)}}\bs e}\) until the residual time expires which occurs with intensity \(\bs A(s)(-\bs{Q})\bs e\).
-%
-%An alternative representation of the spread is found by applying the similarity transform \(\bs{D} = \left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]\), where \(\Delta = \diag((-\bs{Q}^{-1})\bs e)\).
-The new representation is
-%\[ME(\bs \beta, U)\] 
-%where 
-%\[\bs \beta = \left[\begin{array}{cc}\bs \pi & 0\end{array}\right]\left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]=\left[\begin{array}{cc}\cfrac{\bs \alpha(-\bs{Q}^{-1})}{\bs \alpha(-\bs{Q}^{-1})\bs e} & 0\end{array}\right]\left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]=\left[\begin{array}{cc}\cfrac{\bs \alpha\Delta}{\bs \alpha(-\bs{Q}^{-1})\bs e} & 0\end{array}\right], \]
-%and 
-%\[U = \bs{D}^{-1}\bs{Q}\bs{D}=\left[\begin{array}{cc}-\Delta^{-1}\bs{Q}^{-1} & \\ & \bs{I}\end{array}\right] \left[\begin{array}{cc}\bs{Q}&-\bs{Q}\\0&\bs{Q}\end{array}\right] \left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]%=\left[\begin{array}{cc}\Delta^{-1}&\Delta^{-1}\\0&\bs{Q}\end{array}\right] \left[\begin{array}{cc}-\bs{Q}\Delta & \\ & \bs{I}\end{array}\right]
-%=\left[\begin{array}{cc}\Delta^{-1}\bs{Q}\Delta&\Delta^{-1}\\0&\bs{Q}\end{array}\right] \]
-%and the closing vector is unchanged \(\bs u = -U\bs e =  \left[\begin{array}{c} 0 \\ -\bs{Q}\bs e \end{array}\right]\). 
-%
-%Under this representation, the orbit process is initialised according to \(\bs A(0)=\cfrac{\bs \alpha\Delta}{\bs \alpha(-\bs{Q}^{-1})\bs e} \), evolves along \(\bs A(s)=\cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}}{{\bs \alpha\Delta}e^{\bs{Q}s}\bs e}\) and expires with intensity \(\bs A(s)\Delta^{-1}=\cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}}{{\bs \alpha\Delta}e^{\bs{Q}s}\bs e}\Delta^{-1}\bs e\).
-At expiry the age is realised, \(C_t=s\), and the orbit process jumps from \(\bs A(s-) = \cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}}{{\bs \alpha\Delta}e^{\bs{Q}s}\bs e}\) to \(\bs A(s) = \cfrac{{\bs \alpha\Delta} e^{\bs{Q}s}\Delta^{-1}}{{\bs \alpha\Delta}e^{\bs{Q}s}\Delta^{-1}\bs e}\).
-%
-%When the ME inter-arrival times are also Phase-type, then this representation for the spread is also Phase-type.
-
-%\section{SFM approximation}
-%%\paragraph{SFMs} Let \(\left\{\varphi(t)\right\}_{t\geq 0}\) be an irreducible continuous-time Markov Chain (CTMC) with a finite state space \(\mathcal S = \left\{1,...,N\right\}\) and infinitesimal generator \(\bs{T} = \left[T_{ij}\right]_{i,j\in\mathcal S}\).
-Let \(\left\{(X(t),\varphi(t))\right\}_{t\geq0}\) be a Markovian stochastic fluid model (SFM) with phase variable \(\varphi(t)\) and level variable \(X(t)\in (-\infty,\infty)\).
-For each \(i\in\mathcal S\) there are rates \(c_i\in\mathbb R\) such that \(\cfrac{\wrt X(t)}{\wrt t} = c_i\) when \(\varphi(t)=i\).
-The variable \(\left\{X(t)\right\}\) is known as the buffer, or level, of the fluid process.
-%
-%Partition the set of all phases, \(\mathcal S\), into \(\mathcal S_+ = \left\{i\in\mathcal S\mid c_i>0\right\},\, \mathcal S_- = \left\{i\in\mathcal S\mid c_i<0\right\}, \,\mathcal S_0 = \left\{i\in\mathcal S\mid c_i=0\right\},\).
-Use the same partition to partition the generator
-%\[\bs{T} = \left[\begin{array}{ccc} \bs{T}_{++} & \bs{T}_{+-} & \bs{T}_{+0} \\ \bs{T}_{-+} & \bs{T}_{--} & \bs{T}_{-0} \\ \bs{T}_{0+} & \bs{T}_{0-} & \bs{T}_{00} \end{array}\right].\]
-%Let \(\bs{C}_+ = \diag(c_i,i\in\mathcal S_+)\), and \(\bs{C}_- = \diag(|c_i|,i\in\mathcal S_-)\).
-
-%Let \(Z\sim ME_p(\bs \alpha, \bs{S}, \bs s)\) of dimension \(p\) with \(\bs\alpha\bs e = 1\), \(\bs s = -\bs{S}\bs e\) and mean \(\Delta\).
-Let \(\{\bs A(t)\} = \left\{\bs A(t)\right\}_{t\geq 0}\) be a \(N\times p\)-dimensional vector-valued stochastic process which is a piecewise-deterministic Markov process.
-The sample path behaviour of \(\{\bs A(t)\}\) is as follows.
-For some \(i\in \mathcal \bs{S}\), let \(\bs a_i\) be a vector such that \(\bs a_i\bs e = 1\) and \((\bs a_i,\bs{S},\bs s)\) is a representation of a ME distribution.
-Let
-%\[Q = \left[\begin{array}{ccc}  \bs{C}_+ \otimes \bs{S} &&\\ &   \bs{C}_- \otimes \bs{S} & \\ &&0\end{array}\right],\]
-%and \(\bs q = -Q\bs e\). 
-%Given \(\bs A_0 = \bs a = (\bs 0_{(i-1)p},\bs a_i,\bs 0_{(N-i-1)p})\), \(\{\bs A(t)\}\) evolves deterministically according to
-%\[\bs A(t) = \cfrac{\bs a e^{Q t}}{\bs a e^{Q t} \bs e}\]
-
-%Let \(Z\sim ME_p(\bs \alpha, \bs{S}, \bs s)\) of dimension \(p\) with \(\bs\alpha\bs e = 1\), \(\bs s = -\bs{S}\bs e\).
-For each \(i\in\mathcal S\), define \(\bs{S}_i = |c_i|\bs{S}\) and \(\bs s_i = |c_i|\bs s\).
-Let \(\{\bs A(t)\} = \left\{\bs A(t)\right\}_{t\geq 0}\) be a \(p\)-dimensional vector-valued stochastic process which is piecewise-deterministic and c\`adl\`ag.
-Additionally, let \(\left\{{\phi}(t)\right\}_{t\geq 0}\) be a real-valued stochastic process.
-Together, \(\left\{(\bs A(t),{\phi}(t))\right\}_{t\geq 0}\), form a Markov process, however, \(\{\bs A(t)\}\) alone is not Markovian.
-We will show later that \(\left\{{\phi}(t)\right\}\) is a CTMC the same as \(\left\{\varphi(t)\right\}\).
-The sample path behaviour of \(\{\bs A(t)\}\) and \({\phi}(t)\) is as follows.
-\section{The dynamics of the QBD-RAP approximation}\label{sec: dynamics}
-We now have all the elements we need to describe the dynamics of the QBD-RAP approximation.
-Since the phase dynamics are a CTMC, we choose to use an alternate notation to the standard presented in Section \ref{sec: qbd-rap}.
-Let \(\{(L(t),\bs A(t),\phi(t))\}_{t\geq0}\) be the QBD-RAP approximation of a fluid queue, where \(\{L(t)\}\) is the level, \(\{\bs A(t)\}\) is the orbit process and \(\{\phi(t)\}\) is the phase process.
-We use this representation to make explicit how the approximation captures the phase dynamics of the fluid queue.
-We will show later that \(\phi(t)\) captures the phase dynamics of the fluid queue exactly, provided the phase process is independent of the fluid level \(\{X(t)\}\).
-The level \(L(t)=\ell\) approximates which band \(\calD_\ell\) that \(X(t)\) is in at time \(t\), and the orbit \(\bs A(t)\) can be use to obtain an approximation of where \(X(t)\) is within the interval \(\calD_\ell\).
-
-We now proceed to describe the evolution of the orbit and phase processes, before introducing the level variable later.
-
-\paragraph{The orbit process and phase dynamics} On \({\phi}(t)=i\), between event epochs, the process \(\{\bs A(t)\}\) evolves deterministically according to the differential equation 
-\begin{align}\label{eqn: At DE}
-\cfrac{\wrt}{\wrt t}\bs A(t) = \bs A(t)(\bs{S}_i - \bs A(t) (\bs{S}_i + T_{ii}\bs{I}_p)\bs e)\bs{I}_p.
-\end{align}
-Let \(\bs a \in \mathcal A\) be an arbitrary vector in \(\mathcal A\).
-On the event that no events occur before time \(t+u\), \(\bs A(u) = \bs a\in\mathcal A\) and \({\phi}(u)=i\), the solution to (\ref{eqn: At DE}) states that \(\bs A({t+u})\) evolves deterministically according to
-\[\bs A({t+u}) =  \cfrac{\bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}}{\bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t} \bs e} = \cfrac{\bs a e^{\bs{S}_i t}}{\bs a e^{\bs{S}_i t} \bs e}.\]
-
-At time \(t\) an event occurs at rate 
-\[\bs A(t) (\bs{S}_i - T_{ii}\bs{I}_p)\bs e = \bs A(t)\bs s_i - T_{ii}.\]
-More precisely, an event corresponding to a change in phase for \({\phi}(t)\) occurs at rate \(-\bs A(t) T_{ii}\bs e = -T_{ii}\) and an event corresponding to a \emph{change of level} occurs at rate \(-\bs A(t) \bs{S}_i\bs e = \bs A(t) \bs s_i\).
-Later, we will make clear why we say that the latter event corresponds to a change of level.
-Upon an event occurring at time \(t\), with probability \({-T_{ii}}/{\left(-T_{ii} + \bs A({t^-})\bs s_i\right)}\) the event corresponds to a change of phase and with probability \({\bs A({t^-})\bs s_i}/{\left(-T_{ii} + \bs A({t^-})\bs s_i\right)}\) the event corresponds a change of level.
-
-Upon an event corresponding to a change of level occurring at time \(t\) the process \(\{\bs A(t)\}\) jumps to
-\(\bs A(t) = \bs \alpha\).
-
-Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), there are two possibilities; either \(sign(c_i)=sign(c_j)\), or \(sign(c_i)\neq sign(c_j)\).
-As discussed earlier, for states \(i\in\mathcal S_0\) we must specify some association with either \(\mathcal S_+\) or \(\mathcal S_-\).
-Here we choose the augmented state space approach and duplicate \(\calS_0\) and associate one copy with \(i\in\calS_+\) and one copy with \(\calS_-\); call these \(\calS_{+0}\) and \(\calS_{-0}\), respectively.
-We take \(sign(c_i)=+\) for \(i\in\calS_{+0}\) and \(sign(c_i)=-\) for \(i\in\calS_{-0}\).
-
-Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)=sign(c_j)\), at the time of the event \(\bs A(t)\) is unchanged but immediately begins to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), while \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
-
-Upon an event corresponding to a change of phase from \(i\) to \(j\neq i\) occurring at time \(t\), in the case \(sign(c_i)\neq sign(c_j)\), the process \(\{\bs A(t)\}\) jumps to 
-\[\bs A(t) = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{\bs A({t^-}) T_{ij} \bs{D} \bs e} = \cfrac{\bs A({t^-}) T_{ij}\bs{D}}{T_{ij} } = \bs A({t^-}) \bs{D} \] 
-and then immediately proceeds to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), and \(\left\{{\phi}(t)\right\}\) jumps from \({\phi}(t^-)=i\) to \({\phi}(t)=j\).
-
-The two scenarios, \(sign(c_i)\neq sign(c_j)\) and \(sign(c_i)= sign(c_j)\), can be written succinctly by stating that, at the time of an event corresponding to a change of phase from \(i\) to \(j\), \(\{\bs A(t)\}\) jumps to \(\bs A(t)\bs{D}^{1(sign(c_i)\neq sign(c_j))}\) and begins to evolve according to (\ref{eqn: At DE}) with \(i\) replaced by \(j\), meanwhile \(\left\{{\phi}(t)\right\}\) jumps from \(\phi(t^-)=i\) to \(\phi(t)=j\).
-% Moreover, given a change of phase event occurs at time \(t\) with \(\phi(t^-)=i\) then, with probability \(\cfrac{\bs A(t^-)T_{ij}\bs{D}^{1(sign(c_i)\neq sign(c_j))}\bs e}{-\bs A(t^{-})T_{ii}\bs e} = \cfrac{T_{ij}}{-T_{ii}}\), the event corresponds to a change of phase from \(i\) to \(j\).
-
-The following result states that \(\left\{{\phi}(t)\right\}\) has the same distribution as \(\left\{{\varphi}(t)\right\}\) when the fluid queue is unbounded, the latter being the phase process of the fluid queue.
-\begin{thm}\label{thm: 1}
-	Let \(\Theta_i\) be the time at which the first jump of the phase process of the unbounded QBD-RAP, \(\left\{{\phi}(t)\right\}\), occurs given \(\phi(0)=i\).
-For any initial orbit \(\bs a\in\mathcal A\), then \(\Theta_i\) has an exponential distribution with rate parameter \(|T_{ii}|\).
-Furthermore, given \({\phi}(t)\) leaves state \(i\), it jumps to state \(j\) with probability \(T_{ij}/|T_{ii}|\).
-Hence \({\phi}(t)\) and \(\varphi(t)\) have the same probability law.
-\end{thm}
-\begin{proof}
-	Let \(\{\tau_n\}_{n\geq 0}\) with \(\tau_0=0\) and \(\tau_n\) the time of the \(n\)-th change of level of the QBD-RAP.
-Consider partitioning \(\left\{\Theta_i> t \right\}\) with respect to \(\left\{\tau_{n-1}< t \leq\tau_n\right\},\,n=1,2,\dots\).
-For \(n=1\) we can write
-	\begin{align*}
-		&\mathbb P(\Theta_i>  t, \tau_0< t\leq \tau_1 \mid \bs A(0) = \bs a) 
-		= \bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t}\bs e 
-		\intertext{and since \(T_{ii}\bs{I}_p\) commutes with \(\bs{S}_i\) and \(e^{T_{ii}t}\) is a scalar, then this is equal to}
-		&\bs a e^{\bs{S}_it}e^{T_{ii}\bs{I}_pt}\bs e  
-		= \bs a e^{\bs{S}_it}\bs e e^{T_{ii}t} 
-		= \mathbb P(\tau_0 < t\leq \tau_1)e^{T_{ii}t} .
-	\end{align*}
-	For \(n>1\), by partitioning on the times of the first \(n-1\) level changes, \(\tau_1,\dots,\tau_n\), we get 
-	\begin{align*}
-		&\mathbb P(\Theta_i> t, \tau_{n-1}< t \leq \tau_n \mid \bs A(0) = \bs a) 
-		\\&= \int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t \mathbb P(\Theta_i> t, t \leq \tau_n, \tau_{n-1}\in \wrt t_{n-1},\dots \tau_2\in\wrt t_2, \tau_1\in\wrt t_1 \mid \bs A(0) = \bs a)
-		\\&=\displaystyle{\int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t} \bs a e^{(\bs{S}_i+T_{ii}\bs{I}_p)t_1}\bs s_i
-			\left(\prod_{k=2}^{n-1} \bs \alpha e^{(\bs{S}_i+T_{ii}\bs{I}_p)(t_k-t_{k-1})} \bs s_i\right) 
-			\\&\qquad\displaystyle{\times\bs \alpha e^{(\bs{S}_i+T_{ii}\bs{I}_p)(t-t_{n-1})}\bs e}\wrt t_{n-1}\wrt t_{n-2}\hdots\wrt t_{1}.
-		\intertext{Since \(T_{ii}\bs{I}_p\) commutes with \(\bs{S}_i\), \(e^{T_{ii}t_k},\,k=1,...,n-1\) are scalars, and \(t_1+(t_2-t_1)+\hdots+(t_{n-1}-t_{n-2})+(t-t_{n-1})=t\), then this is equal to}
-		& \displaystyle{\int_{t_1=0}^t\int_{t_2=t_1}^t\hdots\int_{t_{n-1}=t_{n-2}}^t} \bs a e^{\bs{S}_it_1}\bs s_i 
-			\left(\prod_{k=2}^{n-1} \bs \alpha e^{\bs{S}_i(t_k-t_{k-1})} \bs s_i\right) \bs \alpha e^{\bs{S}_i(t-t_{n-1})}\bs e
-			\times e^{T_{ii}t} \wrt t_{n-1}\wrt t_{n-2}\hdots\wrt t_{1}
-		\\&= \mathbb P(\tau_{n-1}< t \leq \tau_n)\displaystyle{  e^{T_{ii}t} }.
-	\end{align*}
-	Hence, by the law of total probability, 
-	\begin{align*}
-		\mathbb P(\Theta_i> t) &= \sum_{n=1}^\infty \mathbb P(\Theta_i> t, \tau_{n-1} < t \leq \tau_n)
-		\\&= \sum_{n=1}^\infty \mathbb P(\tau_{n-1} < t \leq \tau_n)\displaystyle{  e^{T_{ii}t} }
-		\\&= e^{T_{ii}t},
-	\end{align*}
-	and therefore \(\Theta_i\) has an exponential distribution with rate \(|T_{ii}|\).
-
-	Upon leaving state \(i\) at time \(t\), \({\phi}(t)\) transitions to state \(j\) with probability 
-	\begin{align*}
-            	\cfrac{\left(\cfrac{\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e}{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e + \bs A(t) \bs s_i}\right)}
-            	{\left(\cfrac{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e}{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs{D}^{1(sign(c_i)\neq sign(c_j))}T_{ij}\bs e + \bs A(t) \bs s_i}\right)}
-            	&=\cfrac{\bs A(t) \bs eT_{ij}}
-            	{\displaystyle\sum_{j\in\mathcal S}\bs A(t) \bs eT_{ij}}
-	= \cfrac{T_{ij}}
-            	{-T_{ii}}.
-	 \end{align*}
-	 Therefore the process \(\left\{{\phi}(t)\right\}\) is has the same probability law as \(\left\{\varphi(t)\right\}\).
-\end{proof}
-\begin{rem}
-The same result can be shown for a regulated boundary.
-For boundary conditions which interact with the phase dynamics, such as a reflecting boundary, the result does not hold.
-The cause is the fact that the phase dynamics are level dependent -- we may see a forced change of phase upon a boundary being hit -- and the QBD-RAP can only approximate the level process of the fluid queue.
-However, until a boundary is hit (by either the fluid queue or QBD-RAP) then the phase processes match.
-We show later that, in the limit as the variance of the the matrix exponential distribution used in the construction of the QBD-RAP goes to zero, then the dynamics of the level process of the fluid queue, \(X(t)\), are captured by the QBD-RAP, and boundary behaviour which interacts with the phase dynamics can be captured too.
-\end{rem}
-
-Since the phase processes \(\left\{{\phi}(t)\right\}\) and \(\left\{\varphi(t)\right\}\) have the same law when boundaries are not present, henceforth, we shall assume \(\left\{{\phi}(t)\right\}\) and \(\left\{\varphi(t)\right\}\) are coupled when possible (they share the same sample path).
-Specifically, in Section~\ref{sec: no change}, we will analyse the QBD-RAP on the event that it remains in the same level, \(\ell\), say, and we compare this to the fluid queue on the event that the level remains in the band \(\calD_\ell\).
-No boundary behaviour is involved in this calculation, so we treat \(\left\{{\phi}(t)\right\}\) and \(\left\{\varphi(t)\right\}\) as coupled and use the latter notation.
-When we must distinguish the two processes, we use \(\phi(t)\) for the phase of the QBD-RAP and \(\varphi(t)\) for the phase of the fluid.
-
-\paragraph{The level process}
-To event epochs of \(\left\{(\bs A(t),{\varphi}(t))\right\}_{t\geq0}\) we associate marks \(\left\{-1,0,+1\right\}\) in the following way.
-\begin{itemize}
-	\item To events epochs corresponding to a change of phase of \( {\varphi}(t)\) we associate the mark \(0\).
-	\item To event epochs at time \(t\) which correspond to a change in level and for which \( {\varphi}(t^-)=i\in\mathcal S_-\) we associate the mark \(-1\).
-	\item To event epochs at time \(t\) which correspond to a change in level and for which \( {\varphi}(t^-)=i\in\mathcal S_+\) we associate the mark \(+1\).
-\end{itemize}
-Now define \(N_+(t)\) (\(N_-(t)\)) as the simple point process which counts the number of event epochs with marks \(+1\) (\(-1\)) which have occurred in the time up to and including time \(t\).
-The level process of the QBD-RAP is given by \( L(t) = N_+(t)-N_-(t)\).
-The process \(\left\{(L(t),\bs A(t),\varphi(t))\right\}_{t\geq 0}\) forms a QBD with RAP components.
-
-One way to specify the QBD with RAP components, \(\left\{(L(t),\bs A(t),{\varphi}(t))\right\}\), is to describe its generator:
-\begin{align*}
-\bs{B} &= \left[\begin{array}{cccccc}
- \ddots & \ddots & \ddots & & & \\
- & \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & & \\ 
- & &  \bs{B}_{-1} & \bs{B}_0 & \bs{B}_{+1} & \\
- & & & \ddots & \ddots & \ddots 
-\end{array}\right],
-\end{align*}
-{where,}
-\begin{align*}
-	\bs B &= \left[\begin{array}{ccc} \bs B_{++} & \bs B_{+-} \\ \bs B_{-+} & \bs B_{--}\end{array}\right]\\
-\end{align*}
-\begin{align*}
-	\bs{B}_{-1} &= \left[\begin{array}{cccc}
-		\bs 0 & & & \\
-		& \bs 0 & & \\
-		& & \bs{C}_-\otimes (\bs s\bs\alpha) & \\ 
-		& & & \bs 0
-		\end{array}\right],
-	%
-	&& \bs{B}_{+1} = \left[\begin{array}{cccc}
-		\bs{C}_+\otimes (\bs s\bs\alpha) & & & \\
-		& \bs 0 & & \\
-		& & \bs 0 & \\ 
-		& & & \bs 0
-		\end{array}\right],
-\end{align*}
-and
-\begin{align*}
-	\bs{B}_{++}  &= \left[\begin{array}{cc} \bs{C}_+\otimes \bs{S}  + \bs{T}_{++}\otimes \bs{I} & \bs T_{+0}\otimes \bs I \\ \bs T_{0+}\otimes \bs I & \bs T_{00}\otimes \bs I\end{array}\right] ,
-	%
-	\bs{B}_{+-}  = \left[\begin{array}{cc} \bs{T}_{+-}\otimes \bs{D}  & \bs 0 \\ \bs T_{0-}\otimes \bs D  & \bs 0 \end{array}\right],
-	%
-	\\ \bs{B}_{-+}  &=  \left[\begin{array}{cc} \bs{T}_{-+}\otimes \bs{D}  & \bs 0 \\ \bs T_{0+}\otimes \bs D  & \bs 0 \end{array}\right] ,
-	% 
-	\bs{B}_{--}  = \left[\begin{array}{cc} \bs{C}_-\otimes \bs{S}  - \bs{T}_{--}\otimes \bs{I} & \bs T_{-0}\otimes \bs I \\ \bs T_{0-}\otimes \bs I & \bs T_{00}\otimes \bs I\end{array}\right].
-\end{align*}
-
-%The generator acts on infinite dimensional row-vectors 
-%\[\bs v = \left( \begin{array}{ccccc} \dots & \bs v_{-1} & \bs v_0 & \bs v_{1} & \dots \end{array} \right),\]
-%via right-multiplying the vector. The vector \(\bs v\) represents an initial distribution on the QBD-RAP. The vectors \(\bs v_k\), \(k\in\mathbb Z\) are 
-%\[\bs v_k = \left( \begin{array}{ccc} \bs v_{k,1} & \dots & \bs v_{k,N} \end{array} \right),\]
-%where each \(\bs v_{k,i}\) is such that \((\bs v_{k,i}, \bs{S}_i,\bs s_i)\) is a valid representation of an ME.
-The initial distribution of the phase and level is given by the product \(\bs v_{k,i}\bs e = \mathbb P( L(0)=k, \varphi(0)=i)\) and, given \(\left\{ L(0)=k, \varphi(0)=i\right\}\), the initial position of \(\left\{\bs A(t)\right\}\) is \(\cfrac{\bs v_{k,i}}{\bs v_{k,i}\bs e} \).
-%
-%The product 
-%\[\bs v(t) := \bs v e^{\bs{B}t} = \left( \begin{array}{ccccc} \dots & \bs v_{-1}(t) & \bs v_0(t) & \bs v_{1}(t) & \dots \end{array} \right),\]
-%where 
-%\[\bs v_k(t) = \left( \begin{array}{ccc} \bs v_{k,1}(t) & \dots & \bs v_{k,N}(t) \end{array} \right),\quad k\in\mathbb Z,\]
-%gives the distribution at time \(t\): for \(k\in\mathbb Z\),  \(\mathbb P({L}(t) = k, \varphi(t)=i) = \bs v_{k,1}(t)\bs e\) and, given \(\left\{{L}(t) = k, \varphi(t)=i\right\}\), the position of \(\left\{\bs A(t)\right\}\) is \(\bs A(t) =\cfrac{\bs v_{k,i}(t)}{\bs v_{k,i}(t)\bs e}.
-\)
-
-\section{Boundary conditions}\label{sec: boundary conditions}
-In the study and practical application of fluid queues, regulated, reflecting, sticky, or a mixture of these boundary conditions may be imposed.
-We now present the intuition as to how one may include such boundary conditions in the QBD-RAP approximation scheme.
-
-Without loss of generality, assume that there is a boundary for the fluid level at \(y_0 = 0\).
-This boundary can only be hit from above in phases \(i\in\mathcal S_-\).
-%For a regulated boundary, upon hitting the boundary in phase \(i\in\mathcal S_-\) the fluid level remains at \(0\), while the phase continues to evolve among states in the set \(\calS_-\cup\calS_0\).
-Upon the first transition of the phase process from \(\calS_-\cup \calS_+\) to \(\calS_+\) the fluid level leaves the boundary.
-For a reflecting boundary we suppose that, upon hitting the boundary in some phase \(i\in\mathcal S_-\) the phase of the process immediately jumps to \(j\in\mathcal S_+\) with probability \(p_{ij}\).
-Suppose that, upon hitting the boundary in phase \(i\in\calS_-\), the phase jumps from \(i\) to \(j\in\calS\) with probability \(p_{ij}\).
-If \(j\in\calS_-\cup\calS_0\) the process remains at the boundary and the phase process evolves among states in \(\calS_-\cup\calS_0\) until the first transition to a state in \(\calS_+\), at which point the level \(\{X(t)\}\) immediately leaves the boundary -- we will call this a \emph{sticky} boundary.
-If \(j\in\calS_+\) the process immediately leaves the boundary -- we will call this a refecting boundary.
-We collect the probabilities \(p_{ij}\) into the matrices \(\bs P_{-+} = [p_{ij}]_{i\in\calS_-,j\in\calS_+},\) \(\bs P_{-0} = [p_{ij}]_{i\in\calS_-,j\in\calS_0},\) and \(\bs P_{--} = [p_{ij}]_{i\in\calS_-,j\in\calS_-}.\)
-
-Adjacent to the boundary at \(y_0=0\) is the band \(\mathcal D_0=[0,\Delta]\) which corresponds to level \(0\) for the QBD-RAP.
-We denote the boundary of the QBD-RAP as level \(-1\), which corresponds to \(\{0\}\) for the fluid queue.
-Modelling the behaviour of the fluid queue at boundaries can be broken down into three components.
-1) Modelling the time and phase when the fluid level hits the boundary.
-2) Modelling the phase whilst the fluid level remains at the boundary.
-3) Modelling the fluid level and phase at the exit from the boundary.
-
-We claim that the event \(\{L(t)=\ell,\phi(t)=i\}\) models the event \(\{X(t)\in\mathcal D_{\ell,i},\varphi(t)=i\}\), and further that instants \(u\) with \(L(u^-)\neq L(u), \phi(u)=i\) model the events \(\{X(u^-)\in \mathcal D_{\ell,i}, X(u)\notin\mathcal D_{\ell,i},\varphi(u)=i\}\).
-With this in mind, we suppose that when the QBD-RAP is in level \(0\) in phase \(i\in\mathcal S_-\) and there is a change of level, this corresponds to approximating the event that the fluid level \(\{X(t)\}\) hits the boundary at \(0\).
-We refer to this as the QBD-RAP hitting the boundary also.
-We show later that, using matrix exponentials with sufficiently small variance in the construction of the QBD-RAP, then the distribution of time until the QBD-RAP first hits the boundary closely models the time until the fluid level hits the boundary.
-This addresses component 1).
-
-%Assuming the fluid process \(\{X(t)\}\) is continuous then we may use the QBD-RAP to approximate the fluid queue hitting the boundary.
-For example, with a regulated lower boundary at \(x=0\), place \(x=0\) at the boundary of the first of the bands \(\mathcal D_0=[0,\Delta]\) which corresponds to level \(\ell =0\).
-When the QBD-RAP process, \(\{(L(t), \bs A(t), \varphi(t))\}\) exits level \(0\) in a phase \(j\in\mathcal S_-\), then we correspond this with the event that fluid queue is hits level \(x=0\) in phase \(j\).
-Corollary~\ref{cor: aln222} states that, for sufficiently large \(p\), the construction described here models distribution of time until \(\{X(t)\}\) leaves \(\mathcal D_0\) arbitrarily closely, and therefore we claim that the QBD-RAP approximation models hitting the boundary at \(x=0\) arbitrarily closely.
-
-For a sticky lower boundary at \(0\), given the fluid level reaches zero at time \(u\) in phase \(i\in\mathcal S_-\), then the phase transitions to some phase \(j\in\calS_-\cup\calS_0\) with probability \(p_{ij}\), and the distribution of the phase \(\varphi(t)\) (on the event that the process has not left the boundary at or before time \(t\)) is given by the elements of the vector
-\[p_{ij} \bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right).\] 
-The distribution for the time until \(\{X(t)\}\) leaves \(x=0\) for the first time after \(u\) is 
-\[1-p_{ij}\bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right)\bs e.\] 
-Therefore, in the case of a sticky boundary, we can capture the behaviour of the process at the boundary exactly.
-For a reflecting boundary there is no time spent at the boundary.
-This addresses component 2) of the modelling problem.
-
-For both sticky and reflecting boundaries, given the QBD-RAP is in the correct phase at the instant upon which it hits the boundary, then upon exiting the boundary the phase can be captured exactly too.
-For a sticky boundary, given the fluid/QBD-RAP hits the boundary in phase \(i\in\mathcal S_-\) at time \(t\), it then jumps to phase \(j\in\calS_+\cup\calS_0\) and exits the boundary in phase \(k\in\mathcal S_+\) at time \((t-u)\) with density
-\[p_{ij}\bs e_j \exp\left({\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}\end{array}\right]}(t-u)\right)\left[\begin{array}{c} \bs T_{-+} \\ \bs T_{0+}\end{array}\right]\bs e_k.\]
-For a reflecting boundary, given the boundary is hit in phase \(i\in\calS_-\), the phase upon leaving the boundary is \(j\in\calS_+\) with probability \(p_{ij}\).
-Upon leaving the boundary, the appropriate orbit position for the QBD-RAP is \(\bs \alpha\), as this corresponds to the fluid level \(X=0\).
-
-To summarise, at time \(t\) the rate at which probability mass accumulates at the sticky boundary in phase \(j \in\calS_-\cup\calS_0\) and level \(-1\), upon hitting the boundary in phase \(i\in\calS_-\) and level \(0\), is
-\[ c_ip_{ij} \bs A(t)\bs s.\]
-The rate at which density from phase \(i\in\calS_-\) and level \(0\) transitions to density in phase \(j\in\calS_+\) and level \(0\) upon hitting a boundary is 
-\[ c_ip_{ij} \bs A(t)\bs s,\]
-and upon this transition, the orbit jumps to \(\bs \alpha\). 
-The rate at which mass leaves the sticky boundary from phase \(i\in\calS_-\cup\calS_0\) and level \(-1\) into phases \(j\in\calS_+\) and level \(0\) is 
-\[T_{ij},\]
-and upon leaving the boundary the orbit is \(\bs\alpha\).
-
-This information can be inscribed in the generator of the QBD-RAP.
-For example, using the augmented state-space scheme to account for phases \(\calS_0\), the generator of the QBD-RAP described above has boundary conditions included as
-\begin{align*}
-&\left[\begin{array}{cc|cccc|cc}
-	\bs T_{--} & \bs T_{0-} & \bs T_{-+} \otimes\bs \alpha & \bs 0 & \bs 0 & \bs 0 &&\\
-	\bs T_{-0} & \bs T_{00} & \bs T_{0+} \otimes\bs \alpha & \bs 0 & \bs 0 & \bs 0 &&\\\hline
-	\bs 0 & \bs 0 &\multicolumn{4}{c|}{\multirow{4}{*}{$\check{\bs B}_{0}$}} & \multirow{4}{*}{$\bs B_{+1}$}& \\
-	\bs 0 & \bs 0 & \multicolumn{4}{c|}{} & &\\
-	(\bs C_-\bs P_{--})\otimes \bs s &(\bs C_-\bs P_{-0}) \otimes \bs s & \multicolumn{4}{c|}{} &&\\
-	\bs 0 & \bs 0 & \multicolumn{4}{c|}{} && \ddots\\ \hline
-	& & \multicolumn{4}{c|}{\bs B_{-1}} & \bs B_0 &\\
-	& & \multicolumn{3}{c}{}& \ddots &  & \ddots
-	\end{array}\right],
-\end{align*}
-where 
-\[\check{\bs B}_{0} = \left[\begin{array}{cccc}
-	 \bs{C}_+\otimes \bs{S} + \bs{T}_{++}\otimes \bs{I} & \bs{T}_{+0}\otimes \bs{I} & \bs{T}_{+-}\otimes \bs{D} & \\
-	 \bs{T}_{0+}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} & \bs{T}_{0-}\otimes \bs{D} & \\
-	 (\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha + \bs T_{-+}\otimes \bs D  &  & \bs{C}_-\otimes \bs{S} + \bs{T}_{--}\otimes \bs{I} & \bs{T}_{-0}\otimes \bs{I} \\
-	 \bs{T}_{0+}\otimes \bs{D} & & \bs{T}_{0-}\otimes \bs{I} & \bs{T}_{00}\otimes \bs{I} 
-	\end{array}\right].\]
-The top-left block represents the point masses due to the sticky boundary.
-The bottom-left block is the transition of density into point masses due to the sticky boundary.
-The top-right block is the transition of point masses into density as the process leaves the boundary.
-The term \((\bs C_-\bs P_{-+})\otimes \bs s \bs \alpha\) in \(\check{\bs B}_{0}\) incorporates the reflecting boundary behaviour.
-
-Upper boundaries can be included in an analogous manner.
-
-The orbit process \(\{\bs A(t)\}\) is not required to model the behaviour at the sticky boundary.
-However, we suppose that \(\bs A(t)=1\) at the boundaries.
-This choice is arbitrary but allows us to generalise some notation later when we are describing the evolution of the QBD-RAP.
-Further, for boundaries at \(y_0=0\) and \(y_{K+1} = (K+1)\Delta\), we let the level process take the value \(L(t)=-1\) at the lower boundary, and \(L(t)=K+1\) at the upper boundary.
-Denote the set of levels (including boundary levels) by \(\mathcal K = \{-1,0,\dots,K,K+1\}\).
-
-\begin{rem}
-	It may be possible to extend the QBD-RAP approximation to model jumps into and out of the boundary, provided that the jumps into/out of the boundary happen at an intensity which can be described by a matrix exponential distribution.
-
-	Furthermore, it may be possible to model more general boundary behaviours which can be approximated by a limit of matrix exponentials.
-For example, a boundary condition where the fluid queue spends a deterministic time at the boundary.
-\end{rem}
-
-\section{Initial conditions}\label{sec: initial conditions}
-%To this point, we have assumed that \((X(0),\varphi(0))=(y_\ell,i), \,i\in\mathcal S_+\) or \((X(0),\varphi(0))=(y_{\ell+1},i), \,i\in\mathcal S_-\), i.e., the initial position of the fluid is at the edge of the cell.
-In these cases, the appropriate initial orbit position is \(\bs \alpha\).
-We now generalise this so that we can approximate arbitrary initial conditions.
-
-%To motivate the construction, first consider the initial condition for the QBD-RAP approximation, \((L(0),\bs A(0), \varphi(0))=(\ell,\bs \alpha,i)\), corresponding to the initial condition of the fluid queue \((X(0),\varphi(0))=(y_\ell,i),\) for some \(i\in\mathcal S_+\).
-After time \(t\), \(c_it\in[0,\Delta)\), and assuming that no events of the QBD-RAP have occurred by this time, the position of the QBD-RAP is \((L(t),\bs A(t),\varphi(t)) = (\ell, \bs A(t), i),\) where \(\bs A(t) = \cfrac{\bs \alpha e^{c_i\bs{S}t}}{\bs \alpha e^{c_i\bs{S}t}\bs e}\).
-The level of the fluid at the same time is \(X(t) = y_\ell + c_i t\).
-We argued earlier that we can think of the orbit \(\bs k(x)\) as corresponding to the fluid being a distance of \(x\) from the left boundary of an interval when \(i\in\calS_+\), or from the right boundary of an interval when \(i\in\calS_-\).
-With this interpretation, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i),\,x_0\in\mathcal D_{\ell,i},\,i\in\mathcal S_+\cup\calS_{+0}\) is to set the initial orbit to \(\bs k(x_0-y_{\ell})\).
-Similarly, an approximation to the initial condition \((X(0),\varphi(0))=(x_0,i), \,x_0\in\mathcal D_{\ell,i}, x_0\neq y_{\ell+1},\,i\in\mathcal S_-\cup\calS_{-0}\) is to set the initial orbit position to \(\bs k(y_{\ell+1}-x_0).\)
-
-We use the notation \(\bs a_{\ell,i}(x_0) = \bs a_{\ell,+}(x_0) = \bs k(x_0-y_\ell)\) for the initial orbit position corresponding to \(x_0\) when the initial phase is \(i\in\calS_+\) and \(x_0\in\calD_{\ell,i}\).
-Similarly, for \(i\in\calS_-\) and \(x_0\in\calD_{\ell,i}\) define the notation \(\bs a_{\ell,i}(x_0) = \bs a_{\ell,-}(x_0) = \bs k(y_{\ell+1}-x_0)\).
-
-More generally, given an initial measure \(\mu_i(\cdot):= \mathbb P(X(0)\in\cdot,\varphi(0)=i)\), an approximation to this initial condition is to set the orbit to \(\displaystyle\int_{x\in\calD_{\ell,i}}\bs a_{\ell,i} (x) \wrt \mu_i\), for \(i\in\mathcal S,\, \ell\in\mathcal K\setminus\{-1,K+1\}\).
-
-\paragraph{Initial conditions for phases \(i\in\calS_0\)}
-While the augmented state-space scheme described above is a convergent one, the analysis of the QBD-RAP scheme is \emph{greatly} simplified if we allow initial conditions which allocate \(0\) probability to phases in \(i\in\calS_{+0}\cup\calS_{-0}\) only.
-To see why initial mass in \(\mathcal S_{+0}\cup\calS_{-0}\) might introduce further complexity, consider a sample path with \(\varphi(0)=k\in\calS_{+0}\) and \(\bs A(0)=\bs a\in\mathcal A\).
-On \(\varphi(u)\in\calS_{+0},\) for all \(u\in[0,t)\), then \(\bs A(u) = \bs a \) is constant.
-Upon the phase leaving \(\calS_{+0}\), at time \(t\), say, the phase will either jump to \(\calS_+\) or \(\calS_-\).
-If \(\varphi(t)\) jumps to \(i\in\calS_+\), the orbit process at time \(t\) will be \(\bs A(t)=\bs a \) and the process evolve from time \(t\) as if it has started in phase \(i\in\calS_+\) with initial orbit \(\bs a \) -- there are no major issues in this case.
-If, however, \(\varphi(t)\) jumps to \(i\in\calS_-\), then the orbit process will jump to \(\bs A(t) = \bs a \bs D\) at time \(t\).
-From time \(t\) onward the QBD-RAP process will evolve as if it has started in phase \(i\in\calS_-\) with initial orbit \(\bs a \bs D\).
-With the way we have structured the proof of convergence, we then have to complete two analyses of the QBD-RAP process -- one for the initial condition \(\bs a\) and another for the initial condition \(\bs a\bs D\).
-
-A solution to this problem is to augment a set of ephemeral states to the QBD-RAP which capture the initial sojourn in \(\calS_0\) only.
-Let us denote this set of ephemeral states \(\calS_0^*\) which has the same phase dynamics as \(\calS_0\).
-Suppose we wish to approximate a fluid queue with the initial condition \(\mu_k\) where \(\mu_k\) has mass 1.
-At time \(0\), we start the QBD-RAP in the ephemeral set of states \(\calS_0^*\) in phase \(k^*\in\calS_0^*\).
-While in \(\calS_0^*\), the phase process evolves according to the generator \(\bs T_{00}\).
-Upon exiting the ephemeral set \(\calS_0^*\) at time \(t\), say, the QBD-RAP process jumps to \(L(t)=\ell,\, \bs A(t)=\int_{\calD_{\ell,i}}\bs a_{\ell,i}(x)\wrt \mu_k,\,\varphi(t)=i\), where \(\ell\in\mathcal K\), \(i\in\calS_+\cup\calS_-\).
-
-Notice that we have not mentioned anything about the level or orbit process during the sojourn in \(\calS_0^*\).
-This is because all information about the level and phase during this time is contained within the initial condition \(\mu_k\), and the only thing that changes is the phase.
-
-To include the ephemeral states within the generator of the QBD-RAP we can do the following, 
-\begin{align*}
-	&\left[\begin{array}{c|cccc}
-		\bs T_{00} & \bs B_{-1}^* &\bs B_{0}^*&\bs B_{1}^* & \hdots \\\hline
-		&\check{\bs B}_{0} &  \check{\bs B}_{+1} &&\\
-		& \check{\bs B}_{-1}&\check{\bs B}_{0}&\bs B_{+1}& \\
-		&&\bs B_{-1} & \bs B_0&\ddots \\
-		&&&\ddots&\ddots 
-	\end{array}\right],
-\end{align*}
-where 
-\begin{align*}
-	\bs B_{-1}^* &= \vligne{\bs T_{0-}\mu_k(\{0\}) & \bs 0} & \bs B_{k}^* &= \vligne{\bs T_{0+}\otimes \bs a_k^+ & \bs 0 & \bs T_{0-}\otimes \bs a_k^- & \bs 0},
-	%
-	\\\check{\bs B}_{0} &= \left[\begin{array}{cc}\bs T_{--} & \bs T_{-0}\\\bs T_{0-} & \bs T_{00}\end{array}\right], 
-	& \check{\bs B}_{+1} &= \left[\begin{array}{cccc}\bs T_{-+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \\ \bs T_{0+}\otimes \bs \alpha & \bs 0 & \bs 0 & \bs 0 \end{array}\right],
-	%
-	\\\check{\bs B}_{-1} &= \left[\begin{array}{cc} \bs0&\bs0\\\bs0&\bs0\\(\bs C_-\bs P_{--})\otimes \bs s & (\bs C_-\bs P_{-0})\otimes \bs s \\\bs0&\bs0\end{array}\right], &&
-\end{align*}
-and 
-\begin{align*}
-	\bs a_\ell^r = \int_{\calD_{\ell,r}} \bs a_{\ell,r}(x)\wrt \mu_k,\, \ell\in\mathcal K\setminus\{-1,K+1\},\, r\in\{+,-\}. 
-\end{align*}
-
-%For technical reasons we must restrict the initial conditions so that there are no point masses at \(x_0=y_{\ell+1}\) when \(i\in\mathcal S_+\) or \(x_0=y_{\ell}\) when \(i\in\mathcal S_-\).
-In either of these cases, the initial orbit position would be \(\cfrac{\bs \alpha e^{\bs{S}\Delta}}{\bs \alpha e^{\bs{S}\Delta}\bs e}\) and the denominator is \(\bs \alpha e^{\bs{S}\Delta}\bs e\).
-To our knowledge, there are currently no results that ensure that \(\bs \alpha e^{\bs{S}\Delta}\bs e\) is not arbitrarily small for the class of matrix exponential distributions with which we work.
-For point masses at any other value of \(x_0\in\mathcal D_\ell, x_0\neq y_{\ell+1}\) and \(i\in\mathcal S_+\) the initial orbit position is \(\cfrac{\bs \alpha e^{\bs{S}(x_0-y_\ell)}}{\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e}\) and the denominator is \(\bs \alpha e^{\bs{S}(x_0-y_\ell)}\bs e\geq 1-\cfrac{\var\left(Z\right)}{(\Delta-(x_0-y_\ell))}\).
-Therefore, by choosing a matrix exponential \(Z\) with sufficiently small variance, then the denominator is arbitrarily close to 1.
-Similarly for initial point masses \(x_0\in\mathcal D_\ell\neq y_{\ell}\), \(i\in\mathcal S_-\).
-This is, however, no real practical loss as point masses at \(x_0=y_{\ell+1}\) when \(i\in\mathcal S_+\) (\(x_0=y_{\ell}\) when \(i\in\mathcal S_-\)) can, and should, have the initial conditions \(\bs v_{\ell+1,i}=\bs \alpha\) (respectively, \(\bs v_{\ell-1,i}\)), corresponding to the fluid starting at the left (right) of the adjacent interval which shares the boundary \(y_{\ell+1}\) (\(y_\ell\)).
-
-\section{At time \(t\) -- closing operators}\label{sec: closing}
-So far we have described how we can use a QBD-RAP to approximate a fluid queue. The level process \(\{L(t)\}\) of the QBD-RAP approximates the process 
-\begin{align}
-	\displaystyle \left\{\sum_{k\in\mathcal K}\sum_{i\in\calS} k1(X(t)\in\calD_{k,i})\right\}_{t\geq 0}. \label{eqn: kfafj}
-\end{align}
-This may be of interest in its own right.
-However, for some applications, this approximation may be too coarse -- the level process tells us nothing about where in the intervals \(\calD_{k,i}\) the fluid queue might be.
-
-We now describe how the orbit position \(\bs A(t)\) can be used to approximate the density of the level of the fluid queue \(X(t)\).
-This reconstruction is entirely post-hoc in the sense that it does not affect the dymaics of the QBD-RAP itself.
-For any time \(t\) we only take the information contained in \(\mathbb E[A(t)1(L(t)=\ell,\varphi(t)=i)]\) and rewrite it in a descriptive way.
-
-Suppose that the QBD-RAP is in level \(L(t)=\ell\), phase \(\phi(t)=j\in\calS_+\), and the orbit is \(\bs A(t)=\bs a\).
-If the QBD-RAP remains in phase \(j\), the QBD-RAP approximation will transition out of level \(\ell\) in the infinitesimal time interval \(t+\wrt u\) with density \(\bs a e^{|c_j|\bs{S}u}|c_j|\bs s\wrt u\).
-At the time of the change of level we estimate the position of \(X(t+u)\) by \(X(t+u)\approx y_{\ell+1}\).
-Tracing this back to time \(t\), we estimate the position of \(X(t)\) as \(X(t)\approx y_{\ell+1} - |c_j|u\).
-Reverse engineering this logic, the approximation to the density of \(X(t) \in\wrt x\) is \(\bs a e^{|c_j|\bs S(y_{\ell+1}-x)/|c_j|}|c_j|\bs s \wrt x/|c_j|\), since \(\wrt x = |c_j|\wrt u\) where \(\wrt x\) is an infinitesimal interval in space and \(\wrt u\) is an infinitesimal interval with respect to time.
-
-Similarly, for \(j\in\mathcal S_-\), if the QBD-RAP remains in phase \(j\) then the QBD-RAP approximation will transition out of level \(\ell\) in the infinitesimal time interval \(t+\wrt u\) with density \(\bs a e^{|c_j|\bs{S}u}|c_j|\bs s\wrt u\).
-At the time of the transition of level, we estimate the position of \(X(t+u)\) by \(X(t+u)\approx y_{\ell}\).
-Tracing this back to time \(t\), we estimate the position of \(X(t)\) as \(X(t)\approx y_{\ell} + |c_j|u\).
-Reverse engineering this logic, the approximation to the density \(X(t) \in\wrt x\) is \(\bs a e^{|c_j|\bs S(x-y_{\ell})/|c_j|}|c_j|\bs s \wrt x/|c_j|\), since \(\wrt x = |c_j|\wrt u\).
-
-For phases \(j\in\mathcal S_0\), if phase \(j\) is associated with \(\mathcal S_+\), we use the same approximation as if \(j\in \mathcal S_+\), and if phase \(j\in\mathcal S_0\) is associated with \(\mathcal S_-\), we use the same approximation as if \(j\in \mathcal S_-\).
-
-This reasoning leads to an approximation of the distribution of the fluid at time \(t\), \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \), as
-\begin{align}
-		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i)  
-		\bs a e^{\bs{S}z} \bs s \wrt x,\label{eqn: density approx aug}
-		% \intertext{for \(j\in\mathcal S_+\cup\calS_{+0}\) and}
-		% &\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \nonumber 
-		% \\&\qquad\times\bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x,\label{eqn: density approx aug2}
-\end{align}
-where \(x\in\mathcal D_{\ell,j}\), \(x_0\in\mathcal D_{\ell_0,i}\) and 
-\[
-	z = \begin{cases} 
-		x - y_\ell & i\in\calS_- \\
-		y_{\ell+1} - x & i\in\calS_+.
-	\end{cases}
-\]
-%For example, for the augmented scheme in Section~\ref{subsec: augmented }, we approximate \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i)\) by 
-%\begin{align}
-%	\label{eqn: density approx aug}
-%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(y_{\ell+1}-x)} \bs s \wrt x & j\in\mathcal S_+\cup \mathcal S_{+0},
-%		\\\label{eqn: density approx aug2}
-%		&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\varphi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a, \varphi(0) = i) \bs a e^{\bs{S}(x-y_\ell)} \bs s \wrt x & j\in\mathcal S_-\cup \mathcal S_{-0},
-%\end{align}
-
-While the estimate (\ref{eqn: density approx aug}) is appealing, it is, however, defective -- it does not integrate to 1 for any finite-dimensional matrix exponential distribution.
-To see this, compute
-\begin{align*}
-	&\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{z\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0),
-	\\&\qquad{}\qquad{}\phi(0) = i) \bs a e^{ \bs{S}z} \bs s \wrt z
-	% \\&\qquad{}{}+\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S_-\cup \calS_{-0}}\int_{x\in\mathcal D_{\ell,j}}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0,
-	% \\&\qquad{}\qquad{}\bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{ \bs{S}(x-y_\ell)} \bs s \wrt x
-	\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a, \phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), 
-	\\&\qquad{}\qquad{}\phi(0) = i) (1-\bs a e^{ \bs{S}\Delta}\bs e  )
-	\\&<1,
-\end{align*}
-since \(\bs a e^{\bs{S}u} \bs e>0\) for any \(u>0\). For an orbit position \(\bs A(t)=\bs a\), the amount of mass missing from each level and phase is 
-\[\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a e^{\bs{S}\Delta} \bs e \wrt x.\]
-The problem is that the triple \((\bs a,\bs S, \bs s)\) defines a distribution on \(x\in[0,\infty)\), however, it only makes sense to take \(x\in[0,\Delta)\) in the approximation scheme.
-
-%One way to rectify this is to renormalise the estimates in the right-hand side of (\ref{eqn: density approx aug})-(\ref{eqn: density approx aug2}) with the missing mass, i.e.~the approximation to \(\mathbb P(X(t)\in \wrt x , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) is
-%\begin{align}
-%	\label{eqn: density approx 2}
-%		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(y_{\ell+1}-x)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x \intertext{for \(j\in\mathcal S_+\cup \calS_{+0}\), and}
-%		\label{eqn: density approx 22}
-%		\displaystyle\int_{\bs a \in\mathcal A} \mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \cfrac{\bs a e^{ \bs{S}(x-y_\ell)} \bs s}{1-\bs a e^{ \bs{S}\Delta} \bs e} \wrt x ,
-%\end{align}
-%for \(j\in\mathcal S_-\cup \calS_{-0}.\) This renormalisation distributes any missing mass over the interval \(\calD_{\ell,j}\) proportional to the estimated density in the region.
-
-To partially rectify this we can instead approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
-\begin{align}
-%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
-	\label{eqn: density approx 3+}&
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a u_i^\ell(x) \wrt x ,
-\end{align}
-{where}
-\begin{align}
-		\bs a u_i^\ell(x) = \begin{cases}
-			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} \bs s + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \bs s\right), & i\in\calS_+, \\ 
-			\bs a \left(e^{ \bs{S}(x-y_\ell)} \bs s+e^{ \bs{S}(2\Delta-(x-y_\ell))} \bs s\right), & i\in\calS_-. \\ 
-		\end{cases}\label{eqn: density approx 3+a}
-\end{align}
-Intuitively, we take the density function \(\bs ae^{\bs{S}x}\bs s\) on \(x \in [0,2\Delta)\), and `fold' it back on itself around \(\Delta\), to create a density function on \([0,\Delta)\), 
-\[\bs ae^{\bs{S}x}\bs s +\bs ae^{\bs{S}(2\Delta-x)}\bs s,\, x \in [0,\Delta).\]
-The missing mass is now proportional to 
-\[\bs a e^{\bs S 2\Delta}\bs e,\]
-which is less than the quantity \(\bs a e^{\bs S \Delta}\bs e\) from the approximation scheme (\ref{eqn: density approx aug}).
-There is nothing special about the choice of truncating the distribution at \(2\Delta\) in this construction, and we could choose any truncation value greater than \(\Delta\).
-
-A third option is to approximate \(\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \) by
-\begin{align}
-%	&\mathbb P(X(t)\in \wrt x  , \varphi(t) = j\mid X(0)=x_0, \varphi(0) = i) \nonumber
-	\label{eqn: density approx 4+}&
-		\displaystyle\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs  a_{\ell,i}(x_0), \phi(0) = i) \bs a v_i^\ell(x) \wrt x ,
-\end{align}
-{where}
-\begin{align}
-		\bs a v_i^\ell(x) = \begin{cases}
-			\bs a \left(e^{ \bs{S}(y_{\ell+1}-x)} + e^{ \bs{S}(2\Delta-(y_{\ell+1}-x))} \right)\left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_+, \\ 
-			\bs a \left(e^{ \bs{S}(x-y_\ell)}+e^{ \bs{S}(2\Delta-(x-y_\ell))}\right) \left[I-e^{\bs S 2\Delta}\right]^{-1}\bs s, & i\in\calS_-. \\ 
-		\end{cases}
-\end{align}
-Here, we take the density function \(\bs a e^{\bs S x}\bs s\), defined on \([0,\infty)\), and map it to a density function on \([0,\Delta)\) by \(\Delta - |(x \mod 2\Delta )-\Delta|\).
-The resulting density function is
-\begin{align*}
-	v(x) &= \sum_{m=0}^\infty \bs a \left(e^{\bs S (x+2\Delta m)} + e^{\bs S (2\Delta -x+2\Delta m)} \right)\bs s
-	%
-	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\sum_{m=0}^\infty e^{\bs S 2\Delta m} \bs s
-	%
-	\\&= \bs a \left(e^{\bs S x} + e^{\bs S (2\Delta -x)} \right)\left[I-e^{S2\Delta}\right]^{-1} \bs s.
-\end{align*}
-
-%The advantages of this approximation are;
-%\begin{itemize}
-%	\item the denominator \(1-\bs ae^{\bs{S}2\Delta}\bs s\) is provably close to \(1\) for concentrated matrix exponential distributions.
-%	\item in (\ref{eqn: density approx 2}) any of the once missing mass is spread across the interval \(\calD_{\ell,j}\) proportionally to the estimated density in the region, whereas the missing mass should be more concentrated around \(\Delta\), as this is the region around which the function \(\bs \alpha e^{\bs Sx}\bs s\) fails to be like the delta function most serverely.
-This is what (\ref{eqn: density approx 3+})-(\ref{eqn: density approx 3-}) do.
-%\end{itemize}
-%Meanwhile, (\ref{eqn: density approx 3+})-(\ref{eqn: density approx 3-}) maintains positivity and smoothness.
-
-The difference between the approximation schemes is the way in which the vector \(\bs a\) is used.
-We generalise this idea with the concept of \emph{closing operators} which is a linear operator \(V(x):\mathcal A \to \mathbb R,\) for each \(x\in[0,\Delta)\).
-For example, the closing operator in Equations (\ref{eqn: density approx aug}) is the operator \(V(x),\,x\in[0,\Delta)\) such that for any \(\bs a \in \mathcal A\),
-\begin{align}
-	\bs a V(x) &= \bs a e^{\bs Sx}\bs s.\label{eqn: alal}
-	%\\\bs a V_-(x-y_\ell) &= \bs a e^{\bs S(x-y_{\ell})}\bs s,\label{eqn:ndaenf}
-\end{align}
-Similarly, in (\ref{eqn: density approx 3+})
-\begin{align}
-	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right),\label{eqn:546}
-\end{align}
-and in (\ref{eqn: density approx 4+})
-\begin{align}
-	\bs a V(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)}\left[I-e^{S2\Delta}\right]^{-1} \bs s\right).\label{eqn:5462}
-\end{align}
-The operator \(V(x)\) is a choice which forms part of the definition of the approximation scheme.
-As we shall see, given certain properties of \(V(x)\), we can prove that the approximation scheme converges, and ensures positivity.
-In the cases above, all the closing operators (\ref{eqn: alal})-(\ref{eqn:546}) lead to an approximation which converges and, due to their interpretation as probability densities, ensure positivity.
-
-% \begin{rem}
-% Since we require \(V\) to be linear, then we actually have 
-% \begin{align}
-% 	&\int_{\bs a \in\mathcal A}\mathbb P(L(t) = \ell, \bs A(t)\in\wrt \bs a,\phi(t) = j\mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i) \bs a V(x) \nonumber 
-% 	\\&= \mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=j) \mid L(0)=\ell_0, \bs A(0) = \bs a_{\ell_0,i}(x_0), \phi(0) = i]V(x).
-% \end{align}
-% \end{rem}
-
-% \begin{rem}
-	\paragraph{Comments on linearity and normalisation}
-	We considered taking the closing operators to be 
-	\begin{align}
-		\bs a V(x) &= \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e},\label{eqn: nonlin}
-	\end{align}
-	which also ensures that the solution always preserves mass (the result of integrating (\ref{eqn: nonlin}) over \([0,\Delta)\) is 1).
-However, (\ref{eqn: nonlin}) is not a linear operator.
-The importance of this fact is that for a linear operator,
-	\[\mathbb E[\bs A(t) V(x)] = \mathbb E[\bs A(t)]V(x),\]
-	thus, the computation of \(\mathbb E[\bs A(t) V(x)]\) can be achieved by first computing \(\mathbb E[\bs A(t)]\) and then applying \(V(x)\) to the result.
-In this sense, \(\mathbb E[\bs A(t)]\) contains all of the information about the history of the process up to time \(t\) needed to compute \(\mathbb E[\bs A(t) V(x)]\).
-This is much the same as in RAPs and QBD-RAPs where \(\mathbb E[\bs A(t)]\) is all that is required to compute probabilities about the future of the processes from time \(t\) onwards.
-
-	When \(V(x)\) is not linear then, in general, 
-	\[\mathbb E[\bs A(t) V(x)] \neq \mathbb E[\bs A(t)]V(x).\]
-	For example, with (\ref{eqn: nonlin}), 
-	\begin{align}
-		\mathbb E[\bs A(t) V(x)] = \int_{\bs a \in\mathcal A} \mathbb P(\bs A(t) \in \wrt \bs a) \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e}.\label{eqn:nonlin2}
-	\end{align}
-	While the calculation in (\ref{eqn:nonlin2}) is theoretically possible, practically it is not.
-The seemingly innocuous integral over \(\bs a \in\mathcal A\) actually amounts to an integral over all possible sample paths of the QBD-RAP (i.e.~all possible event times of the QBD-RAP before time \(t\)) and in general, there is no `nice' way to do this computation.
-
-	In the linear case, computation of
-	\[\mathbb E[\bs A(t)] = \int_{\bs a\in\mathcal A} \mathbb P(\bs A(t) \in \wrt \bs a) \bs a,\]
-	amounts to a matrix-exponential calculation.
-
-	In practice we use 
-	\begin{align}
-		\cfrac{\mathbb E\left[\bs A(t)\right] \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\mathbb E\left[\bs A(t)\right] e^{\bs S 2\Delta}\bs e}
-	\end{align}
-	as a normalised version of (\ref{eqn:546}), and this seems to work well. However, this is not the same as using the closing operator (\ref{eqn: nonlin}).
-% \end{rem}
-
-% \begin{rem}
-% 	Consider approximating a function \(f(x)\) by a linear combination of the elements of \(e^{\bs S x}\bs s\), i.e.
-% 	\begin{align}
-% 		f(x) \approx \bs b e^{\bs S x} \bs s.
-% 	\end{align}
-% 	Further, suppose that we require the residual \(f(x) - \bs b e^{\bs S x} \bs s\) to be orthogonal to the space of functions spanned by the elements of 
-% 	\begin{align}
-% 		\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}.
-% 	\end{align}
-% 	This amounts to the requirement  
-% 	\begin{align}
-% 		\int_{0}^\infty \left(f(x) - \bs b e^{\bs S x} \bs s\right)\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}\wrt x = \bs 0%\int \bs e^{\bs Sx}\bs s \cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e} \bs c \wrt x
-% 	\end{align}
-% 	which implies
-% 	\begin{align}
-% 		\bs b \int_{0}^\infty e^{\bs S x} \bs s\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}\wrt x = \int_{0}^\infty f(x)\cfrac{\bs \alpha e^{\bs Sx}}{\bs \alpha e^{\bs Sx}\bs e}\wrt x
-% 	\end{align}
-% 	and therefore 
-% 	\begin{align}
-% 		\bs b = \int_{0}^\infty f(x)\bs k(x)\wrt x \bs D^{-1}.
-% 	\end{align}
-
-% 	This suggests that we may be able to reconstruct solutions as 
-% 	\[\mathbb E[\bs A(t)1(L(t)=\ell, \varphi(t)=j)]\bs D^{-1} e^{\bs S x}\bs s.\] 
-% 	Some basic numerical experiments suggest this might work and could be investigated as further research.
-Since this reconstruction does not have a stochastic interpretation, and does not enforce positivity, we do not persue it further here.
-% \end{rem}
+\chapter{Approximating fluid queues with the discontinuous Galerkin method \label{ch:galerkin}} 
+% PREFACE WITH CONTRIBUTIONS 
+% TUTORIAL: HOW TO -- AS IN PEPER
+% PROBLEMS AND SOLUTIONS 
+% 	NON-NEGATIVITY AND OSCILLATIONS 
+%	FILTERING AND LIMITING 
+%	NOT A SOLUTION IN OUR CONTEXT & MOTIVATION FOR A NEW METHOD -- THE QBD-RAP

### 2021-11-29 chapter5/chapter.tex
first steps of numerics chapter

%!TEX root = ../thesis.tex
\chapter{Numerical investigations\label{sec: numerics}}
We have now established multiple methods for approximating fluid queues which suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}. In this chapter we numerically investigate various aspects of the approximation schemes. Throughout, we compare the QBD-RAP scheme (from Chapters~\ref{sec: conv} and \ref{sec: construction and modelling}), the discontinuous Galerkin method (without limiters or filtering of the solution) (as decribed in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013} (which we will refere to a just the \emph{uniformisation scheme}). Recall that the uniformisation scheme is equivalent to a discontinuous Galerking scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux. Like the QBD-RAP scheme, due to its stochastic interpretation, the uniformisation scheme is positivity preserving.

Where appropriate, we also compare the above methods with the discontinuous Galerkin method with a slope-limiter implemented, which avoids oscilations and negative solutions. However, recall that, in the context of approximating operators of fluid-fluid queues with these schemes, slope-limiting is not possible, and even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.

The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased. For discontinuous Galerkin method, this means increasing the number of basis functions used to approximate the solution within each cell. For the QBD-RAP scheme, this means increasing the order of the CME distribution used to constrcut the scheme. For the uniformisation scheme, this means dividing each cell into smaller sub-cells over which we approximate the solution. As such, each approximation scheme leads to approximations which are matrices of the same size (although not necessarily the same number of non-zero components). For example, if we use 3 basis functions in the discontinuous Galerkin method, we approximate the solution by a quadratic on each cell; the comparable QBD-RAP scheme is to use a CME of order 3; and the comparable uniformisation scheme is to divide each cell into 3 sub-cells of width \(\Delta/3\).

To keep the content of this chapter contained, we do not investigate the performance of the schemes as the cell sizes are decreased, other than for the uniformisation scheme. As such, for each numerical experiment, we keep the cell size fixed. As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next and link all the cells together. We investigate schemes with an upwind flux only.
We do not investigate the stability of the schemes with respect to the step-size of the time-integration or time-integration scheme (where required). Instead, we fix the time-integration step size for each numerical experiment at a suitably small value so as to obey a certain stability criterion (a CFL-like condition, REFs, appendix?). For schemes which require us to integrate over time we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (see appendix and REF). This scheme claims to introduce no more oscilations into the solution as we integrate over time (REFs). Where a slope limiter is implemented, we always implement the \emph{Generalised MUSCL} limiter (REF and Appendix). We do not investigate filtering for the DG scheme (REFs).

The performance of the discontinuous Galerkin method has been well-studied (REFs), and it is well-known that the discontinuous Galerkin method performs remarkably well on problems with smooth solutions (REF). Here, we mostly focus on investigating the numerical performance of the methods on problems with non-smooth solutions, the purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP method. In the stochastic modelling community it is very common to have problems with discontinuous solutions, such as a non-smooth initial conditions (REFs). Even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities into the problem. There is a very specific set of conditions on the inital density and point masses which must hold for the distribution of the fluid queue to remain smooth as it evolves over time (REFs).Limiting distributions of fluid queues are smooth, however. Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues.

As we shall see, if the solution is smooth, then the Galerkin method is highly-effective and displays rapid convergence to the solution as the number polynomial basis functions is increased. However, when the solution is non-smooth oscilations and negative approximations may occur when using the discontinuous Galerkin method, leading to non-sensical solutions. In these cases, the QBD-RAP approximation performs relatively well compared to the other positivity preserving schemes investigated.

% THE STRUCTURE OF THE CHAPTER IS AS FOLLOWS.

% 1. FUNCTION APPROX./RECONSTRUCTION

% 2. 1d wave eqn

% 3. STATIONARY DISTRIBUTIONS
	% i. SIMPLER
	% ii. MORE COMPLEX

% 4. TRANSIENT DISTRIBUTIONS 
	% i. SIMPLER
	% ii. MORE COMPLEX

% 5. HITTING TIMES 
	% i. SIMPLER
	% ii. MORE COMPLEX

% 6. FLUID-FLUID QUEUES

% 7. CONCLUDING REMARKS

\section{Function approximation/reconstruction}
We start our investigation by looking at how well the methods perform at approximating an initial condition. For the discontinuous Galerkin method, this means projecting the solution on to the set of polynomial functions which define the scheme, then reconstructing the solution. For the spatially-coherent uniformisation scheme, this means looking at the sub-cell averages. For the QBD-RAP scheme, this means computing the intial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}. First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}) performs reasonably compared to the others, and we are able to theoretically prove that it converges. Therefore we use this reconstruction method throughout the rest of the chapter.

For this section we cosider approximating the initial condition over a single cell of width \(\Delta = 1\). To numerically evaluate intergals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations. Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points overwhich to compute the KS statistics.

First we briefly investigate the possible choices of closing operators as introduced in Section~\ref{sec: closing}.

\subsection{QBD-RAP closing operators}
The three closing operators we investgate are the following. The unnormalised closing vector in Equation~(\ref{eqn: alal}). A normalised version of the closing operator in Equation~(\ref{eqn:546}), with the normalised version given by (\ref{eqn: nonlin}). We refer to the closing vector in (\ref{eqn: nonlin}) as the \emph{naive normalised} closing vector. Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme, though it clearly appears to. The third closing vector we investigate is that in Equation~(\ref{eqn:5462}).

First consider approximating the initial condition with density function \(1(x<0.5)\). In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation. Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders. It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interfers with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme. Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection. The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward). Intuitively, there should be very little mass in the tail of the distribution as we are using conentrated matrix exponential distributions in the reconstruction.

If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs}, then we see that the unnormalised reconstruction performs the poorest. Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests naive normalised and normalised reconstruction are very similar. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(2x1(x<0.5)+1(x\geq 0.5)\), and the approximations (left) and \(L^2\) error between the true PDF, \(2\times 1(x<0.5)\), and the approximations (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 2 ks error qbdrap closing vecs}
\end{figure}

Now consider approximating the initial density \(f(x)=1\). Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error, and \(L^2\) error between the true initial distribution and the reconstructions, we now see that the normalised reconstructions outperform the unnormalised reconstruction. This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(x\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
\end{figure}

Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised and normalised closing operators for order 1, 3, 5 and 7, as well as the true PDF. Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots and for all orders. This feature is due to a significant amount of mass being lost due to the truncation. In comparison the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator. In the reconstruction we also need to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}. Here we suppose that the phase is positive, so the `folding' around \(\Delta\) in the normalised closing operators appears appears at the left of the plots. In general, reconstructions via the unnormalised closing operator typically underestimate the value of the function in this region.

Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval. This is a common feature, and drawback, of the QBD-RAP method. Perhaps there is a different reconstruction method or yet another closing operator which could aleviate this issue. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=1\) (orange dashed line).}
	\label{fig: pdf reconstructed}
\end{figure}

Next we consider approximating the initial distribution with density \(-6x^2+6x\). Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing opeartor performs the worst, while the performance of the two normalised reconstructions is indistinguishable. However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, we see that with this metric, then unnormalised closing operator performs better than the two normalised ones for orders 5 and above. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(-2x^3+3x^2\), and the approximation (left) and \(L^2\) error between the true PDF \(-6x^2+6x\) and the approximation (right), for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 6 ks error qbdrap closing vecs}
\end{figure}

The fact that the unnormalised closing operator outperforms the two normalised ones can be explain, once again, by the fact that, for the normalised operators we have `folded' the ME density function back on itself. In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\). In Figure~\ref{fig: pdf reconstructed quadratic} we observe that, at the left of the plot, the normalised reconstructions over estimate the density function in this region, where as the unnormalised reconstruction looks to be doing better. Recall that the `folding' in the normalised closing operatot manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=-6x^2+6x\) (orange dashed line).}
	\label{fig: pdf reconstructed quadratic}
\end{figure}

Lastly, we consider approximating the intial distribution with PDF \(3e^{-3x}/(1-e^{-3})\). This density function is at a maximum at the left of the region. Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case. Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that the normalised reconstructions perform relatively well compared to the unnormalised reconstruction as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF). If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \((1-e^{-3x})/(1-e^{-3})\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 7 ks error qbdrap closing vecs}
\end{figure}

### 2021-11-30 chapter5/chapter.tex
.

@@
 For the spatially-coherent uniformisation scheme, this means looking at the sub-cell averages.
-For the QBD-RAP scheme, this means computing the intial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}.
+For the QBD-RAP scheme, this means computing the initial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}.
 First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}) performs reasonably compared to the others, and we are able to theoretically prove that it converges.
@@
 
-For this section we cosider approximating the initial condition over a single cell of width \(\Delta = 1\).
-To numerically evaluate intergals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations.
-Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points overwhich to compute the KS statistics.
+For this section we consider approximating the initial condition over a single cell of width \(\Delta = 1\).
+To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations.
+Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points over which to compute the KS statistics.
 
@@
 Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders.
-It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interfers with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme.
+It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interferes with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme.
 Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection.
 The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward).
-Intuitively, there should be very little mass in the tail of the distribution as we are using conentrated matrix exponential distributions in the reconstruction.
+Intuitively, there should be very little mass in the tail of the distribution as we are using concentrated matrix exponential distributions in the reconstruction.
 
@@
 This is a common feature, and drawback, of the QBD-RAP method.
-Perhaps there is a different reconstruction method or yet another closing operator which could aleviate this issue.
+Perhaps there is a different reconstruction method or yet another closing operator which could alleviate this issue.
 \begin{figure}
@@
 Next we consider approximating the initial distribution with density \(-6x^2+6x\).
-Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing opeartor performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
+Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
 However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, we see that with this metric, then unnormalised closing operator performs better than the two normalised ones for orders 5 and above.
@@
 In Figure~\ref{fig: pdf reconstructed quadratic} we observe that, at the left of the plot, the normalised reconstructions over estimate the density function in this region, where as the unnormalised reconstruction looks to be doing better.
-Recall that the `folding' in the normalised closing operatot manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions.
+Recall that the `folding' in the normalised closing operator manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions.
 \begin{figure}
@@
 
-Lastly, we consider approximating the intial distribution with PDF \(3e^{-3x}/(1-e^{-3})\).
+Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\).
 This density function is at a maximum at the left of the region.
@@
 \end{figure}
+
+\subsection{Comparison of methods}
+Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.
+
+\section{Travelling wave}
+Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with varying initial conditions.
+Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\).
+The PDE (if/when it exists) which describes this system is
+\[\cfrac{\partial}{\partial t} f(x,t) = \cfrac{\partial}{\partial x} f(x,t),\]
+where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
+\[f(x,t) = f(x-t,0)\]
+so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).
+
+We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space. We discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We then

### 2021-11-30 chapter5/chapter.tex
sec 5.1

@@
 The naive normalised (orange) an normalised closing vectors are coincident.}
+	\label{fig: fun 4 ks error qbdrap closing vecs}
 \end{figure}
@@
 
+Given the evidence above we choose to use the normalised closing operator to reconstruct solutions.
+Further, unlike the naive normalised operator, the normalised operator is theoretically justified in that we proved the closing operator leads to a convergent scheme and is linear.
+
 \subsection{Comparison of methods}
 Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.
+
+First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5.
+This distribution does not have a PDF so we can compare the CDFs only.
+In Figure~\ref{fig: fun 1 comp} we plot the KS metric (left) and \(L^1\) metric (right) between the true CDF and the reconstructed approximations.
+Observing the KS metric, it appears that none of the methods converge and the KS error sits around 0.5.
+This reflects the fact that convergence in distribution implies pointwise convergence of the CDFs except at points of discontinuity.
+None of the methods appear to converge at this point.
+Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the methds appear to show the convergent behaviour we expect.
+Here, the uniformisation method appears to perform the best, while the QBD-RAP method performs the worst.
+The rate of convergence of the QBD-RAP method appears to be similar to the rate of convergence of the DG scheme.
+
+Perhaps it is no surprise that the uniformisation method performs best.
+In the uniformisation method as we increase the order, we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the function.
+As such, the uniformisation method can produce a piecewise continous, linear approximation to the CDF.
+In contrast, both the DG and QBD-RAP methods result in a smooth approximation to the CDF.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
+	\caption{KS error between the true CDF, \(1(x\geq 0.5)\), and the approximations (left) and \(L^1\) error between the true CDF and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: fun 1 comp} 
+\end{figure}
+
+In Figure~\ref{fig: pdf comp fun 1} we plot the approximated CDFs for the DG, uniformisation and QBD-RAP methods alonside the true CDF.
+The DG method displays undesireable fetures for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1.
+On the other hand, although the QBD-RAP method converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1.
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{chapter5/figs/comp/fun1/cdfs_formatted.pdf}
+	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (green dot-dashed line) methods.
+The true distribution function is \(1(x\geq 0.5)\) (grey dotted line).}
+	\label{fig: pdf comp fun 1}
+\end{figure}
+
+Now consider approximaing the initial distribution with PDF \(1(x\leq 0.5)\).
+Figure~\ref{fig: fun 2 comp} plots the KS error (left) and \(L^2\) error between the true and approximate PDFs (right).
+Figure~\ref{fig: fun 2 comp} suggests that all methods converge at a similar rate for this problem.
+Once again, the QBD-RAP method performs worst, the uniformisation method second, and the DG method the best.
+However, once again the DG method exhibits undesireable properties -- the approximation to the CDF is, at some points above 1, is not monotonic, though these violations are not as severe in this case as they were for the case above.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
+	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(x\geq 0.5)\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: fun 2 comp} 
+\end{figure}
+
+So far we have considered two problems which exhibit discontinuities.
+At the other extreme we now consider an initial distribution with density \(-6z^2+6x\).
+In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDF.
+Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG method can theoretically approximate the initial condition exactly.
+This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from 1-3, then plateaus.
+Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, and due to machine arithmetic, the errors for the DG scheme are not 0.
+Regarding the other two methods, they too appear to be convergent at approximately the same rate, with the uniformisation method perform better for the KS error, but very similarly in terms of the \(L^2\) error.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun6/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun6/meshs_l2_pdf_error_formatted.pdf}
+	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(-6z^2+6x\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: fun 6 comp} 
+\end{figure}
+
+Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\).
+Figure~\ref{fig: fun 9 comp} shows the errors.
+Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs tell a similar story.
+For sufficiently high order, the DG method approximates the inital condition very well.
+The uniformisation method performs second best, while the QBD-RAP method is slowest to converge.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun9/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun9/meshs_l2_pdf_error_formatted.pdf}
+	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(cos(4\pi(x+0.5))+1\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: fun 9 comp} 
+\end{figure}
+
+\FloatBarrier
 
@@
 The PDE (if/when it exists) which describes this system is
-\[\cfrac{\partial}{\partial t} f(x,t) = \cfrac{\partial}{\partial x} f(x,t),\]
+\[\cfrac{\partial}{\partial t} f(x,t) = -\cfrac{\partial}{\partial x} f(x,t),\]
 where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
@@
 
-We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space. We discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We then
+We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space.
+We discretise the interval \([0,10]\) into \(10\) cell, each of width 1.
+We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric.
+We use the SSPRK4 method to integrate over time using a step size of \(0.005\), and we evolve the system until time \(t=4\).
+For the DG method we also implement the Generalise MUSCL (REF, appendix?) slope limiter to stabilise the integration.
+
+First consider the initial condition with PDF \(1(x<1)\) so the level of the fluid queue is uniformly distributed over the first cell.
+For the DG and uniformisation methods, the initial condition can be represented exatly, whereas, for the QBD-RAP method, it cannot.
+Thus, in this case, there is no discretisation error in constructing the initial conition for the DG and uniformisation methods.

### 2021-12-01 chapter5/chapter.tex
.:

@@
 In this chapter we numerically investigate various aspects of the approximation schemes.
-Throughout, we compare the QBD-RAP scheme (from Chapters~\ref{sec: conv} and \ref{sec: construction and modelling}), the discontinuous Galerkin method (without limiters or filtering of the solution) (as decribed in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013} (which we will refere to a just the \emph{uniformisation scheme}).
-Recall that the uniformisation scheme is equivalent to a discontinuous Galerking scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux.
+Throughout, we compare the QBD-RAP scheme (from Chapters~\ref{sec: conv} and \ref{sec: construction and modelling}), the discontinuous Galerkin method (without limiters or filtering of the solution) (as decribed in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013} (which we will refer to a just the \emph{uniformisation scheme}).
+Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux.
 Like the QBD-RAP scheme, due to its stochastic interpretation, the uniformisation scheme is positivity preserving.
 
-Where appropriate, we also compare the above methods with the discontinuous Galerkin method with a slope-limiter implemented, which avoids oscilations and negative solutions.
+Where appropriate, we also compare the above methods with the discontinuous Galerkin method with a slope-limiter implemented, which avoids oscillations and negative solutions.
 However, recall that, in the context of approximating operators of fluid-fluid queues with these schemes, slope-limiting is not possible, and even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.
@@
 For discontinuous Galerkin method, this means increasing the number of basis functions used to approximate the solution within each cell.
-For the QBD-RAP scheme, this means increasing the order of the CME distribution used to constrcut the scheme.
+For the QBD-RAP scheme, this means increasing the order of the CME distribution used to construct the scheme.
 For the uniformisation scheme, this means dividing each cell into smaller sub-cells over which we approximate the solution.
@@
 For schemes which require us to integrate over time we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (see appendix and REF).
-This scheme claims to introduce no more oscilations into the solution as we integrate over time (REFs).
+This scheme claims to introduce no more oscillations into the solution as we integrate over time (REFs).
 Where a slope limiter is implemented, we always implement the \emph{Generalised MUSCL} limiter (REF and Appendix).
@@
 Even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities into the problem.
-There is a very specific set of conditions on the inital density and point masses which must hold for the distribution of the fluid queue to remain smooth as it evolves over time (REFs).Limiting distributions of fluid queues are smooth, however.
+There is a very specific set of conditions on the initial density and point masses which must hold for the distribution of the fluid queue to remain smooth as it evolves over time (REFs).Limiting distributions of fluid queues are smooth, however.
 Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues.
@@
 As we shall see, if the solution is smooth, then the Galerkin method is highly-effective and displays rapid convergence to the solution as the number polynomial basis functions is increased.
-However, when the solution is non-smooth oscilations and negative approximations may occur when using the discontinuous Galerkin method, leading to non-sensical solutions.
+However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin method, leading to non-sensical solutions.
 In these cases, the QBD-RAP approximation performs relatively well compared to the other positivity preserving schemes investigated.
@@
 \subsection{QBD-RAP closing operators}
-The three closing operators we investgate are the following.
+The three closing operators we investigate are the following.
 The unnormalised closing vector in Equation~(\ref{eqn: alal}).
@@
 We refer to the closing vector in (\ref{eqn: nonlin}) as the \emph{naive normalised} closing vector.
-Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme, though it clearly appears to.
+Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme, though it clearly appears to converge.
 The third closing vector we investigate is that in Equation~(\ref{eqn:5462}).
@@
 Observing the KS metric, it appears that none of the methods converge and the KS error sits around 0.5.
-This reflects the fact that convergence in distribution implies pointwise convergence of the CDFs except at points of discontinuity.
+This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except at points of discontinuity.
 None of the methods appear to converge at this point.
-Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the methds appear to show the convergent behaviour we expect.
+Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the methods appear to show the convergent behaviour we expect.
 Here, the uniformisation method appears to perform the best, while the QBD-RAP method performs the worst.
@@
 In the uniformisation method as we increase the order, we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the function.
-As such, the uniformisation method can produce a piecewise continous, linear approximation to the CDF.
+As such, the uniformisation method can produce a piecewise continuous, linear approximation to the CDF.
 In contrast, both the DG and QBD-RAP methods result in a smooth approximation to the CDF.
@@
 
-In Figure~\ref{fig: pdf comp fun 1} we plot the approximated CDFs for the DG, uniformisation and QBD-RAP methods alonside the true CDF.
-The DG method displays undesireable fetures for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1.
+In Figure~\ref{fig: pdf comp fun 1} we plot the approximated CDFs for the DG, uniformisation and QBD-RAP methods alongside the true CDF.
+The DG method displays undesirable features for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1.
 On the other hand, although the QBD-RAP method converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1.
@@
 
-Now consider approximaing the initial distribution with PDF \(1(x\leq 0.5)\).
+Now consider approximating the initial distribution with PDF \(1(x\leq 0.5)\).
 Figure~\ref{fig: fun 2 comp} plots the KS error (left) and \(L^2\) error between the true and approximate PDFs (right).
@@
 Once again, the QBD-RAP method performs worst, the uniformisation method second, and the DG method the best.
-However, once again the DG method exhibits undesireable properties -- the approximation to the CDF is, at some points above 1, is not monotonic, though these violations are not as severe in this case as they were for the case above.
+However, once again the DG method exhibits undesirable properties -- the approximation to the CDF is, at some points above 1, is not monotonic, though these violations are not as severe in this case as they were for the case above.
 \begin{figure}
@@
 First consider the initial condition with PDF \(1(x<1)\) so the level of the fluid queue is uniformly distributed over the first cell.
-For the DG and uniformisation methods, the initial condition can be represented exatly, whereas, for the QBD-RAP method, it cannot.
-Thus, in this case, there is no discretisation error in constructing the initial conition for the DG and uniformisation methods.
+For the DG and uniformisation methods, the initial condition can be represented exactly, whereas, for the QBD-RAP method, it cannot.
+Thus, in this case, there is no discretisation error in constructing the initial condition for the DG and uniformisation methods.

### 2021-12-06 chapter5/chapter.tex
rought redraft of chapters 4/5, started on one D wave examples

--- 2021-11-30 chapter5/chapter.tex
@@
 %!TEX root = ../thesis.tex
-\chapter{Numerical investigations\label{sec: numerics}}
-We have now established multiple methods for approximating fluid queues which suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}.
-In this chapter we numerically investigate various aspects of the approximation schemes.
-Throughout, we compare the QBD-RAP scheme (from Chapters~\ref{sec: conv} and \ref{sec: construction and modelling}), the discontinuous Galerkin method (without limiters or filtering of the solution) (as decribed in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013} (which we will refere to a just the \emph{uniformisation scheme}).
-Recall that the uniformisation scheme is equivalent to a discontinuous Galerking scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux.
-Like the QBD-RAP scheme, due to its stochastic interpretation, the uniformisation scheme is positivity preserving.
+\chapter{Corollaries to Chapter~\ref{sec: conv}}\label{ch: global results}
+In this chapter we prove a number of results stemming from Theorem~\ref{thm: a thm!} in Chapter~\ref{sec: conv}.
+We consider the discrete-time embedded process formed by observing the QBD-RAP at changes of level only.
+In Theorem~?
+we prove that the embedded process of the QBD-RAP converges in distribution to the discrete-time embedded process of the fluid queue formed by observing the fluid queue when it crosses the boundaries \(\{y_k\}\).
+In Theorem~?
+we prove that distribution of the sojourn time of the QBD-RAP in a given level converges to distribution of the sojourn time of the fluid queue in the corresponding interval.
+In Theorem~?, we prove that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.
+Initially we set out only to prove Theorem~?, but Theorems~?
+and ?
+appeared as helpful results along the way.
 
-Where appropriate, we also compare the above methods with the discontinuous Galerkin method with a slope-limiter implemented, which avoids oscilations and negative solutions.
-However, recall that, in the context of approximating operators of fluid-fluid queues with these schemes, slope-limiting is not possible, and even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.
+% In this chapter we prove a global convergence result.
+We partition the Laplace tarnsforms with respect to time of the QBD-RAP on the number of level changes.
+Then, by considering the process at changes of level, we are able to work with the discrete-time Markov chain embedded at changes of level.
+First, we show that the transition probabilities of the embedded chain of the QBD-RAP convergence to the transition probabilities of the chain embedded in the fluid queue.
+We show that for a fixed number of level changes, \(n\), say, the embedded process of the QBD-RAP converges to an embedded discrete time process of the fluid queue, embedded at the times when the fluid queue leaves the bands \(\mathcal D_{\ell}\).
+Now, from Chapter~\ref{sec: conv} we have a result on the convergence of the QBD-RAP up to the first change of level (THM) and at the first change of level (COR).
+By time-homogeneity and the strong Markov property of the QBD-RAP then a Corollary to (COR) is the convergence of the transition probabilities of the embedded process of the QBD-RAP to the transition probabilities of the embedded process of the fluid queue.
+Similarly, a Corollary to (THM) is the convegence of the QBD-RAP to the fluid queue between the \(n\)th and \(n+1\)th change of level.
+Putting these Corollaries together, we can than claim a convergence of the distribution of QBD-RAP to the distribution of the fluid queue between the \(n\)th and \(n+1\)th change of level.
+Next, by summing over the partition on the number of level changes, \(n\), and applying the Dominated Convergence Theorem, we can then claim that Laplace transform with respect to time of the fluid queue converges to the Laplace transform with respect to time of the fluid queue.
+Finally, by the Extended Continuity Theorem, we can then claim that the distributions of the QBD-RAP and fluid queue converge for almost all \(t\geq 0\).
 
-The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased.
-For discontinuous Galerkin method, this means increasing the number of basis functions used to approximate the solution within each cell.
-For the QBD-RAP scheme, this means increasing the order of the CME distribution used to constrcut the scheme.
-For the uniformisation scheme, this means dividing each cell into smaller sub-cells over which we approximate the solution.
-As such, each approximation scheme leads to approximations which are matrices of the same size (although not necessarily the same number of non-zero components).
-For example, if we use 3 basis functions in the discontinuous Galerkin method, we approximate the solution by a quadratic on each cell; the comparable QBD-RAP scheme is to use a CME of order 3; and the comparable uniformisation scheme is to divide each cell into 3 sub-cells of width \(\Delta/3\).
+% \begin{enumerate}
+% 	\item Define the embedded process at level changes of the QBD-RAP and the correspoding embedded process of the fluid queue, then use the results of Chapter~\ref{sec: conv} to claim convergence of the embedded processes; that is, convergence at the \(n\)th level change, \(n\geq 1\).
+% 	\item Extend the convergence of the embedded processes to convergence of the Laplace transforms between the \(n\)th and \(n+1\)th change of level, \(n\geq 0\).
+% 	\item Apply the Dominated Convergence Theorem to claim that the Laplace transform with respect to time of the QBD-RAP converges to the Laplace transform with respect to time of the fluid queue. 
+% 	\item Apply the Extended Continuity Theorem for Laplace transforms to claim the the distributions of the processes converge for almost all \(t\geq 0\). 
+% \end{enumerate}
 
-To keep the content of this chapter contained, we do not investigate the performance of the schemes as the cell sizes are decreased, other than for the uniformisation scheme.
-As such, for each numerical experiment, we keep the cell size fixed.
-As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next and link all the cells together.
-We investigate schemes with an upwind flux only.
-We do not investigate the stability of the schemes with respect to the step-size of the time-integration or time-integration scheme (where required).
-Instead, we fix the time-integration step size for each numerical experiment at a suitably small value so as to obey a certain stability criterion (a CFL-like condition, REFs, appendix?).
-For schemes which require us to integrate over time we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (see appendix and REF).
-This scheme claims to introduce no more oscilations into the solution as we integrate over time (REFs).
-Where a slope limiter is implemented, we always implement the \emph{Generalised MUSCL} limiter (REF and Appendix).
-We do not investigate filtering for the DG scheme (REFs).
+% The structure of this chapter is as follows.
+We start by showing weak convergence of the approximation at the \(1\)st and then \(n\)th change of level, \(n> 1\) in Sections~\ref{sec: 1st change} and \ref{sec: nth change}.
+Section~\ref{sec: between n and np1} generalises this to show weak convergence between the \(n\)th and \(n+1\)th change of level.
+Via the Dominated Convergence Theorem (again), Section~\ref{sec: local to global} proves the weak convergence of the QBD-RAP approximation to the fluid queue.
+So as to not obscure the logical flow of the presentation, technical results are reserved for the Appendices.
 
-The performance of the discontinuous Galerkin method has been well-studied (REFs), and it is well-known that the discontinuous Galerkin method performs remarkably well on problems with smooth solutions (REF).
-Here, we mostly focus on investigating the numerical performance of the methods on problems with non-smooth solutions, the purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP method.
-In the stochastic modelling community it is very common to have problems with discontinuous solutions, such as a non-smooth initial conditions (REFs).
-Even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities into the problem.
-There is a very specific set of conditions on the inital density and point masses which must hold for the distribution of the fluid queue to remain smooth as it evolves over time (REFs).Limiting distributions of fluid queues are smooth, however.
-Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues.
+\section{Convergence of an embedded process}\label{sec: 1st change}
+% Before we move on to global convergence we need another result about convergence of the QBD-RAP scheme and the fluid queue at changes of level.
 
-As we shall see, if the solution is smooth, then the Galerkin method is highly-effective and displays rapid convergence to the solution as the number polynomial basis functions is increased.
-However, when the solution is non-smooth oscilations and negative approximations may occur when using the discontinuous Galerkin method, leading to non-sensical solutions.
-In these cases, the QBD-RAP approximation performs relatively well compared to the other positivity preserving schemes investigated.
+% Regarding Remark~\ref{rem: point wies}, we do require convergence at certain boundary points which correspond to changes of level of the QBD-RAP. To this end, we have the following result.
 
-% THE STRUCTURE OF THE CHAPTER IS AS FOLLOWS.
+In this section we consider the embedded process formed by observing the QBD-RAP at changes of level.
+We start by analysing the QBD-RAP at the time of the first change of level, \(\tau_1^{(p)}\).
+Following a similar method to Chapter~\ref{sec: conv} we can prove the following result.
 
-% 1. FUNCTION APPROX./RECONSTRUCTION
+\begin{cor}\label{cor: aln222} For \(\ell,\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\)
+	then, for \(j\in\calS_+\),
+	\begin{align}
+		&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0+1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
+            	 \mid \bs Y^{(p)}(0) = \bs y_0^{(p)}) \nonumber
+	 	%
+		\\&\to \mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0,i))\label{eqn: 1421}
+		 %
+		 %
+		 \intertext{and for \(j\in\calS_-\)}
+		 &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
+            	 \mid \bs Y^{(p)}(0) = \bs y_0^{(p)}) \nonumber
+	 	%
+		\\&\to \mathbb P(\bs X(\tau_1^X) = (y_{\ell_0}, j), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, i)). \nonumber
+		%
+	\end{align}
+	% \begin{align}
+	% 	&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0+1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
+    %         	 \mid L^{(p)}(0) = \ell_{0},\bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \varphi(0) = i) \nonumber
+	%  	%
+	% 	\\&\to \mathbb P(X(\tau_1^X) = y_{\ell_0+1}, \varphi(\tau_1^X) = j, \tau_{1}^X\leq E^\lambda 
+    %         	 \mid X(0) = x_0, \varphi(0) = i)\label{eqn: 1421}
+	% 	 %
+	% 	 %
+	% 	 \intertext{and for \(j\in\calS_-\)}
+	% 	 &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
+    %         	 \mid L^{(p)}(0) = \ell_{0},\bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
+	% 	\varphi(0) = i) \nonumber
+	%  	%
+	% 	\\&\to \mathbb P(X(\tau_1^X) = y_{\ell_0}, \varphi(\tau_1^X) = j, \tau_{1}^X\leq E^\lambda 
+    %         	 \mid X(0) = x_0, \varphi(0) = i). \nonumber
+	% 	%
+	% \end{align}
+	At a boundary, for \(\ell_0\in\{-1,K+1\}\), 
+	\begin{align}
+		&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
+            	 \mid L^{(p)}(0) = \ell_{0},\bs A^{(p)}(0)=1, \varphi(0) = i) \nonumber
+	 	%
+		\\&= \mathbb P(\bs X(\tau_1^X) = (y_\ell, j), \tau_{1}\leq E^\lambda 
+            	 \mid \bs X(0) = (0,i)), \label{eqn: lk78GHJK}
+	\end{align}
+	where \(\ell = 0\), \(i\in\calS_-\cup\calS_{-0}\) and \(j\in\mathcal S_+\) if \(\ell_0=-1\), and \(\ell = K\), \(i\in\calS_+\cup\calS_{+0}\) and \(j\in\calS_-\) if \(\ell_0=K+1\).  
+\end{cor}
+\begin{proof}
+	The proof follows the same structure as the proof of Theorem~\ref{thm: a thm!} however changes are required in all the results used in the proof, as here we do not need to integrate a function \(\psi\).
+Here we only give an outline of the proof.
 
-% 2. 1d wave eqn
+	At a boundary we can model the fluid queue exactly, hence (\ref{eqn: lk78GHJK}) holds.
 
-% 3. STATIONARY DISTRIBUTIONS
-	% i. SIMPLER
-	% ii. MORE COMPLEX
+	Consider first \(i\in \calS_+,j\in\calS_+\).
+Partition the probability (\ref{eqn: 1421}) on the times \(\{\Sigma_n\}_{n\geq 1}\) and \(\{\Gamma_n\}_{n\geq 1}\), and specifically, partition on the event that there are exactly \(m\) events \(\{\Sigma_n\}_{n=1}^m\) and exactly \(m\) events \(\{\Gamma_n\}_{n=1}^m\).
+%Partition further by the phases at these times \(\varphi(\Sigma_n)=j_n\in\calS_-\), \(n=1,\dots,m\) and \(\varphi(\Gamma_n)=k_n\in\calS_+\), \(n=1,...,m\).
+	The resulting partitioned probabilities are
+	\begin{align}
+                 &\int_{x_1=0}^\infty \left(\bs e_i\bs H^{+-}(\lambda,x_1)\otimes \bs a_{\ell_0,i}^{(p)}(x_0)e^{\bs{S}^{(p)}x_1}\bs{D}^{(p)}\right)\wrt x_1 \nonumber
+            	\\&\nonumber \quad \Bigg[\prod_{r=1}^{m-1} \int_{x_{2r}=0}^\infty \left(\bs H^{-+}(\lambda,x_{2r}) \otimes e^{\bs{S}^{(p)}x_{2r}}\bs D^{(p)}\right)\wrt x_{2r} \\&\quad \int_{x_{2r+1}=0}^\infty \left(\bs H^{+-}(\lambda,x_{2r+1}) \otimes e^{\bs{S}^{(p)}x_{2r+1}}\bs D^{(p)}\right) \wrt x_{2r+1}\Bigg] \nonumber
+            	\\&
+            	\quad \int_{x_{2m}=0}^\infty \left(\bs H^{-+}(\lambda, x_{2m}) \otimes e^{\bs{S}^{(p)}x_{2m}}\bs{D}^{(p)}\right) \wrt x_{2m} \nonumber
+				\\&\quad \int_{x_{2m+1}=0}^\infty \left( \bs H^{++}\bs e_j \otimes (\lambda,x_{2m+1}) e^{\bs{S}^{(p)}x_{2m+1}}\bs s^{(p)} \right) \wrt x_{2m+1}.  \label{eqn: prob ofjaiv}
+	\end{align}
 
-% 4. TRANSIENT DISTRIBUTIONS 
-	% i. SIMPLER
-	% ii. MORE COMPLEX
+	To show that the terms (\ref{eqn: prob ofjaiv}) converge to 
+	\begin{align}
+		&\mathbb P(\bs X(\tau_1^X)=(y_{\ell_0+1},j),\tau_1^X\leq E^\lambda, \Sigma_{m}\leq \tau_1^X<\Gamma_{m+1}, \mid \bs X(0)=(x_0, i))
+	\end{align}
+	we can use the bounds from Corollary~\ref{cor: cond bnd 2} and Corollry~\ref{cor: a cor}.
+For \(m=0\) we recognise (\ref{eqn: prob ofjaiv}) as the same form as that appearing in Corollary~\ref{cor: cond bnd 2} upon choosing \(v=0\).
+For \(m\geq 1\), choose the closing operator to be \(\bs v(x)=e^{\bs S x}\bs s\) and set \(x=0\) in Corollary~\ref{cor: a cor}.
+Now take the bound from Corollary~\ref{cor: a cor} and extend it to the case of matrix functions in the same way we extended Corollary~\ref{cor: asjdajaaaaa} to the matrix case in Lemma~\ref{lem: boobies}.
+In this way, we have a bound for (\ref{eqn: prob ofjaiv}) which tends to \(0\) as \(p\to\infty\).
 
-% 5. HITTING TIMES 
-	% i. SIMPLER
-	% ii. MORE COMPLEX
+	% Corollary~\ref{cor: cond bnd} and Lemma~\ref{lem:tttttt} are analogous to Lemma~\ref{lem: Dcoajc} and Corollary~\ref{lem: Dcoajc} used in the proof of Lemma~\ref{lem:vn4}.
+They provide relevant bounds for the difference between terms like (\ref{eqn: prob ofjaiv}) for \(m=0\) and
+	% \begin{align*}
+	% 	&\mathbb P(X(\tau_1^X)=y_{\ell_0+1},\varphi(\tau_1^X)=j,\tau_1^X\leq E^\lambda, \tau_1^X<\Sigma_1 \mid X(0)=x_0, \varphi(0)=i)  
+	% 	\\&= h_{ij}^{++}(\lambda,y_{\ell_0+1}-x_0),
+	% \end{align*}
+	% for example.
 
-% 6. FLUID-FLUID QUEUES
+	% Corollaries~\ref{cor: a cor} and~\ref{cor: aaaaa} are analogous to Corollaries~\ref{eqn: lafkjebjcbbalbvvbrb} and~\ref{cor: fljm7778} used in the proof of Lemma~\ref{lem:vn4}.
+Corollaries~\ref{cor: a cor} and~\ref{cor: aaaaa} give error bounds for the difference between terms such as (\ref{eqn: prob ofjaiv}) for \(m\geq 1\) and the corresponding probability for the fluid queue,
+	% \begin{align}
+	% 	&\mathbb P(X(\tau_1^X)=y_{\ell_0+1},\varphi(\tau_1^X)=j,\tau_1^X\leq E^\lambda, \Sigma_{m}\leq \tau_1^X<\Gamma_{m+1}, \varphi(\Sigma_\ell) = j_\ell, \varphi(\Gamma_\ell)=k_\ell, \nonumber
+	% 	\\&\qquad{} \ell = 1,\dots, m  \mid X(0)=x_0, \varphi(0)=i)
+	% \end{align}
+	% which is given by (\ref{eqn: lst herwe}).
 
-% 7. CONCLUDING REMARKS
+	What reamins is a domination condition so that we may apply the Dominated Convergence Theorem to claim that the sum over the number of up-down and down-up transition converges (i.e.~the sum over \(m\) in (\ref{eqn: prob ofjaiv}) converges).
+To this end, since \(\left[\bs H^{++}(\lambda,x_{m+1})\right]_{ij}\leq G\), then
+	\begin{align}
+			&\int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
+		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
+		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
+		\int_{x_{m+1}=0}^\infty \bs H^{++}(\lambda,x_{m+1}) \bs e 
+		\\&\quad\bs   a_{\ell_0,i}(x_0)e^{\bs{S}x_1}\wrt x_1 \bs{D}e^{\bs{S}x_2} \wrt x_2 \bs{D}\dots e^{\bs{S}x_m} \wrt x_m \bs{D}e^{\bs{S}x_{m+1}}{\bs s} \wrt x_{m+1}\nonumber 
+		%
+			\\&\leq \int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
+		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
+		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
+		\int_{x_{m+1}=0}^\infty \bs e G
+		\\&\quad\bs   a_{\ell_0,i}(x_0)e^{\bs{S}x_1}\wrt x_1 \bs{D}e^{\bs{S}x_2} \wrt x_2 \bs{D}\dots e^{\bs{S}x_m} \wrt x_m \bs{D}e^{\bs{S}x_{m+1}}{\bs s} \wrt x_{m+1}\nonumber
+		%
+		\\&=\int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
+		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
+		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
+		 \bs e G
+		\\&\quad\bs   a_{\ell_0,i}(x_0)e^{\bs{S}x_1}\wrt x_1 \bs{D}e^{\bs{S}x_2} \wrt x_2 \bs{D}\dots e^{\bs{S}x_m} \wrt x_m \bs{D}{\bs e}\nonumber
+		%
+			\\&\leq\int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
+		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
+		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
+		 \bs e G
+		\\&\quad \wrt x_1 \wrt x_2\dots \wrt x_m \label{eqn :ejrvn}
+	\end{align}
+	where the last inequality holds as we claimed in the discussion after (\ref{eqn :NNeeaefjn}) in the proof of Lemma~\ref{lem: gkjljklgagjklagsjlk}.
+Equation~(\ref{eqn :ejrvn}) is of a similar form to (\ref{eqn: bound ggggaaaa}) (they differ by a constant only), hence the same arguments used to bound (\ref{eqn: bound ggggaaaa}) can be applied to get the desired domination result.
 
-\section{Function approximation/reconstruction}
-We start our investigation by looking at how well the methods perform at approximating an initial condition.
-For the discontinuous Galerkin method, this means projecting the solution on to the set of polynomial functions which define the scheme, then reconstructing the solution.
-For the spatially-coherent uniformisation scheme, this means looking at the sub-cell averages.
-For the QBD-RAP scheme, this means computing the initial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}.
-First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}) performs reasonably compared to the others, and we are able to theoretically prove that it converges.
-Therefore we use this reconstruction method throughout the rest of the chapter.
+	Ultimately, we can apply the Dominated Convergence Theorem to prove that the sum of the partitioned probabilities (\ref{eqn: prob ofjaiv}) converges as \(p\to\infty\). The sum of the limits is 
+	\[\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j), \tau_{1}\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0,i)).\]
 
-For this section we consider approximating the initial condition over a single cell of width \(\Delta = 1\).
-To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations.
-Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points over which to compute the KS statistics.
+	 The results for all other cases of \(i,j\in\calS\) follow analogously.
+\end{proof}
 
-First we briefly investigate the possible choices of closing operators as introduced in Section~\ref{sec: closing}.
+Now, let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
+\[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
+be the (stopping) times at which \(\{L^{(p)}(t)\}\), the level process of the QBD-RAP, changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary.
+To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed.
+Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the change-of-level epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
+The subscript \(\bs \alpha^{(p)}\) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\).
+The process \(\{\Ydp(n)\}_{n\geq 0}\) is a discrete-time Markov chain, which is time-homogeneous for \(n\geq 1\).
 
-\subsection{QBD-RAP closing operators}
-The three closing operators we investgate are the following.
-The unnormalised closing vector in Equation~(\ref{eqn: alal}).
-A normalised version of the closing operator in Equation~(\ref{eqn:546}), with the normalised version given by (\ref{eqn: nonlin}).
-We refer to the closing vector in (\ref{eqn: nonlin}) as the \emph{naive normalised} closing vector.
-Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme, though it clearly appears to.
-The third closing vector we investigate is that in Equation~(\ref{eqn:5462}).
+Let \(\{\tau_n^X\}_{n\geq 0}\), be the sequence of (stopping) times with \(\tau_0^X=0\), and 
+\[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(0)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(0)=y_{K+1}\right\} \end{array} \right\}.\]
+For \(n\geq 1\), \(\tau_n^X\) is the time at which \(X(t)\) either changes band, or hits a boundary, or the process leaves a boundary, for the \(n\)th time.
+The embedded process \(\{\bs X(\tau_n)\}\) is a discrete-time Markov chain which is time-homogeneous for \(n\geq 1\).
 
-First consider approximating the initial condition with density function \(1(x<0.5)\).
-In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation.
-Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders.
-It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interferes with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme.
-Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection.
-The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward).
-Intuitively, there should be very little mass in the tail of the distribution as we are using concentrated matrix exponential distributions in the reconstruction.
+We have the following result on the convergence of the embedded processes \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\), which we will utilise later to prove a global result.
+\begin{cor}\label{cor: aln222222} For \(\ell,\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
+	for \(n=0\), then
+	\begin{align}
+		&\mathbb P(\Ydp(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j), \tau_1^{(p)}\leq E^\lambda
+            	 \mid \bs Y^{(p)}(0) = (\ell_0,\bs a_{\ell_0,i}^{(p)}(x_0),i)) \nonumber
+	 	%
+		\\&\to  
+			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_1^{X}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\label{eqn: lim 1}
+	\end{align}
+	and for \(n\geq 1\), then, 
+	\begin{align}
+		&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j), \tau_{n+1}^{(p)}\leq E^\lambda
+            	 \mid \Ydp(n) = (\ell_0,i), \tau_n^{(p)}\leq E^\lambda) \nonumber
+	 	%
+		\\&\to  
+			\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd}
+	\end{align}
+\end{cor}
+\begin{proof}
+The case for \(n=0\) is a direct result of Corollary~\ref{cor: aln222}.
+% For the case with \(n=0\), since \(\tau_1^{(p)}<\infty\) almost surely for all \(p\geq 1\), as is \(\tau_1^X\), then upon taking \(\lambda \to 0\) in the results in Corollary~\ref{cor: aln222} we get the convergence in (\ref{eqn: lim 1}).
 
-If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs}, then we see that the unnormalised reconstruction performs the poorest.
-Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests naive normalised and normalised reconstruction are very similar.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(2x1(x<0.5)+1(x\geq 0.5)\), and the approximations (left) and \(L^2\) error between the true PDF, \(2\times 1(x<0.5)\), and the approximations (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
-	\label{fig: fun 2 ks error qbdrap closing vecs}
-\end{figure}
+For \(n\geq 1\), consider the transition probabilities of the embedded process from the QBD-RAP, 
+\begin{align}
+	&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j),\tau_{n+1}^{(p)}\leq E^\lambda
+	\mid \Ydp(n) = (\ell_0,i),\tau_{n}^{(p)}\leq E^\lambda) \nonumber 
+	\\&=\mathbb P(\Ydp(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+), j), \tau_{1}^{(p)}\leq E^\lambda
+	\mid \bs Y^{(p)}(0)=(\ell_0 ,\bs\alpha, i)) \label{eqn: aaannnn}
+\end{align}
+% since the process is time-homogeneous and the exponential random variable \(E^\lambda\) is memoryless. The probability on the right of (\ref{eqn: aaannnn}) is equal to
+% \begin{align}
+% 	&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1(j\in\calS_-)+1(j\in\calS_+), \varphi(\tau_1^{(p)}) = j
+% 			 \mid \bs Y^{(p)}(0) = (\ell_0 ,\bs\alpha, i)).\nonumber
+% 	%  \intertext{for \(j\in\calS_+\)}
+% 	%  &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \varphi(\tau_1^{(p)}) = j
+% 	% 		 \mid \bs Y^{(p)}(0) = \bs y_0^{(p)}) \nonumber
+% \end{align}
+% Thus, once again noting that \(\tau_1^{(p)}<\infty\) almost surely, then taking \(\lambda \to 0\) in 
+Now, applying Corollary~\ref{cor: aln222} to (\ref{eqn: aaannnn}) states that 
+\begin{align}
+	&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j),\tau_{n+1}^{(p)}\leq E^\lambda
+	\mid \Ydp(n) = (\ell_0,i),\tau_{n}^{(p)}\leq E^\lambda) \nonumber 
+	\\&\to \mathbb P(\bs X(\tau_{1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(\tau_{1}^X) = (y_{\ell_0+1(i\in\calS_-)},i)).\label{eqnL sdfdfaaa}
+\end{align}
+Since the fluid queue is time-homogeneous, and \(E^\lambda\) memoryless, then 
+\begin{align}
+	&\mathbb P(\bs X(\tau_{1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(\tau_{1}^X) = (y_{\ell_0+1(i\in\calS_-)},i))
+	\\&=\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_{n}^{X}\leq E^\lambda ),
+\end{align}
+which prove the result. 
+\end{proof}
+A direct corollary of Corollary~\ref{cor: aln222222} is the convergence of the transition probabilties of the embedded process. 
+\begin{cor}\label{cor: aln22222212} For \(\ell,\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
+		for \(n=0\), then
+		\begin{align}
+			&\mathbb P(\Ydp(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j) 
+					 \mid \bs Y^{(p)}(0) = (\ell_0,\bs a_{\ell_0,i}^{(p)}(x_0),i)) \nonumber
+			 %
+			\\&\to  
+				\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1(j\in\calS_+)}, j)  \mid \bs X(0) = (x_0,i)).\label{eqn: lim 2}
+		\end{align}
+		and for \(n\geq 1\), then, 
+		\begin{align}
+			&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j) 
+					 \mid \Ydp(n) = (\ell_0,i) ) \nonumber
+			 %
+			\\&\to  
+				\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j) \mid \bs X(\tau_{n}) = (y_{\ell_0+1(i\in\calS_-)},i) ).\label{eqn: ddddddd2}
+		\end{align}
+\end{cor}
+\begin{proof}
+	Since \(\tau_1^{(p)}<\infty\) almost surely, as is \(\tau_1^X\), then taking \(\lambda \to 0\) in Corollary~\ref{cor: aln222222} yields the result. 
+\end{proof}
+Corollary~\ref{cor: aln22222212} states that the transition probabilities of the embedded processes converge.
+Thus the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the distribution of \(\{\Ydp(n)\}\) is tight.
+Thus we can establish a convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).
 
-Now consider approximating the initial density \(f(x)=1\).
-Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error, and \(L^2\) error between the true initial distribution and the reconstructions, we now see that the normalised reconstructions outperform the unnormalised reconstruction.
-This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(x\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
-\end{figure}
+Another direct corollary of Corollary~\ref{cor: aln222} is the convergence in distribution of the random variables \(\{\tau_1^{(p)}\}_{p\geq 1}\) to \(\tau_1^X\). 
+\begin{cor}
+	The random variables \(\{\tau_1^{(p)}\}_{p\geq 1}\) converge in distribution to \(\tau_1^X\). 
+\end{cor}
+\begin{proof}
+	By Corollary~\ref{cor: aln222} the probabilities 
+	\begin{align}
+		&\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j),\tau_1^{(p)}\leq E^\lambda
+            	 \mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
+	 	%
+		\\&\to  
+			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1(j\in\calS_+)}, j),\tau_1^{(p)}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\nonumber
+	\end{align}
+	By the law of total probability and the convergence above, 
+	\begin{align}
+		&\mathbb P( \tau_1^{(p)}\leq E^\lambda
+		\mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
+		\\&=\sum_{\ell\in\{\ell_0-1,\ell_0+1\}\cap\mathcal K}\sum_{j\in\calS}\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell,j),\tau_1^{(p)}\leq E^\lambda
+		\mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
+	 	%
+		\\&\to  
+		\sum_{\ell\in\{\ell_0,\ell_0+1\}\cap\{0,1,\dots,K+1\}}\sum_{j\in\calS}\mathbb P(\bs X(\tau_1^X) = (y_{\ell}, j),\tau_1^{(p)}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\nonumber
+		\\&=\mathbb P(\tau^X\leq E^\lambda \mid \bs X(0) = (x_0,i) ).
+	\end{align}
+	Thus, the Laplace transform of \(\tau_1^{(p)}\) converges to the Laplace transform of \(\tau_1^X\).
+By the Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957}, then \(\{\tau_1^{(p)}\}\) converges in distribution to \(\tau_1^X\).
+\end{proof}
 
-Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised and normalised closing operators for order 1, 3, 5 and 7, as well as the true PDF.
-Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots and for all orders.
-This feature is due to a significant amount of mass being lost due to the truncation.
-In comparison the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator.
-In the reconstruction we also need to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}.
-Here we suppose that the phase is positive, so the `folding' around \(\Delta\) in the normalised closing operators appears appears at the left of the plots.
-In general, reconstructions via the unnormalised closing operator typically underestimate the value of the function in this region.
+\section{A global result}\label{sec: local to global}
+The rest of this Chapter is dedicated to proving global convergence results of the QBD-RAP approximation to the fluid queue.
+The structure of the argument is to first show a convergence result of the Laplace transform with respect to time of the distribution of the QBD-RAP and fluid queue and the \(n\)th change of level.
+We then prove a convergence result for the Laplace trasforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue between the \(n\)th and \(n+1\)th change of level.
+Summing over the number of changes of level, \(n\), and via the Dominated Convergence Theorem, we claim that the Laplace trasforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue converge.
+Lastly, we apply the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} to claim a weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
 
-Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval.
-This is a common feature, and drawback, of the QBD-RAP method.
-Perhaps there is a different reconstruction method or yet another closing operator which could alleviate this issue.
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=1\) (orange dashed line).}
-	\label{fig: pdf reconstructed}
-\end{figure}
+\subsection{At the \(n\)th change of level}\label{sec: nth change}
+% So far we have shown error bounds for the QBD-RAP approximation on the event that the fluid queue and QBD-RAP remain in the same band/level (%Corollary~\ref{cor: lst diff} and 
+% Theorem~\ref{thm: a thm!}) and also immediately upon exiting the band (Corollary~\ref{cor: aln222}).
+In the next three subsections we extend this to global convergence of the approximation.
+To do so, we partition the approximation and the fluid queue by the number of level changes.
+The local convergence on a given level, which we just proved, is then used to claim that each term in the partition converges.
+What remains is to argue that we can swap a limit and countable sum which we do by the Dominated Convergence Theorem.
 
-Next we consider approximating the initial distribution with density \(-6x^2+6x\).
-Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
-However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, we see that with this metric, then unnormalised closing operator performs better than the two normalised ones for orders 5 and above.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(-2x^3+3x^2\), and the approximation (left) and \(L^2\) error between the true PDF \(-6x^2+6x\) and the approximation (right), for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
-	\label{fig: fun 6 ks error qbdrap closing vecs}
-\end{figure}
+% Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
+% \[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
+% be the (stopping) times at which \(\{L^{(p)}(t)\}\), the level process of the QBD-RAP, changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary.
+To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed.
+Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi^{(p)}(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the value of the level and phase of the discrete-time process embedded at the change-of-level epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
+The subscript \(\bs \alpha^{(p)}\) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\).
+The process \(\{\Ydp(n)\}_{n\geq 0}\) is a discrete-time Markov chain, which is time-homogeneous for \(n\geq 1\).
 
-The fact that the unnormalised closing operator outperforms the two normalised ones can be explain, once again, by the fact that, for the normalised operators we have `folded' the ME density function back on itself.
-In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\).
-In Figure~\ref{fig: pdf reconstructed quadratic} we observe that, at the left of the plot, the normalised reconstructions over estimate the density function in this region, where as the unnormalised reconstruction looks to be doing better.
-Recall that the `folding' in the normalised closing operator manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions.
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{chapter5/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=-6x^2+6x\) (orange dashed line).}
-	\label{fig: pdf reconstructed quadratic}
-\end{figure}
+% % The process \(\{L^{(p)}(\tau_n^{(p)})\}_{n\geq 0}\) is the level process of the (continuous-time) QBD-RAP process observed at changes of level.
+When \(\tau_n^{(p)}\) is a change of level, or the time when the process leaves a boundary, the value of orbit process at these times is \(\bs A^{(p)}(\tau_n^{(p)}) = \bs \alpha^{(p)}\).
+At time \(\tau_0^{(p)} =0\), \(\bs A^{(p)}(\tau_0^{(p)})=\bs A^{(p)}(0) = \bs   a_{\ell_0,i}^{(p)}(x_0)\).
+The process \(\{\Ydp(t)\} = \{(L^{(p)}(\tau_n^{(p)}),\phi^{(p)}(\tau_n^{(p)}))\}_{n\geq 0}\) is a time-inhomogeneous discrete-time QBD.
+Further, the process \(\{\phi^{(p)}(\tau_n^{(p)})\}_{n\geq 0}\) is a time-inhomogeneous discrete-time Markov chain on the state space \(\mathcal S\).
+For \(\bs   a_{\ell_0,i}^{(p)}(x_0) = \bs \alpha^{(p)}\), or for index \(n\geq 1\), then \(\{L^{(p)}(\tau_n^{(p)}),\phi^{(p)}(\tau_n^{(p)})\}_{n\geq 1}\), and \(\{\phi^{(p)}(\tau_n^{(p)})\}_{n\geq 1}\) are time-homogenous.
 
-Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\).
-This density function is at a maximum at the left of the region.
-Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case.
-Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that the normalised reconstructions perform relatively well compared to the unnormalised reconstruction as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF).
-If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \((1-e^{-3x})/(1-e^{-3})\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
-	\label{fig: fun 7 ks error qbdrap closing vecs}
-\end{figure}
+% Let \(\{\tau_n^X\}_{n\geq 0}\), be the sequence of (stopping) times with \(\tau_0^X=0\), and 
+% \[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(0)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(0)=y_{K+1}\right\} \end{array} \right\}.\]
+% For \(n\geq 1\), \(\tau_n^X\) is the time at which \(X(t)\) either changes band, or hits a boundary, or the process leaves a boundary, for the \(n\)th time.
 
-\subsection{Comparison of methods}
-Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.
+For \(n\geq 1\), consider the Laplace transform 
+\begin{align}
+	&\int_{t=0}^\infty e^{-\lambda t} \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\in \wrt t
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \wrt t \nonumber 
+	 %
+	 \\&=\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^{\lambda}
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}),\nonumber 
+	%
+\end{align}
+% \begin{align}
+% 	&\int_{t=0}^\infty e^{-\lambda t} \mathbb P(L^{(p)}(\tau_n^{(p)}) = \ell, \phi^{(p)}(\tau_n^{(p)}) = j_n, \tau_{n}^{(p)}\in \wrt t
+% 	 \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
+% 	 \\&\qquad{} \phi^{(p)}(0)=i) \wrt t \nonumber 
+% 	 %
+% 	 \\&=\mathbb P(L^{(p)}(\tau_n^{(p)}) = \ell, \phi^{(p)}(\tau_n^{(p)}) = j_n, \tau_{n}^{(p)}\leq E^{\lambda}
+% 	 \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
+% 	 \\&\qquad{} \phi^{(p)}(0)=i),\nonumber 
+% 	%
+% \end{align}
+which is the Laplace transform of the time until the \(n\)th change of level of the QBD-RAP on the event that the level and phase at the \(n\)th change of level are \(\ell\) and \(j_n\), respectively, given that the initial level and phases are \(\ell_0\) and \(i\), respectively and the initial orbit is \(\bs a_{\ell_0,i}^{(p)}(x_0)\).
+Partitioning on the time of the first change of level, \(\tau_1\), and the level and phase at this time gives
+\begin{align}
+	&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \mid \Ydp(1)=(\ell_1, j_1), \tau_1^{(p)}\leq E^\lambda ) \nonumber
+	 %
+	 \\&\times \mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
+	 \mid \bs Y(0)=\bs y_0^{(p)}).
+	%  \\&\times \mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
+	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
+	%  \\&\qquad{} \phi^{(p)}(0)=i).
+	 \label{eqn: mnebrb2}
+\end{align}
+An application of Corollary \ref{cor: aln222} to the expression on the second line of (\ref{eqn: mnebrb2}) states, for \(i\in\calS\), \(j\in\calS\), \(\ell_0\in\mathcal K\), 
+\begin{align}
+	&\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \nonumber
+	%  &\lim_{p\to\infty}\mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
+	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
+	%  \\&\qquad{} \phi^{(p)}(0)=i) \nonumber
+	 %
+	 	\\&\to \begin{cases}
+			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j_1), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, i)) 
+	 %
+	 &\text{ for }j_1\in\calS_+,\, \ell_1=\ell_0+1\\ 
+	%  &\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
+	%  \mid \bs Y^{(p)}(0)=(\ell_0, \bs  a_{\ell_0,i}^{(p)}(x_0), i)) \nonumber
+	%  &\lim_{p\to\infty}\mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
+	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
+	%  \\&\qquad{} \phi^{(p)}(0)=i) \nonumber
+	 %
+	 	\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0}, j_1), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, i)) & \text{for }j_1\in\calS_-,\, \ell_1=\ell_0-1. 
+\end{cases}\nonumber 
+\end{align}
 
-\section{Travelling wave}
-Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with varying initial conditions.
-Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\).
-The PDE (if/when it exists) which describes this system is
-\[\cfrac{\partial}{\partial t} f(x,t) = \cfrac{\partial}{\partial x} f(x,t),\]
-where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
-\[f(x,t) = f(x-t,0)\]
-so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).
+Now, for a given \(j_1\) and \(\ell_1\) consider 
+\begin{align}
+	&\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \mid \Ydp(1)=(\ell_1, j_1), \tau_1^{(p)}\leq E^\lambda) \nonumber 
+	 % 
+	\\&=\mathbb P(\Ydp(n-1) = (\ell, j_n), \tau_{n-1}^{(p)}\leq E^\lambda 
+	 \mid \bs Y(0)=(\ell_1,\bs \alpha, j_1))\label{eqn: dskvnaSF}
+\end{align}
+by the time-homogeneous property of the QBD-RAP and the the memoryless property of the exponential distribution.
+The left-hand side of Equation (\ref{eqn: dskvnaSF}) appears as the first factor in the summands of (\ref{eqn: mnebrb2}).
 
-We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space. We discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We then
+Let 
+\begin{align}\label{eqn: paths set1}
+	\mathcal P^n(\ell_0,\ell_n)&=\{(\ell_1,\dots,\ell_{n-1}) \in \mathcal K^{n-1}\mid |\ell_{r-1}-\ell_r|=1,r = 1,\dots,n\}.
+\end{align}
+The set \(\mathcal P^n(\ell_0,\ell)\) contains all of the possible sequences of levels which \(\{L(t)\}\) or \(\{X(t)\}\) may visit on a sample path which starts in level \(\ell_0\), ends in level \(\ell\) and changes level \(n\) times.
+By partitioning on the times \(\tau_m\), \(m=2,\dots,n-1\), and the phases and the levels at these times and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is
+	\begin{align}
+	 &\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}(\ell_1,\ell)}}\prod_{m=2}^{n}\mathbb P(\Ydp(1) =(\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
+            	 \mid  
+	 	\bs Y(0) = (\ell_{m-1},\bs \alpha, j_{m-1})),  \label{eqn: 161222}
+\end{align}
+where we define \(\ell_n=\ell\). %Each factor in the product is 
+% \begin{align}
+% 	&\mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda 
+%             	 \mid \Ydp(m-1) = (\ell_{m-1}, \nonumber
+% 	 	 j_{m-1}), \tau_{m-1}^{(p)}\leq E^\lambda) 
+% 	\\&=\mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda 
+%             	 \mid \Ydp(0) = (\ell_{m-1}, j_{m-1}), \tau_{m-1^{(p)}}=0)\nonumber
+% 	\\&=\mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
+%             	 \mid \Ydp(0) = (\ell_{m-1}, j_{m-1}))\label{eqn: kk}
+% \end{align}
+% by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution. 
+We can apply Corollary (\ref{cor: aln222}) to the terms (\ref{eqn: 161222}) and conclude 
+\begin{align}
+	& \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
+            	 \mid \bs Y(0) = (\ell_{m-1},\bs \alpha, j_{m-1}))\nonumber
+	%
+	\\&\to\begin{cases}
+	 	\mathbb P(\bs X(\tau_1^X) =( y_{\ell_{m-1}+1},j_m), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, j_{m-1}))  
+	 &\text{ for }\begin{cases} j_m\in\calS_+,\\ \ell_m=\ell_{m-1}+1,\end{cases}\\
+	 %
+	 %
+	%  & \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
+    %         	 \mid \bs Y(0) = (\ell_{m-1},\bs \alpha, j_{m-1}))\nonumber
+	% %
+	% \\&\to
+	 	\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}}, j_m), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, j_{m-1}))  & \text{ for }\begin{cases}j_m\in\calS_-,&\\ \ell_m=\ell_{m-1}-1.&\end{cases}
+\end{cases}\label{eqn:kknvf fawoprgj v2}
+\end{align}
+
+By the time-homogeneous property of the fluid queue and the memoryless property of the exponential distribution, 
+\begin{align}
+	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
+	%  \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{m}^X\leq E^\lambda 
+    %         	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}))\nonumber 
+	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{m}^X\leq E^\lambda 
+            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \nonumber 
+	\tau_{m-1}^X\leq E^{\lambda}).
+\end{align}
+Similarly, 
+\begin{align}
+	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}}, j_m), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
+	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}}, j_m), \tau_{m}^X\leq E^\lambda 
+            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda).
+\end{align}
+
+Now, taking the limit of (\ref{eqn: 161222}) gives 
+\begin{align}
+	&\lim_{p\to\infty}\sum_{j_2,\dots,j_{n-1}\in\mathcal S}\sum_{(\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}(\ell_1,\ell)}\prod_{m=2}^{n}\mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda \nonumber
+            	 \\&\qquad\mid \Ydp(m-1) = (\ell_{m-1}, j_{m-1}), \tau_{m-1}^{(p)}\leq E^\lambda) \nonumber 
+	%
+	\\ &=\sum_{j_2,\dots,j_{n-1}\in\mathcal S}\sum_{(\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}(\ell_1,\ell)}\prod_{m=2}^{n}\lim_{p\to\infty} \mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda \nonumber
+            	 \\&\qquad \mid \Ydp(m-1) = (\ell_{m-1}, 
+	 	 j_{m-1}), \tau_{m-1}^{(p)}\leq E^\lambda),
+	 \label{eqn: 161222b}
+\end{align}
+where we may swap the limit and the sums as they are finite, and we can swap the limit and the product since all the limits exist and the product is finite.
+Substituting the limits into (\ref{eqn: 161222b}) gives
+\begin{align}
+	&\sum_{j_2,\dots,j_{n-1}\in\mathcal S}\sum_{(\ell_2,\dots,\ell_{n-1})  \in\mathcal P^{n-1}(\ell_1,\ell)} \prod_{m=2}^{n}\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1(j_{m}\in\mathcal S_+)}, j_m), \tau_{m}^X\leq E^\lambda \nonumber
+            	\\&\qquad \mid \bs X(\tau_{m-1}^X)=(y_{\ell_{m-2}+1(j_{m-1}\in\mathcal S_+)},j_{m-1}),\tau_{m-1}^X\leq E^\lambda),\nonumber
+		%
+		\\&= \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
+		j_n), \tau_{n}^X\leq E^\lambda \mid \bs X(0)=(y_{\ell_{0}+1(j_{1}\in\mathcal S_+)},
+		j_{1}), 
+		\tau_1^X\leq E^\lambda).
+\end{align}
+
+Therefore, taking the limit as \(p\to \infty \) of (\ref{eqn: mnebrb2}) 
+\begin{align}
+	&\lim_{p\to\infty}\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \mid \Ydp(1)=(\ell_1,j_1), \tau_{1}^{(p)}\leq E^{\lambda}) \nonumber 
+	 %
+	 \\& \quad\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \nonumber
+	 %
+	 \\&= \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K} \lim_{p\to\infty} \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \mid \Ydp(1)=(\ell_1, j_1), \tau_1^{(p)}\leq E^\lambda) \nonumber
+	 %
+	 \\& \quad\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)})\nonumber
+	 %%
+	 %%
+	 \\&= \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K} \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
+		j_n), \tau_{n}^X\leq E^\lambda \mid \nonumber
+		\bs X(\tau_1^X)=(y_{\ell_{0}+1(j_{1}\in\mathcal S_+)},
+		j_{1}),
+		\\&\qquad{}\tau_1^X\leq E^\lambda) \nonumber
+	 %
+	 \mathbb P(\bs X(\tau_1^X)=(y_{\ell_{0}+1(j_{1}\in\mathcal S_+)},
+		j_{1}),\tau_1^X\leq E^\lambda
+		\mid \bs X(0)=(x_0, i)) \nonumber
+	%
+	\\&=\mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
+		j_n), \tau_{n}^X\leq E^\lambda
+		\mid \bs X(0)=(x_0, i))
+	 \label{eqn: mnebrb22}
+\end{align}
+where the swapping of limits and sums is justified as the sums are finite, and swapping limits and products is justified as the product is finite and all limits exist.
+
+Therefore we have proved the following result 
+\begin{lem}\label{lem: kKKJJJF}
+	For all \(\ell,\ell_0\in\mathcal K,\) \(i,j_n\in\mathcal S\), \(x_0\in\mathcal D_{\ell_0,i}\), \(n\geq 1\), then, as \(p\to\infty\),
+	\begin{align}
+		&\mathbb P(\Ydp(n) = (\ell,j_n), \tau_{n}^{(p)}\leq E^{\lambda}
+		 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}),\nonumber 
+		%  &\mathbb P(L^{(p)}(\tau_n^{(p)}) = \ell, \phi^{(p)}(\tau_n^{(p)}) = j_n, \tau_{n}^{(p)}\leq E^{\lambda}
+		%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \phi^{(p)}(0)=i),\nonumber  
+		 %
+		 \\&\to \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
+		j_n), \tau_{n}^X\leq E^\lambda
+		\mid \bs X(0)=(x_0, i)).
+	\end{align}
+\end{lem}
+
+\subsection{Between the \(n\)th and \(n+1\)th change of level}\label{sec: between n and np1}
+Let \(\mathcal T_n^{(p)} = (\tau_{n}^{(p)},\tau_{n+1}^{(p)}]\) and \(\mathcal T_n^X=(\tau_n^X,\tau_{n+1}^X]\) be the interval of time between the \(n\)th and \(n+1\)th change of level of the QBD-RAP and fluid queue, respectively.
+Consider the Laplace transform
+\begin{align}
+	\nonumber&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid 
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\psi(x)\wrt t.
+	% \nonumber&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_\ell}\mathbb P(L^{(p)}(t) = \ell, \widetilde X^{(p)}(t)\in\wrt x, \phi^{(p)}(t) = j, \tau_{n}^{(p)}\leq t<\tau_{n+1}^{(p)} \mid L^{(p)}(0)=\ell_0,
+	% 	\\&\qquad \bs A^{(p)}(0)=\bs a_{\ell_0,i}^{(p)}(x_0), \phi^{(p)}(0)=i)\psi(x)\wrt t.
+\end{align}
+Partitioning on the time of the \(n\)th change of level and the phase and level at this time gives
+\begin{align}
+	&\int_{t=0}^\infty e^{-\lambda t} \int_{u_n=0}^t\sum_{j_n\in\mathcal S}
+	\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
+	t \in(u_n, \tau_{n+1}^{(p)}] \mid\Ydp(n) = (\ell,j_n), \nonumber 
+	\\&\quad \tau_{n}^{(p)}= u_n)\psi(x) \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\in \wrt u_n 
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) 
+	  \wrt t \nonumber 
+	 %
+	 \\&= \sum_{j_n\in\mathcal S}
+	\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), 
+	t\in\mathcal T_{0}^{(p)} \mid\nonumber 
+	  \bs Y(0) = (\ell,\bs\alpha,j_n))\wrt t \psi(x)  
+	  \\&\quad{} \mathbb P(\Ydp(n) = (\ell,j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \label{eqn: l whkrqvlkjbrd}
+\end{align}
+by the time homogenous property of the QBD-RAP, the memoryless property of the exponential distribution, and the convolution theorem of Laplace transforms.
+The swap of integrals and sums is justified by the Fubini-Tonelli Theorem.
+We recognise the probability
+\begin{align}
+	\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \label{eqn: Kkk nj aa w}
+\end{align}
+as that appearing in Lemma \ref{lem: kKKJJJF}. Hence (\ref{eqn: Kkk nj aa w}) converges to 
+\begin{align}
+	\mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
+		j_n), \tau_{n}^X\leq E^\lambda
+		\mid \bs X(0)=(x_0, i)) \label{eqn: ashverkjbvjhkhdsknacva}
+\end{align}
+as \(p\to\infty\).
+
+Now consider the expression 
+\begin{align}
+	&\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j),  
+	t\in\mathcal T_0^{(p)} \mid \bs Y(0) = (\ell,\bs\alpha, j_n)
+	 	 )\wrt t  \psi(x)  \label{eqn: vajkl;vJK:GSKJ}
+\end{align}
+which appears as part of (\ref{eqn: l whkrqvlkjbrd}). We can rewrite (\ref{eqn: vajkl;vJK:GSKJ}) as  
+\begin{align}
+	 &\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
+	t\in\mathcal T_0^{(p)}\mid \Ydp(0) = (\ell,j_n))\wrt t \psi(x) \nonumber 
+	 \\&= \int_{x\in\calD_{\ell,j}} \widehat f^{\ell,(p)} (\lambda)(x,j;x_0,j_n) \psi(x)\wrt x \label{eqn: Jk jJ KK}
+\end{align}
+Applying Theorem~\ref{thm: a thm!}, then (\ref{eqn: Jk jJ KK}) converges to 
+\begin{align}
+	&\int_{x\in\calD_{\ell,j}} \widehat \mu^{\ell} (\lambda)(x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x)\wrt x \nonumber 
+	\\&= \int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_0^X \mid \bs X(0) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n) ) \psi(x).
+\label{eqn: lkHAHJFHKJ J J H}
+\end{align}
+Since the fluid queue is time-homogeneous and \(E^\lambda\) memoryless, then we can write (\ref{eqn: lkHAHJFHKJ J J H}) as 
+\begin{align}
+	&\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x,j), t\in\mathcal T_n^{X} \mid  
+	\bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)},j_n), \tau_n^X\leq E^\lambda ) \psi(x)  \label{eqn: lkHAHJFHKJ J J 22}
+\end{align}
+Therefore, we have shown (\ref{eqn: vajkl;vJK:GSKJ}) converges to (\ref{eqn: lkHAHJFHKJ J J 22}) as \(p\to\infty\).
+
+Returning to the right-hand side of (\ref{eqn: l whkrqvlkjbrd}) and taking the limit as \(p\to\infty\), 
+\begin{align}
+	&\lim_{p\to\infty} \sum_{j_n\in\mathcal S}
+	\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), 
+	t\in\mathcal T_0^{(p)} \mid \nonumber 
+	 \bs Y(0) = (\ell,\bs\alpha,j_n))\wrt t 
+	 \\&\qquad{}\psi(x)  \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}).\label{eqn: eieieie}
+	 %
+	%  \\&= \sum_{j_n\in\mathcal S}
+	% \lim_{p\to\infty} \int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t}  \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
+	% t\in\mathcal T_n^{(p)} \mid \nonumber 
+	%  \Ydp(n) = (\ell,j_n), 
+	%   \tau_{n}^{(p)}= 0)\wrt t \\&\qquad{} \psi(x) 
+	%  \lim_{p\to\infty} \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	%  \mid \bs Y^{(p)}(0)=(\ell_0, \bs  a_{\ell_0,i}^{(p)}(x_0),i))\nonumber
+\end{align}
+Since the sum is finite and the limits exist then substitute in the limiting values, (\ref{eqn: lkHAHJFHKJ J J 22}), and (\ref{eqn: ashverkjbvjhkhdsknacva}), to get 
+\begin{align}
+	 &\sum_{j_n\in\mathcal S}\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} \mid \bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n),\nonumber
+	 \tau_n^X\leq E^\lambda ) \psi(x)
+	 \\&\quad{} \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
+		j_n), \tau_{n}^X\leq E^\lambda
+		\mid \bs X(0)=(x_0,i))\nonumber
+	%
+	\\&= \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} 
+	\mid X(0)=x_0, \varphi(0)=i)\psi(x)\wrt t.
+\end{align}
+
+Hence we have shown the following result 
+\begin{lem}\label{lem: LAkAKFnvnb mav h}
+	For \(\ell,\ell_0\in\mathcal K\), \(i,j\in\calS\), \(n\geq 0\), then, as \(p\to\infty\), 
+	\begin{align} 
+		&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), t\in\mathcal T_n^{(p)} \mid  \nonumber 
+		\bs Y(0) = \bs y_0)\psi(x)\wrt t \nonumber
+		%
+		\\&\to\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j_n}}\mathbb P(\bs X(t)\in(\wrt x, j_n), t\in\mathcal T_n^{X} 
+		\mid \bs X(0)=(x_0,i))\psi(x)\wrt x\wrt t.
+	\end{align}
+\end{lem}
+
+\subsection{Another domination condition}
+Our aim is to prove convergence of 
+\begin{align} 
+	&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j) \mid  \nonumber 
+	\bs Y(0) = \bs y_0)\psi(x)\wrt t \nonumber
+	\\&=\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), t\in\mathcal T_n^{(p)} \mid  \nonumber 
+	\bs Y(0) = \bs y_0)\psi(x)\wrt t \nonumber
+	\\&=\sum_{n=0}^\infty\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), t\in\mathcal T_n^{(p)} \mid   
+	\bs Y(0) = \bs y_0)\psi(x)\wrt t \label{eqn:GHJKLKJHGHJKL}
+\end{align}
+where the swap of the intergals and sums is justified by the Fubini-Tonelli Theorem.
+Now, (\ref{eqn:GHJKLKJHGHJKL}) is an infinite sum of terms appearing in Lemma~\ref{lem: LAkAKFnvnb mav h}.
+Hence, upon taking the limit of (\ref{eqn:GHJKLKJHGHJKL}), if we can justify the swap of the sum and the limit, we will obstain the result we desire.
+The main result of this subsection is Lemma~\ref{lem: another bound 2} in which we show a bound which can serve as a dominating function so that we may apply the Dominated Convergence Theorem to justify the swap.
+
+% To extend Lemma~\ref{lem: LAkAKFnvnb mav h} to show a global convergence result we use the Dominated Convergence Theorem.
+Here we show some geometric bounds on the probability of \(n\) level changes of the QBD-RAP and fluid queue, which ultimately serve as a dominating function in the Dominated Convergence Theorem.
+
+%Let \(c_{min} = \min\limits_{i\in\mathcal S_+\cup \mathcal S_-}|c_i|\).
+
+\begin{lem}\label{lem: another bound}
+	For all \(i\in\mathcal S_+\cup \calS_-\), and \(n\geq 2\),
+	\begin{align}
+		&\mathbb P(\tau_n\leq E^\lambda \mid \phi(\tau_{n-1})=i, \tau_{n-1}\leq  E^\lambda ) \leq b,
+	\end{align}
+	where 
+	\[b = \min\left\{1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_1|,\cfrac{q}{q+\lambda}\right\}\]
+	and  
+	\[|r_1|\leq 2G\cfrac{\var \left(Z\right)}{\varepsilon^2} + 2L\varepsilon.\]
+\end{lem}
+Note that \(b\) and \(r_1\) depend on \(p\) which has been suppressed to simplify notation. When explicitly needed, we use a superscript \(p\) to denote this dependence.  
+\begin{proof}
+	For the QBD-RAP, changes of level can only occur when \(i\in\mathcal S_+\cup\calS_-\).
+
+	Suppose that the phase at time \(\tau_{n-1}\) is \(i\in\mathcal S_+\) and that at time \(\tau_{n-1}\) the QBD-RAP is not at a boundary.
+The arguments for an initial phase \(i\in\mathcal S_-\) are analogous.
+For the QBD-RAP to change level or hit a boundary one of two things must happen, either;
+	\begin{enumerate}
+		\item the fluid remains in phase \(i\) until there is a change of level or boundary hit, or
+		\item the fluid changes phase before there is a change of level or a boundary is hit. 
+	\end{enumerate}
+
+	Hence, for sample paths which contribute to the Laplace transform, one of two things must happen, either; 
+	\begin{enumerate}
+		\item the fluid remains in phase \(i\) until there is a change of level or a boundary is hit and \(E^\lambda\) does not occur before the change of level, or, \label{item: 1}
+		\item the fluid changes phase before there is a change of level or a boundary is hit and \(E^\lambda\) does not occur before the change of phase. \label{item: 2}
+	\end{enumerate}
+
+	The probability of~\ref{item: 1}~is 
+	\begin{align}
+		\int_{x=0}^\infty \bs \alpha e^{\bs{S}x}\bs s e^{(T_{ii}-\lambda)x/|c_i|}\wrt x 
+		&= e^{(T_{ii}-\lambda)\Delta/|c_i|} + r_1,\label{eqn: bnd this 12345}
+	\end{align}
+	by Lemma~\ref{lemma:bound}.
+
+	The probability of~\ref{item: 2}~is 
+	\begin{align}
+		\nonumber\int_{x=0}^\infty \bs \alpha e^{\bs{S}x}\bs e e^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x 
+		&= \int_{x=0}^{\Delta+\varepsilon} \bs \alpha e^{\bs{S}x}\bs ee^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x 
+%		\\\nonumber&{}+\int_{x=\Delta}^{\Delta+\varepsilon}\bs \alpha e^{\bs{S}x}\bs ee^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x 
+		\\&{}+\int_{x=\Delta+\varepsilon}^\infty \bs \alpha e^{\bs{S}x}\bs e e^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x.\label{eqn: bnd this 1234}
+	\end{align}
+	Now, since \(\bs \alpha e^{\bs{S}x}\bs e\leq 1\) for \(x\leq \Delta+\varepsilon\) then the first term on the right-hand side of (\ref{eqn: bnd this 1234}) is less than or equal to 
+	\[\displaystyle \int_{x=0}^{\Delta+\varepsilon} e^{(T_{ii}-\lambda)/|c_i|x}(-T_{ii}/|c_i|)\wrt x \leq \int_{x=0}^{\Delta+\varepsilon} e^{T_{ii}/|c_i|x}(-T_{ii}/|c_i|)\wrt x = 1-e^{T_{ii}/|c_i|(\Delta+\varepsilon)}.\]
+	By Chebyshev's inequality, \(\bs \alpha e^{\bs{S}x}\bs e\leq \cfrac{\var(Z)}{\varepsilon^2}\) for \(x> \Delta+\varepsilon\), hence the second term on the right-hand side of (\ref{eqn: bnd this 1234}) is less than or equal to
+	\[\displaystyle \int_{x=\Delta+\varepsilon}^\infty \cfrac{\var(Z)}{\varepsilon^2} e^{(T_{ii}-\lambda)x/|c_i|}(-T_{ii}/|c_i|)\wrt x \leq  \cfrac{\var(Z)}{\varepsilon^2}.\]
+	Putting these together, then the right-hand side of (\ref{eqn: bnd this 1234}) is less than or equal to
+	\begin{align}
+		1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|} + \cfrac{\var(Z)}{\varepsilon^2}.\label{eqn: the big dog}
+	\end{align}	
+	Combining (\ref{eqn: bnd this 12345}) and (\ref{eqn: the big dog}), then \(\mathbb P(\tau_n\leq E^\lambda  \mid \phi(\tau_{n-1})=i , \tau_{n-1}\leq  E^\lambda)\) is less than or equal to 
+	\begin{align}
+		&e^{(T_{ii}-\lambda)\Delta/|c_i|} + |r_1| + 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|} + \cfrac{\var(Z)}{\varepsilon^2}\nonumber
+		\\&= 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|}\left[1-e^{(-T_{ii}\varepsilon-\lambda \Delta)/|c_i|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_1| \nonumber 
+		\\&= 1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_1|,
+	\end{align}
+	since \(-T_{ii}/|c_i|\leq q\) and \(\lambda \Delta/|c_i| \leq \lambda \Delta/c_{min}\) for all \(i\in\mathcal S_+\cup \calS_-\).
+
+	Now consider the QBD-RAP at a boundary.
+To leave the boundary there must be at-least one change of phase before \(E^\lambda\).
+By a uniformisation argument, the probability of at-least one change of phase before \(E^\lambda\) is less than or equal to \(q/(q+\lambda)\).
+\end{proof}
+
+\begin{lem}\label{lem: another bound 2}
+	For \(n\geq 2\), \(i\in\mathcal S_+\cup \calS_-\), 
+	\begin{align}
+		\mathbb P(\tau_n \leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \leq b^{n-1}.
+	\end{align}
+\end{lem}
+\begin{proof}
+	The proof is by induction.
+
+	For the base case, set \(n=2\) and apply Lemma~\ref{lem: another bound}.
+
+	Now, assume the induction hypothesis \(\mathbb P(\tau_{n-1} \leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \leq b^{n-2}\) for arbitrary \(n\geq 3\).
+
+	Since \(\{\tau_{n-1}\leq E^\lambda\}\) is a subset of \(\{\tau_{n}\leq E^\lambda\}\), then 
+	\begin{align}
+		 \mathbb P(   \tau_n \leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)
+		&= \mathbb P(   \tau_n \leq E^\lambda,  \tau_{n-1} \leq E^\lambda  \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda). \label{eqn: eheh}
+	\end{align}
+	Now partition (\ref{eqn: eheh}) on the phase at time \(\tau_{n-1}\),
+	\begin{align}
+		\nonumber&\sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_n \leq E^\lambda,  \tau_{n-1} \leq E^\lambda,\phi(\tau_{n-1}) = j_{n-1} \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)
+		\\&\nonumber=  \sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_{n}\leq E^\lambda\mid   \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1} \leq E^\lambda)
+		\\&\qquad\times\mathbb P( \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda), \label{eqn: kekew}
+	\end{align}
+	by the strong Markov property of the QBD-RAP and the fact that \(\bs A(\tau_{n-1})=\bs \alpha\).
+
+	By Lemma~\ref{lem: another bound} (\ref{eqn: kekew}) is less than or equal to 
+	\begin{align}
+		&\sum_{j_{n-1}\in\mathcal S} b
+		\mathbb P(  \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda\mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \nonumber
+		\\&= b
+		\mathbb P( \tau_{n-1}\leq E^\lambda\mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)\nonumber
+		\\&\leq b\cdot b^{n-2},
+	\end{align}
+	by the induction hypothesis, and this completes the proof. 
+\end{proof}
+%
+
+\begin{cor}\label{vcor: cdks d}
+	\begin{align}
+		\Bigg|&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}(t) \in (\ell,\wrt x, j), t\in\mathcal T_n \mid 
+		\bs Y(0)=\bs y_0)
+		\psi(x) \wrt t \Bigg|
+		\leq \cfrac{F b^{n-1}}{\lambda}\label{eq: JjbCS  B k }
+	\end{align}
+\end{cor}
+\begin{proof}
+	First, since \(|\psi(x)|\leq F\), then the left-hand side of (\ref{eq: JjbCS  B k }) is less than or equal to 
+	\begin{align}
+		&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}(t) \in (\ell,\wrt x,j), t\in\mathcal T_n \mid  \nonumber 
+		\bs Y(0)=\bs y_0)F\wrt t \nonumber
+		\\&=\int_{t=0}^\infty e^{-\lambda t} \mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \mid \bs Y(0)=\bs y_0)F\wrt t. \label{eqn: alalal}
+	\end{align}
+	Partitioning on the time of the \(1\)st change of level, \(\tau_1\), and the phase and level at time \(\tau_1\), then (\ref{eqn: alalal}) is equal to 
+	\begin{align}
+		&\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \nonumber
+		\mid \Yd(1)=(\ell_1,j_1),  \tau_1=u_1)\nonumber
+		\\&\quad\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
+		\mid \bs Y(0)=\bs y_0)F  \wrt t\nonumber
+		%
+		\\&=\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t-u_1\in\mathcal T_{n-1} \nonumber
+		\mid \bs Y(0)=(\ell_1,\bs \alpha,j_1))\nonumber
+		\\&\quad\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
+		\mid \bs Y(0)=\bs y_0)F  \wrt t, \label{eqn fLks}
+	\end{align}
+	by the time-homogeneous property of the QBD-RAP. By the convolution theorem for Laplace transforms, (\ref{eqn fLks}) is equal to 
+	\begin{align}
+		&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
+		\mid \bs Y(0)=(\ell_1,\bs \alpha, j_1)) \wrt t 
+		\\&\quad{}\int_{u_1=0}^\infty e^{-\lambda u_1}\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
+		\mid \bs Y(0)=\bs y_0)F  \nonumber 
+		%
+		%
+		\\&=\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
+		\mid \bs Y(0)=(\ell_1,\bs\alpha,j_1)) \wrt t 
+		\\&\quad{}\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\leq E^\lambda
+		\mid \bs Y(0)=\bs y_0)F. \label{eqn fLksoioi}
+	\end{align}
+
+	The expression 
+	\begin{align}
+		&\int_{t=0}^\infty e^{-\lambda t} \mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n\nonumber
+		\mid \bs Y(0)=(\ell_1,\bs \alpha,j_1)) \wrt t
+		%
+		\\&\leq \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\tau_{n}\leq t \nonumber
+		\mid \bs Y(0)=(\ell_1,\bs\alpha,j_1)) \wrt t
+		%
+		%
+		\\&\leq b^{n-1}\int_{t=0}^\infty e^{-\lambda t} \wrt t \nonumber 
+		\\&=b^{n-1}\cfrac{1}{\lambda}, \label{eq: bndhs vja sdh}
+	\end{align}
+	by Lemma~\ref{lem: another bound 2}.
+
+	Using the bound (\ref{eq: bndhs vja sdh}) in (\ref{eqn fLks}) gives 
+	\begin{align}
+		&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K} b^{n-1}\cfrac{1}{\lambda} 
+		\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\leq E^\lambda
+		\mid \bs Y(0)=\bs y_0)F\leq b^{n-1}\cfrac{1}{\lambda}F, \label{eqn fLksoioi2}
+	\end{align}
+	by the law of total probability. This concludes the proof. 
+\end{proof}
+
+\subsection{Global convergence}
+Finally, we combine the convergence result of Lemma~\ref{lem: LAkAKFnvnb mav h} and the domination condition from Corollary~\ref{vcor: cdks d} via the Dominated Convergence Theorem to claim convergence of the Laplace transform of the QBD-RAP given by
+\begin{align}
+	&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y}^{(p)}(t) = (\ell,\wrt x,j)\mid \bs Y^{(p)}(0)=\bs y_0^{(p)})  \psi(x)\wrt t. \label{eqn: LLLLaaakkkss}
+\end{align}
+Partitioning (\ref{eqn: LLLLaaakkkss}) on the number of level changes by time \(t\) gives
+\begin{align}
+	& \int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\sum_{n=0}^\infty \mathbb P(  {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid\nonumber 
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \psi(x) \nonumber
+	\\&=\sum_{n=0}^\infty\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}} \mathbb P( {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid  
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \psi(x). \label{eqn: KLDNVnav}
+\end{align}
+We can just the swap of the sum and integrals since 
+\begin{align}
+	& \int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\sum_{n=0}^\infty \mathbb P(  {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid\nonumber 
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t |\psi(x)| \nonumber
+	\\&\leq \int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\sum_{n=0}^\infty \mathbb P(  {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid\nonumber 
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t F \nonumber
+	\\&\leq \int_{t=0}^\infty e^{-\lambda t}\wrt t F \nonumber
+	\\&\leq \cfrac{1}{\lambda} F<\infty \nonumber
+\end{align}
+so the Fubini-Tonelli Theorem applies.
+
+By Lemma (\ref{lem: LAkAKFnvnb mav h}), each term in the sum (\ref{eqn: KLDNVnav}) converges.
+Furthermore, for \(n\geq 1\), each term is dominated by \(\left(b^{(p)}\right)^{n-1}F/\lambda\), from Corollary~\ref{vcor: cdks d}.
+The dominating terms \(\left(b^{(p)}\right)^{n-1}F/\lambda\) depend on \(p\) and may not be summable.
+However, for \(p\) sufficiently large, there exists a \(p_0<\infty\) and a \(B\) with \(B<1\) such that \(b^{(p)}<B\) for all \(p>p_0\).
+Hence we can apply the Dominated Convergence Theorem to (\ref{eqn: KLDNVnav}) and claim that
+\begin{align}
+	&\lim_{p\to\infty} \sum_{n=0}^\infty\int_{x\in\mathcal D_{\ell,j}}\int_{t=0}^\infty e^{-\lambda t} \mathbb P( {\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid \nonumber 
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \psi(x)  \nonumber
+	%
+	% \\&= \sum_{n=0}^\infty\int_{x\in\mathcal D_{\ell,j}}\lim_{p\to\infty} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\widetilde{\bs Y}^{(p)}(t) \in (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid \nonumber
+	% \bs Y^{(p)}(0)=(\ell_0,  \bs  a_{\ell_0,i}^{(p)}(x_0), i))\wrt t \psi(x) \nonumber
+	%
+	\\&= \sum_{n=0}^\infty\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
+	\mid \bs X(0)=(x_0,i)) 
+	\psi(x)\wrt t. \label{eqn:Alakskjs}
+\end{align}
+where the limit is given by Lemma~\ref{lem: LAkAKFnvnb mav h}. Swapping the sum and integrals and by the law of total probability, then (\ref{eqn:Alakskjs}) is equal to 
+\begin{align}
+	& \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
+	\mid \bs X(0)=(x_0, i))\psi(x)\wrt t \nonumber
+	%
+	\\&= \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}} \mathbb P(\bs X(t)\in(\wrt x, j)  
+	\mid \bs X(0)=(x_0, i))\psi(x)\wrt t. \nonumber
+\end{align}
+The swap of the sum and integrals is justfied as 
+\begin{align}
+	& \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
+	\mid \bs X(0)=(x_0, i))|\psi(x)|\wrt t \nonumber
+	%
+	\\&\leq\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\sum_{n=0}^\infty\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^X 
+	\mid \bs X(0)=(x_0, i))\wrt t  F\nonumber
+	\\&\leq\int_{t=0}^\infty e^{-\lambda t} \wrt t  F\nonumber
+	\\&=\cfrac{1}{\lambda}F<\infty
+\end{align}
+so the Fubini-Tonelli Theorem applies.
+
+Thus, we have shown the following result.
+\begin{lem}\label{lem: KajPOw}
+	For all \(\ell_0,\ell\in\mathcal K\), \(i,j\in\mathcal S\), \(x_0\in\mathcal D_{\ell_0,i}\), as \(p\to\infty\), 
+	\begin{align}
+		&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y}^{(p)}(t) = (\ell,\wrt x,j) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)})  \psi(x)\wrt t \nonumber
+		\\&\to \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}} \mathbb P(\bs X(t)\in(\wrt x, j)  
+		\mid \bs X(0)=(x_0, i))\psi(x)\wrt t. \nonumber
+	\end{align}
+\end{lem}
+
+Let \(\mathcal R(L(t),\bs A(t),\varphi(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\varphi(t)}(x)\), \(x\in\calD_{L(t),\varphi(t)}\). 
+\begin{cor}\label{cor: lk}
+	Let \(\psi: \mathbb R\times \mathcal S \to \mathbb R\) be an arbitrary bounded function with \(|\psi(\cdot)|\leq F\).
+%Further, let \(\mathcal L(X(t)) = \sum_{k\in\mathcal K}k1(X(t)\in\mathcal D_k)\) and \(\widetilde{\bs X}(t) = (\mathcal L(X(t)),X(t),\varphi(t)))\).
+	For each \(x_0\in\mathbb R\), \(i\in\mathcal S\), \(\ell_0\in\mathcal K\), 
+	\begin{align}
+		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \nonumber
+		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber
+		\\&\to \int_{t=0}^\infty e^{-\lambda t}  \mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0, i) \right] \wrt t.\nonumber
+	\end{align}
+\end{cor}
+
+\begin{proof}
+	Consider the left-hand side 
+	\begin{align}
+		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid 
+		  \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
+		\\&= \int_{t=0}^\infty e^{-\lambda t}\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A(t),j),j )  1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
+		\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[ \psi( \mathcal R(\ell,\bs A(t),j),j )1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid 
+		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t, \label{eqnL AKJF cjk ajhJK}
+	\end{align}
+	where the swap of the summations and integrals is justified since \(\psi\) is bounded and by the Fubini-Tonelli Theorem.
+By Lemma~\ref{lem: KajPOw}, for each \(\ell\in\mathcal K\), \(j\in\calS\), the terms
+	 \begin{align}
+	 	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A(t),j), j )\wrt x 1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t
+	\end{align}
+		converge to 
+	\[\int_{t=0}^\infty e^{-\lambda t}  \mathbb E\left[\psi({\bs X}(t) )1(\bs X(t)\in(\mathcal D_{\ell_j},j))\mid \bs X(0)=(x_0,i) \right] \wrt t.\]
+
+	If \(\mathcal K\) is finite, we are done upon taking the limit of (\ref{eqnL AKJF cjk ajhJK}) as \(p\to\infty\) and swapping the limit and the sums.
+
+	If \(\mathcal K\) is countably infinite, then for a given \(k\in\mathcal K\), since \(\psi\) is bounded,
+	\begin{align}
+		&\Bigg|\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A(t),j),\varphi(t) )  1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t\Bigg| \nonumber 
+		%
+		\\&\leq F \sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+		\bs Y^{(p)}(0)=\bs y_0^{(p)}\Bigg] \wrt t \nonumber 
+		%
+		\\&\leq F  \int_{t=0}^\infty e^{-\lambda t}\mathbb P( {\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j) \mid   
+		\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \nonumber 
+		%
+		\\&\leq F \mathbb P(\tau_{|\ell-\ell_0|}^{(p)}\leq E^\lambda \mid
+			\bs Y^{(p)}(0)=\bs y_0^{(p)}), \label{eqn:SKAM. fqj}
+	\end{align}
+	since, to be in level \(\ell\) after starting in level \(\ell_0\), there must be at least \(|\ell_0-\ell|\) changes of level.
+By Lemma~\ref{lem: another bound 2} then (\ref{eqn:SKAM.
+fqj}) is bounded by \(\left(b^{(p)}\right)^{|\ell-\ell_0|-1}\) for \(|\ell-\ell_0|\geq 2\) and by \(1\) otherwise.
+Now, choose \(p_0\) sufficiently large so that \(b^{(p)}<B<1\) for all \(p>p_0\).
+Therefore, for all \(p>p_0\), the terms in (\ref{eqnL AKJF cjk ajhJK}) are dominated by \(F\min\{B^{|\ell-\ell_0|-1},1\}\).
+Moreover
+	\begin{align*}
+		F\sum_{\ell\in\mathcal K} \min\{B^{|\ell-\ell_0|-1},1\} 
+		&\leq2\sum_{n=1}^\infty B^{n-1}+1
+		\\&=\cfrac{2}{1-B}+1
+		\\&<\infty,
+	\end{align*}
+	hence the dominating terms are summable. Hence we may apply the Dominated Convergence Theorem to swap the necessary limits and sums. 
+\end{proof}
+
+The Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} can now be use to claim that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.
+\begin{thm}[Extended Continuity Theorem \citep{feller1957}]\label{thm: ext cont thm}
+	For \(p=1,2,\dots\) let \(U_p\) be a measure with Laplace transform \(\omega_p\).
+If \(\omega_p(\lambda)\to\omega(\lambda)\) for \(\lambda > a\geq 0\), then \(\omega\) is the Laplace transform of a measure \(U\) and \(U_p\to U\).
+
+	Conversely, if \(U_p\to U\) and the sequence \(\{\omega_p(a)\}\) is bounded, then \(\omega_p(\lambda)\to\omega(\lambda)\) for \(\lambda >a\). 
+\end{thm}
+Thus, by the the Extended Continuity Theorem~\ref{thm: ext cont thm}
+\begin{align}
+		&\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] \nonumber 
+		\to \mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0, i) \right] \nonumber
+\end{align}
+weakly in \(t\) as \(p\to \infty\). Now, 
+\[\mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0,i) \right]\] 
+is a continuous function of \(t\) (it is a Feller semi-group), moreover, for \(p<\infty\), 
+\[\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \]
+ is also continuous in \(t\). However, the limit 
+ \[\lim\limits_{p\to\infty}\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) ) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \] need not be continuous in \(t\).
+Nonetheless, we can claim that
+\begin{align}
+		&\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )    \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \nonumber
+		\to \mathbb E\left[\psi({\bs X}(t) )\mid \bs X(0)=(x_0, i) \right] \nonumber
+\end{align}
+for almost all \(t\geq 0\).
+At such values of \(t\), since \(\psi\) is an arbitrary bounded, then the Portmanteau Theorem \citep[Theorem 2.1, page 16]{billingsleyconvergence} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
+
+A sufficient condition to upgrade the convergence from weak to point-wise (in the variable \(t\)) is to show that for \(t\geq 0\) 
+\[\sup_{p}\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \leq M(t)<\infty\]
+and the sequence \(\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \) is eventually equicontinuous in \(t\).
+That is, for every \(\varepsilon>0\) there exists a \(\delta(t,\varepsilon)>0\) and an \(p_0(t,\varepsilon)\) such that \(|t-u|<\delta(t,\varepsilon)\) implies that
+\begin{align}
+	&\nonumber \Bigg|\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
+	\\&{}-\mathbb E\left[\psi( \mathcal R(L(u),\bs A(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| \nonumber 
+	<\varepsilon
+\end{align}
+for all \(p\geq p_0(t,\varepsilon)\). This is an area for future work. Alternatively, the theory of \emph{killing} of a stochastic process may be applicable and provide the convergence we seek.
+
+% \begin{rem}
+% 	While the convergence result implies a loose bound on the rate of convergence, it applies to a wide class of QBD-RAP constructions.
+The main cause of the loose bound on the convergence rate is due to our reliance on Chebyshev's inequality.
+This is necessary as we try to make minimal assumptions about the matrix-exponential distributions used in the construction.
+In practice, we use a class of \emph{concentrated matrix-exponential distributions} found numerically in \citep{hht2020}, and for which there is relatively little known about their properties.
+This necessitates the generality of the convergence result.
+% \end{rem}
+
+\section{Extension to arbitrary (but fixed) discretisation structures}
+To conclude this chapter we include some remarks as to how to extend the convergence results to arbitrary discretisation structures.
+
+Throughout, we have assumed that all intervals are of width \(\Delta\), i.e.~\(|y_{\ell+1}-y_\ell|=\Delta\), and that on every interval the dynamics of the fluid queue are modelled based on the same matrix exponential representation \((\bs \alpha, \bs S, \bs s)\).
+These assumptions are, in fact, not necessary, but they do serve to simplify the presentation.
+The convergence results can be extended to use different sequences of matrix exponential representations on each interval, provided that for each sequence of matrix exponential distributions, the variance tends to \(0\).
+Moreover, we can extend the results to intervals of arbitrary width, provided that the width of the intervals is not arbitrarily small.
+Here we describe how one would prove such results.
+
+The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, i.e.~the hypotheses of the Lemma depend only on the interval \(\calD_{\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval and not on any other interval.
+Thus Lemma~\ref{thm: a thm!} holds independently on each interval, as does Corollary~\ref{cor: aln222}.
+
+Let the width of an interval \(\mathcal D_{\ell_0}\) be \(\Delta_{\ell_0}=y_{\ell_0+1}-y_{\ell_0}\) and suppose that sequence of matrix exponential random variables used to model the dynamics of the fluid queue on the interval \(\calD_{\ell_0}\) is \(Z_{\ell_0}^{(p)}\).
+Regarding Lemma~\ref{lem: another bound}, we can extend it the following version,
+\begin{lem}\label{lem: another bound sdfg}
+	Assume \(\inf_{\ell_0}\Delta_{\ell_0}>0\) and \(\sup_{\ell_0}\var(Z_{\ell_0})<\infty\). Then, for all \(i\in\mathcal S_+\cup \calS_-\), \(\ell_0\in\mathcal K\setminus\{-1,K+1\},\) and \(n\geq 2\), 
+	\begin{align}
+		&\mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \Ydp(n-1) = (\ell_0,i), \tau_{n-1}^{(p)}\leq  E^\lambda ) \leq b_{\ell_0}^{(p)},\label{eqn: ell0 version}
+	\end{align}
+	where 
+	\[b_{\ell_0}^{(p)} = 1-e^{-q(\Delta_{\ell_0}+\varepsilon_{\ell_0}^{(p)})}\left[1-e^{q\varepsilon_{\ell_0}^{(p)}-\lambda \Delta_{\ell_0}/|c_{min}|}\right] + \cfrac{\var(Z_{\ell_0}^{(p)})}{\left(\varepsilon^{(p)}\right)^2} + |r_{1,\ell_0}^{(p)}| \]
+	and  
+	\[|r_{1,\ell_0}^{(p)}|\leq 2G\cfrac{\var \left(Z_{\ell_0}^{(p)}\right)}{\left(\varepsilon_{\ell_0}^{(p)}\right)^2} + 2L\varepsilon_{\ell_0}^{(p)}.\]
+	Hence, for all \(i\in\mathcal S_+\cup \calS_-\), and \(n\geq 2\), 
+	\begin{align}
+		&\mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \phi^{(p)}(\tau_{n-1}^{(p)})=i, \tau_{n-1}^{(p)}\leq  E^\lambda ) \leq b^{(p)}, \label{eqn: skjhg}
+	\end{align}
+	where 
+	\[b^{(p)} = \max\left\{\sup_{\ell_0}b_{\ell_0}^{(p)}, \cfrac{q}{\lambda + q}\right\}.\]
+\end{lem}
+\begin{proof}
+	For the proof of (\ref{eqn: ell0 version}) follow the same arguments as in the proof of Lemma~\ref{lem: another bound}.
+The bound in (\ref{eqn: skjhg}) follows by the assumptions in the statement of the result.
+With the bound (\ref{eqn: skjhg}), then
+	\begin{align*}
+		&\mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \phi^{(p)}(\tau_{n-1}^{(p)})=i, \tau_{n-1}^{(p)}\leq  E^\lambda ) 
+		\\&\leq \sup_{\ell_0} \mathbb P(\tau_n^{(p)}\leq E^\lambda \mid \Ydp(n-1) = (\ell_0, i), \tau_{n-1}^{(p)}\leq  E^\lambda ) 
+		\\&\leq \max\left\{\sup_{\ell_0}b_{\ell_0}^{(p)}, \cfrac{q}{\lambda + q}\right\}.
+	\end{align*}
+\end{proof}
+Given Lemma~\ref{lem: another bound sdfg}, then an equivalent of Lemma~\ref{lem: another bound 2} remains true, the proof of which follows verbatim except with the use of Lemma~\ref{lem: another bound} replaced by Lemma~\ref{lem: another bound sdfg}.
+Corollary~\ref{vcor: cdks d} remains true without modification.
+Lemma~\ref{lem: KajPOw} and Corollary~\ref{cor: lk} remain true without modification provided that \(\lim\limits_{p\to\infty}\var(Z_\ell^{(p)})\to 0\) for all \(\ell\).
+
+\begin{rem}
+	I suspect that the approximation results can also be extended to approximating so-called \emph{multi-layer fluid queues}, as described in \cite{bo2008}. 
+\end{rem}

### 2021-12-06 chapter6/chapter.tex
rought redraft of chapters 4/5, started on one D wave examples

%!TEX root = ../thesis.tex
\chapter{Numerical investigations\label{sec: numerics}}
We have now established multiple methods for approximating fluid queues which suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}. In this chapter we numerically investigate various aspects of the approximation schemes. Throughout, we compare the QBD-RAP scheme (from Chapters~\ref{sec: conv} and \ref{sec: construction and modelling}), the discontinuous Galerkin method (without limiters or filtering of the solution) (as decribed in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013} (which we will refer to a just the \emph{uniformisation scheme}). Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux. Like the QBD-RAP scheme, due to its stochastic interpretation, the uniformisation scheme is positivity preserving.

Where appropriate, we also compare the above methods with the discontinuous Galerkin method with a slope-limiter implemented, which avoids oscillations and negative solutions. However, recall that, in the context of approximating operators of fluid-fluid queues with these schemes, slope-limiting is not possible, and even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.

The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased. For discontinuous Galerkin method, this means increasing the number of basis functions used to approximate the solution within each cell. For the QBD-RAP scheme, this means increasing the order of the CME distribution used to construct the scheme. For the uniformisation scheme, this means dividing each cell into smaller sub-cells over which we approximate the solution. As such, each approximation scheme leads to approximations which are matrices of the same size (although not necessarily the same number of non-zero components). For example, if we use 3 basis functions in the discontinuous Galerkin method, we approximate the solution by a quadratic on each cell; the comparable QBD-RAP scheme is to use a CME of order 3; and the comparable uniformisation scheme is to divide each cell into 3 sub-cells of width \(\Delta/3\).

To keep the content of this chapter contained, we do not investigate the performance of the schemes as the cell sizes are decreased, other than for the uniformisation scheme. As such, for each numerical experiment, we keep the cell size fixed. As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next and link all the cells together. We investigate schemes with an upwind flux only.
We do not investigate the stability of the schemes with respect to the step-size of the time-integration or time-integration scheme (where required). Instead, we fix the time-integration step size for each numerical experiment at a suitably small value so as to obey a certain stability criterion (a CFL-like condition, REFs, appendix?). For schemes which require us to integrate over time we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (see appendix and REF). This scheme claims to introduce no more oscillations into the solution as we integrate over time (REFs). Where a slope limiter is implemented, we always implement the \emph{Generalised MUSCL} limiter (REF and Appendix). We do not investigate filtering for the DG scheme (REFs).

The performance of the discontinuous Galerkin method has been well-studied (REFs), and it is well-known that the discontinuous Galerkin method performs remarkably well on problems with smooth solutions (REF). Here, we mostly focus on investigating the numerical performance of the methods on problems with non-smooth solutions, the purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP method. In the stochastic modelling community it is very common to have problems with discontinuous solutions, such as a non-smooth initial conditions (REFs). Even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities into the problem. There is a very specific set of conditions on the initial density and point masses which must hold for the distribution of the fluid queue to remain smooth as it evolves over time (REFs).Limiting distributions of fluid queues are smooth, however. Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues.

As we shall see, if the solution is smooth, then the Galerkin method is highly-effective and displays rapid convergence to the solution as the number polynomial basis functions is increased. However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin method, leading to non-sensical solutions. In these cases, the QBD-RAP approximation performs relatively well compared to the other positivity preserving schemes investigated.

% THE STRUCTURE OF THE CHAPTER IS AS FOLLOWS.

% 1. FUNCTION APPROX./RECONSTRUCTION

% 2. 1d wave eqn

% 3. STATIONARY DISTRIBUTIONS
	% i. SIMPLER
	% ii. MORE COMPLEX

% 4. TRANSIENT DISTRIBUTIONS 
	% i. SIMPLER
	% ii. MORE COMPLEX

% 5. HITTING TIMES 
	% i. SIMPLER
	% ii. MORE COMPLEX

% 6. FLUID-FLUID QUEUES

% 7. CONCLUDING REMARKS

\section{Function approximation/reconstruction}
We start our investigation by looking at how well the methods perform at approximating an initial condition. For the discontinuous Galerkin method, this means projecting the solution on to the set of polynomial functions which define the scheme, then reconstructing the solution. For the spatially-coherent uniformisation scheme, this means looking at the sub-cell averages. For the QBD-RAP scheme, this means computing the initial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}. First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}) performs reasonably compared to the others, and we are able to theoretically prove that it converges. Therefore we use this reconstruction method throughout the rest of the chapter.

For this section we consider approximating the initial condition over a single cell of width \(\Delta = 1\). To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations. Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points over which to compute the KS statistics.

First we briefly investigate the possible choices of closing operators as introduced in Section~\ref{sec: closing}.

\subsection{QBD-RAP closing operators}
The three closing operators we investigate are the following. The unnormalised closing vector in Equation~(\ref{eqn: alal}). A normalised version of the closing operator in Equation~(\ref{eqn:546}), with the normalised version given by (\ref{eqn: nonlin}). We refer to the closing vector in (\ref{eqn: nonlin}) as the \emph{naive normalised} closing vector. Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme, though it clearly appears to converge. The third closing vector we investigate is that in Equation~(\ref{eqn:5462}).

\begin{example}First consider approximating the initial condition with density function \(1(x<0.5)\). In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation. Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders. It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interferes with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme. Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection. The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward). Intuitively, there should be very little mass in the tail of the distribution as we are using concentrated matrix exponential distributions in the reconstruction.

If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs}, then we see that the unnormalised reconstruction performs the poorest. Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests naive normalised and normalised reconstruction are very similar. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(2x1(x<0.5)+1(x\geq 0.5)\), and the approximations (left) and \(L^2\) error between the true PDF, \(2\times 1(x<0.5)\), and the approximations (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 2 ks error qbdrap closing vecs}
\end{figure}
\end{example}

\begin{example}Now consider approximating the initial density \(f(x)=1\). Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error, and \(L^2\) error between the true initial distribution and the reconstructions, we now see that the normalised reconstructions outperform the unnormalised reconstruction. This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(x\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 4 ks error qbdrap closing vecs}
\end{figure}

Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised and normalised closing operators for order 1, 3, 5 and 7, as well as the true PDF. Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots and for all orders. This feature is due to a significant amount of mass being lost due to the truncation. In comparison the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator. In the reconstruction we also need to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}. Here we suppose that the phase is positive, so the `folding' around \(\Delta\) in the normalised closing operators appears appears at the left of the plots. In general, reconstructions via the unnormalised closing operator typically underestimate the value of the function in this region.

Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval. This is a common feature, and drawback, of the QBD-RAP method. Perhaps there is a different reconstruction method or yet another closing operator which could alleviate this issue. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=1\) (orange dashed line).}
	\label{fig: pdf reconstructed}
\end{figure} 
\end{example}

\begin{example}Next we consider approximating the initial distribution with density \(-6x^2+6x\). Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable. However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, we see that with this metric, then unnormalised closing operator performs better than the two normalised ones for orders 5 and above. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(-2x^3+3x^2\), and the approximation (left) and \(L^2\) error between the true PDF \(-6x^2+6x\) and the approximation (right), for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 6 ks error qbdrap closing vecs}
\end{figure}

The fact that the unnormalised closing operator outperforms the two normalised ones can be explain, once again, by the fact that, for the normalised operators we have `folded' the ME density function back on itself. In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\). In Figure~\ref{fig: pdf reconstructed quadratic} we observe that, at the left of the plot, the normalised reconstructions over estimate the density function in this region, where as the unnormalised reconstruction looks to be doing better. Recall that the `folding' in the normalised closing operator manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=-6x^2+6x\) (orange dashed line).}
	\label{fig: pdf reconstructed quadratic}
\end{figure} 
\end{example}

\begin{example}Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\). This density function is at a maximum at the left of the region. Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case. Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that the normalised reconstructions perform relatively well compared to the unnormalised reconstruction as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF). If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \((1-e^{-3x})/(1-e^{-3})\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 7 ks error qbdrap closing vecs}
\end{figure}
\end{example}

Given the evidence above we choose to use the normalised closing operator to reconstruct solutions. Further, unlike the naive normalised operator, the normalised operator is theoretically justified in that we proved the closing operator leads to a convergent scheme and is linear.

\subsection{Comparison of methods}
Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.

\begin{example}First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5. This distribution does not have a PDF so we can compare the CDFs only. In Figure~\ref{fig: fun 1 comp} we plot the KS metric (left) and \(L^1\) metric (right) between the true CDF and the reconstructed approximations. Observing the KS metric, it appears that none of the methods converge and the KS error sits around 0.5. This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except at points of discontinuity. None of the methods appear to converge at this point. Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the methods appear to show the convergent behaviour we expect. Here, the uniformisation method appears to perform the best, while the QBD-RAP method performs the worst. The rate of convergence of the QBD-RAP method appears to be similar to the rate of convergence of the DG scheme.

Perhaps it is no surprise that the uniformisation method performs best. In the uniformisation method as we increase the order, we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the function. As such, the uniformisation method can produce a piecewise continuous, linear approximation to the CDF. In contrast, both the DG and QBD-RAP methods result in a smooth approximation to the CDF. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(1(x\geq 0.5)\), and the approximations (left) and \(L^1\) error between the true CDF and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 1 comp} 
\end{figure}

In Figure~\ref{fig: pdf comp fun 1} we plot the approximated CDFs for the DG, uniformisation and QBD-RAP methods alongside the true CDF. The DG method displays undesirable features for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1. On the other hand, although the QBD-RAP method converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter6/figs/comp/fun1/cdfs_formatted.pdf}
	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (green dot-dashed line) methods. The true distribution function is \(1(x\geq 0.5)\) (grey dotted line).}
	\label{fig: pdf comp fun 1}
\end{figure} 
\end{example}

\begin{example}Now consider approximating the initial distribution with PDF \(1(x\leq 0.5)\). Figure~\ref{fig: fun 2 comp} plots the KS error (left) and \(L^2\) error between the true and approximate PDFs (right). Figure~\ref{fig: fun 2 comp} suggests that all methods converge at a similar rate for this problem. Once again, the QBD-RAP method performs worst, the uniformisation method second, and the DG method the best. However, once again the DG method exhibits undesirable properties -- the approximation to the CDF is, at some points above 1, is not monotonic, though these violations are not as severe in this case as they were for the case above. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(x\geq 0.5)\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 2 comp} 
\end{figure}
\end{example}

\begin{example}So far we have considered two problems which exhibit discontinuities. At the other extreme we now consider an initial distribution with density \(-6z^2+6x\). In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDF. Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG method can theoretically approximate the initial condition exactly. This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from 1-3, then plateaus. Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, and due to machine arithmetic, the errors for the DG scheme are not 0. Regarding the other two methods, they too appear to be convergent at approximately the same rate, with the uniformisation method perform better for the KS error, but very similarly in terms of the \(L^2\) error. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(-6z^2+6x\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 6 comp} 
\end{figure}
\end{example}

\begin{example}Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\). Figure~\ref{fig: fun 9 comp} shows the errors. Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs tell a similar story. For sufficiently high order, the DG method approximates the inital condition very well. The uniformisation method performs second best, while the QBD-RAP method is slowest to converge. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(cos(4\pi(x+0.5))+1\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 9 comp} 
\end{figure}
\end{example}
\FloatBarrier

\section{Travelling wave}
Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with varying initial conditions. Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\). The PDE (if/when it exists) which describes this system is 
\[\cfrac{\partial}{\partial t} f(x,t) = -\cfrac{\partial}{\partial x} f(x,t),\]
where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
\[f(x,t) = f(x-t,0)\]
so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).

We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space. We discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric. We use the SSPRK4 method to integrate over time using a step size of \(0.005\), and we evolve the system until time \(t=4\). For the DG method we also implement the Generalise MUSCL (REF, appendix?) slope limiter to stabilise the integration.

\begin{example}First consider the initial condition with PDF \(1(x<1)\). The level of the fluid queue is uniformly distributed over the first cell. For the DG and uniformisation methods, the initial condition can be represented exactly, whereas, for the QBD-RAP method, it cannot. Thus, in this case, there is no discretisation error in constructing the initial condition for the DG and uniformisation methods. Furthermore, at time \(t=4\), the solution in \(1(x\in[4,5)\), and the projections related to the DG and uniformisation methods can represent this solution exactly too.

Figure~\ref{fig: fun 4 wave} plots the KS error between the approximated and true CDFs and \(L^2\) error between the approximated and true PDFs. The DG method with the MUSCL limiter does not appear to converge for this problem. This is caused by the limiter reducing the method to linear in when oscilations are deteced, so increasing the order does not increase the accuracy of the method. The other two positivity preserving methods, the uniformisation method and QBD-RAP methods, both appear to be converging, with the QBD-RAP method converging faster. With these error metrics, the DG method also appears to converge. However, if we observe the solutions resulting from the DG method (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscilations and negative values.

Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG with MUSCL limiter and QBD-RAP methods. In the first row we observe that the DG method (sans limiter) produces an approximation to the PDF with negative values for orders greater than 1. In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter the approximation does not change significantly after order 3. This is due the the fact that the DG method is at best linear. In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation method, and in the last row is the solution approximated using the QBD-RAP method. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(4\leq x < 5)\), and the approximations for the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).} 
	\label{fig: fun 4 wave} 
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG with MUSCL limiter (third row) and QBD-RAP (bottom row) methods, for order 1, 3, 5, and 7 (columns). The true density function is \(1(4\leq x<5)\).}
	\label{fig: pdf wave fun 4}
\end{figure}

This is a particularly interesting example. It shows that, for the DG method, even though there is no discretisation error for the initial condition, we have use a strong stability preserving time-intergration method, and the projection off which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as in the top row of Figure~\ref{fig: pdf wave fun 4}. 
\end{example}

\begin{example}Another interesting example occurs with the initial condition with CDF \(1(z\geq 0.5)\), i.e.~a point mass at 0.5. The exact solution at time \(t=4\) is therefore a point mass at \(4.5\). No PDF exists for the true distribution, so instead we compare the CDFs. Moreover, when we analysed reconstruction of this initial condition we saw that using the KS metric may be uninformative. Instead, for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs). We also compare the probabilities
\begin{align}
	\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs}
\end{align}
for \(\ell=1,\dots,K\), and also the boundary mass
\begin{align}
	\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs boundary}
\end{align}
by computing 
\begin{align}
	&\sum_{\ell=1}^{10} \left|\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,\ell,1)\right|
	\\&+\left|\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,10,1)\right| \label{eqn: cell errors}
\end{align}
where \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,10,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).

Figure~\ref{fig: fun 2 wave} (left) plots the \(L^1\) error between the true and approximated CDFs (left). The \(L^1\) metric tells a similar story to the previous analysis: the DG method with the limiter does not converge as order increases, the other two positivity preserving methods, the uniformisation and QBD-RAP methods, appear to converge, with the QBD-RAP method converging faster, while the DG method appears to converge the fastest. However, if we plot the approximations to the CDFs from the DG method (not shown) we would once again see an oscilating non-monotonic approximation. Another interesting observation is to compare the performance of the uniformisation and QBD-RAP methods with respect to the \(L^1\) metric on the CDFs in Figure~\ref{fig: fun 2 wave} (left) to that in Figure~\ref{fig: fun 1 comp} (right). Recall that with Figure~\ref{fig: fun 1 comp} we investigate the ability of the different approximation methods to represent the initial distribution with CDF \(1(x\geq 1)\), which is the initial condition of the problem we are consdiering here. In Figure~\ref{fig: fun 1 comp} (right) the uniformisation method out-performed the QBD-RAP method at reconstructing the inital condition. However, in Figure~\ref{fig: fun 2 wave} (left) we see that the QBD-RAP method out-performs the uniformisation method. This suggest that the QBD-RAP method is better able to resolve movement of mass across the domain when integrating over time, compared to the uniformisation method.

Figure~\ref{fig: fun 2 wave} (right) plots the cell-based error metric in (\ref{eqn: cell errors}). Once again, the DG method with MUSCL limiter does not appear to converge as we expect. The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP method converging faster. Most interestingly though is the DG method. It is unclear as to whether this method will converge in this case as the errors in the plot are not obviously decreasing. Moreover, with this metric, the QBD-RAP method out-performs the DG method. Furthermore, the DG method can result in negative estimates of the cell probabilities in (\ref{eqn: cell probs}) and (\ref{eqn: cell probs boundary})
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/meshs_l1_cdf_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/L1_cell_probs.pdf}
	\caption{\(L^1\) error between the true CDF, \(1(4.5\leq x)\), and the approximations (left) and the error metric based on the cell masses as described in (\ref{eqn: cell errors}) for the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).} 
	\label{fig: fun 2 wave} 
\end{figure}
\end{example}

### 2021-12-07 chapter6/chapter.tex
more numerical investigations

@@
 \begin{align}
-	&\sum_{\ell=1}^{10} \left|\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,\ell,1)\right|
-	\\&+\left|\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,10,1)\right| \label{eqn: cell errors}
+	&\sum_{\ell=1}^{10} \left|\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,\ell,1)\right|\nonumber 
+	\\&+\left|\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,11,1)\right| \label{eqn: cell errors}
 \end{align}
-where \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,10,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).
+where \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,11,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).
 
@@
 \end{example}
+
+\begin{example}
+	Now consider an initial distribution which is truncated Gaussian with mean parameter 2.5 and standard deviation parameter 0.5, and trucated at the boundaries, 0 and 10;
+	\begin{align}
+		\mu([0,x)) := \mathbb P(X(0)\leq x, \varphi(0)=1)=\cfrac{\Phi((x-2.5)/0.5)1(0 \leq x < 10.0)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10\leq x), \label{eqn: kdjfksdf}
+	\end{align}
+	where \(\Phi(x)\) is the CDF of the standard normal distribution.
+
+	Hence, at time \(t=4\) the distribution is truncated Gaussian with mean parameter 6.5, standard deviation parameter 0.5, and is truncated below at 4.0 and above at 10, and there is also a small mass at the upper boundary;
+	\begin{align}
+		&\mathbb P(X(4)\leq x, \varphi(4)=1\mid X(0)\sim\mu, \varphi(0)=1) \nonumber 
+		\\&=\cfrac{\Phi((x-6.5)/0.5)1(4.0\leq x)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10 \leq x)\label{eqn: bbbbbaabbaa}%\cfrac{\Phi(7.5/0.5)-\Phi(3.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}.
+	\end{align}
+	The mass at the boundary is approximately \(1.28\times 10^{-12}\).
+There is a small discontinuity at \(x=4\) where the CDF jumps from \(0\) to \(\approx 2.87\times 10^{-7}\), due to the truncation of the initial distribution at \(0\).
+
+	Since the discontinuity at \(x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG method will perform well for this example.
+Figure~\ref{fig: fun 6 wave} confirms that this is indeed the case.
+For both metrics the error obtained by the DG method rapidly decreases to a point where it is swamped by other numerical errors.
+This is charaterisitc of the DG method for smooth problems.
+For low order DG schemes there are regions where the approximated PDF is negative, however, as the order of the DG scheme increases, these regions dissipate.
+Interestingly, even though this is a relatively smooth problem, the DG method with MUSCL limiter does not perform well.
+It must be that the initial and transient distributions are sufficiently `pointy' that small oscillations in the numerical solutions occur, and the limiter reduces the order of the scheme to linear.
+Both the uniformisation and QBD-RAP schemes perform similarly to previous examples.
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_l2_pdf_error_formatted.pdf}
+		\caption{KS error between the true CDF (Equation (\ref{eqn: bbbbbaabbaa})) and the approximations (left) and \(L^2\) error between the true PDF and the approximations for the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+		\label{fig: fun 6 wave} 
+	\end{figure}
+\end{example}
+
+\begin{example}
+	We now want to look at how the methods might handle a point mass.
+We introduce an ephemeral second phase into the model with phase transition rate \(T_{22}=-1\), and fluid rate \(c_2=0\).
+The generator is therefore
+	\[T=\left[\begin{array}{cc} 0 & 0 \\ 1 & -1 \end{array}\right].\]
+
+	We suppose that the initial condition is a point mass at the boundary in phase \(2\), i.e.~
+	\[\mathbb P(X(0)\leq x, \varphi(0)=2)=1(0\leq x).\] 
+	With this inital condition, the transient distribution at time \(t=4\) is 
+	\begin{align}
+		&\mathbb P(X(4)\leq x,\varphi(4)=1 \mid X(0)=0,\varphi(0)=2) \nonumber 
+		\\&= e^{T_{22}4}\left(e^{-T_{22}x}-1\right)1(4< x) + (1-e^{T_{22}4})1(4\leq x) \nonumber
+		\\&= e^{-4}\left(e^{4x}-1\right)1(4< x) + (1-e^{-4})1(4\leq x) \label{eqn: asjda}
+	\end{align}
+	and 
+	\begin{align}
+		\mathbb P(X(4)\leq x,\varphi(4)=2 \mid X(0)=0,\varphi(0)=2) = e^{-4}1(0\leq x).
+	\end{align}
+	The PDF at \(t=4\) is discontinuous at \(x=4\). For this problem, all methods can represent the initial condition exactly.
+
+	Figure~\ref{fig: fun 1 wave} plots the KS error metric between the true and approximated CDFs (right) and the \(L^2\) error metrics between the true and approximated PDFs.
+Observing Figure~\ref{fig: fun 1 wave}, as expected, since the solution is discontinuous, the DG method with MUSCL limiter does perform well.
+The uniformisation and QBD-RAP methods appear to converge, with the QBD-RAP method converging faster.
+DG method converges fastest, however, produces approximations with negative and oscillatory solutions as we might expect given the discontinuity.
+A selection of approximations to the transient PDF are shown in Figure~\ref{fig: pdf wave fun 1}
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_l2_pdf_error_formatted.pdf}
+		\caption{KS error between the true CDF (Equation (\ref{eqn: asjda})) and the approximations (left) and \(L^2\) error between the true PDF and the approximations for the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+		\label{fig: fun 1 wave} 
+	\end{figure}
+	\begin{figure}
+		\centering
+		\includegraphics[width=\textwidth]{chapter6/figs/wave/fun1/pdf_formatted.pdf}
+		\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG with MUSCL limiter (third row) and QBD-RAP (bottom row) methods, for order 1, 3, 5, and 7 (columns).
+The true density function is \(e^{-4}e^{x}1(x<4)\).}
+		\label{fig: pdf wave fun 1}
+	\end{figure} 
+\end{example}
+
+\FloatBarrier
+
+\section{Stationary distributions} \label{sec:stat}
+We briefly turn our attention to stationary distributions of fluid queue.
+Since the stationary distribution is smooth, apart from at the boundaries, we expect that the DG method will work well here.
+The stationary distributions are conveninent as they can be evaluated analytically (CITE) and do not require us to approximate initial conditions or integrate over time.
+
+Here we analyse two simple models which are versions of \cite[Example 2]{bean2009}.
+They are versions of \cite[Example 2]{bean2009} since there no lower boundary behaviour was specified (as it is inconsequential to their analysis), whereas here, we introduce behaviour at the lower boundary.
+The two models differ only by their behaviour at the boundaries.
+
+\begin{model}\label{model: simple}
+	Consider a fluid queue where the driving process is a CTMC with state space \(\calS=\{1,2,3,4\}\), generator 
+	\[\bs T = \left[\begin{array}{cccc}
+		-1.1 & 1.1 & 0 & 0 \\
+		1.1 & -1.1 & 0 & 0 \\ 
+		0.01 & 0 & -0.01 & 0 \\
+		0 & 0.01 & 0 & -0.01 
+	\end{array}\right],\]
+	and there are associated rates \(c_1=1, c_2 = -1, c_3=0, c_4=0\) and boundaries at \(x=0\) and \(x=10\). We specify two types of behaviour at the boundary.
+	\begin{enumerate}[a)]
+		\item (Absorbing model) Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
+		\[\vligne{p_{21} & p_{22} & p_{23} & p_{24}} = \vligne{0 & 0 & 1 & 0}.\]
+		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
+		\[\vligne{p_{11} & p_{12} & p_{13} & p_{14}} = \vligne{0 & 0 & 0 & 1}.\]
+		\item (Reflecting model) Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
+		\[\vligne{p_{21} & p_{22} & p_{23} & p_{24}} = \vligne{1 & 0 & 0 & 0}.\]
+		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
+		\[\vligne{p_{11} & p_{12} & p_{13} & p_{14}} = \vligne{0 & 1 & 0 & 0}.\]
+	\end{enumerate}
+\end{model}
+For Model~\ref{model: simple}~a), upon hitting the boundary, the process stays at the boundary for an exponentially distributed amount of time with mean 100.
+For Model~\ref{model: simple}~b), upon hitting the boundary, the process is immediately reflected.
+In Model~\ref{model: simple}~b) phases 3 and 4 are ephemeral.
+
+The models were dicretised using the DG, uniformisation and QBD-RAP methods using ten cells of width \(1\).
+We compute the coefficients for the stationary distribution in the following way.
+Suppose that a given discretisation method results in an approximation to the generator of the fluid queue as a matrix \(\bs B\).
+Then the stationary coefficients are found by solving
+\begin{align}
+	\bs b \bs B = 0,\\
+	\mbox{such that } \bs b\bs 1=1,
+\end{align}
+for the coefficients \(\bs b\).
+
+\paragraph{Model~\ref{model: simple}~a) Absorbing model}
+Figure~\ref{fig: absorbing stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
+Clearly the DG method is superior here as its error rapidly decrease to a point where is become insignificant compared to other numerical errors.
+The QBD-RAP method and uniformisation method both appear to be converging, with the errors for the QBD-RAP method decreasing faster.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/l2_pdf_error_formatted.pdf}
+	\caption{KS error between the true stationary CDF  and the approximations (left) and \(L^2\) error between the true stationary PDF and the approximations for Model~\ref{model: simple}~a) using the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: absorbing stationary} 
+\end{figure}
+
+\paragraph{Model~\ref{model: simple}~b) Reflecting model}
+Figure~\ref{fig: reflecting stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
+Once again, the DG method is superior, the QBD-RAP method and uniformisation method both appear to be converging, with the errors for the QBD-RAP method decreasing faster.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/stationary_distribution/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/stationary_distribution/l2_pdf_error_formatted.pdf}
+	\caption{KS error between the true stationary CDF and the approximations (left) and \(L^2\) error between the true stationary PDF and the approximations for Model~\ref{model: simple}~b) using the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: reflecting stationary} 
+\end{figure}
+
+\section{Transient distributions}
+Once again we consider Model~\ref{model: simple} and use the same spatial discretisation as described in Section~\ref{sec:stat} (ten cells of width 1).
+We consider two initial conditions, a point mass at 0.0 in phase 1, and the initial distribution with PDF
+\begin{align}
+	\cfrac{1}{2}e^{-x}/(1-e^{-10})\label{eqn: exp init cond}
+\end{align}
+in phases 1 and 2, with no mass at the boundaries.
+We integrate the approximations numerically over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.
+We apply the DG method with, and without, the Generalised MUSCL slope limiter.
+
+To obtain a \emph{ground truth} we simulated 5,000,000 realisations of the fluid queue and computed the empirical CDF, and also the mass within each cell and at the boundaries.
+We do not attempt to numerically approximate the true PDF via simulation.
+We then compute the KS and \(L^1\) error metrics between the approximated CDF and simulated CDF, as well as the cell-wise error metric, given as follows.
+\begin{align}
+	&\sum_{j\in\{1,2\}}\sum_{\ell=1}^{10} \left|\mathbb P(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0)=0.5,\varphi(0)=1)-p(2,\ell,j)\right|\nonumber 
+	\\&+\left|\mathbb P(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))-p(2,11,4)\right|\nonumber 
+	\\&+\left|\mathbb P(X(2)\in\{0\}, \varphi(2)=3\mid X(0),\varphi(0))-p(2,0,3)\right| \label{eqn: cell errors 2}
+\end{align}
+where \(p(2,\ell,j)\) is an approximation to \(\mathbb P(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0),\varphi(0))\), \(p(2,11,4)\) is an approximation to \(\mathbb P(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))\), and \(p(2,0,3)\) is an approximation to \(\mathbb P(X(2)\in\{10\}, \varphi(2)=3\mid X(0),\varphi(0))\).
+
+To examine possible Monte-Carlo error, we used a bootstrap with 1,000 bootstrap samples.
+That is, we sample, with replacement, 5,000,000 realisations of the fluid queue from the original 5,000,000 samples, then compute error metrics for each of the 1,000 bootstrap samples.
+Via the bootstrap, we report the \(5\)th and \(95\)th percentile of the sampling distribribution of the errors.
+
+To evaluate error metrics, we use a grid of 10,001 evenly spaced points for each phase.
+
+To approximate the point mass inital condition we compute the inital coefficients for each scheme exactly.
+For the exponential initial condition (\ref{eqn: exp init cond}) we compute the initial coefficients via Gauss-Lobatto quadrature for the DG method, by using the mid-point rule for the uniformisation method, and by using a trapezoidal rule with 2,001 points on each cell for the QBD-RAP method.
+
+\paragraph{Model~\ref{model: simple}~b) Reflecting model}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple}~b) with the point-mass initial condition, and the corresponding approximations obtained via the DG method (blue solid line with crosses), DG method with Generalised MUSCL limiter (green dashed line with circles), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: reflecting transient pm} 
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple}~b) with the expponential initial condition, and the corresponding approximations obtained via the DG method (blue solid line with crosses), DG method with Generalised MUSCL limiter (green dashed line with circles), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\label{fig: reflecting transient exp} 
+\end{figure}

### 2021-12-22 chapter5/chapter.tex
ch 4 updates, moved appendix A to main section. pre Peter's feedback

@@
 
-\section{Convergence of an embedded process}\label{sec: 1st change}
+\section{Convergence of an embedded process}\label{sec: embedded}
 % Before we move on to global convergence we need another result about convergence of the QBD-RAP scheme and the fluid queue at changes of level.
@@
 
-In this section we consider the embedded process formed by observing the QBD-RAP at changes of level.
-We start by analysing the QBD-RAP at the time of the first change of level, \(\tau_1^{(p)}\).
-Following a similar method to Chapter~\ref{sec: conv} we can prove the following result.
-
-\begin{cor}\label{cor: aln222} For \(\ell,\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\)
-	then, for \(j\in\calS_+\),
-	\begin{align}
-		&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0+1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
-            	 \mid \bs Y^{(p)}(0) = \bs y_0^{(p)}) \nonumber
-	 	%
-		\\&\to \mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0,i))\label{eqn: 1421}
-		 %
-		 %
-		 \intertext{and for \(j\in\calS_-\)}
-		 &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
-            	 \mid \bs Y^{(p)}(0) = \bs y_0^{(p)}) \nonumber
-	 	%
-		\\&\to \mathbb P(\bs X(\tau_1^X) = (y_{\ell_0}, j), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0, i)). \nonumber
-		%
-	\end{align}
-	% \begin{align}
-	% 	&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0+1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
-    %         	 \mid L^{(p)}(0) = \ell_{0},\bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \varphi(0) = i) \nonumber
-	%  	%
-	% 	\\&\to \mathbb P(X(\tau_1^X) = y_{\ell_0+1}, \varphi(\tau_1^X) = j, \tau_{1}^X\leq E^\lambda 
-    %         	 \mid X(0) = x_0, \varphi(0) = i)\label{eqn: 1421}
-	% 	 %
-	% 	 %
-	% 	 \intertext{and for \(j\in\calS_-\)}
-	% 	 &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
-    %         	 \mid L^{(p)}(0) = \ell_{0},\bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
-	% 	\varphi(0) = i) \nonumber
-	%  	%
-	% 	\\&\to \mathbb P(X(\tau_1^X) = y_{\ell_0}, \varphi(\tau_1^X) = j, \tau_{1}^X\leq E^\lambda 
-    %         	 \mid X(0) = x_0, \varphi(0) = i). \nonumber
-	% 	%
-	% \end{align}
-	At a boundary, for \(\ell_0\in\{-1,K+1\}\), 
-	\begin{align}
-		&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell, \varphi(\tau_1^{(p)}) = j, \tau_{1}^{(p)}\leq E^\lambda 
-            	 \mid L^{(p)}(0) = \ell_{0},\bs A^{(p)}(0)=1, \varphi(0) = i) \nonumber
-	 	%
-		\\&= \mathbb P(\bs X(\tau_1^X) = (y_\ell, j), \tau_{1}\leq E^\lambda 
-            	 \mid \bs X(0) = (0,i)), \label{eqn: lk78GHJK}
-	\end{align}
-	where \(\ell = 0\), \(i\in\calS_-\cup\calS_{-0}\) and \(j\in\mathcal S_+\) if \(\ell_0=-1\), and \(\ell = K\), \(i\in\calS_+\cup\calS_{+0}\) and \(j\in\calS_-\) if \(\ell_0=K+1\).  
-\end{cor}
-\begin{proof}
-	The proof follows the same structure as the proof of Theorem~\ref{thm: a thm!} however changes are required in all the results used in the proof, as here we do not need to integrate a function \(\psi\).
-Here we only give an outline of the proof.
-
-	At a boundary we can model the fluid queue exactly, hence (\ref{eqn: lk78GHJK}) holds.
-
-	Consider first \(i\in \calS_+,j\in\calS_+\).
-Partition the probability (\ref{eqn: 1421}) on the times \(\{\Sigma_n\}_{n\geq 1}\) and \(\{\Gamma_n\}_{n\geq 1}\), and specifically, partition on the event that there are exactly \(m\) events \(\{\Sigma_n\}_{n=1}^m\) and exactly \(m\) events \(\{\Gamma_n\}_{n=1}^m\).
-%Partition further by the phases at these times \(\varphi(\Sigma_n)=j_n\in\calS_-\), \(n=1,\dots,m\) and \(\varphi(\Gamma_n)=k_n\in\calS_+\), \(n=1,...,m\).
-	The resulting partitioned probabilities are
-	\begin{align}
-                 &\int_{x_1=0}^\infty \left(\bs e_i\bs H^{+-}(\lambda,x_1)\otimes \bs a_{\ell_0,i}^{(p)}(x_0)e^{\bs{S}^{(p)}x_1}\bs{D}^{(p)}\right)\wrt x_1 \nonumber
-            	\\&\nonumber \quad \Bigg[\prod_{r=1}^{m-1} \int_{x_{2r}=0}^\infty \left(\bs H^{-+}(\lambda,x_{2r}) \otimes e^{\bs{S}^{(p)}x_{2r}}\bs D^{(p)}\right)\wrt x_{2r} \\&\quad \int_{x_{2r+1}=0}^\infty \left(\bs H^{+-}(\lambda,x_{2r+1}) \otimes e^{\bs{S}^{(p)}x_{2r+1}}\bs D^{(p)}\right) \wrt x_{2r+1}\Bigg] \nonumber
-            	\\&
-            	\quad \int_{x_{2m}=0}^\infty \left(\bs H^{-+}(\lambda, x_{2m}) \otimes e^{\bs{S}^{(p)}x_{2m}}\bs{D}^{(p)}\right) \wrt x_{2m} \nonumber
-				\\&\quad \int_{x_{2m+1}=0}^\infty \left( \bs H^{++}\bs e_j \otimes (\lambda,x_{2m+1}) e^{\bs{S}^{(p)}x_{2m+1}}\bs s^{(p)} \right) \wrt x_{2m+1}.  \label{eqn: prob ofjaiv}
-	\end{align}
-
-	To show that the terms (\ref{eqn: prob ofjaiv}) converge to 
-	\begin{align}
-		&\mathbb P(\bs X(\tau_1^X)=(y_{\ell_0+1},j),\tau_1^X\leq E^\lambda, \Sigma_{m}\leq \tau_1^X<\Gamma_{m+1}, \mid \bs X(0)=(x_0, i))
-	\end{align}
-	we can use the bounds from Corollary~\ref{cor: cond bnd 2} and Corollry~\ref{cor: a cor}.
-For \(m=0\) we recognise (\ref{eqn: prob ofjaiv}) as the same form as that appearing in Corollary~\ref{cor: cond bnd 2} upon choosing \(v=0\).
-For \(m\geq 1\), choose the closing operator to be \(\bs v(x)=e^{\bs S x}\bs s\) and set \(x=0\) in Corollary~\ref{cor: a cor}.
-Now take the bound from Corollary~\ref{cor: a cor} and extend it to the case of matrix functions in the same way we extended Corollary~\ref{cor: asjdajaaaaa} to the matrix case in Lemma~\ref{lem: boobies}.
-In this way, we have a bound for (\ref{eqn: prob ofjaiv}) which tends to \(0\) as \(p\to\infty\).
-
-	% Corollary~\ref{cor: cond bnd} and Lemma~\ref{lem:tttttt} are analogous to Lemma~\ref{lem: Dcoajc} and Corollary~\ref{lem: Dcoajc} used in the proof of Lemma~\ref{lem:vn4}.
-They provide relevant bounds for the difference between terms like (\ref{eqn: prob ofjaiv}) for \(m=0\) and
-	% \begin{align*}
-	% 	&\mathbb P(X(\tau_1^X)=y_{\ell_0+1},\varphi(\tau_1^X)=j,\tau_1^X\leq E^\lambda, \tau_1^X<\Sigma_1 \mid X(0)=x_0, \varphi(0)=i)  
-	% 	\\&= h_{ij}^{++}(\lambda,y_{\ell_0+1}-x_0),
-	% \end{align*}
-	% for example.
-
-	% Corollaries~\ref{cor: a cor} and~\ref{cor: aaaaa} are analogous to Corollaries~\ref{eqn: lafkjebjcbbalbvvbrb} and~\ref{cor: fljm7778} used in the proof of Lemma~\ref{lem:vn4}.
-Corollaries~\ref{cor: a cor} and~\ref{cor: aaaaa} give error bounds for the difference between terms such as (\ref{eqn: prob ofjaiv}) for \(m\geq 1\) and the corresponding probability for the fluid queue,
-	% \begin{align}
-	% 	&\mathbb P(X(\tau_1^X)=y_{\ell_0+1},\varphi(\tau_1^X)=j,\tau_1^X\leq E^\lambda, \Sigma_{m}\leq \tau_1^X<\Gamma_{m+1}, \varphi(\Sigma_\ell) = j_\ell, \varphi(\Gamma_\ell)=k_\ell, \nonumber
-	% 	\\&\qquad{} \ell = 1,\dots, m  \mid X(0)=x_0, \varphi(0)=i)
-	% \end{align}
-	% which is given by (\ref{eqn: lst herwe}).
-
-	What reamins is a domination condition so that we may apply the Dominated Convergence Theorem to claim that the sum over the number of up-down and down-up transition converges (i.e.~the sum over \(m\) in (\ref{eqn: prob ofjaiv}) converges).
-To this end, since \(\left[\bs H^{++}(\lambda,x_{m+1})\right]_{ij}\leq G\), then
-	\begin{align}
-			&\int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
-		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
-		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
-		\int_{x_{m+1}=0}^\infty \bs H^{++}(\lambda,x_{m+1}) \bs e 
-		\\&\quad\bs   a_{\ell_0,i}(x_0)e^{\bs{S}x_1}\wrt x_1 \bs{D}e^{\bs{S}x_2} \wrt x_2 \bs{D}\dots e^{\bs{S}x_m} \wrt x_m \bs{D}e^{\bs{S}x_{m+1}}{\bs s} \wrt x_{m+1}\nonumber 
-		%
-			\\&\leq \int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
-		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
-		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
-		\int_{x_{m+1}=0}^\infty \bs e G
-		\\&\quad\bs   a_{\ell_0,i}(x_0)e^{\bs{S}x_1}\wrt x_1 \bs{D}e^{\bs{S}x_2} \wrt x_2 \bs{D}\dots e^{\bs{S}x_m} \wrt x_m \bs{D}e^{\bs{S}x_{m+1}}{\bs s} \wrt x_{m+1}\nonumber
-		%
-		\\&=\int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
-		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
-		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
-		 \bs e G
-		\\&\quad\bs   a_{\ell_0,i}(x_0)e^{\bs{S}x_1}\wrt x_1 \bs{D}e^{\bs{S}x_2} \wrt x_2 \bs{D}\dots e^{\bs{S}x_m} \wrt x_m \bs{D}{\bs e}\nonumber
-		%
-			\\&\leq\int_{x_1=0}^\infty \bs H^{+-}(\lambda,x_1)\nonumber
-		\int_{x_2=0}^\infty \bs H^{-+}(\lambda,x_2) 
-		\hdots \int_{x_m=0}^\infty \bs H^{-+}(\lambda, x_m) 
-		 \bs e G
-		\\&\quad \wrt x_1 \wrt x_2\dots \wrt x_m \label{eqn :ejrvn}
-	\end{align}
-	where the last inequality holds as we claimed in the discussion after (\ref{eqn :NNeeaefjn}) in the proof of Lemma~\ref{lem: gkjljklgagjklagsjlk}.
-Equation~(\ref{eqn :ejrvn}) is of a similar form to (\ref{eqn: bound ggggaaaa}) (they differ by a constant only), hence the same arguments used to bound (\ref{eqn: bound ggggaaaa}) can be applied to get the desired domination result.
-
-	Ultimately, we can apply the Dominated Convergence Theorem to prove that the sum of the partitioned probabilities (\ref{eqn: prob ofjaiv}) converges as \(p\to\infty\). The sum of the limits is 
-	\[\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j), \tau_{1}\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0,i)).\]
-
-	 The results for all other cases of \(i,j\in\calS\) follow analogously.
-\end{proof}
-
-Now, let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
+In this section we consider the embedded process formed by observing the QBD-RAP at changes of level. Now, let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
 \[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
@@
 Corollary~\ref{cor: aln22222212} states that the transition probabilities of the embedded processes converge.
-Thus the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the distribution of \(\{\Ydp(n)\}\) is tight.
+Thus the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the sequence of distributions of \(\{\Ydp(n)\}\) is tight.
 Thus we can establish a convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).

### 2021-12-22 chapter1/chapter.tex
first steps of numerics chapter
changes to ch4 done, sent to peter for review

@@
 %!TEX root = ../thesis.tex
-\chapter{Introduction \label{ch:intro}} 
+\chapter{Introduction \& Preliminaries \label{ch:intro}} 
+% THE ROUGH IDEA & CONTEXT
+% 	THE CONTEXT: FLUID-FLUID QUEUES
+%	PDEs 
+%	NON-NEGATIVITY & CONSERVATION 
+% MATHEMATICAL PRELIMINARIES
 A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\).
@@
 
-For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal A\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a matrix exponential distribution.
+For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal A\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a (possibility defective) matrix exponential distribution.
 %By CITE it is known that \(\mathcal A\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).

### 2022-01-13 chapter5/chapter.tex
.

@@
 %!TEX root = ../thesis.tex
-\chapter{Corollaries to Chapter~\ref{sec: conv}}\label{ch: global results}
-In this chapter we prove a number of results stemming from Theorem~\ref{thm: a thm!} in Chapter~\ref{sec: conv}.
-We consider the discrete-time embedded process formed by observing the QBD-RAP at changes of level only.
+\chapter{Global convergence results~\ref{sec: conv}}\label{ch: global results}
+In this chapter we prove a number of results which stem from the main results of Chapter~\ref{sec: conv}.
+We consider the discrete-time embedded processes formed, by observing the QBD-RAP at changes of level and, by observing the fluid queue at the hitting times of the level process at the of points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
 In Theorem~?
-we prove that the embedded process of the QBD-RAP converges in distribution to the discrete-time embedded process of the fluid queue formed by observing the fluid queue when it crosses the boundaries \(\{y_k\}\).
+we prove that the embedded process of the QBD-RAP converges in distribution to the embedded process of the fluid queue.
 In Theorem~?
@@
 
-In this section we consider the embedded process formed by observing the QBD-RAP at changes of level. Now, let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
+In this section we consider the embedded process formed by observing the QBD-RAP at changes of level. Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
 \[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
@@
 Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the change-of-level epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
-The subscript \(\bs \alpha^{(p)}\) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\).
+The subscript \(\bs \alpha \) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\).
 The process \(\{\Ydp(n)\}_{n\geq 0}\) is a discrete-time Markov chain, which is time-homogeneous for \(n\geq 1\).
@@
 Let \(\{\tau_n^X\}_{n\geq 0}\), be the sequence of (stopping) times with \(\tau_0^X=0\), and 
-\[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(0)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(0)=y_{K+1}\right\} \end{array} \right\}.\]
-For \(n\geq 1\), \(\tau_n^X\) is the time at which \(X(t)\) either changes band, or hits a boundary, or the process leaves a boundary, for the \(n\)th time.
+\[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(0)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(0)=y_{K+1}\right\} \end{array} \right\},\]
+for \(n\geq 0\).
+For \(n\geq 1\), \(\tau_n^X\) is the time at which \(X(t)\) either changes band, or hits a boundary, or leaves a boundary, for the \(n\)th time.
 The embedded process \(\{\bs X(\tau_n)\}\) is a discrete-time Markov chain which is time-homogeneous for \(n\geq 1\).
@@
 We have the following result on the convergence of the embedded processes \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\), which we will utilise later to prove a global result.
-\begin{cor}\label{cor: aln222222} For \(\ell,\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
+\begin{cor}\label{cor: aln222222} For \(\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
 	for \(n=0\), then
@@
 	\end{align}
+
+	Further, for \(i\in\mathcal S_-\cup\calS_{-0}\), \(j\in\calS_+\), \(n\geq 0\), 
+	\begin{align}
+		&\mathbb P(\Ydp(n+1) = (0,j), \tau_{n+1}^{(p)}\leq E^\lambda
+            	 \mid \Ydp(n) = (-1,i), \tau_n^{(p)}\leq E^\lambda) \nonumber
+	 	%
+		\\&\to  
+			\mathbb P(\bs X(\tau_{n+1}^X) = (0^+, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (0,i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd22}
+	\end{align}
+	and for \(i\in\mathcal S_+\cup\calS_{+0}\), \(j\in\calS_-\), \(n\geq 0\),
+	\begin{align}
+		&\mathbb P(\Ydp(n+1) = (K,j), \tau_{n+1}^{(p)}\leq E^\lambda
+            	 \mid \Ydp(n) = (K+1,i), \tau_n^{(p)}\leq E^\lambda) \nonumber
+	 	%
+		\\&\to  
+			\mathbb P(\bs X(\tau_{n+1}^X) = (y_{K+1}^-, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{K+1},i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd3}
+	\end{align}
 \end{cor}
 \begin{proof}
+The convergence in (\ref{eqn: ddddddd22}) and (\ref{eqn: ddddddd3}) holds trivially as the QBD-RAP and fluid queue have the same behaviour at the sticky boundary.
+
 The case for \(n=0\) is a direct result of Corollary~\ref{cor: aln222}.
@@
 
-For \(n\geq 1\), consider the transition probabilities of the embedded process from the QBD-RAP, 
+All that is left is to prove (\ref{eqn: ddddddd}). For \(n\geq 1\), consider the transition probabilities of the embedded process from the QBD-RAP, 
 \begin{align}
@@
 	\\&=\mathbb P(\Ydp(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+), j), \tau_{1}^{(p)}\leq E^\lambda
-	\mid \bs Y^{(p)}(0)=(\ell_0 ,\bs\alpha, i)) \label{eqn: aaannnn}
+	\mid \bs Y^{(p)}(0)=(\ell_0 ,\bs\alpha, i)), \label{eqn: aaannnn}
 \end{align}
-% since the process is time-homogeneous and the exponential random variable \(E^\lambda\) is memoryless. The probability on the right of (\ref{eqn: aaannnn}) is equal to
+since the QBD-RAP is time-homogeneous and the exponential random variable \(E^\lambda\) is memoryless. %The probability on the right of (\ref{eqn: aaannnn}) is equal to
 % \begin{align}
@@
 % Thus, once again noting that \(\tau_1^{(p)}<\infty\) almost surely, then taking \(\lambda \to 0\) in 
-Now, applying Corollary~\ref{cor: aln222} to (\ref{eqn: aaannnn}) states that 
+Applying Corollary~\ref{cor: aln222} to (\ref{eqn: aaannnn}) then
 \begin{align}
@@
 \end{align}
-Since the fluid queue is time-homogeneous, and \(E^\lambda\) memoryless, then 
+Since the fluid queue is time-homogeneous, and \(E^\lambda\) memoryless, then (\ref{eqnL sdfdfaaa}) is equal to 
 \begin{align}
@@
 \end{align}
-which prove the result. 
+which proves the result. 
 \end{proof}
 A direct corollary of Corollary~\ref{cor: aln222222} is the convergence of the transition probabilties of the embedded process. 
-\begin{cor}\label{cor: aln22222212} For \(\ell,\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
+\begin{cor}\label{cor: aln22222212} For \(\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
 		for \(n=0\), then
@@
 Thus the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the sequence of distributions of \(\{\Ydp(n)\}\) is tight.
-Thus we can establish a convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).
+Thus, we can establish the convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).
 
 Another direct corollary of Corollary~\ref{cor: aln222} is the convergence in distribution of the random variables \(\{\tau_1^{(p)}\}_{p\geq 1}\) to \(\tau_1^X\). 
-\begin{cor}
+\begin{cor}\label{cor: short cor}
 	The random variables \(\{\tau_1^{(p)}\}_{p\geq 1}\) converge in distribution to \(\tau_1^X\). 
@@
 
-\section{A global result}\label{sec: local to global}
-The rest of this Chapter is dedicated to proving global convergence results of the QBD-RAP approximation to the fluid queue.
-The structure of the argument is to first show a convergence result of the Laplace transform with respect to time of the distribution of the QBD-RAP and fluid queue and the \(n\)th change of level.
-We then prove a convergence result for the Laplace trasforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue between the \(n\)th and \(n+1\)th change of level.
-Summing over the number of changes of level, \(n\), and via the Dominated Convergence Theorem, we claim that the Laplace trasforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue converge.
+\section{Convergence of the QBD-RAP scheme}\label{sec: local to global}
+The rest of this chapter is dedicated to proving global convergence results of the QBD-RAP approximation scheme to the fluid queue.
+Ultimately, we prove the weak convergence (in space and time) of the QBD-RAP approximation scheme to the distribution of the fluid queue.
+The structure of the argument is to first show a convergence result of the Laplace transform with respect to time of the distribution of the QBD-RAP and fluid queue at the \(n\)th change of level.
+We then prove a convergence result for the Laplace transforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue between the \(n\)th and \(n+1\)th change of level.
+Summing over the number of changes of level, \(n\), and via the Dominated Convergence Theorem, we claim that the Laplace transforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue converge.
 Lastly, we apply the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} to claim a weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
@@
 	 \\&=\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^{\lambda}
-	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}),\nonumber 
+	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}),\label{eqn: aaaaaaccccc} 
 	%
@@
 % \end{align}
-which is the Laplace transform of the time until the \(n\)th change of level of the QBD-RAP on the event that the level and phase at the \(n\)th change of level are \(\ell\) and \(j_n\), respectively, given that the initial level and phases are \(\ell_0\) and \(i\), respectively and the initial orbit is \(\bs a_{\ell_0,i}^{(p)}(x_0)\).
-Partitioning on the time of the first change of level, \(\tau_1\), and the level and phase at this time gives
+which is the Laplace transform of the time until the \(n\)th change of level of the QBD-RAP on the event that the level and phase at the \(n\)th change of level are \(\ell\) and \(j_n\), respectively, given that the initial level and phases are \(\ell_0\) and \(i\), respectively, and the initial orbit is \(\bs a_{\ell_0,i}^{(p)}(x_0)\).
+Partitioning on the time of the first change of level, \(\tau_1\), and the level and phase at this time, then (\ref{eqn: aaaaaaccccc}) is equal to
 \begin{align}
@@
 \end{align}
-An application of Corollary \ref{cor: aln222} to the expression on the second line of (\ref{eqn: mnebrb2}) states, for \(i\in\calS\), \(j\in\calS\), \(\ell_0\in\mathcal K\), 
+An application of Corollary \ref{cor: aln222} to the expression on the second line of (\ref{eqn: mnebrb2}) states, 
 \begin{align}
@@
             	 \mid \bs X(0) = (x_0, i)) & \text{for }j_1\in\calS_-,\, \ell_1=\ell_0-1. 
-\end{cases}\nonumber 
+\end{cases},\nonumber 
 \end{align}
+for \(i\in\calS\), \(j_1\in\calS_+\cup\calS_-\), \(\ell_0\in\mathcal K\), \(x_0\in\calD_{\ell_0}\).
 
-Now, for a given \(j_1\) and \(\ell_1\) consider 
+We now turn our attention to the first factor in the summands of (\ref{eqn: mnebrb2}). For a given \(j_1\in\calS_+\cup\calS_-\) and \(\ell_1\in\mathcal K\),  consider
 \begin{align}
@@
 \end{align}
-by the time-homogeneous property of the QBD-RAP and the the memoryless property of the exponential distribution.
+by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution.
 The left-hand side of Equation (\ref{eqn: dskvnaSF}) appears as the first factor in the summands of (\ref{eqn: mnebrb2}).
@@
 \begin{align}\label{eqn: paths set1}
-	\mathcal P^n(\ell_0,\ell_n)&=\{(\ell_1,\dots,\ell_{n-1}) \in \mathcal K^{n-1}\mid |\ell_{r-1}-\ell_r|=1,r = 1,\dots,n\}.
+	\mathcal P^n(\ell_0,\ell_n)&=\{(\ell_1,\dots,\ell_{n-1}) \in \mathcal K^{n-1}\mid |\ell_{r-1}-\ell_r|=1 \mbox{ or } \ell_{r-1}=\ell_r=0 \mbox{ or } \ell_{r-1}=\ell_r=K,r = 1,\dots,n\}.
 \end{align}
@@
 % by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution. 
-We can apply Corollary (\ref{cor: aln222}) to the terms (\ref{eqn: 161222}) and conclude 
+We can apply Corollary (\ref{cor: aln222}) to the factors of the summands in (\ref{eqn: 161222}) and conclude 
 \begin{align}
@@
 	 % 
+	 \mathbb P(\bs X(\tau_1^X) =( 0,j_m), \tau_{1}^X\leq E^\lambda 
+            	 \mid \bs X(0) = (x_0, j_{m-1}))  
+	 &\text{ for }\begin{cases} j_m\in\calS_+,\\ \ell_m=\ell_{m-1}=0,\end{cases}\\
+	 %
 	 %
@@
 	\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}}, j_m), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0, j_{m-1}))  & \text{ for }\begin{cases}j_m\in\calS_-,&\\ \ell_m=\ell_{m-1}-1.&\end{cases}
+		\mid \bs X(0) = (x_0, j_{m-1}))  & \text{ for }\begin{cases}j_m\in\calS_-,&\\ \ell_m=\ell_{m-1}-1.&\end{cases}\\
+	% 
+	\mathbb P(\bs X(\tau_1^X) = (y_{K+1}, j_m), \tau_{1}^X\leq E^\lambda 
+		\mid \bs X(0) = (x_0, j_{m-1}))  & \text{ for }\begin{cases}j_m\in\calS_-,&\\ \ell_m=\ell_{m-1}=K.&\end{cases}
 \end{cases}\label{eqn:kknvf fawoprgj v2}
@@
 
-By the time-homogeneous property of the fluid queue and the memoryless property of the exponential distribution, 
+By the time-homogeneous property of the fluid queue and the memoryless property of the exponential distribution, (\ref{eqn:kknvf fawoprgj v2}) is equal to 
 \begin{align}
-	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
+	\mathbb P(\bs X(\tau_m^X) =( y_{\ell_{m-1}+1},j_m), \tau_{m}^X\leq E^\lambda 
+            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda )  \label{eqn: lim 1}
+\end{align}
+for \(j_m\in\calS_+, \ell_m=\ell_{m-1}+1\), 
+\begin{align}			
+	 % 
+	 \mathbb P(\bs X(\tau_m^X) =( 0,j_m), \tau_{m}^X\leq E^\lambda 
+            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda )
+\end{align}
+for \(j_m\in\calS_+, \ell_m=\ell_{m-1}=0\), 
+\begin{align}
+	%
+	\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}}, j_m), \tau_{m}^X\leq E^\lambda 
+		\mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda ) 
+\end{align}
+for \(j_m\in\calS_-, \ell_m=\ell_{m-1}-1\), and 
+\begin{align}
+	% 
+	\mathbb P(\bs X(\tau_m^X) = (y_{K+1}, j_m), \tau_{m}^X\leq E^\lambda 
+		\mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda )   \label{eqn: lim 4}
+\end{align}
+for \(j_m\in\calS_-, \ell_m=\ell_{m-1}=K\).
+% \begin{align}
+% 	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{1}^X\leq E^\lambda 
+%             	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
+% 	%  \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{m}^X\leq E^\lambda 
+%     %         	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}))\nonumber 
 % 	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{m}^X\leq E^\lambda 
-    %         	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}))\nonumber 
-	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1}, j_m), \tau_{m}^X\leq E^\lambda 
-            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \nonumber 
-	\tau_{m-1}^X\leq E^{\lambda}).
-\end{align}
-Similarly, 
-\begin{align}
-	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}}, j_m), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
-	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}}, j_m), \tau_{m}^X\leq E^\lambda 
-            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda).
-\end{align}
+%             	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \nonumber 
+% 	\tau_{m-1}^X\leq E^{\lambda}).
+% \end{align}
+% Similarly, 
+% \begin{align}
+% 	&\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}}, j_m), \tau_{1}^X\leq E^\lambda 
+%             	 \mid \bs X(0) = (x_0, j_{m-1})) \nonumber
+% 	 \\&=\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}}, j_m), \tau_{m}^X\leq E^\lambda 
+%             	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda).
+% \end{align}
 
-Now, taking the limit of (\ref{eqn: 161222}) gives 
+Taking the limit of (\ref{eqn: 161222}) gives 
 \begin{align}
@@
 where we may swap the limit and the sums as they are finite, and we can swap the limit and the product since all the limits exist and the product is finite.
-Substituting the limits into (\ref{eqn: 161222b}) gives
+Substituting the limits (\ref{eqn: lim 1})-(\ref{eqn: lim 4}) into (\ref{eqn: 161222b}) gives
 \begin{align}
@@
 		j_{1}), 
-		\tau_1^X\leq E^\lambda).
+		\tau_1^X\leq E^\lambda),
 \end{align}
+by the Law of total probability.
 
-Therefore, taking the limit as \(p\to \infty \) of (\ref{eqn: mnebrb2}) 
+Returning to (\ref{eqn: mnebrb2}) and taking the limit as \(p\to \infty \),  
 \begin{align}
@@
 
-Therefore we have proved the following result 
+Hence, we have proved the following result 
 \begin{lem}\label{lem: kKKJJJF}
@@
 		\mid \bs Y^{(p)}(0)=\bs y_0^{(p)}),\nonumber 
-		%  &\mathbb P(L^{(p)}(\tau_n^{(p)}) = \ell, \phi^{(p)}(\tau_n^{(p)}) = j_n, \tau_{n}^{(p)}\leq E^{\lambda}
-		%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \phi^{(p)}(0)=i),\nonumber  
 		%
@@
 \end{lem} 
+Lemma~\ref{lem: kKKJJJF} is what is required to proceed with the proof of convergence of the QBD-RAP scheme.
+However, at this point we also point out the following corollary, which may be of interest in other contexts.
+\begin{cor}
+	The random variables \(\{\tau_n^{(p)}\}_{p\geq 1}\) converge in distribution to \(\tau_n^X\). 
+\end{cor}
+\begin{proof}
+	The proof follows the same arguments as the proof of Corollary~\ref{cor: short cor} -- in Lemma~\ref{lem: kKKJJJF} sum over \(\ell\) and \(j_n\), then apply the Extended Continuity Theorem \cite[Chapter XIII, Theorem 2a]{feller1957}.
+\end{proof}
 
@@
 	\nonumber&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid 
-	\bs Y^{(p)}(0)=\bs y_0^{(p)})\psi(x)\wrt t.
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\psi(x)\wrt t,
 	% \nonumber&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_\ell}\mathbb P(L^{(p)}(t) = \ell, \widetilde X^{(p)}(t)\in\wrt x, \phi^{(p)}(t) = j, \tau_{n}^{(p)}\leq t<\tau_{n+1}^{(p)} \mid L^{(p)}(0)=\ell_0,
@@
 \end{align}
+where \(\psi:\mathbb R\to \mathbb R\) is a bounded function.
+
 Partitioning on the time of the \(n\)th change of level and the phase and level at this time gives
 \begin{align}
-	&\int_{t=0}^\infty e^{-\lambda t} \int_{u_n=0}^t\sum_{j_n\in\mathcal S}
-	\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
+	&\int_{t=0}^\infty e^{-\lambda t} 
+	\int_{x\in\calD_{\ell,j}}
+	\int_{u_n=0}^t\sum_{j_n\in\mathcal S} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
 	t \in(u_n, \tau_{n+1}^{(p)}] \mid\Ydp(n) = (\ell,j_n), \nonumber 
@@
 \end{align}
-Since the fluid queue is time-homogeneous and \(E^\lambda\) memoryless, then we can write (\ref{eqn: lkHAHJFHKJ J J H}) as 
+Since the fluid queue is time-homogeneous and \(E^\lambda\) memoryless, then (\ref{eqn: lkHAHJFHKJ J J H}) is equal to 
 \begin{align}
@@
 	t\in\mathcal T_0^{(p)} \mid \nonumber 
-	 \bs Y(0) = (\ell,\bs\alpha,j_n))\wrt t 
-	 \\&\qquad{}\psi(x)  \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	 \bs Y(0) = (\ell,\bs\alpha,j_n))\wrt t \psi(x)
+	 \\&\qquad{}  \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
 	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}).\label{eqn: eieieie}
-	 %
-	%  \\&= \sum_{j_n\in\mathcal S}
-	% \lim_{p\to\infty} \int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t}  \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x,j), 
-	% t\in\mathcal T_n^{(p)} \mid \nonumber 
-	%  \Ydp(n) = (\ell,j_n), 
-	%   \tau_{n}^{(p)}= 0)\wrt t \\&\qquad{} \psi(x) 
-	%  \lim_{p\to\infty} \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
-	%  \mid \bs Y^{(p)}(0)=(\ell_0, \bs  a_{\ell_0,i}^{(p)}(x_0),i))\nonumber
-\end{align}
-Since the sum is finite and the limits exist then substitute in the limiting values, (\ref{eqn: lkHAHJFHKJ J J 22}), and (\ref{eqn: ashverkjbvjhkhdsknacva}), to get 
-\begin{align}
-	 &\sum_{j_n\in\mathcal S}\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} \mid \bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n),\nonumber
-	 \tau_n^X\leq E^\lambda ) \psi(x)
-	 \\&\quad{} \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
+	 \\&=\sum_{j_n\in\mathcal S}\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} \mid \bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n),\nonumber
+	 \tau_n^X\leq E^\lambda ) 
+	 \\&\quad{} \psi(x)\mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
 		j_n), \tau_{n}^X\leq E^\lambda
@@
 	\\&= \int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} 
-	\mid X(0)=x_0, \varphi(0)=i)\psi(x)\wrt t.
+	\mid X(0)=x_0, \varphi(0)=i)\psi(x)\wrt t,
 \end{align}
+where the first equality holds since the sum is finite and the limits exist, and the second equality holds from the law of total probability.
 
-Hence we have shown the following result 
+Hence, we have shown the following result 
 \begin{lem}\label{lem: LAkAKFnvnb mav h}
@@
 Now, (\ref{eqn:GHJKLKJHGHJKL}) is an infinite sum of terms appearing in Lemma~\ref{lem: LAkAKFnvnb mav h}.
-Hence, upon taking the limit of (\ref{eqn:GHJKLKJHGHJKL}), if we can justify the swap of the sum and the limit, we will obstain the result we desire.
-The main result of this subsection is Lemma~\ref{lem: another bound 2} in which we show a bound which can serve as a dominating function so that we may apply the Dominated Convergence Theorem to justify the swap.
+Hence, upon taking the limit of (\ref{eqn:GHJKLKJHGHJKL}), if we can justify the swap of the sum and the limit, we will obtain the desired result.
+%The main result of this subsection is Lemma~\ref{lem: another bound 2} in which we show a bound which can serve as a dominating function so that we may apply the Dominated Convergence Theorem to justify the aforementioned swap of sum and limit.
 
@@
 
-Let \(\mathcal R(L(t),\bs A(t),\varphi(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\varphi(t)}(x)\), \(x\in\calD_{L(t),\varphi(t)}\). 
+Let \(\mathcal R(L(t),\bs A(t),\varphi(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\varphi(t)}(x)\), \(x\in\calD_{L(t),\varphi(t)}\). Observe that 
+\begin{align}
+	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \nonumber
+	\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t 
+	\\&=\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y}^{(p)}(t) = (\ell,\wrt x,j) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)})  \psi(x)\wrt t.
+\end{align}
+
 \begin{cor}\label{cor: lk}
-	Let \(\psi: \mathbb R\times \mathcal S \to \mathbb R\) be an arbitrary bounded function with \(|\psi(\cdot)|\leq F\).
+	Let \(\psi: \mathbb R\times \mathcal S \to \mathbb R\) be an arbitrary, bounded function with \(|\psi(\cdot)|\leq F\).
 %Further, let \(\mathcal L(X(t)) = \sum_{k\in\mathcal K}k1(X(t)\in\mathcal D_k)\) and \(\widetilde{\bs X}(t) = (\mathcal L(X(t)),X(t),\varphi(t)))\).
@@
 \end{thm}
-Thus, by the the Extended Continuity Theorem~\ref{thm: ext cont thm}
+Thus, by the Extended Continuity Theorem~\ref{thm: ext cont thm}
 \begin{align}
@@
 \end{align}
-weakly in \(t\) as \(p\to \infty\). Now, 
+weakly in \(t\) as \(p\to \infty\).
+
+Ideally, we would like to obtain a point-wise convergence result in the variable \(t\). However, to date, this has eluded me. Although 
 \[\mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0,i) \right]\] 
-is a continuous function of \(t\) (it is a Feller semi-group), moreover, for \(p<\infty\), 
+is a continuous function of \(t\) (it is a Feller semigroup), and, for \(p<\infty\), 
 \[\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \]
- is also continuous in \(t\). However, the limit 
+ is continuous in \(t\), the limit 
  \[\lim\limits_{p\to\infty}\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) ) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \] need not be continuous in \(t\).
@@
 \end{align}
-for all \(p\geq p_0(t,\varepsilon)\). This is an area for future work. Alternatively, the theory of \emph{killing} of a stochastic process may be applicable and provide the convergence we seek.
+for all \(p\geq p_0(t,\varepsilon)\).
 
@@
 Throughout, we have assumed that all intervals are of width \(\Delta\), i.e.~\(|y_{\ell+1}-y_\ell|=\Delta\), and that on every interval the dynamics of the fluid queue are modelled based on the same matrix exponential representation \((\bs \alpha, \bs S, \bs s)\).
-These assumptions are, in fact, not necessary, but they do serve to simplify the presentation.
+These assumptions are, in fact, not necessary, but they do serve to simplify the presentation slightly.
 The convergence results can be extended to use different sequences of matrix exponential representations on each interval, provided that for each sequence of matrix exponential distributions, the variance tends to \(0\).
@@
 
-The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, i.e.~the hypotheses of the Lemma depend only on the interval \(\calD_{\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval and not on any other interval.
-Thus Lemma~\ref{thm: a thm!} holds independently on each interval, as does Corollary~\ref{cor: aln222}.
+The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, i.e.~the hypotheses of the Theorem depend only on the interval \(\calD_{\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval, and not on any other interval.
+Thus, Lemma~\ref{thm: a thm!} holds independently on each interval, as does Corollary~\ref{cor: aln222}.
 
 Let the width of an interval \(\mathcal D_{\ell_0}\) be \(\Delta_{\ell_0}=y_{\ell_0+1}-y_{\ell_0}\) and suppose that sequence of matrix exponential random variables used to model the dynamics of the fluid queue on the interval \(\calD_{\ell_0}\) is \(Z_{\ell_0}^{(p)}\).
-Regarding Lemma~\ref{lem: another bound}, we can extend it the following version,
+Regarding the domination condition in Lemma~\ref{lem: another bound}, we can extend it the following version,
 \begin{lem}\label{lem: another bound sdfg}
@@
 	For the proof of (\ref{eqn: ell0 version}) follow the same arguments as in the proof of Lemma~\ref{lem: another bound}.
-The bound in (\ref{eqn: skjhg}) follows by the assumptions in the statement of the result.
-With the bound (\ref{eqn: skjhg}), then
+The bound in (\ref{eqn: skjhg}) follows by the assumptions in the statement of the result, since
 	\begin{align*}

### 2022-01-21 appendix4/appendix.tex
.

@@
 \chapter{Kronecker properties}\label{appendix: kronecker}
-Here we detail some properties of Kronecker sum, products, and exponentials (see \citep{MEinAP}, Appendix A.4).
+Here we detail some properties of Kronecker sums, products, and exponential (see \citep{MEinAP}, Appendix A.4).
 
@@
 \bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
-be matrices. The operator \(\otimes\) is the Kronecker product of two matrices; 
+be matrices with dimensions \(n\times m\) and \(n'\times m'\), respectively. The operator \(\otimes\) is the Kronecker product of two matrices; 
 \[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
@@
 \[\bs A\oplus \bs{B} := \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]
+
+The exponential of a square matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}\bs B^n.\]
 
@@
 	%
-	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}  \label{eqn:lstsimplify}\end{align}
+	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}.  \label{eqn:lstsimplify}\end{align}
 \end{lem}
@@
 
-The exponential of a matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}B^n.\]
-
-\cite{ln2015} show the following.
-\begin{lem}\label{lem: sfkjgn}
+\begin{lem}[\cite{ln2015}]\label{lem: sfkjgn}
 	Let \(\bs B\) be the block-partitioned matrix
@@
 \end{lem}
+\begin{proof}
+	See \cite{ln2015}.
+\end{proof}
 
@@
 
+Now, recall the matrix-functions
+\begin{align*}
+	% \bs Q_{+0}(\lambda) &= \bs C_+^{-1}\bs T_{+0}\left[\lambda \bs I - \bs T_{00}\right]^{-1},
+	% %
+	% \\\bs Q_{-0}(\lambda) &= \bs C_-^{-1}\bs T_{-0}\left[\lambda \bs I - \bs T_{00}\right]^{-1},
+	%
+	\bs Q_{++}(\lambda) &= \bs C_+^{-1} \left(\bs T_{++} - \lambda \bs I + \bs T_{+0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0+}\right),
+	%
+	\\\bs Q_{+-}(\lambda) &= \bs C_+^{-1} \left(\bs T_{+-} + \bs T_{+0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0-} \right) ,
+	%
+	\\\bs Q_{--}(\lambda) &= \bs C_-^{-1} \left(\bs T_{--}  - \lambda \bs I + \bs T_{-0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0-}\right),
+	%
+	\\\bs Q_{-+}(\lambda) &=\bs C_-^{-1} \left(\bs T_{-+}+ \bs T_{-0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0+}\right) ,
+\end{align*}
+from Chapter~\ref{sec: conv}.
 \begin{cor}\label{cor: mpr B}
@@
 	%
-	\\&= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x,\vligne{\bs I_{pn} & \bs 0_{pn\times p|\calS_0|}} \label{eqn: akgj987ad}
+	\\&= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x,\vligne{\bs I_{pn} & \bs 0_{pn\times p|\calS_0|}}, \label{eqn: akgj987ad}
 \end{align}

### 2022-01-21 chapter5/chapter.tex
.

@@
 %!TEX root = ../thesis.tex
-\chapter{Global convergence results~\ref{sec: conv}}\label{ch: global results}
+\chapter{Global convergence results}\label{ch: global results}
 In this chapter we prove a number of results which stem from the main results of Chapter~\ref{sec: conv}.
-We consider the discrete-time embedded processes formed, by observing the QBD-RAP at changes of level and, by observing the fluid queue at the hitting times of the level process at the of points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
-In Theorem~?
-we prove that the embedded process of the QBD-RAP converges in distribution to the embedded process of the fluid queue.
-In Theorem~?
-we prove that distribution of the sojourn time of the QBD-RAP in a given level converges to distribution of the sojourn time of the fluid queue in the corresponding interval.
-In Theorem~?, we prove that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.
-Initially we set out only to prove Theorem~?, but Theorems~?
-and ?
-appeared as helpful results along the way.
+We consider the discrete-time embedded processes formed by observing the QBD-RAP at the orbit restart epochs and by observing the fluid queue at the hitting times of the level process at the of points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
+In Corollary~\ref{cor: aln22222212} we prove that the transition probability of the embedded process of the QBD-RAP converge those of the embedded process of the fluid queue.
+In Corollary~\ref{cor: short cor} we prove that distribution of the sojourn time of the QBD-RAP in a given level converges to distribution of the sojourn time of the fluid queue in the corresponding interval.
+In Section~\ref{sec: sub global}, we state global results on the weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
 
@@
 
-In this section we consider the embedded process formed by observing the QBD-RAP at changes of level. Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
+In this section we consider the embedded process formed by observing the QBD-RAP at the orbit restart epochs. Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
 \[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
-be the (stopping) times at which \(\{L^{(p)}(t)\}\), the level process of the QBD-RAP, changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary.
+be orbit restart epochs.
+These are the (stopping) times at which \(\{L^{(p)}(t)\}\), the level process of the QBD-RAP, changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary.
 To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed.
-Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the change-of-level epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
+Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the orbit restart epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
 The subscript \(\bs \alpha \) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\).
@@
 We have the following result on the convergence of the embedded processes \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\), which we will utilise later to prove a global result.
-\begin{cor}\label{cor: aln222222} For \(\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
+\begin{cor}\label{cor: aln222222} For \(\ell_0\in\mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
 	for \(n=0\), then
 	\begin{align}
-		&\mathbb P(\Ydp(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j), \tau_1^{(p)}\leq E^\lambda
+		&\mathbb P(\Ydp(1) = (\ell(\ell_0,j),j), \tau_1^{(p)}\leq E^\lambda
             	 \mid \bs Y^{(p)}(0) = (\ell_0,\bs a_{\ell_0,i}^{(p)}(x_0),i)) \nonumber
@@
 		\\&\to  
-			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_1^{X}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\label{eqn: lim 1}
+			\mathbb P(\bs X(\tau_1^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_1^{X}\leq E^\lambda \mid \bs X(0) = (x_0,i)),\label{eqn: lim 1}
 	\end{align}
@@
 	\begin{align}
-		&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j), \tau_{n+1}^{(p)}\leq E^\lambda
+		&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j), \tau_{n+1}^{(p)}\leq E^\lambda
             	 \mid \Ydp(n) = (\ell_0,i), \tau_n^{(p)}\leq E^\lambda) \nonumber
@@
 		\\&\to  
-			\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd}
+			\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd}
 	\end{align}
-
-	Further, for \(i\in\mathcal S_-\cup\calS_{-0}\), \(j\in\calS_+\), \(n\geq 0\), 
+	where \(\ell(\ell_0,j)\) can take values
 	\begin{align}
-		&\mathbb P(\Ydp(n+1) = (0,j), \tau_{n+1}^{(p)}\leq E^\lambda
-            	 \mid \Ydp(n) = (-1,i), \tau_n^{(p)}\leq E^\lambda) \nonumber
+		\ell(\ell_0,j) = \begin{cases}
+			\ell_0-1, & \ell_0\in\{0,1,\dots,K+1\},\, j\in\calS_-\\
 			%
-		\\&\to  
-			\mathbb P(\bs X(\tau_{n+1}^X) = (0^+, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (0,i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd22}
+			\ell_0, & \ell_0 = 0, j\in\mathcal S_+, \mbox{ or }\ell_0 = K, j\in\mathcal S_-,\\
+			%
+			\ell_0+1, & \ell_0\in\{-1,0,1,\dots,K\},\, j\in\calS_+.
+		\end{cases}
 	\end{align}
-	and for \(i\in\mathcal S_+\cup\calS_{+0}\), \(j\in\calS_-\), \(n\geq 0\),
-	\begin{align}
-		&\mathbb P(\Ydp(n+1) = (K,j), \tau_{n+1}^{(p)}\leq E^\lambda
-            	 \mid \Ydp(n) = (K+1,i), \tau_n^{(p)}\leq E^\lambda) \nonumber
-	 	%
-		\\&\to  
-			\mathbb P(\bs X(\tau_{n+1}^X) = (y_{K+1}^-, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{K+1},i), \tau_n^{X}\leq E^\lambda).\label{eqn: ddddddd3}
-	\end{align}
+	% and 
+	% \begin{align}
+	% 	y(\ell,\ell_0,j) = \begin{cases}
+	% 		y_{\ell_0}, & \ell=\ell_0-1,\, j\in\calS_-, \mbox{ or } \ell=\ell_0=0,\, j \in\calS_+, \\
+	% 		y_{\ell_0+1}, & \ell=\ell_0+1,\, j\in\calS_+, \mbox{ or } \ell=\ell_0=K,\, j \in\calS_-.
+	% 	\end{cases}
+	% \end{align}
 \end{cor}
 \begin{proof}
-The convergence in (\ref{eqn: ddddddd22}) and (\ref{eqn: ddddddd3}) holds trivially as the QBD-RAP and fluid queue have the same behaviour at the sticky boundary.
+The convergence for \(\ell_0\in\{-1,K+1\}\) holds trivially as the QBD-RAP and fluid queue have the same behaviour at the sticky boundary.
 
@@
 \begin{align}
-	&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j),\tau_{n+1}^{(p)}\leq E^\lambda
+	&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j),\tau_{n+1}^{(p)}\leq E^\lambda
 	\mid \Ydp(n) = (\ell_0,i),\tau_{n}^{(p)}\leq E^\lambda) \nonumber 
-	\\&=\mathbb P(\Ydp(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+), j), \tau_{1}^{(p)}\leq E^\lambda
+	\\&=\mathbb P(\Ydp(1) = (\ell(\ell_0,j),j), \tau_{1}^{(p)}\leq E^\lambda
 	\mid \bs Y^{(p)}(0)=(\ell_0 ,\bs\alpha, i)), \label{eqn: aaannnn}
@@
 \begin{align}
-	&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j),\tau_{n+1}^{(p)}\leq E^\lambda
+	&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j),\tau_{n+1}^{(p)}\leq E^\lambda
 	\mid \Ydp(n) = (\ell_0,i),\tau_{n}^{(p)}\leq E^\lambda) \nonumber 
-	\\&\to \mathbb P(\bs X(\tau_{1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(\tau_{1}^X) = (y_{\ell_0+1(i\in\calS_-)},i)).\label{eqnL sdfdfaaa}
+	\\&\to \mathbb P(\bs X(\tau_{1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(0) = (y_{\ell_0+1(i\in\calS_-)},i)).\label{eqnL sdfdfaaa}
 \end{align}
@@
 \begin{align}
-	&\mathbb P(\bs X(\tau_{1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(\tau_{1}^X) = (y_{\ell_0+1(i\in\calS_-)},i))
-	\\&=\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_{n}^{X}\leq E^\lambda ),
+	&\mathbb P(\bs X(\tau_{1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{1}^{X}\leq E^\lambda \mid \bs X(\tau_{1}^X) = (y_{\ell_0+1(i\in\calS_-)},i)) \nonumber
+	\\&=\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j), \tau_{n+1}^{X}\leq E^\lambda \mid \bs X(\tau_{n}^X) = (y_{\ell_0+1(i\in\calS_-)},i), \tau_{n}^{X}\leq E^\lambda ),
 \end{align}
@@
 \end{proof}
-A direct corollary of Corollary~\ref{cor: aln222222} is the convergence of the transition probabilties of the embedded process. 
-\begin{cor}\label{cor: aln22222212} For \(\ell_0\in \mathcal K \setminus \{-1,K+1\},\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
+A direct corollary of Corollary~\ref{cor: aln222222} is the convergence of the transition probabilities of the embedded process. 
+\begin{cor}\label{cor: aln22222212} For \(\ell_0\in \mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
 		for \(n=0\), then
 		\begin{align}
-			&\mathbb P(\Ydp(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j) 
+			&\mathbb P(\Ydp(1) = (\ell(\ell_0,j),j)
 					 \mid \bs Y^{(p)}(0) = (\ell_0,\bs a_{\ell_0,i}^{(p)}(x_0),i)) \nonumber
@@
 			\\&\to  
-				\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1(j\in\calS_+)}, j)  \mid \bs X(0) = (x_0,i)).\label{eqn: lim 2}
+				\mathbb P(\bs X(\tau_1^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j)  \mid \bs X(0) = (x_0,i)).\label{eqn: lim 2}
 		\end{align}
-		and for \(n\geq 1\), then, 
+		and for \(n\geq 1\),  
 		\begin{align}
-			&\mathbb P(\Ydp(n+1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j) 
+			&\mathbb P(\Ydp(n+1) = (\ell(\ell_0,j),j) 
 					 \mid \Ydp(n) = (\ell_0,i) ) \nonumber
@@
 			\\&\to  
-				\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell_0+1(j\in\calS_+)}, j) \mid \bs X(\tau_{n}) = (y_{\ell_0+1(i\in\calS_-)},i) ).\label{eqn: ddddddd2}
+				\mathbb P(\bs X(\tau_{n+1}^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j) \mid \bs X(\tau_{n}) = (y_{\ell_0+1(i\in\calS_-)},i) ).\label{eqn: ddddddd2}
 		\end{align}
@@
 Corollary~\ref{cor: aln22222212} states that the transition probabilities of the embedded processes converge.
-Thus the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the sequence of distributions of \(\{\Ydp(n)\}\) is tight.
+Thus, the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the sequence of distributions of \(\{\Ydp(n)\}\) is tight.
 Thus, we can establish the convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).
@@
 	\begin{align}
-		&\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell_0-1(j\in\calS_-)+1(j\in\calS_+),j),\tau_1^{(p)}\leq E^\lambda
+		&\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell(\ell_0,j),j),\tau_1^{(p)}\leq E^\lambda
             	 \mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
@@
 		\\&\to  
-			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1(j\in\calS_+)}, j),\tau_1^{(p)}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\nonumber
+			\mathbb P(\bs X(\tau_1^X) = (y_{\ell(\ell_0,j)+1(j\in\calS_-)}, j),\tau_1^{(p)}\leq E^\lambda \mid \bs X(0) = (x_0,i)).\nonumber
 	\end{align}
@@
 		\mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
-		\\&=\sum_{\ell\in\{\ell_0-1,\ell_0+1\}\cap\mathcal K}\sum_{j\in\calS}\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell,j),\tau_1^{(p)}\leq E^\lambda
+		\\&=\sum_{\ell\in\{\ell_0-1,\ell_0,\ell_0+1\}\cap\mathcal K}\sum_{j\in\calS}\mathbb P(\bs Y_{\alpha}^{(p)}(1) = (\ell,j),\tau_1^{(p)}\leq E^\lambda
 		\mid \bs Y^{(p)}(0) = \bs y_0) \nonumber
@@
 \section{Convergence of the QBD-RAP scheme}\label{sec: local to global}
-The rest of this chapter is dedicated to proving global convergence results of the QBD-RAP approximation scheme to the fluid queue.
+The rest of this chapter is dedicated to proving a global convergence results of the QBD-RAP approximation scheme to the fluid queue.
 Ultimately, we prove the weak convergence (in space and time) of the QBD-RAP approximation scheme to the distribution of the fluid queue.
-The structure of the argument is to first show a convergence result of the Laplace transform with respect to time of the distribution of the QBD-RAP and fluid queue at the \(n\)th change of level.
-We then prove a convergence result for the Laplace transforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue between the \(n\)th and \(n+1\)th change of level.
-Summing over the number of changes of level, \(n\), and via the Dominated Convergence Theorem, we claim that the Laplace transforms with respect to time of expectations with respect to the distribution of the QBD-RAP approximation and fluid queue converge.
+The structure of the argument is to first show a convergence result for the Laplace transform with respect to time of the distributions of the QBD-RAP and fluid queue at the \(n\)th orbit restart epoch.
+We then prove a convergence result of the QBD-RAP and fluid queue between the \(n\)th and \(n+1\)th orbit restart epoch.
+Summing over the number of orbit restart epochs, \(n\), and via the Dominated Convergence Theorem, we claim a convergence of the Laplace transforms with respect to time of the QBD-RAP approximation and fluid queue.
 Lastly, we apply the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} to claim a weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
 
-\subsection{At the \(n\)th change of level}\label{sec: nth change}
+\subsection{At the \(n\)th orbit restart epoch}\label{sec: nth change}
 % So far we have shown error bounds for the QBD-RAP approximation on the event that the fluid queue and QBD-RAP remain in the same band/level (%Corollary~\ref{cor: lst diff} and 
@@
 % \end{align}
-which is the Laplace transform of the time until the \(n\)th change of level of the QBD-RAP on the event that the level and phase at the \(n\)th change of level are \(\ell\) and \(j_n\), respectively, given that the initial level and phases are \(\ell_0\) and \(i\), respectively, and the initial orbit is \(\bs a_{\ell_0,i}^{(p)}(x_0)\).
-Partitioning on the time of the first change of level, \(\tau_1\), and the level and phase at this time, then (\ref{eqn: aaaaaaccccc}) is equal to
+which is the Laplace transform of the time until the \(n\)th orbit restart epoch of the QBD-RAP on the event that the level and phase at the \(n\)th orbit restart epoch are \(\ell\) and \(j_n\), respectively, given that the initial level and phases are \(\ell_0\) and \(i\), respectively, and the initial orbit is \(\bs a_{\ell_0,i}^{(p)}(x_0)\).
+Partitioning on the time of the first orbit restart epoch, \(\tau_1\), and the level and phase at this time, then (\ref{eqn: aaaaaaccccc}) is equal to
 \begin{align}
-	&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
 	 \mid \Ydp(1)=(\ell_1, j_1), \tau_1^{(p)}\leq E^\lambda ) \nonumber
@@
 	 %
-	 	\\&\to \begin{cases}
-			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j_1), \tau_{1}^X\leq E^\lambda 
+% 	 	\\&\to \begin{cases}
+% 			\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0+1}, j_1), \tau_{1}^X\leq E^\lambda 
+%             	 \mid \bs X(0) = (x_0, i)) 
+% 	 %
+% 	 &\text{ for }j_1\in\calS_+,\, \ell_1=\ell_0+1\\ 
+% 	%  &\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
+% 	%  \mid \bs Y^{(p)}(0)=(\ell_0, \bs  a_{\ell_0,i}^{(p)}(x_0), i)) \nonumber
+% 	%  &\lim_{p\to\infty}\mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
+% 	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
+% 	%  \\&\qquad{} \phi^{(p)}(0)=i) \nonumber
+% 	 %
+% 	 	\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0}, j_1), \tau_{1}^X\leq E^\lambda 
+%             	 \mid \bs X(0) = (x_0, i)) & \text{for }j_1\in\calS_-,\, \ell_1=\ell_0-1. 
+% \end{cases},\nonumber 
+	\\&\to \mathbb P(\bs X(\tau_1^X) = (y_{\ell+1(j_1\in\calS_-)}, j_1), \tau_{1}^X\leq E^\lambda 
 	\mid \bs X(0) = (x_0, i)) 
-	 %
-	 &\text{ for }j_1\in\calS_+,\, \ell_1=\ell_0+1\\ 
-	%  &\lim_{p\to\infty}\mathbb P(\Ydp(1)=(\ell_1, j_1), \tau_{1}^{(p)}\leq E^\lambda
-	%  \mid \bs Y^{(p)}(0)=(\ell_0, \bs  a_{\ell_0,i}^{(p)}(x_0), i)) \nonumber
-	%  &\lim_{p\to\infty}\mathbb P(L^{(p)}(\tau_1^{(p)})=\ell_1, \phi^{(p)}(\tau_1^{(p)}) = j_1, \tau_{1}^{(p)}\leq E^\lambda
-	%  \mid L^{(p)}(0)=\ell_0, \bs A^{(p)}(0)=\bs  a_{\ell_0,i}^{(p)}(x_0), \nonumber 
-	%  \\&\qquad{} \phi^{(p)}(0)=i) \nonumber
-	 %
-	 	\mathbb P(\bs X(\tau_1^X) = (y_{\ell_0}, j_1), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0, i)) & \text{for }j_1\in\calS_-,\, \ell_1=\ell_0-1. 
-\end{cases},\nonumber 
 \end{align}
@@
 by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution.
-The left-hand side of Equation (\ref{eqn: dskvnaSF}) appears as the first factor in the summands of (\ref{eqn: mnebrb2}).
 
@@
 \begin{align}\label{eqn: paths set1}
-	\mathcal P^n(\ell_0,\ell_n)&=\{(\ell_1,\dots,\ell_{n-1}) \in \mathcal K^{n-1}\mid |\ell_{r-1}-\ell_r|=1 \mbox{ or } \ell_{r-1}=\ell_r=0 \mbox{ or } \ell_{r-1}=\ell_r=K,r = 1,\dots,n\}.
+	\mathcal P^n_{\ell_0,\ell_n}&=\left\{(\ell_1,\dots,\ell_{n-1}) \in \mathcal K^{n-1}\Bigg| \begin{array}{c}  \ell_{r-1}=\ell_r=0 \mbox{ or } \ell_{r-1}=\ell_r=K
+		\\ \mbox{ or }|\ell_{r-1}-\ell_r|=1,r = 1,\dots,n\end{array}\right\}.
 \end{align}
-The set \(\mathcal P^n(\ell_0,\ell)\) contains all of the possible sequences of levels which \(\{L(t)\}\) or \(\{X(t)\}\) may visit on a sample path which starts in level \(\ell_0\), ends in level \(\ell\) and changes level \(n\) times.
-By partitioning on the times \(\tau_m\), \(m=2,\dots,n-1\), and the phases and the levels at these times and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is
+The set \(\mathcal P^n_{\ell_0,\ell}\) contains all the possible values which \(\{L(\tau_m)\}_{m=2}^{n-1}\) may take on a sample path which starts in level \(\ell_0\), ends in level \(\ell\) and has \(n\) orbit restart epochs.
+Similarly, \(\mathcal P^n_{\ell_0,\ell}\) contains all the possible values which \(\{X(\tau_m^X)\}_{m=2}^{n-1}\) may take on a sample path which starts in band \(\ell_0\), ends in band \(\ell\) at time \(\tau_n^X\).
+
+Thus, by partitioning on the times \(\tau_m\), \(m=2,\dots,n-1\), and the phases and the levels at these times and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is  
 	\begin{align}
-	 &\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}(\ell_1,\ell)}}\prod_{m=2}^{n}\mathbb P(\Ydp(1) =(\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
+		\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}&\mathbb P(\Ydp(m) =(\ell_m, j_m), 
+		\tau_{m}^{(p)}\leq E^\lambda 
+		\mid  \nonumber 
+\Ydp (m-1) = (\ell_{m-1}, j_{m-1}), \\&\tau_{m-1}^{(p)}\leq E^\lambda ) \nonumber
+		%
+		\\=\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}&\mathbb P(\Ydp(1) =(\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
             	 \mid  
-	 	\bs Y(0) = (\ell_{m-1},\bs \alpha, j_{m-1})),  \label{eqn: 161222}
+	 	\bs Y^{(p)}(0) = (\ell_{m-1},\bs \alpha, j_{m-1})),  \label{eqn: 161222}
 \end{align}
-where we define \(\ell_n=\ell\). %Each factor in the product is 
+by the time-homogeneity property of the QBD-RAP and the memoryless property of \(E^\lambda\) and where we define \(\ell_n=\ell\) for notational convenience. %Each factor in the product is 
 % \begin{align}
@@
 	& \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
-            	 \mid \bs Y(0) = (\ell_{m-1},\bs \alpha, j_{m-1}))\nonumber
+            	 \mid \bs Y^{(p)}(0) = (\ell_{m-1},\bs \alpha, j_{m-1}))\nonumber
 	%
-	\\&\to\begin{cases}
-	 	\mathbb P(\bs X(\tau_1^X) =( y_{\ell_{m-1}+1},j_m), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0, j_{m-1}))  
-	 &\text{ for }\begin{cases} j_m\in\calS_+,\\ \ell_m=\ell_{m-1}+1,\end{cases}\\
-	 % 
-	 \mathbb P(\bs X(\tau_1^X) =( 0,j_m), \tau_{1}^X\leq E^\lambda 
-            	 \mid \bs X(0) = (x_0, j_{m-1}))  
-	 &\text{ for }\begin{cases} j_m\in\calS_+,\\ \ell_m=\ell_{m-1}=0,\end{cases}\\
-	 %
-	 %
-	%  & \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda 
-    %         	 \mid \bs Y(0) = (\ell_{m-1},\bs \alpha, j_{m-1}))\nonumber
-	% %
-	% \\&\to
-	\mathbb P(\bs X(\tau_1^X) = (y_{\ell_{m-1}}, j_m), \tau_{1}^X\leq E^\lambda 
-		\mid \bs X(0) = (x_0, j_{m-1}))  & \text{ for }\begin{cases}j_m\in\calS_-,&\\ \ell_m=\ell_{m-1}-1.&\end{cases}\\
-	% 
-	\mathbb P(\bs X(\tau_1^X) = (y_{K+1}, j_m), \tau_{1}^X\leq E^\lambda 
-		\mid \bs X(0) = (x_0, j_{m-1}))  & \text{ for }\begin{cases}j_m\in\calS_-,&\\ \ell_m=\ell_{m-1}=K.&\end{cases}
-\end{cases}\label{eqn:kknvf fawoprgj v2}
+	\\&\to \mathbb P(\bs X(\tau_1^X) =( y_{\ell_{m}+1(j_m\in\calS_-)},j_m), \tau_{1}^X\leq E^\lambda 
+	\mid \bs X(0) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1}))  \label{eqn:kknvf fawoprgj v2}
 \end{align}
@@
 \begin{align}
-	\mathbb P(\bs X(\tau_m^X) =( y_{\ell_{m-1}+1},j_m), \tau_{m}^X\leq E^\lambda 
-            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda )  \label{eqn: lim 1}
+	\mathbb P(\bs X(\tau_m^X) =( y_{\ell_{m}+1(j_m\in\calS_-)},j_m), \tau_{m}^X\leq E^\lambda 
+		\mid \bs X(\tau_{m-1}) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1})).\label{eqn: lim 12}
 \end{align}
-for \(j_m\in\calS_+, \ell_m=\ell_{m-1}+1\), 
-\begin{align}			
-	 % 
-	 \mathbb P(\bs X(\tau_m^X) =( 0,j_m), \tau_{m}^X\leq E^\lambda 
-            	 \mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda )
-\end{align}
-for \(j_m\in\calS_+, \ell_m=\ell_{m-1}=0\), 
-\begin{align}
-	%
-	\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}}, j_m), \tau_{m}^X\leq E^\lambda 
-		\mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda ) 
-\end{align}
-for \(j_m\in\calS_-, \ell_m=\ell_{m-1}-1\), and 
-\begin{align}
-	% 
-	\mathbb P(\bs X(\tau_m^X) = (y_{K+1}, j_m), \tau_{m}^X\leq E^\lambda 
-		\mid \bs X(\tau_{m-1}^X) = (x_0, j_{m-1}), \tau_{m-1}^X\leq E^\lambda )   \label{eqn: lim 4}
-\end{align}
-for \(j_m\in\calS_-, \ell_m=\ell_{m-1}=K\).
 % \begin{align}
@@
 
-Taking the limit of (\ref{eqn: 161222}) gives 
+Thus, returning to (\ref{eqn: 161222}) and taking the limit as \(p\to\infty\),  
 \begin{align}
-	&\lim_{p\to\infty}\sum_{j_2,\dots,j_{n-1}\in\mathcal S}\sum_{(\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}(\ell_1,\ell)}\prod_{m=2}^{n}\mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda \nonumber
-            	 \\&\qquad\mid \Ydp(m-1) = (\ell_{m-1}, j_{m-1}), \tau_{m-1}^{(p)}\leq E^\lambda) \nonumber 
+	&\lim_{p\to\infty}\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}\mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda \nonumber
+            	 \mid \bs Y^{(p)}(0) = (\ell_{m-1},\bs\alpha, j_{m-1})) \nonumber 
 	%
-	\\ &=\sum_{j_2,\dots,j_{n-1}\in\mathcal S}\sum_{(\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}(\ell_1,\ell)}\prod_{m=2}^{n}\lim_{p\to\infty} \mathbb P(\Ydp(m) = (\ell_m, j_m), \tau_{m}^{(p)}\leq E^\lambda \nonumber
-            	 \\&\qquad \mid \Ydp(m-1) = (\ell_{m-1}, 
-	 	 j_{m-1}), \tau_{m-1}^{(p)}\leq E^\lambda),
+	\\ &=\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}\lim_{p\to\infty} \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda \nonumber
+            	 \mid \bs Y^{(p)}(0) = (\ell_{m-1}, \bs\alpha,
+	 	 j_{m-1}) ),
 	 \label{eqn: 161222b}
@@
 where we may swap the limit and the sums as they are finite, and we can swap the limit and the product since all the limits exist and the product is finite.
-Substituting the limits (\ref{eqn: lim 1})-(\ref{eqn: lim 4}) into (\ref{eqn: 161222b}) gives
+Substituting the limits (\ref{eqn: lim 12}) into (\ref{eqn: 161222b}) gives
 \begin{align}
-	&\sum_{j_2,\dots,j_{n-1}\in\mathcal S}\sum_{(\ell_2,\dots,\ell_{n-1})  \in\mathcal P^{n-1}(\ell_1,\ell)} \prod_{m=2}^{n}\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m-1}+1(j_{m}\in\mathcal S_+)}, j_m), \tau_{m}^X\leq E^\lambda \nonumber
-            	\\&\qquad \mid \bs X(\tau_{m-1}^X)=(y_{\ell_{m-2}+1(j_{m-1}\in\mathcal S_+)},j_{m-1}),\tau_{m-1}^X\leq E^\lambda),\nonumber
+	&\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}} \prod_{m=2}^{n}\mathbb P(\bs X(\tau_m^X) = (y_{\ell_{m}+1(j_{m}\in\mathcal S_-)}, j_m), \tau_{m}^X\leq E^\lambda\mid \nonumber
+            	\\&\qquad\qquad  \bs X(\tau_{m-1}^X)=(y_{\ell_{m-1}+1(j_{m-1}\in\mathcal S_-)},j_{m-1}),\tau_{m-1}^X\leq E^\lambda),\nonumber
 		%
 		\\&= \mathbb P(\bs X(\tau_n^X) = (y_{\ell+1(j_{n}\in\mathcal S_-)}, 
-		j_n), \tau_{n}^X\leq E^\lambda \mid \bs X(0)=(y_{\ell_{0}+1(j_{1}\in\mathcal S_+)},
+		j_n), \tau_{n}^X\leq E^\lambda \mid \bs X(0)=(y_{\ell_{1}+1(j_{1}\in\mathcal S_-)},
 		j_{1}), 
@@
 \end{align}
-by the Law of total probability.
+by the strong Markov property of the fluid queue and the Law of total probability.
 
-Returning to (\ref{eqn: mnebrb2}) and taking the limit as \(p\to \infty \),  
+Returning now to (\ref{eqn: mnebrb2}) and taking the limit as \(p\to \infty \),  
 \begin{align}
-	&\lim_{p\to\infty}\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	&\lim_{p\to\infty}\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap \mathcal K}\mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
 	 \mid \Ydp(1)=(\ell_1,j_1), \tau_{1}^{(p)}\leq E^{\lambda}) \nonumber 
@@
 		j_n), \tau_{n}^X\leq E^\lambda \mid \nonumber
-		\bs X(\tau_1^X)=(y_{\ell_{0}+1(j_{1}\in\mathcal S_+)},
+		\bs X(\tau_1^X)=(y_{\ell_{1}+1(j_{1}\in\mathcal S_-)},
 		j_{1}),
@@
 	 %
-	 \mathbb P(\bs X(\tau_1^X)=(y_{\ell_{0}+1(j_{1}\in\mathcal S_+)},
+	 \mathbb P(\bs X(\tau_1^X)=(y_{\ell_{1}+1(j_{1}\in\mathcal S_-)},
 		j_{1}),\tau_1^X\leq E^\lambda
@@
 \end{align}
-where the swapping of limits and sums is justified as the sums are finite, and swapping limits and products is justified as the product is finite and all limits exist.
+where the swapping of limits and sums in the first equality is justified as the sums are finite, the swapping limits and products in the first equality is justified as the product is finite and all limits exist, and the last inequality is the Law of total probability.
 
@@
 \begin{proof}
-	The proof follows the same arguments as the proof of Corollary~\ref{cor: short cor} -- in Lemma~\ref{lem: kKKJJJF} sum over \(\ell\) and \(j_n\), then apply the Extended Continuity Theorem \cite[Chapter XIII, Theorem 2a]{feller1957}.
+	The proof follows the same arguments as the proof of Corollary~\ref{cor: short cor}.
+In Lemma~\ref{lem: kKKJJJF} sum over \(\ell\) and \(j_n\), then apply the Extended Continuity Theorem \cite[Chapter XIII, Theorem 2a]{feller1957}.
 \end{proof}
 
-\subsection{Between the \(n\)th and \(n+1\)th change of level}\label{sec: between n and np1}
-Let \(\mathcal T_n^{(p)} = (\tau_{n}^{(p)},\tau_{n+1}^{(p)}]\) and \(\mathcal T_n^X=(\tau_n^X,\tau_{n+1}^X]\) be the interval of time between the \(n\)th and \(n+1\)th change of level of the QBD-RAP and fluid queue, respectively.
-Consider the Laplace transform
+\subsection{Between the \(n\)th and \(n+1\)th orbit restart epochs}\label{sec: between n and np1}
+Let \(\mathcal T_n^{(p)} = (\tau_{n}^{(p)},\tau_{n+1}^{(p)}]\) and \(\mathcal T_n^X=(\tau_n^X,\tau_{n+1}^X]\). Consider the Laplace transform 
 \begin{align}
@@
 
-Partitioning on the time of the \(n\)th change of level and the phase and level at this time gives
+Partitioning on the time of the \(n\)th orbit restart epoch and the phase and level at this time gives
 \begin{align}
@@
 \end{align}
-as that appearing in Lemma \ref{lem: kKKJJJF}. Hence (\ref{eqn: Kkk nj aa w}) converges to 
+as that appearing in Lemma \ref{lem: kKKJJJF}, hence (\ref{eqn: Kkk nj aa w}) converges to 
 \begin{align}
@@
 	t\in\mathcal T_0^{(p)}\mid \Ydp(0) = (\ell,j_n))\wrt t \psi(x) \nonumber 
-	 \\&= \int_{x\in\calD_{\ell,j}} \widehat f^{\ell,(p)} (\lambda)(x,j;x_0,j_n) \psi(x)\wrt x \label{eqn: Jk jJ KK}
+	 \\&= \int_{x\in\calD_{\ell,j}} \widehat f^{\ell,(p)} (\lambda)(x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x)\wrt x \label{eqn: Jk jJ KK}
 \end{align}
@@
 		%
-		\\&\to\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j_n}}\mathbb P(\bs X(t)\in(\wrt x, j_n), t\in\mathcal T_n^{X} 
+		\\&\to\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j_n}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} 
 		\mid \bs X(0)=(x_0,i))\psi(x)\wrt x\wrt t.
@@
 
-\subsection{Another domination condition}
+\subsection{A domination condition}
 Our aim is to prove convergence of 
@@
 Hence, upon taking the limit of (\ref{eqn:GHJKLKJHGHJKL}), if we can justify the swap of the sum and the limit, we will obtain the desired result.
+To this end, we now show a domination condition so tht we may apply the Dominated Convergence Theorem.
 %The main result of this subsection is Lemma~\ref{lem: another bound 2} in which we show a bound which can serve as a dominating function so that we may apply the Dominated Convergence Theorem to justify the aforementioned swap of sum and limit.
@@
 	where 
-	\[b = \min\left\{1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_1|,\cfrac{q}{q+\lambda}\right\}\]
+	\[b = \min\left\{1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,\cfrac{q}{q+\lambda}\right\}\]
 	and  
-	\[|r_1|\leq 2G\cfrac{\var \left(Z\right)}{\varepsilon^2} + 2L\varepsilon.\]
+	\[|r_2|\leq 2G\cfrac{\var \left(Z\right)}{\varepsilon^2} + 2L\varepsilon.\]
 \end{lem}
@@
 \begin{proof}
-	For the QBD-RAP, changes of level can only occur when \(i\in\mathcal S_+\cup\calS_-\).
+	For the QBD-RAP, orbit restart epochs can only occur when \(i\in\mathcal S_+\cup\calS_-\).
 
@@
 The arguments for an initial phase \(i\in\mathcal S_-\) are analogous.
-For the QBD-RAP to change level or hit a boundary one of two things must happen, either;
+For an orbit restart epochs to occur, either;
 	\begin{enumerate}
-		\item the fluid remains in phase \(i\) until there is a change of level or boundary hit, or
-		\item the fluid changes phase before there is a change of level or a boundary is hit. 
+		\item the QBD-RAP remains in phase \(i\) until the orbit restart epoch occurs, or
+		\item the QBD-RAP changes phase before there is an orbit restart epoch, after which the orbit restart epoch occurs eventually. 
 	\end{enumerate}
@@
 	\begin{enumerate}
-		\item the fluid remains in phase \(i\) until there is a change of level or a boundary is hit and \(E^\lambda\) does not occur before the change of level, or, \label{item: 1}
-		\item the fluid changes phase before there is a change of level or a boundary is hit and \(E^\lambda\) does not occur before the change of phase. \label{item: 2}
+		\item the phase remains \(i\) until there is an orbit restart epoch and \(E^\lambda\) does not occur before the orbit restart epoch, or, \label{item: 1}
+		\item the phase changes before there is an orbit restart epoch and \(E^\lambda\) does not occur before the orbit restart epoch. \label{item: 2}
 	\end{enumerate}
@@
 		\int_{x=0}^\infty \bs \alpha e^{\bs{S}x}\bs s e^{(T_{ii}-\lambda)x/|c_i|}\wrt x 
-		&= e^{(T_{ii}-\lambda)\Delta/|c_i|} + r_1,\label{eqn: bnd this 12345}
+		&= e^{(T_{ii}-\lambda)\Delta/|c_i|} + r_2,\label{eqn: bnd this 12345}
 	\end{align}
@@
 	\begin{align}
-		&e^{(T_{ii}-\lambda)\Delta/|c_i|} + |r_1| + 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|} + \cfrac{\var(Z)}{\varepsilon^2}\nonumber
-		\\&= 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|}\left[1-e^{(-T_{ii}\varepsilon-\lambda \Delta)/|c_i|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_1| \nonumber 
-		\\&= 1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_1|,
+		&e^{(T_{ii}-\lambda)\Delta/|c_i|} + |r_2| + 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|} + \cfrac{\var(Z)}{\varepsilon^2}\nonumber
+		\\&= 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|}\left[1-e^{(-T_{ii}\varepsilon-\lambda \Delta)/|c_i|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2| \nonumber 
+		\\&= 1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,
 	\end{align}
@@
 	\end{align}
-	Partitioning on the time of the \(1\)st change of level, \(\tau_1\), and the phase and level at time \(\tau_1\), then (\ref{eqn: alalal}) is equal to 
+	Partitioning on the time of the \(1\)st orbit restart epoch, \(\tau_1\), and the phase and level at time \(\tau_1\), then (\ref{eqn: alalal}) is equal to 
 	\begin{align}
-		&\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \nonumber
+		&\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \nonumber
 		\mid \Yd(1)=(\ell_1,j_1),  \tau_1=u_1)\nonumber
@@
 		%
-		\\&=\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t-u_1\in\mathcal T_{n-1} \nonumber
+		\\&=\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t-u_1\in\mathcal T_{n-1} \nonumber
 		\mid \bs Y(0)=(\ell_1,\bs \alpha,j_1))\nonumber
@@
 	\begin{align}
-		&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
+		&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
 		\mid \bs Y(0)=(\ell_1,\bs \alpha, j_1)) \wrt t 
@@
 		%
-		\\&=\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
+		\\&=\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
 		\mid \bs Y(0)=(\ell_1,\bs\alpha,j_1)) \wrt t 
@@
 
-\subsection{Global convergence}
+\subsection{Global convergence}\label{sec: sub global}
 Finally, we combine the convergence result of Lemma~\ref{lem: LAkAKFnvnb mav h} and the domination condition from Corollary~\ref{vcor: cdks d} via the Dominated Convergence Theorem to claim convergence of the Laplace transform of the QBD-RAP given by
@@
 \end{align}
-Partitioning (\ref{eqn: LLLLaaakkkss}) on the number of level changes by time \(t\) gives
+Partitioning (\ref{eqn: LLLLaaakkkss}) on the number of orbit restart epochs by time \(t\) gives
 \begin{align}
@@
 \end{align}
-where the limit is given by Lemma~\ref{lem: LAkAKFnvnb mav h}. Swapping the sum and integrals and by the law of total probability, then (\ref{eqn:Alakskjs}) is equal to 
+where we have used Lemma~\ref{lem: LAkAKFnvnb mav h}. Swapping the sum and integrals and by the law of total probability, then (\ref{eqn:Alakskjs}) is equal to 
 \begin{align}
@@
 	\end{align}
-	since, to be in level \(\ell\) after starting in level \(\ell_0\), there must be at least \(|\ell_0-\ell|\) changes of level.
+	since, to be in level \(\ell\) after starting in level \(\ell_0\), there must be at least \(|\ell_0-\ell|\) orbit restart epochs.
 By Lemma~\ref{lem: another bound 2} then (\ref{eqn:SKAM.
@@
 Therefore, for all \(p>p_0\), the terms in (\ref{eqnL AKJF cjk ajhJK}) are dominated by \(F\min\{B^{|\ell-\ell_0|-1},1\}\).
-Moreover
+Moreover,
 	\begin{align*}
@@
 	\end{align*}
-	hence the dominating terms are summable. Hence we may apply the Dominated Convergence Theorem to swap the necessary limits and sums. 
+	hence the dominating terms are summable. Hence, we may apply the Dominated Convergence Theorem to swap the necessary limits and sums, from which the result follows.
 \end{proof}
 
-The Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} can now be use to claim that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.
+The Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} can now be used to claim that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.
 \begin{thm}[Extended Continuity Theorem \citep{feller1957}]\label{thm: ext cont thm}
@@
 for almost all \(t\geq 0\).
-At such values of \(t\), since \(\psi\) is an arbitrary bounded, then the Portmanteau Theorem \citep[Theorem 2.1, page 16]{billingsleyconvergence} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
+At such values of \(t\), since \(\psi\) is arbitrary and bounded, then the Portmanteau Theorem \citep[Theorem 2.1, page 16]{billingsleyconvergence} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
 

### 2022-01-21 chapter6/chapter.tex
.

@@
 
-\subsection{Comparison of methods}
+\subsection{Comparison of methods}\label{sec: comp}
 Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.
@@
 
-Here we analyse two simple models which are versions of \cite[Example 2]{bean2009}.
-They are versions of \cite[Example 2]{bean2009} since there no lower boundary behaviour was specified (as it is inconsequential to their analysis), whereas here, we introduce behaviour at the lower boundary.
-The two models differ only by their behaviour at the boundaries.
+Here we analyse two simple models which are based on Example~2 in \citep{bean2009}, except here we introduce add a lower boundary to the model (no lower boundary is specified in Example~2 of \citep{bean2009} as it is inconsequential to their analysis).
 
@@
 		-1.1 & 1.1 & 0 & 0 \\
-		1.1 & -1.1 & 0 & 0 \\ 
+		1 & -1 & 0 & 0 \\ 
 		0.01 & 0 & -0.01 & 0 \\
@@
 \end{model}
+Phases~3 and 4 are included only to capture behaviour at the boundary.
+
 For Model~\ref{model: simple}~a), upon hitting the boundary, the process stays at the boundary for an exponentially distributed amount of time with mean 100.
 For Model~\ref{model: simple}~b), upon hitting the boundary, the process is immediately reflected.
-In Model~\ref{model: simple}~b) phases 3 and 4 are ephemeral.
+In Model~\ref{model: simple}~b) phases 3 and 4 are ephemeral and are included only for notational convenience.
 
-The models were dicretised using the DG, uniformisation and QBD-RAP methods using ten cells of width \(1\).
+The models were discretised using the DG, uniformisation and QBD-RAP methods using ten cells of width \(1\).
 We compute the coefficients for the stationary distribution in the following way.
@@
 \section{Transient distributions}
-Once again we consider Model~\ref{model: simple} and use the same spatial discretisation as described in Section~\ref{sec:stat} (ten cells of width 1).
-We consider two initial conditions, a point mass at 0.0 in phase 1, and the initial distribution with PDF
+Once again we consider Model~\ref{model: simple}~b)\footnote{We do not include the results for Model~\ref{model: simple}~a) as, for the purposes of the comparisons here, they are much the same as those reported for Model~\ref{model: simple}~b)} and use the same spatial discretisation as described in Section~\ref{sec:stat} (ten cells with width \(\Delta=1\)).
+Two initial conditions are considered, a point mass at 0.0 in phase 1, and the initial distribution with PDF
 \begin{align}
@@
 in phases 1 and 2, with no mass at the boundaries.
-We integrate the approximations numerically over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.
+We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.
 We apply the DG method with, and without, the Generalised MUSCL slope limiter.
 
-To obtain a \emph{ground truth} we simulated 5,000,000 realisations of the fluid queue and computed the empirical CDF, and also the mass within each cell and at the boundaries.
+To obtain a \emph{ground truth} 5,000,000 realisations of the fluid queue were simulated until \(t=2\), then the empirical CDF and the mass within each cell and at the boundaries were computed from the simulations.
 We do not attempt to numerically approximate the true PDF via simulation.
@@
 
-To examine possible Monte-Carlo error, we used a bootstrap with 1,000 bootstrap samples.
-That is, we sample, with replacement, 5,000,000 realisations of the fluid queue from the original 5,000,000 samples, then compute error metrics for each of the 1,000 bootstrap samples.
-Via the bootstrap, we report the \(5\)th and \(95\)th percentile of the sampling distribribution of the errors.
+To account for possible Monte-Carlo error, we used a bootstrap with 1,000 bootstrap samples.
+That is, we sample, with replacement, 5,000,000 realisations of the fluid queue from the original 5,000,000 samples, then compute error metrics with the resampled data.
+We resample 1,000 times.
+Via the bootstrap, we report the \(5\)th and \(95\)th percentile of the sampling distribution of the errors.
 
@@
 
-To approximate the point mass inital condition we compute the inital coefficients for each scheme exactly.
+To approximate the point mass initial condition we compute the initial coefficients for each scheme exactly.
 For the exponential initial condition (\ref{eqn: exp init cond}) we compute the initial coefficients via Gauss-Lobatto quadrature for the DG method, by using the mid-point rule for the uniformisation method, and by using a trapezoidal rule with 2,001 points on each cell for the QBD-RAP method.
 
-\paragraph{Model~\ref{model: simple}~b) Reflecting model}
+\paragraph{Model~\ref{model: simple}~b) Reflecting model with exponential initial condition}
+Figure~\ref{fig: reflecting transient exp} shows the error metrics for the four different spatial discretisation schemes (DG, DG with MUSCL limiter, uniformisation and QBD-RAP).
+For both error metrics the DG scheme converges rapidly until computational errors become significant.
+The uniformisation and QBD-RAP schemes converge at a slower rate, with the QBD-RAP scheme converging at a faster rate.
+The DG scheme with MUSCL limiter does not appear to be converging, which suggests there is at least one iteration during the numerical integration over time at which the numerical solution displays oscillations.
+However, when the DG approximations for the distribution at time \(t=2\) are plotted, they do not display oscillations.
+This suggests that the oscillations which might occur with the DG scheme must be transient, and have dissipated by time \(t=2\).%
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple}~b) with the exponential initial condition, and the corresponding approximations obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).}
+	\label{fig: reflecting transient exp} 
+\end{figure}
 
+\paragraph{Model~\ref{model: simple}~b) Reflecting model with a point-mass initial condition}
+Figure~\ref{fig: reflecting transient pm} shows the error metrics for the four different spatial discretisation schemes (DG, DG with MUSCL limiter, uniformisation and QBD-RAP).
+Comparing the error metrics in Figure~\ref{fig: reflecting transient pm} for the point mass initial condition, with the ones in Figure~\ref{fig: reflecting transient exp} for the exponential initial condition, all schemes perform worse for the point mass initial condition.
+Regarding comparative rates of convergence, we come to similar conclusions as we did for the exponential initial condition.
+The DG scheme converges fastest, followed by the QBD-RAP scheme, then the uniformisation scheme, while the DG scheme with MUSCL limiter does not appear to converge due to the limiter reducing the scheme to a linear approximation in the spatial variable.
+%
 \begin{figure}
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple}~b) with the point-mass initial condition, and the corresponding approximations obtained via the DG method (blue solid line with crosses), DG method with Generalised MUSCL limiter (green dashed line with circles), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple}~b) with the point-mass initial condition, and the corresponding approximations obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).}
 	\label{fig: reflecting transient pm} 
 \end{figure}
+\FloatBarrier
 
+\section{Hitting/Exit times}
+\begin{model}\label{model: simple2}
+Consider an unbounded fluid queue with two phases, generator 
+\[\bs T = \left[\begin{array}{cc}
+	-1.1 & 1.1 \\
+	1 & -1
+\end{array}\right],\]
+and associated rates \(c_1=1, c_2=-1\). We approximate the first exit time of the interval \([0,1]\). 
+\end{model}
+Let \(\theta = \{\inf t>0 \mid X(t)=0, \mbox{ or }X(t)=1\},\) then the distribution of the exit time in phase \(i\in\{1,2\}\) is 
+\begin{equation}\label{eqn: hit cdf}\mathbb P(\theta < t, \varphi(t)=i\mid \bs X(0)\sim \mu),\end{equation}
+for some initial distribution \(\mu\). We look at two initial conditions; an exponential with equal mass in each phase, 
+\[\mathbb P(X(0)\in\wrt x,\varphi(0)=i) = exp(-x)/(1-exp(-1))/2,\]
+and a point mass at \(X(0)=0\) in phase \(\varphi(0)=1\).
+
+To do so, we partition \([0,1]\) into three intervals of width \(1/3\), and use the DG, uniformisation and QBD-RAP methods to discretise the fluid queue.
+To capture the mass which has left the interval \([0,1]\), we suppose that when the process hits the boundary it is absorbed forever at the boundary and remains in the phase which it hit the boundary.
+
+We integrate the discretisations until time \(t=10\) using the SSPRK4 method with t-step size 0.005.
+We use the DG scheme with and without a slope limiter during the time-integration.
+At each time-step of the numerical integration, we record the amount of mass at the absorbing boundaries in each phase, which gives us an approximation of the cumulative distribution function of the exit time in each phase up to time \(t=10\).
+
+For comparison, we simulated 5,000,000 realisations and recorded the exit time from the interval \([0,1]\) and the phase at the time of exit.
+We then compute the empirical CDF of the hitting probabilities
+\[\mathbb P(\theta < t, \varphi(t)=i\mid \bs X(0)\sim \mu),\] 
+for \(t=0.005\times k\), \(k=0,...,2000\).
+
+To account for Monte-Carlo errors, we took 1,000 bootstrap resamples of the original 5,000,000 samples and computed the empirical CDF of the hitting probabilities for each bootstrap sample.
+For each bootstrap sample, we resampled 5,000,000 points.
+For each bootstrap sample we compute error metrics between the empirical and approximated CDFs and recorded the 5th and 95th percentile of the distribution of the errors.
+
+\paragraph{Exponential initial condition}
+Figure~\ref{fig: hitting time exp} shows the error metrics recorded for the four different numerical approximation schemes.
+The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP converging at a faster rate.
+For both error metrics, the DG scheme converges rapidly to order 5, after which there is no further improvement.
+The DG scheme with MUSCL limiter does not appear to converge, suggesting oscillations in the DG approximation.
+However, there are no obvious signs of this oscillation in the exit-time CDF.
+For the KS-metric, the QBD-RAP scheme outperforms the DG scheme at orders 15, 17 and 21, which, to me, is unexpected.
+Perhaps the oscillations in the DG scheme are preventing more accurate solutions, or it could be that certain numerical errors in the QBD-RAP scheme have had a favourable cancellation and caused a lower error.
 \begin{figure}
 	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple}~b) with the expponential initial condition, and the corresponding approximations obtained via the DG method (blue solid line with crosses), DG method with Generalised MUSCL limiter (green dashed line with circles), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
-	\label{fig: reflecting transient exp} 
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first exit time CDFs (Equation~\ref{eqn: hit cdf}) for Model~\ref{model: simple2} with the exponential initial condition.
+The approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (orange dashed line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: hitting time exp} 
 \end{figure}
+
+\paragraph{Point mass initial condition}
+Figure~\ref{fig: hitting time pm} shows the error metrics recorded for the four different numerical approximation schemes.
+The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP converging at a faster rate.
+The DG scheme with MUSCL limiter does not appear to converge, suggesting oscillations in the DG approximation.
+The DG scheme appears to converge fastest up to the order-19 scheme, at which point the error increases.
+This is caused by oscillations in the numerical solution.
+In this case we can observe the oscillation in the exit time CDF.
+
+In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the exit time for phase 1, obtained via the DG scheme and QBD-RAP scheme, each using an order-21 basis, and also via simulation.
+Observing Figure~\ref{fig: hitting time oscillation} the DG approximation clearly displays oscillations around \(t=0\) and \(t=1\).
+The largest oscillations are around \(t=0\).
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first exit time CDFs (Equation~\ref{eqn: hit cdf}) for Model~\ref{model: simple2} with the point mass initial condition.
+The approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (orange dashed line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: hitting time pm} 
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{chapter6/figs/hitting_times_model/hitting_times/point_mass/cdf_order21DG_and_sims.pdf}%
+	\caption{Approximations of the CDF of the first exit time in phase 1 for Model~\ref{model: simple2} with the point mass initial condition.
+The blue line was obtained from the order-21 DG scheme, the purple dotted line from the order-21 QBD-RAP scheme, and the gold dashed line is the empirical CDF obtained via simulation.
+The DG scheme displays oscillations.
+}
+	\label{fig: hitting time oscillation} 
+\end{figure}
+
+\section{First-return times of fluid-fluid queues}
+Here we approximate a model considered in \cite{blnos2022} which is a modified version of a model first presented in \cite{lnp13}. 
+\begin{model}\label{model: ffq}
+	Consider a stochastic fluid-fluid queue $\{(X(t),Y(t),\varphi(t))\}_{t\geq0},$ where $\{X(t)\}$ and $\{Y(t)\}$ represent the workloads in Buffers~1 and~2 at time $t \geq 0$, respectively, both driven by the phase $\{\varphi(t)\},$ which is a Markov chain on the state space $\mathcal{S} = \{11,10,01,00\}$.
+Both $\{X(t)\}$ and $\{Y(t)\}$ have a regulated boundary at 0.
+Here, the state $11$ indicates inputs to both buffers being \textsc{on}, the state $00$ indicates both being \textsc{off}, the state $10$ is when only the first input is \textsc{on}, and the state $01$ is when only the second is \textsc{on}.
+The input of Buffer~$k$ is switched from \textsc{on} to \textsc{off} with rate $\gamma_k$, and from \textsc{off} to \textsc{on} with rate $\beta_k$, for $k = 1, 2$.
+Thus, the infinitesimal generator $T$ for $\varphi(t)$ is given by
+	\begin{align*} 
+		T = \left[ \begin{array}{cccc} -(\gamma_1 + \gamma_2) & \gamma_2 & \gamma_1 & 0 \\
+							\beta_2 & -(\gamma_1 + \beta_2) & 0 & \gamma_1 \\
+							\beta_1 & 0 & -(\gamma_2 + \beta_1) & \gamma_2 \\
+							0 & \beta_1 &\beta_2 &-(\beta_1 + \beta_2)
+	\end{array}\right].
+	\end{align*}
+
+	The net rates of change for $X(t)$, denoted $c_i$, are given by 
+	% 
+	\begin{align*} 
+	& (c_{11},c_{10},c_{01},c_{00})   =  \begin{array}{rrrr} 
+	(\lambda_1-\theta_1,  & \lambda_1 -\theta_1, &  -\theta_1, & -\theta_1),
+	\end{array}  
+	\end{align*} 
+	% 
+	and the net rates of change for $Y(t)$, denoted $r_i$, are as follows  
+	%  
+	\begin{align*} 
+	(r_{11},r_{10},r_{01},r_{00})  & = \left\{ \begin{array}{lrrrll}  
+	% (\lambda_2 -\kappa, & \;\;\;\;0, & \lambda_2 - \kappa, &  \;\;\;\;0) & \text{if } X_t = 0,  Y_t = 0,\\
+	%  \vspace*{-0.3cm} \\
+												(\lambda_2 - \kappa, & \;\;\;-\kappa,  & \lambda_2 - \kappa, & -\kappa) &\text{if } X_t = 0, \\
+												\vspace*{-0.3cm} \\
+	%	(\lambda_2 -\theta_2,  &  0,  & \lambda_2 - \theta_2,  & \;\;\;\; 0) & \text{if } X_t \in (0,x^*), Y_t = 0,\\
+	%											  \vspace*{-0.3cm} \\
+		(\lambda_2 - \theta_2, & -\theta_2, & \lambda_2 - \theta_2, & -\theta_2) & \text{if } X_t \in (0,x^*),\\
+												\vspace*{-0.3cm} \\
+		(\;\;\;\;\;\;\; \lambda_2, & \;\;\;\;0, & \lambda_2, &  \;\;\;\;0) & \text{if } X_t \geq x^*. \end{array}
+												\right.
+	\end{align*}
+
+	For our numerical experiments, we use the parameter choices given in~\citep{lnp13}: 
+	% 
+		\begin{align} 
+			\label{eqn:parameters}
+		\gamma_1 & =11, \quad  \beta_1 = 1, \quad \lambda_1 = 12.48, \quad  \theta_1 = 1.6, \quad  \kappa = 2.6, \\
+			\label{eqn:parameters-2}
+		\gamma_2 & = 22, \quad \beta_2  = 1, \quad  \lambda_2 = 16.25, \quad \theta_2 = 1.0, \quad x^* = 1.6.
+		\end{align} 
+\end{model}
+
+While the true problem has an unbounded domain $[0,\infty)$, the approximations require the domain of approximation to be a finite interval.
+Here we choose an upper bound of \(48\) and place a regulated boundary at the upper boundary.
+The effect of this truncation can be partly quantified by evaluating \(\lim\limits_{t\to\infty}\mathbb P\left(X(t) > 48\right)\approx 5.83 \times 10^{-9}\), \(i\in\calS\).
+
+We obtained approximations to the infinitesimal generator of the fluid queue \(\{(X(t),\varphi(t))\}_{t\geq 0}\) via the DG method, QBD-RAP method and uniformisation method, all of which used a cell width of \(\Delta=0.4\).
+The matrix resulting from the approximation is then used to approximate the first-return operator \(\Psi(s)\) as discussed in Section~\ref{SOME INTRO SECTION}.
+Due to the stochastic interpretation of the uniformisation and QBD-RAP schemes, the approximations to \(\Psi(s)\) have a stochastic interpretation as the first-return probabilities of a fluid queue driven by a CTMC and QBD-RAP, respectively.
+For the DG method the approximation of \(\Psi(s)\) is not well-understood.
+However, I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\theta_2)\) on to a set of polynomial basis functions, much like the DG method itself.
+
+Ultimately, we want to approximate the first-return distribution 
+\begin{align}\label{eqn: first return Y 1}
+	\mathbb P(X(\theta_2)\leq x, \varphi(\theta_2)=i\mid \bs X(0)\sim \mu).
+\end{align}
+For Model~\ref{model: ffq}, it is only possible for the process \(Y(t)\) to return to \(0\) at time \(t\) when \((X(t),\varphi(t))\in[0,1.6)\times \{10,00\}\).
+As such, evaluate the approximations over this region only.
+We use a grid of 10,001 points at which to evaluate the approximations of the CDF in each phase.
+We consider first the initial distribution which is a point mass at \(Y(0)=0,\, X(0)=5,\, \varphi(0)=01\).
+
+For comparison, we simulated 5,000,000 realisations of the fluid-fluid queue and recorded the value of \(X\) and \(\varphi\) at the time of first return of the second fluid, \(Y\).
+The empirical approximation of (\ref{eqn: first return Y 1}) was then constructed, and error metrics for the difference between the empirical CDF and the approximations was computed.
+To account for Monte-Carlo errors, we used a bootstrap with 1,000 bootstrap samples to construct 1,000 bootstrap samples of the error estimates and recorded the 5th and 95th percentiles of the error distribution.
+Each of the 1,000 bootstrap samples was constructed by resampling the original 5,000,000 realisations 5,000,000 times, with replacement.
+
+In Figure~\ref{fig: ffq return cts} we plot the error metrics for the approximations to the distribution (\ref{eqn: first return Y 1}).
+The DG method performs best converging rapidly until the error in the approximation scheme is swamped by other numerical errors.
+Here, the first return distribution appears to be smooth, hence we might expect the DG method to perform well.
+Note that there is a significant number other sources of error here; machine precision errors, errors in solving the Ricatti equation to approximate \(\Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), and truncation errors.
+Furthermore, for the QBD-RAP, since the parameters \(\bs \alpha,\,\bs S,\,\bs s\,\) and \(\bs D\) are found numerically, then there is another source of error from this.
+
+Figure~\ref{fig: ffq return cts} also shows that the uniformisation and QBD-RAP methods also appear to be converging, albeit, at a slower rate. 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\theta_2)\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}.
+The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: ffq return cts} 
+\end{figure}
+
+By modifying slightly Model~\ref{model: ffq}, we can construct a first return distribution which is discontinuous. 
+\begin{model}\label{model: ffq2}
+	Consider a fluid-fluid queue which is the same as Model~\ref{model: ffq} except 
+	\begin{align}
+		r_{00}(X(t)) = \begin{cases}
+			-\kappa, & \mbox{ if }X(t)=0,\\
+			-\theta_2, & \mbox{ if }X(t)=\in(0,x^*),\\
+			\theta_2, & \mbox{ if }X(t)\geq x^*.
+		\end{cases}
+	\end{align}
+	Further, consider the initial distribution which is a point-mass at \(Y(0)=0,\, X(0)=2, \varphi(0)=00\).
+\end{model}
+As before, we use the DG, uniformisation and QBD-RAP methods to approximate the model, and compare to simulations.
+
+For Model~\ref{model: ffq2} there is a point mass at \(X(\theta_2)=1.2\) of magnitude \(e^{-(\beta_1+\beta_2)}\times 0.5\), which occurs when the phase remains \(\varphi(t)=00\) until this \(\theta_2\).
+
+Figure~\ref{fig: ffq return discts} plots the error metrics for the first return distribution.
+Observing Figure~\ref{fig: ffq return discts} we see that the DG method performs well, followed by the QBD-RAP method, then the uniformisation method performs worst.
+However, the DG method results in an oscillatory solution with the CDF taking impossible values (decreasing at points) as shown in Figure~\ref{fig: ffq2 oscillation}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\theta_2)\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}.
+The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: ffq return discts} 
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{chapter6/figs/ffq/discts/phase_4_cdf.pdf}%
+	\caption{Approximations of the CDF of the first exit time in phase 1 for Model~\ref{model: simple2} with the point mass initial condition.
+The blue line was obtained from the order-21 DG scheme, the purple dotted line from the order-21 QBD-RAP scheme, and the gold dashed line is the empirical CDF obtained via simulation.
+The DG scheme displays oscillations.
+}
+	\label{fig: ffq2 oscillation} 
+\end{figure}
+
+\FloatBarrier
+\section{Discussion}
+In this chapter we have numerically investigated some properties of the QBD-RAP approximation scheme and compared it to existing methods; the uniformisation scheme of \cite{bo2013} and the discontinuous Galerkin scheme.
+In general, the numerical experiments show that, for problems with discontinuities, the DG approximation can exhibit oscillations and negativity, while the QBD-RAP and uniformisation approximations avoid this and, of the latter two, the QBD-RAP scheme often converges faster.
+To avoid the problems of oscillations and negativity, we can sometimes employ a \emph{slope limiter} with the DG scheme, which effectively reduces the scheme to linear in the regions where oscillations are detected.
+The numerical experiments demonstrate the loss of accuracy in the approximation when a slope limiter is used for a discontinuous problem.
+Moreover, for the application of the methods to fluid-fluid queues, there is no useful way to apply the concept of a slope limiter.
+%When the slope limiter detects oscillations in the approximate solution, it reduces the DG scheme to linear in the region surrounding the oscillations, otherwise, the slope limiter leaves the solution unchanged.
+Thus, the slope limiter permits high-order approximations away from oscillations, while also removing oscillations.
+In general, we observe that the smoother the problem the better the performance of the DG method, and it rapidly surpasses the performance of the other two methods.
+For smooth problems the DG method is almost unbeatable.
+
+As a first step in the numerical experiments, we examined the ability for each method to approximate different initial conditions.
+For the DG method, this is equivalent to a projection of the initial condition on to a set of basis polynomials.
+For the uniformisation method this is equivalent to projecting the initial condition on to a basis of constant functions.
+Section~\ref{sec: comp} demonstrates that the DG (projection) scheme can result in oscillations and negative regions in the approximation when the initial condition is discontinuous.
+The uniformisation and QBD-RAP methods avoid this problem, but appear to have higher errors and the QBD-RAP method appears to have the largest errors.
+For discontinuous initial conditions the rates of convergence are comparable for all three methods.
+However, when the initial condition to be approximated is sufficiently smooth, then the DG approximation is \emph{far} superior.
+
+Next, we instrumented the performance of approximations for a simple travelling wave problem with various initial conditions.
+For this problem the solution is given in terms of the initial condition by \(f(x,t) = f(x-t,0)\).
+Once again, for discontinuous problems, the DG method can display oscillations and negativity, while the other methods (DG with slope limiter, uniformisation and QBD-RAP methods) avoid this.
+Further, for discontinuous problems the rates of convergence of the QBD-RAP and DG methods can be similar.
+For smooth problems the DG method is superior.
+
+We then instrument the performance of the approximations on a simple fluid queue.
+We first look at the stationary distribution, which is known to be smooth.
+Since the problem is smooth, then the DG method is superior as expected.
+Of the uniformisation and QBD-RAP methods, the QBD-RAP method gives more accurate solutions.
+We then turn our attention to approximating transient distributions for the same models and consider two different initial conditions, a point-mass and an exponential initial condition.
+The discontinuous initial condition results in a discontinuous transient distribution.
+As for the exponential initial condition, this example demonstrates that, even if the initial condition appears `nice', it can still result in discontinuous, or non-differetiable solutions.
+The numerical evidence suggests that the DG method can display oscillations, while the other methods do not.
+The DG method with slope limiter detects the oscillations and reduces the method to linear.
+Of the uniformisation and QBD-RAP methods, the latter performs better.
+
+Next we look at exit times for the same fluid queue with two initial conditions, an exponential initial condition and a point-mass.
+We look at the exit time of the fluid level from the interval \((0,1)\).
+For this problem there is never any in-flow of mass at the boundaries of the interval and so, for a solution to be continuous, the initial condition needs to be chosen carefully, otherwise discontinuities in the \textit{transient distribution} may result, as is the case for both initial conditions here.
+The numerical results suggest that, due to the discontinuities in the problems, the DG method can perform not as well as we might expect.
+Since the uniformisation and QBD-RAP methods can handle discontinuities, they perform as expected, with the QBD-RAP method performing better than the uniformisation.
+
+Lastly, we apply the DG, uniformisation and QBD-RAP methods to some simple fluid-fluid queues.
+In the first example, which appears to have a smooth solution, the DG method performs very well.
+Of the other two methods the QBD-RAP method performs better than the uniformisation.
+In the second example, which has a discontinuity, the DG method produces the lowest errors, but exhibits oscillations in the solution.
+The other two methods do not produce oscillations, and of the two the QBD-RAP method performs best.
+
+In conclusion, when the problem is known to be smooth, the DG method is very likely to produce excellent results.
+However, for discontinuous problems, the method can show oscillations and infeasible or negative regions of the solution.
+The slope limiter overcomes this, but reduces the accuracy of the DG method to linear near discontinuities, sometimes severely affecting the quality of the approximation.
+The uniformisation and QBD-RAP methods are alternative approximation schemes which exhibit larger errors, but avoid oscillatory solutions and ensure positivity.
+Of uniformisation and QBD-RAP methods, the latter seems to produce lower errors.

### 2022-01-28 chapter6/figs/cp_figs.sh
ch 6, numerics, first draft done.

@@ -1,0 +1,81 @@
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/fluidfluid_discontinuous/figs/ks_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/ffq/discts
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/fluidfluid_discontinuous/figs/l1_cell_probs_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/ffq/discts
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/fluidfluid_discontinuous/figs/l1_cdf_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/ffq/discts
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/fluidfluid_discontinuous/figs/phase_4_cdf.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/ffq/discts
+
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/fluidfluid_continuous/figs/ks_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/ffq/cts
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/fluidfluid_continuous/figs/l1_cell_probs_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/ffq/cts
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/fluidfluid_continuous/figs/l1_cdf_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/ffq/cts
+
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/hitting_times_model/hitting_times/figs/exp/l1_cdf_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/hitting_times_model/hitting_times/exp
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/hitting_times_model/hitting_times/figs/exp/pdf_order21.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/hitting_times_model/hitting_times/exp
[…]

### 2022-01-28 appendix_dg_toy/appendix.tex
same draft

%!TEX root = ../thesis.tex
\chapter{DG applied to a toy example}\label{appendix:example}
\begin{center}
    \begin{minipage}{0.8\textwidth}
        \textit{This appendix has been taken from Appendix~2 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis. I am a co-author of the paper \cite{blnos2022}. The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara. I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme. I contributed Sections~4.4 and 5.1. I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6. Appendix~A is also my original work. I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
        }
    \end{minipage}
    \end{center}
\renewcommand\thefigure{\arabic{figure}}
Here we include a small toy example to show how we construct a DG approximation and to help clarify the notation.

Consider a process \(\{(\overline X_t,Y_t,\varphi_t)\}_{t\geq 0}\) with two phases, \(\varphi_t\in\calS=\{1,2\}\) and generator matrix \(\bs T\). Let \(\mathcal I = 1.8\), and partition into two intervals \(\calD_1=[0,1]\) and \(\calD_2=[1,1.8]\), hence \(x_1=0,\,x_2=1,\,x_3=1.8\). We choose a basis of Lagrange polynomials of order 1 to define our approximation space. That is, 
\begin{align*}
	\phi_{1}^1(x) &= 1-x,& \phi_2^1(x) &= x, & &x \in\calD_1,\\
	\phi_{1}^2(x) &= \cfrac{1.8-x}{0.8},& \phi_2^2(x) &= \cfrac{x-1}{0.8}, & & x \in\calD_2.
\end{align*}
The mesh and basis functions are shown in Figure~\ref{fig: mesh}. 
\begin{figure}
\centering
\begin{tikzpicture}[scale=4]
    % Draw axes
    \draw [<->,thick] (0,1.1) node (yaxis) [above] {}
        |- (2,0) node (xaxis) [right] {$x$};
    % Draw two intersecting lines
    \draw[red] (0,0) coordinate (a_1) -- (1,1) coordinate (a_2);
    \draw[blue] (0,1) coordinate (b_1) -- (1,0) coordinate (b_2);
    \draw[red] (1,0) coordinate (a_3) -- (1.8,1) coordinate (a_4);
    \draw[blue] (1,1) coordinate (b_3) -- (1.8,0) coordinate (b_4);
    \draw[dashed] (1,0) coordinate (c_1) -- (1,1.1) coordinate (c_2);
    \draw[dashed] (1.8,0) coordinate (d_1) -- (1.8,1.1) coordinate (d_2);
    \filldraw (0.5,0) circle (0pt) node[align=center, below] {\(\calD_1\)};
    \filldraw (1.4,0) circle (0pt) node[align=center, below] {\(\calD_2\)};
    \filldraw (0,0) circle (0pt) node[align=center, below] {\(x_1=0\)};
    \filldraw (1,0) circle (0pt) node[align=center, below] {\(x_2=1\)};
    \filldraw (1.8,0) circle (0pt) node[align=center, below] {\(x_3=1.8\)};  
    \filldraw (0,1) circle (0pt) node[align=center, left] {\(1\)};    
    \filldraw (0,1) circle (0pt) node[align=center, right] {\color{blue}\(\phi_1^1(x)\)};    
    \filldraw (1,1) circle (0pt) node[align=center, left] {\color{red}\(\phi_2^1(x)\)};    
    \filldraw (1,1) circle (0pt) node[align=center, right] {\color{blue}\(\phi_1^2(x)\)};    
    \filldraw (1.8,1) circle (0pt) node[align=center, left] {\color{red}\(\phi_2^2(x)\)};    
\end{tikzpicture}
\caption{A mesh with nodes \(x_1=0\), \(x_2=1\) and \(x_3=1.8\) and cells \(\calD_1=[0,1]\), \(\calD_2=[1,1.8]\). There are two basis functions on each cell. Point masses are located at \(x_1=0\) and \(x_3=1.8\).\label{fig: mesh}}
\end{figure}
We can verify that the matrices \(\bs M\) and \(\bs G\) are given by 
\[\bs M = \left[\begin{array}{cc|cc} 1/3 & 1/6 & 0 & 0 \\ 1/6 & 1/3 & 0 & 0 \\\hline 0 & 0 & 4/15 & 4/30 \\ 0 & 0 & 4/30 & 4/15 \end{array}\right],\, \bs G = \left[\begin{array}{cc|cc} -1/2 & 1/2 & 0 & 0 \\ -1/2 & 1/2 & 0 & 0 \\\hline 0 & 0 & -1/2 & 1/2 \\ 0 & 0 & -1/2 & 1/2 \end{array}\right].\] 
The matrix \(\bs P\) is given by 
\[\bs P = \left[\begin{array}{cc|cc} 1/2 & 0 & 0 & 0 \\ 0 & 1/2 & 0 & 0 \\\hline 0 & 0 & 2/5 & 0 \\ 0 & 0 & 0 & 2/5 \end{array}\right].\]

Let \(c_1=1\) and \(c_2=-2\). Then the flux matrices are given by 
\[\bs F_1 = \left[\begin{array}{cc|cc} 0 & 0 & 0 & 0 \\ 0 & -1 & 1 & 0 \\\hline 0 & 0 & 0 & 0 \\ 0 & 0 & 0 & -1 \end{array}\right],\quad \bs F_2 = \left[\begin{array}{cc|cc} -1 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \\\hline 0 & 1 & -1 & 0 \\ 0 & 0 & 0 & 0 \end{array}\right].\]

Suppose that \(r_1(x)>0\) on \(\calD_1=\mathcal F_1^+\) and \(r_1(x)<0\) on \(\calD_2\cup\{1\}=\mathcal F_1^-\), and further, that \(r_2(x)<0\) on \(\{0\}\cup\calD_1=\mathcal F_2^-\) and \(r_2(x)>0\) on \(\calD_2=\mathcal F_2^+\). Specifically, let 
\[r_1(x) = \begin{cases} 1 & x \in [0,1], \\ -1 & x \in [1,1.8],\end{cases}\quad r_2(x) = \begin{cases} -1 & x = 0, \\ -2 & x \in (0,1], \\ 1 & x \in [1,1.8]. \end{cases}\]

%Therefore, on the interior of the space we have 
%\begin{align*}
%  B &= \left[\begin{array}{cc|cc|cc|cc} 
%T_{11}/3 & T_{11}/6 & 0 & 0 & T_{12}/3 & T_{12}/6 & 0 & 0 \\
%T_{11}/6 & T_{11}/3 & 0 & 0 & T_{12}/6 & T_{12}/3 & 0 & 0 \\\hline 
%0 & 0 & T_{11}/15 & T_{11}/30 & 0 & 0 & T_{12}/15 & T_{12}/30 \\ 
%0 & 0 & T_{11}/30 & T_{11}/15 & 0 & 0 & T_{12}/30 & T_{12}/15 \\\hline
%T_{21}/3 & T_{21}/6 & 0 & 0 & T_{22}/3 & T_{22}/6 & 0 & 0 \\
%T_{21}/6 & T_{21}/3 & 0 & 0 & T_{22}/6 & T_{22}/3 & 0 & 0 \\\hline
%0 & 0 & T_{21}/15 & T_{21}/30 & 0 & 0 & T_{22}/15 & T_{22}/30 \\
%0 & 0 & T_{21}/30 & T_{21}/15 & 0 & 0 & T_{22}/30 & T_{22}/15 
%\end{array}\right] 
%\\&{}+\left[\begin{array}{cc|cc|cc|cc} 
%-1/2 & 1/2 & 0 & 0 & 0 & 0 & 0 & 0 \\
%-1/2 & -1/2 & 1 & 0 & 0 & 0 & 0 & 0 \\\hline 
%0 & 0 & -1/2 & 1/2 & 0 & 0 & 0 & 0 \\ 
%0 & 0 & -1/2 & -1/2 & 0 & 0 & 0 & 0 \\\hline
%0 & 0 & 0 & 0 & -1 & -1 & 0 & 0 \\
%0 & 0 & 0 & 0 & 1 & -1 & 0 & 0 \\\hline
%0 & 0 & 0 & 0 & 0 & 2 & -1 & -1 \\
%0 & 0 & 0 & 0 & 0 & 0 & 1 & -1 
%\end{array}\right].
%\end{align*}
%Reordering and adding the boundary conditions gives 
%\begin{align*}
%\widehat{\mathfrak B} &= \left[\begin{array}{c|cc|cc|cc|cc|c} 
%T_{22} & T_{21} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\\hline 
%0 & T_{11}/3 & T_{11}/6 & T_{12}/3 & T_{12}/6 & 0 & 0 & 0 & 0 & 0 \\
%0 & T_{11}/6 & T_{11}/3 & T_{12}/6 & T_{12}/3 & 0 & 0 & 0 & 0 & 0 \\\hline 
%0 & T_{21}/3 & T_{21}/6 & T_{22}/3 & T_{22}/6 & 0 & 0 & 0 & 0 & 0 \\ 
%0 & T_{21}/6 & T_{21}/3 & T_{22}/6 & T_{22}/3 & 0 & 0 & 0 & 0 & 0\\\hline
%0 & 0 & 0 & 0 & 0 & T_{11}/15 & T_{11}/30 & T_{12}/15 & T_{12}/30 & 0 \\ 
%0 & 0 & 0 & 0 & 0 & T_{11}/30 & T_{11}/15 & T_{12}/30 & T_{12}/15 & 0\\\hline
%0 & 0 & 0 & 0 & 0 & T_{21}/15 & T_{21}/30 & T_{22}/15 & T_{22}/30 & 0 \\
%0 & 0 & 0 & 0 & 0 & T_{21}/30 & T_{21}/15 & T_{22}/30 & T_{22}/15 & 0 \\\hline
%0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & T_{12} & T_{11}
%\end{array}\right] 
%\\&{}+\left[\begin{array}{c|cc|cc|cc|cc|c} 
%0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\\hline 
%0 & -1/2 & 1/2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
%0 & -1/2 & -1/2 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\\hline 
%2 & 0 & 0 & -1 & -1 & 0 & 0 & 0 & 0 & 0 \\ 
%0 & 0 & 0 & 1 & -1 & 0 & 0 & 0 & 0 & 0 \\\hline
%0 & 0 & 0 & 0 & 0 & -1/2 & 1/2 & 0 & 0 & 0 \\
%0 & 0 & 0 & 0 & 0 & -1/2 & -1/2 & 0 & 0 & 1 \\\hline
%0 & 0 & 0 & 0 & 2 & 0 & 0 & -1 & -1 & 0 \\
%0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & -1 & 0 \\
%0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 
%\end{array}\right].
%\end{align*}

Then, constructing \({  \bs B}\) we get 
\[
\begin{blockarray}{c|cc|cc|cc|cc|c}
\nabla & \BAmulticolumn{4}{c|}{\calD_1} & \BAmulticolumn{4}{c|}{\calD_2} & \Delta\\
\calF_2^- & \BAmulticolumn{2}{c|}{\calF_1^+} & \BAmulticolumn{2}{c|}{\calF_2^-} & \BAmulticolumn{2}{c|}{\calF_1^-} & \BAmulticolumn{2}{c|}{\calF_2^+} & \calF_1^-\\
q_{\nabla,1} & a_{1,1}^1 & a_{1,2}^1 & a_{2,1}^1 & a_{2,2}^1 & a_{1,1}^2 & a_{1,2}^2 & a_{2,1}^2 & a_{2,2}^2 & q_{\Delta,1}\\\BAhline
\begin{block}{[c|cc|cc|cc|cc|c]} 
T_{22} & 4T_{21} & -2T_{21} & 0 & 0 & 0 & 0 & 0 & 0 & 0\\\BAhline
0 & T_{11}-3 & 3 & T_{12} & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & -1 & T_{11}-1 & 0 & T_{12} & 5 & -2.5 & 0 & 0 & 0 \\\BAhline 
2 & T_{21} & 0 & T_{22}-2 & -2 & 0 & 0 & 0 & 0 & 0 \\ 
0 & 0 & T_{21} & 6 & T_{22}-6 & 0 & 0 & 0 & 0 & 0\\\BAhline
0 & 0 & 0 & 0 & 0 & T_{11}-\frac{15}{4} & \frac{15}{4} & T_{12} & 0 & 0 \\ 
0 & 0 & 0 & 0 & 0 & -\frac{5}{4} & T_{11}-\frac{5}{4} & 0 & T_{12} &1\\\BAhline
0 & 0 & 0 & -4 & 8 & T_{21} & 0 & T_{22}-\frac{5}{2} & -\frac{5}{2} & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & T_{21} & \frac{15}{2} & T_{22}-\frac{15}{2} & 0 \\\BAhline
0 & 0 & 0 & 0 & 0 & 0 & 0 & -2T_{12} & 4T_{12} & T_{11} \\
\end{block}
\end{blockarray}.
%\\&{}+\left[\begin{array}{c|cc|cc|cc|cc|c} 
%0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\\hline 
%0 & -3 & 3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
%0 & -1 & -1 & 0 & 0 & 4 & -2 & 0 & 0 & 0 \\\hline 
%4 & 0 & 0 & -2 & -2 & 0 & 0 & 0 & 0 & 0 \\ 
%0 & 0 & 0 & 6 & -6 & 0 & 0 & 0 & 0 & 0 \\\hline
%0 & 0 & 0 & 0 & 0 & -3.75 & 3.75 & 0 & 0 & 0 \\
%0 & 0 & 0 & 0 & 0 & -1.25 & -1.25 & 0 & 0 & 2.5 \\\hline
%0 & 0 & 0 & -5 & 10 & 0 & 0 & -2.5 & -2.5 & 0 \\
%0 & 0 & 0 & 0 & 0 & 0 & 0 & 7.5 & -7.5 & 0 \\\hline
%0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 
%\end{array}\right].
\]

We also have sub-matrices
\begin{align*}
	&{  \bs B}_{11}^{++} = \left[\begin{array}{cc}
		T_{11}-3 & 3 \\
		-1 & T_{11}-1
	\end{array}\right],
	\,
	{  \bs B}_{11}^{+-} = \left[\begin{array}{cc|c}
		0 & 0 & 0\\
		5 & -2.5 & 0
	\end{array}\right], 
	%
%	\\&{  B}_{11}^{-+} = \left[\begin{array}{cc}
%		0 & 0 \\
%		0 & 0 \\ \hline
%		0 & 0 
%	\end{array}\right], 
	\,
	{  \bs B}_{11}^{--} = \left[\begin{array}{cc|c}
		T_{11}-\frac{15}{2} & \frac{15}{2} & 0\\
		-\frac{5}{2} & T_{11}-\frac{5}{2} & 1\\\hline
		0 & 0 & T_{11}
	\end{array}\right],
	%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%
%	\\&{  B}_{12}^{++} = \left[\begin{array}{cc}
%		0 & 0 \\
%		0 & 0 
%	\end{array}\right],
%	\,
	\\&{  \bs B}_{12}^{+-} = \left[\begin{array}{c|cc}
		0 & T_{12} & 0 \\
		0 & 0 & T_{12}
	\end{array}\right], \,
	%
	{  \bs B}_{12}^{-+} = \left[\begin{array}{cc}
		T_{12} & 0 \\
		0 & T_{12} \\\hline
		-2T_{12} & 4T_{12}
	\end{array}\right],
	\, 
%	{  B}_{12}^{--} = \left[\begin{array}{c|cc}
%		0 & 0 & 0 \\
%		0 & 0 & 0
%	\end{array}\right],
	%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%
%	\\&{  B}_{21}^{++} = \left[\begin{array}{cc}
%		0 & 0 \\
%		0 & 0
%	\end{array}\right],
%	\,
	\\&{  \bs B}_{21}^{+-} = \left[\begin{array}{cc|c}
		T_{21} & 0 & 0 \\
		0 & T_{21} & 0 
	\end{array}\right], \,
	%
	{  \bs B}_{21}^{-+} = \left[\begin{array}{cc}
		4T_{21} & -2T_{21} \\ \hline
		T_{21} & 0 \\
		0 & T_{21} 
	\end{array}\right],
	\, 
%	{  B}_{21}^{--} = \left[\begin{array}{cc|c}
%		0 & 0 & 0 \\\hline
%		0 & 0 & 0 \\ 
%		0 & 0 & 0 
%	\end{array}\right],
	%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%
	\\&{  \bs B}_{22}^{++} = \left[\begin{array}{cc}
		T_{22}-\frac{5}{2} & -\frac{5}{2} \\
		\frac{15}{2} & T_{22} -\frac{15}{2}
	\end{array}\right],
	\,
	{  \bs B}_{22}^{+-} = \left[\begin{array}{c|cc}
		0 & -4 & 8 \\
		0 & 0 & 0 
	\end{array}\right], \,
	%
%	\\&{  B}_{22}^{-+} = \left[\begin{array}{cc}
%		0 & 0 \\\hline
%		0 & 0 \\
%		0 & 0 
%	\end{array}\right], 
%	\,
	{  \bs B}_{22}^{--} = \left[\begin{array}{c|cc}
		T_{22} & 0 & 0\\\hline
		2 & T_{22}-2 & -2\\
		0 & 6 & T_{22}-6
	\end{array}\right],
	%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%
\end{align*}
and \({  \bs B}_{11}^{-+} = \bs 0_{3\times2},\, {  \bs B}_{12}^{++} = \bs 0_{2\times2},\, {  \bs B}_{12}^{--} = \bs 0_{2\times3},\, {  \bs B}_{21}^{++} = \bs 0_{2\times 2},\, {  \bs B}_{21}^{--} = \bs 0_{3\times3},\, {  \bs B}_{22}^{-+} = \bs 0_{3\times2},\) where \(\bs 0_{n\times m}\) denotes an \(n\times m\) matrix of zeros. Furthermore, 
\begin{align*}
	{  \bs B}^{++} &= \left[\begin{array}{cc|cc}
		T_{11}-3 & 3 & 0 & 0 \\
		-1 & T_{11}-1 & 0 & 0 \\\hline
		0 & 0 & T_{22}-\frac{5}{2} & -\frac{5}{2}\\
		0 & 0 & \frac{15}{2} & -\frac{15}{2}
	\end{array}\right],
	\\
	{  \bs B}^{+-} &= \left[\begin{array}{c|cc|cc|c}
		0 & T_{12} & 0 & 0 & 0 & 0\\
		0 & 0 & T_{12} & 5 & -2.5 & 0\\\hline 
		0 & -4 & 8 & T_{21} & 0 & 0 \\
		0 & 0 & 0 & 0 & T_{21} & 0 
	\end{array}\right], 
	\\
	{  \bs B}^{-+} &= \left[\begin{array}{cc|cc}
		4T_{21} & -2T_{21} & 0 & 0 \\\hline
		T_{21} & 0 & 0 & 0\\
		0 & T_{21} & 0 & 0 \\\hline
		0 & 0 & T_{12} & 0 \\ 
		0 & 0 & 0 & T_{12} \\\hline
		0 & 0 & -2T_{12} & 4T_{12}
	\end{array}\right],
	\intertext{}
	{  \bs B}^{--} &= \left[\begin{array}{c|cc|cc|c}
		T_{22} & 0 & 0 & 0 & 0 & 0 \\\hline 
		2 & T_{22}-2 & -2 & 0 & 0 & 0 \\
		0 & 6 & T_{22}-6 & 0 & 0 & 0 \\\hline 
		0 & 0 & 0 & T_{11}-\frac{15}{4} & -\frac{15}{4} & 0 \\
		0 & 0 & 0 & -\frac{5}{4} & T_{11}-\frac{5}{4} & 1 \\\hline 
		0 & 0 & 0 & 0 & 0 & T_{11}
	\end{array}\right].
\end{align*}

Since \(r_1(x)\) and \(r_2(x)\) are constant on each cell then \(  \bs R^+\) and \(  \bs R^-\) take a particularly simple form. We have 
\[  \bs R^+ = \left[\begin{array}{cc|cc}1&0&0&0\\0&1&0&0\\\hline0&0&1&0\\0&0&0&1\end{array}\right], \quad   \bs R^- = \left[\begin{array}{c|cc|cc|c}1&0&0&0&0&0\\\hline0&1/2&0&0&0&0\\0&0&1/2&0&0&0\\\hline0&0&0&1&0&0\\0&0&0&0&1&0\\\hline0&0&0&0&0&1\end{array}\right].\]

The DG approximations \(  \bs D^{m n}(s),\, m,n\in\{+,-\}\) can now be constructed as 
\[  \bs D^{m n}(s) = \begin{cases}   \bs R^m\left({  \bs B}^{mm} - s \bs I\right) & n=m,
	\\   \bs R^m {  \bs B}^{mn} & n \neq m. \end{cases}\]

For a given value of \(s\), we construct and solve the matrix Riccati equation,
\begin{align*}
	\bs D^{+-}(s)
+ \bs \Psi(s)   \bs D^{-+}(s)\Psi(s)
+   \bs D^{++}(s)\bs \Psi(s)
+ \bs \Psi(s)  \bs D^{--}(s)
= \bs 0,
\end{align*}
for the matrix \(\bs \Psi(s)\) using, for example, Newtons method \citep{bot08}. To obtain the stationary distribution we require \(\bs \Psi(0)\).

Now, to find \( {\bs\xi}\), we solve the linear system in Equations \eqref{eqn:xi1}-\eqref{eqn:xi2}. The result is a vector which we denote, 
\[ {\bs\xi} = \left[\begin{array}{c|cc|cc|c} {\xi}^\nabla_{2} &  {\xi}_{2,1}^1 &  {\xi}_{2,2}^1 &  {\xi}_{1,1}^2 &  {\xi}_{1,2}^2 &  {\xi}_{1}^\Delta \end{array}\right],\]
where \( {\xi}^\nabla_{2}\) is an approximation to \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =0, \varphi_{\theta_n} = 2\right)\) and \( {\xi}^\Delta_{1}\) is an approximation to the artificial point mass \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =1.8, \varphi_{\theta_n} = 1\right)\). For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 2\right)\), is constructed as \( {\xi}_{2,1}^1\phi_1^1(x) +  {\xi}_{2,2}^1\phi_2^1(x)\). For \(x\in\calD_2\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 1\right)\), is constructed as \( {\xi}_{1,1}^2\phi_1^2(x) +  {\xi}_{1,2}^2\phi_2^2(x).\)

Next, given a value of \(y\), we solve the system \eqref{eqn:pisystem1}-\eqref{eqn:pisystem2} to find \( {\bs p}= {\bs p}^-\) and \( {\bs \pi}(y)\).

For the point masses we have 
\[ {\bs p}^- = \left[\begin{array}{c|cc|cc|c} {p}_{2}^\nabla &  {p}_{2,1}^1 &  {p}_{2,2}^1 &  {p}_{1,1}^2 &  {p}_{1,2}^2 &  {p}_{1}^\Delta \end{array}\right],\]
where \( {p}_{2}^\nabla\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_{t} =0, Y_t=0, \varphi_{t} = 2\right)\) and \( {p}_{1}^\Delta\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t=0, \varphi_{t} = 1\right)\). For \(x\in\calD_1\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 2\right)\), is constructed as \( {p}_{2,1}^1\phi_1^1(x) +  {p}_{2,2}^1\phi_2^1(x)\). For \(x\in\calD_2\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 1\right)\), is constructed as \( {p}_{1,1}^2\phi_1^2(x) +  {p}_{1,2}^2\phi_2^2(x).\)

Similarly, for \( {\bs\pi}^-(y)\), we have 
\[ {\bs \pi}^-(y) = \left[\begin{array}{c|cc|cc|c} {\pi}_{2}^\nabla(y) &  {\pi}_{2,1}^1(y) &  {\pi}_{2,2}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{1,2}^2(y) &  {\pi}_{1}^\Delta(y) \end{array}\right],\]
where \( {\pi}_{2}^\nabla(y)\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t\in\wrt y, \varphi_{t} = 2\right)\) and \( {\pi}_{1}^\Delta(y)\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t\in \wrt y, \varphi_{t} = 1\right)\). For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\), is constructed as \( {\pi}_{2,1}^1(y)\phi_1^1(x) +  {\pi}_{2,2}^1(y)\phi_2^1(x)\). For \(x\in\calD_2\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\), is constructed as \( {\pi}_{1,1}^2(y)\phi_1^2(x) +  {\pi}_{1,2}^2(y)\phi_2^2(x).\)

For \( {\bs\pi}^+(y)\), we have 
\[ {\bs \pi}^+(y) = \left[\begin{array}{cc|cc}  {\pi}_{1,1}^1(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{2,1}^2(y) &  {\pi}_{2,2}^2(y) \end{array}\right].\]
For \(x\in\calD_1\) an approximation to the joint density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\) is constructed as \( {\pi}_{1,1}^1(y)\phi_1^1(x) +  {\pi}_{1,2}^1(y)\phi_2^1(x)\). For \(x\in\calD_2\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\) is constructed as \( {\pi}_{2,1}^2(y)\phi_1^2(x) +  {\pi}_{2,2}^2(y)\phi_2^2(x).\)

In summary, for \(i\in\calS\), a global approximation of the joint stationary distribution is
\begin{align*}
	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = i\right) &\approx \displaystyle \sum_{r\in\{1,2\},k\in\{1,2\}}  {\pi}_{i,r}^k(y)\phi_r^k(x)\wrt x \wrt y ,\quad x\in(0,1.8),\, y>0, \\
	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t=0, \varphi_{t} = i\right) & \approx \displaystyle \sum_{r\in\{1,2\},k\in\{1,2\}}  {p}_{i,r}^k\phi_r^k(x)\wrt x ,\quad x\in(0,1.8),\\
	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx{\pi}_{i}^\nabla(y)\wrt y,\quad y>0, \\
	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 0, Y_t =0, \varphi_{t} = i\right) & \approx {p}_{i}^\nabla,\quad \\ 
	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx {\pi}_{i}^\Delta(y)\wrt y,\quad y>0, \\
	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t = 0, \varphi_{t} = i\right) & \approx {p}_{i}^\Delta.\quad
\end{align*}
%\[\left[\begin{array}{c|cc|cc|cc|cc|c} {\pi}_{2,\nabla}(y) &  {\pi}_{1,1}^1(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{2,1}^1(y) &  {\pi}_{2,2}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{1,2}^2(y) &  {\pi}_{2,1}^2(y) &  {\pi}_{2,2}^2(y) &  {\pi}_{1,\Delta}(y) \end{array}\right]\]

### 2022-01-28 appendix_math_background/appendix.tex
same draft

%!TEX root = ../thesis.tex
\chapter{More mathematical background}\label{app: math background}
\section{Time-integration schemes}\label{sec: time integration}
% SSPRK
% CFL

### 2022-01-28 appendix_properties_of_DG_B/appendix.tex
same draft

%!TEX root = ../thesis.tex
\chapter{Properties of DG operator \(  \bs B\)}\label{appendix:properties}
\begin{center}
    \begin{minipage}{0.8\textwidth}
        \textit{This appendix has been taken from Appendix~1 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis. I am a co-author of the paper \cite{blnos2022}. The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara. I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme. I contributed Sections~4.4 and 5.1. I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6. Appendix~A is also my original work. I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
        }
    \end{minipage}
    \end{center}
Recall that the coefficients \({\boldsymbol a}_i^k(t)\) can be used to construct an approximate solution to a differential equation at time \(t\) as \(u_i^k(x,t)={\boldsymbol a}_i^k(t)\boldsymbol \phi^k(x)\tr{}.\) For \(i\in\calS,\,k\in\{\nabla,1,...,K,\Delta\},\,r\in\mathcal N_k\), define \(\alpha_{i,r}^k(t) := a_{i,r}^k(t)\displaystyle\int_{x\in\calD_k}\phi_r^k(x)\wrt x,\) and row-vectors \(\boldsymbol \alpha_i^k(t)=(\alpha_{i,r}^k(t))_{r\in\mathcal N_k}\). Motivated by the fact that we may be interested in approximations of the probabilities \(\mathbb P(X(t)\in\calD_k,\varphi(t)=i)\) rather than the function \(u_i^k\) itself, we can pose the problem equivalently in terms of the integrals 
\[\mathbb P\left(X(t)\in\calD_k,\varphi(t)=i\right)\approx{\boldsymbol a}_i^k(t)\int_{x\in\calD_k}\boldsymbol \phi^k(x)\tr{}\wrt x={\boldsymbol \alpha}_i^k(t)\boldsymbol 1.\] 
Define
\[\boldsymbol \alpha^k(t) = (\boldsymbol \alpha_i^k(t))_{i\in\calS},\mbox{ and } \boldsymbol \alpha(t)=(\boldsymbol \alpha^k(t)))_{k\in\{\nabla,1,...,K,\Delta\}},\]
and matrices 
\begin{align*}
\bs P_k &= \diag\left(\int_{x\in\calD_k}\phi_r^k(x)\wrt x\right)_{ r\in\mathcal N_k},\, k\in\{1,...,K\},
\\\bs P &=\left[\begin{array}{ccc}
		\bs I_{N_{|\calS|}}\otimes \bs P_1 & &\\
		& \ddots &\\
		& & \bs I_{N_{|\calS|}}\otimes \bs P_K
	\end{array}\right].
\end{align*}
By choosing the basis \(\{\phi_{r}^k\}_{r\in\mathcal N_k,k\in\{1,...,K\}}\) such that \(\int_{x\in\calD_k}\phi_r^k(x)\wrt x\neq 0\) for all \(r,k\), then \(\bs P\) is invertible. This is the case for the Lagrange polynomials, but not, for example, for the Legendre polynomials. We can (loosely) interpret the new coefficients \(\alpha_{i,r}^k(t)\) as representing the amount of probability captured by the basis function \(\phi_r^k(x)\) in phase \(i\).

The differential equation (\ref{eqn: DG ODE w BCs}) can be equivalently expressed as 
\(\label{eqn: DG alpha ODE w BCs}
	\cfrac{\wrt}{\wrt t} \boldsymbol \alpha(t)
	% 
	= \boldsymbol \alpha(t)  { \mathfrak{\bs B}},
\)
where 
\[ {\mathfrak{\bs B}} = \left[\begin{array}{ccc}
		\bs I_{|\calS_\nabla|} & & \\
		& \bs P^{-1} &  \\
		& & \bs I_{|\calS_\Delta|}
	\end{array}\right]
	\bs B
	\left[\begin{array}{ccc}
		\bs I_{|\calS_\nabla|} & & \\
		& \bs P&  \\
		& & \bs I_{|\calS_\Delta|}
	\end{array}\right].\] 
Let
\begin{align*}
 {  {\mathfrak{\bs B}}}^{\nabla1} &:= \bs T_{\nabla+}\otimes\left( \boldsymbol\phi^1(0)\bs M_1^{-1}\bs P_1\right),
 %
 \\ {  {\mathfrak{\bs B}}}^{1\nabla} &:=-\diag(c_i\mathbb 1_{(c_i<0)})_{i\in\calS}\otimes \bs P_1^{-1}\boldsymbol \phi^1(0)\tr{},
 %
 \\ {  {\mathfrak{\bs B}}}^{\Delta K} &:= \bs T_{\Delta-}\otimes \left(\boldsymbol\phi^K(\mathcal I)\bs M_K^{-1}\bs P_K\right)
 %
 \\ {  {\mathfrak{\bs B}}}^{K,\Delta} &:= \diag(c_i\mathbb 1_{(c_i>0)})_{i\in\calS} \otimes  \bs P_K^{-1} \boldsymbol\phi^K(\mathcal I)\tr{},
 %
 \intertext{}\mathfrak{\bs B}^{kk}_{ij} &:= \begin{cases}
    	T_{ii}\bs I_{N_k} + c_i\bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k & i = j, \\
	T_{ij}\bs I_{N_k} & i \neq j,
    \end{cases}\quad \mbox{ for \(k=1,\dots K\),}
 %
 \\\mathfrak{\bs B}^{k,k+1}_{ij} &:= \begin{cases}
	c_i\bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k+1}^{-1}\bs P_{k+1} & i = j, \\
	\bs 0_{N_k} & i \neq j,
    \end{cases}\quad \mbox{ for \(k=1,\dots K-1\),}
 %
 \\\mathfrak{\bs B}^{k-1,k}_{ij} &:= \begin{cases}
    	c_i\bs P_k^{-1}\bs F_i^{k,k-1}\bs M_{k-1}^{-1}\bs P_{k-1} & i = j, \\
	T_{ij}\bs I_{N_k} & i \neq j,
    \end{cases},\quad \mbox{ for \(k=2,\dots K\),}
\\
     {  {\mathfrak{\bs B}}}^{kk}%&=\left[\begin{array}{ccc}T_{11}I_{N_k} + c_1P_k^{-1}(F_1^{kk}+G_k)M_k^{-1}P_k & T_{12}I_{N_k} & T_{1N_\calS}I_{N_k}  \\ T_{21}I_{N_k} & & \\ \vdots &\ddots & \vdots \\ & &   T_{N_\calS-1,N_\calS}I_{N_k} \\  T_{N_\calS1}I_{N_k} &  T_{N_\calS,N_\calS-1}I_{N_k} & T_{N_\calS,N_\calS}I_{N_k} +c_{N_\calS}P_k^{-1}(F_{N_\calS}^{kk}+G_k)M_k^{-1}P_k\end{array}\right]
%    %
%    &=:\left[\begin{array}{ccc} {  {\mathfrak B}}_{11}^{kk} &  {  {\mathfrak B}}^{kk}_{12} &  {  {\mathfrak B}}^{kk}_{1N_\calS}  \\  {  {\mathfrak B}}^{kk}_{21} & & \\ \vdots &\ddots & \vdots \\ & &    {  {\mathfrak B}}^{kk}_{N_\calS-1,N_\calS} \\   {  {\mathfrak B}}^{kk}_{N_\calS1} &   {  {\mathfrak B}}^{kk}_{N_\calS,N_\calS-1} &  {  {\mathfrak B}}^{kk}_{N_\calS,N_\calS} \end{array}\right], 
    %
    &=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}_{11}^{kk} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{1N_\calS}  \\ \vdots &\ddots & \vdots \\   {  {\mathfrak{\bs B}}}^{kk}_{N_\calS1} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{N_\calS,N_\calS} \end{array}\right], \mbox{ for \(k=1,\dots K\),}
    %
\\ {  {\mathfrak{\bs B}}}^{k,k+1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k+1}M_{k+1}^{-1}P_{k+1}&  & \\  &\ddots &  \\ & &   \\  &  &c_{N_\calS}P_k^{-1}F_{N_\calS}^{k,k+1}M_{k+1}^{-1}P_{k+1} \end{array}\right] 
%
&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k+1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k+1}_{1,N_\calS}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k+1}_{N_\calS,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k+1}_{N_\calS,N_\calS} \end{array}\right],\mbox{ for \(k=1,\dots K-1\),}
%
\\ {  {\mathfrak{\bs B}}}^{k,k-1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k-1}M_{k-1}^{-1}P_{k-1}&  & \\  &\ddots &  \\  &  &c_{N_\calS}P_k^{-1}F_{N_\calS}^{k,k-1}M_{k-1}^{-1}P_{k-1} \end{array}\right]
%
&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k-1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k-1}_{1,N_\calS}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k-1}_{N_\calS,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k-1}_{N_\calS,N_\calS} \end{array}\right], \mbox{ for \(k=2,\dots K\).}
\end{align*} 
Then
\begin{align*}
{  {\mathfrak{\bs B}}} &= \left[\begin{array}{llllll}
	\bs T_{\nabla\nabla}&  {  {\mathfrak{\bs B}}}^{\nabla1} & & & & \\
	 {  {\mathfrak{\bs B}}}^{1\nabla} &  {  {\mathfrak{\bs B}}}^{11} &  {  {\mathfrak{\bs B}}}^{12} & & & \\
	&  {  {\mathfrak{\bs B}}}^{21} &  {  {\mathfrak{\bs B}}}^{22} &  {  {\mathfrak{\bs B}}}^{23} & & \\
	& & \ddots & \ddots & \ddots & \\
	& &  {   {\mathfrak{\bs B}}}^{K-1,K-2} & {   {\mathfrak{\bs B}}}^{K-1,K-1} &  {  {\mathfrak{\bs B}}}^{K-1,K} & \\
	& & & {   {\mathfrak{\bs B}}}^{K,K-1} &  {  {\mathfrak{\bs B}}}^{K,K} &  {  {\mathfrak{\bs B}}}^{K,\Delta} \\
	& & & &  {  {\mathfrak{\bs B}}}^{\Delta, K} & \bs T_{\Delta\Delta}
\end{array}\right].\end{align*}

	\begin{rem}
	One may recognise the structure of \( {  {\mathfrak{\bs B}}}\) as the structure of a quasi-birth-and-death process (QBD), with levels \(k=1,...,K\). This raises the question of whether \( {  {\mathfrak{\bs B}}}\) is indeed a representation of the generator matrix of a QBD, or QBD-like process. In the case of a constant basis function on each cell, i.e.~\(N_k=1\) and \(\phi_1^k(x)\propto1\), \(k=1,...,K\), then \( {  {\mathfrak{\bs B}}}\) is the generator of a QBD: it has zero row-sums, negative diagonal entires, and non-negative off-diagonal entries, the QBD-phase variable is \(\{\varphi_t\}\) and the level is \(k=\nabla,1,...,K,\Delta\). In fact, if \(h_k\) is the same for every \(k=1,...,K\), then this is the same QBD discretisation of a stochastic fluid process analysed by \cite{bo2013}. However, for higher-degree polynomials \(\mathfrak{\bs B}\) is not necessarily the generator of a QBD process. We conjecture that, using polynomial basis functions, then \(N_k=1\) and \(\phi_1^k(x)\propto1\), \(k=1,...,K\) is the only DG approximation which has an interpretation as a QBD-like process -- not even as a QBD-RAP \citep{bn2010}.
	\end{rem}

	In the following lemma, we use the following properties of the Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature nodes. 
	\paragraph{Property 1} \(\displaystyle\sum_{s=1}^{N_k} \phi_s^k(x) = \begin{cases} 1&x\in\calD_k,\\ 0 & x\notin \calD_k.\end{cases}\)

	For \(k\in\{1,...,K\}\), let \(\boldsymbol e_n^k\) be a row-vector of length \(N_k\) with a 1 in the \(n\)th position and zeros elsewhere.
	\paragraph{Property 2} At the cell edges, \(\bs \phi^k(x_k) = \bs e_1^k\) and \(\bs\phi^k(x_{k+1})=\bs e_{N_k}^k\), \(k=1,...,K\).

\begin{lem}
	If \(\{\phi_r^k(x)\}_{r\in\mathcal N_k}\), are chosen as the Lagrange interpolating polynomials on \(\mathcal D_k\), \(k\in\{1,\dots,K\}\), then the matrix \( {  {\mathfrak{\bs B}}}\) has zero row-sums. 
\end{lem}
\begin{proof}
	Let \(\boldsymbol 1\)\, and \(\boldsymbol 0\) be column vectors of ones and zeros, respectively, with an appropriate length depending on the context. Using Property 1, observe that 
	\begin{align*}
        \bs M_k\boldsymbol 1 &= \left(\sum_{s=1}^{N_k} \int_{x\in\calD_k}\phi_r^k(x)\phi_s^k(x)\wrt x\right)_{r\in\mathcal N_k}\tr{}
	%
        \\&= \left(\int_{x\in\calD_k}\phi_r^k(x)\sum_{s=1}^{N_k} \phi_s^k(x)\wrt x \right)_{r\in\mathcal N_k} \tr{}
        %
        \\&= \left(\int_{x\in\calD_k}\phi_r^k(x)\wrt x\right)_{r\in\mathcal N_k} \tr{}
        %
        \\&= \bs P_k\boldsymbol 1,
	\end{align*}
	hence \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\). Also 
	\begin{align*}
        \bs G_k\boldsymbol 1 &= \left(\sum_{s=1}^{N_k} \int_{x\in\calD_k}\phi_r^k(x)\cfrac{\wrt }{\wrt x}\phi_s^k(x)\wrt x\right)_{r\in\mathcal N_k} \tr{}
	%
        \\&= \left(\int_{x\in\calD_k}\phi_r^k(x)\cfrac{\wrt }{\wrt x}\sum_{s=1}^{N_k} \phi_s^k(x)\wrt x \right)_{r\in\mathcal N_k} \tr{}
	%
        \\&= \left(\int_{x\in\calD_k}\phi_r^k(x)\cfrac{\wrt }{\wrt x}1\wrt x\right)_{r\in\mathcal N_k} \tr{}
	%
        \\&= \boldsymbol 0,
	\end{align*}
	where we have again used Property 1.

	Consider first \(c_i>0\). Let \(\boldsymbol b\) and \(\boldsymbol d\) be arbitrary row-vectors of length \(N_k\) and \(N_{k+1}\), respectively. By Property 2, for \(k=1,...,K-1,\)
	\begin{align*}
        \bs F_i^{kk}\boldsymbol b &= -\boldsymbol \phi^k(x_{k+1})\tr{} \boldsymbol \phi^k(x_{k+1})\boldsymbol b \\&= -(\bs e^k_{N_k})\tr{} \bs e_{N_k}^k\bs b \\&= -b_{N_k}(\boldsymbol e_{N_k}^k)\tr{},
        %
        \\\bs F_i^{k,k+1}\boldsymbol d &= \boldsymbol \phi^{k}(x_{k+1})\tr{} \boldsymbol \phi^{k+1}(x_{k+1})\boldsymbol d \\&= (\bs e^{k}_{N_k})\tr{} \bs e_{1}^{k+1}\bs d \\&= d_{1}(\boldsymbol e_{N_{k}}^{k})\tr{}.
	\end{align*}
	Therefore, for \(c_i>0\), we claim
	\begin{align*}
		&\sum_{j\in\calS}T_{ij}\bs I_{N_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_k^{-1}\bs P_k\boldsymbol 1 = \bs 0.
	\end{align*}
	The first sum is zero since \(\bs T\) is a generator of a continuous-time Markov chain. This leaves the other two terms, which, using our previous observations, we get 
	\begin{align*}
		& c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k+1}^{-1}\bs P_{k+1}\boldsymbol 1
		\\&= c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\boldsymbol 1
		\\&= c_i \bs P_k^{-1}\bs F_i^{kk}\boldsymbol 1 + c_i \bs P_k^{-1}\bs G_k\boldsymbol 1 +  c_i \bs P_k^{-1}F_i^{k,k+1}\boldsymbol 1
		\\&= c_i \bs P_k^{-1}(-\boldsymbol e_{N_k}^k)\tr{} + \boldsymbol 0 + c_i \bs P_k^{-1}(\boldsymbol e_{N_k}^k)\tr{}
		\\&= \boldsymbol 0.
	\end{align*}

	Similarly, for \(c_i<0\), and row-vectors \(\boldsymbol b\) and \(\boldsymbol d\) of length \(N_k\) and \(N_{k-1}\), respectively, 
	\begin{align*}
		\bs F_i^{kk}\boldsymbol b &= \boldsymbol \phi^k(x_{k})\tr{} \boldsymbol \phi^k(x_{k})\boldsymbol b \\&= (\bs e^{k}_{1})\tr{} \bs e_{1}^{k}\bs b  \\&= b_{1}(\boldsymbol e_{1}^k)\tr{} 
		%
		\\\bs F_i^{k,k-1}\boldsymbol d &= -\boldsymbol \phi^{k}(x_{k})\tr{} \boldsymbol \phi^{k-1}(x_{k})\boldsymbol d \\&= -(\bs e^{k}_{1})\tr{} \bs e_{N_{k-1}}^{k-1}\bs d \\&= -d_{N_{k-1}}(\boldsymbol e_{1}^k)\tr{}.
		\end{align*} 
		Therefore, for \(c_i<0\) and \(k=2,...,K\), using the same arguments as before we have
	\begin{align*}
		&\sum_{j\in\calS}T_{ij}\bs I_{N_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\bs M_k^{-1}\bs P_k\boldsymbol 1
		\\&= \boldsymbol 0 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k-1}^{-1}\bs P_{k-1}\boldsymbol 1
		\\&= c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\boldsymbol 1
		\\&= c_i \bs P_k^{-1}\bs F_i^{kk}\boldsymbol 1 + c_i \bs P_k^{-1}\bs G_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\boldsymbol 1
		\\&= c_i \bs P_k^{-1}(\boldsymbol e_{1}^k)\tr{} + \boldsymbol 0 + c_i \bs P_k^{-1}(-\boldsymbol e_{1}^k)\tr{}
		\\&= \boldsymbol 0.
	\end{align*}

	For the lower boundary,
	\begin{align*}
		\bs T_{\nabla\nabla}\boldsymbol 1 +  {  {\mathfrak{\bs B}}}^{\nabla1}\boldsymbol 1 
		&= \bs T_{\nabla\nabla}\boldsymbol 1 + \left[\bs T_{\nabla+}\otimes\left( \boldsymbol\phi^1(0)\bs M_1^{-1}\bs P_1\right)\right]\boldsymbol 1.
		\intertext{Swapping the order of summation and recalling \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\) then this is equal to}
		&\bs T_{\nabla\nabla}\boldsymbol 1 + \left[\bs T_{\nabla+}\otimes\left( \boldsymbol\phi^1(0)\bs M_1^{-1}\bs P_1\right)\boldsymbol 1\right]\boldsymbol 1 
		%
		\\&= \bs T_{\nabla\nabla}\boldsymbol 1 + \left[\bs T_{\nabla+}\otimes \boldsymbol e_1^1\bs 1\right]\bs 1
		\\&= \bs T_{\nabla\nabla}\boldsymbol 1 + \left[\bs T_{\nabla+}\otimes \boldsymbol 1\right]\bs 1
		\\&= \bs T_{\nabla\nabla}\boldsymbol 1 + \bs T_{\nabla+}\bs 1
		%
		\\&= \boldsymbol 0.
	\end{align*}
	Also, for \(c_i<0\), 
	\begin{align*}
		-c_i \bs P_1^{-1}\boldsymbol \phi^1(0)\tr{} + c_i \bs P_1^{-1}(\bs F_i^{1,1}+\bs G_k)\bs M_1^{-1}\bs P_1\boldsymbol 1
		&= -c_i \bs P_1^{-1}\boldsymbol \phi^1(0)\tr{} + c_i \bs P_1^{-1}(\bs F_i^{1,1}+\bs G_k)\boldsymbol 1
		%
		\\&=-c_i \bs P_1^{-1}\boldsymbol \phi^1(0)\tr{} + c_i \bs P_1^{-1}\bs F_i^{1,1}\boldsymbol 1
		%
		\\&=-c_i \bs P_1^{-1}(\boldsymbol e_1^1)\tr{} + c_i \bs P_1^{-1}(\boldsymbol e_1^1)\tr{} 
		\\&= \bs 0.
	\end{align*}

	For the upper boundary,
	\begin{align*}
		\bs T_{\Delta\Delta}\boldsymbol 1 +  {  {\mathfrak{\bs B}}}^{\Delta K}\boldsymbol 1 
		&= \bs T_{\Delta\Delta}\boldsymbol 1 + [\bs T_{\Delta-}\otimes( \boldsymbol\phi^K(\mathcal I)\bs M_K^{-1}\bs P_K)]\boldsymbol 1.
		\intertext{Swapping the order of summation and recalling \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\) then this is equal to}
        \bs T_{\Delta\Delta}\boldsymbol 1 + [\bs T_{\Delta-}\otimes( \boldsymbol\phi^K(\mathcal I)\bs M_K^{-1}\bs P_K)\bs 1]\boldsymbol 1 
		%
		%\\&%= T_{\Delta\Delta}\boldsymbol 1 + [T_{\Delta-}\otimes \boldsymbol\phi^K(\mathcal I)\bs 1]\boldsymbol 1
		&= \bs T_{\Delta\Delta}\boldsymbol 1 + [\bs T_{\Delta-}\otimes\bs e^K_{N_K}\bs 1]\boldsymbol 1
		\\&= \bs T_{\Delta\Delta}\boldsymbol 1 + [\bs T_{\Delta-}\otimes \bs e^K_{N_K}]\bs 1
		\\&= \bs T_{\Delta\Delta}\boldsymbol 1 + \bs T_{\Delta-}\bs 1
		%
		\\&= \boldsymbol 0.
	\end{align*}
	Also, for \(c_i>0\), 
	\begin{align*}
		c_i \bs P_K^{-1}\boldsymbol \phi^K(\mathcal I)\tr{} + c_i \bs P_K^{-1}(\bs F_i^{K,K}+\bs G_K)\bs M_K^{-1}\bs P_K\boldsymbol 1
		&= c_i \bs P_K^{-1}\boldsymbol \phi^K(\mathcal I)\tr{} + c_i \bs P_K^{-1}(\bs F_i^{K,K}+\bs G_K)\boldsymbol 1
		%
		\\&=c_i \bs P_K^{-1}\boldsymbol \phi^K(\mathcal I)\tr{} + c_i \bs P_K^{-1}\bs F_i^{K,K}\boldsymbol 1
		%
		\\&=c_i \bs P_K^{-1}(\boldsymbol e_{N_K}^K)\tr{} + c_i \bs P_K^{-1}(-\boldsymbol e_{N_K}^K)\tr{} 
		\\&= \bs 0.
	\end{align*}

	Combining all of the above we have shown that the row sums of \( {  {\mathfrak B}}\) are zero. 
\end{proof}

\begin{cor}
	The DG approximation to the generator \(  \bs B\) conserves probability. That is, for all \(t\geq 0\), 
	\begin{align*}
	&\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(t)+\sum_{i\in\calS} \int_{x\in[0,\mathcal I]}u_i(x,t)\wrt x 
	%
	\\&= \sum_{i\in\calS_\nabla}q_{\nabla,i}(0)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(0)+\sum_{i\in\calS} \int_{x\in[0,\mathcal I]}u_i(x,0)\wrt x.
	\end{align*}
\end{cor}
\begin{proof}
%For a basis of Lagrange polynomials the DG operator \( {  {\mathfrak B}}\) has zero row-sums, therefore 
%\begin{align*}
%	\boldsymbol \alpha(t)\boldsymbol 1 
%	%
%	=  \boldsymbol \alpha(0)\exp( {  {\mathfrak B}}t)\boldsymbol 1 
%	%
%	= \boldsymbol \alpha(0)\boldsymbol 1. 
%\end{align*}
Let \(\{\psi_r^k(x)\}_{r\in\mathcal N_k,k\in\{1,...,K\}},\) be a basis for \(span(\phi_r^k(x),r\in\mathcal N_k,k\in\{1,...,K\})\), where \(\{\phi_r^k(x)\}_{r\in\mathcal N_k,k\in\{1,...,K\}}\) are the Lagrange polynomials. Also define \(\psi_1^\nabla(x)=\delta(x)\) and \(\psi_1^\Delta(x)=\delta(x-\mathcal I)\) to capture the point masses at the boundaries. Let us use the same vector notation for the basis \(\psi_r^k(x)\) as we do for \(\phi_r^k(x)\). For \(k\in\{1,...,K\}\), since \(\{\psi_r^k(x)\}_{r\in\mathcal N_k}\) and \(\{\phi_r^k(x)\}_{r\in\mathcal N_k}\) have the same span, then there is a matrix \(\bs V^k\) such that  \(\bs \psi^k(x)\tr{} = \bs V^k\bs \phi^k(x)\tr{}\). Trivially, this also holds for \(k=\nabla,\Delta\).

Let 
\[ \bs W = \left[\begin{array}{ccc}
		\bs I_{|\calS_\nabla|} & & \\
		& \bs P &  \\
		& & \bs I_{|\calS_\Delta|}
	\end{array}\right]
\mbox{ and }
  \bs V = \left[\begin{array}{ccccc}
		\bs I_{|\calS_\nabla|} & & & &  \\
		& \bs V^1 & & &  \\
		& & \ddots & & \\
		& & & \bs V^K \\
		& & & & \bs I_{|\calS_\Delta|}
	\end{array}\right]. \]
For a DG approximation, \(\bs B\), constructed from \(\{\psi_r^k\}_{r\in\mathcal N_k,k\in\{\nabla,1,...,K,\Delta\}}\), it can be shown that \(\bs B\) is similar to \( {  {\mathfrak{\bs B}}}\) with similarity matrix, \(\bs V\bs W\), such that \(  \bs B_{ij} =  \bs V\bs W{ \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V^{-1},\,i,j\in\calS\).
Therefore 
\begin{align*} 
	\int_{x\in[0,\mathcal I]}  \bs B_{ij}\bs\psi(x)\tr{} \wrt x &= \bs V\bs W { \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V ^{-1}\int_{x\in[0,\mathcal I]} \bs V\bs \phi(x)\tr{}\wrt x 
	%
	\\&=\bs V\bs W { \mathfrak{\bs B}}_{ij}\bs W^{-1} \bs W\bs 1 \\&= \bs V \bs W { \mathfrak{\bs B}}_{ij}\bs 1,
\end{align*}
since \(\displaystyle\int_{x\in[0,\mathcal I]} \bs \phi(x)\tr{}\wrt x =\bs W\bs 1\).
The row sums of \(\mathfrak{\bs B}\) are 0, hence 
\begin{align}\label{eqn:Bsums0} 
	\int_{x\in[0,\mathcal I]}  \sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\tr{} \wrt x  &=  \bs V\bs W \sum_{j\in\calS}{ \mathfrak{\bs B}}_{ij}\bs 1 \\&=  \bs V\bs W \bs 0 \\&= \bs 0.
\end{align}

Let \( {}^\psi \bs a_i(t)\), \(i\in\calS\) denote the coefficients related to the DG approximation constructed with the basis \(\{\psi_r^k\}_{r\in\mathcal N_k,k\in\{\nabla,1,...,K,\Delta\}}\) (to distinguish them from \(\bs a\) and \(\bs \alpha\) used above). The DE constructed by the DG method is
\begin{align*}
	\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x) = \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x).
\end{align*}
Integrating over \(x\in [0,\mathcal I]\) and summing over \(j\in\calS\) we get
\begin{align*}
	\int_{x\in[0,\mathcal I]}\sum_{j\in\calS}\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x)\wrt x = \int_{x\in[0,\mathcal I]}\sum_{j\in\calS}\sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x).
\end{align*}
Exchanging the order of operations gives 
\begin{align}\label{eqn:totalprobDE}
	\cfrac{\wrt}{\wrt t}\sum_{j\in\calS} \left({}^\psi \boldsymbol a_j (t)\right)\int_{x\in[0,\mathcal I]} \boldsymbol \psi(x)\wrt x =  \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\int_{x\in[0,\mathcal I]}\sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\wrt x = 0,
\end{align}
where the right-hand side is \(0\) due to Equation (\ref{eqn:Bsums0}). This holds for all \(t\geq 0\). The left-hand side of (\ref{eqn:totalprobDE}) is the rate of change (with respect to time) of the total mass of the system. Since this is \( 0\) for all \(t\geq 0\), there is no change in the total mass of the system and thus probability is conserved. 
%\begin{align*}
%	&\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(t)+\sum_{i\in\calS}\int_{x\in[0,\mathcal I]} u_i(x,t)\wrt x 
%	%
%	\\&=\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(t)+\sum_{i\in\calS}\int_{x\in[0,\mathcal I]}\boldsymbol a_i(t) \boldsymbol \psi(x)\wrt x 
%	%
%	\\&=\sum_{j\in\calS}\int_{x\in[0,\mathcal I]}\left[\vligne{\bs q_{\nabla}(0) & \boldsymbol a(0) & \bs q_{\Delta}(0)} \exp(  Bt)\right]_{j}\boldsymbol \psi(x)\wrt x 
%	%
%	\\&=\sum_{j\in\calS}\int_{x\in[0,\mathcal I]}\left[\vligne{\bs q_{\nabla}(0) & \boldsymbol a(0) & \bs q_{\Delta}(0)} \left(I+  Bt + \cfrac{  B^2t^2}{2!}+...\right)\right]_{j}\boldsymbol \psi(x)\wrt x
%	%
%	\\&=\sum_{i\in\calS_\nabla}q_{\nabla,i}(0)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(0)+\sum_{i,j\in\calS}\int_{x\in[0,\mathcal I]}\boldsymbol a_i(0)\boldsymbol \psi(x)\wrt x
%	%
%	\\&=\sum_{i\in\calS_\nabla}q_{\nabla,i}(0)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(0)+\sum_{i\in\calS}\int_{x\in[0,\mathcal I]} u_i(x,0)\wrt x.
%\end{align*}
\end{proof}
%\begin{remark}
%	For any choice of basis \(\{\phi_r^k(x)\}_{r\in\{1,...,N_k\},k\in\{1,...,K\}}\) which spans the set of polynomials of degree \(N_k-1\), the DG approximation constructed with this basis will conserve probability. 
%\end{remark}

### 2022-01-28 chapter1/chapter.tex
same draft

@@
 \chapter{Introduction \& Preliminaries \label{ch:intro}} 
+% intro to intro
+% contextual background
+	% Fluid queues 
+		% definition
+		% applications 
+		% existing results 
+	% Fluid-fluid queues
+		% definition 
+		% basic analysis in terms of generators
+		% the need for approximation and how the old work doesn't fit here
+	% The DG method
+		% pertition into cells
+		% project on to a basis
+		% problems: oscillations
+		% soutions and why they dont fit
+		% observation about constant-basis methods
+		% this thesis asks the question: motivated by the constant basis being a probability model, can we dream up a more accurate probability model to approximate a fluid queue?
+	% On the structure of the constant basis/uniformisation method (QBD)
+	% least variable is PH, so what about MEs?
+% structure of the thesis
+	% rest of chapter 1 gives mathematical preliminaries
+	% chapter 2 explains the DG method, problems/oscillations, slope limiting, and appllication to SFFMs
+	% inspired by the structure of the order-1 scheme, and to solve the negativity problem, chapter 3 introduces a new approximation method; The QBD-RAP.
+In this chapter we derive the intuitively of the model and explain it's dynamics.
+	% we then move to convergence, with chapter 4 proving a convergence of the QBD-RAP up to the first hitting time of the fluid queue on the edges of an interval.
+technical results and extensions are left to the appendix: certain properties of closing vectors, convergence without ephemeral phases, and certain matrix algebraic manipulations.
+	% chapter 5 stitches together the results of chapter 4 to prove a global result about convergence
+	% chapter 6 investigate, numerically, the performance of the DG, DG with limiter, uniformisation and QBD-RAP schemes. 
+	% chapter 7  makes concluding remarks.
+% mathematical preliminaries 
+	% CTMCs
+	% fluid queues
+	% fluid-fluid queues and operator-analytic expressions (all of them)
+		% the equations of bo2014
+		% meets the partition of the DG/QBD-RAP, and reconstructing the Fil sets 
+    % why we cant solve the equations, what we need to approximate to solve it, i.e. B, R, then D, Psi
+	% projections
+	% phase-type distributions 
+		% definition
+		% least variable property
+	% matrix exponentials
+		% definition 
+		% properties 
+		% verification that parameters give an ME
+		% CMEs
+	% QBD-RAPs, then QBDs as a special case
+		% orbit processes, their interpretation for PH and how they differ for MEs
+	% convergence theorems 
+		% DCT 
+		% Continuity of Laplace transforms
+
 % THE ROUGH IDEA & CONTEXT
@@
 Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \citep[Section 6.5]{nodalDGBook} and references therein).
-Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in theses regions.
+Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in these regions.
 Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.
 
-\section{Fluid queues}
+\section{Mathematical Preliminaries}
+\subsection{Fluid queues}
 A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer.
@@
 \begin{equation}
-	-\int_x\int_t \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= \int_{x}\int_t\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x + \int_x\int_t\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
+	-\int_x\int_t \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= \int_{x}\int_t\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x + \int_x\int_t\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs \psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
 \end{equation}
@@
 
-\section{Matrix-exponential distributions}
+\subsection{Fluid-fluid Queues}
+	\label{sec:prelim}
+	\begin{center}
+		\begin{minipage}{0.8\textwidth}
+			\textit{This subsection has been taken from Sections~2 and 3 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis.
+I am a co-author of the paper \cite{blnos2022}.
+The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara.
+I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme.
+I contributed Sections~4.4 and 5.1.
+I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6.
+Appendix~A is also my original work.
+I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
+			}
+		\end{minipage}
+		\end{center}
+
+An unbounded stochastic fluid-fluid model (SFFM) is a Markov process $\{(\widehat{X}_t, \widehat{Y}_t, \varphi(t)),t \geq 0\},$ where the phase $\{\varphi(t)\}$ is a continuous-time Markov chain on a finite state space $\mathcal{S}$;  $\{\widehat{X}_t\}$ is the first fluid, which varies linearly at rate~$c_{\varphi(t)}$
+% 
+\begin{align*} 
+	\widehat X(t) := \widehat X(0) + \int_0^t c_{\varphi(s)} \wrt s;
+\end{align*} 
+% 
+and $\widehat{Y}_t$ is the second fluid, which varies at rate $r_{\varphi(t)}(\widehat{X}_t)$:
+% 
+\begin{align*} 
+	\widehat{Y}_t := \widehat Y(0)) + \int_0^t r_{\varphi(s)}(\widehat{X}_s) \wrt s.
+\end{align*} 
+% 
+Regulated boundaries may also be included for both fluids.
+To distinguish between unbounded and bounded processes, we use the notations \(\widehat X(t)\) and \(\widehat Y(t)\) to denote unbounded processes, and \(X(t)\) and \(Y(t)\) to denote fluid levels with a regulated lower boundary at 0.
+
+As classic fluid processes, $\{(\widehat X(t), \varphi(t)),t \geq 0\}$, or unbounded analogues, are used extensively in many areas, such as insurance and environmental modelling, it is clear that stochastic fluid-fluid models have an even wider range of applicability.
+
+We assume that $X(t), Y(t) \in [0,\infty)$ and that there is a regulated boundary at level $0$ for both buffers: 
+% 
+	\begin{align*} 
+		\frac{\wrt}{\wrt t} X(t) & := \max\{0, c_i\} \quad \mbox{if } X(t) = 0 \mbox{ and } \varphi(t) = i, \\
+          	\frac{\wrt}{\wrt t} Y(t) & := \max\{0, r_i(x)\} \quad \mbox{if } Y(t) = 0,\,X(t) = x \mbox{ and } \varphi(t) = i, 	
+	\end{align*} 
+	% 
+for $i \in \mathcal{S} = \{1,...,N_\calS\}$. Let $T$ be the irreducible generator for the finite-state Markov chain $\{\varphi(t)\}$. 
+%  on $\mathcal{S} = \{1, \ldots, L\}$. 
+We denote by $\bs R(x) := \diag(r_i(x))_{i \in \mathcal{S}}$ the diagonal fluid-rate matrix of functions for $\{Y(t)\}$.
+
+\begin{rem}[Notation]\label{remark: notation 1}For future reference we require some notation regarding the elements of the model introduced above.
+We use the notation \(\bs u = (u_h)_{h\in \mathcal H}\) to denote a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some countable index set.
+Similarly, \(\bs u = (\bs u_h)_{h\in\mathcal H}\), is a row-vector defined by a collection of row-vectors \(\bs u_h\).
+The notation \(\bs u_m=(u_h)_{h\in\mathcal H_m}\) refers to the vector containing the subset of elements corresponding to \(\mathcal H_m\subseteq \mathcal H\).
+When the index set is empty, the resulting vector \(\bs u_m\) is a vector of dimension 0.
+In cases when there are two indices, we order the elements of the vector according to the first index, then the second; i.e.~\(\bs u = (u_{g}^h)_{g\in\mathcal G,h\in\mathcal H} = ((u_g^h)_{g\in\mathcal G})_{h\in\mathcal H}\).
+Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\).
+The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).
+\end{rem}
+
+Let \(\mathcal S_-=\{i\in\calS\mid c_i<0\},\, \mathcal S_+=\{i\in\calS\mid c_i>0\},\, \mathcal S_0=\{i\in\calS\mid c_i=0\},\, \mathcal S_\nabla=\{i\in\calS\mid c_i\leq0\},\, \mathcal S_\Delta=\{i\in\calS\mid c_i\geq0\}\).
+
+For the remainder of this section, we summarise the findings of~\citep{bo2014} on the joint stationary distribution of $\{(X(t), Y(t), \varphi(t)),t \geq 0\}$.
+For each Markovian state $i \in \mathcal{S}$, we partition the state space of \(X(t)\), \([0,\infty)\), according to the rates of change $r_i(\cdot)$ for the second fluid $\{Y(t)\}$: $[0,\infty) := \mathcal{F}^{+}_i \cup \mathcal{F}^{-}_i \cup \mathcal{F}^{0}_i,$
+where 
+% 
+		\begin{align} 
+%				\label{eqn:Fi1}
+			\mathcal{F}^{+}_i & := \{u \in \mathcal{F} : r_i(u) > 0\},  \;
+%				\label{eqn:Fi2}
+			\mathcal{F}^{-}_i :=  \{u \in \mathcal{F}:  r_i(u) < 0\}, \;
+%				\label{eqn:Fi3}
+			\mathcal{F}^{0}_i := \{u \in \mathcal{F}: r_i(u) = 0\}.\label{eqn:fil}
+		\end{align} 
+% 
+For all $i \in \mathcal{S}$, the functions $r_i(\cdot)$ are assumed to be sufficiently well-behaved that $\mathcal{F}^{m}_i$, $m \in \{+, -, 0\}$, is a finite union of intervals and isolated points.
+
+We assume that the process $\{(X(t), Y(t), \varphi(t)),t\geq 0\}$ is positive recurrent, in order to guarantee the existence of the joint stationary density. Define stationary operators 
+\begin{align} 
+		\label{eqn:jointpi} 
+	\overline \bbpi_i(y)(\mathcal{A}) & := \lim_{t \rightarrow \infty} \frac{\partial}{\partial y} \mathbb{P}\left(X(t) \in \mathcal{A}, Y(t) \leq y, \varphi(t) = i\right), \quad y>0,\\
+% 
+		\label{eqn:jointmass}
+		\overline{\mathbb p}_i(\mathcal{A}) & := \lim_{t \rightarrow \infty}  \mathbb{P}(X(t) \in \mathcal{A}, Y(t) = 0, \varphi(t) = i]),
+	\end{align} 
+	where $\mathcal{A} \subset [0,\infty)$.
+Then let $\overline{{\bbpi}}(y) = (\overline \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint stationary density operators and $\overline{\mathbb{p}} = (\overline{\mathbb p}_i)_{i \in \mathcal{S}}$ a vector containing the joint stationary mass operators.
+
+The determination of $\overline{{\bbpi}}(y)$ involves two important matrices of operators, $\overline{\mathbb B}$ and $\overline{\mathbb \Psi}$.
+The operator \(\overline{\mathbb B}\) is the \textit{infinitesimal generator} of the process \(\{(X(t),\varphi(t))\}\).
+%Intuitively, for a set $\mathcal{A} \subset \mathcal{F}$ and a measure vector $\boldsymbol{\mu} = (\mu_i)_{i \in \mathcal{S}}$, $\bs{\mu}V(t)(\mathcal{A})$, where \(V(t)\) is a semigroup with infinitesimal generator \(B\), gives the conditional probability of $X(t) \in \mathcal{A}$,
+The operator \(\overline{\mathbb \Psi}\) is such that $\boldsymbol{\mu} \overline{\mathbb \Psi} (\mathcal{A})$ is the conditional probability of $\{Y(t)\}$ returning to level zero and doing so when $X(t) \in \mathcal{A}$, given that the initial distribution is $\boldsymbol{\mu}$.
+
+\subsubsection{Matrix $\overline{\mathbb B}$ of Operators}
+\label{subsec:B_operators}
+Since \(\{(X(t),\varphi(t)),t\geq 0\}\) is a Markov process, the evolution of probability can be described by a semigroup.
+Let $\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$ be the set of integrable complex-valued Borel measures on the Borel $\sigma$-algebra $\mathcal{B}_{\mathcal{S} \times \mathbb{R}_{+}}$.
+For $\bs{\mu} \in \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$, we can write \(\bs{\mu} = (\mu_i)_{i \in \mathcal{S}}\).
+The measures \(\mu_i(\cdot)\) represent an initial distribution, \(\mu_i(\cdot) = \mathbb P(X(0))\in\cdot, \varphi(0) = i)\).
+Let \(\{\overline{\mathbb V}(t)\}_{t\geq 0},\, \overline{\mathbb V}(t):\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\mapsto \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\) be the semigroup describing the evolution of probability for \(\{(X(t),\varphi(t)),t\geq 0\}\) structured as a matrix of operators, \(\left[ \overline{\mathbb V}(t)\right]_{ij}= \overline{\mathbb V}_{ij}(t)\) where,
+\[\mu_i \overline{ \mathbb V}_{ij}(t)(\mathcal A) = \int_{x\in[0,\infty)} \wrt \mu_i(x)\mathbb P (X(t) \in\mathcal A,\varphi(t) = j \mid X(0)) = x, \varphi(0) = i).\]
+Intuitively, the operator \( \overline{\mathbb V}(t)\) maps an initial measure \(\boldsymbol \mu\) on \((X(0),\varphi(0))\) to the measure \(\mathbb P(X(t)\in \mathcal A, \varphi(t)=j)=:\mu_j(t)(\mathcal A)\).
+The matrix of operators \( \overline{\mathbb B}:=[\overline{\mathbb B}_{ij}]_{i,j\in\mathcal S}\) is the \textit{infinitesimal generator} of the semigroup \(\{ \overline{\mathbb V}(t)\}\) defined by 
+\[ \overline{\mathbb B} =\left.\cfrac{\wrt}{\wrt t} \overline{\mathbb V}(t)\right|_{t=0},\]
+with domain the set of measures for which this limit exists.
+Specifically, the domain of \(\overline{\mathbb B}\) is the set of measures, \(\bs \mu = (\mu_i)_{i\in\calS}\), for which each \(\mu_i\) admits an absolutely continuous density on \((0,\infty)\), and can have a point mass at \(0\) if \(i\in\mathcal S_\nabla\); call this set of measures \(\mathcal M_0\).
+That is, the measures \(\mu_i,\, i\in\calS,\) such that \((\mu_i)_{i\in\calS}\in \mathcal M_0\) are absolutely continuous on \((0,\infty)\) and may have point masses at \(0\) if \(i\in\calS_\nabla\).
+The measure \(\mu_i\) cannot have a point mass at 0 if \(i\notin \mathcal S_\nabla\).
+In the sequel we write \(v_i(x),\, x>0\), as the density of \(\mu_i\) and \(q_i\) as the point mass of \(\mu_i\) at \(x=0\) (if such a point mass exists).
+
+To use the operators \(\{ \overline{\mathbb  V}(t)\}\) and \( \overline{\mathbb  B}\) to analyse the fluid-fluid model, \cite{bo2014} explicitly track when \((X(t),\varphi(t))\in(\mathcal F_i^m,i)\) for \(i\in\mathcal S,\, m \in \{+,-,0\}\) by partitioning the operators \( \overline{\mathbb  V}(t)\) and \( \overline{\mathbb  B}\) into \( \overline{\mathbb  V}_{ij}^{m n}\) and \( \overline{\mathbb  B}_{ij}^{m n}\), for \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\), where
+\[\left.\mu_i\right|_{\calF_i^m}  \overline{\mathbb  V}_{ij}^{m n}(t)(\mathcal A) := \int_{x\in[0,\infty)} \left.
+\wrt \mu_i\right|_{\calF_i^m}(x)\mathbb P (X(t) \in\mathcal A\cap \calF_j^n,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
+and \(\left.\mu_i\right|_{E}\) is the restriction of \(\mu_i\) to \(E\). Similarly, for \( \overline{\mathbb  B}_{ij}^{m n},\) \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\).
+
+We claim that numerical schemes are needed to approximate the analytic operator equations introduced in \cite{bo2014}.
+The DG scheme we choose to use here works by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \(\calD_k=[x_k,x_{k+1}]\) then, on each interval, the operator \(\overline{\mathbb B}\) is projected onto a basis of polynomials.
+So, to help elucidate the connection between the operators \(\{ \overline{\mathbb  V}(t)\}\), \( \overline{\mathbb  B}\) and their DG approximation counterparts, we take a slightly different approach to partitioning these operators than that taken in \cite{bo2014}.
+Rather than partition according to the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\), we use the same partition as that in the construction of the DG scheme.
+By doing so, we can directly correspond elements of the partitioned operators to their approximation counterparts.
+Since the partition used to construct the DG scheme is finer, then we can always reconstruct the partition in terms of the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\).
+
+Let us first partition the space \([0,\infty)\) into \(\calD_\nabla = \{0\}\), and non-trivial intervals \(\calD_k=[x_k,x_{k+1}]\setminus \{0\},\) with \(x_1=0,\, x_k<x_{k+1},\,k=1,2,...\).
+The symbol \(\nabla\) is used to refer to sets and quantities which are relevant to boundary at \(x=0\).
+For $\bs{\mu} \in \mathcal{M}_0(\mathcal{S} \times \mathbb{R}_{+})$ we write \(\bs{\mu} = (\mu^{k}_i)_{i \in \mathcal{S},k\in\{\nabla,1,2,...\}},\) where $\mu^{k}_i(\cdot) = \mu_i(\cdot \cap \calD_k),\,k=\nabla,1,2,\dots.$ We also have densities, \(v_i^k(x),\,x>0\), associated with each measure, \(\mu_i^k\).
+For \( i,j\in\calS,\,k,\ell\in\{\nabla,1,2,\dots\}\) define the operators
+\[\mu_i^k \mathbb V_{ij}^{k \ell}(t)(\mathcal A) := \int_{x\in\calD_k} \wrt \mu_i^k(x)\mathbb P (X(t) \in\mathcal A\cap \calD_\ell,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
+and the matrices of operators \(\mathbb V^{k\ell}(t) := \vligne{\mathbb V_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\{\nabla,1,2,\dots\}\) and write 
+\[\mathbb V(t) = \left[
+	\begin{array}{llll}
+		\mathbb V^{\nabla,\nabla}(t)&\mathbb V^{\nabla,1}(t)& \mathbb V^{\nabla,2}(t) &\hdots \\
+		\mathbb V^{1,\nabla}(t)&\mathbb V^{1,1}(t)&\mathbb V^{1,2}(t)&\hdots\\
+		\mathbb V^{2,\nabla}(t)&\mathbb V^{2,1}(t)&\mathbb V^{2,2}(t)&\hdots\\
+		\vdots & \vdots & \vdots & \ddots 
+%		&\ddots &&&&\reflectbox{\(\ddots\)}\\
+%		&&\mathbb V^{k-1,k-1}(t) & \mathbb V^{k-1,k}(t) & \mathbb V^{k-1,k+1}(t)& \\
+%		&&\mathbb V^{k,k-1}(t) & \mathbb V^{k,k}(t) & \mathbb V^{k,k+1}(t)& \\
+%		&&\mathbb V^{k+1,k-1}(t) & \mathbb V^{k+1,k}(t) & \mathbb V^{k+1,k+1}(t)& \\
+%		&\reflectbox{\(\ddots\)}&&&&\ddots
+	\end{array}\right].\]
+Now define \(\mathbb B=\left.\cfrac{\wrt}{\wrt t}\mathbb V(t)\right|_{t=0}\) as the infinitesimal generator of \(\{\mathbb V(t)\}\), resulting in the tridiagonal matrix of operators 
+\[\mathbb B(t) = \left[
+	\begin{array}{llll}
+		\mathbb B^{\nabla,\nabla}(t)&\mathbb B^{\nabla,1}(t)& & \\
+		\mathbb B^{1,\nabla}(t)&\mathbb B^{1,1}(t)&\mathbb B^{1,2}(t)&\\
+		&\mathbb B^{2,1}(t)&\mathbb B^{2,2}(t)&\ddots\\
+		& & \ddots & \ddots 
+	\end{array}\right],\]
+%\begin{align*}
+%\mathbb B &= \left[\begin{array}{lllll}
+%		\ddots &&&&\\
+%		\mathbb B^{k-2,k-1} &\mathbb B^{k-1,k-1} & \mathbb B^{k-1,k} & & \\
+%		&\mathbb B^{k,k-1} & \mathbb B^{k,k} & \mathbb B^{k,k+1}& \\
+%		& & \mathbb B^{k+1,k} & \mathbb B^{k+1,k+1}&\mathbb B^{k+1,k+2} \\
+%		&&&&\ddots
+%	\end{array}\right],
+%\end{align*}
+where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\{\nabla,1,2,\dots\}\). 
+The tridiagonal structure arises since, for \(|k-\ell|\geq2\) (where we take \(\nabla = 0\) if it appears in the differences) it is impossible for \(\{X(t)\}\) to move from \(\mathcal D_k\) to \(\mathcal D_\ell\) in an infinitesimal amount of time.
+
+\begin{rem}[More notation]\label{remark: notation 2} We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
+The operators denoted with an overline play a minor role in the introductory sections of this paper, but do not appear again.
+We use a blackboard font sans overline (e.g.~\(\mathbb V(t)\) and \(\mathbb B\)) to represent the same operators but which are constructed with the finer partition defined by \(\mathcal D_k,\,k=\nabla,1,2,\dots\).
+We use the letters \(i,j\in\mathcal S\) to represent states of the phase process, letters \(m,n,\in\mathcal \{+,-,0\}\) to refer to the partition in terms of the sets in Equations (\ref{eqn:fil}), and the letters \(k,\ell\in\{\nabla,1,2,...\}\) to refer to the finer partition into sets \(\{\calD_k\}_k\).
+With a slight abuse of notation, whenever we use the dummy variables \(k,\ell\) we imply \(k,\ell\in\{\nabla,1,2,\dots\}\), the dummy variables \(m,n\) imply \(m,n\in\{+,-,0\}\) and the dummy variables \(i,j\) imply \(i,j\in\calS\).
+E.g.~\(\mathbb B_{ij}^{k\ell}\) means \(\mathbb B_{ij}^{k\ell},\, i,j\in\calS, k,\ell\in\{\nabla,1,2,\dots\}\) and \(\mathbb B_{ij}^{mn}\) means \(\mathbb B_{ij}^{mn},\, i,j\in\calS, m,n\in\{+,-,0\}\).
+\end{rem}
+
+By an appropriate choice of the intervals \(\{\mathcal D_k\},\, k\in\{\nabla,1,2,\dots,\}\), the partition used in \citep{bo2014} can be recovered.
+Intuitively, we must ensure that each of the boundaries of \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), align with a boundary of a cell \(\mathcal D_k=[x_k,x_{k+1}]\setminus\{0\}\).
+Then, each set \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), can be written as a union of cells, \(\mathcal D_k,\,k=\nabla,1,2,\dots\), sans a collection of points which have measure zero for all measures in \(\mathcal M_0\), and this collection of points is inconsequential for the purposes of the approximations presented here.
+
+Formally, to recover the partition used in \citep{bo2014} we choose the intervals \(\calD_k\) such that \(   l (\calD_k\cap\calF_i^m) \in \{   l (\calD_k), 0\}\) for all \(i\in\calS,\,m\in\{+,-,0\},\,k\in\{\nabla,1,2,\dots\}\), for all \( l \in\mathcal M_0\).
+That is, we choose \(\calD_k\) such that it is contained (up to sets of measure 0 with respect to measures in \(\mathcal M_0\)) within \(\calF_i^m\) for some \(m\in\{+,-,0\}\) and all \(i\in\calS\).
+We assume such a partition for the rest of the paper.
+For \(i\in\calS,\,m\in\{+,-,0\}\), let \(\mathcal K^m_i = \{k\in\{\nabla,1,2,\dots\}\mid  l (\calD_k\cap\calF_i^m) =   l (\calD_k),\,l\in\mathcal M_0 \}\), so that \(\bigcup\limits_{k\in\mathcal K_i^m} \calD_k\) and \(\mathcal F_i^m\) are equal up to a set of \(l\)-measure 0.
+Define \(\mathcal K^m = \bigcup\limits_{i\in\calS}\mathcal K_i^m\), \(m\in\{+,-,0\}\).
+%Then for \(m\in\{+,-,0\}\), \(\mu_i^m(\cdot) = \sum\limits_{k\in\mathcal K_i^m} \mu_i^k(\cdot)\).
+
+To recover the partition defined by (\ref{eqn:fil}) we bundle together the elements of \(\mathbb V(t)\) which correspond to \(\calF_i^m\) and \(\calF_j^n\).
+That is, for \(m,n\in\{+,-,0\}\), define \(\mathbb V_{ij}^{m n}(t)\) as the matrix of operators
+\[\mathbb V_{ij}^{m n}(t) = \left[\mathbb V_{ij}^{k \ell}(t)\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n}.\] 
+Then, for \(i,j\in\calS,\,m,n\in\{+,-,0\}\), we can write \(\overline{\mathbb  V}_{ij}^{mn}(t) = \boldsymbol 1_{|\mathcal K_i^m|} \mathbb V_{ij}^{m n}(t) \boldsymbol 1_{|\mathcal K_j^n|}\tr{}\) where \(\boldsymbol 1_{|\mathcal K_i^m|}\) and \(\bs 1_{|\mathcal K_j^n|}\) are row-vectors of 1's of length \(|\mathcal K_i^m|\) and \(|\mathcal K_j^n|\), respectively, and \({}\tr{}\) denotes the transpose.
+The same construction can be achieved with \(\mathbb B\).
+
+Let \(\calS_k^+=\{i\in\calS\mid r_i(x)>0,\,\forall x \in\calD_k\}\), \(\calS_k^0=\{i\in\calS\mid r_i(x)=0,\,\forall x \in\calD_k\}\), \(\calS_k^-=\{i\in\calS\mid r_i(x)<0,\,\forall x \in\calD_k\}\) and \(\calS_k^\bullet=\{i\in\calS\mid r_i(x)\neq0,\,\forall x \in\calD_k\}\) for \(k\in\{\nabla,1,\dots,K,\Delta\}\).
+For later reference, we need the following constructions.
+For \(k,\ell\in\{\nabla,1,2,...\}\)
+\begin{align}
+	\mathbb B^{k\ell} & = \left[\mathbb B_{ij}^{k\ell}\right]_{i,j\in\mathcal S},\label{eqn:ref1here}
+\end{align}
+for \(i,j\in\calS\) 
+\begin{align}
+	\mathbb B_{ij} & = \left[\mathbb B_{ij}^{k\ell}\right]_{k,\ell \in \{\nabla,1,2,\dots\}},
+\end{align}
+%for \(i,j\in\calS, \, m,n\in\{+,-,0\}\) 
+%\begin{align}
+%	\mathbb B_{ij}^{m n} &= \left[\mathbb B_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\label{eqn: Bmn etc}
+%	\\ \mathbb B_{ij}^{k n} &= \left[\mathbb B_{ij}^{k \ell}\right]_{\ell\in\mathcal K_j^n} \mbox{ for }k\in\{\nabla,1,2,...\},
+%	\\ \mathbb B_{ij}^{m \ell} &= \left[\mathbb B_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m} \mbox{ for } \ell\in\{\nabla,1,2,...\},
+%	\label{eqn: Bl0 etc}
+%\end{align}
+and for \(m,n\in\{+,-,0\}\) 
+\begin{align}
+	\mathbb B^{m n} &= \left[\left[\mathbb B_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n},\label{eqn: Bmn2 etc}
+	\\ \mathbb B^{k n} &= \left[\left[\mathbb B_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{\ell\in\mathcal K^n} \mbox{ for }k\in\{\nabla,1,2,...\},
+	\\ \mathbb B^{m \ell} &= \left[\left[\mathbb B_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{k\in\mathcal K^m} \mbox{ for } \ell\in\{\nabla,1,2,...\}.
+	\label{eqn:ref2here}
+\end{align}
+We persist with the partition \(\calD_k,\,k\in\nabla,1,2,\dots\) throughout this paper, as this is consistent with the partition used in the DG method, and note that for all the operators defined with this partition, the partitioning used in \citep{bo2014} can always be recovered by the above construction.
+
+We can write \(\mu_i^k\mathbb B_{ij}^{k\ell}(\mathcal A)\) in kernel form as \(\displaystyle\int_{x\in\calD_k,y\in\mathcal A}\wrt \mu_i^k(x)\mathbb B_{ij}^{k\ell}(x,dy)\). It is known that
+\[\mu_i^k\mathbb B_{ij}^{kk}(\wrt y):=\int_{x\in\calD_k}\wrt\mu_i^k(x)\mathbb B_{ij}^{kk}(x,\wrt y)=\begin{cases}
+v_i^k(y)T_{ij}\wrt y, & i\neq j,\\
+v_i^k(y)T_{ii}\wrt y - c_i\cfrac{\wrt}{\wrt y}v_i^k(y)\wrt y, & i=j, 
+\end{cases}\]
+on the interior of \(\calD_k\) \citep{kk1995}.
+Intuitively, \(v_i^k(y)T_{ij}\wrt y\) represents the instantaneous rate of transition from phase \(i\) to \(j\) in the infinitesimal interval \(\wrt y\), \(v_i^k(y)T_{ii}\wrt y\) represents no such transition occurring, and \(- c_i\cfrac{\wrt}{\wrt y}v_i^k(y)\wrt y\) represents the drift across the interval \(\wrt y\) when the phase is \(i\).
+
+Translating the results of \cite{bo2014} to use the partition \(\{\calD_k\}\) we may state that, for all \(i,j\in\calS\), \(k\in\{1,2,\dots\}\),
+\begin{align*}
+\mu_i^{k}\mathbb B_{ij}^{kk}(\calD_k)&=\int_{x\in\calD_k}v_i^k(x)T_{ij}\wrt x - c_iv_i^{k}(x_{k+1})\mathbb 1_{(c_i>0)} + c_iv_i^{k}(x_{k})\mathbb 1_{(c_i<0)},
+\end{align*}
+where \(\mathbb 1_{}\) is the indicator function. 
+Intuitively, the first term represents the instantaneous rate of the stochastic transitions of the phase process \(\{\varphi(t)\}\), the second term represents the flux out of the right-hand edge of \(\calD_k\) which occurs when \(c_i>0\) only, and the last term represents the flux out of the left-hand edge of \(\calD_k\) which occurs when \(c_i<0\) only.
+
+The results of \cite{bo2014} also imply that, 
+\begin{align*}
+\mu_i^{k}\mathbb B_{ii}^{k,k+1}(\calD_{k+1})&= c_iv_i^{k}(x_{k+1})\mathbb 1_{(c_i>0)},\mbox{ for all \(i\in\calS\), \(k\in\{1,2,...\}\),}
+\\\mu_i^{k}\mathbb B_{ii}^{k,k-1}(\calD_{k-1})&= -c_iv_i^{k}(x_{k})\mathbb 1_{(c_i<0)}, \mbox{  for all \(i\in\calS\), \(k\in\{2,3,...\}\)}.
+\end{align*}
+Intuitively, the first equation represents the flux from \(\calD_k\) to \(\calD_{k+1}\) across the shared boundary at \(x_{k+1}\) which occurs when \(c_i>0\) only.
+The second expression represents the flux from \(\calD_k\) to \(\calD_{k-1}\) across the shared boundary at \(x_{k}\) which occurs when \(c_i<0\) only.
+
+\cite{bo2014} also state that, at the boundary \(x=0\), for states \(i\in\calS\) with \(c_i\leq0\) such that a point mass at \(0\) is possible, we have 
+\begin{align*}
+	\mu_i^\nabla\mathbb B_{ii}^{\nabla,\nabla} &= \mu_i^\nabla(\{0\})T_{ii},\\
+	\mu_i^\nabla\mathbb B_{ij}^{\nabla,\nabla} &= \mu_i^\nabla(\{0\})T_{ij},\,j\in\calS,\,c_j\leq 0,\\
+	\mu_i^\nabla\mathbb B_{ij}^{\nabla,1} &= \mu_i^\nabla(\{0\})T_{ij},\,j\in\calS,\,c_j> 0,\\
+	\mu_i^1\mathbb B_{ii}^{1,\nabla} &= -c_iv_i^1(0^+),\,j\in\calS,\,c_j< 0,
+\end{align*}
+where \(0^+\) is the right limit at \(0\).
+Otherwise, \(\mu_i^{k}\mathbb B_{ij}^{k\ell}=0,\,\mbox{ for } |k-\ell|\geq 2,\,i,j\in\mathcal S \mbox{ or }|k-\ell|=1,\,i,j\in\mathcal S,i\neq j\), where we take \(\nabla = 0\) if it appears in the differences, capturing the facts that the process \(\{X(t)\}\) is continuous and that drift across boundaries occurs only when \(\{\varphi(t)\}\) remains in the same phase.
+
+Note that we have not presented \(\mathbb B\) in its full detail here and refer the reader to \citep{bo2014} for the details.
+The main goal here is to show how \(\mathbb B\) is used to construct the stationary distribution of the SFFM and to illustrate the link between the operator \(\mathbb B\) and the DG approximation of the same object.
+As we shall see later, these expressions closely resemble the DG approximations to the same quantities.
+
+\subsubsection{Matrix $\mathbb D(s)$ of Operators}
+Let $b(t) := \int_0^t \left| r_{\varphi(z)}(X(z)) \right|  \wrt z$ be the total unregulated amount of fluid that has flowed into or out of the second buffer $\{Y(t)\}$ during $[0,t]$, and let $\eta(y) := \inf \{t > 0: b(t) = y\}$ be the first time this accumulated in-out amount hits level $y$.
+Note that at the stopping time \(\eta(y)\) it must be that \((X({\eta(y)}),\varphi({\eta(y)}))\in(\calF_i^m,i)\) for some \(i\in\calS\) and \(m\in\{+,-\}\), i.e.~\(m\neq0\).
+We define the operators $\mathbb{U}_{ij}^{k\ell}(y,s): \mathcal{M}_0(\mathcal{D}_{k}\cap\calF_i^m) \mapsto \mathcal{M}_0 (\calD_\ell\cap\mathcal{F}_j^n)$, for $k\in\mathcal K_i^+\cup\mathcal K_i^-$,  $\ell\in\mathcal K_j^+\cup\mathcal K_j^-$, and $i \in \mathcal{S}_k^\bullet,\,j \in \mathcal{S}_k^\bullet$, by
+% 
+	\begin{align*} 
+		&\mu_i^{k}\mathbb{U}_{ij}^{k\ell} (y,s) (\mathcal{A}) 
+		\\&:= \int_{x \in \calD_k} \wrt  \mu_i^{k}(x) \mathbb{E}\left[e^{-s\eta(y)}{1}\left\{\varphi({\eta(y)}) = j, \; X({\eta(y)}) \in \mathcal{A}\cap\calD_\ell\right\} \mid \varphi(0) = i, X(0) = x\right].
+	\end{align*} 
+Then, construct the matrix of operators 
+\begin{align*}
+%\mathbb{U}^{k \ell}(y,s)&:=[\mathbb U_{ij}^{k\ell}(y,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}, \, k,\ell\in \mathcal K^+\cup\mathcal K^-,
+\mathbb U(y,s) &:= \left[[\mathbb U_{ij}^{k\ell}(y,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}\right]_{k,\ell\in\mathcal K^+\cup\mathcal K^-}.\end{align*}
+%\[\mathbb U(t) = \left[
+%	\begin{array}{llll}
+%		\mathbb U^{\nabla,\nabla}(t)&\mathbb U^{\nabla,1}(t)& \mathbb U^{\nabla,2}(t) &\hdots \\
+%		\mathbb U^{1,\nabla}(t)&\mathbb U^{1,1}(t)&\mathbb U^{1,2}(t)&\hdots\\
+%		\mathbb U^{2,\nabla}(t)&\mathbb U^{2,1}(t)&\mathbb U^{2,2}(t)&\hdots\\
+%		\vdots & \vdots & \vdots & \ddots 
+%	\end{array}\right].\]
+%	\begin{align*} 
+%		\mathbb{U}(y,s) = \left[\begin{array}{llllll} 
+%		\ddots & & & &~\reflectbox{$\ddots$}\\
+%            		& \mathbb{U}^{k-1,k-1}(y,s) & \mathbb{U}^{k-1,k}(y,s) & \mathbb{U}^{k-1,k+1}(y,s) \\
+%			& \mathbb{U}^{k,k-1}(y,s) & \mathbb{U}^{k,k}(y,s) & \mathbb{U}^{k,k+1}(y,s) \\
+%			& \mathbb{U}^{k+1,k-1}(y,s) & \mathbb{U}^{k+1,k}(y,s) & \mathbb{U}^{k+1,k+1}(y,s) \\	
+%			\reflectbox{$\ddots$}& & & & \ddots		
+%		\end{array}\right], 
+%	\end{align*} 
+	% 
+%	where $\mathbb{U}^{k \ell}:=\mathbb U_{ij}^{k\ell}$ is an $|\mathcal{S}| \times |\mathcal{S}|$ matrix of operators for $y > 0$, $s \in \mathbb{C}$, and Re$(s) \geq 0$. 
+The matrix of operators \(\mathbb D(s)\) is the infinitesimal generator of the semigroup \(\{\mathbb U(y,s)\}_{y\geq 0}\) defined by 
+\[\mathbb D(s) = \cfrac{\wrt}{\wrt y}\mathbb U(y,s)|_{y=0},\]
+whenever this limit exists.
+
+	Recalling the constructions in Equations (\ref{eqn:ref1here})-(\ref{eqn:ref2here}) and using Lemma~$4$ of \citep{bo2014} gives the following expression for $\mathbb{D}(s)$. 
+\begin{lem}\label{lemma: D(s)}
+	For $y \geq 0$, $s \in \mathbb{C}$ with \textit{Re}$(s) \geq 0$, $i,j \in \mathcal{S}$, $k\in \mathcal K_i^+\cup\mathcal K_i^-$, $\ell\in \mathcal K_j^+\cup\mathcal K_j^-$,
+	% 
+	\begin{align*}
+		\mathbb{D}_{ij}^{k\ell}(s) = [\mathbb{R}^{k}(
+		\mathbb{B}^{k\ell } - s\mathbb{I} + \mathbb{B}^{k0 }(s \mathbb{I} - \mathbb{B}^{00})^{-1}\mathbb{B}^{0\ell})]_{ij}, 
+	\end{align*} 
+	% 
+	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} := \diag(\mathbb{R}_i^{k})_{i \in \mathcal{S}}$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
+	\begin{align*} 
+		{\mu}_i^{k}\mathbb{R}_i^{k}(\mathcal{A}) := \int_{x \in \mathcal{A} \cap \mathcal{D}_k} \frac{1}{r_i(x)}\wrt  \mu_i^{k}(x),\quad k\in \mathcal K_i^+\cup\mathcal K_i^-.
+	\end{align*} 
+\end{lem}
+
+	Also, construct the matrices of operators 
+	\begin{align*}
+            	%\mathbb D^{k \ell} &:= \left[\mathbb D_{ij}^{k \ell}\right]_{i,j\in\mathcal S},\quad k,\ell\in\{\nabla,1,2,...\},
+		\mathbb D^{m n} &:= \left[\left[\mathbb D_{ij}^{k \ell}\right]_{i\in\calS_k^m,j\in\calS_k^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n}.
+		%\\ \mathbb D_{ij}^{m n} &:= \left[\mathbb D_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\quad i,j\in\calS,\,m,n\in\{+,-,0\}.
+        \end{align*}
+
+\subsubsection{Matrix $\mathbb\Psi(s)$ of Operators}\label{sec: intro Psi}
+We denote by $\mathbb \Psi(s)$ the matrix of operators with the same dimensions as \(\mathbb D^{+-}\), recording the Laplace-Stieltjes transforms of the time for $\{Y(t)\}$ to return, for the first time, to the initial level of zero as introduced in \citep{bo2014} but constructed with respect to the finer partition \(\{\calD_k\}\).
+Define the stopping time $\zeta_Y(\{E\}):= \inf \{t > 0: Y(t) \in E\}$ to be the first time $\{Y(t)\}$ hits the set $E$, then each component $\mathbb \Psi_{ij}^{k\ell}(s): \mathcal{M}_0(\mathcal D_k) \mapsto \mathcal{M}_0(\mathcal D_\ell), i,j \in \mathcal{S},\,k\in\mathcal K_i^+$ and $\ell \in \mathcal K_j^-$, is given by
+% 
+\begin{align*} 
+    % 
+	 \mu_i^{k}\mathbb\Psi_{ij}^{k\ell}(s) (\mathcal{A}) 
+	& := \int_{x \in \calD_k} \wrt \mu_i^k(x)
+	% 
+	 \mathbb{E}\left[e^{-s\zeta_Y(\{0\})}1\left\{\varphi({\zeta_Y(\{0\})}) = j, \; X({\zeta_Y(\{0\})}) \in \mathcal{A}\cap\calD_\ell\right\} \mid X(0) = x, Y(0) = 0, \varphi(0) = i\right].
+	 % 
+\end{align*}
+
+Bean and O'Reilly \cite[Theorem~1]{bo2014} give the following result which characterises \(\mathbb\Psi(s)\).
+\begin{theo} 
+	\label{theo:Psi} 
+	For \textit{Re}$(s) \geq 0$, $\mathbb\Psi(s)$ satisfies the  equation: 
+	% 
+	\begin{align*} 
+		\mathbb{D}^{+-}(s) + \mathbb\Psi(s)\mathbb{D}^{-+}(s)\mathbb\Psi(s) + \mathbb{D}^{++}(s)\mathbb\Psi(s) + \mathbb\Psi(s)\mathbb{D}^{--}(s) = 0. 
+	\end{align*} 
+	% 
+	Furthermore, if $s$ is real then $\mathbb\Psi(s)$ is the minimal nonnegative solution. 
+\end{theo}
+
+\subsubsection{Stationary Distribution}
+
+Let $\mathbb\Psi := \mathbb\Psi(0)$.
+We define $\zeta_Y^n(\{0\}) := \inf\{t \geq \zeta_Y^{n - 1}(\{0\}): Y(t) = 0\}$, for $n \geq  2$, to be the sequence of hitting times to level $0$ of $Y(t)$, with $\zeta_Y^1(\{0\}): = \zeta_Y(\{0\})$.
+Consider a discrete-time Markov process $\{X({\zeta_Y^n(\{0\})}), \varphi(\zeta_Y^n(\{0\})),n \geq 1\}$, and for $i \in \mathcal{S},\,k\in\mathcal K_i^-$ define the measures $\bbxi_i^k$ as follows
+	%
+	\begin{align*}
+		\bbxi_i^k(\mathcal{A}) := \lim_{n \rightarrow \infty} \mathbb{P}\left(X(\zeta_Y^n(\{0\})) \in \mathcal{A}\cap\calD_k, \varphi(\zeta_Y^n(\{0\})) = i\right).
+	\end{align*} 
+	By \citep{bo2014}, the vector of measures $\boldsymbol{\bbxi} := (\bbxi_i^k)_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}$ satisfies the following set of equations 
+ % 
+ 	\begin{align}
+		\vligne{\boldsymbol{\bbxi}  & \boldsymbol{0}}\left(-\left[\begin{array}{ll} 
+			\mathbb{B}^{--} & \mathbb{B}^{-0} \\
+			\mathbb{B}^{0-} & \mathbb{B}^{00} 
+		\end{array} \right]\right)^{-1}\left[\begin{array}{l} 
+			\mathbb{B}^{-+} \\ 
+			\mathbb{B}^{0+}
+		\end{array} \right]\mathbb\Psi & = \boldsymbol{\bbxi}, \label{eqn:xi1}\\ 
+		\sum_{k\in\mathcal K^-}\sum_{i \in \mathcal{S}_k^-}\bbxi_i^k(\mathcal{F}^-_i) & = 1. \label{eqn:xi2}
+	\end{align}
+
+We reproduce Theorem 2 of \citep{bo2014} below, which gives the joint stationary distribution of $\{X(t), Y(t), \varphi(t)\}$.
+Recall that the joint stationary density operator ${\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{X(t), Y(t), \varphi(t)\}$ and the joint stationary mass operator ${\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$ are defined by~\eqref{eqn:jointpi} and \eqref{eqn:jointmass}, respectively.
+%Following out notational convention, define an equivalent joint stationary density operator $\boldsymbol{\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{X(t), Y(t), \varphi(t)\}$ and an equivalent joint stationary mass operator $\boldsymbol{\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$, but partitioned accprsing .
+We can partition \(\bbpi\) as follows 
+% 
+	\begin{align*} 
+		\bbpi(y) &= \vligne{\bbpi^{+}(y) & \bbpi^{-}(y) & \bbpi^{0}(y)} 
+		\\&= \vligne{\left(\bbpi^{k}_i(y)\right)_{i \in \mathcal{S}_k^+,k\in\mathcal K^+} & \left(\bbpi^{k}_i(y)\right)_{i \in \mathcal{S}_k^-,k\in\mathcal K^-} & \left(\bbpi^{k}_i(y)\right)_{i \in \mathcal{S}_k^0,k\in\mathcal K^0}},
+	\end{align*} 
+	%
+	where  
+% 
+	\begin{align*} 
+		\bbpi_i^{k}(y)(\mathcal{A}) = \bbpi_i(y)(\mathcal{A}\cap \mathcal D_k).
+	\end{align*} 
+%
+Similarly we can write 
+	\begin{align*} 
+		\mathbb{p} &= \vligne{\mathbb{p}^{-} & \mathbb{p}^{0}} 
+		= \vligne{\left(\mathbb p^{k}_i\right)_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}  & \left(\mathbb p^{k}_i\right)_{i \in \mathcal{S}_k^0,k\in\mathcal K^0} },
+	\end{align*} 
+	%
+	where  
+% 
+	\( 
+		\mathbb p_i^{k}(\mathcal{A}) = \mathbb p_i(\mathcal{A}\cap \mathcal D_k).
+	\) 
+%
+\begin{theo} 
+	\label{theo:density} 
+The density ${\bbpi}^{m}(y)$, for $m \in \{+,-,0\}$ and $y > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the following set of equations:
+% 	 
+	\begin{align} 
+	&\label{11} \; \bbpi^{0}(y) = \vligne{\bbpi^{+}(y) & \bbpi^{-}(y)}\left[\begin{array}{l} \mathbb{B}^{+0} \\ \mathbb{B}^{-0} \end{array} \right]\left(-\mathbb{B}^{00}\right)^{-1}, \\
+	% 
+	&  \vligne{\bbpi^{+}(y) & \bbpi^{-}(y)} = \vligne{\mathbb{p}^{-} & \mathbb{p}^{0}}\left[\begin{array}{l} \mathbb{B}^{-+} \\ \mathbb{B}^{0+} \end{array} \right]\vligne{e^{\mathbb{K}y} & e^{\mathbb{K}y}\mathbb\Psi}\left[\begin{array}{cc} \mathbb{R}^{+} & 0 \\ 0 & \mathbb{R}^{-}\end{array}\right], \\
+	% 
+	&  \vligne{\mathbb{p}^{-}  & \mathbb{p}^{0}} = z \vligne{{\bbxi} & \bs{0}} 
+	\left(-\left[\begin{array}{ll} 
+		\mathbb{B}^{--} & \mathbb{B}^{-0} \\
+		\mathbb{B}^{0-} & \mathbb{B}^{00} 
+		\end{array} \right] \right)^{-1},  \label{eqn:mass}\\
+	%  
+	& \sum_{m \in \{+,-,0\}}\sum_{i \in \mathcal{S}} \int_{y = 0}^{\infty} \bbpi_i^{m}(y)(\mathcal{F}^{m}_i)\wrt y + \sum_{m \in \{-,0\}} \sum_{i \in \mathcal{S}}\mathbb p^{m}_i(\mathcal{F}^{m}_i) = 1, \label{14}
+	\end{align}
+	% 
+	where $\mathbb{K} := \mathbb{D}^{++}(0) + \mathbb\Psi\mathbb{D}^{(-+)}(0)$ and $z$ is a normalising constant. 
+\end{theo}
+
+At this point we reiterate that Equations~(\ref{11})-(\ref{14}) are operator equations and are only amenable to numerical evaluation in the simplest of cases.
+Sources of this intractability come from, for example, the need to find the inverse operator \(( - \mathbb B^{00})^{-1}\), and the need to find the solution, \(\mathbb \Psi(s)\), of the operator equation in Theorem~\ref{theo:Psi}.
+There is also the complexity of the partition of the operators defined by the sets \(\mathcal F_{i}^m,\, i\in\calS,\,m\in\{+,-,0\}\).
+Therefore, there is the need for approximation schemes such as the DG scheme we introduce next.
+
+\subsection{Matrix-exponential distributions}
 Here we recount some facts about matrix exponential distributions.
@@
 
-%\section{Rational arrival processes}
+%\subsection{Rational arrival processes}
 %Let \(N\) be a simple point process, with event time \(Y_0=0<Y_1<Y_2<\cdot\).
@@
 
-\section{QBD-RAPs}\label{sec: qbd-rap}
+\subsection{QBD-RAPs}\label{sec: qbd-rap}
 As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes.
@@
 Now consider a Marked RAP, \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\{-1,0,+1\}).\) The process \(\{(L(t),\bs A(t))\}_{t\geq0}\) formed by letting \(L(t) = N_{+1}(t) - N_{-1}(t)\) is a QBD-RAP.
+
+\subsection{Convergence theorems}
+\begin{thm}[Extended Continuity Theorem \citep{feller1957}]\label{thm: ext cont thm}
+	For \(p=1,2,\dots\) let \(U_p\) be a measure with Laplace transform \(\zeta_p\).
+If \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda > a\geq 0\), then \(\zeta\) is the Laplace transform of a measure \(U\) and \(U_p\to U\).
+
+	Conversely, if \(U_p\to U\) and the sequence \(\{\zeta_p(a)\}\) is bounded, then \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda >a\). 
+\end{thm}

### 2022-01-28 chapter2/chapter.tex
same draft

@@
 \chapter{Approximating fluid queues with the discontinuous Galerkin method \label{ch:galerkin}} 
-% PREFACE WITH CONTRIBUTIONS 
-% TUTORIAL: HOW TO -- AS IN PEPER
-% PROBLEMS AND SOLUTIONS 
-% 	NON-NEGATIVITY AND OSCILLATIONS 
-%	FILTERING AND LIMITING 
-%	NOT A SOLUTION IN OUR CONTEXT & MOTIVATION FOR A NEW METHOD -- THE QBD-RAP
+\begin{center}
+\begin{minipage}{0.8\textwidth}
+    \textit{Sections~\ref{sec:DG} and \ref{sec: approx r} have been taken from Sections~4 and 5 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis.
+I am a co-author of the paper \cite{blnos2022}.
+The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara.
+I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme.
+I contributed Sections~4.4 and 5.1.
+I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6.
+Appendix~A is also my original work.
+I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
+    }
+\end{minipage}
+\end{center}
+% The DG method for B 
+    % direct copy paste from the paper
+% Problems with DG, non-neg/oscillations
+    % solutions: filtering and limiting
+% Approximating R: projection, interpolation, cell averages (eluding to QBD-RAP and unif method)
+% Approximating Psi (and the stationary distribution, but this isnt really necessary)
+    % constructing D
+    % solving the matrix-Riccati equation, iteration/algorithm.
+
+% Appendix: the toy model. 
+% Appendix: some properties of B?
+
+\section{Discontinuous Galerkin Approximation of the Generator of a Fluid Queue}
+	\label{sec:DG}
+Discontinuous Galerkin (DG) methods can be used to approximate the solutions to systems of partial differential equations (PDEs).
+For a more thorough description of these methods see \citep{nodalDGBook}.
+The domain of approximation is partitioned into intervals, referred to individually as \textit{cells} and collectively as a \textit{mesh}.
+On each cell, we have a finite element approximation, which constructs a finite-dimensional smooth Sobolev space using piecewise-polynomial basis functions, and then projects the partial differential equations onto this space.
+This projection leads to a new system of equations, referred to as the \textit{weak form} of the original system of PDEs.
+Next, we must approximate the \textit{flux operator} which moves probability from one cell to another, in a manner similar to the underlying principle of a finite-volume approximation.
+This method conserves probability, and can handle discontinuities, such as jumps and point masses.
+Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the stationary distribution of an SFFM.
+
+\subsection{The Partial Differential Equation}
+We start by introducing the PDE from which we will extract the approximation to the generator \(\mathbb B\).
+
+Let $f_i(x,t)$ be the joint density of $\{(X_t, \varphi_t)\}$: 
+	% 
+	\begin{align*} 
+	    f_i(x,t) := \frac{\partial}{\partial x} 	\mathbb{P}\left(X_t \leq x, \varphi_t = i\right),\quad x>0,\, i \in\calS,
+	\end{align*} 
+which satisfies the system of partial differential equations 
+%
+\begin{align*}
+\frac{\partial}{\partial t} f_i(x,t)& = \sum_{j\in \mathcal{S}}  f_j(x,t)T_{ji} - c_i \frac{\partial}{\partial x} f_i(x,t), \quad x>0, i\in\mathcal S
+\end{align*}
+% 
+subject to suitable boundary conditions~\citep{bo2014}. In matrix form, 
+\begin{align}
+\cfrac{\partial}{\partial t} \boldsymbol f(x,t) &= \boldsymbol f(x,t)\bs T -  \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\bs C\label{eqn:pde_density_matrix}, 
+\end{align}
+where \(\boldsymbol f(x,t) = \left(f_i(x,t)\right)_{i\in\mathcal S}\) is a row-vector. 
+This system of PDEs is closely related to the generator \(\mathbb B\).
+Write \(\boldsymbol \mu(t)(\cdot) = (\mu_j(t)(\cdot))_{j\in\mathcal S}\), then, for \(\mathcal A\subset (0,\infty)\), and assuming \(\boldsymbol \mu(t)\) admits a density,
+\begin{align*}
+	\boldsymbol \mu(t)(\mathcal A) = \int_{x\in\mathcal A}\boldsymbol f(x,t)\wrt x.
+\end{align*}
+That is, \(f_i(x,t)\) is the density of \(\mu_i(t)\). 
+Then \(\boldsymbol \mu(t)\) satisfies the operator differential equation
+\begin{align*}
+	\cfrac{\wrt}{\wrt t}\boldsymbol \mu(t)(\wrt x) = \boldsymbol \mu(t)\mathbb B(\wrt x) = \boldsymbol f(x,t)\bs T\wrt x - \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\bs C\wrt x,
+\end{align*}
+on the interior of the space \([0,\infty)\).
+Thus, by approximating the operator on the right-hand side of Equation (\ref{eqn:pde_density_matrix}) we can approximate the infinitesimal operator \(\mathbb B\).
+The DG method does exactly this, by approximating the operator with a matrix.
+
+\subsection{Cells, Test Functions, and Weak Formulation}
+To begin with, consider an unbounded first fluid level \(\{\widehat X_t,t\geq0\}\), \(\widehat X_t\in(-\infty,\infty)\).
+We will eventually truncate this space so that we have a finite dimensional approximation; however, this requires a discussion on boundary conditions which we save for later.
+Let \(\mathcal D_k = [x_k,x_{k+1}],\, k\in\mathbb Z\) partition the domain \((-\infty,\infty)\).
+We call the \(\calD_k\) \textit{cells}.
+
+On each cell \(\calD_k\) we choose \(N_k\) linearly independent functions \(\{\phi_r^k\}_{r=1}^{N_k}\), compactly supported on \(\calD_k\) (i.e.~\(\phi_r^k(x)=0\) for \(x\notin\calD_k\)) to form a basis for the space \(W_k\), in which we formulate the approximation.
+Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi_r^k\}_{r=1}^{N_k}\) to be the space of polynomials of degree \(N_k-1\).
+It is convenient in this work to take \(\{\phi_r^k\}_{r=1}^{N_k}\) as a basis of Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature points, since our approximations inherit nice properties from this \citep{nodalDGBook}.
+However, the constructions presented here are general and any basis can be used.
+For the sake of illustration, the reader may think of \(\{\phi_r^k\}_{r=1}^{N_k}\) as the Lagrange polynomials.
+On each cell \(\mathcal D_k\) we approximate
+\[f_i(x,t)\approx u_i^k(x,t)=\sum\limits_{r=1}^{N_k}a_{i,r}^k(t)\phi_r^k(x),\] 
+where \(a_{i,r}^k(t)\) are yet-to-be-determined time-dependent coefficients.
+We refer to \(u_i^k\) as the \textit{local} approximation on \(\calD_k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_i^k\) on the whole domain.
+The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\).
+
+Let \(\mathcal N_k := \left\{1,\dots,N_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
+\[\boldsymbol \phi^k(x) = (\phi_r^k(x))_{r\in\mathcal N_k}, \quad \boldsymbol a_i^k(x) = (a_{i,r}^k(x))_{r\in\mathcal N_k},\,i\in\mathcal S.\]
+%Also define \textit{global} row-vectors 
+%\[\boldsymbol \phi(x) = (\boldsymbol\phi^k(x))_{k\in\{1,...,K\}},\quad \boldsymbol a_i(x) = (\boldsymbol a_{i}^k(x))_{k\in\{1,...,K\}},\,i\in\mathcal S.\]
+Note that we will always use the letter \(r\) to index the basis function within each cell.
+
+The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation (\ref{eqn:pde_density_matrix}).
+In general, to construct the weak-form we need a set of test functions, say \(W\).
+Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to the derivative with respect to \(x\), to get
+\begin{align}
+\begin{multlined}[t]
+	\int_{x\in\mathbb R}\cfrac{\partial}{\partial t} f_j(x,t)\psi(x)\wrt x = \int_{x\in\mathbb R} \sum_{i\in\calS}f_i(x,t)T_{ij}\cfrac{\wrt}{\wrt x}\psi(x)\wrt x 
+	%
+	+  \int_{x\in\mathbb R} f_j(x,t)c_j\psi(x)\wrt x \\{}- [ f_j(x,t)c_j\psi(x)]_{x=-\infty}^{x=\infty}, \end{multlined}\label{eqn:weak form}
+\end{align}
+for \(j\in\calS\).
+It is common to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero.
+Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE.
+For a sufficiently rich set of test functions \(W\) the weak and strong forms of the PDE are equivalent.
+Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE.
+For example, this may allow discontinuities with respect to \(x\) in the solution -- something which is ill-defined for the strong form.
+
+For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\), the same as the set of basis functions of our solution space.
+Proceeding as described above, the weak formulation is
+\begin{align*}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j(x,t)\phi_r^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i(x,t)T_{ij}\phi_r^k(x)\wrt x  
+	%
+	+  \int_{x\in\calD_k} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x \\ {}- [f_j(x,t)c_j\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}, \end{multlined}
+\end{align*}
+since \(\phi_r^k\) is compactly supported on \(\calD_k\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\).
+Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi_r^k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\).
+Also, note that \(\cfrac{\wrt}{\wrt x}\phi_r^k(x)\in W\).
+Using this, we can write
+\begin{align*}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi_r^k(x)\wrt x 
+	= \int_{x\in\calD_k}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi_r^k(x)\wrt x  
+	%
+	\\ {}
+	+  \int_{x\in\calD_k} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x - [f_j(x,t)c_j\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}, \end{multlined}
+\end{align*}
+	which is equivalent to
+\begin{align}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi_r^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi_r^k(x)\wrt x  
+	%
+	+  \int_{x\in\calD_k} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x \\{} - [f_j(x,t)c_j\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}. \end{multlined} \label{eqn:hash}
+\end{align}
+Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_k\), it can be expressed as \(\boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}\), which we now substitute into (\ref{eqn:hash})
+%\begin{align*}
+%	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}\phi_r^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_i^k(t) \boldsymbol \phi^k(x)\tr{}T_{ij}\phi_r^k(x)\wrt x
+%	%
+%	\\ {}+  \int_{x\in\calD_k}\boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\phi_r^k(x)\wrt x - c_j[f_j(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}.\end{multlined}
+%\end{align*}
+and repeat this for all test functions \(\phi_r^k(x)\), \(r=1,...,N_k\), to get the following system of equations,
+\begin{align}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}\boldsymbol \phi^k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_i^k(t) \boldsymbol \phi^k(x)\tr{}T_{ij}\boldsymbol\phi^k (x)\wrt x
+	%
+	\\{}+  \int_{x\in\calD_k}\boldsymbol a_j^k(t) \boldsymbol \phi^k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\boldsymbol\phi^k(x)\wrt x - c_j[f_j(x,t) \boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}},\quad k\in\mathbb Z.
+\end{multlined} \label{eqn:DG system}
+\end{align}
+
+\subsection{Mass, Stiffness, and Flux}
+For \(k\in\mathbb Z\), define local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
+\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi^k(x)\tr{}\boldsymbol \phi^k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi^k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi^k(x)\wrt x.\]
+%and global mass and stiffness matrices by
+%\begin{align*}
+%	M &:= \left[\begin{array}{ccc} M_1 & & \\ & \ddots & \\ & & M_K \end{array}\right] = \int_{x\in\mathcal I}\boldsymbol \phi(x)\tr{}\boldsymbol \phi(x)\wrt x,
+%	\\G &:= \left[\begin{array}{ccc} G_1 & & \\ & \ddots & \\ & & G_K \end{array}\right] = \int_{x\in\mathcal I}\boldsymbol \phi(x)\tr{}\cfrac{\wrt}{\wrt x}\boldsymbol \phi(x)\wrt x,
+%\end{align*}
+We can write (\ref{eqn:DG system}) as 
+\begin{align}
+	&\cfrac{\wrt}{\wrt t} \boldsymbol a_j^k(t) \bs M_k = \sum_{i\in\calS} \boldsymbol a_i^k(t)M_kT_{ij} 
+	%
+	+  c_j\boldsymbol a_j^k(t) \bs G_k - c_j[f_j(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}}.\label{eqn:DG matrix}
+\end{align}
+
+It remains to approximate the \textit{flux}, \(f_j(x,t)\) at the cell edges \(x_k,\,k\in\mathbb Z\), so that we may evaluate the terms \([f_j(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}\), \(r=1,...,N_k,\,k\in\mathbb Z\).
+This is the key for DG -- it joins the local approximations on each cell \(\calD_k\), into a global approximation on the whole domain of approximation.
+The flux is the instantaneous rate (with respect to time) at which density moves across the boundaries \(x_k,\,k\in\mathbb Z\).
+There are different choices for the flux, and we refer the reader to \citep{c99,nodalDGBook}, and references therein, for some discussion of the topic.
+Here, we choose the \textit{upwind} scheme, which, as we shall see, closely resembles the flux terms from the generator \(\mathbb B\).
+The approximate flux, also known as the \textit{numerical flux}, is given by
+\[f^*_j(x,t) = sign(c_j)\lim_{\varepsilon\to0^+}u_j(x-\varepsilon c_j,t),\]
+at each \(x=x_k,\,k\in\mathbb Z\). 
+Intuitively, the upwind flux takes the value of the density immediately on the upwind side of each \(x_k\).
+
+Denote by \(x^-\) and \(x^+\) the left and right limits at \(x\), respectively. Assume first \(c_j>0\), then 
+\begin{align*}
+	-c_j[f_j(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}} & \approx -c_j[f_j^*(x,t)\phi_r^k(x)]_{x=x_k}^{x=x_{k+1}}
+	\\& = -c_jf_j^*(x_{k+1},t)\phi_r^k(x_{k+1}) + c_jf_j^*(x_{k},t)\phi_r^k(x_{k})
+	\\& = -c_ju_j(x_{k+1}^-,t)\phi_r^k(x_{k+1}) + c_ju_j(x_{k}^-,t)\phi_r^k(x_{k})
+	\\& = -c_ju_j^k(x_{k+1}^-,t)\phi_r^k(x_{k+1}) + c_ju_j^{k-1}(x_{k}^-,t)\phi_r^k(x_{k})
+	\\& = -c_j\boldsymbol a_j^k(t)\boldsymbol \phi^k(x_{k+1}^-)\tr{}\phi_r^k(x_{k+1}) + c_j\boldsymbol a_j^{k-1}(t)\boldsymbol \phi^{k-1}(x_{k}^-)\tr{}\phi_r^k(x_{k}).
+\end{align*}
+In matrix form,  
+\begin{align*}
+	-c_j[f_j(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}} & \approx -c_j[f_j^*(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}}
+	\\& = -c_j\boldsymbol a_j^k(t)\boldsymbol \phi^k(x_{k+1}^-)\tr{}\boldsymbol\phi^k(x_{k+1}) + c_j\boldsymbol a_j^{k-1}(t)\boldsymbol \phi^{k-1}(x_{k}^-)\tr{}\boldsymbol\phi^k(x_{k})
+	\\& = c_j\boldsymbol a_j^k(t)F_j^{k,k}+c_j\boldsymbol a_j^{k-1}(t)F_j^{k-1,k},
+\end{align*}
+where, for \(j\in \calS\) with \(c_j>0\), we define \(\bs F_j^{k,k} := -\boldsymbol \phi^k(x_{k+1}^-)\tr{}\boldsymbol\phi^k(x_{k+1}),\,k\in\mathbb Z\) and \(\bs F_j^{k-1,k} := \boldsymbol \phi^{k-1}(x_{k}^-)\tr{}\boldsymbol\phi^k(x_{k}),\,k\in\mathbb Z\).
+
+Now proceed similarly for \(c_j<0\) to get the approximation 
+\begin{align*}
+	-c_j[f_j(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}} & \approx -c_j[f_j^*(x,t)\boldsymbol\phi^k(x)]_{x=x_k}^{x=x_{k+1}}
+	\\& = -c_j\boldsymbol a_j^{k+1}(t)\boldsymbol \phi^{k+1}(x_{k+1}^+)\tr{}\boldsymbol\phi^k(x_{k+1}) + c_j\boldsymbol a_j^{k}(t)\boldsymbol \phi^{k}(x_{k}^+)\tr{}\boldsymbol\phi^k(x_{k})
+	\\& = c_j\boldsymbol a_j^{k+1}(t)\bs F_j^{k+1,k}+c_j\boldsymbol a_j^{k}(t)\bs F_j^{k,k},
+\end{align*}
+where, for \(j\in \calS\) with \(c_j<0\), we define \(\bs F_j^{k+1,k} := -\boldsymbol \phi^{k+1}(x_{k+1}^+)\tr{}\boldsymbol\phi^k(x_{k+1}),\,k\in\mathbb Z,\) and \(\bs F_j^{k,k} := \boldsymbol \phi^{k}(x_{k}^+)\tr{}\boldsymbol\phi^k(x_{k}),\,k\in\mathbb Z\).
+
+The matrices \(\bs F_j^{k-1,k},\,\bs F_j^{k,k},\) and \(\bs F_j^{k+1,k}\) are the local flux matrices.
+For convenience, we also define the matrices \(\bs F_j^{k,k+1}=0\) for \(c_j<0\) and \(\bs F_j^{k,k-1}=0\) for \(c_j>0\), \(k\in\mathbb Z\).
+
+% from which we can assemble the global flux matrices
+%\begin{align*}
+%	F_j &= \left[\begin{array}{cccccc}
+%		F_j^{1,1} & F_j^{1,2} & & & & \\
+%		 & F_j^{2,2} & F_j^{2,3} & & & \\
+%		& & \ddots & \ddots & & \\
+%		& & & \ddots & \ddots & \\
+%		& & & & F_j^{K-1,K-1} & F_j^{K-1,K} \\
+%		& & & & & F_j^{K,K}
+%	\end{array}\right],\quad c_j\geq0,
+%	%
+%	\\F_j &= \left[\begin{array}{cccccc}
+%		F_j^{1,1} & & & & & \\
+%		F_j^{2,1} & F_j^{2,2} & & & & \\
+%		& \ddots & \ddots & & & \\
+%		& & \ddots & \ddots & & \\
+%		& & & F_j^{K-1,K-2} & F_j^{K-1,K} & \\
+%		& & & & F_j^{K-1,K} & F_j^{K,K}
+%	\end{array}\right],\quad c_j<0.
+%\end{align*}
+%
+%We can now write 
+%\begin{align*}
+%	&\cfrac{\wrt}{\wrt t} \boldsymbol a_j(t) M = \sum_{i\in\calS} \boldsymbol a_j(t)MT_{ij} 
+%	%
+%	+  c_j\boldsymbol a_j(t) (G + F_j).
+%\end{align*}
+%
+%Defining \(\boldsymbol a(t) = (\boldsymbol a_j(t))_{i\in\calS}\), then 
+%\begin{align}\label{eqn:DG ODE}
+%	&\cfrac{\wrt}{\wrt t} \boldsymbol a(t)(I_{N_\calS}\otimes M) = \boldsymbol a(t){  B,}
+%\end{align}
+%where 
+%\[{  B} = \left[(T\otimes M) +
+%	\left[\begin{array}{ccc}
+%		c_1(G+F_1) & &  \\
+%		& \ddots & \\
+%		& & c_{N_\calS}(G+F_{N_\calS})\\
+%	\end{array}\right] \right],\]
+%\(\otimes\) is the kronecker product and \(I_{N_\calS}\) is the \(N_\calS\times N_\calS\) identity matrix.
+
+%So far, this DG approximation has been constructed according to the lexicographic ordering of \(\calS\times\{1,...,K\}\).
+A reordering of this construction according to the lexicographic ordering of \(\{1,...,K\}\times\calS\) helps elucidate the connection with the partitioned operator \(\mathbb B\) and eases notation slightly for the next discussion on boundary conditions.
+
+%Also define \textit{global} row-vectors 
+%\[\boldsymbol \phi(x) = (\boldsymbol\phi^k(x))_{k\in\{1,...,K\}},\quad \boldsymbol a_j(x) = (\boldsymbol a_{i}^k(x))_{k\in\{1,...,K\}},\,i\in\mathcal S.\]
+
+To write this out as a \textit{global} system, define the row-vectors 
+\[\boldsymbol a^k(t) = (\boldsymbol a_{i}^k(t))_{i\in\mathcal S},\quad \boldsymbol a(t) = (\boldsymbol a^k(t))_{k\in\mathbb Z},\]
+and the block-diagonal matrix 
+\begin{align*}
+\widetilde{\bs M} &= \left[\begin{array}{ccc}\ddots&&\\&\bs I_{N_\calS}\otimes \bs M_k&\\&&\ddots\end{array}\right],
+\intertext{where \(N_\calS=|\calS|\), \(\otimes\) is the Kronecker product, and the block-tridiagonal matrix}
+\widetilde{\mathfrak B} &= \left[\begin{array}{ccccc}
+	\ddots & \ddots & \ddots & & \\
+	& \widetilde{\mathfrak B}^{k,k-1} & \widetilde{\mathfrak B}^{k,k} & \widetilde{\mathfrak B}^{k,k+1} & \\
+	& & \ddots & \ddots & \ddots 
+\end{array}\right],
+\end{align*}
+where, for \(k\in\mathbb Z\), 
+\begin{align*}
+    \widetilde{\mathfrak{\bs B}}^{kk}&=\left[\begin{array}{ccc}T_{11}\bs M_k + c_1(\bs F_1^{kk}+\bs G_k) & T_{12}{\bs M_k} & T_{1N_\calS}{\bs M_k}   \\ T_{21}\bs M_k & & \\ \vdots &\ddots & \vdots \\ & &   T_{N_\calS-1,N_\calS}{\bs M_k} \\  T_{N_\calS1}{\bs M_k} &  T_{N_\calS,N_\calS-1}{\bs M_k} & T_{N_\calS,N_\calS}{\bs M_k} +c_{N_\calS}(\bs F_{N_\calS}^{kk}+\bs G_k)\end{array}\right],\\
+	\widetilde{\mathfrak{\bs B}}^{k,k+1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k+1}&  & \\  &\ddots & \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k+1} \end{array}\right],\\
+	\widetilde{\mathfrak{\bs B}}^{k,k-1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k-1}&  & \\  &\ddots &  \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k-1} \end{array}\right].
+\end{align*} 
+The matrices \(\widetilde{\mathfrak{\bs B}}^{k\ell}\) are defined by sub-blocks; denote these sub-blocks by \(\widetilde{\mathfrak{\bs B}}^{k\ell}_{ij}\):
+\begin{align*}
+	\widetilde{\mathfrak{\bs B}}^{kk}_{ij} &= \begin{cases}T_{ij}\bs M_k + c_i(\bs F_i^{kk}+\bs G_k) & i=j,\\T_{ij}\bs M_k & i\neq j,\end{cases}
+\\	\widetilde{\mathfrak{\bs B}}^{k\ell}_{ij} &= \begin{cases}c_i\bs F_i^{k\ell} & i=j,\\0 & i\neq j,\end{cases}\quad \ell \in \{k-1,k\}.
+\end{align*}
+The global system of equations is 
+\begin{align}\label{eqn:DG ODE}
+	&\cfrac{\wrt}{\wrt t} \boldsymbol a(t)\widetilde{\bs M} = \boldsymbol a(t){\widetilde{\mathfrak{\bs B}}}.
+\end{align}
+
+\subsection{Boundary conditions}
+To enable computation, this numerical approximation has to take place on a finite interval, which means we must consider a bounded domain and specify boundary conditions.
+Recall from Section~\ref{sec:prelim} that we wish to impose a regulated boundary at \(x=0\).
+To apply the DG method, we must truncate the state space of the first fluid at some finite interval upper bound, \([0,\mathcal I]\), for some \(\mathcal I<\infty\), and specify the boundary behaviour at \(x=\mathcal I\).
+Here we consider \(\mathcal I\) to be a regulated boundary.
+Let us denote the doubly-bounded fluid level by \(\overline X_t\).
+Ultimately, we wish to approximate a fluid-fluid queue where the first fluid level, \(X_t\), is bounded below at 0, only.
+Thus, the first step in the approximation scheme is to approximate \(X_t\) by \(\overline X_t\).
+The truncation of \(X_t\) to \(\overline X_t\) will result in an artificial point mass at the upper bound, which we have to address properly.
+It is important to choose an \(\mathcal I\) sufficiently large to control the error induced by the artificial upper bound, however, with larger \(\mathcal I\) there comes increased computational burden.
+We shall further comment on this in Section~\ref{sec:numexp}, where we report our numerical experiments.
+
+Let $[0,\mathcal{I}]$ be the domain of the approximation, where $\mathcal{I} < \infty$, and assume there is a regulated boundary for \(\{\overline X_t\}\) at \(x=\mathcal I\).
+We partition the space $[0,\mathcal{I}]$ into \(\mathcal D_\nabla=\{0\},\) \(\mathcal D_\Delta=\{\mathcal I\},\) and \(K\) non-trivial intervals, \(\calD_k=[x_k,x_{k+1}]\setminus \{\{0\}\cup\{\mathcal I\}\},\, x_k<x_{k+1},\, k=1,...,K\), \(x_1=0,\,x_{K+1}=\mathcal I\) and define \(h_k := x_{k+1}-x_k\).
+The notation \(\Delta\) refers to quantities and sets which are relevant to the boundary at \(\mathcal I\).
+
+For states with \(c_i\leq 0\), there is the possibility of point mass accumulating at the boundary at~\(0\).
+Denote these point masses by \(q_{\nabla,i}(t)\) for \(i\in\mathcal S_\nabla\).
+For states with \(c_i>0\) there is no possibility of a point mass at \(0\).
+Similarly, for \(c_i\geq 0\) there is the possibility of a point mass at \(\mathcal I\).
+Denote these point masses by \(q_{\Delta,i}(t)\), for \(i\in\mathcal S_\Delta\).
+For states with \(c_i<0\) there is no possibility of a point mass at \(\mathcal I\).
+Let \(\bs q_\nabla(t)=(q_{\nabla,i}(t))_{i\in\calS_\nabla}\) and \(\bs q_\Delta(t) = (q_{\Delta,i}(t))_{i\in\calS_\Delta}\) and \(\bs f_m(x,t) = (f_i(x,t))_{i\in\calS_m}\), \(m\in\{+,-,0\}\).
+
+Let us first consider the boundary at \(\overline X_t=0\).
+\cite{bo2014} show the following boundary conditions describe the evolution of probability/density of a stochastic fluid model with a regulated boundary at \(0\);
+\begin{align}\label{eqn:BC1}
+\cfrac{\wrt}{\wrt t}\bs q_\nabla(t) &= \bs q_\nabla(t) T_{\nabla \nabla} - \boldsymbol f_\nabla(0,t)C_\nabla,
+\\\label{eqn:BC2}
+\bs q_\nabla(t)T_{\nabla+}&=\bs f_+(0,t)C_+.
+\end{align}
+Equation (\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(T_{\nabla\nabla}\), and that the flux of probability density into the point masses is \(- \boldsymbol f_\nabla(0,t)C_\nabla\).
+Substituting the DG approximation for \(\boldsymbol f_\nabla(0,t)\) into (\ref{eqn:BC1}) gives, for \(j\in\calS_\nabla\),
+\begin{equation*}\label{eqn:DGBC1}
+\cfrac{\wrt}{\wrt t} q_{\nabla,j}(t) = \sum_{i\in\calS_\nabla} q_{\nabla,i}(t) T_{i j} - \boldsymbol a^1_j(t)\boldsymbol \phi^1(0)\tr{}c_j.
+\end{equation*}
+Equation (\ref{eqn:BC2}) describes the flux of probability mass to density upon a transition from a phase in \(\calS_\nabla\) to a phase in \(\calS_+\).
+Thus, the flux into the left-hand edge of \(\calD_1\) in phase \(j\in\calS_+\) is, \(\sum\limits_{i\in\calS_\nabla} q_{\nabla,i}(t)T_{ij}\).
+Therefore, we can now evaluate
+\begin{align*}
+	-c_j[f_j(x,t)\boldsymbol\phi^1(x)]_{x=0}^{x=x_{1}} & =  -c_jf_j(x_1,t)\boldsymbol\phi^1(x_1)+c_jf_j(0,t)\boldsymbol\phi^1(0)
+	\\& \approx -c_j(f_j^*(x_1,t)\boldsymbol\phi^1(x_1)+\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)T_{ij}\boldsymbol\phi^1(0)
+	\\& = c_j\boldsymbol a_j^{1}(t)F_j^{1,1}+\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)T_{ij}\boldsymbol\phi^1(0),
+\end{align*}
+for \(i \in \calS_+\). 
+Thus, the DG approximation of the flux into point masses \(q_{\nabla,j}(t)\) is \(-\boldsymbol a_j^1(t)\boldsymbol \phi^1(0)\tr{} c_j,\,j\in\calS_-\), the rate of transition of point mass within \(\bs q_{\nabla}(t)\) is \(T_{\nabla\nabla}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_\nabla}q_{\nabla,i}(t)T_{ij}\boldsymbol\phi^1(0),\,j\in\calS_+\).
+
+Similarly, for the boundary at \(X_t=\mathcal I\) the boundary conditions are 
+\begin{align*}
+\cfrac{\wrt}{\wrt t}\bs q_\Delta(t) &= \bs q_\Delta(t) T_{\Delta \Delta} + \boldsymbol f_\Delta(\mathcal I,t)C_\Delta,\\
+\bs q_\Delta(t)T_{\Delta-}&=-\bs f_-(\mathcal I,t)C_-.
+\end{align*}
+Using the same arguments as above, 
+\begin{align*}
+\cfrac{\wrt}{\wrt t} q_{\Delta,j}(t) &= \sum_{i\in\calS_\Delta}q_{\Delta,i}(t) T_{ij} + \boldsymbol a^K_j(t)\boldsymbol \phi^K(\mathcal I)\tr{}c_j,
+\\-c_j[f_j(x,t)\boldsymbol\phi^K(x)]_{x=x_K}^{x=\mathcal I} & \approx c_j\boldsymbol a_j^{K}(t)F_j^{K,K}+\sum_{i\in\calS_\Delta}q_{\Delta,i}(t)T_{ij}\boldsymbol\phi^K(\mathcal I),
+\end{align*}
+for \(j\in\calS_\Delta\). 
+Thus, the DG approximation of the flux into the point mass \(q_{\Delta,j}(t)\) is \(\boldsymbol a_j^K(t)\boldsymbol \phi^K(0)\tr{} c_j\), \(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{\Delta}(t)\) is \(T_{\Delta\Delta}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_\Delta}q_{\Delta,i}(t)T_{ij}\boldsymbol\phi^K(\mathcal I)\), \(j\in\calS_-\).
+
+To include this behaviour in the DG generator we truncate the doubly-infinite matrix \(\widetilde{\mathfrak{\bs B}}\) at \(k=1\) and \(k=K\), then append \(|\mathcal S_\nabla|\) rows and columns to the top and left, and \(|\mathcal S_\Delta|\) rows and columns to the bottom and right.
+These represent the point masses \(\bs q_\nabla(t)\) and \(\bs q_\Delta(t)\), respectively.
+Given the discussion above, the truncated matrix is
+\[\widehat{\mathfrak{\bs B}} = \left[\begin{array}{llllll}
+	\bs T_{\nabla\nabla}& \widetilde{\mathfrak{\bs B}}^{\nabla1} & & & & \\
+	\widetilde{\mathfrak{\bs B}}^{1\nabla} & \widetilde{\mathfrak{\bs B}}^{11} & \widetilde{\mathfrak{\bs B}}^{12} & & & \\
+	& \widetilde{\mathfrak{\bs B}}^{21} & \widetilde{\mathfrak{\bs B}}^{22} & \widetilde{\mathfrak{\bs B}}^{23} & & \\
+	& & \ddots & \ddots & \ddots & \\
+	& & \widetilde{\mathfrak{\bs B}}^{K-1,K-2} &\widetilde{\mathfrak{\bs B}}^{K-1,K-1} & \widetilde{\mathfrak{\bs B}}^{K-1,K} & \\
+	& & &\widetilde{\mathfrak{\bs B}}^{K,K-1} & \widetilde{\mathfrak{\bs B}}^{K,K} & \widetilde{\mathfrak{\bs B}}^{K,\Delta} \\
+	& & & & \widetilde{\mathfrak{\bs B}}^{\Delta, K} & \bs T_{\Delta\Delta}
+\end{array}\right],\]
+where \(\widetilde{\mathfrak{\bs B}}^{\nabla1} := \bs T_{\nabla+}\otimes \boldsymbol\phi^1(0)\), \(\widetilde{\mathfrak{\bs B}}^{1\nabla} :=-\diag(c_i\mathbb 1_{(c_i<0)})_{i\in\calS}\otimes \boldsymbol \phi^1(0)\tr{}\), \(\widetilde{\mathfrak{\bs B}}^{\Delta K} := \bs T_{\Delta-}\otimes \boldsymbol\phi^K(\mathcal I)\) and \(\widetilde{\mathfrak{\bs B}}^{K,\Delta} := \diag(c_i\mathbb 1_{(c_i>0)})_{i\in\calS} \otimes \boldsymbol \phi^K(\mathcal I)\tr{},\) and \(\otimes\) is the Kronecker product.
+Where we have used the same sub-block notation as we have for \(\widetilde{\mathfrak{\bs B}}\).
+
+After the addition of the boundary conditions, the system of ODEs (\ref{eqn:DG ODE}) can now be written as 
+\begin{align}\label{eqn: DG ODE w BCs}
+	\cfrac{\wrt}{\wrt t} \vligne{\boldsymbol q_{\nabla}(t) & {\boldsymbol a}(t) & \boldsymbol q_\Delta(t)} 
+	% 
+	= \vligne{\boldsymbol q_{\nabla}(t) & {\boldsymbol a}(t) & \boldsymbol q_\Delta(t)} \widehat{\mathfrak{\bs B}}\widehat{\bs M}^{-1},
+\end{align}
+where \(\widehat{\bs M} = \left[\begin{array}{ccccc}
+		\bs I_{|\calS_\nabla|} & & & & \\
+		& \bs I_{N_S}\otimes \bs M_1 & & & \\
+		& & \ddots & & \\
+		& & & \bs I_{N_S}\otimes \bs M_K & \\
+		& & & & \bs I_{|\calS_\Delta|}
+	\end{array}\right]\).
+Define \( \bs B = \widehat{\mathfrak{\bs B}}\widehat{\bs M}^{-1}\), with the sub-block as we used for \(\widetilde{\mathfrak{\bs B}}\).
+
+Regarding our notational convention, the matrices in fraktur fonts (e.g.~\(\mathfrak{\bs B}\)) are intermediary constructions that are not directly referred to for the rest of the paper (but do appear again in the appendix).
+We use regular mathematics fonts to represent DG approximations to operators, i.e.~\(\bs B\) is a DG approximation to \(\mathbb B\) and \(\bs \Psi\) is an approximation to \(\mathbb \Psi\).
+
+We prove the following result in Appendix\ref{appendix:properties}.
+%This transformation is not strictly necessary and all the following calculations can be done with either form, as long as week keep a consistent interpretation of the DG generator and related coefficients.
+However, it is convenient to prove properties of the generator.
+\begin{cor}
+	The approximate generator \( \bs B\) conserves probability. That is, for all \(t\geq 0\),
+	\begin{align*}
+	\begin{multlined}[t]\sum_{i\in\calS_\nabla}q_{\nabla,i}(t)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(t)+\sum_{i\in\calS} \int_{x\in[0,\mathcal I]}u_i(x,t)\wrt x 
+	%
+	\\= \sum_{i\in\calS_\nabla}q_{\nabla,i}(0)+\sum_{i\in\calS_\Delta}q_{\Delta,i}(0)+\sum_{i\in\calS} \int_{x\in[0,\mathcal I]}u_i(x,0)\wrt x.\end{multlined}
+	\end{align*}
+\end{cor}
+
+\subsection{Putting it all together}
+Recall that the ultimate goal for our DG approximation is to approximate the operator \(\mathbb B\).
+We have that \({  \bs B}^{k\ell}\) is an approximation to \(\mathbb B^{k\ell}\), \(k,\ell \in \{\nabla,1,...,K,\Delta\}\).
+%\[\mu_i^k\mathbb B_{ii}^{kk}(\wrt x)\approx\bs a_i^k   B_{ii}^{kk} \bs \phi^k(x)\tr{} = \bs a_i^k \left[T_{ii}M_k + c_i(F_i^{kk}+G_k)\right] \bs \phi^k(x)\tr{}\wrt x. \]
+%
+%Substituting in the approximations to the flux into Equation \eqref{eqn:DG matrix} and post multiplying by \(M_k^{-1}\) and \(\bs \phi^k(x)\tr{}\) we get 
+%\begin{align*}
+%	&\cfrac{\wrt}{\wrt t} \boldsymbol a_i^k(t)\bs \phi^k(x)\tr{}  
+%	%
+%	\\&= \begin{cases}\displaystyle\sum_{i\in\calS} \boldsymbol a_i^k(t)\left[T_{ij}  
+%	+  c_i G_kM_k^{-1} + c_i F_i^{k,k}M_k^{-1}\right] \bs \phi^k(x)\tr{} + c_i\bs a_i^{k-1}(t)F_i^{k-1,k}M_k^{-1}\bs \phi^k(x)\tr{} & c_i>0 \\
+%	%
+%	%
+%	\displaystyle\sum_{i\in\calS} \boldsymbol a_i^k(t)\left[T_{ij}  
+%	+  c_i G_kM_k^{-1} + c_i F_i^{k,k}M_k^{-1}\right] \bs \phi^k(x)\tr{} + c_i\bs a_i^{k+1}(t)F_i^{k+1,k}M_k^{-1}\bs \phi^k(x)\tr{} & c_i<0\\
+%	%
+%	%
+%	\displaystyle\sum_{i\in\calS} \boldsymbol a_i^k(t)T_{ij}\bs \phi^k(x)\tr{} &c_i=0.
+%	\end{cases}
+%\end{align*}
+%Intuitively, \(\boldsymbol a_i^k(t)T_{ii}\bs \phi^k(x)\tr{} +  c_i\boldsymbol a_i^k(t) G_kM_k^{-1}\bs \phi^k(x)\tr{} + c_i\bs a_i^k(t)F_i^{k,k}M_k^{-1}\bs \phi^k(x)\tr{} \) can be seen to approximate the density of \(\mu_i^k\mathbb B_{ii}^{kk}(\wrt x)\) -- the first term \(\boldsymbol a_i^k(t)T_{ii}\bs \phi^k(x)\tr{}\) represents stochastic jumps out of phase \(i\), the second term \(c_i\boldsymbol a_i^k(t) G_kM_k^{-1}\bs \phi^k(x)\tr{}\) represents the movement of density within cell \(\calD_k\) by moving the density between basis functions, and the last term \(c_i\bs a_i^k(t)F_i^{k,k}M_k^{-1}\bs \phi^k(x)\tr{}\) represents the flow of density out of the cell \(\calD_k\).
+Similarly \(\boldsymbol a_i^k(t)T_{ij}\bs \phi^k(x)\tr{}\) can be seen to approximate the density of \(\mu_i^k\mathbb B_{ij}^{kk}(\wrt x)\), \(i\neq j\), stochastic jumps from phase \(i\) to phase \(j\) within cell \(\calD_k\).
+Also \(c_i\bs a_i^{k-1}(t)F_i^{k-1,k}M_k^{-1} \bs \phi^k(x)\tr{}\) and \(c_i\bs a_i^{k+1}(t)F_i^{k+1,k}M_k^{-1}\bs \phi^k(x)\tr{} \) approximate the densities of \(\mu_i^{k-1}\mathbb B_{ij}^{k-1,k}(\wrt x)\) and \(\mu_i^{k+1}\mathbb B_{ij}^{k+1,k}(\wrt x)\), respectively.
+
+Given we have now truncated the space and added boundaries, let us define \(\mathcal M_{0,\mathcal I}\) as the set of measures, \(\mu_i\), which admit an absolutely continuous density on \((0,\mathcal I)\), may have a point mass at \(x=0\) if \(i\in\mathcal S_\nabla\), and another at \(x=\mathcal I\) if \(i\in\calS_{\Delta}\).
+The set \(\mathcal M_{0,\mathcal I}\) is the domain of the operator \(\mathbb B\) truncated to the interval \([0,\mathcal I]\) with regulated boundaries at \(x=0\) and \(x=\mathcal I\).
+Also, redefine \(\mathcal K^m_i = \{k\in\{\nabla,1,...,K,\Delta\}\mid   l  (\calD_k\cap\calF_i^m) =    l (\calD_k) \}\) for \(i\in\calS,m\in\{+,-,0\}\) for all \(   l\in\mathcal M_{0,\mathcal I} \).
+
+Approximations \( \bs B^{mn}_{ij}\), \( \bs B_{ij}\), and \( \bs B^{mn}\) to \(\mathbb B^{mn}_{ij}\), \(\mathbb B_{ij}\), and \(\mathbb B^{mn}\), \(i,j\in\calS,\,m,n\in\{+,-,0\}\), are constructed from the block-matrices \({  \bs B}^{k\ell}_{ij}\), \(i,j\in\calS\), \(k,\ell\in\{\nabla,1,\dots,K,\Delta\}\), as
+\begin{align*}
+	{  \bs B}_{ij}^{m n} &= \left[{  \bs B}_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\quad i,j\in\calS,\,m,n\in\{+,-,0\},
+%\intertext{an approximation \({  B}_{ij}\) to \(\mathbb B_{ij}\), \({i,j\in\calS},\) is}
+\\	{  \bs B}_{ij} &= \left[{  \bs B}_{ij}^{k\ell}\right]_{k,\ell\in\{\nabla,1,...,K,\Delta\}},\,{i,j\in\calS},
+%\intertext{and an approximation \({  B}^{mn}\) to \(\mathbb B^{mn}\), \(m,n\in\{+,-,0\}\) is}
+\\	{  \bs B}^{m n} &= \left[\left[{  \bs B}_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n},\,m,n\in\{+,-,0\}.
+\end{align*}
+
+\section{Application to an SFFM}\label{sec:DGSFFM}
+Given our approximation \(\bs B\) to the generator \(\mathbb B\) we now follow the recipe from \citep{bo2014}, replacing the actual generator \(\mathbb B\) with its approximation \({  \bs B}\), to construct approximations, \(\bs \pi\) and \(\bs p\), to the stationary operators, \(\bbpi\) and \(\mathbb p\).
+
+It may be convenient to think of our approximations in terms of approximations of kernels.
+Recall that the operators in \citep{bo2014} can be thought of in terms of kernels.
+That is, for some function \(\bs g = (g_i(x))_{i\in\calS}\), we can write \(\bs \mu \mathbb B \boldsymbol g\tr{} = \displaystyle \sum\limits_{k,\ell\in\{\nabla,1,\dots,K,\Delta\}}\sum\limits_{i,j\in\calS} \displaystyle \int_{x,y}\wrt \mu_i(x) \mathbb B_{ij}^{k\ell}(x,\wrt y)g_j(y)\) where \(\mathbb B_{ij}^{k\ell}(x,\wrt y)\) is the kernel of the operator \(\mathbb B_{ij}^{k\ell}\).
+
+Let \(\boldsymbol a^\nabla(t):=\boldsymbol q_\nabla(t)\) and \(\boldsymbol a^\Delta(t):=\boldsymbol q_\Delta(t)\), and define basis functions \(\bs\phi^\nabla(x) = \phi^\nabla_1(x) = \delta(x)\) and \(\bs\phi^\Delta(x) = \phi^\Delta_1(x) = \delta(x-\mathcal I)\), where \(\delta\) is the Dirac delta function, \(N_\nabla = N_\Delta = 1\), and \(\mathcal N_\nabla = \mathcal N_\Delta = \{1\}\).
+Also define \(\widehat M_\nabla=I_{|\calS_\nabla|}\) and \(\widehat M_\Delta=I_{|\calS_\Delta|}\) and row-vectors
+\[\boldsymbol \phi(x) = (\bs\phi^k(x))_{k\in\{\nabla,1,...,K,\Delta\}}, \quad \boldsymbol a_i(t) = (\bs a_{i}^k(t))_{k\in\{\nabla,1,...,K,\Delta\}},\,i\in\calS.\]
+
+To pose the approximation \(\mathbb B\) in kernel form let \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in W,\,i\in\calS\) be the initial density of the process, and \(\boldsymbol \phi(x)\bs b_i\tr{}\in W,\,i\in\calS\) be a test function.
+Observe that, from our DG construction earlier and the definition of \(\widehat{\bs M}\),
+\[\sum_{i,j\in\calS}\int_{x,y\in[0,\mathcal I]} \boldsymbol a_i \boldsymbol \phi(x)\tr{} \boldsymbol \phi (x) \wrt x \widehat{\bs M}^{-1}{  \bs B}_{ij} \bs \phi(y)\tr{} \bs \phi(y)\bs b_j \wrt y= \sum_{i,j\in\calS} \boldsymbol a_i {  \bs B}_{ij} \widehat{\bs M} \bs b_j .\]
+Thus, we can think of 
+\[\boldsymbol \phi (x) \widehat{\bs M}^{-1}{  \bs B}_{ij} \bs \phi(y)\tr{}\wrt y,\]
+as an approximation to the kernel \(\mathbb B_{ij}(x,\wrt y)\). This concept can be extended to all the approximations of operators considered in this work.
+
+\subsection{Approximating the operator \(\mathbb R\)}\label{sec: approx r}
+Recall the operator \(\mathbb R^k\) from Lemma~\ref{lemma: D(s)}.
+Essentially, the operator \(\mathbb R^k\) takes an initial measure \(\boldsymbol \mu^k\) and multiplies each element by \(1/|r_i(x)|\) on cells \(\calD_k\) where \(r_i(x)\neq 0\).
+In the context of DG the initial distribution is given by \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in W,\,i\in\calS\).
+Thus, for \(k\in\{\nabla,1,...,K,\Delta\}\) such that \(r_i(x)\neq0\) on \(\calD_k\), we have
+\[\boldsymbol a_i^k \boldsymbol \phi^k(x)\tr{}\mathbb R^k_i = \cfrac{\boldsymbol a_i^k \boldsymbol \phi^k(x)\tr{}}{|r_i(x)|}.\]
+Decompose the right-hand side into a component which lies in \(W\) and another orthogonal to W: 
+\[\cfrac{\boldsymbol a_i^k \boldsymbol \phi^k(x)\tr{}}{|r_i(x)|} = \bs \rho^k_i \bs \phi^k(x)\tr{} + g_i^\perp(x),\] where \(\bs \rho^k_i \bs\phi^k(x)\tr{}\in W\), \(g_i^\perp \in W^\perp\).
+Now, multiply by test functions \(\{\phi_r^k(x)\}_{r=1}^{N_k}\) and integrate over \([0,\mathcal I]\):
+\begin{align*}
+	\boldsymbol a_i^k\int_{x\in[0,\mathcal I]} \cfrac{ \boldsymbol \phi^k(x)\tr{}\bs \phi^k(x)}{|r_i(x)|}\wrt x
+	&=\bs \rho^k_i \int_{x\in[0,\mathcal I]}\bs \phi^k(x)\tr{}\bs \phi^k(x)\wrt x + \int_{x\in[0,\mathcal I]} g_i^\perp(x)\bs\phi^k(x)\wrt x 
+	%
+	\\&= \bs \rho^k_i \int_{x\in[0,\mathcal I]}\bs \phi^k(x)\tr{}\bs \phi^k(x)\wrt x = \bs \rho_i^k\bs M_k,
+\end{align*}
+since \(g_i(x)^\perp\in W^\perp\). Define the matrix \(\bs M_k^r := \displaystyle\int_{x\in[0,\mathcal I]} \cfrac{ \boldsymbol \phi^k(x)\tr{}\bs \phi^k(x)}{|r_i(x)|}\wrt x\), then 
+\(
+	\boldsymbol a_i^k\bs M_k^r
+	= \bs \rho^k_i M_k,
+\)
+which implies
+\(
+	\bs \rho^k_i  = \boldsymbol a_i^k\bs M_k^r\bs M_k^{-1}.
+\)
+Thus, we have the approximation 
+\[\boldsymbol a_i^k \boldsymbol \phi^k(x)\tr{}\mathbb R^k_i = \cfrac{\boldsymbol a_i^k \boldsymbol \phi^k(x)\tr{}}{|r_i(x)|}\approx \boldsymbol a_i^k\bs M_k^r\bs M_k^{-1}\bs\phi^k(x)\tr{}.\]
+Since \(\boldsymbol a_i^k\) is arbitrary, we see that we approximate \(\mathbb R_i^k\) by \(  \bs R_i^k = \bs M_k^r\bs M_k^{-1},\)
+and \(\mathbb R^k\) by \(  \bs R^k = \diag(  \bs R_i^k)_{i\in\calS^\bullet_k}\).
+
+In practice, we implement a Gauss-Lobatto quadrature approximation to compute the elements of \(\bs M_k^r\).
+
+\subsection{Approximating the operator \(\bs D\) and the DG Riccati equation}
+Recalling Lemma~\ref{lemma: D(s)} and replacing the operators \(\mathbb R^k\) and \(\mathbb B^{\ell m}\), by their approximations we have the following approximation to \(\mathbb D^{mn}(s)\)
+\begin{align*}
+		 {\bs D}^{mn}(s) = \left[ \bs {R}^{m}\left(
+		{ {\bs B}}^{mn} - s{\bs I} + { {\bs B}}^{m0 }\left({ {\bs B}}^{00}- s\bs I \right)^{-1} { {\bs B}}^{0n}\right)\right],\quad m,n \in\{+,-\}.
+\end{align*}
+
+Let \(\bs\phi^k(x){\bs M}_k^{-1}\bs \Psi_{ij}^{k\ell}(s)\bs\phi^\ell(y)\tr{}\wrt y\), \(i,j\in\calS,\) \(k\in\mathcal K_i^+\,, \ell\in\mathcal K_j^-\) be a finite-dimensional approximation of the operator kernel \(\mathbb\Psi_{ij}^{k\ell}(s)(x,\wrt y)\), where \(\bs \Psi_{ij}^{k\ell}(s)\) is a matrix of constants for a given \(s\).
+Construct an approximation to \(\mathbb\Psi(s)(x,\wrt y)\) by
+\[\bs\phi^+(x)\widehat{\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt y = \left[\left[\bs\phi^k(x){M}_k^{-1}\bs \Psi_{ij}^{k\ell}(s)\bs\phi^\ell(y)\tr{}\wrt y\right]_{i\in\calS_k^+,j\in\calS_\ell^-}\right]_{ k\in\mathcal K^+, \ell\in\mathcal K^-},\]
+where \(\bs\phi^+(x) = (\bs\phi^k(x))_{i\in\calS_k^+,k\in\mathcal K^+}\) and \(\bs\phi^-(y) = (\bs\phi^k(y))_{i\in\calS_k^-,k\in\mathcal K^-}\) are row-vectors, \(\bs \Psi(s)\) is a matrix of constants for a given \(s\) with the same size as \(\bs D^{+-}\), and \(\widehat{\bs M}_m,\) \(m\in\{+,-,0\}\) is a block diagonal matrix \(\widehat{\bs M}_m = \diag\left( \left( \bs M_k\right)_{i\in\calS_k^m}\right)_{k\in\mathcal K^m}\), \(m\in\{+,-,0\}\).
+Now replace the theoretical kernels in Theorem~\ref{theo:Psi} by their DG approximations to get
+\begin{align*}
+&\bs\phi^+(x)\widehat{\bs M}_+^{-1}  \bs D^{+-}(s)\bs\phi^-(y)\tr{}\wrt y
+\\&{}+ \int_{z_1,z_2}\bs\phi^+(x)\widehat{\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(z_1)\tr{}\bs\phi^-(z_1) \widehat{\bs M}_-^{-1}  \bs D^{-+}(s)\bs\phi^+(z_2)\bs\phi^+(z_2)\widehat{\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt z_1\wrt z_2\wrt y
+\\&{}+ \int_{z_1}\bs\phi^+(x)\widehat{\bs M}_+^{-1}  \bs D^{++}(s)\bs\phi^+(z_1)\tr{}\bs\phi^+(z_1) \widehat{\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt z_1\wrt y
+\\&{}+ \int_{z_1}\bs\phi^+(x)\widehat{\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(z_1)\tr{}\bs\phi^-(z_1)\widehat{\bs M}_-^{-1}  \bs D^{--}(s)\bs\phi^-(y)\tr{}\wrt z_1\wrt y
+= 0.
+\end{align*}
+Multiplying on the left by \(\bs\phi^+(x)\tr{}\) and on the right by \(\bs\phi^-(y)\), integrating over both \(x\) and \(y\), then post-multiplying by \(\widehat{\bs M}^{-1}_-\) gives the following matrix Ricatti equation
+\begin{align}\label{eqn:RiccatiPsi}
+    \bs D^{+-}(s)
++ \bs \Psi(s)   \bs D^{-+}(s)\bs \Psi(s)
++   \bs D^{++}(s)\bs \Psi(s)
++ \bs \Psi(s)  \bs D^{--}(s)
+= 0.
+\end{align}
+Thus, we may find \(\bs \Psi(s)\) by solving \eqref{eqn:RiccatiPsi}, using one of the methods in \citep{bot08}. Here, we use the Newtons method.
+
+Given the stochastic interpretation of \(\mathbb\Psi(0)\) we know that \( \bbnu \mathbb\Psi(0)([0,\infty))=1\) for every vector of measures \( \bbnu\) such that \( \bbnu([0,\infty)\boldsymbol 1 = 1\), when an SFFM is recurrent.
+It appears that this result carries over to the matrix \(\bs \Psi(0)\) giving the property that \(\displaystyle\int_{y\in[0,\mathcal I]} \bs \Psi(0)\bs\phi^-(y)\tr{}\wrt y = \bs 1\).
+However, we have only observed this numerically and have no proof of this property.
+
+\subsection{Putting it all together: constructing an approximation to the stationary distribution}
+We find an approximation to the stationary distribution by replacing the theoretical operators in Theorem~\ref{theo:density} with their approximations.
+Table \ref{table:notations} defines the notation we use for the DG approximations to stationary operators.
+
+ \begin{table}[h!]
+ \centering
+ \begin{tabular}{c|c|c|c}
+	\begin{tabular}{c}Exact \\operator\end{tabular} & Operator indices & \begin{tabular}{c}Approximation \\ notation\end{tabular} & Approximations \\\hline 
+	%
+	\( \bbxi_i^k \) & \(i\in\calS_k^-,\,k\in\mathcal K^-\)  & \(\bs \xi_i^k := (\xi_{i,r}^k)_{r\in\mathcal N_k}\) & 
+	\(%\begin{array}{c}
+	\bbxi_i^k(\wrt x)\approx  \bs{\xi}_{i}^k \bs \phi^k(x)\tr{}\wrt x,
+	 %\\\bbxi_i^\nabla(\{0\}) %:=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
+	%\approx \bs\xi_{i}^\nabla,
+	%\\\bbxi_i^\Delta(\{\mathcal I\}) %:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = \mathcal I, \varphi_{\theta_n} = i\right]
+	%\approx \bs\xi_{i}^\Delta.%\end{array}
+	\)
+	 \\\hline
+%
+%
+	\(\mathbb p_{i}^k\) & \(\begin{array}{c}i\in\calS_k^-\cup\calS_k^0,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_m\end{array}\) & \(\bs p_{i}^k := (p_{i,r}^k)_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
+	\mathbb p_{i}^k(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+	\approx\bs{p}_{i}^k \bs\phi^k(x)\tr{}\wrt x
+	%\\ \mathbb p_{i}^\nabla(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
+	%\approx \bs p_{i}^\nabla,\\ \mathbb p_{i}^\Delta(\{\mathcal I\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = \mathcal I, \varphi_{t} = i\right]
+	%\approx \bs p_{i}^\Delta.% \end{array}
+	\)\\\hline
+%
+%
+	 \(\bbpi_{i}^k(y)\)  & \(\begin{array}{c}i\in\calS,\\k\in\{\nabla,1,\dots,K,\Delta\}\end{array}\) & \(\bs \pi_{i}^k(y) := (\pi_{i,r}^k(y))_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
+	 \bbpi_{i}^k(y)(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+	\approx\bs{\pi}_{i}^k(y) \bs \phi^k(x)\tr{}\wrt x
+	%\\ \bbpi_{i}^\nabla(y)(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
+	%\approx \bs \pi_{i}^\nabla(y),\\ \bbpi_{i}^\Delta(y)(\{\mathcal I\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = \mathcal I, \varphi_{t} = i\right]
+	%\approx \bs \pi_{i}^\Delta(y). \end{array}
+	\)\\\hline
+ \end{tabular}
+ \caption{Notation for the approximation of the stationary operators of an SFFM.
+The first column contains the operators which we are approximating, the second column contains indices for which the operators are defined, the third column defines the notation we use for the coefficients of the approximation, and the last column defines how the coefficients and basis functions are used to approximate the operators.
+\label{table:notations}}
+ \end{table}
+
+%  \begin{table}[h!]
+% \centering
+% \begin{tabular}{c|c|c|c}
+%	\begin{tabular}{c}Exact \\Operator\end{tabular} & Notation & Set of indices & Approximation \\\hline 
+%	%
+%	\( \bbxi_i^k(\wrt x) \) & \(\xi_{i,r}^k\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\mathcal K_i^-\setminus \{\nabla\cup\Delta\}.\end{array}\) & \(%:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\\
+%	 \displaystyle\sum\limits_{r=1}^{N_k} {\xi}_{i,r}^k \phi_r^k(x)\wrt x,\,x\in\calD_k.\)\\\hline
+%
+%	&\(\xi_{i,r}^k\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\mathcal K_i^-\cap \{\nabla\cup\Delta\}.\end{array}\) & \(\begin{array}{l}\bbxi_i^k(\{0\}) %:=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
+%	\approx \xi_{i,r}^\nabla,
+%	\\\bbxi_i^k(\{\mathcal I\}) %:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = \mathcal I, \varphi_{\theta_n} = i\right]
+%	\approx \xi_{i,r}^\Delta.\end{array}\)\\\hline 
+%
+%	&\(p_{i,r}^k\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\setminus \{\nabla\cup\Delta\}.\end{array}\) & \(\mathbb p_{i}^k(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+%	\approx\displaystyle\sum\limits_{r=1}^{N_k} {p}_{i,r}^k \phi_r^k(x)\wrt x,\,x\in\calD_k.\)\\\hline
+%
+%	&\(p_{i,r}^k\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{\nabla\cup\Delta\}.\end{array}\) & \(\begin{array}{l} \mathbb p_{i}^k(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
+%	\approx p_{i,r}^\nabla,\\ \mathbb p_{i}^k(\{\mathcal I\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = \mathcal I, \varphi_{t} = i\right]
+%	\approx p_{i,r}^\Delta. \end{array}\) \\\hline
+%
+%	&\(\pi_{i,r}^k(y)\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\setminus \{\nabla\cup\Delta\}.\end{array}\) & \(\bbpi_{i}^k(y)(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+%	\approx\displaystyle\sum\limits_{r=1}^{N_k} {\pi}_{i,r}^k(y) \phi_r^k(x)\wrt x,\,y>0,\, x\in\calD_k.\)\\\hline
+% \end{tabular}
+% \caption{Notation for the approximation of the stationary operators of an SFFM.\label{table:notations}}
+% \end{table}
+
+With the notation in Table \ref{table:notations} define row-vectors 
+ \begin{align*}
+	%{\bs{\xi}}_i^k &= ( \xi_{i,r}^k)_{r\in\{1,...,N_k\}}, \quad  { i\in\calS,\,k\in\mathcal K_i^-},
+	%
+	{\bs{\xi}}^k &:= ( \bs\xi_{i}^k)_{i\in\calS_k^-}, \quad  {k\in\mathcal K_i^-},
+	%
+	\\ {\bs{\xi}}& := ( \bs \xi^k)_{k\in\mathcal K^-},
+	% 
+	%\\{\bs{p}}_i^k &= (  p_{i,r}^k)_{r\in\{1,...,N_k\}}, \quad  {i\in\calS,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m},
+	%
+	\\\bs{p}^{k,m} &:= (  \bs p_{i}^k)_{i\in\calS_k^m}, \quad  k\in{\mathcal K^m},\, m\in\{-,0\},
+	%
+	\\ {\bs{p}^m} &:= (  \bs p^{k,m})_{k\in\mathcal K^m},\quad m\in\{-,0\},
+	%
+	\\ {\bs{p}} &:= (  \bs p^m)_{m\in\{-,0\}},
+	%
+	%\\{\bs{\pi}}_i^k(y) &= (  \pi_{i,r}^k(y))_{r\in\{1,...,N_k\}},\quad  i\in\calS,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m,
+	%
+	\\{\bs{\pi}}^{k,m}(y) &:= (  \bs \pi_{i}^k(y))_{i\in\calS_k^m},\quad  k\in\{\nabla,1,\dots,K,\Delta\},\,m\in\{+,-,0\},
+	%
+	\\{\bs{\pi}}^m(y) &:= (\bs \pi^{k,m}(y))_{k\in\mathcal K^m},\quad  m\in\{+,-,0\},
+	%
+	\\ {\bs{\pi}}(y) &:= (  \bs \pi^m(y))_{m\in\{+,-,0\}}.
+ \end{align*}
+
+Proceeding similarly to the derivation of the Ricatti equation (\ref{eqn:RiccatiPsi}), we can argue that the coefficients \( {\boldsymbol{\xi}}\) are the solution to the matrix system 
+% 	 
+	\begin{align*}
+		\vligne{ {\boldsymbol{\xi}}  & \boldsymbol{0}}\left(-\left[\begin{array}{ll} 
+			 { {\bs B}}^{--} &  { {\bs B}}^{-0} \\
+                         { {\bs B}}^{0-} &  { {\bs B}}^{00} 
+		\end{array} \right]\right)^{-1}\left[\begin{array}{l} 
+			 { {\bs B}}^{-+} \\ 
+			 { {\bs B}}^{0+}
+		\end{array} \right]\bs \Psi(0) & =  {\boldsymbol{\xi}}, \\ 
+		\int_{x\in[0,\mathcal I]} {\bs \xi}\left[\begin{array}{c}\bs \phi^-(x)\tr{} \\ \bs \phi^0(x)\tr{}\end{array}\right]\wrt x \boldsymbol 1 & = 1. 
+	\end{align*} 
+Essentially, we replace the theoretical operators in (\ref{eqn:xi1}) and (\ref{eqn:xi2}) with their DG counterparts.
+
+Similarly, the coefficients \( {\boldsymbol{p}}\) are given by 
+	\begin{equation}\vligne{\bs{p}^{-}  & \bs{p}^{0}} = z \vligne{{\bs\xi} & \bs{0}} 
+	\left(-\left[\begin{array}{ll} 
+		{\bs B}^{--} & {\bs B}^{-0} \\
+		{\bs B}^{0-} & {\bs B}^{00} 
+		\end{array} \right] \right)^{-1},\label{eqn:pisystem1}\end{equation}
+		where \(z\) is a normalising constant. The coefficients \(\bs \pi(y)\) are given by 
+\begin{align} 
+	\;  {\bs{\pi}}^{0}(y) &= \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)}\left[\begin{array}{l} { {\bs B}}^{+0} \\ { {\bs B}}^{-0} \end{array} \right]\left(-{ {\bs B}}^{00}\right)^{-1}, \\ 
+	% 
+	 \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)}& = \vligne{ {\bs{p}}^{-} &  {\bs{p}}^{0}}\left[\begin{array}{l} { {\bs B}}^{-+} \\ { {\bs B}}^{0+} \end{array} \right]\vligne{e^{ {\bs K}y} & e^{ {\bs K}y}\bs \Psi(0)}\left[\begin{array}{cc}  {\bs R}^{+} & 0 \\ 0 &  {\bs R}^{-}\end{array}\right], \\
+	%  
+	 \sum_{i \in \mathcal{S}}\sum_{k \in \{\nabla,1,...,K,\Delta\}} & \int_{y = 0}^{\infty} \int_{x \in [0,\mathcal I]}  \bs \pi_{i}^{k}(y)\bs\phi^k(x)\tr{}\wrt x\wrt y 
+	 %
+	 \\&{}+  \sum_{i \in \mathcal{S}}\sum_{\ell \in \{-,0\}}\sum_{k\in \mathcal K_i^\ell}  \int_{x \in [0,\mathcal I]}  \bs p^{k}_{i}\bs\phi^k(x)\tr{}\wrt x = 1,\label{eqn:pisystem2}
+	\end{align}
+	% 
+	where $ {\bs K} :=  {\bs D}^{++}(0) + \bs \Psi(0) {\bs D}^{(-+)}(0)$, and $z$ is a normalising constant.
+
+%For \(i,k,r\) such that \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\mathcal K_i^-\setminus \{\nabla\cup\Delta\}\), define coefficients \(  \xi_{i,r}^k\) to be determined later such that 
+%\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}} {\xi}_{i,r}^k \phi_r^k(x),\quad x \in \calD_k,\]
+%is an approximation to the density of \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\mathcal K_i^-\cap \{\nabla\cup\Delta\},\) define point masses \(  \xi_{i,r}^k\) such that \( \xi_{i,r}^\nabla\) and \( \xi_{i,r}^\Delta\) approximate the point masses \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]\) and \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = \mathcal I, \varphi_{\theta_n} = i\right]\), respectively.
+In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\mathcal K_i^-\), let \( {\bs{\xi}}_i^k = ( \xi_{i,r}^k)_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\mathcal K_i^-\), let \( {\bs{\xi}}^k = ( \xi_{i}^k)_{\{i\in\calS\mid r_i(x)<0,x\in\calD_k\}}\) and let \( {\bs{\xi}} = ( \xi^k)_{k\in\bigcup\limits_{i\in\calS}\mathcal K_i^-}\).
+%
+%Proceeding similarly to the derivation of the Ricatti equation (\ref{eqn:RiccatiPsi}), we can argue that the DG approximation to \( {\boldsymbol{\xi}}\) is the solution to the matrix system 
+%% 	 
+%	\begin{align}
+%		\vligne{ {\boldsymbol{\xi}}  & \boldsymbol{0}}\left(-\left[\begin{array}{ll} 
+%			 { {B}}^{--} &  { {B}}^{-0} \\
+%                         { {B}}^{0-} &  { {B}}^{00} 
+%		\end{array} \right]\right)^{-1}\left[\begin{array}{l} \label{eqn:xi1}
+%			 { {B}}^{-+} \\ 
+%			 { {B}}^{0+}
+%		\end{array} \right]\Psi(0) & =  {\boldsymbol{\xi}}, \\ 
+%		\int_y  {\bs \xi}\left[\begin{array}{c}\bs \phi^-(x)\tr{} \\ \bs \phi^0(x)\tr{}\end{array}\right]\wrt x \boldsymbol 1 & = 1. \label{eqn:xi2}
+%	\end{align} 
+%Essentially, we replace the theoretical operators in (\ref{eqn:xi1}) and (\ref{eqn:xi2}) with their DG counterparts. 
+%%	We write 
+%%	\( {\bs\xi} = \vligne{ {\bs\xi}_\nabla &  {\bs\xi}^\circ &  {\bs\xi}_\Delta}\), where \( {\bs\xi}_\nabla=( {\xi}_{\nabla,i})_{\{\i\in\calS_\nabla\mid r_i(0)<0\}}\) and \( {\bs\xi}_\Delta=( {\xi}_{\Delta,i})_{\{\i\in\calS_\Delta\mid r_i(0)<0\}}\) are associated with the boundaries at \(0\) and \(\mathcal I\), respectively, and \( {\bs\xi}^\circ\) is associated with the interior \(x\in(0,\mathcal I)\).
+Let us denote the elements of \( {\bs{\xi}}^\circ\) as \( {\xi}_{i,r}^k\), \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\mathcal K_i^-\) where the elements are ordered by \(k\) then \(i\) and then \(r\).
+%%	We construct an approximation to the density of \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\) as \(\displaystyle\sum\limits_{\substack{k\in\mathcal K_i^- \\r\in\{1,...,N_k\}}} {\xi}_{i,r}^k \phi_r^k(x)\).
+An approximation to \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]\) is \( {\xi}_{\nabla,i}\) and an approximation to the artificial point mass \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = \mathcal I, \varphi_{\theta_n} = i\right]\) is \( {\xi}_{\Delta,i}\).
+%	
+%	Next, we seek to approximate the stationary operators \(\bs p\) defined in (\ref{eqn:jointmass}).
+For \(i,k,r\) such that \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\setminus \{\nabla\cup\Delta\}\), coefficients \(p_{i,r}^k\) to be determined later, such that
+%\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}}p_{i,r}^k \phi_r^k(x),\quad x \in \calD_k,\]
+%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{\nabla\cup\Delta\},\) define point masses \(p_{i,r}^k\) such that \(p_{i,r}^\nabla\) and \(p_{i,r}^\Delta\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} = \mathcal I, \varphi_{t} = i\right]\), respectively.
+In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\), let \( {\bs{p}}_i^k = (  p_{i,r}^k)_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\), let \( {\bs{p}}^k = (  p_{i}^k)_{\{i\in\calS\mid r_i(x)\leq 0,x\in\calD_k\}}\) and let \( {\bs{p}} = (  p^k)_{k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m}\).
+Again, by proceeding similarly to the derivation of (\ref{eqn:RiccatiPsi}), that the DG approximation to \( {\boldsymbol{p}}\) is
+%	\[\vligne{\bs{p}^{-}  & \bs{p}^{0}} = z \vligne{{\bbxi} & \bs{0}} 
+%	\left(-\left[\begin{array}{ll} 
+%		{B}^{--} & {B}^{-0} \\
+%		{B}^{0-} & {B}^{00} 
+%		\end{array} \right] \right)^{-1},\]
+%		where \(z\) is a normalising constant. 
+%That is, we substitute the DG approximations of the operators into the corresponding equation in Theorem~\ref{theo:density}.
+%%	
+%%	For \(i\in\calS_\nabla, m\in\{-,0\}\) define \(  p^m_{\nabla,i}\) as the approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, Y_t=0, \varphi_{t} = i\right]\).
+%%	
+%%	 \( {\bs p}^-_{\nabla}=(  p^-_{\nabla,i})_{\{i\in\calS_\nabla\mid r_i(0)<0\}}\) and \( {\bs p}^0_{\nabla}=(  p^0_{\nabla,i})_{\{i\in\calS_\nabla\mid r_i(0)=0\}}\). 
+%%	Now define \( {\bs p}^m_{\nabla},\, {\bs p}^m_{\circ},\, {\bs p}^m_{\Delta}\) as the approximation to \(\bs p\) at the left-hand boundary, interior, and on the right-hand boundary of \([0,\mathcal I]\), respectively.
+%%	\[ {\bs p}^m := \vligne{ {\bs p}^m_\nabla(y) &  {\bs p}^m_\circ(y) &  {\bs p}^m_\Delta(y)},\quad m\in\{-,0\},\]
+%%where Specifically,  
+%
+%	
+%	Next we approximate the operator \({{\bbpi}}(y)\).
+For \(i,k,r\) such that \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\setminus \{\nabla\cup\Delta\}\), define \(\pi_{i,r}^k(y)\), to be determined later, such that
+%\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}}\pi_{i,r}^k(y) \phi_r^k(x),\quad x \in \calD_k,\]
+%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in\wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\cap \{\nabla\cup\Delta\}\) define point masses \(\pi_{i,r}^k(y)\) such that \(\pi_{i,r}^\nabla(y)\) and \(\pi_{i,r}^\Delta(y)\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t\in\wrt y, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in\wrt y, X_{t} = \mathcal I, \varphi_{t} = i\right]\), respectively.
+In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\), let \( {\bs{\pi}}_i^k(y) = (  \pi_{i,r}^k(y))_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\), let \( {\bs{\pi}}^k(y) = (  \pi_{i}^k(y))_{\{i\in\calS\mid r_i(x)\leq 0,x\in\calD_k\}}\) and let \( {\bs{\pi}}(y) = (  \pi^k(y))_{k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m}\).
+Again, we can argue that the DG approximation to \( {\boldsymbol{\pi}}(y)\) is given by substituting the DG operators for their theoretical counterparts in Theorem~\ref{theo:density}
+%	\begin{align} 
+%	& \;  {\bs{\pi}}^{0}(y) = \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)}\left[\begin{array}{l} { {B}}^{+0} \\ { {B}}^{-0} \end{array} \right]\left(-{ {B}}^{00}\right)^{-1}, \label{eqn:pisystem1} \\
+%	% 
+%	&  \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)} = \vligne{ {\bs{p}}^{-} &  {\bs{p}}^{0}}\left[\begin{array}{l} { {B}}^{-+} \\ { {B}}^{0+} \end{array} \right]\vligne{e^{ {K}y} & e^{ {K}y}\Psi(0)}\left[\begin{array}{cc}  {R}^{+} & 0 \\ 0 &  {R}^{-}\end{array}\right], \\
+%	%  
+%	 &\sum_{r \in \{\nabla,1,...,K,\Delta\}}\sum_{i \in \mathcal{S}_{\ell}} \int_{y = 0}^{\infty}  \pi_{i,r}^{k}(y)\phi_{r}^k(x)\wrt x\wrt y + \sum_{\ell \in \{-,0\}} \sum_{i \in \mathcal{S}}\sum_{r\in \mathcal K_i^\ell}   p^{r}_i(\mathcal{F}^{\ell}_i) = 1,\label{eqn:pisystemend}
+%	\end{align}
+%	% 
+%	where $ {K} :=  {D}^{++}(0) + \Psi(0) {D}^{(-+)}(0)$, and $z$ is a normalising constant.
+%	%
+%%Also \( {\bs p}^-_{\Delta}=(  p^-_{\Delta,i})_{\{i\in\calS_\Delta\mid r_i(\mathcal I)<0\}}\) and \( {\bs p}^0_{\Delta}=(  p^0_{\Delta,i})_{\{i\in\calS_\Delta\mid r_i(\mathcal I)=0\}}\) where \(  p^m_{\Delta,i},\,m\in\{-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\mathcal I, Y_t=0, \varphi_{t} = i\right]\).
+On the interior, we have
+%%\( {\bs p}^m_{\circ}=( {\bs p}_{i}^k)_{i\in\calS,k\in\mathcal K_i^m}\) where \( {\bs p}_i^k=( {p}_{i,r}^k)_{r=1,...,N_k}\).
+On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, Y_t=0, \varphi_{t} = i\right]\) is given by \( {\bs p}_i^k\bs \phi^k(x)\tr{}\).
+%%
+%%Similarly for \( {\bs\pi}\) we write \[ {\bs \pi}^m =\vligne{ {\bs \pi}^m_\nabla(y) &  {\bs \pi}^m_\circ(y) &  {\bs \pi}^m_\Delta(y)},\quad m\in\{+,-,0\},\]
+%%where \( {\bs \pi}^m_{\nabla}(y),\, {\bs \pi}^m_{\circ}(y),\, {\bs \pi}^m_{\Delta}(y)\) correspond to the approximation at the left-boundary, interior, and on the right boundary of \(x\in[0,\mathcal I]\), respectively.
+%%%
+%%Specifically, \( {\bs \pi}^+_{\nabla}(y)=(  \pi^+_{\nabla,i}(y))_{\{i\in\calS_\nabla\mid r_i(0)>0\}}\), \( {\bs \pi}^-_{\nabla}(y)=(  \pi^-_{\nabla,i}(y))_{\{i\in\calS_\nabla\mid r_i(0)<0\}}\) and \( {\bs \pi}^0_{\nabla}(y)=(  \pi^0_{\nabla,i}(y))_{\{i\in\calS_\nabla\mid r_i(0)=0\}}\) where \(  \pi^m_{\nabla,i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, Y_t=\wrt y, \varphi_{t} = i\right]\).
+%%%
+%%Also \( {\bs \pi}^+_{\Delta}(y)=(  \pi^+_{\Delta,i}(y))_{\{i\in\calS_\Delta\mid r_i(\mathcal I)>0\}}\), \( {\bs \pi}^-_{\Delta}(y)=(  \pi^-_{\Delta,i}(y))_{\{i\in\calS_\Delta\mid r_i(\mathcal I)<0\}}\) and \( {\bs \pi}^0_{\Delta}(y)=(  \pi^0_{\Delta,i}(y))_{\{i\in\calS_\Delta\mid r_i(\mathcal I)=0\}}\) where \(  \pi^m_{\Delta,i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\mathcal I, Y_t\in\wrt y, \varphi_{t} = i\right]\).
+%%%
+%%On the interior, we have 
+%%\( {\bs \pi}^m_{\circ}(y)=( {\bs \pi}_{i}^k(y))_{i\in\calS,k\in\mathcal K_i^m}\) where \( {\bs \pi}_i^k(y)=( {\pi}_{i,r}^k(y))_{r=1,...,N_k}\).
+On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, Y_t\in\wrt y, \varphi_{t} = i\right]\) is given by \( {\bs \pi}_i^k(y)\bs \phi^k(x)\tr{}\).
+
+To assist the reader in understanding these constructions and the notation we provide an explicitly worked toy-example in Appendix~\ref{appendix:example}.
+
+\section{A stochastic interpretation of the simplest DG scheme}
+% links between DG and the uniformisation scheme
+% positivity preservation
+% stochastic interpretations
+% how I use/analyse it in the numerics section
+
+\section{Oscillations and slope limiting}\label{sec: slope limiting}
+% why the MUSCL limiter?
+%   it requires no tuning and guarentees non-negativity
+%   some other limiting methods require tuning and can still produce negative results if not tuned properly
+%   ^ same goes for filtering

### 2022-01-28 chapter5/chapter.tex
same draft

@@
 \chapter{Global convergence results}\label{ch: global results}
-In this chapter we prove a number of results which stem from the main results of Chapter~\ref{sec: conv}.
+In this chapter we prove some results which stem from the main results of Chapter~\ref{sec: conv}.
 We consider the discrete-time embedded processes formed by observing the QBD-RAP at the orbit restart epochs and by observing the fluid queue at the hitting times of the level process at the of points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
@@
 In Section~\ref{sec: sub global}, we state global results on the weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
+
+In this chapter we work with the augmented state space scheme to model phases with rates \(c_i=0\) as described in Section~\ref{sec: zero states}.
+The results of this chapter rely on results of Chapter~\ref{sec: conv} and therefore apply to the QBD-RAP scheme which uses ephemeral states to model a fluid queue which starts in a phase with rate \(0\), as described in Section~\ref{sec: initial conditions}.
+However, supplementing the results of Chapter~\ref{sec: conv} with the results from Appendix~\ref{app:extend conv} and using the same arguments from this chapter with only slight modifications, then the results of this chapter can be extended to the augmented state-space QBD-RAP scheme without the initial ephemeral phases.
 
@@
 \section{Convergence of the QBD-RAP scheme}\label{sec: local to global}
-The rest of this chapter is dedicated to proving a global convergence results of the QBD-RAP approximation scheme to the fluid queue.
+This section is dedicated to proving a global convergence results of the QBD-RAP approximation scheme to the fluid queue.
 Ultimately, we prove the weak convergence (in space and time) of the QBD-RAP approximation scheme to the distribution of the fluid queue.
@@
 The set \(\mathcal P^n_{\ell_0,\ell}\) contains all the possible values which \(\{L(\tau_m)\}_{m=2}^{n-1}\) may take on a sample path which starts in level \(\ell_0\), ends in level \(\ell\) and has \(n\) orbit restart epochs.
-Similarly, \(\mathcal P^n_{\ell_0,\ell}\) contains all the possible values which \(\{X(\tau_m^X)\}_{m=2}^{n-1}\) may take on a sample path which starts in band \(\ell_0\), ends in band \(\ell\) at time \(\tau_n^X\).
-
-Thus, by partitioning on the times \(\tau_m\), \(m=2,\dots,n-1\), and the phases and the levels at these times and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is  
+%Similarly, \(\mathcal P^n_{\ell_0,\ell}\) contains all the possible values which \(\{X(\tau_m^X)\}_{m=2}^{n-1}\) may take on a sample path which starts in band \(\ell_0\), ends in band \(\ell\) at time \(\tau_n^X\).
+
+Thus, by partitioning on the phases and the levels at the times \(\tau_m\), \(m=2,\dots,n-1\),and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is  
 	\begin{align}
@@
 % by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution. 
-We can apply Corollary (\ref{cor: aln222}) to the factors of the summands in (\ref{eqn: 161222}) and conclude 
+We can apply Corollary \ref{cor: aln222} to the factors of the summands in (\ref{eqn: 161222}) and conclude 
 \begin{align}
@@
 	\\&\to \mathbb P(\bs X(\tau_1^X) =( y_{\ell_{m}+1(j_m\in\calS_-)},j_m), \tau_{1}^X\leq E^\lambda 
-	\mid \bs X(0) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1}))  \label{eqn:kknvf fawoprgj v2}
+	\mid \bs X(0) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1})).  \label{eqn:kknvf fawoprgj v2}
 \end{align}
@@
 \begin{align}
-	\mathbb P(\bs X(\tau_m^X) =( y_{\ell_{m}+1(j_m\in\calS_-)},j_m), \tau_{m}^X\leq E^\lambda 
-		\mid \bs X(\tau_{m-1}) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1})).\label{eqn: lim 12}
+	&\mathbb P(\bs X(\tau_m^X) =( y_{\ell_{m}+1(j_m\in\calS_-)},j_m), \tau_{m}^X\leq E^\lambda 
+		\mid \bs X(\tau_{m-1}) = (y_{\ell_{m-1}+1(j_{m-1}\in\calS_-)}, j_{m-1}), \nonumber
+	\\&\tau_{m-1}^X\leq E^\lambda ).\label{eqn: lim 12}
 \end{align}
@@
 	%
-	\\ &=\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}\lim_{p\to\infty} \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda \nonumber
+	\\ &=\sum_{\substack{j_2,\dots,j_{n-1}\in\mathcal S \\ (\ell_2,\dots,\ell_{n-1}) \in\mathcal P^{n-1}_{\ell_1,\ell}}}\prod_{m=2}^{n}\lim_{p\to\infty} \mathbb P(\Ydp(1) = (\ell_m, j_m), \tau_{1}^{(p)}\leq E^\lambda
             	 \mid \bs Y^{(p)}(0) = (\ell_{m-1}, \bs\alpha,
@@
 \subsection{Between the \(n\)th and \(n+1\)th orbit restart epochs}\label{sec: between n and np1}
+In the last section we proved a convergence of the QBD-RAP to the fluid queue \emph{at} the \(n\)th orbit restart epoch.
+However, we ultimately want to make a convergence statement about the QBD-RAP scheme at any time.
+The next step is therefore to show a convergence of the QBD-RAP to the fluid queue between the \(n\)th and \(n+1\)th orbit restart epoch, after which we sum over \(n\) to prove a convergence result independent of the number of orbit restart epochs.
+
 Let \(\mathcal T_n^{(p)} = (\tau_{n}^{(p)},\tau_{n+1}^{(p)}]\) and \(\mathcal T_n^X=(\tau_n^X,\tau_{n+1}^X]\). Consider the Laplace transform 
 \begin{align}
-	\nonumber&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid 
-	\bs Y^{(p)}(0)=\bs y_0^{(p)})\psi(x)\wrt t,
+	&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_{\ell,j}}\mathbb P({\bs Y}^{(p)}(t) = (\ell, \wrt x, j), t\in\mathcal T_n^{(p)} \mid 
+	\bs Y^{(p)}(0)=\bs y_0^{(p)})\psi(x)\wrt t, \label{eqn: aaabbababbaa}
 	% \nonumber&\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\calD_\ell}\mathbb P(L^{(p)}(t) = \ell, \widetilde X^{(p)}(t)\in\wrt x, \phi^{(p)}(t) = j, \tau_{n}^{(p)}\leq t<\tau_{n+1}^{(p)} \mid L^{(p)}(0)=\ell_0,
@@
 
-Partitioning on the time of the \(n\)th orbit restart epoch and the phase and level at this time gives
+Partitioning (\ref{eqn: aaabbababbaa}) on the time of the \(n\)th orbit restart epoch and the phase and level at this time gives
 \begin{align}
@@
 	&\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x,j), t\in\mathcal T_n^{X} \mid  
-	\bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)},j_n), \tau_n^X\leq E^\lambda ) \psi(x)  \label{eqn: lkHAHJFHKJ J J 22}
+	\bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)},j_n), \tau_n^X\leq E^\lambda ) \psi(x).  \label{eqn: lkHAHJFHKJ J J 22}
 \end{align}
@@
 \subsection{A domination condition}
+% We now want to sum the expressions in Lemma~\ref{lem: LAkAKFnvnb mav h} over \(n\) to obtain a result which is independent of the number of orbit restart epochs.
+To do so, we use the Dominated Convergence Theorem.
+We prove an appropriate domination condition below.
 Our aim is to prove convergence of 
@@
 Hence, upon taking the limit of (\ref{eqn:GHJKLKJHGHJKL}), if we can justify the swap of the sum and the limit, we will obtain the desired result.
-To this end, we now show a domination condition so tht we may apply the Dominated Convergence Theorem.
+To this end, we now show a domination condition in Corollary~\ref{vcor: cdks d} so that we may apply the Dominated Convergence Theorem.
 %The main result of this subsection is Lemma~\ref{lem: another bound 2} in which we show a bound which can serve as a dominating function so that we may apply the Dominated Convergence Theorem to justify the aforementioned swap of sum and limit.
@@
 \begin{cor}\label{vcor: cdks d}
+	For \(\bs y_0=(\ell_0,\bs a_{\ell_0,i}(x_0),i),\, \lambda>0,\, \ell_0,\ell\in\mathcal K,\,i,j\in\mathcal S\) and any bounded function \(\psi\leq F\), then
 	\begin{align}
@@
 		\psi(x) \wrt t \Bigg|
-		\leq \cfrac{F b^{n-1}}{\lambda}\label{eq: JjbCS  B k }
+		\leq \cfrac{F b^{n-1}}{\lambda}.\label{eq: JjbCS  B k }
 	\end{align}
@@
 	\begin{align}
-		&\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \nonumber
+		&\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_n \nonumber
 		\mid \Yd(1)=(\ell_1,j_1),  \tau_1=u_1)\nonumber
@@
 		%
-		\\&=\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t-u_1\in\mathcal T_{n-1} \nonumber
-		\mid \bs Y(0)=(\ell_1,\bs \alpha,j_1))\nonumber
+		\\&=\int_{t=0}^\infty e^{-\lambda t} \int_{u_1=0}^t \sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\mathbb P( {\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t-u_1\in\mathcal T_{n-1} \nonumber
+		\mid 
+		\bs Y(0)=(\ell_1,\bs \alpha,j_1))
 		\\&\quad\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
@@
 	\begin{align}
-		&\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
+		&\sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
 		\mid \bs Y(0)=(\ell_1,\bs \alpha, j_1)) \wrt t 
@@
 		%
-		\\&=\sum_{j_1\in\mathcal S}\sum_{\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
+		\\&=\sum_{\substack{j_1\in\mathcal S\\\ell_1\in\{\ell_0+1,\ell_0,\ell_0-1\}\cap\mathcal K}}\int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}(t) \in (\ell,\mathcal D_{\ell,j},j), t\in\mathcal T_{n-1} \nonumber
 		\mid \bs Y(0)=(\ell_1,\bs\alpha,j_1)) \wrt t 
@@
 
-	Using the bound (\ref{eq: bndhs vja sdh}) in (\ref{eqn fLks}) gives 
+	Using the bound (\ref{eq: bndhs vja sdh}) in (\ref{eqn fLksoioi}) gives 
 	\begin{align}
@@
 \end{align}
-We can just the swap of the sum and integrals since 
+We can justify the swap of the sum and integrals since 
 \begin{align}
@@
 
-Let \(\mathcal R(L(t),\bs A(t),\varphi(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\varphi(t)}(x)\), \(x\in\calD_{L(t),\varphi(t)}\). Observe that 
+Lemma~\ref{lem: KajPOw} establishes a convergence for a given interval \(\mathcal D_\ell,\, \ell\in\mathcal K\), and phase \(j\in\mathcal S\).
+We now formally extend this to a global result.
+To do so, we find it convenient to re-write the problem in terms of expectations.
+
+Let \(\mathcal R(L(t),\bs A(t),\varphi(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\varphi(t)}(x)\), \(x\in\calD_{L(t),\varphi(t)}\), then
 \begin{align}
 	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \nonumber
-	\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t 
-	\\&=\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y}^{(p)}(t) = (\ell,\wrt x,j) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)})  \psi(x)\wrt t.
-\end{align}
+	\bs Y (0)=\bs y_0 \Bigg] \wrt t 
+	% \\&=\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\int_{x\in\calD_{L(t),\varphi(t)}}\bs A(t)\bs v_{L(t),\varphi(t)}(x)\psi(x,\varphi(t)),\varphi(t) )\wrt x  \mid \nonumber
+	% \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t 
+	\\&=\int_{t=0}^\infty e^{-\lambda t}\int_{\bs a\in\mathcal A}\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{x\in\calD_{\ell,j}}\mathbb P( {\bs Y}(t) \in (\ell,\wrt \bs a,j) \mid \bs Y(0)=\bs y_0)\bs a\bs v_{\ell,j}(x)  \psi(x)\wrt x\wrt t \nonumber
+	\\&=\int_{t=0}^\infty e^{-\lambda t}\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y} (t) \in (\ell,\wrt x,j) \mid \bs Y (0)=\bs y_0 )  \psi(x)\wrt t,
+\end{align}
+and the terms in the last line are those in Lemma~\ref{lem: KajPOw}.
 
@@
 %Further, let \(\mathcal L(X(t)) = \sum_{k\in\mathcal K}k1(X(t)\in\mathcal D_k)\) and \(\widetilde{\bs X}(t) = (\mathcal L(X(t)),X(t),\varphi(t)))\).
-	For each \(x_0\in\mathbb R\), \(i\in\mathcal S\), \(\ell_0\in\mathcal K\), 
-	\begin{align}
-		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \nonumber
+	For each \(i\in\mathcal S\), \(\ell_0\in\mathcal K\), \(x_0\in\mathcal D_{\ell_0,i}\),
+	\begin{align}
+		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )  \mid \nonumber
 		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber
@@
 	\begin{align}
-		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid 
+		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid 
 		  \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
-		\\&= \int_{t=0}^\infty e^{-\lambda t}\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A(t),j),j )  1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+		\\&= \int_{t=0}^\infty e^{-\lambda t}\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )  1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
 		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
-		\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[ \psi( \mathcal R(\ell,\bs A(t),j),j )1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid 
+		\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[ \psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid 
 		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t, \label{eqnL AKJF cjk ajhJK}
 	\end{align}
-	where the swap of the summations and integrals is justified since \(\psi\) is bounded and by the Fubini-Tonelli Theorem.
+	where the swap of the summations and integrals is on the last line justified since \(\psi\) is bounded and by the Fubini-Tonelli Theorem.
 By Lemma~\ref{lem: KajPOw}, for each \(\ell\in\mathcal K\), \(j\in\calS\), the terms
 	 \begin{align}
-	 	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A(t),j), j )\wrt x 1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+	 	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j), j )\wrt x 1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
 		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t
@@
 	\begin{align}
-		&\Bigg|\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A(t),j),\varphi(t) )  1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+		&\Bigg|\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),\varphi(t) )  1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
 		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t\Bigg| \nonumber 
 		%
-		\\&\leq F \sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[1({\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j))  \mid \nonumber
+		\\&\leq F \sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
 		\bs Y^{(p)}(0)=\bs y_0^{(p)}\Bigg] \wrt t \nonumber 
@@
 The Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} can now be used to claim that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.
-\begin{thm}[Extended Continuity Theorem \citep{feller1957}]\label{thm: ext cont thm}
-	For \(p=1,2,\dots\) let \(U_p\) be a measure with Laplace transform \(\omega_p\).
-If \(\omega_p(\lambda)\to\omega(\lambda)\) for \(\lambda > a\geq 0\), then \(\omega\) is the Laplace transform of a measure \(U\) and \(U_p\to U\).
-
-	Conversely, if \(U_p\to U\) and the sequence \(\{\omega_p(a)\}\) is bounded, then \(\omega_p(\lambda)\to\omega(\lambda)\) for \(\lambda >a\). 
+
+\begin{thm}
+	For all \(\bs y_0^{(p)}=(\ell,\bs a_{\ell_0,i}^{(p)}(x_0),i)\), \(\ell_0\in\mathcal K\), \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\calS\), and any bounded function \(\psi:\mathbb R\times \mathcal S \to \mathbb R\), then
+	\begin{align}
+			&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] \nonumber 
+			\to \mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0, i) \right] \nonumber
+	\end{align}
+	weakly in \(t\) as \(p\to \infty\). 
 \end{thm}
-Thus, by the Extended Continuity Theorem~\ref{thm: ext cont thm}
-\begin{align}
-		&\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] \nonumber 
-		\to \mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0, i) \right] \nonumber
-\end{align}
-weakly in \(t\) as \(p\to \infty\).
+\begin{proof}
+	Combine the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} with the convergence of Laplace transforms in Corollary~\ref{cor: lk}. 
+\end{proof}
 
@@
 is a continuous function of \(t\) (it is a Feller semigroup), and, for \(p<\infty\), 
-\[\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \]
+\[\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \]
  is continuous in \(t\), the limit 
- \[\lim\limits_{p\to\infty}\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) ) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \] need not be continuous in \(t\).
+ \[\lim\limits_{p\to\infty}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) ) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \] need not be continuous in \(t\).
 Nonetheless, we can claim that
 \begin{align}
-		&\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )    \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \nonumber
+		&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )    \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \nonumber
 		\to \mathbb E\left[\psi({\bs X}(t) )\mid \bs X(0)=(x_0, i) \right] \nonumber
@@
 for almost all \(t\geq 0\).
-At such values of \(t\), since \(\psi\) is arbitrary and bounded, then the Portmanteau Theorem \citep[Theorem 2.1, page 16]{billingsleyconvergence} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
+At such values of \(t\), since \(\psi\) is arbitrary and bounded, then the Portmanteau Theorem \citep[Theorem 2.1]{billingsleyconvergence} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
 
 A sufficient condition to upgrade the convergence from weak to point-wise (in the variable \(t\)) is to show that for \(t\geq 0\) 
-\[\sup_{p}\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \leq M(t)<\infty\]
-and the sequence \(\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \) is eventually equicontinuous in \(t\).
+\[\sup_{p}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )  \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \leq M(t)<\infty\]
+and the sequence \(\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \) is eventually equicontinuous in \(t\).
 That is, for every \(\varepsilon>0\) there exists a \(\delta(t,\varepsilon)>0\) and an \(p_0(t,\varepsilon)\) such that \(|t-u|<\delta(t,\varepsilon)\) implies that
 \begin{align}
-	&\nonumber \Bigg|\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
-	\\&{}-\mathbb E\left[\psi( \mathcal R(L(u),\bs A(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| \nonumber 
+	&\nonumber \Bigg|\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
+	\\&{}-\mathbb E\left[\psi( \mathcal R(L^{(p)}(u),\bs A^{(p)}(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| \nonumber 
 	<\varepsilon
@@
 for all \(p\geq p_0(t,\varepsilon)\).
+
+% \begin{lem}
+% 	\begin{align}
+% 		&\Bigg|\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
+% 		\\&{}-\mathbb E\left[\psi( \mathcal R(L(u),\bs A(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| < Lc_{max}|t-u| \nonumber
+% 	\end{align}
+% 	i.e.~uniformly Lipschitz in \(t\), and this does not depend on \(p\). 
+% \end{lem}
+% \begin{proof}
+% 	Suppose \(i\in\calS_-\). 
+% 	\begin{align*}
+% 		&\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\varphi(t+u)),\varphi(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
+% 		%
+% 		\\&=\mathbb E\left[\int_{x\in\calD_{L(t+u),\varphi(t+u)}}\bs A(t+u)\bs v_{L(t+u),\varphi(t+u)}\psi(x,\varphi(t+u) )\wrt x   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
+% 		%
+% 		\\&=\int_{L(t+u),\varphi(t+u)}\bs a e^{(c_i\bs S + T_{ii}\bs I)u}\bs v_{\ell,i}(x)\psi(x,i)\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
+% 		%
+% 		\\&\quad{}+\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}\bs v_{\ell+1,i}(x)\psi(x,i)\wrt x 
+% 		%
+% 		\\&\quad{}+o(u^2)
+% 		%
+% 		\\&=\int_{x\in[0,\infty)} \bs a e^{(c_i\bs S + T_{ii}\bs I)u}e^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x
+% 		%
+% 		\\&\quad{}+\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt x 
+% 		%
+% 		\\&\quad{}+o(u^2).
+% 	\end{align*}
+% 	Now, 
+% 	\begin{align*}
+% 		&\int_{x\in[0,\infty)} \bs a e^{(c_i\bs S + T_{ii}\bs I)u}e^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x
+% 		%
+% 		\\&=\int_{x\in[0,\infty)} \bs a e^{c_i\bs S u} e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x),i)\wrt x
+% 		%
+% 		\\&=\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
+% 	\end{align*}
+% 	and 
+% 	\begin{align*}
+% 		&\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt v\wrt x 
+% 		%
+% 		\\&=\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s \int_{x\in[0,\infty)} \bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt x \wrt v
+% 		%
+% 		\\&=\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s \int_{x\in[c_i(u-v),\infty)} \bs \alpha e^{\bs S x}\bs se^{T_{ii}(u-v)}\psi((\ell-1)\Delta+m(x),i)\wrt x \wrt v
+% 		%
+% 		\\&=\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s e^{T_{ii}(u-v)} \left(\psi((\ell-1)\Delta+m(\Delta-c_i(u-v)),i) + r_2\right) \wrt v
+% 		%
+% 		\\&=\int_{v=0}^u\bs a e^{c_i\bs Sv}c_i\bs s e^{T_{ii}u} \left(\psi((\ell-1)\Delta+m(\Delta-c_i(u-v)),i) + r_2 \right) \wrt v
+% 		%
+% 		\\&=\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \left( \psi((\ell-1)\Delta+m(\Delta+x-c_iu)),i) + r_2 \right) \wrt x.
+% 	\end{align*}
+% 	Also, for \(v\leq c_iu\leq \Delta \), 
+% 	\begin{align*}
+% 		\psi((\ell-1)\Delta+m(\Delta+x-c_iu)) &= \psi((\ell-1)\Delta+\Delta - |mod((\Delta+x-c_iu),2\Delta) - \Delta|)
+% 		% 
+% 		\\&=\psi(\ell\Delta - |x-c_iu|).
+% 	\end{align*}
+% 	Now, 
+% 	\begin{align*}
+% 		&\left|\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x \right|
+% 		%
+% 		\\&\leq \left|\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\wrt v\wrt x \right|F
+% 		%
+% 		\\&\leq \left|\sum_{j\in\calS_-} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)} \bs e\wrt v \right|F
+% 		%
+% 		\\&\leq \left|\sum_{j\in\calS_-} \int_{v=0}^uT_{ij}\wrt v\right|F,
+% 		% 
+% 		\\&= \left|\sum_{j\in\calS_-} uT_{ij} \right|F,
+% 	\end{align*}
+% 	similarly, 
+% 	\begin{align*}
+% 		&\left|\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x \right|
+% 		%
+% 		\\&\leq\left|\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\wrt v\wrt x \right|F
+% 		%
+% 		\\&\leq\left|\sum_{j\in\calS_+}\int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs e\wrt v \right|F
+% 		%
+% 		\\&\leq\left|\sum_{j\in\calS_+}\int_{v=0}^uT_{ij}\wrt v \right|F
+% 		%
+% 		\\&=\left|\sum_{j\in\calS_+}uT_{ij}\right|F
+% 	\end{align*}
+
+% 	Also note
+% 	\begin{align*}
+% 		\int_{x\in\calD_{\ell,i}}\bs a\bs v_{\ell,i}(x)\psi(x,i)\wrt x = \int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x.
+% 	\end{align*}
+
+% 	Therefore, consider the derivative
+% 	\begin{align*}
+% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\varphi(t+u)),\varphi(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
+% 		\\&{}-\int_{x\in\calD_{\ell,i}}\bs a\bs v_{\ell,i}(x)\psi(x,i)\wrt x \Bigg)
+% 		\\&= \lim_{u\to 0^+} u^{-1}\Bigg(\int_{x\in[0,\infty)} \bs a e^{(c_i\bs S + T_{ii}\bs I)u}e^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x
+% 		%
+% 		\\&\quad{}+\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell-1)\Delta+m(x),i)\wrt x 
+% 		%
+% 		\\&\quad{}+o(u^2)- \int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg).
+% 		%
+% 		\\&= \lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
+% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi((\ell-1)\Delta+m(\Delta+x-c_iu)),i)\wrt x
+% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
+% 		%
+% 		\\&\quad{}+\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} r_2  \wrt x
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi(\ell\Delta+m(x),j)\wrt v\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{x\in[0,\infty)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}e^{\bs S x}\bs s\psi((\ell+1)\Delta-m(x),j)\wrt v\wrt x
+% 		%
+% 		\\&\quad{}+o(u^2)\Bigg).
+% 	\end{align*}
+% 	The second limit is less than or equal to 
+% 	\begin{align*}
+% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s\wrt xe^{T_{ii}u}r_2 + \left|\sum_{j\in\calS_+}uT_{ij}\right|F + \left|\sum_{j\in\calS_i}uT_{ij}\right|F + o(u^2)\Bigg) = \bs a \bs sr_2 + |T_{jj}|F.
+% 	\end{align*}
+% 	The first limit is 
+% 	\begin{align*}
+% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
+% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi((\ell-1)\Delta+m(\Delta+x-c_iu)),i)\wrt x
+% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
+% 		%
+% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
+% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi(\ell\Delta - |x-c_iu|)\wrt x
+% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
+% 		%
+% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[c_iu,\Delta)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i)\wrt x
+% 		\\&\quad{}+\int_{x=0}^{c_iu}\bs a e^{\bs Sx}\bs s e^{T_{ii}u} \psi(\ell\Delta - |x-c_iu|)\wrt x
+% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
+% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
+% 		%
+% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[0,\Delta)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i)\wrt x
+% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x
+% 		\\&\quad{}-\int_{x=0}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
+% 		%
+% 		\\&=\lim_{u\to 0^+}u^{-1}\Bigg(\int_{x\in[0,\Delta)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i)\wrt x-\int_{x=0}^\Delta\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+x,i)\wrt x
+% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\psi(\ell\Delta+m(x-c_iu),i)\wrt x -\int_{x=\Delta}^\infty\bs ae^{\bs Sx}\bs s\psi(\ell\Delta+m(x),i)\wrt x\Bigg)
+% 		%
+% 		\\&= \lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^\Delta\bs ae^{\bs Sx}\bs s\left|e^{T_{ii}u}\psi(\ell\Delta+(x-c_iu),i) - \psi(\ell\Delta+x,i)\wrt x\right|
+% 		\\&\quad{}+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}\left|\psi(\ell\Delta+m(x-c_iu),i) -\psi(\ell\Delta+m(x),i)\right|\wrt x\Bigg)
+% 		%
+% 		\\&\leq \lim_{u\to 0^+}u^{-1}\Bigg(\int_{x=0}^\Delta\bs ae^{\bs Sx}\bs s L^{\psi}|T_{ii}||c_iu|
+% 		+\int_{x\in[\Delta,\infty)} \bs a e^{\bs Sx}\bs se^{T_{ii}u}L^{\psi}|T_{ii}||c_iu|\wrt x\Bigg)
+% 		%
+% 		\\&= \lim_{u\to 0^+}u^{-1} L^{\psi}|T_{ii}||c_iu|
+% 	\end{align*}
+% \end{proof}
 
@@
 \section{Extension to arbitrary (but fixed) discretisation structures}
-To conclude this chapter we include some remarks as to how to extend the convergence results to arbitrary discretisation structures.
+To conclude this chapter we include some remarks on how to extend the convergence results to arbitrary discretisation structures.
 
@@
 \begin{rem}
-	I suspect that the approximation results can also be extended to approximating so-called \emph{multi-layer fluid queues}, as described in \cite{bo2008}. 
+	I suspect that the convergence results can also be extended to approximating so-called \emph{multi-layer fluid queues}, as described in \cite{bo2008}. 
 \end{rem}

### 2022-01-28 chapter6/chapter.tex
same draft

--- 2021-12-06 chapter6/chapter.tex
@@
 \chapter{Numerical investigations\label{sec: numerics}}
-We have now established multiple methods for approximating fluid queues which suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}.
+We have now established multiple methods for approximating fluid queues which are suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}.
 In this chapter we numerically investigate various aspects of the approximation schemes.
-Throughout, we compare the QBD-RAP scheme (from Chapters~\ref{sec: conv} and \ref{sec: construction and modelling}), the discontinuous Galerkin method (without limiters or filtering of the solution) (as decribed in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013} (which we will refer to a just the \emph{uniformisation scheme}).
-Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux.
-Like the QBD-RAP scheme, due to its stochastic interpretation, the uniformisation scheme is positivity preserving.
+Throughout, we compare the QBD-RAP scheme (from Chapter~\ref{sec: construction and modelling}), the discontinuous Galerkin method (as described in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013}, which we will refer to a just the \emph{uniformisation scheme}.
+%Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux.
+Recall that, due to their stochastic interpretation, the uniformisation and QBD-RAP schemes are positivity preserving.
 
-Where appropriate, we also compare the above methods with the discontinuous Galerkin method with a slope-limiter implemented, which avoids oscillations and negative solutions.
-However, recall that, in the context of approximating operators of fluid-fluid queues with these schemes, slope-limiting is not possible, and even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.
+Where possible, we implement the discontinuous Galerkin method with and without a slope-limiter implemented \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
+The slope limiter avoids oscillations and negative solutions.
+However, we note that in the context of approximating the operator \(\mathbb\Psi\) for a fluid-fluid queue, it is not obvious how one might apply the concept of slope-limiting, other than to post-process the solution with a limiter or filter.
+Even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.
+Moreover, there is a computational cost in applying a limiter.
+However, high-order accuracy of the DG scheme is maintained away from the discontinuities \citep{c99}, \citep[Section~5.6.2]{nodalDGBook}.
 
 The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased.
-For discontinuous Galerkin method, this means increasing the number of basis functions used to approximate the solution within each cell.
-For the QBD-RAP scheme, this means increasing the order of the CME distribution used to construct the scheme.
-For the uniformisation scheme, this means dividing each cell into smaller sub-cells over which we approximate the solution.
-As such, each approximation scheme leads to approximations which are matrices of the same size (although not necessarily the same number of non-zero components).
-For example, if we use 3 basis functions in the discontinuous Galerkin method, we approximate the solution by a quadratic on each cell; the comparable QBD-RAP scheme is to use a CME of order 3; and the comparable uniformisation scheme is to divide each cell into 3 sub-cells of width \(\Delta/3\).
+For simplicity, the number of basis functions is kept constant across all cells.
+We refer to the number of basis functions on a cell as the \emph{dimension}.
+For discontinuous Galerkin method, increasing the dimension of the scheme means increasing the number of polynomial basis functions used to approximate the solution within each cell.
+E.g.~if we use 3 basis functions in the discontinuous Galerkin method, we approximate the solution by a quadratic on each cell.
+For the QBD-RAP scheme, increasing the dimension of the scheme means increasing the order of the CME distribution used to construct the scheme.
+To make a comparable equivalent for the uniformisation scheme with a fixed cell size we divide each cell into smaller sub-cells over which we approximate the solution.
+I.e.~for a dimension \(p\) uniformisation scheme we divide each cell into \(p\) sub-cells.
+Equivalently, we may think of a dimension \(p\) uniformisation scheme as using \(p\) piecewise constant functions to approximate the solution on each cell.
+For all schemes (DG, QBD-RAP and uniformisation), if we construct and order \(p\) approximation, there are \(K\) cells, and \(N\) phases then the resulting approximation to the generator \(\mathbb B\) is a matrix of dimension \(pKN\).
+In this sense each approximation scheme leads to matrices of the same size (although not necessarily the same number of non-zero elements).
 
-To keep the content of this chapter contained, we do not investigate the performance of the schemes as the cell sizes are decreased, other than for the uniformisation scheme.
-As such, for each numerical experiment, we keep the cell size fixed.
-As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next and link all the cells together.
+To keep the content of this chapter contained, we do not investigate all aspects of the schemes.
+For each numerical experiment we keep the cell size fixed for the DG and QBD-RAP schemes.
+As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next \citep{nodalDGBook}.
 We investigate schemes with an upwind flux only.
-We do not investigate the stability of the schemes with respect to the step-size of the time-integration or time-integration scheme (where required).
-Instead, we fix the time-integration step size for each numerical experiment at a suitably small value so as to obey a certain stability criterion (a CFL-like condition, REFs, appendix?).
-For schemes which require us to integrate over time we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (see appendix and REF).
-This scheme claims to introduce no more oscillations into the solution as we integrate over time (REFs).
-Where a slope limiter is implemented, we always implement the \emph{Generalised MUSCL} limiter (REF and Appendix).
-We do not investigate filtering for the DG scheme (REFs).
+For schemes which require us to integrate over time we do not investigate the stability of the schemes with respect to the \(t\)-step-size of the time-integration or time-integration scheme itself (where required).
+Instead, we fix the time-integration step size for each numerical experiment at a suitably small value to obey a certain stability criterion (a CFL-like condition, \citep[Section~4.8]{nodalDGBook}).
+Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages \citep{sr2002}, \citep[Section~5.7]{nodalDGBook} (see also Appendix~\ref{sec: time integration}), which claims to introduce no more oscillations into the solution as we integrate over time.
+Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
+We do not investigate filtering for the DG scheme (see \citep[Section~5.6.1]{nodalDGBook} and references therein).
 
-The performance of the discontinuous Galerkin method has been well-studied (REFs), and it is well-known that the discontinuous Galerkin method performs remarkably well on problems with smooth solutions (REF).
+The performance of the discontinuous Galerkin method has been well-studied in some contexts \cite{c99}, \cite[Section~5.5]{nodalDGBook}, and it is well-known that the discontinuous Galerkin method performs remarkably well on problems with smooth solutions.
 Here, we mostly focus on investigating the numerical performance of the methods on problems with non-smooth solutions, the purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP method.
-In the stochastic modelling community it is very common to have problems with discontinuous solutions, such as a non-smooth initial conditions (REFs).
-Even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities into the problem.
-There is a very specific set of conditions on the initial density and point masses which must hold for the distribution of the fluid queue to remain smooth as it evolves over time (REFs).Limiting distributions of fluid queues are smooth, however.
-Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues.
+In the stochastic modelling community it is very common to have problems with discontinuities, such as a non-smooth initial conditions.
+Even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities or non-smooth behaviour into the problem (see, for example, Section~\ref{sec: transient approx}, below).
+A very specific set of conditions must hold for the initial density and point masses for the distribution to remain continuous as it evolves over time \citep{bo2014,bop2020}.
+Limiting distributions of fluid queues are smooth, however.
+Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues (see, for example, Section~\ref{sec: return approx}).
 
-As we shall see, if the solution is smooth, then the Galerkin method is highly-effective and displays rapid convergence to the solution as the number polynomial basis functions is increased.
-However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin method, leading to non-sensical solutions.
+As we shall see, if the solution is smooth, then the Galerkin method is highly-effective and displays rapid convergence to the solution as the number of polynomial basis functions is increased.
+However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin method, leading to nonsense solutions.
 In these cases, the QBD-RAP approximation performs relatively well compared to the other positivity preserving schemes investigated.
+
+The structure of this chapter is as follows.
+In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different methods and observe their performance at approximating the initial condition, which allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
+In Section~\ref{sec: wave num} we investigate a simple travelling wave problem with various initial conditions.
+For this problem the dynamics are deterministic, which allows us to instrument the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
+Next, we investigate a simple fluid queue with two phases.
+Section~\ref{sec:stat} investigates the ability of the methods to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the fluid queue for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the methods to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
+Section~\ref{sec: ffq num} approximates the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\) for two simple fluid-fluid queue models.
 
@@
 
-\section{Function approximation/reconstruction}
+\section{Function approximation/reconstruction}\label{sec: recon num}
 We start our investigation by looking at how well the methods perform at approximating an initial condition.
-For the discontinuous Galerkin method, this means projecting the solution on to the set of polynomial functions which define the scheme, then reconstructing the solution.
-For the spatially-coherent uniformisation scheme, this means looking at the sub-cell averages.
-For the QBD-RAP scheme, this means computing the initial vector for the approximation, then reconstructing the solutions as described in Section~\ref{sec: closing}.
-First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}) performs reasonably compared to the others, and we are able to theoretically prove that it converges.
-Therefore we use this reconstruction method throughout the rest of the chapter.
+By approximating the initial condition only, we aim to instrument the performance of the approximation schemes without any dynamics.
 
-For this section we consider approximating the initial condition over a single cell of width \(\Delta = 1\).
-To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a simple trapezoidal rule with 10,001 function evaluations.
-Similarly, we use 10,001 function evaluations to approximate \(L^2\) norms, and also as a finite set of points over which to compute the KS statistics.
+For the discontinuous Galerkin method, this we project the initial condition on to the set of polynomial functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstructing the solutions as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
+For the purposes of approximation and reconstruction for the QBD-RAP scheme we must orientate the initial condition in a positive or negative direction; here, we suppose that the initial conditions belong to a positive phase.
+First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}).
+From the investigation, we decide to use the closing vector (\ref{eqn:5462}) throughout the rest of the chapter.
 
-First we briefly investigate the possible choices of closing operators as introduced in Section~\ref{sec: closing}.
+For this investigation we consider approximating initial conditions over a single cell of width \(\Delta = 1\).
+To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a trapezoidal rule with 10,001 function evaluations.
+Similarly, we use 10,001 function evaluations to approximate \(L^p\) norms, and also as a finite set of points over which to compute the KS statistics.
+
+% First we briefly investigate the possible choices of closing operators as introduced in Section~\ref{sec: closing}.
 
@@
 Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme, though it clearly appears to converge.
-The third closing vector we investigate is that in Equation~(\ref{eqn:5462}).
+The third closing vector we investigate is that in Equation~(\ref{eqn:5462}) and we refer to this as the \emph{normalised closing operator}.
 
-\begin{example}First consider approximating the initial condition with density function \(1(x<0.5)\).
-In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation.
+\begin{example}First consider approximating the initial condition with density function \(2\times 1(x<0.5)\).
+In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation with the three different closing vectors.
 Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders.
-It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interferes with other errors in the reconstruction to cause the error to be lower at low orders for the unnormalised scheme.
+It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interferes with other errors to cause the error to be lower at low orders for the unnormalised scheme.
 Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection.
@@
 If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs}, then we see that the unnormalised reconstruction performs the poorest.
-Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests naive normalised and normalised reconstruction are very similar.
-\begin{figure}
+Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests the naive normalised and normalised reconstruction are very similar.
+\begin{figure}[h]
 	\centering
@@
 \end{figure}
+\exampleFloatBarrier
 \end{example}
@@
 \begin{example}Now consider approximating the initial density \(f(x)=1\).
-Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error, and \(L^2\) error between the true initial distribution and the reconstructions, we now see that the normalised reconstructions outperform the unnormalised reconstruction.
+Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error and \(L^2\) error of the PDF we now see that the normalised reconstructions outperform the unnormalised reconstruction.
 This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution.
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 
-Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised and normalised closing operators for order 1, 3, 5 and 7, as well as the true PDF.
-Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots and for all orders.
+Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised and normalised closing operators for dimension 1, 3, 5 and 7, as well as the true PDF.
+Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots.
 This feature is due to a significant amount of mass being lost due to the truncation.
-In comparison the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator.
-In the reconstruction we also need to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}.
+In comparison, the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator.
+The `folding' in the normalised closing operator manifest itself as more mass at the left of these plots compared to the unnormalised reconstructions.
+% Recall that, as discussed in Section~\ref{}, for the reconstruction we must to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}.
 Here we suppose that the phase is positive, so the `folding' around \(\Delta\) in the normalised closing operators appears appears at the left of the plots.
@@
 Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval.
-This is a common feature, and drawback, of the QBD-RAP method.
 Perhaps there is a different reconstruction method or yet another closing operator which could alleviate this issue.
-\begin{figure}
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=\textwidth]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=1\) (orange dashed line).}
+	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
+	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various dimensions, and the true PDF which is \(f(x)=1\) (orange dashed line).}
 	\label{fig: pdf reconstructed}
 \end{figure} 
+\exampleFloatBarrier
 \end{example}
@@
 \begin{example}Next we consider approximating the initial distribution with density \(-6x^2+6x\).
-Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the order of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
-However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, we see that with this metric, then unnormalised closing operator performs better than the two normalised ones for orders 5 and above.
-\begin{figure}
+Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the dimension of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
+However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, then the unnormalised closing operator performs better than the two normalised ones for dimension 5 and above.
+\begin{figure}[h]
 	\centering
@@
 
-The fact that the unnormalised closing operator outperforms the two normalised ones can be explain, once again, by the fact that, for the normalised operators we have `folded' the ME density function back on itself.
+The fact that the unnormalised closing operator outperforms the two normalised ones can be explained by the `folding' of the normalised operators around \(\Delta\).
 In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\).
-In Figure~\ref{fig: pdf reconstructed quadratic} we observe that, at the left of the plot, the normalised reconstructions over estimate the density function in this region, where as the unnormalised reconstruction looks to be doing better.
-Recall that the `folding' in the normalised closing operator manifest itself as extra mass at the left of these plots compared to the unnormalised reconstructions.
-\begin{figure}
+In Figure~\ref{fig: pdf reconstructed quadratic} at the left of the plots, we observe that the normalised reconstructions over-estimate the density function, whereas the unnormalised reconstruction looks to be doing better.
+The `folding' in the normalised closing operator manifest itself as more mass at the left of these plots compared to the unnormalised reconstructions.
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=\textwidth]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
+	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
 	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=-6x^2+6x\) (orange dashed line).}
@@
 \end{figure} 
+\exampleFloatBarrier
 \end{example}
@@
 Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case.
-Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that the normalised reconstructions perform relatively well compared to the unnormalised reconstruction as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF).
+Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that this is the case; the normalised reconstructions perform relatively well compared to the unnormalised reconstruction, as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF).
 If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation.
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 	\caption{KS error between the true CDF, \((1-e^{-3x})/(1-e^{-3})\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
+The naive normalised (orange) and normalised closing vectors are coincident.}
 	\label{fig: fun 7 ks error qbdrap closing vecs}
 \end{figure}
+\exampleFloatBarrier
 \end{example}
@@
 
-\subsection{Comparison of methods}
+\FloatBarrier
+\subsection{Comparison of methods}\label{sec: comp}
 Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.
@@
 \begin{example}First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5.
-This distribution does not have a PDF so we can compare the CDFs only.
+This distribution does not have a PDF, so we compare the CDFs only.
 In Figure~\ref{fig: fun 1 comp} we plot the KS metric (left) and \(L^1\) metric (right) between the true CDF and the reconstructed approximations.
 Observing the KS metric, it appears that none of the methods converge and the KS error sits around 0.5.
-This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except at points of discontinuity.
-None of the methods appear to converge at this point.
+This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except, perhaps, at points of discontinuity.
+None of the methods appear to converge at the discontinuity at \(x=0.5\).
 Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the methods appear to show the convergent behaviour we expect.
@@
 Perhaps it is no surprise that the uniformisation method performs best.
-In the uniformisation method as we increase the order, we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the function.
+In the uniformisation method as we increase the order we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the initial distribution.
 As such, the uniformisation method can produce a piecewise continuous, linear approximation to the CDF.
 In contrast, both the DG and QBD-RAP methods result in a smooth approximation to the CDF.
-\begin{figure}
+Given the initial distribution is far from smooth (it's a point mass), then we might expect that the uniformisation method will perform relatively well.
+\begin{figure}[h]
 	\centering
@@
 	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(1(x\geq 0.5)\), and the approximations (left) and \(L^1\) error between the true CDF and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\caption{KS error between the true CDF, \(1(x\geq 0.5)\), and the approximations (left) and \(L^1\) error between the true CDF and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
 	\label{fig: fun 1 comp} 
@@
 
-In Figure~\ref{fig: pdf comp fun 1} we plot the approximated CDFs for the DG, uniformisation and QBD-RAP methods alongside the true CDF.
+In Figure~\ref{fig: pdf comp fun 1} we plot approximated CDFs from the DG, uniformisation and QBD-RAP methods alongside the true CDF.
 The DG method displays undesirable features for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1.
 On the other hand, although the QBD-RAP method converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1.
-\begin{figure}
+The uniformisation scheme approximates the CDF exceptionally well.
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=\textwidth]{chapter6/figs/comp/fun1/cdfs_formatted.pdf}
-	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (green dot-dashed line) methods.
+	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/comp/fun1/cdfs_formatted.pdf}
+	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) methods.
 The true distribution function is \(1(x\geq 0.5)\) (grey dotted line).}
@@
 Figure~\ref{fig: fun 2 comp} suggests that all methods converge at a similar rate for this problem.
-Once again, the QBD-RAP method performs worst, the uniformisation method second, and the DG method the best.
-However, once again the DG method exhibits undesirable properties -- the approximation to the CDF is, at some points above 1, is not monotonic, though these violations are not as severe in this case as they were for the case above.
-\begin{figure}
+Here, the QBD-RAP method performs worst, the uniformisation method second, and the DG method the best.
+However, once again the DG method exhibits undesirable properties (plots not shown) -- the approximation to the CDF is at some points above 1 and is not monotonic (although these violations do not appear to be as severe in this case as they were for the example above).
+\begin{figure}[h]
 	\centering
@@
 	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(x\geq 0.5)\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(x\geq 0.5)\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
 	\label{fig: fun 2 comp} 
 \end{figure}
+\exampleFloatBarrier
 \end{example}
@@
 \begin{example}So far we have considered two problems which exhibit discontinuities.
-At the other extreme we now consider an initial distribution with density \(-6z^2+6x\).
-In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDF.
-Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG method can theoretically approximate the initial condition exactly.
-This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from 1-3, then plateaus.
+At the other extreme we now consider an initial distribution with density \(-6x^2+6x\).
+In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDFs.
+Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG method can approximate the initial condition exactly.
+This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from dimension 1 to 3, then plateaus.
 Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, and due to machine arithmetic, the errors for the DG scheme are not 0.
-Regarding the other two methods, they too appear to be convergent at approximately the same rate, with the uniformisation method perform better for the KS error, but very similarly in terms of the \(L^2\) error.
-\begin{figure}
+Regarding the other two methods, they too appear to be convergent at approximately the same rate, with the uniformisation method performing better for the KS error, but very similarly in terms of the \(L^2\) error.
+\begin{figure}[h]
 	\centering
@@
 	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(-6z^2+6x\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(-6z^2+6x\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
 	\label{fig: fun 6 comp} 
 \end{figure}
+\exampleFloatBarrier
 \end{example}
@@
 Figure~\ref{fig: fun 9 comp} shows the errors.
-Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs tell a similar story.
-For sufficiently high order, the DG method approximates the inital condition very well.
+Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs (right) tell a similar story; for sufficiently high order, the DG method approximates the initial condition very well.
 The uniformisation method performs second best, while the QBD-RAP method is slowest to converge.
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(cos(4\pi(x+0.5))+1\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(cos(4\pi(x+0.5))+1\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
 	\label{fig: fun 9 comp} 
 \end{figure}
+\exampleFloatBarrier
 \end{example}
+
+In conclusion, we have observed that the DG method performs well with respect to the error metrics considered, but can display oscillations and result in CDFs which are not non-decreasing.
+On the other hand, the uniformisation and QBD-RAP schemes result in non-decreasing approximations to the CDFs, but can converge slower with respect to the error metrics considered, with the QBD-RAP scheme often converging the slowest.
 \FloatBarrier
 
-\section{Travelling wave}
-Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with varying initial conditions.
+\section{Travelling wave}\label{sec: wave num}
+Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with various initial conditions.
 Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\).
@@
 \[f(x,t) = f(x-t,0)\]
-so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).
+so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right.
+We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).
+This example is convenient as it has a known solution and no stochastic dynamics, hence we can instrument the ability of the schemes to approximate the flow of mass, without any stochastic dynamics.
 
@@
 We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric.
-We use the SSPRK4 method to integrate over time using a step size of \(0.005\), and we evolve the system until time \(t=4\).
-For the DG method we also implement the Generalise MUSCL (REF, appendix?) slope limiter to stabilise the integration.
+We use the SSPRK4 method to integrate over time with a \(t\)-step size of \(0.005\), and we evolve the system until time \(t=4\).\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} For the DG method we also implement the Generalised MUSCL slope limiter to stabilise the integration \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
 
-\begin{example}First consider the initial condition with PDF \(1(x<1)\).
+To investigate the performance of the methods without the need to reconstruct the function within each cell we use the cell-wise error metric obtained by computing 
+\begin{align}
+	&\sum_{\ell=1}^{10} \left|\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,\ell,1)\right|\nonumber 
+	\\&+\left|\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,11,1)\right| \label{eqn: cell errors}
+\end{align}
+where 
+\begin{align}
+	\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs}
+\end{align}
+for \(\ell=1,\dots,K\), and 
+\begin{align}
+	\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs boundary}
+\end{align}
+is the mass at the boundary and where \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,11,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).
+
+\begin{example}\label{ex: wave 4}
+First consider the initial condition with PDF \(1(x<1)\).
 The level of the fluid queue is uniformly distributed over the first cell.
@@
 Thus, in this case, there is no discretisation error in constructing the initial condition for the DG and uniformisation methods.
-Furthermore, at time \(t=4\), the solution in \(1(x\in[4,5)\), and the projections related to the DG and uniformisation methods can represent this solution exactly too.
+At time \(t=4\), the solution is \(f(x,4)=1(x\in[4,5))\).
+The projections related to the DG and uniformisation methods can represent this solution exactly too, hence we might expect these methods to work well here.
+
+We plot the cell-wise error metric in Figure~\ref{fig: fun 4 wave cp} and observe that the DG method appears to converge fastest, followed by the QBD-RAP method, then the uniformisation method, and the DG scheme with a slope limiter does not appear to converge.
+The DG scheme with a slope limiter has detected oscillations in the approximate solution and reduced the scheme to linear where oscillations are detected, thereby decreasing the accuracy of the scheme to linear where oscillations are present.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/L1_cell_probs.pdf}
+	\caption{Cell-wise error defined in (\ref{eqn: cell errors}) for the travelling wave in Example~\ref{ex: wave 4}.
+Plotted are the cell-wise errors for the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line) versus the dimension of the approximation.}
+	\label{fig: fun 4 wave cp} 
+\end{figure}
 
 Figure~\ref{fig: fun 4 wave} plots the KS error between the approximated and true CDFs and \(L^2\) error between the approximated and true PDFs.
-The DG method with the MUSCL limiter does not appear to converge for this problem.
-This is caused by the limiter reducing the method to linear in when oscilations are deteced, so increasing the order does not increase the accuracy of the method.
-The other two positivity preserving methods, the uniformisation method and QBD-RAP methods, both appear to be converging, with the QBD-RAP method converging faster.
-With these error metrics, the DG method also appears to converge.
-However, if we observe the solutions resulting from the DG method (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscilations and negative values.
+Comparing Figure~\ref{fig: fun 4 wave cp} with Figure~\ref{fig: fun 4 wave}, each method seems to show a similar rate of decay for error metrics with and without reconstruction.
 
-Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG with MUSCL limiter and QBD-RAP methods.
-In the first row we observe that the DG method (sans limiter) produces an approximation to the PDF with negative values for orders greater than 1.
-In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter the approximation does not change significantly after order 3.
-This is due the the fact that the DG method is at best linear.
+Returning to Figure~\ref{fig: fun 4 wave}, the DG method with the MUSCL limiter does not appear to converge with these error metrics for this problem (due to the limiter reducing the approximation to linear near the discontinuity), while the other two positivity preserving methods, the uniformisation and QBD-RAP schemes, both appear to be converging with the QBD-RAP method converging faster.
+With these error metrics, the DG method appears to be converging fastest.
+However, if we observe the approximations resulting from the DG method (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscillations and negative values.
+
+Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG with MUSCL limiter and QBD-RAP methods for various dimension schemes.
+In the first row we observe that the DG method (sans limiter) produces an approximation to the PDF with negative values when the dimension is greater than 1.
+In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter, the DG approximation does not change significantly after order 3.
+This is due the fact that the DG method is at best linear in the presence of oscillations.
 In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation method, and in the last row is the solution approximated using the QBD-RAP method.
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(4\leq x < 5)\), and the approximations for the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\caption{KS error (left) and \(L^2\) error of the PDFs for the travelling eave problem in Example~\ref{ex: wave 4} using the approximations from the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
 	\label{fig: fun 4 wave} 
 \end{figure}
-\begin{figure}
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=\textwidth]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG with MUSCL limiter (third row) and QBD-RAP (bottom row) methods, for order 1, 3, 5, and 7 (columns).
+	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
+	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG with MUSCL limiter (third row) and QBD-RAP (bottom row) methods, for order 1, 3, 5, and 7 (columns) for the travelling eave problem in Example~\ref{ex: wave 4}.
 The true density function is \(1(4\leq x<5)\).}
@@
 This is a particularly interesting example.
-It shows that, for the DG method, even though there is no discretisation error for the initial condition, we have use a strong stability preserving time-intergration method, and the projection off which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as in the top row of Figure~\ref{fig: pdf wave fun 4}.
+It shows that, for the DG method, even though there is no discretisation error for the initial condition, we use a strong stability preserving time-integration method, and the projection off which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as in the top row of Figure~\ref{fig: pdf wave fun 4}.
+\exampleFloatBarrier
 \end{example}
 
-\begin{example}Another interesting example occurs with the initial condition with CDF \(1(z\geq 0.5)\), i.e.~a point mass at 0.5.
+\begin{example}\label{ex: wave 2}
+Another interesting example occurs with the initial condition with CDF \(1(x\geq 0.5)\), i.e.~a point mass at 0.5.
 The exact solution at time \(t=4\) is therefore a point mass at \(4.5\).
@@
 Moreover, when we analysed reconstruction of this initial condition we saw that using the KS metric may be uninformative.
-Instead, for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs).
-We also compare the probabilities
-\begin{align}
-	\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs}
-\end{align}
-for \(\ell=1,\dots,K\), and also the boundary mass
-\begin{align}
-	\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs boundary}
-\end{align}
-by computing 
-\begin{align}
-	&\sum_{\ell=1}^{10} \left|\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,\ell,1)\right|
-	\\&+\left|\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)-p(4,10,1)\right| \label{eqn: cell errors}
-\end{align}
-where \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,10,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).
+Instead, for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs) and the cell-wise error from (\ref{eqn: cell probs})-(\ref{eqn: cell errors}).
 
-Figure~\ref{fig: fun 2 wave} (left) plots the \(L^1\) error between the true and approximated CDFs (left).
-The \(L^1\) metric tells a similar story to the previous analysis: the DG method with the limiter does not converge as order increases, the other two positivity preserving methods, the uniformisation and QBD-RAP methods, appear to converge, with the QBD-RAP method converging faster, while the DG method appears to converge the fastest.
-However, if we plot the approximations to the CDFs from the DG method (not shown) we would once again see an oscilating non-monotonic approximation.
+Figure~\ref{fig: fun 2 wave} (left) plots the \(L^1\) error between the true and approximated CDFs.
+The \(L^1\) metric tells a similar story to the previous analysis: the DG method with the limiter does not converge as order increases, the other two positivity preserving methods (the uniformisation and QBD-RAP methods) appear to converge, with the QBD-RAP method converging faster, while the DG method appears to converge the fastest.
+However, if we plot the approximations to the CDFs from the DG method (not shown) we would once again see an oscillating, non-monotonic function.
 Another interesting observation is to compare the performance of the uniformisation and QBD-RAP methods with respect to the \(L^1\) metric on the CDFs in Figure~\ref{fig: fun 2 wave} (left) to that in Figure~\ref{fig: fun 1 comp} (right).
-Recall that with Figure~\ref{fig: fun 1 comp} we investigate the ability of the different approximation methods to represent the initial distribution with CDF \(1(x\geq 1)\), which is the initial condition of the problem we are consdiering here.
-In Figure~\ref{fig: fun 1 comp} (right) the uniformisation method out-performed the QBD-RAP method at reconstructing the inital condition.
+Recall, in Figure~\ref{fig: fun 1 comp} we investigated the ability of the different approximation methods to represent the initial distribution with CDF \(1(x\geq 0.5)\), which is the initial condition of the problem we are considering here.
+In Figure~\ref{fig: fun 1 comp} (right) the uniformisation method out-performed the QBD-RAP method at reconstructing the initial condition.
 However, in Figure~\ref{fig: fun 2 wave} (left) we see that the QBD-RAP method out-performs the uniformisation method.
-This suggest that the QBD-RAP method is better able to resolve movement of mass across the domain when integrating over time, compared to the uniformisation method.
+This suggests that the QBD-RAP method is better able to resolve movement of mass across the domain when integrating over time, compared to the uniformisation method.
 
 Figure~\ref{fig: fun 2 wave} (right) plots the cell-based error metric in (\ref{eqn: cell errors}).
-Once again, the DG method with MUSCL limiter does not appear to converge as we expect.
+Once again, the DG method with MUSCL limiter does not appear to converge.
 The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP method converging faster.
 Most interestingly though is the DG method.
-It is unclear as to whether this method will converge in this case as the errors in the plot are not obviously decreasing.
-Moreover, with this metric, the QBD-RAP method out-performs the DG method.
-Furthermore, the DG method can result in negative estimates of the cell probabilities in (\ref{eqn: cell probs}) and (\ref{eqn: cell probs boundary})
-\begin{figure}
+It is unclear whether this method will converge in this case as the errors in the plot are not obviously decreasing.
+Moreover, with this metric, the QBD-RAP method out-performs the DG method when the dimension is large enough.
+Furthermore, the DG method can result in negative estimates of the cell probabilities in (\ref{eqn: cell probs}) and (\ref{eqn: cell probs boundary}) (plots not shown).
+\begin{figure}[h]
 	\centering
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/L1_cell_probs.pdf} 
-	\caption{\(L^1\) error between the true CDF, \(1(4.5\leq x)\), and the approximations (left) and the error metric based on the cell masses as described in (\ref{eqn: cell errors}) for the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+	\caption{\(L^1\) error between the true CDF and the approximations (left) and the cell-wise error metric in (\ref{eqn: cell errors}) for the travelling wave problem in Example~\ref{ex: wave 2} where approximation were constructed via the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
 	\label{fig: fun 2 wave} 
 \end{figure}
+\exampleFloatBarrier
 \end{example}
+
+\begin{example}\label{ex: wave 6}
+	Now consider an initial distribution which is truncated Gaussian with mean parameter 2.5 and standard deviation parameter 0.5, and truncated at the boundaries, 0 and 10;
+	\begin{align}
+		\mu([0,x)) := \mathbb P(X(0)\leq x, \varphi(0)=1)=\cfrac{\Phi((x-2.5)/0.5)1(0 \leq x < 10.0)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10\leq x), \label{eqn: kdjfksdf}
+	\end{align}
+	where \(\Phi(x)\) is the CDF of the standard normal distribution.
+
+	At time \(t=4\) the distribution is truncated Gaussian with mean parameter 6.5, standard deviation parameter 0.5, and is truncated below at 4.0 and above at 10, and there is also a small mass at the upper boundary;
+	\begin{align}
+		&\mathbb P(X(4)\leq x, \varphi(4)=1\mid X(0)\sim\mu, \varphi(0)=1)  
+		=\cfrac{\Phi((x-6.5)/0.5)1(4.0\leq x)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10 \leq x).\label{eqn: bbbbbaabbaa}%\cfrac{\Phi(7.5/0.5)-\Phi(3.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}.
+	\end{align}
+	At time \(t=4\) the mass at the boundary is approximately \(1.28\times 10^{-12}\) and there is a small discontinuity at \(x=4\) where the CDF jumps from \(0\) to \(\sim 2.87\times 10^{-7}\), due to the truncation of the initial distribution at \(0\).
+
+	Since the discontinuity at \(x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG method will perform well for this example.
+Figures~\ref{fig: fun 6 wave cp} and~\ref{fig: fun 6 wave} confirms that this is indeed the case.
+For all three metrics the error obtained by the DG method rapidly decreases to a point where it is swamped by other numerical errors.
+This is characteristic of the DG method for the smooth problems we investigate throughout this chapter.
+For low order DG schemes there are regions where the approximated PDF is negative, however, as the order of the DG scheme increases, these regions disappear.
+Interestingly, even though this is a relatively smooth problem, the DG method with MUSCL limiter does not perform well.
+It must be that the initial and transient distributions are `sufficiently pointy' that oscillations in the numerical solutions occur and the limiter reduces the order of the scheme to linear.
+Both the uniformisation and QBD-RAP schemes perform similarly to previous examples.
+Comparing the cell-wise error plots in Figure~\ref{fig: fun 6 wave cp} with the higher-resolution error metrics in Figure~\ref{fig: fun 6 wave}, all the approximation schemes show a similar rate of decay of errors as a function of the dimension of the approximation.
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/L1_cell_probs.pdf}
+		\caption{Cell-wise error defined in (\ref{eqn: cell errors}) for the travelling wave problem in Example~\ref{ex: wave 6}.
+Plotted are the cell-wise errors for the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line) versus the dimension of the approximation.}
+		\label{fig: fun 6 wave cp} 
+	\end{figure}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error for the PDFs for Example~\ref{ex: wave 6} where approximations were constructed via the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+		\label{fig: fun 6 wave} 
+	\end{figure}
+\exampleFloatBarrier
+\end{example}
+
+\begin{example}\label{ex: wave 1}
+	We now want to look at how the methods might handle a mass at the boundary.
+We introduce an ephemeral second phase into the model with phase transition rate \(T_{22}=-1\), and fluid rate \(c_2=0\).
+The generator is therefore
+	\[T=\left[\begin{array}{cc} 0 & 0 \\ 1 & -1 \end{array}\right].\]
+
+	We suppose that the initial condition is a point mass at the boundary in phase \(2\), i.e. \[\mathbb P(X(0)\leq x, \varphi(0)=2)=1(0\leq x).\] 
+	With this initial condition, the transient distribution at time \(t=4\) is 
+	\begin{align}
+		&\mathbb P(X(4)\leq x,\varphi(4)=1 \mid X(0)=0,\varphi(0)=2) \nonumber 
+		\\&= e^{T_{22}4}\left(e^{-T_{22}x}-1\right)1(4< x) + (1-e^{T_{22}4})1(4\geq x) \nonumber
+		\\&= e^{-4}\left(e^{4x}-1\right)1(4< x) + (1-e^{-4})1(4\geq x) \label{eqn: asjda}
+	\end{align}
+	and 
+	\begin{align}
+		\mathbb P(X(4)\leq x,\varphi(4)=2 \mid X(0)=0,\varphi(0)=2) = e^{-4}1(0\leq x).
+	\end{align}
+	The PDF at \(t=4\) is discontinuous at \(x=4\), which is at the edge of a cell. For this problem, all methods can represent the initial condition exactly.
+
+	Figure~\ref{fig: fun 1 wave cp} plots the cell-wise error metric and Figure~\ref{fig: fun 1 wave} plots the KS error (right) and the \(L^2\) error between of the PDFs.
+Once again, the cell-wise error metrics in Figure~\ref{fig: fun 1 wave cp} show similar characteristics to the finer-resolution error metrics in Figure~\ref{fig: fun 1 wave}.
+Observing Figures~\ref{fig: fun 1 wave cp} and \ref{fig: fun 1 wave}, the DG method with MUSCL limiter does not perform well, which might be expected due to the discontinuity in the transient distribution at \(x=4\) in Phase~1.
+The uniformisation and QBD-RAP methods appear to converge, with the QBD-RAP method converging faster.
+The DG method converges fastest, however, produces approximations with negative and oscillatory solutions as we might expect given the discontinuity.
+A selection of approximations to the transient PDF are shown in Figure~\ref{fig: pdf wave fun 1}.
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/L1_cell_probs.pdf} 
+		\caption{Cell-wise error for the travelling wave problem in Example~\ref{ex: wave 1}.
+Plotted are the cell-wise errors for the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line) versus the dimension of the approximation.}
+		\label{fig: fun 1 wave cp} 
+	\end{figure}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error of the PDFs for the travelling wave problem in Example~\ref{ex: wave 1} where approximations were constructed via the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+		\label{fig: fun 1 wave} 
+	\end{figure}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun1/pdf_formatted.pdf}
+		\caption{Approximate PDFs for Example~\ref{ex: wave 1} using the DG (top row), uniformisation (second row), DG with MUSCL limiter (third row) and QBD-RAP (bottom row) methods, with dimensions 1, 3, 5, and 7 (columns).
+The true density function is \(e^{-4}e^{x}1(x<4)\).}
+		\label{fig: pdf wave fun 1}
+	\end{figure} 
+	\exampleFloatBarrier
+\end{example}
+
+\FloatBarrier
+\section{Stationary distributions} \label{sec:stat}
+We briefly turn our attention to stationary distributions of fluid queues.
+Since the stationary distribution is smooth we expect that the DG method will work well here.
+The stationary distribution is convenient as we can evaluate them analytically \cite{s2017} and do not require us to approximate initial conditions or integrate over time.
+Hence, analysing the performance of the approximation methods with the stationary distribution allows us to instrument the ability of the methods to capture the stochastic dynamics, without numerical integration in time.
+
+Here we analyse a simple model which is based on Example~2 in \citep{bean2009}, except here we add a lower boundary to the model (no lower boundary is specified in Example~2 of \citep{bean2009} as it is inconsequential to their analysis).
+
+% \begin{model}\label{model: simple}
+% 	Consider a fluid queue where the driving process is a CTMC with state space \(\calS=\{1,2,3,4\}\), generator 
+% 	\[\bs T = \left[\begin{array}{cccc}
+% 		-1.1 & 1.1 & 0 & 0 \\
+% 		1 & -1 & 0 & 0 \\ 
+% 		0.01 & 0 & -0.01 & 0 \\
+% 		0 & 0.01 & 0 & -0.01 
+% 	\end{array}\right],\]
+% 	and there are associated rates \(c_1=1, c_2 = -1, c_3=0, c_4=0\) and boundaries at \(x=0\) and \(x=10\). We specify two types of behaviour at the boundary.
+% 	\begin{enumerate}[a)]
+% 		\item (Absorbing model) Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
+% 		\[\vligne{p_{21} & p_{22} & p_{23} & p_{24}} = \vligne{0 & 0 & 1 & 0}.\]
+% 		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
+% 		\[\vligne{p_{11} & p_{12} & p_{13} & p_{14}} = \vligne{0 & 0 & 0 & 1}.\]
+% 		\item (Reflecting model) Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
+% 		\[\vligne{p_{21} & p_{22} & p_{23} & p_{24}} = \vligne{1 & 0 & 0 & 0}.\]
+% 		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
+% 		\[\vligne{p_{11} & p_{12} & p_{13} & p_{14}} = \vligne{0 & 1 & 0 & 0}.\]
+% 	\end{enumerate}
+% \end{model}
+% Phases~3 and 4 are included only to capture behaviour at the boundary.
+
+% For Model~\ref{model: simple}~a), upon hitting the boundary, the process stays at the boundary for an exponentially distributed amount of time with mean 100.
+For Model~\ref{model: simple}~b), upon hitting the boundary, the process is immediately reflected.
+In Model~\ref{model: simple}~b) phases 3 and 4 are ephemeral and are included only for notational convenience.
+
+\begin{model}\label{model: simple}
+	Consider a fluid queue where the driving process is a CTMC with state space \(\calS=\{1,2,3,4\}\), generator 
+	\[\bs T = \left[\begin{array}{cccc}
+		-1.1 & 1.1 \\
+		1 & -1 \\ 
+	\end{array}\right],\]
+	and there are associated rates \(c_1=1, c_2 = -1,\) and boundaries at \(x=0\) and \(x=10\). We specify two types of behaviour at the boundary.
+
+	Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
+		\[\vligne{p_{21} & p_{22} } = \vligne{1 & 0}.\]
+		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
+		\[\vligne{p_{11} & p_{12} } = \vligne{0 & 1}.\]
+	Thus, upon hitting the boundary, the fluid queue immediately transitions to the other phase and is reflected.
+\end{model}
+
+The model was discretised using the DG, uniformisation and QBD-RAP methods using ten cells of width \(1\).
+We compute the coefficients for the stationary distribution in the following way.
+Suppose that a given discretisation method results in an approximation to the generator of the fluid queue as a matrix \(\bs B\).
+Then the stationary coefficients are found by solving
+\begin{align}
+	\bs b \bs B = 0,\\
+	\mbox{such that } \bs b\bs 1=1,
+\end{align}
+for the coefficients \(\bs b\).
+
+% \paragraph{Model~\ref{model: simple}~a) Absorbing model}
+% Figure~\ref{fig: absorbing stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
+Clearly the DG method is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
+The QBD-RAP method and uniformisation method both appear to be converging, with the errors for the QBD-RAP method decreasing faster.
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/ks_error_formatted.pdf}%
+% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/l2_pdf_error_formatted.pdf}
+% 	\caption{KS error between the true stationary CDF  and the approximations (left) and \(L^2\) error between the true stationary PDF and the approximations for Model~\ref{model: simple}~a) using the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+% 	\label{fig: absorbing stationary} 
+% \end{figure}
+
+\paragraph{Model~\ref{model: simple}.}
+Figure~\ref{fig: reflecting stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
+Clearly the DG method is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
+The QBD-RAP method and uniformisation method both appear to be converging, with the errors for the QBD-RAP method decreasing faster.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/stationary_distribution/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/stationary_distribution/l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDFs for Model~\ref{model: simple} using the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+	\label{fig: reflecting stationary} 
+\end{figure}
+
+\FloatBarrier
+\section{Transient distributions}\label{sec: transient approx}
+Once again we consider Model~\ref{model: simple} and use the same spatial discretisation as described in Section~\ref{sec:stat} (ten cells with width \(\Delta=1\)).
+Two initial conditions are considered, a point mass at 0 in phase 1, and the initial distribution with PDF
+\begin{align}
+	\cfrac{1}{2}e^{-x}/(1-e^{-10})\label{eqn: exp init cond}
+\end{align}
+in phases 1 and 2, with no mass at the boundaries.
+We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{Once again, the \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We apply the DG method with, and without, the Generalised MUSCL slope limiter.
+
+To obtain a \emph{ground truth} 5,000,000 realisations of the fluid queue were simulated until \(t=2\), then the empirical CDF and the masses within each cell and at the boundaries were computed from the simulations.
+We do not attempt to numerically approximate the true PDF via simulation.
+We then compute the KS and \(L^1\) error metrics between the approximated CDF and simulated CDF, as well as the cell-wise error metric, given as follows.
+\begin{align}
+	&\sum_{j\in\{1,2\}}\sum_{\ell=1}^{10} \left|\mathbb P_{sim}(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0)=0.5,\varphi(0)=1)-p(2,\ell,j)\right|\nonumber 
+	\\&+\left|\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))-p(2,11,4)\right|\nonumber 
+	\\&+\left|\mathbb P_{sim}X(2)\in\{0\}, \varphi(2)=3\mid X(0),\varphi(0))-p(2,0,3)\right| \label{eqn: cell errors 2}
+\end{align}
+where \(p(2,\ell,j)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0),\varphi(0))\), \(p(2,11,4)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))\), and \(p(2,0,3)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=3\mid X(0),\varphi(0))\), and \(\mathbb P_{sim}\) denotes an empirical probability evaluated using the simulations of the fluid queue.
+
+To account for possible Monte-Carlo error, we used a bootstrap with 1,000 bootstrap samples.
+We sample 5,000,000 realisations of the fluid queue with replacement from the original 5,000,000 samples, then compute error metrics with the resampled data.
+We resample 1,000 times.
+Via the bootstrap, we report the \(5\)th and \(95\)th percentile of the sampling distribution of the errors.
+
+To evaluate error metrics, we use a grid of 10,001 evenly spaced points for each phase.
+
+To approximate the point mass initial condition we compute the initial coefficients for each scheme exactly.
+For the exponential initial condition in (\ref{eqn: exp init cond}) we compute the initial coefficients via Gauss-Lobatto quadrature for the DG method, by using the mid-point rule for the uniformisation method, and by using a trapezoidal rule with 2,001 points on each cell for the QBD-RAP method.
+
+\paragraph{Model~\ref{model: simple} with exponential initial condition}
+Figure~\ref{fig: reflecting transient exp} shows the error metrics for the four different spatial discretisation schemes (DG, DG with MUSCL limiter, uniformisation and QBD-RAP).
+For both error metrics the DG scheme converges rapidly until computational errors become significant.
+The uniformisation and QBD-RAP schemes converge at a slower rate than the DG method, with the QBD-RAP scheme converging faster than the uniformisation method.
+The DG scheme with MUSCL limiter does not appear to be converging, which suggests there is at least one iteration during the numerical integration over time at which the numerical solution displays oscillations.
+However, when the DG approximations for the distribution at time \(t=2\) are plotted, they do not display oscillations (not shown).
+This suggests that the oscillations which might occur with the DG scheme must be transient, and have dissipated by time \(t=2\).
+The oscillations which the limiter detects in the DG scheme are likely from the reflecting boundary at \(x=0\).
+In the early stages of the evolution of the model the majority of the density in Phase~2 near zero will move to the left at rate 1 and hit the boundary at \(x=0\).
+Upon hitting the boundary, the density is reflected into Phase~1.
+Meanwhile, the initial density in Phase~1 moves to the right at rate 1 and the density reflected at the boundary from Phase~2 fills in to the region between \(x=0\) and \(x=t\) in Phase~1.
+Intuitively, in the early evolution of the model, there will be a sharp peak in the transient density at \(x=t\) -- I suspect that the transient density is continuous at this point, but not differentiable.
+%
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: reflecting transient exp} 
+\end{figure}
+
+Figure~\ref{fig: reflecting transient exp cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition.
+Since the cell-wise errors do not require us to reconstruct the value of the solution within each cell, then this metric allows us to observe the error characteristics of the methods without reconstruction.
+Figure~\ref{fig: reflecting transient exp cp} shows similar convergence characteristics to the other error metrics in Figure~\ref{fig: reflecting transient exp}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/L1_cell_probs_error_formatted.pdf}
+	\caption{Cell-wise error metric from (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: reflecting transient exp cp} 
+\end{figure}
+\exampleFloatBarrier
+\paragraph{Model~\ref{model: simple} with a point-mass initial condition}
+Figure~\ref{fig: reflecting transient pm} shows the error metrics for the four different spatial discretisation schemes (DG, DG with MUSCL limiter, uniformisation and QBD-RAP).
+Comparing the error metrics in Figure~\ref{fig: reflecting transient pm} for the point mass initial condition with the ones in Figure~\ref{fig: reflecting transient exp} for the exponential initial condition all schemes perform worse for the point mass initial condition.
+Regarding comparative rates of convergence for this problem, the DG scheme converges fastest, followed by the QBD-RAP scheme, then the uniformisation scheme, while the DG scheme with MUSCL limiter does not appear to converge due to the limiter reducing the scheme to a linear approximation around the discontinuity.
+%
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) errors between of the CDF (right) at time \(t=2\) for Model~\ref{model: simple} with the point-mass initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: reflecting transient pm} 
+\end{figure}
+
+Figure~\ref{fig: reflecting transient pm cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the point mass initial condition.
+Figure~\ref{fig: reflecting transient pm cp} suggests that approximating the cell-wise error seems to be a difficult problem.
+This is likely caused by the discontinuity at \(x=2\) in phase \(1\), which lies exactly on a cell boundary.
+To investigate how the position of the discontinuity might affect the error metrics we evolved the model to time \(t=2.1\) and computed the KS error, \(L^1\) error between the CDFs and the cell-wise error.
+Once again, we use simulation and bootstrapping to approximate the \emph{ground truth}.
+The plots (not shown) of KS error and \(L^1\) error between the CDFs for the model at times \(t=2\) and \(t=2.1\) are relatively similar and not noteworthy.
+The plot of the cell-wise error metric at time \(t=2.1\) in Figure~\ref{fig: reflecting transient pm t2.1} is somewhat more interesting.
+Figure~\ref{fig: reflecting transient pm t2.1} shows that the cell-wise error metric is somewhat volatile for the DG scheme; I suspect that this is due to the oscillations in the DG approximation.
+In contrast, in Figure~\ref{fig: reflecting transient pm t2.1} the uniformisation and QBD-RAP schemes have monotonically decreasing error curves.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/L1_cell_probs_error_formatted.pdf}
+	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2\) with the point mass initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: reflecting transient pm cp} 
+\end{figure}%
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1/L1_cell_probs_error_formatted.pdf}
+	\caption{\(L^1\) errors between the simulated and approximated cell masses for Model~\ref{model: simple} at time \(t=2.1\) with the point mass initial condition, and the corresponding approximations obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: reflecting transient pm t2.1} 
+\end{figure}
+
+\FloatBarrier
+\section{Hitting times}\label{sec: return approx}
+Once again we consider Model~\ref{model: simple}.
+Let \(\zeta_{X}(\{0,1\}) = \{\inf t>0 \mid X(t)=0, \mbox{ or }X(t)=1\},\) be the first hitting time of \(\{X(t)\}\) on the set \(\{0,1\}\).
+The distribution of the hitting time in phase \(i\in\{1,2\}\) is
+\begin{equation}\label{eqn: hit cdf}\mathbb P(\zeta_{X}(\{0,1\}) < t, \varphi(t)=i\mid \bs X(0)\sim \mu),\end{equation}
+for some initial distribution \(\mu\). We look at two initial conditions; an exponential with equal mass in each phase, 
+\[\mathbb P(X(0)\in\wrt x,\varphi(0)=i) = e^{-x}/(1-e^{-1})/2,\,i\in\{1,2\},\]
+and a point mass at \(X(0)=0\) in phase \(\varphi(0)=1\).
+
+We use the DG, uniformisation and QBD-RAP methods to discretise the fluid queue and partition \([0,1]\) into three intervals of width \(1/3\).
+To capture the mass which has left the interval \([0,1]\), we suppose that when the process hits the boundary it is absorbed forever at the boundary and in the phase in which the process first hit the boundary.
+
+We integrate the schemes until time \(t=10\) using the SSPRK4 method with \(t\)-step size 0.005/3.\footnote{Since we use a smaller cell-width in this example than in previous examples we need to reduce the \(t\)-step size accordingly to ensure that numerical integration is stable for schemes up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We use the DG scheme with and without a slope limiter during the time-integration.
+At each time-step of the numerical integration, we record the amount of mass at the absorbing boundaries in each phase, which gives us an approximation of the cumulative distribution function of the hitting time in each phase up to time \(t=10\).
+
+As a \emph{ground truth} we simulated 5,000,000 realisations and recorded the hitting time on the boundary of the interval \([0,1]\) and the phase at the time of hitting.
+We then compute the empirical CDF of the hitting probabilities
+\[\mathbb P(\zeta_{X}(\{0,1\}) < t, \varphi(t)=i\mid \bs X(0)\sim \mu),\] 
+for \(t=0.005/3\times k\), \(k=0,...,6000\).
+
+To account for Monte-Carlo errors we took 1,000 bootstrap resamples of the original 5,000,000 samples and computed the empirical CDF of the hitting probabilities for each bootstrap sample.
+For each bootstrap sample, we resampled 5,000,000 points with replacement from the original 5,000,000 realisations.
+For each bootstrap sample we computed the error metrics between the empirical and approximated CDFs and recorded the 5th and 95th percentile of the distribution of the errors.
+
+\paragraph{Exponential initial condition}
+Figure~\ref{fig: hitting time exp} shows the error metrics recorded for the four different numerical approximation schemes.
+The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP converging at a faster rate.
+For both error metrics the DG scheme performs the best.
+For the \(L^1\) error metric the DG scheme converges up to order 5 after which there is no improvement in the error; I believe that this is due to other numerical errors dominating.
+The DG scheme with MUSCL limiter does not appear to converge, suggesting oscillations in the DG approximation.
+Intuitively, with this initial condition we suspect that the transient distribution of the fluid queue on the event that it remains in the interval \((0,1)\) is discontinuous at \(x=t\) in Phase~1 and \(x=1-t\) in Phase~2 for \(x<1\).
+We also expect there is a discontinuity in the hitting time PDF at time \(t=1\).
+Indeed, small oscillations are visible in hitting time PDF for both phases, see Figure~\ref{fig: hitting time oscillation exp}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the exponential initial condition.
+The approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (orange dashed line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: hitting time exp} 
+\end{figure}
+\begin{figure}[h]
+	\centering 
+	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/pdf_order21.pdf}%
+	\caption{Approximations of the PDFs of the first hitting time for Model~\ref{model: simple} with the exponential initial condition.
+The blue line were obtained from the dimension-21 DG scheme and the purple lines were obtained from the dimension-21 QBD-RAP scheme.
+The DG scheme displays oscillations around the discontinuities at \(t=1\).
+}
+	\label{fig: hitting time oscillation exp} 
+\end{figure}
+\exampleFloatBarrier
+\paragraph{Point mass initial condition}
+Figure~\ref{fig: hitting time pm} shows the KS error (left) and \(L^1\) error for the CDF for the four different numerical approximation schemes applied to the hitting time problem with the point mass initial condition.
+The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP converging at a faster rate.
+The DG scheme with MUSCL limiter does not appear to converge, suggesting oscillations in the DG approximation.
+The DG scheme appears to converge fastest, however, oscillations are present in the DG approximation.
+In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis.
+Observing Figure~\ref{fig: hitting time oscillation} the DG approximation clearly displays oscillations around and the discontinuity at \(t=1\).
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
+The approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (orange dashed line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: hitting time pm} 
+\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/cdf_order21DG_and_sims.pdf}%
+	\caption{Approximations of the CDF of the first hitting time in Phase~1 for Model~\ref{model: simple} with the point mass initial condition.
+The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dashed line is the empirical CDF obtained via simulation.
+The DG scheme displays oscillations.
+}
+	\label{fig: hitting time oscillation} 
+\end{figure}
+
+\FloatBarrier
+\section{First-return distributions of fluid-fluid queues}\label{sec: ffq num}
+Now consider the fluid-fluid queue model first analysed in \cite{blnos2022} which is a modified version of a model first presented in \cite{lnp13}.
+We refer the reader to \cite{blnos2022} for more on the DG method applied to this model.
+\begin{model}\label{model: ffq}
+	Consider a stochastic fluid-fluid queue $\{(X(t),Y(t),\varphi(t))\}_{t\geq0},$ where $\{X(t)\}$ and $\{Y(t)\}$ represent the workloads in Buffers~1 and~2 at time $t \geq 0$, respectively, both driven by the phase $\{\varphi(t)\},$ which is a Markov chain on the state space $\mathcal{S} = \{11,10,01,00\}$.
+Both $\{X(t)\}$ and $\{Y(t)\}$ have a regulated boundary at 0.
+Here, the state $11$ indicates inputs to both buffers being \textnormal{\textsc{on}}, the state $00$ indicates both being \textnormal{\textsc{off}}, the state $10$ is when only the first input is \textnormal{\textsc{on}}, and the state $01$ is when only the second is \textnormal{\textsc{on}}.
+The input of Buffer~$k$ is switched from \textnormal{\textsc{on}} to \textnormal{\textsc{off}} with rate $\gamma_k$, and from \textnormal{\textsc{off}} to \textnormal{\textsc{on}} with rate $\beta_k$, for $k = 1, 2$.
+Thus, the infinitesimal generator $T$ for $\varphi(t)$ is given by
+	\begin{align*} 
+		T = \left[ \begin{array}{cccc} -(\gamma_1 + \gamma_2) & \gamma_2 & \gamma_1 & 0 \\
+							\beta_2 & -(\gamma_1 + \beta_2) & 0 & \gamma_1 \\
+							\beta_1 & 0 & -(\gamma_2 + \beta_1) & \gamma_2 \\
+							0 & \beta_1 &\beta_2 &-(\beta_1 + \beta_2)
+	\end{array}\right].
+	\end{align*}
+
+	The net rates of change for $X(t)$, denoted $c_i$, are given by 
+	% 
+	\begin{align*} 
+	& (c_{11},c_{10},c_{01},c_{00})   =  \begin{array}{rrrr} 
+	(\lambda_1-\theta_1,  & \lambda_1 -\theta_1, &  -\theta_1, & -\theta_1),
+	\end{array}  
+	\end{align*} 
+	% 
+	and the net rates of change for $Y(t)$, denoted $r_i$, are as follows  
+	%  
+	\begin{align*} 
+	(r_{11},r_{10},r_{01},r_{00})  & = \left\{ \begin{array}{lrrrll}  
+	% (\lambda_2 -\kappa, & \;\;\;\;0, & \lambda_2 - \kappa, &  \;\;\;\;0) & \text{if } X_t = 0,  Y_t = 0,\\
+	%  \vspace*{-0.3cm} \\
+												(\lambda_2 - \kappa, & \;\;\;-\kappa,  & \lambda_2 - \kappa, & -\kappa) &\text{if } X_t = 0, \\
+												\vspace*{-0.3cm} \\
+	%	(\lambda_2 -\theta_2,  &  0,  & \lambda_2 - \theta_2,  & \;\;\;\; 0) & \text{if } X_t \in (0,x^*), Y_t = 0,\\
+	%											  \vspace*{-0.3cm} \\
+		(\lambda_2 - \theta_2, & -\theta_2, & \lambda_2 - \theta_2, & -\theta_2) & \text{if } X_t \in (0,x^*),\\
+												\vspace*{-0.3cm} \\
+		(\;\;\;\;\;\;\; \lambda_2, & \;\;\;\;0, & \lambda_2, &  \;\;\;\;0) & \text{if } X_t \geq x^*. \end{array}
+												\right.
+	\end{align*}
+
+	For our numerical experiments, we use the parameter choices given in~\citep{lnp13}: 
+	% 
+		\begin{align} 
+			\label{eqn:parameters}
+		\gamma_1 & =11, \quad  \beta_1 = 1, \quad \lambda_1 = 12.48, \quad  \theta_1 = 1.6, \quad  \kappa = 2.6, \\
+			\label{eqn:parameters-2}
+		\gamma_2 & = 22, \quad \beta_2  = 1, \quad  \lambda_2 = 16.25, \quad \theta_2 = 1.0, \quad x^* = 1.6.
+		\end{align} 
+\end{model}
+
+While the true problem has an unbounded domain $[0,\infty)$, the approximations require the domain of approximation to be a finite interval.
+Here we choose an upper bound of \(48\) and place a regulated boundary at the upper boundary.
+The effect of this truncation can be partly quantified by evaluating \(\lim\limits_{t\to\infty}\mathbb P\left(X(t) > 48\right)\approx 5.83 \times 10^{-9}\), \(i\in\calS\).
+
+We obtained approximations to the infinitesimal generator of the fluid queue \(\{(X(t),\varphi(t))\}_{t\geq 0}\) via the DG method, QBD-RAP method and uniformisation method, all of which used a cell width of \(\Delta=0.4\).
+The matrix resulting from the approximation is then used to approximate the first-return operator \(\mathbb\Psi(s)\) as discussed in Section~\ref{sec: intro Psi}.
+Due to the stochastic interpretation of the uniformisation and QBD-RAP schemes the approximations to \(\Psi(s)\) have a stochastic interpretation as the first-return probabilities of a fluid queue driven by a CTMC and QBD-RAP, respectively (see \cite[Chapter~7]{p2019} and \cite{bgnp2021} for details on the latter).
+For the DG method the approximation of \(\mathbb \Psi(s)\) is not as well-understood.\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{Y}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}
+
+Ultimately, we approximate the first-return distribution 
+\begin{align}\label{eqn: first return Y 1}
+	\mathbb P(X(\zeta_{Y}(\{0\}))\leq x, \varphi(\zeta_{Y}(\{0\}))=i\mid \bs X(0)\sim \mu),
+\end{align}
+where we recall \(\zeta_{Y}(E) = \inf\{t>0\mid Y(t)\in E\}\) is the first hitting time of \(\{Y(t)\}\) on the set \(E\).
+For Model~\ref{model: ffq}, it is only possible for the process \(Y(t)\) to return to \(0\) at time \(t\) when \((X(t),\varphi(t))\in[0,1.6)\times \{10,00\}\), so we evaluate the approximations over this region only.
+We use a grid of 10,001 points on which we evaluate the approximations of the CDF in each phase.
+We consider first the initial distribution which is a point mass at \(Y(0)=0,\, X(0)=5,\, \varphi(0)=01\).
+
+For comparison, we simulated 5,000,000 realisations of the fluid-fluid queue and recorded the value of \(X\) and \(\varphi\) at the time of first return of the second fluid, \(Y\).
+The empirical approximation of (\ref{eqn: first return Y 1}) was then constructed, and error metrics for the difference between the empirical CDF and the approximations was computed.
+To account for Monte-Carlo errors we used a bootstrap with 1,000 bootstrap samples to construct 1,000 bootstrap samples of the error estimates and recorded the 5th and 95th percentiles of the error distribution.
+Each of the 1,000 bootstrap samples was constructed by resampling the original 5,000,000 realisations 5,000,000 times with replacement.
+
+In Figure~\ref{fig: ffq return cts} we plot the error metrics for the approximations to the distribution (\ref{eqn: first return Y 1}).
+The DG method performs best converging rapidly until the error in the approximation scheme is swamped by other numerical errors.
+The QBD-RAP method is second best and the uniformisation scheme appears to be the slowest to converge.
+Here, the first return distribution appears to be smooth, hence we might expect the DG method to perform well.
+Note that there is a significant number other sources of error here; machine precision errors, errors in solving the matrix Ricatti equation to approximate \(\mathbb \Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), and truncation errors.
+Furthermore, for the QBD-RAP, since the parameters \(\bs \alpha,\,\bs S,\,\bs s\,\) and \(\bs D\) are found numerically, then there is another source of error from this.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) in (\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}.
+The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: ffq return cts} 
+\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cell_probs_error_formatted.pdf}%
+	\caption{Total error between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}.
+The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: ffq cell probs} 
+\end{figure}
+\exampleFloatBarrier
+By modifying slightly Model~\ref{model: ffq}, we can construct a first return distribution which is discontinuous. 
+\begin{model}\label{model: ffq2}
+	Consider a fluid-fluid queue which is the same as Model~\ref{model: ffq} except 
+	\begin{align}
+		r_{00}(X(t)) = \begin{cases}
+			-\kappa, & \mbox{ if }X(t)=0,\\
+			-\theta_2, & \mbox{ if }X(t)=\in(0,x^*),\\
+			\theta_2, & \mbox{ if }X(t)\geq x^*,
+		\end{cases}
+	\end{align}
+	with an initial distribution which is a point-mass at \(Y(0)=0,\, X(0)=2, \varphi(0)=00\). 
+\end{model}
+As before, we use the DG, uniformisation and QBD-RAP methods to approximate the model, and compare to simulations.
+
+For Model~\ref{model: ffq2} there is a point mass at \(X(\zeta_{Y}(\{0\}))=1.2\) of magnitude \(e^{-(\beta_1+\beta_2)}\times 0.5\), which occurs when the phase remains in \(\varphi(t)=00\) for all \(t\in [0,\zeta_{Y}(\{0\})]\).
+
+Figure~\ref{fig: ffq return discts} plots the error metrics for the first return distribution.
+Observing Figure~\ref{fig: ffq return discts} we see that the DG method performs well, followed by the QBD-RAP method, then the uniformisation method performs worst.
+However, the DG method results in an oscillatory solution with the CDF taking impossible values (decreasing at points) as shown in Figure~\ref{fig: ffq2 oscillation}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}.
+The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: ffq return discts} 
+\end{figure} 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cell_probs_error_formatted.pdf}%
+	\caption{Cell-wise error between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing in each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq2}.
+The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\label{fig: ffq2 cell probs} 
+\end{figure}
+\begin{figure}[h] 
+	\centering
+	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/ffq/discts/phase_4_cdf.pdf}%
+	\caption{Approximations of the CDF of the distribution of \(X\) at the time \(\zeta_Y(\{0\})\) in Phase~1 for Model~\ref{model: ffq2}.
+The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dashed line is the empirical CDF obtained via simulation.
+The DG scheme displays oscillations around the discontinuity.
+}
+	\label{fig: ffq2 oscillation} 
+\end{figure}
+\FloatBarrier
+
+\section{Discussion}
+
+The structure of this chapter is as follows.
+In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different methods and observe their performance at approximating the initial condition, which allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
+In Section~\ref{sec: wave num} we investigate a simple travelling wave problem with various initial conditions.
+For this problem the dynamics are deterministic, which allows us to instrument the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
+Next, we investigate a simple fluid queue with two phases.
+Section~\ref{sec:stat} investigates the ability of the methods to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the fluid queue for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the methods to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
+Section~\ref{sec: ffq num} approximates the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\) for two simple fluid-fluid queue models.
+
+In this chapter we have numerically investigated some properties of the QBD-RAP approximation scheme and compared it to existing methods; the uniformisation scheme of \cite{bo2013} and the discontinuous Galerkin scheme.
+In general, the numerical experiments show that, for problems with discontinuities, the DG approximation can exhibit oscillations, while the QBD-RAP and uniformisation approximations avoid this and, of the latter two, the QBD-RAP scheme often converges faster.
+To avoid the problems of oscillations we can sometimes employ a \emph{slope limiter} with the DG scheme, which effectively reduces the scheme to linear in the regions where oscillations are detected.
+The numerical experiments demonstrate a loss of accuracy in the DG approximation when a slope limiter is used for a purely discontinuous problem.
+Moreover, for the application of the methods to fluid-fluid queues, there is no obvious way to apply the concept of a slope limiter.
+%When the slope limiter detects oscillations in the approximate solution, it reduces the DG scheme to linear in the region surrounding the oscillations, otherwise, the slope limiter leaves the solution unchanged.
+Thus, the slope limiter permits high-order approximations away from oscillations, while also removing oscillations.
+In general, we observe that the smoother the problem is the better the performance of the DG method, and it emphatically outperforms the other two methods.
+
+As a first step in the numerical experiments, in Section~\ref{sec: recon num}, we examined the ability of each method to approximate different initial conditions.
+For the DG method, this is equivalent to a projection of the initial condition on to a set of basis polynomials.
+For the uniformisation method this is equivalent to projecting the initial condition on to a basis of constant functions.
+For the QBD-RAP method the approximation of the initial condition is as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
+Section~\ref{sec: comp} demonstrates that the DG scheme (projection) can result in oscillations and negative regions in the approximation when the initial condition is discontinuous.
+The uniformisation and QBD-RAP methods avoid this problem, but appear to have higher errors and the QBD-RAP method appears to have the largest errors.
+For discontinuous initial conditions the rates of convergence are comparable for all three methods.
+However, when the initial condition to be approximated is sufficiently smooth, then the DG approximation is superior.
+
+Next, in Section~\ref{sec: wave num}, we instrumented the performance of approximations for a simple travelling wave problem with various initial conditions.
+For this problem the solution is given in terms of the initial condition by \(f(x,t) = f(x-t,0)\).
+This model is useful as it allows us to instrument the ability of the methods to capture the flow of probability without stochastic dynamics.
+Once again, for discontinuous problems, the DG method can display oscillations, while the other methods (DG with slope limiter, uniformisation and QBD-RAP methods) avoid oscillations.
+Further, for discontinuous problems the rates of convergence of the QBD-RAP and DG methods can be similar.
+For smooth problems the DG method is superior.
+Interestingly, even though the QBD-RAP method performed worst for approximating initial conditions in the previous section, it outperformed the uniformisation method for this model, demonstrating that the QBD-RAP method can capture the dynamics of the flow of probability better than the uniformisation method.
+
+We then instrumented the performance of the approximations on a simple fluid queue with two phases.
+First, in Section~\ref{sec:stat}, we looked at the stationary distribution, which is known to be smooth.
+Since the problem is smooth, then the DG method was superior as expected.
+Of the uniformisation and QBD-RAP methods, the QBD-RAP method gives more accurate solutions.
+In Section~\ref{sec: transient approx} we turned our attention to approximating transient distributions for the same models and consider two different initial conditions, a point-mass and an exponential initial condition.
+The discontinuous initial condition results in a discontinuous transient distribution.
+As for the exponential initial condition, this example demonstrates that, even if the initial condition appears `nice', it can still result in discontinuous, or non-differetiable solutions.
+The numerical evidence suggests that the DG method can display oscillations, while the other methods do not.
+The DG method with slope limiter detects the oscillations and reduces the method to linear.
+Of the uniformisation and QBD-RAP methods, the latter performs better.
+
+Next, Section~\ref{sec: return approx} looked at hitting times for the same fluid queue with two initial conditions, an exponential initial condition and a point-mass.
+We looked at the hitting time of the fluid level on the points \(\{0\}\) and \(\{1\}\).
+For this problem there is never any in-flow of mass at the boundaries of the interval and so, for a solution to be continuous, the initial condition needs to be chosen carefully, otherwise discontinuities in the transient distribution may result, as is the case for both initial conditions here.
+The numerical results suggest that, due to the discontinuities in the problems, the DG method may display oscillations.
+Since the uniformisation and QBD-RAP methods can handle discontinuities, they perform as expected, with the QBD-RAP method performing better than the uniformisation.
+
+Lastly, we applied the DG, uniformisation and QBD-RAP methods to two simple fluid-fluid queues in Section~\ref{sec: ffq num}.
+In the first fluid-fluid queue, which appears to have a smooth solution, the DG method performs very well.
+Of the two positivity preserving schemes the QBD-RAP scheme performs better than the uniformisation scheme.
+In the second example, which has a discontinuity, the DG method produces the lowest errors, but exhibits oscillations in the solution.
+The QBD-RAP and uniformisation schemes do not produce oscillations and, of the two, the QBD-RAP method performs best.
+
+In conclusion, when the problem is known to be smooth, the DG method is very likely to produce excellent results.
+However, for discontinuous problems, the method can show oscillations and infeasible or negative regions of the solution.
+The slope limiter overcomes this, but reduces the accuracy of the DG method to linear near discontinuities, sometimes severely affecting the quality of the approximation.
+The uniformisation and QBD-RAP methods are alternative approximation schemes which exhibit larger errors, but avoid oscillatory solutions and guarantee a non-decreasing approximation of the CDF.
+Of the uniformisation and QBD-RAP methods, the latter often to produced lower errors.

### 2022-02-10 chapter6/figs/cp_figs.sh
first draft :D

@@ -32,6 +32,7 @@ cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/hitting_times_model/re
 cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/hitting_times_model/reflecting_model/transient_distribution/point_mass/datat2/figs/l1_cdf_error_formatted.pdf '/Users/anguslewis/Documents/PhDThesis/chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1'
 cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/hitting_times_model/reflecting_model/transient_distribution/point_mass/datat2/figs/L1_cell_probs_error_formatted.pdf '/Users/anguslewis/Documents/PhDThesis/chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1'
 cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/hitting_times_model/reflecting_model/transient_distribution/point_mass/datat2/figs/ks_error_formatted.pdf '/Users/anguslewis/Documents/PhDThesis/chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1'
+cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/hitting_times_model/reflecting_model/transient_distribution/point_mass/datat2/figs/order_21_CDFs_formatted.pdf '/Users/anguslewis/Documents/PhDThesis/chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1'
 
 cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/wave/func_count_1/mesh_comp/figs/pdf_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/wave/fun1
 cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/wave/func_count_1/mesh_comp/figs/meshs_l1_cdf_error_formatted.pdf /Users/anguslewis/Documents/PhDThesis/chapter6/figs/wave/fun1
@@ -51,6 +52,7 @@ cp /Users/anguslewis/Documents/SFFMProject/DFQExamples.jl/wave/func_count_4/mesh
[…]

### 2022-02-10 appendix_dg_toy/appendix.tex
.

@@
 I am a co-author of the paper \cite{blnos2022}.
-The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara.
-I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme.
-I contributed Sections~4.4 and 5.1.
-I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6.
-Appendix~A is also my original work.
-I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
         }
@@
 Consider a process \(\{(\overline X_t,Y_t,\varphi_t)\}_{t\geq 0}\) with two phases, \(\varphi_t\in\calS=\{1,2\}\) and generator matrix \(\bs T\).
-Let \(\mathcal I = 1.8\), and partition into two intervals \(\calD_1=[0,1]\) and \(\calD_2=[1,1.8]\), hence \(x_1=0,\,x_2=1,\,x_3=1.8\).
+Let \(\mathcal I = 1.8\), and partition into two intervals \(\calD_0=[0,1]\) and \(\calD_1=[1,1.8]\), hence \(x_0=0,\,x_1=1,\,x_2=1.8\).
 We choose a basis of Lagrange polynomials of order 1 to define our approximation space.
@@
 \begin{align*}
-	\phi_{1}^1(x) &= 1-x,& \phi_2^1(x) &= x, & &x \in\calD_1,\\
-	\phi_{1}^2(x) &= \cfrac{1.8-x}{0.8},& \phi_2^2(x) &= \cfrac{x-1}{0.8}, & & x \in\calD_2.
+	\phi_{0}^1(x) &= 1-x,& \phi_0^2(x) &= x, & &x \in\calD_0,\\
+	\phi_{1}^1(x) &= \cfrac{1.8-x}{0.8},& \phi_1^2(x) &= \cfrac{x-1}{0.8}, & & x \in\calD_1.
 \end{align*}
@@
     \draw[dashed] (1.8,0) coordinate (d_1) -- (1.8,1.1) coordinate (d_2);
-    \filldraw (0.5,0) circle (0pt) node[align=center, below] {\(\calD_1\)};
-    \filldraw (1.4,0) circle (0pt) node[align=center, below] {\(\calD_2\)};
-    \filldraw (0,0) circle (0pt) node[align=center, below] {\(x_1=0\)};
-    \filldraw (1,0) circle (0pt) node[align=center, below] {\(x_2=1\)};
-    \filldraw (1.8,0) circle (0pt) node[align=center, below] {\(x_3=1.8\)};  
+    \filldraw (0.5,0) circle (0pt) node[align=center, below] {\(\calD_0\)};
+    \filldraw (1.4,0) circle (0pt) node[align=center, below] {\(\calD_1\)};
+    \filldraw (0,0) circle (0pt) node[align=center, below] {\(x_0=0\)};
+    \filldraw (1,0) circle (0pt) node[align=center, below] {\(x_1=1\)};
+    \filldraw (1.8,0) circle (0pt) node[align=center, below] {\(x_2=1.8\)};  
     \filldraw (0,1) circle (0pt) node[align=center, left] {\(1\)};    
-    \filldraw (0,1) circle (0pt) node[align=center, right] {\color{blue}\(\phi_1^1(x)\)};    
-    \filldraw (1,1) circle (0pt) node[align=center, left] {\color{red}\(\phi_2^1(x)\)};    
-    \filldraw (1,1) circle (0pt) node[align=center, right] {\color{blue}\(\phi_1^2(x)\)};    
-    \filldraw (1.8,1) circle (0pt) node[align=center, left] {\color{red}\(\phi_2^2(x)\)};    
+    \filldraw (0,1) circle (0pt) node[align=center, right] {\color{blue}\(\phi_0^1(x)\)};    
+    \filldraw (1,1) circle (0pt) node[align=center, left] {\color{red}\(\phi_0^2(x)\)};    
+    \filldraw (1,1) circle (0pt) node[align=center, right] {\color{blue}\(\phi_1^1(x)\)};    
+    \filldraw (1.8,1) circle (0pt) node[align=center, left] {\color{red}\(\phi_1^2(x)\)};    
 \end{tikzpicture}
-\caption{A mesh with nodes \(x_1=0\), \(x_2=1\) and \(x_3=1.8\) and cells \(\calD_1=[0,1]\), \(\calD_2=[1,1.8]\).
+\caption{A mesh with nodes \(x_0=0\), \(x_1=1\) and \(x_2=1.8\) and cells \(\calD_0=[0,1]\), \(\calD_1=[1,1.8]\).
 There are two basis functions on each cell.
-Point masses are located at \(x_1=0\) and \(x_3=1.8\).\label{fig: mesh}}
+Point masses are located at \(x_0=0\) and \(x_2=1.8\).\label{fig: mesh}}
 \end{figure}
@@
 
-Suppose that \(r_1(x)>0\) on \(\calD_1=\mathcal F_1^+\) and \(r_1(x)<0\) on \(\calD_2\cup\{1\}=\mathcal F_1^-\), and further, that \(r_2(x)<0\) on \(\{0\}\cup\calD_1=\mathcal F_2^-\) and \(r_2(x)>0\) on \(\calD_2=\mathcal F_2^+\).
+Suppose that \(r_1(x)>0\) on \(\calD_0=\mathcal F_1^+\) and \(r_1(x)<0\) on \(\calD_1\cup\{1\}=\mathcal F_1^-\), and further, that \(r_2(x)<0\) on \(\{0\}\cup\calD_0=\mathcal F_2^-\) and \(r_2(x)>0\) on \(\calD_1=\mathcal F_2^+\).
 Specifically, let
@@
 \begin{blockarray}{c|cc|cc|cc|cc|c}
-\nabla & \BAmulticolumn{4}{c|}{\calD_1} & \BAmulticolumn{4}{c|}{\calD_2} & \Delta\\
+{-1} & \BAmulticolumn{4}{c|}{\calD_0} & \BAmulticolumn{4}{c|}{\calD_2} & {K+1}\\
 \calF_2^- & \BAmulticolumn{2}{c|}{\calF_1^+} & \BAmulticolumn{2}{c|}{\calF_2^-} & \BAmulticolumn{2}{c|}{\calF_1^-} & \BAmulticolumn{2}{c|}{\calF_2^+} & \calF_1^-\\
-q_{\nabla,1} & a_{1,1}^1 & a_{1,2}^1 & a_{2,1}^1 & a_{2,2}^1 & a_{1,1}^2 & a_{1,2}^2 & a_{2,1}^2 & a_{2,2}^2 & q_{\Delta,1}\\\BAhline
+q_{{-1},0} & a_{0,1}^1 & a_{0,1}^2 & a_{0,2}^1 & a_{0,2}^2 & a_{1,1}^1 & a_{1,1}^2 & a_{1,2}^1 & a_{1,2}^2 & q_{{K+1},1}\\\BAhline
 \begin{block}{[c|cc|cc|cc|cc|c]} 
@@
 Now, to find \( {\bs\xi}\), we solve the linear system in Equations \eqref{eqn:xi1}-\eqref{eqn:xi2}. The result is a vector which we denote, 
-\[ {\bs\xi} = \left[\begin{array}{c|cc|cc|c} {\xi}^\nabla_{2} &  {\xi}_{2,1}^1 &  {\xi}_{2,2}^1 &  {\xi}_{1,1}^2 &  {\xi}_{1,2}^2 &  {\xi}_{1}^\Delta \end{array}\right],\]
-where \( {\xi}^\nabla_{2}\) is an approximation to \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =0, \varphi_{\theta_n} = 2\right)\) and \( {\xi}^\Delta_{1}\) is an approximation to the artificial point mass \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =1.8, \varphi_{\theta_n} = 1\right)\).
-For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 2\right)\), is constructed as \( {\xi}_{2,1}^1\phi_1^1(x) +  {\xi}_{2,2}^1\phi_2^1(x)\).
-For \(x\in\calD_2\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 1\right)\), is constructed as \( {\xi}_{1,1}^2\phi_1^2(x) +  {\xi}_{1,2}^2\phi_2^2(x).\)
+\[ {\bs\xi} = \left[\begin{array}{c|cc|cc|c} {\xi}_{-1,2} &  {\xi}_{0,2}^1 &  {\xi}_{0,2}^2 &  {\xi}_{1,1}^1 &  {\xi}_{1,1}^2 &  {\xi}_{K+1,1} \end{array}\right],\]
+where \( {\xi}_{-1,2}\) is an approximation to \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =0, \varphi_{\theta_n} = 2\right)\) and \( {\xi}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =1.8, \varphi_{\theta_n} = 1\right)\).
+For \(x\in\calD_0\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 2\right)\), is constructed as \( {\xi}_{0,2}^0\phi_0^1(x) +  {\xi}_{0,2}^0\phi_0^1(x)\).
+For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 1\right)\), is constructed as \( {\xi}_{1,1}^1\phi_1^1(x) +  {\xi}_{1,1}^2\phi_1^2(x).\)
 
@@
 For the point masses we have 
-\[ {\bs p}^- = \left[\begin{array}{c|cc|cc|c} {p}_{2}^\nabla &  {p}_{2,1}^1 &  {p}_{2,2}^1 &  {p}_{1,1}^2 &  {p}_{1,2}^2 &  {p}_{1}^\Delta \end{array}\right],\]
-where \( {p}_{2}^\nabla\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_{t} =0, Y_t=0, \varphi_{t} = 2\right)\) and \( {p}_{1}^\Delta\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t=0, \varphi_{t} = 1\right)\).
-For \(x\in\calD_1\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 2\right)\), is constructed as \( {p}_{2,1}^1\phi_1^1(x) +  {p}_{2,2}^1\phi_2^1(x)\).
-For \(x\in\calD_2\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 1\right)\), is constructed as \( {p}_{1,1}^2\phi_1^2(x) +  {p}_{1,2}^2\phi_2^2(x).\)
+\[ {\bs p}^- = \left[\begin{array}{c|cc|cc|c} {p}_{-1,2} &  {p}_{0,2}^1 &  {p}_{0,2}^2 &  {p}_{1,1}^1 &  {p}_{1,1}^2 &  {p}_{K+1,1} \end{array}\right],\]
+where \( {p}_{-1,2}\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_{t} =0, Y_t=0, \varphi_{t} = 2\right)\) and \( {p}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t=0, \varphi_{t} = 1\right)\).
+For \(x\in\calD_0\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 2\right)\), is constructed as \( {p}_{0,2}^1\phi_0^1(x) +  {p}_{0,2}^2\phi_0^2(x)\).
+For \(x\in\calD_1\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 1\right)\), is constructed as \( {p}_{1,1}^1\phi_1^1(x) +  {p}_{1,1}^2\phi_1^2(x).\)
 
 Similarly, for \( {\bs\pi}^-(y)\), we have 
-\[ {\bs \pi}^-(y) = \left[\begin{array}{c|cc|cc|c} {\pi}_{2}^\nabla(y) &  {\pi}_{2,1}^1(y) &  {\pi}_{2,2}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{1,2}^2(y) &  {\pi}_{1}^\Delta(y) \end{array}\right],\]
-where \( {\pi}_{2}^\nabla(y)\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t\in\wrt y, \varphi_{t} = 2\right)\) and \( {\pi}_{1}^\Delta(y)\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t\in \wrt y, \varphi_{t} = 1\right)\).
-For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\), is constructed as \( {\pi}_{2,1}^1(y)\phi_1^1(x) +  {\pi}_{2,2}^1(y)\phi_2^1(x)\).
-For \(x\in\calD_2\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\), is constructed as \( {\pi}_{1,1}^2(y)\phi_1^2(x) +  {\pi}_{1,2}^2(y)\phi_2^2(x).\)
+\[ {\bs \pi}^-(y) = \left[\begin{array}{c|cc|cc|c} {\pi}_{-1,2}(y) &  {\pi}_{0,2}^1(y) &  {\pi}_{0,2}^2(y) &  {\pi}_{1,1}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{K+1,1}(y) \end{array}\right],\]
+where \( {\pi}_{-1,2}(y)\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t\in\wrt y, \varphi_{t} = 2\right)\) and \( {\pi}_{K+1,1}(y)\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t\in \wrt y, \varphi_{t} = 1\right)\).
+For \(x\in\calD_0\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\), is constructed as \( {\pi}_{0,2}^1(y)\phi_0^1(x) +  {\pi}_{0,2}^2(y)\phi_0^2(x)\).
+For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\), is constructed as \( {\pi}_{1,1}^1(y)\phi_1^1(x) +  {\pi}_{1,1}^2(y)\phi_1^2(x).\)
 
 For \( {\bs\pi}^+(y)\), we have 
-\[ {\bs \pi}^+(y) = \left[\begin{array}{cc|cc}  {\pi}_{1,1}^1(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{2,1}^2(y) &  {\pi}_{2,2}^2(y) \end{array}\right].\]
-For \(x\in\calD_1\) an approximation to the joint density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\) is constructed as \( {\pi}_{1,1}^1(y)\phi_1^1(x) +  {\pi}_{1,2}^1(y)\phi_2^1(x)\).
-For \(x\in\calD_2\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\) is constructed as \( {\pi}_{2,1}^2(y)\phi_1^2(x) +  {\pi}_{2,2}^2(y)\phi_2^2(x).\)
+\[ {\bs \pi}^+(y) = \left[\begin{array}{cc|cc}  {\pi}_{0,1}^1(y) &  {\pi}_{0,1}^2(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{1,2}^2(y) \end{array}\right].\]
+For \(x\in\calD_0\) an approximation to the joint density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\) is constructed as \( {\pi}_{0,1}^1(y)\phi_0^1(x) +  {\pi}_{0,1}^2(y)\phi_0^2(x)\).
+For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\) is constructed as \( {\pi}_{1,2}^1(y)\phi_1^1(x) +  {\pi}_{1,2}^2(y)\phi_1^2(x).\)
 
@@
 \begin{align*}
-	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = i\right) &\approx \displaystyle \sum_{r\in\{1,2\},k\in\{1,2\}}  {\pi}_{i,r}^k(y)\phi_r^k(x)\wrt x \wrt y ,\quad x\in(0,1.8),\, y>0, \\
-	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t=0, \varphi_{t} = i\right) & \approx \displaystyle \sum_{r\in\{1,2\},k\in\{1,2\}}  {p}_{i,r}^k\phi_r^k(x)\wrt x ,\quad x\in(0,1.8),\\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx{\pi}_{i}^\nabla(y)\wrt y,\quad y>0, \\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 0, Y_t =0, \varphi_{t} = i\right) & \approx {p}_{i}^\nabla,\quad \\ 
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx {\pi}_{i}^\Delta(y)\wrt y,\quad y>0, \\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t = 0, \varphi_{t} = i\right) & \approx {p}_{i}^\Delta.\quad
+	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = i\right) &\approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {\pi}_{i,k}^r(y)\phi_k^r(x)\wrt x \wrt y ,\quad x\in(0,1.8),\, y>0, \\
+	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t=0, \varphi_{t} = i\right) & \approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {p}_{k,i}^r\phi_k^r(x)\wrt x ,\quad x\in(0,1.8),\\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx{\pi}_{-1,i}(y)\wrt y,\quad y>0, \\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 0, Y_t =0, \varphi_{t} = i\right) & \approx {p}_{-1,i},\quad \\ 
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx {\pi}_{K+1,i}(y)\wrt y,\quad y>0, \\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t = 0, \varphi_{t} = i\right) & \approx {p}_{K+1,i}.\quad
 \end{align*}
-%\[\left[\begin{array}{c|cc|cc|cc|cc|c} {\pi}_{2,\nabla}(y) &  {\pi}_{1,1}^1(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{2,1}^1(y) &  {\pi}_{2,2}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{1,2}^2(y) &  {\pi}_{2,1}^2(y) &  {\pi}_{2,2}^2(y) &  {\pi}_{1,\Delta}(y) \end{array}\right]\]
+%\[\left[\begin{array}{c|cc|cc|cc|cc|c} {\pi}_{2,{-1}}(y) &  {\pi}_{1,1}^1(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{2,1}^1(y) &  {\pi}_{2,2}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{1,2}^2(y) &  {\pi}_{2,1}^2(y) &  {\pi}_{2,2}^2(y) &  {\pi}_{1,{K+1}}(y) \end{array}\right]\]

### 2022-02-10 appendix_properties_of_B/appendix.tex
.

\section{Properties of the DG operator \(\bs B\)}\label{sec:properties}
In this section we prove that the DG approximation \(\bs B\) conserves probability.

Recall that the coefficients \({\boldsymbol a}_{k,i}(t)\) can be used to construct an approximate solution to a differential equation at time \(t\) as \(u_{k,i}(x,t)={\boldsymbol a}_{k,i}(t)\boldsymbol \phi_k(x)\tr{}.\) For \(i\in\calS,\,k\in\mathcal K,\,r\in\mathcal N_k\), define \(\alpha_{k,i}^r(t) := a_{k,i}^r(t)\displaystyle\int_{x\in\calD_k}\phi^r_k(x)\wrt x,\) and row-vectors \(\boldsymbol \alpha_{k,i}(t)=(\alpha_{k,i}^r(t))_{r\in\mathcal N_k}\). Motivated by the fact that we may be interested in approximations of the probabilities \(\mathbb P(X(t)\in\calD_k,\varphi(t)=i)\) rather than the function \(u_{k,i}\) itself, we can pose the problem equivalently in terms of the integrals 
\[\mathbb P\left(X(t)\in\calD_k,\varphi(t)=i\right)\approx{\boldsymbol a}_{k,i}(t)\int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\wrt x={\boldsymbol \alpha}_{k,i}(t)\boldsymbol 1.\] 
Define
\[\boldsymbol \alpha_k(t) = (\boldsymbol \alpha_{k,i}(t))_{i\in\calS},\mbox{ and } \boldsymbol \alpha(t)=(\boldsymbol \alpha_k(t)))_{k\in\mathcal K},\]
and matrices 
\begin{align*}
\bs P_k &= \diag\left(\int_{x\in\calD_k}\phi^r_k(x)\wrt x\right)_{ r\in\mathcal N_k},\, k\in\mathcal K^\circ,
\\\bs P &=\left[\begin{array}{ccc}
		\bs I_{N_{|\calS|}}\otimes \bs P_0 & &\\
		& \ddots &\\
		& & \bs I_{N_{|\calS|}}\otimes \bs P_K
	\end{array}\right].
\end{align*}
By choosing the basis \(\{\phi_{k}^r\}_{r\in\mathcal N_k,k\in\mathcal K^\circ}\) such that \(\int_{x\in\calD_k}\phi^r_k(x)\wrt x\neq 0\) for all \(r,k\), then \(\bs P\) is invertible. This is the case for the Lagrange polynomials, but not, for example, for the Legendre polynomials. We can (loosely) interpret the new coefficients \(\alpha_{k,i}^r(t)\) as representing the amount of probability captured by the basis function \(\phi^r_k(x)\) in phase \(i\).

The differential equation (\ref{eqn: DG ODE w BCs}) can be equivalently expressed as 
\(\label{eqn: DG alpha ODE w BCs}
	\cfrac{\wrt}{\wrt t} \boldsymbol \alpha(t)
	% 
	= \boldsymbol \alpha(t)  { \mathfrak{\bs B}},
\)
where 
\[ {\mathfrak{\bs B}} = \left[\begin{array}{ccc}
		\bs I_{|\calS_{-1}|} & & \\
		& \bs P^{-1} &  \\
		& & \bs I_{|\calS_{K+1}|}
	\end{array}\right]
	\bs B
	\left[\begin{array}{ccc}
		\bs I_{|\calS_{-1}|} & & \\
		& \bs P&  \\
		& & \bs I_{|\calS_{K+1}|}
	\end{array}\right].\] 
Let
\begin{align*}
 {  {\mathfrak{\bs B}}}^{{-1}0} &:= \bs T_{{-1}+}\otimes\left( \boldsymbol\phi_0(0)\bs M_0^{-1}\bs P_0\right),
 %
 \\ {  {\mathfrak{\bs B}}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1}\mathbb 1_{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}}\otimes \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{},
 %
 \\ {  {\mathfrak{\bs B}}}^{{K+1} K} &:= \bs T_{{K+1}-}\otimes \left(\boldsymbol\phi_K(b)\bs M_K^{-1}\bs P_K\right)
 %
 \\ {  {\mathfrak{\bs B}}}^{K {K+1}} &:= \left[c_ip_{ij}^{K+1}\mathbb 1_{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes  \bs P_K^{-1} \boldsymbol\phi_K(b)\tr{},
 %
 \\\mathfrak{\bs B}^{kk}_{ij} &:= \begin{cases}
    	T_{ii}\bs I_{N_k} + c_i\bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k & i = j, \\
	T_{ij}\bs I_{N_k} & i \neq j,
    \end{cases}\quad \mbox{ for \(k=1,2,\dots K-1\),}
 %
 \\\mathfrak{\bs B}^{00}_{ij} &:= \begin{cases}
	T_{ii}\bs I_{N_0} + c_i\bs P_0^{-1}(\bs F_i^{00}+\bs G_0)\bs M_0^{-1}\bs P_0 & i = j, \\
T_{ij}\bs I_{N_0} - c_ip_{ij}^{-1} 1(c_i<0)\bs F_i^{00}\bs M_0^{-1}\bs P_0 & i \neq j,
\end{cases}
%
\\\mathfrak{\bs B}^{KK}_{ij} &:= \begin{cases}
	T_{ii}\bs I_{N_K} + c_i\bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K & i = j, \\
T_{ij}\bs I_{N_K} + c_ip_{ij}^{K+1} 1(c_i>0)\bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K & i \neq j,
\end{cases}
%
 \\\mathfrak{\bs B}^{k,k+1}_{ij} &:= \begin{cases}
	c_i\bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k+1}^{-1}\bs P_{k+1} & i = j, \\
	\bs 0_{N_k} & i \neq j,
    \end{cases}\quad \mbox{ for \(k=0,1,\dots K-1\),}
 %
 \\\mathfrak{\bs B}^{k-1,k}_{ij} &:= \begin{cases}
    	c_i\bs P_k^{-1}\bs F_i^{k,k-1}\bs M_{k-1}^{-1}\bs P_{k-1} & i = j, \\
	\bs 0_{N_k} & i \neq j,
    \end{cases},\quad \mbox{ for \(k=2,\dots K\),}
\\
     {  {\mathfrak{\bs B}}}^{kk}%&=\left[\begin{array}{ccc}T_{11}I_{N_k} + c_1P_k^{-1}(F_1^{kk}+G_k)M_k^{-1}P_k & T_{12}I_{N_k} & T_{1N_\calS}I_{N_k}  \\ T_{21}I_{N_k} & & \\ \vdots &\ddots & \vdots \\ & &   T_{N_\calS-1,N_\calS}I_{N_k} \\  T_{N_\calS1}I_{N_k} &  T_{N_\calS,N_\calS-1}I_{N_k} & T_{N_\calS,N_\calS}I_{N_k} +c_{N_\calS}P_k^{-1}(F_{N_\calS}^{kk}+G_k)M_k^{-1}P_k\end{array}\right]
%    %
%    &=:\left[\begin{array}{ccc} {  {\mathfrak B}}_{11}^{kk} &  {  {\mathfrak B}}^{kk}_{12} &  {  {\mathfrak B}}^{kk}_{1N_\calS}  \\  {  {\mathfrak B}}^{kk}_{21} & & \\ \vdots &\ddots & \vdots \\ & &    {  {\mathfrak B}}^{kk}_{N_\calS-1,N_\calS} \\   {  {\mathfrak B}}^{kk}_{N_\calS1} &   {  {\mathfrak B}}^{kk}_{N_\calS,N_\calS-1} &  {  {\mathfrak B}}^{kk}_{N_\calS,N_\calS} \end{array}\right], 
    %
    &=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}_{11}^{kk} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{1N_\calS}  \\ \vdots &\ddots & \vdots \\   {  {\mathfrak{\bs B}}}^{kk}_{N_\calS1} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{N_\calS,N_\calS} \end{array}\right], \mbox{ for \(k\in\mathcal K^\circ\),}
    %
\\ {  {\mathfrak{\bs B}}}^{k,k+1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k+1}M_{k+1}^{-1}P_{k+1}&  & \\  &\ddots &  \\ & &   \\  &  &c_{N_\calS}P_k^{-1}F_{N_\calS}^{k,k+1}M_{k+1}^{-1}P_{k+1} \end{array}\right] 
%
&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k+1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k+1}_{1,N_\calS}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k+1}_{N_\calS,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k+1}_{N_\calS,N_\calS} \end{array}\right],\mbox{ for \(k=0,1,\dots K-1\),}
%
\\ {  {\mathfrak{\bs B}}}^{k,k-1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k-1}M_{k-1}^{-1}P_{k-1}&  & \\  &\ddots &  \\  &  &c_{N_\calS}P_k^{-1}F_{N_\calS}^{k,k-1}M_{k-1}^{-1}P_{k-1} \end{array}\right]
%
&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k-1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k-1}_{1,N_\calS}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k-1}_{N_\calS,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k-1}_{N_\calS,N_\calS} \end{array}\right], \mbox{ for \(k=1,2,\dots K\).}
\end{align*} 
Then
\begin{align*}
{  {\mathfrak{\bs B}}} &= \left[\begin{array}{llllll}
	\bs T_{{-1},{-1}}&  {  {\mathfrak{\bs B}}}^{{-1}0} & & & & \\
	 {  {\mathfrak{\bs B}}}^{0{-1}} &  {  {\mathfrak{\bs B}}}^{00} &  {  {\mathfrak{\bs B}}}^{01} & & & \\
	&  {  {\mathfrak{\bs B}}}^{10} &  {  {\mathfrak{\bs B}}}^{11} &  {  {\mathfrak{\bs B}}}^{12} & & \\
	& & \ddots & \ddots & \ddots & \\
	& &  {   {\mathfrak{\bs B}}}^{K-1,K-2} & {   {\mathfrak{\bs B}}}^{K-1,K-1} &  {  {\mathfrak{\bs B}}}^{K-1,K} & \\
	& & & {   {\mathfrak{\bs B}}}^{K,K-1} &  {  {\mathfrak{\bs B}}}^{K,K} &  {  {\mathfrak{\bs B}}}^{K,{K+1}} \\
	& & & &  {  {\mathfrak{\bs B}}}^{{K+1}, K} & \bs T_{{K+1},{K+1}}
\end{array}\right].\end{align*}

	\begin{rem}
	One may recognise the structure of \( {  {\mathfrak{\bs B}}}\) as the structure of a quasi-birth-and-death process (QBD), with levels \(k\in\mathcal K^\circ\). This raises the question of whether \( {  {\mathfrak{\bs B}}}\) is indeed a representation of the generator matrix of a QBD, or QBD-like process. In the case of a constant basis function on each cell, i.e.~\(N_k=1\) and \(\phi_k^1(x)\propto1\), \(k\in\mathcal K^\circ\), then \( {  {\mathfrak{\bs B}}}\) is the generator of a QBD: it has zero row-sums, negative diagonal entires, and non-negative off-diagonal entries, the QBD-phase variable is \(\{\varphi_t\}\) and the level is \(k\in\mathcal K\). In fact, if \(\Delta_k\) is the same for every \(k\in\mathcal K^\circ\), then this is the same QBD discretisation of a stochastic fluid process analysed by \cite{bo2013}. However, for higher-degree polynomials \(\mathfrak{\bs B}\) is not necessarily the generator of a QBD process. We conjecture that, using polynomial basis functions, then \(N_k=1\) and \(\phi_k^1(x)\propto1\), \(k\in\mathcal K^\circ\) is the only DG approximation which has an interpretation as a QBD-like process -- not even as a QBD-RAP \citep{bn2010}.
	\end{rem}

	In the following lemma, we use the following properties of the Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature nodes. 
	\paragraph{Property 1} \(\displaystyle\sum_{s=1}^{N_k} \phi_k^s(x) = \begin{cases} 1&x\in\calD_k,\\ 0 & x\notin \calD_k.\end{cases}\)

	For \(k\in\mathcal K^\circ\), let \(\boldsymbol e_n^k\) be a row-vector of length \(N_k\) with a 1 in the \(n\)th position and zeros elsewhere.
	\paragraph{Property 2} At the cell edges, \(\bs \phi_k(y_k) = \bs e_1^k\) and \(\bs\phi_k(y_{k+1})=\bs e_{N_k}^k\), \(k\in\mathcal K^\circ\).

\begin{lem}
	If \(\{\phi^r_k(x)\}_{r\in\mathcal N_k}\), are chosen as the Lagrange interpolating polynomials on \(\mathcal D_k\), \(k\in\mathcal K^\circ\), then the matrix \( {  {\mathfrak{\bs B}}}\) has zero row-sums. 
\end{lem}
\begin{proof}
	First we make numerous algebraic observations. Let \(\boldsymbol 1\) and \(\boldsymbol 0\) be column vectors of ones and zeros, respectively, with an appropriate length depending on the context. Using Property 1, observe that 
	\begin{align*}
        \bs M_k\boldsymbol 1 &= \left(\sum_{s=1}^{N_k} \int_{x\in\calD_k}\phi^r_k(x)\phi_s^k(x)\wrt x\right)_{r\in\mathcal N_k}\tr{}
	%
        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\sum_{s=1}^{N_k} \phi_s^k(x)\wrt x \right)_{r\in\mathcal N_k} \tr{}
        %
        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\wrt x\right)_{r\in\mathcal N_k} \tr{}
        %
        \\&= \bs P_k\boldsymbol 1,
	\end{align*}
	hence \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\). Also, 
	\begin{align*}
        \bs G_k\bs M_k^{-1}\bs P_k\boldsymbol 1 = \bs G_k\boldsymbol 1 &= \left(\sum_{s=1}^{N_k} \int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}\phi_k^s(x)\wrt x\right)_{r\in\mathcal N_k} \tr{}
	%
        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}\sum_{s=1}^{N_k} \phi_k^s(x)\wrt x \right)_{r\in\mathcal N_k} \tr{}
	%
        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}1\wrt x\right)_{r\in\mathcal N_k} \tr{}
	%
        \\&= \boldsymbol 0,
	\end{align*}
	where we have again used Property 1. Consider \(c_i>0\) and let \(\boldsymbol b\) and \(\boldsymbol d\) be arbitrary row-vectors of length \(N_k\) and \(N_{k+1}\), respectively. By Property 2, for \(k\in\mathcal K^\circ,\)
	\begin{align*}
        \bs F_i^{kk}\boldsymbol b &= -\boldsymbol \phi_k(y_{k+1})\tr{} \boldsymbol \phi_k(y_{k+1})\boldsymbol b \\&= -(\bs e^k_{N_k})\tr{} \bs e_{N_k}^k\bs b \\&= -b_{N_k}(\boldsymbol e_{N_k}^k)\tr{},
        %
        \\\bs F_i^{k,k+1}\boldsymbol d &= \boldsymbol \phi_{k}(y_{k+1})\tr{} \boldsymbol \phi_{k+1}(y_{k+1})\boldsymbol d \\&= (\bs e^{k}_{N_k})\tr{} \bs e_{1}^{k+1}\bs d \\&= d_{1}(\boldsymbol e_{N_{k}}^{k})\tr{}.
	\end{align*}
	Hence, observe that 
	\begin{align*}
		&\bs P_k^{-1}\bs F_i^{kk} \bs M_k^{-1} \bs P_k \bs 1 = \bs P_k^{-1} \bs \phi_k(y_k)'\bs 1 = -\bs P_k^{-1}\bs F_i^{k,k+1} \bs M_k^{-1} \bs P_k \bs 1 = - \bs P_k^{-1}(\bs e_{N_k}^k)',
	\end{align*}
	for \(k=0,1,\dots,K-1\). 
	Similarly, for \(c_i<0\), 
	\begin{align*}
		&\bs P_k^{-1}\bs F_i^{kk} \bs M_k^{-1} \bs P_k \bs 1 = \bs P_k^{-1} \bs \phi_k(y_{k+1})'\bs 1 = -\bs P_k^{-1}\bs F_i^{k,k-1} \bs M_k^{-1} \bs P_k \bs 1 = - \bs P_k^{-1}(\bs e_{1}^k)',
	\end{align*}
	for \(k=1,\dots,K\).

	Now, with the above observations made, we first claim that, for \(c_i>0\), and \(k=0,1,...,K-1\), 
	\begin{align*}
		\sum_{j\in\calS}\mathfrak{\bs B}^{kk}_{ij}\bs 1 + \mathfrak{\bs B}^{kk+1}_{ij}\bs 1
		&=\sum_{j\in\calS}T_{ij}\bs I_{N_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_k^{-1}\bs P_k\boldsymbol 1 
		\\&=\bs 0.
	\end{align*}
	The first sum is zero since \(\bs T\) is a generator of a continuous-time Markov chain. This leaves the other two terms, which, using our observations, we get 
	\begin{align*}
		& c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k+1}^{-1}\bs P_{k+1}\boldsymbol 1
		\\&= c_i \bs P_k^{-1}\bs F_i^{kk} \bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs G_k\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k+1}^{-1}\bs P_{k+1}\boldsymbol 1 
		\\&= \boldsymbol 0.
	\end{align*}
	For \(c_i>0\) and \(k=K\), 
	\begin{align*}
		&\sum_{j\in\calS}\mathfrak{\bs B}^{KK}_{ij}\bs 1 + \mathfrak{\bs B}^{K{K+1}}_{ij}\bs 1
		\\&= \sum_{j\in\calS}T_{ij}\bs I_{N_K}\bs 1-\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 -\sum_{j\in\calS_{{K+1} }}c_ip_{ij}^{K+1} \bs P_K^{-1}\boldsymbol \phi_K(0)\tr{}\bs 1 
		\\&\quad{} + c_i \bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K\boldsymbol 1 
		\\&= -\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 -\sum_{j\in\calS_{{K+1}}}c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\bs 1+ c_i \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\boldsymbol 1
		\\&\quad{}+ c_i \bs P_K^{-1}\bs G_K\bs M_K^{-1}\bs P_K\boldsymbol 1
		%
		\\&= - c_i \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 + c_i \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\boldsymbol 1 + c_i \bs P_K^{-1}\bs G_K\boldsymbol 1
		\\&= \bs 0.
	\end{align*}
	% Similarly, for \(c_i<0\), and row-vectors \(\boldsymbol b\) and \(\boldsymbol d\) of length \(N_k\) and \(N_{k-1}\), respectively, 
	% \begin{align*}
	% 	\bs F_i^{kk}\boldsymbol b &= \boldsymbol \phi^k(y_{k})\tr{} \boldsymbol \phi^k(y_{k})\boldsymbol b \\&= (\bs e^{k}_{1})\tr{} \bs e_{1}^{k}\bs b  \\&= b_{1}(\boldsymbol e_{1}^k)\tr{} 
	% 	%
	% 	\\\bs F_i^{k,k-1}\boldsymbol d &= -\boldsymbol \phi^{k}(y_{k})\tr{} \boldsymbol \phi^{k-1}(y_{k})\boldsymbol d \\&= -(\bs e^{k}_{1})\tr{} \bs e_{N_{k-1}}^{k-1}\bs d \\&= -d_{N_{k-1}}(\boldsymbol e_{1}^k)\tr{}.
	% 	\end{align*}

		Similarly, for \(c_i<0\) and \(k=1,...,K\), using the same arguments as before we have
	\begin{align*}
		\sum_{j\in\calS}\mathfrak{\bs B}^{kk}_{ij}\bs 1 + \mathfrak{\bs B}^{kk-1}_{ij}\bs 1
		&=\sum_{j\in\calS}T_{ij}\bs I_{N_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\bs M_k^{-1}\bs P_k\boldsymbol 1
		% \\&= \boldsymbol 0 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k-1}^{-1}\bs P_{k-1}\boldsymbol 1
		% \\&= c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\boldsymbol 1
		% \\&= c_i \bs P_k^{-1}\bs F_i^{kk}\boldsymbol 1 + c_i \bs P_k^{-1}\bs G_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\boldsymbol 1
		% \\&= c_i \bs P_k^{-1}(\boldsymbol e_{1}^k)\tr{} + \boldsymbol 0 + c_i \bs P_k^{-1}(-\boldsymbol e_{1}^k)\tr{}
		\\&= \boldsymbol 0.
	\end{align*}
	For \(c_i<0\) and \(k=K\), 
	\begin{align*}
		\sum_{j\in\calS}\mathfrak{\bs B}^{00}_{ij}\bs 1 + \mathfrak{\bs B}^{0{-1}}_{ij}\bs 1
		&= \sum_{j\in\calS}T_{ij}\bs I_{N_0}\bs 1-\sum_{j\in\calS_+} c_ip_{ij}^{-1} \bs P_0^{-1}\bs F_i^{00}\bs M_0^{-1}\bs P_0 \bs 1 -\sum_{j\in\calS_{{-1} }}c_ip_{ij}^{-1} \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{}\bs 1 
		\\&\quad{} + c_i \bs P_0^{-1}(\bs F_i^{0,0}+\bs G_0)\bs M_0^{-1}\bs P_0\boldsymbol 1 
		% \\&= -\sum_{j\in\calS_+} c_ip_{ij}^{-1} \bs P_1^{-1}\bs F_i^{11}\bs M_1^{-1}\bs P_1 \bs 1 -\sum_{j\in\calS_{{-1} }}c_ip_{ij}^{-1} \bs P_1^{-1}\bs F_i^{11}\bs M_1^{-1}\bs P_1\bs 1+ c_i \bs P_1^{-1}\bs F_i^{1,1}\bs M_1^{-1}\bs P_1\boldsymbol 1
		% \\&\quad{}+ c_i \bs P_1^{-1}\bs G_1\bs M_1^{-1}\bs P_1\boldsymbol 1
		% %
		% \\&= - c_i \bs P_1^{-1}\bs F_i^{11}\bs M_1^{-1}\bs P_1 \bs 1 + c_i \bs P_1^{-1}\bs F_i^{1,1}\bs M_1^{-1}\bs P_1\boldsymbol 1 + c_i \bs P_1^{-1}\bs G_1\boldsymbol 1
		\\&= \bs 0.
	\end{align*}

	This just leaves the boundaries. For the lower boundary,
	\begin{align*}
		\mathfrak{\bs B}^{{-1},{-1}}\boldsymbol 1 +  {  {\mathfrak{\bs B}}}^{{-1}0}\boldsymbol 1 
		&= \bs T_{{-1},{-1}}\boldsymbol 1 + \left[\bs T_{{-1}+}\otimes\left( \boldsymbol\phi_0(0)\bs M_0^{-1}\bs P_0\right)\right]\boldsymbol 1. 
		\intertext{Swapping the order of summation and recalling \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\) then this is equal to}
		\bs T_{{-1},{-1}}\boldsymbol 1 + \left[\bs T_{{-1}+}\otimes\left( \boldsymbol\phi_0(0)\bs M_0^{-1}\bs P_0\right)\boldsymbol 1\right]\boldsymbol 1 
		%
		&= \bs T_{{-1},{-1}}\boldsymbol 1 + \left[\bs T_{{-1}+}\otimes \boldsymbol e_1^0\bs 1\right]\bs 1
		\\&= \bs T_{{-1},{-1}}\boldsymbol 1 + \left[\bs T_{{-1}+}\otimes \boldsymbol 1\right]\bs 1
		\\&= \bs T_{{-1},{-1}}\boldsymbol 1 + \bs T_{{-1}+}\bs 1
		%
		\\&= \boldsymbol 0.
	\end{align*}	
	For the upper boundary,
	\begin{align*}
		\mathfrak{\bs B}^{{K+1},{K+1}}\boldsymbol 1 +  {  {\mathfrak{\bs B}}}^{{K+1} K}\boldsymbol 1 
		&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes( \boldsymbol\phi_K(b)\bs M_K^{-1}\bs P_K)]\boldsymbol 1.
		\intertext{Swapping the order of summation and recalling \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\) then this is equal to}
        \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes( \boldsymbol\phi_K(b)\bs M_K^{-1}\bs P_K)\bs 1]\boldsymbol 1 
		%
		%\\&%= T_{{K+1},{K+1}}\boldsymbol 1 + [T_{{K+1}-}\otimes \boldsymbol\phi^K(b)\bs 1]\boldsymbol 1
		&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes\bs e^K_{N_K}\bs 1]\boldsymbol 1
		\\&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes \bs e^K_{N_K}]\bs 1
		\\&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + \bs T_{{K+1}-}\bs 1
		%
		\\&= \boldsymbol 0.
	\end{align*}

	Combining all the above we have shown that the row sums of \( {  {\mathfrak B}}\) are zero. 
\end{proof}

\begin{cor}
	The DG approximation to the generator \(  \bs B\) conserves probability. That is, for all \(t\geq 0\), 
	\begin{align*}
	&\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS} \int_{x\in[0,b]}u_i(x,t)\wrt x 
	%
	\\&= \sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i\in\calS} \int_{x\in[0,b]}u_i(x,0)\wrt x.
	\end{align*}
\end{cor}
\begin{proof}
%For a basis of Lagrange polynomials the DG operator \( {  {\mathfrak B}}\) has zero row-sums, therefore 
%\begin{align*}
%	\boldsymbol \alpha(t)\boldsymbol 1 
%	%
%	=  \boldsymbol \alpha(0)\exp( {  {\mathfrak B}}t)\boldsymbol 1 
%	%
%	= \boldsymbol \alpha(0)\boldsymbol 1. 
%\end{align*}
Let \(\{\psi^r_k(x)\}_{r\in\mathcal N_k,k\in\mathcal K^\circ},\) be a basis for \(span(\phi^r_k(x),r\in\mathcal N_k,k\in\mathcal K^\circ)\), where \(\{\phi^r_k(x)\}_{r\in\mathcal N_k,k\in\mathcal K^\circ}\) are the Lagrange polynomials. Also define \(\psi_1^{-1}(x)=\delta(x)\) and \(\psi_1^{K+1}(x)=\delta(x-b)\) to capture the point masses at the boundaries. Let us use the same vector notation for the basis \(\psi^r_k(x)\) as we do for \(\phi^r_k(x)\). For \(k\in\mathcal K^\circ\), since \(\{\psi^r_k(x)\}_{r\in\mathcal N_k}\) and \(\{\phi^r_k(x)\}_{r\in\mathcal N_k}\) have the same span, then there is a matrix \(\bs V^k\) such that  \(\bs \psi_k(x)\tr{} = \bs V^k\bs \phi_k(x)\tr{}\). Trivially, this also holds for \(k={-1},{K+1}\).

Let 
\[ \bs W = \left[\begin{array}{ccc}
		\bs I_{|\calS_{-1}|} & & \\
		& \bs P &  \\
		& & \bs I_{|\calS_{K+1}|}
	\end{array}\right]
\mbox{ and }
  \bs V = \left[\begin{array}{ccccc}
		\bs I_{|\calS_{-1}|} & & & &  \\
		& \bs V^0 & & &  \\
		& & \ddots & & \\
		& & & \bs V^K \\
		& & & & \bs I_{|\calS_{K+1}|}
	\end{array}\right]. \]
For a DG approximation, \(\bs B\), constructed from the basis \(\{\psi^r_k\}_{r\in\mathcal N_k,k\in\mathcal K}\), it can be shown that \(\bs B\) is similar to \( {  {\mathfrak{\bs B}}}\) with similarity matrix, \(\bs V\bs W\), such that \(  \bs B_{ij} =  \bs V\bs W{ \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V^{-1},\,i,j\in\calS\).
Therefore 
\begin{align*} 
	\int_{x\in[0,b]}  \bs B_{ij}\bs\psi(x)\tr{} \wrt x &= \bs V\bs W { \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V^{-1}\int_{x\in[0,b]} \bs V\bs \phi(x)\tr{}\wrt x 
	%
	\\&=\bs V\bs W { \mathfrak{\bs B}}_{ij}\bs W^{-1} \bs W\bs 1 \\&= \bs V \bs W { \mathfrak{\bs B}}_{ij}\bs 1,
\end{align*}
since \(\displaystyle\int_{x\in[0,b]} \bs \phi(x)\tr{}\wrt x =\bs W\bs 1\).
The row sums of \(\mathfrak{\bs B}\) are 0, hence 
\begin{align}\label{eqn:Bsums0} 
	\int_{x\in[0,b]}  \sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\tr{} \wrt x  &=  \bs V\bs W \sum_{j\in\calS}{ \mathfrak{\bs B}}_{ij}\bs 1 \\&=  \bs V\bs W \bs 0 \\&= \bs 0.
\end{align}

Let \( {}^\psi \bs a_i(t)\), \(i\in\calS\) denote the coefficients related to the DG approximation constructed with the basis \(\{\psi^r_k\}_{r\in\mathcal N_k,k\in\mathcal K}\) (to distinguish them from \(\bs a\) and \(\bs \alpha\) used above). The DE constructed by the DG method is
\begin{align*}
	\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x) = \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x).
\end{align*}
Integrating over \(x\in [0,b]\) and summing over \(j\in\calS\) we get
\begin{align*}
	\int_{x\in[0,b]}\sum_{j\in\calS}\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x)\wrt x = \int_{x\in[0,b]}\sum_{j\in\calS}\sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x).
\end{align*}
Exchanging the order of operations gives 
\begin{align}\label{eqn:totalprobDE}
	\cfrac{\wrt}{\wrt t}\sum_{j\in\calS} \left({}^\psi \boldsymbol a_j (t)\right)\int_{x\in[0,b]} \boldsymbol \psi(x)\wrt x =  \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\int_{x\in[0,b]}\sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\wrt x = 0,
\end{align}
where the right-hand side is \(0\) due to Equation (\ref{eqn:Bsums0}). This holds for all \(t\geq 0\). The left-hand side of (\ref{eqn:totalprobDE}) is the rate of change (with respect to time) of the total mass of the system. Since this is \( 0\) for all \(t\geq 0\), there is no change in the total mass of the system and thus probability is conserved. 
%\begin{align*}
%	&\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS}\int_{x\in[0,b]} u_i(x,t)\wrt x 
%	%
%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS}\int_{x\in[0,b]}\boldsymbol a_i(t) \boldsymbol \psi(x)\wrt x 
%	%
%	\\&=\sum_{j\in\calS}\int_{x\in[0,b]}\left[\vligne{\bs q_{{-1}}(0) & \boldsymbol a(0) & \bs q_{{K+1}}(0)} \exp(  Bt)\right]_{j}\boldsymbol \psi(x)\wrt x 
%	%
%	\\&=\sum_{j\in\calS}\int_{x\in[0,b]}\left[\vligne{\bs q_{{-1}}(0) & \boldsymbol a(0) & \bs q_{{K+1}}(0)} \left(I+  Bt + \cfrac{  B^2t^2}{2!}+...\right)\right]_{j}\boldsymbol \psi(x)\wrt x
%	%
%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i,j\in\calS}\int_{x\in[0,b]}\boldsymbol a_i(0)\boldsymbol \psi(x)\wrt x
%	%
%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i\in\calS}\int_{x\in[0,b]} u_i(x,0)\wrt x.
%\end{align*}
\end{proof}
%\begin{remark}
%	For any choice of basis \(\{\phi^r_k(x)\}_{r\in\{1,...,N_k\},k\in\{1,...,K\}}\) which spans the set of polynomials of degree \(N_k-1\), the DG approximation constructed with this basis will conserve probability. 
%\end{remark}

% \section{Putting it all together}
% Recall that the ultimate goal for our DG approximation is to approximate the operator \(\mathbb B\). We have that \({  \bs B}^{k\ell}\) is an approximation to \(\mathbb B^{k\ell}\), \(k,\ell \in \{{-1},1,...,K,{K+1}\}\). 
% %\[\mu_{k,i}\mathbb B_{ii}^{kk}(\wrt x)\approx\bs a_{k,i}   B_{ii}^{kk} \bs \phi^k(x)\tr{} = \bs a_{k,i} \left[T_{ii}M_k + c_i(F_i^{kk}+G_k)\right] \bs \phi^k(x)\tr{}\wrt x. \]
% %
% %Substituting in the approximations to the flux into Equation \eqref{eqn:DG matrix} and post multiplying by \(M_k^{-1}\) and \(\bs \phi^k(x)\tr{}\) we get 
% %\begin{align*}
% %	&\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,i}(t)\bs \phi^k(x)\tr{}  
% %	%
% %	\\&= \begin{cases}\displaystyle\sum_{i\in\calS} \boldsymbol a_{k,i}(t)\left[T_{ij}  
% %	+  c_i G_kM_k^{-1} + c_i F_i^{k,k}M_k^{-1}\right] \bs \phi^k(x)\tr{} + c_i\bs a_i^{k-1}(t)F_i^{k-1,k}M_k^{-1}\bs \phi^k(x)\tr{} & c_i>0 \\
% %	%
% %	%
% %	\displaystyle\sum_{i\in\calS} \boldsymbol a_{k,i}(t)\left[T_{ij}  
% %	+  c_i G_kM_k^{-1} + c_i F_i^{k,k}M_k^{-1}\right] \bs \phi^k(x)\tr{} + c_i\bs a_i^{k+1}(t)F_i^{k+1,k}M_k^{-1}\bs \phi^k(x)\tr{} & c_i<0\\
% %	%
% %	%
% %	\displaystyle\sum_{i\in\calS} \boldsymbol a_{k,i}(t)T_{ij}\bs \phi^k(x)\tr{} &c_i=0.
% %	\end{cases}
% %\end{align*}
% %Intuitively, \(\boldsymbol a_{k,i}(t)T_{ii}\bs \phi^k(x)\tr{} +  c_i\boldsymbol a_{k,i}(t) G_kM_k^{-1}\bs \phi^k(x)\tr{} + c_i\bs a_{k,i}(t)F_i^{k,k}M_k^{-1}\bs \phi^k(x)\tr{} \) can be seen to approximate the density of \(\mu_{k,i}\mathbb B_{ii}^{kk}(\wrt x)\) -- the first term \(\boldsymbol a_{k,i}(t)T_{ii}\bs \phi^k(x)\tr{}\) represents stochastic jumps out of phase \(i\), the second term \(c_i\boldsymbol a_{k,i}(t) G_kM_k^{-1}\bs \phi^k(x)\tr{}\) represents the movement of density within cell \(\calD_k\) by moving the density between basis functions, and the last term \(c_i\bs a_{k,i}(t)F_i^{k,k}M_k^{-1}\bs \phi^k(x)\tr{}\) represents the flow of density out of the cell \(\calD_k\). Similarly \(\boldsymbol a_{k,i}(t)T_{ij}\bs \phi^k(x)\tr{}\) can be seen to approximate the density of \(\mu_{k,i}\mathbb B_{ij}^{kk}(\wrt x)\), \(i\neq j\), stochastic jumps from phase \(i\) to phase \(j\) within cell \(\calD_k\). Also \(c_i\bs a_i^{k-1}(t)F_i^{k-1,k}M_k^{-1} \bs \phi^k(x)\tr{}\) and \(c_i\bs a_i^{k+1}(t)F_i^{k+1,k}M_k^{-1}\bs \phi^k(x)\tr{} \) approximate the densities of \(\mu_i^{k-1}\mathbb B_{ij}^{k-1,k}(\wrt x)\) and \(\mu_i^{k+1}\mathbb B_{ij}^{k+1,k}(\wrt x)\), respectively.

% Given we have now truncated the space and added boundaries, let us define \(\mathcal M_{0,b}\) as the set of measures, \(\mu_i\), which admit an absolutely continuous density on \((0,b)\), may have a point mass at \(x=0\) if \(i\in\mathcal S_{-1}\), and another at \(x=b\) if \(i\in\calS_{{K+1}}\). The set \(\mathcal M_{0,b}\) is the domain of the operator \(\mathbb B\) truncated to the interval \([0,b]\) with regulated boundaries at \(x=0\) and \(x=b\). Also, redefine \(\mathcal K^m_i = \{k\in\{{-1},1,...,K,{K+1}\}\mid   l  (\calD_k\cap\calF_i^m) =    l (\calD_k) \}\) for \(i\in\calS,m\in\{+,-,0\}\) for all \(   l\in\mathcal M_{0,b} \).

### 2022-02-10 chapter1/chapter.tex
.

--- 2021-12-22 chapter1/chapter.tex
@@
 %!TEX root = ../thesis.tex
-\chapter{Introduction \& Preliminaries \label{ch:intro}} 
+\chapter{The Existing Literature \& Mathematical Preliminaries \label{ch:math}} 
+% intro to intro
+% contextual background
+	% Fluid queues 
+		% definition
+		% applications 
+		% existing results 
+	% Fluid-fluid queues
+		% definition 
+		% basic analysis in terms of generators
+		% the need for approximation and how the old work doesn't fit here
+	% The DG method
+		% pertition into cells
+		% project on to a basis
+		% problems: oscillations
+		% soutions and why they dont fit
+		% observation about constant-basis methods
+		% this thesis asks the question: motivated by the constant basis being a probability model, can we dream up a more accurate probability model to approximate a fluid queue?
+	% On the structure of the constant basis/uniformisation method (QBD)
+	% least variable is PH, so what about MEs?
+% structure of the thesis
+	% rest of chapter 1 gives mathematical preliminaries
+	% chapter 2 explains the DG method, problems/oscillations, slope limiting, and appllication to SFFMs
+	% inspired by the structure of the order-1 scheme, and to solve the negativity problem, chapter 3 introduces a new approximation method; The QBD-RAP.
+In this chapter we derive the intuitively of the model and explain it's dynamics.
+	% we then move to convergence, with chapter 4 proving a convergence of the QBD-RAP up to the first hitting time of the fluid queue on the edges of an interval.
+technical results and extensions are left to the appendix: certain properties of closing vectors, convergence without ephemeral phases, and certain matrix algebraic manipulations.
+	% chapter 5 stitches together the results of chapter 4 to prove a global result about convergence
+	% chapter 6 investigate, numerically, the performance of the DG, DG with limiter, uniformisation and QBD-RAP schemes. 
+	% chapter 7  makes concluding remarks.
+% mathematical preliminaries 
+	% CTMCs
+	% fluid queues
+	% fluid-fluid queues and operator-analytic expressions (all of them)
+		% the equations of bo2014
+		% meets the partition of the DG/QBD-RAP, and reconstructing the Fil sets 
+    % why we cant solve the equations, what we need to approximate to solve it, i.e. B, R, then D, Psi
+	% projections
+	% phase-type distributions 
+		% definition
+		% least variable property
+	% matrix exponentials
+		% definition 
+		% properties 
+		% verification that parameters give an ME
+		% CMEs
+	% QBD-RAPs, then QBDs as a special case
+		% orbit processes, their interpretation for PH and how they differ for MEs
+	% convergence theorems 
+		% DCT 
+		% Continuity of Laplace transforms
+
 % THE ROUGH IDEA & CONTEXT
@@
 % MATHEMATICAL PRELIMINARIES
-A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\).
-The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC).
-The level process, \(\{X(t)\}_{t\geq0}\), is a real-valued, continuous, and piecewise linear.
+\section{Stochastic processes}
+% \subsubsection{Random variables}
+% Let \(\mathcal S\) be a set.
 
-Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982} as a canonical application in this area), power systems \citep{hydro}, risk processes \citep{betal2005} and environmental modelling \citep{wurm2020}.
-Fluid queues are relatively well studied.
-Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods \citep{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \citep{anick1982,kk1995,beanetal2019}.
-%For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest.
-Anick Mitra Sondhi CITE, analysed fluid queues using a more direct differential equation-based method.
-Since Ramaswami's CITE initial work, there has been significant developments in the analysis of fluid queues CITE and related algorithms CITE.
+% Roughly following \cite{billingsleyconvergence}, a random variable \(Z\), is a mapping between a \emph{sample space}, \(\mathcal S\), say, to the real numbers.
 
-More recently, \cite{bo2014} extended fluid queues to \emph{so-called} stochastic fluid-fluid queues.
-In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
-The analysis,  \citep{bo2014}, is in principal similar to the matrix-analytic methods of \citep{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
-For practical computation of the results of \cite{bo2014}, a matrix-discretisation of the infinitesimal generator of the fluid queue can be used.
-To this end, to date, two possible discretisation have been suggested.
-Taking a differential equations-based approach, \cite{beanetal2019} use the discontinuous Galerkin (DG) method to discretise this operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process.
-Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations.
-It turns out that the latter approach is a sub-class of the former; the QBD can be viewed as the simplest DG scheme where the operator is projected onto a basis of piecewise constant functions.
+% An \emph{event} is a set \(\{\omega \in \mathcal S \mid Z(\omega)\in E\}\equiv \{\omega\in Z^{-1}(E)\}\) where \(E\subseteq \mathcal S\) and \(Z^{-1}\) is the inverse image of the mapping \(Z\).
+Associated with a \(Z\) is a \emph{probability measure}, \(P\), which maps events to \([0,1]\) and has the properties \(P(\emptyset)=0\), \(P(S)=1\) and if .
+The notation \(P(\{\omega\in Z^{-1}(E)\})\) is cumbersome, so we write \(\mathbb P(Z\in E)\) in its place.
 
-%A QBD can be viewed as a two-dimensional CTMC, \(\{(L(t),\varphi(t))\}_{t\geq0}\), where \(\{L(t)\}\) is the discrete level process, and \(\{\varphi(t)\}_{t\geq0}\) is the phase process.
-The level process \(\{L(t)\}\) is skip free, meaning that, given the process is at level \(L(t)=\ell\), is may only jump to \(\ell+1\) or \(\ell-1\) at jump epochs.
-The sojourn time of \(\{L(t)\}\) in a given level follows a phase-type distribution
+% The sample space, \(\mathcal S\), is the set of all possible outcomes. Associated with the random variable \(Z\) is a \emph{probability measure} which assigns
 
-In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \citep[Section 3.3]{koltai2011}, see also \citep{nodalDGBook} and references therein.
-One justification for the positivity preserving property is from the interpretation of the discretisation as a stochastic process thereby ensuring positivity.
-For higher order DG schemes there is no such interpretation.
-Moreover, higher-order DG approximation schemes may produce negative, or highly oscillatory solutions, particularly when discontinuities or steep gradients are present.
-Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \citep[Section 6.5]{nodalDGBook} and references therein).
-Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in theses regions.
-Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.
+A stochastic process is a sequence, \(\{X(t)\}_{t\in\mathcal T}\), of random variables or random vectors indexed by some index set \(\mathcal T\).
+In this thesis in the case that the index set is \(\mathcal T=\{0,1,2,...\}\) we say that the process \(\{X(n)\}_{n\in \mathcal T}\) is a discrete-time process, and we will typically use the dummy variable \(n\).
+When the index set is \(\mathcal T=[0,\infty)\) we say write \(\{X(t)\}_{t\in\mathcal T}\) is a continuous-time process, and we will typically use the dummy variable \(t\).
+We may omit the index set and write \(\{X(t)\}\) in place of \(\{X(t)\}_{t\in\mathcal T}\) when it is not explicitly needed, or we may write \(\{X(t)\}_{t\geq t_0}\) to mean \(\{X(t)\}_{t\in[t_0,\infty)}\).
+The \emph{sample space} of \(X(t)\) is the set of possible values that \(X(t)\) can take at any \(t\in\mathcal T\).
 
-\section{Fluid queues}
-A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer.
-The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we we assume to be \(\mathcal S=\{1,2,\dots,N\}\) without loss of generality, and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\).
+The \emph{initial distribution}, \(\mu\), of a stochastic process is the distribution of \(X(0)\).
+More generally, for a random variable, \(Z\), we write \(Z\sim \nu\) when \(Z\) has the distribution given by the probability measure \(\nu\).
+For the probability that \(X(t)\) lies in some set \(E\subset \mathcal S\) we write \(\mathbb P(X(t)\in E \mid X(0)\sim \mu)\).
+When \(\mu\) assigns probability 1 to a single point, \(x\in\calS\), say, we write \(\mathbb P(X(t)\in E \mid X(0)=x).\) A stochastic process is \emph{stationary} if
+\begin{align}\label{eqn: stationary}
+	&\nonumber\mathbb P(X(t_n)\in E_n,X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
+	\\&= \mathbb P(X(t_n+t)=i,X(t_{n-1}+t)\in E_{n-1},...,X(t_0+t)\in E_0),
+\end{align}
+for any \(t,t_0,...,t_n\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\).
+A random variable, \(\tau\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\tau\in\sigma(X(s), s\leq t),\) where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s), s\leq t\}\).
+
+We call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable
+\begin{align}\label{eqn: markov property}
+	\nonumber &\mathbb P(X(n+1)=j \mid X(n)=i,X(n-1)=i_{n-1},...,X(0)=i_0) 
+	\\&= \mathbb P(X(n+1)=j \mid X(n)=i) 
+\end{align}
+for all \(n\in\{0,1,2,...\}\) and \(i,i_0,...,i_{n-1},j\in\calS,\) and refer to (\ref{eqn: markov property}) as the Markov property.
+The process \(\{X(n)\}_{n\in\{0,1,2,...\}}\) is said to be time-homogeneous if
+\(\mathbb P(X(n+1)=j \mid X(n)=i)\) does not depend on \(n\).
+The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(n)\}_{n\in\{0,1,2,...\}}\), on the event that \(\{\tau<\infty\}\), then
+\begin{align}\label{eqn: strong markov property}
+	\mathbb P(X(\tau+1)=j \mid \sigma(X(0),X(1),...,X(\tau))) = \mathbb P(X(\tau+1)=j \mid X(\tau)).
+\end{align}
+
+We call \(\{X(t)\}_{t\geq 0}\) a \emph{continuous-time Markov chain} if \(\calS\) is countable and for all \(t_{n+1} > t_n > t_{n-1} > ... > t_0 > 0\) and \(i,i_0,...,i_{n-1},j\in\calS,\)
+\begin{align}\label{eqn: markov property cmtc}
+	\mathbb P(X(t_{n+1})=j \mid X(n)=i, X(n-1)=i_{n-1},...,X(0)=i_0) = \mathbb P(X(t_{n+1})=j \mid X(t_n)=i).
+\end{align}
+The process \(\{X(t)\}_{t\geq 0}\) is said to be time-homogeneous if 
+\(\mathbb P(X(t+s)=j \mid X(s)=i)\) does not depend on \(s\).
+The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(t)\}_{t\geq 0}\), on the event that \(\{\tau<\infty\}\), then
+\begin{align}\label{eqn: strong markov property ctmc}
+	\mathbb P(X(\tau + t)=j \mid \sigma(X(s),s\leq \tau)) = \mathbb P(X(\tau+t)=j \mid X(\tau)).
+\end{align}
+
+Most generally, we call \(\{X(t)\}_{t\geq 0}\) a \emph{Markov process} if for all \(t_n > t_{n-1} > ... > t_0 > 0,\, t_k\in\mathcal T,\, k=0,...,n+1,\) and \(E_0,...,E_n\subseteq \mathcal S\)
+\begin{align}\label{eqn: markov process}
+	\mathbb P(X(t_n)\in E_n,X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
+	= \mathbb P(X(t_n)\in E_n\mid X(t_{n-1})),
+\end{align}
+and is time homogeneous if \(\mathbb P(X(t+s)\in E_n\mid X(s))\) does not depend on \(s\).
+
+The Chapman-Kolmogorov equations state that 
+\begin{align}\label{eqn: CK}
+	&\nonumber \mathbb P(X(t+s)\in E\mid X(0)\sim \mu) 
+	\\&= \int_{x\in\calS}\mathbb P(X(t+s)\in E\mid X(t)=x)\mathbb P(X(t)\in \wrt x\mid X(0)\sim \mu).
+\end{align}
+% Markov processes, time-homogeneity, strong markov prperty, partitioning and LOTP
+
+\section{Some semigroup theory}
+The evolution of Markov processes can be described by an \emph{operator semigroup}.
+Furthermore, semigroups arise in other contexts related to the analysis of fluid queues and fluid-fluid queues.
+Semigroups (and therefore Markov processes) can be characterised by their \emph{infinitesimal generator}, or generator for short.
+One of the aims of this thesis is to approximate the generator of a type of Markov process called a fluid queue, which we introduce more formally in Section~\ref{sec: fqs}.
+Here we very briefly introduce semigroups and their \emph{infinitesimal generator}.
+
+\subsubsection{Operators}
+Let \(L\) be a Banach space (a vector space with a metric and where Cauchy sequences converge to a point within the space).
+An operator is a mapping which takes elements of a vector space and sends them to elements of a vector space (not necessarily the same space).
+That is, let \(V,W\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\subset V\to W\) is an operator.
+The set \(\mathcal D(\mathbb B)\) is the domain of the operator, and is not necessarily all of \(L\), when specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to W\).
+Let \(c\in \mathbb R\) and \(v\in V\), \(w\in W\).
+Then \(\mathbb B\) is a linear operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\).
+We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms attached to \(V\) and \(W\) respectively.
+An operator is said to be bounded if \(||\mathbb B||<M\) for some \(\infty > M\in \mathbb R\), for ever \(v\in V\).
+
+% For example, let \(L\) be the space of real, smooth functions for which \(f(x)\to 0\) as \(x\to \infty\), and consider the sup norm, \(||f||=\sup\limits_{x\in\mathbb R}f(x)\).
+Then differentiation is an operator which maps \(L\) to it self.
+That is \(B=\cfrac{\wrt}{\wrt x}:L\to L\).
+In addition \(B\) is linear since for any \(f,g\in L\), \(B(f+g) = \cfrac{\wrt}{\wrt x}(f(x)+g(x)) =  \cfrac{\wrt}{\wrt x}f(x) +  \cfrac{\wrt}{\wrt x}g(x)\) and \(B(cf)= \cfrac{\wrt}{\wrt x}cf(x)=c \cfrac{\wrt}{\wrt x}f(x)\).
+The operator \(B\) is, however, not bounded.
+
+% The \emph{graph} of \(\mathbb B\) is defined as the linear subspace of \(L\times L\), \(\mathcal G(B)=\{(f,Bf):f\in \mathcal D(B)\}\subset V\times W\}\).
+The operator \(B\) is said to be closed if \(\mathcal G(B)\) is a closed subset of \(V\times W\).
+That is, for every sequence \(\{f_n\}\) in \(\mathcal D(B)\) such that \(f_n\to f\) and \(Bf_n=g_n\to g \), then \(f\in\mathcal D(B)\) and \(Bf = g\).
+Note that we must have both \(\{f_n\}\) and \(\{g_n\}\) converge.
+
+% For example, let \(B=\cfrac{\wrt}{\wrt x}:C^1[0,1]\subset C^0[0,1]\to C^0[0,1]\), where \(C^n[0,1]\) is the set of all \(n\)-times continuously differentiable functions on the interval \([0,1]\).
+The graph of \(B\) is the set \(G(B)=\{(f,Bf); f\in C^1[0,1]\}\).
+The operator \(B\) is closed since for any \(f_n\in C^1[0,1]\) with \(f_n\to f\) and \(Bf_n = \cfrac{\wrt f_n}{\wrt x}=g_n\to g\), then \(f\in C^1[0,1]=\mathcal D(B)\) and \(Bf=g\).
+Notice that we can construct \(\{f_n\}\) such that \(f_n\to f\notin \mathcal D(B)\) (e.g.~from the Weierstrass Approximation Theorem there is a sequence of polynomials in \(C^1[0,1]\) converging to \(f(x)=|x-1/2|\notin C^1[0,1]\)) but in these cases \(Bf_n = \cfrac{\wrt f_n}{\wrt x}=g_n\) must not converge.
+
+\subsubsection{Semigroups}  A family of linear operators is a collection of indexed linear operators \(\{\mathbb B(i):i\in I\}\) where \(I\) is some index set.
+That is, for each \(i\in I\), \(\mathbb B(i)\) is a linear operator.
+A family of operators \(\{T(t):L\to L:t\geq 0\}\) is said to have the \emph{semigroup} property if \(\mathbb T(t+s)=\mathbb T(t)T(s)\).
+If, furthermore, \(\mathbb T(0)=I\), the identity operator, then the family \(\{\mathbb T(t):t\geq 0\}\) is known as a semigroup.
+If \(\mathbb T(t)f\to f\) as \(t\to 0^+\), then \(\{\mathbb T(t):L\to L:t\geq 0\}\) is said to be strongly continuous.
+If the \(||\mathbb T(t)||\leq 1\) for all \(t\geq 0\) then \(\{\mathbb T(t):L\to L:t\geq 0\}\) is known as a contraction semigroup.
+
+For example, for a bounded linear operator \(\mathbb B\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\), then it can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), hence \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup.
+As a more concrete example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\).
+On \(C^\omega(\mathbb R)\), \(\cfrac{\wrt}{\wrt x}\) is bounded.
+Then \(e^{\frac{\wrt}{\wrt x} t } f(x) = \sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is just the Taylor series of \(f\) about the point \(x\).
+
+\subsubsection{Infinitesimal generators} The \emph{infinitesimal generator} (or just \emph{generator} for short) of a semigroup \(\mathbb T(t):L\to L\) is the linear operator defined by 
+\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(T(t)f-I),\]
+and the domain, \(\mathcal D(\mathbb B)\) is the subspace of \(L\) for which this limit exists.
+Note that an essential part of the definition of the generator is its domain.
+For this reason it is common to denote the generator as the pair \((\mathbb B,\mathcal D(\mathbb B))\).
+Here we will denote the generator as just \(\mathbb B\) and explicitly state its domain where relevant.
+
+To determine the generator \(\mathbb B\) we can proceed by either differentiation of the semigroup, 
+\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(T(t)f-I),\]
+or integration;
+\[\int_{t=0}^\infty e^{-st}T(t)\wrt t = (sI-\mathbb B)^{-1}=:R_s.\]
+The operator \(R_s\) is known as the resolvent.
+
+% \section{Markov processes and semigroups}
+% Semigroups arise naturally in Markov processes.
+Let \(\{X(t):t\geq 0\}\) be a Markov process with state space \(\mathcal S\).
+The Markov property states that for \(s,t\geq 0\), \(X(t+s)\) is independent of \(X(u),\, 0\leq u<t\) given \(X(t)\).
+That is, \(P(X(t+s)\in\mathcal A \mid X(t), X(u), 0\leq u<t)=P(X(t+s)\in\mathcal A \mid X(t))\).
+If \(P(X(t+s)\in\mathcal A \mid X(t))=P(X(s)\in\mathcal A \mid X(0))\), that is the distributions are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.
+Where necessary, we will use the notation \(X_\mu(t)\) to denote the Markov process with initial distribution \(\mu\), i.e.~\(P(X(0)\in\cdot)=\mu(\cdot)\), and \(X_x(t)\) to be the Markov process with initial distribution \(\delta(x)\).
+Markov processes can be characterised by \emph{probability kernels}.
+
+% Let \((U,\mathcal U)\) and \((V,\mathcal V)\) be two measure spaces, then a probability kernel \(\nu\), is a mapping \(\nu:U\times \mathcal V\to \mathbb R_+\) such that \(\nu(u,\mathcal A)\) is \(\mathcal U\)-measurable in \(u\in U\) for fixed \(\mathcal A\in \mathcal V\), and a probability measure in \(\mathcal A\in \mathcal V\) for fixed \(u\in U\).
+We say that \(\nu\) is a probability kernel from \(U\) to \(V\).
+
+% With a probability kernel comes associated integral operators. For a measure \(\mu\) on \((U,\mathcal U)\), then 
+% \[\mu \nu (\mathcal A)=\int_{U} \mu(\wrt x) \nu(x,\mathcal A)\]
+% is also a measure on \((U,\mathcal U)\). For a function \(f:V\to \mathbb R\) then 
+% \[\nu f (x)=\int_{V}  \nu(x,\wrt y)f(y).\] 
+% We can also compose two transition kernels.
+Let \(\nu\) be a transition kernel from \(U\) to \(V\) and \(\eta\) be a transition kernel from \(V\) to \(W\).
+Then \(\nu\eta\) is also a probability kernel from \(U\) to \(W\);
+% \[\nu\eta(x,\mathcal A) = \int_{V}\nu(x,\wrt y)\eta(y,\mathcal A).\]
+
+% Let \(\mu_{t}(\cdot)=P(X(t)\in \cdot)\) for \(t\geq 0\) be a one-parameter family of measures on \((\mathcal S,\mathcal F)\), and \(\nu_{t}(x,\cdot)=P(X(t)\in\cdot \mid X(0)=x)\) for \(t\geq 0\) be a one-parameter family of probability kernels from \(\mathcal S\) to \(\mathcal S\).
+Then \(\mu_t\) and \(\nu_{t}\) characterise the time-homogeneous Markov process \(X(t)\).
+The probability kernel \(\nu_t\) is referred to as a transition kernel.
+The kernels \(\nu_t\) form a semigroup since, using the law of total probability and time-homogeneity,
+% \begin{align*}
+% \nu_{t+s}(x,\mathcal A)&=P(X(t+s)\in\mathcal A\mid X(0)=x)
+% \\&=\int_{y\in\mathcal S}P(X(t)\in\wrt y\mid X(0)=x)P(X(t+s)\in\mathcal A\mid X(t)=y)
+% \\&=\int_{y\in\mathcal S}P(X(t)\in\wrt y\mid X(0)=x)P(X(s)\in\mathcal A\mid X(0)=y)
+% \\&=\int_{y\in\mathcal S}\nu_t(x,\wrt y)\nu_s(y,\mathcal A)=\nu_t\nu_s(x,\mathcal A).
+% \end{align*}
+
+% We can also construct other semigroups from \(\nu_t\).
+First we need dome more definitions.
+Let \((\mathcal S,\mathcal F)\) be a measure space.
+A \emph{Markov operator on measures} is an operator \(M\) which takes finite measures on \((\mathcal S,\mathcal F)\) to finite measures on \((\mathcal S,\mathcal F)\) such that \(M\) is linear and for a finite measure \(\mu\), \(\mu M(\mathcal S)=\mu (\mathcal S)\).
+We also have the following.
+Let \((\mathcal S,\mathcal F, \mu)\) be a measure space and let \(L_1\) be the class of all \(\mu\)-integrable generalised functions on \(\mathcal S\).
+A linear operator \(P:L_1\to L_1\) is a \emph{Markov operator on densities} when; if \(f\geq 0\) almost everywhere with respect to \(\mu\), then \(fP \geq 0\) almost everywhere with respect to \(\mu\); and if \(f\geq 0\) almost everywhere with respect to \(\mu\), then \(||fP|| = ||f||\).
+% \begin{lem}
+% For a Markov operator \(M\) and measure \(\mu <<\lambda\) (absolutely continuous with respect to Lebesgue measure), if \(\mu M <<\lambda\), then \(M\) induces an almost-unique Markov operator on densities \(P\) with respect to \(\mu\).
+% \end{lem}
+% Let \((\mathcal S,\mathcal F)\) be a measure space and let \(B(\mathcal S)\) be the class of bounded measurable functions.
+An operator \(K:B(\mathcal S)\to B(\mathcal S)\) is a \emph{transition operator} when \(K\) is linear and for \(f\), \(\{f_n\}\in B(\mathcal S)\), if \(f\geq 0\) almost everywhere with respect to a measure \(\mu\), \(Kf\geq 0\) almost everywhere with respect to \(\mu\), \(K1(\mathcal S)=1(\mathcal S)\) and if \(f_n\to 0^+\), then \(Kf_n\to 0^+\).
+% \begin{lem}
+% Every transition probability kernel \(\nu\) induces a Markov operator on measures, \(\mu M(\mathcal A) = \mu \nu(\mathcal A)\), and a transition operator on functions, \(Kf(x)=\nu f(x)\).
+In the other direction, every Markov operator induces a probability kernel \(\delta(x)M(\mathcal A)=\nu(x,\mathcal A)\), and every transition operator induces a transition kernel \(\nu(x,\mathcal A)=K1(\mathcal A)(x)\).
+% \end{lem}
+
+% We can now use Markov operators and transition operators to construct semigroups from the transition kernel semigroup.
+A transition semigroup is the semigroup \(K(t)f(x) = \nu_tf(x)\).
+Similarly we can construct the Markov semigroup, \(\mu M(t)(\mathcal A) = \mu \nu_t (\mathcal A)\) for measures and \(fM(t)(\wrt y) = f\nu_t (\wrt y)\) for densities.
+
+% \subsubsection{Feller processes}
+% Feller processes are Markov processes with particularly nice transition semigroups.
+A continuous-time homogeneous Markov process \(\{X(t):t\geq 0\}\) is a Feller process when, for all \(x\in\mathcal S\)
+% \begin{align}
+% y\to x&\implies X_y(t)\xrightarrow[]{d} X_x(t),\, \forall t\geq 0,
+% \\t \to 0 &\implies X_x(t)\xrightarrow[]{P}x, 
+% \end{align}
+% where \(\xrightarrow[]{d}\) denotes convergence in distribution and \(\xrightarrow[]{P}\) denotes convergence in probability.
+It can be shown that these probabilistic properties are equivalent to conditions on the transition semigroup.
+A semigroup of linear, positive (maps positive functions to positive functions), conservative (\(K(t)1(\mathcal S) = 1(\mathcal S)\)), contractions, is a Feller semigroup if for every \(f\in C_0\) (continuous function for which \(|x|\to \infty  \implies f(x)\to 0\)), and every \(x\in \mathcal S\),
+% \begin{align*}
+% K(t)f&\in C_0,\\
+% \lim_{t\to 0^+}K(t)f(x) &= f(x).
+% \end{align*}
+% \begin{thm}
+% A Markov process is a Feller process if and only if its evolution operators form a Feller semigroup. 
+% \end{thm}
+
+% It is a fact that transition operators and Markov operators on densities are \emph{adjoint} operators.
+Therefore in many cases we can work with transition operators and carry the results over to Markov operators on densities, or \emph{vice-versa}.
+
+% \subsubsection{A long story short}
+% A time homogeneous Markov processes can be defined by a transition kernel, \(\nu_t(x,\mathcal A)=P(X(t)\in\mathcal A\mid X(0)=x)\) and an initial distribution \(\mu(\mathcal A)=P(X(0)\in\mathcal A)\).
+The operators \(\nu_t\) form a semigroup due to the Markov property and time homogeneity.
+We can construct other semigroups from \(\nu_t\) by either integrating measures on the left, \(\mu\nu_t(\mathcal A)=:\mu M(t)(\mathcal A)\) (which we will call Markov semigroups), or by integrating functions on the right, \(\nu_tf(x) =: K(t)f(x)\).
+The latter is sometimes known as the transition semigroup, \(K(t):L\to L\) where \(L\) is a Banach space.
+These semigroup operators are adjoint, so results from transition semigroups may be able to be translated to results for Markov semigroups.
+
+% Feller processes are Markov processes with nice properties.
+It can be shown that a process is a Feller process if and only if its transition semigroup is a Feller semigroup.
+A Feller semigroup is a semigroup which maps continuous functions to continuous functions and is strongly continuous, \(\lim\limits_{t\to 0^+}K(t)f(x)=f(x)\).
+
+% Under certain conditions, a semigroup, (and hence also a Feller process), can be characterised by its generator, \(B\) defined by;
+% \[Bf:=\lim_{t\to 0^+} \frac{1}{t}(K(t)-I)f,\]
+% on the set of functions \(f\) for which this limit exists.
+This set is known as the domain of the generator \(\mathcal D(B)\), which is not necessarily all of \(L\).
+The domain of \(B\) is a very important part of its definition, and some authors highlight this by writing the generator as the pair \((B,\mathcal D(B))\).
+
+% On \(\mathcal D\left(\bigcup\limits_{n=0}^\infty B^n\right)\) the semigroup with generator \(B\) is given by the operator exponential 
+% \[e^{Bt}=\sum_{k=0}^\infty \frac{t^kB^k}{k!} \mbox{ for }t\geq 0.\]
+
+% It may be surprising that the semigroup \(K(t)\), which is defined on all of \(L\), can be characterised entirely in term of its generator \(B\), which is only defined on \(\mathcal D(B)\) and furthermore, only defines \(K(t)\) in terms of an operator exponential on \(\mathcal D\left(\bigcup\limits_{n=0}^\infty B^n\right)\).
+A step in the direction of characterisation of \(K(t)\) in terms of its generator is the fact that, when \(\{K(t)\}\) is strongly continuous, \(\mathcal D(B)\) is dense in \(L\).
+
+% The Hille-Yosida Theorem gives necessary and sufficient conditions for \((B,\mathcal D(B))\) to be the generator of a strongly continuous semigroup.
+A key element in the proof of the Hille-Yosida Theorem is the Yosida approximation given by
+% \[B_\lambda = \lambda B(\lambda - B)^{-1},\]
+% where \((\lambda - B)^{-1}\) is the resolvent of \(B\).
+It can be shown that the domain of \((\lambda - B)^{-1}=L\) for every \(\lambda >0\), that \(\{e^{tB_\lambda}\}\) is a strongly continuous semigroup of \(L\), and that \(B_\lambda \to B\) as \(\lambda \to \infty\).
+Hille and Yosida were able to use these facts to determine necessary and sufficient conditions in term of \(B\), for \((B,\mathcal D(B))\) to generate a strongly continuous semigroup.
+Furthermore, since \(R_\lambda\) is one-to-one with domain \(L\) and range \(\mathcal D(B)\), then there is a correspondence between elements in \(\mathcal D(B)\) and \(L\).
+That is, we can represent any element of \(\mathcal D(B)\) as \(R_\lambda f\) for some \(f\in L\).
+This can be useful to characterise the domain, \(\mathcal D(B)\).
+
+\section{Fluid queues}\label{sec: fqs}
+An unbounded fluid queue is a two-dimensional stochastic process \(\{\ddot{\bs X}(t)\} = \{(\ddot X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{\ddot X(t)\}_{t\geq0}\) is known as the level process or buffer.
+The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\) without loss of generality, and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\).
 We assume that \(\bs T\) is \emph{conservative}.
@@
 
-Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{i\in\calS\mid c_i<0\}\) and \(\calS_0 = \{i\in\calS\mid c_i=0\}\).
+Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{i\in\calS\mid c_i<0\}\), \(\calS_0 = \{i\in\calS\mid c_i=0\}\), \(\mathcal S_{-1}=\{i\in\calS\mid c_i\leq0\},\, \mathcal S_{K+1}=\{i\in\calS\mid c_i\geq0\}\).
 We assume, without loss of generality, that the generator \(\bs T\) is partitioned into sub-matrices
 \[\bs T = \left[\begin{array}{ccc}\bs T_{++} & \bs T_{+-} & \bs T_{+0} \\ \bs T_{-+} & \bs T_{--} & \bs T_{-0} \\ \bs T_{0+} & \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
-where \(\bs T_{mn} = [T_{ij}]_{i\in\mathcal S_m, j\in\mathcal S_n}\), \(m,n\in\{+,-,0\}\). Also define the diagonal matrices 
+where \(\bs T_{mn} = [T_{ij}]_{i\in\mathcal S_m, j\in\mathcal S_n}\), \(m,n\in\{+,-,0\}\).\footnote{Let's clarify some notation.
+We use the notation \(\bs u = (u_h)_{h\in \mathcal H}\) to denote a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some index set.
+Similarly, \(\bs u = (\bs u_h)_{h\in\mathcal H}\), is a row-vector defined by a collection of row-vectors \(\bs u_h\).
+The notation \(\bs u_m=(u_h)_{h\in\mathcal H_m}\) refers to the vector containing the subset of elements corresponding to \(\mathcal H_m\subseteq \mathcal H\).
+When the index set is empty, the resulting vector \(\bs u_m\) is a vector of dimension 0.
+In cases when there are two indices, we order the elements of the vector according to the first index, then the second; i.e.~\(\bs u = (u_{g}^h)_{g\in\mathcal G,h\in\mathcal H} = ((u_g^h)_{g\in\mathcal G})_{h\in\mathcal H}\).
+Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\).
+The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).} Also define the diagonal matrices
 \begin{align*}
@@
 
-When no boundary conditions are imposed, the level process is given by 
-\[X(t) = X(0) + \int_{s=0}^t c_{\varphi(s)}\wrt s.\]
-Sample paths of \(\{X(t)\}\) are continuous and piecewise linear, with \(\cfrac{\wrt }{\wrt t} X(t) = c_\varphi(t)\), when \(X(t)\) is differentiable.
-Given sample paths of \(\{\varphi(t)\}\), then \(\{X(t)\}\) is deterministic, and in this sense, \(\{\varphi(t)\}\) is the only stochastic element of the fluid queue.
+The level process is given by 
+\[\ddot X(t) = \ddot X(0) + \int_{s=0}^t c_{\varphi(s)}\wrt s.\]
+Sample paths of \(\{\ddot X (t)\}\) are continuous and piecewise linear, with \(\cfrac{\wrt }{\wrt t} \ddot X(t) = c_\varphi(t)\), when \(\ddot X(t)\) is differentiable.
+Given sample paths of \(\{\varphi(t)\}\), then \(\{\ddot X(t)\}\) is deterministic, and in this sense, \(\{\varphi(t)\}\) is the only stochastic element of the fluid queue.
 
 Often, boundary conditions are imposed.
+We denote a fluid queue bounded below at \(0\) and unbounded above by \(\{\dot{\bs X}(t)\} = \{(\dot X(t),\varphi(t))\}_{t\geq0}\), and a fluid queue bounded below at \(0\) and above by \(b<\infty\) by \(\{{\bs X}(t)\} = \{( X(t),\varphi(t))\}_{t\geq0}\).
 Here, we consider a mixture of \emph{regulated} and \emph{reflecting} boundary conditions.
@@
 In summary, the evolution of the level can be expressed as 
-\[\cfrac{\wrt}{\wrt t} X(t) = \begin{cases} c_{\varphi(t)}, & \mbox{ if } X(t)>0, \\ \max\{0,c_{\varphi(t)}\}, & \mbox{ if } X(t)=0, \\ \min\{0,c_{\varphi(t)}\}, & \mbox{ if } X(t)=M.  \end{cases}\]
+\[\cfrac{\wrt}{\wrt t} X(t) = \begin{cases} c_{\varphi(t)}, & \mbox{ if } X(t)>0, \\ \max\{0,c_{\varphi(t)}\}, & \mbox{ if } X(t)=0, \\ \min\{0,c_{\varphi(t)}\}, & \mbox{ if } X(t)=b.  \end{cases}\]
 
-Let \(\bs f(x,t) = \vligne{f_i(x,t)}_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i)\), assuming it exists.
+Let \(\bs f(x,t) = (f_i(x,t))_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i)\), assuming it exists.
 When a differentiable density exists, the system of partial differential equation which describes the evolution of the densities \(\bs f(x,t)\) is
-\begin{equation}
-	\cfrac{\partial}{\partial t} \bs f(x,t) = \bs f(x,t)\bs T - \cfrac{\partial}{\partial x}\bs f(x,t)\bs C.\label{eqn: pde}
-\end{equation}
+\begin{align}
+	\cfrac{\partial}{\partial t} \bs f(x,t) = \bs f(x,t)\bs T - \cfrac{\partial}{\partial x}\bs f(x,t)\bs C,\label{eqn: pde}
+\end{align}
+on the interior \(x\in(0,b)\), with appropriate boundary conditions (see Section~\ref{subsec: boundary DG}).
 The initial condition is the initial distribution of the fluid queue, \(f_i(x,0)\).
 Often a differentiable density function does not exist and therefore the partial differential equation (\ref{eqn: pde}) is not well-defined.
-For example, for a fluid queue with a regulated boundary, if the initial distribution of the fluid queue is a point mass at any point \(x_0\geq 0\) and in phase \(i\in\calS_+\cup\calS_0\), then a density function \(f_i(x,t)\) will not exist for any finite \(t\).
+For example, for a fluid queue with no upper boundary, if the initial distribution of the fluid queue is a point mass at any point \(x_0\geq 0\) and in phase \(i\in\calS_+\cup\calS_0\), then a density function \(f_i(x,t)\) will not exist for any finite \(t\).
 Specifically, a point mass will persist along the ray \(x_0+c_it\), \(t\geq 0\).
@@
 A weak solution satisfies
-\begin{equation}
-	-\int_x\int_t \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= \int_{x}\int_t\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x + \int_x\int_t\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
-\end{equation}
-for every vector of test functions, \(\bs \psi(x,t)\), wich are smooth and have compact support.
+\begin{align}
+	-\int_{x=0}^b\int_{t=0}^\infty \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= &\int_{x=0}^b\int_{t=0}^\infty\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x \nonumber 
+	\\&{} + \int_{x=0}^b \int_{t=0}^\infty\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs \psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
+\end{align}
+for every row-vector of test functions, \(\bs \psi(x,t) = ( \psi_i(x,t))_{i\in\calS}\), which are smooth and have compact support and \(\bs \psi(x,0)=\bs \psi(0,t) = \bs \psi(b,t)=\bs 0\).
 
-Boundary conditions may also be imposed on (\ref{eqn: pde}).
+\subsection{Transient analysis of fluid queues}
+The transient analysis of fluid queues has been relatively well studied (see, for example, \cite{ar2004,bean2005,dasilva2005,bean2009} which are matrix-analytics-methods-based, and also \cite{rs2003} which is differential-equations-based and treats a slightly more complex model than the ones considered here, but is none-the-less relevant).
+Given the interest in discontinuous and point-mass initial conditions, transient analysis of fluid queues rarely relies on solving governing differential equations numerically, although it is quite possible to do so.
+Instead, expressions for transient distributions of fluid queues are derived in terms of Laplace transforms with respect to time, and/or moments of the distributions are derived \cite{ar2004,bean2005,bean2009}.
+Moreover, these techniques often lead to quantities which have direct probabilistic interpretations which further aids in their analysis and also applications to other problems \cite{ar2003,dasilva2005,bnp2018} (we also leverage these stochastic interpretations in Chapter~\ref{sec: conv} to derive expressions for certain the Laplace transforms of the fluid queue).
+In some context, the actual transient distributions are not required, and Laplace transforms or moments are all that are required.
+In other cases, we can invert the Laplace transforms using known methods (\cite{aw2006}, or in the case of the discontinuities we might prefer \cite{hhat2020}, which uses the same concentrated matrix exponential distribution that we do).
 
-Discretisation methods approximate the operator on the right-hand side of (\ref{eqn: pde}) by a matrix, in our case, by the generator of a QBD-RAP.
+One such analysis of fluid queues derives the Laplace transform of the time taken for the level of an unbounded fluid queue to return to its initial level in a certain phase, given it started in a phase with positive rate \cite{bean2005}.
+The principles underlying the derivation of this first return operator are the same as those applied in \cite{bo2014} to derive the first return operator for fluid-fluid queues.
+Moreover, certain matrices appearing in the analysis have a stochastic interpretation which we leverage and write down certain Laplace transforms in Chapter~\ref{sec: conv} in terms of these matrices.
+Given its relevance we briefly recount the analysis of \cite{bean2005} here.
 
-\section{Matrix-exponential distributions}
+Consider an unbounded fluid queue \((\{\ddot X(t), \varphi(t)\})\) as described in \ref{sec: fqs}.
+Let \(\zeta_X(E)\) be the random variable which is the first hitting time of \(\{\ddot X(t)\}\) on the set \(E\) and define the matrix \(\bs \Psi_X(s)\) with elements \(\left[\bs \Psi_X(s)\right]_{ij}\) where \(i\in\calS_+\), \(j\in\calS_-\),
+\begin{align}
+	\left[\bs \Psi_X(\lambda)\right]_{ij} = \mathbb E\left[ e^{-\zeta_X(z)\lambda} 1(\zeta_X(z)<\infty, \varphi(\zeta_X(z)=j)\mid \ddot X(0)=z, \varphi(0)=i\right],
+\end{align}
+i.e.~the Laplace-Stieltjes transform of the time taken for the fluid queue to first return to level \(z\) and do so in phase \(j\), given it started at level \(z\) in phase \(i\).
+Define $\beta_X(t) := \int_0^t \left| c_{\varphi(z)} \right|  \wrt z$, which is the total amount of fluid to flow in or out of the buffer \(\{\ddot X(t)\}\) by time \(t\), and $\eta_X(y) := \inf \{t > 0: \beta_X(t) = y\}$ as the first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\).
+Further, let \(\bs H(\lambda,y)\) be the matrix with elements \(\left[\hat{\bs{\Delta}}^y(\lambda)\right]_{ij}\), \(i,j\in\calS_+\cup\calS_-\), given by
+\begin{align}
+	\mathbb E\left[e^{-\lambda \eta_X(y)}1(\eta_X(y)<\infty,\varphi(\eta_X(y))=j)\mid \ddot X(0)=0, \varphi(0)=i\right],
+\end{align}
+which is the Laplace-Stieltjes transform of the time taken for \(y\) amount of fluid to flow in or out of the fluid queue and to be in phase \(j\) at this time, given the initial level of the fluid queue was \(0\) and the initial phase was \(i\).
+
+It turns out that, for fixed \(\lambda \geq 0\), \(\bs H (\lambda,y)\) is a semigroup (with variable \(y\)) \cite{bean2005}.
+As such, \cite{bean2005} find the infinitesimal generator, \(\bs Q(\lambda)\) of \(\bs H(\lambda,y)\) and can therefore write \(\bs H(\lambda,y)=e^{\bs Q(\lambda)y}\).
+The derivation of the generator \(\bs Q(\lambda)=[Q_{ij}(\lambda)]_{ij\in\calS_+\cup\calS_-}\) is based on a direct analysis of sample paths of the fluid queue over an infinitesimal time, \(u\) say, and leveraging the fact that complex sample paths occur with probability \(\mathcal O(u^2)\).
+As a result, there are only three types of sample paths to consider.
+Further, \cite{bean2005} similarly argue that the generator \(\bs Q(\lambda)\) can be partitioned into blocks \(\bs Q_{mn}(\lambda) = \left[Q_{ij}(\lambda)\right]_{i\in\calS_m,j\in\calS_n}\), \(m\in\{+,-\},\,n\in\{+,-,0\}\) where
+\begin{align*}
+	\bs Q_{+0}(\lambda) &= \bs C_+^{-1}\bs T_{+0}\left[\lambda \bs I - \bs T_{00}\right]^{-1},
+	%
+	\\\bs Q_{-0}(\lambda) &= \bs C_-^{-1}\bs T_{-0}\left[\lambda \bs I - \bs T_{00}\right]^{-1},
+	%
+	\\\bs Q_{++}(\lambda) &= \bs C_+^{-1} \left(\bs T_{++} - \lambda \bs I + \bs T_{+0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0+}\right),
+	%
+	\\\bs Q_{+-}(\lambda) &= \bs C_+^{-1} \left(\bs T_{+-} + \bs T_{+0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0-} \right) ,
+	%
+	\\\bs Q_{--}(\lambda) &= \bs C_-^{-1} \left(\bs T_{--}  - \lambda \bs I + \bs T_{-0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0-}\right),
+	%
+	\\\bs Q_{-+}(\lambda) &=\bs C_-^{-1} \left(\bs T_{-+}+ \bs T_{-0}\left[\lambda \bs I - \bs T_{00}\right]^{-1}\bs T_{0+}\right).
+\end{align*}
+The functions 
+\begin{align}
+	\bs H^{++}(\lambda,y)&= \left[h_{ij}^{++}(\lambda,y)\right]_{i\in \mathcal S_+,j\in\mathcal S_{+}\cup\calS_{+0}} := e^{\bs{Q}_{++}(\lambda)y}\vligne{\bs C_+^{-1} & \bs Q_{+0}(\lambda)},  \label{eqn: lst 1a}
+	\\\bs H^{--}(\lambda,y) &= \left[h_{ij}^{--}(\lambda,y)\right]_{i\in\mathcal S_-,j\in\mathcal S_{-}\cup\calS_{-0}}:= e^{\bs{Q}_{--}(\lambda)y}\vligne{\bs C_-^{-1} & \bs Q_{-0}(\lambda)},
+	\\\bs H^{+-}(\lambda,y)  &= \left[h_{ij}^{+-}(\lambda,y)\right]_{i\in\mathcal S_+,\,j\in\mathcal S_-}:= e^{\bs{Q}_{++}(\lambda)y}\bs{Q}_{+-}(\lambda), 
+	\\\bs H^{-+}(\lambda,y)&= \left[h_{ij}^{-+}(\lambda,y)\right]_{i\in\mathcal S_-,\, j\in\mathcal S_+} := e^{\bs{Q}_{--}(\lambda)y}\bs{Q}_{-+}(\lambda), \label{eqn: lst 4a}
+\end{align}
+for \(y,\lambda\geq 0\) have the following stochastic interpretations.
+The function \(h_{ij}^{++}(\lambda,y)\) (\(h_{ij}^{--}(\lambda,y)\)) is the Laplace transform with respect to time of the time taken for the fluid level to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), given the phase was initially \(i\in\mathcal S_+\) (\(i\in\mathcal S_-\)) \citep{bean2005}.
+The function \(h_{ij}^{+-}(\lambda,y)\) (\(h_{ij}^{-+}(\lambda,y)\)) is the Laplace transform with respect to time of the time taken for the fluid level, \(\{y(t)\}\) to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), after which time the phase instantaneously changes to \(j\in\mathcal S_-\) (\(\mathcal S_+\)), given the phase was initially \(i\in\mathcal S_+\) (\(\mathcal S_-\)) \citep{bean2005}.
+
+Thus, \cite{bean2005} are able to characterise by the above matrix expressions, segments of sample paths of the fluid queue where the fluid level is non-decreasing and non-increasing.
+\cite{bean2005} then partition the sample paths which contribute to \(\Psi_X\) as those which either, (a), have a single transition from \(\calS_+\) to \(\calS_-\) (perhaps via \(\calS_0\)), and, (b), those which have more than one transition from \(\calS_+\) to \(\calS_-\) (perhaps via \(\calS_0\)).
+An expression for Laplace-Stieltjes transform of (a) is
+\[\int_{y=0}^\infty e^{\bs Q_{++}(\lambda)y}\bs Q_{+-}(\lambda)e^{\bs Q_{--}y}\wrt y.\]
+For the sample paths (b), there must be at least on point at which the phase process transitions from \(\calS_-\) to \(\calS_+\) (perhaps via \(\calS_0\)) before the first return time.
+Further, one of the transitions from \(\calS_-\) to \(\calS_+\) (perhaps via \(\calS_0\)) must occur lower than all others.
+By considering the lowest level, \(y\), at which a phase transitions from \(\calS_-\) to \(\calS_+\) (perhaps via \(\calS_0\)) occurs, \cite{bean2005} characterise the Laplace transform of the paths (b) by the expression
+\begin{align}
+	\int_{y=0}^\infty e^{\bs Q_{++}(\lambda)y} \bs \Psi_x(\lambda)\bs Q_{-+}(\lambda)\bs \Psi_X(\lambda) e^{\bs Q_{--}(\lambda)y}\wrt y.
+\end{align}
+
+Adding the expressions for (a) and (b) together, then \cite{bean2005} state 
+\begin{align}
+	\bs \Psi_X(\lambda) = \int_{y=0}^\infty e^{\bs Q_{++}(\lambda)y}\bs Q_{+-}(\lambda)e^{\bs Q_{--}y} + e^{\bs Q_{++}(\lambda)y} \bs \Psi_x(\lambda)\bs Q_{-+}(\lambda)\bs \Psi_X(\lambda) e^{\bs Q_{--}(\lambda)y}\wrt y,
+\end{align}
+which can be shown to be equivalent to \citep[Lemma~3 and Theorem~9.2]{br1997}
+\begin{align}
+	\bs Q_{++}(\lambda)\bs \Psi_X(\lambda) + \bs \Psi_X(\lambda)\bs Q_{--}(\lambda)\bs \Psi_X(\lambda) + \bs Q_{+-}(\lambda) + \bs \Psi_X(\lambda)\bs Q_{-+}(\lambda)\bs \Psi_X(\lambda) = \bs 0.
+\end{align}
+
+The key concepts of this argument are; (1) to partition the state space of the driving process into sets on which the fluid level is increasing, decreasing, or constant; (2) to characterise the sections of sample paths of the fluid level which are non-decreasing and non-increasing as semigroups and derive their generators; (3) to partition the sample paths which comprise \(\bs \Psi_X(\lambda)\) such that we can write down an expression for \(\bs \Psi_X(\lambda)\) in terms of the expressions derived in (2) and \(\bs \Psi_X(\lambda)\) and solve the resulting expression.
+
+% DO THE \(\bs Q(s)\) equation for \(\Psi(s)\); this is all possible because we have \(\bs T\) and the process can be partitioned in to up down sections.
+For FFQs the same can be done.
+The first step towards this is finding the generator of the fluid queue and partitioning it appropriately, which is what we do next.
+
+% Fluid models (with or without boundaries) are Feller processes, hence their associated semigroups are Feller semigroups.
+Let \(\boldsymbol f(x,t)\) be the density of a fluid model at time \(t\).
+Denote the Markov semigroup associated with the fluid model as \(V(t)\).
+Therefore, \(\boldsymbol f(x,t)\wrt x=\boldsymbol \mu V(t)(\wrt x)\), where \(\boldsymbol \mu(\wrt x)=\boldsymbol f(x,0)\wrt x\) is the initial probability density.
+
+% Let \(L=C(\{1,2,...,N\}\times[-\infty,\infty])\) be the space of bounded continuous functions on \(\{1,2,...,N\}\times[-\infty,\infty]\) equipped with the sup norm, \(||\boldsymbol f||_\infty = \sup\limits_{x\in[-\infty,\infty],i\in\mathcal S}|f_i(x)|\).
+Continuity at \(\infty\) means that \(\lim\limits_{x\to \infty} f(x)=\lim\limits_{x\to -\infty} f(x)\) and the limit exists and is finite.
+That is, for \(\boldsymbol f\in L\), \(\boldsymbol f(x)=(f_1(x),\dots,f_N(x))\), where each \(f_i(x)\) is a bounded continuous function on the extended reals.
+Denote by \(V^*(t):L\to L\) the transition semigroup of a SFM: \(V^*(t)\boldsymbol f(x) = \int_{y}P(X(t)\in\wrt y, \varphi(t)=k\mid X(0)=x,\varphi(0)=i)f_j(y)\).
+The superscript \(^*\) alludes to the fact that \(V^*(t)\) is the adjoint of \(V(t)\).
+% \begin{lem}\label{lem: adjoint generator}
+% For an unbounded SFM the generator of the semigroup \(V^*(t)\) is 
+% \[A = T+\cfrac{\wrt}{\wrt x}C.\]
+% \end{lem}
+%The weak form, (\ref{eqn: weak pde}) is derived by multiplying (\ref{eqn: pde}) by \(\bs \psi(x,t)\), then integrating the last term by parts and assuming \( \psi(x,t)
+
+%%
+%%
+% On the analysis fluid queues, MAM and why the methods dont apply here.
+
+\section{Fluid-fluid Queues}
+	\label{sec:prelim}
+	\begin{center}
+		\begin{minipage}{0.8\textwidth}
+			\textit{This subsection was largely taken from Sections~2 and 3 of \cite{blnos2022} with changes, such as notations, so that this chapter is consistent with the rest of the thesis.
+I am a co-author of the paper \cite{blnos2022}.
+%The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara.
+I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme.
+I contributed Sections~4.4 and 5.1.
+I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6.
+Appendix~A is also my original work.
+I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
+			}
+		\end{minipage}
+		\end{center}
+
+An unbounded stochastic fluid-fluid queue \citep{bo2013} is a Markov process with three elements, $\{(\ddot{X}(t), \ddot{Y}(t), \varphi(t))\}_{t \geq 0},$ where $\{(X(t),\varphi(t))\}_{t\geq 0}$ is a classical fluid queue and $\ddot{Y}(t)$ is the second fluid, which varies at rate $r_{\varphi(t)}(\ddot{X}(t))$:
+% 
+\begin{align*} 
+	\ddot{Y}(t) := \ddot Y(0)) + \int_0^t r_{\varphi(s)}(\ddot{X}(s)) \wrt s.
+\end{align*} 
+% 
+Regulated boundaries may also be included for the second fluid level.
+In this thesis we consider the second fluid to have a regulated boundary at \(y=0\) and unbounded above.
+To distinguish between unbounded and bounded processes, we use the notation \(\ddot Y(t)\) to denote unbounded processes and \(\dot Y(t)\) to denote a second fluid level with a regulated lower boundary at 0.
+
+% As classic fluid processes, $\{(\widehat X(t), \varphi(t)),t \geq 0\}$, or unbounded analogues, are used extensively in many areas, such as insurance and environmental modelling, it is clear that stochastic fluid-fluid models have an even wider range of applicability.
+
+In the following, we assume that $\dot Y(t) \in [0,\infty)$ and that there is a boundary at level $0$ for both the first and second fluid levels and the first fluid level is also bounded above at \(b\):
+% 
+	\begin{align*} 
+		\frac{\wrt}{\wrt t}  X(t) & := \max\{0, c_i\} \quad \mbox{if }  X(t) = 0 \mbox{ and } \varphi(t) = i, \\
+		\frac{\wrt}{\wrt t}  X(t) & := \min\{0, c_i\} \quad \mbox{if }  X(t) = b \mbox{ and } \varphi(t) = i, \\
+          	\frac{\wrt}{\wrt t} \dot Y(t) & := \max\{0, r_i(x)\} \quad \mbox{if } \dot Y(t) = 0,\, X(t) = x \mbox{ and } \varphi(t) = i, 	
+	\end{align*} 
+	% 
+for $i \in \mathcal{S} = \{1,...,N_\calS\}$. Let $\bs R(x) := \diag(r_i(x))_{i \in \mathcal{S}}$ be the diagonal fluid-rate matrix of functions for $\{\dot Y(t)\}$.
+
+For the remainder of this section, we summarise the findings of~\cite{bo2014} on the joint stationary distribution of $\{( \dot Y(t), X(t), \varphi(t))\}_{t \geq 0}$.
+The derivation of the stationary distribution relies on obtaining the operator \(\mathbb \Psi\) which gives the distribution of the process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to the level 0, given \(\dot Y(0)=0\).
+
+The concepts leading to the derivation of \( {\mathbb \Psi}\) for the fluid-fluid queue are much the same as the derivation of the analogous quantity for a classical fluid queue.
+For a classical fluid queue the infinitesimal generator of the driving process is the matrix \(\bs T\).
+We partition \(\calS\) into the sets \(\calS_+,\,\calS_-,\, \calS_0\) according to whether \(c_i\) is positive, negative or zero, respectively, and partition \(\bs T\) similarly;
+\[\bs T = \left[\begin{array}{ccc}\bs T_{++} & \bs T_{+-} & \bs T_{+0} \\ \bs T_{-+} & \bs T_{--} & \bs T_{-0} \\ \bs T_{0+} & \bs T_{0-} & \bs T_{00}  \end{array}\right].\]
+Using the block matrices \(\bs T_{mn},\,m,n\in\{+,-,0\}\), and the rates \(c_i\), we can write down the distribution of the phase on sample paths where \(\{X(t)\}\) is increasing or decreasing.
+Further, we can also determine the distribution of the phase process at the time when the fluid level goes from increasing to decreasing or decreasing to increasing.
+Ultimately, we are able to derive a matrix-Riccati equation where the solution is the unique matrix with \((i,j)\)th entry, \(i\in\calS_+\), \(j\in\calS_-\), which is the probability that \(\{X(t)\}\) first returns to zero in phase \(j\) given it started at zero in phase \(i\).
+That is, the solution of the Riccati equation gives us the distribution of the driving process at the time when the level process first returns to \(0\).
+
+For the fluid-fluid queue we can use analogous arguments.
+First, we need the infinitesimal generator of the driving process, \(\{(X(t),\varphi(t))\}\), which we denote by \(\mathbb B\).
+Then we need to partition \(\mathbb B\) on sets for which \(\{\dot Y(t)\}\) is increasing, decreasing or constant -- these are the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\).
+We can then derive an operator-Riccati equation for the first return operator for the fluid-fluid queue, \(\mathbb \Psi\), in terms of the partitioned generator \(\mathbb B\), the rates \(r_i(x)\).
+The solution to the operator-Riccati equation gives the distribution of the driving process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to \(0\).
+
+The problem we have is to solve the operator-Riccati equation -- something which is only possibly in the very simplest of cases.
+This is where the methods in this thesis comes in.
+Here, we approximate \(\mathbb B\) by a finite-dimensional matrix, which we then partition according to the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\).
+We then substitute the resulting matrices into the operator-Riccati equation and this gives us a matrix-Riccati equation, which we can then solve using known methods.
+
+We use finite-element methods for the approximation of \(\mathbb B\).
+A key element of finite-element schemes is the partition of the state space of \(\{(X(t),\varphi(t))\}\) into sets of the form \(([y_k,y_{k+1}],i)\), where \([y_k,y_{k+1}]=:\calD_k\) are known as cells.
+We must choose the cells wisely so that the partition into the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\) can be recovered from the partition into cells.
+This is not terribly difficult to do, but does require some notation and explanation.
+We now proceed to introduce the work of \cite{bo2014} and explain how we can recover the partition into the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\) from the cells as determined by the finite-element approximation.
+
+For each Markovian state $i \in \mathcal{S}$, we partition the state space of \( X(t)\), \([0,\infty)\), according to the rates of change $r_i(\cdot)$ for the second fluid $\{\dot Y(t)\}$: $[0,\infty) := \mathcal{F}^{+}_i \cup \mathcal{F}^{-}_i \cup \mathcal{F}^{0}_i,$
+where 
+% 
+		\begin{align} 
+%				\label{eqn:Fi1}
+			\mathcal{F}^{+}_i & := \{u \in \mathcal{F} : r_i(u) > 0\},  \;
+%				\label{eqn:Fi2}
+			\mathcal{F}^{-}_i :=  \{u \in \mathcal{F}:  r_i(u) < 0\}, \;
+%				\label{eqn:Fi3}
+			\mathcal{F}^{0}_i := \{u \in \mathcal{F}: r_i(u) = 0\}.\label{eqn:fil}
+		\end{align} 
+% 
+For all $i \in \mathcal{S}$, the functions $r_i(\cdot)$ are assumed to be sufficiently well-behaved that $\mathcal{F}^{m}_i$, $m \in \{+, -, 0\}$, is a finite union of intervals and isolated points.
+
+We assume that the process $\{(\dot Y(t), X(t), \varphi(t))\}_{t\geq 0}$ is positive recurrent, in order to guarantee the existence of the joint stationary density. Define stationary operators 
+\begin{align} 
+		\label{eqn:jointpi} 
+		\bbpi_i(y)(\mathcal{A}) & := \lim_{t \rightarrow \infty} \frac{\partial}{\partial y} \mathbb{P}\left( \dot Y(t) \leq y,  X(t) \in \mathcal{A}, \varphi(t) = i\right), \quad y>0,\\
+% 
+		\label{eqn:jointmass}
+		{\mathbb p}_i(\mathcal{A}) & := \lim_{t \rightarrow \infty}  \mathbb{P}( \dot Y(t) = 0,  X(t) \in \mathcal{A}, \varphi(t) = i]),
+	\end{align} 
+	where $\mathcal{A} \subset [0,\infty)$.
+Then let ${{\bbpi}}(y) = ( \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint stationary density operators and ${\mathbb{p}} = ( {\mathbb p}_i)_{i \in \mathcal{S}}$ a vector containing the joint stationary mass operators.
+The determination of ${{\bbpi}}(y)$ involves two important matrices of operators, $ {\mathbb D}$ and $ {\mathbb \Psi}$ which we now introduce.
+%The operator \( {\mathbb B}\) is the \textit{infinitesimal generator} of the process \(\{(X(t),\varphi(t))\}\).
+%Intuitively, for a set $\mathcal{A} \subset \mathcal{F}$ and a measure vector $\boldsymbol{\mu} = (\mu_i)_{i \in \mathcal{S}}$, $\bs{\mu}V(t)(\mathcal{A})$, where \(V(t)\) is a semigroup with infinitesimal generator \(B\), gives the conditional probability of $X(t) \in \mathcal{A}$,
+% The operator \( {\mathbb \Psi}\) is such that ${\boldsymbol{\mu}}^{+} {\mathbb \Psi} (\mathcal{A})$ gives the conditional probability of $\{\dot Y(t)\}$ returning to level zero and doing so when $X(t) \in \mathcal{A}$, given that the initial distribution of \((X(0),\varphi(0))\) in the set \(\bigcup_{i\in\calS}(\mathcal F_i^+,i)\), is ${\boldsymbol{\mu}}^{+}$.
+
+\subsection{The infinitesimal generator, \(\mathbb B\), of the driving process}
+Given the discussion above, if we are to replicate the arguments of \cite{bean2005} to derive the Laplace-Stieltjes transform of the first return operator of a fluid-fluid queue, we first need an expression for the generator of the driving process, which, for a fluid-fluid queue is a fluid queue.
+We then need to partition it according to whether the second fluid is increasing, decreasing or constant.
+
+The generator of a fluid queue is a differential operator, and to enable computation approximation method are needed.
+The approximation schemes which we discuss in this thesis are all cell-based methods which discretise the level of the fluid queue into cells.
+Thus, one complexity in approximation is to reconcile the partition according to whether the second fluid is increasing, decreasing or constant and the partition which the approximation method uses.
+As long as the rates of the fluid-fluid queue are not too-badly behaved, this is not terribly difficult to do, but it does require some discussion and notation.
+
+\label{subsec:B_operators}
+\subsubsection{General definition}
+Since \(\{( X(t),\varphi(t))\}_{t\geq 0}\) is a Markov process, the evolution of probability can be described by a semigroup.
+Let $\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$ be the set of integrable complex-valued Borel measures on the Borel $\sigma$-algebra $\mathcal{B}_{\mathcal{S} \times \mathbb{R}_{+}}$.
+For $\overline{\bs{\mu}} \in \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$, we can write \(\overline{\bs{\mu}} = (\overline{\mu_i})_{i \in \mathcal{S}}\).
+The measures \(\overline{\mu_i}(\cdot)\) represent an initial distribution, \(\overline{\mu_i}(\cdot) = \mathbb P(X(0))\in\cdot, \varphi(0) = i)\).
+Let \(\{\overline{\mathbb V}(t)\}_{t\geq 0},\, \overline{\mathbb V}(t):\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\mapsto \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\) be the semigroup describing the evolution of probability for \(\{(X(t),\varphi(t))\}_{t\geq 0}\) structured as a matrix of operators, \(\left[ \overline{\mathbb V}(t)\right]_{ij}= \overline{\mathbb V}_{ij}(t)\) where,
+\[\overline\mu_i \overline{ \mathbb V}_{ij}(t)(\mathcal A) = \int_{x\in[0,\infty)} \wrt \overline\mu_i(x)\mathbb P (X(t) \in\mathcal A,\varphi(t) = j \mid X(0)) = x, \varphi(0) = i).\]
+Intuitively, the operator \( \overline{\mathbb V}(t)\) maps an initial measure \(\overline{\boldsymbol \mu}\) on \((X(0),\varphi(0))\) to the measure \(\mathbb P(X(t)\in \mathcal A, \varphi(t)=j)=:\overline\mu_j(t)(\mathcal A)\).
+The matrix of operators \( \overline{\mathbb B}:=[\overline{\mathbb B}_{ij}]_{i,j\in\mathcal S}\) is the \textit{infinitesimal generator} of the semigroup \(\{ \overline{\mathbb V}(t)\}\) defined by 
+\[ \overline{\mathbb B} =\left.\cfrac{\wrt}{\wrt t} \overline{\mathbb V}(t)\right|_{t=0},\]
+with domain the set of measures for which this limit exists.
+Specifically, the domain of \(\overline{\mathbb B}\) is the set of measures, \(\overline{\bs \mu} = (\overline\mu_i)_{i\in\calS}\), for which each \(\overline\mu_i\) admits an absolutely continuous density on \((0,\infty)\), and can have a point mass at \(0\) if \(i\in\mathcal S_{-1}\) or a point mass at \(b\) if \(i\in\mathcal S_{K+1}\); call this set of measures \(\mathcal M_{0,b}\).
+The measure \(\overline\mu_i\) cannot have a point mass at 0 if \(i\notin \mathcal S_{-1}\), nor can they have a point mass at \(b\) if \(i\notin\calS_{{K+1}}\).
+In the sequel we write \(v_i(x),\, x>0\), as the density of \(\overline\mu_i\), and \(q_{{-1},i}\) and \(q_{{K+1},i}\) as the point masses of \(\overline\mu_i\) at \(0\) and \(b\), respectively (if such point masses exist).
+
+\subsubsection{A partition with respect to rates of the second fluid}
+To use the operators \(\{ \overline{\mathbb  V}(t)\}\) and \( \overline{\mathbb  B}\) to analyse the fluid-fluid model, \cite{bo2014} explicitly track when \((X(t),\varphi(t))\in(\mathcal F_i^m,i)\) for \(i\in\mathcal S,\, m \in \{+,-,0\}\) by partitioning the operators \( \overline{\mathbb  V}(t)\) and \( \overline{\mathbb  B}\) into \( \overline{\mathbb  V}_{ij}^{m n}\) and \( \overline{\mathbb  B}_{ij}^{m n}\), for \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\), where
+\[\left.\mu_i\right|_{\calF_i^m}  \overline{\mathbb  V}_{ij}^{m n}(t)(\mathcal A) := \int_{x\in[0,\infty)} \left.
+\wrt \mu_i\right|_{\calF_i^m}(x)\mathbb P (X(t) \in\mathcal A\cap \calF_j^n,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
+and \(\left.\mu_i\right|_{E}\) is the restriction of \(\mu_i\) to \(E\). Similarly, for \( \overline{\mathbb  B}_{ij}^{m n},\) \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\).
+
+\subsubsection{A partition as dictated by the finite-element method}
+We claim that numerical schemes are needed to approximate the analytic operator equations introduced in \cite{bo2014}.
+The schemes we choose to use here work by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \(\calD_k=[y_k,y_{k+1}]\) then constructing an appropriate approximation to \(\overline{\mathbb B}\) on each interval.
+To help elucidate the connection between the operators \(\{ \overline{\mathbb  V}(t)\}\), \( \overline{\mathbb  B}\) and their approximation counterparts we take a slightly different approach to partitioning these operators than that taken in \cite{bo2014}.
+Rather than partition according to the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\), we use the same partition as that in the construction of the approximation schemes.
+By doing so, we can directly correspond elements of the partitioned operators to their approximation counterparts.
+Since the partition used to construct the schemes is finer, then we can always reconstruct the partition in terms of the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\).
+
+Let us first partition the space \([0,\infty)\) into \(\calD_{-1} = \{0\}\), \(\calD_{{K+1}}=\{b\}\), and non-trivial intervals \(\calD_k=[y_k,y_{k+1}]\setminus (\{0\}\cup\{b\}),\) with \(y_0=0,\,y_{K+1}=b,\, y_k<y_{k+1},\,k\in\mathcal K^\circ:=\{0,1,2,...,K\}\) and define \(\mathcal K = \{-1,K+1\}\cup\mathcal K^\circ\).
+For $\bs{\mu} \in \mathcal{M}_{0,b}(\mathcal{S} \times \mathbb{R}_{+})$ we write \(\bs{\mu} = (\mu_{k,i})_{i \in \mathcal{S},k\in\mathcal K},\) where $\mu_{k,i}(\cdot) = \mu_i(\cdot \cap \calD_k),\,k\in\mathcal K.$ When it exists, we denote by \(v_{k,i}(x),\,x>0\), the densities associated with each measure, \(\mu_{k,i},\,k\in \mathcal K^\circ\).
+For \( i,j\in\calS,\,k,\ell\in\mathcal K\) define the operators
+\[\mu_{k,i} \mathbb V_{ij}^{k \ell}(t)(\mathcal A) := \int_{x\in\calD_k} \wrt \mu_{k,i}(x)\mathbb P (X(t) \in\mathcal A\cap \calD_\ell,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
+and the matrices of operators \(\mathbb V^{k\ell}(t) := \vligne{\mathbb V_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\) and write 
+\[\mathbb V(t) = \left[
+	\begin{array}{ccccc}
+		\mathbb V^{{-1},{-1}}(t)&\mathbb V^{{-1},0}(t)& \mathbb V^{{-1},1}(t) &\hdots & \mathbb V^{{-1},{K+1}}(t)\\
+		\mathbb V^{0,{-1}}(t)&\mathbb V^{0,0}(t)&\mathbb V^{0,1}(t)&\hdots&\mathbb V^{0,{K+1}}(t)\\
+		\mathbb V^{1,{-1}}(t)&\mathbb V^{1,0}(t)&\mathbb V^{1,1}(t)&\hdots&\mathbb V^{1,{K+1}}(t)\\
+		\vdots & \vdots & \vdots & \ddots & \vdots \\
+		\mathbb V^{{K+1},{-1}}(t) &\mathbb V^{{K+1},0}(t) &\mathbb V^{{K+1},1}(t) & \hdots & \mathbb V^{{K+1},{K+1}}(t) 
+%		&\ddots &&&&\reflectbox{\(\ddots\)}\\
+%		&&\mathbb V^{k-1,k-1}(t) & \mathbb V^{k-1,k}(t) & \mathbb V^{k-1,k+1}(t)& \\
+%		&&\mathbb V^{k,k-1}(t) & \mathbb V^{k,k}(t) & \mathbb V^{k,k+1}(t)& \\
+%		&&\mathbb V^{k+1,k-1}(t) & \mathbb V^{k+1,k}(t) & \mathbb V^{k+1,k+1}(t)& \\
+%		&\reflectbox{\(\ddots\)}&&&&\ddots
+	\end{array}\right].\]
+Now define \(\mathbb B=\left.\cfrac{\wrt}{\wrt t}\mathbb V(t)\right|_{t=0}\) as the infinitesimal generator of \(\{\mathbb V(t)\}\), resulting in the tridiagonal matrix of operators 
+\[\mathbb B(t) = \left[
+	\begin{array}{ccccc}
+		\mathbb B^{{-1},{-1}}(t)&\mathbb B^{{-1},0}(t)& & & \\
+		\mathbb B^{0,{-1}}(t)&\mathbb B^{0,0}(t)&\mathbb B^{0,1}(t)&& \\
+		&\mathbb B^{1,0}(t)&\mathbb B^{1,1}(t)&\ddots& \\
+		& & \ddots & \ddots & \mathbb B^{K,{K+1}}\\
+		& & & \mathbb B^{{K+1},K} & \mathbb B^{{K+1},{K+1}}
+	\end{array}\right],\]
+%\begin{align*}
+%\mathbb B &= \left[\begin{array}{lllll}
+%		\ddots &&&&\\
+%		\mathbb B^{k-2,k-1} &\mathbb B^{k-1,k-1} & \mathbb B^{k-1,k} & & \\
+%		&\mathbb B^{k,k-1} & \mathbb B^{k,k} & \mathbb B^{k,k+1}& \\
+%		& & \mathbb B^{k+1,k} & \mathbb B^{k+1,k+1}&\mathbb B^{k+1,k+2} \\
+%		&&&&\ddots
+%	\end{array}\right],
+%\end{align*}
+where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\). 
+The tridiagonal structure arises since, for \(|k-\ell|\geq2\) (where we take \({-1} = 0\) and \({K+1} = K+1\) if they appear in the difference) it is impossible for \(\{X(t)\}\) to move from \(\mathcal D_k\) to \(\mathcal D_\ell\) in an infinitesimal amount of time.\footnote{We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
+The operators denoted with an overline play a minor role in the introductory sections of this thesis, but do not appear again.
+We use a blackboard font sans overline (e.g.~\(\mathbb V(t)\) and \(\mathbb B\)) to represent the same operators but which are constructed with the finer partition defined by \(\mathcal D_k,\,k\in\mathcal K\).
+We use the letters \(i,j\in\mathcal S\) to represent states of the phase process, letters \(m,n,\in\mathcal \{+,-,0\}\) to refer to the partition in terms of the sets in Equations (\ref{eqn:fil}), and the letters \(k,\ell\in\mathcal K\) to refer to the finer partition into sets \(\{\calD_k\}_k\).
+With a slight abuse of notation, whenever we use the dummy variables \(k,\ell\) without qualification we imply \(k,\ell\in\mathcal K\), the dummy variables without qualification \(m,n\) imply \(m,n\in\{+,-,0\}\) and the dummy variables \(i,j\) without qualification imply \(i,j\in\calS\).
+E.g.~\(\mathbb B_{ij}^{k\ell}\) means \(\mathbb B_{ij}^{k\ell},\, i,j\in\calS, k,\ell\in\mathcal K\) and \(\mathbb B_{ij}^{mn}\) means \(\mathbb B_{ij}^{mn},\, i,j\in\calS, m,n\in\{+,-,0\}\).
+}
+
+By an appropriate choice of the intervals \(\{\mathcal D_k\},\, k\in\mathcal K\), the partition used in \cite{bo2014} can be recovered.
+Intuitively, we must ensure that each of the boundaries of \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), align with a boundary of a cell \(\mathcal D_k=[y_k,y_{k+1}]\setminus(\{0\}\cup\{b\})\).
+Then, each set \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), can be written as a union of cells, \(\mathcal D_k,\,k\in\mathcal K\), sans a collection of points which have measure zero for all measures in \(\mathcal M_{0,b}\), and this collection of points is inconsequential for the purposes of the approximations presented here.
+
+Formally, to recover the partition used in \citep{bo2014} we choose the intervals \(\calD_k\) such that \(   l (\calD_k\cap\calF_i^m) \in \{   l (\calD_k), 0\}\) for all \(i\in\calS,\,m\in\{+,-,0\},\,k\in\mathcal K\), for all \( l \in\mathcal M_{0,b}\).
+That is, we choose \(\calD_k\) such that it is contained, up to sets of measure 0 with respect to measures in \(\mathcal M_{0,b}\), within one of the sets \(\calF_i^m\) for \(m\in\{+,-,0\}\) and \(i\in\calS\).
+We assume such a partition for the rest of the paper.
+For \(i\in\calS,\,m\in\{+,-,0\}\), let \(\mathcal K^m_i = \{k\in\mathcal K\mid  l (\calD_k\cap\calF_i^m) =   l (\calD_k),\,l\in\mathcal M_0 \}\), so that \(\bigcup\limits_{k\in\mathcal K_i^m} \calD_k\) and \(\mathcal F_i^m\) are equal up to a set of \(\mathcal M_{0,b}\)-measure 0.
+Define \(\mathcal K^m = \bigcup\limits_{i\in\calS}\mathcal K_i^m\), \(m\in\{+,-,0\}\).
+%Then for \(m\in\{+,-,0\}\), \(\mu_i^m(\cdot) = \sum\limits_{k\in\mathcal K_i^m} \mu_i^k(\cdot)\).
+
+To recover the partition defined by (\ref{eqn:fil}) we bundle together the elements of \(\mathbb V(t)\) which correspond to \(\calF_i^m\) and \(\calF_j^n\).
+That is, for \(m,n\in\{+,-,0\}\), define \(\mathbb V_{ij}^{m n}(t)\) as the matrix of operators
+\[\mathbb V_{ij}^{m n}(t) = \left[\mathbb V_{ij}^{k \ell}(t)\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n}.\] 
+%Then, for \(i,j\in\calS,\,m,n\in\{+,-,0\}\), we can write \(\overline{\mathbb  V}_{ij}^{mn}(t) = \boldsymbol 1_{|\mathcal K_i^m|} \mathbb V_{ij}^{m n}(t) \boldsymbol 1_{|\mathcal K_j^n|}\tr{}\) where \(\boldsymbol 1_{|\mathcal K_i^m|}\) and \(\bs 1_{|\mathcal K_j^n|}\) are row-vectors of 1's of length \(|\mathcal K_i^m|\) and \(|\mathcal K_j^n|\), respectively, and \({}\tr{}\) denotes the transpose.
+The same construction can be achieved with \(\mathbb B\).
+
+Let \(\calS_k^+=\{i\in\calS\mid r_i(x)>0,\,\forall x \in\calD_k\}\), \(\calS_k^0=\{i\in\calS\mid r_i(x)=0,\,\forall x \in\calD_k\}\), \(\calS_k^-=\{i\in\calS\mid r_i(x)<0,\,\forall x \in\calD_k\}\) and \(\calS_k^\bullet=\{i\in\calS\mid r_i(x)\neq0,\,\forall x \in\calD_k\}\) for \(k\in\mathcal K\).
+For later reference, we need the following constructions.
+For \(k,\ell\in\mathcal K\)
+\begin{align}
+	\mathbb B^{k\ell} & = \left[\mathbb B_{ij}^{k\ell}\right]_{i,j\in\mathcal S},\label{eqn:ref1here}
+\end{align}
+for \(i,j\in\calS\) 
+\begin{align}
+	\mathbb B_{ij} & = \left[\mathbb B_{ij}^{k\ell}\right]_{k,\ell \in \mathcal K},
+\end{align}
+%for \(i,j\in\calS, \, m,n\in\{+,-,0\}\) 
+%\begin{align}
+%	\mathbb B_{ij}^{m n} &= \left[\mathbb B_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\label{eqn: Bmn etc}
+%	\\ \mathbb B_{ij}^{k n} &= \left[\mathbb B_{ij}^{k \ell}\right]_{\ell\in\mathcal K_j^n} \mbox{ for }k\in\{{-1},1,2,...\},
+%	\\ \mathbb B_{ij}^{m \ell} &= \left[\mathbb B_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m} \mbox{ for } \ell\in\{{-1},1,2,...\},
+%	\label{eqn: Bl0 etc}
+%\end{align}
+and for \(m,n\in\{+,-,0\}\) 
+\begin{align}
+	\mathbb B^{m n} &= \left[\left[\mathbb B_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n},\label{eqn: Bmn2 etc}
+	\\ \mathbb B^{k n} &= \left[\left[\mathbb B_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{\ell\in\mathcal K^n} \mbox{ for }k\in\mathcal K,
+	\\ \mathbb B^{m \ell} &= \left[\left[\mathbb B_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{k\in\mathcal K^m} \mbox{ for } \ell\in\mathcal K.
+	\label{eqn:ref2here}
+\end{align}
+We persist with the partition \(\calD_k,\,k\in\mathcal K\) throughout this thesis as this is consistent with the partition used in the finite-element approximation schemes considered in this thesis.
+Note that for all the operators defined with this partition, the partitioning used in \citep{bo2014} can always be recovered by the above construction.
+
+\subsubsection{The details}
+We can write \(\mu_{k,i}\mathbb B_{ij}^{k\ell}(\mathcal A)\) in kernel form as \(\displaystyle\int_{x\in\calD_k,y\in\mathcal A}\wrt \mu_{k,i}(x)\mathbb B_{ij}^{k\ell}(x,dy)\). It is known that
+\[\mu_{k,i}\mathbb B_{ij}^{kk}(\wrt y):=\int_{x\in\calD_k}\wrt\mu_{k,i}(x)\mathbb B_{ij}^{kk}(x,\wrt y)=\begin{cases}
+v_{k,i}(y)T_{ij}\wrt y, & i\neq j,\\
+v_{k,i}(y)T_{ii}\wrt y - c_i\cfrac{\wrt}{\wrt y}v_{k,i}(y)\wrt y, & i=j, 
+\end{cases}\]
+on the interior of \(\calD_k\), \(k\in\mathcal K^\circ\), \citep{kk1995}.
+Intuitively, \(v_{k,i}(y)T_{ij}\wrt y\) represents the instantaneous rate of transition from phase \(i\) to \(j\) in the infinitesimal interval \(\wrt y\), \(v_{k,i}(y)T_{ii}\wrt y\) represents no such transition occurring, and \(- c_i\cfrac{\wrt}{\wrt y}v_{k,i}(y)\wrt y\) represents the drift across the interval \(\wrt y\) when the phase is \(i\).
+
+Translating the results of \cite{bo2014} to use the partition \(\{\calD_k\}\) we may state that, for all \(i,j\in\calS\), \(k\in\{1,\dots,K-1\}\),
+\begin{align*}
+	\mu_{k,i}\mathbb B_{ij}^{kk}(\calD_k)&=\int_{x\in\calD_k}v_{k,i}(x)T_{ij}\wrt x,\\
+	\mu_{k,i}\mathbb B_{ii}^{kk}(\calD_k)&=\int_{x\in\calD_k}v_{k,i}(x)T_{ii}\wrt x - c_iv_{k,i}(y_{k+1})\mathbb 1_{(c_i>0)} + c_iv_{k,i}(y_{k})\mathbb 1_{(c_i<0)},
+\end{align*}
+where \(\mathbb 1_{}\) is the indicator function. 
+Intuitively, the first expression represents the instantaneous rate of the stochastic transitions of the phase process \(\{\varphi(t)\}\) while \(\{X(t)\}\) remains in \(\calD_k\).
+The first term in the second expression represents the net rate of transition out of phase \(i\) while \(\{X(t)\}\) remains in \(\calD_k\), while the second and third terms in the second expression represent the flux out of the right-hand edge of \(\calD_k\) when \(c_i>0\) and the flux out of the left-hand edge of \(\calD_k\) when \(c_i<0\), respectively.
+
+The results of \cite{bo2014} also imply that, 
+\begin{align*}
+\mu_{k,i}\mathbb B_{ii}^{k,k+1}(\calD_{k+1})&= c_iv_{k,i}(y_{k+1})\mathbb 1_{(c_i>0)},\mbox{ for all \(i\in\calS\), \(k\in\{0,1,2,...,K-1\}\),}
+\\\mu_{k,i}\mathbb B_{ii}^{k,k-1}(\calD_{k-1})&= -c_iv_{k,i}(y_{k})\mathbb 1_{(c_i<0)}, \mbox{  for all \(i\in\calS\), \(k\in\{1,2,3,...,K\}\)}.
+\end{align*}
+Intuitively, the first equation represents the flux from \(\calD_k\) to \(\calD_{k+1}\) across the shared boundary at \(y_{k+1}\) which occurs when \(c_i>0\) only.
+The second expression represents the flux from \(\calD_k\) to \(\calD_{k-1}\) across the shared boundary at \(y_{k}\) which occurs when \(c_i<0\) only.
+
+At the boundary \(x=0\), point mass moves between phases according to 
+\begin{align*}
+	\mu_{-1,i}\mathbb B_{ii}^{{-1},{-1}} &= \mu_{-1,i}(\{0\})T_{ii},\mbox{ if } c_i\leq 0,\,\\
+	\mu_{-1,i}\mathbb B_{ij}^{{-1},{-1}} &= \mu_{-1,i}(\{0\})T_{ij},\mbox{ if }c_i\leq 0,\,c_j\leq 0,
+	\end{align*}
+	point mass becomes density at rate
+	\begin{align*}
+	\mu_{-1,i}\mathbb B_{ij}^{{-1},0} &= \mu_{-1,i}(\{0\})T_{ij},\mbox{ if }c_i\leq 0,\,c_j> 0,
+	\end{align*}
+	and density becomes point mass at rate
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ij}^{0,{-1}} &= -c_ip^{{-1}}_{ij}v_{0,i}(0^+),\mbox{ if }c_i< 0,\,c_j\leq 0.
+	\end{align*}
+	In \(\calD_0\), if \(c_i<0\) and \(c_j>0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{-1}\) to phase \(j\),
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_0)&=\int_{x\in\calD_0}v_{0,i}(x)T_{ij}\wrt x - c_iv_{0,i}(0^+)p^{-1}_{ij}.
+	\end{align*}
+	Also in \(\calD_0\), if \(i\neq j\) and \(c_i\geq 0\) or \(i\neq j\), \(c_i< 0\) and \(c_j\leq 0\), the phase changes from \(i\) to \(j\) while \(\{X(t)\}\) remains in \(\calD_0\) at rate
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_0)&=\int_{x\in\calD_0}v_{0,i}(x)T_{ij}\wrt x,
+	\end{align*}
+	else, \(\{(X(t),\varphi(t))\}\) leaves \((\calD_0,i)\) at rate
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ii}^{0,0}(\calD_0)&=\int_{x\in\calD_0}v_{0,i}(x)T_{ii}\wrt x - c_iv_{0,i}(y_0)\mathbb 0_{(c_i> 0)} + c_iv_{0,i}(0^+)\mathbb 1_{(c_i< 0)},
+\end{align*}
+where \(0^+\) is the right limit at \(0\). Similarly, at the boundary \(x=b\), point mass moves between phases according to 
+\begin{align*}
+	\mu_{K+1,i}\mathbb B_{ii}^{{K+1},{K+1}} &= \mu_{K+1,i}(\{0\})T_{ii}, \mbox{ if }c_i\geq 0,\,\\
+	\mu_{K+1,i}\mathbb B_{ij}^{{K+1},{K+1}} &= \mu_{K+1,i}(\{0\})T_{ij},\mbox{ if }c_i\geq 0,\,c_j\geq 0,
+	\end{align*}
+	point mass becomes density at rate
+	\begin{align*}
+	\mu_{K+1,i}\mathbb B_{ij}^{{K+1},K} &= \mu_{K+1,i}(\{0\})T_{ij},\mbox{ if }c_i\geq 0,\,c_j<0,
+	\end{align*}
+	and density becomes point mass at rate
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ij}^{K,{K+1}} &= c_ip^{{K+1}}_{ij}v_{K,i}(0^+),\mbox{ if }c_i> 0,\,c_j\geq 0.
+	\end{align*}
+	In \(\calD_K\), if \(c_i>0\) and \(c_j<0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{K+1}\) to phase \(j\),
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_K)&=\int_{x\in\calD_K}v_{K,i}(x)T_{ij}\wrt x + c_iv_{K,i}(b^-)p^{K+1}_{ij},\,c_i>0,\,c_j<0.
+	\end{align*}
+	If \(i\neq j\) and \(c_i\leq 0\) or \(i\neq j\), \(c_i>0\) and \(c_j\geq 0\), the phase changes from \(i\) to \(j\) within \(\calD_K\) at rate
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_K)&=\int_{x\in\calD_K}v_{K,i}(x)T_{ij}\wrt x,
+	\end{align*}
+	else, \(\{(X(t),\varphi(t))\}\) leaves \((\calD_K,i)\) at rate
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ii}^{K,K}(\calD_K)&=\int_{x\in\calD_K}v_{K,i}(x)T_{ii}\wrt x + c_iv_{K,i}(y_K)\mathbb 1_{(c_i< 0)} - c_iv_{K,i}(b^-)\mathbb 1_{(c_i> 0)},
+\end{align*}
+where \(b^-\) is the left limit at \(b\). Otherwise, \(\mu_{k,i}\mathbb B_{ij}^{k\ell}=0\).
+
+Note that we have not presented \(\mathbb B\) in its full detail here and refer the reader to \citep{bo2014} for the details.
+The main goal here is to show how \(\mathbb B\) is used to construct the stationary distribution of the fluid-fluid queue and to illustrate the link between the operator \(\mathbb B\) and the approximations of the same object.
+As we shall see later, these expressions closely resemble the approximations to the same quantities.
+
+\subsection{The infinitesimal generator, \(\mathbb D\), of an in-out process}
+Let $\beta(t) := \int_0^t \left| r_{\varphi(z)}(X(z)) \right|  \wrt z$ be the total unregulated amount of fluid that has flowed into or out of the second buffer during $[0,t]$, and let $\eta(y) := \inf \{t > 0: \beta(t) = y\}$ be the first time this accumulated in-out amount hits level $y$.
+Note that at the stopping time \(\eta(y)\) it must be that \((X({\eta(y)}),\varphi({\eta(y)}))\in(\calF_i^m,i)\) for some \(i\in\calS\) and \(m\in\{+,-\}\), i.e.~\(m\neq0\).
+We define the operators $\mathbb{U}_{ij}^{k\ell}(y,s): \mathcal{M}_{0,b}(\mathcal{D}_{k}\cap\calF_i^m) \mapsto \mathcal{M}_{0,b} (\calD_\ell\cap\mathcal{F}_j^n)$, for $k\in\mathcal K_i^+\cup\mathcal K_i^-$,  $\ell\in\mathcal K_j^+\cup\mathcal K_j^-$, and $i \in \mathcal{S}_k^\bullet,\,j \in \mathcal{S}_k^\bullet$, by
+% 
+	\begin{align*} 
+		&\mu_{k,i}\mathbb{U}_{ij}^{k\ell} (y,s) (\mathcal{A}) 
+		\\&:= \int_{x \in \calD_k} \wrt  \mu_{k,i}(x) \mathbb{E}\left[e^{-s\eta(y)}{1}\left\{X({\eta(y)}) \in \mathcal{A}\cap\calD_\ell,\;\varphi({\eta(y)}) = j\right\} \mid \varphi(0) = i, X(0) = x\right].
+	\end{align*} 
+Then, construct the matrix of operators 
+\begin{align*}
+%\mathbb{U}^{k \ell}(y,s)&:=[\mathbb U_{ij}^{k\ell}(y,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}, \, k,\ell\in \mathcal K^+\cup\mathcal K^-,
+\mathbb U(y,s) &:= \left[[\mathbb U_{ij}^{k\ell}(y,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}\right]_{k,\ell\in\mathcal K^+\cup\mathcal K^-}.\end{align*}
+%\[\mathbb U(t) = \left[
+%	\begin{array}{llll}
+%		\mathbb U^{{-1},{-1}}(t)&\mathbb U^{{-1},1}(t)& \mathbb U^{{-1},2}(t) &\hdots \\
+%		\mathbb U^{1,{-1}}(t)&\mathbb U^{1,1}(t)&\mathbb U^{1,2}(t)&\hdots\\
+%		\mathbb U^{2,{-1}}(t)&\mathbb U^{2,1}(t)&\mathbb U^{2,2}(t)&\hdots\\
+%		\vdots & \vdots & \vdots & \ddots 
+%	\end{array}\right].\]
+%	\begin{align*} 
+%		\mathbb{U}(y,s) = \left[\begin{array}{llllll} 
+%		\ddots & & & &~\reflectbox{$\ddots$}\\
+%            		& \mathbb{U}^{k-1,k-1}(y,s) & \mathbb{U}^{k-1,k}(y,s) & \mathbb{U}^{k-1,k+1}(y,s) \\
+%			& \mathbb{U}^{k,k-1}(y,s) & \mathbb{U}^{k,k}(y,s) & \mathbb{U}^{k,k+1}(y,s) \\
+%			& \mathbb{U}^{k+1,k-1}(y,s) & \mathbb{U}^{k+1,k}(y,s) & \mathbb{U}^{k+1,k+1}(y,s) \\	
+%			\reflectbox{$\ddots$}& & & & \ddots		
+%		\end{array}\right], 
+%	\end{align*} 
+	% 
+%	where $\mathbb{U}^{k \ell}:=\mathbb U_{ij}^{k\ell}$ is an $|\mathcal{S}| \times |\mathcal{S}|$ matrix of operators for $y > 0$, $s \in \mathbb{C}$, and Re$(s) \geq 0$. 
+The matrix of operators \(\mathbb D(s)\) is the infinitesimal generator of the semigroup \(\{\mathbb U(y,s)\}_{y\geq 0}\) defined by 
+\[\mathbb D(s) = \cfrac{\wrt}{\wrt y}\mathbb U(y,s)|_{y=0},\]
+whenever this limit exists.
+
+	Recalling the constructions in Equations (\ref{eqn:ref1here})-(\ref{eqn:ref2here}) and using Lemma~$4$ of \cite{bo2014} gives the following expression for $\mathbb{D}(s)$. 
+\begin{lem}\label{lemma: D(s)}
+	For $y \geq 0$, $s \in \mathbb{C}$ with \textit{Re}$(s) \geq 0$, $i,j \in \mathcal{S}$, $k\in \mathcal K_i^+\cup\mathcal K_i^-$, $\ell\in \mathcal K_j^+\cup\mathcal K_j^-$,
+	% 
+	\begin{align*}
+		\mathbb{D}_{ij}^{k\ell}(s) = [\mathbb{R}^{k}(
+		\mathbb{B}^{k\ell } - s\mathbb{I} + \mathbb{B}^{k0 }(s \mathbb{I} - \mathbb{B}^{00})^{-1}\mathbb{B}^{0\ell})]_{ij}, 
+	\end{align*} 
+	% 
+	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} := \diag(\mathbb{R}_i^{k})_{i \in \mathcal{S}}$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
+	\begin{align*} 
+		{\mu}_{k,i}\mathbb{R}_{k,i}(\mathcal{A}) := \int_{x \in \mathcal{A} \cap \mathcal{D}_k} \frac{1}{r_i(x)}\wrt  \mu_{k,i}(x),\quad k\in \mathcal K_i^+\cup\mathcal K_i^-.
+	\end{align*} 
+\end{lem}
+
+	Also, construct the matrices of operators 
+	\begin{align*}
+            	%\mathbb D^{k \ell} &:= \left[\mathbb D_{ij}^{k \ell}\right]_{i,j\in\mathcal S},\quad k,\ell\in\{{-1},1,2,...\},
+		\mathbb D^{m n} &:= \left[\left[\mathbb D_{ij}^{k \ell}\right]_{i\in\calS_k^m,j\in\calS_k^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n}.
+		%\\ \mathbb D_{ij}^{m n} &:= \left[\mathbb D_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\quad i,j\in\calS,\,m,n\in\{+,-,0\}.
+        \end{align*}
+
+\subsection{The first-return operator, $\mathbb\Psi(s)$}\label{sec: intro Psi}
+We denote by $\mathbb \Psi(s)$ the matrix of operators with the same dimensions as \(\mathbb D^{+-}\), recording the Laplace-Stieltjes transforms of the time for $\{\dot Y(t)\}$ to return, for the first time, to the initial level of zero as introduced in \cite{bo2014} but constructed with respect to the finer partition \(\{\calD_k\}\).
+Define the stopping time $\zeta_Y(E):= \inf \{t > 0: \dot Y(t) \in E\}$ to be the first time $\{\dot Y(t)\}$ hits the set $E$, then each component $\mathbb \Psi_{ij}^{k\ell}(s): \mathcal{M}_{0,b}(\mathcal D_k,i) \mapsto \mathcal{M}_{0,b}(\mathcal D_\ell,j), i,j \in \mathcal{S},\,k\in\mathcal K_i^+$ and $\ell \in \mathcal K_j^-$, is given by
+% 
+\begin{align*} 
+    % 
+	&\mu_{k,i}\mathbb\Psi_{ij}^{k\ell}(s) (\mathcal{A}) 
+	\\& := \int_{x \in \calD_k} \wrt \mu_{k,i}(x)
+	% 
+	 \mathbb{E}\big[e^{-s\zeta_Y(\{0\})}1\left\{X({\zeta_Y(\{0\})}) \in \mathcal{A}\cap\calD_\ell,\; \varphi({\zeta_Y(\{0\})}) = j \right\} \mid X(0) = x, 
+	 \\&\qquad{} \dot Y(0) = 0, \varphi(0) = i\big].
+	 % 
+\end{align*} 
+\cite{bo2014}~Theorem~1 provides the following result which characterises \(\mathbb\Psi(s)\).
+\begin{theo} 
+	\label{theo:Psi} 
+	For \textit{Re}$(s) \geq 0$, $\mathbb\Psi(s)$ satisfies the  equation: 
+	% 
+	\begin{align*} 
+		\mathbb{D}^{+-}(s) + \mathbb\Psi(s)\mathbb{D}^{-+}(s)\mathbb\Psi(s) + \mathbb{D}^{++}(s)\mathbb\Psi(s) + \mathbb\Psi(s)\mathbb{D}^{--}(s) = 0. 
+	\end{align*} 
+	% 
+	Furthermore, if $s$ is real then $\mathbb\Psi(s)$ is the minimal nonnegative solution. 
+\end{theo}
+
+\subsection{Stationary Distribution}
+
+Let $\mathbb\Psi := \mathbb\Psi(0)$.
+We define $\zeta_Y^n(\{0\}) := \inf\{t \geq \zeta_Y^{n - 1}(\{0\}): \dot Y(t) = 0\}$, for $n \geq  2$, to be the sequence of hitting times to level $0$ of $\dot Y(t)$, with $\zeta_Y^1(\{0\}): = \zeta_Y(\{0\})$.
+Consider the discrete-time Markov process $\{X({\zeta_Y^n(\{0\})}), \varphi(\zeta_Y^n(\{0\}))\}_{n \geq 1}$, and for $i \in \mathcal{S},\,k\in\mathcal K_i^-$ define the measures $\bbxi_{k,i}$ as follows
+	%
+	\begin{align*}
+		\bbxi_{k,i}(\mathcal{A}) := \lim_{n \rightarrow \infty} \mathbb{P}\left(X(\zeta_Y^n(\{0\})) \in \mathcal{A}\cap\calD_k, \varphi(\zeta_Y^n(\{0\})) = i\right).
+	\end{align*} 
+	By \cite{bo2014}, the vector of measures $\boldsymbol{\bbxi} := (\bbxi_{k,i})_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}$ satisfies the following set of equations 
+ % 
+ 	\begin{align}
+		\vligne{\boldsymbol{\bbxi}  & \boldsymbol{0}}\left(-\left[\begin{array}{ll} 
+			\mathbb{B}^{--} & \mathbb{B}^{-0} \\
+			\mathbb{B}^{0-} & \mathbb{B}^{00} 
+		\end{array} \right]\right)^{-1}\left[\begin{array}{l} 
+			\mathbb{B}^{-+} \\ 
+			\mathbb{B}^{0+}
+		\end{array} \right]\mathbb\Psi & = \boldsymbol{\bbxi}, \label{eqn:xi1}\\ 
+		\sum_{k\in\mathcal K^-}\sum_{i \in \mathcal{S}_k^-}\bbxi_{k,i}(\mathcal{F}^-_i) & = 1. \label{eqn:xi2}
+	\end{align}
+
+We reproduce Theorem 2 of \cite{bo2014} below, which gives the joint stationary distribution of $\{(\dot Y(t),X(t),  \varphi(t))\}$.
+Recall that the joint stationary density operator ${\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{(\dot Y(t), X(t), \varphi(t))\}$ and the joint stationary mass operator ${\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$ are defined by~\eqref{eqn:jointpi} and \eqref{eqn:jointmass}, respectively.
+%Following out notational convention, define an equivalent joint stationary density operator $\boldsymbol{\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{X(t), Y(t), \varphi(t)\}$ and an equivalent joint stationary mass operator $\boldsymbol{\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$, but partitioned accprsing .
+We can partition \(\bbpi\) as follows 
+% 
+	\begin{align*} 
+		\bbpi(y) &= \vligne{\bbpi^{+}(y) & \bbpi^{-}(y) & \bbpi^{0}(y)} 
+		\\&= \vligne{\left(\bbpi_{k,i}(y)\right)_{i \in \mathcal{S}_k^+,k\in\mathcal K^+} & \left(\bbpi_{k,i}(y)\right)_{i \in \mathcal{S}_k^-,k\in\mathcal K^-} & \left(\bbpi_{k,i}(y)\right)_{i \in \mathcal{S}_k^0,k\in\mathcal K^0}},
+	\end{align*} 
+	%
+	where  
+% 
+	\begin{align*} 
+		\bbpi_{k,i}(y)(\mathcal{A}) = \bbpi_i(y)(\mathcal{A}\cap \mathcal D_k).
+	\end{align*} 
+%
+Similarly, we can write 
+	\begin{align*} 
+		\mathbb{p} &= \vligne{\mathbb{p}^{-} & \mathbb{p}^{0}} 
+		= \vligne{\left(\mathbb p_{k,i}\right)_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}  & \left(\mathbb p_{k,i}\right)_{i \in \mathcal{S}_k^0,k\in\mathcal K^0} },
+	\end{align*} 
+	%
+	where  
+% 
+	\( 
+		\mathbb p_{k,i}(\mathcal{A}) = \mathbb p_i(\mathcal{A}\cap \mathcal D_k).
+	\) 
+%
+\begin{theo} 
+	\label{theo:density} 
+The density ${\bbpi}^{m}(y)$, for $m \in \{+,-,0\}$ and $y > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the following set of equations:
+% 	 
+	\begin{align} 
+	&\label{11} \; \bbpi^{0}(y) = \vligne{\bbpi^{+}(y) & \bbpi^{-}(y)}\left[\begin{array}{l} \mathbb{B}^{+0} \\ \mathbb{B}^{-0} \end{array} \right]\left(-\mathbb{B}^{00}\right)^{-1}, \\
+	% 
+	&  \vligne{\bbpi^{+}(y) & \bbpi^{-}(y)} = \vligne{\mathbb{p}^{-} & \mathbb{p}^{0}}\left[\begin{array}{l} \mathbb{B}^{-+} \\ \mathbb{B}^{0+} \end{array} \right]\vligne{e^{\mathbb{K}y} & e^{\mathbb{K}y}\mathbb\Psi}\left[\begin{array}{cc} \mathbb{R}^{+} & 0 \\ 0 & \mathbb{R}^{-}\end{array}\right], \\
+	% 
+	&  \vligne{\mathbb{p}^{-}  & \mathbb{p}^{0}} = z \vligne{{\bbxi} & \bs{0}} 
+	\left(-\left[\begin{array}{ll} 
+		\mathbb{B}^{--} & \mathbb{B}^{-0} \\
+		\mathbb{B}^{0-} & \mathbb{B}^{00} 
+		\end{array} \right] \right)^{-1},  \label{eqn:mass}\\
+	%  
+	& \sum_{m \in \{+,-,0\}}\sum_{i \in \mathcal{S}} \int_{y = 0}^{\infty} \bbpi_i^{m}(y)(\mathcal{F}^{m}_i)\wrt y + \sum_{m \in \{-,0\}} \sum_{i \in \mathcal{S}}\mathbb p^{m}_i(\mathcal{F}^{m}_i) = 1, \label{14}
+	\end{align}
+	% 
+	where $\mathbb{K} := \mathbb{D}^{++}(0) + \mathbb\Psi\mathbb{D}^{(-+)}(0)$ and $z$ is a normalising constant. 
+\end{theo}
+
+At this point we reiterate that Equations~(\ref{11})-(\ref{14}) are operator equations and are only amenable to numerical evaluation in the simplest of cases.
+Sources of this intractability come from, for example, the need to find the inverse operator \(( - \mathbb B^{00})^{-1}\), and the need to find the solution, \(\mathbb \Psi(s)\), of the operator equation in Theorem~\ref{theo:Psi}.
+There is also the complexity of the partition of the operators defined by the sets \(\mathcal F_{i}^m,\, i\in\calS,\,m\in\{+,-,0\}\).
+Therefore, there is the need for approximation schemes such as those presented in this thesis.
+We now proceed to some mathematical preliminaries required to construct our approximations.
+
+\section{Quasi-birth-and-death processes with rational arrival process components}
+One of the approximation schemes we develop in this thesis is a quasi-birth-and-death process with rational-arrival-process components (QBD-RAP).
+The class of QBD-RAP processes are built from matrix-exponentially distributed inter-event times.
+We now introduce the class of matrix exponential distributions and recount some important properties before formally introducing the QBD-RAP.
+
+\subsection{Matrix-exponential distributions}
 Here we recount some facts about matrix exponential distributions.
-See \citep{MEinAP} for a more detailed exposition.
+See \cite{MEinAP} for a more detailed exposition.
 A random variable, \(Z\), is said to have a matrix-exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
@@
 %Here \(\bs s\) can be taken to be \(\bs D e\).
-The \(n\)th inter-arrival, \(T_n-T_{n-1}\), has a matrix exponential distribution with density \(\bs \alpha(-\bs S\bs D )^{n-1}e^{\bs S x_n}\bs s\).
+The \(n\)th inter-arrival, \(T_n-T_{n-1}\), has a matrix exponential distribution with density \(\bs \alpha(-\bs S\bs D )^{n-1}e^{\bs S y_n}\bs s\).
 Denote such a process \(N \sim RAP(\bs \alpha, \bs S, \bs D)\).
@@
 
-\section{QBD-RAPs}\label{sec: qbd-rap}
-As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes.
+\subsection{QBD-RAPs}\label{sec: qbd-rap}
+% As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes.
+
 To define a QBD-RAP we first need a Batch (Marked) RAP.
-
 Let \(\mathscr K\subset \mathbb Z\) be a set of \emph{marks}.
@@
 Now consider a Marked RAP, \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\{-1,0,+1\}).\) The process \(\{(L(t),\bs A(t))\}_{t\geq0}\) formed by letting \(L(t) = N_{+1}(t) - N_{-1}(t)\) is a QBD-RAP.
+
+% vector space? Polynomial basis?
+% me basis? semigroup? generator? 
+\section{Laplace Transforms}
+In Chapters~\ref{sec: conv} and \ref{ch: global results} we work with an object called the \emph{Laplace transform}.
+For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transforms of \(\mu\) to be
+\[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
+where the \emph{region on convergences} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
+\[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
+When \(\mu\) is the probability measure associated with a random variable, \(Z\), say, then we may write 
+\[\widehat \mu(\lambda)=\mathbb E[e^{-\lambda Z}],\]
+and the region of convergence is at least \([0,\infty)\). 
+Further, letting \(E^\lambda\) be an exponentially distributed random variable with rate \(\lambda\) and noting that \(\mathbb P(E^\lambda >t ) = e^{-\lambda t}\), then 
+\[\widehat \mu(\lambda)=\mathbb P(Z<E^\lambda),\]
+which gives a probabilistic interpretation of the Laplace transform.
+
+A convenient property of the Laplace transform which we utilise is the Convolution Theorem. 
+\begin{thm}[Convolution Theorem]
+	Let \(f,g: [0,\infty) \to \mathbb R\) be integrable functions, then 
+	\[\mathcal L\left(\int_{u=0}^tf(u)g(t-u)\wrt u\right)(\lambda) = \mathcal L\left(f\right)\cdot \mathcal L\left(g\right).\]
+\end{thm}
+The Convolution Theorem states that the Laplace transform of the convolution, given by \(\displaystyle \int_{u=0}^tf(u)g(t-u)\wrt u\), is equal to the product of the Laplace transform of \(f\) and \(g\).
+
+The Laplace transform is unique in the sense that, if \(\mu\) and \(\nu\) are two measures on the Borel sets of \([0,\infty)\) and 
+\[\widehat \mu(\lambda) = \widehat \nu(\lambda)\] 
+for all \(\lambda > a\) with \(a<\infty\), then \(\mu\) and \(\nu\) are the same. In terms of functions, \(f,g: [0,\infty) \to \mathbb R\), if 
+\[\mathcal L(f)(\lambda) = \mathcal L(g)(\lambda),\]
+for all \(\lambda > a\) with \(a<\infty\), and \(f\) and \(g\) are continuous, then \(f(t)=g(t)\) for all \(t\geq 0\). Without knowing \(f\) and \(g\) are continuous, then we can only claim that 
+\(f(t)=g(t)\) for all \(t\geq 0, t\notin \mathcal N,\) where \(\mathcal N\) is a \emph{null set} with respect to Lebesgue measure.
+
+\section{Convergence theorems}
+We use the following convergence theorems to help us prove that the QBD-RAP scheme converges weakly to the distribution of the fluid queue.
+The first result we state is a sweeping statement about convergence of measures.
+First, let's define the notion of weak convergence.
+We follow \cite{portmanteaubook}.
+
+Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra.
+Further, denote by \(C_b(E)\) the set of continuous bounded functions of \(E\).
+Let \(\mu,\,\mu_1,\,\mu_2,...\in \mathcal M_f(E)\).
+We say that \(\{\mu_n\}_{n\in\mathbb N}\) converges weakly to \(\mu\), formally \(\mu_n\to \mu\) weakly as \(n\to\infty\), if
+\[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
+
+We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}.
+Let \(\mathcal M_{\leq 1}(E):=\{\mu \in \mathcal M_f(E)\mid \mu(E)\leq 1\},\) the set of all sub-probability measures on \((E,\mathcal E)\).
+\begin{thm}[(Part of) The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
+	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
+	\begin{enumerate}
+		\item[(i)] \(\mu_n\to \mu\) weakly as \(n\to \infty\). 
+		\item[(ii)] \(\int f\wrt \mu_n \to \int f\wrt \mu\) for all bounded, Lipschitz continuous \(f\).
+		\\ \(\vdots\) 
+	\end{enumerate}
+\end{thm}
+There are 8 parts to The Portmanteau Theorem in \cite{portmanteaubook}, Theorem~13.16.
+Here we only quote to relevant parts; the other 6 parts require us to define other concepts which are not relevant to this thesis, so they are omitted.
+See \cite{portmanteaubook}, Theorem~13.16 for details.
+
+Another tool we can use to show convergence of measures is to show that the Laplace transforms converge, as stated in the following theorem.
+\begin{thm}[Extended Continuity Theorem, \cite{feller1957}, Theorem 2a]\label{thm: ext cont thm}
+	For \(p=1,2,...,\) let \(U_p\) be a measure with Laplace transform \(\zeta_p\).
+If \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda > a\geq 0\), then \(\zeta\) is the Laplace transform of a measure \(U\) and \(U_p\to U\) [weakly].
+
+	Conversely, if \(U_p\to U\)[weakly] and the sequence \(\{\zeta_p(a)\}\) is bounded, then \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda >a\). 
+\end{thm}
+
+In Chapters~\ref{sec: conv} and \ref{ch: global results} we use the Dominated Convergence Theorem to aid our convergence arguments.
+In applied probability we often want to prove convergence of certain probabilistic expressions but the expressions themselves are too difficult to characterise.
+An approach which can simplify matters is to partition the expression on certain events where we have a simpler characterisation of the expression on each event, thereby enabling us to prove convergence on each element in the partition.
+The original expression can be written as a sum over the partition.
+To establish the convergence result we initially desired, we can use the convergence of each element of the partition and the Dominated Convergence Theorem.
+I have taken the following from Theorem~1.13~\cite{steinreal}.
+Further, \cite{steinreal} use the notation
+\[\int f = \int f \wrt x = \int f \wrt m(x),\]
+where \(m\) denotes Lebesgue measure, to denote the Lebesgue integral. With their notation clarified we now quote their results.
+\begin{thm}[Dominated Convergence Theorem]
+	Suppose \(\{f_n\}\) is a sequence of measurable functions such that \(f_n(x)\to f(x)\) almost everywhere with respect to \(x\), as \(n\) tends to infinity.
+If \(|f_n(x)|\leq g(x)\), where \(g\) is integrable, then
+	\[\int|f_n-f|\to 0 \mbox{ as } n \to \infty,\]
+	and consequently 
+	\[\int f_n\to\int f \mbox{ as } n\to \infty.\]
+\end{thm}
+
+Also in Chapters~\ref{sec: conv} and \ref{ch: global results} we want to manipulate infinite sums or integrals and rearrange the order of summation or integration.
+However, things can go awry when we swap the order of integration/summation if we are not careful.
+The next few results give some conditions under which we have equality under before and after swapping the order of summation/integration.
+Once again, we follow \cite{steinreal} quoting their Theorem~2.13.
+If \(f\) is a function in \(\mathbb R^{d_1}\times \mathbb R^{d_2}\), the \emph{slice} of \(f\) corresponding to \(y\in\mathbb R^{d_2}\) is the function \(f^y\) of the \(x\in\mathbb R^{d_1}\) variable, given by
+\[f^y(x)=f(x,y).\]
+Similarly, the slice of \(f\) for a fixed \(x\in\mathbb R^{d_1}\) is \(f_x(y)=f(x,y)\). 
+\begin{thm}[Fubini's Theorem]\label{Fubini}
+	Suppose \(f(x,y)\) is integrable on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
+	\begin{itemize}
+		\item The slice \(f^y\) is integrable on \(\mathbb R^{d_1}\).
+		\item The function defined by \(\int_{R^{d_1}} f^y(x)\wrt x\) is integrable on \(\mathbb R^{d_2}\). 
+	\end{itemize}
+	Moreover:
+	\begin{itemize}
+		\item[(iii)] \(\displaystyle\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y = \int_{\mathbb R^d}f.\)
+	\end{itemize}
+\end{thm}
+\cite{steinreal} then follow ``Clearly, the [Fubini] theorem is symmetric in \(x\) and \(y\) so that we also may conclude that the slice \(f_x\) is integrable on \(\mathbb R^{d_2}\) for almost every \(x\).
+Moreover, \(\int_{\mathbb R^{d_2}}f_x(y)\wrt y\) is integrable and
+\[\displaystyle\int_{R^{d_1}}\left(\int_{R^{d_2}}f(x,y)\wrt y\right)\wrt x = \int_{\mathbb R^d}f.\]
+In particular, Fubini's theorem states that the integral of \(f\) on \(\mathbb R^d\) can be computed by iterating lower-dimensional integrals, and that the iterations can be taken in any order
+\[\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y=\displaystyle\int_{R^{d_1}}\left(\int_{R^{d_2}}f(x,y)\wrt y\right)\wrt x = \int_{\mathbb R^d}f.\mbox{''}\]
+It is this last statement which is most powerful. Effectively, if either
+\[\int_{R^{d_2}}\left(\int_{R^{d_1}}|f(x,y)|\wrt x\right)\wrt y<\infty,\]
+or
+\[\int_{R^{d_1}}\left(\int_{R^{d_2}}|f(x,y)|\wrt y\right)\wrt x<\infty,\]
+then we can swap the order of integration.
+
+A closely related theorem which is often used alongside Fubini's Theorem is the following 
+\begin{thm}[Tonelli's Theorem]\label{Tonelli}
+	Suppose \(f(x,y)\) a non-negative measurable function on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
+	\begin{itemize}
+		\item The slice \(f^y\) is integrable on \(\mathbb R^{d_1}\).
+		\item The function defined by \(\int_{R^{d_1}} f^y(x)\wrt x\) is integrable on \(\mathbb R^{d_2}\). 
+	\end{itemize}
+	Moreover:
+	\begin{itemize}
+		\item[(iii)] \(\displaystyle\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y = \int_{\mathbb R^d}f(x,y)\wrt x\wrt y \mbox{ in the extended sense}.\)
+	\end{itemize}
+\end{thm}
+
+Where the \emph{extended Lebesgue integral} of an extended valued (it can take the value \(+\infty\)) non-negative function \(f\) is defined by 
+\[\int f(x)\wrt x = \sup_{g}\int g(x)\wrt x.\]
+Once again, we note that the theorem is symmetric in \(x\) and \(y\), so we can establish that we may swap the order of integration provided that \(f\) is non-negative.
+
+Collectively, we refer to Theorems~\ref{Fubini} and \ref{Tonelli} together as the Fubini-Tonelli Theorem, but they are otherwise known collectively as just Fubini's Theorem.
+Often, they are used in conjunction.
+Since \(|f|\) is a non-negative function, then we may use Tonelli's Theorem and compute (or bound) the integral \(\int |f|\) via computing an iterated integral.
+If this is found to be finite, then Fubini's Theorem applies so \(f\) is integrable, and we may evaluate \(\int f\) via an iterated integral.
+
+In the context of probability, the function \(f\) which we are integrating is often positive, so Tonelli's Theorem is all that is required to justify a swap of iterated integrals.
+
+\section{Other existing literature review and the context of this thesis} 
+We have already covered some technical concepts from the literature already.
+Here, we elucidate exactly what the contribution of this thesis is.
+The focus of this thesis is on developing numerical approximations of fluid queues such that we may approximate the operator-analytic analysis of fluid-fluid queues form \cite{bo2014}.
+%The analysis of fluid-fluid queues in \cite{bo2014} relies on a certain transient analysis of the driving process, which is a fluid queue.
+
+Related to the analysis of fluid-fluid queues are the works of \cite{mz2012,lnp13,bo2013b,bop2020}.
+\cite{mz2012} analyse more general models that the fluid-fluid model and derive operator-analytic expressions for the stationary distribution of multidimensional Markov additive processes.
+Although markedly different in their approach \cite{mz2012}, like  \cite{bo2014}, is inspired by a matrix-analytic approach.
+The work of \cite{lnp13} considers a fluid-fluid queue where the driving process \(\{(X(t),\varphi(t))\}\) is level-dependent, and derives the marginal probability distribution of the first level, and bounds for that of the second level.
+The analysis of \cite{bo2014} does not consider a level-dependent driving process such as this, but I believe that it would be relatively straight forward to extend the results to the level-dependent case where the rates \(c_i\) depend on \(X(t)\) in a peicewise-constant way.
+\cite{bo2013b} analyses a model which is simple special case of the fluid-fluid queue.
+They consider a model where the second level process \(\{Y(t)\}\) has constant rates which do not depend on \(\ddot X(t)\) (but do depend on \(\varphi(t)\)), while \(\{(\ddot X(t),\varphi(t))\}\) is an \emph{unbounded} fluid queue.
+They derive matrix-analytic expressions for certain Laplace transforms with respect to the position of the first fluid \(\{X(t)\}\) relative to its starting position \(X(0)\) of the probability of certain transient events (such as first return and draining/filling times) of \(\{Y(t)\}\).
+The yet-unpublished work of \cite{bop2020} treats a similar model to that of \cite{bo2013b}, except that the driving process has a regulated lower boundary at \(0\); this significantly complicated the analysis.
+They derive matrix-analytic expressions for the first return operator of the second fluid given a specific exponential form of the initial distribution and a certain boundary condition is met.
+
+Returning now to \cite{bo2013} which introduces the fluid-fluid queue in its generality and derives operator-analytic expressions for the first-return operator of the second fluid and the stationary distribution.
+Their analysis is inspired by the analysis of classical fluid queues in \cite{bean2005} whereby matrix-analytic expressions for the fluid level are derived in terms of the generator of the driving process.
+The driving process of a fluid-fluid queue is a fluid queue.
+
+The analysis of fluid queues can be classified broadly into two approaches, matrix-analytic methods (for example, \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}) and differential equations based approaches \cite{anick1982,kk1995,blnos2022}.
+Often, it is the stationary distribution of the fluid queue which is of interest.
+In the context of the analysis of fluid-fluid queue as in \cite{bo2014}, however, we require information about the transient distribution of the fluid queue.
+Specifically, on the event that the second fluid is non-decreasing (non-increasing), we need to know the amount of fluid to have flowed into or out of the second level process, the time (or Laplace transform of time) taken to do so, and the position of the first fluid and phase at the time when the in-out process of the second fluid reaches a given height.
+Moreover, we want the resulting expressions to be readily computable.
+None of the existing literature gives exact expressions which have all the aforementioned properties.
+Even if we consider a simpler model where the rates \(r_i(x)\) are constant on intervals \(\calD_k\), then the existing results are not satisfactory.
+Perhaps the closest is the work of \cite{bean2009} who compute expressions for the Laplace transform with respect to the time take for the fluid to exit an interval, \(\calD_k\), say, on the even that fluid exits in some phase \(j\), given it started within the interval at some point \(x_0\) in some phase \(i\).
+These expressions could perhaps be used to analyse a fluid-fluid queue where the rates \(r_i(x)=r_k\) for all \(i\in\calS\), and \(x\in\calD_k\).
+Even with this simplified fluid-fluid queue, it is unclear how we could use these expressions to compute the position of the second fluid.
+
+An approximate method which is appropriate for the analysis of fluid-fluid queues has been proposed.
+The method \citep{bo2013}, which we refer to as the uniformisation method as it is derived via uniformisation arguments, approximates a fluid queue by a discrete-time Markov chain.
+\cite{bo2013} approximate the fluid queue \(\{(\dot X(t),\varphi(t))\}\) by the quasi-birth-and-death-process \(\{(L(t),\varphi(t))\}\).
+In both processes, \(\{\varphi(t)\}\) is the same.
+The level process \(\{L(t)\}\) approximates the level process \(\{\dot X(t)\}\).
+Specifically, \cite{bo2013} derive the level process such that the event \(L(t)=k\) approximates the event \(\dot X(t)\in \calD_k\), where all the levels \(\{\calD_k\}_{k}\) have the same width, \(\Delta\).
+By taking \(\Delta\to 0\), they prove that the quasi-birth-and-death-process \(\{(L(t),\varphi(t))\}\) converges to the fluid queue \(\{(\dot X(t),\varphi(t))\}\).
+Replacing the driving process of the fluid-fluid queue, \(\{(\dot X(t),\varphi(t))\}\), by the approximation \(\{(L(t),\varphi(t))\}\) and approximating the rate functions \(r_i(x)\) by constants on each interval \(\calD_k\) yields a classical fluid queue which, intuitively, approximates the fluid-fluid queue.
+To date, the uniformisation approximation has not been applied in the context of fluid-fluid queues, and its ability to approximate transient quantities of fluid queues has not been well studied.
+In Chapter~\ref{sec: numerics} we investigate the numerical performance of the uniformisation approximation and show that it is effective.
+However, of the methods considered in Chapter~\ref{sec: numerics}, its performance is the poorest and its rate of convergence is the slowest for the numerical experiments we conduct.
+Though we should not rush to discount the uniformisation approximation.
+Due to its stochastic interpretation, approximations to probabilities are guaranteed to be non-negative, we can employ probabilistic techniques to analyse the approximation (as \cite{bo2013} did), and is simplicity lends itself more easily to analysis.
+Further, when we use the uniformisation technique to approximate a fluid-fluid queue the resultant approximation is itself a fluid queue.
+We can leverage the stochastic interpretation of the approximation of the fluid-fluid queue to help analyse this approximation.
+
+It turns out that the uniformisation approximation of \cite{bo2013} is equivalent to a certain finite-element approximation which is a specific application of the \emph{discontinuous Galerkin (DG) method}.
+We introduce the discontinuous Galerkin method and its application to fluid queues, and fluid-fluid queues, in Chapter~\ref{ch:galerkin}.
+DG approximation schemes have found great success for approximating solutions of partial differential equations, however, to my knowledge, they have not been applied to fluid queues, or fluid-fluid queue to date (except for our paper \cite{blnos2022}, of which I am a co-author and the relevant parts of that work are included in this thesis, in Section~\ref{appendix: sec: 2} and Chapter~\ref{ch:galerkin}).
+The DG scheme has the advantage that is rate of convergence is rapid for smooth problems, however, it can produce oscillatory approximations and result in negative probabilities.
+
+In Chapter~\ref{sec: construction and modelling} we attempt to derive a new approximation scheme which converges faster than the uniformisation scheme and still has a stochastic interpretation so that estimates of probability are guaranteed to be positive.
+
+% Even if we assume \(r_i(x)\) is piecewise constant on intervals \(\calD_k\), say, then we still need expressions for the amount of time spent in each phase while the fluid level remains in \(\calD_k\).
+Analyses, such as \cite{bean2009}, derive expressions for the time spent in the intervals \(\calD_k\), but do not track how much time is spent in each phase.
+Moreover,
+
+% For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest.
+Anick Mitra Sondhi CITE, analysed fluid queues using a more direct differential equation-based method.
+Since Ramaswami's CITE initial work, there has been significant developments in the analysis of fluid queues CITE and related algorithms CITE.
+
+% The analysis of fluid-fluid queues of \cite{bo2014} is inspired by the matrix-analytic approach to fluid queues.
+In particular, the approach taken by Bean CITE THEIR FIRST HITTING TIME PAPER AND LATOUCHE AND DA SILVA SOARES.
+The approach requires
+
+% ...

### 2022-02-10 chapter2/chapter.tex
.

--- 2021-11-29 chapter2/chapter.tex
@@
 \chapter{Approximating fluid queues with the discontinuous Galerkin method \label{ch:galerkin}}
-% PREFACE WITH CONTRIBUTIONS 
-% TUTORIAL: HOW TO -- AS IN PEPER
-% PROBLEMS AND SOLUTIONS 
-% 	NON-NEGATIVITY AND OSCILLATIONS 
-%	FILTERING AND LIMITING 
-%	NOT A SOLUTION IN OUR CONTEXT & MOTIVATION FOR A NEW METHOD -- THE QBD-RAP
+
+In this chapter we introduce the discontinuous Galerkin (DG) method as applied to fluid queues and to approximate the operator-analytic expressions for fluid-fluid queues in \cite{bo2014}. 
+\begin{center}
+	\begin{minipage}{0.8\textwidth}
+		\textit{Apart from Section~\ref{sec: slope limiting}, this chapter has been taken from Section~4 and Appendix~1 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis.
+I am a co-author of the paper \cite{blnos2022}.
+%The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara.
+I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme.
+I contributed Sections~4.4 and 5.1.
+I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6.
+Appendix~A is also my original work.
+I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
+		}
+	\end{minipage}
+	\end{center}
+% The DG method for B 
+    % direct copy paste from the paper
+% Problems with DG, non-neg/oscillations
+    % solutions: filtering and limiting
+% Approximating R: projection, interpolation, cell averages (eluding to QBD-RAP and unif method)
+% Approximating Psi (and the stationary distribution, but this isnt really necessary)
+    % constructing D
+    % solving the matrix-Riccati equation, iteration/algorithm.
+
+% Appendix: the toy model. 
+% Appendix: some properties of B?
+
+% \section{Discontinuous Galerkin Approximation of the Generator of a Fluid Queue}
+	% \label{sec:DG}
+Discontinuous Galerkin (DG) methods can be used to approximate the solutions to systems of partial differential equations (PDEs).
+For a more thorough description of these methods see \citep{nodalDGBook}.
+The domain of approximation is partitioned into intervals, referred to individually as \textit{cells} and collectively as a \textit{mesh}.
+On each cell, we have a finite element approximation, which constructs a finite-dimensional smooth Sobolev space using piecewise-polynomial basis functions, and then projects the partial differential equations onto this space.
+This projection leads to a new system of equations, referred to as the \textit{weak form} of the original system of PDEs.
+Next, we must approximate the \textit{flux operator} which moves probability from one cell to another, in a manner similar to the underlying principle of a finite-volume approximation.
+This method conserves probability, and can handle discontinuities, such as jumps and point masses.
+Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the stationary distribution of an SFFM.
+
+\section{The Partial Differential Equation}
+We start by introducing the PDE from which we will extract the approximation to the generator \(\mathbb B\).
+
+Let $f_i(x,t)$ be the joint density of $\{(X_t, \varphi_t)\}$: 
+	% 
+	\begin{align*} 
+	    f_i(x,t) := \cfrac{\partial}{\partial x} 	\mathbb{P}\left(X_t \leq x, \varphi_t = i\right),\quad 0<x<b,\, i \in\calS,
+	\end{align*} 
+which satisfies the system of partial differential equations 
+%
+\begin{align*}
+\cfrac{\partial}{\partial t} f_i(x,t)& = \sum_{j\in \mathcal{S}}  f_j(x,t)T_{ji} - c_i \cfrac{\partial}{\partial x} f_i(x,t), \quad 0<x<b,\, i\in\mathcal S,
+\end{align*}
+% 
+subject to suitable boundary conditions. In matrix form, 
+\begin{align}
+\cfrac{\partial}{\partial t} \boldsymbol f(x,t) &= \boldsymbol f(x,t)\bs T -  \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\bs C\label{eqn:pde_density_matrix}, 
+\end{align}
+where \(\boldsymbol f(x,t) = \left(f_i(x,t)\right)_{i\in\mathcal S}\) is a row-vector. 
+% This system of PDEs is closely related to the generator \(\mathbb B\); \(\boldsymbol \mu(t)\) satisfies the operator differential equation
+% \begin{align*}
+% 	\cfrac{\wrt}{\wrt t}\boldsymbol \mu(t)(\wrt x)  = \boldsymbol \mu(t)\mathbb B(\wrt x) = \boldsymbol f(x,t)\bs T\wrt x - \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\bs C\wrt x,
+% \end{align*}
+% on the interior of the space \([0,\infty)\).
+Thus, by approximating the operator on the right-hand side of Equation (\ref{eqn:pde_density_matrix}) we can approximate the infinitesimal operator \(\mathbb B\).
+The DG method does exactly this, by approximating the operator with a matrix.
+
+\section{Cells, Test Functions, and Weak Formulation}
+To begin with, consider an unbounded first fluid level \(\{\ddot X_t,t\geq0\}\), \(\ddot X_t\in(-\infty,\infty)\).
+We will eventually truncate this space so that we have a finite dimensional approximation; however, this requires a discussion on boundary conditions which we save for later.
+Let \(\mathcal D_k = [y_k,y_{k+1}],\, k\in\mathbb Z\) partition the domain \((-\infty,\infty)\).
+We call the \(\calD_k\) \textit{cells}.
+
+On each cell \(\calD_k\) we choose \(N_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{N_k}\), compactly supported on \(\calD_k\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_k\)) to form a basis for the space \(W_k\), in which we formulate the approximation.
+Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{N_k}\) to be the space of polynomials of degree \(N_k-1\).
+It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{N_k}\) as a basis of Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature points, since our approximations inherit nice properties from this \citep{nodalDGBook}.
+However, the constructions presented here are general, and any basis can be used.
+For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{N_k}\) as the Lagrange polynomials.
+On each cell \(\mathcal D_k\) we approximate
+\[f_i(x,t)\approx u_{k,i}(x,t)=\sum\limits_{r=1}^{N_k}a_{k,i}^r(t)\phi^r_k(x),\] 
+where \(a_{k,i}^r(t)\) are yet-to-be-determined time-dependent coefficients.
+We refer to \(u_{k,i}\) as the \textit{local} approximation on \(\calD_k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_{k,i}\) on the whole domain.
+The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\).
+
+Let \(\mathcal N_k := \left\{1,\dots,N_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
+\[\boldsymbol \phi_k(x) = (\phi^r_k(x))_{r\in\mathcal N_k}, \quad \boldsymbol a_{k,i}(x) = (a_{k,i}^r(x))_{r\in\mathcal N_k},\,i\in\mathcal S.\]
+%Also define \textit{global} row-vectors 
+%\[\boldsymbol \phi(x) = (\boldsymbol\phi^k(x))_{k\in\{1,...,K\}},\quad \boldsymbol a_i(x) = (\boldsymbol a_{k,i}(x))_{k\in\{1,...,K\}},\,i\in\mathcal S.\]
+Note that we will always use the letter \(r\) to index the basis function within each cell.
+
+The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation (\ref{eqn:pde_density_matrix}).
+In general, to construct the weak-form we need a set of test functions, say \(W\).
+Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to the derivative with respect to \(x\), to get
+\begin{align}
+\begin{multlined}[t]
+	\int_{x\in\mathbb R}\cfrac{\partial}{\partial t} f_j(x,t)\psi(x)\wrt x = \int_{x\in\mathbb R} \sum_{i\in\calS}f_i(x,t)T_{ij}\psi(x)\wrt x 
+	%
+	+  \int_{x\in\mathbb R} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\psi(x)\wrt x \\{}- [ f_j(x,t)c_j\psi(x)]_{x=-\infty}^{x=\infty}, \end{multlined}\label{eqn:weak form}
+\end{align}
+for \(j\in\calS\).
+It is common to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero.
+Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE.
+For a sufficiently rich set of test functions \(W\) the weak and strong forms of the PDE are equivalent.
+Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE.
+For example, this may allow discontinuities with respect to \(x\) in the solution -- something which is ill-defined for the strong form.
+
+For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\), the same as the set of basis functions of our solution space.
+Proceeding as described above, the weak formulation is
+\begin{align*}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j(x,t)\phi^r_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i(x,t)T_{ij}\phi^r_k(x)\wrt x  
+	%
+	+  \int_{x\in\calD_k} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x \\ {}- [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
+\end{align*}
+since \(\phi^r_k\) is compactly supported on \(\calD_k\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\).
+Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi^r_k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\).
+Also, note that \(\cfrac{\wrt}{\wrt x}\phi^r_k(x)\in W\).
+Using this, we can write
+\begin{align*}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi^r_k(x)\wrt x 
+	= \int_{x\in\calD_k}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi^r_k(x)\wrt x  
+	%
+	\\ {}
+	+  \int_{x\in\calD_k} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
+\end{align*}
+	which is equivalent to
+\begin{align}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi^r_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi^r_k(x)\wrt x  
+	%
+	+  \int_{x\in\calD_k} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x \\{} - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}. \end{multlined} \label{eqn:hash}
+\end{align}
+Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_k\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), which we now substitute into (\ref{eqn:hash})
+%\begin{align*}
+%	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi^k(x)\tr{}\phi^r_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi^k(x)\tr{}T_{ij}\phi^r_k(x)\wrt x
+%	%
+%	\\ {}+  \int_{x\in\calD_k}\boldsymbol a_{k,j}(t) \boldsymbol \phi^k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - c_j[f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}.\end{multlined}
+%\end{align*}
+and repeat this for all test functions \(\phi^r_k(x)\), \(r=1,...,N_k\), to get the following system of equations,
+\begin{align}
+	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi_k(x)\tr{}T_{ij}\boldsymbol\phi_k (x)\wrt x
+	%
+	\\{}+  \int_{x\in\calD_k}\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\boldsymbol\phi_k(x)\wrt x - c_j[f_j(x,t) \boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}},\quad k\in\mathbb Z.
+\end{multlined} \label{eqn:DG system}
+\end{align}
+
+\section{Mass, Stiffness, and Flux}
+For \(k\in\mathbb Z\), define local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
+\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x.\]
+%and global mass and stiffness matrices by
+%\begin{align*}
+%	M &:= \left[\begin{array}{ccc} M_1 & & \\ & \ddots & \\ & & M_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\boldsymbol \phi(x)\wrt x,
+%	\\G &:= \left[\begin{array}{ccc} G_1 & & \\ & \ddots & \\ & & G_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\cfrac{\wrt}{\wrt x}\boldsymbol \phi(x)\wrt x,
+%\end{align*}
+We can write (\ref{eqn:DG system}) as 
+\begin{align}
+	&\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \bs M_k = \sum_{i\in\calS} \boldsymbol a_{k,i}(t)\bs M_kT_{ij} 
+	%
+	+  c_j\boldsymbol a_{k,j}(t) \bs G_k - c_j[f_j(x,t)\boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}}.\label{eqn:DG matrix}
+\end{align}
+
+It remains to approximate the \textit{flux}, \(f_j(x,t)\) at the cell edges \(y_k,\,k\in\mathbb Z\), so that we may evaluate the terms \([f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}\), \(r=1,...,N_k,\,k\in\mathbb Z\).
+This is the key for DG -- it joins the local approximations on each cell \(\calD_k\), into a global approximation on the whole domain of approximation.
+The flux is the instantaneous rate (with respect to time) at which density moves across the boundaries \(y_k,\,k\in\mathbb Z\).
+There are different choices for the flux, and we refer the reader to \citep{c99,nodalDGBook}, and references therein, for some discussion of the topic.
+Here, we choose the \textit{upwind} scheme, which, as we shall see, closely resembles the flux terms from the generator \(\mathbb B\).
+The approximate flux, also known as the \textit{numerical flux}, is given by
+\[f^*_j(x,t) = sign(c_j)\lim_{\varepsilon\to0^+}u_j(x-\varepsilon c_j,t),\]
+at each \(x=y_k,\,k\in\mathbb Z\). 
+Intuitively, the upwind flux takes the value of the density immediately on the upwind side of each \(y_k\).
+
+Denote by \(x^-\) and \(x^+\) the left and right limits at \(x\), respectively. Assume first \(c_j>0\), then 
+\begin{align*}
+	-c_j[f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}} & \approx -c_j[f_j^*(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}
+	\\& = -c_jf_j^*(y_{k+1},t)\phi^r_k(y_{k+1}) + c_jf_j^*(y_{k},t)\phi^r_k(y_{k})
+	\\& = -c_ju_j(y_{k+1}^-,t)\phi^r_k(y_{k+1}) + c_ju_j(y_{k}^-,t)\phi^r_k(y_{k})
+	\\& = -c_ju_{k,j}(y_{k+1}^-,t)\phi^r_k(y_{k+1}) + c_ju_{k-1,j}(y_{k}^-,t)\phi^r_k(y_{k})
+	\\& = -c_j\boldsymbol a_{k,j}(t)\boldsymbol \phi_k(y_{k+1}^-)\tr{}\phi^r_k(y_{k+1}) + c_j\boldsymbol a_{k-1,j}(t)\boldsymbol \phi_{k-1}(y_{k}^-)\tr{}\phi^r_k(y_{k}).
+\end{align*}
+In matrix form,  
+\begin{align*}
+	-c_j[f_j(x,t)\boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}} & \approx -c_j[f_j^*(x,t)\boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}}
+	\\& = -c_j\boldsymbol a_{k,j}(t)\boldsymbol \phi_k(y_{k+1}^-)\tr{}\boldsymbol\phi_k(y_{k+1}) + c_j\boldsymbol a_{k-1,j}(t)\boldsymbol \phi_{k-1}(y_{k}^-)\tr{}\boldsymbol\phi_k(y_{k})
+	\\& = c_j\boldsymbol a_{k,j}(t)F_j^{k,k}+c_j\boldsymbol a_{k-1,j}(t)F_j^{k-1,k},
+\end{align*}
+where, for \(j\in \calS\) with \(c_j>0\), we define \(\bs F_j^{k,k} := -\boldsymbol \phi_k(y_{k+1}^-)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z\) and \(\bs F_j^{k-1,k} := \boldsymbol \phi_{k-1}(y_{k}^-)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
+
+Now proceed similarly for \(c_j<0\) to get the approximation 
+\begin{align*}
+	-c_j[f_j(x,t)\boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}} & \approx -c_j[f_j^*(x,t)\boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}}
+	\\& = -c_j\boldsymbol a_{k+1,j}(t)\boldsymbol \phi_{k+1}(y_{k+1}^+)\tr{}\boldsymbol\phi_k(y_{k+1}) + c_j\boldsymbol a_{k,j}(t)\boldsymbol \phi_{k}(y_{k}^+)\tr{}\boldsymbol\phi_k(y_{k})
+	\\& = c_j\boldsymbol a_{k+1,j}(t)\bs F_j^{k+1,k}+c_j\boldsymbol a_{k,j}(t)\bs F_j^{k,k},
+\end{align*}
+where, for \(j\in \calS\) with \(c_j<0\), we define \(\bs F_j^{k+1,k} := -\boldsymbol \phi_{k+1}(y_{k+1}^+)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z,\) and \(\bs F_j^{k,k} := \boldsymbol \phi_{k}(y_{k}^+)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
+
+The matrices \(\bs F_j^{k-1,k},\,\bs F_j^{k,k},\) and \(\bs F_j^{k+1,k}\) are the local flux matrices.
+For convenience, we also define the matrices \(\bs F_j^{k,k+1}=0\) for \(c_j<0\) and \(\bs F_j^{k,k-1}=0\) for \(c_j>0\), \(k\in\mathbb Z\).
+
+% from which we can assemble the global flux matrices
+%\begin{align*}
+%	F_j &= \left[\begin{array}{cccccc}
+%		F_j^{1,1} & F_j^{1,2} & & & & \\
+%		 & F_j^{2,2} & F_j^{2,3} & & & \\
+%		& & \ddots & \ddots & & \\
+%		& & & \ddots & \ddots & \\
+%		& & & & F_j^{K-1,K-1} & F_j^{K-1,K} \\
+%		& & & & & F_j^{K,K}
+%	\end{array}\right],\quad c_j\geq0,
+%	%
+%	\\F_j &= \left[\begin{array}{cccccc}
+%		F_j^{1,1} & & & & & \\
+%		F_j^{2,1} & F_j^{2,2} & & & & \\
+%		& \ddots & \ddots & & & \\
+%		& & \ddots & \ddots & & \\
+%		& & & F_j^{K-1,K-2} & F_j^{K-1,K} & \\
+%		& & & & F_j^{K-1,K} & F_j^{K,K}
+%	\end{array}\right],\quad c_j<0.
+%\end{align*}
+%
+%We can now write 
+%\begin{align*}
+%	&\cfrac{\wrt}{\wrt t} \boldsymbol a_j(t) M = \sum_{i\in\calS} \boldsymbol a_j(t)MT_{ij} 
+%	%
+%	+  c_j\boldsymbol a_j(t) (G + F_j).
+%\end{align*}
+%
+%Defining \(\boldsymbol a(t) = (\boldsymbol a_j(t))_{i\in\calS}\), then 
+%\begin{align}\label{eqn:DG ODE}
+%	&\cfrac{\wrt}{\wrt t} \boldsymbol a(t)(I_{N_\calS}\otimes M) = \boldsymbol a(t){  B,}
+%\end{align}
+%where 
+%\[{  B} = \left[(T\otimes M) +
+%	\left[\begin{array}{ccc}
+%		c_1(G+F_1) & &  \\
+%		& \ddots & \\
+%		& & c_{N_\calS}(G+F_{N_\calS})\\
+%	\end{array}\right] \right],\]
+%\(\otimes\) is the kronecker product and \(I_{N_\calS}\) is the \(N_\calS\times N_\calS\) identity matrix.
+
+%So far, this DG approximation has been constructed according to the lexicographic ordering of \(\calS\times\{1,...,K\}\).
+A reordering of this construction according to the lexicographic ordering of \(\{1,...,K\}\times\calS\) helps elucidate the connection with the partitioned operator \(\mathbb B\) and eases notation slightly for the next discussion on boundary conditions.
+
+%Also define \textit{global} row-vectors 
+%\[\boldsymbol \phi(x) = (\boldsymbol\phi^k(x))_{k\in\{1,...,K\}},\quad \boldsymbol a_j(x) = (\boldsymbol a_{k,i}(x))_{k\in\{1,...,K\}},\,i\in\mathcal S.\]
+
+To write this out as a \textit{global} system, define the row-vectors 
+\[\boldsymbol a_k(t) = (\boldsymbol a_{k,i}(t))_{i\in\mathcal S},\quad \boldsymbol a(t) = (\boldsymbol a_k(t))_{k\in\mathbb Z},\]
+and the block-tridiagonal matrix 
+\begin{align*}
+% \ddot{\bs M} &= \left[\begin{array}{ccc}\ddots&&\\&\bs I_{N_\calS}\otimes \bs M_k&\\&&\ddots\end{array}\right],
+% \intertext{where \(N_\calS=|\calS|\), \(\otimes\) is the Kronecker product, and the block-tridiagonal matrix}
+\ddot{\bs B} &= \left[\begin{array}{ccccc}
+	\ddots & \ddots & \ddots & & \\
+	& \ddot{\bs B}^{k,k-1} & \ddot{\bs B}^{k,k} & \ddot{\bs B}^{k,k+1} & \\
+	& & \ddots & \ddots & \ddots 
+\end{array}\right],
+\end{align*}
+where, for \(k\in\mathbb Z\), 
+\begin{align*}
+    \ddot{{\bs B}}^{kk}&=\left[\begin{array}{ccc}T_{11}\bs I + c_1(\bs F_1^{kk}+\bs G_k)\bs M_k^{-1} & T_{12}\bs I & T_{1N_\calS}\bs I  \\ T_{21}\bs I & & \\ \vdots &\ddots & \vdots \\ & &   T_{N_\calS-1,N_\calS}\bs I \\  T_{N_\calS1}\bs I &  T_{N_\calS,N_\calS-1}\bs I & T_{N_\calS,N_\calS}\bs I +c_{N_\calS}(\bs F_{N_\calS}^{kk}+\bs G_k)\bs M_k^{-1}\end{array}\right],\\
+	\ddot{{\bs B}}^{k,k+1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k+1}\bs M_{k+1}^{-1}&  & \\  &\ddots & \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k+1}\bs M_{k+1}^{-1} \end{array}\right],\\
+	\ddot{{\bs B}}^{k,k-1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k-1}\bs M_{k-1}^{-1}&  & \\  &\ddots &  \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k-1}\bs M_{k-1}^{-1} \end{array}\right].
+\end{align*} 
+The global system of equations is 
+\begin{align}\label{eqn:DG ODE}
+	&\cfrac{\wrt}{\wrt t} \boldsymbol a(t) = \boldsymbol a(t){\ddot{{\bs B}}}.
+\end{align}
+
+\section{Boundary conditions}\label{subsec: boundary DG}
+To enable computation, this numerical approximation has to take place on a finite interval, which means we must consider a bounded domain and specify boundary conditions.
+Recall that we wish to impose lower and upper boundaries at \(0\) and \(b\), respectively.
+%In the numerics chapter, Chapter~\ref{sec: numerics}, we wish to approximate a fluid-fluid queue where the first fluid level, \(\dot X_t\), is bounded below at 0, only: in that case the first step in the approximation scheme is to approximate \(\dot X_t\) by \( X_t\).
+The truncation of \(X_t\) to \(\overline X_t\) will result in an artificial point mass at the upper bound, which we have to address properly.
+It is important to choose an \(b\) sufficiently large to control the error induced by the artificial upper bound, however, with larger \(b\) there comes increased computational burden.
+
+Let $[0,b]$ be the domain of the approximation, where $b < \infty$.
+We partition the space $[0,b]$ into \(\mathcal D_{-1}=\{0\},\) \(\mathcal D_{K+1}=\{b\},\) and \(K\) non-trivial intervals, \(\calD_k=[y_k,y_{k+1}]\setminus \{\{0\}\cup\{b\}\},\, y_k<y_{k+1},\, k\in\mathcal K^\circ\), \(y_0=0,\,y_{K+1}=b\) and define \(\Delta_k := y_{k+1}-y_k\).
+
+For states with \(c_i\leq 0\), there is the possibility of point mass accumulating at the boundary at~\(0\).
+Denote these point masses by \(q_{{-1},i}(t)\) for \(i\in\mathcal S_{-1}\).
+For states with \(c_i>0\) there is no possibility of a point mass at \(0\).
+Similarly, for \(c_i\geq 0\) there is the possibility of a point mass at \(b\).
+Denote these point masses by \(q_{{K+1},i}(t)\), for \(i\in\mathcal S_{K+1}\).
+For states with \(c_i<0\) there is no possibility of a point mass at \(b\).
+Let \(\bs q_{-1}(t)=(q_{{-1},i}(t))_{i\in\calS_{-1}}\) and \(\bs q_{K+1}(t) = (q_{{K+1},i}(t))_{i\in\calS_{K+1}}\) and \(\bs f_m(x,t) = (f_i(x,t))_{i\in\calS_m}\), \(m\in\{+,-,0\}\).
+
+Let us first consider the boundary at \( X_t=0\). The following boundary conditions describe the evolution of probability/density of a stochastic fluid model with a boundary at \(0\);
+\begin{align}\label{eqn:BC1}
+\cfrac{\wrt}{\wrt t}\bs q_{-1}(t) &= \bs q_{-1}(t) T_{{-1} {-1}} - \boldsymbol f_-(0,t)\bs C_-\bs P_{-{-1}},
+\\\label{eqn:BC2}
+\bs q_{-1}(t)T_{{-1}+}&=\bs f_+(0,t)C_+,
+\end{align}
+where \(\bs P_{-{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\).
+Equation (\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(T_{{-1},{-1}}\), and that the flux of probability density into the point masses is \(- \boldsymbol f_-(0,t)\bs P_{-{-1}}C_{-1}\).
+Substituting the DG approximation for \(\boldsymbol f_-(0,t)\) into (\ref{eqn:BC1}) gives, for \(j\in\calS_{-1}\),
+\begin{equation*}\label{eqn:DGBC1}
+\cfrac{\wrt}{\wrt t} q_{{-1},j}(t) = \sum_{i\in\calS_{-1}} q_{{-1},i}(t) T_{i j} - \sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0)\tr{}p_{ij}^{-1} c_i.
+\end{equation*}
+Equation (\ref{eqn:BC2}) describes the flux of probability mass to density upon a transition from a phase in \(\calS_{-1}\) to a phase in \(\calS_+\).
+Thus, the flux into the left-hand edge of \(\calD_0\) in phase \(j\in\calS_+\) is, \(\sum\limits_{i\in\calS_{-1}} q_{{-1},i}(t)T_{ij}\).
+Therefore, we can now evaluate
+\begin{align*}
+	-c_j[f_j(x,t)\boldsymbol\phi_0(x)]_{x=0}^{x=y_{0}} & =  -c_jf_j(y_0,t)\boldsymbol\phi_0(y_0)+c_jf_j(0,t)\boldsymbol\phi_0(0)
+	\\& \approx -c_j(f_j^*(y_0,t)\boldsymbol\phi_0(y_0)+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0)
+	\\& = c_j\boldsymbol a_{0,j}(t)F_j^{0,0}+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0),
+\end{align*}
+for \(j \in \calS_+\). 
+Thus, the DG approximation of the flux into point masses \(q_{{-1},j}(t)\) is \[-\sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0)\tr{} p_{ij}^{-1} c_i,\,j\in\calS_-,\] the rate of transition of point mass within \(\bs q_{{-1}}(t)\) is \(T_{{-1},{-1}}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0),\,j\in\calS_+\).
+
+Similarly, for the upper boundary at \(b\) the boundary conditions are 
+\begin{align*}
+\cfrac{\wrt}{\wrt t}\bs q_{K+1}(t) &= \bs q_{K+1}(t) T_{{K+1} {K+1}} + \boldsymbol f_+(b,t)\bs C_+\bs P_{+{K+1}},\\
+\bs q_{K+1}(t)T_{{K+1}-}&=-\bs f_-(b,t)C_-.
+\end{align*}
+Using the same arguments as above, 
+\begin{align*}
+\cfrac{\wrt}{\wrt t} q_{{K+1},j}(t) &= \sum_{i\in\calS_{K+1}}q_{{K+1},i}(t) T_{ij} + \sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(b)\tr{}p_{ij}^{K+1} c_i,
+\\-c_j[f_j(x,t)\boldsymbol\phi_K(x)]_{x=y_K}^{x=b} & \approx c_j\boldsymbol a_{K,j}(t)F_j^{K,K}+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b),
+\end{align*}
+for \(j\in\calS_-\). 
+Thus, the DG approximation of the flux into the point mass \(q_{{K+1},j}(t)\) is \(\sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(0)\tr{} p_{ij}^{K+1} c_i\), \(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{{K+1}}(t)\) is \(T_{{K+1},{K+1}}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b)\), \(j\in\calS_-\).
+
+To include this behaviour in the DG generator we truncate the doubly-infinite matrix \(\ddot{{\bs B}}\) at \(k=0\) and \(k=K\), then append \(|\mathcal S_{-1}|\) rows and columns to the top and left, and \(|\mathcal S_{K+1}|\) rows and columns to the bottom and right.
+These represent the point masses \(\bs q_{-1}(t)\) and \(\bs q_{K+1}(t)\), respectively.
+Given the discussion above, the truncated matrix is
+\[{{\bs B}} = \left[\begin{array}{llllll}
+	\bs T_{{-1},{-1}}& {{\bs B}}^{{-1}0} & & & & \\
+	{{\bs B}}^{0{-1}} & {{\bs B}}^{00} & {{\bs B}}^{01} & & & \\
+	& {{\bs B}}^{10} & {{\bs B}}^{11} & {{\bs B}}^{12} & & \\
+	& & \ddots & \ddots & \ddots & \\
+	& & {{\bs B}}^{K-1,K-2} &{{\bs B}}^{K-1,K-1} & {{\bs B}}^{K-1,K} & \\
+	& & &{{\bs B}}^{K,K-1} & {{\bs B}}^{K,K} & {{\bs B}}^{K,{K+1}} \\
+	& & & & {{\bs B}}^{{K+1}, K} & \bs T_{{K+1},{K+1}}
+\end{array}\right],\]
+where 
+\begin{align*}
+	{{\bs B}}^{k\ell} &= \ddot{{\bs B}}^{k\ell}, \mbox{ for \(k\in\mathcal K^\circ\), \(\ell\in\{k-1,k,k+1\}\), \(k=\ell\neq 0\) or \(k=\ell\neq K\),}
+	\\ \bs B^{00} &= \ddot{\bs B}^{00} + \left[-c_ip_{ij}^{-1} 1(c_i<0, c_j>0)\right]_{i\in\calS,j\in\calS}\otimes \bs \phi^0(0)\tr{}\bs \phi^0(0)\bs M_0^{-1},
+	\\ \bs B^{KK} &= \ddot{\bs B}^{KK} + \left[c_ip_{ij}^{-1} 1(c_i>0, c_j<0)\right]_{i\in\calS,j\in\calS}\otimes \bs \phi^K(b)\tr{}\bs \phi^K(b)\bs M_K^{-1}, 
+	\\ {{\bs B}}^{{-1}0} &:= \bs T_{{-1}+}\otimes \boldsymbol\phi^0(0), 
+	\\ {{\bs B}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1}\mathbb 1_{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}} \otimes \boldsymbol \phi^0(0)\tr{}, 
+	\\ {{\bs B}}^{{K+1} K} &:= \bs T_{{K+1}-}\otimes \boldsymbol\phi^K(b),
+	\\ {{\bs B}}^{K,{K+1}} &:= \left[c_ip_{ij}^{K+1}\mathbb 1_{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes \boldsymbol \phi^K(b)\tr{},
+\end{align*} 
+and \(\otimes\) is the Kronecker product.
+
+For future reference, we also define the matrices \({{\bs B}}^{k\ell}_{ij}\) for \(k\in\{2,\dots,K-1\}\) by
+\begin{align*}
+	{{\bs B}}^{kk}_{ij} &= \begin{cases}T_{ij}\bs I_{N_k} + c_i(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1} & i=j,\\T_{ij} \bs I_{N_k}& i\neq j,\end{cases}
+\\	{{\bs B}}^{k\ell}_{ij} &= \begin{cases}c_i\bs F_i^{k\ell}\bs M_\ell^{-1} & i=j,\\\bs 0 & i\neq j,\end{cases}\quad \ell \in \{k-1,k\}
+\end{align*}
+and
+\begin{align*}
+	{{\bs B}}^{00}_{ij} &= \begin{cases}T_{ij}\bs I_{N_k} + c_i(\bs F_i^{00}+\bs G_0)\bs M_0^{-1} & i=j,\\T_{ij}\bs I_{N_k} - 1(c_i<0,c_j>0)c_ip_{ij}^{-1} \bs F_i^{00}\bs M_0^{-1} & i\neq j,\end{cases}
+\\	{{\bs B}}^{KK}_{ij} &= \begin{cases}T_{ij}\bs I_{N_k} + c_i(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1} & i=j,\\T_{ij}\bs I_{N_K} + 1(c_i>0,c_j<0)c_ip_{ij}^{K+1} \bs F_i^{KK}\bs M_K^{-1} & i\neq j.\end{cases}
+\end{align*}
+
+After the addition of the boundary conditions, the system of ODEs (\ref{eqn:DG ODE}) can now be written as 
+\begin{align}\label{eqn: DG ODE w BCs}
+	\cfrac{\wrt}{\wrt t} \vligne{\boldsymbol q_{{-1}}(t) & {\boldsymbol a}(t) & \boldsymbol q_{K+1}(t)} 
+	% 
+	= \vligne{\boldsymbol q_{{-1}}(t) & {\boldsymbol a}(t) & \boldsymbol q_{K+1}(t)} \bs B.
+\end{align}
+
+Approximations \( \bs B^{mn}_{ij}\), \( \bs B_{ij}\), and \( \bs B^{mn}\) to \(\mathbb B^{mn}_{ij}\), \(\mathbb B_{ij}\), and \(\mathbb B^{mn}\), \(i,j\in\calS,\,m,n\in\{+,-,0\}\), are constructed from the block-matrices \({  \bs B}^{k\ell}_{ij}\), \(i,j\in\calS\), \(k,\ell\in\mathcal K\), as
+\begin{align*}
+	{  \bs B}_{ij}^{m n} &= \left[{  \bs B}_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\quad i,j\in\calS,\,m,n\in\{+,-,0\},
+%\intertext{an approximation \({  B}_{ij}\) to \(\mathbb B_{ij}\), \({i,j\in\calS},\) is}
+\\	{  \bs B}_{ij} &= \left[{  \bs B}_{ij}^{k\ell}\right]_{k,\ell\in\mathcal K},\,{i,j\in\calS},
+%\intertext{and an approximation \({  B}^{mn}\) to \(\mathbb B^{mn}\), \(m,n\in\{+,-,0\}\) is}
+\\	{  \bs B}^{m n} &= \left[\left[{  \bs B}_{ij}^{k\ell}\right]_{i\in\calS_k^m,j\in\calS_\ell^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n},\,m,n\in\{+,-,0\}.
+\end{align*}
+% where \(\bs B = \widehat{{\bs B}}\widehat{\bs M}^{-1}\) and \(\widehat{\bs M} = \left[\begin{array}{ccccc}
+% 	\bs I_{|\calS_{-1}|} & & & & \\
+% 	& \bs I_{N_S}\otimes \bs M_1 & & & \\
+% 	& & \ddots & & \\
+% 	& & & \bs I_{N_S}\otimes \bs M_K & \\
+% 	& & & & \bs I_{|\calS_{K+1}|}
+% \end{array}\right]\), that is, 
+% \[\bs B = \left[\begin{array}{llllll}
+% 	\bs T_{{-1},{-1}}& \widehat{{\bs B}}^{{-1}1}\bs M_1^{-1} & & & & \\
+% 	\widehat{{\bs B}}^{1{-1}} & \ddot{{\bs B}}^{11}\bs M_1^{-1} & \ddot{{\bs B}}^{12}\bs M_2^{-1} & & & \\
+% 	& \ddot{{\bs B}}^{21}\bs M_1^{-1} & \ddot{{\bs B}}^{22}\bs M_2^{-1} & \ddot{{\bs B}}^{23}\bs M_3^{-1} & & \\
+% 	& & \ddots & \ddots & \ddots & \\
+% 	& & \ddot{{\bs B}}^{K-1,K-2}\bs M_{K-2}^{-1} &\ddot{{\bs B}}^{K-1,K-1}\bs M_{K-1}^{-1} & \ddot{{\bs B}}^{K-1,K}\bs M_{K}^{-1} & \\
+% 	& & &\ddot{{\bs B}}^{K,K-1}\bs M_{K-1}^{-1} & \ddot{{\bs B}}^{K,K}\bs M_{K}^{-1} & \widehat{{\bs B}}^{K,{K+1}} \\
+% 	& & & & \widehat{{\bs B}}^{{K+1}, K}\bs M_{K}^{-1} & \bs T_{{K+1},{K+1}}
+% \end{array}\right].\]
+
+% Regarding our notational convention, we use regular mathematics fonts to represent DG approximations to operators, i.e.~\(\bs B\) is a DG approximation to \(\mathbb B\) and \(\bs \Psi\) is an approximation to \(\mathbb \Psi\).
+
+We prove the following result in Appendix\ref{sec:properties}.
+%This transformation is not strictly necessary and all the following calculations can be done with either form, as long as week keep a consistent interpretation of the DG generator and related coefficients.
+However, it is convenient to prove properties of the generator.
+\begin{cor}
+	The approximate generator \( \bs B\) conserves probability. That is, for all \(t\geq 0\),
+	\begin{align*}
+	\begin{multlined}[t]\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS} \int_{x\in[0,b]}u_i(x,t)\wrt x 
+	%
+	\\= \sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i\in\calS} \int_{x\in[0,b]}u_i(x,0)\wrt x.\end{multlined}
+	\end{align*}
+\end{cor}
+
+\section{Application to an SFFM}\label{sec:DGSFFM}
+Given our approximation \(\bs B\) to the generator \(\mathbb B\) we now follow the recipe from \cite{bo2014}, replacing the actual generator \(\mathbb B\) with its approximation \({  \bs B}\), to construct approximations, \(\bs \pi\) and \(\bs p\), to the stationary operators, \(\bbpi\) and \(\mathbb p\).
+
+It may be convenient to think of our approximations in terms of approximations of kernels.
+Recall that the operators in \citep{bo2014} can be thought of in terms of kernels.
+That is, for some function \(\bs g = (g_i(x))_{i\in\calS}\), we can write \(\bs \mu \mathbb B \boldsymbol g\tr{} = \displaystyle \sum\limits_{k,\ell\in\mathcal K}\sum\limits_{i,j\in\calS} \displaystyle \int_{x,y}\wrt \mu_i(x) \mathbb B_{ij}^{k\ell}(x,\wrt y)g_j(y)\) where \(\mathbb B_{ij}^{k\ell}(x,\wrt y)\) is the kernel of the operator \(\mathbb B_{ij}^{k\ell}\).
+
+Let \(\boldsymbol a_{-1}(t):=\boldsymbol q_{-1}(t)\) and \(\boldsymbol a_{K+1}(t):=\boldsymbol q_{K+1}(t)\), and define basis functions \(\bs\phi_{-1}(x) = \phi_{-1}^1(x) = \delta(x)\) and \(\bs\phi_{K+1}(x) = \phi_{K+1}^1(x) = \delta(x-b)\), where \(\delta\) is the Dirac delta, \(N_{-1} = N_{K+1} = 1\), and \(\mathcal N_{-1} = \mathcal N_{K+1} = \{1\}\).
+Also define \({\bs M}_{-1}=\bs I_{|\calS_{-1}|}\), \({\bs M}_{K+1}=\bs I_{|\calS_{K+1}|}\), the block-diagonal matrix \(\bs M=diag(\bs M_k,k\in\mathcal K)\), and row-vectors
+\[\boldsymbol \phi(x) = (\bs\phi_k(x))_{k\in\mathcal K}, \quad \boldsymbol a_i(t) = (\bs a_{k,i}(t))_{k\in\mathcal K},\,i\in\calS.\]
+
+To pose the approximation \(\bs B\) in kernel form let \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in W,\,i\in\calS\) be the initial density of the process, and \(\boldsymbol \phi(x)\bs b_i\tr{}\in W,\,i\in\calS\) be a test function.
+Observe that, from our DG construction earlier and the definition of \({\bs M}\),
+\[\sum_{i,j\in\calS}\int_{x,y\in[0,b]} \boldsymbol a_i \boldsymbol \phi(x)\tr{} \boldsymbol \phi (x) \wrt x {\bs M}^{-1}{  \bs B}_{ij} \bs \phi(y)\tr{} \bs \phi(y)\bs b_j \wrt y= \sum_{i,j\in\calS} \boldsymbol a_i {  \bs B}_{ij} {\bs M} \bs b_j .\]
+Thus, we can think of 
+\[\boldsymbol \phi (x) {\bs M}^{-1}{  \bs B}_{ij} \bs \phi(y)\tr{}\wrt y,\]
+as an approximation to the kernel \(\mathbb B_{ij}(x,\wrt y)\). This concept can be extended to all the approximations of operators considered in this work.
+
+\subsection{Approximating the operator \(\mathbb R\)}\label{sec: approx r}
+Recall the operator \(\mathbb R^k\) from Lemma~\ref{lemma: D(s)}.
+Essentially, the operator \(\mathbb R^k\) takes an initial measure \(\boldsymbol \mu_k\) and multiplies each element by \(1/|r_i(x)|\) on cells \(\calD_k\) where \(r_i(x)\neq 0\).
+In the context of DG the initial distribution is given by \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in W,\,i\in\calS\).
+Thus, for \(k\in\mathcal K\) such that \(r_i(x)\neq0\) on \(\calD_k\), we have
+\[\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}\mathbb R^k_i = \cfrac{\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}}{|r_i(x)|}.\]
+Decompose the right-hand side into a component which lies in \(W\) and another orthogonal to \(W\): 
+\[\cfrac{\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}}{|r_i(x)|} = \bs \rho_{k,i} \bs \phi_k(x)\tr{} + g_i^\perp(x),\] where \(\bs \rho_{k,i} \bs\phi_k(x)\tr{}\in W\), \(g_i^\perp \in W^\perp\).
+Now, multiply by test functions \(\{\phi^r_k(x)\}_{r=1}^{N_k}\) and integrate over \([0,b]\):
+\begin{align*}
+	\boldsymbol a_{k,i}\int_{x\in[0,b]} \cfrac{ \boldsymbol \phi_k(x)\tr{}\bs \phi_k(x)}{|r_i(x)|}\wrt x
+	&=\bs \rho_{k,i} \int_{x\in[0,b]}\bs \phi_k(x)\tr{}\bs \phi_k(x)\wrt x + \int_{x\in[0,b]} g_i^\perp(x)\bs\phi_k(x)\wrt x 
+	%
+	\\&= \bs \rho_{k,i} \int_{x\in[0,b]}\bs \phi_k(x)\tr{}\bs \phi_k(x)\wrt x = \bs \rho_{k,i}\bs M_k,
+\end{align*}
+since \(g_i(x)^\perp\in W^\perp\). Define the matrix \(\bs M_k^r := \displaystyle\int_{x\in[0,b]} \cfrac{ \boldsymbol \phi_k(x)\tr{}\bs \phi_k(x)}{|r_i(x)|}\wrt x\), then 
+\(
+	\boldsymbol a_{k,i}\bs M_k^r
+	= \bs \rho_{k,i} M_k,
+\)
+which implies
+\(
+	\bs \rho_{k,i}  = \boldsymbol a_{k,i}\bs M_k^r\bs M_k^{-1}.
+\)
+Thus, we have the approximation 
+\[\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}\mathbb R^k_i = \cfrac{\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}}{|r_i(x)|}\approx \boldsymbol a_{k,i}\bs M_k^r\bs M_k^{-1}\bs\phi_k(x)\tr{}.\]
+Since \(\boldsymbol a_{k,i}\) is arbitrary, we see that we approximate \(\mathbb R_{k,i}\) by \(  \bs R_{k,i} = \bs M_k^r\bs M_k^{-1},\)
+and \(\mathbb R^k\) by \(  \bs R^k = \diag(  \bs R_{k,i})_{i\in\calS^\bullet_k}\).
+
+In practice, we implement a Gauss-Lobatto quadrature approximation to compute the elements of \(\bs M_k^r\).
+
+\begin{rem}
+	We could also use interpolation to approximate \(\mathbb R\). 
+\end{rem}
+
+\subsection{Approximating the operator \(\bs D\) and the Riccati equation}
+Recalling Lemma~\ref{lemma: D(s)} and replacing the operators \(\mathbb R^k\) and \(\mathbb B^{\ell m}\), by their approximations we have the following approximation to \(\mathbb D^{mn}(s)\)
+\begin{align*}
+		 {\bs D}^{mn}(s) = \left[ \bs {R}^{m}\left(
+		{ {\bs B}}^{mn} - s{\bs I} + { {\bs B}}^{m0 }\left({ {\bs B}}^{00}- s\bs I \right)^{-1} { {\bs B}}^{0n}\right)\right],\quad m,n \in\{+,-\}.
+\end{align*}
+
+Let \(\bs\phi_k(x){\bs M}_k^{-1}\bs \Psi_{ij}^{k\ell}(s)\bs\phi_\ell(y)\tr{}\wrt y\), \(i,j\in\calS,\) \(k\in\mathcal K_i^+\,, \ell\in\mathcal K_j^-\) be a finite-dimensional approximation of the operator kernel \(\mathbb\Psi_{ij}^{k\ell}(s)(x,\wrt y)\), where \(\bs \Psi_{ij}^{k\ell}(s)\) is a matrix of constants for a given \(s\).
+Construct an approximation to \(\mathbb\Psi(s)(x,\wrt y)\) by
+\[\bs\phi^+(x){\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt y = \bs \phi^+(x)\bs M_+ \left[\left[\bs \Psi_{ij}^{k\ell}\right]_{i\in\calS_k^+,j\in\calS_\ell^-}\right]_{k\in\mathcal K^+,\ell\in\mathcal K^-}\bs \phi^-(y)'\wrt y\]%\left[\left[\bs\phi^k(x)\bs {M}_k^{-1}\bs \Psi_{ij}^{k\ell}(s)\bs\phi^\ell(y)\tr{}\wrt y\right]_{i\in\calS_k^+,j\in\calS_\ell^-}\right]_{ k\in\mathcal K^+, \ell\in\mathcal K^-},\]
+where \(\bs\phi^+(x) = (\bs\phi_k(x))_{i\in\calS_k^+,k\in\mathcal K^+}\) and \(\bs\phi^-(y) = (\bs\phi_k(y))_{i\in\calS_k^-,k\in\mathcal K^-}\) are row-vectors, \(\bs \Psi(s)\) is a matrix of constants for a given \(s\) with the same size as \(\bs D^{+-}\), and \({\bs M}_m,\) \(m\in\{+,-,0\}\) is a block diagonal matrix \({\bs M}_m = \diag\left( \left( \bs M_k\right)_{i\in\calS_k^m}\right)_{k\in\mathcal K^m}\), \(m\in\{+,-,0\}\).
+Now replace the theoretical kernels in Theorem~\ref{theo:Psi} by their DG approximations to get
+\begin{align*}
+&\bs\phi^+(x){\bs M}_+^{-1}  \bs D^{+-}(s)\bs\phi^-(y)\tr{}\wrt y
+\\&{}+ \int_{z_1,z_2}\bs\phi^+(x){\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(z_1)\tr{}\bs\phi^-(z_1) {\bs M}_-^{-1}  \bs D^{-+}(s)\bs\phi^+(z_2)
+\\&{}\quad\times\bs\phi^+(z_2){\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt z_1\wrt z_2\wrt y
+\\&{}+ \int_{z_1}\bs\phi^+(x){\bs M}_+^{-1}  \bs D^{++}(s)\bs\phi^+(z_1)\tr{}\bs\phi^+(z_1) {\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt z_1\wrt y
+\\&{}+ \int_{z_1}\bs\phi^+(x){\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(z_1)\tr{}\bs\phi^-(z_1){\bs M}_-^{-1}  \bs D^{--}(s)\bs\phi^-(y)\tr{}\wrt z_1\wrt y
+= 0.
+\end{align*}
+Multiplying on the left by \(\bs\phi^+(x)\tr{}\) and on the right by \(\bs\phi^-(y)\), integrating over both \(x\) and \(y\), then post-multiplying by \({\bs M}^{-1}_-\) gives the following matrix Ricatti equation
+\begin{align}\label{eqn:RiccatiPsi}
+    \bs D^{+-}(s)
++ \bs \Psi(s)   \bs D^{-+}(s)\bs \Psi(s)
++   \bs D^{++}(s)\bs \Psi(s)
++ \bs \Psi(s)  \bs D^{--}(s)
+= \bs 0.
+\end{align}
+Thus, we may find \(\bs \Psi(s)\) by solving \eqref{eqn:RiccatiPsi}, using one of the methods in \citep{bot08}. Here, we use the Newtons method.
+
+\begin{rem}
+	Given the stochastic interpretation of \(\mathbb\Psi(0)\) we know that \( \bs \mu^+ \mathbb\Psi(0)([0,\infty))=1\) for every vector of measures \( \bs \mu^+\) such that \( \bs \mu^+([0,\infty)\boldsymbol 1 = 1\), when an SFFM is recurrent.
+It appears that this result carries over to the matrix \(\bs \Psi(0)\) giving the property that \(\displaystyle\int_{y\in[0,b]} \bs \Psi(0)\bs\phi^-(y)\tr{}\wrt y = \bs 1\).
+However, we have only observed this numerically and have no proof of this property.
+\end{rem}
+
+\subsection{Putting it all together: constructing an approximation to the stationary distribution}
+We find an approximation to the stationary distribution by replacing the theoretical operators in Theorem~\ref{theo:density} with their approximations.
+Table \ref{table:notations} defines the notation we use for the DG approximations to stationary operators.
+
+ \begin{table}[h!]
+ \centering
+ \begin{tabular}{c|c|c|c}
+	\begin{tabular}{c}Exact \\operator\end{tabular} & Operator indices & \begin{tabular}{c}Approximation \\ notation\end{tabular} & Approximations \\\hline 
+	%
+	\( \bbxi_{k,i} \) & \(i\in\calS_k^-,\,k\in\mathcal K^-\)  & \(\bs \xi_{k,i} := (\xi_{k,i}^r)_{r\in\mathcal N_k}\) & 
+	\(%\begin{array}{c}
+	\bbxi_{k,i}(\wrt x)\approx  \bs{\xi}_{k,i} \bs \phi^k(x)\tr{}\wrt x,
+	 %\\\bbxi_i^{-1}(\{0\}) %:=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
+	%\approx \bs\xi_{i}^{-1},
+	%\\\bbxi_i^{K+1}(\{b\}) %:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]
+	%\approx \bs\xi_{i}^{K+1}.%\end{array}
+	\)
+	 \\\hline
+%
+%
+	\(\mathbb p_{k,i}\) & \(\begin{array}{c}i\in\calS_k^-\cup\calS_k^0,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_m\end{array}\) & \(\bs p_{k,i} := (p_{k,i}^r)_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
+	\mathbb p_{k,i}(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+	\approx\bs{p}_{k,i} \bs\phi_k(x)\tr{}\wrt x
+	%\\ \mathbb p_{i}^{-1}(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
+	%\approx \bs p_{i}^{-1},\\ \mathbb p_{i}^{K+1}(\{b\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = b, \varphi_{t} = i\right]
+	%\approx \bs p_{i}^{K+1}.% \end{array}
+	\)\\\hline
+%
+%
+	 \(\bbpi_{k,i}(y)\)  & \(\begin{array}{c}i\in\calS,\\k\in\mathcal K\end{array}\) & \(\bs \pi_{k,i}(y) := (\pi_{k,i}^r(y))_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
+	 \bbpi_{k,i}(y)(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+	\approx\bs{\pi}_{k,i}(y) \bs \phi_k(x)\tr{}\wrt x
+	%\\ \bbpi_{i}^{-1}(y)(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
+	%\approx \bs \pi_{i}^{-1}(y),\\ \bbpi_{i}^{K+1}(y)(\{b\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = b, \varphi_{t} = i\right]
+	%\approx \bs \pi_{i}^{K+1}(y). \end{array}
+	\)\\\hline
+ \end{tabular}
+ \caption{Notation for the approximation of the stationary operators of an SFFM.
+The first column contains the operators which we are approximating, the second column contains indices for which the operators are defined, the third column defines the notation we use for the coefficients of the approximation, and the last column defines how the coefficients and basis functions are used to approximate the operators.
+\label{table:notations}}
+ \end{table}
+
+%  \begin{table}[h!]
+% \centering
+% \begin{tabular}{c|c|c|c}
+%	\begin{tabular}{c}Exact \\Operator\end{tabular} & Notation & Set of indices & Approximation \\\hline 
+%	%
+%	\( \bbxi_{k,i}(\wrt x) \) & \(\xi_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\mathcal K_i^-\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(%:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\\
+%	 \displaystyle\sum\limits_{r=1}^{N_k} {\xi}_{k,i}^r \phi^r_k(x)\wrt x,\,x\in\calD_k.\)\\\hline
+%
+%	&\(\xi_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\mathcal K_i^-\cap \{{-1}\cup{K+1}\}.\end{array}\) & \(\begin{array}{l}\bbxi_{k,i}(\{0\}) %:=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
+%	\approx \xi_{i,r}^{-1},
+%	\\\bbxi_{k,i}(\{b\}) %:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]
+%	\approx \xi_{i,r}^{K+1}.\end{array}\)\\\hline 
+%
+%	&\(p_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(\mathbb p_{k,i}(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+%	\approx\displaystyle\sum\limits_{r=1}^{N_k} {p}_{k,i}^r \phi^r_k(x)\wrt x,\,x\in\calD_k.\)\\\hline
+%
+%	&\(p_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\}.\end{array}\) & \(\begin{array}{l} \mathbb p_{k,i}(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
+%	\approx p_{i,r}^{-1},\\ \mathbb p_{k,i}(\{b\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = b, \varphi_{t} = i\right]
+%	\approx p_{i,r}^{K+1}. \end{array}\) \\\hline
+%
+%	&\(\pi_{k,i}^r(y)\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(\bbpi_{k,i}(y)(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+%	\approx\displaystyle\sum\limits_{r=1}^{N_k} {\pi}_{k,i}^r(y) \phi^r_k(x)\wrt x,\,y>0,\, x\in\calD_k.\)\\\hline
+% \end{tabular}
+% \caption{Notation for the approximation of the stationary operators of an SFFM.\label{table:notations}}
+% \end{table}
+
+With the notation in Table \ref{table:notations} define row-vectors 
+ \begin{align*}
+	%{\bs{\xi}}_{k,i} &= ( \xi_{k,i}^r)_{r\in\{1,...,N_k\}}, \quad  { i\in\calS,\,k\in\mathcal K_i^-},
+	%
+	{\bs{\xi}}_k &:= ( \bs\xi_{k,i})_{i\in\calS_k^-}, \quad  {k\in\mathcal K_i^-},
+	%
+	\\ {\bs{\xi}}& := ( \bs \xi_k)_{k\in\mathcal K^-},
+	% 
+	%\\{\bs{p}}_{k,i} &= (  p_{k,i}^r)_{r\in\{1,...,N_k\}}, \quad  {i\in\calS,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m},
+	%
+	\\\bs{p}^{m}_k &:= (  \bs p_{k,i})_{i\in\calS_k^m}, \quad  k\in{\mathcal K^m},\, m\in\{-,0\},
+	%
+	\\ {\bs{p}^m} &:= (  \bs p^{m}_k)_{k\in\mathcal K^m},\quad m\in\{-,0\},
+	%
+	\\ {\bs{p}} &:= (  \bs p^m)_{m\in\{-,0\}},
+	%
+	%\\{\bs{\pi}}_{k,i}(y) &= (  \pi_{k,i}^r(y))_{r\in\{1,...,N_k\}},\quad  i\in\calS,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m,
+	%
+	\\{\bs{\pi}}^{k}_m(y) &:= (  \bs \pi_{k,i}(y))_{i\in\calS_k^m},\quad  k\in\mathcal K,\,m\in\{+,-,0\},
+	%
+	\\{\bs{\pi}}^m(y) &:= (\bs \pi^{k}_m(y))_{k\in\mathcal K^m},\quad  m\in\{+,-,0\},
+	%
+	\\ {\bs{\pi}}(y) &:= (  \bs \pi^m(y))_{m\in\{+,-,0\}}.
+ \end{align*}
+
+Proceeding similarly to the derivation of the Ricatti equation (\ref{eqn:RiccatiPsi}), we can argue that the coefficients \( {\boldsymbol{\xi}}\) are the solution to the matrix system 
+% 	 
+	\begin{align*}
+		\vligne{ {\boldsymbol{\xi}}  & \boldsymbol{0}}\left(-\left[\begin{array}{ll} 
+			 { {\bs B}}^{--} &  { {\bs B}}^{-0} \\
+                         { {\bs B}}^{0-} &  { {\bs B}}^{00} 
+		\end{array} \right]\right)^{-1}\left[\begin{array}{l} 
+			 { {\bs B}}^{-+} \\ 
+			 { {\bs B}}^{0+}
+		\end{array} \right]\bs \Psi(0) & =  {\boldsymbol{\xi}}, \\ 
+		\int_{x\in[0,b]} {\bs \xi}\left[\begin{array}{c}\bs \phi^-(x)\tr{} \\ \bs \phi^0(x)\tr{}\end{array}\right]\wrt x \boldsymbol 1 & = 1. 
+	\end{align*} 
+Essentially, we replace the theoretical operators in (\ref{eqn:xi1}) and (\ref{eqn:xi2}) with their DG counterparts.
+
+Similarly, the coefficients \( {\boldsymbol{p}}\) are given by 
+	\begin{equation}\vligne{\bs{p}^{-}  & \bs{p}^{0}} = z \vligne{{\bs\xi} & \bs{0}} 
+	\left(-\left[\begin{array}{ll} 
+		{\bs B}^{--} & {\bs B}^{-0} \\
+		{\bs B}^{0-} & {\bs B}^{00} 
+		\end{array} \right] \right)^{-1},\label{eqn:pisystem1}\end{equation}
+		where \(z\) is a normalising constant. The coefficients \(\bs \pi(y)\) are given by 
+\begin{align} 
+	\;  {\bs{\pi}}^{0}(y) &= \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)}\left[\begin{array}{l} { {\bs B}}^{+0} \\ { {\bs B}}^{-0} \end{array} \right]\left(-{ {\bs B}}^{00}\right)^{-1}, \\ 
+	% 
+	 \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)}& = \vligne{ {\bs{p}}^{-} &  {\bs{p}}^{0}}\left[\begin{array}{l} { {\bs B}}^{-+} \\ { {\bs B}}^{0+} \end{array} \right]\vligne{e^{ {\bs K}y} & e^{ {\bs K}y}\bs \Psi(0)}\left[\begin{array}{cc}  {\bs R}^{+} & 0 \\ 0 &  {\bs R}^{-}\end{array}\right], \\
+	%  
+	 \sum_{i \in \mathcal{S}}\sum_{k \in \mathcal K} & \int_{y = 0}^{\infty} \int_{x \in [0,b]}  \bs \pi_{k,i}(y)\bs\phi_k(x)\tr{}\wrt x\wrt y 
+	 %
+	 \\&{}+  \sum_{i \in \mathcal{S}}\sum_{\ell \in \{-,0\}}\sum_{k\in \mathcal K_i^\ell}  \int_{x \in [0,b]}  \bs p_{k,i}\bs\phi_k(x)\tr{}\wrt x = 1,\label{eqn:pisystem2}
+	\end{align}
+	% 
+	where $ {\bs K} :=  {\bs D}^{++}(0) + \bs \Psi(0) {\bs D}^{(-+)}(0)$, and $z$ is a normalising constant.
+
+%For \(i,k,r\) such that \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\mathcal K_i^-\setminus \{{-1}\cup{K+1}\}\), define coefficients \(  \xi_{k,i}^r\) to be determined later such that 
+%\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}} {\xi}_{k,i}^r \phi^r_k(x),\quad x \in \calD_k,\]
+%is an approximation to the density of \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\mathcal K_i^-\cap \{{-1}\cup{K+1}\},\) define point masses \(  \xi_{k,i}^r\) such that \( \xi_{i,r}^{-1}\) and \( \xi_{i,r}^{K+1}\) approximate the point masses \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]\) and \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]\), respectively.
+In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\mathcal K_i^-\), let \( {\bs{\xi}}_{k,i} = ( \xi_{k,i}^r)_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\mathcal K_i^-\), let \( {\bs{\xi}}^k = ( \xi_{k,i})_{\{i\in\calS\mid r_i(x)<0,x\in\calD_k\}}\) and let \( {\bs{\xi}} = ( \xi^k)_{k\in\bigcup\limits_{i\in\calS}\mathcal K_i^-}\).
+%
+%Proceeding similarly to the derivation of the Ricatti equation (\ref{eqn:RiccatiPsi}), we can argue that the DG approximation to \( {\boldsymbol{\xi}}\) is the solution to the matrix system 
+%% 	 
+%	\begin{align}
+%		\vligne{ {\boldsymbol{\xi}}  & \boldsymbol{0}}\left(-\left[\begin{array}{ll} 
+%			 { {B}}^{--} &  { {B}}^{-0} \\
+%                         { {B}}^{0-} &  { {B}}^{00} 
+%		\end{array} \right]\right)^{-1}\left[\begin{array}{l} \label{eqn:xi1}
+%			 { {B}}^{-+} \\ 
+%			 { {B}}^{0+}
+%		\end{array} \right]\Psi(0) & =  {\boldsymbol{\xi}}, \\ 
+%		\int_y  {\bs \xi}\left[\begin{array}{c}\bs \phi^-(x)\tr{} \\ \bs \phi^0(x)\tr{}\end{array}\right]\wrt x \boldsymbol 1 & = 1. \label{eqn:xi2}
+%	\end{align} 
+%Essentially, we replace the theoretical operators in (\ref{eqn:xi1}) and (\ref{eqn:xi2}) with their DG counterparts. 
+%%	We write 
+%%	\( {\bs\xi} = \vligne{ {\bs\xi}_{-1} &  {\bs\xi}^\circ &  {\bs\xi}_{K+1}}\), where \( {\bs\xi}_{-1}=( {\xi}_{{-1},i})_{\{\i\in\calS_{-1}\mid r_i(0)<0\}}\) and \( {\bs\xi}_{K+1}=( {\xi}_{{K+1},i})_{\{\i\in\calS_{K+1}\mid r_i(0)<0\}}\) are associated with the boundaries at \(0\) and \(b\), respectively, and \( {\bs\xi}^\circ\) is associated with the interior \(x\in(0,b)\).
+Let us denote the elements of \( {\bs{\xi}}^\circ\) as \( {\xi}_{k,i}^r\), \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\mathcal K_i^-\) where the elements are ordered by \(k\) then \(i\) and then \(r\).
+%%	We construct an approximation to the density of \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\) as \(\displaystyle\sum\limits_{\substack{k\in\mathcal K_i^- \\r\in\{1,...,N_k\}}} {\xi}_{k,i}^r \phi^r_k(x)\).
+An approximation to \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]\) is \( {\xi}_{{-1},i}\) and an approximation to the artificial point mass \(\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]\) is \( {\xi}_{{K+1},i}\).
+%	
+%	Next, we seek to approximate the stationary operators \(\bs p\) defined in (\ref{eqn:jointmass}).
+For \(i,k,r\) such that \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}\), coefficients \(p_{k,i}^r\) to be determined later, such that
+%\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}}p_{k,i}^r \phi^r_k(x),\quad x \in \calD_k,\]
+%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\},\) define point masses \(p_{k,i}^r\) such that \(p_{i,r}^{-1}\) and \(p_{i,r}^{K+1}\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} = b, \varphi_{t} = i\right]\), respectively.
+In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\), let \( {\bs{p}}_{k,i} = (  p_{k,i}^r)_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\), let \( {\bs{p}}^k = (  p_{k,i})_{\{i\in\calS\mid r_i(x)\leq 0,x\in\calD_k\}}\) and let \( {\bs{p}} = (  p^k)_{k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m}\).
+Again, by proceeding similarly to the derivation of (\ref{eqn:RiccatiPsi}), that the DG approximation to \( {\boldsymbol{p}}\) is
+%	\[\vligne{\bs{p}^{-}  & \bs{p}^{0}} = z \vligne{{\bbxi} & \bs{0}} 
+%	\left(-\left[\begin{array}{ll} 
+%		{B}^{--} & {B}^{-0} \\
+%		{B}^{0-} & {B}^{00} 
+%		\end{array} \right] \right)^{-1},\]
+%		where \(z\) is a normalising constant. 
+%That is, we substitute the DG approximations of the operators into the corresponding equation in Theorem~\ref{theo:density}.
+%%	
+%%	For \(i\in\calS_{-1}, m\in\{-,0\}\) define \(  p^m_{{-1},i}\) as the approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, Y_t=0, \varphi_{t} = i\right]\).
+%%	
+%%	 \( {\bs p}^-_{{-1}}=(  p^-_{{-1},i})_{\{i\in\calS_{-1}\mid r_i(0)<0\}}\) and \( {\bs p}^0_{{-1}}=(  p^0_{{-1},i})_{\{i\in\calS_{-1}\mid r_i(0)=0\}}\). 
+%%	Now define \( {\bs p}^m_{{-1}},\, {\bs p}^m_{\circ},\, {\bs p}^m_{{K+1}}\) as the approximation to \(\bs p\) at the left-hand boundary, interior, and on the right-hand boundary of \([0,b]\), respectively.
+%%	\[ {\bs p}^m := \vligne{ {\bs p}^m_{-1}(y) &  {\bs p}^m_\circ(y) &  {\bs p}^m_{K+1}(y)},\quad m\in\{-,0\},\]
+%%where Specifically,  
+%
+%	
+%	Next we approximate the operator \({{\bbpi}}(y)\).
+For \(i,k,r\) such that \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}\), define \(\pi_{k,i}^r(y)\), to be determined later, such that
+%\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}}\pi_{k,i}^r(y) \phi^r_k(x),\quad x \in \calD_k,\]
+%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in\wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\}\) define point masses \(\pi_{k,i}^r(y)\) such that \(\pi_{i,r}^{-1}(y)\) and \(\pi_{i,r}^{K+1}(y)\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t\in\wrt y, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in\wrt y, X_{t} = b, \varphi_{t} = i\right]\), respectively.
+In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\), let \( {\bs{\pi}}_{k,i}(y) = (  \pi_{k,i}^r(y))_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\), let \( {\bs{\pi}}^k(y) = (  \pi_{k,i}(y))_{\{i\in\calS\mid r_i(x)\leq 0,x\in\calD_k\}}\) and let \( {\bs{\pi}}(y) = (  \pi^k(y))_{k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m}\).
+Again, we can argue that the DG approximation to \( {\boldsymbol{\pi}}(y)\) is given by substituting the DG operators for their theoretical counterparts in Theorem~\ref{theo:density}
+%	\begin{align} 
+%	& \;  {\bs{\pi}}^{0}(y) = \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)}\left[\begin{array}{l} { {B}}^{+0} \\ { {B}}^{-0} \end{array} \right]\left(-{ {B}}^{00}\right)^{-1}, \label{eqn:pisystem1} \\
+%	% 
+%	&  \vligne{ {\bs{\pi}}^{+}(y) &  {\bs{\pi}}^{-}(y)} = \vligne{ {\bs{p}}^{-} &  {\bs{p}}^{0}}\left[\begin{array}{l} { {B}}^{-+} \\ { {B}}^{0+} \end{array} \right]\vligne{e^{ {K}y} & e^{ {K}y}\Psi(0)}\left[\begin{array}{cc}  {R}^{+} & 0 \\ 0 &  {R}^{-}\end{array}\right], \\
+%	%  
+%	 &\sum_{r \in \{{-1},1,...,K,{K+1}\}}\sum_{i \in \mathcal{S}_{\ell}} \int_{y = 0}^{\infty}  \pi_{i,r}^{k}(y)\phi_{k}^r(x)\wrt x\wrt y + \sum_{\ell \in \{-,0\}} \sum_{i \in \mathcal{S}}\sum_{r\in \mathcal K_i^\ell}   p^{r}_i(\mathcal{F}^{\ell}_i) = 1,\label{eqn:pisystemend}
+%	\end{align}
+%	% 
+%	where $ {K} :=  {D}^{++}(0) + \Psi(0) {D}^{(-+)}(0)$, and $z$ is a normalising constant.
+%	%
+%%Also \( {\bs p}^-_{{K+1}}=(  p^-_{{K+1},i})_{\{i\in\calS_{K+1}\mid r_i(b)<0\}}\) and \( {\bs p}^0_{{K+1}}=(  p^0_{{K+1},i})_{\{i\in\calS_{K+1}\mid r_i(b)=0\}}\) where \(  p^m_{{K+1},i},\,m\in\{-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =b, Y_t=0, \varphi_{t} = i\right]\).
+On the interior, we have
+%%\( {\bs p}^m_{\circ}=( {\bs p}_{k,i})_{i\in\calS,k\in\mathcal K_i^m}\) where \( {\bs p}_{k,i}=( {p}_{k,i}^r)_{r=1,...,N_k}\).
+On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, Y_t=0, \varphi_{t} = i\right]\) is given by \( {\bs p}_{k,i}\bs \phi^k(x)\tr{}\).
+%%
+%%Similarly for \( {\bs\pi}\) we write \[ {\bs \pi}^m =\vligne{ {\bs \pi}^m_{-1}(y) &  {\bs \pi}^m_\circ(y) &  {\bs \pi}^m_{K+1}(y)},\quad m\in\{+,-,0\},\]
+%%where \( {\bs \pi}^m_{{-1}}(y),\, {\bs \pi}^m_{\circ}(y),\, {\bs \pi}^m_{{K+1}}(y)\) correspond to the approximation at the left-boundary, interior, and on the right boundary of \(x\in[0,b]\), respectively.
+%%%
+%%Specifically, \( {\bs \pi}^+_{{-1}}(y)=(  \pi^+_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)>0\}}\), \( {\bs \pi}^-_{{-1}}(y)=(  \pi^-_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)<0\}}\) and \( {\bs \pi}^0_{{-1}}(y)=(  \pi^0_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)=0\}}\) where \(  \pi^m_{{-1},i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, Y_t=\wrt y, \varphi_{t} = i\right]\).
+%%%
+%%Also \( {\bs \pi}^+_{{K+1}}(y)=(  \pi^+_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)>0\}}\), \( {\bs \pi}^-_{{K+1}}(y)=(  \pi^-_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)<0\}}\) and \( {\bs \pi}^0_{{K+1}}(y)=(  \pi^0_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)=0\}}\) where \(  \pi^m_{{K+1},i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =b, Y_t\in\wrt y, \varphi_{t} = i\right]\).
+%%%
+%%On the interior, we have 
+%%\( {\bs \pi}^m_{\circ}(y)=( {\bs \pi}_{k,i}(y))_{i\in\calS,k\in\mathcal K_i^m}\) where \( {\bs \pi}_{k,i}(y)=( {\pi}_{k,i}^r(y))_{r=1,...,N_k}\).
+On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, Y_t\in\wrt y, \varphi_{t} = i\right]\) is given by \( {\bs \pi}_{k,i}(y)\bs \phi^k(x)\tr{}\).
+
+To assist the reader in understanding these constructions and the notation we provide an explicitly worked toy-example in Appendix~\ref{appendix:example}.
+
+% \section{A stochastic interpretation of the simplest DG scheme}
+% links between DG and the uniformisation scheme
+% positivity preservation
+% stochastic interpretations
+% how I use/analyse it in the numerics section
+
+\section{Other problems we can solve with the DG method}
+The utility of the DG approximation scheme for fluid queues in not limited to approximating the first return and stationary operators of fluid-fluid queues.
+For example, we can use a DG scheme to approximate the transient distribution of the fluid queue at time \(t\), which requires us to find the coefficients \(\vligne{\bs q_{-1}(t) & \bs a(t) & \bs q_{{K+1}}(t)}\) via the differential equation (\ref{eqn: DG ODE w BCs}) given an initial condition, typically by numerically integrating (\ref{eqn: DG ODE w BCs}).
+The stationary distribution of the fluid queue can be found by solving
+\begin{align*}
+	\bs b \bs B &= \bs 0,
+	\\ \bs b \bs 1 &= \bs 1,
+\end{align*}
+for the coefficients \(\bs b\).
+We can also approximate first hitting times.
+For example, given an initial condition of the cell \(\calD_k\), we can project the initial condition onto the polynomial basis with coefficients \(\bs a_k(0)\), then approximate the probability that the level process first hits \(\{y_k,y_{k+1}\}\) after time \(t_0\) by finding \(\bs a_k(0)e^{B^{kk} t_0}\int_{\calD_k}\bs \phi_k(x)\wrt x\) where the coefficients \(\bs a_k(t_0)\) can be found bi integrating the differential equation
+\[\cfrac{\wrt}{\wrt t}\bs a_k(t)=\bs a_k(t)\bs B^{kk},\]
+over time. We now briefly introduce time integration schemes which we use in this thesis, and some issues pertaining to oscillatory approximations.
+
+% APPLICATION TO INTERGRATING OVER TIME
+\section{Time-integration schemes}\label{sec: time integration}
+In this thesis we numerically integrate ODEs of the form
+\begin{align}\label{eqn: DG ODE w BCs 3}
+	\cfrac{\wrt}{\wrt t} \boldsymbol a(t)
+	% 
+	= \boldsymbol a(t) \bs Q
+\end{align}
+where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix.
+To do so, we employ the strong stability preserving Runge-Kutta (SSPRK) scheme of order 4 with 5 stages \citep{sr2002}.
+SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
+
+In our context, the SSPRK method with \(s\) stages has the form 
+\begin{align*}
+	&\bs v^{(0)} = \bs a(t),\\
+	%
+	&\bs v^{(\ell)} = \sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q,\, \ell = 1,\dots,s,\\
+	%
+	&\bs a(t+h) =\bs v^{(s)},
+\end{align*}
+where \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are coefficients which define the scheme and \(h\) is the \(t\)-step size.
+The coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are chosen to be positive so that \(\bs v^{(s)}\) is a convex combination of forward-Euler steps.
+Moreover, the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are optimised so that the allowable \(t\)-step size is as large as possible.
+The maximum allowable \(t\)-step size is \(\Delta_{RK}\), and we require
+\[\Delta_{RK} \leq \min_{\ell k}\cfrac{\alpha_{\ell k}}{\beta_{\ell k}}\Delta_E,\]
+where \(\Delta_E\) is the maximum allowable \(t\)-step size for the forward-Euler scheme.
+
+The \(t\)-step size needs to be chosen to be sufficiently small so that the time integration is stable. Here, for an order \(p\) DG scheme we choose the time step to be less than 
+\[\max_{i\in\calS}|c_i|\min_{r\in{2,...,p+1}}\left(z_{r}-z_{r-1}\right) \cfrac{\min_{k}{K+1}_k}{2}\]
+where \(z_{r-1}\leq z_r\), \(z_1=-1\), \(z_{p+1}=1\) and \(z_r, r=2,...,p,\) are the \(p-1\) zeros of the first derivative of \(P_{p}(x)\), the order \(p\) Legendre polynomial.
+This follows advice from \cite{nodalDGBook}.
+
+\section{Slope limiters}\label{sec: slope limiting}
+A well-know problem with DG schemes is that, in the presence of steep gradients or discontinuities, spurious oscillations in the approximate solution can occur and positivity is not guaranteed (see \cite{nodalDGBook}~Section~5.6 and \cite{koltai2011}~Section~3.3, for example).
+To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
+Slope limiting effectively alters the DG operator in regions where oscillations are detected to ensure non-oscillatory and non-negativity by reducing the order of the approximation to linear in these regions.
+However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
+As a result, the limiter may unnecessarily reduce the accuracy of the scheme in the presence of natural oscillations (see \citep{nodalDGBook},~Example~5.8).
+Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear operators.
+
+We now describe a simple slope limiter known as the \emph{Generalised MUSCL} scheme \citep{c99} (see also \cite[Section~5.6.1]{nodalDGBook}). Define the \emph{minmod} function 
+\begin{align*}
+	m(a_1,a_2,a_3)=\begin{cases}
+		s\min_{i\in\{1,2,3\}}|a_i|, & |s|=1, \\
+		0, &\mbox{otherwise},
+	\end{cases}
+\end{align*}
+where \(s=\frac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
+When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
+In the context of slope-limiting, the three arguments \(a_1,\,a_2\) and \(a_3\) are the gradients of the approximate solution in three cells.
+When the signs of \(a_1,\,a_2\) and \(a_3\) differ, this suggests an oscillation in the approximate solution.
+
+Now, define \(\overline u_{k}\) as the average value of the approximate solution on cell \(k\), then the slope limited solution on cell \(k\) is the linear approximant 
+\begin{align*}
+	\Pi_{k}^{lim}(u_k)=\overline u_{k} + (x-\overline y_k)m\left(\Pi^1_{\partial x}u_k,\cfrac{\overline u_{k+1}-\overline u_{k}}{\Delta_k},\cfrac{\overline u_{k}-\overline u_{k-1}}{\Delta_k}\right),
+\end{align*}
+where \(\Pi^1_{\partial x}u_k\) is the slope of the linear projection of \(u_k\) on \(\calD_k\) and \(\overline y_k=(y_k+y_{k+1})/2\) is the centre of \(\calD_k\).
+
+In non-oscillatory regions of the solution the slope limiter is unnecessary and reduces accuracy, so we only apply the limiter to cells \(k\) where oscillations are detected.
+To determine which cells need limiting we compute
+\begin{align*}
+	v_k^L &:= \overline{u}_{k} - m(\overline{u}_{k} - {u}_{k}^L,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+	\\v_k^R &:= \overline{u}_{k} + m(\overline{u}_{k} - {u}_{k}^R,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+\end{align*}
+where \({u}_{k}^L\) and \({u}_{k}^R\) are the values of the approximate solution on cell \(k\) evaluated at the left-hand and right-hand edges of the cell.
+We apply the slope limiter to cell \(k\) if \(v_k^L\neq u_k^L\) or \(v_k^R\neq u_k^R\), otherwise the approximation on cell \(k\) is left unaltered.
+
+In the context of approximating the first return operator of a fluid-fluid queue there is no direct equivalent of slope limiting, other than, perhaps, to re-compute the solution with constant basis function if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
+
+\subsection{Slope limiters and time integration}
+To use slope limiters within the SSPRK scheme we first project the initial condition on to \(W\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
+
+For example, to find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation (\ref{eqn: DG ODE w BCs 3}).
+At each stage of the time-integration we apply the slope limiter, i.e.~the scheme above with a slope limiter is
+\begin{align*}
+	&\bs v^{(0)} = \vligne{\bs q_{-1}(t) & \bs a(t) & \bs q_{{K+1}}(t)},\\
+	%
+	&\bs v^{(\ell)} = \Pi^{lim}\left(\sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k} \bs B \bs v^{(k)}\right),\, \ell = 1,\dots,s,\\
+	%
+	&\vligne{\bs q_{-1}(t+h) & \bs a(t+h) & \bs q_{{K+1}}(t+h)} =\bs v^{(s)},
+\end{align*} 
+where \(\Pi^{lim}\) is the slope limiter function.
+
+\paragraph{Consequences of slope limiting} We have already mentioned that the Generalised MUSCL slope limiter reduces the order to linear in regions where oscillatory solutions are detected.
+When slope limiting is used in conjunction with a time-integration scheme the reduction in order may not be local in time and may persist past the time when the limiter first detected oscillations.
+For example, consider a problem with an initial condition which introduces oscillations into the numerical approximation.
+If the slope limiter is applied to the initial condition then, in the region around the oscillations, the approximation to the initial condition will be linear.
+Even if the slope limiter is not applied (or not needed) after this point, the initial condition has still been altered, perhaps significantly, from the original approximation and this can affect the accuracy of transient approximations.
+This is not to say that the slope-limited-regions of the approximation will \emph{always} remain linear.
+When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of Generalised slope limiter described.
+Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will effect the accuracy of subsequent computations.
+
+\subsection{Another slope limiting scheme}\label{sec: anoth pos pres}
+The advantage of the approach above is that, in areas of the approximation where the solution is smooth, then the high-order accuracy of the method can be retained, while maintaining positivity where necessary.
+However, if there are no regions of the solution of interest which are smooth, then there is no real advantage in this approach as the limiter will reduce the approximate solution to linear in all regions of interest.
+Hence, we may as well just use a linear scheme and save on some computation.
+Perhaps a better approach would be to use a linear scheme with a smaller cell width, such that the computational work remains approximately the same.
+
+For example, for a DG scheme with \(2p\) basis functions on a cell, \(\mathcal D_k\), of width \(\Delta\), say, we construct block matrices of dimension \(2p\times 2p\) (i.e.~the mass, stiffness and flux matrices, \(\bs M_k\), \(\bs G_k\), and \(\bs F^{k\ell}\), respectively) and compute the approximate solution using these matrices.
+Instead, we could break the cell, \(\mathcal D_k\), up into \(p\) sub-cells of width \(\Delta/p\) and use a DG scheme with \(2\) basis functions on each sub-cell (i.e.~a linear basis on each cell).
+With this approach, to approximate the solution on the original interval \(\mathcal D_k\), we would also construct block matrices of dimension \(2p\times 2p\) and compute the approximate solution on \(\calD_k\) using these matrices in the same way as before.
+If we know \emph{a priori} that we will apply the slope limiter to cell \(\calD_k\), then intuitively we expect that the latter scheme will be far more accurate; it will approximate the solution on \(\calD_k\) by a piecewise linear function with \(p\) pieces, whereas the DG scheme with \(2p\) basis functions on cell \(\calD_k\) will, after limiting, represent the solution as a single linear function on \(\calD_k\).
+
+\subsection{Briefly on accuracy}
+As a rough guide on accuracy we now paraphrase some know results on the accuracy of the DG method for a similar kind of problem to the one under study here.
+For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the approximation to the flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution at time \(t\) provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
+Applying this result to the type of positivity preserving scheme in \ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
+
+% To quote Section~5.7 of \cite{nodalDGBook} SSPRK methods ``can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
+
+% why the MUSCL limiter?
+%   it requires no tuning and guarentees non-negativity
+%   some other positivity-preserving methods require tuning and can still produce negative results if not tuned properly
+%   ^ same goes for filtering

### 2022-02-10 chapter5/chapter.tex
.

@@
 
-\begin{thm}
+\begin{thm}\label{thm: big thm}
 	For all \(\bs y_0^{(p)}=(\ell,\bs a_{\ell_0,i}^{(p)}(x_0),i)\), \(\ell_0\in\mathcal K\), \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\calS\), and any bounded function \(\psi:\mathbb R\times \mathcal S \to \mathbb R\), then

### 2022-02-10 chapter6/chapter.tex
.

@@
 \chapter{Numerical investigations\label{sec: numerics}}
-We have now established multiple methods for approximating fluid queues which are suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}.
+We have now established multiple schemes for approximating fluid queues which are suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}.
 In this chapter we numerically investigate various aspects of the approximation schemes.
-Throughout, we compare the QBD-RAP scheme (from Chapter~\ref{sec: construction and modelling}), the discontinuous Galerkin method (as described in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013}, which we will refer to a just the \emph{uniformisation scheme}.
-%Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume method with an upwind flux.
+Throughout, we compare the QBD-RAP scheme (from Chapter~\ref{sec: construction and modelling}), the discontinuous Galerkin scheme (as described in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013}, which we will refer to a just the \emph{uniformisation scheme}.
+%Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume scheme with an upwind flux.
 Recall that, due to their stochastic interpretation, the uniformisation and QBD-RAP schemes are positivity preserving.
 
-Where possible, we implement the discontinuous Galerkin method with and without a slope-limiter implemented \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
+For some numerical experiments we implement the discontinuous Galerkin scheme with and without a slope-limiter implemented \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
 The slope limiter avoids oscillations and negative solutions.
@@
 Moreover, there is a computational cost in applying a limiter.
-However, high-order accuracy of the DG scheme is maintained away from the discontinuities \citep{c99}, \citep[Section~5.6.2]{nodalDGBook}.
+Another issue is that the need for a slope-limiter, or filter, may depend on the initial condition.
+In contrast, for the uniformisation and QBD-RAP the approximate solution is guaranteed to produce positive probabilities without any postprocessing or extra computation.
 
@@
 For simplicity, the number of basis functions is kept constant across all cells.
-We refer to the number of basis functions on a cell as the \emph{dimension}.
-For discontinuous Galerkin method, increasing the dimension of the scheme means increasing the number of polynomial basis functions used to approximate the solution within each cell.
-E.g.~if we use 3 basis functions in the discontinuous Galerkin method, we approximate the solution by a quadratic on each cell.
+We refer to the number of basis functions on a cell \(\calD_k\) as the \emph{dimension} and consider dimensions \(p=2r+1\) for \(r=0,1,2,3,...,10\).
+For discontinuous Galerkin scheme, increasing the dimension of the scheme means increasing the number of polynomial basis functions used to approximate the solution within each cell.
+E.g.~if we use 3 basis functions in the discontinuous Galerkin scheme, we approximate the solution by a quadratic on each cell.
 For the QBD-RAP scheme, increasing the dimension of the scheme means increasing the order of the CME distribution used to construct the scheme.
@@
 Equivalently, we may think of a dimension \(p\) uniformisation scheme as using \(p\) piecewise constant functions to approximate the solution on each cell.
-For all schemes (DG, QBD-RAP and uniformisation), if we construct and order \(p\) approximation, there are \(K\) cells, and \(N\) phases then the resulting approximation to the generator \(\mathbb B\) is a matrix of dimension \(pKN\).
+For all schemes (DG, QBD-RAP and uniformisation), if we construct and order \(p\) approximation, there are \(K\) cells, \(N\) phases and \(c_i\neq 0\) for all \(i\in\calS\), then the resulting approximation to the generator \(\mathbb B\) is a matrix of dimension \(pKN + N\).
 In this sense each approximation scheme leads to matrices of the same size (although not necessarily the same number of non-zero elements).
@@
 To keep the content of this chapter contained, we do not investigate all aspects of the schemes.
-For each numerical experiment we keep the cell size fixed for the DG and QBD-RAP schemes.
+For each numerical experiment we keep the cell size, \(\Delta\), fixed for the DG and QBD-RAP schemes.
 As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next \citep{nodalDGBook}.
@@
 Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages \citep{sr2002}, \citep[Section~5.7]{nodalDGBook} (see also Appendix~\ref{sec: time integration}), which claims to introduce no more oscillations into the solution as we integrate over time.
+
 Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
+We investigate two schemes with slope limiting.
+In the first scheme, which we refer to the \emph{DG-lim} scheme, we use a DG scheme of dimension \(p\) and apply the Generalised MUSCL slope limiter as necessary.
+In the second scheme, which we refer to as the \emph{DG-lin-lim} scheme, we follow the construction described in \ref{sec: anoth pos pres}, which is effectively a DG scheme with cells of width \(2\Delta/(p+1)\), for \(p=1,3,5,...,21\), and a basis of dimension \(2\) (linear) on each cell and apply the Generalised MUSCL limiter as necessary.
+Thus, the DG-lin-lim scheme uses \(p+1\), \(p=1,3,5,...,21\), basis functions to represent the solution on each of the original cells \(\calD_k\), and so we define the dimension of the scheme to be \(p+1\) to be comparable with the other schemes.
 We do not investigate filtering for the DG scheme (see \citep[Section~5.6.1]{nodalDGBook} and references therein).
 
-The performance of the discontinuous Galerkin method has been well-studied in some contexts \cite{c99}, \cite[Section~5.5]{nodalDGBook}, and it is well-known that the discontinuous Galerkin method performs remarkably well on problems with smooth solutions.
-Here, we mostly focus on investigating the numerical performance of the methods on problems with non-smooth solutions, the purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP method.
+The performance of the discontinuous Galerkin scheme has been well-studied in some contexts \cite{c99}, \cite[Section~5.5]{nodalDGBook}, and it is well-known that the discontinuous Galerkin scheme performs remarkably well on problems with smooth solutions.
+Here, we mostly focus on investigating the numerical performance of the schemes on problems with non-smooth solutions, the purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP scheme.
 In the stochastic modelling community it is very common to have problems with discontinuities, such as a non-smooth initial conditions.
@@
 
-As we shall see, if the solution is smooth, then the Galerkin method is highly-effective and displays rapid convergence to the solution as the number of polynomial basis functions is increased.
-However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin method, leading to nonsense solutions.
-In these cases, the QBD-RAP approximation performs relatively well compared to the other positivity preserving schemes investigated.
+As we shall see, if the solution is smooth, then the Galerkin scheme is highly-effective and displays rapid convergence to the solution as the number of polynomial basis functions is increased.
+However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin scheme, leading to nonsense solutions.
+In these cases, the QBD-RAP and DG-lin-lim approximations perform relatively well compared to the other positivity preserving schemes investigated.
+
+All the following error plots are on a \(\log_{10}\)-\(\log_{10}\) scale.
+If the error function takes the form \(error = \beta_0^* p^{\beta_1}\), where \(p\) is the dimension of the scheme, then the \(\log_{10}\)-\(\log_{10}\) will show a linear trend; \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\), where \(\beta_0=\log_{10}(\beta_0^*)\).
+Where relevant, to estimate the asymptotic rate of convergence, \(\beta_1\), we estimate the of the line \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\) using ordinary least squares and the last eight data points on each plot.
+We plot the estimated lines of best fit and quote their slope to the right of the plots.
 
 The structure of this chapter is as follows.
-In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different methods and observe their performance at approximating the initial condition, which allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
+In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different schemes and observe their performance at approximating the initial condition, which allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
 In Section~\ref{sec: wave num} we investigate a simple travelling wave problem with various initial conditions.
@@
 Next, we investigate a simple fluid queue with two phases.
-Section~\ref{sec:stat} investigates the ability of the methods to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the fluid queue for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the methods to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
+Section~\ref{sec:stat} investigates the ability of the schemes to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the fluid queue for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the schemes to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
 Section~\ref{sec: ffq num} approximates the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\) for two simple fluid-fluid queue models.
@@
 \section{Function approximation/reconstruction}\label{sec: recon num}
-We start our investigation by looking at how well the methods perform at approximating an initial condition.
+We start our investigation by looking at how well the schemes perform at approximating an initial condition.
 By approximating the initial condition only, we aim to instrument the performance of the approximation schemes without any dynamics.
 
-For the discontinuous Galerkin method, this we project the initial condition on to the set of polynomial functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstructing the solutions as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
+For the discontinuous Galerkin method, we project the initial condition on to the set of polynomial functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstructing the solutions as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
 For the purposes of approximation and reconstruction for the QBD-RAP scheme we must orientate the initial condition in a positive or negative direction; here, we suppose that the initial conditions belong to a positive phase.
@@
 
-\subsection{QBD-RAP closing operators}
+\subsection{QBD-RAP closing operators}\label{sec: closing vecs numerics}
 The three closing operators we investigate are the following.
@@
 
-\begin{example}First consider approximating the initial condition with density function \(2\times 1(x<0.5)\).
+\begin{example}\label{ex: another example}
+	First consider approximating the initial condition with density function \(2\times 1(x<0.5)\).
 In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation with the three different closing vectors.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(2x1(x<0.5)+1(x\geq 0.5)\), and the approximations (left) and \(L^2\) error between the true PDF, \(2\times 1(x<0.5)\), and the approximations (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: another example} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves of the naive normalised (orange) and normalised closing vectors are almost coincident.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}%To estimate the asymptotic rate of convergence, we fit a linear model of \(\log_{10}(error) = \beta_0 + \beta_1\log_{10}(Dim)\) to the the last 6 data points on the curves above, which corresponds to a rate of convergence \(error \propto N^{-\beta_1}\).
+The estimate of \(\beta_1\) was approximately -1.0 in all cases.}
 	\label{fig: fun 2 ks error qbdrap closing vecs}
@@
 
-\begin{example}Now consider approximating the initial density \(f(x)=1\).
+\begin{example}\label{ex: al}
+	Now consider approximating the initial density \(f(x)=1\).
 Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error and \(L^2\) error of the PDF we now see that the normalised reconstructions outperform the unnormalised reconstruction.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(x\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: al} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves of the error curves of the naive normalised (orange) and normalised closing vectors are almost coincident.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 4 ks error qbdrap closing vecs}
@@
 Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval.
-Perhaps there is a different reconstruction method or yet another closing operator which could alleviate this issue.
+Perhaps there is a different reconstruction method or another closing operator which could alleviate this issue.
 \begin{figure}[h]
@@
 	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various dimensions, and the true PDF which is \(f(x)=1\) (orange dashed line).}
+	\caption{Approximations to the PDF \(f(x)=1\) (orange dotted line) for Example~\ref{ex: al}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
 	\label{fig: pdf reconstructed}
@@
 
-\begin{example}Next we consider approximating the initial distribution with density \(-6x^2+6x\).
+\begin{example}\label{ex: aasmm}
+	Next we consider approximating the initial distribution with density \(-6x^2+6x\).
 Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the dimension of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(-2x^3+3x^2\), and the approximation (left) and \(L^2\) error between the true PDF \(-6x^2+6x\) and the approximation (right), for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) an normalised closing vectors are coincident.}
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error for the PDFs (right) of Example~\ref{ex: aasmm}, for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves for the naive normalised (orange) and normalised closing vectors are almost coincident.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 6 ks error qbdrap closing vecs}
@@
 	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the unnormalised closing operator (blue solid line), normalised closing operator (green dash-dotted line), for various orders, and the true PDF which is \(f(x)=-6x^2+6x\) (orange dashed line).}
+	\caption{Approximations to the PDF \(f(x)=-6x^2+6x\) (orange dotted line) for Example~\ref{ex: aasmm}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
 	\label{fig: pdf reconstructed quadratic}
@@
 
-\begin{example}Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\).
+\begin{example}\label{ex: kjfk}
+	Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\).
 This density function is at a maximum at the left of the region.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \((1-e^{-3x})/(1-e^{-3})\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles).
-The naive normalised (orange) and normalised closing vectors are coincident.}
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: kjfk} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves for the naive normalised (orange) and normalised closing vectors are almost coincident.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 7 ks error qbdrap closing vecs}
@@
 
-Given the evidence above we choose to use the normalised closing operator to reconstruct solutions.
-Further, unlike the naive normalised operator, the normalised operator is theoretically justified in that we proved the closing operator leads to a convergent scheme and is linear.
+In Table~\ref{table: error rates for QBD-RAP closing} we summarise the rates of convergence of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots).
+\begin{table}
+	\centering
+	\begin{tabular}{c|c|c|c}
+			Example & PDF & \(\beta_1\), unnormalised & \(\beta_1\), normalised \\\hline
+			\ref{ex: another example} & \(2\times 1(x<0.5)\) & -1.07 & -0.98 \\
+			\ref{ex: al} & 1 & -0.99 & -0.91 \\
+			\ref{ex: aasmm} & \(-6x^2+6x\) & -2.72 & -2.57 \\ 
+			\ref{ex: kjfk} & \(3e^{-3x}/(1-e^-3)\) & -1.11 & -2.35 
+	\end{tabular}
+	\caption{\label{table: error rates for QBD-RAP closing} Estimated rates of convergence, \(\beta_1\), of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots).}
+\end{table}
+
+In summary, the naive normalised and normalised closing operators perform almost identically.
+In some cases the unnormalised closing operator can outperform the normalised closing operators, typically when the mass of the function to be approximated at the left-hand edge of the cell is low.
+Further, the rates of convergence for all three schemes was similar, except for in Example~\ref{ex: kjfk} where we approximated the initial condition \(3e^{-3x}/(1-e^-3)\); in this case the rate of convergence of the normalised closing operators was approximately twice as large as the unnormalised version.
+Given the results above and its theoretical tractability, we will opt to use the normalised closing operator for the rest of the numerical experiments.
 
 \FloatBarrier
-\subsection{Comparison of methods}\label{sec: comp}
-Here we compare the ability of the QBD-RAP, uniformisation, and DG methods to reconstruct initial conditions.
-
-\begin{example}First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5.
+\subsection{Comparison of schemes}\label{sec: comp}
+Here we compare the ability of the QBD-RAP, uniformisation, and DG schemes to reconstruct initial conditions.
+
+\begin{example}\label{ex: exa}
+	First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5.
 This distribution does not have a PDF, so we compare the CDFs only.
 In Figure~\ref{fig: fun 1 comp} we plot the KS metric (left) and \(L^1\) metric (right) between the true CDF and the reconstructed approximations.
-Observing the KS metric, it appears that none of the methods converge and the KS error sits around 0.5.
+Observing the KS metric, it appears that none of the schemes converge and the KS error sits around 0.5.
 This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except, perhaps, at points of discontinuity.
-None of the methods appear to converge at the discontinuity at \(x=0.5\).
-Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the methods appear to show the convergent behaviour we expect.
-Here, the uniformisation method appears to perform the best, while the QBD-RAP method performs the worst.
-The rate of convergence of the QBD-RAP method appears to be similar to the rate of convergence of the DG scheme.
-
-Perhaps it is no surprise that the uniformisation method performs best.
-In the uniformisation method as we increase the order we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the initial distribution.
-As such, the uniformisation method can produce a piecewise continuous, linear approximation to the CDF.
-In contrast, both the DG and QBD-RAP methods result in a smooth approximation to the CDF.
-Given the initial distribution is far from smooth (it's a point mass), then we might expect that the uniformisation method will perform relatively well.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
-	\caption{KS error between the true CDF, \(1(x\geq 0.5)\), and the approximations (left) and \(L^1\) error between the true CDF and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
+None of the schemes appear to converge at the discontinuity at \(x=0.5\).
+Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the schemes appear to show the convergent behaviour we expect.
+Here, the uniformisation scheme appears to perform the best, while the QBD-RAP scheme performs the worst.
+The rate of convergence of the QBD-RAP scheme appears to be similar to the rate of convergence of the DG scheme.
+
+Perhaps it is no surprise that the uniformisation scheme performs best.
+In the uniformisation scheme as we increase the order we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the initial distribution.
+As such, the uniformisation scheme can produce a piecewise continuous, linear approximation to the CDF.
+In contrast, both the DG and QBD-RAP schemes result in a smooth approximation to the CDF.
+Given the initial distribution is far from smooth (it's a point mass), then we might expect that the uniformisation scheme will perform relatively well.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^1\) error of the CDF (right) of Example~\ref{ex: exa} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 1 comp} 
@@
 
-In Figure~\ref{fig: pdf comp fun 1} we plot approximated CDFs from the DG, uniformisation and QBD-RAP methods alongside the true CDF.
-The DG method displays undesirable features for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1.
-On the other hand, although the QBD-RAP method converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1.
+In Figure~\ref{fig: pdf comp fun 1} we plot approximated CDFs from the DG, uniformisation and QBD-RAP schemes alongside the true CDF.
+The DG scheme displays undesirable features for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1.
+On the other hand, although the QBD-RAP scheme converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1.
 The uniformisation scheme approximates the CDF exceptionally well.
@@
 	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/comp/fun1/cdfs_formatted.pdf}
-	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) methods.
+	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes for Example~\ref{ex: exa}.
 The true distribution function is \(1(x\geq 0.5)\) (grey dotted line).}
@@
 
-\begin{example}Now consider approximating the initial distribution with PDF \(1(x\leq 0.5)\).
+\begin{example}\label{eqn: example something}
+	Now consider approximating the initial distribution with PDF \(1(x\leq 0.5)\).
 Figure~\ref{fig: fun 2 comp} plots the KS error (left) and \(L^2\) error between the true and approximate PDFs (right).
-Figure~\ref{fig: fun 2 comp} suggests that all methods converge at a similar rate for this problem.
-Here, the QBD-RAP method performs worst, the uniformisation method second, and the DG method the best.
-However, once again the DG method exhibits undesirable properties (plots not shown) -- the approximation to the CDF is at some points above 1 and is not monotonic (although these violations do not appear to be as severe in this case as they were for the example above).
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(x\geq 0.5)\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
+Figure~\ref{fig: fun 2 comp} suggests that all schemes converge at a similar rate for this problem.
+Here, the QBD-RAP scheme performs worst, the uniformisation scheme second, and the DG scheme the best.
+However, once again the DG scheme exhibits undesirable properties (plots not shown) -- the approximation to the CDF is at some points above 1 and is not monotonic (although these violations do not appear to be as severe in this case as they were for the example above).
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{eqn: example something} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 2 comp} 
@@
 
-\begin{example}So far we have considered two problems which exhibit discontinuities.
+\begin{example}\label{ex: alllll}
+	So far we have considered two problems which exhibit discontinuities.
 At the other extreme we now consider an initial distribution with density \(-6x^2+6x\).
 In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDFs.
-Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG method can approximate the initial condition exactly.
+Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG scheme can approximate the initial condition exactly.
 This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from dimension 1 to 3, then plateaus.
 Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, and due to machine arithmetic, the errors for the DG scheme are not 0.
-Regarding the other two methods, they too appear to be convergent at approximately the same rate, with the uniformisation method performing better for the KS error, but very similarly in terms of the \(L^2\) error.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(-6z^2+6x\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
+Regarding the other two schemes, they too appear to be convergent at approximately the same rate, with the uniformisation scheme performing better for the KS error, but very similarly in terms of the \(L^2\) error.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun6/meshs_l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: alllll} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 6 comp} 
@@
 
-\begin{example}Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\).
+\begin{example}\label{ex: aalllkkdd}Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\).
 Figure~\ref{fig: fun 9 comp} shows the errors.
-Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs (right) tell a similar story; for sufficiently high order, the DG method approximates the initial condition very well.
-The uniformisation method performs second best, while the QBD-RAP method is slowest to converge.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(cos(4\pi(x+0.5))+1\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dashed line with circles).}
+Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs (right) tell a similar story; for sufficiently high order, the DG scheme approximates the initial condition very well.
+The uniformisation scheme performs second best, while the QBD-RAP scheme is slowest to converge.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={1.1cm 0.8cm 0.1cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={1.1cm 0.8cm 0.1cm 1.25cm},clip]{chapter6/figs/comp/fun9/meshs_l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: aalllkkdd} for the DG scheme (blue solid line), uniformisation scheme (orange dashed line) and QBD-RAP scheme (purple dashed line).
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 9 comp} 
@@
 
-In conclusion, we have observed that the DG method performs well with respect to the error metrics considered, but can display oscillations and result in CDFs which are not non-decreasing.
+In conclusion, we have observed that the DG scheme performs well with respect to the error metrics considered, but can display oscillations and result in CDFs which are not non-decreasing.
 On the other hand, the uniformisation and QBD-RAP schemes result in non-decreasing approximations to the CDFs, but can converge slower with respect to the error metrics considered, with the QBD-RAP scheme often converging the slowest.
@@
 
-We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space.
-We discretise the interval \([0,10]\) into \(10\) cell, each of width 1.
+We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space and discretise the interval \([0,10]\) into \(10\) cell, each of width 1.
 We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric.
-We use the SSPRK4 method to integrate over time with a \(t\)-step size of \(0.005\), and we evolve the system until time \(t=4\).\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} For the DG method we also implement the Generalised MUSCL slope limiter to stabilise the integration \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
-
-To investigate the performance of the methods without the need to reconstruct the function within each cell we use the cell-wise error metric obtained by computing 
+Furthermore, we use the SSPRK4 method to integrate over time with a \(t\)-step size of \(0.005\), and we evolve the system until time \(t=4\).\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} For the DG scheme we also implement the Generalised MUSCL slope limiter to stabilise the integration \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
+
+To investigate the performance of the schemes without the need to reconstruct the function within each cell we use the cell-wise error metric obtained by computing 
 \begin{align}
@@
 The level of the fluid queue is uniformly distributed over the first cell.
-For the DG and uniformisation methods, the initial condition can be represented exactly, whereas, for the QBD-RAP method, it cannot.
-Thus, in this case, there is no discretisation error in constructing the initial condition for the DG and uniformisation methods.
+For the DG-based schemes (including the uniformisation scheme) the initial condition can be represented exactly whereas, for the QBD-RAP scheme, it cannot.
+Thus, in this case, there is no discretisation error in constructing the initial condition for the DG and uniformisation schemes.
 At time \(t=4\), the solution is \(f(x,4)=1(x\in[4,5))\).
-The projections related to the DG and uniformisation methods can represent this solution exactly too, hence we might expect these methods to work well here.
-
-We plot the cell-wise error metric in Figure~\ref{fig: fun 4 wave cp} and observe that the DG method appears to converge fastest, followed by the QBD-RAP method, then the uniformisation method, and the DG scheme with a slope limiter does not appear to converge.
+The projections related to the DG-based schemes can represent this solution exactly too, hence we might expect these schemes to work well here.
+
+We plot the cell-wise error metric in Figure~\ref{fig: fun 4 wave cp} and observe that the DG scheme appears to converge fastest, followed by the QBD-RAP and DG-lin-lim schemes which perform comparably, then the uniformisation scheme.
+The worst performer is the DG-lim scheme which does not appear to converge.
 The DG scheme with a slope limiter has detected oscillations in the approximate solution and reduced the scheme to linear where oscillations are detected, thereby decreasing the accuracy of the scheme to linear where oscillations are present.
@@
 	\caption{Cell-wise error defined in (\ref{eqn: cell errors}) for the travelling wave in Example~\ref{ex: wave 4}.
-Plotted are the cell-wise errors for the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line) versus the dimension of the approximation.}
+Plotted are the cell-wise errors for the DG (blue solid line), DG-lim scheme (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim scheme (gold dashed line) schemes, versus the dimension of the approximation.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 4 wave cp} 
@@
 Figure~\ref{fig: fun 4 wave} plots the KS error between the approximated and true CDFs and \(L^2\) error between the approximated and true PDFs.
-Comparing Figure~\ref{fig: fun 4 wave cp} with Figure~\ref{fig: fun 4 wave}, each method seems to show a similar rate of decay for error metrics with and without reconstruction.
-
-Returning to Figure~\ref{fig: fun 4 wave}, the DG method with the MUSCL limiter does not appear to converge with these error metrics for this problem (due to the limiter reducing the approximation to linear near the discontinuity), while the other two positivity preserving methods, the uniformisation and QBD-RAP schemes, both appear to be converging with the QBD-RAP method converging faster.
-With these error metrics, the DG method appears to be converging fastest.
-However, if we observe the approximations resulting from the DG method (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscillations and negative values.
-
-Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG with MUSCL limiter and QBD-RAP methods for various dimension schemes.
-In the first row we observe that the DG method (sans limiter) produces an approximation to the PDF with negative values when the dimension is greater than 1.
+Comparing Figure~\ref{fig: fun 4 wave cp} with Figure~\ref{fig: fun 4 wave}, each scheme seems to show a similar rate of decay for these two error metrics as they did for the cell-wise errors.
+
+Returning to Figure~\ref{fig: fun 4 wave}, the DG-lim scheme does not appear to converge with these error metrics.
+Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes, all appear to be converging with the QBD-RAP and  DG-lin-lim schemes converging similarly and faster than the uniformisation scheme.
+With these error metrics, the DG scheme is converging fastest.
+However, if we observe the approximations resulting from the DG scheme (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscillations and negative values.
+
+Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG with MUSCL limiter and QBD-RAP schemes for various dimension schemes.
+In the first row we observe that the DG scheme (sans limiter) produces an approximation to the PDF with negative values when the dimension is greater than 1.
 In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter, the DG approximation does not change significantly after order 3.
-This is due the fact that the DG method is at best linear in the presence of oscillations.
-In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation method, and in the last row is the solution approximated using the QBD-RAP method.
+This is due the fact that the DG scheme is at best linear in the presence of oscillations.
+In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation scheme, and in the last row is the solution approximated using the QBD-RAP scheme.
 \begin{figure}[h]
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error of the PDFs for the travelling eave problem in Example~\ref{ex: wave 4} using the approximations from the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 4} using the approximations from the DG (blue solid line), DG-lim scheme (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 4 wave} 
@@
 	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG with MUSCL limiter (third row) and QBD-RAP (bottom row) methods, for order 1, 3, 5, and 7 (columns) for the travelling eave problem in Example~\ref{ex: wave 4}.
+	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row) and DG-lin-lim schemes, for order 1, 3, 5, and 7 (columns) for the travelling eave problem in Example~\ref{ex: wave 4}.
 The true density function is \(1(4\leq x<5)\).}
@@
 This is a particularly interesting example.
-It shows that, for the DG method, even though there is no discretisation error for the initial condition, we use a strong stability preserving time-integration method, and the projection off which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as in the top row of Figure~\ref{fig: pdf wave fun 4}.
+It shows that, for the DG scheme, even though there is no discretisation error for the initial condition, we use a strong stability preserving time-integration method, and the projection off which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as in the top row of Figure~\ref{fig: pdf wave fun 4}.
 \exampleFloatBarrier
@@
 No PDF exists for the true distribution, so instead we compare the CDFs.
-Moreover, when we analysed reconstruction of this initial condition we saw that using the KS metric may be uninformative.
-Instead, for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs) and the cell-wise error from (\ref{eqn: cell probs})-(\ref{eqn: cell errors}).
+Moreover, when we analysed reconstruction of this initial condition we saw that using the KS metric may be uninformative ue to the lack of point-wise convergence at the discontinuity.
+So, for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs) instead, and also with the cell-wise error from (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}).
 
 Figure~\ref{fig: fun 2 wave} (left) plots the \(L^1\) error between the true and approximated CDFs.
-The \(L^1\) metric tells a similar story to the previous analysis: the DG method with the limiter does not converge as order increases, the other two positivity preserving methods (the uniformisation and QBD-RAP methods) appear to converge, with the QBD-RAP method converging faster, while the DG method appears to converge the fastest.
-However, if we plot the approximations to the CDFs from the DG method (not shown) we would once again see an oscillating, non-monotonic function.
-Another interesting observation is to compare the performance of the uniformisation and QBD-RAP methods with respect to the \(L^1\) metric on the CDFs in Figure~\ref{fig: fun 2 wave} (left) to that in Figure~\ref{fig: fun 1 comp} (right).
-Recall, in Figure~\ref{fig: fun 1 comp} we investigated the ability of the different approximation methods to represent the initial distribution with CDF \(1(x\geq 0.5)\), which is the initial condition of the problem we are considering here.
-In Figure~\ref{fig: fun 1 comp} (right) the uniformisation method out-performed the QBD-RAP method at reconstructing the initial condition.
-However, in Figure~\ref{fig: fun 2 wave} (left) we see that the QBD-RAP method out-performs the uniformisation method.
-This suggests that the QBD-RAP method is better able to resolve movement of mass across the domain when integrating over time, compared to the uniformisation method.
-
-Figure~\ref{fig: fun 2 wave} (right) plots the cell-based error metric in (\ref{eqn: cell errors}).
-Once again, the DG method with MUSCL limiter does not appear to converge.
-The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP method converging faster.
-Most interestingly though is the DG method.
-It is unclear whether this method will converge in this case as the errors in the plot are not obviously decreasing.
-Moreover, with this metric, the QBD-RAP method out-performs the DG method when the dimension is large enough.
-Furthermore, the DG method can result in negative estimates of the cell probabilities in (\ref{eqn: cell probs}) and (\ref{eqn: cell probs boundary}) (plots not shown).
+The \(L^1\) metric tells a similar story to the previous analysis: the DG-lim scheme does not converge as order increases, the other three positivity preserving schemes (the uniformisation, DG-lin-lim and QBD-RAP schemes) appear to converge, with the QBD-RAP and DG-lin-lim schemes converging at a similar rate and faster than the uniformisation scheme.
+The DG scheme appears to converge the fastest.
+However, if we plot the approximations to the CDFs from the DG scheme (not shown) we would once again see an oscillating, non-monotonic function.
+
+Another interesting observation is to compare the performance of the uniformisation and QBD-RAP schemes with respect to the \(L^1\) metric on the CDFs in Figure~\ref{fig: fun 2 wave} (left) to that in Figure~\ref{fig: fun 1 comp} (right).
+Recall, in Figure~\ref{fig: fun 1 comp} we investigated the ability of the approximation schemes to represent the initial distribution with CDF \(1(x\geq 0.5)\), which is the initial condition of the problem we are considering here.
+In Figure~\ref{fig: fun 1 comp} (right) the uniformisation scheme out-performed the QBD-RAP scheme at reconstructing the initial condition.
+However, in Figure~\ref{fig: fun 2 wave} (left) we see that the QBD-RAP scheme out-performs the uniformisation scheme.
+This suggests that the QBD-RAP scheme is better able to resolve movement of mass across the domain when integrating over time, compared to the uniformisation scheme.
+
+Figure~\ref{fig: fun 2 wave} (right) plots the cell-based error metric in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}).
+Once again, the DG-lim scheme does not appear to converge.
+Interestingly, the error curve for the DG scheme is not monotonic.
+Furthermore, the DG scheme can result in negative estimates of the cell probabilities in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (plots not shown).
+Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme all appear to converge, with the QBD-RAP and DG-lin-lim schemes converging at a similar rate and faster than the DG scheme when the dimension of the basis used is sufficiently large.
+The estimated rate of convergence of the QBD-RAP and DG-lin-lim schemes is approximately -4 for this example, which is significantly faster than the rate of convergence of these schemes in all other examples in this section.
 \begin{figure}[h]
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/L1_cell_probs.pdf} 
-	\caption{\(L^1\) error between the true CDF and the approximations (left) and the cell-wise error metric in (\ref{eqn: cell errors}) for the travelling wave problem in Example~\ref{ex: wave 2} where approximation were constructed via the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+	\caption{\(L^1\) error of CDFs (left) and the cell-wise error metric in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (right) for the travelling wave problem in Example~\ref{ex: wave 2} where approximation were constructed via the DG (blue solid line), DG-lin-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: fun 2 wave} 
@@
 
-	Since the discontinuity at \(x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG method will perform well for this example.
+	Since the discontinuity at \(t=4,\,x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG scheme will perform well for this example.
 Figures~\ref{fig: fun 6 wave cp} and~\ref{fig: fun 6 wave} confirms that this is indeed the case.
-For all three metrics the error obtained by the DG method rapidly decreases to a point where it is swamped by other numerical errors.
-This is characteristic of the DG method for the smooth problems we investigate throughout this chapter.
+For all three error metrics the error obtained by the DG scheme rapidly decreases to a point where it is swamped by other numerical errors.
+This is characteristic of the DG scheme for the smooth problems we investigate throughout this chapter.
 For low order DG schemes there are regions where the approximated PDF is negative, however, as the order of the DG scheme increases, these regions disappear.
-Interestingly, even though this is a relatively smooth problem, the DG method with MUSCL limiter does not perform well.
+
+	Interestingly, even though this is a relatively smooth problem, the DG-lim scheme does not perform well.
 It must be that the initial and transient distributions are `sufficiently pointy' that oscillations in the numerical solutions occur and the limiter reduces the order of the scheme to linear.
-Both the uniformisation and QBD-RAP schemes perform similarly to previous examples.
-Comparing the cell-wise error plots in Figure~\ref{fig: fun 6 wave cp} with the higher-resolution error metrics in Figure~\ref{fig: fun 6 wave}, all the approximation schemes show a similar rate of decay of errors as a function of the dimension of the approximation.
+
+	The other three positivity preserving schemes appear to converge, with the uniformisation scheme converging the slowest and the DG-lin-lim scheme converging the fastest of the three.
+For the KS and \(L^2\) error on the PDFs in Figure~\ref{fig: fun 6 wave} the DG-lin-lim and QBD-RAP schemes perform similarly, however, for the cell-wise error in Figures~\ref{fig: fun 6 wave cp} the DG-lin-lim scheme performs significantly better than the QBD-RAP scheme.
+Given that the DG-lim-lim scheme performed similarly to the QBD-RAP scheme for the previous examples, and also for the other error metrics for this example, this is unexpected.
+
+	To investigate the performance of the QBD-RAP and DG-lin-lim schemes, in Figure~\ref{fig: fun 6 wave pdfs} we plot the dimensions 21 QBD-RAP and dimensions 22 DG-lin-lim approximations to the PDF.
+Figure~\ref{fig: fun 6 wave pdfs} shows that the DG-lin-lim scheme approximates the PDF well, except for at the peak located at \(x=6.5\) -- I suspect the limiter has affected the solution in this region.
+Specifically, the DG-lin-lim scheme underestimates the density in the peak, but overestimates the density either side of the peak.
+As a result, when we integrate over \(x\in[6,7]\) to compute the cell-wise probabilities these errors cancel to some extent and the DG-lin-lim scheme produces an accurate estimate of the probability in this cell.
+In comparison, the QBD-RAP scheme also underestimates the density at the peak, but also within the entire interval \([6,7]\), albeit less so at the edges.
+The QBD-RAP scheme also overestimates the density in the adjacent cells, \([5,7]\) and \([7,8]\).
 	\begin{figure}[h]
@@
 		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/L1_cell_probs.pdf}
-		\caption{Cell-wise error defined in (\ref{eqn: cell errors}) for the travelling wave problem in Example~\ref{ex: wave 6}.
-Plotted are the cell-wise errors for the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line) versus the dimension of the approximation.}
+		\caption{Cell-wise error defined in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) for the travelling wave problem in Example~\ref{ex: wave 6}.
+Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 		\label{fig: fun 6 wave cp} 
@@
 		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error for the PDFs for Example~\ref{ex: wave 6} where approximations were constructed via the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+		\caption{KS error (left) and \(L^2\) error for the PDFs (right) for Example~\ref{ex: wave 6} where approximations were constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 		\label{fig: fun 6 wave} 
 	\end{figure}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun6/order_21_PDFs_formatted.pdf}
+		\caption{Dimension 21 QBD-RAP (purple) and dimension 22 DG-lin-lim (gold) approximations to the transient PDF at time \(t=4\) for Example~\ref{ex: wave 6}. The true PDF is plotted in grey.}  
+		\label{fig: fun 6 wave pdfs} 
+	\end{figure}
 \exampleFloatBarrier
@@
 \begin{example}\label{ex: wave 1}
-	We now want to look at how the methods might handle a mass at the boundary.
+	We now want to look at how the schemes might handle a mass at the boundary.
 We introduce an ephemeral second phase into the model with phase transition rate \(T_{22}=-1\), and fluid rate \(c_2=0\).
@@
 	\end{align}
-	The PDF at \(t=4\) is discontinuous at \(x=4\), which is at the edge of a cell. For this problem, all methods can represent the initial condition exactly.
-
-	Figure~\ref{fig: fun 1 wave cp} plots the cell-wise error metric and Figure~\ref{fig: fun 1 wave} plots the KS error (right) and the \(L^2\) error between of the PDFs.
+	The PDF at \(t=4\) is discontinuous at \(x=4\), which is at the edge of a cell. For this problem, all schemes can represent the initial condition exactly.
+
+	Figure~\ref{fig: fun 1 wave cp} plots the cell-wise error metric and Figure~\ref{fig: fun 1 wave} plots the KS error (right) and the \(L^2\) error between the PDFs (right).
 Once again, the cell-wise error metrics in Figure~\ref{fig: fun 1 wave cp} show similar characteristics to the finer-resolution error metrics in Figure~\ref{fig: fun 1 wave}.
-Observing Figures~\ref{fig: fun 1 wave cp} and \ref{fig: fun 1 wave}, the DG method with MUSCL limiter does not perform well, which might be expected due to the discontinuity in the transient distribution at \(x=4\) in Phase~1.
-The uniformisation and QBD-RAP methods appear to converge, with the QBD-RAP method converging faster.
-The DG method converges fastest, however, produces approximations with negative and oscillatory solutions as we might expect given the discontinuity.
+Observing Figures~\ref{fig: fun 1 wave cp} and \ref{fig: fun 1 wave}, the DG-lim scheme does not perform well, which might be expected due to the discontinuity in the transient distribution at \(x=4\) in Phase~1.
+The uniformisation, DG-lin-lim and QBD-RAP schemes appear to converge, with the QBD-RAP scheme converging fastest, but performing similarly to the DG-lin-lim scheme, and the uniformisation scheme performs the poorest of the three.
+The DG scheme converges fastest, however, produces approximations with negative and oscillatory solutions as we might expect given the discontinuity.
 A selection of approximations to the transient PDF are shown in Figure~\ref{fig: pdf wave fun 1}.
@@
 		\caption{Cell-wise error for the travelling wave problem in Example~\ref{ex: wave 1}.
-Plotted are the cell-wise errors for the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line) versus the dimension of the approximation.}
+Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 		\label{fig: fun 1 wave cp}  
@@
 		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error of the PDFs for the travelling wave problem in Example~\ref{ex: wave 1} where approximations were constructed via the DG method (blue solid line with crosses), DG method with the MUSCL limiter (green dash-dotted line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 1} where approximations were constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 		\label{fig: fun 1 wave} 
@@
 		\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun1/pdf_formatted.pdf}
-		\caption{Approximate PDFs for Example~\ref{ex: wave 1} using the DG (top row), uniformisation (second row), DG with MUSCL limiter (third row) and QBD-RAP (bottom row) methods, with dimensions 1, 3, 5, and 7 (columns).
+		\caption{Approximate transient PDFs at time \(t=4\) for Example~\ref{ex: wave 1} using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row), and DG-lin-lim schemes of dimensions 1, 3, 5, and 7 (columns).
 The true density function is \(e^{-4}e^{x}1(x<4)\).}
@@
 We briefly turn our attention to stationary distributions of fluid queues.
-Since the stationary distribution is smooth we expect that the DG method will work well here.
+Since the stationary distribution is smooth we expect that the DG scheme will work well here.
 The stationary distribution is convenient as we can evaluate them analytically \cite{s2017} and do not require us to approximate initial conditions or integrate over time.
-Hence, analysing the performance of the approximation methods with the stationary distribution allows us to instrument the ability of the methods to capture the stochastic dynamics, without numerical integration in time.
+Hence, analysing the performance of the approximation schemes with the stationary distribution allows us to instrument the ability of the schemes to capture the stochastic dynamics, without numerical integration in time.
 
@@
 
-The model was discretised using the DG, uniformisation and QBD-RAP methods using ten cells of width \(1\).
+The model was discretised using the DG, uniformisation and QBD-RAP schemes using ten cells of width \(1\).
 We compute the coefficients for the stationary distribution in the following way.
-Suppose that a given discretisation method results in an approximation to the generator of the fluid queue as a matrix \(\bs B\).
+Suppose that a given discretisation scheme results in an approximation to the generator of the fluid queue as a matrix \(\bs B\).
 Then the stationary coefficients are found by solving
@@
 % Figure~\ref{fig: absorbing stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
-Clearly the DG method is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
-The QBD-RAP method and uniformisation method both appear to be converging, with the errors for the QBD-RAP method decreasing faster.
+Clearly the DG scheme is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
+The QBD-RAP scheme and uniformisation scheme both appear to be converging, with the errors for the QBD-RAP scheme decreasing faster.
 % \begin{figure}[h]
@@
 % 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/l2_pdf_error_formatted.pdf}
-% 	\caption{KS error between the true stationary CDF  and the approximations (left) and \(L^2\) error between the true stationary PDF and the approximations for Model~\ref{model: simple}~a) using the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
+% 	\caption{KS error between the true stationary CDF  and the approximations (left) and \(L^2\) error between the true stationary PDF and the approximations for Model~\ref{model: simple}~a) using the DG scheme (blue solid line), uniformisation scheme (orange dashed line) and QBD-RAP scheme (green dash-dotted line).}
 % 	\label{fig: absorbing stationary} 
@@
 
-\paragraph{Model~\ref{model: simple}.}
+% \paragraph{Model~\ref{model: simple}.}
 Figure~\ref{fig: reflecting stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
-Clearly the DG method is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
-The QBD-RAP method and uniformisation method both appear to be converging, with the errors for the QBD-RAP method decreasing faster.
+Clearly the DG scheme is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
+The QBD-RAP and uniformisation schemes both appear to be converging, with the errors for the QBD-RAP scheme decreasing faster.
 \begin{figure}[h]
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/stationary_distribution/l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error of the PDFs for Model~\ref{model: simple} using the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (purple dotted line with circles).}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Model~\ref{model: simple} using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dotted line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: reflecting stationary} 
@@
 in phases 1 and 2, with no mass at the boundaries.
-We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{Once again, the \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We apply the DG method with, and without, the Generalised MUSCL slope limiter.
+We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{Once again, the \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We apply the DG without a limiter and also the DG-lim and DG-lin-lim schemes which utilise a Generalised MUSCL slope limiter.
 
@@
 To approximate the point mass initial condition we compute the initial coefficients for each scheme exactly.
-For the exponential initial condition in (\ref{eqn: exp init cond}) we compute the initial coefficients via Gauss-Lobatto quadrature for the DG method, by using the mid-point rule for the uniformisation method, and by using a trapezoidal rule with 2,001 points on each cell for the QBD-RAP method.
+For the exponential initial condition in (\ref{eqn: exp init cond}) we compute the initial coefficients via Gauss-Lobatto quadrature for the DG scheme, by using the mid-point rule for the uniformisation scheme, and by using a trapezoidal rule with 2,001 points on each cell for the QBD-RAP scheme.
 
 \paragraph{Model~\ref{model: simple} with exponential initial condition}
-Figure~\ref{fig: reflecting transient exp} shows the error metrics for the four different spatial discretisation schemes (DG, DG with MUSCL limiter, uniformisation and QBD-RAP).
+Figure~\ref{fig: reflecting transient exp} shows the KS and \(L^1\) metric on the CDF  for the five different schemes (DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes).
 For both error metrics the DG scheme converges rapidly until computational errors become significant.
-The uniformisation and QBD-RAP schemes converge at a slower rate than the DG method, with the QBD-RAP scheme converging faster than the uniformisation method.
-The DG scheme with MUSCL limiter does not appear to be converging, which suggests there is at least one iteration during the numerical integration over time at which the numerical solution displays oscillations.
+The uniformisation, DG-lin-lim and QBD-RAP schemes converge at a slower rate than the DG scheme, with the QBD-RAP and DG-lin-lim schemes converging faster than the uniformisation scheme.
+For the KS metric, the DG-lin-lim scheme outperforms the QBD-RAP scheme at all orders, while for the \(L^1\) metric, the two schemes perform similarly.
+
+The DG-lim does not appear to be converging, which suggests there is at least one iteration during the numerical integration over time at which the numerical solution displays oscillations.
 However, when the DG approximations for the distribution at time \(t=2\) are plotted, they do not display oscillations (not shown).
-This suggests that the oscillations which might occur with the DG scheme must be transient, and have dissipated by time \(t=2\).
+This suggests that the oscillations which might occur with the DG scheme are transient, and have dissipated by time \(t=2\).
 The oscillations which the limiter detects in the DG scheme are likely from the reflecting boundary at \(x=0\).
@@
 Meanwhile, the initial density in Phase~1 moves to the right at rate 1 and the density reflected at the boundary from Phase~2 fills in to the region between \(x=0\) and \(x=t\) in Phase~1.
-Intuitively, in the early evolution of the model, there will be a sharp peak in the transient density at \(x=t\) -- I suspect that the transient density is continuous at this point, but not differentiable.
+Intuitively, in the early evolution of the model, there will be a sharp peak in the transient density at \(x=t\).
+I suspect that the true transient density is continuous, but not differentiable, at the point \(x=t\) for \(t<10\).
 %
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: reflecting transient exp} 
@@
 Figure~\ref{fig: reflecting transient exp cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition.
-Since the cell-wise errors do not require us to reconstruct the value of the solution within each cell, then this metric allows us to observe the error characteristics of the methods without reconstruction.
-Figure~\ref{fig: reflecting transient exp cp} shows similar convergence characteristics to the other error metrics in Figure~\ref{fig: reflecting transient exp}.
+Since the cell-wise errors do not require us to reconstruct the value of the solution within each cell, then this metric allows us to observe the error characteristics of the schemes without reconstruction.
+Figure~\ref{fig: reflecting transient exp cp} shows similar convergence characteristics to the KS error metric in Figure~\ref{fig: reflecting transient exp} (left).
 \begin{figure}[h]
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/L1_cell_probs_error_formatted.pdf}
-	\caption{Cell-wise error metric from (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\caption{Cell-wise error metric from (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: reflecting transient exp cp} 
@@
 \paragraph{Model~\ref{model: simple} with a point-mass initial condition}
-Figure~\ref{fig: reflecting transient pm} shows the error metrics for the four different spatial discretisation schemes (DG, DG with MUSCL limiter, uniformisation and QBD-RAP).
+Figure~\ref{fig: reflecting transient pm} shows the KS and \(L^1\) metric on the CDF  for the five different schemes (DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes).
 Comparing the error metrics in Figure~\ref{fig: reflecting transient pm} for the point mass initial condition with the ones in Figure~\ref{fig: reflecting transient exp} for the exponential initial condition all schemes perform worse for the point mass initial condition.
-Regarding comparative rates of convergence for this problem, the DG scheme converges fastest, followed by the QBD-RAP scheme, then the uniformisation scheme, while the DG scheme with MUSCL limiter does not appear to converge due to the limiter reducing the scheme to a linear approximation around the discontinuity.
+Regarding comparative convergence of error for this problem, for both error metrics in the DG scheme converges fastest, followed by the QBD-RAP and DG-lin-lim schemes which converge comparatively, then the uniformisation scheme, while the DG-lim does not appear to converge due to the limiter reducing the scheme to a linear approximation around the discontinuity.
 %
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) errors between of the CDF (right) at time \(t=2\) for Model~\ref{model: simple} with the point-mass initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\caption{KS (left) and \(L^1\) (right) errors between the CDF at time \(t=2\) for Model~\ref{model: simple} with the point-mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: reflecting transient pm} 
@@
 To investigate how the position of the discontinuity might affect the error metrics we evolved the model to time \(t=2.1\) and computed the KS error, \(L^1\) error between the CDFs and the cell-wise error.
-Once again, we use simulation and bootstrapping to approximate the \emph{ground truth}.
+Once again, we use simulation and bootstrapping to approximate the true distribution.
 The plots (not shown) of KS error and \(L^1\) error between the CDFs for the model at times \(t=2\) and \(t=2.1\) are relatively similar and not noteworthy.
 The plot of the cell-wise error metric at time \(t=2.1\) in Figure~\ref{fig: reflecting transient pm t2.1} is somewhat more interesting.
-Figure~\ref{fig: reflecting transient pm t2.1} shows that the cell-wise error metric is somewhat volatile for the DG scheme; I suspect that this is due to the oscillations in the DG approximation.
+Firstly, Figure~\ref{fig: reflecting transient pm t2.1} shows that all schemes improved from time \(t=2\) to time \(t=2.1\) with respect to this metric.
+Figure~\ref{fig: reflecting transient pm t2.1} also shows that the cell-wise error metric is somewhat volatile for the DG scheme; I suspect that this is due to the oscillations in the DG approximation.
 In contrast, in Figure~\ref{fig: reflecting transient pm t2.1} the uniformisation and QBD-RAP schemes have monotonically decreasing error curves.
+Also interesting is the DG-lin-lim scheme, which does not improve past dimension 7.
+I suspect that the discontinuity at \(x=t\) is not easily resolved by the scheme given the presence of the limiter.
+Figure~\ref{fig: reflecting transient pm t2.1 cdfs} plots the CDFs in Phase~1 around the discontinuity in the transient distribution at time \(t=2.1\) for the dimension 21 QBD-RAP and dimension 22 DG-lin-lim schemes.
+Neither scheme does a particularly good job of capturing the discontinuity, but the QBD-RAP scheme appears to perform slightly better.
 \begin{figure}[h]
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/L1_cell_probs_error_formatted.pdf}
-	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2\) with the point mass initial condition, where the approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
+	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2\) with the point mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
 Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1/L1_cell_probs_error_formatted.pdf}
-	\caption{\(L^1\) errors between the simulated and approximated cell masses for Model~\ref{model: simple} at time \(t=2.1\) with the point mass initial condition, and the corresponding approximations obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (green dashed line), uniformisation method (orange dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2.1\) with the point mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: reflecting transient pm t2.1} 
+\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mas_t_2.1/order_21_CDFs_formatted.pdf}
+	\caption{Dimension 21 QBD-RAP and dimension 22 DG-lin-lim approximations to the CDF in Phase~1 for Model~\ref{model: simple} with a point-mass initial condition.
+A CDF estimated via simulation is plotted in grey.}
+	\label{fig: reflecting transient pm t2.1 cdfs} 
 \end{figure}
@@
 
-We use the DG, uniformisation and QBD-RAP methods to discretise the fluid queue and partition \([0,1]\) into three intervals of width \(1/3\).
+We use the DG, uniformisation and QBD-RAP schemes to discretise the fluid queue and partition \([0,1]\) into three intervals of width \(1/3\).
 To capture the mass which has left the interval \([0,1]\), we suppose that when the process hits the boundary it is absorbed forever at the boundary and in the phase in which the process first hit the boundary.
 
-We integrate the schemes until time \(t=10\) using the SSPRK4 method with \(t\)-step size 0.005/3.\footnote{Since we use a smaller cell-width in this example than in previous examples we need to reduce the \(t\)-step size accordingly to ensure that numerical integration is stable for schemes up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We use the DG scheme with and without a slope limiter during the time-integration.
+We integrate the schemes until time \(t=10\) using the SSPRK4 method with \(t\)-step size 0.005/3.\footnote{Since we use a smaller cell-width in this example than in previous examples we need to reduce the \(t\)-step size accordingly to ensure that numerical integration is stable for schemes up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We use the DG scheme with and without a slope limiter during the time-integration and implement both the DG-lim and DG-lin-lim schemes.
 At each time-step of the numerical integration, we record the amount of mass at the absorbing boundaries in each phase, which gives us an approximation of the cumulative distribution function of the hitting time in each phase up to time \(t=10\).
@@
 \paragraph{Exponential initial condition}
-Figure~\ref{fig: hitting time exp} shows the error metrics recorded for the four different numerical approximation schemes.
-The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP converging at a faster rate.
+Figure~\ref{fig: hitting time exp} shows the error metrics recorded for the five numerical approximation schemes applied to the hitting time problem with the exponential initial condition.
 For both error metrics the DG scheme performs the best.
-For the \(L^1\) error metric the DG scheme converges up to order 5 after which there is no improvement in the error; I believe that this is due to other numerical errors dominating.
-The DG scheme with MUSCL limiter does not appear to converge, suggesting oscillations in the DG approximation.
-Intuitively, with this initial condition we suspect that the transient distribution of the fluid queue on the event that it remains in the interval \((0,1)\) is discontinuous at \(x=t\) in Phase~1 and \(x=1-t\) in Phase~2 for \(x<1\).
+For the \(L^1\) error metric the DG scheme converges up to order 5, after which there is no improvement in the error; I believe that this is due to other numerical errors dominating.
+The DG-lim does not appear to converge, suggesting oscillations in the DG approximation.
+Intuitively, with this initial condition we suspect that the transient distribution of the fluid queue on the event that it remains in the interval \((0,1)\) is discontinuous at \(x=t\) in Phase~1 and \(x=1-t\) in Phase~2 for \(x,t<1\).
 We also expect there is a discontinuity in the hitting time PDF at time \(t=1\).
-Indeed, small oscillations are visible in hitting time PDF for both phases, see Figure~\ref{fig: hitting time oscillation exp}.
+Regarding the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme, all appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly and converging at a faster rate than the uniformisation scheme.
+
+In Figure~\ref{fig: hitting time oscillation exp} we plot the PDFs of the hitting times \(\xi_x(\{0,1\})\), in each phase, for the dimension 21 DG, dimension 21, QBD-RAP and dimension 22 DG-lin-lim schemes.
+Figure~\ref{fig: hitting time oscillation exp} demonstrates oscillations in the DG approximation, though the PDF is not negative in this case.
+All schemes seem to capture the discontinuity at \(t=1\) relatively well.
+There is an interesting artefact in the QBD-RAP approximation around \(t=0\) where the scheme has generated an oscillation.
+I suspect that this artefact should disappear as the dimension of the scheme is increased, unlike the DG scheme where oscillations can persist when the dimension is increased.
+Let's investigate this artefact further.
+
+Intuitively, most sample paths which exit near \(t=0\) start near the boundary then see a change of phase shortly after \(t=0\) and remain in that phase until hitting \(\{0,1\}\).
+Consider such a sample path which starts at \(x_0=0\) in Phase~1, then exits at time \(v\) with a single change of phase at time \(v-u>0\).
+The QBD-RAP approximation to this sample paths has density\footnote{One advantage of the QBD-RAP and uniformisation approaches is that, from their stochastic interpretation, we can use sample-path arguments to analyse the scheme.}
+\begin{align}
+	\bs \alpha e^{(\bs S - \bs I)(v-u)} \bs D e^{(\bs S - 1.1\bs I)u}\bs s = \bs \alpha e^{\bs S(v-u)} \bs D e^{\bs Su}\bs s e^{-v-0.1u},\label{eqn :lf}
+\end{align}
+and since the sample path will hit \(\{0\}\) at \(v=2u\), intuitively, we would expect (\ref{eqn :lf}) to approximate a point mass at \(v=2u\).
+To see this, recall that we can approximate a point mass at \((v-u)\) by
+\begin{equation}\bs k(\Delta - (v-u))\bs e^{\bs Sx}\bs s,\label{eqn:aafba}\end{equation}
+where \(\bs k(z)=\cfrac{\bs \alpha e^{\bs Sz}}{\bs \alpha e^{\bs Sz\bs e}}\),
+and recall that the matrix \(\bs D\) is an approximation; 
+\begin{align}
+	\bs \alpha e^{\bs S(v-u)}\bs D &= \mathbb E\left[\bs k(Z-(v-u))1(Z>v-u)\right]\nonumber 
+	% \\&\approx \bs \alpha e^{\bs S(v-u)}\bs e \bs k(\Delta-(v-u))
+	\\&\approx\mathbb P(Z>v-u) \bs k(\Delta-(v-u)), \label{eqn: aegpb}
+\end{align} where \(Z\sim ME(\bs \alpha,\bs S)\)
+and \(Z\) is concentrated around \(\Delta\).
+
+Thus, (\ref{eqn :lf}) is approximately  
+\[e^{-v-0.1u} \mathbb P(Z>v-u) \bs k(\Delta-(v-u))\bs e^{\bs Su}\bs s,\]
+which approximates a point mass at \(u=v-u\), or \(2u=v\).
+Retracing our arguments, sources of error in this approximation might come from the approximation in (\ref{eqn: aegpb}) and from the approximation of a point mass by (\ref{eqn:aafba}), which we commented on in Section~\ref{sec: closing vecs numerics} where we conducted numerical experiments about the unnormalised closing operator given by \(\bs v(x) = e^{\bs S x}\bs s\).
+
+% To further comment on the approximation of a point mass by (\ref{eqn:aafba}), recall the . We found that this closing operator has difficulty approximating mass at .
+
 \begin{figure}[h]
@@
 	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the exponential initial condition.
-The approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (orange dashed line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: hitting time exp} 
@@
 	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/pdf_order21.pdf}%
-	\caption{Approximations of the PDFs of the first hitting time for Model~\ref{model: simple} with the exponential initial condition.
-The blue line were obtained from the dimension-21 DG scheme and the purple lines were obtained from the dimension-21 QBD-RAP scheme.
+	\caption{Approximations of the PDF of the first hitting time for Model~\ref{model: simple} with the exponential initial condition.
+The blue lines were obtained from the dimension-21 DG scheme, the purple lines were obtained from the dimension-21 QBD-RAP scheme and the gold lines were obtained from the dimension 22 DG-lin-lim scheme.
 The DG scheme displays oscillations around the discontinuities at \(t=1\).
+The QBD-RAP scheme has an oscillation near \(t=0\).
 }
@@
 \paragraph{Point mass initial condition}
-Figure~\ref{fig: hitting time pm} shows the KS error (left) and \(L^1\) error for the CDF for the four different numerical approximation schemes applied to the hitting time problem with the point mass initial condition.
-The uniformisation and QBD-RAP method both appear to converge, with the QBD-RAP converging at a faster rate.
-The DG scheme with MUSCL limiter does not appear to converge, suggesting oscillations in the DG approximation.
-The DG scheme appears to converge fastest, however, oscillations are present in the DG approximation.
-In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis.
-Observing Figure~\ref{fig: hitting time oscillation} the DG approximation clearly displays oscillations around and the discontinuity at \(t=1\).
+Figure~\ref{fig: hitting time pm} shows the KS error (left) and \(L^1\) error for the CDF (right) for the five numerical approximation schemes applied to the hitting time problem with the point mass initial condition.
+The DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation as we might expect given the discontinuity.
+Of the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes all appear to converge, with the QBD-RAP converging fastest and the DG-lin-lim and uniformisation schemes performing similarly.
+The DG scheme appears to perform best, however, oscillations are present in the DG approximation.
+In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis and for the DG-lin-lim scheme with dimension 22.
+Observing Figure~\ref{fig: hitting time oscillation} the DG approximation displays oscillations around and the discontinuity at \(t=1\) while the other two schemes do not.
+Of the two positivity preserving schemes in Figure~\ref{fig: hitting time oscillation}, then QBD-RAP scheme appears to approximate the discontinuity best.
 \begin{figure}[h]
@@
 	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
-The approximations were obtained via the DG method (blue solid line), DG method with Generalised MUSCL limiter (orange dashed line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: hitting time pm} 
@@
 	\caption{Approximations of the CDF of the first hitting time in Phase~1 for Model~\ref{model: simple} with the point mass initial condition.
-The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dashed line is the empirical CDF obtained via simulation.
+The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dashed line from the dimension 22 DG-lin-lim scheme.
+The empirical CDF obtained via simulation is plotted as the teal dashed line.
 The DG scheme displays oscillations.
@@
 Now consider the fluid-fluid queue model first analysed in \cite{blnos2022} which is a modified version of a model first presented in \cite{lnp13}.
-We refer the reader to \cite{blnos2022} for more on the DG method applied to this model.
+We refer the reader to \cite{blnos2022} for more on the DG scheme applied to this model.
 \begin{model}\label{model: ffq}
@@
 
-We obtained approximations to the infinitesimal generator of the fluid queue \(\{(X(t),\varphi(t))\}_{t\geq 0}\) via the DG method, QBD-RAP method and uniformisation method, all of which used a cell width of \(\Delta=0.4\).
+We obtained approximations to the generator of the fluid queue \(\{(X(t),\varphi(t))\}\) via the DG, QBD-RAP and uniformisation schemes, all of which used a cell width of \(\Delta=0.4\).
 The matrix resulting from the approximation is then used to approximate the first-return operator \(\mathbb\Psi(s)\) as discussed in Section~\ref{sec: intro Psi}.
-Due to the stochastic interpretation of the uniformisation and QBD-RAP schemes the approximations to \(\Psi(s)\) have a stochastic interpretation as the first-return probabilities of a fluid queue driven by a CTMC and QBD-RAP, respectively (see \cite[Chapter~7]{p2019} and \cite{bgnp2021} for details on the latter).
-For the DG method the approximation of \(\mathbb \Psi(s)\) is not as well-understood.\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{Y}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}
+Due to the stochastic interpretation of the uniformisation and QBD-RAP schemes the approximations to \(\mathbb \Psi(s)\) have a stochastic interpretation as the first-return probabilities of a fluid queue driven by a CTMC and QBD-RAP, respectively (see \cite[Chapter~7]{p2019} and \cite{bgnp2021} for details on the latter).
+For the DG scheme the approximation of \(\mathbb \Psi(s)\) is not as well-understood.\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{Y}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}
 
@@
 In Figure~\ref{fig: ffq return cts} we plot the error metrics for the approximations to the distribution (\ref{eqn: first return Y 1}).
-The DG method performs best converging rapidly until the error in the approximation scheme is swamped by other numerical errors.
-The QBD-RAP method is second best and the uniformisation scheme appears to be the slowest to converge.
-Here, the first return distribution appears to be smooth, hence we might expect the DG method to perform well.
-Note that there is a significant number other sources of error here; machine precision errors, errors in solving the matrix Ricatti equation to approximate \(\mathbb \Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), and truncation errors.
+The DG scheme performs best converging rapidly until the error in the approximation scheme is swamped by other numerical errors.
+The QBD-RAP scheme is second best and the uniformisation scheme appears to be the slowest to converge.
+Here, the first return distribution appears to be smooth, hence we might expect the DG scheme to perform well.
+Note that there is a significant number other sources of error here; machine precision errors, errors in solving the matrix Riccati equation to approximate \(\mathbb \Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), and truncation errors.
 Furthermore, for the QBD-RAP, since the parameters \(\bs \alpha,\,\bs S,\,\bs s\,\) and \(\bs D\) are found numerically, then there is another source of error from this.
@@
 	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) in (\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}.
-The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: ffq return cts} 
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cell_probs_error_formatted.pdf}%
-	\caption{Total error between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}.
-The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+	\caption{Cell-wise errors between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}.
+The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: ffq cell probs} 
@@
 \end{model}
-As before, we use the DG, uniformisation and QBD-RAP methods to approximate the model, and compare to simulations.
+As before, we use the DG, uniformisation and QBD-RAP schemes to approximate the model, and compare to simulations.
 
@@
 Figure~\ref{fig: ffq return discts} plots the error metrics for the first return distribution.
-Observing Figure~\ref{fig: ffq return discts} we see that the DG method performs well, followed by the QBD-RAP method, then the uniformisation method performs worst.
-However, the DG method results in an oscillatory solution with the CDF taking impossible values (decreasing at points) as shown in Figure~\ref{fig: ffq2 oscillation}
+Observing Figure~\ref{fig: ffq return discts} we see that the DG scheme performs well, followed by the QBD-RAP scheme, then the uniformisation scheme performs worst.
+However, the DG scheme results in an oscillatory solution with the CDF taking impossible values (decreasing at points) as shown in Figure~\ref{fig: ffq2 oscillation}
 \begin{figure}[h]
@@
 	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}.
-The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: ffq return discts} 
@@
 	\caption{Cell-wise error between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing in each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq2}.
-The approximations were obtained via the DG method (blue solid line), uniformisation method (green dashed line) and QBD-RAP method (purple dotted line).
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
+The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
 	\label{fig: ffq2 cell probs} 
@@
 	\caption{Approximations of the CDF of the distribution of \(X\) at the time \(\zeta_Y(\{0\})\) in Phase~1 for Model~\ref{model: ffq2}.
-The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dashed line is the empirical CDF obtained via simulation.
+The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dotted line is the empirical CDF obtained via simulation.
 The DG scheme displays oscillations around the discontinuity.
@@
 \section{Discussion}
-
-The structure of this chapter is as follows.
-In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different methods and observe their performance at approximating the initial condition, which allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
-In Section~\ref{sec: wave num} we investigate a simple travelling wave problem with various initial conditions.
-For this problem the dynamics are deterministic, which allows us to instrument the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
-Next, we investigate a simple fluid queue with two phases.
-Section~\ref{sec:stat} investigates the ability of the methods to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the fluid queue for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the methods to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
-Section~\ref{sec: ffq num} approximates the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\) for two simple fluid-fluid queue models.
-
-In this chapter we have numerically investigated some properties of the QBD-RAP approximation scheme and compared it to existing methods; the uniformisation scheme of \cite{bo2013} and the discontinuous Galerkin scheme.
-In general, the numerical experiments show that, for problems with discontinuities, the DG approximation can exhibit oscillations, while the QBD-RAP and uniformisation approximations avoid this and, of the latter two, the QBD-RAP scheme often converges faster.
+In this chapter we have numerically investigated some properties of the QBD-RAP approximation scheme and compared it to existing schemes; the uniformisation scheme of \cite{bo2013} and the discontinuous Galerkin scheme.
+In general, the numerical experiments demonstrate the smoother the problem is the better the performance of the DG scheme, and it emphatically outperforms the other two schemes.
+However, for problems with discontinuities the DG approximation can exhibit oscillations, while the QBD-RAP and uniformisation approximations avoid this and, of the latter two, the QBD-RAP scheme often converges faster.
 To avoid the problems of oscillations we can sometimes employ a \emph{slope limiter} with the DG scheme, which effectively reduces the scheme to linear in the regions where oscillations are detected.
-The numerical experiments demonstrate a loss of accuracy in the DG approximation when a slope limiter is used for a purely discontinuous problem.
-Moreover, for the application of the methods to fluid-fluid queues, there is no obvious way to apply the concept of a slope limiter.
+In some contexts, we implemented two slope-limited DG schemes, the DG-lim scheme which takes a high order DG scheme and limits the solution as necessary, and the DG-lin-lim, which is a linear approximation on a finer grid and is designed to use approximately the same computational resources as the other schemes we consider.
+The numerical experiments demonstrate a loss of accuracy in the approximation when a DG-lim scheme is used for a purely discontinuous problem.
+The numerical experiments suggest that the DG-lin-lim scheme can perform well, and is similar to the performance of the QBD-RAP scheme, in the presence of discontinuities.
+For highly discontinuous problems (i.e.~problems with point masses) the QBD-RAP scheme can outperform the DG-lin-lim scheme.
+%For the application of the schemes to fluid-fluid queues, there is no obvious way to apply the concept of a slope limiter.
 %When the slope limiter detects oscillations in the approximate solution, it reduces the DG scheme to linear in the region surrounding the oscillations, otherwise, the slope limiter leaves the solution unchanged.
 Thus, the slope limiter permits high-order approximations away from oscillations, while also removing oscillations.
-In general, we observe that the smoother the problem is the better the performance of the DG method, and it emphatically outperforms the other two methods.
-
-As a first step in the numerical experiments, in Section~\ref{sec: recon num}, we examined the ability of each method to approximate different initial conditions.
-For the DG method, this is equivalent to a projection of the initial condition on to a set of basis polynomials.
-For the uniformisation method this is equivalent to projecting the initial condition on to a basis of constant functions.
-For the QBD-RAP method the approximation of the initial condition is as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
+
+As a first step in the numerical experiments, in Section~\ref{sec: recon num}, we examined the ability of each scheme to approximate different initial conditions.
+For the DG scheme, this is equivalent to a projection of the initial condition on to a set of basis polynomials, for the uniformisation scheme this is equivalent to projecting the initial condition on to a basis of constant functions, and for the QBD-RAP scheme the approximation of the initial condition is as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
 Section~\ref{sec: comp} demonstrates that the DG scheme (projection) can result in oscillations and negative regions in the approximation when the initial condition is discontinuous.
-The uniformisation and QBD-RAP methods avoid this problem, but appear to have higher errors and the QBD-RAP method appears to have the largest errors.
-For discontinuous initial conditions the rates of convergence are comparable for all three methods.
+The uniformisation and QBD-RAP schemes avoid this problem, but appear to have higher errors and the QBD-RAP scheme appears to have the largest errors.
+For discontinuous initial conditions the rates of convergence are comparable for all three schemes.
 However, when the initial condition to be approximated is sufficiently smooth, then the DG approximation is superior.
 
-Next, in Section~\ref{sec: wave num}, we instrumented the performance of approximations for a simple travelling wave problem with various initial conditions.
-For this problem the solution is given in terms of the initial condition by \(f(x,t) = f(x-t,0)\).
-This model is useful as it allows us to instrument the ability of the methods to capture the flow of probability without stochastic dynamics.
-Once again, for discontinuous problems, the DG method can display oscillations, while the other methods (DG with slope limiter, uniformisation and QBD-RAP methods) avoid oscillations.
-Further, for discontinuous problems the rates of convergence of the QBD-RAP and DG methods can be similar.
-For smooth problems the DG method is superior.
-Interestingly, even though the QBD-RAP method performed worst for approximating initial conditions in the previous section, it outperformed the uniformisation method for this model, demonstrating that the QBD-RAP method can capture the dynamics of the flow of probability better than the uniformisation method.
+Next, in Section~\ref{sec: wave num}, we instrumented the performance of approximations for a simple travelling wave model with various initial conditions.
+For this model the solution is given in terms of the initial condition by \(f(x,t) = f(x-t,0)\).
+This model is useful as it allows us to instrument the ability of the schemes to capture the flow of probability without stochastic dynamics, and because the solution is known.
+With the travelling wave model we demonstrate that, for problems with discontinuities, the DG scheme can display oscillations, while the other schemes (DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes) avoid oscillations.
+For discontinuous problems the rates of convergence of the QBD-RAP and DG-lin-lim schemes is similar.
+Interestingly, even though the QBD-RAP scheme performed worst for approximating initial conditions in the previous section, it outperformed the uniformisation scheme for this model, demonstrating that the QBD-RAP scheme can capture the dynamics of the flow of probability better than the uniformisation scheme.
+For smooth problems the DG scheme is superior.
 
@@
 First, in Section~\ref{sec:stat}, we looked at the stationary distribution, which is known to be smooth.
-Since the problem is smooth, then the DG method was superior as expected.
-Of the uniformisation and QBD-RAP methods, the QBD-RAP method gives more accurate solutions.
+Since the problem is smooth, then the DG scheme was superior as expected.
+Of the uniformisation and QBD-RAP schemes, the QBD-RAP scheme gives more accurate solutions.
 In Section~\ref{sec: transient approx} we turned our attention to approximating transient distributions for the same models and consider two different initial conditions, a point-mass and an exponential initial condition.
 The discontinuous initial condition results in a discontinuous transient distribution.
-As for the exponential initial condition, this example demonstrates that, even if the initial condition appears `nice', it can still result in discontinuous, or non-differetiable solutions.
-The numerical evidence suggests that the DG method can display oscillations, while the other methods do not.
-The DG method with slope limiter detects the oscillations and reduces the method to linear.
-Of the uniformisation and QBD-RAP methods, the latter performs better.
+As for the exponential initial condition, this example demonstrates that, even if the initial condition appears `nice', it can still result in non-smooth (i.e.~non-differetiable) solutions.
+The numerical evidence suggests that the DG scheme can display oscillations, while the other schemes do not.
+The DG-lim scheme detects the oscillations and reduces the approximation to linear.
+Of the uniformisation, DG-lin-lim and QBD-RAP schemes, the latter two performs similarly and better than the uniformisation scheme.
+Of the DG-lin-lim and QBD-RAP schemes the QBD-RAP scheme performed better in the presence of a point mass while the DG-lin-lim scheme performed better with the exponential initial condition.
 
@@
 For this problem there is never any in-flow of mass at the boundaries of the interval and so, for a solution to be continuous, the initial condition needs to be chosen carefully, otherwise discontinuities in the transient distribution may result, as is the case for both initial conditions here.
-The numerical results suggest that, due to the discontinuities in the problems, the DG method may display oscillations.
-Since the uniformisation and QBD-RAP methods can handle discontinuities, they perform as expected, with the QBD-RAP method performing better than the uniformisation.
-
-Lastly, we applied the DG, uniformisation and QBD-RAP methods to two simple fluid-fluid queues in Section~\ref{sec: ffq num}.
-In the first fluid-fluid queue, which appears to have a smooth solution, the DG method performs very well.
+The numerical results suggest that, due to the discontinuities in the problems, the DG scheme may display oscillations.
+Since the uniformisation DG-lin-lim, and QBD-RAP schemes can handle discontinuities, they perform as expected, with the QBD-RAP and DG-lin-lim schemes performing similarly and better than the uniformisation.
+
+Lastly, we applied the DG, uniformisation and QBD-RAP schemes to two simple fluid-fluid queues in Section~\ref{sec: ffq num}.
+In the first fluid-fluid queue, which appears to have a smooth solution, the DG scheme performs very well.
 Of the two positivity preserving schemes the QBD-RAP scheme performs better than the uniformisation scheme.
-In the second example, which has a discontinuity, the DG method produces the lowest errors, but exhibits oscillations in the solution.
-The QBD-RAP and uniformisation schemes do not produce oscillations and, of the two, the QBD-RAP method performs best.
-
-In conclusion, when the problem is known to be smooth, the DG method is very likely to produce excellent results.
-However, for discontinuous problems, the method can show oscillations and infeasible or negative regions of the solution.
-The slope limiter overcomes this, but reduces the accuracy of the DG method to linear near discontinuities, sometimes severely affecting the quality of the approximation.
-The uniformisation and QBD-RAP methods are alternative approximation schemes which exhibit larger errors, but avoid oscillatory solutions and guarantee a non-decreasing approximation of the CDF.
-Of the uniformisation and QBD-RAP methods, the latter often to produced lower errors.
+In the second example, which has a discontinuity, the DG scheme produces the lowest errors, but exhibits oscillations in the solution.
+The QBD-RAP and uniformisation schemes do not produce oscillations and, of the two, the QBD-RAP scheme performs best.
+
+In conclusion, when the problem is known to be smooth, the DG scheme is very likely to produce excellent results.
+However, for discontinuous problems, the scheme can show oscillations and infeasible or negative regions of the solution.
+The slope limiter overcomes this, but reduces the accuracy of the DG scheme to linear near discontinuities, sometimes severely affecting the quality of the approximation.
+A peicewise-linear DG scheme with a limiter, such as the DG-lin-lim scheme considered here, can produce satisfactory results.
+The uniformisation and QBD-RAP schemes are other alternative approximation schemes which can exhibit larger errors, but avoid oscillatory solutions and guarantee a non-decreasing approximation of the CDF and have a stochastic interpretation.
+Of the uniformisation and QBD-RAP schemes, the latter often produced lower errors and performed similarly to the DG-lin-lim scheme.

### 2022-02-10 chapter7/chapter.tex
.

%!TEX root = ../thesis.tex
\chapter{Conclusions \label{ch: conclusion}}

### 2022-02-10 intro/chapter.tex
.

%!TEX root = ../thesis.tex
\chapter{Introduction \label{ch:intro}} 
% intro to intro
% contextual background
	% Fluid queues 
		% definition
		% applications 
		% existing results 
	% Fluid-fluid queues
		% definition 
		% basic analysis in terms of generators
		% the need for approximation and how the old work doesn't fit here
	% The DG method
		% pertition into cells
		% project on to a basis
		% problems: oscillations
		% soutions and why they dont fit
		% observation about constant-basis methods
		% this thesis asks the question: motivated by the constant basis being a probability model, can we dream up a more accurate probability model to approximate a fluid queue?
	% On the structure of the constant basis/uniformisation method (QBD)
	% least variable is PH, so what about MEs?
% structure of the thesis
	% rest of chapter 1 gives mathematical preliminaries
	% chapter 2 explains the DG method, problems/oscillations, slope limiting, and appllication to SFFMs
	% inspired by the structure of the order-1 scheme, and to solve the negativity problem, chapter 3 introduces a new approximation method; The QBD-RAP. In this chapter we derive the intuitively of the model and explain it's dynamics.
	% we then move to convergence, with chapter 4 proving a convergence of the QBD-RAP up to the first hitting time of the fluid queue on the edges of an interval. technical results and extensions are left to the appendix: certain properties of closing vectors, convergence without ephemeral phases, and certain matrix algebraic manipulations. 
	% chapter 5 stitches together the results of chapter 4 to prove a global result about convergence
	% chapter 6 investigate, numerically, the performance of the DG, DG with limiter, uniformisation and QBD-RAP schemes. 
	% chapter 7  makes concluding remarks.
% mathematical preliminaries 
	% CTMCs
	% fluid queues
	% fluid-fluid queues and operator-analytic expressions (all of them)
		% the equations of bo2014
		% meets the partition of the DG/QBD-RAP, and reconstructing the Fil sets 
    % why we cant solve the equations, what we need to approximate to solve it, i.e. B, R, then D, Psi
	% projections
	% phase-type distributions 
		% definition
		% least variable property
	% matrix exponentials
		% definition 
		% properties 
		% verification that parameters give an ME
		% CMEs
	% QBD-RAPs, then QBDs as a special case
		% orbit processes, their interpretation for PH and how they differ for MEs
	% convergence theorems 
		% DCT 
		% Continuity of Laplace transforms

% THE ROUGH IDEA & CONTEXT
% 	THE CONTEXT: FLUID-FLUID QUEUES
%	PDEs 
%	NON-NEGATIVITY & CONSERVATION 
% MATHEMATICAL PRELIMINARIES

An fluid queue is a two-dimensional stochastic process \(\{{\bs X}(t)\} = \{( X(t),\varphi(t))\}_{t\geq0}\). The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC) and determines the rate at which \(\{ X(t)\}\) moves. The level process, \(\{ X(t)\}_{t\geq0}\), is real-valued, continuous, piecewise linear and deterministic given \(\{\varphi(t)\}\).

Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982}, a canonical application in this area), power systems \citep{hydro}, risk processes \citep{betal2005} and environmental modelling \citep{wurm2020}. Fluid queues are relatively well studied. Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods e.g.~\cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \cite{anick1982,kk1995,blnos2022}. %For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest. Anick Mitra Sondhi CITE, analysed fluid queues using a more direct differential equation-based method. Since Ramaswami's CITE initial work, there has been significant developments in the analysis of fluid queues CITE and related algorithms CITE.

More recently, \cite{bo2014} extended fluid queues to so-called \emph{stochastic fluid-fluid queues}. In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\), so \((X(t),\varphi(t))\) determines the rate at which \(\{Y(t)\}\) moves. The analysis of \cite{bo2014}, is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used. To this end, to date, two possible discretisation have been suggested. Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise this operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process. The QBD approximation of \cite{bo2013} is derived via a uniformisation argument, so we refer to it as the uniformisation approximation scheme throughout this thesis. Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations. It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the former; the uniformisation scheme can be viewed as the simplest DG scheme where the operator is projected onto a basis of piecewise constant functions with an upwind flux.

%A QBD can be viewed as a two-dimensional CTMC, \(\{(L(t),\varphi(t))\}_{t\geq0}\), where \(\{L(t)\}\) is the discrete level process, and \(\{\varphi(t)\}_{t\geq0}\) is the phase process. The level process \(\{L(t)\}\) is skip free, meaning that, given the process is at level \(L(t)=\ell\), is may only jump to \(\ell+1\) or \(\ell-1\) at jump epochs. The sojourn time of \(\{L(t)\}\) in a given level follows a phase-type distribution

In the context of approximating fluid queues, one advantage of the uniformisation scheme and, equivalently, a DG scheme with constant basis functions, is that is guarantees probabilities computed from the approximation are positive \citep[Section~3.3]{koltai2011}. One justification for the positivity preserving property of the uniformisation scheme is from its interpretation as a stochastic process. For higher order DG schemes there is no such interpretation and positivity is not guaranteed \citep[Section~3.3]{koltai2011}. Moreover, higher-order DG approximation schemes may produce negative and oscillatory solutions, particularly when discontinuities or steep gradients are present. Methods to navigate the problem of negative and oscillatory solutions have been developed, such as filtering and slope limiting (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6 and references therein). Slope limiting alters the discretised operator in regions where oscillations are detected and reduces the order of the approximation to linear in these regions. Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.

Depending on the context, filtering of the approximate solution to remove oscillations may not necessarily guarantee a strictly non-negative approximation, or may result in severe smearing of the solution at discontinuities or regions with steep gradients \citep[Section~5.6.1]{nodalDGBook}. Moreover, filtering requires us to make a trade-off between filtering enough of the oscillations away while maintaining sufficient accuracy -- a choice which may not be obvious \emph{a priori}. Slope limiting does guarantee positivity but reduces the approximation to linear where oscillations in the approximate solution are detected \citep[Section~5.6.1]{nodalDGBook}. Furthermore, limiting and filtering do not distinguish between natural oscillations which are a fundamental feature of the solution and spurious oscillations which are caused by the approximation scheme, and they may remove both from the approximation. This can lead to an unnecessary loss of accuracy in the approximation (see \citep{nodalDGBook},~Example~5.8).

In the context of approximating the first return operator of a fluid-fluid queue the application of a slope limiter would amount to post-processing the solution, reducing the order of the approximation in areas where oscillations are detected to linear. Otherwise, perhaps another approach would be to re-compute the solution with lower-order basis if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions. Post-hoc filtering of the solution is also possible, but we would still need to tune the filter appropriately. Both filtering and slope limiting are also dependent on the initial condition.

In approximating the first return operator of a fluid-fluid queue we first approximate an operator-Riccati equation by a matrix-Riccati equation by substituting in matrix approximations to the infinitesimal generator of the fluid queue which are constructed via the DG scheme \cite{blnos2022}. We then solve the matrix-Riccati equation via an iterative procedure \citep{bean2005b,blnos2022}. For the DG method there is a limited theoretical backing as to why we may approximate the operator-Riccati equation with the matrix-Riccati equation other than it is a sensible thing to do, and it seems to work in practice. In the case of the QBD approximation of \cite{bo2013} this is more justifiable as in this case the resulting matrix-Riccati equation can be derived by considering the first-return probabilities for a fluid queue driven by the QBD \citep{bean2005} where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\).

Motivated by this, this thesis derives a new approximation to a fluid queue. The approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributions to model the sojourn time in a given interval on the event that the phase of the fluid is constant. The sojourn time in a given interval on the event that the phase of the fluid is constant is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution so, in this sense, the best approximation to this deterministic sojourn time. Thus, it appears that the approximation of \cite{bo2013} is the best-possible Markov chain approximation. Recently, there has been much work on a class of concentrated matrix exponential distributions \cite{hhat2020} which are postulated to be the least-variable matrix exponential distribution. Matrix exponential distributions generalise Phase-type distributions; they have the same functional form, without the restriction that the distribution has an interpretation in terms of the absorption time of a continuous-time Markov chain. A class of stochastic processes, known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \cite{bn2010}, extend QBDs, which have Phase-type inter-event times, to allow matrix-exponentially distributed inter-event times. Thus, by using matrix exponential distributions, we attempt to construct a QBD-RAP which better captures the dynamics of the fluid queue than the QBD approximation in \cite{bo2013}, while retaining a stochastic interpretation.

As the QBD-RAP has a stochastic interpretation then the approximations it produces are guaranteed to have non-negative density functions. Moreover, the matrix-Riccati equation we solve to approximate the first-return distribution of a fluid-fluid queue can be derived by considering the first-return probabilities for a RAP-modulated fluid queue driven by the QBD-RAP \citep{p2019,bgnp2021}. Another attractive property is that the QBD-RAP (and uniformisation) method are guaranteed to produce non-negative estimates of probabilities for any initial condition without any further computation or post-processing, which also means that the approximations to operators we get from the QBD-RAP (and uniformisation) scheme are linear; a desirable mathematical property.

The structure of this thesis is as follows. The rest of this chapter is dedicated to mathematical preliminaries and introduces the main mathematical objects and tools which we will need. In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way to make clear how the approximations we develop correspond to the theoretical operators.

Chapter~\ref{ch:galerkin} demonstrates a way that we can use the discontinuous Galerkin (DG) method to approximate certain operators and distributions of fluid and fluid-fluid queues. Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:galerkin} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and non-monotonic CDFs. However, one of the problems we are interested in solving is how to approximate the first-return operator of a fluid-fluid queue, which is a non-standard problem for the DG method. It is not clear how to use the concept of \emph{slope limiting} in the context of approximating the first-return operator of a fluid-fluid queue.

In the following chapter, Chapter~\ref{sec: construction and modelling}, we develop a new approximation scheme which, due to its interpretation as a stochastic process (a QBD-RAP), ensures all approximations to CDFs are non-decreasing. The chapter takes a stochastic modelling approach to developing the approximation scheme. Once we have established the new approximation scheme, we then attempt to prove convergence of the approximation scheme in Chapters~\ref{sec: conv}~and~\ref{ch: global results} via certain Laplace transforms with respect to time. Chapter~\ref{sec: conv} utilises methods and ideas relating specifically to QBD-RAPs and matrix-exponential distributions. Ultimately, Chapter~\ref{sec: conv} proves that, on the event that the QBD-RAP has not-yet seen an \emph{orbit restart epoch} (which is much like a change of level, but not always), certain Laplace transforms with respect to time of the QBD-RAP scheme converge to corresponding Laplace transforms with respect to time of the fluid queue on the event that the level of the fluid queue remains in a given interval. In this sense, this is a local convergence result as it relates only to convergence to the fluid queue in an interval.

Chapter~\ref{ch: global results} then looks to extend the local convergence result to a global convergence results on the whole domain of approximation. Chapter~\ref{ch: global results} uses more traditional Markov process arguments which rely on properties such as the strong Markov property, time-homogeneity and the Law of total probability. In Chapter~\ref{ch: global results} we first consider the discrete-time process which the process embedded in the QBD-RAP at times when the QBD-RAP sees an orbit restart epoch and prove that the embedded process converges in distribution to a corresponding embedded process in the fluid queue. The rest of Chapter~\ref{ch: global results} then attempt to prove a global converge result. The main result being Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (weakly with respect to the spatial and temporal variable) to the distribution of the fluid queue.

Once we have established a few methods for approximating fluid queues, Chapter~\ref{sec: numerics} then numerically explores some properties of the approximations numerically. Given that the QBD-RAP scheme was developed so that it guarantees positivity of the approximation, we largely focus on problems with discontinuities.

Finally, Chapter~\ref{ch: conclusion} makes concluding remarks.

To help the reader understand the notations used in the DG scheme, we provide a small toy model example in Appendix~\ref{appendix:example}. The rest of the appendices contain technical results relating to proving the convergence of the QBD-RAP approximation scheme. Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operator} introduced as part of the QBD-RAP scheme have the properties we claim they do to prove convergence in Chapter~\ref{sec: conv}. Appendix~\ref{app:extend conv} extends some results from Chapter~\ref{sec: conv} to a setting which requires slightly less computation. Lastly, Appendix~\ref{appendix: kronecker} provides some algebraic results which help us to manipulate certain Laplace transforms from Chapter~\ref{sec: conv}.

### 2022-02-10 summary/summary.tex
.

\documentclass[a4paper]{article}
\usepackage{amsmath}
\title{Summary of thesis}
\author{Angus Lewis}

\begin{document}
\maketitle
The stochastic fluid-fluid model (SFFM) is a Markov process {\(\{(X_t,Y_t,\varphi_t)\}_{t\geq 0}\),} where {\(\{\varphi_t\}_{t\geq 0}\)} is a continuous-time Markov chain, the first fluid, {\(\{X_t\}_{t\geq 0}\),} is a classical stochastic fluid process driven by {\(\{\varphi_t\}_{t\geq0}\),} and the second fluid, {\(\{Y_t\}_{t\geq0}\),} is driven by the pair {\(\{(X_t,\varphi_t)\}_{t\geq 0}\).} Operator-analytic expressions for the stationary distribution of the SFFM, in terms of the infinitesimal generator of the process {\(\{(X_t,\varphi_t)\}_{t\geq 0}\),} are known \cite{bo2014}. However, these operator-analytic expressions do not lend themselves to direct computation.

In this thesis we investigate approximations to the infinitesimal generator of the process {\(\{(X_t,\varphi_t)\}_{t\geq 0}\)} so that one may approximate the fluid-fluid performance measures derived in \cite{bo2014}. As a first step we first introduce modified versions of the operators derived in \cite{bo2014} which are partitioned so that we can directly correspond elements of the true and approximated operators.

Perhaps the simplest approximation is to approximate the continuous fluid process by a discrete-time Markov chain. This was considered in \cite{bo2013} and results in a quasi-birth-and-death process (QBD). Another approach is to use the finite-element method -- in Chapter~2 we apply the discontinuous Galerkin (DG) scheme. It turns out that the simplest DG method is equivalent to the Markov chain approximation of \cite{bo2013}. In some circumstances the DG method can result in approximations to the cumulative distribution function (CDF) which oscillate and are non-monotonic. This is clearly undesirable as we know that CDFs must be monotonic and non-decreasing. Due to its interpretation as a stochastic process, the Markov chain approximation of \cite{bo2013} ensures that the resulting approximations to the CDF are monotonic and non-decreasing. However, the rate of convergence of the Markov chain approximation of \cite{bo2013} is, in general, much slower than the DG scheme.

Motivated by this, the thesis proceeds to derive a new approximation to a fluid queue. The approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributions to model the sojourn time in a given interval on the event that the phase of the fluid is constant. The sojourn time in a given interval on the event that the phase of the fluid is constant is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution so, in this sense, the best approximation to this deterministic sojourn time. Thus, it appears that the approximation of \cite{bo2013} is the best-possible Markov chain approximation. Recently, there has been much work on a class of concentrated matrix exponential distributions \cite{hhat2020} which are postulated to be the least-variable matrix exponential distribution. Matrix exponential distributions have the same functional form as Phase-type distributions without the restriction that the distribution has an interpretation in terms of the absorption time of a continuous-time Markov chain. A class of stochastic processes, known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \cite{bn2010}, extend QBDs, which have Phase-type inter-event times, to allow matrix-exponentially distributed inter-event times. Thus, by using matrix exponential distributions, Chapter~3 attempts to construct a QBD-RAP which better captures the dynamics of the fluid queue than the QBD approximation in \cite{bo2013}. As the QBD-RAP has a stochastic interpretation then the approximations of to CDFs it produces are guaranteed to be monotonic and non-decreasing.

The thesis then moves on to mathematical analysis of the QBD-RAP process. Chapter~4 proves a convergence of the QBD-RAP process to the fluid queue up to the time that the fluid queue leaves a given interval. Chapter~5 then uses the results of Chapter~4 to prove a global convergence results of the QBD-RAP approximation scheme to the fluid queue. Chapter~4 relies more on matrix-exponential specific arguments, while Chapter~5 uses more traditional Markov process arguments, such as partitioning on the position of the process at convenient stopping times, then exploiting the strong Markov property and time homogeneity.

Chapter~6 numerically investigates the performance of the DG, QBD-RAP and Markov chain approximations. The numerics demonstrate that, if the solution is smooth enough, then the discontinuous Galerkin method is highly-effective and displays rapid convergence to the solution as the number of basis functions is increased. However, when the solution is sufficiently non-smooth, oscillations and negative approximations may occur when using the discontinuous Galerkin method, leading to nonsense solutions. In these cases, the QBD-RAP approximation performs relatively well compared to the other positivity preserving schemes investigated.

\begin{thebibliography}{9}
    \bibitem{bn2010}
    Bean, N.~G.~and Nielsen, B.~F. (2010). {`Quasi-birth-and-death processes with rational arrival process components.}' \textit{Stochastic Models}, 26(3), 309-334.

    \bibitem{bo2013}
    Bean, N.~G.~and O'Reilly, M.~M. (2013). {`Spatially-coherent uniformisation of a stochastic fluid model to a quasi-birth-and-death-process.'} \textit{Performance Evaluation}, 70(9):578 -- 592.

    \bibitem{bo2014}
    Bean, N.~G.~and O'Reilly, M.~M. (2014). {`The stochastic fluid-fluid model: A stochastic fluid model driven by an uncountable-state process, which is a stochastic fluid itself.'} \textit{Stochastic Processes and their Applications}, 124:1741 -- 1772.

    \bibitem{hhat2020}
    Horv\'ath, I.~and S\'af\'ar, O.~and Telek, M.~and Zamb\'o, B. (2020). `{High order concentrated matrix-exponential distributions.}' \textit{Stochastic Models}, 36(2), 176 -- 192.
\end{thebibliography}
\end{document}

### 2022-02-17 appendix_dg_toy/appendix.tex
first draft of the whole thing! :D

@@
 
-Consider a process \(\{(\overline X_t,Y_t,\varphi_t)\}_{t\geq 0}\) with two phases, \(\varphi_t\in\calS=\{1,2\}\) and generator matrix \(\bs T\).
-Let \(\mathcal I = 1.8\), and partition into two intervals \(\calD_0=[0,1]\) and \(\calD_1=[1,1.8]\), hence \(x_0=0,\,x_1=1,\,x_2=1.8\).
+Consider a process \(\{( \dot Y(t),X(t),\varphi(t))\}_{t\geq 0}\) with two phases, \(\varphi(t)\in\calS=\{1,2\}\) and generator matrix \(\bs T\).
+Let \(b = 1.8\), and the cell edges be \(y_0=0,\,y_1=1,\,y_2=1.8\).
 We choose a basis of Lagrange polynomials of order 1 to define our approximation space.
@@
 \begin{align*}
-	\phi_{0}^1(x) &= 1-x,& \phi_0^2(x) &= x, & &x \in\calD_0,\\
-	\phi_{1}^1(x) &= \cfrac{1.8-x}{0.8},& \phi_1^2(x) &= \cfrac{x-1}{0.8}, & & x \in\calD_1.
+	\phi_{0}^1(x) &= 1-x,& \phi_0^2(x) &= x, & &x \in(0,1),\\
+	\phi_{1}^1(x) &= \cfrac{1.8-x}{0.8},& \phi_1^2(x) &= \cfrac{x-1}{0.8}, & & x \in(1,1.8).
 \end{align*}
@@
 \end{tikzpicture}
-\caption{A mesh with nodes \(x_0=0\), \(x_1=1\) and \(x_2=1.8\) and cells \(\calD_0=[0,1]\), \(\calD_1=[1,1.8]\).
-There are two basis functions on each cell.
-Point masses are located at \(x_0=0\) and \(x_2=1.8\).\label{fig: mesh}}
+\caption{A mesh with nodes \(y_0=0\), \(y_1=1\) and \(y_2=1.8\). There are two basis functions on each cell. Boundaries are located at \(y_0=0\) and \(y_2=1.8\).\label{fig: mesh}}
 \end{figure}
 We can verify that the matrices \(\bs M\) and \(\bs G\) are given by 
-\[\bs M = \left[\begin{array}{cc|cc} 1/3 & 1/6 & 0 & 0 \\ 1/6 & 1/3 & 0 & 0 \\\hline 0 & 0 & 4/15 & 4/30 \\ 0 & 0 & 4/30 & 4/15 \end{array}\right],\, \bs G = \left[\begin{array}{cc|cc} -1/2 & 1/2 & 0 & 0 \\ -1/2 & 1/2 & 0 & 0 \\\hline 0 & 0 & -1/2 & 1/2 \\ 0 & 0 & -1/2 & 1/2 \end{array}\right].\]
-The matrix \(\bs P\) is given by 
-\[\bs P = \left[\begin{array}{cc|cc} 1/2 & 0 & 0 & 0 \\ 0 & 1/2 & 0 & 0 \\\hline 0 & 0 & 2/5 & 0 \\ 0 & 0 & 0 & 2/5 \end{array}\right].\]
+\begin{align*}
+	\bs M &= \left[\begin{array}{c|c} \bs M_1 & \bs 0 \\\hline \bs 0 & \bs M_2 \end{array}\right] = \left[\begin{array}{cc|cc} 1/3 & 1/6 & 0 & 0 \\ 1/6 & 1/3 & 0 & 0 \\\hline 0 & 0 & 4/15 & 4/30 \\ 0 & 0 & 4/30 & 4/15 \end{array}\right],
+	\\ \bs G &= \left[\begin{array}{c|c} \bs G_1 & \bs 0 \\ \hline \bs 0 & \bs G_2 \end{array}\right] = \left[\begin{array}{cc|cc} -1/2 & 1/2 & 0 & 0 \\ -1/2 & 1/2 & 0 & 0 \\\hline 0 & 0 & -1/2 & 1/2 \\ 0 & 0 & -1/2 & 1/2 \end{array}\right] .
+\end{align*}
+% The matrix \(\bs P\) is given by 
+% \[\bs P = \left[\begin{array}{cc|cc} 1/2 & 0 & 0 & 0 \\ 0 & 1/2 & 0 & 0 \\\hline 0 & 0 & 2/5 & 0 \\ 0 & 0 & 0 & 2/5 \end{array}\right].\]
 
-Let \(c_1=1\) and \(c_2=-2\). Then the flux matrices are given by 
-\[\bs F_1 = \left[\begin{array}{cc|cc} 0 & 0 & 0 & 0 \\ 0 & -1 & 1 & 0 \\\hline 0 & 0 & 0 & 0 \\ 0 & 0 & 0 & -1 \end{array}\right],\quad \bs F_2 = \left[\begin{array}{cc|cc} -1 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \\\hline 0 & 1 & -1 & 0 \\ 0 & 0 & 0 & 0 \end{array}\right].\]
+Let \(c_1=1\) and \(c_2=-2\). Then the cells are \(\mathcal D_{0,1} = (0,1),\,\mathcal D_{1,1}=[1,1.8),\, \mathcal D_{0,2}=(0,1],\, \mathcal D_{1,2}=(1,1.8)\) and the flux matrices are given by 
+\begin{align*}
+	\bs F_1 &= \left[\begin{array}{c|c} \bs F_1^{00} & \bs F_1^{01} \\\hline \bs 0 & \bs F_1^{11} \end{array}\right] = \left[\begin{array}{cc|cc} 0 & 0 & 0 & 0 \\ 0 & -1 & 1 & 0 \\\hline 0 & 0 & 0 & 0 \\ 0 & 0 & 0 & -1 \end{array}\right],
+	\\\bs F_2 &= \left[\begin{array}{c|c} \bs F_2^{00} & \bs 0 \\\hline \bs  F_2^{10} & \bs F_2^{11} \end{array}\right] = \left[\begin{array}{cc|cc} -1 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \\\hline 0 & 1 & -1 & 0 \\ 0 & 0 & 0 & 0 \end{array}\right].
+\end{align*}
 
-Suppose that \(r_1(x)>0\) on \(\calD_0=\mathcal F_1^+\) and \(r_1(x)<0\) on \(\calD_1\cup\{1\}=\mathcal F_1^-\), and further, that \(r_2(x)<0\) on \(\{0\}\cup\calD_0=\mathcal F_2^-\) and \(r_2(x)>0\) on \(\calD_1=\mathcal F_2^+\).
+Suppose that \(r_1(x)>0\) on \(\calD_{0,1}=\mathcal F_1^+\) and \(r_1(x)<0\) on \(\calD_{1,1}\cup\{1.8\}=\mathcal F_1^-\), and further, that \(r_2(x)<0\) on \(\{0\}\cup\calD_{0,2}=\mathcal F_2^-\) and \(r_2(x)>0\) on \(\calD_{1,2}=\mathcal F_2^+\).
 Specifically, let
-\[r_1(x) = \begin{cases} 1 & x \in [0,1], \\ -1 & x \in [1,1.8],\end{cases}\quad r_2(x) = \begin{cases} -1 & x = 0, \\ -2 & x \in (0,1], \\ 1 & x \in [1,1.8]. \end{cases}\]
+\[r_1(x) = \begin{cases} 1 & x \in (0,1), \\ -1 & x \in [1,1.8],\end{cases}\quad r_2(x) = \begin{cases} -1 & x = 0, \\ -2 & x \in (0,1], \\ 1 & x \in (1,1.8). \end{cases}\]
 
@@
 Then, constructing \({  \bs B}\) we get 
+\begin{center}
 \[
@@
 \]
+\end{center}
 
@@
 	\,
-	{  \bs B}_{11}^{--} = \left[\begin{array}{cc|c}
+	\\&{  \bs B}_{11}^{--} = \left[\begin{array}{cc|c}
 		T_{11}-\frac{15}{2} & \frac{15}{2} & 0\\
@@
 %	\,
-	\\&{  \bs B}_{12}^{+-} = \left[\begin{array}{c|cc}
+	{  \bs B}_{12}^{+-} = \left[\begin{array}{c|cc}
 		0 & T_{12} & 0 \\
@@
 	%
-	{  \bs B}_{12}^{-+} = \left[\begin{array}{cc}
+	\\&{  \bs B}_{12}^{-+} = \left[\begin{array}{cc}
 		T_{12} & 0 \\
@@
 %	\,
-	\\&{  \bs B}_{21}^{+-} = \left[\begin{array}{cc|c}
+	{  \bs B}_{21}^{+-} = \left[\begin{array}{cc|c}
 		T_{21} & 0 & 0 \\
@@
 	%
-	{  \bs B}_{21}^{-+} = \left[\begin{array}{cc}
+	\\&{  \bs B}_{21}^{-+} = \left[\begin{array}{cc}
 		4T_{21} & -2T_{21} \\ \hline
@@
 	%%%%%%%%%%%%%%%%%%%%
-	\\&{  \bs B}_{22}^{++} = \left[\begin{array}{cc}
+	{  \bs B}_{22}^{++} = \left[\begin{array}{cc}
 		T_{22}-\frac{5}{2} & -\frac{5}{2} \\
@@
 	\,
-	{  \bs B}_{22}^{+-} = \left[\begin{array}{c|cc}
+	\\&{  \bs B}_{22}^{+-} = \left[\begin{array}{c|cc}
 		0 & -4 & 8 \\
@@
 \[ {\bs\xi} = \left[\begin{array}{c|cc|cc|c} {\xi}_{-1,2} &  {\xi}_{0,2}^1 &  {\xi}_{0,2}^2 &  {\xi}_{1,1}^1 &  {\xi}_{1,1}^2 &  {\xi}_{K+1,1} \end{array}\right],\]
-where \( {\xi}_{-1,2}\) is an approximation to \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =0, \varphi_{\theta_n} = 2\right)\) and \( {\xi}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} =1.8, \varphi_{\theta_n} = 1\right)\).
-For \(x\in\calD_0\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 2\right)\), is constructed as \( {\xi}_{0,2}^0\phi_0^1(x) +  {\xi}_{0,2}^0\phi_0^1(x)\).
-For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(\overline X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = 1\right)\), is constructed as \( {\xi}_{1,1}^1\phi_1^1(x) +  {\xi}_{1,1}^2\phi_1^2(x).\)
+where \( {\xi}_{-1,2}\) is an approximation to \( \lim\limits_{n\to\infty}\mathbb{P}\left( X{(\theta_n)} =0, \varphi({\theta_n}) = 2\right)\) and \( {\xi}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{n\to\infty}\mathbb{P}\left( X({\theta_n}) =1.8, \varphi({\theta_n}) = 1\right)\).
+For \(x\in\calD_{0,2}\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left( X({\theta_n}) \in \wrt x, \varphi({\theta_n}) = 2\right)\), is constructed as \( {\xi}_{0,2}^0\phi_0^1(x) +  {\xi}_{0,2}^0\phi_0^1(x)\).
+For \(x\in\calD_{1,1}\) an approximation to the density of \( \lim\limits_{n\to\infty}\mathbb{P}\left(X({\theta_n}) \in \wrt x, \varphi({\theta_n}) = 1\right)\), is constructed as \( {\xi}_{1,1}^1\phi_1^1(x) +  {\xi}_{1,1}^2\phi_1^2(x).\)
 
@@
 \[ {\bs p}^- = \left[\begin{array}{c|cc|cc|c} {p}_{-1,2} &  {p}_{0,2}^1 &  {p}_{0,2}^2 &  {p}_{1,1}^1 &  {p}_{1,1}^2 &  {p}_{K+1,1} \end{array}\right],\]
-where \( {p}_{-1,2}\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_{t} =0, Y_t=0, \varphi_{t} = 2\right)\) and \( {p}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t=0, \varphi_{t} = 1\right)\).
-For \(x\in\calD_0\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 2\right)\), is constructed as \( {p}_{0,2}^1\phi_0^1(x) +  {p}_{0,2}^2\phi_0^2(x)\).
-For \(x\in\calD_1\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t = 0, \varphi_{t} = 1\right)\), is constructed as \( {p}_{1,1}^1\phi_1^1(x) +  {p}_{1,1}^2\phi_1^2(x).\)
+where \( {p}_{-1,2}\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t)=0,X({t}) =0, \varphi_{t} = 2\right)\) and \( {p}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t)=0, X(t) =1.8, \varphi({t}) = 1\right)\).
+For \(x\in\calD_{0,2}\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t)=0, X(t) \in \wrt x, \varphi({t}) = 2\right)\), is constructed as \( {p}_{0,2}^1\phi_0^1(x) +  {p}_{0,2}^2\phi_0^2(x)\).
+For \(x\in\calD_{1,1}\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t)=0, X(t) \in \wrt x, \varphi({t}) = 1\right)\), is constructed as \( {p}_{1,1}^1\phi_1^1(x) +  {p}_{1,1}^2\phi_1^2(x).\)
 
@@
 \[ {\bs \pi}^-(y) = \left[\begin{array}{c|cc|cc|c} {\pi}_{-1,2}(y) &  {\pi}_{0,2}^1(y) &  {\pi}_{0,2}^2(y) &  {\pi}_{1,1}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{K+1,1}(y) \end{array}\right],\]
-where \( {\pi}_{-1,2}(y)\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t\in\wrt y, \varphi_{t} = 2\right)\) and \( {\pi}_{K+1,1}(y)\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =1.8, Y_t\in \wrt y, \varphi_{t} = 1\right)\).
-For \(x\in\calD_0\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\), is constructed as \( {\pi}_{0,2}^1(y)\phi_0^1(x) +  {\pi}_{0,2}^2(y)\phi_0^2(x)\).
-For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\), is constructed as \( {\pi}_{1,1}^1(y)\phi_1^1(x) +  {\pi}_{1,1}^2(y)\phi_1^2(x).\)
+where \( {\pi}_{-1,2}(y)\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(  Y(t)\in\wrt y, X(t) =0, \varphi({t}) = 2\right)\) and \( {\pi}_{K+1,1}(y)\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t)\in \wrt y, X(t) =1.8, \varphi({t}) = 1\right)\).
+For \(x\in\calD_{0,2}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y, X(t) \in \wrt x, \varphi({t}) = 2\right)\), is constructed as \( {\pi}_{0,2}^1(y)\phi_0^1(x) +  {\pi}_{0,2}^2(y)\phi_0^2(x)\).
+For \(x\in\calD_{1,1}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t) \in \wrt y, X(t) \in \wrt x, \varphi({t}) = 1\right)\), is constructed as \( {\pi}_{1,1}^1(y)\phi_1^1(x) +  {\pi}_{1,1}^2(y)\phi_1^2(x).\)
 
@@
 \[ {\bs \pi}^+(y) = \left[\begin{array}{cc|cc}  {\pi}_{0,1}^1(y) &  {\pi}_{0,1}^2(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{1,2}^2(y) \end{array}\right].\]
-For \(x\in\calD_0\) an approximation to the joint density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 1\right)\) is constructed as \( {\pi}_{0,1}^1(y)\phi_0^1(x) +  {\pi}_{0,1}^2(y)\phi_0^2(x)\).
-For \(x\in\calD_1\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = 2\right)\) is constructed as \( {\pi}_{1,2}^1(y)\phi_1^1(x) +  {\pi}_{1,2}^2(y)\phi_1^2(x).\)
+For \(x\in\calD_{0,1}\) an approximation to the joint density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y, X_t \in \wrt x, \varphi({t}) = 1\right)\) is constructed as \( {\pi}_{0,1}^1(y)\phi_0^1(x) +  {\pi}_{0,1}^2(y)\phi_0^2(x)\).
+For \(x\in\calD_{1,2}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y,  X(t) \in \wrt x, \varphi({t}) = 2\right)\) is constructed as \( {\pi}_{1,2}^1(y)\phi_1^1(x) +  {\pi}_{1,2}^2(y)\phi_1^2(x).\)
 
@@
 \begin{align*}
-	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t \in \wrt y, \varphi_{t} = i\right) &\approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {\pi}_{i,k}^r(y)\phi_k^r(x)\wrt x \wrt y ,\quad x\in(0,1.8),\, y>0, \\
-	\lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t \in \wrt x, Y_t=0, \varphi_{t} = i\right) & \approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {p}_{k,i}^r\phi_k^r(x)\wrt x ,\quad x\in(0,1.8),\\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t =0, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx{\pi}_{-1,i}(y)\wrt y,\quad y>0, \\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 0, Y_t =0, \varphi_{t} = i\right) & \approx {p}_{-1,i},\quad \\ 
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t \in \wrt y, \varphi_{t} = i\right) & \approx {\pi}_{K+1,i}(y)\wrt y,\quad y>0, \\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(\overline X_t = 1.8, Y_t = 0, \varphi_{t} = i\right) & \approx {p}_{K+1,i}.\quad
+	\lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y,  X(t) \in \wrt x, \varphi({t}) = i\right) &\approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {\pi}_{i,k}^r(y)\phi_k^r(x)\wrt x \wrt y ,\intertext{ for \(x\in(0,1.8),\, y>0, and\)}
+	\lim\limits_{t\to\infty}\mathbb{P}\left( Y(t)=0, X(t) \in \wrt x, \varphi({t}) = i\right) & \approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {p}_{k,i}^r\phi_k^r(x)\wrt x ,\quad x\in(0,1.8),\\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y,X(t) =0 ,\varphi({t}) = i\right) & \approx{\pi}_{-1,i}(y)\wrt y,\quad y>0, \\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t) =0, X(t) =0,\varphi({t}) = i\right) & \approx {p}_{-1,i},\quad \\ 
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(  Y(t) \in \wrt y, X(t) = 1.8, \varphi({t}) = i\right) & \approx {\pi}_{K+1,i}(y)\wrt y,\quad y>0, \\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t) = 0, X(t) = 1.8, \varphi({t}) = i\right) & \approx {p}_{K+1,i}.\quad
 \end{align*}

### 2022-02-17 appendix_kronecker/appendix.tex
first draft of the whole thing! :D

--- 2022-01-21 appendix4/appendix.tex
@@
 \begin{proof}
-	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence 
+	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
 	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
@@
 	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
-	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A\otimes \bs{I}_m\right) \nonumber
+	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
 	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
@@
 \begin{proof}
-	Computing the integral on the left-hand side and then factorising the result and using the \ref{eqn:mpr} multiple times gives
+	Computing the integral on the left-hand side and then factorising the result and using the~\ref{eqn:mpr} multiple times gives
 	\begin{align}
@@
 	\end{align}
-	{Using the \ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
+	{Using the~\ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
 	\begin{align}
@@
 
-	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore (\ref{eqn: is an integral}) is 
+	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore, (\ref{eqn: is an integral}) is 
 	\begin{align*}
@@
 Using Lemma~\ref{lem: sfkjgn} we can show the following result.
-\begin{lem}
+\begin{lem} Assuming \(\lambda >0\) and \(\lambda\bs I-\bs B_{22}\) is invertible, then 
 	\begin{align}
@@
 \begin{cor}\label{cor: mpr B}
-	For \(m\in\{+,-\}\) the top-left quadrant of size \(m_1\times m_1=|\calS_m|\cdot p \times |\calS_m|\cdot p\) of \(e^{\bs B_{mm}t}\), 
+	For \(m\in\{+,-\}\) the top-left quadrant of size \(m_1\times m_1=|\calS_m|  p \times |\calS_m|  p\) of \(e^{\bs B_{mm}t}\), 
 	\begin{align}
-		&\vligne{\bs I & \bs 0}\int_{t=0}^\infty e^{-\lambda t} \exp\left\{\left[\begin{array}{cc} \bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S & \bs T_{m0}\otimes \bs I \\ \bs T_{0m} \otimes \bs I & \bs T_{00}\otimes \bs I \end{array}\right]t\right\} \wrt t\left[\begin{array}{c} \bs I \\ \bs 0 \end{array}\right], \nonumber
+		&\vligne{\bs I_{|\calS_m|  p\times |\calS_m|  p} & \bs 0_{|\calS_m|  p\times |\calS_0|  p}}\int_{t=0}^\infty e^{-\lambda t} \exp\left\{\left[\begin{array}{cc} \bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S & \bs T_{m0}\otimes \bs I \\ \bs T_{0m} \otimes \bs I & \bs T_{00}\otimes \bs I \end{array}\right]t\right\} \wrt t
+		\\&\quad{}\times \left[\begin{array}{c} \bs I_{|\calS_m|  p\times |\calS_m|  p} \\ \bs 0_{|\calS_0|  p\times |\calS_m|  p} \end{array}\right], \nonumber
 	\end{align}
@@
 	\end{align}
-	For \(m\in\{+,-\}\) the top-right quadrant of size \(m_1\times m_2=|\calS_m|\cdot p\times |\calS_0|\cdot p\) of  \(e^{\bs B_{mm}t}\), 
+	For \(m\in\{+,-\}\) the top-right quadrant of size \(m_1\times m_2=|\calS_m|  p\times |\calS_0|  p\) of  \(e^{\bs B_{mm}t}\), 
 	\begin{align}
-		&\vligne{\bs I & \bs 0}\int_{t=0}^\infty e^{-\lambda t} \exp\left\{\left[\begin{array}{cc} \bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S & \bs T_{m0}\otimes \bs I \\ \bs T_{0m} \otimes \bs I & \bs T_{00}\otimes \bs I \end{array}\right]t\right\} \wrt t \left[\begin{array}{c}\bs 0 \\ \bs I\end{array}\right] ,\nonumber
+		&\vligne{\bs I_{|\calS_m|  p\times |\calS_m|  p} & \bs 0_{|\calS_m|  p\times |\calS_0|  p}}\int_{t=0}^\infty e^{-\lambda t} \exp\left\{\left[\begin{array}{cc} \bs T_{mm}\otimes \bs I + \bs C_m\otimes \bs S & \bs T_{m0}\otimes \bs I \\ \bs T_{0m} \otimes \bs I & \bs T_{00}\otimes \bs I \end{array}\right]t\right\} \wrt t
+		\\&\quad{}\times \left[\begin{array}{c}\bs 0_{|\calS_m|  p\times |\calS_m|  p} \\ \bs I_{|\calS_0|  p\times |\calS_0|  p}\end{array}\right] ,\nonumber
 	\end{align}
@@
 	\begin{align}
-		&\vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t \bs{B}_{m{n}} %= \vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t  \left[\begin{array}{cc} \bs{T}_{mn}\otimes \bs{D} & \bs 0 \\ \bs T_{0n}\otimes \bs D & \bs 0 \end{array}\right]
-	= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x\left(\vligne{\bs I_{pn} & \bs 0_{pn\times p|\calS_0|}}\right), \label{eqn: akgj987adKLDJ}
+		&\vligne{\bs I_{|\calS_m|  p\times |\calS_m|  p} & \bs 0_{|\calS_m|  p\times |\calS_0|  p} }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t \bs{B}_{m{n}} %= \vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t  \left[\begin{array}{cc} \bs{T}_{mn}\otimes \bs{D} & \bs 0 \\ \bs T_{0n}\otimes \bs D & \bs 0 \end{array}\right]
+		\\&= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x\left(\vligne{\bs I_{|\calS_n|p} & \bs 0_{|\calS_n|p\times |\calS_0|p}}\right), \label{eqn: akgj987adKLDJ}
 \end{align}
@@
 \begin{proof}
-	From Lemma~\ref{lem: sfkjgn} the top-left quadrant of size \(m_1\times m_1=|\calS_m|\cdot p \times |\calS_m|\cdot p\) of the integral with respect to \(t\) on the left-hand side of (\ref{eqn: skagh87}) is
+	From Lemma~\ref{lem: sfkjgn} the top-left quadrant of size \(m_1\times m_1=|\calS_m|  p \times |\calS_m|  p\) of the integral with respect to \(t\) on the left-hand side of (\ref{eqn: skagh87}) is 
 	\begin{align}
@@
 
-	Now, from Lemma~\ref{lem: sfkjgn} the top-right quadrant of size \(m_1\times m_2=|\calS_m|\cdot p\times |\calS_0|\cdot p\) of the integral with respect to \(t\) on the left-hand side of (\ref{eqn: skagh873}) is
+	Now, from Lemma~\ref{lem: sfkjgn} the top-right quadrant of size \(m_1\times m_2=|\calS_m|  p\times |\calS_0|  p\) of the integral with respect to \(t\) on the left-hand side of (\ref{eqn: skagh873}) is
 	\begin{align}
@@
 		\displaystyle (\lambda \bs I - \bs T_{00}\otimes \bs I)^{-1} &= \int_{u=0}^\infty e^{-(\lambda \bs I - \bs T_{00}\otimes \bs I)u}\wrt u \\&= \int_{u=0}^\infty e^{-\lambda u} e^{(\bs T_{00}\otimes \bs I)u}\wrt u \\&= \int_{u=0}^\infty e^{-\lambda u} e^{\bs T_{00}u}\otimes \bs I \wrt u,
-	\end{align*} by (\ref{eqn:09ksdjgah}). Using this and the \ref{eqn:mpr} we can write  
+	\end{align*} by (\ref{eqn:09ksdjgah}). Using this and the~\ref{eqn:mpr} we can write  
 	\begin{align}
@@
 \begin{align}
-	\nonumber&\vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t \bs{B}_{m{n}} = \vligne{\bs I & \bs 0 }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t  \left[\begin{array}{cc} \bs{T}_{mn}\otimes \bs{D} & \bs 0 \\ \bs T_{0n}\otimes \bs D & \bs 0 \end{array}\right]
+	\nonumber&\vligne{\bs I_{|\calS_m|p\times |\calS_m|p} & \bs 0_{|\calS_m|p\times |\calS_0|p} }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t \bs{B}_{m{n}} 
+	\\\nonumber &= \vligne{\bs I_{|\calS_m|p\times |\calS_m|p} & \bs 0_{|\calS_m|p\times |\calS_0|p} }\int_{t=0}^\infty e^{-\lambda t} e^{\bs{B}_{mm}t} \wrt t  \left[\begin{array}{cc} \bs{T}_{mn}\otimes \bs{D} & \bs 0 \\ \bs T_{0n}\otimes \bs D & \bs 0 \end{array}\right]
 	%
@@
 	%
-	\\&= \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x (\bs C_m^{-1}(\bs{T}_{mn} + \bs T_{m0}(\lambda \bs I - \bs T_{00})^{-1}\bs T_{0n}) \vligne{\bs I_n & \bs 0_{n\times |\calS_0|}} \otimes \bs D) \nonumber
+	\\&= \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x (\bs C_m^{-1}(\bs{T}_{mn} + \bs T_{m0}(\lambda \bs I - \bs T_{00})^{-1}\bs T_{0n}) \vligne{\bs I_{|\calS_n|\times |\calS_n|} & \bs 0_{|\calS_n|\times |\calS_0|}} \otimes \bs D) \nonumber
 	%
-	\\&= \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x \left( \left( \bs{Q}_{mn} (\lambda) \vligne{\bs I_n & \bs 0_{n\times |\calS_0|}} \right) \otimes \bs D\right)  \nonumber
+	\\&= \int_{x=0}^\infty e^{\bs Q_{mm}(\lambda)x}\otimes  e^{\bs S x}\wrt x \left( \left( \bs{Q}_{mn} (\lambda) \vligne{\bs I_{|\calS_n|} & \bs 0_{|\calS_n|\times |\calS_0|}} \right) \otimes \bs D\right)  \nonumber
 	%
-	\\&= \int_{x=0}^\infty \left(\bs H^{mn}(\lambda,x) \vligne{\bs I_n & \bs 0_{n\times |\calS_0|}}\right) \otimes  e^{\bs S x}\bs D\wrt x,\nonumber
+	\\&= \int_{x=0}^\infty \left(\bs H^{mn}(\lambda,x) \vligne{\bs I_{|\calS_n|} & \bs 0_{|\calS_n|\times |\calS_0|}}\right) \otimes  e^{\bs S x}\bs D\wrt x,\nonumber
 	%
-	\\&= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x,\vligne{\bs I_{pn} & \bs 0_{pn\times p|\calS_0|}}, \label{eqn: akgj987ad}
+	\\&= \int_{x=0}^\infty \bs H^{mn}(\lambda,x)  \otimes  e^{\bs S x}\bs D\wrt x,\vligne{\bs I_{|\calS_m|p\times |\calS_n|p} & \bs 0_{|\calS_n|p\times |\calS_0|p}}, \label{eqn: akgj987ad}
 \end{align}
-for \(m,n\in\{+,-\}\), \(m\neq n\) which is (\ref{eqn: akgj987adKLDJ}), where the last line holds from the \ref{eqn:mpr}. 
+for \(m,n\in\{+,-\}\), \(m\neq n\) which is (\ref{eqn: akgj987adKLDJ}), where the last line holds from the~\ref{eqn:mpr}. 
 \end{proof}

### 2022-02-17 appendix_properties_of_B/appendix.tex
first draft of the whole thing! :D

@@
-\section{Properties of the DG operator \(\bs B\)}\label{sec:properties}
+\chapter{Properties of the DG operator \(\bs B\)\label{sec:properties}}
 In this section we prove that the DG approximation \(\bs B\) conserves probability.
 
-Recall that the coefficients \({\boldsymbol a}_{k,i}(t)\) can be used to construct an approximate solution to a differential equation at time \(t\) as \(u_{k,i}(x,t)={\boldsymbol a}_{k,i}(t)\boldsymbol \phi_k(x)\tr{}.\) For \(i\in\calS,\,k\in\mathcal K,\,r\in\mathcal N_k\), define \(\alpha_{k,i}^r(t) := a_{k,i}^r(t)\displaystyle\int_{x\in\calD_k}\phi^r_k(x)\wrt x,\) and row-vectors \(\boldsymbol \alpha_{k,i}(t)=(\alpha_{k,i}^r(t))_{r\in\mathcal N_k}\).
-Motivated by the fact that we may be interested in approximations of the probabilities \(\mathbb P(X(t)\in\calD_k,\varphi(t)=i)\) rather than the function \(u_{k,i}\) itself, we can pose the problem equivalently in terms of the integrals
-\[\mathbb P\left(X(t)\in\calD_k,\varphi(t)=i\right)\approx{\boldsymbol a}_{k,i}(t)\int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\wrt x={\boldsymbol \alpha}_{k,i}(t)\boldsymbol 1.\] 
+Recall that the coefficients \({\boldsymbol a}_{k,i}(t)\) can be used to construct an approximate solution to a differential equation at time \(t\) as \(u_{k,i}(x,t)={\boldsymbol a}_{k,i}(t)\boldsymbol \phi_k(x)\tr{}.\) For \(i\in\calS,\,k\in\mathcal K,\,r\in\{1,...,p_k\}\), define \(\alpha_{k,i}^r(t) := a_{k,i}^r(t)\displaystyle\int_{x=y_k}^{y_{k+1}}\phi^r_k(x)\wrt x,\) and row-vectors \(\boldsymbol \alpha_{k,i}(t)=(\alpha_{k,i}^r(t))_{r\in\{1,...,p_k\}}\).
+Motivated by the fact that we may be interested in approximations of the probabilities \(\mathbb P(X(t)\in\calD_{k,i},\varphi(t)=i)\) rather than the function \(u_{k,i}\) itself, we can pose the problem equivalently in terms of the integrals
+\[\mathbb P\left(X(t)\in\calD_{k,i},\varphi(t)=i\right)\approx{\boldsymbol a}_{k,i}(t)\int_{x\in\calD_{k,i}}\boldsymbol \phi_k(x)\tr{}\wrt x={\boldsymbol \alpha}_{k,i}(t)\boldsymbol 1.\] 
 Define
@@
 \begin{align*}
-\bs P_k &= \diag\left(\int_{x\in\calD_k}\phi^r_k(x)\wrt x\right)_{ r\in\mathcal N_k},\, k\in\mathcal K^\circ,
+\bs P_k &= \diag\left(\int_{x=y_k}^{y_{k+1}}\phi^r_k(x)\wrt x\right)_{ r\in\{1,...,p_k\}},\, k\in\mathcal K^\circ,
 \\\bs P &=\left[\begin{array}{ccc}
-		\bs I_{N_{|\calS|}}\otimes \bs P_0 & &\\
+		\bs I_{N}\otimes \bs P_0 & &\\
 		& \ddots &\\
-		& & \bs I_{N_{|\calS|}}\otimes \bs P_K
+		& & \bs I_{N}\otimes \bs P_K
 	\end{array}\right].
 \end{align*}
-By choosing the basis \(\{\phi_{k}^r\}_{r\in\mathcal N_k,k\in\mathcal K^\circ}\) such that \(\int_{x\in\calD_k}\phi^r_k(x)\wrt x\neq 0\) for all \(r,k\), then \(\bs P\) is invertible.
+By choosing the basis \(\{\phi_{k}^r\}_{r\in\{1,...,p_k\},k\in\mathcal K^\circ}\) such that \(\int_{x=y_k}^{y_{k+1}}\phi^r_k(x)\wrt x\neq 0\) for all \(r,k\), then \(\bs P\) is invertible.
 This is the case for the Lagrange polynomials, but not, for example, for the Legendre polynomials.
@@
 
-The differential equation (\ref{eqn: DG ODE w BCs}) can be equivalently expressed as 
+The differential equation~(\ref{eqn: DG ODE w BCs}) can be equivalently expressed as 
 \(\label{eqn: DG alpha ODE w BCs}
@@
  %
- \\ {  {\mathfrak{\bs B}}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1}\mathbb 1_{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}}\otimes \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{},
+ \\ {  {\mathfrak{\bs B}}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1} 1{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}}\otimes \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{},
  %
@@
  %
- \\ {  {\mathfrak{\bs B}}}^{K {K+1}} &:= \left[c_ip_{ij}^{K+1}\mathbb 1_{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes  \bs P_K^{-1} \boldsymbol\phi_K(b)\tr{},
+ \\ {  {\mathfrak{\bs B}}}^{K {K+1}} &:= \left[c_ip_{ij}^{K+1} 1{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes  \bs P_K^{-1} \boldsymbol\phi_K(b)\tr{},
  %
  \\\mathfrak{\bs B}^{kk}_{ij} &:= \begin{cases}
-    	T_{ii}\bs I_{N_k} + c_i\bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k & i = j, \\
-	T_{ij}\bs I_{N_k} & i \neq j,
+    	T_{ii}\bs I_{p_k} + c_i\bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k & i = j, \\
+	T_{ij}\bs I_{p_k} & i \neq j,
     \end{cases}\quad \mbox{ for \(k=1,2,\dots K-1\),}
@@
 	c_i\bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k+1}^{-1}\bs P_{k+1} & i = j, \\
-	\bs 0_{N_k} & i \neq j,
+	\bs 0_{p_k} & i \neq j,
     \end{cases}\quad \mbox{ for \(k=0,1,\dots K-1\),}
@@
     	c_i\bs P_k^{-1}\bs F_i^{k,k-1}\bs M_{k-1}^{-1}\bs P_{k-1} & i = j, \\
-	\bs 0_{N_k} & i \neq j,
+	\bs 0_{p_k} & i \neq j,
     \end{cases},\quad \mbox{ for \(k=2,\dots K\),}
 \\
-     {  {\mathfrak{\bs B}}}^{kk}%&=\left[\begin{array}{ccc}T_{11}I_{N_k} + c_1P_k^{-1}(F_1^{kk}+G_k)M_k^{-1}P_k & T_{12}I_{N_k} & T_{1N_\calS}I_{N_k}  \\ T_{21}I_{N_k} & & \\ \vdots &\ddots & \vdots \\ & &   T_{N_\calS-1,N_\calS}I_{N_k} \\  T_{N_\calS1}I_{N_k} &  T_{N_\calS,N_\calS-1}I_{N_k} & T_{N_\calS,N_\calS}I_{N_k} +c_{N_\calS}P_k^{-1}(F_{N_\calS}^{kk}+G_k)M_k^{-1}P_k\end{array}\right]
+     {  {\mathfrak{\bs B}}}^{kk}%&=\left[\begin{array}{ccc}T_{11}I_{p_k} + c_1P_k^{-1}(F_1^{kk}+G_k)M_k^{-1}P_k & T_{12}I_{p_k} & T_{1N}I_{p_k}  \\ T_{21}I_{p_k} & & \\ \vdots &\ddots & \vdots \\ & &   T_{N-1,N}I_{p_k} \\  T_{N1}I_{p_k} &  T_{N,N-1}I_{p_k} & T_{N,N}I_{p_k} +c_{N}P_k^{-1}(F_{N}^{kk}+G_k)M_k^{-1}P_k\end{array}\right]
 %    %
-%    &=:\left[\begin{array}{ccc} {  {\mathfrak B}}_{11}^{kk} &  {  {\mathfrak B}}^{kk}_{12} &  {  {\mathfrak B}}^{kk}_{1N_\calS}  \\  {  {\mathfrak B}}^{kk}_{21} & & \\ \vdots &\ddots & \vdots \\ & &    {  {\mathfrak B}}^{kk}_{N_\calS-1,N_\calS} \\   {  {\mathfrak B}}^{kk}_{N_\calS1} &   {  {\mathfrak B}}^{kk}_{N_\calS,N_\calS-1} &  {  {\mathfrak B}}^{kk}_{N_\calS,N_\calS} \end{array}\right],
+%    &=:\left[\begin{array}{ccc} {  {\mathfrak B}}_{11}^{kk} &  {  {\mathfrak B}}^{kk}_{12} &  {  {\mathfrak B}}^{kk}_{1N}  \\  {  {\mathfrak B}}^{kk}_{21} & & \\ \vdots &\ddots & \vdots \\ & &    {  {\mathfrak B}}^{kk}_{N-1,N} \\   {  {\mathfrak B}}^{kk}_{N1} &   {  {\mathfrak B}}^{kk}_{N,N-1} &  {  {\mathfrak B}}^{kk}_{N,N} \end{array}\right],
     %
-    &=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}_{11}^{kk} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{1N_\calS}  \\ \vdots &\ddots & \vdots \\   {  {\mathfrak{\bs B}}}^{kk}_{N_\calS1} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{N_\calS,N_\calS} \end{array}\right], \mbox{ for \(k\in\mathcal K^\circ\),}
+    &=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}_{11}^{kk} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{1N}  \\ \vdots &\ddots & \vdots \\   {  {\mathfrak{\bs B}}}^{kk}_{N1} &  \hdots &  {  {\mathfrak{\bs B}}}^{kk}_{N,N} \end{array}\right], \mbox{ for \(k\in\mathcal K^\circ\),}
     %
-\\ {  {\mathfrak{\bs B}}}^{k,k+1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k+1}M_{k+1}^{-1}P_{k+1}&  & \\  &\ddots &  \\ & &   \\  &  &c_{N_\calS}P_k^{-1}F_{N_\calS}^{k,k+1}M_{k+1}^{-1}P_{k+1} \end{array}\right]
+\\ {  {\mathfrak{\bs B}}}^{k,k+1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k+1}M_{k+1}^{-1}P_{k+1}&  & \\  &\ddots &  \\ & &   \\  &  &c_{N}P_k^{-1}F_{N}^{k,k+1}M_{k+1}^{-1}P_{k+1} \end{array}\right]
 %
-&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k+1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k+1}_{1,N_\calS}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k+1}_{N_\calS,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k+1}_{N_\calS,N_\calS} \end{array}\right],\mbox{ for \(k=0,1,\dots K-1\),}
+&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k+1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k+1}_{1,N}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k+1}_{N,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k+1}_{N,N} \end{array}\right],\mbox{ for \(k=0,1,\dots K-1\),}
 %
-\\ {  {\mathfrak{\bs B}}}^{k,k-1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k-1}M_{k-1}^{-1}P_{k-1}&  & \\  &\ddots &  \\  &  &c_{N_\calS}P_k^{-1}F_{N_\calS}^{k,k-1}M_{k-1}^{-1}P_{k-1} \end{array}\right]
+\\ {  {\mathfrak{\bs B}}}^{k,k-1}%&=\left[\begin{array}{ccc}c_1P_k^{-1}F_1^{k,k-1}M_{k-1}^{-1}P_{k-1}&  & \\  &\ddots &  \\  &  &c_{N}P_k^{-1}F_{N}^{k,k-1}M_{k-1}^{-1}P_{k-1} \end{array}\right]
 %
-&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k-1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k-1}_{1,N_\calS}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k-1}_{N_\calS,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k-1}_{N_\calS,N_\calS} \end{array}\right], \mbox{ for \(k=1,2,\dots K\).}
+&=:\left[\begin{array}{ccc} {  {\mathfrak{\bs B}}}^{k,k-1}_{1,1}& \hdots &  {  {\mathfrak{\bs B}}}^{k,k-1}_{1,N}\\  \vdots &\ddots & \vdots \\  {  {\mathfrak{\bs B}}}^{k,k-1}_{N,1} & \hdots & {  {\mathfrak{\bs B}}}^{k,k-1}_{N,N} \end{array}\right], \mbox{ for \(k=1,2,\dots K\).}
 \end{align*} 
@@
 	One may recognise the structure of \( {  {\mathfrak{\bs B}}}\) as the structure of a quasi-birth-and-death process (QBD), with levels \(k\in\mathcal K^\circ\).
-This raises the question of whether \( {  {\mathfrak{\bs B}}}\) is indeed a representation of the generator matrix of a QBD, or QBD-like process.
-In the case of a constant basis function on each cell, i.e.~\(N_k=1\) and \(\phi_k^1(x)\propto1\), \(k\in\mathcal K^\circ\), then \( {  {\mathfrak{\bs B}}}\) is the generator of a QBD: it has zero row-sums, negative diagonal entires, and non-negative off-diagonal entries, the QBD-phase variable is \(\{\varphi_t\}\) and the level is \(k\in\mathcal K\).
+This raises whether \( {  {\mathfrak{\bs B}}}\) is indeed a representation of the generator matrix of a QBD, or QBD-like process.
+In the case of a constant basis function on each cell, i.e.~\(p_k=1\) and \(\phi_k^1(x)\propto1\), \(k\in\mathcal K^\circ\), then \( {  {\mathfrak{\bs B}}}\) is the generator of a QBD: it has zero row-sums, negative diagonal entries, and non-negative off-diagonal entries, the QBD-phase variable is \(\{\varphi(t)\}\) and the level is \(k\in\mathcal K\).
 In fact, if \(\Delta_k\) is the same for every \(k\in\mathcal K^\circ\), then this is the same QBD discretisation of a stochastic fluid process analysed by \cite{bo2013}.
 However, for higher-degree polynomials \(\mathfrak{\bs B}\) is not necessarily the generator of a QBD process.
-We conjecture that, using polynomial basis functions, then \(N_k=1\) and \(\phi_k^1(x)\propto1\), \(k\in\mathcal K^\circ\) is the only DG approximation which has an interpretation as a QBD-like process -- not even as a QBD-RAP \citep{bn2010}.
+We conjecture that, using polynomial basis functions, then \(p_k=1\) and \(\phi_k^1(x)\propto1\), \(k\in\mathcal K^\circ\) is the only DG approximation which has an interpretation as a QBD-like process -- not even as a QBD-RAP \citep{bn2010}.
 	\end{rem}
 
-	In the following lemma, we use the following properties of the Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature nodes. 
-	\paragraph{Property 1} \(\displaystyle\sum_{s=1}^{N_k} \phi_k^s(x) = \begin{cases} 1&x\in\calD_k,\\ 0 & x\notin \calD_k.\end{cases}\)
+	For simplicity, define \(\calD_k=[y_k,y_{k+1}]\). In the following lemma, we use the following properties of the Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature nodes. 
+	\paragraph{Property 1} \(\displaystyle\sum_{s=1}^{p_k} \phi_k^s(x) = \begin{cases} 1&x\in\calD_k,\\ 0 & x\notin \calD_k.\end{cases}\)
 
-	For \(k\in\mathcal K^\circ\), let \(\boldsymbol e_n^k\) be a row-vector of length \(N_k\) with a 1 in the \(n\)th position and zeros elsewhere.
-	\paragraph{Property 2} At the cell edges, \(\bs \phi_k(y_k) = \bs e_1^k\) and \(\bs\phi_k(y_{k+1})=\bs e_{N_k}^k\), \(k\in\mathcal K^\circ\).
+	For \(k\in\mathcal K^\circ\), let \(\boldsymbol e_n^k\) be a row-vector of length \(p_k\) with a 1 in the \(n\)th position and zeros elsewhere.
+	\paragraph{Property 2} At the cell edges, \(\bs \phi_k(y_k) = \bs e_1^k\) and \(\bs\phi_k(y_{k+1})=\bs e_{p_k}^k\), \(k\in\mathcal K^\circ\).
 
 \begin{lem}
-	If \(\{\phi^r_k(x)\}_{r\in\mathcal N_k}\), are chosen as the Lagrange interpolating polynomials on \(\mathcal D_k\), \(k\in\mathcal K^\circ\), then the matrix \( {  {\mathfrak{\bs B}}}\) has zero row-sums.
+	If \(\{\phi^r_k(x)\}_{r\in\{1,...,p_k\}}\), are chosen as the Lagrange interpolating polynomials on \(\mathcal D_k\), \(k\in\mathcal K^\circ\), then the matrix \( {  {\mathfrak{\bs B}}}\) has zero row-sums.
 \end{lem}
@@
 	\begin{align*}
-        \bs M_k\boldsymbol 1 &= \left(\sum_{s=1}^{N_k} \int_{x\in\calD_k}\phi^r_k(x)\phi_s^k(x)\wrt x\right)_{r\in\mathcal N_k}\tr{}
+        \bs M_k\boldsymbol 1 &= \left(\sum_{s=1}^{p_k} \int_{x\in\calD_k}\phi^r_k(x)\phi_s^k(x)\wrt x\right)_{r\in\{1,...,p_k\}}\tr{}
 	%
-        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\sum_{s=1}^{N_k} \phi_s^k(x)\wrt x \right)_{r\in\mathcal N_k} \tr{}
+        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\sum_{s=1}^{p_k} \phi_s^k(x)\wrt x \right)_{r\in\{1,...,p_k\}} \tr{}
         %
-        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\wrt x\right)_{r\in\mathcal N_k} \tr{}
+        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\wrt x\right)_{r\in\{1,...,p_k\}} \tr{}
         %
@@
 	\begin{align*}
-        \bs G_k\bs M_k^{-1}\bs P_k\boldsymbol 1 = \bs G_k\boldsymbol 1 &= \left(\sum_{s=1}^{N_k} \int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}\phi_k^s(x)\wrt x\right)_{r\in\mathcal N_k} \tr{}
+        \bs G_k\bs M_k^{-1}\bs P_k\boldsymbol 1 = \bs G_k\boldsymbol 1 &= \left(\sum_{s=1}^{p_k} \int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}\phi_k^s(x)\wrt x\right)_{r\in\{1,...,p_k\}} \tr{}
 	%
-        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}\sum_{s=1}^{N_k} \phi_k^s(x)\wrt x \right)_{r\in\mathcal N_k} \tr{}
+        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}\sum_{s=1}^{p_k} \phi_k^s(x)\wrt x \right)_{r\in\{1,...,p_k\}} \tr{}
 	%
-        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}1\wrt x\right)_{r\in\mathcal N_k} \tr{}
+        \\&= \left(\int_{x\in\calD_k}\phi^r_k(x)\cfrac{\wrt }{\wrt x}1\wrt x\right)_{r\in\{1,...,p_k\}} \tr{}
 	%
@@
 	where we have again used Property 1.
-Consider \(c_i>0\) and let \(\boldsymbol b\) and \(\boldsymbol d\) be arbitrary row-vectors of length \(N_k\) and \(N_{k+1}\), respectively.
+Consider \(c_i>0\) and let \(\boldsymbol b\) and \(\boldsymbol d\) be arbitrary row-vectors of length \(p_k\) and \(p_{k+1}\), respectively.
 By Property 2, for \(k\in\mathcal K^\circ,\)
 	\begin{align*}
-        \bs F_i^{kk}\boldsymbol b &= -\boldsymbol \phi_k(y_{k+1})\tr{} \boldsymbol \phi_k(y_{k+1})\boldsymbol b \\&= -(\bs e^k_{N_k})\tr{} \bs e_{N_k}^k\bs b \\&= -b_{N_k}(\boldsymbol e_{N_k}^k)\tr{},
+        \bs F_i^{kk}\boldsymbol b &= -\boldsymbol \phi_k(y_{k+1})\tr{} \boldsymbol \phi_k(y_{k+1})\boldsymbol b \\&= -(\bs e^k_{p_k})\tr{} \bs e_{p_k}^k\bs b \\&= -b_{p_k}(\boldsymbol e_{p_k}^k)\tr{},
         %
-        \\\bs F_i^{k,k+1}\boldsymbol d &= \boldsymbol \phi_{k}(y_{k+1})\tr{} \boldsymbol \phi_{k+1}(y_{k+1})\boldsymbol d \\&= (\bs e^{k}_{N_k})\tr{} \bs e_{1}^{k+1}\bs d \\&= d_{1}(\boldsymbol e_{N_{k}}^{k})\tr{}.
+        \\\bs F_i^{k,k+1}\boldsymbol d &= \boldsymbol \phi_{k}(y_{k+1})\tr{} \boldsymbol \phi_{k+1}(y_{k+1})\boldsymbol d \\&= (\bs e^{k}_{p_k})\tr{} \bs e_{1}^{k+1}\bs d \\&= d_{1}(\boldsymbol e_{N_{k}}^{k})\tr{}.
 	\end{align*}
@@
 	\begin{align*}
-		&\bs P_k^{-1}\bs F_i^{kk} \bs M_k^{-1} \bs P_k \bs 1 = \bs P_k^{-1} \bs \phi_k(y_k)'\bs 1 = -\bs P_k^{-1}\bs F_i^{k,k+1} \bs M_k^{-1} \bs P_k \bs 1 = - \bs P_k^{-1}(\bs e_{N_k}^k)',
+		&\bs P_k^{-1}\bs F_i^{kk} \bs M_k^{-1} \bs P_k \bs 1 = \bs P_k^{-1} \bs \phi_k(y_k)'\bs 1 = -\bs P_k^{-1}\bs F_i^{k,k+1} \bs M_k^{-1} \bs P_k \bs 1 = - \bs P_k^{-1}(\bs e_{p_k}^k)',
 	\end{align*}
@@
 		\sum_{j\in\calS}\mathfrak{\bs B}^{kk}_{ij}\bs 1 + \mathfrak{\bs B}^{kk+1}_{ij}\bs 1
-		&=\sum_{j\in\calS}T_{ij}\bs I_{N_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_k^{-1}\bs P_k\boldsymbol 1 
+		&=\sum_{j\in\calS}T_{ij}\bs I_{p_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_k^{-1}\bs P_k\boldsymbol 1 
 		\\&=\bs 0.
@@
 		\\&\quad{} + c_i \bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K\boldsymbol 1 
-		\\&= -\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 -\sum_{j\in\calS_{{K+1}}}c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\bs 1+ c_i \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\boldsymbol 1
+		\\&= -\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 
+		\\&{}\quad-\sum_{j\in\calS_{{K+1}}}c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\bs 1+ c_i \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\boldsymbol 1
 		\\&\quad{}+ c_i \bs P_K^{-1}\bs G_K\bs M_K^{-1}\bs P_K\boldsymbol 1
@@
 	\end{align*}
-	% Similarly, for \(c_i<0\), and row-vectors \(\boldsymbol b\) and \(\boldsymbol d\) of length \(N_k\) and \(N_{k-1}\), respectively, 
+	% Similarly, for \(c_i<0\), and row-vectors \(\boldsymbol b\) and \(\boldsymbol d\) of length \(p_k\) and \(N_{k-1}\), respectively, 
 	% \begin{align*}
@@
 		\sum_{j\in\calS}\mathfrak{\bs B}^{kk}_{ij}\bs 1 + \mathfrak{\bs B}^{kk-1}_{ij}\bs 1
-		&=\sum_{j\in\calS}T_{ij}\bs I_{N_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\bs M_k^{-1}\bs P_k\boldsymbol 1
+		&=\sum_{j\in\calS}T_{ij}\bs I_{p_k}\bs 1 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k-1}\bs M_k^{-1}\bs P_k\boldsymbol 1
 		% \\&= \boldsymbol 0 + c_i \bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k\boldsymbol 1 + c_i \bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k-1}^{-1}\bs P_{k-1}\boldsymbol 1
@@
 		&= \bs T_{{-1},{-1}}\boldsymbol 1 + \left[\bs T_{{-1}+}\otimes\left( \boldsymbol\phi_0(0)\bs M_0^{-1}\bs P_0\right)\right]\boldsymbol 1. 
-		\intertext{Swapping the order of summation and recalling \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\) then this is equal to}
+		\intertext{ Swapping the order of summation and recalling \(\bs M_k^{-1}\bs P_k\boldsymbol 1=\boldsymbol 1\) then this is equal to}
 		\bs T_{{-1},{-1}}\boldsymbol 1 + \left[\bs T_{{-1}+}\otimes\left( \boldsymbol\phi_0(0)\bs M_0^{-1}\bs P_0\right)\boldsymbol 1\right]\boldsymbol 1 
@@
 	\begin{align*}
-	&\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS} \int_{x\in[0,b]}u_i(x,t)\wrt x 
+	&\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS} \int_{x\in(0,b)}u_i(x,t)\wrt x 
 	%
-	\\&= \sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i\in\calS} \int_{x\in[0,b]}u_i(x,0)\wrt x.
+	\\&= \sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i\in\calS} \int_{x\in(0,b)}u_i(x,0)\wrt x.
 	\end{align*}
@@
 %\end{align*}
-Let \(\{\psi^r_k(x)\}_{r\in\mathcal N_k,k\in\mathcal K^\circ},\) be a basis for \(span(\phi^r_k(x),r\in\mathcal N_k,k\in\mathcal K^\circ)\), where \(\{\phi^r_k(x)\}_{r\in\mathcal N_k,k\in\mathcal K^\circ}\) are the Lagrange polynomials.
+Let \(\{\psi^r_k(x)\}_{r\in\{1,...,p_k\},k\in\mathcal K^\circ},\) be a basis for \(span(\phi^r_k(x),r\in\{1,...,p_k\},k\in\mathcal K^\circ)\), where \(\{\phi^r_k(x)\}_{r\in\{1,...,p_k\},k\in\mathcal K^\circ}\) are the Lagrange polynomials.
 Also define \(\psi_1^{-1}(x)=\delta(x)\) and \(\psi_1^{K+1}(x)=\delta(x-b)\) to capture the point masses at the boundaries.
 Let us use the same vector notation for the basis \(\psi^r_k(x)\) as we do for \(\phi^r_k(x)\).
-For \(k\in\mathcal K^\circ\), since \(\{\psi^r_k(x)\}_{r\in\mathcal N_k}\) and \(\{\phi^r_k(x)\}_{r\in\mathcal N_k}\) have the same span, then there is a matrix \(\bs V^k\) such that  \(\bs \psi_k(x)\tr{} = \bs V^k\bs \phi_k(x)\tr{}\).
+For \(k\in\mathcal K^\circ\), since \(\{\psi^r_k(x)\}_{r\in\{1,...,p_k\}}\) and \(\{\phi^r_k(x)\}_{r\in\{1,...,p_k\}}\) have the same span, then there is a matrix \(\bs V^k\) such that  \(\bs \psi_k(x)\tr{} = \bs V^k\bs \phi_k(x)\tr{}\).
 Trivially, this also holds for \(k={-1},{K+1}\).
@@
 Let 
-\[ \bs W = \left[\begin{array}{ccc}
+\begin{align*}
+	\bs W &= \left[\begin{array}{ccc}
 		\bs I_{|\calS_{-1}|} & & \\
@@
 	\end{array}\right]
-\mbox{ and }
-  \bs V = \left[\begin{array}{ccccc}
+\end{align*}{ and }
+\begin{align*}
+  \bs V &= \left[\begin{array}{ccccc}
 		\bs I_{|\calS_{-1}|} & & & &  \\
@@
 		& & & & \bs I_{|\calS_{K+1}|}
-	\end{array}\right]. \]
-For a DG approximation, \(\bs B\), constructed from the basis \(\{\psi^r_k\}_{r\in\mathcal N_k,k\in\mathcal K}\), it can be shown that \(\bs B\) is similar to \( {  {\mathfrak{\bs B}}}\) with similarity matrix, \(\bs V\bs W\), such that \(  \bs B_{ij} =  \bs V\bs W{ \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V^{-1},\,i,j\in\calS\).
-Therefore 
+	\end{array}\right]. 
+\end{align*}
+For a DG approximation, \(\bs B\), constructed from the basis \(\{\psi^r_k\}_{r\in\{1,...,p_k\},k\in\mathcal K}\), it can be shown that \(\bs B\) is similar to \( {  {\mathfrak{\bs B}}}\) with similarity matrix, \(\bs V\bs W\), such that
+\[  \bs B_{ij} =  \bs V\bs W{ \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V^{-1},\,i,j\in\calS.\]
+Therefore, 
 \begin{align*} 
-	\int_{x\in[0,b]}  \bs B_{ij}\bs\psi(x)\tr{} \wrt x &= \bs V\bs W { \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V^{-1}\int_{x\in[0,b]} \bs V\bs \phi(x)\tr{}\wrt x 
+	\int_{x\in(0,b)}  \bs B_{ij}\bs\psi(x)\tr{} \wrt x &= \bs V\bs W { \mathfrak{\bs B}}_{ij}\bs W^{-1}\bs V^{-1}\int_{x\in(0,b)} \bs V\bs \phi(x)\tr{}\wrt x 
 	%
@@
 \end{align*}
-since \(\displaystyle\int_{x\in[0,b]} \bs \phi(x)\tr{}\wrt x =\bs W\bs 1\).
+since \(\displaystyle\int_{x\in(0,b)} \bs \phi(x)\tr{}\wrt x =\bs W\bs 1\).
 The row sums of \(\mathfrak{\bs B}\) are 0, hence 
 \begin{align}\label{eqn:Bsums0} 
-	\int_{x\in[0,b]}  \sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\tr{} \wrt x  &=  \bs V\bs W \sum_{j\in\calS}{ \mathfrak{\bs B}}_{ij}\bs 1 \\&=  \bs V\bs W \bs 0 \\&= \bs 0.
+	\int_{x\in(0,b)}  \sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\tr{} \wrt x  &=  \bs V\bs W \sum_{j\in\calS}{ \mathfrak{\bs B}}_{ij}\bs 1 \\&=  \bs V\bs W \bs 0 \\&= \bs 0.
 \end{align}
 
-Let \( {}^\psi \bs a_i(t)\), \(i\in\calS\) denote the coefficients related to the DG approximation constructed with the basis \(\{\psi^r_k\}_{r\in\mathcal N_k,k\in\mathcal K}\) (to distinguish them from \(\bs a\) and \(\bs \alpha\) used above).
+Let \( {}^\psi \bs a_i(t)\), \(i\in\calS\) denote the coefficients related to the DG approximation constructed with the basis \(\{\psi^r_k\}_{r\in\{1,...,p_k\},k\in\mathcal K}\) (to distinguish them from \(\bs a\) and \(\bs \alpha\) used above).
 The DE constructed by the DG method is
 \begin{align*}
-	\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x) = \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x).
+	\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x)\tr{} = \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x)\tr{}.
 \end{align*}
-Integrating over \(x\in [0,b]\) and summing over \(j\in\calS\) we get
+Integrating over \(x\in (0,b)\) and summing over \(j\in\calS\) we get
 \begin{align*}
-	\int_{x\in[0,b]}\sum_{j\in\calS}\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x)\wrt x = \int_{x\in[0,b]}\sum_{j\in\calS}\sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x).
+	\int_{x\in(0,b)}\sum_{j\in\calS}\cfrac{\wrt}{\wrt t}\left({}^\psi \boldsymbol a_j (t)\right)\boldsymbol \psi(x)\tr{}\wrt x = \int_{x\in(0,b)}\sum_{j\in\calS}\sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\bs B_{ij}\bs\psi(x)\tr{}.
 \end{align*}
@@
 \begin{align}\label{eqn:totalprobDE}
-	\cfrac{\wrt}{\wrt t}\sum_{j\in\calS} \left({}^\psi \boldsymbol a_j (t)\right)\int_{x\in[0,b]} \boldsymbol \psi(x)\wrt x =  \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\int_{x\in[0,b]}\sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\wrt x = 0,
+	\cfrac{\wrt}{\wrt t}\sum_{j\in\calS} \left({}^\psi \boldsymbol a_j (t)\right)\int_{x\in(0,b)} \boldsymbol \psi(x)\tr{}\wrt x =  \sum_{i\in\calS}\left({}^\psi \boldsymbol a_i (t)\right)\int_{x\in(0,b)}\sum_{j\in\calS}\bs B_{ij}\bs\psi(x)\tr{}\wrt x = 0,
 \end{align}
-where the right-hand side is \(0\) due to Equation (\ref{eqn:Bsums0}).
+where the right-hand side is \(0\) due to Equation~(\ref{eqn:Bsums0}).
 This holds for all \(t\geq 0\).
@@
 %\begin{align*}
-%	&\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS}\int_{x\in[0,b]} u_i(x,t)\wrt x 
+%	&\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS}\int_{x\in(0,b)} u_i(x,t)\wrt x 
 %	%
-%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS}\int_{x\in[0,b]}\boldsymbol a_i(t) \boldsymbol \psi(x)\wrt x 
+%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(t)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)+\sum_{i\in\calS}\int_{x\in(0,b)}\boldsymbol a_i(t) \boldsymbol \psi(x)\wrt x 
 %	%
-%	\\&=\sum_{j\in\calS}\int_{x\in[0,b]}\left[\vligne{\bs q_{{-1}}(0) & \boldsymbol a(0) & \bs q_{{K+1}}(0)} \exp(  Bt)\right]_{j}\boldsymbol \psi(x)\wrt x 
+%	\\&=\sum_{j\in\calS}\int_{x\in(0,b)}\left[\vligne{\bs q_{{-1}}(0) & \boldsymbol a(0) & \bs q_{{K+1}}(0)} \exp(  Bt)\right]_{j}\boldsymbol \psi(x)\wrt x 
 %	%
-%	\\&=\sum_{j\in\calS}\int_{x\in[0,b]}\left[\vligne{\bs q_{{-1}}(0) & \boldsymbol a(0) & \bs q_{{K+1}}(0)} \left(I+  Bt + \cfrac{  B^2t^2}{2!}+...\right)\right]_{j}\boldsymbol \psi(x)\wrt x
+%	\\&=\sum_{j\in\calS}\int_{x\in(0,b)}\left[\vligne{\bs q_{{-1}}(0) & \boldsymbol a(0) & \bs q_{{K+1}}(0)} \left(I+  Bt + \cfrac{  B^2t^2}{2!}+...\right)\right]_{j}\boldsymbol \psi(x)\wrt x
 %	%
-%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i,j\in\calS}\int_{x\in[0,b]}\boldsymbol a_i(0)\boldsymbol \psi(x)\wrt x
+%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i,j\in\calS}\int_{x\in(0,b)}\boldsymbol a_i(0)\boldsymbol \psi(x)\wrt x
 %	%
-%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i\in\calS}\int_{x\in[0,b]} u_i(x,0)\wrt x.
+%	\\&=\sum_{i\in\calS_{-1}}q_{{-1},i}(0)+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(0)+\sum_{i\in\calS}\int_{x\in(0,b)} u_i(x,0)\wrt x.
 %\end{align*}
@@
 %\begin{remark}
-%	For any choice of basis \(\{\phi^r_k(x)\}_{r\in\{1,...,N_k\},k\in\{1,...,K\}}\) which spans the set of polynomials of degree \(N_k-1\), the DG approximation constructed with this basis will conserve probability.
+%	For any choice of basis \(\{\phi^r_k(x)\}_{r\in\{1,...,p_k\},k\in\{1,...,K\}}\) which spans the set of polynomials of degree \(p_k-1\), the DG approximation constructed with this basis will conserve probability.
 %\end{remark}
@@
 % Given we have now truncated the space and added boundaries, let us define \(\mathcal M_{0,b}\) as the set of measures, \(\mu_i\), which admit an absolutely continuous density on \((0,b)\), may have a point mass at \(x=0\) if \(i\in\mathcal S_{-1}\), and another at \(x=b\) if \(i\in\calS_{{K+1}}\).
-The set \(\mathcal M_{0,b}\) is the domain of the operator \(\mathbb B\) truncated to the interval \([0,b]\) with regulated boundaries at \(x=0\) and \(x=b\).
+The set \(\mathcal M_{0,b}\) is the domain of the operator \(\mathbb B\) truncated to the interval \((0,b)\) with regulated boundaries at \(x=0\) and \(x=b\).
 Also, redefine \(\mathcal K^m_i = \{k\in\{{-1},1,...,K,{K+1}\}\mid   l  (\calD_k\cap\calF_i^m) =    l (\calD_k) \}\) for \(i\in\calS,m\in\{+,-,0\}\) for all \(   l\in\mathcal M_{0,b} \).

### 2022-02-17 chapter1/chapter.tex
first draft of the whole thing! :D

@@
 %!TEX root = ../thesis.tex
-\chapter{The Existing Literature \& Mathematical Preliminaries \label{ch:math}} 
+\chapter{The existing literature \& mathematical preliminaries\label{ch:math}} 
 % intro to intro
@@
 
-A stochastic process is a sequence, \(\{X(t)\}_{t\in\mathcal T}\), of random variables or random vectors indexed by some index set \(\mathcal T\).
-In this thesis in the case that the index set is \(\mathcal T=\{0,1,2,...\}\) we say that the process \(\{X(n)\}_{n\in \mathcal T}\) is a discrete-time process, and we will typically use the dummy variable \(n\).
-When the index set is \(\mathcal T=[0,\infty)\) we say write \(\{X(t)\}_{t\in\mathcal T}\) is a continuous-time process, and we will typically use the dummy variable \(t\).
+A stochastic process is a sequence, \(\{X(t)\}_{t\in\mathcal T}\), of random variables (or random vectors) indexed by some index set \(\mathcal T\).
+In this thesis, in the case that the index set is \(\mathcal T=\{0,1,2,...\}\) we say that the process \(\{X(n)\}_{n\in \mathcal T}\) is a discrete-time process, and we will typically use the dummy variable \(n\) for the `time' index.
+When the index set is \(\mathcal T=[0,\infty)\) we say \(\{X(t)\}_{t\in\mathcal T}\) is a continuous-time process, and we will typically use the dummy variable \(t\) for the `time' index.
 We may omit the index set and write \(\{X(t)\}\) in place of \(\{X(t)\}_{t\in\mathcal T}\) when it is not explicitly needed, or we may write \(\{X(t)\}_{t\geq t_0}\) to mean \(\{X(t)\}_{t\in[t_0,\infty)}\).
-The \emph{sample space} of \(X(t)\) is the set of possible values that \(X(t)\) can take at any \(t\in\mathcal T\).
+The \emph{sample space} of \(X(t)\) is the set of possible values that \(X(t)\) can take at any time \(t\in\mathcal T\).
 
@@
 More generally, for a random variable, \(Z\), we write \(Z\sim \nu\) when \(Z\) has the distribution given by the probability measure \(\nu\).
-For the probability that \(X(t)\) lies in some set \(E\subset \mathcal S\) we write \(\mathbb P(X(t)\in E \mid X(0)\sim \mu)\).
-When \(\mu\) assigns probability 1 to a single point, \(x\in\calS\), say, we write \(\mathbb P(X(t)\in E \mid X(0)=x).\) A stochastic process is \emph{stationary} if
+For the probability that \(X(t)\) lies in some set \(E\subset \mathcal S\) given \(X(0)\sim \mu\), we write \(\mathbb P(X(t)\in E \mid X(0)\sim \mu)\).
+When \(\mu\) assigns probability 1 to a single point, \(x\in\calS\), say, we write \(\mathbb P(X(t)\in E \mid X(0)=x).\)
+
+A stochastic process is \emph{stationary} if
 \begin{align}\label{eqn: stationary}
@@
 \end{align}
-for any \(t,t_0,...,t_n\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\).
+for any \(t,t_0,...,t_n\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\), i.e.~the joint distribution~(\ref{eqn: stationary}) does not depend on \(t\).
 A random variable, \(\tau\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\tau\in\sigma(X(s), s\leq t),\) where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s), s\leq t\}\).
 
-We call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable
-\begin{align}\label{eqn: markov property}
+We call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable and
+\begin{align}\label{eqn: Markov property}
 	\nonumber &\mathbb P(X(n+1)=j \mid X(n)=i,X(n-1)=i_{n-1},...,X(0)=i_0) 
@@
 \end{align}
-for all \(n\in\{0,1,2,...\}\) and \(i,i_0,...,i_{n-1},j\in\calS,\) and refer to (\ref{eqn: markov property}) as the Markov property.
+for all \(n\in\{0,1,2,...\}\) and \(i,i_0,...,i_{n-1},j\in\calS,\) and refer to (\ref{eqn: Markov property}) as the Markov property.
 The process \(\{X(n)\}_{n\in\{0,1,2,...\}}\) is said to be time-homogeneous if
@@
 The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(n)\}_{n\in\{0,1,2,...\}}\), on the event that \(\{\tau<\infty\}\), then
-\begin{align}\label{eqn: strong markov property}
+\begin{align}\label{eqn: strong Markov property}
 	\mathbb P(X(\tau+1)=j \mid \sigma(X(0),X(1),...,X(\tau))) = \mathbb P(X(\tau+1)=j \mid X(\tau)).
@@
 We call \(\{X(t)\}_{t\geq 0}\) a \emph{continuous-time Markov chain} if \(\calS\) is countable and for all \(t_{n+1} > t_n > t_{n-1} > ... > t_0 > 0\) and \(i,i_0,...,i_{n-1},j\in\calS,\)
-\begin{align}\label{eqn: markov property cmtc}
-	\mathbb P(X(t_{n+1})=j \mid X(n)=i, X(n-1)=i_{n-1},...,X(0)=i_0) = \mathbb P(X(t_{n+1})=j \mid X(t_n)=i).
-\end{align}
+\begin{align}\label{eqn: Markov property cmtc}
+	\mathbb P(X(t_{n+1})=j \mid X(n)=i, X(n-1)=i_{n-1},...,X(0)=i_0) = \mathbb P(X(t_{n+1})=j \mid X(t_n)=i),
+\end{align}
+and refer to (\ref{eqn: Markov property cmtc}) as the \emph{Markov property}.
 The process \(\{X(t)\}_{t\geq 0}\) is said to be time-homogeneous if 
@@
 The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(t)\}_{t\geq 0}\), on the event that \(\{\tau<\infty\}\), then
-\begin{align}\label{eqn: strong markov property ctmc}
+\begin{align}\label{eqn: strong Markov property ctmc}
 	\mathbb P(X(\tau + t)=j \mid \sigma(X(s),s\leq \tau)) = \mathbb P(X(\tau+t)=j \mid X(\tau)).
@@
 Most generally, we call \(\{X(t)\}_{t\geq 0}\) a \emph{Markov process} if for all \(t_n > t_{n-1} > ... > t_0 > 0,\, t_k\in\mathcal T,\, k=0,...,n+1,\) and \(E_0,...,E_n\subseteq \mathcal S\)
-\begin{align}\label{eqn: markov process}
+\begin{align}\label{eqn: Markov process}
 	\mathbb P(X(t_n)\in E_n,X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
@@
 \end{align}
-% Markov processes, time-homogeneity, strong markov prperty, partitioning and LOTP
+% Markov processes, time-homogeneity, strong Markov prperty, partitioning and LOTP
 
@@
 The evolution of Markov processes can be described by an \emph{operator semigroup}.
-Furthermore, semigroups arise in other contexts related to the analysis of fluid queues and fluid-fluid queues.
+Semigroups also arise in other contexts related to the analysis of fluid queues and fluid-fluid queues (see, for example, Sections~\ref{sec: ffq intro} and~\ref{sec: transient ffq intro}).
 Semigroups (and therefore Markov processes) can be characterised by their \emph{infinitesimal generator}, or generator for short.
-One of the aims of this thesis is to approximate the generator of a type of Markov process called a fluid queue, which we introduce more formally in Section~\ref{sec: fqs}.
-Here we very briefly introduce semigroups and their \emph{infinitesimal generator}.
+One of the aims of this thesis is to approximate the generator of a fluid queue.
+Here we very briefly introduce semigroups and their \emph{infinitesimal generators}.
 
@@
 An operator is a mapping which takes elements of a vector space and sends them to elements of a vector space (not necessarily the same space).
-That is, let \(V,W\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\subset V\to W\) is an operator.
-The set \(\mathcal D(\mathbb B)\) is the domain of the operator, and is not necessarily all of \(L\), when specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to W\).
+That is, let \(V\) and \(W\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\subset V\to W\) is an operator.
+The set \(\mathcal D(\mathbb B)\) is the domain of the operator, and is not necessarily all of \(V\).
+When specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to W\).
 Let \(c\in \mathbb R\) and \(v\in V\), \(w\in W\).
-Then \(\mathbb B\) is a linear operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\).
-We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms attached to \(V\) and \(W\) respectively.
-An operator is said to be bounded if \(||\mathbb B||<M\) for some \(\infty > M\in \mathbb R\), for ever \(v\in V\).
+Then \(\mathbb B\) is a \emph{linear} operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\).
+We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms attached to \(V\) and \(W\), respectively.
+An operator is said to be bounded if \(||\mathbb B||<M\) for some \(M<\infty, M\in \mathbb R\), for every \(v\in V\).
 
@@
 
-\subsubsection{Semigroups}  A family of linear operators is a collection of indexed linear operators \(\{\mathbb B(i):i\in I\}\) where \(I\) is some index set.
+\subsubsection{Semigroups}  A family of linear operators is a collection of indexed linear operators \(\{\mathbb B(i)\}_{i\in I}\) where \(I\) is some index set.
 That is, for each \(i\in I\), \(\mathbb B(i)\) is a linear operator.
-A family of operators \(\{T(t):L\to L:t\geq 0\}\) is said to have the \emph{semigroup} property if \(\mathbb T(t+s)=\mathbb T(t)T(s)\).
-If, furthermore, \(\mathbb T(0)=I\), the identity operator, then the family \(\{\mathbb T(t):t\geq 0\}\) is known as a semigroup.
-If \(\mathbb T(t)f\to f\) as \(t\to 0^+\), then \(\{\mathbb T(t):L\to L:t\geq 0\}\) is said to be strongly continuous.
-If the \(||\mathbb T(t)||\leq 1\) for all \(t\geq 0\) then \(\{\mathbb T(t):L\to L:t\geq 0\}\) is known as a contraction semigroup.
+A family of operators \(\{\mathbb T(t)\}_{t\geq 0}\), \(\mathbb T(t):L\to L\), is said to have the \emph{semigroup} property if \(\mathbb T(t+s)=\mathbb T(t)\mathbb T(s)\).
+If \(\mathbb T(0)=I\), the identity operator, then the family \(\{\mathbb T(t)_{t\geq 0}\) is known as a semigroup.
+If \(\mathbb T(t)f\to f\) as \(t\to 0^+\), then \(\{\mathbb T(t)\}_{t\geq 0}\) is said to be strongly continuous.
+If \(||\mathbb T(t)||\leq 1\) for all \(t\geq 0\) then \(\{\mathbb T(t)\}_{t\geq 0}\) is known as a contraction semigroup.
 
 For example, for a bounded linear operator \(\mathbb B\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\), then it can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), hence \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup.
-As a more concrete example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\).
-On \(C^\omega(\mathbb R)\), \(\cfrac{\wrt}{\wrt x}\) is bounded.
-Then \(e^{\frac{\wrt}{\wrt x} t } f(x) = \sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is just the Taylor series of \(f\) about the point \(x\).
-
-\subsubsection{Infinitesimal generators} The \emph{infinitesimal generator} (or just \emph{generator} for short) of a semigroup \(\mathbb T(t):L\to L\) is the linear operator defined by 
-\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(T(t)f-I),\]
-and the domain, \(\mathcal D(\mathbb B)\) is the subspace of \(L\) for which this limit exists.
+As another example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\).
+Then \(\sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is just the Taylor series of \(f\) about the point \(x\).
+
+\subsubsection{Infinitesimal generators} The \emph{infinitesimal generator} (or just \emph{generator} for short) of a semigroup \(\{\mathbb T(t)\}\) is the linear operator defined by 
+\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(\mathbb T(t)f-I),\]
+and the domain, \(\mathcal D(\mathbb B)\), is the subspace of \(L\) for which this limit exists.
 Note that an essential part of the definition of the generator is its domain.
 For this reason it is common to denote the generator as the pair \((\mathbb B,\mathcal D(\mathbb B))\).
-Here we will denote the generator as just \(\mathbb B\) and explicitly state its domain where relevant.
 
 To determine the generator \(\mathbb B\) we can proceed by either differentiation of the semigroup, 
-\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(T(t)f-I),\]
+\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(\mathbb T(t)f-I),\]
 or integration;
-\[\int_{t=0}^\infty e^{-st}T(t)\wrt t = (sI-\mathbb B)^{-1}=:R_s.\]
+\[\int_{t=0}^\infty e^{-st}\mathbb T(t)\wrt t = (sI-\mathbb B)^{-1}=:R_s.\]
 The operator \(R_s\) is known as the resolvent.
+
+One of the fundamental results of semigroup theory is the Hille-Yosida Theorem which essentially states that a semigroup is entirely characterised by its generator.
+This is not immediately obvious as the generator is defined on a subset of \(L\) only, whereas \(\{\mathbb T(t)\}\) is defined on all of \(L\).
 
@@
 The Markov property states that for \(s,t\geq 0\), \(X(t+s)\) is independent of \(X(u),\, 0\leq u<t\) given \(X(t)\).
-That is, \(P(X(t+s)\in\mathcal A \mid X(t), X(u), 0\leq u<t)=P(X(t+s)\in\mathcal A \mid X(t))\).
-If \(P(X(t+s)\in\mathcal A \mid X(t))=P(X(s)\in\mathcal A \mid X(0))\), that is the distributions are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.
+That is, \(P(X(t+s)\in\mathcal{E} \mid X(t), X(u), 0\leq u<t)=P(X(t+s)\in\mathcal{E} \mid X(t))\).
+If \(P(X(t+s)\in\mathcal{E} \mid X(t))=P(X(s)\in\mathcal{E} \mid X(0))\), that is the distributions are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.
 Where necessary, we will use the notation \(X_\mu(t)\) to denote the Markov process with initial distribution \(\mu\), i.e.~\(P(X(0)\in\cdot)=\mu(\cdot)\), and \(X_x(t)\) to be the Markov process with initial distribution \(\delta(x)\).
@@
 
-% Let \((U,\mathcal U)\) and \((V,\mathcal V)\) be two measure spaces, then a probability kernel \(\nu\), is a mapping \(\nu:U\times \mathcal V\to \mathbb R_+\) such that \(\nu(u,\mathcal A)\) is \(\mathcal U\)-measurable in \(u\in U\) for fixed \(\mathcal A\in \mathcal V\), and a probability measure in \(\mathcal A\in \mathcal V\) for fixed \(u\in U\).
+% Let \((U,\mathcal U)\) and \((V,\mathcal V)\) be two measure spaces, then a probability kernel \(\nu\), is a mapping \(\nu:U\times \mathcal V\to \mathbb R_+\) such that \(\nu(u,\mathcal{E})\) is \(\mathcal U\)-measurable in \(u\in U\) for fixed \(\mathcal{E}\in \mathcal V\), and a probability measure in \(\mathcal{E}\in \mathcal V\) for fixed \(u\in U\).
 We say that \(\nu\) is a probability kernel from \(U\) to \(V\).
@@
 % With a probability kernel comes associated integral operators. For a measure \(\mu\) on \((U,\mathcal U)\), then 
-% \[\mu \nu (\mathcal A)=\int_{U} \mu(\wrt x) \nu(x,\mathcal A)\]
+% \[\mu \nu (\mathcal{E})=\int_{U} \mu(\wrt x) \nu(x,\mathcal{E})\]
 % is also a measure on \((U,\mathcal U)\). For a function \(f:V\to \mathbb R\) then 
@@
 Then \(\nu\eta\) is also a probability kernel from \(U\) to \(W\);
-% \[\nu\eta(x,\mathcal A) = \int_{V}\nu(x,\wrt y)\eta(y,\mathcal A).\]
+% \[\nu\eta(x,\mathcal{E}) = \int_{V}\nu(x,\wrt y)\eta(y,\mathcal{E}).\]
 
@@
 % \begin{align*}
-% \nu_{t+s}(x,\mathcal A)&=P(X(t+s)\in\mathcal A\mid X(0)=x)
-% \\&=\int_{y\in\mathcal S}P(X(t)\in\wrt y\mid X(0)=x)P(X(t+s)\in\mathcal A\mid X(t)=y)
-% \\&=\int_{y\in\mathcal S}P(X(t)\in\wrt y\mid X(0)=x)P(X(s)\in\mathcal A\mid X(0)=y)
-% \\&=\int_{y\in\mathcal S}\nu_t(x,\wrt y)\nu_s(y,\mathcal A)=\nu_t\nu_s(x,\mathcal A).
+% \nu_{t+s}(x,\mathcal{E})&=P(X(t+s)\in\mathcal{E}\mid X(0)=x)
+% \\&=\int_{y\in\mathcal S}P(X(t)\in\wrt y\mid X(0)=x)P(X(t+s)\in\mathcal{E}\mid X(t)=y)
+% \\&=\int_{y\in\mathcal S}P(X(t)\in\wrt y\mid X(0)=x)P(X(s)\in\mathcal{E}\mid X(0)=y)
+% \\&=\int_{y\in\mathcal S}\nu_t(x,\wrt y)\nu_s(y,\mathcal{E})=\nu_t\nu_s(x,\mathcal{E}).
 % \end{align*}
@@
 % \begin{lem}
-% Every transition probability kernel \(\nu\) induces a Markov operator on measures, \(\mu M(\mathcal A) = \mu \nu(\mathcal A)\), and a transition operator on functions, \(Kf(x)=\nu f(x)\).
-In the other direction, every Markov operator induces a probability kernel \(\delta(x)M(\mathcal A)=\nu(x,\mathcal A)\), and every transition operator induces a transition kernel \(\nu(x,\mathcal A)=K1(\mathcal A)(x)\).
+% Every transition probability kernel \(\nu\) induces a Markov operator on measures, \(\mu M(\mathcal{E}) = \mu \nu(\mathcal{E})\), and a transition operator on functions, \(Kf(x)=\nu f(x)\).
+In the other direction, every Markov operator induces a probability kernel \(\delta(x)M(\mathcal{E})=\nu(x,\mathcal{E})\), and every transition operator induces a transition kernel \(\nu(x,\mathcal{E})=K1(\mathcal{E})(x)\).
 % \end{lem}
@@
 A transition semigroup is the semigroup \(K(t)f(x) = \nu_tf(x)\).
-Similarly we can construct the Markov semigroup, \(\mu M(t)(\mathcal A) = \mu \nu_t (\mathcal A)\) for measures and \(fM(t)(\wrt y) = f\nu_t (\wrt y)\) for densities.
+Similarly we can construct the Markov semigroup, \(\mu M(t)(\mathcal{E}) = \mu \nu_t (\mathcal{E})\) for measures and \(fM(t)(\wrt y) = f\nu_t (\wrt y)\) for densities.
 
@@
 % \subsubsection{A long story short}
-% A time homogeneous Markov processes can be defined by a transition kernel, \(\nu_t(x,\mathcal A)=P(X(t)\in\mathcal A\mid X(0)=x)\) and an initial distribution \(\mu(\mathcal A)=P(X(0)\in\mathcal A)\).
+% A time homogeneous Markov processes can be defined by a transition kernel, \(\nu_t(x,\mathcal{E})=P(X(t)\in\mathcal{E}\mid X(0)=x)\) and an initial distribution \(\mu(\mathcal{E})=P(X(0)\in\mathcal{E})\).
 The operators \(\nu_t\) form a semigroup due to the Markov property and time homogeneity.
-We can construct other semigroups from \(\nu_t\) by either integrating measures on the left, \(\mu\nu_t(\mathcal A)=:\mu M(t)(\mathcal A)\) (which we will call Markov semigroups), or by integrating functions on the right, \(\nu_tf(x) =: K(t)f(x)\).
+We can construct other semigroups from \(\nu_t\) by either integrating measures on the left, \(\mu\nu_t(\mathcal{E})=:\mu M(t)(\mathcal{E})\) (which we will call Markov semigroups), or by integrating functions on the right, \(\nu_tf(x) =: K(t)f(x)\).
 The latter is sometimes known as the transition semigroup, \(K(t):L\to L\) where \(L\) is a Banach space.
@@
 \section{Fluid queues}\label{sec: fqs}
-An unbounded fluid queue is a two-dimensional stochastic process \(\{\ddot{\bs X}(t)\} = \{(\ddot X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{\ddot X(t)\}_{t\geq0}\) is known as the level process or buffer.
-The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\) without loss of generality, and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\).
+An unbounded fluid queue is a two-dimensional stochastic process which we denote by \(\{\ddot{\bs X}(t)\} = \{(\ddot X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{\ddot X(t)\}_{t\geq0}\) is known as the level process or buffer.
+The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\), and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\).
 We assume that \(\bs T\) is \emph{conservative}.
-Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\).
-
-Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{i\in\calS\mid c_i<0\}\), \(\calS_0 = \{i\in\calS\mid c_i=0\}\), \(\mathcal S_{-1}=\{i\in\calS\mid c_i\leq0\},\, \mathcal S_{K+1}=\{i\in\calS\mid c_i\geq0\}\).
+Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\) which determine the rate at which \(\{X(t)\}\) moves.
+
+Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{i\in\calS\mid c_i<0\}\), \(\calS_0 = \{i\in\calS\mid c_i=0\}\), \(\mathcal S_{-1}=\{i\in\calS\mid c_i\leq0\},\, \mathcal S_{K+1}=\{i\in\calS\mid c_i\geq0\}\)\footnote{The notation \(\mathcal S_{-1}\) and \(\mathcal S_{K+1}\) will make sense later when we introduce the partition into cells for the approximation schemes into \(K\) cells, plus a lower and upper boundary which we represent with indices \(-1\) and \(K+1\) respectively.}.
 We assume, without loss of generality, that the generator \(\bs T\) is partitioned into sub-matrices
@@
 \end{align*}
+and 
+\(
+	\widehat{\bs C} = diag(c_i,i\in\calS),
+\)
 where \(diag(a_i,i\in\mathcal I)\) denotes a diagonal matrix with entries \(a_i\) down the diagonal.
@@
 Here, we consider a mixture of \emph{regulated} and \emph{reflecting} boundary conditions.
-Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition).
+Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(sign(c_i)=sign(c_j)\) or \(sign(c_j)=0\) then the process is absorbed in the boundary, otherwise it is reflected.
 At a lower boundary, if \(j\in\calS_0\cup\calS_-\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\), and the phase process continues to evolve according to the sub-generator
 \[\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
-until such a time that \(\varphi(t)\) transitions to a phase \(k\in\calS_+\), at which time \(X(t)\) leaves the boundary.
+until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_+\), at which time \(\{X(t)\}\) leaves the boundary.
 Similarly, at an upper boundary if \(j\in\calS_0\cup\calS_+\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\) and the phase process continues to evolve according to the sub-generator
 \[\left[\begin{array}{cc} \bs T_{++} & \bs T_{+0} \\ \bs T_{0+} & \bs T_{00}  \end{array}\right],\]
-until such a time that \(\varphi(t)\) transitions to a phase \(k\in\calS_-\) at which time \(X(t)\) leaves the boundary.
-Without loss of generality, we assume the lower and upper boundaries (when present) are at \(x=0\) and \(x=M>0\), respectively.
+until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_-\) at which time \(\{X(t)\}\) leaves the boundary.
+Without loss of generality, we assume the lower and upper boundaries (when present) are at \(x=0\) and \(x=b>0\), respectively.
 
@@
 
-Let \(\bs f(x,t) = (f_i(x,t))_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i)\), assuming it exists.
+Let \(\bs f(x,t) = (f_i(x,t))_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i\mid \bs X(0)\sim \bs \mu )\), assuming it exists.
 When a differentiable density exists, the system of partial differential equation which describes the evolution of the densities \(\bs f(x,t)\) is
 \begin{align}
-	\cfrac{\partial}{\partial t} \bs f(x,t) = \bs f(x,t)\bs T - \cfrac{\partial}{\partial x}\bs f(x,t)\bs C,\label{eqn: pde}
+	\cfrac{\partial}{\partial t} \bs f(x,t) = \bs f(x,t)\bs T - \cfrac{\partial}{\partial x}\bs f(x,t)\widehat{\bs C},\label{eqn: pde}
 \end{align}
 on the interior \(x\in(0,b)\), with appropriate boundary conditions (see Section~\ref{subsec: boundary DG}).
-The initial condition is the initial distribution of the fluid queue, \(f_i(x,0)\).
-Often a differentiable density function does not exist and therefore the partial differential equation (\ref{eqn: pde}) is not well-defined.
+The initial condition is the initial distribution of the fluid queue which, when it has a density, we write as \(f_i(x,0)\).
+Often a differentiable density function does not exist and therefore the partial differential equation~(\ref{eqn: pde}) is not well-defined.
 For example, for a fluid queue with no upper boundary, if the initial distribution of the fluid queue is a point mass at any point \(x_0\geq 0\) and in phase \(i\in\calS_+\cup\calS_0\), then a density function \(f_i(x,t)\) will not exist for any finite \(t\).
@@
 	-\int_{x=0}^b\int_{t=0}^\infty \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= &\int_{x=0}^b\int_{t=0}^\infty\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x \nonumber 
-	\\&{} + \int_{x=0}^b \int_{t=0}^\infty\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs \psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
-\end{align}
-for every row-vector of test functions, \(\bs \psi(x,t) = ( \psi_i(x,t))_{i\in\calS}\), which are smooth and have compact support and \(\bs \psi(x,0)=\bs \psi(0,t) = \bs \psi(b,t)=\bs 0\).
-
-\subsection{Transient analysis of fluid queues}
+	\\&{} + \int_{x=0}^b \int_{t=0}^\infty\bs f(x,t)\widehat{\bs C} \cfrac{\partial}{\partial x} \bs \psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
+\end{align}
+for every row-vector of test functions, \(\bs \psi(x,t) = ( \psi_i(x,t))_{i\in\calS}\), which are smooth, have compact support and \(\bs \psi(x,0)=\bs \psi(0,t) = \bs \psi(b,t)=\bs 0\).
+
+\subsection{Transient analysis of fluid queues}\label{sec: transient ffq intro}
 The transient analysis of fluid queues has been relatively well studied (see, for example, \cite{ar2004,bean2005,dasilva2005,bean2009} which are matrix-analytics-methods-based, and also \cite{rs2003} which is differential-equations-based and treats a slightly more complex model than the ones considered here, but is none-the-less relevant).
-Given the interest in discontinuous and point-mass initial conditions, transient analysis of fluid queues rarely relies on solving governing differential equations numerically, although it is quite possible to do so.
-Instead, expressions for transient distributions of fluid queues are derived in terms of Laplace transforms with respect to time, and/or moments of the distributions are derived \cite{ar2004,bean2005,bean2009}.
-Moreover, these techniques often lead to quantities which have direct probabilistic interpretations which further aids in their analysis and also applications to other problems \cite{ar2003,dasilva2005,bnp2018} (we also leverage these stochastic interpretations in Chapter~\ref{sec: conv} to derive expressions for certain the Laplace transforms of the fluid queue).
+Given the interest in discontinuous and point-mass initial conditions, transient analysis of fluid queues rarely relies on solving governing differential equations numerically, although it is quite possible to do so\footnote{We have already alluded to some difficulties with differential-equation-based approaches to problems with discontinuous solutions such as ill-defined PDEs, the need for weak solutions, and oscillatory approximations with possibly infeasible approximate solutions, for example negative probability approximations.}.
+Instead, expressions for transient distributions of fluid queues are derived in terms of Laplace transforms with respect to time, and/or moments of the transient distributions are derived \citep{ar2004,bean2005,bean2009}.
+Moreover, these techniques often lead to quantities which have direct probabilistic interpretations which further aids in their analysis and also applications to other problems \citep{ar2003,dasilva2005,bnp2018} (we also leverage these stochastic interpretations in Chapter~\ref{sec: conv} to derive expressions for certain the Laplace transforms of the fluid queue).
 In some context, the actual transient distributions are not required, and Laplace transforms or moments are all that are required.
@@
 
-One such analysis of fluid queues derives the Laplace transform of the time taken for the level of an unbounded fluid queue to return to its initial level in a certain phase, given it started in a phase with positive rate \cite{bean2005}.
+One such analysis of fluid queues derives the Laplace transform of the time taken for the level of an unbounded fluid queue to return to its initial level in a certain phase, given it started in a phase with positive rate \citep{bean2005}.
 The principles underlying the derivation of this first return operator are the same as those applied in \cite{bo2014} to derive the first return operator for fluid-fluid queues.
-Moreover, certain matrices appearing in the analysis have a stochastic interpretation which we leverage and write down certain Laplace transforms in Chapter~\ref{sec: conv} in terms of these matrices.
+Moreover, certain matrices appearing in the analysis have a stochastic interpretation which we leverage to write down certain Laplace transforms in Chapter~\ref{sec: conv} in terms of these matrices.
 Given its relevance we briefly recount the analysis of \cite{bean2005} here.
 
-Consider an unbounded fluid queue \((\{\ddot X(t), \varphi(t)\})\) as described in \ref{sec: fqs}.
-Let \(\zeta_X(E)\) be the random variable which is the first hitting time of \(\{\ddot X(t)\}\) on the set \(E\) and define the matrix \(\bs \Psi_X(s)\) with elements \(\left[\bs \Psi_X(s)\right]_{ij}\) where \(i\in\calS_+\), \(j\in\calS_-\),
-\begin{align}
-	\left[\bs \Psi_X(\lambda)\right]_{ij} = \mathbb E\left[ e^{-\zeta_X(z)\lambda} 1(\zeta_X(z)<\infty, \varphi(\zeta_X(z)=j)\mid \ddot X(0)=z, \varphi(0)=i\right],
-\end{align}
-i.e.~the Laplace-Stieltjes transform of the time taken for the fluid queue to first return to level \(z\) and do so in phase \(j\), given it started at level \(z\) in phase \(i\).
-Define $\beta_X(t) := \int_0^t \left| c_{\varphi(z)} \right|  \wrt z$, which is the total amount of fluid to flow in or out of the buffer \(\{\ddot X(t)\}\) by time \(t\), and $\eta_X(y) := \inf \{t > 0: \beta_X(t) = y\}$ as the first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\).
-Further, let \(\bs H(\lambda,y)\) be the matrix with elements \(\left[\hat{\bs{\Delta}}^y(\lambda)\right]_{ij}\), \(i,j\in\calS_+\cup\calS_-\), given by
+Consider an unbounded fluid queue \(\{(\ddot X(t), \varphi(t))\}\).
+Let \(\zeta_X(E)\) be the random variable which is the first hitting time of \(\{\ddot X(t)\}\) on the set \(E\) and define the matrix \(\bs \Psi_X(s)\) with elements \(\left[\bs \Psi_X(s)\right]_{ij}\), \(i\in\calS_+\), \(j\in\calS_-\), given by
+\begin{align}
+	\left[\bs \Psi_X(\lambda)\right]_{ij} = \mathbb E\left[ e^{-\zeta_X(z)\lambda} 1(\zeta_X(z)<\infty, \varphi(\zeta_X(z)=j)\mid \ddot X(0)=z, \varphi(0)=i\right].
+\end{align}
+\(\left[\bs \Psi_X(\lambda)\right]_{ij}\) is the Laplace-Stieltjes transform of the time taken for the fluid queue to first return to level \(z\) and do so in phase \(j\in\calS_-\), given it started at level \(z\) in phase \(i\in\calS_+\).
+Define the in-out fluid level by $\beta_X(t) := \int_0^t \left| c_{\varphi(z)} \right|  \wrt z$, which is the total amount of fluid to flow in to or out of the buffer \(\{\ddot X(t)\}\) by time \(t\), and also define $\eta_X(y) := \inf \{t > 0: \beta_X(t) = y\}$ as the first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\).
+Further, let \(\bs H(\lambda,y)\) be the matrix with elements \(h_{ij}(\lambda,y)\), \(i,j\in\calS_+\cup\calS_-\), given by
 \begin{align}
@@
 
-It turns out that, for fixed \(\lambda \geq 0\), \(\bs H (\lambda,y)\) is a semigroup (with variable \(y\)) \cite{bean2005}.
-As such, \cite{bean2005} find the infinitesimal generator, \(\bs Q(\lambda)\) of \(\bs H(\lambda,y)\) and can therefore write \(\bs H(\lambda,y)=e^{\bs Q(\lambda)y}\).
-The derivation of the generator \(\bs Q(\lambda)=[Q_{ij}(\lambda)]_{ij\in\calS_+\cup\calS_-}\) is based on a direct analysis of sample paths of the fluid queue over an infinitesimal time, \(u\) say, and leveraging the fact that complex sample paths occur with probability \(\mathcal O(u^2)\).
+It turns out that, for fixed \(\lambda \geq 0\), \(\bs H (\lambda,y)\) is a semigroup (with variable \(y\)) \citep{bean2005}.
+\cite{bean2005} find the infinitesimal generator, \(\bs Q(\lambda)\) of \(\bs H(\lambda,y)\) and, since the generator is bounded, we can write \(\bs H(\lambda,y)=e^{\bs Q(\lambda)y}\).
+The derivation of the generator \(\bs Q(\lambda)=[Q_{ij}(\lambda)]_{ij\in\calS_+\cup\calS_-}\) is based on a direct analysis of sample paths of the fluid queue over an infinitesimal time, \(u\) say, and by leveraging the fact that complex sample paths occur with probability \(\mathcal O(u^2)\).
 As a result, there are only three types of sample paths to consider.
@@
 \end{align*}
-The functions 
+\cite{bean2005} also define the functions 
 \begin{align}
@@
 \end{align}
-for \(y,\lambda\geq 0\) have the following stochastic interpretations.
-The function \(h_{ij}^{++}(\lambda,y)\) (\(h_{ij}^{--}(\lambda,y)\)) is the Laplace transform with respect to time of the time taken for the fluid level to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), given the phase was initially \(i\in\mathcal S_+\) (\(i\in\mathcal S_-\)) \citep{bean2005}.
-The function \(h_{ij}^{+-}(\lambda,y)\) (\(h_{ij}^{-+}(\lambda,y)\)) is the Laplace transform with respect to time of the time taken for the fluid level, \(\{y(t)\}\) to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), after which time the phase instantaneously changes to \(j\in\mathcal S_-\) (\(\mathcal S_+\)), given the phase was initially \(i\in\mathcal S_+\) (\(\mathcal S_-\)) \citep{bean2005}.
-
-Thus, \cite{bean2005} are able to characterise by the above matrix expressions, segments of sample paths of the fluid queue where the fluid level is non-decreasing and non-increasing.
-\cite{bean2005} then partition the sample paths which contribute to \(\Psi_X\) as those which either, (a), have a single transition from \(\calS_+\) to \(\calS_-\) (perhaps via \(\calS_0\)), and, (b), those which have more than one transition from \(\calS_+\) to \(\calS_-\) (perhaps via \(\calS_0\)).
+for \(y,\lambda\geq 0,\) which have the following stochastic interpretations.
+The function \(h_{ij}^{++}(\lambda,y)\) (\(h_{ij}^{--}(\lambda,y)\)) is the Laplace transform (with respect to time) of the time taken for the fluid level to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), given the phase was initially \(i\in\mathcal S_+\) (\(i\in\mathcal S_-\)).
+The function \(h_{ij}^{+-}(\lambda,y)\) (\(h_{ij}^{-+}(\lambda,y)\)) is the Laplace transform (with respect to time) of the time taken for the fluid level, \(\{Y(t)\}\) to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), after which time the phase instantaneously changes to \(j\in\mathcal S_-\) (\(\mathcal S_+\)), given the phase was initially \(i\in\mathcal S_+\) (\(\mathcal S_-\)) \citep{bean2005}.
+
+Thus, \cite{bean2005} are able to characterise, by the above matrix expressions, segments of sample paths of the fluid queue where the fluid level is non-decreasing and non-increasing.
+\cite{bean2005} then partition the sample paths which contribute to \(\bs \Psi_X\) as those which either, (a) have a single transition from \(\calS_+\) to \(\calS_-\) (perhaps via \(\calS_0\)), and, (b) those which have more than one transition from \(\calS_+\) to \(\calS_-\) (perhaps via \(\calS_0\)).
 An expression for Laplace-Stieltjes transform of (a) is
 \[\int_{y=0}^\infty e^{\bs Q_{++}(\lambda)y}\bs Q_{+-}(\lambda)e^{\bs Q_{--}y}\wrt y.\]
-For the sample paths (b), there must be at least on point at which the phase process transitions from \(\calS_-\) to \(\calS_+\) (perhaps via \(\calS_0\)) before the first return time.
+For the sample paths (b), there must be at least one point at which the phase process transitions from \(\calS_-\) to \(\calS_+\) (perhaps via \(\calS_0\)) before the first return time.
 Further, one of the transitions from \(\calS_-\) to \(\calS_+\) (perhaps via \(\calS_0\)) must occur lower than all others.
@@
 
-Adding the expressions for (a) and (b) together, then \cite{bean2005} state 
+Adding the expressions for (a) and (b) together, \cite{bean2005} state 
 \begin{align}
@@
 
-The key concepts of this argument are; (1) to partition the state space of the driving process into sets on which the fluid level is increasing, decreasing, or constant; (2) to characterise the sections of sample paths of the fluid level which are non-decreasing and non-increasing as semigroups and derive their generators; (3) to partition the sample paths which comprise \(\bs \Psi_X(\lambda)\) such that we can write down an expression for \(\bs \Psi_X(\lambda)\) in terms of the expressions derived in (2) and \(\bs \Psi_X(\lambda)\) and solve the resulting expression.
+The key concepts of this argument are; (1) to partition the state space of the driving process into sets on which the fluid level is increasing, decreasing, or constant; (2) to characterise the sections of sample paths of the fluid level which are non-decreasing and non-increasing as semigroups and derive their generators; (3) to partition the sample paths which comprise \(\bs \Psi_X(\lambda)\) such that we can write down an expression for \(\bs \Psi_X(\lambda)\) in terms of the expressions derived in (2) and \(\bs \Psi_X(\lambda)\) itself and then to solve the resulting expression.
 
@@
 
-% Fluid models (with or without boundaries) are Feller processes, hence their associated semigroups are Feller semigroups.
-Let \(\boldsymbol f(x,t)\) be the density of a fluid model at time \(t\).
-Denote the Markov semigroup associated with the fluid model as \(V(t)\).
+% fluid queues (with or without boundaries) are Feller processes, hence their associated semigroups are Feller semigroups.
+Let \(\boldsymbol f(x,t)\) be the density of a fluid queue at time \(t\).
+Denote the Markov semigroup associated with the fluid queue as \(V(t)\).
 Therefore, \(\boldsymbol f(x,t)\wrt x=\boldsymbol \mu V(t)(\wrt x)\), where \(\boldsymbol \mu(\wrt x)=\boldsymbol f(x,0)\wrt x\) is the initial probability density.
@@
 
-\section{Fluid-fluid Queues}
+\section{Fluid-fluid Queues}\label{sec: ffq intro}
 	\label{sec:prelim}
@@
 
-An unbounded stochastic fluid-fluid queue \citep{bo2013} is a Markov process with three elements, $\{(\ddot{X}(t), \ddot{Y}(t), \varphi(t))\}_{t \geq 0},$ where $\{(X(t),\varphi(t))\}_{t\geq 0}$ is a classical fluid queue and $\ddot{Y}(t)$ is the second fluid, which varies at rate $r_{\varphi(t)}(\ddot{X}(t))$:
-% 
-\begin{align*} 
-	\ddot{Y}(t) := \ddot Y(0)) + \int_0^t r_{\varphi(s)}(\ddot{X}(s)) \wrt s.
+A stochastic fluid-fluid queue \citep{bo2013} is a Markov process with three elements, $\{( \ddot{Y}(t),{X}(t), \varphi(t))\}_{t \geq 0},$ where $\{(X(t),\varphi(t))\}_{t\geq 0}$ is a classical fluid queue\footnote{We assume \(X(t)\) is doubly-bounded as this results in a finite-dimensional approximation, but this is not necessary} and $\ddot{Y}(t)$ is the second fluid, which varies at rate $r_{\varphi(t)}(\dot{X}(t))$:
+% 
+\begin{align*} 
+	\ddot{Y}(t) := \ddot Y(0)) + \int_0^t r_{\varphi(s)}({X}(s)) \wrt s.
 \end{align*} 
@@
 In this thesis we consider the second fluid to have a regulated boundary at \(y=0\) and unbounded above.
-To distinguish between unbounded and bounded processes, we use the notation \(\ddot Y(t)\) to denote unbounded processes and \(\dot Y(t)\) to denote a second fluid level with a regulated lower boundary at 0.
-
-% As classic fluid processes, $\{(\widehat X(t), \varphi(t)),t \geq 0\}$, or unbounded analogues, are used extensively in many areas, such as insurance and environmental modelling, it is clear that stochastic fluid-fluid models have an even wider range of applicability.
-
-In the following, we assume that $\dot Y(t) \in [0,\infty)$ and that there is a boundary at level $0$ for both the first and second fluid levels and the first fluid level is also bounded above at \(b\):
+To distinguish between unbounded and bounded processes, we use the notation \(\ddot Y(t)\) to denote the unbounded process and \(\dot Y(t)\) to denote the second fluid level process with a regulated lower boundary at 0.
+
+% As classic fluid processes, $\{(\widehat X(t), \varphi(t)),t \geq 0\}$, or unbounded analogues, are used extensively in many areas, such as insurance and environmental modelling, it is clear that stochastic fluid-fluid queues have an even wider range of applicability.
+
+In the following, we assume that $\dot Y(t) \in [0,\infty)$ and that there is a boundary at level $0$ for both the first and second fluid levels and the first fluid level is also bounded above at \(b\).
+Thus, the rates of change of the fluid levels can be summarised as:
 % 
@@
 	% 
-for $i \in \mathcal{S} = \{1,...,N_\calS\}$. Let $\bs R(x) := \diag(r_i(x))_{i \in \mathcal{S}}$ be the diagonal fluid-rate matrix of functions for $\{\dot Y(t)\}$.
+for $i \in \mathcal{S} = \{1,...,N\}$. Let $\bs R(x) := \diag(r_i(x))_{i \in \mathcal{S}}$ be the diagonal fluid-rate matrix of functions for $\{\dot Y(t)\}$.
 
@@
 
-The concepts leading to the derivation of \( {\mathbb \Psi}\) for the fluid-fluid queue are much the same as the derivation of the analogous quantity for a classical fluid queue.
-For a classical fluid queue the infinitesimal generator of the driving process is the matrix \(\bs T\).
-We partition \(\calS\) into the sets \(\calS_+,\,\calS_-,\, \calS_0\) according to whether \(c_i\) is positive, negative or zero, respectively, and partition \(\bs T\) similarly;
-\[\bs T = \left[\begin{array}{ccc}\bs T_{++} & \bs T_{+-} & \bs T_{+0} \\ \bs T_{-+} & \bs T_{--} & \bs T_{-0} \\ \bs T_{0+} & \bs T_{0-} & \bs T_{00}  \end{array}\right].\]
-Using the block matrices \(\bs T_{mn},\,m,n\in\{+,-,0\}\), and the rates \(c_i\), we can write down the distribution of the phase on sample paths where \(\{X(t)\}\) is increasing or decreasing.
-Further, we can also determine the distribution of the phase process at the time when the fluid level goes from increasing to decreasing or decreasing to increasing.
-Ultimately, we are able to derive a matrix-Riccati equation where the solution is the unique matrix with \((i,j)\)th entry, \(i\in\calS_+\), \(j\in\calS_-\), which is the probability that \(\{X(t)\}\) first returns to zero in phase \(j\) given it started at zero in phase \(i\).
-That is, the solution of the Riccati equation gives us the distribution of the driving process at the time when the level process first returns to \(0\).
-
-For the fluid-fluid queue we can use analogous arguments.
+The concepts leading to the derivation of \( {\mathbb \Psi}\) for the fluid-fluid queue are much the same as the derivation of the analogous quantity, \(\bs\Psi_X\), for a classical fluid queue.
 First, we need the infinitesimal generator of the driving process, \(\{(X(t),\varphi(t))\}\), which we denote by \(\mathbb B\).
-Then we need to partition \(\mathbb B\) on sets for which \(\{\dot Y(t)\}\) is increasing, decreasing or constant -- these are the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\).
+Then we need to partition \(\mathbb B\) on sets for which \(\{\dot Y(t)\}\) is increasing, decreasing or constant -- these are the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), below.
 We can then derive an operator-Riccati equation for the first return operator for the fluid-fluid queue, \(\mathbb \Psi\), in terms of the partitioned generator \(\mathbb B\), the rates \(r_i(x)\).
@@
 The problem we have is to solve the operator-Riccati equation -- something which is only possibly in the very simplest of cases.
-This is where the methods in this thesis comes in.
-Here, we approximate \(\mathbb B\) by a finite-dimensional matrix, which we then partition according to the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\).
-We then substitute the resulting matrices into the operator-Riccati equation and this gives us a matrix-Riccati equation, which we can then solve using known methods.
-
-We use finite-element methods for the approximation of \(\mathbb B\).
-A key element of finite-element schemes is the partition of the state space of \(\{(X(t),\varphi(t))\}\) into sets of the form \(([y_k,y_{k+1}],i)\), where \([y_k,y_{k+1}]=:\calD_k\) are known as cells.
+This is where the methods in this thesis come in.
+Here, we approximate \(\mathbb B\) by a finite-dimensional matrix which we then partition according to the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\).
+We then substitute the resulting matrices into the operator-Riccati equation and this gives us a matrix-Riccati equation which we can then solve using known methods \cite{bot08}.
+
+A key element of the cell-based approximation schemes is the partition of the state space of \(\{(X(t),\varphi(t))\}\) into sets of the form \((\calD_{k,i},i)\), where \(\calD_{k,i}\) are intervals are known as cells.
 We must choose the cells wisely so that the partition into the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\) can be recovered from the partition into cells.
 This is not terribly difficult to do, but does require some notation and explanation.
-We now proceed to introduce the work of \cite{bo2014} and explain how we can recover the partition into the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\) from the cells as determined by the finite-element approximation.
-
-For each Markovian state $i \in \mathcal{S}$, we partition the state space of \( X(t)\), \([0,\infty)\), according to the rates of change $r_i(\cdot)$ for the second fluid $\{\dot Y(t)\}$: $[0,\infty) := \mathcal{F}^{+}_i \cup \mathcal{F}^{-}_i \cup \mathcal{F}^{0}_i,$
+We now proceed to introduce the work of \cite{bo2014} and explain how we can recover the partition into the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\) from the cells as determined by the cell-based approximations.
+
+For each Markovian state $i \in \mathcal{S}$, we partition the state space of \( X(t)\), \([0,b]\), according to the rates of change $r_i(\cdot)$ for the second fluid $\{\dot Y(t)\}$: $[0,b] := \mathcal{F}^{+}_i \cup \mathcal{F}^{-}_i \cup \mathcal{F}^{0}_i,$
 where 
@@
 %				\label{eqn:Fi1}
-			\mathcal{F}^{+}_i & := \{u \in \mathcal{F} : r_i(u) > 0\},  \;
+			\mathcal{F}^{+}_i & := \{u \in [0,b] : r_i(u) > 0\},  \; \nonumber
 %				\label{eqn:Fi2}
-			\mathcal{F}^{-}_i :=  \{u \in \mathcal{F}:  r_i(u) < 0\}, \;
+			\\\mathcal{F}^{-}_i &:=  \{u \in [0,b]:  r_i(u) < 0\}, \; \nonumber
 %				\label{eqn:Fi3}
-			\mathcal{F}^{0}_i := \{u \in \mathcal{F}: r_i(u) = 0\}.\label{eqn:fil}
-		\end{align} 
-% 
-For all $i \in \mathcal{S}$, the functions $r_i(\cdot)$ are assumed to be sufficiently well-behaved that $\mathcal{F}^{m}_i$, $m \in \{+, -, 0\}$, is a finite union of intervals and isolated points.
-
-We assume that the process $\{(\dot Y(t), X(t), \varphi(t))\}_{t\geq 0}$ is positive recurrent, in order to guarantee the existence of the joint stationary density. Define stationary operators 
+			\\\mathcal{F}^{0}_i &:= \{u \in [0,b]: r_i(u) = 0\}.\label{eqn:fil}
+		\end{align} 
+% 
+For all $i \in \mathcal{S}$, the functions $r_i(\cdot)$ are assumed to be sufficiently well-behaved so that $\mathcal{F}^{m}_i$, $m \in \{+, -, 0\}$, is a finite union of intervals and isolated points.
+
+We assume that the process $\{(\dot Y(t), X(t), \varphi(t))\}_{t\geq 0}$ is positive recurrent, in order to guarantee the existence of the joint stationary distribution. Define stationary operators 
 \begin{align} 
 		\label{eqn:jointpi} 
-		\bbpi_i(y)(\mathcal{A}) & := \lim_{t \rightarrow \infty} \frac{\partial}{\partial y} \mathbb{P}\left( \dot Y(t) \leq y,  X(t) \in \mathcal{A}, \varphi(t) = i\right), \quad y>0,\\
+		\bbpi_i(y)(\mathcal{E}) & := \lim_{t \rightarrow \infty} \frac{\partial}{\partial y} \mathbb{P}\left( \dot Y(t) \leq y,  X(t) \in \mathcal{E}, \varphi(t) = i\right), \quad y>0,\\
 % 
 		\label{eqn:jointmass}
-		{\mathbb p}_i(\mathcal{A}) & := \lim_{t \rightarrow \infty}  \mathbb{P}( \dot Y(t) = 0,  X(t) \in \mathcal{A}, \varphi(t) = i]),
-	\end{align} 
-	where $\mathcal{A} \subset [0,\infty)$.
-Then let ${{\bbpi}}(y) = ( \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint stationary density operators and ${\mathbb{p}} = ( {\mathbb p}_i)_{i \in \mathcal{S}}$ a vector containing the joint stationary mass operators.
+		{\mathbb p}_i(\mathcal{E}) & := \lim_{t \rightarrow \infty}  \mathbb{P}( \dot Y(t) = 0,  X(t) \in \mathcal{E}, \varphi(t) = i]),
+	\end{align} 
+	where $\mathcal{E} \subset [0,b]$.
+Then let ${{\bbpi}}(y) = ( \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint stationary density operators and ${\mathbb{p}} = ( {\mathbb p}_i)_{i \in \mathcal{S}}$ be a vector containing the joint stationary mass operators.
 The determination of ${{\bbpi}}(y)$ involves two important matrices of operators, $ {\mathbb D}$ and $ {\mathbb \Psi}$ which we now introduce.
 %The operator \( {\mathbb B}\) is the \textit{infinitesimal generator} of the process \(\{(X(t),\varphi(t))\}\).
-%Intuitively, for a set $\mathcal{A} \subset \mathcal{F}$ and a measure vector $\boldsymbol{\mu} = (\mu_i)_{i \in \mathcal{S}}$, $\bs{\mu}V(t)(\mathcal{A})$, where \(V(t)\) is a semigroup with infinitesimal generator \(B\), gives the conditional probability of $X(t) \in \mathcal{A}$,
-% The operator \( {\mathbb \Psi}\) is such that ${\boldsymbol{\mu}}^{+} {\mathbb \Psi} (\mathcal{A})$ gives the conditional probability of $\{\dot Y(t)\}$ returning to level zero and doing so when $X(t) \in \mathcal{A}$, given that the initial distribution of \((X(0),\varphi(0))\) in the set \(\bigcup_{i\in\calS}(\mathcal F_i^+,i)\), is ${\boldsymbol{\mu}}^{+}$.
+%Intuitively, for a set $\mathcal{E} \subset \mathcal{F}$ and a measure vector $\boldsymbol{\mu} = (\mu_i)_{i \in \mathcal{S}}$, $\bs{\mu}V(t)(\mathcal{E})$, where \(V(t)\) is a semigroup with infinitesimal generator \(B\), gives the conditional probability of $X(t) \in \mathcal{E}$,
+% The operator \( {\mathbb \Psi}\) is such that ${\boldsymbol{\mu}}^{+} {\mathbb \Psi} (\mathcal{E})$ gives the conditional probability of $\{\dot Y(t)\}$ returning to level zero and doing so when $X(t) \in \mathcal{E}$, given that the initial distribution of \((X(0),\varphi(0))\) in the set \(\bigcup_{i\in\calS}(\mathcal F_i^+,i)\), is ${\boldsymbol{\mu}}^{+}$.
 
 \subsection{The infinitesimal generator, \(\mathbb B\), of the driving process}
-Given the discussion above, if we are to replicate the arguments of \cite{bean2005} to derive the Laplace-Stieltjes transform of the first return operator of a fluid-fluid queue, we first need an expression for the generator of the driving process, which, for a fluid-fluid queue is a fluid queue.
-We then need to partition it according to whether the second fluid is increasing, decreasing or constant.
-
-The generator of a fluid queue is a differential operator, and to enable computation approximation method are needed.
+Given the discussion above, if we are to replicate the arguments of \cite{bean2005} to derive the Laplace-Stieltjes transform of the first return operator of a fluid-fluid queue, we first need an expression for the generator of the driving process.
+We also need to partition it according to whether the second fluid is increasing, decreasing or constant.
+
+The generator of a fluid queue is a differential operator and to enable computation of the first-return operator, approximation method are needed.
 The approximation schemes which we discuss in this thesis are all cell-based methods which discretise the level of the fluid queue into cells.
 Thus, one complexity in approximation is to reconcile the partition according to whether the second fluid is increasing, decreasing or constant and the partition which the approximation method uses.
-As long as the rates of the fluid-fluid queue are not too-badly behaved, this is not terribly difficult to do, but it does require some discussion and notation.
+As long as the rates of the fluid-fluid queue are not `too-badly' behaved, this is not terribly difficult to do, but it does require some discussion and notation.
 
 \label{subsec:B_operators}
-\subsubsection{General definition}
+\subsubsection{The transition semigroup of a fluid queue}
 Since \(\{( X(t),\varphi(t))\}_{t\geq 0}\) is a Markov process, the evolution of probability can be described by a semigroup.
-Let $\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$ be the set of integrable complex-valued Borel measures on the Borel $\sigma$-algebra $\mathcal{B}_{\mathcal{S} \times \mathbb{R}_{+}}$.
-For $\overline{\bs{\mu}} \in \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$, we can write \(\overline{\bs{\mu}} = (\overline{\mu_i})_{i \in \mathcal{S}}\).
-The measures \(\overline{\mu_i}(\cdot)\) represent an initial distribution, \(\overline{\mu_i}(\cdot) = \mathbb P(X(0))\in\cdot, \varphi(0) = i)\).
-Let \(\{\overline{\mathbb V}(t)\}_{t\geq 0},\, \overline{\mathbb V}(t):\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\mapsto \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\) be the semigroup describing the evolution of probability for \(\{(X(t),\varphi(t))\}_{t\geq 0}\) structured as a matrix of operators, \(\left[ \overline{\mathbb V}(t)\right]_{ij}= \overline{\mathbb V}_{ij}(t)\) where,
-\[\overline\mu_i \overline{ \mathbb V}_{ij}(t)(\mathcal A) = \int_{x\in[0,\infty)} \wrt \overline\mu_i(x)\mathbb P (X(t) \in\mathcal A,\varphi(t) = j \mid X(0)) = x, \varphi(0) = i).\]
-Intuitively, the operator \( \overline{\mathbb V}(t)\) maps an initial measure \(\overline{\boldsymbol \mu}\) on \((X(0),\varphi(0))\) to the measure \(\mathbb P(X(t)\in \mathcal A, \varphi(t)=j)=:\overline\mu_j(t)(\mathcal A)\).
+Let $\mathcal{M}(\mathcal{S} \times [0,b])$ be the set of integrable complex-valued Borel measures on the Borel $\sigma$-algebra $\mathcal{B}_{\mathcal{S} \times [0,b]}$.
+For $\overline{\bs{\mu}} \in \mathcal{M}(\mathcal{S} \times [0,b])$, we can write \(\overline{\bs{\mu}} = (\overline{\mu_i})_{i \in \mathcal{S}}\).
+The measures \(\overline{\mu_i}(\cdot)\) represent an initial distribution, \(\overline{\mu_i}(\cdot) = \mathbb P(X(0)\in\cdot, \varphi(0) = i)\).
+Let \(\{\overline{\mathbb V}(t)\}_{t\geq 0},\, \overline{\mathbb V}(t):\mathcal{M}(\mathcal{S} \times [0,b]\mapsto \mathcal{M}(\mathcal{S} \times [0,b]\) be the semigroup describing the evolution of probability for \(\{(X(t),\varphi(t))\}_{t\geq 0}\) structured as a matrix of operators, \(\left[ \overline{\mathbb V}(t)\right]_{ij}= \overline{\mathbb V}_{ij}(t)\) where,
+\[\overline\mu_i \overline{ \mathbb V}_{ij}(t)(\mathcal{E}) = \int_{x\in[0,b]} \wrt \overline\mu_i(x)\mathbb P (X(t) \in\mathcal{E},\varphi(t) = j \mid X(0) = x, \varphi(0) = i).\]
+Intuitively, the operator \( \overline{\mathbb V}(t)\) maps an initial measure \(\overline{\boldsymbol \mu}\) on \((X(0),\varphi(0))\) to the measure \(\mathbb P(X(t)\in \mathcal{E}, \varphi(t)=j)=:\overline\mu_j(t)(\mathcal{E})\).
 The matrix of operators \( \overline{\mathbb B}:=[\overline{\mathbb B}_{ij}]_{i,j\in\mathcal S}\) is the \textit{infinitesimal generator} of the semigroup \(\{ \overline{\mathbb V}(t)\}\) defined by 
@@
 with domain the set of measures for which this limit exists.
-Specifically, the domain of \(\overline{\mathbb B}\) is the set of measures, \(\overline{\bs \mu} = (\overline\mu_i)_{i\in\calS}\), for which each \(\overline\mu_i\) admits an absolutely continuous density on \((0,\infty)\), and can have a point mass at \(0\) if \(i\in\mathcal S_{-1}\) or a point mass at \(b\) if \(i\in\mathcal S_{K+1}\); call this set of measures \(\mathcal M_{0,b}\).
+Specifically, the domain of \(\overline{\mathbb B}\) is the set of measures, \(\overline{\bs \mu} = (\overline\mu_i)_{i\in\calS}\), for which each \(\overline\mu_i\) admits an absolutely continuous density on \((0,b)\), and can have a point mass at \(0\) if \(i\in\mathcal S_{-1}\) or a point mass at \(b\) if \(i\in\mathcal S_{K+1}\); call this set of measures \(\mathcal M_{0,b}\).
 The measure \(\overline\mu_i\) cannot have a point mass at 0 if \(i\notin \mathcal S_{-1}\), nor can they have a point mass at \(b\) if \(i\notin\calS_{{K+1}}\).
-In the sequel we write \(v_i(x),\, x>0\), as the density of \(\overline\mu_i\), and \(q_{{-1},i}\) and \(q_{{K+1},i}\) as the point masses of \(\overline\mu_i\) at \(0\) and \(b\), respectively (if such point masses exist).
+In the sequel we write \(\overline v_i(x)\) as the density of \(\overline\mu_i\), and \(q_{{-1},i}\) and \(q_{{K+1},i}\) as the point masses of \(\overline\mu_i\) at \(0\) and \(b\), respectively (if such point masses exist).
 
 \subsubsection{A partition with respect to rates of the second fluid}
-To use the operators \(\{ \overline{\mathbb  V}(t)\}\) and \( \overline{\mathbb  B}\) to analyse the fluid-fluid model, \cite{bo2014} explicitly track when \((X(t),\varphi(t))\in(\mathcal F_i^m,i)\) for \(i\in\mathcal S,\, m \in \{+,-,0\}\) by partitioning the operators \( \overline{\mathbb  V}(t)\) and \( \overline{\mathbb  B}\) into \( \overline{\mathbb  V}_{ij}^{m n}\) and \( \overline{\mathbb  B}_{ij}^{m n}\), for \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\), where
-\[\left.\mu_i\right|_{\calF_i^m}  \overline{\mathbb  V}_{ij}^{m n}(t)(\mathcal A) := \int_{x\in[0,\infty)} \left.
-\wrt \mu_i\right|_{\calF_i^m}(x)\mathbb P (X(t) \in\mathcal A\cap \calF_j^n,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
-and \(\left.\mu_i\right|_{E}\) is the restriction of \(\mu_i\) to \(E\). Similarly, for \( \overline{\mathbb  B}_{ij}^{m n},\) \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\).
-
-\subsubsection{A partition as dictated by the finite-element method}
+To use the operators \(\{ \overline{\mathbb  V}(t)\}\) and \( \overline{\mathbb  B}\) to analyse the fluid-fluid queue, \cite{bo2014} explicitly track when \((X(t),\varphi(t))\in(\mathcal F_i^m,i)\) for \(i\in\mathcal S,\, m \in \{+,-,0\}\) by partitioning the operators \( \overline{\mathbb  V}(t)\) and \( \overline{\mathbb  B}\) into \( \overline{\mathbb  V}_{ij}^{m n}\) and \( \overline{\mathbb  B}_{ij}^{m n}\), for \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\), where
+\[\left.\overline\mu_i\right|_{\calF_i^m}  \overline{\mathbb  V}_{ij}^{m n}(t)(\mathcal{E}) := \int_{x\in[0,b]} \left.
+\wrt \overline\mu_i\right|_{\calF_i^m}(x)\mathbb P (X(t) \in\mathcal{E}\cap \calF_j^n,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
+and \(\left.\overline\mu_i\right|_{E}\) is the restriction of \(\overline\mu_i\) to \(E\), \(\left.\overline\mu_i\right|_{E}(\mathcal E)=\overline\mu_i(\mathcal E\cap E)\).
+Similarly, for \( \overline{\mathbb  B}_{ij}^{m n},\) \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\).
+
+\subsubsection{A partition as dictated by the approximation schemes}
 We claim that numerical schemes are needed to approximate the analytic operator equations introduced in \cite{bo2014}.
-The schemes we choose to use here work by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \(\calD_k=[y_k,y_{k+1}]\) then constructing an appropriate approximation to \(\overline{\mathbb B}\) on each interval.
+The schemes we choose to use here work by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \({\calD_k,i}\) then constructing an approximation to \(\overline{\mathbb B}\) on each interval.
 To help elucidate the connection between the operators \(\{ \overline{\mathbb  V}(t)\}\), \( \overline{\mathbb  B}\) and their approximation counterparts we take a slightly different approach to partitioning these operators than that taken in \cite{bo2014}.
@@
 By doing so, we can directly correspond elements of the partitioned operators to their approximation counterparts.
-Since the partition used to construct the schemes is finer, then we can always reconstruct the partition in terms of the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\).
-
-Let us first partition the space \([0,\infty)\) into \(\calD_{-1} = \{0\}\), \(\calD_{{K+1}}=\{b\}\), and non-trivial intervals \(\calD_k=[y_k,y_{k+1}]\setminus (\{0\}\cup\{b\}),\) with \(y_0=0,\,y_{K+1}=b,\, y_k<y_{k+1},\,k\in\mathcal K^\circ:=\{0,1,2,...,K\}\) and define \(\mathcal K = \{-1,K+1\}\cup\mathcal K^\circ\).
-For $\bs{\mu} \in \mathcal{M}_{0,b}(\mathcal{S} \times \mathbb{R}_{+})$ we write \(\bs{\mu} = (\mu_{k,i})_{i \in \mathcal{S},k\in\mathcal K},\) where $\mu_{k,i}(\cdot) = \mu_i(\cdot \cap \calD_k),\,k\in\mathcal K.$ When it exists, we denote by \(v_{k,i}(x),\,x>0\), the densities associated with each measure, \(\mu_{k,i},\,k\in \mathcal K^\circ\).
+Since the partition used to construct the approximation schemes is finer, then we can reconstruct the partition in terms of the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\).
+
+Let us first partition the space \([0,b]\) into \(\calD_{-1,i}=\calD_{-1} = \{0\}\), \(i\in\calS_{-1}\), \(\calD_{{K+1},i}=\calD_{{K+1}}=\{b\}\), \(i\in\calS_{K+1}\), and non-trivial intervals \(\calD_{k,i}=[y_k,y_{k+1})\setminus \{0\},\) \(i\in\calS_+\cup\calS_0\) and \(\calD_{k,i}=(y_k,y_{k+1}]\setminus \{b\}\), \(i\in\calS_-\), with \(y_0=0,\,y_{K+1}=b,\, y_k<y_{k+1},\,k\in\mathcal K^\circ:=\{0,1,2,...,K\}\) and define \(\mathcal K = \{-1,K+1\}\cup\mathcal K^\circ\).
+For $\bs{\mu} \in \mathcal{M}_{0,b}(\mathcal{S} \times [0,b])$ we write \(\bs{\mu} = (\mu_{k,i})_{i \in \mathcal{S},k\in\mathcal K},\) where $\mu_{k,i}(\cdot) = \mu_i(\cdot \cap \calD_{k,i}),\,k\in\mathcal K.$ When it exists, we denote by \(v_{k,i}(x),\,x>0\), the densities associated with each measure, \(\mu_{k,i},\,k\in \mathcal K^\circ\).
 For \( i,j\in\calS,\,k,\ell\in\mathcal K\) define the operators
-\[\mu_{k,i} \mathbb V_{ij}^{k \ell}(t)(\mathcal A) := \int_{x\in\calD_k} \wrt \mu_{k,i}(x)\mathbb P (X(t) \in\mathcal A\cap \calD_\ell,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
+\[\mu_{k,i} \mathbb V_{ij}^{k \ell}(t)(\mathcal{E}) := \int_{x\in\calD_{k,i}} \wrt \mu_{k,i}(x)\mathbb P (X(t) \in\mathcal{E}\cap \calD_{\ell,i},\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
 and the matrices of operators \(\mathbb V^{k\ell}(t) := \vligne{\mathbb V_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\) and write 
@@
 %\end{align*}
-where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\). 
-The tridiagonal structure arises since, for \(|k-\ell|\geq2\) (where we take \({-1} = 0\) and \({K+1} = K+1\) if they appear in the difference) it is impossible for \(\{X(t)\}\) to move from \(\mathcal D_k\) to \(\mathcal D_\ell\) in an infinitesimal amount of time.\footnote{We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
+where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\).\footnote{We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
 The operators denoted with an overline play a minor role in the introductory sections of this thesis, but do not appear again.
@@
 E.g.~\(\mathbb B_{ij}^{k\ell}\) means \(\mathbb B_{ij}^{k\ell},\, i,j\in\calS, k,\ell\in\mathcal K\) and \(\mathbb B_{ij}^{mn}\) means \(\mathbb B_{ij}^{mn},\, i,j\in\calS, m,n\in\{+,-,0\}\).
-}
+} The tridiagonal structure arises since, for \(|k-\ell|\geq2\) it is impossible for \(\{X(t)\}\) to move from \(\mathcal D_{k,i}\) to \(\mathcal D_{\ell,j}\) in an infinitesimal amount of time.
 
 By an appropriate choice of the intervals \(\{\mathcal D_k\},\, k\in\mathcal K\), the partition used in \cite{bo2014} can be recovered.
-Intuitively, we must ensure that each of the boundaries of \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), align with a boundary of a cell \(\mathcal D_k=[y_k,y_{k+1}]\setminus(\{0\}\cup\{b\})\).
+Intuitively, we must ensure that each of the boundaries of \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), align with a boundary of a cell \(\mathcal D_{k,i}\).
 Then, each set \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), can be written as a union of cells, \(\mathcal D_k,\,k\in\mathcal K\), sans a collection of points which have measure zero for all measures in \(\mathcal M_{0,b}\), and this collection of points is inconsequential for the purposes of the approximations presented here.
 
-Formally, to recover the partition used in \citep{bo2014} we choose the intervals \(\calD_k\) such that \(   l (\calD_k\cap\calF_i^m) \in \{   l (\calD_k), 0\}\) for all \(i\in\calS,\,m\in\{+,-,0\},\,k\in\mathcal K\), for all \( l \in\mathcal M_{0,b}\).
-That is, we choose \(\calD_k\) such that it is contained, up to sets of measure 0 with respect to measures in \(\mathcal M_{0,b}\), within one of the sets \(\calF_i^m\) for \(m\in\{+,-,0\}\) and \(i\in\calS\).
-We assume such a partition for the rest of the paper.
-For \(i\in\calS,\,m\in\{+,-,0\}\), let \(\mathcal K^m_i = \{k\in\mathcal K\mid  l (\calD_k\cap\calF_i^m) =   l (\calD_k),\,l\in\mathcal M_0 \}\), so that \(\bigcup\limits_{k\in\mathcal K_i^m} \calD_k\) and \(\mathcal F_i^m\) are equal up to a set of \(\mathcal M_{0,b}\)-measure 0.
+Formally, to recover the partition used in \citep{bo2014} we choose the intervals \(\calD_{k,i}\) such that \(   l (\calD_{k,i}\cap\calF_i^m) \in \{   l (\calD_{k,i}), 0\}\) for all \(i\in\calS,\,m\in\{+,-,0\},\,k\in\mathcal K\), for all \( l \in\mathcal M_{0,b}\).
+That is, we choose \(\calD_{k,i}\) such that it is contained, up to sets of measure 0 with respect to measures in \(\mathcal M_{0,b}\), within one of the sets \(\calF_i^m\) for \(m\in\{+,-,0\}\) and \(i\in\calS\).
+We assume such a partition for the rest of the thesis.
+For \(i\in\calS,\,m\in\{+,-,0\}\), let \(\mathcal K^m_i = \{k\in\mathcal K\mid  l (\calD_{k,i}\cap\calF_i^m) =   l (\calD_{k,i}),\,l\in\mathcal M_0 \}\), so that \(\bigcup\limits_{k\in\mathcal K_i^m} \calD_{k,i}\) and \(\mathcal F_i^m\) are equal up to a set of \(\mathcal M_{0,b}\)-measure 0.
 Define \(\mathcal K^m = \bigcup\limits_{i\in\calS}\mathcal K_i^m\), \(m\in\{+,-,0\}\).
@@
 
-Let \(\calS_k^+=\{i\in\calS\mid r_i(x)>0,\,\forall x \in\calD_k\}\), \(\calS_k^0=\{i\in\calS\mid r_i(x)=0,\,\forall x \in\calD_k\}\), \(\calS_k^-=\{i\in\calS\mid r_i(x)<0,\,\forall x \in\calD_k\}\) and \(\calS_k^\bullet=\{i\in\calS\mid r_i(x)\neq0,\,\forall x \in\calD_k\}\) for \(k\in\mathcal K\).
+Let \(\calS_k^+=\{i\in\calS\mid r_i(x)>0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^0=\{i\in\calS\mid r_i(x)=0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^-=\{i\in\calS\mid r_i(x)<0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^\bullet=\{i\in\calS\mid r_i(x)\neq0,\,\forall x \in\calD_{k,i}\}\) for \(k\in\mathcal K^\circ\) and \(\calS_{k}^+=\{i\in\calS_{k}\mid r_i(y_{k})>0\}\), \(\calS_{k}^-=\{i\in\calS_{k}\mid r_i(y_{k})<0\}\), \(\calS_{k}^0=\{i\in\calS_{k}\mid r_i(y_{k})=0\}\), \(\calS_{k}^\bullet=\{i\in\calS_{k}\mid r_i(y_{k})\neq 0\}\) for \(k\in\{-1,K+1\}\).
 For later reference, we need the following constructions.
-For \(k,\ell\in\mathcal K\)
+For \(k,\ell\in\mathcal K\) let
 \begin{align}
@@
 \end{align}
-for \(i,j\in\calS\) 
+for \(i,j\in\calS\) let 
 \begin{align}
@@
 %\end{align}
-and for \(m,n\in\{+,-,0\}\) 
+and for \(m,n\in\{+,-,0\}\) let
 \begin{align}
@@
 \end{align}
-We persist with the partition \(\calD_k,\,k\in\mathcal K\) throughout this thesis as this is consistent with the partition used in the finite-element approximation schemes considered in this thesis.
+We persist with the partition \(\calD_{k,i},\,k\in\mathcal K\) throughout this thesis as this is consistent with the partition used in the approximation schemes considered in this thesis.
 Note that for all the operators defined with this partition, the partitioning used in \citep{bo2014} can always be recovered by the above construction.
 
-\subsubsection{The details}
-We can write \(\mu_{k,i}\mathbb B_{ij}^{k\ell}(\mathcal A)\) in kernel form as \(\displaystyle\int_{x\in\calD_k,y\in\mathcal A}\wrt \mu_{k,i}(x)\mathbb B_{ij}^{k\ell}(x,dy)\). It is known that
-\[\mu_{k,i}\mathbb B_{ij}^{kk}(\wrt y):=\int_{x\in\calD_k}\wrt\mu_{k,i}(x)\mathbb B_{ij}^{kk}(x,\wrt y)=\begin{cases}
+\subsubsection{Definition of \(\mathbb B\)}
+When \(v_k\) is differentiable we can write \(\mu_{k,i}\mathbb B_{ij}^{k\ell}(\mathcal{E})\) in kernel form as \[\displaystyle\int_{x\in\calD_{k,i},y\in\mathcal{E}}\wrt \mu_{k,i}(x)\mathbb B_{ij}^{k\ell}(x,dy).\] It is known that
+\begin{align}\label{eqn: lfj22}\mu_{k,i}\mathbb B_{ij}^{kk}(\wrt y):=\int_{x\in\calD_{k,i}}\wrt\mu_{k,i}(x)\mathbb B_{ij}^{kk}(x,\wrt y)=\begin{cases}
 v_{k,i}(y)T_{ij}\wrt y, & i\neq j,\\
 v_{k,i}(y)T_{ii}\wrt y - c_i\cfrac{\wrt}{\wrt y}v_{k,i}(y)\wrt y, & i=j, 
-\end{cases}\]
-on the interior of \(\calD_k\), \(k\in\mathcal K^\circ\), \citep{kk1995}.
+\end{cases}\end{align}
+on the interior of \(\calD_{k,i}\), \(k\in\mathcal K^\circ\), \citep{kk1995}.
 Intuitively, \(v_{k,i}(y)T_{ij}\wrt y\) represents the instantaneous rate of transition from phase \(i\) to \(j\) in the infinitesimal interval \(\wrt y\), \(v_{k,i}(y)T_{ii}\wrt y\) represents no such transition occurring, and \(- c_i\cfrac{\wrt}{\wrt y}v_{k,i}(y)\wrt y\) represents the drift across the interval \(\wrt y\) when the phase is \(i\).
 
-Translating the results of \cite{bo2014} to use the partition \(\{\calD_k\}\) we may state that, for all \(i,j\in\calS\), \(k\in\{1,\dots,K-1\}\),
-\begin{align*}
-	\mu_{k,i}\mathbb B_{ij}^{kk}(\calD_k)&=\int_{x\in\calD_k}v_{k,i}(x)T_{ij}\wrt x,\\
-	\mu_{k,i}\mathbb B_{ii}^{kk}(\calD_k)&=\int_{x\in\calD_k}v_{k,i}(x)T_{ii}\wrt x - c_iv_{k,i}(y_{k+1})\mathbb 1_{(c_i>0)} + c_iv_{k,i}(y_{k})\mathbb 1_{(c_i<0)},
-\end{align*}
-where \(\mathbb 1_{}\) is the indicator function. 
-Intuitively, the first expression represents the instantaneous rate of the stochastic transitions of the phase process \(\{\varphi(t)\}\) while \(\{X(t)\}\) remains in \(\calD_k\).
-The first term in the second expression represents the net rate of transition out of phase \(i\) while \(\{X(t)\}\) remains in \(\calD_k\), while the second and third terms in the second expression represent the flux out of the right-hand edge of \(\calD_k\) when \(c_i>0\) and the flux out of the left-hand edge of \(\calD_k\) when \(c_i<0\), respectively.
+Translating the results of \cite{bo2014} to use the partition \(\{\calD_{k,i}\}\) we may state that, for all \(i,j\in\calS\), \(i\neq j\), \(k\in\{1,\dots,K-1\}\),
+\begin{align*}
+	\mu_{k,i}\mathbb B_{ij}^{kk}(\calD_{k,i})&=\int_{x\in\calD_k}v_{k,i}(x)T_{ij}\wrt x,\\
+	\mu_{k,i}\mathbb B_{ii}^{kk}(\calD_{k,i})&=\int_{x\in\calD_k}v_{k,i}(x)T_{ii}\wrt x - c_iv_{k,i}(y_{k+1}^-) 1{(c_i>0)} + c_iv_{k,i}(y_{k}^+) 1{(c_i<0)},
+\end{align*}
+where \( 1(\cdot)\) is the indicator function and \(x^+\) and \(x^-\) denote the right and left limits at \(x\). 
+Intuitively, the first expression represents the instantaneous rate of the stochastic transitions of the phase process \(\{\varphi(t)\}\) from \(i\) to \(j\) and \(\{X(t)\}\) remains in \(\calD_{k,i}\).
+The first term in the second expression represents the net rate of transition out of phase \(i\) while \(\{X(t)\}\) remains in \(\calD_{k,i}\), while the second and third terms in the second expression represent the flux out of the right-hand edge of \(\calD_{k,i}\) when \(c_i>0\) and the flux out of the left-hand edge of \(\calD_{k,i}\) when \(c_i<0\), respectively.
+Essentially, this is just a rewriting of (\ref{eqn: lfj22}) as an integral equation and including the boundary conditions.
 
@@
 \begin{align*}
-\mu_{k,i}\mathbb B_{ii}^{k,k+1}(\calD_{k+1})&= c_iv_{k,i}(y_{k+1})\mathbb 1_{(c_i>0)},\mbox{ for all \(i\in\calS\), \(k\in\{0,1,2,...,K-1\}\),}
-\\\mu_{k,i}\mathbb B_{ii}^{k,k-1}(\calD_{k-1})&= -c_iv_{k,i}(y_{k})\mathbb 1_{(c_i<0)}, \mbox{  for all \(i\in\calS\), \(k\in\{1,2,3,...,K\}\)}.
-\end{align*}
-Intuitively, the first equation represents the flux from \(\calD_k\) to \(\calD_{k+1}\) across the shared boundary at \(y_{k+1}\) which occurs when \(c_i>0\) only.
-The second expression represents the flux from \(\calD_k\) to \(\calD_{k-1}\) across the shared boundary at \(y_{k}\) which occurs when \(c_i<0\) only.
+\mu_{k,i}\mathbb B_{ii}^{k,k+1}(\calD_{k+1})&= c_iv_{k,i}(y_{k+1}^-) 1{(c_i>0)},\mbox{ for all \(i\in\calS\), \(k\in\{0,1,2,...,K-1\}\),}
+\\\mu_{k,i}\mathbb B_{ii}^{k,k-1}(\calD_{k-1})&= -c_iv_{k,i}(y_{k}^+) 1{(c_i<0)}, \mbox{  for all \(i\in\calS\), \(k\in\{1,2,3,...,K\}\)}.
+\end{align*}
+Intuitively, the first equation represents the flux from \(\calD_{k,i}\) to \(\calD_{k+1,i}\) across the shared boundary at \(y_{k+1}\) which occurs when \(c_i>0\) only.
+The second expression represents the flux from \(\calD_{k,i}\) to \(\calD_{k-1,i}\) across the shared boundary at \(y_{k}\) which occurs when \(c_i<0\) only.
 
@@
 	\end{align*}
-	In \(\calD_0\), if \(c_i<0\) and \(c_j>0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{-1}\) to phase \(j\),
-	\begin{align*}
-	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_0)&=\int_{x\in\calD_0}v_{0,i}(x)T_{ij}\wrt x - c_iv_{0,i}(0^+)p^{-1}_{ij}.
-	\end{align*}
-	Also in \(\calD_0\), if \(i\neq j\) and \(c_i\geq 0\) or \(i\neq j\), \(c_i< 0\) and \(c_j\leq 0\), the phase changes from \(i\) to \(j\) while \(\{X(t)\}\) remains in \(\calD_0\) at rate
-	\begin{align*}
-	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_0)&=\int_{x\in\calD_0}v_{0,i}(x)T_{ij}\wrt x,
-	\end{align*}
-	else, \(\{(X(t),\varphi(t))\}\) leaves \((\calD_0,i)\) at rate
-	\begin{align*}
-	\mu_{0,i}\mathbb B_{ii}^{0,0}(\calD_0)&=\int_{x\in\calD_0}v_{0,i}(x)T_{ii}\wrt x - c_iv_{0,i}(y_0)\mathbb 0_{(c_i> 0)} + c_iv_{0,i}(0^+)\mathbb 1_{(c_i< 0)},
-\end{align*}
-where \(0^+\) is the right limit at \(0\). Similarly, at the boundary \(x=b\), point mass moves between phases according to 
+	In \(\calD_{0,i}\), if \(c_i<0\) and \(c_j>0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{-1}\) to phase \(j\), hence
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_{0,i})&=\int_{x\in\calD_{0,i}}v_{0,i}(x)T_{ij}\wrt x - c_iv_{0,i}(0^+)p^{-1}_{ij}.
+	\end{align*}
+	Also in \(\calD_{0,i}\), if \(i\neq j\) and \(c_i\geq 0\) or \(i\neq j\), \(c_i< 0\) and \(c_j\leq 0\), the phase changes from \(i\) to \(j\) while \(\{X(t)\}\) remains in \(\calD_{0,i}\) at rate
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_{0,i})&=\int_{x\in\calD_{0,i}}v_{0,i}(x)T_{ij}\wrt x,
+	\end{align*}
+	else, \(\{(X(t),\varphi(t))\}\) leaves \((\calD_{0,i})\) at rate
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ii}^{0,0}(\calD_{0,i})&=\int_{x\in\calD_{0,i}}v_{0,i}(x)T_{ii}\wrt x - c_iv_{0,i}(y_1^-) 1{(c_i> 0)} + c_iv_{0,i}(0^+)1{(c_i< 0)}.
+\end{align*}
+
+Similarly, at the boundary \(x=b\), point mass moves between phases according to 
 \begin{align*}
@@
 	\end{align*}
-	In \(\calD_K\), if \(c_i>0\) and \(c_j<0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{K+1}\) to phase \(j\),
-	\begin{align*}
-	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_K)&=\int_{x\in\calD_K}v_{K,i}(x)T_{ij}\wrt x + c_iv_{K,i}(b^-)p^{K+1}_{ij},\,c_i>0,\,c_j<0.
-	\end{align*}
-	If \(i\neq j\) and \(c_i\leq 0\) or \(i\neq j\), \(c_i>0\) and \(c_j\geq 0\), the phase changes from \(i\) to \(j\) within \(\calD_K\) at rate
-	\begin{align*}
-	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_K)&=\int_{x\in\calD_K}v_{K,i}(x)T_{ij}\wrt x,
-	\end{align*}
-	else, \(\{(X(t),\varphi(t))\}\) leaves \((\calD_K,i)\) at rate
-	\begin{align*}
-	\mu_{K,i}\mathbb B_{ii}^{K,K}(\calD_K)&=\int_{x\in\calD_K}v_{K,i}(x)T_{ii}\wrt x + c_iv_{K,i}(y_K)\mathbb 1_{(c_i< 0)} - c_iv_{K,i}(b^-)\mathbb 1_{(c_i> 0)},
-\end{align*}
-where \(b^-\) is the left limit at \(b\). Otherwise, \(\mu_{k,i}\mathbb B_{ij}^{k\ell}=0\).
+	In \(\calD_{K,i}\), if \(c_i>0\) and \(c_j<0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{K+1}\) to phase \(j\),
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_{K,i})&=\int_{x\in\calD_{K,i}}v_{K,i}(x)T_{ij}\wrt x + c_iv_{K,i}(b^-)p^{K+1}_{ij},\,c_i>0,\,c_j<0.
+	\end{align*}
+	If \(i\neq j\) and \(c_i\leq 0\) or \(i\neq j\), \(c_i>0\) and \(c_j\geq 0\), the phase changes from \(i\) to \(j\) within \(\calD_{K,i}\) at rate
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_{K,i})&=\int_{x\in\calD_{K,i}}v_{K,i}(x)T_{ij}\wrt x,
+	\end{align*}
+	else, \(\{(X(t),\varphi(t))\}\) leaves \((\calD_{K,i},i)\) at rate
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ii}^{K,K}(\calD_{K,i})&=\int_{x\in\calD_{K,i}}v_{K,i}(x)T_{ii}\wrt x + c_iv_{K,i}(y_K^+)1{(c_i< 0)} - c_iv_{K,i}(b^-)1{(c_i> 0)}.
+\end{align*}
+Otherwise, \(\mu_{k,i}\mathbb B_{ij}^{k\ell}=0\).
 
@@
 Note that at the stopping time \(\eta(y)\) it must be that \((X({\eta(y)}),\varphi({\eta(y)}))\in(\calF_i^m,i)\) for some \(i\in\calS\) and \(m\in\{+,-\}\), i.e.~\(m\neq0\).
-We define the operators $\mathbb{U}_{ij}^{k\ell}(y,s): \mathcal{M}_{0,b}(\mathcal{D}_{k}\cap\calF_i^m) \mapsto \mathcal{M}_{0,b} (\calD_\ell\cap\mathcal{F}_j^n)$, for $k\in\mathcal K_i^+\cup\mathcal K_i^-$,  $\ell\in\mathcal K_j^+\cup\mathcal K_j^-$, and $i \in \mathcal{S}_k^\bullet,\,j \in \mathcal{S}_k^\bullet$, by
-% 
-	\begin{align*} 
-		&\mu_{k,i}\mathbb{U}_{ij}^{k\ell} (y,s) (\mathcal{A}) 
-		\\&:= \int_{x \in \calD_k} \wrt  \mu_{k,i}(x) \mathbb{E}\left[e^{-s\eta(y)}{1}\left\{X({\eta(y)}) \in \mathcal{A}\cap\calD_\ell,\;\varphi({\eta(y)}) = j\right\} \mid \varphi(0) = i, X(0) = x\right].
+We define the operators $\mathbb{U}_{ij}^{k\ell}(y,s): \mathcal{M}_{0,b}(\mathcal{D}_{k,i}\cap\calF_i^m) \mapsto \mathcal{M}_{0,b} (\calD_{\ell,j}\cap\mathcal{F}_j^n)$, for $k\in\mathcal K_i^+\cup\mathcal K_i^-$,  $\ell\in\mathcal K_j^+\cup\mathcal K_j^-$, and $i \in \mathcal{S}_k^\bullet,\,j \in \mathcal{S}_k^\bullet$, by
+% 
+	\begin{align*} 
+		&\mu_{k,i}\mathbb{U}_{ij}^{k\ell} (y,s) (\mathcal{E}) 
+		\\&:= \int_{x \in \calD_{k,i}} \wrt  \mu_{k,i}(x) \mathbb{E}\left[e^{-s\eta(y)}{1}\left\{X({\eta(y)}) \in \mathcal{E}\cap\calD_{\ell,j},\;\varphi({\eta(y)}) = j\right\} \mid \varphi(0) = i, X(0) = x\right].
 	\end{align*} 
@@
 	% 
-	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} := \diag(\mathbb{R}_i^{k})_{i \in \mathcal{S}}$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
-	\begin{align*} 
-		{\mu}_{k,i}\mathbb{R}_{k,i}(\mathcal{A}) := \int_{x \in \mathcal{A} \cap \mathcal{D}_k} \frac{1}{r_i(x)}\wrt  \mu_{k,i}(x),\quad k\in \mathcal K_i^+\cup\mathcal K_i^-.
+	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} := diag(\mathbb{R}_i^{k},\,i \in \mathcal{S})$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
+	\begin{align*} 
+		{\mu}_{k,i}\mathbb{R}_{k,i}(\mathcal{E}) := \int_{x \in \mathcal{E} \cap \mathcal{D}_{k,i}} \frac{1}{|r_i(x)|}\wrt  \mu_{k,i}(x),\quad k\in \mathcal K_i^+\cup\mathcal K_i^-.
 	\end{align*} 
@@
 \subsection{The first-return operator, $\mathbb\Psi(s)$}\label{sec: intro Psi}
-We denote by $\mathbb \Psi(s)$ the matrix of operators with the same dimensions as \(\mathbb D^{+-}\), recording the Laplace-Stieltjes transforms of the time for $\{\dot Y(t)\}$ to return, for the first time, to the initial level of zero as introduced in \cite{bo2014} but constructed with respect to the finer partition \(\{\calD_k\}\).
-Define the stopping time $\zeta_Y(E):= \inf \{t > 0: \dot Y(t) \in E\}$ to be the first time $\{\dot Y(t)\}$ hits the set $E$, then each component $\mathbb \Psi_{ij}^{k\ell}(s): \mathcal{M}_{0,b}(\mathcal D_k,i) \mapsto \mathcal{M}_{0,b}(\mathcal D_\ell,j), i,j \in \mathcal{S},\,k\in\mathcal K_i^+$ and $\ell \in \mathcal K_j^-$, is given by
-% 
-\begin{align*} 
-    % 
-	&\mu_{k,i}\mathbb\Psi_{ij}^{k\ell}(s) (\mathcal{A}) 
-	\\& := \int_{x \in \calD_k} \wrt \mu_{k,i}(x)
-	% 
-	 \mathbb{E}\big[e^{-s\zeta_Y(\{0\})}1\left\{X({\zeta_Y(\{0\})}) \in \mathcal{A}\cap\calD_\ell,\; \varphi({\zeta_Y(\{0\})}) = j \right\} \mid X(0) = x, 
+We denote by $\mathbb \Psi(s)$ the matrix of operators with the same dimensions as \(\mathbb D^{+-}\), recording the Laplace-Stieltjes transforms of the time for $\{\dot Y(t)\}$ to return, for the first time, to the initial level of zero as introduced in \cite{bo2014} but constructed with respect to the finer partition \(\{\calD_{k,i}\}\).
+Define the stopping time $\zeta_Y(E):= \inf \{t > 0: \dot Y(t) \in E\}$ to be the first time $\{\dot Y(t)\}$ hits the set $E$, then each component $\mathbb \Psi_{ij}^{k\ell}(s): \mathcal{M}_{0,b}(\mathcal D_{k,i}) \mapsto \mathcal{M}_{0,b}(\mathcal D_{\ell,j}), i,j \in \mathcal{S},\,k\in\mathcal K_i^+$ and $\ell \in \mathcal K_j^-$, is given by
+% 
+\begin{align*} 
+    % 
+	&\mu_{k,i}\mathbb\Psi_{ij}^{k\ell}(s) (\mathcal{E}) 
+	\\& := \int_{x \in \calD_{k,i}} \wrt \mu_{k,i}(x)
+	% 
+	 \mathbb{E}\big[e^{-s\zeta_Y(\{0\})}1\left(X({\zeta_Y(\{0\})}) \in \mathcal{E}\cap\calD_{\ell,j},\; \varphi({\zeta_Y(\{0\})}) = j \right) \mid X(0) = x, 
 	 \\&\qquad{} \dot Y(0) = 0, \varphi(0) = i\big].
@@
 	% 
-	Furthermore, if $s$ is real then $\mathbb\Psi(s)$ is the minimal nonnegative solution. 
+	Furthermore, if $s$ is real then $\mathbb\Psi(s)$ is the minimal non-negative solution. 
 \end{theo}
@@
 	\begin{align*}
-		\bbxi_{k,i}(\mathcal{A}) := \lim_{n \rightarrow \infty} \mathbb{P}\left(X(\zeta_Y^n(\{0\})) \in \mathcal{A}\cap\calD_k, \varphi(\zeta_Y^n(\{0\})) = i\right).
+		\bbxi_{k,i}(\mathcal{E}) := \lim_{n \rightarrow \infty} \mathbb{P}\left(X(\zeta_Y^n(\{0\})) \in \mathcal{E}\cap\calD_{k,i}, \varphi(\zeta_Y^n(\{0\})) = i\right).
 	\end{align*} 
@@
 	\begin{align*} 
-		\bbpi_{k,i}(y)(\mathcal{A}) = \bbpi_i(y)(\mathcal{A}\cap \mathcal D_k).
+		\bbpi_{k,i}(y)(\mathcal{E}) = \bbpi_i(y)(\mathcal{E}\cap \mathcal D_{k,i}).
 	\end{align*} 
@@
 	\( 
-		\mathbb p_{k,i}(\mathcal{A}) = \mathbb p_i(\mathcal{A}\cap \mathcal D_k).
+		\mathbb p_{k,i}(\mathcal{E}) = \mathbb p_i(\mathcal{E}\cap \mathcal D_{k,i}).
 	\) 
@@
 	\label{theo:density} 
-The density ${\bbpi}^{m}(y)$, for $m \in \{+,-,0\}$ and $y > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the following set of equations:
+The operator ${\bbpi}^{m}(y)$, for $m \in \{+,-,0\}$ and $y > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the following set of equations:
 % 	 
@@
 Therefore, there is the need for approximation schemes such as those presented in this thesis.
-We now proceed to some mathematical preliminaries required to construct our approximations.
-
-\section{Quasi-birth-and-death processes with rational arrival process components}
+
+\section{Quasi-birth-and-death processes with rational arrival process components}\label{qbd-rap intro}
 One of the approximation schemes we develop in this thesis is a quasi-birth-and-death process with rational-arrival-process components (QBD-RAP).
-The class of QBD-RAP processes are built from matrix-exponentially distributed inter-event times.
-We now introduce the class of matrix exponential distributions and recount some important properties before formally introducing the QBD-RAP.
-
-\subsection{Matrix-exponential distributions}
+QBD-RAP processes are built from matrix-exponentially distributed inter-event times.
+We now introduce the class of matrix exponential distributions and recount some important properties before introducing the QBD-RAP.
+
+\subsection{Matrix exponential distributions}
 Here we recount some facts about matrix exponential distributions.
 See \cite{MEinAP} for a more detailed exposition.
-A random variable, \(Z\), is said to have a matrix-exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
+A random variable, \(Z\), is said to have a matrix exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
 The density function of \(Z\) is given by \(f_Z(x) = \bs\alpha e^{\bs S x}\bs s\).
@@
 These condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\).
-However, in general there is no way to determine whether, given a triplet \((\bs \alpha, \bs S, \bs s)\) whether it is a representation of a matrix-exponential distribution, or not.
+However, in general there is no way to determine whether, a given triplet \((\bs \alpha, \bs S, \bs s)\) is a representation of a matrix exponential distribution, or not.
 Nonetheless, some properties of a triplet \((\bs \alpha, \bs S, \bs s)\) are known, such as the following, which is used in the characterisation of QBD-RAPs.
 \begin{thm}[Theorem 4.1.3, \cite{MEinAP}]
-	The density function of a matrix-exponential distribution with representation \((\bs\alpha, \bs S, \bs s)\) can be expressed in terms of real-valued constants as 
+	The density function of a matrix exponential distribution with representation \((\bs\alpha, \bs S, \bs s)\) can be expressed in terms of real-valued constants as 
 	\begin{align}
@@
 Here \(\mu_j,\, j=1,\dots,m_1\) are the real eigenvalues of \(\bs S\), while \(\eta_j+i\sigma_j, \, \eta_j-i\sigma_j,\, j=1,\dots,m_2\) denote its complex eigenvalues, which come in conjugate pairs.
-Thus \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m+1}p_j + 2\sum_{j=1}^{m_2}q_j\).
+Thus, \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m+1}p_j + 2\sum_{j=1}^{m_2}q_j\).
 \end{thm}
@@
 \begin{cor}[Corollary 4.1.5, \cite{MEinAP}]
-If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix-exponential distribution, then \(\bs S\) has a real dominating eigenvalues.
+If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix exponential distribution, then \(\bs S\) has a real dominating eigenvalues.
 \end{cor}
@@
 
-The class of matrix-exponential distribution is characterised as the class of probability distributions which have a rational Laplace transform.
+The class of matrix exponential distribution is characterised as the class of probability distributions which have a rational Laplace transform.
 That is, \(\displaystyle \int_{x=0}^\infty e^{-\lambda x}\bs \alpha e^{\bs S x}\bs s \wrt x\) is a ratio of two polynomial functions in \(\lambda\).
@@
 
-A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix-exponential distribution with this representation.
+A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix exponential distribution with this representation.
 The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\).
@@
 
-For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal A\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a (possibility defective) matrix exponential distribution.
-%By CITE it is known that \(\mathcal A\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).
+For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal{A}\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a (possibility defective) matrix exponential distribution.
+%By CITE it is known that \(\mathcal{E}\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).
 
@@
 
-To define a QBD-RAP we first need a Batch (Marked) RAP.
+To define a QBD-RAP we first need a Batch (Marked) rational-arrival-process (RAP).
 Let \(\mathscr K\subset \mathbb Z\) be a set of \emph{marks}.
@@
 For \(i\in\mathscr K\), let \(N_i\) be simple point processes associated with events with marks of type \(i\) only, and let \(\{N_i(t)\}_{t\geq 0}\) be the associated counting processes of events of mark \(i\).
-For a matrix \(\bs B\) let \(dev(\bs B)\) denote the real part of the dominant eigenvalue of \(\bs B\) (the one with maximal real part).
 Denote by \(f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n)\) the joint density, probability mass function of the first \(n\) inter-arrival times, \(Y_1,Y_2-Y_1,\dots,Y_n-Y_{n-1}\), and the associated marks \(M_n\).
@@
 
-Also from \cite{bn2010}, associated with a Marked RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\)
+Also, from \cite{bn2010}, associated with a Marked RAP is a row-vector-valued \emph{orbit} process, \(\{\bs A(t)\}_{t\geq0},\)
 \[\bs A(t) = \cfrac{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S (t-Y_{N(t)})}}{\bs \alpha\left( \prod\limits_{i=1}^{N(t)} e^{\bs S(Y_{i}-Y_{i-1})}\bs D_{M_i}\right)e^{\bs S( t-Y_{N(t)})}\bs e}.\]
@@
 where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
-The process \(\{\bs A(t)\}\) "jumps" at event times of \(N\) (the process may not always jump at these times, but typically the dynamics change discontinuously at this point).
-At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\).
+The process \(\{\bs A(t)\}\) `jumps' at event times of \(N\) (the process may not actually jump at these times, but we still call it a `jump' and typically the dynamics change at this point).
+At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\), i.e.~\(\mathbb P(N(t)=n,N(t+dt)=n+1)=\bs A(t) \bs D\bs e\wrt t\).
 Upon an event the event is associated with mark \(i\) with probability \(\bs A(t) \bs D_i\bs e/\bs A(t) \bs D\bs e\).
@@
 
-Intuitively, \(\bs A(t)\) encodes all of the information about the event times of the Marked RAP and associated marks up to time \(t\) that is needed to determine the future behaviour of the point process.
+Intuitively, \(\bs A(t)\) encodes all the information about the event times of the Marked RAP and associated marks up to time \(t\) that is needed to determine the future behaviour of the point process.
 Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\).
 Then \(N\mid \mathcal F_t \equiv N\mid \bs A(t) \sim BRAP(\bs A(t), \bs S, \bs D_i,i\in\mathscr K)\).
-In words, the future of the point process after time \(t\) given all of the information about the process up to and including time \(t\), is distributed as a Marked RAP with initial vector \(\bs A(t)\).
+In words, the future of the point process after time \(t\) given all the information about the process up to and including time \(t\), is distributed as a Marked RAP with initial vector \(\bs A(t)\).
 
 Now consider a Marked RAP, \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\{-1,0,+1\}).\) The process \(\{(L(t),\bs A(t))\}_{t\geq0}\) formed by letting \(L(t) = N_{+1}(t) - N_{-1}(t)\) is a QBD-RAP.
+QBDs are QBD-RAPs where the inter-event times are of Phase-type.
 
@@
 \section{Laplace Transforms}
-In Chapters~\ref{sec: conv} and \ref{ch: global results} we work with an object called the \emph{Laplace transform}.
-For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transforms of \(\mu\) to be
+In Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with an object called the \emph{Laplace transform}.
+For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transform of \(\mu\) to be
 \[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
-where the \emph{region on convergences} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
+where the \emph{region on convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
 \[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
@@
 We use the following convergence theorems to help us prove that the QBD-RAP scheme converges weakly to the distribution of the fluid queue.
-The first result we state is a sweeping statement about convergence of measures.
+The first result we state, the Portmanteau Theorem, is a sweeping statement about convergence of measures.
 First, let's define the notion of weak convergence.
@@
 Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra.
-Further, denote by \(C_b(E)\) the set of continuous bounded functions of \(E\).
+Further, denote by \(C_b(E)\) the set of continuous bounded functions on \(E\).
 Let \(\mu,\,\mu_1,\,\mu_2,...\in \mathcal M_f(E)\).
@@
 \[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
-
 We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}.
 Let \(\mathcal M_{\leq 1}(E):=\{\mu \in \mathcal M_f(E)\mid \mu(E)\leq 1\},\) the set of all sub-probability measures on \((E,\mathcal E)\).
-\begin{thm}[(Part of) The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
+\begin{thm}[[Part of] The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
 	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
@@
 
-In Chapters~\ref{sec: conv} and \ref{ch: global results} we use the Dominated Convergence Theorem to aid our convergence arguments.
-In applied probability we often want to prove convergence of certain probabilistic expressions but the expressions themselves are too difficult to characterise.
+In Chapters~\ref{sec: conv} and~\ref{ch: global results} we use the Dominated Convergence Theorem to aid our convergence arguments.
+In applied probability we often want to prove convergence of certain probabilistic expressions but the expressions themselves are too difficult to characterise as one.
 An approach which can simplify matters is to partition the expression on certain events where we have a simpler characterisation of the expression on each event, thereby enabling us to prove convergence on each element in the partition.
@@
 I have taken the following from Theorem~1.13~\cite{steinreal}.
-Further, \cite{steinreal} use the notation
+\cite{steinreal} use the notation
 \[\int f = \int f \wrt x = \int f \wrt m(x),\]
@@
 
-Also in Chapters~\ref{sec: conv} and \ref{ch: global results} we want to manipulate infinite sums or integrals and rearrange the order of summation or integration.
+Also in Chapters~\ref{sec: conv} and~\ref{ch: global results} we want to manipulate infinite sums or integrals and rearrange the order of summation or integration.
 However, things can go awry when we swap the order of integration/summation if we are not careful.
@@
 \begin{thm}[Tonelli's Theorem]\label{Tonelli}
-	Suppose \(f(x,y)\) a non-negative measurable function on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
+	Suppose \(f(x,y)\) is a non-negative measurable function on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
 	\begin{itemize}
@@
 
-Collectively, we refer to Theorems~\ref{Fubini} and \ref{Tonelli} together as the Fubini-Tonelli Theorem, but they are otherwise known collectively as just Fubini's Theorem.
+Collectively, we refer to Theorems~\ref{Fubini} and~\ref{Tonelli} together as the Fubini-Tonelli Theorem, but they are otherwise known collectively as just Fubini's Theorem.
 Often, they are used in conjunction.
@@
 
-\section{Other existing literature review and the context of this thesis} 
-We have already covered some technical concepts from the literature already.
-Here, we elucidate exactly what the contribution of this thesis is.
+\section{More on relevant literature \& the context of the thesis} 
+In the process of introducing the technical concepts we have already covered some the relevant literature already.
+Here, we provide some more context for, and elucidate exactly what, the contributions of this thesis.
 The focus of this thesis is on developing numerical approximations of fluid queues such that we may approximate the operator-analytic analysis of fluid-fluid queues form \cite{bo2014}.
@@
 
-Related to the analysis of fluid-fluid queues are the works of \cite{mz2012,lnp13,bo2013b,bop2020}.
-\cite{mz2012} analyse more general models that the fluid-fluid model and derive operator-analytic expressions for the stationary distribution of multidimensional Markov additive processes.
+\paragraph{Fluid-fluid queues and related models} Related to the analysis of fluid-fluid queues are the works of \cite{mz2012,lnp13,bo2013b} and \cite{bop2020}.
+\cite{mz2012} analyse related discrete-time multidimensional Markov additive processes and derive operator-analytic expressions for the stationary distribution.
 Although markedly different in their approach \cite{mz2012}, like  \cite{bo2014}, is inspired by a matrix-analytic approach.
-The work of \cite{lnp13} considers a fluid-fluid queue where the driving process \(\{(X(t),\varphi(t))\}\) is level-dependent, and derives the marginal probability distribution of the first level, and bounds for that of the second level.
-The analysis of \cite{bo2014} does not consider a level-dependent driving process such as this, but I believe that it would be relatively straight forward to extend the results to the level-dependent case where the rates \(c_i\) depend on \(X(t)\) in a peicewise-constant way.
-\cite{bo2013b} analyses a model which is simple special case of the fluid-fluid queue.
-They consider a model where the second level process \(\{Y(t)\}\) has constant rates which do not depend on \(\ddot X(t)\) (but do depend on \(\varphi(t)\)), while \(\{(\ddot X(t),\varphi(t))\}\) is an \emph{unbounded} fluid queue.
-They derive matrix-analytic expressions for certain Laplace transforms with respect to the position of the first fluid \(\{X(t)\}\) relative to its starting position \(X(0)\) of the probability of certain transient events (such as first return and draining/filling times) of \(\{Y(t)\}\).
-The yet-unpublished work of \cite{bop2020} treats a similar model to that of \cite{bo2013b}, except that the driving process has a regulated lower boundary at \(0\); this significantly complicated the analysis.
+The work of \cite{lnp13} considers a fluid-fluid queue where the driving process \(\{(X(t),\varphi(t))\}\) is level-dependent, and derives computable expressions for the marginal probability distribution of the first level, and bounds for that of the second level.
+Their analysis is somewhat specific to the data handling model considered and only arrives at bounds for the marginal limiting distribution of the second fluid level.
+In contrast, the work of \cite{bo2014} treats more general models and derives operator-analytic expressions for the joint stationary distribution.
+Therefore, compared to \cite{lnp13}, the methods of this thesis therefore apply to a general class of models and enable to approximation of the joint stationary distribution.
+However, unlike \cite{lnp13}, the analysis of \cite{bo2014} does not consider a level-dependent driving process such as this, but I believe that it would be relatively straight forward to extend their results (and the results of this thesis) to the level-dependent case where the rates \(c_i\) depend on \(X(t)\) in a piecewise-constant way.
+\cite{bo2013b} analyse a model which is simple special case of the fluid-fluid queue.
+They consider a model where the second level process \(\{\dot Y(t)\}\) has constant rates which do not depend on \(\ddot X(t)\) (but do depend on \(\varphi(t)\)), while \(\{(\ddot X(t),\varphi(t))\}\) is an \emph{unbounded} fluid queue.
+They derive matrix-analytic expressions for certain Laplace transforms, with respect to the position of the first fluid \(\{X(t)\}\) relative to its starting position \(X(0)\), of the probability of certain transient events (such as first return and draining/filling times) of \(\{Y(t)\}\).
+In contrast, \cite{bo2013b} together with this thesis, treat a more general model and together, give computational methods for various performance measures of fluid-fluid queues.
+The yet-unpublished work of \cite{bop2020} treats a similar model to that of \cite{bo2013b}, except that the driving process has a regulated lower boundary at \(0\); this significantly complicates the analysis.
 They derive matrix-analytic expressions for the first return operator of the second fluid given a specific exponential form of the initial distribution and a certain boundary condition is met.
-
-Returning now to \cite{bo2013} which introduces the fluid-fluid queue in its generality and derives operator-analytic expressions for the first-return operator of the second fluid and the stationary distribution.
+Once again, \cite{bo2013b} together with this thesis, treat a more general model and together, give computational methods for various performance measures of fluid-fluid queues.
+
+Returning now to \cite{bo2014} which introduces the fluid-fluid queue in its generality and derives operator-analytic expressions for the first-return operator of the second fluid and the joint stationary distribution.
 Their analysis is inspired by the analysis of classical fluid queues in \cite{bean2005} whereby matrix-analytic expressions for the fluid level are derived in terms of the generator of the driving process.
 The driving process of a fluid-fluid queue is a fluid queue.
-
-The analysis of fluid queues can be classified broadly into two approaches, matrix-analytic methods (for example, \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}) and differential equations based approaches \cite{anick1982,kk1995,blnos2022}.
+The analysis of \cite{bo2014} requires specific expressions about the transient distribution of the fluid queue which cannot, in general, be obtained from the existing literature.
+
+\paragraph{Transient analysis of fluid queues} The analysis of fluid queues can be classified broadly into two approaches, matrix-analytic methods (for example, \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}) and differential equations based approaches (such as \cite{anick1982,kk1995,blnos2022}).
 Often, it is the stationary distribution of the fluid queue which is of interest.
@@
 None of the existing literature gives exact expressions which have all the aforementioned properties.
-Even if we consider a simpler model where the rates \(r_i(x)\) are constant on intervals \(\calD_k\), then the existing results are not satisfactory.
-Perhaps the closest is the work of \cite{bean2009} who compute expressions for the Laplace transform with respect to the time take for the fluid to exit an interval, \(\calD_k\), say, on the even that fluid exits in some phase \(j\), given it started within the interval at some point \(x_0\) in some phase \(i\).
-These expressions could perhaps be used to analyse a fluid-fluid queue where the rates \(r_i(x)=r_k\) for all \(i\in\calS\), and \(x\in\calD_k\).
+Even if we consider a simpler model where the rates \(r_i(x)\) are constant on intervals \(\calD_{k,i}\), then the existing results are not satisfactory.
+Perhaps the closest is the work of \cite{bean2009} who compute expressions for the Laplace transform with respect to the time take for the fluid to exit an interval, \(\calD_{k,i}\), say, on the even that fluid exits in some phase \(j\), given it started within the interval at some point \(x_0\) and in some phase \(i\).
+These expressions could perhaps be used to analyse a fluid-fluid queue where the rates \(r_i(x)=r_k\) for all \(i\in\calS\), and \(x\in\calD_{k,i}\).
 Even with this simplified fluid-fluid queue, it is unclear how we could use these expressions to compute the position of the second fluid.
@@
 An approximate method which is appropriate for the analysis of fluid-fluid queues has been proposed.
-The method \citep{bo2013}, which we refer to as the uniformisation method as it is derived via uniformisation arguments, approximates a fluid queue by a discrete-time Markov chain.
+The method of \cite{bo2013}, which we refer to as the uniformisation method as it is derived via uniformisation arguments, approximates a fluid queue by a discrete-time Markov chain.
 \cite{bo2013} approximate the fluid queue \(\{(\dot X(t),\varphi(t))\}\) by the quasi-birth-and-death-process \(\{(L(t),\varphi(t))\}\).
-In both processes, \(\{\varphi(t)\}\) is the same.
+In both processes, \(\{\varphi(t)\}\) is the same so there is no approximation in phase process.
 The level process \(\{L(t)\}\) approximates the level process \(\{\dot X(t)\}\).
-Specifically, \cite{bo2013} derive the level process such that the event \(L(t)=k\) approximates the event \(\dot X(t)\in \calD_k\), where all the levels \(\{\calD_k\}_{k}\) have the same width, \(\Delta\).
-By taking \(\Delta\to 0\), they prove that the quasi-birth-and-death-process \(\{(L(t),\varphi(t))\}\) converges to the fluid queue \(\{(\dot X(t),\varphi(t))\}\).
+Specifically, \cite{bo2013} derive the level process such that the event that \(L(t)=k\) and \(\varphi(t)\) approximates the event \(\dot X(t)\in \calD_{k,i}\) and \(\varphi(t)=i\), where all the levels \(\{\calD_{k,i}\}_{k}\) have the same width, \(\Delta=y_{k+1}-y_k\).
+By considering arbitrarily small \(\Delta\), they prove that the distribution of the quasi-birth-and-death-process \(\{(L(t),\varphi(t))\}\) can be made arbitrarily close to the distribution of the fluid queue \(\{(\dot X(t),\varphi(t))\}\).
 Replacing the driving process of the fluid-fluid queue, \(\{(\dot X(t),\varphi(t))\}\), by the approximation \(\{(L(t),\varphi(t))\}\) and approximating the rate functions \(r_i(x)\) by constants on each interval \(\calD_k\) yields a classical fluid queue which, intuitively, approximates the fluid-fluid queue.
-To date, the uniformisation approximation has not been applied in the context of fluid-fluid queues, and its ability to approximate transient quantities of fluid queues has not been well studied.
+\cite{bo2013} numerically investigate the ability of their approximation to approximate \(\bs \Psi_X\).
+To date, the uniformisation approximation has not been applied in the context of fluid-fluid queues, and its ability to approximate transient quantities of fluid queues (other than the aforementioned investigation of \(\bs\Psi_X\)) has not been well studied.
 In Chapter~\ref{sec: numerics} we investigate the numerical performance of the uniformisation approximation and show that it is effective.
@@
 Though we should not rush to discount the uniformisation approximation.
-Due to its stochastic interpretation, approximations to probabilities are guaranteed to be non-negative, we can employ probabilistic techniques to analyse the approximation (as \cite{bo2013} did), and is simplicity lends itself more easily to analysis.
-Further, when we use the uniformisation technique to approximate a fluid-fluid queue the resultant approximation is itself a fluid queue.
-We can leverage the stochastic interpretation of the approximation of the fluid-fluid queue to help analyse this approximation.
+Due to the stochastic interpretation of the uniformisation approximation as a QBD, approximations to probabilities are guaranteed to be non-negative, and we can employ probabilistic techniques to analyse the approximation (as \cite{bo2013} did).
+Moreover, its simplicity lends itself more easily to mathematical analysis.
+Further, when the uniformisation scheme is used in the context of approximating a fluid-fluid queue the resultant approximation is itself a fluid queue, and this stochastic interpretation could aid in the analysis of this approximation.
 
@@
 We introduce the discontinuous Galerkin method and its application to fluid queues, and fluid-fluid queues, in Chapter~\ref{ch:galerkin}.
-DG approximation schemes have found great success for approximating solutions of partial differential equations, however, to my knowledge, they have not been applied to fluid queues, or fluid-fluid queue to date (except for our paper \cite{blnos2022}, of which I am a co-author and the relevant parts of that work are included in this thesis, in Section~\ref{appendix: sec: 2} and Chapter~\ref{ch:galerkin}).
-The DG scheme has the advantage that is rate of convergence is rapid for smooth problems, however, it can produce oscillatory approximations and result in negative probabilities.
+DG approximation schemes have found great success for approximating solutions of partial differential equations, however, to my knowledge, they have not been applied to fluid queues, or fluid-fluid queue to date (except for our paper \cite{blnos2022}, of which I am a co-author and the relevant parts of that work are included in this thesis).
+The DG scheme has the advantage that its rate of convergence is rapid for smooth problems, however, it can produce oscillatory approximations and result in negative probabilities.
 

### 2022-02-17 chapter2/chapter.tex
first draft of the whole thing! :D

@@
 %!TEX root = ../thesis.tex
-\chapter{Approximating fluid queues with the discontinuous Galerkin method \label{ch:galerkin}}
+\chapter{Approximating fluid queues with the discontinuous Galerkin method\label{ch:galerkin}}
 
-In this chapter we introduce the discontinuous Galerkin (DG) method as applied to fluid queues and to approximate the operator-analytic expressions for fluid-fluid queues in \cite{bo2014}. 
+In this chapter we introduce the discontinuous Galerkin (DG) method applied to fluid queues to approximate the operator-analytic expressions for fluid-fluid queues in \cite{bo2014}. 
 \begin{center}
 	\begin{minipage}{0.8\textwidth}
-		\textit{Apart from Section~\ref{sec: slope limiting}, this chapter has been taken from Section~4 and Appendix~1 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis.
+		\textit{Apart from Sections~\ref{sec: other applications}-\ref{sec: slope limiting}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis.
 I am a co-author of the paper \cite{blnos2022}.
@@
 Discontinuous Galerkin (DG) methods can be used to approximate the solutions to systems of partial differential equations (PDEs).
-For a more thorough description of these methods see \citep{nodalDGBook}.
-The domain of approximation is partitioned into intervals, referred to individually as \textit{cells} and collectively as a \textit{mesh}.
-On each cell, we have a finite element approximation, which constructs a finite-dimensional smooth Sobolev space using piecewise-polynomial basis functions, and then projects the partial differential equations onto this space.
-This projection leads to a new system of equations, referred to as the \textit{weak form} of the original system of PDEs.
-Next, we must approximate the \textit{flux operator} which moves probability from one cell to another, in a manner similar to the underlying principle of a finite-volume approximation.
-This method conserves probability, and can handle discontinuities, such as jumps and point masses.
-Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the stationary distribution of an SFFM.
+For a most readable and practical text on DG methods we refer the reader to \cite{nodalDGBook}.
+Here we describe how one can apply the DG method to approximate fluid queues and fluid-fluid queues.
+In finite-element methods, such as the DG method, the domain of approximation is partitioned into intervals, referred to individually as \textit{cells} and collectively as a \textit{mesh}.
+On each cell we construct a finite-dimensional smooth Sobolev space using piecewise-polynomial basis functions, and then project the partial differential equations onto this space.
+This projection leads to a new system of equations which represent a \textit{weak form} of the original system of PDEs.
+Next, we approximate the \textit{flux operator} which moves probability from one cell to another, in a manner similar to the underlying principle of a finite-volume approximation.
+This creates a system of ordinary differential equations which is known as a \emph{semi-discrete system} (discrete in the spatial variable but continuous in time).
+We can solve the semi-discrete system to approximate transient distributions of the fluid queue or, as we do in Section~\ref{sec:DGSFFM}, we can extract and manipulate an approximation to the generator of the fluid queue to approximate performance measures of fluid-fluid queues.
+The DG method conserves probability.
+Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the stationary distribution of a stochastic fluid-fluid queue.
 
-\section{The Partial Differential Equation}
+\section{The partial differential equation}
 We start by introducing the PDE from which we will extract the approximation to the generator \(\mathbb B\).
 
-Let $f_i(x,t)$ be the joint density of $\{(X_t, \varphi_t)\}$: 
+Let $f_i(x,t)$ be the joint density of $(X_t, \varphi_t)$: 
 	% 
 	\begin{align*} 
-	    f_i(x,t) := \cfrac{\partial}{\partial x} 	\mathbb{P}\left(X_t \leq x, \varphi_t = i\right),\quad 0<x<b,\, i \in\calS,
+	    f_i(x,t) := \cfrac{\partial}{\partial x} 	\mathbb{P}\left(X(t) \leq x, \varphi(t) = i\right),\quad 0<x<b,\, i \in\calS,
 	\end{align*} 
@@
 \begin{align}
-\cfrac{\partial}{\partial t} \boldsymbol f(x,t) &= \boldsymbol f(x,t)\bs T -  \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\bs C\label{eqn:pde_density_matrix}, 
+\cfrac{\partial}{\partial t} \boldsymbol f(x,t) &= \boldsymbol f(x,t)\bs T -  \cfrac{\partial}{\partial x}\boldsymbol f(x,t)\widehat{\bs C}\label{eqn:pde_density_matrix}, 
 \end{align}
@@
 % on the interior of the space \([0,\infty)\).
-Thus, by approximating the operator on the right-hand side of Equation (\ref{eqn:pde_density_matrix}) we can approximate the infinitesimal operator \(\mathbb B\).
+Thus, by approximating the operator on the right-hand side of Equation~(\ref{eqn:pde_density_matrix}) we can approximate the infinitesimal operator \(\mathbb B\).
 The DG method does exactly this, by approximating the operator with a matrix.
 
-\section{Cells, Test Functions, and Weak Formulation}
+\section{Cells, test functions, and weak formulation}
 To begin with, consider an unbounded first fluid level \(\{\ddot X_t,t\geq0\}\), \(\ddot X_t\in(-\infty,\infty)\).
 We will eventually truncate this space so that we have a finite dimensional approximation; however, this requires a discussion on boundary conditions which we save for later.
-Let \(\mathcal D_k = [y_k,y_{k+1}],\, k\in\mathbb Z\) partition the domain \((-\infty,\infty)\).
-We call the \(\calD_k\) \textit{cells}.
+Let \(\mathcal D_{k,i} = [y_k,y_{k+1}),\, k\in\mathbb Z\) for \(i\in\calS_+\cup\calS_0\) and \(\mathcal D_{k,i} = (y_k,y_{k+1}],\, k\in\mathbb Z\) for \(i\in\calS_-\), \(y_k<y_{k+1}\), partition the domain \((-\infty,\infty)\).
+We call the \(\calD_{k,i}\) \textit{cells} and define \(\Delta_k=y_{k+1}-y_k\).
 
-On each cell \(\calD_k\) we choose \(N_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{N_k}\), compactly supported on \(\calD_k\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_k\)) to form a basis for the space \(W_k\), in which we formulate the approximation.
+For each \(k\) we choose \(N_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{N_k}\), compactly supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) to form a basis for the space \(W_k\), in which we formulate the approximation.
 Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{N_k}\) to be the space of polynomials of degree \(N_k-1\).
@@
 For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{N_k}\) as the Lagrange polynomials.
-On each cell \(\mathcal D_k\) we approximate
+On each cell \(\mathcal D_{k,i}\) we approximate
 \[f_i(x,t)\approx u_{k,i}(x,t)=\sum\limits_{r=1}^{N_k}a_{k,i}^r(t)\phi^r_k(x),\] 
 where \(a_{k,i}^r(t)\) are yet-to-be-determined time-dependent coefficients.
-We refer to \(u_{k,i}\) as the \textit{local} approximation on \(\calD_k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_{k,i}\) on the whole domain.
+We refer to \(u_{k,i}\) as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_{k,i}\) on the whole domain.
 The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\).
@@
 
-The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation (\ref{eqn:pde_density_matrix}).
+The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation~(\ref{eqn:pde_density_matrix}).
 In general, to construct the weak-form we need a set of test functions, say \(W\).
@@
 \begin{align*}
-	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j(x,t)\phi^r_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i(x,t)T_{ij}\phi^r_k(x)\wrt x  
+	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS}  f_i(x,t)T_{ij}\phi^r_k(x)\wrt x  
 	%
-	+  \int_{x\in\calD_k} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x \\ {}- [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
+	+  \int_{x=y_k}^{y_{k+1}} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x \\ {}- [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
 \end{align*}
-since \(\phi^r_k\) is compactly supported on \(\calD_k\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\).
+since \(\phi^r_k\) is compactly supported on \(\calD_{k,j}\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\).
 Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi^r_k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\).
@@
 \begin{align*}
-	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi^r_k(x)\wrt x 
-	= \int_{x\in\calD_k}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi^r_k(x)\wrt x  
+	\begin{multlined}[t]\int_{x=y_{k}}^{y_{k+1}}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi^r_k(x)\wrt x 
+	= \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi^r_k(x)\wrt x  
 	%
 	\\ {}
-	+  \int_{x\in\calD_k} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
+	+  \int_{x=y_k}^{y_{k+1}} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
 \end{align*}
@@
 \begin{align}
-	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi^r_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi^r_k(x)\wrt x  
+	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi^r_k(x)\wrt x  
 	%
-	+  \int_{x\in\calD_k} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x \\{} - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}. \end{multlined} \label{eqn:hash}
+	\\{} + \int_{x=y_k}^{y_{k+1}} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}. \end{multlined} \label{eqn:hash}
 \end{align}
-Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_k\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), which we now substitute into (\ref{eqn:hash})
+Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), which we now substitute into (\ref{eqn:hash})
 %\begin{align*}
@@
 \begin{align}
-	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi_k(x)\tr{}T_{ij}\boldsymbol\phi_k (x)\wrt x
+	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi_k(x)\tr{}T_{ij}\boldsymbol\phi_k (x)\wrt x
 	%
-	\\{}+  \int_{x\in\calD_k}\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\boldsymbol\phi_k(x)\wrt x - c_j[f_j(x,t) \boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}},\quad k\in\mathbb Z.
+	\\{}+  \int_{x=y_k}^{y_{k+1}}\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\boldsymbol\phi_k(x)\wrt x - c_j[f_j(x,t) \boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}},\quad k\in\mathbb Z.
 \end{multlined} \label{eqn:DG system}
@@
 
-\section{Mass, Stiffness, and Flux}
+\section{Mass, stiffness, and flux matrices}
 For \(k\in\mathbb Z\), define local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
@@
 It remains to approximate the \textit{flux}, \(f_j(x,t)\) at the cell edges \(y_k,\,k\in\mathbb Z\), so that we may evaluate the terms \([f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}\), \(r=1,...,N_k,\,k\in\mathbb Z\).
-This is the key for DG -- it joins the local approximations on each cell \(\calD_k\), into a global approximation on the whole domain of approximation.
+This is the key for DG -- it joins the local approximations on each cell into a global approximation on the whole domain of approximation.
 The flux is the instantaneous rate (with respect to time) at which density moves across the boundaries \(y_k,\,k\in\mathbb Z\).
@@
 \end{align*}
-where, for \(j\in \calS\) with \(c_j>0\), we define \(\bs F_j^{k,k} := -\boldsymbol \phi_k(y_{k+1}^-)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z\) and \(\bs F_j^{k-1,k} := \boldsymbol \phi_{k-1}(y_{k}^-)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
+where, for \(j\in \calS\) with \(c_j>0\), we define \(\bs F_+^{k,k}=\bs F_j^{k,k} := -\boldsymbol \phi_k(y_{k+1}^-)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z\) and \(\bs F_j^{k-1,k} := \boldsymbol \phi_{k-1}(y_{k}^-)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
 
@@
 \end{align*}
-where, for \(j\in \calS\) with \(c_j<0\), we define \(\bs F_j^{k+1,k} := -\boldsymbol \phi_{k+1}(y_{k+1}^+)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z,\) and \(\bs F_j^{k,k} := \boldsymbol \phi_{k}(y_{k}^+)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
+where, for \(j\in \calS\) with \(c_j<0\), we define \(\bs F_j^{k+1,k} := -\boldsymbol \phi_{k+1}(y_{k+1}^+)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z,\) and \(\bs F_-^{k,k}=\bs F_j^{k,k} := \boldsymbol \phi_{k}(y_{k}^+)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
 
 The matrices \(\bs F_j^{k-1,k},\,\bs F_j^{k,k},\) and \(\bs F_j^{k+1,k}\) are the local flux matrices.
-For convenience, we also define the matrices \(\bs F_j^{k,k+1}=0\) for \(c_j<0\) and \(\bs F_j^{k,k-1}=0\) for \(c_j>0\), \(k\in\mathbb Z\).
+For convenience, we also define the matrices \(\bs F_j^{k,k+1}=0\) for \(j\in\calS_-\) and \(\bs F_j^{k,k-1}=0\) for \(j\in\calS_+\), \(k\in\mathbb Z\).
 
@@
 \[\boldsymbol a_k(t) = (\boldsymbol a_{k,i}(t))_{i\in\mathcal S},\quad \boldsymbol a(t) = (\boldsymbol a_k(t))_{k\in\mathbb Z},\]
-and the block-tridiagonal matrix 
+and the double-infinite block-tridiagonal matrix 
 \begin{align*}
@@
 Let $[0,b]$ be the domain of the approximation, where $b < \infty$.
-We partition the space $[0,b]$ into \(\mathcal D_{-1}=\{0\},\) \(\mathcal D_{K+1}=\{b\},\) and \(K\) non-trivial intervals, \(\calD_k=[y_k,y_{k+1}]\setminus \{\{0\}\cup\{b\}\},\, y_k<y_{k+1},\, k\in\mathcal K^\circ\), \(y_0=0,\,y_{K+1}=b\) and define \(\Delta_k := y_{k+1}-y_k\).
+We partition the space $[0,b]$ into \(\mathcal D_{-1}=\{0\},\) \(\mathcal D_{K+1}=\{b\},\) and \(K\) non-trivial intervals, \(\calD_{k,i}=[y_k,y_{k+1})\setminus \{0\},\,i\in\calS_+\cup\calS_0,\calD_{k,i}=(y_k,y_{k+1}]\setminus \{0\},\,i\in\calS_-,\, y_k<y_{k+1},\, k\in\mathcal K^\circ\), \(y_0=0,\,y_{K+1}=b\) and define \(\Delta_k := y_{k+1}-y_k\).
 
@@
 
-Let us first consider the boundary at \( X_t=0\). The following boundary conditions describe the evolution of probability/density of a stochastic fluid model with a boundary at \(0\);
+Let us first consider the boundary at \( X_t=0\). The following boundary conditions describe the evolution of probability/density of a stochastic fluid queue with a boundary at \(0\);
 \begin{align}\label{eqn:BC1}
-\cfrac{\wrt}{\wrt t}\bs q_{-1}(t) &= \bs q_{-1}(t) T_{{-1} {-1}} - \boldsymbol f_-(0,t)\bs C_-\bs P_{-{-1}},
+\cfrac{\wrt}{\wrt t}\bs q_{-1}(t) &= \bs q_{-1}(t) \bs T_{{-1},{-1}} + \boldsymbol f_-(0,t)\bs C_-\bs P_{-,{-1}},
 \\\label{eqn:BC2}
-\bs q_{-1}(t)T_{{-1}+}&=\bs f_+(0,t)C_+,
+\bs f_-(0^+,t)\bs C_-\bs P_{-,+} + \bs q_{-1}(t)\bs T_{{-1},+}&=\bs f_+(0,t)\bs C_+,
 \end{align}
-where \(\bs P_{-{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\).
-Equation (\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(T_{{-1},{-1}}\), and that the flux of probability density into the point masses is \(- \boldsymbol f_-(0,t)\bs P_{-{-1}}C_{-1}\).
+where \(\bs P_{-,+} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{+}}\) and \(\bs P_{-,{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\).
+Equation~(\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(\bs T_{{-1},{-1}}\), and that the flux of probability density into the point masses is \(\boldsymbol f_-(0,t)\bs P_{-,{-1}}\bs C_{-1}\).
 Substituting the DG approximation for \(\boldsymbol f_-(0,t)\) into (\ref{eqn:BC1}) gives, for \(j\in\calS_{-1}\),
 \begin{equation*}\label{eqn:DGBC1}
-\cfrac{\wrt}{\wrt t} q_{{-1},j}(t) = \sum_{i\in\calS_{-1}} q_{{-1},i}(t) T_{i j} - \sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0)\tr{}p_{ij}^{-1} c_i.
+\cfrac{\wrt}{\wrt t} q_{{-1},j}(t) = \sum_{i\in\calS_{-1}} q_{{-1},i}(t) T_{i j} - \sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0^+)\tr{}p_{ij}^{-1} c_i.
 \end{equation*}
-Equation (\ref{eqn:BC2}) describes the flux of probability mass to density upon a transition from a phase in \(\calS_{-1}\) to a phase in \(\calS_+\).
-Thus, the flux into the left-hand edge of \(\calD_0\) in phase \(j\in\calS_+\) is, \(\sum\limits_{i\in\calS_{-1}} q_{{-1},i}(t)T_{ij}\).
+Equation~(\ref{eqn:BC2}) describes the reflection of process at \(0\), plus the flux of probability mass to density upon a transition from a phase in \(\calS_{-1}\) to a phase in \(\calS_+\).
+Thus, the flux into the left-hand edge of \(\calD_{0,j}\) in phase \(j\in\calS_+\) is, \(\sum\limits_{i\in\calS_{-1}} q_{{-1},i}(t)T_{ij}\).
 Therefore, we can now evaluate
 \begin{align*}
-	-c_j[f_j(x,t)\boldsymbol\phi_0(x)]_{x=0}^{x=y_{0}} & =  -c_jf_j(y_0,t)\boldsymbol\phi_0(y_0)+c_jf_j(0,t)\boldsymbol\phi_0(0)
-	\\& \approx -c_j(f_j^*(y_0,t)\boldsymbol\phi_0(y_0)+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0)
-	\\& = c_j\boldsymbol a_{0,j}(t)F_j^{0,0}+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0),
+	&-c_j[f_j(x,t)\boldsymbol\phi_0(x)]_{x=0}^{x=y_{0}} 
+	\\& =  -c_jf_j(y_0^-,t)\boldsymbol\phi_0(y_0)+c_jf_j(0^+,t)\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_if_i(0^+,t)p_{ij}^{-1}\bs \phi_0(0)
+	\\& \approx -c_j f_j^*(y_0,t)\boldsymbol\phi_0(y_0)+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs \phi_0(0)\tr{}\bs \phi_0(0)
+	\\& = c_j\boldsymbol a_{0,j}(t)\bs F_j^{0,0}+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs F_j^{0,0},
 \end{align*}
 for \(j \in \calS_+\).
-Thus, the DG approximation of the flux into point masses \(q_{{-1},j}(t)\) is \[-\sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0)\tr{} p_{ij}^{-1} c_i,\,j\in\calS_-,\] the rate of transition of point mass within \(\bs q_{{-1}}(t)\) is \(T_{{-1},{-1}}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0),\,j\in\calS_+\).
+
+Thus, the DG approximation of the flux into point masses \(q_{{-1},j}(t)\) is \[-\sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0)\tr{} p_{ij}^{-1} c_i,\,j\in\calS_-,\] the rate of transition of point mass within \(\bs q_{{-1}}(t)\) is \(\bs T_{{-1},{-1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0),\,j\in\calS_+\), and the DG approximation to density reflected at the lower boundary is \(\sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs F_j^{0,0}\).
 
@@
 \begin{align*}
-\cfrac{\wrt}{\wrt t}\bs q_{K+1}(t) &= \bs q_{K+1}(t) T_{{K+1} {K+1}} + \boldsymbol f_+(b,t)\bs C_+\bs P_{+{K+1}},\\
-\bs q_{K+1}(t)T_{{K+1}-}&=-\bs f_-(b,t)C_-.
+\cfrac{\wrt}{\wrt t}\bs q_{K+1}(t) &= \bs q_{K+1}(t) T_{{K+1}, {K+1}} + \boldsymbol f_+(b,t)\bs C_+\bs P_{+,{K+1}},\\
+\bs f_+(b^-,t)\bs C_+\bs P_{+,-} + \bs q_{K+1}(t)T_{{K+1}-}&=\bs f_-(b,t)\bs C_-,
 \end{align*}
-Using the same arguments as above, 
+where \(\bs P_{+,-} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{-}}\) and \(\bs P_{+,K+1} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{K+1}}\). Using the same arguments as above, 
 \begin{align*}
 \cfrac{\wrt}{\wrt t} q_{{K+1},j}(t) &= \sum_{i\in\calS_{K+1}}q_{{K+1},i}(t) T_{ij} + \sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(b)\tr{}p_{ij}^{K+1} c_i,
-\\-c_j[f_j(x,t)\boldsymbol\phi_K(x)]_{x=y_K}^{x=b} & \approx c_j\boldsymbol a_{K,j}(t)F_j^{K,K}+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b),
+\\-c_j[f_j(x,t)\boldsymbol\phi_K(x)]_{x=y_K}^{x=b} & \approx c_j\boldsymbol a_{K,j}(t)F_j^{K,K}+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b) 
+\\&\quad{} - \sum_{i\in\calS_+}c_ip_{ij}^{K+1}\bs a_{0,i}(t)\bs F_j^{0,0}
 \end{align*}
 for \(j\in\calS_-\).
-Thus, the DG approximation of the flux into the point mass \(q_{{K+1},j}(t)\) is \(\sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(0)\tr{} p_{ij}^{K+1} c_i\), \(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{{K+1}}(t)\) is \(T_{{K+1},{K+1}}\), and the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b)\), \(j\in\calS_-\).
 
-To include this behaviour in the DG generator we truncate the doubly-infinite matrix \(\ddot{{\bs B}}\) at \(k=0\) and \(k=K\), then append \(|\mathcal S_{-1}|\) rows and columns to the top and left, and \(|\mathcal S_{K+1}|\) rows and columns to the bottom and right.
+Thus, the DG approximation of the flux into the point mass \(q_{{K+1},j}(t)\) is 
+\[\sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(0)\tr{} p_{ij}^{K+1} c_i,\]
+\(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{{K+1}}(t)\) is \(\bs T_{{K+1},{K+1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b)\), \(j\in\calS_-\) and the DG approximation to reflection of density at the upper boundary is \(- \sum_{i\in\calS_+}c_ip_{ij}^{K+1}\bs a_{0,i}(t)\bs F_j^{0,0}\).
+
+To include the behaviour in the DG generator we truncate the doubly-infinite matrix \(\ddot{{\bs B}}\) at \(k=0\) and \(k=K\), then append \(|\mathcal S_{-1}|\) rows and columns to the top and left, and \(|\mathcal S_{K+1}|\) rows and columns to the bottom and right.
 These represent the point masses \(\bs q_{-1}(t)\) and \(\bs q_{K+1}(t)\), respectively.
@@
 	{{\bs B}}^{k\ell} &= \ddot{{\bs B}}^{k\ell}, \mbox{ for \(k\in\mathcal K^\circ\), \(\ell\in\{k-1,k,k+1\}\), \(k=\ell\neq 0\) or \(k=\ell\neq K\),}
-	\\ \bs B^{00} &= \ddot{\bs B}^{00} + \left[-c_ip_{ij}^{-1} 1(c_i<0, c_j>0)\right]_{i\in\calS,j\in\calS}\otimes \bs \phi^0(0)\tr{}\bs \phi^0(0)\bs M_0^{-1},
-	\\ \bs B^{KK} &= \ddot{\bs B}^{KK} + \left[c_ip_{ij}^{-1} 1(c_i>0, c_j<0)\right]_{i\in\calS,j\in\calS}\otimes \bs \phi^K(b)\tr{}\bs \phi^K(b)\bs M_K^{-1}, 
+	\\ \bs B^{00} &= \ddot{\bs B}^{00} - \left[c_ip_{ij}^{-1} 1(c_i<0, c_j>0)\right]_{i\in\calS,j\in\calS}\otimes \bs F_-^{00},
+	\\ \bs B^{KK} &= \ddot{\bs B}^{KK} + \left[c_ip_{ij}^{-1} 1(c_i>0, c_j<0)\right]_{i\in\calS,j\in\calS}\otimes \bs F_+^{00}, 
 	\\ {{\bs B}}^{{-1}0} &:= \bs T_{{-1}+}\otimes \boldsymbol\phi^0(0), 
-	\\ {{\bs B}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1}\mathbb 1_{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}} \otimes \boldsymbol \phi^0(0)\tr{}, 
+	\\ {{\bs B}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1} 1{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}} \otimes \boldsymbol \phi^0(0)\tr{}, 
 	\\ {{\bs B}}^{{K+1} K} &:= \bs T_{{K+1}-}\otimes \boldsymbol\phi^K(b),
-	\\ {{\bs B}}^{K,{K+1}} &:= \left[c_ip_{ij}^{K+1}\mathbb 1_{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes \boldsymbol \phi^K(b)\tr{},
+	\\ {{\bs B}}^{K,{K+1}} &:= \left[c_ip_{ij}^{K+1} 1{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes \boldsymbol \phi^K(b)\tr{},
 \end{align*} 
@@
 
-We prove the following result in Appendix\ref{sec:properties}.
+We prove the following result in Appendix~\ref{sec:properties}.
 %This transformation is not strictly necessary and all the following calculations can be done with either form, as long as week keep a consistent interpretation of the DG generator and related coefficients.
@@
 
-\section{Application to an SFFM}\label{sec:DGSFFM}
+\section{Application to a fluid-fluid queue}\label{sec:DGSFFM}
 Given our approximation \(\bs B\) to the generator \(\mathbb B\) we now follow the recipe from \cite{bo2014}, replacing the actual generator \(\mathbb B\) with its approximation \({  \bs B}\), to construct approximations, \(\bs \pi\) and \(\bs p\), to the stationary operators, \(\bbpi\) and \(\mathbb p\).
@@
 	\boldsymbol a_{k,i}\bs M_k^r
-	= \bs \rho_{k,i} M_k,
+	= \bs \rho_{k,i} \bs M_k,
 \)
@@
 Since \(\boldsymbol a_{k,i}\) is arbitrary, we see that we approximate \(\mathbb R_{k,i}\) by \(  \bs R_{k,i} = \bs M_k^r\bs M_k^{-1},\)
-and \(\mathbb R^k\) by \(  \bs R^k = \diag(  \bs R_{k,i})_{i\in\calS^\bullet_k}\).
+and \(\mathbb R^k\) by \(  \bs R^k = diag(  \bs R_{k,i},{i\in\calS^\bullet_k})\).
 
@@
 
-\subsection{Approximating the operator \(\bs D\) and the Riccati equation}
+\subsection{Approximating the operator \(\mathbb D\) and the Riccati equation}
 Recalling Lemma~\ref{lemma: D(s)} and replacing the operators \(\mathbb R^k\) and \(\mathbb B^{\ell m}\), by their approximations we have the following approximation to \(\mathbb D^{mn}(s)\)
@@
 \[\bs\phi^+(x){\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt y = \bs \phi^+(x)\bs M_+ \left[\left[\bs \Psi_{ij}^{k\ell}\right]_{i\in\calS_k^+,j\in\calS_\ell^-}\right]_{k\in\mathcal K^+,\ell\in\mathcal K^-}\bs \phi^-(y)'\wrt y\]%\left[\left[\bs\phi^k(x)\bs {M}_k^{-1}\bs \Psi_{ij}^{k\ell}(s)\bs\phi^\ell(y)\tr{}\wrt y\right]_{i\in\calS_k^+,j\in\calS_\ell^-}\right]_{ k\in\mathcal K^+, \ell\in\mathcal K^-},\]
-where \(\bs\phi^+(x) = (\bs\phi_k(x))_{i\in\calS_k^+,k\in\mathcal K^+}\) and \(\bs\phi^-(y) = (\bs\phi_k(y))_{i\in\calS_k^-,k\in\mathcal K^-}\) are row-vectors, \(\bs \Psi(s)\) is a matrix of constants for a given \(s\) with the same size as \(\bs D^{+-}\), and \({\bs M}_m,\) \(m\in\{+,-,0\}\) is a block diagonal matrix \({\bs M}_m = \diag\left( \left( \bs M_k\right)_{i\in\calS_k^m}\right)_{k\in\mathcal K^m}\), \(m\in\{+,-,0\}\).
+where \(\bs\phi^+(x) = (\bs\phi_k(x))_{i\in\calS_k^+,k\in\mathcal K^+}\) and \(\bs\phi^-(y) = (\bs\phi_k(y))_{i\in\calS_k^-,k\in\mathcal K^-}\) are row-vectors, \(\bs \Psi(s)\) is a matrix of constants for a given \(s\) with the same size as \(\bs D^{+-}\), and \({\bs M}_m,\) \(m\in\{+,-,0\}\) is a block diagonal matrix \({\bs M}_m = diag\left(  \bs M_k, {i\in\calS_k^m}, {k\in\mathcal K^m}\right)\), \(m\in\{+,-,0\}\).
 Now replace the theoretical kernels in Theorem~\ref{theo:Psi} by their DG approximations to get
@@
 \end{align*}
-Multiplying on the left by \(\bs\phi^+(x)\tr{}\) and on the right by \(\bs\phi^-(y)\), integrating over both \(x\) and \(y\), then post-multiplying by \({\bs M}^{-1}_-\) gives the following matrix Ricatti equation
+Multiplying on the left by \(\bs\phi^+(x)\tr{}\) and on the right by \(\bs\phi^-(y)\), integrating over both \(x\) and \(y\), then post-multiplying by \({\bs M}^{-1}_-\) gives the following matrix Riccati equation
 \begin{align}\label{eqn:RiccatiPsi}
@@
 \begin{rem}
-	Given the stochastic interpretation of \(\mathbb\Psi(0)\) we know that \( \bs \mu^+ \mathbb\Psi(0)([0,\infty))=1\) for every vector of measures \( \bs \mu^+\) such that \( \bs \mu^+([0,\infty)\boldsymbol 1 = 1\), when an SFFM is recurrent.
+	Given the stochastic interpretation of \(\mathbb\Psi(0)\) we know that \( \bs \mu^+ \mathbb\Psi(0)([0,\infty))=1\) for every vector of measures \( \bs \mu^+\) such that \( \bs \mu^+([0,\infty)\boldsymbol 1 = 1\), when a fluid-fluid queue is recurrent.
 It appears that this result carries over to the matrix \(\bs \Psi(0)\) giving the property that \(\displaystyle\int_{y\in[0,b]} \bs \Psi(0)\bs\phi^-(y)\tr{}\wrt y = \bs 1\).
@@
 We find an approximation to the stationary distribution by replacing the theoretical operators in Theorem~\ref{theo:density} with their approximations.
-Table \ref{table:notations} defines the notation we use for the DG approximations to stationary operators.
+Table~\ref{table:notations} defines the notation we use for the DG approximations to stationary operators.
 
@@
  \begin{tabular}{c|c|c|c}
-	\begin{tabular}{c}Exact \\operator\end{tabular} & Operator indices & \begin{tabular}{c}Approximation \\ notation\end{tabular} & Approximations \\\hline 
+	\begin{tabular}{c}Exact\end{tabular} & Operator indices & \begin{tabular}{c}Approximation \\ notation\end{tabular} & Approximations \\\hline 
 	%
@@
  \end{tabular}
- \caption{Notation for the approximation of the stationary operators of an SFFM.
+ \caption{Notation for the approximation of the stationary operators of a fluid-fluid queue.
 The first column contains the operators which we are approximating, the second column contains indices for which the operators are defined, the third column defines the notation we use for the coefficients of the approximation, and the last column defines how the coefficients and basis functions are used to approximate the operators.
@@
 
-With the notation in Table \ref{table:notations} define row-vectors 
+With the notation in Table~\ref{table:notations} define row-vectors 
  \begin{align*}
@@
 
-Proceeding similarly to the derivation of the Ricatti equation (\ref{eqn:RiccatiPsi}), we can argue that the coefficients \( {\boldsymbol{\xi}}\) are the solution to the matrix system 
+Proceeding similarly to the derivation of the Riccati equation~(\ref{eqn:RiccatiPsi}), we can argue that the coefficients \( {\boldsymbol{\xi}}\) are the solution to the matrix system 
 % 	 
@@
 %
-%Proceeding similarly to the derivation of the Ricatti equation (\ref{eqn:RiccatiPsi}), we can argue that the DG approximation to \( {\boldsymbol{\xi}}\) is the solution to the matrix system 
+%Proceeding similarly to the derivation of the Riccati equation~(\ref{eqn:RiccatiPsi}), we can argue that the DG approximation to \( {\boldsymbol{\xi}}\) is the solution to the matrix system 
 %% 	 
@@
 
-\section{Other problems we can solve with the DG method}
-The utility of the DG approximation scheme for fluid queues in not limited to approximating the first return and stationary operators of fluid-fluid queues.
-For example, we can use a DG scheme to approximate the transient distribution of the fluid queue at time \(t\), which requires us to find the coefficients \(\vligne{\bs q_{-1}(t) & \bs a(t) & \bs q_{{K+1}}(t)}\) via the differential equation (\ref{eqn: DG ODE w BCs}) given an initial condition, typically by numerically integrating (\ref{eqn: DG ODE w BCs}).
-The stationary distribution of the fluid queue can be found by solving
+\section{Other problems we can solve with cell-based approximation schemes}\label{sec: other applications}
+The utility of the DG scheme (or any of the approximation schemes considered in this thesis, for that matter) for fluid queues is not limited to approximating the first return and stationary operators of fluid-fluid queues.
+For example, we can use a DG scheme to approximate the transient distribution of the fluid queue at time \(t\), which requires us to find the coefficients \(\vligne{\bs q_{-1}(t) & \bs a(t) & \bs q_{{K+1}}(t)}\), given an initial condition, typically by numerically integrating (\ref{eqn: DG ODE w BCs}).
+An approximation to the stationary distribution of the fluid queue can be found by solving
 \begin{align*}
@@
 We can also approximate first hitting times.
-For example, given an initial condition of the cell \(\calD_k\), we can project the initial condition onto the polynomial basis with coefficients \(\bs a_k(0)\), then approximate the probability that the level process first hits \(\{y_k,y_{k+1}\}\) after time \(t_0\) by finding \(\bs a_k(0)e^{B^{kk} t_0}\int_{\calD_k}\bs \phi_k(x)\wrt x\) where the coefficients \(\bs a_k(t_0)\) can be found bi integrating the differential equation
+For example, given an initial condition on cell \(k\), we first find the initial coefficients \(\bs a_k(0)\), then approximate the probability that the level process first hits \(\{y_k,y_{k+1}\}\) after time \(t_0\) by finding \(\bs a_k(0)e^{B^{kk} t_0}\int_{y_k}^{y_{k+1}}\bs \phi_k(x)\wrt x\) where the coefficients \(\bs a_k(t_0)\) can be found by integrating the differential equation
 \[\cfrac{\wrt}{\wrt t}\bs a_k(t)=\bs a_k(t)\bs B^{kk},\]
-over time. We now briefly introduce time integration schemes which we use in this thesis, and some issues pertaining to oscillatory approximations.
+over time.
+
+Indeed, the DG scheme has the potential to answer many mathematical questions for fluid queues.
+However, one reason it may not have found such widespread use in the context of fluid queues is the aforementioned difficulties surrounding discontinuous solutions, which we often want to consider.
+Moreover, there are sometimes other techniques which can give exact solutions, for example exact expressions for transient quantities are derived in \cite{rs2003,ar2004,bean2005,bean2009}, and for the stationary distribution are derived in \cite{ajr2005,dasilva2005,lnp13,s2017}.
+
+Nonetheless, the DG scheme can be used to answer a very broad range of questions about fluid queues (such as transient distributions, hitting times on complex geometries, and of course analysis of fluid-fluid queues) which have no exact, readily computable solution.
+Moreover, we will numerically investigate the performance of the DG scheme (and other numerical schemes which we will develop) in Chapter~\ref{sec: numerics} which require numerical integration with respect to time.
+Hence, we now briefly introduce the numerical time integration scheme which we use in this thesis, as well as some issues pertaining to oscillatory approximations.
 
@@
 \end{align}
-where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix.
+where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix, given an initial condition \(\bs a(0)\).
 To do so, we employ the strong stability preserving Runge-Kutta (SSPRK) scheme of order 4 with 5 stages \citep{sr2002}.
-SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
+SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``...can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
 
@@
 where \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are coefficients which define the scheme and \(h\) is the \(t\)-step size.
-The coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are chosen to be positive so that \(\bs v^{(s)}\) is a convex combination of forward-Euler steps.
+To ensure the stability-preserving property the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are chosen to be positive so that \(\bs v^{(s)}\) is a convex combination of forward-Euler steps.
 Moreover, the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are optimised so that the allowable \(t\)-step size is as large as possible.
-The maximum allowable \(t\)-step size is \(\Delta_{RK}\), and we require
-\[\Delta_{RK} \leq \min_{\ell k}\cfrac{\alpha_{\ell k}}{\beta_{\ell k}}\Delta_E,\]
-where \(\Delta_E\) is the maximum allowable \(t\)-step size for the forward-Euler scheme.
+The maximum allowable \(t\)-step size is \(h_{RK}\), and we require
+\[h_{RK} \leq \min_{\ell k}\cfrac{\alpha_{\ell k}}{\beta_{\ell k}}h_E,\]
+where \(h_E\) is the maximum allowable \(t\)-step size for the forward-Euler scheme.
 
 The \(t\)-step size needs to be chosen to be sufficiently small so that the time integration is stable. Here, for an order \(p\) DG scheme we choose the time step to be less than 
-\[\max_{i\in\calS}|c_i|\min_{r\in{2,...,p+1}}\left(z_{r}-z_{r-1}\right) \cfrac{\min_{k}{K+1}_k}{2}\]
+\[\max_{i\in\calS}|c_i|\min_{r\in{2,...,p+1}}\left(z_{r}-z_{r-1}\right) \cfrac{\min_{k}{\Delta}_k}{2}\]
 where \(z_{r-1}\leq z_r\), \(z_1=-1\), \(z_{p+1}=1\) and \(z_r, r=2,...,p,\) are the \(p-1\) zeros of the first derivative of \(P_{p}(x)\), the order \(p\) Legendre polynomial.
@@
 \section{Slope limiters}\label{sec: slope limiting}
-A well-know problem with DG schemes is that, in the presence of steep gradients or discontinuities, spurious oscillations in the approximate solution can occur and positivity is not guaranteed (see \cite{nodalDGBook}~Section~5.6 and \cite{koltai2011}~Section~3.3, for example).
+A well-know problem with DG schemes is that, in the presence of steep gradients or discontinuities, spurious oscillations in the approximate solution can occur and positivity is not guaranteed (see \cite{nodalDGBook}~Section~5.6, and \cite{koltai2011}~Section~3.3, for example).
 To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
-Slope limiting effectively alters the DG operator in regions where oscillations are detected to ensure non-oscillatory and non-negativity by reducing the order of the approximation to linear in these regions.
+Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory and non-negativity by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
 However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
 As a result, the limiter may unnecessarily reduce the accuracy of the scheme in the presence of natural oscillations (see \citep{nodalDGBook},~Example~5.8).
-Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear operators.
+Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear.
 
-We now describe a simple slope limiter known as the \emph{Generalised MUSCL} scheme \citep{c99} (see also \cite[Section~5.6.1]{nodalDGBook}). Define the \emph{minmod} function 
+We now describe a simple slope limiter known as the \emph{Generalised MUSCL} scheme \citep{c99} (see also \citep{nodalDGBook},~Section~5.6.1). Define the \emph{minmod} function 
 \begin{align*}
@@
 When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
-In the context of slope-limiting, the three arguments \(a_1,\,a_2\) and \(a_3\) are the gradients of the approximate solution in three cells.
+In the context of DG the three arguments \(a_1,\,a_2\) and \(a_3\) are the estimates of the average gradient of the approximate solution near a given cell.
 When the signs of \(a_1,\,a_2\) and \(a_3\) differ, this suggests an oscillation in the approximate solution.
@@
 \end{align*}
-where \(\Pi^1_{\partial x}u_k\) is the slope of the linear projection of \(u_k\) on \(\calD_k\) and \(\overline y_k=(y_k+y_{k+1})/2\) is the centre of \(\calD_k\).
+where \(\Pi^1_{\partial x}u_k\) is the slope of the linear projection of \(u_k\) on cell \(k\) and \(\overline y_k=(y_k+y_{k+1})/2\) is the centre of the \(k\)th cell.
 
@@
 
-In the context of approximating the first return operator of a fluid-fluid queue there is no direct equivalent of slope limiting, other than, perhaps, to re-compute the solution with constant basis function if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
+% Notice that the process of limiting is essentially a post-processing of the approximate solution after applying
 
-\subsection{Slope limiters and time integration}
+% In the context of approximating the first return operator of a fluid-fluid queue there is no direct equivalent of slope limiting, other than, perhaps, to re-compute the solution with constant basis function if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
+
+\subsection{Slope limiters and time integration}\label{subsec: slope lim and int}
 To use slope limiters within the SSPRK scheme we first project the initial condition on to \(W\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
-
-For example, to find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation (\ref{eqn: DG ODE w BCs 3}).
+To find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the (limited) initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation~(\ref{eqn: DG ODE w BCs 3}).
 At each stage of the time-integration we apply the slope limiter, i.e.~the scheme above with a slope limiter is
 \begin{align*}
-	&\bs v^{(0)} = \vligne{\bs q_{-1}(t) & \bs a(t) & \bs q_{{K+1}}(t)},\\
+	&\bs v^{(0)} = \bs a(t),\\
 	%
-	&\bs v^{(\ell)} = \Pi^{lim}\left(\sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k} \bs B \bs v^{(k)}\right),\, \ell = 1,\dots,s,\\
+	&\bs v^{(\ell)} = \Pi^{lim}\left(\sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q\right),\, \ell = 1,\dots,s,\\
 	%
-	&\vligne{\bs q_{-1}(t+h) & \bs a(t+h) & \bs q_{{K+1}}(t+h)} =\bs v^{(s)},
+	&\bs a(t+h) =\bs v^{(s)},
 \end{align*} 
-where \(\Pi^{lim}\) is the slope limiter function.
+where \(\Pi^{lim}\) is the slope limiter function which determines on which cells the solution needs limiting, and if so, applies the limiter \(\Pi_k^{lim}\).
+
+We refer to this (vanilla) application of a limiter to the DG scheme as the DG-lim scheme.
 
@@
 When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of Generalised slope limiter described.
-Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will effect the accuracy of subsequent computations.
+Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy of subsequent computations.
 
@@
 The advantage of the approach above is that, in areas of the approximation where the solution is smooth, then the high-order accuracy of the method can be retained, while maintaining positivity where necessary.
-However, if there are no regions of the solution of interest which are smooth, then there is no real advantage in this approach as the limiter will reduce the approximate solution to linear in all regions of interest.
+However, if the problem at hand is dominates by discontinuities and/or if there are no regions of the solution of interest which are smooth, then there is no real advantage in this approach as the limiter will reduce the approximate solution to linear in all regions of interest.
 Hence, we may as well just use a linear scheme and save on some computation.
-Perhaps a better approach would be to use a linear scheme with a smaller cell width, such that the computational work remains approximately the same.
 
+Perhaps a better approach would be to use a linear scheme with a smaller cell width such that the computational work remains approximately the same.
 For example, for a DG scheme with \(2p\) basis functions on a cell, \(\mathcal D_k\), of width \(\Delta\), say, we construct block matrices of dimension \(2p\times 2p\) (i.e.~the mass, stiffness and flux matrices, \(\bs M_k\), \(\bs G_k\), and \(\bs F^{k\ell}\), respectively) and compute the approximate solution using these matrices.
-Instead, we could break the cell, \(\mathcal D_k\), up into \(p\) sub-cells of width \(\Delta/p\) and use a DG scheme with \(2\) basis functions on each sub-cell (i.e.~a linear basis on each cell).
-With this approach, to approximate the solution on the original interval \(\mathcal D_k\), we would also construct block matrices of dimension \(2p\times 2p\) and compute the approximate solution on \(\calD_k\) using these matrices in the same way as before.
-If we know \emph{a priori} that we will apply the slope limiter to cell \(\calD_k\), then intuitively we expect that the latter scheme will be far more accurate; it will approximate the solution on \(\calD_k\) by a piecewise linear function with \(p\) pieces, whereas the DG scheme with \(2p\) basis functions on cell \(\calD_k\) will, after limiting, represent the solution as a single linear function on \(\calD_k\).
+Alternatively, consider breaking the cell, \(\mathcal D_k\), up into \(p\) sub-cells of width \(\Delta/p\) and using a DG scheme with \(2\) basis functions on each sub-cell (i.e.~a linear basis on each cell).
+On each of the \(p\) sub-cells we construct block matrices of size \(2\times 2\).
+To approximate the solution on the original interval \(\mathcal D_k\) we form the \(p\) approximations on each sub-cell into larger matrices which are of dimension \(2p\times 2p\), and we then use in the same way as before.
+Thus, the scheme with \(2p\) basis functions on the cell \(\calD_k\), and the scheme with a 2 basis functions on \(p\) sub-cell have approximately the same computational cost (the latter scheme results in a sparser matrix).
+
+If we know \emph{a priori} that we will apply the slope limiter to cell \(\calD_k\), then intuitively we suspect that the latter scheme may be far more accurate; it will approximate the solution on \(\calD_k\) by a piecewise linear function with \(p\) pieces, whereas the DG scheme with \(2p\) basis functions on cell \(\calD_k\) will, after limiting, represent the solution as a single linear function on \(\calD_k\).
+
+We refer to this scheme as the DG-lin-lim scheme, as it is a linear DG scheme with a limiter.
+
+\subsection{Slope limiters and linear operators}\label{sec: limiting and linearity}
+The process of slope limiting, as described above, essentially amounts to a non-linear post-processing of the approximate solution after applying the operator \(\bs B\).
+Thus, the slope-limited scheme arrives at an approximation which is the action of a non-linear operator applied to the initial condition.
+The fact that the approximate operator is non-linear makes solving the Riccati equation~\ref{eqn:RiccatiPsi} difficult/impossible, whereas there are efficient techniques to solve matrix-Riccati equations \cite{bot08}.
+Thus, for the application to fluid-fluid queues slope limiting is not a viable option.
+We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{Y(t)\}\) first return to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation.
+The application of the limiter/filter is dependent on the initial condition, so the post-processing needs to be done for every initial condition under consideration.
+If there are many initial distributions to consider, this could result in an increased computational cost.
+We do not investigate this post-processing approach in this thesis.
+
+On a related note, the advantage of a linear approximation can also be useful if we want to approximate transient distributions of fluid queues for many initial conditions.
+For example, consider solving (\ref{eqn: DG ODE w BCs 3}) numerically with a forward-Euler scheme to integrate over time.
+To advance the coefficients from time \(t\) to time \(t+h\) we compute
+\begin{align*}
+	&\bs a(t+h) = \bs a(t) + h \bs a(t)\bs Q = \bs a(t)\left[\bs I + h\bs Q\right].
+\end{align*} 
+Thus, the coefficients at time \(t_0=nh\) are given by 
+\begin{align*}
+	&\bs a(nh) = \bs a(0)\left[\bs I + h\bs Q\right]^n,
+\end{align*} 
+which is a linear transformation of the coefficients \(\bs a(0)\).
+Thus, if we determine the matrix \(\left[\bs I + h\bs Q\right]^n\), then we may apply it to any initial condition to approximate the solution at time \(t_0=nh\).
+In the presence of a limiter, this is not possible as the coefficients at time \(t_0=nh\) are not a linear transformation of the initial coefficients.
+
+Another point to note is that limiters incur a computational cost.
+There is a computational cost in reducing the solution to linear when oscillations are detected, and also in determining whether there are oscillations or not.
+Thus, even if the solution needs no limiting, there is an added computational cost in determining this is the case.
+Given the limiter is applied for each stage and each time-step of the time integration scheme in Section~\ref{subsec: slope lim and int}, then this computational cost may not be insignificant.
+
+One advantage of the QBD-RAP (and uniformisation) scheme we develop in this thesis is that the resulting operator is positivity-preserving and linear.
+% This raises two points.
+First, the concept of limiting applies only to problems with the concept of an approximate solution, and not to problems where the object of interest is the approximate operator itself.
+For example, we solve the Ricatti equation~\ref{eqn:RiccatiPsi} for the matrix \(\bs \Psi\) to find an approximation to \(\mathbb \Psi\).
+There is no concept of an approximate solution in this problem, only the concepts of the approximate operators.
+Thus, there is no concept of slope limiting
+% 
+% added computational cost when lots of limiting is needed
+%
 
@@
 As a rough guide on accuracy we now paraphrase some know results on the accuracy of the DG method for a similar kind of problem to the one under study here.
-For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the approximation to the flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution at time \(t\) provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
-Applying this result to the type of positivity preserving scheme in \ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
+For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
+Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
 

### 2022-02-17 chapter5/chapter.tex
first draft of the whole thing! :D

@@
 \chapter{Global convergence results}\label{ch: global results}
-In this chapter we prove some results which stem from the main results of Chapter~\ref{sec: conv}.
-We consider the discrete-time embedded processes formed by observing the QBD-RAP at the orbit restart epochs and by observing the fluid queue at the hitting times of the level process at the of points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
+In this chapter we prove convergence results which build on the main results of Chapter~\ref{sec: conv}.
+We consider the discrete-time embedded processes formed by observing the QBD-RAP at the orbit restart epochs and by observing the fluid queue at the hitting times of the level process at the on points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
 In Corollary~\ref{cor: aln22222212} we prove that the transition probability of the embedded process of the QBD-RAP converge those of the embedded process of the fluid queue.
-In Corollary~\ref{cor: short cor} we prove that distribution of the sojourn time of the QBD-RAP in a given level converges to distribution of the sojourn time of the fluid queue in the corresponding interval.
-In Section~\ref{sec: sub global}, we state global results on the weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
+In Corollary~\ref{cor: short cor} we prove that the distribution of the sojourn time of the QBD-RAP in a given level converges to the distribution of the sojourn time of the fluid queue in the corresponding interval.
+In Theorem~\ref{thm: big thm}, we state global results on the weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
 
@@
 The results of this chapter rely on results of Chapter~\ref{sec: conv} and therefore apply to the QBD-RAP scheme which uses ephemeral states to model a fluid queue which starts in a phase with rate \(0\), as described in Section~\ref{sec: initial conditions}.
-However, supplementing the results of Chapter~\ref{sec: conv} with the results from Appendix~\ref{app:extend conv} and using the same arguments from this chapter with only slight modifications, then the results of this chapter can be extended to the augmented state-space QBD-RAP scheme without the initial ephemeral phases.
+However, supplementing the results of Chapter~\ref{sec: conv} with the results from Appendix~\ref{app:extend conv} and using the same arguments from this chapter with only slight modifications, then the results of this chapter can be extended to the augmented state space QBD-RAP scheme without the initial ephemeral phases.
 
@@
 First, we show that the transition probabilities of the embedded chain of the QBD-RAP convergence to the transition probabilities of the chain embedded in the fluid queue.
-We show that for a fixed number of level changes, \(n\), say, the embedded process of the QBD-RAP converges to an embedded discrete time process of the fluid queue, embedded at the times when the fluid queue leaves the bands \(\mathcal D_{\ell}\).
+We show that for a fixed number of level changes, \(n\), say, the embedded process of the QBD-RAP converges to an embedded continuous-time process of the fluid queue, embedded at the times when the fluid queue leaves the bands \(\mathcal D_{\ell}\).
 Now, from Chapter~\ref{sec: conv} we have a result on the convergence of the QBD-RAP up to the first change of level (THM) and at the first change of level (COR).
@@
 % The structure of this chapter is as follows.
-We start by showing weak convergence of the approximation at the \(1\)st and then \(n\)th change of level, \(n> 1\) in Sections~\ref{sec: 1st change} and \ref{sec: nth change}.
+We start by showing weak convergence of the approximation at the \(1\)st and then \(n\)th change of level, \(n> 1\) in Sections~\ref{sec: 1st change} and~\ref{sec: nth change}.
 Section~\ref{sec: between n and np1} generalises this to show weak convergence between the \(n\)th and \(n+1\)th change of level.
@@
 
-In this section we consider the embedded process formed by observing the QBD-RAP at the orbit restart epochs. Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
-\[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)})\right\},\]
+Consider the embedded process formed by observing the QBD-RAP at the orbit restart epochs. Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
+\[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)}), \mbox{or \(\{\bs Y(t)\}\) hits a boundary}\right\},\]
 be orbit restart epochs.
-These are the (stopping) times at which \(\{L^{(p)}(t)\}\), the level process of the QBD-RAP, changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary.
+These are the (stopping) times at which the level process of the QBD-RAP, \(\{L^{(p)}(t)\}\), changes, or the boundary is hit, or, if the process is at the boundary, the process leaves the boundary.
 To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed.
 Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the orbit restart epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
-The subscript \(\bs \alpha \) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\).
+The subscript \(\bs \alpha \) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\), unless the process is temporarily absorbed in a sticky boundary, in which case \(\bs A^{(p)}(\tau_n^{(p)})=1\).
 The process \(\{\Ydp(n)\}_{n\geq 0}\) is a discrete-time Markov chain, which is time-homogeneous for \(n\geq 1\).
@@
 Let \(\{\tau_n^X\}_{n\geq 0}\), be the sequence of (stopping) times with \(\tau_0^X=0\), and 
-\[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(0)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(0)=y_{K+1}\right\} \end{array} \right\},\]
+\[\tau_{n+1}^X = \min\left\{\begin{array}{c}\inf\left\{t>\tau_n^X\mid X(t)=y_{\ell}, \ell\in\mathcal K\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq 0, X(\tau_{n-1}^X)=0\right\}, \\ \inf\left\{t>\tau_n^X \mid X(t) \neq y_{K+1}, X(\tau_{n-1}^X)=y_{K+1}\right\} \end{array} \right\},\]
 for \(n\geq 0\).
@@
 
-We have the following result on the convergence of the embedded processes \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\), which we will utilise later to prove a global result.
+We have the following result on the convergence of the embedded processes \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\), which we also will utilise later to prove a global result.
 \begin{cor}\label{cor: aln222222} For \(\ell_0\in\mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
@@
 	\end{align}
-	and for \(n\geq 1\), then, 
+	and for \(n\geq 1\), 
 	\begin{align}
@@
 Ultimately, we prove the weak convergence (in space and time) of the QBD-RAP approximation scheme to the distribution of the fluid queue.
-The structure of the argument is to first show a convergence result for the Laplace transform with respect to time of the distributions of the QBD-RAP and fluid queue at the \(n\)th orbit restart epoch.
-We then prove a convergence result of the QBD-RAP and fluid queue between the \(n\)th and \(n+1\)th orbit restart epoch.
-Summing over the number of orbit restart epochs, \(n\), and via the Dominated Convergence Theorem, we claim a convergence of the Laplace transforms with respect to time of the QBD-RAP approximation and fluid queue.
+This is done in stages.
+First, by considering the Laplace transform with respect to time of the distribution of the QBD-RAP \emph{at the time of the \(n\)th orbit restart epoch}, \(\tau_n\).
+Second, considering the Laplace transform with respect to time of the distribution of the QBD-RAP \emph{between} the \(n\)th and \((n+1)\)th orbit restart epoch.
+Third, summing over the number of orbit restart epochs, \(n\), and via the Dominated Convergence Theorem, we claim a convergence of the Laplace transforms with respect to time of the QBD-RAP and fluid queue.
 Lastly, we apply the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} to claim a weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
@@
 % For \(n\geq 1\), \(\tau_n^X\) is the time at which \(X(t)\) either changes band, or hits a boundary, or the process leaves a boundary, for the \(n\)th time. 
-
-For \(n\geq 1\), consider the Laplace transform 
+The idea for the analysis of the QBD-RAP at the \(n\)th orbit restart epoch is to partition the distribution of the QBD-RAP on the phase and level at the times \(\tau_1^{(p)}\) and \(\tau_2^{(p)},...,\tau_{n-1}^{(p)}\) and then use the strong Markov property to write the expression in terms of the distributions of the QBD-RAP at time \(\tau_k^{p}\) given the distribution of the QBD-RAP at time \(\tau_{k-1}^{(p)}\), \(k=1,...,n\).
+The distribution at time \(\tau_1^{(p)}\) depends on the initial condition, so we treat it separately and use Corollary~\ref{cor: aln222} to claim convergence to the fluid queue.
+By time-homogeneity, the distributions of the QBD-RAP at times \(\tau_k^{p}\) given the position of the QBD-RAP at time \(\tau_{k-1}^{(p)}\) is equal to the distribution of the QBD-RAP at time \(\tau_1^{(p)}\) given an initial distribution of the QBD-RAP.
+Hence, we may also use Corollary~\ref{cor: aln222} to claim convergence to the fluid queue for each of these distributions.
+Putting all the elements together yields the convergence at the \(n\)th orbit restart epoch.
+
+Formally, for \(n\geq 1\), consider the Laplace transform 
 \begin{align}
@@
 \end{align}
-An application of Corollary \ref{cor: aln222} to the expression on the second line of (\ref{eqn: mnebrb2}) states, 
+An application of Corollary~\ref{cor: aln222} to the expression on the second line of (\ref{eqn: mnebrb2}) states, 
 \begin{align}
@@
 	\\&\to \mathbb P(\bs X(\tau_1^X) = (y_{\ell+1(j_1\in\calS_-)}, j_1), \tau_{1}^X\leq E^\lambda 
-	\mid \bs X(0) = (x_0, i)) 
+	\mid \bs X(0) = (x_0, i)) \label{eqn: aMMMMM}
 \end{align}
@@
 
-Thus, by partitioning on the phases and the levels at the times \(\tau_m\), \(m=2,\dots,n-1\),and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is  
+Partitioning on the phases and the levels at the times \(\tau_m\), \(m=2,\dots,n-1\), and using the strong Markov property of the QBD-RAP, then (\ref{eqn: dskvnaSF}) is  
 	\begin{align}
@@
 % by the time-homogeneous property of the QBD-RAP and the memoryless property of the exponential distribution. 
-We can apply Corollary \ref{cor: aln222} to the factors of the summands in (\ref{eqn: 161222}) and conclude 
+We can apply Corollary~\ref{cor: aln222} to the factors of the summands in (\ref{eqn: 161222}) and conclude 
 \begin{align}
@@
 
-Thus, returning to (\ref{eqn: 161222}) and taking the limit as \(p\to\infty\),  
+Thus, returning to (\ref{eqn: 161222}), taking the limit as \(p\to\infty\), and using the convergence established in (\ref{eqn: aMMMMM}) and (\ref{eqn: lim 12}), then 
 \begin{align}
@@
 
-\subsection{Between the \(n\)th and \(n+1\)th orbit restart epochs}\label{sec: between n and np1}
+\subsection{Between the \(n\)th and \((n+1)\)th orbit restart epochs}\label{sec: between n and np1}
 In the last section we proved a convergence of the QBD-RAP to the fluid queue \emph{at} the \(n\)th orbit restart epoch.
 However, we ultimately want to make a convergence statement about the QBD-RAP scheme at any time.
-The next step is therefore to show a convergence of the QBD-RAP to the fluid queue between the \(n\)th and \(n+1\)th orbit restart epoch, after which we sum over \(n\) to prove a convergence result independent of the number of orbit restart epochs.
-
-Let \(\mathcal T_n^{(p)} = (\tau_{n}^{(p)},\tau_{n+1}^{(p)}]\) and \(\mathcal T_n^X=(\tau_n^X,\tau_{n+1}^X]\). Consider the Laplace transform 
+The next step is to show a convergence of the QBD-RAP to the fluid queue between the \(n\)th and \((n+1)\)th orbit restart epoch, after which we sum over \(n\) to prove a convergence result independent of the number of orbit restart epochs.
+
+The idea is to partition the distribution of the QBD-RAP between time \(\tau_n^{(p)}\) and \(\tau_{n+1}^p\) on the position of the QBD-RAP at time \(\tau_n^{(p)}\), then use the strong Markov property to write the distribution of the QBD-RAP between time \(\tau_n^{(p)}\) and \(\tau_{n+1}^p\) in terms of the distribution of the QBD-RAP at time \(\tau_n^{(p)}\) given the initial position of the QBD-RAP, and the distribution of the QBD-RAP at between time \(\tau_n^{(p)}\) and \(\tau_{n+1}^{(p)}\) given the position of the QBD-RAP at time \(\tau_n^{(p)}\).
+In the previous section we showed convergence of the distribution of the QBD-RAP at time \(\tau_n^{(p)}\) given the initial position of the QBD-RAP.
+By a time-homogeneity argument, the distribution of the QBD-RAP between time \(\tau_n^{(p)}\) and \(\tau_{n+1}^{(p)}\) given the position of the QBD-RAP at time \(\tau_n^{(p)}\) is equal to the distribution of the QBD-RAP before time \(\tau_1^{(p)}\) given an initial condition on the QBD-RAP.
+Thus, Theorem~\ref{thm: a thm!} establishes the convergence of this term.
+Together, we have convergence of the QBD-RAP between the \(n\)th and \((n+1)\)th orbit restart epoch.
+
+More formally, let \(\mathcal T_n^{(p)} = (\tau_{n}^{(p)},\tau_{n+1}^{(p)}]\) and \(\mathcal T_n^X=(\tau_n^X,\tau_{n+1}^X]\). Consider the Laplace transform 
 \begin{align}
@@
 \end{align}
-as that appearing in Lemma \ref{lem: kKKJJJF}, hence (\ref{eqn: Kkk nj aa w}) converges to 
+as that appearing in Lemma~\ref{lem: kKKJJJF}, hence (\ref{eqn: Kkk nj aa w}) converges to 
 \begin{align}
@@
 \begin{align}
-	&\int_{x\in\calD_{\ell,j}} \widehat \mu^{\ell} (\lambda)(x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x)\wrt x \nonumber 
+	&\int_{x\in\calD_{\ell,j}} \widehat \mu^{\ell} (\lambda)(\wrt x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x) \nonumber 
 	\\&= \int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_0^X \mid \bs X(0) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n) ) \psi(x).
@@
 
-Returning to the right-hand side of (\ref{eqn: l whkrqvlkjbrd}) and taking the limit as \(p\to\infty\), 
+Returning to the right-hand side of (\ref{eqn: l whkrqvlkjbrd}), taking the limit as \(p\to\infty\), and recalling the limits in (\ref{eqn: ashverkjbvjhkhdsknacva}) and (\ref{eqn: lkHAHJFHKJ J J 22}), then
 \begin{align}
@@
 	\\&\qquad{}  \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
-	 \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}).\label{eqn: eieieie}
+	\mid \bs Y^{(p)}(0)=\bs y_0^{(p)}) \nonumber 
+	\\&= \sum_{j_n\in\mathcal S}\lim_{p\to\infty}
+	\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P({\bs Y}^{(p)}(t) = (\ell,\wrt x, j), 
+	t\in\mathcal T_0^{(p)} \mid \nonumber 
+	\bs Y(0) = (\ell,\bs\alpha,j_n))\wrt t \psi(x)
+	\\&\qquad{} \lim_{p\to\infty} \mathbb P(\Ydp(n) = (\ell, j_n), \tau_{n}^{(p)}\leq E^\lambda 
+	\mid \bs Y^{(p)}(0)=\bs y_0^{(p)})\nonumber 
 	\\&=\sum_{j_n\in\mathcal S}\int_{x\in\calD_{\ell,j}} \int_{t=0}^\infty e^{-\lambda t} \mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} \mid \bs X(\tau_n^X) = (y_{\ell+1(j_n\in\mathcal S_-)}, j_n),\nonumber
@@
 \end{align}
-where the first equality holds since the sum is finite and the limits exist, and the second equality holds from the law of total probability.
+where the first equality holds since the sum is finite and the limits exist, and the last equality holds from the law of total probability.
 
@@
 		\\&\to\int_{t=0}^\infty e^{-\lambda t}  \int_{x\in\calD_{\ell,j_n}}\mathbb P(\bs X(t)\in(\wrt x, j), t\in\mathcal T_n^{X} 
-		\mid \bs X(0)=(x_0,i))\psi(x)\wrt x\wrt t.
+		\mid \bs X(0)=(x_0,i))\psi(x)\wrt t.
 	\end{align}
@@
 
-\subsection{A domination condition}
+\subsection{Intermezzo: A domination condition}
 % We now want to sum the expressions in Lemma~\ref{lem: LAkAKFnvnb mav h} over \(n\) to obtain a result which is independent of the number of orbit restart epochs.
@@
 Hence, upon taking the limit of (\ref{eqn:GHJKLKJHGHJKL}), if we can justify the swap of the sum and the limit, we will obtain the desired result.
-To this end, we now show a domination condition in Corollary~\ref{vcor: cdks d} so that we may apply the Dominated Convergence Theorem.
+To this end, we show a domination condition in Corollary~\ref{vcor: cdks d} so that we may apply the Dominated Convergence Theorem.
+First, we need a couple of intermediate results.
 %The main result of this subsection is Lemma~\ref{lem: another bound 2} in which we show a bound which can serve as a dominating function so that we may apply the Dominated Convergence Theorem to justify the aforementioned swap of sum and limit.
@@
 	where 
-	\[b = \min\left\{1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,\cfrac{q}{q+\lambda}\right\}\]
-	and  
-	\[|r_2|\leq 2G\cfrac{\var \left(Z\right)}{\varepsilon^2} + 2L\varepsilon.\]
+	\[b = \max\left\{1-e^{-\gamma(\Delta+\varepsilon)}\left[1-e^{\gamma\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,\cfrac{\gamma}{\gamma+\lambda}\right\},\]  
+	\[|r_2|\leq 2G\cfrac{\var \left(Z\right)}{\varepsilon^2} + 2L\varepsilon,\]
+	and \(\gamma = \max_{i\in\calS_+\cup\calS_-}-T_{ii}/|c_i|\). 
 \end{lem}
@@
 		\\&= 1-e^{T_{ii}(\Delta+\varepsilon)/|c_i|}\left[1-e^{(-T_{ii}\varepsilon-\lambda \Delta)/|c_i|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2| \nonumber 
-		\\&= 1-e^{-q(\Delta+\varepsilon)}\left[1-e^{q\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,
-	\end{align}
-	since \(-T_{ii}/|c_i|\leq q\) and \(\lambda \Delta/|c_i| \leq \lambda \Delta/c_{min}\) for all \(i\in\mathcal S_+\cup \calS_-\).
+		\\&= 1-e^{-\gamma(\Delta+\varepsilon)}\left[1-e^{\gamma\varepsilon-\lambda \Delta/|c_{min}|}\right] + \cfrac{\var(Z)}{\varepsilon^2} + |r_2|,
+	\end{align}
+	since \(-T_{ii}/|c_i|\leq \gamma\) and \(\lambda \Delta/|c_i| \leq \lambda \Delta/c_{min}\) for all \(i\in\mathcal S_+\cup \calS_-\).
 
@@
 To leave the boundary there must be at-least one change of phase before \(E^\lambda\).
-By a uniformisation argument, the probability of at-least one change of phase before \(E^\lambda\) is less than or equal to \(q/(q+\lambda)\).
+By a uniformisation argument, the probability of at-least one change of phase before \(E^\lambda\) is less than or equal to \(\gamma/(\gamma+\lambda)\).
 \end{proof}
@@
 	\end{align}
-	Now partition (\ref{eqn: eheh}) on the phase at time \(\tau_{n-1}\),
-	\begin{align}
-		\nonumber&\sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_n \leq E^\lambda,  \tau_{n-1} \leq E^\lambda,\phi(\tau_{n-1}) = j_{n-1} \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)
-		\\&\nonumber=  \sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_{n}\leq E^\lambda\mid   \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1} \leq E^\lambda)
-		\\&\qquad\times\mathbb P( \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda), \label{eqn: kekew}
+	Now partition~(\ref{eqn: eheh}) on the phase and level at time \(\tau_{n-1}\),
+	\begin{align}
+		\nonumber&\sum_{\ell_{n-1}\in\mathcal K}\sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_n \leq E^\lambda,  \tau_{n-1} \leq E^\lambda, L(\tau_{n-1}) = \ell_{n-1},\phi(\tau_{n-1}) = j_{n-1} \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda)
+		\\&\nonumber=  \sum_{\ell_{n-1}\in\mathcal K}\sum_{j_{n-1}\in\mathcal S}\mathbb P(  \tau_{n}\leq E^\lambda\mid  L(\tau_{n-1}) = \ell_{n-1}, \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1} \leq E^\lambda)
+		\\&\qquad\times\mathbb P( L(\tau_{n-1}) = \ell_{n-1}, \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda \mid \phi(\tau_1) = i, \tau_1\leq E^\lambda), \label{eqn: kekew}
 	\end{align}
@@
 	\begin{align}
-		&\sum_{j_{n-1}\in\mathcal S} b
-		\mathbb P(  \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda\mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \nonumber
+		&\sum_{\ell_{n-1}\in\mathcal K}\sum_{j_{n-1}\in\mathcal S} b
+		\mathbb P( L(\tau_{n-1}) = \ell_{n-1} \phi(\tau_{n-1}) = j_{n-1}, \tau_{n-1}\leq E^\lambda\mid \phi(\tau_1) = i, \tau_1\leq E^\lambda) \nonumber
 		\\&= b
@@
 	\end{align}
-	Partitioning on the time of the \(1\)st orbit restart epoch, \(\tau_1\), and the phase and level at time \(\tau_1\), then (\ref{eqn: alalal}) is equal to 
+	Partitioning on the time of the first orbit restart epoch, \(\tau_1\), and the phase and level at time \(\tau_1\), and using the strong Markov property, then (\ref{eqn: alalal}) is equal to 
 	\begin{align}
@@
 		\mid \bs Y(0)=(\ell_1,\bs \alpha, j_1)) \wrt t 
-		\\&\quad{}\int_{u_1=0}^\infty e^{-\lambda u_1}\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
+		\\&\qquad{}\times\int_{u_1=0}^\infty e^{-\lambda u_1}\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\in \wrt u_1
 		\mid \bs Y(0)=\bs y_0)F  \nonumber 
@@
 		\mid \bs Y(0)=(\ell_1,\bs\alpha,j_1)) \wrt t 
-		\\&\quad{}\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\leq E^\lambda
+		\\&\qquad{}\times\mathbb P(\Yd(1)=(\ell_1,j_1), \tau_1\leq E^\lambda
 		\mid \bs Y(0)=\bs y_0)F. \label{eqn fLksoioi}
@@
 
-	The expression 
+	Since \(\{\bs Y(t)\in(\ell,\calD_{\ell,j},j),t\in\mathcal T_n\}\subseteq \{t\in\mathcal T_n\}\), then the expression 
 	\begin{align}
@@
 	\end{align}
-	by Lemma~\ref{lem: another bound 2}.
+	where the second inequality holds by Lemma~\ref{lem: another bound 2}.
 
@@
 \end{align}
-Partitioning (\ref{eqn: LLLLaaakkkss}) on the number of orbit restart epochs by time \(t\) gives
+We start by partitioning (\ref{eqn: LLLLaaakkkss}) on the number of orbit restart epochs by time \(t\) to get
 \begin{align}
@@
 
-By Lemma (\ref{lem: LAkAKFnvnb mav h}), each term in the sum (\ref{eqn: KLDNVnav}) converges.
+By Lemma~\ref{lem: LAkAKFnvnb mav h}, each term in the sum (\ref{eqn: KLDNVnav}) converges.
 Furthermore, for \(n\geq 1\), each term is dominated by \(\left(b^{(p)}\right)^{n-1}F/\lambda\), from Corollary~\ref{vcor: cdks d}.
@@
 However, for \(p\) sufficiently large, there exists a \(p_0<\infty\) and a \(B\) with \(B<1\) such that \(b^{(p)}<B\) for all \(p>p_0\).
-Hence we can apply the Dominated Convergence Theorem to (\ref{eqn: KLDNVnav}) and claim that
+Hence, we can apply the Dominated Convergence Theorem to (\ref{eqn: KLDNVnav}) and claim that
 \begin{align}
@@
 \end{align}
-where we have used Lemma~\ref{lem: LAkAKFnvnb mav h}. Swapping the sum and integrals and by the law of total probability, then (\ref{eqn:Alakskjs}) is equal to 
+where Lemma~\ref{lem: LAkAKFnvnb mav h} gives the limits of the terms in the sum. Swapping the sum and integrals and by the law of total probability, then (\ref{eqn:Alakskjs}) is equal to 
 \begin{align}
@@
 \end{align}
-The swap of the sum and integrals is justfied as 
+The swap of the sum and integrals is justified as 
 \begin{align}
@@
 	% \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t 
-	\\&=\int_{t=0}^\infty e^{-\lambda t}\int_{\bs a\in\mathcal A}\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{x\in\calD_{\ell,j}}\mathbb P( {\bs Y}(t) \in (\ell,\wrt \bs a,j) \mid \bs Y(0)=\bs y_0)\bs a\bs v_{\ell,j}(x)  \psi(x)\wrt x\wrt t \nonumber
-	\\&=\int_{t=0}^\infty e^{-\lambda t}\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y} (t) \in (\ell,\wrt x,j) \mid \bs Y (0)=\bs y_0 )  \psi(x)\wrt t,
+	\\&=\int_{t=0}^\infty e^{-\lambda t}\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{\bs a\in\mathcal A}\int_{x\in\calD_{\ell,j}}\mathbb P( {\bs Y}(t) \in (\ell,\wrt \bs a,j) \mid \bs Y(0)=\bs y_0)\bs a\bs v_{\ell,j}(x)  \psi(x)\wrt x\wrt t \nonumber
+	\\&=\sum_{\substack{\ell\in\mathcal K\\j\in\calS}}\int_{t=0}^\infty e^{-\lambda t}\int_{x\in\mathcal D_{\ell,j}}\mathbb P( {\bs Y} (t) \in (\ell,\wrt x,j) \mid \bs Y (0)=\bs y_0 )  \psi(x)\wrt t,
 \end{align}
@@
 \begin{cor}\label{cor: lk}
-	Let \(\psi: \mathbb R\times \mathcal S \to \mathbb R\) be an arbitrary, bounded function with \(|\psi(\cdot)|\leq F\).
+	Let \(\psi: \mathbb R\times \mathcal S \to \mathbb R\) be an arbitrary, bounded function with \(|\psi(\cdot,\cdot)|\leq F\).
 %Further, let \(\mathcal L(X(t)) = \sum_{k\in\mathcal K}k1(X(t)\in\mathcal D_k)\) and \(\widetilde{\bs X}(t) = (\mathcal L(X(t)),X(t),\varphi(t)))\).
@@
 		%
-		\\&\leq F  \int_{t=0}^\infty e^{-\lambda t}\mathbb P( {\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j) \mid   
+		\\&= F  \int_{t=0}^\infty e^{-\lambda t}\mathbb P( {\bs Y}^{(p)}(t)\in(\ell,\mathcal D_{\ell,j},j) \mid   
 		\bs Y^{(p)}(0)=\bs y_0^{(p)})\wrt t \nonumber 
@@
 for almost all \(t\geq 0\).
-At such values of \(t\), since \(\psi\) is arbitrary and bounded, then the Portmanteau Theorem \citep[Theorem 2.1]{billingsleyconvergence} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
+At such values of \(t\), since \(\psi\) is arbitrary and bounded, then The Portmanteau Theorem (Theorem~13.16 of \cite{portmanteaubook}) states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
 
@@
 and the sequence \(\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \) is eventually equicontinuous in \(t\).
-That is, for every \(\varepsilon>0\) there exists a \(\delta(t,\varepsilon)>0\) and an \(p_0(t,\varepsilon)\) such that \(|t-u|<\delta(t,\varepsilon)\) implies that
+That is, for every \(\varepsilon>0\) there exists a \(\delta(t,\varepsilon)>0\) and a \(p_0(t,\varepsilon)\) such that \(|t-u|<\delta(t,\varepsilon)\) implies that
 \begin{align}
@@
 The main cause of the loose bound on the convergence rate is due to our reliance on Chebyshev's inequality.
-This is necessary as we try to make minimal assumptions about the matrix-exponential distributions used in the construction.
-In practice, we use a class of \emph{concentrated matrix-exponential distributions} found numerically in \citep{hht2020}, and for which there is relatively little known about their properties.
+This is necessary as we try to make minimal assumptions about the matrix exponential distributions used in the construction.
+In practice, we use a class of \emph{concentrated matrix exponential distributions} found numerically in \citep{hht2020}, and for which there is relatively little known about their properties.
 This necessitates the generality of the convergence result.
@@
 
-The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, i.e.~the hypotheses of the Theorem depend only on the interval \(\calD_{\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval, and not on any other interval.
+The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, i.e.~the hypotheses of the theorem depend only on the interval/level \({\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval/level, and not on any other interval/level.
 Thus, Lemma~\ref{thm: a thm!} holds independently on each interval, as does Corollary~\ref{cor: aln222}.
 
-Let the width of an interval \(\mathcal D_{\ell_0}\) be \(\Delta_{\ell_0}=y_{\ell_0+1}-y_{\ell_0}\) and suppose that sequence of matrix exponential random variables used to model the dynamics of the fluid queue on the interval \(\calD_{\ell_0}\) is \(Z_{\ell_0}^{(p)}\).
+Let the width of interval \({\ell_0}\) be \(\Delta_{\ell_0}=y_{\ell_0+1}-y_{\ell_0}\) and suppose that sequence of matrix exponential random variables used to model the dynamics of the fluid queue on the interval \({\ell_0}\) is \(Z_{\ell_0}^{(p)}\).
 Regarding the domination condition in Lemma~\ref{lem: another bound}, we can extend it the following version,
 \begin{lem}\label{lem: another bound sdfg}
-	Assume \(\inf_{\ell_0}\Delta_{\ell_0}>0\) and \(\sup_{\ell_0}\var(Z_{\ell_0})<\infty\). Then, for all \(i\in\mathcal S_+\cup \calS_-\), \(\ell_0\in\mathcal K\setminus\{-1,K+1\},\) and \(n\geq 2\), 
+	Assume \(\inf_{\ell_0}\Delta_{\ell_0}>0\) and \(\sup_{\ell_0}\var(Z_{\ell_0})<\infty\). Then, for all \(i\in\mathcal S_+\cup \calS_-\), \(\ell_0\in\mathcal K^\circ,\) and \(n\geq 2\), 
 	\begin{align}
@@
 	For the proof of (\ref{eqn: ell0 version}) follow the same arguments as in the proof of Lemma~\ref{lem: another bound}.
-The bound in (\ref{eqn: skjhg}) follows by the assumptions in the statement of the result, since
+The bound in (\ref{eqn: skjhg}) follows by the assumptions in the statement of the lemma, since
 	\begin{align*}

### 2022-02-17 chapter6/chapter.tex
first draft of the whole thing! :D

@@
 \chapter{Numerical investigations\label{sec: numerics}}
-We have now established multiple schemes for approximating fluid queues which are suitable for approximating the performance measures for fluid-fluid queues derived in \cite{bo2014}.
+We have now established multiple schemes for approximating fluid queues which are suitable for approximating the performance measures of fluid-fluid queues derived in \cite{bo2014}.
 In this chapter we numerically investigate various aspects of the approximation schemes.
-Throughout, we compare the QBD-RAP scheme (from Chapter~\ref{sec: construction and modelling}), the discontinuous Galerkin scheme (as described in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation as described in \cite{bo2013}, which we will refer to a just the \emph{uniformisation scheme}.
+Throughout, we compare the QBD-RAP scheme (from Chapter~\ref{sec: construction and modelling}), the discontinuous Galerkin scheme (as described in Chapter~\ref{ch:galerkin}), and the spatially-coherent uniformisation scheme introduced by \cite{bo2013}.
 %Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume scheme with an upwind flux.
@@
 
-For some numerical experiments we implement the discontinuous Galerkin scheme with and without a slope-limiter implemented \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
-The slope limiter avoids oscillations and negative solutions.
-However, we note that in the context of approximating the operator \(\mathbb\Psi\) for a fluid-fluid queue, it is not obvious how one might apply the concept of slope-limiting, other than to post-process the solution with a limiter or filter.
-Even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.
-Moreover, there is a computational cost in applying a limiter.
-Another issue is that the need for a slope-limiter, or filter, may depend on the initial condition.
-In contrast, for the uniformisation and QBD-RAP the approximate solution is guaranteed to produce positive probabilities without any postprocessing or extra computation.
-
 The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased.
 For simplicity, the number of basis functions is kept constant across all cells.
-We refer to the number of basis functions on a cell \(\calD_k\) as the \emph{dimension} and consider dimensions \(p=2r+1\) for \(r=0,1,2,3,...,10\).
+We refer to the number of basis functions on a cell \(k\) as the \emph{dimension} and consider dimensions \(p=2r+1\) for \(r=0,1,2,3,...,10\).
 For discontinuous Galerkin scheme, increasing the dimension of the scheme means increasing the number of polynomial basis functions used to approximate the solution within each cell.
@@
 Equivalently, we may think of a dimension \(p\) uniformisation scheme as using \(p\) piecewise constant functions to approximate the solution on each cell.
-For all schemes (DG, QBD-RAP and uniformisation), if we construct and order \(p\) approximation, there are \(K\) cells, \(N\) phases and \(c_i\neq 0\) for all \(i\in\calS\), then the resulting approximation to the generator \(\mathbb B\) is a matrix of dimension \(pKN + N\).
+For all schemes (DG, QBD-RAP and uniformisation), if we construct an order \(p\) approximation, there are \(K\) cells, \(N\) phases and \(c_i\neq 0\) for all \(i\in\calS\), then the resulting approximation to the generator \(\mathbb B\) is a matrix of dimension \(pKN + N\).
 In this sense each approximation scheme leads to matrices of the same size (although not necessarily the same number of non-zero elements).
@@
 We investigate schemes with an upwind flux only.
-For schemes which require us to integrate over time we do not investigate the stability of the schemes with respect to the \(t\)-step-size of the time-integration or time-integration scheme itself (where required).
+For schemes which require us to integrate over time we do not investigate the stability of the schemes with respect to the \(t\)-step-size, or the time-integration scheme itself (where required).
 Instead, we fix the time-integration step size for each numerical experiment at a suitably small value to obey a certain stability criterion (a CFL-like condition, \citep[Section~4.8]{nodalDGBook}).
-Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages \citep{sr2002}, \citep[Section~5.7]{nodalDGBook} (see also Appendix~\ref{sec: time integration}), which claims to introduce no more oscillations into the solution as we integrate over time.
-
-Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
+Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (\cite{sr2002}, \cite{nodalDGBook}~Section~5.7, see also Section~\ref{sec: time integration}), which claims to introduce no more oscillations into the solution as we integrate over time.
+
+For some numerical experiments we implement the discontinuous Galerkin scheme with and without a slope-limiter implemented (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
+The slope limiter avoids oscillations and negative solutions.
+However, we note that in the context of approximating the operator \(\mathbb\Psi\) for a fluid-fluid queue, it is not obvious how one might apply the concept of slope-limiting, other than to post-process the solution with a limiter or filter.
+Even when slope limiting is possible, the resultant approximation is at best linear around discontinuities.
+Moreover, there is a computational cost in applying a limiter, which could be significant, particularly if many initial conditions are to be considered.
+In contrast, for the uniformisation and QBD-RAP the approximate solution is guaranteed to produce positive probabilities without any post-processing or extra computation.
+
+Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
 We investigate two schemes with slope limiting.
-In the first scheme, which we refer to the \emph{DG-lim} scheme, we use a DG scheme of dimension \(p\) and apply the Generalised MUSCL slope limiter as necessary.
-In the second scheme, which we refer to as the \emph{DG-lin-lim} scheme, we follow the construction described in \ref{sec: anoth pos pres}, which is effectively a DG scheme with cells of width \(2\Delta/(p+1)\), for \(p=1,3,5,...,21\), and a basis of dimension \(2\) (linear) on each cell and apply the Generalised MUSCL limiter as necessary.
-Thus, the DG-lin-lim scheme uses \(p+1\), \(p=1,3,5,...,21\), basis functions to represent the solution on each of the original cells \(\calD_k\), and so we define the dimension of the scheme to be \(p+1\) to be comparable with the other schemes.
-We do not investigate filtering for the DG scheme (see \citep[Section~5.6.1]{nodalDGBook} and references therein).
-
-The performance of the discontinuous Galerkin scheme has been well-studied in some contexts \cite{c99}, \cite[Section~5.5]{nodalDGBook}, and it is well-known that the discontinuous Galerkin scheme performs remarkably well on problems with smooth solutions.
-Here, we mostly focus on investigating the numerical performance of the schemes on problems with non-smooth solutions, the purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP scheme.
+In the first scheme, which we refer to the \emph{DG-lim} scheme, we use a DG scheme of dimension \(p\) and apply the Generalised MUSCL slope limiter as described in Section~\ref{sec: slope limiting}.
+In the second scheme, which we refer to as the \emph{DG-lin-lim} scheme, we follow the construction described in~\ref{sec: anoth pos pres}, which is effectively a DG scheme with cells of width \(2\Delta/(p+1)\), for \(p=1,3,5,...,21\), and a basis of dimension \(2\) (linear) on each cell, with an upwind flux, and we apply the Generalised MUSCL limiter.
+Thus, the DG-lin-lim scheme uses \(p+1\), \(p=1,3,5,...,21\), basis functions to represent the solution on each of the original cells \(\calD_k\), and we define the dimension of the scheme to be \(p+1\).
+We do not investigate filtering for the DG scheme.
+
+The performance of the discontinuous Galerkin scheme has been well-studied in some contexts (\cite{c99}, \cite{nodalDGBook}~Section~5.5), and it is well-known that the discontinuous Galerkin scheme performs remarkably well on problems with smooth solutions.
+Here, we mostly focus on investigating the numerical performance of the schemes on problems with non-smooth solutions.
+The purpose of this is to emphasise the positivity-preserving properties of the QBD-RAP scheme.
 In the stochastic modelling community it is very common to have problems with discontinuities, such as a non-smooth initial conditions.
-Even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities or non-smooth behaviour into the problem (see, for example, Section~\ref{sec: transient approx}, below).
-A very specific set of conditions must hold for the initial density and point masses for the distribution to remain continuous as it evolves over time \citep{bo2014,bop2020}.
+As we will see (in Section~\ref{sec: transient approx}), even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities or non-smooth behaviour into the problem.
+A very specific set of conditions on the initial density and point masses must hold for the transient distribution to remain continuous as it evolves over time \citep{bo2014,bop2020}.
+Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues (see, for example, Section~\ref{sec: return approx}).
 Limiting distributions of fluid queues are smooth, however.
-Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues (see, for example, Section~\ref{sec: return approx}).
-
-As we shall see, if the solution is smooth, then the Galerkin scheme is highly-effective and displays rapid convergence to the solution as the number of polynomial basis functions is increased.
+
+As we shall see, if the solution is smooth, then the DG scheme is highly-effective and exhibits rapid convergence to the solution as the number of polynomial basis functions is increased.
 However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin scheme, leading to nonsense solutions.
@@
 
-All the following error plots are on a \(\log_{10}\)-\(\log_{10}\) scale.
+Lastly, before we start the numerical investigation, note that all the following error plots are on a \(\log_{10}\)-\(\log_{10}\) scale.
 If the error function takes the form \(error = \beta_0^* p^{\beta_1}\), where \(p\) is the dimension of the scheme, then the \(\log_{10}\)-\(\log_{10}\) will show a linear trend; \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\), where \(\beta_0=\log_{10}(\beta_0^*)\).
@@
 The structure of this chapter is as follows.
-In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different schemes and observe their performance at approximating the initial condition, which allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
-In Section~\ref{sec: wave num} we investigate a simple travelling wave problem with various initial conditions.
-For this problem the dynamics are deterministic, which allows us to instrument the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
+In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different schemes and observe their performance.
+This allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
+In Section~\ref{sec: wave num} we investigate travelling wave problems with various initial conditions.
+For the travelling wave problem the dynamics are deterministic, which allows us to instrument the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
 Next, we investigate a simple fluid queue with two phases.
-Section~\ref{sec:stat} investigates the ability of the schemes to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the fluid queue for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the schemes to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
-Section~\ref{sec: ffq num} approximates the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\) for two simple fluid-fluid queue models.
+Section~\ref{sec:stat} investigates the ability of the schemes to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the model for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the schemes to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
+Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num} and approximate the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\).
 
@@
 
-For the discontinuous Galerkin method, we project the initial condition on to the set of polynomial functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstructing the solutions as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
+For the discontinuous Galerkin method, we project the initial condition on to the set of polynomial functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstructing the solutions as described in Sections~\ref{sec: initial conditions} and~\ref{sec: closing}.
 For the purposes of approximation and reconstruction for the QBD-RAP scheme we must orientate the initial condition in a positive or negative direction; here, we suppose that the initial conditions belong to a positive phase.
-First we investigate three closing vectors we can use for the reconstruction for the QBD-RAP, from which we find that the closing operator in (\ref{eqn:5462}).
+First we investigate three closing vectors we can use for the reconstruction in the QBD-RAP scheme.
 From the investigation, we decide to use the closing vector (\ref{eqn:5462}) throughout the rest of the chapter.
@@
 To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a trapezoidal rule with 10,001 function evaluations.
-Similarly, we use 10,001 function evaluations to approximate \(L^p\) norms, and also as a finite set of points over which to compute the KS statistics.
+Similarly, we use 10,001 function evaluations to approximate \(L^p\) errors, and also as a finite set of points over which to compute the KS statistics.
 
@@
 We refer to the closing vector in (\ref{eqn: nonlin}) as the \emph{naive normalised} closing vector.
-Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme, though it clearly appears to converge.
+Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme.
 The third closing vector we investigate is that in Equation~(\ref{eqn:5462}) and we refer to this as the \emph{normalised closing operator}.
@@
 In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation with the three different closing vectors.
-Interestingly, in this case, the unnormalised error outperforms the other two reconstructions at low orders.
-It just so happens that, in this case, the error due to truncation of the unnormalised reconstruction destructively interferes with other errors to cause the error to be lower at low orders for the unnormalised scheme.
+Interestingly, in this case, the unnormalised closing operator outperforms the other two at low dimensions.
+In this case, it just so happens that the error due to truncation of the unnormalised reconstruction constructively interferes with other errors to cause the error to be lower at low dimensions for the unnormalised scheme.
+
 	Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection.
@@
 
-If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs}, then we see that the unnormalised reconstruction performs the poorest.
+If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} (right), then we see that the unnormalised reconstruction performs the poorest.
 Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests the naive normalised and normalised reconstruction are very similar.
@@
 
-Figure~\ref{fig: pdf reconstructed} also show that both closing operators do not approximate the initial condition well at the right-hand side of the interval.
+Figure~\ref{fig: pdf reconstructed} also show that both closing operators fail to approximate the initial condition well at the right-hand side of the interval.
 Perhaps there is a different reconstruction method or another closing operator which could alleviate this issue.
@@
 	Next we consider approximating the initial distribution with density \(-6x^2+6x\).
-Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the KS error against the dimension of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
-However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs}, which plots the \(L^2\) error between the reconstructed PDF and the true PDF, then the unnormalised closing operator performs better than the two normalised ones for dimension 5 and above.
+Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the KS error against the dimension of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
+However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the \(L^2\) error between the reconstructed PDF and the true PDF, then the unnormalised closing operator performs better than the two normalised ones for dimension 5 and above.
 \begin{figure}[h]
@@
 In Table~\ref{table: error rates for QBD-RAP closing} we summarise the rates of convergence of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots).
+In general, the normalised closing operators converges at a rate approximately as fast, or faster, than the unnormalised reconstruction.
 \begin{table}
@@
 In Figure~\ref{fig: fun 1 comp} we plot the KS metric (left) and \(L^1\) metric (right) between the true CDF and the reconstructed approximations.
-Observing the KS metric, it appears that none of the schemes converge and the KS error sits around 0.5.
-This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except, perhaps, at points of discontinuity.
-None of the schemes appear to converge at the discontinuity at \(x=0.5\).
+Observing the KS metric, it appears that none of the schemes converge and the KS error sits around 0.5 for all schemes.
+This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except, perhaps, at points of discontinuity and none of the schemes appear to converge at the discontinuity at \(x=0.5\).
 Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the schemes appear to show the convergent behaviour we expect.
@@
 Perhaps it is no surprise that the uniformisation scheme performs best.
-In the uniformisation scheme as we increase the order we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the initial distribution.
+In the uniformisation scheme as we increase the dimension we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the initial distribution.
 As such, the uniformisation scheme can produce a piecewise continuous, linear approximation to the CDF.
@@
 Here, the QBD-RAP scheme performs worst, the uniformisation scheme second, and the DG scheme the best.
-However, once again the DG scheme exhibits undesirable properties (plots not shown) -- the approximation to the CDF is at some points above 1 and is not monotonic (although these violations do not appear to be as severe in this case as they were for the example above).
+However, once again the DG scheme exhibits undesirable properties (plots not shown) -- the approximation to the CDF is at some points above 1 and is not monotonic (although these violations do not appear to be as severe in this case as they were for Example~\ref{ex: exa}).
 \begin{figure}[h]
@@
 In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDFs.
-Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG scheme can approximate the initial condition exactly.
+Since the DG method projects the initial condition onto a polynomial basis, then, for a dimension 3 approximation and above, the DG scheme can approximate the initial condition exactly.
 This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from dimension 1 to 3, then plateaus.
-Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, and due to machine arithmetic, the errors for the DG scheme are not 0.
+Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, the errors for the DG scheme are not 0.
 Regarding the other two schemes, they too appear to be convergent at approximately the same rate, with the uniformisation scheme performing better for the KS error, but very similarly in terms of the \(L^2\) error.
@@
 \begin{example}\label{ex: aalllkkdd}Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\).
-Figure~\ref{fig: fun 9 comp} shows the errors.
+Figure~\ref{fig: fun 9 comp} shows the KS and \(L^2\) errors for the PDF.
 Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs (right) tell a similar story; for sufficiently high order, the DG scheme approximates the initial condition very well.
@@
 
-\section{Travelling wave}\label{sec: wave num}
+\section{Travelling wave problems}\label{sec: wave num}
 Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with various initial conditions.
@@
 We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric.
-Furthermore, we use the SSPRK4 method to integrate over time with a \(t\)-step size of \(0.005\), and we evolve the system until time \(t=4\).\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} For the DG scheme we also implement the Generalised MUSCL slope limiter to stabilise the integration \citep{c99}, \citep[Section~5.6.2]{nodalDGBook} (see also Section~\ref{sec: slope limiting}).
+Furthermore, we use the SSPRK4 method to integrate over time with a \(t\)-step size of \(0.005\), and we evolve the system until time \(t=4\).\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} For the DG scheme we also implement the Generalised MUSCL slope limiter in the form of the DG-lim and DG-lin-lim schemes (see also Section~\ref{sec: slope limiting}).
 
@@
 \begin{align}
-	\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs}
-\end{align}
-for \(\ell=1,\dots,K\), and 
-\begin{align}
 	\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1) \label{eqn: cell probs boundary}
 \end{align}
-is the mass at the boundary and where \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,11,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).
+is the mass at the boundary, \(p(4,\ell,1)\) is an approximation to \(\mathbb P(X(4)\in\calD_{\ell,1}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\), and \(p(4,11,1)\) is an approximation to \(\mathbb P(X(4)\in\{10\}, \varphi(4)=1\mid X(0)=0.5,\varphi(0)=1)\).
 
@@
 Thus, in this case, there is no discretisation error in constructing the initial condition for the DG and uniformisation schemes.
-At time \(t=4\), the solution is \(f(x,4)=1(x\in[4,5))\).
+Moreover, at time \(t=4\), the solution is \(f(x,4)=1(x\in[4,5))\).
 The projections related to the DG-based schemes can represent this solution exactly too, hence we might expect these schemes to work well here.
@@
 
-Returning to Figure~\ref{fig: fun 4 wave}, the DG-lim scheme does not appear to converge with these error metrics.
+Continuing with Figure~\ref{fig: fun 4 wave}.
+The DG-lim scheme does not appear to converge with these error metrics.
 Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes, all appear to be converging with the QBD-RAP and  DG-lin-lim schemes converging similarly and faster than the uniformisation scheme.
@@
 In the first row we observe that the DG scheme (sans limiter) produces an approximation to the PDF with negative values when the dimension is greater than 1.
-In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter, the DG approximation does not change significantly after order 3.
+In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter, the DG approximation does not change significantly after dimension 3.
 This is due the fact that the DG scheme is at best linear in the presence of oscillations.
-In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation scheme, and in the last row is the solution approximated using the QBD-RAP scheme.
+In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation scheme, in the fourth row is the solution approximated using the QBD-RAP scheme, and in the last row is the approximation from the DG-lin-lim scheme.
 \begin{figure}[h]
@@
 	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row) and DG-lin-lim schemes, for order 1, 3, 5, and 7 (columns) for the travelling eave problem in Example~\ref{ex: wave 4}.
+	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row) and DG-lin-lim schemes, for dimension 1, 3, 5, and 7 (columns) for the travelling eave problem in Example~\ref{ex: wave 4}.
 The true density function is \(1(4\leq x<5)\).}
@@
 No PDF exists for the true distribution, so instead we compare the CDFs.
-Moreover, when we analysed reconstruction of this initial condition we saw that using the KS metric may be uninformative ue to the lack of point-wise convergence at the discontinuity.
-So, for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs) instead, and also with the cell-wise error from (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}).
+Also recall that, when we analysed reconstruction of this initial condition we saw that using the KS metric may be uninformative due to the lack of point-wise convergence at the discontinuity, so for this example, we measure errors by looking at the \(L^1\) error between the CDFs (the area between the CDFs) instead, and also with the cell-wise error from (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}).
 
 Figure~\ref{fig: fun 2 wave} (left) plots the \(L^1\) error between the true and approximated CDFs.
-The \(L^1\) metric tells a similar story to the previous analysis: the DG-lim scheme does not converge as order increases, the other three positivity preserving schemes (the uniformisation, DG-lin-lim and QBD-RAP schemes) appear to converge, with the QBD-RAP and DG-lin-lim schemes converging at a similar rate and faster than the uniformisation scheme.
-The DG scheme appears to converge the fastest.
+The \(L^1\) metric tells a similar story to the previous analysis: the DG-lim scheme does not converge as dimension increases, the other three positivity preserving schemes (the uniformisation, DG-lin-lim and QBD-RAP schemes) appear to converge, with the QBD-RAP and DG-lin-lim schemes converging at a similar rate and faster than the uniformisation scheme.
+The DG scheme performs the best.
 However, if we plot the approximations to the CDFs from the DG scheme (not shown) we would once again see an oscillating, non-monotonic function.
@@
 Furthermore, the DG scheme can result in negative estimates of the cell probabilities in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (plots not shown).
-Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme all appear to converge, with the QBD-RAP and DG-lin-lim schemes converging at a similar rate and faster than the DG scheme when the dimension of the basis used is sufficiently large.
-The estimated rate of convergence of the QBD-RAP and DG-lin-lim schemes is approximately -4 for this example, which is significantly faster than the rate of convergence of these schemes in all other examples in this section.
+Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme all appear to converge, with the QBD-RAP and DG-lin-lim schemes perform similarly and better than the DG scheme when the dimension is sufficiently large.
+The estimated rate of convergence of the QBD-RAP and DG-lin-lim schemes is approximately -4 for this example, which is significantly faster than the rate of convergence of these schemes in all other examples in this section (typically around -2 to -1).
 \begin{figure}[h]
@@
 
-	The other three positivity preserving schemes appear to converge, with the uniformisation scheme converging the slowest and the DG-lin-lim scheme converging the fastest of the three.
+	The other three positivity preserving schemes appear to converge, with the uniformisation scheme performing the worst and the DG-lin-lim scheme performing the best of the three for all error metrics.
 For the KS and \(L^2\) error on the PDFs in Figure~\ref{fig: fun 6 wave} the DG-lin-lim and QBD-RAP schemes perform similarly, however, for the cell-wise error in Figures~\ref{fig: fun 6 wave cp} the DG-lin-lim scheme performs significantly better than the QBD-RAP scheme.
-Given that the DG-lim-lim scheme performed similarly to the QBD-RAP scheme for the previous examples, and also for the other error metrics for this example, this is unexpected.
+Given that the DG-lin-lim scheme performed similarly to the QBD-RAP scheme for the previous examples, and also for the other error metrics for this example, this is unexpected.
 
 	To investigate the performance of the QBD-RAP and DG-lin-lim schemes, in Figure~\ref{fig: fun 6 wave pdfs} we plot the dimensions 21 QBD-RAP and dimensions 22 DG-lin-lim approximations to the PDF.
-Figure~\ref{fig: fun 6 wave pdfs} shows that the DG-lin-lim scheme approximates the PDF well, except for at the peak located at \(x=6.5\) -- I suspect the limiter has affected the solution in this region.
+Figure~\ref{fig: fun 6 wave pdfs} shows that the DG-lin-lim scheme approximates the PDF well, except for at the peak located at \(x=6.5\) (I suspect the limiter has affected the solution in this region).
 Specifically, the DG-lin-lim scheme underestimates the density in the peak, but overestimates the density either side of the peak.
-As a result, when we integrate over \(x\in[6,7]\) to compute the cell-wise probabilities these errors cancel to some extent and the DG-lin-lim scheme produces an accurate estimate of the probability in this cell.
-In comparison, the QBD-RAP scheme also underestimates the density at the peak, but also within the entire interval \([6,7]\), albeit less so at the edges.
+As a result, when we integrate over \(x\in[6,7]\) to compute the cell-wise probabilities, errors cancel over this region and the DG-lin-lim scheme produces an accurate estimate of the probability in this cell.
+In comparison, the QBD-RAP scheme also underestimates the density at the peak, but also within the entire interval \([6,7]\), albeit less so at the edges than the middle.
 The QBD-RAP scheme also overestimates the density in the adjacent cells, \([5,7]\) and \([7,8]\).
+Hence, the overall cell-wise error is large.
 	\begin{figure}[h]
@@
 
-	Figure~\ref{fig: fun 1 wave cp} plots the cell-wise error metric and Figure~\ref{fig: fun 1 wave} plots the KS error (right) and the \(L^2\) error between the PDFs (right).
-Once again, the cell-wise error metrics in Figure~\ref{fig: fun 1 wave cp} show similar characteristics to the finer-resolution error metrics in Figure~\ref{fig: fun 1 wave}.
-Observing Figures~\ref{fig: fun 1 wave cp} and \ref{fig: fun 1 wave}, the DG-lim scheme does not perform well, which might be expected due to the discontinuity in the transient distribution at \(x=4\) in Phase~1.
+	Figure~\ref{fig: fun 1 wave cp} plots the cell-wise error metric and Figure~\ref{fig: fun 1 wave} plots the KS error (left) and the \(L^2\) error between the PDFs (right).
+Observing Figures~\ref{fig: fun 1 wave cp} and~\ref{fig: fun 1 wave}, the DG-lim scheme does not perform well, which might be expected due to the discontinuity in the transient distribution at \(x=t<10\) in Phase~1.
 The uniformisation, DG-lin-lim and QBD-RAP schemes appear to converge, with the QBD-RAP scheme converging fastest, but performing similarly to the DG-lin-lim scheme, and the uniformisation scheme performs the poorest of the three.
-The DG scheme converges fastest, however, produces approximations with negative and oscillatory solutions as we might expect given the discontinuity.
+The DG scheme performs the best, however, produces approximations which are negative and oscillatory as we might expect given the discontinuity.
 A selection of approximations to the transient PDF are shown in Figure~\ref{fig: pdf wave fun 1}.
@@
 		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 1} where approximations were constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 1} where approximations are constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
@@
 
+In summary, we applied the DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes to travelling wave problems with various initial conditions.
+All the problems induced oscillations in the DG scheme at some point and as a result the DG-lim scheme reduced the order of the approximation to linear.
+In some cases, the oscillations in the DG scheme appear to be transient and were not material in the final approximation.
+In other cases, the oscillations were present in the final DG approximation causing the approximation to take unphysical values.
+The DG-lin-lim, uniformisation and QBD-RAP schemes are three positivity preserving alternatives.
+In general, the uniformisation scheme performs worse than the DG-lin-lim and QBD-RAP scheme, and appears to converge slowly to the true solution.
+The DG-lin-lim and QBD-RAP schemes perform similarly with the DG-lin-lim scheme performing slightly better for the smoother problems and the QBD-RAP scheme performing slightly better for the problems with more significant discontinuities and point masses.
+
 \FloatBarrier
@@
 Since the stationary distribution is smooth we expect that the DG scheme will work well here.
-The stationary distribution is convenient as we can evaluate them analytically \cite{s2017} and do not require us to approximate initial conditions or integrate over time.
+The stationary distribution is convenient for numerical experiments as we can evaluate it analytically \citep{s2017}, and it does not require us to approximate initial conditions or integrate over time.
 Hence, analysing the performance of the approximation schemes with the stationary distribution allows us to instrument the ability of the schemes to capture the stochastic dynamics, without numerical integration in time.
 
-Here we analyse a simple model which is based on Example~2 in \citep{bean2009}, except here we add a lower boundary to the model (no lower boundary is specified in Example~2 of \citep{bean2009} as it is inconsequential to their analysis).
+Here we analyse a simple model which is based on Example~2 in \cite{bean2009}, except here we add a lower boundary to the model (no lower boundary is specified in Example~2 of \cite{bean2009} as it is inconsequential to their analysis).
 
@@
 
-	Upon hitting the lower boundary, the process transitions from phase \(2\) to phase \(j\) with probability \(p_{2j}\) where 
+	Upon hitting the lower boundary, the process transitions from Phase \(2\) to Phase \(j\) with probability \(p_{2j}\) where 
 		\[\vligne{p_{21} & p_{22} } = \vligne{1 & 0}.\]
-		Upon hitting the upper boundary, the process transitions from phase \(1\) to phase \(j\) with probability \(p_{1j}\) where 
+		Upon hitting the upper boundary, the process transitions from Phase \(1\) to Phase \(j\) with probability \(p_{1j}\) where 
 		\[\vligne{p_{11} & p_{12} } = \vligne{0 & 1}.\]
@@
 % \paragraph{Model~\ref{model: simple}.}
-Figure~\ref{fig: reflecting stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
+Figure~\ref{fig: reflecting stationary} plots KS errors between the true stationary CDF and the approximations and the \(L^2\) error between the true stationary PDF and the approximations.
 Clearly the DG scheme is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
@@
 Once again we consider Model~\ref{model: simple} and use the same spatial discretisation as described in Section~\ref{sec:stat} (ten cells with width \(\Delta=1\)).
-Two initial conditions are considered, a point mass at 0 in phase 1, and the initial distribution with PDF
+Two initial conditions are considered, a point mass at 0 in Phase~1, and the initial distribution with PDF
 \begin{align}
@@
 in phases 1 and 2, with no mass at the boundaries.
-We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{Once again, the \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We apply the DG without a limiter and also the DG-lim and DG-lin-lim schemes which utilise a Generalised MUSCL slope limiter.
-
-To obtain a \emph{ground truth} 5,000,000 realisations of the fluid queue were simulated until \(t=2\), then the empirical CDF and the masses within each cell and at the boundaries were computed from the simulations.
+We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We apply the DG without a limiter and also the DG-lim and DG-lin-lim schemes which utilise a Generalised MUSCL slope limiter.
+
+To obtain a \emph{ground truth} 5,000,000 realisations of the fluid queue were simulated until \(t=2\), then the empirical CDF and the masses within each cell were computed from the simulations.
 We do not attempt to numerically approximate the true PDF via simulation.
-We then compute the KS and \(L^1\) error metrics between the approximated CDF and simulated CDF, as well as the cell-wise error metric, given as follows.
+We then compute the KS and \(L^1\) error metrics between the approximated and simulated CDFs, as well as the cell-wise error metric, given as follows.
 \begin{align}
@@
 To account for possible Monte-Carlo error, we used a bootstrap with 1,000 bootstrap samples.
-We sample 5,000,000 realisations of the fluid queue with replacement from the original 5,000,000 samples, then compute error metrics with the resampled data.
-We resample 1,000 times.
+For the bootstrap we sample 5,000,000 realisations of the fluid queue with replacement from the original 5,000,000 samples, then compute error metrics with the resampled data.
+We resample 1,000 times, resulting in 1,000 estimates of the errors.
 Via the bootstrap, we report the \(5\)th and \(95\)th percentile of the sampling distribution of the errors.
@@
 The uniformisation, DG-lin-lim and QBD-RAP schemes converge at a slower rate than the DG scheme, with the QBD-RAP and DG-lin-lim schemes converging faster than the uniformisation scheme.
-For the KS metric, the DG-lin-lim scheme outperforms the QBD-RAP scheme at all orders, while for the \(L^1\) metric, the two schemes perform similarly.
+For the KS metric, the DG-lin-lim scheme outperforms the QBD-RAP scheme at all dimensions, while for the \(L^1\) metric, the two schemes perform similarly.
 
 The DG-lim does not appear to be converging, which suggests there is at least one iteration during the numerical integration over time at which the numerical solution displays oscillations.
-However, when the DG approximations for the distribution at time \(t=2\) are plotted, they do not display oscillations (not shown).
+However, when the DG approximations for the distribution at time \(t=2\) are plotted, they do not appear to display oscillations (not shown).
 This suggests that the oscillations which might occur with the DG scheme are transient, and have dissipated by time \(t=2\).
@@
 Comparing the error metrics in Figure~\ref{fig: reflecting transient pm} for the point mass initial condition with the ones in Figure~\ref{fig: reflecting transient exp} for the exponential initial condition all schemes perform worse for the point mass initial condition.
-Regarding comparative convergence of error for this problem, for both error metrics in the DG scheme converges fastest, followed by the QBD-RAP and DG-lin-lim schemes which converge comparatively, then the uniformisation scheme, while the DG-lim does not appear to converge due to the limiter reducing the scheme to a linear approximation around the discontinuity.
+Regarding comparative convergence of error for this problem, for both error metrics the DG scheme converges fastest, followed by the QBD-RAP and DG-lin-lim schemes which converge comparatively, then the uniformisation scheme, while the DG-lim does not appear to converge due to the limiter reducing the scheme to a linear approximation around the discontinuity.
 %
@@
 Figure~\ref{fig: reflecting transient pm cp} suggests that approximating the cell-wise error seems to be a difficult problem.
-This is likely caused by the discontinuity at \(x=2\) in phase \(1\), which lies exactly on a cell boundary.
+This is likely caused by the discontinuity at \(x=2\) in Phase \(1\), which lies exactly on a cell boundary.
 To investigate how the position of the discontinuity might affect the error metrics we evolved the model to time \(t=2.1\) and computed the KS error, \(L^1\) error between the CDFs and the cell-wise error.
@@
 
+In summary, for the exponential initial condition the DG scheme performed well, producing the lowest errors.
+This was followed by the DG-lin-lim scheme, then the QBD-RAP scheme, then the uniformisation scheme, while the DG-lim scheme did not perform well at all.
+The fact that the DG-lim scheme did not perform well suggests that there must have been oscillations in the DG solution at some point, although these oscillations are not obvious in the DG approximation at time \(t=2\).
+In general the methods all performed better for the exponential initial condition than for the point mass initial condition.
+For the point mass initial condition the DG scheme produced oscillatory solutions.
+Of the positivity preserving methods, the DG-lim scheme performed poorly and did not appear to be converging, the uniformisation scheme appears to be converging but at a slower rate than the DG-lin-lim and QBD-RAP schemes.
+Of the two most performant positivity preserving schemes, the QBD-RAP scheme performed slightly better for the point mass initial condition, while the DG-lin-lim scheme performed slightly better for the exponential initial condition.
+
 \FloatBarrier
@@
 for some initial distribution \(\mu\). We look at two initial conditions; an exponential with equal mass in each phase, 
-\[\mathbb P(X(0)\in\wrt x,\varphi(0)=i) = e^{-x}/(1-e^{-1})/2,\,i\in\{1,2\},\]
+\[\mathbb P(X(0)\in\wrt x,\varphi(0)=i) = \cfrac{1}{2}\cfrac{e^{-x}}{(1-e^{-1}},\,i\in\{1,2\},\]
 and a point mass at \(X(0)=0\) in phase \(\varphi(0)=1\).
@@
 We use the DG, uniformisation and QBD-RAP schemes to discretise the fluid queue and partition \([0,1]\) into three intervals of width \(1/3\).
-To capture the mass which has left the interval \([0,1]\), we suppose that when the process hits the boundary it is absorbed forever at the boundary and in the phase in which the process first hit the boundary.
+To capture the mass which has hit the set \(\{0,1\}\), we suppose that when the process hits \(\{0\}\) or \(\{1\}\) it is absorbed forever and remains in the phase in which the process first hit \(\{0\}\) or \(\{1\}\).
 
@@
 
-As a \emph{ground truth} we simulated 5,000,000 realisations and recorded the hitting time on the boundary of the interval \([0,1]\) and the phase at the time of hitting.
+As a \emph{ground truth} we simulated 5,000,000 realisations and recorded the hitting time on the set \(\{0,1\}\) and the phase at the time of hitting.
 We then compute the empirical CDF of the hitting probabilities
@@
 To account for Monte-Carlo errors we took 1,000 bootstrap resamples of the original 5,000,000 samples and computed the empirical CDF of the hitting probabilities for each bootstrap sample.
-For each bootstrap sample, we resampled 5,000,000 points with replacement from the original 5,000,000 realisations.
-For each bootstrap sample we computed the error metrics between the empirical and approximated CDFs and recorded the 5th and 95th percentile of the distribution of the errors.
+For each bootstrap sample, we resampled 5,000,000 points with replacement from the original 5,000,000 realisations, then computed the error metrics between the empirical and approximated CDFs and, ultimately, estimated the 5th and 95th percentile of the distribution of the errors.
 
@@
 For both error metrics the DG scheme performs the best.
-For the \(L^1\) error metric the DG scheme converges up to order 5, after which there is no improvement in the error; I believe that this is due to other numerical errors dominating.
-The DG-lim does not appear to converge, suggesting oscillations in the DG approximation.
+For the \(L^1\) error metric the DG scheme converges up to dimension 3, after which there is little improvement in the error; I believe that this is due to other numerical errors dominating.
+The DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation.
 Intuitively, with this initial condition we suspect that the transient distribution of the fluid queue on the event that it remains in the interval \((0,1)\) is discontinuous at \(x=t\) in Phase~1 and \(x=1-t\) in Phase~2 for \(x,t<1\).
 We also expect there is a discontinuity in the hitting time PDF at time \(t=1\).
-Regarding the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme, all appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly and converging at a faster rate than the uniformisation scheme.
-
-In Figure~\ref{fig: hitting time oscillation exp} we plot the PDFs of the hitting times \(\xi_x(\{0,1\})\), in each phase, for the dimension 21 DG, dimension 21, QBD-RAP and dimension 22 DG-lin-lim schemes.
+Regarding the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes, all appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly and converging at a faster rate than the uniformisation scheme.
+
+In Figure~\ref{fig: hitting time oscillation exp} we plot the PDFs of the hitting times \(\zeta_X(\{0,1\})\), in each phase, for the dimension 21 DG, dimension 21, QBD-RAP and dimension 22 DG-lin-lim schemes.
 Figure~\ref{fig: hitting time oscillation exp} demonstrates oscillations in the DG approximation, though the PDF is not negative in this case.
@@
 and since the sample path will hit \(\{0\}\) at \(v=2u\), intuitively, we would expect (\ref{eqn :lf}) to approximate a point mass at \(v=2u\).
-To see this, recall that we can approximate a point mass at \((v-u)\) by
+To see that this is the case, recall that we can approximate a point mass at \((v-u)\) by
 \begin{equation}\bs k(\Delta - (v-u))\bs e^{\bs Sx}\bs s,\label{eqn:aafba}\end{equation}
@@
 which approximates a point mass at \(u=v-u\), or \(2u=v\).
-Retracing our arguments, sources of error in this approximation might come from the approximation in (\ref{eqn: aegpb}) and from the approximation of a point mass by (\ref{eqn:aafba}), which we commented on in Section~\ref{sec: closing vecs numerics} where we conducted numerical experiments about the unnormalised closing operator given by \(\bs v(x) = e^{\bs S x}\bs s\).
+Retracing our arguments, sources of error in this approximation come from the approximation in (\ref{eqn: aegpb}) and from the approximation of a point mass by (\ref{eqn:aafba}), which we commented on in Section~\ref{sec: closing vecs numerics} where we conducted numerical experiments about the unnormalised closing operator given by \(\bs v(x) = e^{\bs S x}\bs s\).
+Indeed, we saw that the unnormalised closing operator performed poorly for the exponential initial condition in Section~\ref{sec: closing vecs numerics}.
 
@@
 The DG scheme appears to perform best, however, oscillations are present in the DG approximation.
-In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis and for the DG-lin-lim scheme with dimension 22.
+In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis, and for the DG-lin-lim scheme with dimension 22.
 Observing Figure~\ref{fig: hitting time oscillation} the DG approximation displays oscillations around and the discontinuity at \(t=1\) while the other two schemes do not.
-Of the two positivity preserving schemes in Figure~\ref{fig: hitting time oscillation}, then QBD-RAP scheme appears to approximate the discontinuity best.
+Of the two positivity preserving schemes in Figure~\ref{fig: hitting time oscillation}, the QBD-RAP scheme appears to approximate the discontinuity best.
 \begin{figure}[h]
@@
 
+In summary, the results of this experiment are somewhat similar to the last experiment on transient distributions; the DG-lim scheme performs poorly, the DG scheme exhibits oscillations, of the three other positivity preserving schemes, the uniformisation scheme performs poorest, while the QBD-RAP and DG-lin-lim schemes perform similarly, with the QBD-RAP scheme performing slightly better for the point mass initial condition and the DG-lin-lim scheme performing slightly better for the exponential initial condition.
+
 \FloatBarrier
 \section{First-return distributions of fluid-fluid queues}\label{sec: ffq num}
-Now consider the fluid-fluid queue model first analysed in \cite{blnos2022} which is a modified version of a model first presented in \cite{lnp13}.
+Now consider the fluid-fluid queue first analysed in \cite{blnos2022} which is a modified version of a model first presented in \cite{lnp13}.
 We refer the reader to \cite{blnos2022} for more on the DG scheme applied to this model.
 \begin{model}\label{model: ffq}
-	Consider a stochastic fluid-fluid queue $\{(X(t),Y(t),\varphi(t))\}_{t\geq0},$ where $\{X(t)\}$ and $\{Y(t)\}$ represent the workloads in Buffers~1 and~2 at time $t \geq 0$, respectively, both driven by the phase $\{\varphi(t)\},$ which is a Markov chain on the state space $\mathcal{S} = \{11,10,01,00\}$.
+	Consider a stochastic fluid-fluid queue $\{(Y(t),X(t),\varphi(t))\}_{t\geq0},$ where $\{X(t)\}$ and $\{Y(t)\}$ represent the workloads in Buffers~1 and~2 at time $t \geq 0$, respectively, both driven by the phase $\{\varphi(t)\},$ which is a Markov chain on the state space $\mathcal{S} = \{11,10,01,00\}$.
 Both $\{X(t)\}$ and $\{Y(t)\}$ have a regulated boundary at 0.
@@
 		\end{align} 
+	We consider the initial distribution which is a point mass at \(Y(0)=0,\, X(0)=5,\, \varphi(0)=01\). 
 \end{model}
@@
 While the true problem has an unbounded domain $[0,\infty)$, the approximations require the domain of approximation to be a finite interval.
-Here we choose an upper bound of \(48\) and place a regulated boundary at the upper boundary.
-The effect of this truncation can be partly quantified by evaluating \(\lim\limits_{t\to\infty}\mathbb P\left(X(t) > 48\right)\approx 5.83 \times 10^{-9}\), \(i\in\calS\).
+Here we choose an upper bound of \(b=48\) and place a regulated boundary at the upper boundary.
+The effect of this truncation can be partly quantified by evaluating \(\lim\limits_{t\to\infty}\mathbb P\left(X(t) > 48\right)\approx 5.83 \times 10^{-9}\).
 
 We obtained approximations to the generator of the fluid queue \(\{(X(t),\varphi(t))\}\) via the DG, QBD-RAP and uniformisation schemes, all of which used a cell width of \(\Delta=0.4\).
-The matrix resulting from the approximation is then used to approximate the first-return operator \(\mathbb\Psi(s)\) as discussed in Section~\ref{sec: intro Psi}.
+The approximate generator matrix resulting from the approximation is then used to approximate the first-return operator \(\mathbb\Psi(s)\) as discussed in Section~\ref{sec: intro Psi}.
 Due to the stochastic interpretation of the uniformisation and QBD-RAP schemes the approximations to \(\mathbb \Psi(s)\) have a stochastic interpretation as the first-return probabilities of a fluid queue driven by a CTMC and QBD-RAP, respectively (see \cite[Chapter~7]{p2019} and \cite{bgnp2021} for details on the latter).
@@
 We use a grid of 10,001 points on which we evaluate the approximations of the CDF in each phase.
-We consider first the initial distribution which is a point mass at \(Y(0)=0,\, X(0)=5,\, \varphi(0)=01\).
-
-For comparison, we simulated 5,000,000 realisations of the fluid-fluid queue and recorded the value of \(X\) and \(\varphi\) at the time of first return of the second fluid, \(Y\).
+
+For a `ground truth' comparison, we simulated 5,000,000 realisations of the fluid-fluid queue and recorded the value of \(\{X(t)\}\) and \(\{\varphi(t)\}\) at the time of first return of the second fluid, \(\{Y(t)\}\).
 The empirical approximation of (\ref{eqn: first return Y 1}) was then constructed, and error metrics for the difference between the empirical CDF and the approximations was computed.
-To account for Monte-Carlo errors we used a bootstrap with 1,000 bootstrap samples to construct 1,000 bootstrap samples of the error estimates and recorded the 5th and 95th percentiles of the error distribution.
+To account for Monte-Carlo errors we used a bootstrap with 1,000 bootstrap samples to construct 1,000 bootstrap samples of the error estimates and ultimately estimated the 5th and 95th percentiles of the error distribution.
 Each of the 1,000 bootstrap samples was constructed by resampling the original 5,000,000 realisations 5,000,000 times with replacement.
 
-In Figure~\ref{fig: ffq return cts} we plot the error metrics for the approximations to the distribution (\ref{eqn: first return Y 1}).
+In Figures~\ref{fig: ffq return cts} and~\ref{fig: ffq cell probs} we plot the error metrics for the approximations to the distribution~(\ref{eqn: first return Y 1}).
 The DG scheme performs best converging rapidly until the error in the approximation scheme is swamped by other numerical errors.
@@
 Here, the first return distribution appears to be smooth, hence we might expect the DG scheme to perform well.
+
 Note that there is a significant number other sources of error here; machine precision errors, errors in solving the matrix Riccati equation to approximate \(\mathbb \Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), and truncation errors.
@@
 \exampleFloatBarrier
+
 By modifying slightly Model~\ref{model: ffq}, we can construct a first return distribution which is discontinuous. 
@@
 \end{model}
-As before, we use the DG, uniformisation and QBD-RAP schemes to approximate the model, and compare to simulations.
+As before, we use the DG, uniformisation and QBD-RAP schemes to approximate the model, and compare the approximations to simulations.
 
@@
 
-Figure~\ref{fig: ffq return discts} plots the error metrics for the first return distribution.
+Figures~\ref{fig: ffq return discts} and~\ref{fig: ffq2 cell probs} plot the error metrics for the first return distribution.
 Observing Figure~\ref{fig: ffq return discts} we see that the DG scheme performs well, followed by the QBD-RAP scheme, then the uniformisation scheme performs worst.
@@
 \end{figure}
+
+In summary, for the smooth problem the DG scheme was highly effective, significantly outperforming the positivity preserving QBD-RAP and uniformisation methods.
+For the discontinuous problem, the DG scheme displayed oscillations, but produced the lowest errors.
+For both problems, of the positivity preserving methods (QBD-RAP and uniformisation scheme), the QBD-RAP scheme performed the best.
+
 \FloatBarrier
@@
 In this chapter we have numerically investigated some properties of the QBD-RAP approximation scheme and compared it to existing schemes; the uniformisation scheme of \cite{bo2013} and the discontinuous Galerkin scheme.
-In general, the numerical experiments demonstrate the smoother the problem is the better the performance of the DG scheme, and it emphatically outperforms the other two schemes.
+In general, the numerical experiments that the smoother the problem is the better the performance of the DG scheme, and it emphatically outperforms the other two schemes for smooth problems.
 However, for problems with discontinuities the DG approximation can exhibit oscillations, while the QBD-RAP and uniformisation approximations avoid this and, of the latter two, the QBD-RAP scheme often converges faster.
+
 To avoid the problems of oscillations we can sometimes employ a \emph{slope limiter} with the DG scheme, which effectively reduces the scheme to linear in the regions where oscillations are detected.
-In some contexts, we implemented two slope-limited DG schemes, the DG-lim scheme which takes a high order DG scheme and limits the solution as necessary, and the DG-lin-lim, which is a linear approximation on a finer grid and is designed to use approximately the same computational resources as the other schemes we consider.
-The numerical experiments demonstrate a loss of accuracy in the approximation when a DG-lim scheme is used for a purely discontinuous problem.
+In some contexts, we implemented two slope-limited DG schemes, the DG-lim scheme which takes a high order DG scheme and limits the solution as necessary, and the DG-lin-lim, which is a linear approximation on a finer grid designed to use approximately the same computational resources as the other schemes we consider.
+The numerical experiments demonstrate a significant loss of accuracy in the approximation when a DG-lim scheme is used for discontinuous problems.
 The numerical experiments suggest that the DG-lin-lim scheme can perform well, and is similar to the performance of the QBD-RAP scheme, in the presence of discontinuities.
@@
 As a first step in the numerical experiments, in Section~\ref{sec: recon num}, we examined the ability of each scheme to approximate different initial conditions.
-For the DG scheme, this is equivalent to a projection of the initial condition on to a set of basis polynomials, for the uniformisation scheme this is equivalent to projecting the initial condition on to a basis of constant functions, and for the QBD-RAP scheme the approximation of the initial condition is as described in Sections~\ref{sec: initial conditions} and \ref{sec: closing}.
+For the DG scheme, this is equivalent to a projection of the initial condition on to a set of basis polynomials, for the uniformisation scheme this is equivalent to projecting the initial condition on to a basis of constant functions, and for the QBD-RAP scheme the approximation of the initial condition is as described in Sections~\ref{sec: initial conditions} and~\ref{sec: closing}.
 Section~\ref{sec: comp} demonstrates that the DG scheme (projection) can result in oscillations and negative regions in the approximation when the initial condition is discontinuous.
@@
 However, when the initial condition to be approximated is sufficiently smooth, then the DG approximation is superior.
+The QBD-RAP scheme performed the poorest for the examples considered.
 
@@
 For this model the solution is given in terms of the initial condition by \(f(x,t) = f(x-t,0)\).
-This model is useful as it allows us to instrument the ability of the schemes to capture the flow of probability without stochastic dynamics, and because the solution is known.
+This model is useful as it allows us to instrument the ability of the schemes to capture the flow of probability without stochastic dynamics, and since the solution is known, so there is no need for simulation.
 With the travelling wave model we demonstrate that, for problems with discontinuities, the DG scheme can display oscillations, while the other schemes (DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes) avoid oscillations.
@@
 Of the uniformisation and QBD-RAP schemes, the QBD-RAP scheme gives more accurate solutions.
-In Section~\ref{sec: transient approx} we turned our attention to approximating transient distributions for the same models and consider two different initial conditions, a point-mass and an exponential initial condition.
+
+In Section~\ref{sec: transient approx} we turned our attention to approximating transient distributions for the same model and consider two different initial conditions, a point-mass and an exponential initial condition.
 The discontinuous initial condition results in a discontinuous transient distribution.
@@
 The numerical results suggest that, due to the discontinuities in the problems, the DG scheme may display oscillations.
-Since the uniformisation DG-lin-lim, and QBD-RAP schemes can handle discontinuities, they perform as expected, with the QBD-RAP and DG-lin-lim schemes performing similarly and better than the uniformisation.
+Since the uniformisation, DG-lin-lim, and QBD-RAP schemes can handle discontinuities, they perform suitably, with the QBD-RAP and DG-lin-lim schemes performing similarly and better than the uniformisation.
 
@@
 The slope limiter overcomes this, but reduces the accuracy of the DG scheme to linear near discontinuities, sometimes severely affecting the quality of the approximation.
-A peicewise-linear DG scheme with a limiter, such as the DG-lin-lim scheme considered here, can produce satisfactory results.
-The uniformisation and QBD-RAP schemes are other alternative approximation schemes which can exhibit larger errors, but avoid oscillatory solutions and guarantee a non-decreasing approximation of the CDF and have a stochastic interpretation.
+A piecewise-linear DG scheme with a limiter, such as the DG-lin-lim scheme considered here, can produce satisfactory results.
+The uniformisation and QBD-RAP schemes are other alternative approximation schemes which avoid oscillatory solutions and guarantee a non-decreasing approximation of the CDF and have a stochastic interpretation.
 Of the uniformisation and QBD-RAP schemes, the latter often produced lower errors and performed similarly to the DG-lin-lim scheme.
+Moreover, the uniformisation and QBD-RAP schemes have interpretations which can aid in their analysis (such as justifying the Riccati equation to approximate \(\mathbb \Psi\) for a fluid-fluid queue).
+The uniformisation and QBD-RAP schemes also result in linear approximations, whereas the DG-lim and DG-lin-lim schemes do not and this can be beneficial in certain contexts.

### 2022-02-17 chapter7/chapter.tex
first draft of the whole thing! :D

@@
 %!TEX root = ../thesis.tex
-\chapter{Conclusions \label{ch: conclusion}}
+\chapter{Conclusion\label{ch: conclusion}} 
+A fluid-fluid queue is a stochastic fluid queue, where the driving process is a fluid queue itself.
+Fluid queues provide a model for a single continuous performance measure of a system in the presence of a random environment.
+Fluid queues have found a wide variety of applications including risk processes, telecommunications, and environmental modelling, among others.
+Given the success of fluid queues it is believed that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success.
+\cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and stationary distribution of a fluid-fluid queue.
+
+Motivated by computation of the theoretical operators for stochastic fluid-fluid queues in \cite{bo2014}, this thesis investigates approximations method for the generator of stochastic fluid queues.
+Cell-based approximation methods are particularly useful as they allow flexible partitioning of the approximate operators, which is required in the analysis of fluid-fluid queues.
+
+In Chapter~\ref{ch:galerkin} we introduced a discontinuous Galerkin (DG) scheme for the approximation of the generator of a fluid queue and show how to approximate the operators from \cite{bo2014}.
+High-order DG schemes are known to perform well for smooth problems, but problems can occur in the presence of discontinuities, such as negative probability estimates.
+In some contexts, slope limiting can be used to rectify this, but they result in approximations to operators which are non-linear which complicates the computation of the operators in \cite{bo2014}.
+The uniformisation scheme of \cite{bo2013}, which is equivalent to the simplest DG scheme, does not give erroneous approximations in the presence of discontinuities but can converge slowly compared to high-order DG schemes.
+
+In Chapter~\ref{sec: construction and modelling}, we introduce a new approximation to a fluid queue.
+Inspired by the uniformisation scheme (which is a QBD process) and its ability to handle discontinuous solutions, we construct a new discretisation of a fluid queue in the form of a QBD-RAP.
+Chapter~\ref{sec: construction and modelling} describes the construction of the approximation and the intuition behind it.
+
+In Chapters~\ref{sec: conv} and~\ref{ch: global results} we prove that the QBD-RAP approximation scheme converges weakly (in the spatial and temporal variables) to the distribution of the fluid queue.
+Chapters~\ref{sec: conv} uses matrix-exponential-specific arguments in its proof and ultimately proves the convergence of the QBD-RAP scheme to the fluid queue when the latter remains within a given interval.
+Chapter~\ref{ch: global results} uses more traditional Markov process arguments to prove a global convergence result for the approximation scheme.
+
+Chapter~\ref{sec: numerics} investigates, via numerical experiments, the performance of the DG, QBD-RAP, and uniformisation schemes.
+In some contexts, we also implement two positivity preserving DG schemes which utilise a limiter.
+For smooth problems the DG scheme was superior, however, for discontinuous problems, the DG scheme exhibited oscillations and produced negative probability estimates.
+One of the slope limited DG schemes, in which a slope limiter is applied to a high-order DG scheme (which we title the DG-lim scheme), performed very poorly due to the limiter reducing the order of the scheme to linear near discontinuities.
+The QBD-RAP scheme performed similarly to the linear DG approximation on a finer grid with a slope limiter (the DG-lin-lim scheme).
+The QBD-RAP scheme performed slightly better on problems with point masses, and the DG-lin-lim scheme performed slightly better on the smoother problems.
+However, the QBD-RAP scheme constructs an approximation to the generator of the fluid queue which is linear and this is advantageous for the application to fluid-fluid queues.
+The uniformisation scheme performed better than the DG-lim scheme but worse than all the others.
+
+In summary, this thesis investigated three main computational frameworks for the analysis of fluid queues and fluid-fluid queues, namely, the uniformisation, DG, and QBD-RAP schemes.
+The uniformisation scheme is from \cite{bo2013}.
+The DG method is well known, but its application to fluid and fluid-fluid queues has not been well-studied.
+This thesis introduces the DG scheme in the context of fluid and fluid-fluid queues.
+The QBD-RAP scheme is a novel approach which overcomes some known issues of the DG scheme, and converges faster than the uniformisation scheme.
+Analysis of the QBD-RAP scheme proves that it is convergent.
+Numerical experiments demonstrate the effectiveness of the approximation schemes for various problems.

### 2022-02-17 front/abstract.tex
first draft of the whole thing! :D

%!TEX root = ../thesis.tex
\chapter{Abstract}
\label{ch:abstract}
A fluid-fluid queue is a stochastic fluid queue, where the driving process is a fluid queue itself. Fluid queues provide a model for a single continuous performance measure of a system in the presence of a random environment. Fluid queues have found a wide variety of applications including risk processes, telecommunications, and environmental modelling, among others. Given the success of fluid queues it is believed that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success. \cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and stationary distribution of a fluid-fluid queue.

This thesis provides approximations to fluid queue so that we can approximate the operators in \cite{bo2014}. It investigates three main approximation schemes; the DG scheme (Chapter~\ref{ch:galerkin}) which is a popular finite-element scheme, the uniformisation scheme of \cite{bo2013} which approximates a fluid queue but a continuous-time Markov chain (specifically, a quasi-birth-and-death-process (QBD)), and the QBD-RAP scheme which is a generalisation of a QBD to allow matrix exponential inter event times. The QBD-RAP scheme is novel; we describe the construction of the scheme in Chapter~\ref{sec: construction and modelling}, and provide an analysis to show that it is convergent in Chapters~\ref{sec: conv} and \ref{ch: global results}. We demonstrate the effectiveness of the approximation schemes in Chapter~\ref{sec: numerics}, focussing on problems with discontinuous solutions. In general, we find that the DG scheme performs remarkably well for smooth problems, but can produce oscillations and negative probability estimates in the presence of discontinuities, the QBD-RAP approximation performs well in the presence of discontinuities, but does not perform as well as the DG scheme for smooth problems, the uniformisation scheme produces reliable approximations in the presence of discontinuities, but converges slowest for all problems.

### 2022-02-17 front/acknowledgements.tex
first draft of the whole thing! :D

%!TEX root = ../thesis.tex
\chapter{Acknowledgements}
\label{ch:acknowledgements}

First and foremost, to my principle supervisor, Nigel Bean. Thanks for all the support, the interest you have taken in this work, and the faith you have shown in me allowing me to essentially do whatever I want for this thesis. The effort you have put in to supervising me has been much appreciated.

To Peter Taylor, my co-supervisor, thanks for joining me on this quest. Thanks for the insightful discussions on maths, cricket and all the other important happenings of the world during our meetings. You've provided an invaluable perspective on my work and on mathematics as a science.

To Giang Nguyen. You were the one who initially inspired me to take up mathematics research which led me to where I am now. You gave me so much of your time and went well above and beyond to support me. Thank you.

To my friend Caitlin, thanks for all the support, coffee, and company, over the last five years. It has been so enjoyable. I will miss it.

To my friends Dennis, Phil and Andrew. I can't understate the value of this group (and Caitlin too). I've been lucky to be able to share ideas, get advice, and receive mentorship and support from all of you.

To my wife, Alice. Thank you for your support over the last five years. I've learnt so much about life from you.

### 2022-02-17 front/declaration.tex
first draft of the whole thing! :D

%!TEX root = ../thesis.tex
\chapter{Declaration}
\label{ch:declaration}

This thesis contains work taken from the publication \cite{blnos2022} which approximates the operator analytic expressions of \cite{bo2014} via the DG method. I am a co-author on this paper. The main conceptualisation and initial writing of the manuscript was done by Vikram Sunkara, Giang Nguyen and Nigel Bean. I extended the manuscript in the following ways.
\begin{itemize}
    \item Significant additions to Section~2 to introduce the partition induced by the approximation scheme to the theoretical operators of \cite{bo2014} to help elucidate the connection between the theoretical operators and the approximation counterparts. 
    \item The addition of Section~4.4 which introduces exact boundary dynamics. Prior to this boundary conditions were approximated by a small interval in the DG scheme (in this thesis, I generalised the boundary conditions further).
    \item Added and proved Corollary~4.1.
    \item Formally deriving, and describing, the approximation to the operator \(\mathbb R\) in Section~5.1. 
    \item Introduced the concept of operator kernels to the paper (used throughout Section~5.5).
    \item Wrote the final version of the code (Vikram did the initial version).
    \item Extended the numerical experiments to consider higher-order bases, smaller cell-widths, elucidating issues around truncation errors, and computing confidence intervals. 
    \item Added Appendix~A on the properties of the DG approximation to the generator of the fluid queue.
    \item Added details to Appendix~B. 
    \item I also made significant contributions to the writing (and rewriting) the manuscript, submission, and responding to reviewers comments. 
\end{itemize}
To clarify which parts of this thesis are related to the paper \cite{blnos2022}, the following sections were taken, almost verbatim except for minor editing, notational changes so that the notation matches that of the thesis, and the addition more general boundary conditions,
\begin{itemize}
    \item Section~\ref{sec: ffq intro},
    \item Chapter~\ref{ch:galerkin}, except for Sections~\ref{sec: other applications}-\ref{sec: slope limiting},
    \item Appendix~\ref{appendix:example},
    \item Appendix~\ref{sec:properties}.
\end{itemize}

### 2022-02-17 front/statement.tex
first draft of the whole thing! :D

%!TEX root = ../thesis.tex
\chapter{Signed Statement}
{\parindent=0pt\parskip=4ex

%
%  Check the University website for the current statements.  The ones below are from 
%    http://calendar.adelaide.edu.au/aprhdr/2016/specifications-thesis
%  and valid for 2016
%
%  The second one is for when you are using material that is already published.  
%

%4.1	For a thesis that does not contain work already in the public domain.

% I certify that this work contains no material which has been accepted for the award of any other degree or diploma in my name in any university or other tertiary institution and, to the best of my knowledge and belief, contains no material previously published or written by another person, except where due reference has been made in the text. In addition, I certify that no part of this work will, in the future, be used in a submission in my name for any other degree or diploma in any university or other tertiary institution without the prior approval of the University of Adelaide and where applicable, any partner institution responsible for the joint award of this degree.

% I give consent to this copy of my thesis, when deposited in the University Library, being made available for loan and photocopying, subject to the provisions of the Copyright Act 1968.

% I also give permission for the digital version of my thesis to be made available on the web, via the University's digital research repository, the Library Search and also through web search engines, unless permission has been granted by the University to restrict access for a period of time.

%
%%4.2	For a thesis that contains publications.
%
% 	I certify that this work contains no material which has been accepted for the award of any other degree or diploma in my name in any university or other tertiary institution and, to the best of my knowledge and belief, contains no material previously published or written by another person, except where due reference has been made in the text. In addition, I certify that no part of this work will, in the future, be used in a submission in my name for any other degree or diploma in any university or other tertiary institution without the prior approval of the University of Adelaide and where applicable, any partner institution responsible for the joint award of this degree.
%
%I give consent to this copy of my thesis when deposited in the University Library, being made available for loan and photocopying, subject to the provisions of the Copyright Act 1968.

%The author acknowledges that copyright of published works contained within this thesis resides with the copyright holder(s) of those works.
%
%I also give permission for the digital version of my thesis to be made available on the web, via the University's digital research repository, the Library %Search and also through web search engines, unless permission has been granted by the University to restrict access for a period of time.
%
I certify that this work contains no material which has been accepted for the award of any other degree or diploma in my name, in any university or other tertiary institution and, to the best of my knowledge and belief, contains no material previously published or written by another person, except where due reference has been made in the text. In addition, I certify that no part of this work will, in the future, be used in a submission in my name, for any other degree or diploma in any university or other tertiary institution without the prior approval of the University of Adelaide and where applicable, any partner institution responsible for the joint award of this degree.

I give permission for the digital version of my thesis to be made available on the web, via the University's digital research repository, the Library Search and also through web search engines, unless permission has been granted by the University to restrict access for a period of time.

I acknowledge the support I have received for my research through the provision of an Australian Government Research Training Program Scholarship.

Signed: \dotfill\quad 
Date: \dotfill

}

### 2022-02-17 intro/chapter.tex
first draft of the whole thing! :D

@@
 %!TEX root = ../thesis.tex
-\chapter{Introduction \label{ch:intro}} 
+\chapter{Introduction\label{ch:intro}} 
 % intro to intro
@@
 
-An fluid queue is a two-dimensional stochastic process \(\{{\bs X}(t)\} = \{( X(t),\varphi(t))\}_{t\geq0}\).
+A fluid queue is a two-dimensional stochastic process \(\{{\bs X}(t)\} = \{( X(t),\varphi(t))\}_{t\geq0}\).
 The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC) and determines the rate at which \(\{ X(t)\}\) moves.
@@
 
-Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982}, a canonical application in this area), power systems \citep{hydro}, risk processes \citep{betal2005} and environmental modelling \citep{wurm2020}.
+Typically, the phase, \(\varphi(t)\), represents the underlying operating state or environment of a system at time \(t\).
+The level variable, \(X(t)\), describes some continuous variable of the system.
+Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982}, a canonical application in this area), power systems \citep{hydro}, risk processes \citep{betal2005}, environmental modelling \citep{wurm2020}, dam storage \citep{loynes1962}, and more.
+For example, \cite{anick1982} model a \emph{data handling switch} in which there are \(N\) information sources which asynchronously switch from ON to OFF, or OFF to ON.
+The phase process \(\varphi(t)\) models the number of data sources which are ON at time \(t\), and the rate at which data is received when the buffer is on is assumed (without loss of generality) to be 1.
+The switch stores data in a buffer and processes the data at rate \(c\), thus the net rate of change of the buffer at time \(t\) when there are \(r\) information sources which are ON is \(r-c\).
+The level process, \(X(t)\), models the amount of data in the buffer at time \(t\) and the rate at which \(X(t)\) changes which \(\varphi(t)=r\) is \(c_r=r-c\).
+The model described above could be used to aid in the analysis and design of the data handling switch, for example, by determining the minimum rate \(r\) for which the system is \emph{stable} in steady state, which practically means that the processing power of the switch is large enough that the amount of data in the buffer does not grow without bound.
+
+The success of fluid queues largely lies in their mix of flexibility, any (finite) number of phases can be used, and their parsimony and analytic tractability which enables effective transient and stationary analysis \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018,bnp2018}.
 Fluid queues are relatively well studied.
@@
 
-More recently, \cite{bo2014} extended fluid queues to so-called \emph{stochastic fluid-fluid queues}.
+More recently, \cite{bo2014} extends fluid queues to so-called \emph{stochastic fluid-fluid queues}.
 In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\), so \((X(t),\varphi(t))\) determines the rate at which \(\{Y(t)\}\) moves.
-The analysis of \cite{bo2014}, is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
+The addition of the second level process \(\{Y(t)\}\) greatly increases the modelling potential.
+For example, \cite{hydro,bo2014} describe a model for a hydroelectric power generator which can be operated `on design' and `off design'.
+When the system operates `off design' there system is less efficient and there is more wear, but may be used to optimise overall system performance.
+Further, maintenance of the generator takes place periodically to improve its performance and lifespan.
+The system can plausibly be modelled as a fluid queue, \(\{(\varphi(t),X(t))\}\), where the state space of \(\{\varphi(t)\}\) consists of states representing the systems states `start', `stop', `on design', `off design', `maintenance', and `idle'.
+The level variable \(\{X(t)\}\) take values in \([0,1]\) and represents the deterioration level of the system with 0 being perfect working order and 1 meaning the system needs replacement.
+The second level process \(\{Y(t)\}\) could be used to model the revenue of the power generator, which, logically, should depend on the state of the system, \(\{\varphi(t)\}\).
+Moreover, the revenue could also depend on the deterioration level \(\{X(t)\}\), as, for example, when the system is brand new it may operate more efficiently, and when it is near broken it may be less efficient.
+
+% In this paper, we are interested in the following modeling potential. Suppose that, besides
+% the level X (t) at time t, we would like to model some other continuous performance measure
+% Y (t), such as the net profit at time t for example. To illustrate this, consider a hydro-power
+% generator, which can be operated “on design” and “off design”. The latter causes increased
+% wear, and is less efficient, but can be useful to optimize overall system operation. The generator
+% must be periodically maintained, in order to improve its performance and prolong its lifespan.
+% An important related problem is the evaluation of maintenance strategies and the impact of
+% [1–6,10,12–16,28,29][7,12,21]
+% N.G. Bean, M.M. O’Reilly / Stochastic Processes and their Applications 124 (2014) 1741–1772 1743
+% maintenance on reliability. This problem can be modeled as a bounded SFM {(φ(t), X (t)), t ≥ 0}
+% with S consisting of operating phases (start, stop, on design, off design, maintenance, idle) and
+% deterioration level X (t) ∈ [0, 1], where 0 means the generator is brand new and 1 means it needs
+% replacement, and it is assumed that once level 1 is reached, the process {(φ(t), X (t)), t ≥ 0}
+% restarts itself from level 0 [12].
+% Now, in order to describe the continuous revenue process {Y (t), t ≥ 0} associated with
+% the process {(φ(t), X (t)), t ≥ 0}, it may be plausible to use a SFM driven by {φ(t), t ≥ 0}.
+% This is because the rate at which the revenue is increasing/decreasing at time t, ideally should
+% depend on the phase φ(t). However, for practical reasons, we may additionally require that the
+% revenue Y (t) at time t depends on the operating level X (t). For example, if the generator is
+% newer, it may operate more efficiently, produce more energy and require less-costly maintenance.
+% Hence, the rates at which the revenue level Y (t) is increasing/decreasing may depend both on
+% the phase i ∈ S and the operating level X (t) = x at time t, and so to reflect this, we denote
+% these rates by ri (x). Summarizing this, we propose a new class of fluid models, denoted by
+% {(φ(t), X (t), Y (t)), t ≥ 0}. We refer to this class as the stochastic fluid–fluid model, since it is
+% a stochastic fluid model with level Y (t) driven by {(φ(t), X (t)), t ≥ 0}, which is also a stochastic
+% fluid model with level X (t). The buffers collecting fluid from the process {X (t), t ≥ 0} and
+% {Y (t), t ≥ 0} are referred to as X and Y , respectively.
+
+The addition of the second level process \(\{Y(t)\}\) greatly increases the flexibility and modelling potential, however comes at some cost to the ease of their analysis.
+The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
 For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used.
 To this end, to date, two possible discretisation have been suggested.
-Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise this operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process.
-The QBD approximation of \cite{bo2013} is derived via a uniformisation argument, so we refer to it as the uniformisation approximation scheme throughout this thesis.
+Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise the operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process.
+The QBD approximation of \cite{bo2013} is derived via a uniformisation argument, so we refer to it as the uniformisation scheme throughout this thesis.
 Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations.
-It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the former; the uniformisation scheme can be viewed as the simplest DG scheme where the operator is projected onto a basis of piecewise constant functions with an upwind flux.
+It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the former; the uniformisation scheme can be viewed as the simplest DG scheme where cells are all have equal width, the operator is projected onto a basis of piecewise constant functions, and an upwind flux is used to approximate the flow of mass between cells.
+This is also equivalent to a finite-volume scheme.
 
@@
 
-In the context of approximating fluid queues, one advantage of the uniformisation scheme and, equivalently, a DG scheme with constant basis functions, is that is guarantees probabilities computed from the approximation are positive \citep[Section~3.3]{koltai2011}.
+One drawback of the uniformisation scheme is that convergence can be relatively slow compared to higher-order DG schemes.
+However, in the context of approximating fluid queues, one advantage of the uniformisation scheme and, equivalently, a DG scheme with constant basis functions, is that the uniformisation scheme guarantees probabilities computed from the approximation are positive \citep[Section~3.3]{koltai2011}.
 One justification for the positivity preserving property of the uniformisation scheme is from its interpretation as a stochastic process.
@@
 Moreover, higher-order DG approximation schemes may produce negative and oscillatory solutions, particularly when discontinuities or steep gradients are present.
+This is problematic for probabilistic problems as we know that probabilities must be positive.
 Methods to navigate the problem of negative and oscillatory solutions have been developed, such as filtering and slope limiting (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6 and references therein).
-Slope limiting alters the discretised operator in regions where oscillations are detected and reduces the order of the approximation to linear in these regions.
+Slope limiting alters the discretised operator in regions where oscillations are detected and reduces the order of the approximation to linear in these regions but ensures the solution will be non-negative.
 Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.
@@
 Slope limiting does guarantee positivity but reduces the approximation to linear where oscillations in the approximate solution are detected \citep[Section~5.6.1]{nodalDGBook}.
-Furthermore, limiting and filtering do not distinguish between natural oscillations which are a fundamental feature of the solution and spurious oscillations which are caused by the approximation scheme, and they may remove both from the approximation.
+Furthermore, limiting and filtering do not distinguish between natural oscillations, which are a fundamental feature of the solution and spurious oscillations which are caused by the approximation scheme, and they may remove both from the approximation.
 This can lead to an unnecessary loss of accuracy in the approximation (see \citep{nodalDGBook},~Example~5.8).
+Other drawbacks of these methods is that they result in non-linear operators, so we cannot use matrix algebra for computations, and they incur an additional computational cost on top of the DG scheme.
 
-In the context of approximating the first return operator of a fluid-fluid queue the application of a slope limiter would amount to post-processing the solution, reducing the order of the approximation in areas where oscillations are detected to linear.
-Otherwise, perhaps another approach would be to re-compute the solution with lower-order basis if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
-Post-hoc filtering of the solution is also possible, but we would still need to tune the filter appropriately.
-Both filtering and slope limiting are also dependent on the initial condition.
+% In the context of approximating the first return operator of a fluid-fluid queue, the application of a slope limiter is not possible since we need the approximated operators to be linear.
 
-In approximating the first return operator of a fluid-fluid queue we first approximate an operator-Riccati equation by a matrix-Riccati equation by substituting in matrix approximations to the infinitesimal generator of the fluid queue which are constructed via the DG scheme \cite{blnos2022}.
+In approximating the first return operator of a fluid-fluid queue we first approximate an operator-Riccati equation by a matrix-Riccati equation by substituting in matrix approximations to the infinitesimal generator of the fluid queue which are constructed via the DG scheme \citep{blnos2022}.
 We then solve the matrix-Riccati equation via an iterative procedure \citep{bean2005b,blnos2022}.
-For the DG method there is a limited theoretical backing as to why we may approximate the operator-Riccati equation with the matrix-Riccati equation other than it is a sensible thing to do, and it seems to work in practice.
-In the case of the QBD approximation of \cite{bo2013} this is more justifiable as in this case the resulting matrix-Riccati equation can be derived by considering the first-return probabilities for a fluid queue driven by the QBD \citep{bean2005} where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\).
+Since the operators constructed with limiters are non-linear (not matrices) then we cannot apply limiters in this case.
+Alternatively, we could apply a post-processing based on slope limiters, or filtering.
+Both of these post-processing methods incur a computational cost and need to be applied on a case-by-case basis depending on the initial condition.
+
+Furthermore, for the DG method there is a limited theoretical backing as to why we may approximate the operator-Riccati equation with the matrix-Riccati equation other than it is a sensible thing to do, and it seems to work in practice.
+In the case of the QBD approximation of \cite{bo2013} the approximate matrix-Riccati is more justified as in it can be derived by considering the first-return probabilities for a fluid queue driven by the approximating QBD where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\) \citep{bean2005}.
+
+Hence, for the approximation of fluid queues with application to fluid-fluid queues, on the one hand we can use the DG scheme which can produce high-order approximations when the problem is smooth, but can display spurious oscillations and result in negative probability estimates in the presence of discontinuities.
+While, on the other hand, we can use the uniformisation scheme which constructs in a linear operator to approximate the generator and guarantees non-negative probability estimates, but convergence is relatively slower.
 
 Motivated by this, this thesis derives a new approximation to a fluid queue.
-The approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributions to model the sojourn time in a given interval on the event that the phase of the fluid is constant.
-The sojourn time in a given interval on the event that the phase of the fluid is constant is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution so, in this sense, the best approximation to this deterministic sojourn time.
+The construction of the approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributions to approximate the sojourn time of the fluid level in a given interval on the event that the phase of the fluid is constant.
+The sojourn time in a given interval on the event that the phase of the fluid is constant is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution, so in this sense, the best approximation to this deterministic sojourn time.
 Thus, it appears that the approximation of \cite{bo2013} is the best-possible Markov chain approximation.
-Recently, there has been much work on a class of concentrated matrix exponential distributions \cite{hhat2020} which are postulated to be the least-variable matrix exponential distribution.
+Recently, there has been much work on a class of concentrated matrix exponential distributions \citep{hhat2020} which are postulated to be the least-variable matrix exponential distributions of a given order.
 Matrix exponential distributions generalise Phase-type distributions; they have the same functional form, without the restriction that the distribution has an interpretation in terms of the absorption time of a continuous-time Markov chain.
-A class of stochastic processes, known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \cite{bn2010}, extend QBDs, which have Phase-type inter-event times, to allow matrix-exponentially distributed inter-event times.
+A class of stochastic processes, known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \citep{bn2010}, extends QBDs, which have Phase-type inter-event times, to allow matrix-exponentially distributed inter-event times.
 Thus, by using matrix exponential distributions, we attempt to construct a QBD-RAP which better captures the dynamics of the fluid queue than the QBD approximation in \cite{bo2013}, while retaining a stochastic interpretation.
@@
 As the QBD-RAP has a stochastic interpretation then the approximations it produces are guaranteed to have non-negative density functions.
-Moreover, the matrix-Riccati equation we solve to approximate the first-return distribution of a fluid-fluid queue can be derived by considering the first-return probabilities for a RAP-modulated fluid queue driven by the QBD-RAP \citep{p2019,bgnp2021}.
-Another attractive property is that the QBD-RAP (and uniformisation) method are guaranteed to produce non-negative estimates of probabilities for any initial condition without any further computation or post-processing, which also means that the approximations to operators we get from the QBD-RAP (and uniformisation) scheme are linear; a desirable mathematical property.
+Further, for the QBD-RAP (and uniformisation) scheme non-negativity is guaranteed for any initial condition without any further computation or post-processing.
+For the approximation of fluid-fluid queues, the matrix-Riccati equation we solve to approximate the first-return distribution of a fluid-fluid queue can be derived by considering the first-return probabilities for a RAP-modulated fluid queue driven by the QBD-RAP \citep{p2019,bgnp2021}.
+The stochastic interpretation can also be leveraged to aid in the analysis of the approximation.
+Another attractive property is that the approximations to operators obtained via the QBD-RAP (and uniformisation) scheme are linear which is necessary for the application to approximating the first-return operator of fluid-fluid queues and may also provide other computational and analytical benefits.
 
 The structure of this thesis is as follows.
-The rest of this chapter is dedicated to mathematical preliminaries and introduces the main mathematical objects and tools which we will need.
-In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way to make clear how the approximations we develop correspond to the theoretical operators.
+The next chapter, Chapter~\ref{ch:math}, is dedicated to mathematical preliminaries and introduces the main mathematical objects and tools which we will need.
+In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way that makes clear exactly how the approximations we develop correspond to the theoretical operators.
+Chapter~\ref{ch:math} also introduces the existing literature and gives context to this thesis.
 
 Chapter~\ref{ch:galerkin} demonstrates a way that we can use the discontinuous Galerkin (DG) method to approximate certain operators and distributions of fluid and fluid-fluid queues.
-Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:galerkin} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and non-monotonic CDFs.
-However, one of the problems we are interested in solving is how to approximate the first-return operator of a fluid-fluid queue, which is a non-standard problem for the DG method.
-It is not clear how to use the concept of \emph{slope limiting} in the context of approximating the first-return operator of a fluid-fluid queue.
+Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:galerkin} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and negativity.
+At the end of Chapter~\ref{ch:galerkin} we also introduce other numerical procedures which we use in our numerical investigations in Chapter~\ref{sec: numerics}.
 
-In the following chapter, Chapter~\ref{sec: construction and modelling}, we develop a new approximation scheme which, due to its interpretation as a stochastic process (a QBD-RAP), ensures all approximations to CDFs are non-decreasing.
+In the following chapter, Chapter~\ref{sec: construction and modelling}, we develop a new QBD-RAP approximation scheme which, due to its interpretation as a stochastic process, ensures all estimates of probabilities will be non-negative.
 The chapter takes a stochastic modelling approach to developing the approximation scheme.
-Once we have established the new approximation scheme, we then attempt to prove convergence of the approximation scheme in Chapters~\ref{sec: conv}~and~\ref{ch: global results} via certain Laplace transforms with respect to time.
-Chapter~\ref{sec: conv} utilises methods and ideas relating specifically to QBD-RAPs and matrix-exponential distributions.
-Ultimately, Chapter~\ref{sec: conv} proves that, on the event that the QBD-RAP has not-yet seen an \emph{orbit restart epoch} (which is much like a change of level, but not always), certain Laplace transforms with respect to time of the QBD-RAP scheme converge to corresponding Laplace transforms with respect to time of the fluid queue on the event that the level of the fluid queue remains in a given interval.
-In this sense, this is a local convergence result as it relates only to convergence to the fluid queue in an interval.
 
-Chapter~\ref{ch: global results} then looks to extend the local convergence result to a global convergence results on the whole domain of approximation.
-Chapter~\ref{ch: global results} uses more traditional Markov process arguments which rely on properties such as the strong Markov property, time-homogeneity and the Law of total probability.
-In Chapter~\ref{ch: global results} we first consider the discrete-time process which the process embedded in the QBD-RAP at times when the QBD-RAP sees an orbit restart epoch and prove that the embedded process converges in distribution to a corresponding embedded process in the fluid queue.
+Once we have established the new approximation scheme, we then prove a convergence of the approximation scheme in Chapters~\ref{sec: conv}~and~\ref{ch: global results} via certain Laplace transforms with respect to time.
+Chapter~\ref{sec: conv} utilises methods and ideas relating specifically to QBD-RAPs and matrix exponential distributions.
+Ultimately, Chapter~\ref{sec: conv} proves that, on the event that the QBD-RAP has not-yet seen an \emph{orbit restart epoch} (which is much like a change of level, but not always), certain Laplace transforms with respect to time of the QBD-RAP scheme converge to corresponding Laplace transforms with respect to time of the fluid queue.
+This is a local convergence result as it relates only to convergence to the fluid queue in an interval.
+
+Chapter~\ref{ch: global results} then looks to extend the local convergence results of Chapter~\ref{sec: conv} to a global convergence result on the whole domain of approximation.
+Chapter~\ref{ch: global results} uses more traditional Markov process arguments which rely on things such as the strong Markov property, time-homogeneity and the Law of total probability.
+In Chapter~\ref{ch: global results} we first consider the discrete-time process which the process embedded in the QBD-RAP at times when the QBD-RAP sees an orbit restart epoch.
+We prove that the embedded process converges in distribution to a corresponding embedded process of the fluid queue.
 The rest of Chapter~\ref{ch: global results} then attempt to prove a global converge result.
@@
 
-Once we have established a few methods for approximating fluid queues, Chapter~\ref{sec: numerics} then numerically explores some properties of the approximations numerically.
+Once we have established methods for approximating fluid queues, Chapter~\ref{sec: numerics} then numerically explores some properties of the approximations.
 Given that the QBD-RAP scheme was developed so that it guarantees positivity of the approximation, we largely focus on problems with discontinuities.
@@
 To help the reader understand the notations used in the DG scheme, we provide a small toy model example in Appendix~\ref{appendix:example}.
+Appendix~\ref{sec:properties} provides a proof that the DG scheme conserves probability.
 The rest of the appendices contain technical results relating to proving the convergence of the QBD-RAP approximation scheme.
-Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operator} introduced as part of the QBD-RAP scheme have the properties we claim they do to prove convergence in Chapter~\ref{sec: conv}.
+Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operators} introduced as part of the QBD-RAP scheme have the properties we claim they do to prove convergence in Chapter~\ref{sec: conv}.
 Appendix~\ref{app:extend conv} extends some results from Chapter~\ref{sec: conv} to a setting which requires slightly less computation.

### 2022-02-17 appendix_properties_of_B/appendix.tex
typos

@@
  \\\mathfrak{\bs B}^{00}_{ij} &:= \begin{cases}
-	T_{ii}\bs I_{N_0} + c_i\bs P_0^{-1}(\bs F_i^{00}+\bs G_0)\bs M_0^{-1}\bs P_0 & i = j, \\
-T_{ij}\bs I_{N_0} - c_ip_{ij}^{-1} 1(c_i<0)\bs F_i^{00}\bs M_0^{-1}\bs P_0 & i \neq j,
+	T_{ii}\bs I_{p_0} + c_i\bs P_0^{-1}(\bs F_i^{00}+\bs G_0)\bs M_0^{-1}\bs P_0 & i = j, \\
+T_{ij}\bs I_{p_0} - c_ip_{ij}^{-1} 1(c_i<0)\bs F_i^{00}\bs M_0^{-1}\bs P_0 & i \neq j,
 \end{cases}
@@
 \\\mathfrak{\bs B}^{KK}_{ij} &:= \begin{cases}
-	T_{ii}\bs I_{N_K} + c_i\bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K & i = j, \\
-T_{ij}\bs I_{N_K} + c_ip_{ij}^{K+1} 1(c_i>0)\bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K & i \neq j,
+	T_{ii}\bs I_{p_K} + c_i\bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K & i = j, \\
+T_{ij}\bs I_{p_K} + c_ip_{ij}^{K+1} 1(c_i>0)\bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K & i \neq j,
 \end{cases}
@@
         %
-        \\\bs F_i^{k,k+1}\boldsymbol d &= \boldsymbol \phi_{k}(y_{k+1})\tr{} \boldsymbol \phi_{k+1}(y_{k+1})\boldsymbol d \\&= (\bs e^{k}_{p_k})\tr{} \bs e_{1}^{k+1}\bs d \\&= d_{1}(\boldsymbol e_{N_{k}}^{k})\tr{}.
+        \\\bs F_i^{k,k+1}\boldsymbol d &= \boldsymbol \phi_{k}(y_{k+1})\tr{} \boldsymbol \phi_{k+1}(y_{k+1})\boldsymbol d \\&= (\bs e^{k}_{p_k})\tr{} \bs e_{1}^{k+1}\bs d \\&= d_{1}(\boldsymbol e_{p_{k}}^{k})\tr{}.
 	\end{align*}
@@
 		&\sum_{j\in\calS}\mathfrak{\bs B}^{KK}_{ij}\bs 1 + \mathfrak{\bs B}^{K{K+1}}_{ij}\bs 1
-		\\&= \sum_{j\in\calS}T_{ij}\bs I_{N_K}\bs 1-\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 -\sum_{j\in\calS_{{K+1} }}c_ip_{ij}^{K+1} \bs P_K^{-1}\boldsymbol \phi_K(0)\tr{}\bs 1
+		\\&= \sum_{j\in\calS}T_{ij}\bs I_{p_K}\bs 1-\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 -\sum_{j\in\calS_{{K+1} }}c_ip_{ij}^{K+1} \bs P_K^{-1}\boldsymbol \phi_K(0)\tr{}\bs 1
 		\\&\quad{} + c_i \bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K\boldsymbol 1 
@@
 	\end{align*}
-	% Similarly, for \(c_i<0\), and row-vectors \(\boldsymbol b\) and \(\boldsymbol d\) of length \(p_k\) and \(N_{k-1}\), respectively, 
+	% Similarly, for \(c_i<0\), and row-vectors \(\boldsymbol b\) and \(\boldsymbol d\) of length \(p_k\) and \(p_{k-1}\), respectively, 
 	% \begin{align*}
@@
 	% 	%
-	% 	\\\bs F_i^{k,k-1}\boldsymbol d &= -\boldsymbol \phi^{k}(y_{k})\tr{} \boldsymbol \phi^{k-1}(y_{k})\boldsymbol d \\&= -(\bs e^{k}_{1})\tr{} \bs e_{N_{k-1}}^{k-1}\bs d \\&= -d_{N_{k-1}}(\boldsymbol e_{1}^k)\tr{}.
+	% 	\\\bs F_i^{k,k-1}\boldsymbol d &= -\boldsymbol \phi^{k}(y_{k})\tr{} \boldsymbol \phi^{k-1}(y_{k})\boldsymbol d \\&= -(\bs e^{k}_{1})\tr{} \bs e_{p_{k-1}}^{k-1}\bs d \\&= -d_{p_{k-1}}(\boldsymbol e_{1}^k)\tr{}.
 	% 	\end{align*}
@@
 		\sum_{j\in\calS}\mathfrak{\bs B}^{00}_{ij}\bs 1 + \mathfrak{\bs B}^{0{-1}}_{ij}\bs 1
-		&= \sum_{j\in\calS}T_{ij}\bs I_{N_0}\bs 1-\sum_{j\in\calS_+} c_ip_{ij}^{-1} \bs P_0^{-1}\bs F_i^{00}\bs M_0^{-1}\bs P_0 \bs 1 -\sum_{j\in\calS_{{-1} }}c_ip_{ij}^{-1} \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{}\bs 1
+		&= \sum_{j\in\calS}T_{ij}\bs I_{p_0}\bs 1-\sum_{j\in\calS_+} c_ip_{ij}^{-1} \bs P_0^{-1}\bs F_i^{00}\bs M_0^{-1}\bs P_0 \bs 1 -\sum_{j\in\calS_{{-1} }}c_ip_{ij}^{-1} \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{}\bs 1
 		\\&\quad{} + c_i \bs P_0^{-1}(\bs F_i^{0,0}+\bs G_0)\bs M_0^{-1}\bs P_0\boldsymbol 1 
@@
 		%\\&%= T_{{K+1},{K+1}}\boldsymbol 1 + [T_{{K+1}-}\otimes \boldsymbol\phi^K(b)\bs 1]\boldsymbol 1
-		&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes\bs e^K_{N_K}\bs 1]\boldsymbol 1
-		\\&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes \bs e^K_{N_K}]\bs 1
+		&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes\bs e^K_{p_K}\bs 1]\boldsymbol 1
+		\\&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + [\bs T_{{K+1}-}\otimes \bs e^K_{p_K}]\bs 1
 		\\&= \bs T_{{K+1},{K+1}}\boldsymbol 1 + \bs T_{{K+1}-}\bs 1

### 2022-02-17 chapter2/chapter.tex
typos

@@
 
-For each \(k\) we choose \(N_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{N_k}\), compactly supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) to form a basis for the space \(W_k\), in which we formulate the approximation.
-Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{N_k}\) to be the space of polynomials of degree \(N_k-1\).
-It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{N_k}\) as a basis of Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature points, since our approximations inherit nice properties from this \citep{nodalDGBook}.
+For each \(k\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), compactly supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) to form a basis for the space \(W_k\), in which we formulate the approximation.
+Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{p_k}\) to be the space of polynomials of degree \(p_k-1\).
+It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{p_k}\) as a basis of Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature points, since our approximations inherit nice properties from this \citep{nodalDGBook}.
 However, the constructions presented here are general, and any basis can be used.
-For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{N_k}\) as the Lagrange polynomials.
+For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials.
 On each cell \(\mathcal D_{k,i}\) we approximate
-\[f_i(x,t)\approx u_{k,i}(x,t)=\sum\limits_{r=1}^{N_k}a_{k,i}^r(t)\phi^r_k(x),\] 
+\[f_i(x,t)\approx u_{k,i}(x,t)=\sum\limits_{r=1}^{p_k}a_{k,i}^r(t)\phi^r_k(x),\] 
 where \(a_{k,i}^r(t)\) are yet-to-be-determined time-dependent coefficients.
@@
 
-Let \(\mathcal N_k := \left\{1,\dots,N_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
+Let \(\mathcal N_k := \left\{1,\dots,p_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
 \[\boldsymbol \phi_k(x) = (\phi^r_k(x))_{r\in\mathcal N_k}, \quad \boldsymbol a_{k,i}(x) = (a_{k,i}^r(x))_{r\in\mathcal N_k},\,i\in\mathcal S.\]
@@
 %\end{align*}
-and repeat this for all test functions \(\phi^r_k(x)\), \(r=1,...,N_k\), to get the following system of equations,
+and repeat this for all test functions \(\phi^r_k(x)\), \(r=1,...,p_k\), to get the following system of equations,
 \begin{align}
@@
 
-It remains to approximate the \textit{flux}, \(f_j(x,t)\) at the cell edges \(y_k,\,k\in\mathbb Z\), so that we may evaluate the terms \([f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}\), \(r=1,...,N_k,\,k\in\mathbb Z\).
+It remains to approximate the \textit{flux}, \(f_j(x,t)\) at the cell edges \(y_k,\,k\in\mathbb Z\), so that we may evaluate the terms \([f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}\), \(r=1,...,p_k,\,k\in\mathbb Z\).
 This is the key for DG -- it joins the local approximations on each cell into a global approximation on the whole domain of approximation.
@@
 \begin{align*}
-    \ddot{{\bs B}}^{kk}&=\left[\begin{array}{ccc}T_{11}\bs I + c_1(\bs F_1^{kk}+\bs G_k)\bs M_k^{-1} & T_{12}\bs I & T_{1N_\calS}\bs I  \\ T_{21}\bs I & & \\ \vdots &\ddots & \vdots \\ & &   T_{N_\calS-1,N_\calS}\bs I \\  T_{N_\calS1}\bs I &  T_{N_\calS,N_\calS-1}\bs I & T_{N_\calS,N_\calS}\bs I +c_{N_\calS}(\bs F_{N_\calS}^{kk}+\bs G_k)\bs M_k^{-1}\end{array}\right],\\
-	\ddot{{\bs B}}^{k,k+1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k+1}\bs M_{k+1}^{-1}&  & \\  &\ddots & \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k+1}\bs M_{k+1}^{-1} \end{array}\right],\\
-	\ddot{{\bs B}}^{k,k-1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k-1}\bs M_{k-1}^{-1}&  & \\  &\ddots &  \\  &  &c_{N_\calS}\bs F_{N_\calS}^{k,k-1}\bs M_{k-1}^{-1} \end{array}\right].
+    \ddot{{\bs B}}^{kk}&=\left[\begin{array}{ccc}T_{11}\bs I + c_1(\bs F_1^{kk}+\bs G_k)\bs M_k^{-1} & T_{12}\bs I & T_{1N}\bs I  \\ T_{21}\bs I & & \\ \vdots &\ddots & \vdots \\ & &   T_{N-1,N}\bs I \\  T_{N1}\bs I &  T_{N,N-1}\bs I & T_{N,N}\bs I +c_{N}(\bs F_{N}^{kk}+\bs G_k)\bs M_k^{-1}\end{array}\right],\\
+	\ddot{{\bs B}}^{k,k+1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k+1}\bs M_{k+1}^{-1}&  & \\  &\ddots & \\  &  &c_{N}\bs F_{N}^{k,k+1}\bs M_{k+1}^{-1} \end{array}\right],\\
+	\ddot{{\bs B}}^{k,k-1}&=\left[\begin{array}{ccc}c_1\bs F_1^{k,k-1}\bs M_{k-1}^{-1}&  & \\  &\ddots &  \\  &  &c_{N}\bs F_{N}^{k,k-1}\bs M_{k-1}^{-1} \end{array}\right].
 \end{align*} 
@@
 \begin{align*}
-	{{\bs B}}^{kk}_{ij} &= \begin{cases}T_{ij}\bs I_{N_k} + c_i(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1} & i=j,\\T_{ij} \bs I_{N_k}& i\neq j,\end{cases}
+	{{\bs B}}^{kk}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1} & i=j,\\T_{ij} \bs I_{p_k}& i\neq j,\end{cases}
 \\	{{\bs B}}^{k\ell}_{ij} &= \begin{cases}c_i\bs F_i^{k\ell}\bs M_\ell^{-1} & i=j,\\\bs 0 & i\neq j,\end{cases}\quad \ell \in \{k-1,k\}
@@
 \begin{align*}
-	{{\bs B}}^{00}_{ij} &= \begin{cases}T_{ij}\bs I_{N_k} + c_i(\bs F_i^{00}+\bs G_0)\bs M_0^{-1} & i=j,\\T_{ij}\bs I_{N_k} - 1(c_i<0,c_j>0)c_ip_{ij}^{-1} \bs F_i^{00}\bs M_0^{-1} & i\neq j,\end{cases}
-\\	{{\bs B}}^{KK}_{ij} &= \begin{cases}T_{ij}\bs I_{N_k} + c_i(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1} & i=j,\\T_{ij}\bs I_{N_K} + 1(c_i>0,c_j<0)c_ip_{ij}^{K+1} \bs F_i^{KK}\bs M_K^{-1} & i\neq j.\end{cases}
+	{{\bs B}}^{00}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{00}+\bs G_0)\bs M_0^{-1} & i=j,\\T_{ij}\bs I_{p_k} - 1(c_i<0,c_j>0)c_ip_{ij}^{-1} \bs F_i^{00}\bs M_0^{-1} & i\neq j,\end{cases}
+\\	{{\bs B}}^{KK}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1} & i=j,\\T_{ij}\bs I_{p_K} + 1(c_i>0,c_j<0)c_ip_{ij}^{K+1} \bs F_i^{KK}\bs M_K^{-1} & i\neq j.\end{cases}
 \end{align*}
@@
 % 	\bs I_{|\calS_{-1}|} & & & & \\
-% 	& \bs I_{N_S}\otimes \bs M_1 & & & \\
+% 	& \bs I_{p_S}\otimes \bs M_1 & & & \\
 % 	& & \ddots & & \\
-% 	& & & \bs I_{N_S}\otimes \bs M_K & \\
+% 	& & & \bs I_{p_S}\otimes \bs M_K & \\
 % 	& & & & \bs I_{|\calS_{K+1}|}
@@
 
-Let \(\boldsymbol a_{-1}(t):=\boldsymbol q_{-1}(t)\) and \(\boldsymbol a_{K+1}(t):=\boldsymbol q_{K+1}(t)\), and define basis functions \(\bs\phi_{-1}(x) = \phi_{-1}^1(x) = \delta(x)\) and \(\bs\phi_{K+1}(x) = \phi_{K+1}^1(x) = \delta(x-b)\), where \(\delta\) is the Dirac delta, \(N_{-1} = N_{K+1} = 1\), and \(\mathcal N_{-1} = \mathcal N_{K+1} = \{1\}\).
+Let \(\boldsymbol a_{-1}(t):=\boldsymbol q_{-1}(t)\) and \(\boldsymbol a_{K+1}(t):=\boldsymbol q_{K+1}(t)\), and define basis functions \(\bs\phi_{-1}(x) = \phi_{-1}^1(x) = \delta(x)\) and \(\bs\phi_{K+1}(x) = \phi_{K+1}^1(x) = \delta(x-b)\), where \(\delta\) is the Dirac delta, \(p_{-1} = p_{K+1} = 1\), and \(\mathcal N_{-1} = \mathcal N_{K+1} = \{1\}\).
 Also define \({\bs M}_{-1}=\bs I_{|\calS_{-1}|}\), \({\bs M}_{K+1}=\bs I_{|\calS_{K+1}|}\), the block-diagonal matrix \(\bs M=diag(\bs M_k,k\in\mathcal K)\), and row-vectors
@@
 \[\cfrac{\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}}{|r_i(x)|} = \bs \rho_{k,i} \bs \phi_k(x)\tr{} + g_i^\perp(x),\] where \(\bs \rho_{k,i} \bs\phi_k(x)\tr{}\in W\), \(g_i^\perp \in W^\perp\).
-Now, multiply by test functions \(\{\phi^r_k(x)\}_{r=1}^{N_k}\) and integrate over \([0,b]\):
+Now, multiply by test functions \(\{\phi^r_k(x)\}_{r=1}^{p_k}\) and integrate over \([0,b]\):
 \begin{align*}

### 2022-06-19 chapter1/chapter.tex
Ch7 2nd draft

@@
 
-A stochastic process is a sequence, \(\{X(t)\}_{t\in\mathcal T}\), of random variables (or random vectors) indexed by some index set \(\mathcal T\).
+Following \cite{rossBook}, a stochastic process is a sequence, \(\{X(t)\}_{t\in\mathcal T}\), of random variables (or random vectors) indexed by some index set \(\mathcal T\).
 In this thesis, in the case that the index set is \(\mathcal T=\{0,1,2,...\}\) we say that the process \(\{X(n)\}_{n\in \mathcal T}\) is a discrete-time process, and we will typically use the dummy variable \(n\) for the `time' index.
@@
 We may omit the index set and write \(\{X(t)\}\) in place of \(\{X(t)\}_{t\in\mathcal T}\) when it is not explicitly needed, or we may write \(\{X(t)\}_{t\geq t_0}\) to mean \(\{X(t)\}_{t\in[t_0,\infty)}\).
-The \emph{sample space} of \(X(t)\) is the set of possible values that \(X(t)\) can take at any time \(t\in\mathcal T\).
+The \emph{state space} of \(X(t)\) is the set of possible values that \(X(t)\) can take at any time \(t\in\mathcal T\).
 
@@
 More generally, for a random variable, \(Z\), we write \(Z\sim \nu\) when \(Z\) has the distribution given by the probability measure \(\nu\).
-For the probability that \(X(t)\) lies in some set \(E\subset \mathcal S\) given \(X(0)\sim \mu\), we write \(\mathbb P(X(t)\in E \mid X(0)\sim \mu)\).
+For the probability that \(X(t)\) lies in some measurable set \(E\subset \mathcal S\) given \(X(0)\sim \mu\), we write \(\mathbb P(X(t)\in E \mid X(0)\sim \mu)\).
 When \(\mu\) assigns probability 1 to a single point, \(x\in\calS\), say, we write \(\mathbb P(X(t)\in E \mid X(0)=x).\)
 
-A stochastic process is \emph{stationary} if
+A stochastic process is \emph{stationary} if \(X(t_n),X(t_{n-1}),...,X(t_0)\) and \(X(t_n+t),X(t_{n-1}+t),...,X(t_0+t)\) have the same distribution for any \(t\). That is, 
 \begin{align}\label{eqn: stationary}
 	&\nonumber\mathbb P(X(t_n)\in E_n,X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
-	\\&= \mathbb P(X(t_n+t)=i,X(t_{n-1}+t)\in E_{n-1},...,X(t_0+t)\in E_0),
-\end{align}
-for any \(t,t_0,...,t_n\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\), i.e.~the joint distribution~(\ref{eqn: stationary}) does not depend on \(t\).
+	\\&= \mathbb P(X(t_n+t)\in E_n,X(t_{n-1}+t)\in E_{n-1},...,X(t_0+t)\in E_0),
+\end{align}
+for any \(t,t_0,...,t_n\in\mathcal T\) with \(t_0+t,...,t_n+t\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\).
 A random variable, \(\tau\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\tau\in\sigma(X(s), s\leq t),\) where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s), s\leq t\}\).
 
-We call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable and
+As in \cite[Chapter~1.2]{MEinAP}, we call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable and
 \begin{align}\label{eqn: Markov property}
@@
 \(\mathbb P(X(n+1)=j \mid X(n)=i)\) does not depend on \(n\).
+The probabilities \(\mathbb P(X(n+1)=j \mid X(n)=i)=:p_{ij}\) are the \emph{transition probabilities}, and \(\bs P=[p_{ij}]_{i,j\in\calS}\) is the \emph{transition matrix}, of the Markov chain.
 The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(n)\}_{n\in\{0,1,2,...\}}\), on the event that \(\{\tau<\infty\}\), then
@@
 
-We call \(\{X(t)\}_{t\geq 0}\) a \emph{continuous-time Markov chain} if \(\calS\) is countable and for all \(t_{n+1} > t_n > t_{n-1} > ... > t_0 > 0\) and \(i,i_0,...,i_{n-1},j\in\calS,\)
+Also, as in \cite[Section~1.3]{MEinAP}, we call \(\{X(t)\}_{t\geq 0}\) a \emph{continuous-time Markov chain} (CTMC) if \(\calS\) is countable and for all \(t_{n+1} > t_n > t_{n-1} > ...
+> t_0 > 0\) and \(i,i_0,...,i_{n-1},j\in\calS,\)
 \begin{align}\label{eqn: Markov property cmtc}
@@
 \end{align}
-and refer to (\ref{eqn: Markov property cmtc}) as the \emph{Markov property}.
-The process \(\{X(t)\}_{t\geq 0}\) is said to be time-homogeneous if 
-\(\mathbb P(X(t+s)=j \mid X(s)=i)\) does not depend on \(s\).
+and refer to (\ref{eqn: Markov property cmtc}) as the \emph{Markov property}. The process \(\{X(t)\}_{t\geq 0}\) is said to be time-homogeneous if 
+\(\mathbb P(X(t+s)=j \mid X(s)=i)\) does not depend on \(s\). For a time-homogeneous CTMC, the \emph{transition function} is the matrix function 
+\[\bs P(t) = [\mathbb P(X(t)=j \mid X(t)=i)]_{i,j\in\calS}.\]
+The Chapman-Kolmogorov equations state that 
+\begin{align} 
+	&\nonumber \bs P(t+s)=\bs P(t)\bs P(s).
+\end{align}
+The \emph{infinitesimal generator} matrix of a CTMC is 
+\begin{align} 
+	&\nonumber \bs T = [T_{ij}]_{i,j\in\calS} = \cfrac{\wrt}{\wrt t}P(t),
+\end{align}
+and the elements \(T_{ij}\) are the \emph{transition rates}. A CTMC is said to be \emph{conservative} if \(\bs P(t)\bs 1=\bs 1\) for all \(t\geq 0\) where \(\bs 1\) is a column vector of ones.
+
 The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(t)\}_{t\geq 0}\), on the event that \(\{\tau<\infty\}\), then 
@@
 
-For example, for a bounded linear operator \(\mathbb B\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\), then it can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), hence \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup.
-As another example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\).
-Then \(\sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is just the Taylor series of \(f\) about the point \(x\).
+For example, for a bounded linear operator \(\mathbb B:\mathcal D(\mathbb B)\to \mathcal D(\mathbb B)\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\) where, for \(f\in \mathcal D(\mathbb B)\), \(\mathbb B^kf\) is defined recursively by \(\mathbb B^k f = \mathbb B^{k-1} (\mathbb B f)\).
+It can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), hence \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup.
+As another example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega(\mathbb R) \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\).
+Then \(\mathbb T(t)f(x)=\sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is the Taylor series of \(f\) about the point \(x\) and the operator \(\mathbb T(t)f(x)=f(x-t)\) is the shift operator.
+The semigroup property follows since
+\[\mathbb T(t)\mathbb T(s)f(x)=\mathbb T(t)f(x-s)=f(x-s-t)=\mathbb T(t+s)f(x).\]
 
 \subsubsection{Infinitesimal generators} The \emph{infinitesimal generator} (or just \emph{generator} for short) of a semigroup \(\{\mathbb T(t)\}\) is the linear operator defined by 
-\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(\mathbb T(t)f-I),\]
+\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(\mathbb T(t)f-f),\]
 and the domain, \(\mathcal D(\mathbb B)\), is the subspace of \(L\) for which this limit exists.
@@
 To determine the generator \(\mathbb B\) we can proceed by either differentiation of the semigroup, 
-\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(\mathbb T(t)f-I),\]
+\[\mathbb Bf = \lim\limits_{t\to 0^+}\frac{1}{t}(\mathbb T(t)f-f),\]
 or integration;
@@
 The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\), and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\).
-We assume that \(\bs T\) is \emph{conservative}.
+We assume that \(\bs T\) is {conservative} and time-homogeneous.
 Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\) which determine the rate at which \(\{X(t)\}\) moves.
@@
 \[\bs T = \left[\begin{array}{ccc}\bs T_{++} & \bs T_{+-} & \bs T_{+0} \\ \bs T_{-+} & \bs T_{--} & \bs T_{-0} \\ \bs T_{0+} & \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
-where \(\bs T_{mn} = [T_{ij}]_{i\in\mathcal S_m, j\in\mathcal S_n}\), \(m,n\in\{+,-,0\}\).\footnote{Let's clarify some notation.
+where \(\bs T_{mn} = [T_{ij}]_{i\in\mathcal S_m, j\in\mathcal S_n}\), \(m,n\in\{+,-,0\}\).
+
+{Let's clarify some notation.
 We use the notation \(\bs u = (u_h)_{h\in \mathcal H}\) to denote a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some index set.
@@
 Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\).
-The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).} Also define the diagonal matrices
+The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).}
+
+Also define the diagonal matrices 
 \begin{align*}
@@
 Often, boundary conditions are imposed.
-We denote a fluid queue bounded below at \(0\) and unbounded above by \(\{\dot{\bs X}(t)\} = \{(\dot X(t),\varphi(t))\}_{t\geq0}\), and a fluid queue bounded below at \(0\) and above by \(b<\infty\) by \(\{{\bs X}(t)\} = \{( X(t),\varphi(t))\}_{t\geq0}\).
+We denote a fluid queue bounded below at \(0\) and unbounded above by \(\{\dot{\bs X}(t)\} = \{(\dot X(t),\varphi(t))\}_{t\geq0}\), and a fluid queue bounded below at \(0\) and above at \(b<\infty\) by \(\{{\bs X}(t)\} = \{( X(t),\varphi(t))\}_{t\geq0}\).
 Here, we consider a mixture of \emph{regulated} and \emph{reflecting} boundary conditions.
@@
 until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_-\) at which time \(\{X(t)\}\) leaves the boundary.
-Without loss of generality, we assume the lower and upper boundaries (when present) are at \(x=0\) and \(x=b>0\), respectively.
+It is without loss of generality that we assume the lower and upper boundaries (when present) are at \(x=0\) and \(x=b>0\), respectively.
 
@@
 Let \(\bs f(x,t) = (f_i(x,t))_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i\mid \bs X(0)\sim \bs \mu )\), assuming it exists.
-When a differentiable density exists, the system of partial differential equation which describes the evolution of the densities \(\bs f(x,t)\) is
+When a differentiable density exists, the system of partial differential equations which describes the evolution of the densities \(\bs f(x,t)\) is
 \begin{align}
@@
 In such situations, it is the \emph{weak solution} to (\ref{eqn: pde}) that we seek.
-A weak solution satisfies
+A weak solution (ignoring boundary conditions) satisfies
 \begin{align}
@@
 \end{align}
-for every row-vector of test functions, \(\bs \psi(x,t) = ( \psi_i(x,t))_{i\in\calS}\), which are smooth, have compact support and \(\bs \psi(x,0)=\bs \psi(0,t) = \bs \psi(b,t)=\bs 0\).
+for every row-vector of test functions, \(\bs \psi(x,t) = ( \psi_i(x,t))_{i\in\calS}\), which are smooth, have compact support and \(\bs \psi(x,0)=\bs \psi(0,t) = \bs \psi(b,t)=\bs 0\) \cite[Chapter~10]{borthwickBook}.
 
@@
 Moreover, these techniques often lead to quantities which have direct probabilistic interpretations which further aids in their analysis and also applications to other problems \citep{ar2003,dasilva2005,bnp2018} (we also leverage these stochastic interpretations in Chapter~\ref{sec: conv} to derive expressions for certain the Laplace transforms of the fluid queue).
-In some context, the actual transient distributions are not required, and Laplace transforms or moments are all that are required.
+In some contexts, the actual transient distributions are not required, and Laplace transforms or moments are all that are required.
 In other cases, we can invert the Laplace transforms using known methods (\cite{aw2006}, or in the case of the discontinuities we might prefer \cite{hhat2020}, which uses the same concentrated matrix exponential distribution that we do).
@@
 		\begin{minipage}{0.8\textwidth}
-			\textit{This subsection was largely taken from Sections~2 and 3 of \cite{blnos2022} with changes, such as notations, so that this chapter is consistent with the rest of the thesis.
+			\textit{This subsection was largely taken from Sections~2 and 3 of \cite{blnos2022} with changes, such as notation, so that this chapter is consistent with the rest of the thesis.
 I am a co-author of the paper \cite{blnos2022}.
@@
 
-For the remainder of this section, we summarise the findings of~\cite{bo2014} on the joint stationary distribution of $\{( \dot Y(t), X(t), \varphi(t))\}_{t \geq 0}$.
-The derivation of the stationary distribution relies on obtaining the operator \(\mathbb \Psi\) which gives the distribution of the process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to the level 0, given \(\dot Y(0)=0\).
+For the remainder of this section, we summarise the findings of~\cite{bo2014} on the joint limiting distribution of $\{( \dot Y(t), X(t), \varphi(t))\}_{t \geq 0}$.
+The derivation of the limiting distribution relies on obtaining the operator \(\mathbb \Psi\) which gives the distribution of the process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to the level 0, given \(\dot Y(0)=0\).
 
@@
 
-We assume that the process $\{(\dot Y(t), X(t), \varphi(t))\}_{t\geq 0}$ is positive recurrent, in order to guarantee the existence of the joint stationary distribution. Define stationary operators 
+We assume that the process $\{(\dot Y(t), X(t), \varphi(t))\}_{t\geq 0}$ is positive recurrent, in order to guarantee the existence of the joint limiting distribution. Define limiting operators 
 \begin{align} 
@@
 where $\mathcal{E} \subset [0,b]$.
-Then let ${{\bbpi}}(y) = ( \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint stationary density operators and ${\mathbb{p}} = ( {\mathbb p}_i)_{i \in \mathcal{S}}$ be a vector containing the joint stationary mass operators.
+The keen observer will notice that (\ref{eqn:jointpi}) assumes a certain differentiability condition on the limiting distribution.
+Indeed, it is not always the case that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) (trivially, consider \(\dot Y(t)=y\) for all \(t\)).
+Conditions on fluid-fluid queues to ensure that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) are not known, and here we just assume that (\ref{eqn:jointpi}) is differentiable.
+
+Now, let ${{\bbpi}}(y) = ( \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint limiting density operators and ${\mathbb{p}} = ( {\mathbb p}_i)_{i \in \mathcal{S}}$ be a vector containing the joint limiting mass operators.
 The determination of ${{\bbpi}}(y)$ involves two important matrices of operators, $ {\mathbb D}$ and $ {\mathbb \Psi}$ which we now introduce.
@@
 
-The generator of a fluid queue is a differential operator and to enable computation of the first-return operator, approximation method are needed.
+The generator of a fluid queue is a differential operator and to enable computation of the first-return operator, approximation methods are needed.
 The approximation schemes which we discuss in this thesis are all cell-based methods which discretise the level of the fluid queue into cells.
@@
 Specifically, the domain of \(\overline{\mathbb B}\) is the set of measures, \(\overline{\bs \mu} = (\overline\mu_i)_{i\in\calS}\), for which each \(\overline\mu_i\) admits an absolutely continuous density on \((0,b)\), and can have a point mass at \(0\) if \(i\in\mathcal S_{-1}\) or a point mass at \(b\) if \(i\in\mathcal S_{K+1}\); call this set of measures \(\mathcal M_{0,b}\).
-The measure \(\overline\mu_i\) cannot have a point mass at 0 if \(i\notin \mathcal S_{-1}\), nor can they have a point mass at \(b\) if \(i\notin\calS_{{K+1}}\).
+The measures \(\overline\mu_i\) cannot have a point masses at 0 if \(i\notin \mathcal S_{-1}\), nor can they have a point masses at \(b\) if \(i\notin\calS_{{K+1}}\), nor can they have point masses in \((0,b)\) for any \(i\in\calS\).
 In the sequel we write \(\overline v_i(x)\) as the density of \(\overline\mu_i\), and \(q_{{-1},i}\) and \(q_{{K+1},i}\) as the point masses of \(\overline\mu_i\) at \(0\) and \(b\), respectively (if such point masses exist).
@@
 The schemes we choose to use here work by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \({\calD_k,i}\) then constructing an approximation to \(\overline{\mathbb B}\) on each interval.
-To help elucidate the connection between the operators \(\{ \overline{\mathbb  V}(t)\}\), \( \overline{\mathbb  B}\) and their approximation counterparts we take a slightly different approach to partitioning these operators than that taken in \cite{bo2014}.
+To help elucidate the connection between the operators \(\{ \overline{\mathbb  V}(t)\}\), \( \overline{\mathbb  B}\) and their approximate counterparts we take a slightly different approach to partitioning these operators than that taken in \cite{bo2014}.
 Rather than partition according to the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\), we use the same partition as that in the construction of the approximation schemes.
@@
 Let us first partition the space \([0,b]\) into \(\calD_{-1,i}=\calD_{-1} = \{0\}\), \(i\in\calS_{-1}\), \(\calD_{{K+1},i}=\calD_{{K+1}}=\{b\}\), \(i\in\calS_{K+1}\), and non-trivial intervals \(\calD_{k,i}=[y_k,y_{k+1})\setminus \{0\},\) \(i\in\calS_+\cup\calS_0\) and \(\calD_{k,i}=(y_k,y_{k+1}]\setminus \{b\}\), \(i\in\calS_-\), with \(y_0=0,\,y_{K+1}=b,\, y_k<y_{k+1},\,k\in\mathcal K^\circ:=\{0,1,2,...,K\}\) and define \(\mathcal K = \{-1,K+1\}\cup\mathcal K^\circ\).
-For $\bs{\mu} \in \mathcal{M}_{0,b}(\mathcal{S} \times [0,b])$ we write \(\bs{\mu} = (\mu_{k,i})_{i \in \mathcal{S},k\in\mathcal K},\) where $\mu_{k,i}(\cdot) = \mu_i(\cdot \cap \calD_{k,i}),\,k\in\mathcal K.$ When it exists, we denote by \(v_{k,i}(x),\,x>0\), the densities associated with each measure, \(\mu_{k,i},\,k\in \mathcal K^\circ\).
+For $\bs{\mu} \in \mathcal{M}_{0,b}(\mathcal{S} \times [0,b])$ we write \(\bs{\mu} = (\mu_{k,i})_{i \in \mathcal{S},k\in\mathcal K},\) where $\mu_{k,i}(\cdot) = \mu_i(\cdot \cap \calD_{k,i}),\,k\in\mathcal K.$ We denote by \(v_{k,i}(x),\,x>0\), the density associated with the measure, \(\mu_{k,i},\,k\in \mathcal K^\circ\).
 For \( i,j\in\calS,\,k,\ell\in\mathcal K\) define the operators
@@
 \end{align}
-We persist with the partition \(\calD_{k,i},\,k\in\mathcal K\) throughout this thesis as this is consistent with the partition used in the approximation schemes considered in this thesis.
+We persist with the partition \(\calD_{k,i},\,k\in\mathcal K\) throughout as this is consistent with the partition used in the approximation schemes considered in this thesis.
 Note that for all the operators defined with this partition, the partitioning used in \citep{bo2014} can always be recovered by the above construction.
@@
 where \( 1(\cdot)\) is the indicator function and \(x^+\) and \(x^-\) denote the right and left limits at \(x\). 
-Intuitively, the first expression represents the instantaneous rate of the stochastic transitions of the phase process \(\{\varphi(t)\}\) from \(i\) to \(j\) and \(\{X(t)\}\) remains in \(\calD_{k,i}\).
+Intuitively, the first expression represents the instantaneous rate of the stochastic transitions of the phase process \(\{\varphi(t)\}\) from \(i\) to \(j\) while \(\{X(t)\}\) remains in \(\calD_{k,i}\).
 The first term in the second expression represents the net rate of transition out of phase \(i\) while \(\{X(t)\}\) remains in \(\calD_{k,i}\), while the second and third terms in the second expression represent the flux out of the right-hand edge of \(\calD_{k,i}\) when \(c_i>0\) and the flux out of the left-hand edge of \(\calD_{k,i}\) when \(c_i<0\), respectively.
@@
 
-At the boundary \(x=0\), point mass moves between phases according to 
+At the boundary \(x=0\), the rate at which point masses move between phases is 
 \begin{align*}
 	\mu_{-1,i}\mathbb B_{ii}^{{-1},{-1}} &= \mu_{-1,i}(\{0\})T_{ii},\mbox{ if } c_i\leq 0,\,\\
-	\mu_{-1,i}\mathbb B_{ij}^{{-1},{-1}} &= \mu_{-1,i}(\{0\})T_{ij},\mbox{ if }c_i\leq 0,\,c_j\leq 0,
-	\end{align*}
-	point mass becomes density at rate
-	\begin{align*}
-	\mu_{-1,i}\mathbb B_{ij}^{{-1},0} &= \mu_{-1,i}(\{0\})T_{ij},\mbox{ if }c_i\leq 0,\,c_j> 0,
-	\end{align*}
-	and density becomes point mass at rate
+	\mu_{-1,i}\mathbb B_{ij}^{{-1},{-1}} &= \mu_{-1,i}(\{0\})T_{ij},\mbox{ if }c_i\leq 0,\,c_j\leq 0. 
+	\end{align*}
+	The rate at which point mass leaves the boundary is
+	\begin{align*}
+	\mu_{-1,i}\mathbb B_{ij}^{{-1},0} &= \mu_{-1,i}(\{0\})T_{ij},\mbox{ if }c_i\leq 0,\,c_j> 0. 
+	\end{align*}
+	The rate at which point masses accumulate at the boundary is 
 	\begin{align*}
@@
 
-Similarly, at the boundary \(x=b\), point mass moves between phases according to 
+Similarly, at the boundary \(x=b\), the rate at which point masses move between phases is 
 \begin{align*}
 	\mu_{K+1,i}\mathbb B_{ii}^{{K+1},{K+1}} &= \mu_{K+1,i}(\{0\})T_{ii}, \mbox{ if }c_i\geq 0,\,\\
-	\mu_{K+1,i}\mathbb B_{ij}^{{K+1},{K+1}} &= \mu_{K+1,i}(\{0\})T_{ij},\mbox{ if }c_i\geq 0,\,c_j\geq 0,
-	\end{align*}
-	point mass becomes density at rate
-	\begin{align*}
-	\mu_{K+1,i}\mathbb B_{ij}^{{K+1},K} &= \mu_{K+1,i}(\{0\})T_{ij},\mbox{ if }c_i\geq 0,\,c_j<0,
-	\end{align*}
-	and density becomes point mass at rate
+	\mu_{K+1,i}\mathbb B_{ij}^{{K+1},{K+1}} &= \mu_{K+1,i}(\{0\})T_{ij},\mbox{ if }c_i\geq 0,\,c_j\geq 0.
+	\end{align*}
+	The rate at which point mass leaves the boundary is
+	\begin{align*}
+	\mu_{K+1,i}\mathbb B_{ij}^{{K+1},K} &= \mu_{K+1,i}(\{0\})T_{ij},\mbox{ if }c_i\geq 0,\,c_j<0.
+	\end{align*}
+	The rate at which point masses accumulate at the boundary is 
 	\begin{align*}
@@
 Note that we have not presented \(\mathbb B\) in its full detail here and refer the reader to \citep{bo2014} for the details.
-The main goal here is to show how \(\mathbb B\) is used to construct the stationary distribution of the fluid-fluid queue and to illustrate the link between the operator \(\mathbb B\) and the approximations of the same object.
+The main goal here is to show how \(\mathbb B\) is used to construct the limiting distribution of the fluid-fluid queue and to illustrate the link between the operator \(\mathbb B\) and the approximations of the same object.
 As we shall see later, these expressions closely resemble the approximations to the same quantities.
@@
 
-\subsection{Stationary Distribution}
+\subsection{Limiting Distribution}
 
@@
 We define $\zeta_Y^n(\{0\}) := \inf\{t \geq \zeta_Y^{n - 1}(\{0\}): \dot Y(t) = 0\}$, for $n \geq  2$, to be the sequence of hitting times to level $0$ of $\dot Y(t)$, with $\zeta_Y^1(\{0\}): = \zeta_Y(\{0\})$.
-Consider the discrete-time Markov process $\{X({\zeta_Y^n(\{0\})}), \varphi(\zeta_Y^n(\{0\}))\}_{n \geq 1}$, and for $i \in \mathcal{S},\,k\in\mathcal K_i^-$ define the measures $\bbxi_{k,i}$ as follows
+Consider the discrete-time Markov process $\{X({\zeta_Y^n(\{0\})}), \varphi(\zeta_Y^n(\{0\}))\}_{n \geq 1}$, and for $i \in \mathcal{S},\,k\in\mathcal K_i^-$ define the measures $\bbxi_{k,i}$ given by
 	%
@@
 	\end{align*} 
-	By \cite{bo2014}, the vector of measures $\boldsymbol{\bbxi} := (\bbxi_{k,i})_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}$ satisfies the following set of equations 
+	By \cite{bo2014}, the vector of measures $\boldsymbol{\bbxi} := (\bbxi_{k,i})_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}$ satisfies the set of equations 
  % 
@@
 
-We reproduce Theorem 2 of \cite{bo2014} below, which gives the joint stationary distribution of $\{(\dot Y(t),X(t),  \varphi(t))\}$.
-Recall that the joint stationary density operator ${\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{(\dot Y(t), X(t), \varphi(t))\}$ and the joint stationary mass operator ${\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$ are defined by~\eqref{eqn:jointpi} and \eqref{eqn:jointmass}, respectively.
-%Following out notational convention, define an equivalent joint stationary density operator $\boldsymbol{\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{X(t), Y(t), \varphi(t)\}$ and an equivalent joint stationary mass operator $\boldsymbol{\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$, but partitioned accprsing .
-We can partition \(\bbpi\) as follows 
+We reproduce Theorem 2 of \cite{bo2014} below, which gives the joint limiting distribution of $\{(\dot Y(t),X(t),  \varphi(t))\}$.
+Recall that the joint limiting density operator ${\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{(\dot Y(t), X(t), \varphi(t))\}$ and the joint limiting mass operator ${\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$ are defined by~\eqref{eqn:jointpi} and \eqref{eqn:jointmass}, respectively.
+%Following out notational convention, define an equivalent joint limiting density operator $\boldsymbol{\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{X(t), Y(t), \varphi(t)\}$ and an equivalent joint limiting mass operator $\boldsymbol{\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$, but partitioned accprsing .
+We can partition \(\bbpi\) so that 
 % 
@@
 	\label{theo:density} 
-The operator ${\bbpi}^{m}(y)$, for $m \in \{+,-,0\}$ and $y > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the following set of equations:
+The operator ${\bbpi}^{m}(y)$, for $m \in \{+,-,0\}$ and $y > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the set of equations:
 % 	 
@@
 \section{Quasi-birth-and-death processes with rational arrival process components}\label{qbd-rap intro}
-One of the approximation schemes we develop in this thesis is a quasi-birth-and-death process with rational-arrival-process components (QBD-RAP).
+One of the approximation schemes we develop in this thesis is based on a suitably defined quasi-birth-and-death process with rational-arrival-process components (QBD-RAP).
 QBD-RAP processes are built from matrix-exponentially distributed inter-event times.
@@
 The density function of \(Z\) is given by \(f_Z(x) = \bs\alpha e^{\bs S x}\bs s\).
-The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) There is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility.
-These condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\).
+The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) In general, in \cite{MEinAP}, there is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility.
+The condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\).
 However, in general there is no way to determine whether, a given triplet \((\bs \alpha, \bs S, \bs s)\) is a representation of a matrix exponential distribution, or not.
@@
 \begin{thm}[Theorem 4.1.4, \cite{MEinAP}]\label{thm: 4.1.4}
-	Consider the nonvanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}=0,\,d_{jk} =0,\) or \(e_{jk}=0\).
+	Consider the nonvanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}\neq 0,\,d_{jk}\neq 0,\) or \(e_{jk}\neq 0\).
 Among the corresponding eigenvalues \(\lambda_j\), there is a real dominating eigenvalue \(\kappa\), say.
@@
 
-The class of matrix exponential distribution is characterised as the class of probability distributions which have a rational Laplace transform.
+The class of matrix exponential distributions is characterised as the class of probability distributions which have a rational Laplace transform.
 That is, \(\displaystyle \int_{x=0}^\infty e^{-\lambda x}\bs \alpha e^{\bs S x}\bs s \wrt x\) is a ratio of two polynomial functions in \(\lambda\).
@@
 where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
-The process \(\{\bs A(t)\}\) `jumps' at event times of \(N\) (the process may not actually jump at these times, but we still call it a `jump' and typically the dynamics change at this point).
+The process \(\{\bs A(t)\}\) can jump at event times of \(N\) (the process may not actually jump at these times, but we may still refer to it a `jump' and typically the dynamics change at this point).
 At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\), i.e.~\(\mathbb P(N(t)=n,N(t+dt)=n+1)=\bs A(t) \bs D\bs e\wrt t\).
@@
 \[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
-where the \emph{region on convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
+where the \emph{region of convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
 \[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
@@
 In Chapters~\ref{sec: conv} and~\ref{ch: global results} we use the Dominated Convergence Theorem to aid our convergence arguments.
-In applied probability we often want to prove convergence of certain probabilistic expressions but the expressions themselves are too difficult to characterise as one.
-An approach which can simplify matters is to partition the expression on certain events where we have a simpler characterisation of the expression on each event, thereby enabling us to prove convergence on each element in the partition.
+In applied probability we often want to prove convergence of certain probabilistic expressions.
+An approach which can simplify matters is to partition the expression on certain events where we have a simpler characterisation, thereby enabling us to prove convergence on each element in the partition.
 The original expression can be written as a sum over the partition.
@@
 Once again, we follow \cite{steinreal} quoting their Theorem~2.13.
-If \(f\) is a function in \(\mathbb R^{d_1}\times \mathbb R^{d_2}\), the \emph{slice} of \(f\) corresponding to \(y\in\mathbb R^{d_2}\) is the function \(f^y\) of the \(x\in\mathbb R^{d_1}\) variable, given by
+If \(f\) is a function in \(\mathbb R^{d}=\mathbb R^{d_1}\times \mathbb R^{d_2}\), the \emph{slice} of \(f\) corresponding to \(y\in\mathbb R^{d_2}\) is the function \(f^y\) of the \(x\in\mathbb R^{d_1}\) variable, given by
 \[f^y(x)=f(x,y).\]
@@
 
-A closely related theorem which is often used alongside Fubini's Theorem is the following 
+A closely related theorem which is often used alongside Fubini's Theorem is Tonelli's Theorem.
+Define the \emph{extended Lebesgue integral} of an extended valued (it can take the value \(+\infty\)) non-negative function \(f\) by
+\[\int f(x)\wrt x = \sup_{g}\int g(x)\wrt x.\]
 \begin{thm}[Tonelli's Theorem]\label{Tonelli}
@@
 
-Where the \emph{extended Lebesgue integral} of an extended valued (it can take the value \(+\infty\)) non-negative function \(f\) is defined by 
-\[\int f(x)\wrt x = \sup_{g}\int g(x)\wrt x.\]
 Once again, we note that the theorem is symmetric in \(x\) and \(y\), so we can establish that we may swap the order of integration provided that \(f\) is non-negative.
+Tonelli's Theorem allows us to exchange the order of integration for any non-negative extended real-valued function, and includes the case where the value of the integral is infinite.
+In contrast, Fubini's Theorem us to exchange the order of integration for any real-valued function, provided that the integral is finite.
 
@@
 \section{More on relevant literature \& the context of the thesis} 
-In the process of introducing the technical concepts we have already covered some the relevant literature already.
-Here, we provide some more context for, and elucidate exactly what, the contributions of this thesis.
+In the process of introducing the technical concepts we have already covered some the relevant literature.
+Here, we provide some more context for the contributions of this thesis.
 The focus of this thesis is on developing numerical approximations of fluid queues such that we may approximate the operator-analytic analysis of fluid-fluid queues form \cite{bo2014}.
@@
 \paragraph{Fluid-fluid queues and related models} Related to the analysis of fluid-fluid queues are the works of \cite{mz2012,lnp13,bo2013b} and \cite{bop2020}.
-\cite{mz2012} analyse related discrete-time multidimensional Markov additive processes and derive operator-analytic expressions for the stationary distribution.
+\cite{mz2012} analyse related discrete-time multidimensional Markov additive processes and derive operator-analytic expressions for the limiting distribution.
 Although markedly different in their approach \cite{mz2012}, like  \cite{bo2014}, is inspired by a matrix-analytic approach.
@@
 Their analysis is somewhat specific to the data handling model considered and only arrives at bounds for the marginal limiting distribution of the second fluid level.
-In contrast, the work of \cite{bo2014} treats more general models and derives operator-analytic expressions for the joint stationary distribution.
-Therefore, compared to \cite{lnp13}, the methods of this thesis therefore apply to a general class of models and enable to approximation of the joint stationary distribution.
-However, unlike \cite{lnp13}, the analysis of \cite{bo2014} does not consider a level-dependent driving process such as this, but I believe that it would be relatively straight forward to extend their results (and the results of this thesis) to the level-dependent case where the rates \(c_i\) depend on \(X(t)\) in a piecewise-constant way.
+In contrast, the work of \cite{bo2014} treats more general models and derives operator-analytic expressions for the joint limiting distribution.
+Therefore, compared to \cite{lnp13}, the methods of this thesis therefore apply to a general class of models and enable to approximation of the joint limiting distribution.
+However, unlike \cite{lnp13}, the analysis of \cite{bo2014} does not consider a level-dependent driving process such as this, but I believe that it would be relatively straightforward to extend their results (and the results of this thesis) to the level-dependent case where the rates \(c_i\) depend on \(X(t)\) in a piecewise-constant way.
 \cite{bo2013b} analyse a model which is simple special case of the fluid-fluid queue.
@@
 
-Returning now to \cite{bo2014} which introduces the fluid-fluid queue in its generality and derives operator-analytic expressions for the first-return operator of the second fluid and the joint stationary distribution.
+Returning now to \cite{bo2014} which introduces the fluid-fluid queue in its generality and derives operator-analytic expressions for the first-return operator of the second fluid and the joint limiting distribution.
 Their analysis is inspired by the analysis of classical fluid queues in \cite{bean2005} whereby matrix-analytic expressions for the fluid level are derived in terms of the generator of the driving process.
@@
 \paragraph{Transient analysis of fluid queues} The analysis of fluid queues can be classified broadly into two approaches, matrix-analytic methods (for example, \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}) and differential equations based approaches (such as \cite{anick1982,kk1995,blnos2022}).
-Often, it is the stationary distribution of the fluid queue which is of interest.
+Often, it is the limiting distribution of the fluid queue which is of interest.
 In the context of the analysis of fluid-fluid queue as in \cite{bo2014}, however, we require information about the transient distribution of the fluid queue.

### 2022-06-19 chapter2/chapter.tex
Ch7 2nd draft

@@
 	\begin{minipage}{0.8\textwidth}
-		\textit{Apart from Sections~\ref{sec: other applications}-\ref{sec: slope limiting}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis.
+		\textit{Apart from Sections~\ref{sec: other applications}-\ref{sec: slope limiting}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notation, so that this chapter is consistent with the rest of the thesis.
 I am a co-author of the paper \cite{blnos2022}.
@@
 % Approximating R: projection, interpolation, cell averages (eluding to QBD-RAP and unif method)
-% Approximating Psi (and the stationary distribution, but this isnt really necessary)
+% Approximating Psi (and the limiting distribution, but this isnt really necessary)
     % constructing D
@@
 The DG method conserves probability.
-Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the stationary distribution of a stochastic fluid-fluid queue.
+Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the limiting distribution of a stochastic fluid-fluid queue.
 
@@
 We refer to \(u_{k,i}\) as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_{k,i}\) on the whole domain.
-The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\).
+The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\mathbb Z\).
 
@@
 In general, to construct the weak-form we need a set of test functions, say \(W\).
-Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to the derivative with respect to \(x\), to get
+Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to get
 \begin{align}
@@
 \end{align}
-Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), which we now substitute into (\ref{eqn:hash})
+Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), where the prime, \({}\tr{}\), denotes the transpose, which we now substitute into (\ref{eqn:hash})
 %\begin{align*}
@@
 %\end{align*}
-and repeat this for all test functions \(\phi^r_k(x)\), \(r=1,...,p_k\), to get the following system of equations,
+and repeat this for all test functions \(\phi^r_k(x)\), \(r=1,...,p_k\), to get the system of equations,
 \begin{align}
@@
 \[\boldsymbol a_k(t) = (\boldsymbol a_{k,i}(t))_{i\in\mathcal S},\quad \boldsymbol a(t) = (\boldsymbol a_k(t))_{k\in\mathbb Z},\]
-and the double-infinite block-tridiagonal matrix 
+and the doubly-infinite block-tridiagonal matrix 
 \begin{align*}
@@
 
-Let us first consider the boundary at \( X_t=0\). The following boundary conditions describe the evolution of probability/density of a stochastic fluid queue with a boundary at \(0\);
+Let us first consider the boundary at \( X_t=0\). The boundary conditions that describe the evolution of probability/density of a stochastic fluid queue with a boundary at \(0\) are;
 \begin{align}\label{eqn:BC1}
@@
 
-For future reference, we also define the matrices \({{\bs B}}^{k\ell}_{ij}\) for \(k\in\{2,\dots,K-1\}\) by
+For future reference, we also define the matrices \({{\bs B}}^{k\ell}_{ij}\) for \(k\in\{2,\dots,K-1\}\), \(\ell\in\{k-1,k\},\) \(i,j\in\calS\), by
 \begin{align*}
@@
 \section{Application to a fluid-fluid queue}\label{sec:DGSFFM}
-Given our approximation \(\bs B\) to the generator \(\mathbb B\) we now follow the recipe from \cite{bo2014}, replacing the actual generator \(\mathbb B\) with its approximation \({  \bs B}\), to construct approximations, \(\bs \pi\) and \(\bs p\), to the stationary operators, \(\bbpi\) and \(\mathbb p\).
+Given our approximation \(\bs B\) to the generator \(\mathbb B\) we now follow the recipe from \cite{bo2014}, replacing the actual generator \(\mathbb B\) with its approximation \({  \bs B}\), to construct approximations, \(\bs \pi\) and \(\bs p\), to the limiting operators, \(\bbpi\) and \(\mathbb p\).
 
@@
 \end{align*}
-Multiplying on the left by \(\bs\phi^+(x)\tr{}\) and on the right by \(\bs\phi^-(y)\), integrating over both \(x\) and \(y\), then post-multiplying by \({\bs M}^{-1}_-\) gives the following matrix Riccati equation
+Multiplying on the left by \(\bs\phi^+(x)\tr{}\) and on the right by \(\bs\phi^-(y)\), integrating over both \(x\) and \(y\), then post-multiplying by \({\bs M}^{-1}_-\) gives the matrix Riccati equation
 \begin{align}\label{eqn:RiccatiPsi}
@@
 \end{align}
-Thus, we may find \(\bs \Psi(s)\) by solving \eqref{eqn:RiccatiPsi}, using one of the methods in \citep{bot08}. Here, we use the Newtons method.
+Thus, we may find \(\bs \Psi(s)\) by solving \eqref{eqn:RiccatiPsi}, using one of the methods in \citep{bot08}. Here, we use Newtons method.
 
@@
 
-\subsection{Putting it all together: constructing an approximation to the stationary distribution}
-We find an approximation to the stationary distribution by replacing the theoretical operators in Theorem~\ref{theo:density} with their approximations.
-Table~\ref{table:notations} defines the notation we use for the DG approximations to stationary operators.
+\subsection{Putting it all together: constructing an approximation to the limiting distribution}
+We find an approximation to the limiting distribution by replacing the theoretical operators in Theorem~\ref{theo:density} with their approximations.
+Table~\ref{table:notations} defines the notation we use for the DG approximations to limiting operators.
 
@@
  \end{tabular}
- \caption{Notation for the approximation of the stationary operators of a fluid-fluid queue.
+ \caption{Notation for the approximation of the limiting operators of a fluid-fluid queue.
 The first column contains the operators which we are approximating, the second column contains indices for which the operators are defined, the third column defines the notation we use for the coefficients of the approximation, and the last column defines how the coefficients and basis functions are used to approximate the operators.
@@
 % \end{tabular}
-% \caption{Notation for the approximation of the stationary operators of an SFFM.\label{table:notations}}
+% \caption{Notation for the approximation of the limiting operators of an SFFM.\label{table:notations}}
 % \end{table}
@@
 %	
-%	Next, we seek to approximate the stationary operators \(\bs p\) defined in (\ref{eqn:jointmass}).
+%	Next, we seek to approximate the limiting operators \(\bs p\) defined in (\ref{eqn:jointmass}).
 For \(i,k,r\) such that \(i\in\calS,\,r\in\{1,...,N_k\},\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}\), coefficients \(p_{k,i}^r\) to be determined later, such that
@@
 \section{Other problems we can solve with cell-based approximation schemes}\label{sec: other applications}
-The utility of the DG scheme (or any of the approximation schemes considered in this thesis, for that matter) for fluid queues is not limited to approximating the first return and stationary operators of fluid-fluid queues.
+The utility of the DG scheme (or any of the approximation schemes considered in this thesis, for that matter) for fluid queues is not limited to approximating the first return and limiting operators of fluid-fluid queues.
 For example, we can use a DG scheme to approximate the transient distribution of the fluid queue at time \(t\), which requires us to find the coefficients \(\vligne{\bs q_{-1}(t) & \bs a(t) & \bs q_{{K+1}}(t)}\), given an initial condition, typically by numerically integrating (\ref{eqn: DG ODE w BCs}).
-An approximation to the stationary distribution of the fluid queue can be found by solving
+An approximation to the limiting distribution of the fluid queue can be found by solving
 \begin{align*}
@@
 Indeed, the DG scheme has the potential to answer many mathematical questions for fluid queues.
-However, one reason it may not have found such widespread use in the context of fluid queues is the aforementioned difficulties surrounding discontinuous solutions, which we often want to consider.
-Moreover, there are sometimes other techniques which can give exact solutions, for example exact expressions for transient quantities are derived in \cite{rs2003,ar2004,bean2005,bean2009}, and for the stationary distribution are derived in \cite{ajr2005,dasilva2005,lnp13,s2017}.
+However, one reason DG schemes may not have found such widespread use in the context of fluid queues is the difficulties surrounding discontinuous solutions, which we elaborate on in Section~\ref{sec: slope limiting} and demonstrate in Chapter~\ref{sec: numerics} (see also Section~5.6~\cite{nodalDGBook}), which we often want to consider.
+Moreover, there are sometimes other techniques which can give exact solutions, for example exact expressions for transient quantities are derived in \cite{rs2003,ar2004,bean2005,bean2009}, and for the limiting distribution are derived in \cite{ajr2005,dasilva2005,lnp13,s2017}.
 
@@
 To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
-Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory and non-negativity by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
+Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
 However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
@@
 This is not to say that the slope-limited-regions of the approximation will \emph{always} remain linear.
-When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of Generalised slope limiter described.
+When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of generalised slope limiter described.
 Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy of subsequent computations.
@@
 The advantage of the approach above is that, in areas of the approximation where the solution is smooth, then the high-order accuracy of the method can be retained, while maintaining positivity where necessary.
-However, if the problem at hand is dominates by discontinuities and/or if there are no regions of the solution of interest which are smooth, then there is no real advantage in this approach as the limiter will reduce the approximate solution to linear in all regions of interest.
+However, if the problem at hand is dominated by discontinuities and/or if there are no regions of the solution of interest which are smooth, then there is no real advantage in this approach as the limiter will reduce the approximate solution to linear in all regions of interest.
 Hence, we may as well just use a linear scheme and save on some computation.
@@
 On each of the \(p\) sub-cells we construct block matrices of size \(2\times 2\).
-To approximate the solution on the original interval \(\mathcal D_k\) we form the \(p\) approximations on each sub-cell into larger matrices which are of dimension \(2p\times 2p\), and we then use in the same way as before.
-Thus, the scheme with \(2p\) basis functions on the cell \(\calD_k\), and the scheme with a 2 basis functions on \(p\) sub-cell have approximately the same computational cost (the latter scheme results in a sparser matrix).
+To approximate the solution on the original interval \(\mathcal D_k\) we form the \(p\) approximations on each sub-cell into larger matrices which are of dimension \(2p\times 2p\), and then use in the same way as before.
+Thus, the scheme with \(2p\) basis functions on the cell \(\calD_k\), and the scheme with a 2 basis functions on each of the \(p\) sub-cell have approximately the same computational cost (the latter scheme results in a sparser matrix).
 
@@
 
-\subsection{Briefly on accuracy}
-As a rough guide on accuracy we now paraphrase some know results on the accuracy of the DG method for a similar kind of problem to the one under study here.
+\subsection{Briefly, on accuracy}
+As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method for a similar kind of problem to the one under study here.
 For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.

### 2022-06-19 chapter6/chapter.tex
Ch7 2nd draft

@@
 We have now established multiple schemes for approximating fluid queues which are suitable for approximating the performance measures of fluid-fluid queues derived in \cite{bo2014}.
+This chapter investigates, numerically, various aspects of the approximation schemes.
+Throughout, we compare the QBD-RAP scheme (from Chapter~\ref{sec: construction and modelling}), the discontinuous Galerkin scheme (as described in Chapter~\ref{ch:galerkin}, with and without slope limiting), and the \emph{spatially-coherent uniformisation} scheme introduced by \cite{bo2013}.
+
+The performance of the discontinuous Galerkin scheme has been well-studied in some contexts (\cite{c99}, \cite{nodalDGBook}~Section~5.5), and it is well-known that the discontinuous Galerkin scheme performs remarkably well on problems with smooth solutions.
+Here, we mostly focus on investigating the numerical performance of the schemes on problems with non-smooth solutions.
+The purpose is to emphasise the probabilistic interpretation and positivity-preserving properties of the QBD-RAP scheme.
+In the stochastic modelling community it is common to have problems with discontinuities, such as non-smooth initial conditions, hence the type of problems we investigate in this chapter are of relevance.
+As we will see (in Section~\ref{sec: transient approx}), even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities or non-smooth behaviour into the problem.
+A very specific set of conditions on the initial density and point masses must hold for the transient distribution to remain continuous as it evolves over time \citep{bo2014,bop2020}.
+Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues (see, for example, Section~\ref{sec: ffq num}).
+Limiting distributions of fluid queues are smooth, however.
+
+The experiments demonstrate that the DG scheme can perform extremely well when the true solution is sufficiently smooth.
+However, when discontinuities are present in the true solution, the DG scheme can lead to approximate solutions which are oscillatory and this may lead to approximate solutions which violate the axioms of probability, such as approximations to PDFs which take negative values (equivalently, approximations to CDFs which are non-increasing), and in the worst cases, approximations to CDFs which take values outside \([0,1]\).
+Hence, there is a need for schemes which ensure the axioms of probability are maintained.
+
+By design, the QBD-RAP scheme, uniformisation scheme, and DG schemes with slope limiting, produce approximations which obey the axioms of probability, i.e.
+produce approximations to PDFs which are non-negative, and CDFs which are non-decreasing, and \([0,1]\)-valued.
+
+We implement two slope limited DG schemes, the DG-lim and DG-lin-lim scheme, which we describe in more detail in the next section.
+The numerical experiments demonstrate that the DG-lim scheme does not lead to a viable approximation scheme in the presence of discontinuities, while the DG-lin-lim scheme results in a relatively performant approximation scheme for the discontinuous problems considered here.
+In general, the QBD-RAP scheme performs similarly to the DG-lin-lim scheme, while the uniformisation scheme is a viable approximation method, but produces larger errors than the QBD-RAP and DG-lin-lim scheme, for the examples considered here.
+
+Furthermore, when slope limiting is used, the resultant approximation is at best linear around discontinuities.
+Moreover, there is a computational cost in applying a limiter (or filter), which could be significant, particularly if many initial conditions are to be considered.
+In contrast, for the uniformisation and QBD-RAP, the approximate solution is guaranteed to produce positive probabilities without any post-processing or extra computation.
+
+Lastly, for the DG scheme I know of no rigorous argument which shows that we can use the DG approximation to the generator of a fluid queue to approximate performance measures of fluid-fluid queues, such as the operator \(\mathbb \Psi\).
+Here, we rely on the intuition that substituting the DG approximation for the true operator \emph{should} give reasonable results, and further, \emph{assume} that the resulting approximations are polynomial.
+For example this means that, other than being the solution to a matrix Riccati equation, the DG approximations to \(\mathbb \Psi\) that we compute in this chapter are not actually rigorously defined.
+
+On the other hand, using the stochastic approximations (QBD-RAP of uniformisation), results in rigorously defined operators as proved in \cite{bgnp2021}.
+
+In summary, the QBD-RAP and uniformisation schemes are the only two schemes which result in rigorously defined approximations, preserve the axioms of probability, and, of the two, the numerical experiments demonstrate that the QBD-RAP methods is superior.
+
+\section{Preliminaries}
+Before presenting the results of the experiments, we need to describe some aspects of the approximation schemes in more detail, which we now proceed to do.
+
+The main focus of the numerical experiments is on the error as the number of basis functions per cell is increase.
+In the numerical experiments the number of basis functions used to approximate the solutions is the same on all cells and the number of basis functions on a cell (and hence all cells) is referred to as the \emph{dimension} of the scheme.
+For the discontinuous Galerkin scheme, increasing the dimension of the scheme means increasing the number of polynomial basis functions used to approximate the solution within each cell.
+For example, if we use 3 basis functions in the discontinuous Galerkin scheme, we approximate the solution by 3 linearly independent quadratics on each cell and hence the approximation is a quadratic on each cell.
+For the QBD-RAP scheme, increasing the dimension of the scheme means increasing the order of the CME distribution used to construct the scheme.
+To make a comparable equivalent for the uniformisation scheme, we divide each cell into smaller sub-cells over which we approximate the solution.
+That is, for a dimension \(p\) uniformisation scheme we divide each cell into \(p\) sub-cells.
+Equivalently, we may think of a dimension \(p\) uniformisation scheme as using \(p\) piecewise constant functions to approximate the solution on each cell.
+For all schemes (DG, QBD-RAP and uniformisation), if we construct a dimension \(p\) approximation, there are \(K\) cells, \(N\) phases and \(c_i\neq 0\) for all \(i\in\calS\), then the resulting approximation to the generator \(\mathbb B\) is a square matrix of dimension \(pKN + N\).\footnote{There are \(pK\) basis functions for each phase which approximate the solution on the interior of the domain, hence \(pKN\) basis functions in total, plus an additional \(N\) dimensions to capture boundary conditions.} In this sense each approximation scheme leads to matrices of the same order (although not necessarily the same number of non-zero elements).
+
+We implement the discontinuous Galerkin scheme with and without a slope-limiter (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
+The slope limiter stops oscillations appearing in the approximations and hence also avoids non-probabilistic approximations to PDFs and CDFs.
+We investigate two slope-limited schemes.
+In the scheme we refer to the \emph{DG-lim} scheme, we use a DG scheme of dimension \(p\) and apply the Generalised MUSCL slope limiter as described in Section~\ref{sec: slope limiting}.
+In the \emph{DG-lin-lim} scheme, we follow the construction described in Section~\ref{sec: anoth pos pres}, which is effectively a DG scheme with cells of width \(2\Delta/(p+1)\), for \(p=1,3,5,...,21\), and a basis of dimension \(2\) (linear) on each cell, with an upwind flux, and we apply the Generalised MUSCL limiter.
+Thus, the DG-lin-lim scheme uses \(p+1\), \(p=1,3,5,...,21\), basis functions to represent the solution on each of the original cells \(\calD_k\), and we define the dimension of the scheme to be \(p+1\).
+In the presence of discontinuities, slope-limiting forces the approximation to be linear, i.e.~an order~1 polynomial, around the discontinuity.
+For DG schemes of dimension greater than 2 with a slope limiter and in the presence of discontinuities, the ability of the scheme to represent higher-order polynomials is not well-utilised as the slope limiter forces the solution to be linear anyway.
+The DG-lin-lim scheme recognises this and instead of using higher-order polynomials, uses piecewise-linear functions and smaller cell widths to approximate the solution.
+
+The concept of slope-limiting originates in the context of numerically integrating (typically with respect to a \emph{time} variable) a so-called \emph{semi-discrete system}, as we do in Section~\ref{sec: transient approx}.
+However, approximating the operator \(\mathbb\Psi\) of a fluid-fluid queue amounts to solving an implicit matrix-equation and there is no concept of numerical integration with respect to time.
+Hence, the concept of slope-limiting, it is not directly transferable.
+One could process the initial condition and post-process the solution with a limiter.
+However, there are possibly better choices for this post-processing, such as filtering, \citep[Section~5.6.1]{nodalDGBook}.
+One could also view the composition of the DG operator and the limiting operator as a one, but the result is a non-linear operator which is of no use for simplifying the Riccati Equation which characterises \(\mathbb \Psi\).
+Considering there is no theory on the use of the DG scheme to approximate \(\mathbb \Psi\) and hence the resultant approximation is not rigorously-defined, it is not known if the use of a limiter in this context is possible.
+
+To keep the content of this chapter contained, we do not investigate all aspects of the schemes.
+For each numerical experiment we keep the number of cells, \(K\) (equivalently, the cell size, \(\Delta\)), fixed for the DG and QBD-RAP schemes.
+As part of deriving a discontinuous Galerkin scheme, one needs to choose the \emph{numerical flux} which is used to approximate the transition of density from one cell to the next \citep{nodalDGBook}.
+We investigate schemes with an upwind flux only.
+We do not investigate filtering for the DG scheme.
+
+For schemes which require us to integrate over time we do not investigate the stability of the schemes with respect to the \(t\)-step-size, or the time-integration scheme itself (where required).
+Instead, we fix the time-integration step size for each numerical experiment at a suitably small value to obey a certain stability criterion (a CFL-like condition, \citep[Section~4.8]{nodalDGBook}).
+Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (\cite{sr2002}, \cite{nodalDGBook}~Section~5.7, see also Section~\ref{sec: time integration}), which claims to introduce no more oscillations into the solution as we integrate over time.
+
+Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
+
+In this chapter, all the error plots are on a \(\log_{10}\)-\(\log_{10}\) scale.
+If the error function takes the form \(error = \beta_0^* p^{\beta_1}\), where \(p\) is the dimension of the scheme, then the \(\log_{10}\)-\(\log_{10}\) will show a linear trend since \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\), where \(\beta_0=\log_{10}(\beta_0^*)\).
+Where relevant, to estimate the asymptotic rate of convergence, \(\beta_1\), we estimate the slope of the line \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\) using ordinary least squares and the last eight data points of each line.
+The estimated linear trends are plotted and the slope of the line is displayed to the right of the plot.
+
+The structure of this chapter is as follows.
+In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different schemes and observe their performance which allows us to investigate the performance of the reconstruction methods without considering a specific model or any dynamics.
+In Section~\ref{sec: wave num} we investigate travelling wave problems with various initial conditions.
+For the travelling wave problem the dynamics are deterministic, which means that the solution is known, and we do not need to resort to simulation as a \emph{ground-truth}.
+Furthermore, the travelling wave problems allow us to investigate the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
+Next, we investigate a simple fluid queue with two phases.
+Section~\ref{sec:stat} investigates approximations to the limiting distribution of the queue, Section~\ref{sec: transient approx} investigates approximations to the transient distribution of the queue for two initial conditions, and Section~\ref{sec: hit approx} investigates approximations the first hitting time of the fluid level on the set \(\{0,1\}\).
+Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num}, compute approximations to \(\mathbb \Psi\) and ultimately, approximate the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\).
+
+% \chapter{Old}
+% We have now established multiple schemes for approximating fluid queues which are suitable for approximating the performance measures of fluid-fluid queues derived in \cite{bo2014}.
 In this chapter we numerically investigate various aspects of the approximation schemes.
@@
 %Recall that the uniformisation scheme is equivalent to a discontinuous Galerkin scheme with a constant basis function in each cell, which is also equivalent to the finite-volume scheme with an upwind flux.
-Recall that, due to their stochastic interpretation, the uniformisation and QBD-RAP schemes are positivity preserving.
-
-The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased.
+% Recall that, due to their stochastic interpretation, the uniformisation and QBD-RAP schemes are positivity preserving.
+
+% The main focus of our numerical experiments is on the convergence properties as the number of basis functions is increased.
 For simplicity, the number of basis functions is kept constant across all cells.
@@
 
-To keep the content of this chapter contained, we do not investigate all aspects of the schemes.
+% To keep the content of this chapter contained, we do not investigate all aspects of the schemes.
 For each numerical experiment we keep the cell size, \(\Delta\), fixed for the DG and QBD-RAP schemes.
@@
 
-For some numerical experiments we implement the discontinuous Galerkin scheme with and without a slope-limiter implemented (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
+% For some numerical experiments we implement the discontinuous Galerkin scheme with and without a slope-limiter implemented (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
 The slope limiter avoids oscillations and negative solutions.
@@
 
-Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
+% Where a slope limiter is implemented, we implement the \emph{Generalised MUSCL} limiter (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
 We investigate two schemes with slope limiting.
@@
 
-The performance of the discontinuous Galerkin scheme has been well-studied in some contexts (\cite{c99}, \cite{nodalDGBook}~Section~5.5), and it is well-known that the discontinuous Galerkin scheme performs remarkably well on problems with smooth solutions.
+% The performance of the discontinuous Galerkin scheme has been well-studied in some contexts (\cite{c99}, \cite{nodalDGBook}~Section~5.5), and it is well-known that the discontinuous Galerkin scheme performs remarkably well on problems with smooth solutions.
 Here, we mostly focus on investigating the numerical performance of the schemes on problems with non-smooth solutions.
@@
 
-As we shall see, if the solution is smooth, then the DG scheme is highly-effective and exhibits rapid convergence to the solution as the number of polynomial basis functions is increased.
+% As we shall see, if the solution is smooth, then the DG scheme is highly-effective and exhibits rapid convergence to the solution as the number of polynomial basis functions is increased.
 However, when the solution is non-smooth oscillations and negative approximations may occur when using the discontinuous Galerkin scheme, leading to nonsense solutions.
@@
 
-Lastly, before we start the numerical investigation, note that all the following error plots are on a \(\log_{10}\)-\(\log_{10}\) scale.
+% Lastly, before we start the numerical investigation, note that all the following error plots are on a \(\log_{10}\)-\(\log_{10}\) scale.
 If the error function takes the form \(error = \beta_0^* p^{\beta_1}\), where \(p\) is the dimension of the scheme, then the \(\log_{10}\)-\(\log_{10}\) will show a linear trend; \(\log_{10}(error) = \beta_0 + \beta_1 \log_{10}(p)\), where \(\beta_0=\log_{10}(\beta_0^*)\).
@@
 
-The structure of this chapter is as follows.
+% The structure of this chapter is as follows.
 In Section~\ref{sec: recon num} we compute approximations to various initial conditions for the different schemes and observe their performance.
-This allows us to instrument the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
+This allows us to investigate the performance of the reconstruction methods without considering a specific model or any dynamics of the problem.
 In Section~\ref{sec: wave num} we investigate travelling wave problems with various initial conditions.
-For the travelling wave problem the dynamics are deterministic, which allows us to instrument the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
+For the travelling wave problem the dynamics are deterministic, which allows us to investigate the ability of the schemes to approximate the flow of probability across cells without any stochastic dynamics.
 Next, we investigate a simple fluid queue with two phases.
-Section~\ref{sec:stat} investigates the ability of the schemes to approximate the stationary distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the model for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the schemes to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
+Section~\ref{sec:stat} investigates the ability of the schemes to approximate the limiting distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the model for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the schemes to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
 Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num} and approximate the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\).
@@
 
-% 3. STATIONARY DISTRIBUTIONS
+% 3. limiting DISTRIBUTIONS
 	% i. SIMPLER
@@
 We start our investigation by looking at how well the schemes perform at approximating an initial condition.
-By approximating the initial condition only, we aim to instrument the performance of the approximation schemes without any dynamics.
-
-For the discontinuous Galerkin method, we project the initial condition on to the set of polynomial functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstructing the solutions as described in Sections~\ref{sec: initial conditions} and~\ref{sec: closing}.
-For the purposes of approximation and reconstruction for the QBD-RAP scheme we must orientate the initial condition in a positive or negative direction; here, we suppose that the initial conditions belong to a positive phase.
-First we investigate three closing vectors we can use for the reconstruction in the QBD-RAP scheme.
+By approximating the initial condition only, we aim to investigate the performance of the approximation schemes without any dynamics.
+
+For the discontinuous Galerkin method, we project the initial condition on to the set of polynomial basis functions which define the scheme, for the spatially-coherent uniformisation scheme, we look at the sub-cell averages of the initial condition, and for the QBD-RAP scheme, we compute the initial vector for the approximation, then reconstruct the solutions as described in Sections~\ref{sec: initial conditions} and~\ref{sec: closing}.
+For the purposes of approximation and reconstruction for the QBD-RAP scheme we must orientate the initial condition in a positive or negative direction; here we suppose that the initial conditions belong to a positive phase.
+First we investigate the three closing vectors from Section~\ref{sec: closing} which we can use for the reconstruction in the QBD-RAP scheme.
 From the investigation, we decide to use the closing vector (\ref{eqn:5462}) throughout the rest of the chapter.
 
-For this investigation we consider approximating initial conditions over a single cell of width \(\Delta = 1\).
+In this section we consider approximating initial conditions over a single cell of width \(\Delta = 1\).
 To numerically evaluate integrals arising in the approximation step (inner products and cell averages) we use a trapezoidal rule with 10,001 function evaluations.
-Similarly, we use 10,001 function evaluations to approximate \(L^p\) errors, and also as a finite set of points over which to compute the KS statistics.
+Similarly, we use 10,001 function evaluations to approximate \(L^p\) errors, and also as a finite set of points over which to compute the Kolmogorov-Smirnov (KS) statistics.
 
@@
 \subsection{QBD-RAP closing operators}\label{sec: closing vecs numerics}
-The three closing operators we investigate are the following.
-The unnormalised closing vector in Equation~(\ref{eqn: alal}).
-A normalised version of the closing operator in Equation~(\ref{eqn:546}), with the normalised version given by (\ref{eqn: nonlin}).
-We refer to the closing vector in (\ref{eqn: nonlin}) as the \emph{naive normalised} closing vector.
-Technically we have not yet proved that the \emph{naive normalised} closing vector leads to a convergent approximation scheme.
-The third closing vector we investigate is that in Equation~(\ref{eqn:5462}) and we refer to this as the \emph{normalised closing operator}.
+The three closing operators we investigate are the following. The unnormalised closing vector, 
+\begin{align}
+	\bs a {\bs v}(x) &= \bs a e^{\bs Sx}\bs s.\tag{\ref{eqn: alal}}
+	%\\\bs a {\bs v}_-(x-y_\ell) &= \bs a e^{\bs S(x-y_{\ell})}\bs s,\label{eqn:ndaenf}
+\end{align} 
+A normalised version of the closing operator in Equation~(\ref{eqn:546}), with the normalised version given by 
+\begin{align}
+	\bs a {\bs v}(x) &= \cfrac{\bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)} \bs s\right)}{1-\bs a e^{\bs S 2\Delta}\bs e}. \tag{\ref{eqn: nonlin}}
+\end{align} 
+We refer to the closing operator in (\ref{eqn: nonlin}) as the \emph{non-linear normalised} closing operator.\footnote{Technically we have not yet proved that the \emph{non-linear} normalised closing operator leads to a convergent approximation scheme.} The third closing vector we investigate is given by
+\begin{align}
+	\bs a {\bs v}(x) &= \bs a \left(e^{ \bs{S}x} \bs s + e^{ \bs{S}(2\Delta-x)}\right)\left[I-e^{S2\Delta}\right]^{-1} \bs s,\tag{\ref{eqn:5462}}
+\end{align}
+which we refer to as the \emph{normalised} closing operator.
 
 \begin{example}\label{ex: another example}
-	First consider approximating the initial condition with density function \(2\times 1(x<0.5)\).
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: another example} for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves of the non-linear normalised (orange) and normalised closing vectors are almost coincident.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}%To estimate the asymptotic rate of convergence, we fit a linear model of \(\log_{10}(error) = \beta_0 + \beta_1\log_{10}(Dim)\) to the the last 6 data points on the curves above, which corresponds to a rate of convergence \(error \propto N^{-\beta_1}\).
+The estimate of \(\beta_1\) was approximately -1.0 in all cases.}
+		\label{fig: fun 2 ks error qbdrap closing vecs}
+	\end{figure} 
+	%\exampleFloatBarrier
+	First consider approximating the initial condition with density function \(2\times 1(x<0.5)\), where \(1(\cdot)\) is the indicator function.
 In the left-hand side of Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} we plot the Kolmogorov-Smirnov error between the true CDF and the CDF constructed via the QBD-RAP approximation with the three different closing vectors.
@@
 
-	Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the naive normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this subsection.
-The difference between the naive normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward).
+	Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the non-linear normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this section.
+The difference between the non-linear normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward).
 Intuitively, there should be very little mass in the tail of the distribution as we are using concentrated matrix exponential distributions in the reconstruction.
@@
 If we instead look at the \(L^2\) error between the true PDF and the reconstructions in Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} (right), then we see that the unnormalised reconstruction performs the poorest.
-Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests the naive normalised and normalised reconstruction are very similar.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: another example} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves of the naive normalised (orange) and normalised closing vectors are almost coincident.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}%To estimate the asymptotic rate of convergence, we fit a linear model of \(\log_{10}(error) = \beta_0 + \beta_1\log_{10}(Dim)\) to the the last 6 data points on the curves above, which corresponds to a rate of convergence \(error \propto N^{-\beta_1}\).
-The estimate of \(\beta_1\) was approximately -1.0 in all cases.}
-	\label{fig: fun 2 ks error qbdrap closing vecs}
-\end{figure} 
-\exampleFloatBarrier
+Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also suggests the non-linear normalised and normalised reconstruction are very similar.
 \end{example}
@@
 \begin{example}\label{ex: al}
-	Now consider approximating the initial density \(f(x)=1\).
-Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error and \(L^2\) error of the PDF we now see that the normalised reconstructions outperform the unnormalised reconstruction.
-This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution.
 	\begin{figure}[h]
@@
 		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: al} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves of the error curves of the naive normalised (orange) and normalised closing vectors are almost coincident.
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: al} for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves of the error curves of the non-linear normalised (orange) and normalised closing vectors are almost coincident.
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
@@
 	\end{figure}
-
-Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised and normalised closing operators for dimension 1, 3, 5 and 7, as well as the true PDF.
+	Now consider approximating the initial density \(f(x)=1\) for \(x\in[0,1)\).
+Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error and \(L^2\) error of the PDF we now see that the normalised reconstructions outperform the unnormalised reconstruction.
+This suggests that, in this case, the `folding' of the closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution.
+
+Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised (\ref{eqn:546}) and normalised (\ref{eqn:5462}) closing operators for dimension 1, 3, 5 and 7, as well as the true PDF.
 Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots.
@@
 In comparison, the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator.
-The `folding' in the normalised closing operator manifest itself as more mass at the left of these plots compared to the unnormalised reconstructions.
+The `folding' in the normalised closing operator manifests itself as more mass at the left of these plots compared to the unnormalised reconstructions.
 % Recall that, as discussed in Section~\ref{}, for the reconstruction we must to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}.
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
+	\caption{Approximations to the PDF \(f(x)=1,\, x\in[0,1)\) (orange dotted line) for Example~\ref{ex: al}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
+	\label{fig: pdf reconstructed}
+\end{figure} 
+%\exampleFloatBarrier
 Figure~\ref{fig: pdf reconstructed} also show that both closing operators fail to approximate the initial condition well at the right-hand side of the interval.
-Perhaps there is a different reconstruction method or another closing operator which could alleviate this issue.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
-	\caption{Approximations to the PDF \(f(x)=1\) (orange dotted line) for Example~\ref{ex: al}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
-	\label{fig: pdf reconstructed}
-\end{figure} 
-\exampleFloatBarrier
+A potential area for further research is to determine if there is a different reconstruction method or another closing operator which could alleviate this issue.
 \end{example}
@@
 \begin{example}\label{ex: aasmm}
-	Next we consider approximating the initial distribution with density \(-6x^2+6x\).
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error for the PDFs (right) of Example~\ref{ex: aasmm}, for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves for the non-linear normalised (orange) and normalised closing vectors are almost coincident.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 6 ks error qbdrap closing vecs}
+	\end{figure}
+	Next we consider approximating the initial distribution with density \(-6x^2+6x\) on \([0,1)\).
 Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the KS error against the dimension of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
 However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the \(L^2\) error between the reconstructed PDF and the true PDF, then the unnormalised closing operator performs better than the two normalised ones for dimension 5 and above.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error for the PDFs (right) of Example~\ref{ex: aasmm}, for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves for the naive normalised (orange) and normalised closing vectors are almost coincident.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: fun 6 ks error qbdrap closing vecs}
-\end{figure}
-
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
+		\caption{Approximations to the PDF \(f(x)=-6x^2+6x\) (orange dotted line) for Example~\ref{ex: aasmm}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
+		\label{fig: pdf reconstructed quadratic}
+	\end{figure} 
+	%\exampleFloatBarrier
 The fact that the unnormalised closing operator outperforms the two normalised ones can be explained by the `folding' of the normalised operators around \(\Delta\).
@@
 In Figure~\ref{fig: pdf reconstructed quadratic} at the left of the plots, we observe that the normalised reconstructions over-estimate the density function, whereas the unnormalised reconstruction looks to be doing better.
-The `folding' in the normalised closing operator manifest itself as more mass at the left of these plots compared to the unnormalised reconstructions.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
-	\caption{Approximations to the PDF \(f(x)=-6x^2+6x\) (orange dotted line) for Example~\ref{ex: aasmm}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
-	\label{fig: pdf reconstructed quadratic}
-\end{figure} 
-\exampleFloatBarrier
+The `folding' in the normalised closing operator manifests itself as more mass at the left of these plots compared to the unnormalised reconstructions.
 \end{example}
@@
 \begin{example}\label{ex: kjfk}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: kjfk} for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
+The error curves for the non-linear normalised (orange) and normalised closing vectors are almost coincident.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 7 ks error qbdrap closing vecs}
+	\end{figure}
+	%\exampleFloatBarrier
 	Lastly, we consider approximating the initial distribution with PDF \(3e^{-3x}/(1-e^{-3})\).
@@
 Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that this is the case; the normalised reconstructions perform relatively well compared to the unnormalised reconstruction, as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF).
-If we observed plots of the PDFs (omitted), we would once again see that this is due to the loss of mass at the left-hand side of the region due to the truncation.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: kjfk} for the three closing vectors considered; unnormalised (blue solid line), naive normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves for the naive normalised (orange) and normalised closing vectors are almost coincident.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: fun 7 ks error qbdrap closing vecs}
-\end{figure}
-\exampleFloatBarrier
+Plots of the PDFs (omitted), confirm that this is due to the loss of mass at the left-hand side of the region due to the truncation.
 \end{example}
 
-In Table~\ref{table: error rates for QBD-RAP closing} we summarise the rates of convergence of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots).
-In general, the normalised closing operators converges at a rate approximately as fast, or faster, than the unnormalised reconstruction.
+Table~\ref{table: error rates for QBD-RAP closing} summarises the rates of convergence of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots).
+In general, the normalised closing operators converge at a rate approximately as fast, or faster, than the unnormalised reconstruction.
 \begin{table}
@@
 	\begin{tabular}{c|c|c|c}
-			Example & PDF & \(\beta_1\), unnormalised & \(\beta_1\), normalised \\\hline
+			Example & PDF on \([0,1)\) & \(\beta_1\), unnormalised & \(\beta_1\), normalised \\\hline
 			\ref{ex: another example} & \(2\times 1(x<0.5)\) & -1.07 & -0.98 \\
@@
 			\ref{ex: aasmm} & \(-6x^2+6x\) & -2.72 & -2.57 \\ 
-			\ref{ex: kjfk} & \(3e^{-3x}/(1-e^-3)\) & -1.11 & -2.35 
+			\ref{ex: kjfk} & \(3e^{-3x}/(1-e^{-3})\) & -1.11 & -2.35 
 	\end{tabular}
@@
 
-In summary, the naive normalised and normalised closing operators perform almost identically.
+In summary, the non-linear normalised and normalised closing operators perform almost identically.
 In some cases the unnormalised closing operator can outperform the normalised closing operators, typically when the mass of the function to be approximated at the left-hand edge of the cell is low.
-Further, the rates of convergence for all three schemes was similar, except for in Example~\ref{ex: kjfk} where we approximated the initial condition \(3e^{-3x}/(1-e^-3)\); in this case the rate of convergence of the normalised closing operators was approximately twice as large as the unnormalised version.
-Given the results above and its theoretical tractability, we will opt to use the normalised closing operator for the rest of the numerical experiments.
+Further, the rates of convergence for all three schemes were similar to each other, except for in Example~\ref{ex: kjfk} where we approximated the initial condition \(3e^{-3x}/(1-e^{-3})\); in this case the rate of convergence of the normalised closing operators was approximately twice as large as the unnormalised version.
+Given the results above and its theoretical tractability, we will opt to use the normalised closing operator in (\ref{eqn:5462}) for the rest of the numerical experiments.
 
@@
 Here we compare the ability of the QBD-RAP, uniformisation, and DG schemes to reconstruct initial conditions.
+With this section we aim to investigate the accuracy of the reconstruction methods for the numerical schemes without any dynamics.
 
 \begin{example}\label{ex: exa}
-	First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5.
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^1\) error of the CDF (right) of Example~\ref{ex: exa} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 1 comp} 
+	\end{figure}
+	First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass of \(1\) at \(0.5\).
 This distribution does not have a PDF, so we compare the CDFs only.
@@
 Observing the KS metric, it appears that none of the schemes converge and the KS error sits around 0.5 for all schemes.
-This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except, perhaps, at points of discontinuity and none of the schemes appear to converge at the discontinuity at \(x=0.5\).
+This reflects the fact that convergence in distribution implies point-wise convergence of the CDFs except, perhaps, at points of discontinuity.
+None of the schemes converge at the discontinuity at \(x=0.5\).
 Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the schemes appear to show the convergent behaviour we expect.
@@
 Perhaps it is no surprise that the uniformisation scheme performs best.
-In the uniformisation scheme as we increase the dimension we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the initial distribution.
+In the uniformisation scheme, as we increase the dimension, we partition the cell \([0,1)\) into smaller sub-cells, and use constant functions on each sub-cell to construct an approximate initial density.
 As such, the uniformisation scheme can produce a piecewise continuous, linear approximation to the CDF.
-In contrast, both the DG and QBD-RAP schemes result in a smooth approximation to the CDF.
-Given the initial distribution is far from smooth (it's a point mass), then we might expect that the uniformisation scheme will perform relatively well.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^1\) error of the CDF (right) of Example~\ref{ex: exa} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: fun 1 comp} 
-\end{figure}
-
-In Figure~\ref{fig: pdf comp fun 1} we plot approximated CDFs from the DG, uniformisation and QBD-RAP schemes alongside the true CDF.
-The DG scheme displays undesirable features for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1.
-On the other hand, although the QBD-RAP scheme converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1.
-The uniformisation scheme approximates the CDF exceptionally well.
+In contrast, both the DG and QBD-RAP schemes result in a smooth approximation to the CDF over the entire cell.
+Given the initial distribution is far from smooth (it is a point mass), then we might expect that the uniformisation scheme will perform relatively well.
+
 \begin{figure}[h]
@@
 \end{figure} 
-\exampleFloatBarrier
+%\exampleFloatBarrier
+In Figure~\ref{fig: pdf comp fun 1} we plot approximated CDFs from the DG, uniformisation and QBD-RAP schemes alongside the true CDF.
+The DG scheme displays undesirable features for an approximation to a CDF -- it is not monotonically increasing and is not constrained to the bounds of probability, \([0,1]\).
+On the other hand, although the QBD-RAP scheme converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1.
+The uniformisation scheme approximates the CDF exceptionally well.
 \end{example}
@@
 \begin{example}\label{eqn: example something}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{eqn: example something} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 2 comp} 
+	\end{figure}
+	%\exampleFloatBarrier
 	Now consider approximating the initial distribution with PDF \(1(x\leq 0.5)\).
@@
 However, once again the DG scheme exhibits undesirable properties (plots not shown) -- the approximation to the CDF is at some points above 1 and is not monotonic (although these violations do not appear to be as severe in this case as they were for Example~\ref{ex: exa}).
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{eqn: example something} for the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (purple dashed line) schemes.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: fun 2 comp} 
-\end{figure}
-\exampleFloatBarrier
 \end{example}
@@
 \begin{example}\label{ex: alllll}
-	So far we have considered two problems which exhibit discontinuities.
-At the other extreme we now consider an initial distribution with density \(-6x^2+6x\).
-In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDFs.
-Since the DG method projects the initial condition onto a polynomial basis, then, for a dimension 3 approximation and above, the DG scheme can approximate the initial condition exactly.
-This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from dimension 1 to 3, then plateaus.
-Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, the errors for the DG scheme are not 0.
-Regarding the other two schemes, they too appear to be convergent at approximately the same rate, with the uniformisation scheme performing better for the KS error, but very similarly in terms of the \(L^2\) error.
 	\begin{figure}[h]
@@
 	\end{figure}
-\exampleFloatBarrier
+	%\exampleFloatBarrier
+	So far we have considered two problems which exhibit discontinuities.
+At the other extreme we now consider an initial distribution with density \(-6x^2+6x\) on \([0,1)\).
+In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDFs.
+Since the DG method projects the initial condition onto a polynomial basis, then, for a dimension 3 approximation and above, the DG scheme can represent the initial condition exactly.
+This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from dimension 1 to 3, then plateaus.
+Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, the errors for the DG scheme are not 0.
+Regarding the other two schemes, they too appear to be convergent at approximately the same rate, with the uniformisation scheme performing better for the KS error, and both performing similarly in terms of the \(L^2\) error.
 \end{example}
 
-\begin{example}\label{ex: aalllkkdd}Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\).
-Figure~\ref{fig: fun 9 comp} shows the KS and \(L^2\) errors for the PDF.
-Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs (right) tell a similar story; for sufficiently high order, the DG scheme approximates the initial condition very well.
-The uniformisation scheme performs second best, while the QBD-RAP scheme is slowest to converge.
+\begin{example}\label{ex: aalllkkdd}
 	\begin{figure}[h]
@@
 	\end{figure}
-\exampleFloatBarrier
+	%\exampleFloatBarrier
+	Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\).
+Figure~\ref{fig: fun 9 comp} shows the KS and \(L^2\) errors for the PDF.
+Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs (right) tell a similar story; for sufficiently high order, the DG scheme approximates the initial condition very well.
+The uniformisation scheme performs second best, while the QBD-RAP scheme performs poorest.
 \end{example}
 
-In conclusion, we have observed that the DG scheme performs well with respect to the error metrics considered, but can display oscillations and result in CDFs which are not non-decreasing.
+\subsection*{Summary}
+First, we investigated the three proposed closing operators for the QBD-RAP scheme proposed in Section~\ref{sec: closing}.
+We concluded that the normalised closing operator is most appropriate, hence we will use it for the rest of this chapter.
+
+Next, we investigated reconstructing approximations to initial conditions for the DG, QBD-RAP and uniformisation schemes.
+We observed that the DG scheme performs well with respect to the error metrics considered, but can display oscillations and result in approximations which violate the axioms of probability.
 On the other hand, the uniformisation and QBD-RAP schemes result in non-decreasing approximations to the CDFs, but can converge slower with respect to the error metrics considered, with the QBD-RAP scheme often converging the slowest.
+
+This section investigated the performance of the schemes without considering any of the dynamics of the problem.
+By doing so, we have gained insights into the performance of the approximations independent of the dynamics.
+In the remaining sections we introduce dynamics into the problem.
+%Throughout the remainder of this chapter, keeping the findings of this section in mind will help us understand which errors are due to the reconstruction of the solution, and which errors are due to the approximation of the dynamics of the problem.
 \FloatBarrier
@@
 \section{Travelling wave problems}\label{sec: wave num}
-Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with various initial conditions.
+Here we investigate the performance of the different schemes for approximating transient distributions of one-dimensional travelling wave problems with various initial conditions.
 Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\).
@@
 \[\cfrac{\partial}{\partial t} f(x,t) = -\cfrac{\partial}{\partial x} f(x,t),\]
-where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
+where \(f(x,t)\) is the density (if/when it exists) at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
 \[f(x,t) = f(x-t,0)\]
@@
 We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).
-This example is convenient as it has a known solution and no stochastic dynamics, hence we can instrument the ability of the schemes to approximate the flow of mass, without any stochastic dynamics.
-
-We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space and discretise the interval \([0,10]\) into \(10\) cell, each of width 1.
-We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric.
+This example is convenient as it has a known solution and no stochastic dynamics, hence we can investigate the ability of the schemes to approximate the flow of mass, without any stochastic dynamics.
+
+We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space and discretise the interval \([0,10]\) into \(10\) cells, each of width 1.
+We use $10,001$ points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which to evaluate the CDFs to approximate the KS metric.
 Furthermore, we use the SSPRK4 method to integrate over time with a \(t\)-step size of \(0.005\), and we evolve the system until time \(t=4\).\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} For the DG scheme we also implement the Generalised MUSCL slope limiter in the form of the DG-lim and DG-lin-lim schemes (see also Section~\ref{sec: slope limiting}).
@@
 \begin{example}\label{ex: wave 4}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/L1_cell_probs.pdf}
+		\caption{Cell-wise error defined in (\ref{eqn: cell errors}) for the travelling wave in Example~\ref{ex: wave 4}.
+Plotted are the cell-wise errors for the DG (blue solid line), DG-lim scheme (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim scheme (gold dashed line) schemes, versus the dimension of the approximation.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 4 wave cp} 
+	\end{figure}
 First consider the initial condition with PDF \(1(x<1)\).
@@
 
-We plot the cell-wise error metric in Figure~\ref{fig: fun 4 wave cp} and observe that the DG scheme appears to converge fastest, followed by the QBD-RAP and DG-lin-lim schemes which perform comparably, then the uniformisation scheme.
-The worst performer is the DG-lim scheme which does not appear to converge.
-The DG scheme with a slope limiter has detected oscillations in the approximate solution and reduced the scheme to linear where oscillations are detected, thereby decreasing the accuracy of the scheme to linear where oscillations are present.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/L1_cell_probs.pdf}
-	\caption{Cell-wise error defined in (\ref{eqn: cell errors}) for the travelling wave in Example~\ref{ex: wave 4}.
-Plotted are the cell-wise errors for the DG (blue solid line), DG-lim scheme (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim scheme (gold dashed line) schemes, versus the dimension of the approximation.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: fun 4 wave cp} 
-\end{figure}
-
+We plot the cell-wise error metric in Figure~\ref{fig: fun 4 wave cp} and observe that the DG scheme clearly produces the lowest errors, followed by the QBD-RAP and DG-lin-lim schemes which perform comparably, then the uniformisation scheme.
+The worst performer is the DG-lim scheme which does not appear to converge; during the integration of the solution over time the DG-lim scheme has detected oscillations and reduced the scheme to linear where the oscillations were detected, thereby decreasing the accuracy of the scheme to linear where oscillations are present.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_l2_pdf_error_formatted.pdf}
+	\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 4} using the approximations from the DG (blue solid line), DG-lim scheme (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+	\label{fig: fun 4 wave} 
+\end{figure}
 Figure~\ref{fig: fun 4 wave} plots the KS error between the approximated and true CDFs and \(L^2\) error between the approximated and true PDFs.
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
+	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row) and DG-lin-lim schemes (fifth row), for dimension 1, 3, 5, and 7 (columns) for the travelling wave problem in Example~\ref{ex: wave 4}.
+The true density function is \(1(4\leq x<5)\).}
+	\label{fig: pdf wave fun 4} 
+\end{figure}
 Continuing with Figure~\ref{fig: fun 4 wave}.
@@
 With these error metrics, the DG scheme is converging fastest.
-However, if we observe the approximations resulting from the DG scheme (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscillations and negative values.
-
-Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG with MUSCL limiter and QBD-RAP schemes for various dimension schemes.
-In the first row we observe that the DG scheme (sans limiter) produces an approximation to the PDF with negative values when the dimension is greater than 1.
-In the third row of Figure~\ref{fig: pdf wave fun 4} we observe that, with the limiter, the DG approximation does not change significantly after dimension 3.
-This is due the fact that the DG scheme is at best linear in the presence of oscillations.
+However, if we observe the approximations resulting from the DG scheme (Figure~\ref{fig: pdf wave fun 4} top-row), we find them to be unsatisfactory due to oscillations and negative densities.
+
+Figure~\ref{fig: pdf wave fun 4} plots the density functions reconstructed using the DG, uniformisation, DG-lim, QBD-RAP and DG-lin-lim schemes for various dimension schemes.
+In the first row observe that, when the dimension is greater than one, the DG scheme (sans limiter) produces an approximation to the PDF which is negative in some places.
+In the third row of Figure~\ref{fig: pdf wave fun 4} observe that, with the limiter, the DG-lim approximation does not change significantly after dimension 3.
+This is due to the fact that the DG-lim scheme is at best linear in the presence of oscillations.
 In the second row of Figure~\ref{fig: pdf wave fun 4} is the solution approximated using the uniformisation scheme, in the fourth row is the solution approximated using the QBD-RAP scheme, and in the last row is the approximation from the DG-lin-lim scheme.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun4/meshs_l2_pdf_error_formatted.pdf}
-	\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 4} using the approximations from the DG (blue solid line), DG-lim scheme (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: fun 4 wave} 
-\end{figure}
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun4/pdfs_formatted.pdf}
-	\caption{Reconstructed PDFs using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row) and DG-lin-lim schemes, for dimension 1, 3, 5, and 7 (columns) for the travelling eave problem in Example~\ref{ex: wave 4}.
-The true density function is \(1(4\leq x<5)\).}
-	\label{fig: pdf wave fun 4}
-\end{figure}
 
 This is a particularly interesting example.
-It shows that, for the DG scheme, even though there is no discretisation error for the initial condition, we use a strong stability preserving time-integration method, and the projection off which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as in the top row of Figure~\ref{fig: pdf wave fun 4}.
-\exampleFloatBarrier
+For the DG scheme, even though there is no discretisation error for the initial condition, we use a strong stability preserving time-integration method, and the projection on which the DG method is based can represent the transient distribution at time \(t=4\) exactly, there is still the possibility of badly behaved solutions as shown in the top row of Figure~\ref{fig: pdf wave fun 4}.
+%\exampleFloatBarrier
 \end{example}
@@
 \begin{example}\label{ex: wave 2}
-Another interesting example occurs with the initial condition with CDF \(1(x\geq 0.5)\), i.e.~a point mass at 0.5.
+Another interesting example occurs with the initial condition with CDF \(1(x\geq 0.5)\), i.e.~a point mass of 1 at \(0.5\).
 The exact solution at time \(t=4\) is therefore a point mass at \(4.5\).
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/meshs_l1_cdf_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/L1_cell_probs.pdf} 
+	\caption{\(L^1\) error of CDFs (left) and the cell-wise error metric in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (right) for the travelling wave problem in Example~\ref{ex: wave 2} where approximation were constructed via the DG (blue solid line), DG-lin-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+	\label{fig: fun 2 wave} 
+\end{figure}
+%\exampleFloatBarrier
 Figure~\ref{fig: fun 2 wave} (left) plots the \(L^1\) error between the true and approximated CDFs.
-The \(L^1\) metric tells a similar story to the previous analysis: the DG-lim scheme does not converge as dimension increases, the other three positivity preserving schemes (the uniformisation, DG-lin-lim and QBD-RAP schemes) appear to converge, with the QBD-RAP and DG-lin-lim schemes converging at a similar rate and faster than the uniformisation scheme.
+The \(L^1\) metric tells a similar story to the previous analysis: the DG-lim scheme does not perform well, the other three positivity preserving schemes (the uniformisation, DG-lin-lim and QBD-RAP schemes) appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly and better than the uniformisation scheme.
 The DG scheme performs the best.
-However, if we plot the approximations to the CDFs from the DG scheme (not shown) we would once again see an oscillating, non-monotonic function.
+However, if we plot the approximations to the CDFs from the DG scheme (not shown) we once again see an oscillating, non-monotonic function.
 
@@
 However, in Figure~\ref{fig: fun 2 wave} (left) we see that the QBD-RAP scheme out-performs the uniformisation scheme.
-This suggests that the QBD-RAP scheme is better able to resolve movement of mass across the domain when integrating over time, compared to the uniformisation scheme.
+This suggests that the QBD-RAP scheme is better able to resolve movement of mass across the domain when integrating over time compared to the uniformisation scheme.
 
@@
 Interestingly, the error curve for the DG scheme is not monotonic.
-Furthermore, the DG scheme can result in negative estimates of the cell probabilities in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (plots not shown).
-Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme all appear to converge, with the QBD-RAP and DG-lin-lim schemes perform similarly and better than the DG scheme when the dimension is sufficiently large.
-The estimated rate of convergence of the QBD-RAP and DG-lin-lim schemes is approximately -4 for this example, which is significantly faster than the rate of convergence of these schemes in all other examples in this section (typically around -2 to -1).
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/meshs_l1_cdf_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun2/L1_cell_probs.pdf} 
-	\caption{\(L^1\) error of CDFs (left) and the cell-wise error metric in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (right) for the travelling wave problem in Example~\ref{ex: wave 2} where approximation were constructed via the DG (blue solid line), DG-lin-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: fun 2 wave} 
-\end{figure}
-\exampleFloatBarrier
+Furthermore, the DG scheme can result in negative estimates of the cell probabilities in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (not shown).
+Of the other three positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP scheme all appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly; even performing better than the DG scheme when the dimension is sufficiently large.
+The estimated rate of convergence of the QBD-RAP and DG-lin-lim schemes is approximately -4 for this example, which is significantly faster than the rate of convergence of any scheme in any of the other examples in this section (typically around -2 to -1).
 \end{example}
@@
 \begin{example}\label{ex: wave 6}
-	Now consider an initial distribution which is truncated Gaussian with mean parameter 2.5 and standard deviation parameter 0.5, and truncated at the boundaries, 0 and 10;
-	\begin{align}
-		\mu([0,x)) := \mathbb P(X(0)\leq x, \varphi(0)=1)=\cfrac{\Phi((x-2.5)/0.5)1(0 \leq x < 10.0)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10\leq x), \label{eqn: kdjfksdf}
-	\end{align}
-	where \(\Phi(x)\) is the CDF of the standard normal distribution.
-
-	At time \(t=4\) the distribution is truncated Gaussian with mean parameter 6.5, standard deviation parameter 0.5, and is truncated below at 4.0 and above at 10, and there is also a small mass at the upper boundary;
-	\begin{align}
-		&\mathbb P(X(4)\leq x, \varphi(4)=1\mid X(0)\sim\mu, \varphi(0)=1)  
-		=\cfrac{\Phi((x-6.5)/0.5)1(4.0\leq x)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)} + 1(10 \leq x).\label{eqn: bbbbbaabbaa}%\cfrac{\Phi(7.5/0.5)-\Phi(3.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}.
-	\end{align}
-	At time \(t=4\) the mass at the boundary is approximately \(1.28\times 10^{-12}\) and there is a small discontinuity at \(x=4\) where the CDF jumps from \(0\) to \(\sim 2.87\times 10^{-7}\), due to the truncation of the initial distribution at \(0\).
-
+	Now consider an initial distribution which is truncated Gaussian with mean 2.5 and standard deviation 0.5, and truncated at the boundaries, 0 and 10;
+	\begin{align}
+		\mu([0,x)) := \mathbb P(X(0)\leq x, \varphi(0)=1)=\cfrac{\Phi((x-2.5)/0.5)-\Phi(-2.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}, \label{eqn: kdjfksdf}
+	\end{align}
+	for \(x\in[0,10]\) and where \(\Phi(x)\) is the CDF of the standard normal distribution.
+
+	At time \(t=4\) the distribution is truncated Gaussian with mean 6.5, standard deviation parameter 0.5, and is truncated below at 4.0 and above at 10, and there is also a small mass at the upper boundary;
+	\begin{align}
+		\nonumber & \mathbb P(X(4)\leq x, \varphi(4)=1\mid X(0)\sim\mu, \varphi(0)=1)  
+		\\=&\cfrac{\Phi((x-4-2.5)/0.5)-\Phi(-2.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}1(x\in[4,10)) 
+		+ 1(x\geq 10).\label{eqn: bbbbbaabbaa}%\cfrac{\Phi(7.5/0.5)-\Phi(3.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}.
+	\end{align}
+	At time \(t=4\) the mass at the boundary is approximately \(1.28\times 10^{-12}\) and there is also a small discontinuity in the density at \(x=4\).
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/L1_cell_probs.pdf}
+		\caption{Cell-wise error defined in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) for the travelling wave problem in Example~\ref{ex: wave 6}.
+Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 6 wave cp} 
+	\end{figure}
 	Since the discontinuity at \(t=4,\,x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG scheme will perform well for this example.
@@
 
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error for the PDFs (right) for Example~\ref{ex: wave 6} where approximations were constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 6 wave} 
+	\end{figure}
 	The other three positivity preserving schemes appear to converge, with the uniformisation scheme performing the worst and the DG-lin-lim scheme performing the best of the three for all error metrics.
@@
 
-	To investigate the performance of the QBD-RAP and DG-lin-lim schemes, in Figure~\ref{fig: fun 6 wave pdfs} we plot the dimensions 21 QBD-RAP and dimensions 22 DG-lin-lim approximations to the PDF.
-Figure~\ref{fig: fun 6 wave pdfs} shows that the DG-lin-lim scheme approximates the PDF well, except for at the peak located at \(x=6.5\) (I suspect the limiter has affected the solution in this region).
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun6/order_21_PDFs_formatted.pdf}
+		\caption{Dimension 21 QBD-RAP (purple) and dimension 22 DG-lin-lim (gold) approximations to the transient PDF at time \(t=4\) for Example~\ref{ex: wave 6}. The true PDF is plotted in grey.}  
+		\label{fig: fun 6 wave pdfs} 
+	\end{figure}
+%\exampleFloatBarrier
+	To investigate this unexpected fact, in Figure~\ref{fig: fun 6 wave pdfs} we plot the dimension 21 QBD-RAP and dimension 22 DG-lin-lim approximations to the PDF.
+Figure~\ref{fig: fun 6 wave pdfs} shows that the DG-lin-lim scheme approximates the PDF well, except at the peak located at \(x=6.5\) (the limiter has affected the solution in this region).
 Specifically, the DG-lin-lim scheme underestimates the density in the peak, but overestimates the density either side of the peak.
@@
 The QBD-RAP scheme also overestimates the density in the adjacent cells, \([5,7]\) and \([7,8]\).
-Hence, the overall cell-wise error is large.
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/L1_cell_probs.pdf}
-		\caption{Cell-wise error defined in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) for the travelling wave problem in Example~\ref{ex: wave 6}.
-Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-		\label{fig: fun 6 wave cp} 
-	\end{figure}
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_ks_error_formatted.pdf}%
-		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun6/meshs_l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error for the PDFs (right) for Example~\ref{ex: wave 6} where approximations were constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-		\label{fig: fun 6 wave} 
-	\end{figure}
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun6/order_21_PDFs_formatted.pdf}
-		\caption{Dimension 21 QBD-RAP (purple) and dimension 22 DG-lin-lim (gold) approximations to the transient PDF at time \(t=4\) for Example~\ref{ex: wave 6}. The true PDF is plotted in grey.}  
-		\label{fig: fun 6 wave pdfs} 
-	\end{figure}
-\exampleFloatBarrier
+Hence, the overall cell-wise error is larger for the QBD-RAP scheme.
 \end{example}
@@
 
-	We suppose that the initial condition is a point mass at the boundary in phase \(2\), i.e. \[\mathbb P(X(0)\leq x, \varphi(0)=2)=1(0\leq x).\] 
+	We suppose that the initial condition is a point mass of 1 at the boundary \(x=0\) in Phase~\(2\), i.e. \[\mathbb P(X(0)\leq x, \varphi(0)=2)=1(0\leq x).\] 
 	With this initial condition, the transient distribution at time \(t=4\) is 
@@
 		&\mathbb P(X(4)\leq x,\varphi(4)=1 \mid X(0)=0,\varphi(0)=2) \nonumber 
-		\\&= e^{T_{22}4}\left(e^{-T_{22}x}-1\right)1(4< x) + (1-e^{T_{22}4})1(4\geq x) \nonumber
-		\\&= e^{-4}\left(e^{4x}-1\right)1(4< x) + (1-e^{-4})1(4\geq x) \label{eqn: asjda}
+		\\&= \left(e^{T_{22}(4-x)}-e^{T_{22}4}\right)1(4< x) + (1-e^{T_{22}4})1(4\geq x) \nonumber
+		\\&= \left(e^{-(4-x)}-e^{-4}\right)1(4< x) + (1-e^{-4})1(4\geq x) \label{eqn: asjda} 
 	\end{align}
@@
 
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/L1_cell_probs.pdf} 
+		\caption{Cell-wise error for the travelling wave problem in Example~\ref{ex: wave 1}.
+Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 1 wave cp}  
+	\end{figure}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_ks_error_formatted.pdf}%
+		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_l2_pdf_error_formatted.pdf}
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 1} where approximations are constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+		\label{fig: fun 1 wave} 
+	\end{figure}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun1/pdf_formatted.pdf}
+		\caption{Approximate transient PDFs at time \(t=4\) for Example~\ref{ex: wave 1} using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row), and DG-lin-lim (fifth row) schemes of dimensions 1, 3, 5, and 7 (columns).
+The true density function is \(e^{-(x-4)}1(x<4)\).}
+		\label{fig: pdf wave fun 1}
+	\end{figure} 
+	%\exampleFloatBarrier
 	Figure~\ref{fig: fun 1 wave cp} plots the cell-wise error metric and Figure~\ref{fig: fun 1 wave} plots the KS error (left) and the \(L^2\) error between the PDFs (right).
@@
 A selection of approximations to the transient PDF are shown in Figure~\ref{fig: pdf wave fun 1}.
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/L1_cell_probs.pdf} 
-		\caption{Cell-wise error for the travelling wave problem in Example~\ref{ex: wave 1}.
-Plotted are the cell-wise errors for the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes, versus the dimension of the approximation.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-		\label{fig: fun 1 wave cp}  
-	\end{figure}
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_ks_error_formatted.pdf}%
-		\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/wave/fun1/meshs_l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error of the PDF (right) for the travelling wave problem in Example~\ref{ex: wave 1} where approximations are constructed via the DG (blue solid line), DG-lim (green dash-dotted line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-		\label{fig: fun 1 wave} 
-	\end{figure}
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/wave/fun1/pdf_formatted.pdf}
-		\caption{Approximate transient PDFs at time \(t=4\) for Example~\ref{ex: wave 1} using the DG (top row), uniformisation (second row), DG-lim (third row), QBD-RAP (fourth row), and DG-lin-lim schemes of dimensions 1, 3, 5, and 7 (columns).
-The true density function is \(e^{-4}e^{x}1(x<4)\).}
-		\label{fig: pdf wave fun 1}
-	\end{figure} 
-	\exampleFloatBarrier
+
+	Interestingly, this solution is discontinuous at a cell-edge and even though the DG scheme\footnote{All the schemes considered allow discontinuities at cell-egdes.} allows for discontinuities in the PDF at cell edges, it is unable to capture this discontinuity without producing oscillations.
 \end{example}
 
-In summary, we applied the DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes to travelling wave problems with various initial conditions.
+\subsection*{Summary}
+We applied the DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes to travelling wave problems with various initial conditions.
 All the problems induced oscillations in the DG scheme at some point and as a result the DG-lim scheme reduced the order of the approximation to linear.
-In some cases, the oscillations in the DG scheme appear to be transient and were not material in the final approximation.
-In other cases, the oscillations were present in the final DG approximation causing the approximation to take unphysical values.
-The DG-lin-lim, uniformisation and QBD-RAP schemes are three positivity preserving alternatives.
-In general, the uniformisation scheme performs worse than the DG-lin-lim and QBD-RAP scheme, and appears to converge slowly to the true solution.
+In some cases, the oscillations in the DG scheme appeared to be transient and were not material in the final approximation.
+In other cases, the oscillations were present in the final DG approximation causing the approximation to violate the axioms of probability.
+Of the three positivity preserving alternatives, the DG-lin-lim, uniformisation and QBD-RAP schemes, in general, the uniformisation scheme performs worse than the other two and converges slowly to the true solution.
 The DG-lin-lim and QBD-RAP schemes perform similarly with the DG-lin-lim scheme performing slightly better for the smoother problems and the QBD-RAP scheme performing slightly better for the problems with more significant discontinuities and point masses.
 
+This section is the first step in analysing the ability of the approximation schemes to capture the dynamics fluid queues.
+This section demonstrates that, although the QBD-RAP scheme may have certain deficiencies for solution reconstruction (as demonstrated in Section~\ref{sec: recon num}), it can resolve the movement of probability between cells better than the uniformisation scheme, and similarly compared DG-lin-lim scheme.
+In the following sections we consider more complicated fluid queue dynamics by introducing some simple stochastic phase changes.
+
 \FloatBarrier
-\section{Stationary distributions} \label{sec:stat}
-We briefly turn our attention to stationary distributions of fluid queues.
-Since the stationary distribution is smooth we expect that the DG scheme will work well here.
-The stationary distribution is convenient for numerical experiments as we can evaluate it analytically \citep{s2017}, and it does not require us to approximate initial conditions or integrate over time.
-Hence, analysing the performance of the approximation schemes with the stationary distribution allows us to instrument the ability of the schemes to capture the stochastic dynamics, without numerical integration in time.
+\section{Limiting distributions} \label{sec:stat}
+We briefly turn our attention to limiting distributions of fluid queues.
+Since the limiting distribution is smooth we expect that the DG scheme will work well here.
+Moreover, since the solution is smooth, the slope limiter has no effect which means that the DG-lim and DG schemes are equivalent.
+Analysing the performance of the approximation schemes with the limiting distribution allows us to explore the ability of the schemes to capture the stochastic dynamics, without numerical integration with respect to time.
+In addition, the limiting distribution is convenient for numerical experiments as we can evaluate it analytically \citep{s2017}, and it does not require us to approximate initial conditions.
 
@@
 \begin{model}\label{model: simple}
-	Consider a fluid queue where the driving process is a CTMC with state space \(\calS=\{1,2,3,4\}\), generator 
+	Consider a fluid queue where the driving process is a CTMC with state space \(\calS=\{1,2\}\), generator 
 	\[\bs T = \left[\begin{array}{cccc}
@@
 
-	Upon hitting the lower boundary, the process transitions from Phase \(2\) to Phase \(j\) with probability \(p_{2j}\) where 
+	Upon hitting the lower boundary, the process immediately transitions from Phase~\(2\) to Phase~\(j\) with probability \(p_{2j}\) where 
 		\[\vligne{p_{21} & p_{22} } = \vligne{1 & 0}.\]
-		Upon hitting the upper boundary, the process transitions from Phase \(1\) to Phase \(j\) with probability \(p_{1j}\) where 
+	Upon hitting the upper boundary, the process immediately transitions from Phase~\(1\) to Phase~\(j\) with probability \(p_{1j}\) where 
 		\[\vligne{p_{11} & p_{12} } = \vligne{0 & 1}.\]
@@
 The model was discretised using the DG, uniformisation and QBD-RAP schemes using ten cells of width \(1\).
-We compute the coefficients for the stationary distribution in the following way.
+We compute the coefficients for the limiting distribution in the following way.
 Suppose that a given discretisation scheme results in an approximation to the generator of the fluid queue as a matrix \(\bs B\).
-Then the stationary coefficients are found by solving
+Then the limiting coefficients are found by solving
 \begin{align}
@@
 % \paragraph{Model~\ref{model: simple}~a) Absorbing model}
-% Figure~\ref{fig: absorbing stationary} plots KS errors between the true stationary CDF and the approximations (left) and the \(L^2\) error between the true stationary PDF and the approximations.
+% Figure~\ref{fig: absorbing limiting} plots KS errors between the true limiting CDF and the approximations (left) and the \(L^2\) error between the true limiting PDF and the approximations.
 Clearly the DG scheme is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
@@
 % 	\centering
-% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/ks_error_formatted.pdf}%
-% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/stationary_distribution/l2_pdf_error_formatted.pdf}
-% 	\caption{KS error between the true stationary CDF  and the approximations (left) and \(L^2\) error between the true stationary PDF and the approximations for Model~\ref{model: simple}~a) using the DG scheme (blue solid line), uniformisation scheme (orange dashed line) and QBD-RAP scheme (green dash-dotted line).}
-% 	\label{fig: absorbing stationary} 
+% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/limiting_distribution/ks_error_formatted.pdf}%
+% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/absorbing_model/limiting_distribution/l2_pdf_error_formatted.pdf}
+% 	\caption{KS error between the true limiting CDF  and the approximations (left) and \(L^2\) error between the true limiting PDF and the approximations for Model~\ref{model: simple}~a) using the DG scheme (blue solid line), uniformisation scheme (orange dashed line) and QBD-RAP scheme (green dash-dotted line).}
+% 	\label{fig: absorbing limiting} 
 % \end{figure}
@@
 % \paragraph{Model~\ref{model: simple}.}
-Figure~\ref{fig: reflecting stationary} plots KS errors between the true stationary CDF and the approximations and the \(L^2\) error between the true stationary PDF and the approximations.
-Clearly the DG scheme is superior here as its error rapidly decreases to a point where is become insignificant compared to other numerical errors.
-The QBD-RAP and uniformisation schemes both appear to be converging, with the errors for the QBD-RAP scheme decreasing faster.
 \begin{figure}[h]
@@
 \end{figure}
+Figure~\ref{fig: reflecting stationary} plots KS errors between the true limiting CDF and the approximations and the \(L^2\) error between the true limiting PDF and the approximations.
+Clearly the DG scheme is superior here as its error rapidly decreases to a point where it becomes insignificant compared to other numerical errors.
+The QBD-RAP and uniformisation schemes both appear to be converging, with the errors for the QBD-RAP scheme decreasing faster.
 
@@
 Once again we consider Model~\ref{model: simple} and use the same spatial discretisation as described in Section~\ref{sec:stat} (ten cells with width \(\Delta=1\)).
-Two initial conditions are considered, a point mass at 0 in Phase~1, and the initial distribution with PDF
+Two initial conditions are considered, a point mass with mass 1 at 0 in Phase~1, and the initial distribution with PDF
 \begin{align}
@@
 \end{align}
-in phases 1 and 2, with no mass at the boundaries.
-We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We apply the DG without a limiter and also the DG-lim and DG-lin-lim schemes which utilise a Generalised MUSCL slope limiter.
-
-To obtain a \emph{ground truth} 5,000,000 realisations of the fluid queue were simulated until \(t=2\), then the empirical CDF and the masses within each cell were computed from the simulations.
-We do not attempt to numerically approximate the true PDF via simulation.
+in Phases 1 and 2, with no mass at the boundaries.
+We numerically integrate over time until time \(t=2.0\) using the SSPRK4 method with t-step size 0.005.\footnote{The \(t\)-step size must be chosen to ensure that numerical integration over time is stable up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} Here we use the DG scheme without a limiter and the DG-lin-lim scheme.
+Error plots for the DG-lim scheme are included for completeness, but we do not comment on them in detail in this section due to their poor performance for discontinuous problems which we noted previously.
+
+To obtain a \emph{ground truth} \(5\times 10^{10}\) realisations of the fluid queue were simulated until \(t=2\), then the empirical CDF and the masses within each cell were computed from the simulations.
+We denote the empirical probability distribution by \(\mathbb P_{sim}\).
 We then compute the KS and \(L^1\) error metrics between the approximated and simulated CDFs, as well as the cell-wise error metric, given as follows.
@@
 \end{align}
-where \(p(2,\ell,j)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0),\varphi(0))\), \(p(2,11,4)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))\), and \(p(2,0,3)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=3\mid X(0),\varphi(0))\), and \(\mathbb P_{sim}\) denotes an empirical probability evaluated using the simulations of the fluid queue.
+where \(p(2,\ell,j)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\calD_{\ell,j}, \varphi(2)=j\mid X(0),\varphi(0))\), \(p(2,11,4)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{10\}, \varphi(2)=4\mid X(0),\varphi(0))\), and \(p(2,0,3)\) is an approximation to \(\mathbb P_{sim}(X(2)\in\{0\}, \varphi(2)=3\mid X(0),\varphi(0))\).
 
 To account for possible Monte-Carlo error, we used a bootstrap with 1,000 bootstrap samples.
-For the bootstrap we sample 5,000,000 realisations of the fluid queue with replacement from the original 5,000,000 samples, then compute error metrics with the resampled data.
+For the bootstrap we sample \(5\times 10^{10}\) realisations of the fluid queue with replacement from the original \(5\times 10^{10}\) samples, then compute error metrics with the resampled data.
 We resample 1,000 times, resulting in 1,000 estimates of the errors.
@@
 \paragraph{Model~\ref{model: simple} with exponential initial condition}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
+	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured strip surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+	\label{fig: reflecting transient exp} 
+\end{figure}
 Figure~\ref{fig: reflecting transient exp} shows the KS and \(L^1\) metric on the CDF for the five different schemes (DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes).
-For both error metrics the DG scheme converges rapidly until computational errors become significant.
+For both error metrics the DG scheme converges rapidly until computational/simulation errors become significant.
 The uniformisation, DG-lin-lim and QBD-RAP schemes converge at a slower rate than the DG scheme, with the QBD-RAP and DG-lin-lim schemes converging faster than the uniformisation scheme.
@@
 Intuitively, in the early evolution of the model, there will be a sharp peak in the transient density at \(x=t\).
-I suspect that the true transient density is continuous, but not differentiable, at the point \(x=t\) for \(t<10\).
+The true transient density is continuous, but not differentiable, at the point \(x=t\) for \(t<10\).
 %
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/exp/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the true transient CDF at time \(t=2\) for Model~\ref{model: simple} with the exponential initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: reflecting transient exp} 
-\end{figure}
-
-Figure~\ref{fig: reflecting transient exp cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition.
-Since the cell-wise errors do not require us to reconstruct the value of the solution within each cell, then this metric allows us to observe the error characteristics of the schemes without reconstruction.
-Figure~\ref{fig: reflecting transient exp cp} shows similar convergence characteristics to the KS error metric in Figure~\ref{fig: reflecting transient exp} (left).
+
 \begin{figure}[h]
@@
 \end{figure}
-\exampleFloatBarrier
+%\exampleFloatBarrier
+Figure~\ref{fig: reflecting transient exp cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the exponential initial condition.
+Since the cell-wise errors do not require us to reconstruct the value of the solution within each cell, this metric allows us to observe the error characteristics of the schemes without reconstruction.
+Figure~\ref{fig: reflecting transient exp cp} shows similar convergence characteristics to the KS error metric in Figure~\ref{fig: reflecting transient exp} (left).
+
 \paragraph{Model~\ref{model: simple} with a point-mass initial condition}
-Figure~\ref{fig: reflecting transient pm} shows the KS and \(L^1\) metric on the CDF  for the five different schemes (DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes).
-Comparing the error metrics in Figure~\ref{fig: reflecting transient pm} for the point mass initial condition with the ones in Figure~\ref{fig: reflecting transient exp} for the exponential initial condition all schemes perform worse for the point mass initial condition.
-Regarding comparative convergence of error for this problem, for both error metrics the DG scheme converges fastest, followed by the QBD-RAP and DG-lin-lim schemes which converge comparatively, then the uniformisation scheme, while the DG-lim does not appear to converge due to the limiter reducing the scheme to a linear approximation around the discontinuity.
+\begin{figure}[h]
+	\centering
+	% \includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/ks_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/l1_cdf_error_formatted.pdf}%
+	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/L1_cell_probs_error_formatted.pdf}
+	% \caption{KS (left) and \(L^1\) (right) errors between the CDF at time \(t=2\) for Model~\ref{model: simple} with the point-mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+	\caption{\(L^1\) errors between the CDF, and cell-wise errors, at time \(t=2\) for Model~\ref{model: simple} with the point-mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+	\label{fig: reflecting transient pm} 
+\end{figure}
+Figure~\ref{fig: reflecting transient pm} (left) shows \(L^1\) error on the CDF for the five different schemes (DG, DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes).
+Comparing the \(L^1\) error metrics in Figure~\ref{fig: reflecting transient pm} for the point mass initial condition with the ones in Figure~\ref{fig: reflecting transient exp} for the exponential initial condition all schemes perform worse for the point mass initial condition.
+Regarding comparative convergence of error for this problem, for both error metrics the DG scheme converges fastest, followed by the QBD-RAP and DG-lin-lim schemes which converge comparatively, then the uniformisation scheme.
+For the DG scheme, the rate at which the error decreases as the dimension of the scheme increases slows significantly after dimension 3.
+This is likely due to the DG scheme approximating smooth regions of the solution well when the dimension is 3 or greater, after which the majority of the error is due to the discontinuity in the solution, for which the rate of convergence is slower.
 %
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/ks_error_formatted.pdf}%
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the CDF at time \(t=2\) for Model~\ref{model: simple} with the point-mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
-Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
-The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: reflecting transient pm} 
-\end{figure}
-
-Figure~\ref{fig: reflecting transient pm cp} plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the point mass initial condition.
-Figure~\ref{fig: reflecting transient pm cp} suggests that approximating the cell-wise error seems to be a difficult problem.
-This is likely caused by the discontinuity at \(x=2\) in Phase \(1\), which lies exactly on a cell boundary.
-To investigate how the position of the discontinuity might affect the error metrics we evolved the model to time \(t=2.1\) and computed the KS error, \(L^1\) error between the CDFs and the cell-wise error.
-Once again, we use simulation and bootstrapping to approximate the true distribution.
-The plots (not shown) of KS error and \(L^1\) error between the CDFs for the model at times \(t=2\) and \(t=2.1\) are relatively similar and not noteworthy.
-The plot of the cell-wise error metric at time \(t=2.1\) in Figure~\ref{fig: reflecting transient pm t2.1} is somewhat more interesting.
-Firstly, Figure~\ref{fig: reflecting transient pm t2.1} shows that all schemes improved from time \(t=2\) to time \(t=2.1\) with respect to this metric.
-Figure~\ref{fig: reflecting transient pm t2.1} also shows that the cell-wise error metric is somewhat volatile for the DG scheme; I suspect that this is due to the oscillations in the DG approximation.
-In contrast, in Figure~\ref{fig: reflecting transient pm t2.1} the uniformisation and QBD-RAP schemes have monotonically decreasing error curves.
-Also interesting is the DG-lin-lim scheme, which does not improve past dimension 7.
-I suspect that the discontinuity at \(x=t\) is not easily resolved by the scheme given the presence of the limiter.
-Figure~\ref{fig: reflecting transient pm t2.1 cdfs} plots the CDFs in Phase~1 around the discontinuity in the transient distribution at time \(t=2.1\) for the dimension 21 QBD-RAP and dimension 22 DG-lin-lim schemes.
-Neither scheme does a particularly good job of capturing the discontinuity, but the QBD-RAP scheme appears to perform slightly better.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/L1_cell_probs_error_formatted.pdf}
-	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2\) with the point mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/reflecting_model/transient_distribution/point_mass/L1_cell_probs_error_formatted.pdf}
+% 	\caption{Cell-wise errors for Model~\ref{model: simple} at time \(t=2\) with the point mass initial condition, where the approximations were obtained via the DG (blue solid line), DG-lim (green dashed line), uniformisation (orange dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
 Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.}
-	\label{fig: reflecting transient pm cp} 
-\end{figure}%
+% 	\label{fig: reflecting transient pm cp} 
+% \end{figure}%
 \begin{figure}[h]
@@
 	\caption{Dimension 21 QBD-RAP and dimension 22 DG-lin-lim approximations to the CDF in Phase~1 for Model~\ref{model: simple} with a point-mass initial condition.
-A CDF estimated via simulation is plotted in grey.}
+An empirical estimate of the CDF obtained via simulation is plotted in grey.}
 	\label{fig: reflecting transient pm t2.1 cdfs} 
 \end{figure}
-
-In summary, for the exponential initial condition the DG scheme performed well, producing the lowest errors.
+Figure~\ref{fig: reflecting transient pm} (right) plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the point mass initial condition.
+Figure~\ref{fig: reflecting transient pm} (right) suggests that approximating the cell-wise error seems to be a difficult problem.
+This is likely caused by the discontinuity at \(x=2\) in Phase \(1\), which lies exactly on a cell boundary.
+To investigate how the position of the discontinuity might affect the error metrics we evolved the model to time \(t=2.1\) and computed the \(L^1\) error between the CDFs and the cell-wise error.
+Once again, we use simulation and bootstrapping to approximate the true distribution.
+The plot (not shown) of the \(L^1\) error between the CDFs for the model at times \(t=2\) and \(t=2.1\) is relatively similar to the one show and not noteworthy.
+The plot of the cell-wise error metric at time \(t=2.1\) in Figure~\ref{fig: reflecting transient pm t2.1} is somewhat more interesting.
+Firstly, Figure~\ref{fig: reflecting transient pm t2.1} shows that all schemes improved from time \(t=2\) to time \(t=2.1\) with respect to this metric.
+Figure~\ref{fig: reflecting transient pm t2.1} also shows that the cell-wise error metric is somewhat volatile for the DG scheme; this is due to the oscillations in the DG approximation.
+In contrast, in Figure~\ref{fig: reflecting transient pm t2.1} the uniformisation and QBD-RAP schemes have monotonically decreasing error curves.
+Also interesting is the DG-lin-lim scheme, which does not improve past dimension 7.
+It seems that the discontinuity at \(x=t\) is not easily resolved by the scheme given the presence of the limiter.
+Figure~\ref{fig: reflecting transient pm t2.1 cdfs} plots the CDFs in Phase~1 around the discontinuity in the transient distribution at time \(t=2.1\) for the dimension 21 QBD-RAP and dimension 22 DG-lin-lim schemes.
+Neither scheme does a particularly good job of capturing the discontinuity, but the QBD-RAP scheme appears to perform slightly better.
+
+\subsection*{Summary}
+For the exponential initial condition the DG scheme performed well, producing the lowest errors.
 This was followed by the DG-lin-lim scheme, then the QBD-RAP scheme, then the uniformisation scheme, while the DG-lim scheme did not perform well at all.
@@
 For the point mass initial condition the DG scheme produced oscillatory solutions.
-Of the positivity preserving methods, the DG-lim scheme performed poorly and did not appear to be converging, the uniformisation scheme appears to be converging but at a slower rate than the DG-lin-lim and QBD-RAP schemes.
-Of the two most performant positivity preserving schemes, the QBD-RAP scheme performed slightly better for the point mass initial condition, while the DG-lin-lim scheme performed slightly better for the exponential initial condition.
+Of the viable positivity preserving methods the uniformisation scheme appears to be converging but at a slower rate than the DG-lin-lim and QBD-RAP schemes.
+Of the two best performing positivity preserving schemes, the QBD-RAP scheme performed slightly better for the point mass initial condition, while the DG-lin-lim scheme performed slightly better for the exponential initial condition.
+
+This section investigated the performance of the approximation schemes for a simple fluid queue.
+In the next section we analyse the approximation schemes in more detail by investigating their ability to approximate hitting times of fluid queues.
 
 \FloatBarrier
-\section{Hitting times}\label{sec: return approx}
+\section{Hitting times}\label{sec: hit approx}
+We can gain further insight into the approximation schemes by investigating how they perform approximating hitting times of fluid queues.
+This is also an important aspect to consider if we wish to apply the schemes to approximate operators from the analysis fluid-fluid queues.
+Recall, in the analysis of fluid-fluid queues in Section~\ref{sec: ffq intro}, we partition sample paths of the second fluid into periods where it is either increasing, decreasing, or constant.
+The position of \((\varphi(t),X(t))\) determines the rate at which the second fluid moves, hence the hitting times on the boundaries of the sets \(\mathcal F_i^+, \mathcal F_i^-,\) and \(\mathcal F_i^0\) determine the periods of time when the second fluid is either increasing, decreasing, or constant.
+
 Once again we consider Model~\ref{model: simple}.
@@
 for some initial distribution \(\mu\). We look at two initial conditions; an exponential with equal mass in each phase, 
-\[\mathbb P(X(0)\in\wrt x,\varphi(0)=i) = \cfrac{1}{2}\cfrac{e^{-x}}{(1-e^{-1}},\,i\in\{1,2\},\]
+\[\mathbb P(X(0)\in\wrt x,\varphi(0)=i) = \cfrac{1}{2}\cfrac{e^{-x}}{(1-e^{-1})},\,i\in\{1,2\},\]
 and a point mass at \(X(0)=0\) in phase \(\varphi(0)=1\).
 
-We use the DG, uniformisation and QBD-RAP schemes to discretise the fluid queue and partition \([0,1]\) into three intervals of width \(1/3\).
+We partition the interval \([0,1]\) into three intervals of width \(1/3\).
 To capture the mass which has hit the set \(\{0,1\}\), we suppose that when the process hits \(\{0\}\) or \(\{1\}\) it is absorbed forever and remains in the phase in which the process first hit \(\{0\}\) or \(\{1\}\).
-
 We integrate the schemes until time \(t=10\) using the SSPRK4 method with \(t\)-step size 0.005/3.\footnote{Since we use a smaller cell-width in this example than in previous examples we need to reduce the \(t\)-step size accordingly to ensure that numerical integration is stable for schemes up to dimension 21, adhering to a CFL-like condition \cite[Section~4.8]{nodalDGBook}.} We use the DG scheme with and without a slope limiter during the time-integration and implement both the DG-lim and DG-lin-lim schemes.
-At each time-step of the numerical integration, we record the amount of mass at the absorbing boundaries in each phase, which gives us an approximation of the cumulative distribution function of the hitting time in each phase up to time \(t=10\).
-
-As a \emph{ground truth} we simulated 5,000,000 realisations and recorded the hitting time on the set \(\{0,1\}\) and the phase at the time of hitting.
+At each time-step of the numerical integration, we record the amount of mass at the absorbing boundaries in each phase, which gives us an approximation of the cumulative distribution function of the hitting time in each phase, up to time \(t=10\).
+
+As a \emph{ground truth} we simulate \(5\times 10^{10}\) realisations and record the hitting time on the set \(\{0,1\}\) and the phase at the time of hitting.
 We then compute the empirical CDF of the hitting probabilities
@@
 
-To account for Monte-Carlo errors we took 1,000 bootstrap resamples of the original 5,000,000 samples and computed the empirical CDF of the hitting probabilities for each bootstrap sample.
-For each bootstrap sample, we resampled 5,000,000 points with replacement from the original 5,000,000 realisations, then computed the error metrics between the empirical and approximated CDFs and, ultimately, estimated the 5th and 95th percentile of the distribution of the errors.
+To account for Monte-Carlo errors we take 1,000 bootstrap resamples of the original \(5\times 10^{10}\) samples and compute the empirical CDF of the hitting times for each bootstrap sample.
+For each bootstrap sample, we resample \(5\times 10^{10}\) points with replacement from the original \(5\times 10^{10}\) realisations, then compute the error metrics between the empirical and approximated CDFs and, ultimately, estimate the 5th and 95th percentile of the distribution of the errors.
 
 \paragraph{Exponential initial condition}
-Figure~\ref{fig: hitting time exp} shows the error metrics recorded for the five numerical approximation schemes applied to the hitting time problem with the exponential initial condition.
-For both error metrics the DG scheme performs the best.
-For the \(L^1\) error metric the DG scheme converges up to dimension 3, after which there is little improvement in the error; I believe that this is due to other numerical errors dominating.
-The DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation.
-Intuitively, with this initial condition we suspect that the transient distribution of the fluid queue on the event that it remains in the interval \((0,1)\) is discontinuous at \(x=t\) in Phase~1 and \(x=1-t\) in Phase~2 for \(x,t<1\).
-We also expect there is a discontinuity in the hitting time PDF at time \(t=1\).
-Regarding the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes, all appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly and converging at a faster rate than the uniformisation scheme.
-
-In Figure~\ref{fig: hitting time oscillation exp} we plot the PDFs of the hitting times \(\zeta_X(\{0,1\})\), in each phase, for the dimension 21 DG, dimension 21, QBD-RAP and dimension 22 DG-lin-lim schemes.
-Figure~\ref{fig: hitting time oscillation exp} demonstrates oscillations in the DG approximation, though the PDF is not negative in this case.
-All schemes seem to capture the discontinuity at \(t=1\) relatively well.
-There is an interesting artefact in the QBD-RAP approximation around \(t=0\) where the scheme has generated an oscillation.
-I suspect that this artefact should disappear as the dimension of the scheme is increased, unlike the DG scheme where oscillations can persist when the dimension is increased.
-Let's investigate this artefact further.
-
-Intuitively, most sample paths which exit near \(t=0\) start near the boundary then see a change of phase shortly after \(t=0\) and remain in that phase until hitting \(\{0,1\}\).
-Consider such a sample path which starts at \(x_0=0\) in Phase~1, then exits at time \(v\) with a single change of phase at time \(v-u>0\).
-The QBD-RAP approximation to this sample paths has density\footnote{One advantage of the QBD-RAP and uniformisation approaches is that, from their stochastic interpretation, we can use sample-path arguments to analyse the scheme.}
-\begin{align}
-	\bs \alpha e^{(\bs S - \bs I)(v-u)} \bs D e^{(\bs S - 1.1\bs I)u}\bs s = \bs \alpha e^{\bs S(v-u)} \bs D e^{\bs Su}\bs s e^{-v-0.1u},\label{eqn :lf}
-\end{align}
-and since the sample path will hit \(\{0\}\) at \(v=2u\), intuitively, we would expect (\ref{eqn :lf}) to approximate a point mass at \(v=2u\).
-To see that this is the case, recall that we can approximate a point mass at \((v-u)\) by
-\begin{equation}\bs k(\Delta - (v-u))\bs e^{\bs Sx}\bs s,\label{eqn:aafba}\end{equation}
-where \(\bs k(z)=\cfrac{\bs \alpha e^{\bs Sz}}{\bs \alpha e^{\bs Sz\bs e}}\),
-and recall that the matrix \(\bs D\) is an approximation; 
-\begin{align}
-	\bs \alpha e^{\bs S(v-u)}\bs D &= \mathbb E\left[\bs k(Z-(v-u))1(Z>v-u)\right]\nonumber 
-	% \\&\approx \bs \alpha e^{\bs S(v-u)}\bs e \bs k(\Delta-(v-u))
-	\\&\approx\mathbb P(Z>v-u) \bs k(\Delta-(v-u)), \label{eqn: aegpb}
-\end{align} where \(Z\sim ME(\bs \alpha,\bs S)\)
-and \(Z\) is concentrated around \(\Delta\).
-
-Thus, (\ref{eqn :lf}) is approximately  
-\[e^{-v-0.1u} \mathbb P(Z>v-u) \bs k(\Delta-(v-u))\bs e^{\bs Su}\bs s,\]
-which approximates a point mass at \(u=v-u\), or \(2u=v\).
-Retracing our arguments, sources of error in this approximation come from the approximation in (\ref{eqn: aegpb}) and from the approximation of a point mass by (\ref{eqn:aafba}), which we commented on in Section~\ref{sec: closing vecs numerics} where we conducted numerical experiments about the unnormalised closing operator given by \(\bs v(x) = e^{\bs S x}\bs s\).
-Indeed, we saw that the unnormalised closing operator performed poorly for the exponential initial condition in Section~\ref{sec: closing vecs numerics}.
-
-% To further comment on the approximation of a point mass by (\ref{eqn:aafba}), recall the . We found that this closing operator has difficulty approximating mass at .
-
 \begin{figure}[h]
@@
 \end{figure}
+Figure~\ref{fig: hitting time exp} shows the error metrics recorded for the five numerical approximation schemes applied to the hitting time problem with the exponential initial condition.
+For both error metrics the DG scheme performs the best.
+
+The DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation.
+With this initial condition the transient distribution of the fluid queue on the event that it remains in the interval \((0,1)\) is discontinuous at \(x=t\) in Phase~1 and \(x=1-t\) in Phase~2 for \(x,t<1\).
+There is also a discontinuity in the hitting time PDF at time \(t=1\).
+
+Regarding the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes, all appear to converge, with the QBD-RAP and DG-lin-lim schemes performing similarly and better than the uniformisation scheme.
+
+In Figure~\ref{fig: hitting time oscillation exp} we plot the PDFs of the hitting times \(\zeta_X(\{0,1\})\), in each phase, for the dimension 21 DG, dimension 21, QBD-RAP and dimension 22 DG-lin-lim schemes.
+Figure~\ref{fig: hitting time oscillation exp} demonstrates oscillations in the DG approximation, though the PDF is not negative in this case.
+All schemes seem to capture the discontinuity at \(t=1\) relatively well.
+There is an interesting artefact in the QBD-RAP approximation around \(t=0\) where the scheme has generated an oscillation.
+Let us investigate this artefact further.
+
+Intuitively, most sample paths which exit near \(t=0\) start near a boundary, they see a change of phase shortly after \(t=0\), then remain in that phase until hitting \(\{0,1\}\).
+Consider such a sample path which starts at \(x_0=0\) in Phase~1 and exits at time \(v\) with a single change of phase at time \(u\in(0,v)\).
+For the fluid queue, \(u=v/2\), and we want the QBD-RAP scheme to approximate this.
+The QBD-RAP approximation to this sample paths has density\footnote{One significant advantage of the QBD-RAP and uniformisation approaches is that, from their stochastic interpretation, we can use sample-path arguments to analyse the scheme.}
+\begin{align}
+	\bs \alpha e^{(\bs S - \bs I)u} \bs D e^{(\bs S - 1.1\bs I)(v-u)}\bs s = \bs \alpha e^{\bs Su} \bs D e^{\bs S(v-u)}\bs s e^{-1.1v+0.1u}.\label{eqn :lf}
+\end{align}
+We want (\ref{eqn :lf}) to approximate a point mass at \(v=2u\). To see that this is the case, recall that we can approximate a point mass at \(u\) by 
+\begin{equation}\bs k(\Delta - u)\bs e^{\bs Sx}\bs s,\label{eqn:aafba}\end{equation}
+where \(\bs k(z)=\cfrac{\bs \alpha e^{\bs Sz}}{\bs \alpha e^{\bs Sz\bs e}}\),
+and recall that the matrix \(\bs D\) is an approximation; 
+\begin{align}
+	\bs \alpha e^{\bs Su}\bs D &= \mathbb E\left[\bs k(Z-u)1(Z>u)\right]\nonumber 
+	% \\&\approx \bs \alpha e^{\bs S(v-u)}\bs e \bs k(\Delta-(v-u))
+	\\&\approx\mathbb P(Z>u) \bs k(\Delta-u), \label{eqn: aegpb}
+\end{align} where \(Z\sim ME(\bs \alpha,\bs S)\)
+and \(Z\) is concentrated around \(\Delta\).
+
+Thus, (\ref{eqn :lf}) is approximately  
+\[e^{-1.1v+0.1u} \mathbb P(Z>u) \bs k(\Delta-u)\bs e^{\bs S(v-u)}\bs s,\]
+which approximates a point mass at \(v-u=u\), or \(2u=v\).
+Retracing our arguments, sources of error in this approximation come from the approximation in (\ref{eqn: aegpb}) and from the approximation of a point mass by (\ref{eqn:aafba}), which we commented on in Section~\ref{sec: closing vecs numerics} where we conducted numerical experiments about the unnormalised closing operator given by \(\bs v(x) = e^{\bs S x}\bs s\).
+Indeed, we saw in Example~\ref{ex: kjfk} that the unnormalised closing operator performed poorly for the exponential initial condition.
+The phenomenon discussed above is also related to Example~\ref{ex: al}, where the reconstructions for the QBD-RAP scheme failed to capture mass near the right-hand edge of the interval.
+
 \begin{figure}[h]
@@
 The DG scheme displays oscillations around the discontinuities at \(t=1\).
-The QBD-RAP scheme has an oscillation near \(t=0\).
+The QBD-RAP scheme has oscillations near \(t=0\) and \(t=0.4\).
 }
@@
 \end{figure}
-\exampleFloatBarrier
+%\exampleFloatBarrier
+In Figure~\ref{fig: hitting time oscillation exp} there is another small oscillation present in the QBD-RAP approximation around \(t=0.35\) in both phases.
+The cause of this oscillation is likely related to the discussion above.
+Consider a sample path which starts at \(x=0.34\), just to the right of the cell-edge at \(1/3\), in Phase~1, changes phase once at time \(u\), then remains in that phase until the boundary at \(x=0\) is hit.
+For the fluid queue, the hitting time of this sample path is \(0.34 + 2u\).
+The QBD-RAP approximation to this sample paths has density
+\begin{align}
+	\int_{v=u}^\infty \bs \alpha e^{(\bs S - \bs I)u} \bs D e^{(\bs S - 1.1\bs I)(v-u)}\bs s \bs \alpha e^{(\bs S-1.1\bs I)w}\bs s \wrt v.\label{eqn :lf2}
+\end{align}
+The first term in (\ref{eqn :lf2}) is the same as the first term in (\ref{eqn :lf}), and this is likely the source of the oscillation.
+
+% To further comment on the approximation of a point mass by (\ref{eqn:aafba}), recall the . We found that this closing operator has difficulty approximating mass at .
+
 \paragraph{Point mass initial condition}
-Figure~\ref{fig: hitting time pm} shows the KS error (left) and \(L^1\) error for the CDF (right) for the five numerical approximation schemes applied to the hitting time problem with the point mass initial condition.
-The DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation as we might expect given the discontinuity.
-Of the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes all appear to converge, with the QBD-RAP converging fastest and the DG-lin-lim and uniformisation schemes performing similarly.
-The DG scheme appears to perform best, however, oscillations are present in the DG approximation.
-In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis, and for the DG-lin-lim scheme with dimension 22.
-Observing Figure~\ref{fig: hitting time oscillation} the DG approximation displays oscillations around and the discontinuity at \(t=1\) while the other two schemes do not.
-Of the two positivity preserving schemes in Figure~\ref{fig: hitting time oscillation}, the QBD-RAP scheme appears to approximate the discontinuity best.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/ks_error_formatted.pdf}%
+\begin{figure}[h]
+	\centering
+	% \includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/ks_error_formatted.pdf}%
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
+	% \caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
+The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
+Bootstrapped 90\% confidence intervals are shown by the lighter coloured bars surrounding the lines.
+The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
+	\caption{\(L^1\) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
 The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
@@
 The empirical CDF obtained via simulation is plotted as the teal dashed line.
-The DG scheme displays oscillations.
+The DG scheme displays negative approximations to probabilities.
 }
@@
 \end{figure}
-
-In summary, the results of this experiment are somewhat similar to the last experiment on transient distributions; the DG-lim scheme performs poorly, the DG scheme exhibits oscillations, of the three other positivity preserving schemes, the uniformisation scheme performs poorest, while the QBD-RAP and DG-lin-lim schemes perform similarly, with the QBD-RAP scheme performing slightly better for the point mass initial condition and the DG-lin-lim scheme performing slightly better for the exponential initial condition.
+Figure~\ref{fig: hitting time pm} shows the \(L^1\) error for the CDF for the five numerical approximation schemes applied to the hitting time problem with the point mass initial condition.
+Due to the discontinuity the KS error is not very insightful for this problem.
+
+Observing Figure~\ref{fig: hitting time pm}, the DG-lim scheme does not appear to converge, suggesting oscillations in the DG approximation as we might expect given the discontinuity.
+Of the other positivity preserving schemes, the uniformisation, DG-lin-lim and QBD-RAP schemes all appear to converge, with the QBD-RAP converging fastest and the DG-lin-lim and uniformisation schemes performing similarly.
+
+The error for the DG scheme reduces significantly between dimension 1 and dimension 3, then reduces at a slower rate after that.
+For dimension 3 and above the DG scheme approximates the smooth regions of the solution well and the majority of the error is from the region around the discontinuity, which reduces at a slower rate.
+The DG scheme appears to perform best, however, oscillations are present in the DG approximation.
+In Figure~\ref{fig: hitting time oscillation} we plot the CDFs of the hitting time for Phase~1 obtained via the DG and QBD-RAP schemes, each using a dimension 21 basis, and for the DG-lin-lim scheme with dimension 22.
+Observing Figure~\ref{fig: hitting time oscillation} the DG approximation displays oscillations around the discontinuity at \(t=1\) while the other two schemes do not.
+Of the two positivity preserving schemes in Figure~\ref{fig: hitting time oscillation}, the QBD-RAP scheme appears to approximate the discontinuity best.
+
+\subsection*{Summary}
+The results of this experiment are somewhat similar to the last experiment on transient distributions; the DG-lim scheme performs poorly, the DG scheme exhibits oscillations, of the three other positivity preserving schemes, the uniformisation scheme performs poorest, while the QBD-RAP and DG-lin-lim schemes perform similarly, with the QBD-RAP scheme performing slightly better for the point mass initial condition and the DG-lin-lim scheme performing slightly better for the exponential initial condition.
 
@@
 
-We obtained approximations to the generator of the fluid queue \(\{(X(t),\varphi(t))\}\) via the DG, QBD-RAP and uniformisation schemes, all of which used a cell width of \(\Delta=0.4\).
-The approximate generator matrix resulting from the approximation is then used to approximate the first-return operator \(\mathbb\Psi(s)\) as discussed in Section~\ref{sec: intro Psi}.
+We obtain approximations to the generator of the fluid queue \(\{(X(t),\varphi(t))\}\) via the DG, QBD-RAP and uniformisation schemes (as discussed at the start of this chapter, we do not apply slope limiting for this calculation).
+All the discretisations use a cell width of \(\Delta=0.4\).
+The approximate generator matrix is then used to approximate the first-return operator \(\mathbb\Psi(s)\) as discussed in Section~\ref{sec: intro Psi}.
+As in Section~\ref{sec: intro Psi}, let \(\bs{B}^{m,n}\), \(m,n\in\{+,-,0\}\) be an approximation to \(\mathbb B^{m,n},\, m,n\in\{+,-,0\}\) of dimension \(p\), and recall that \(\bs{D}^{n,m},\,m,n\in\{+,-\}\) is given by
+\[\bs D^{m,n} = \bs{R}^m (\bs{B}^{m,n} + \bs{B}^{m,0}(\bs B^{0,0})^{-1}\bs{B}^{0,n}), \, m,n\in\{+,-\}, \]
+where \(\bs R^m\) is a diagonal matrix with diagonal blocks \(1/r_i^k\bs{I}_p\), \(i\in\calS_k^m\), \(k\in\mathcal K^m\), where \(r_i^k\) is the (constant) value of \(r_i\) on cell \(k\).
+Then \(\bs{\Psi}\), an approximation to \(\mathbb \Psi\), is given by the solution to the matrix-Ricatti equation 
+\begin{align}\label{eqn:RiccatiPsi2}
+    \bs D^{+-}
++ \bs \Psi   \bs D^{-+}\bs \Psi
++   \bs D^{++}\bs \Psi
++ \bs \Psi  \bs D^{--}
+= \bs 0.
+\end{align}
 Due to the stochastic interpretation of the uniformisation and QBD-RAP schemes the approximations to \(\mathbb \Psi(s)\) have a stochastic interpretation as the first-return probabilities of a fluid queue driven by a CTMC and QBD-RAP, respectively (see \cite[Chapter~7]{p2019} and \cite{bgnp2021} for details on the latter).
-For the DG scheme the approximation of \(\mathbb \Psi(s)\) is not as well-understood.\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{Y}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}
+For the DG scheme the approximation of \(\mathbb \Psi(s)\) is not as well-understood.
+Here, we assume that it is given by the solution to (\ref{eqn:RiccatiPsi2}) and is a polynomial approximation.
+%\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{Y}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}
 
@@
 
-For a `ground truth' comparison, we simulated 5,000,000 realisations of the fluid-fluid queue and recorded the value of \(\{X(t)\}\) and \(\{\varphi(t)\}\) at the time of first return of the second fluid, \(\{Y(t)\}\).
-The empirical approximation of (\ref{eqn: first return Y 1}) was then constructed, and error metrics for the difference between the empirical CDF and the approximations was computed.
+For a `ground truth' comparison, we simulated \(5\times 10^{10}\) realisations of the fluid-fluid queue and recorded the value of \(\{X(t)\}\) and \(\{\varphi(t)\}\) at the time of first return of the second fluid, \(\{Y(t)\}\).
+The empirical approximation of (\ref{eqn: first return Y 1}) was then constructed, and error metrics for the difference between the empirical CDF and the approximations were computed.
 To account for Monte-Carlo errors we used a bootstrap with 1,000 bootstrap samples to construct 1,000 bootstrap samples of the error estimates and ultimately estimated the 5th and 95th percentiles of the error distribution.
-Each of the 1,000 bootstrap samples was constructed by resampling the original 5,000,000 realisations 5,000,000 times with replacement.
-
-In Figures~\ref{fig: ffq return cts} and~\ref{fig: ffq cell probs} we plot the error metrics for the approximations to the distribution~(\ref{eqn: first return Y 1}).
-The DG scheme performs best converging rapidly until the error in the approximation scheme is swamped by other numerical errors.
-The QBD-RAP scheme is second best and the uniformisation scheme appears to be the slowest to converge.
-Here, the first return distribution appears to be smooth, hence we might expect the DG scheme to perform well.
-
-Note that there is a significant number other sources of error here; machine precision errors, errors in solving the matrix Riccati equation to approximate \(\mathbb \Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), and truncation errors.
-Furthermore, for the QBD-RAP, since the parameters \(\bs \alpha,\,\bs S,\,\bs s\,\) and \(\bs D\) are found numerically, then there is another source of error from this.
+Each of the 1,000 bootstrap samples was constructed by resampling the original \(5\times 10^{10}\) realisations \(5\times 10^{10}\) times with replacement.
+
 \begin{figure}[h]
@@
 \end{figure}
-\exampleFloatBarrier
+%\exampleFloatBarrier
+In Figures~\ref{fig: ffq return cts} and~\ref{fig: ffq cell probs} we plot the error metrics for the approximations to the distribution~(\ref{eqn: first return Y 1}).
+The DG scheme performs best converging rapidly until the error in the approximation scheme is swamped by other numerical (simulation) errors\footnote{There are a significant number of other sources of error here; the largest contribution to error is simulation, but also machine precision errors, errors in solving the matrix Riccati equation to approximate \(\mathbb \Psi(s)\), errors from approximating error metrics (numerical integration/finding KS statistic), errors from approximating the initial condition and truncation errors.
+Furthermore, for the QBD-RAP, since the parameters \(\bs \alpha,\,\bs S,\,\bs s\,\) and \(\bs D\) are found numerically, then there is another source of error from this.}.
+The QBD-RAP scheme is second best and the uniformisation scheme appears to be the slowest to converge.
+Here, the first return distribution appears to be smooth, hence we might expect the DG scheme to perform well.
+The initial condition, however, is not smooth and the DG approximation violates the axioms of probability for the initial distribition.
+Interestingly, however, there is no sign of this in the first return distribution.
 
@@
 
-For Model~\ref{model: ffq2} there is a point mass at \(X(\zeta_{Y}(\{0\}))=1.2\) of magnitude \(e^{-(\beta_1+\beta_2)}\times 0.5\), which occurs when the phase remains in \(\varphi(t)=00\) for all \(t\in [0,\zeta_{Y}(\{0\})]\).
-
-Figures~\ref{fig: ffq return discts} and~\ref{fig: ffq2 cell probs} plot the error metrics for the first return distribution.
-Observing Figure~\ref{fig: ffq return discts} we see that the DG scheme performs well, followed by the QBD-RAP scheme, then the uniformisation scheme performs worst.
-However, the DG scheme results in an oscillatory solution with the CDF taking impossible values (decreasing at points) as shown in Figure~\ref{fig: ffq2 oscillation}
+Consider a sample path of Model~\ref{model: ffq2} which begins at \(X(0)=2\) in Phase~\(00\) and remains in Phase~\(00\) until a time \(t>0.5\).
+On this sample path, at time \(u\leq 0.5\) the position of the first fluid level is \(X(u)=2-1.6u\).
+Moreover, for \(u\in[0,0.25]\), \(X(u)\geq x^*\) and \(r_{00}(X(u))=\theta_2=1>0\), and for \(u\in(0.25,0.5]\), \(X(u)\in(0,x^*)\) and \(r_{00}(X(u))=-\theta_2=-1<0\).
+Hence, for \(u\leq 0.5\) the second fluid level is \(Y(u)=t1(t\leq 0.25)+(0.25-t)1(t>0.25)\) and at time \(t=0.5\), the second fluid level first returns to \(0\), i.e.~\(\zeta_{Y}(\{0\})=0.5\) on this sample path.
+The probability of this sample path is \(e^{-(\beta_1+\beta_2)\times 0.5}\).
+Thus, in the distribution of the first return time of the second fluid level there is a point mass at \(t=0.5\) with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\).
+Furthermore, on the same sample path \(X(\zeta_Y(\{0\}))=X(0.5)=1.2\), and hence there is a point mass with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\) in the distribution of \(X(\zeta_Y(\{0\}))\) also.
+
 \begin{figure}[h]
@@
 \end{figure} 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cell_probs_error_formatted.pdf}%
-	\caption{Cell-wise error between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing in each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq2}.
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cell_probs_error_formatted.pdf}%
+% 	\caption{Cell-wise error between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing in each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq2}.
 The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
@@
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\label{fig: ffq2 cell probs} 
-\end{figure}
+% 	\label{fig: ffq2 cell probs} 
+% \end{figure}
 \begin{figure}[h] 
@@
 The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dotted line is the empirical CDF obtained via simulation.
-The DG scheme displays oscillations around the discontinuity.
-}
+The DG scheme displays oscillations around the discontinuity, which mean that is does not represent a CDF.}
 	\label{fig: ffq2 oscillation} 
 \end{figure}
-
-In summary, for the smooth problem the DG scheme was highly effective, significantly outperforming the positivity preserving QBD-RAP and uniformisation methods.
-For the discontinuous problem, the DG scheme displayed oscillations, but produced the lowest errors.
-For both problems, of the positivity preserving methods (QBD-RAP and uniformisation scheme), the QBD-RAP scheme performed the best.
-
+Figure~\ref{fig: ffq return discts} plots error metrics for the first return distribution.
+Note that due to the discontinuity, the KS error is bounded below in how small it can be.
+The approximations all give a continuous approximation to the CDF, hence the smallest possible KS error is 1/2 the size of the discontinuity, which is \(\approx 0.184\) here.
+Observing Figure~\ref{fig: ffq return discts} we see that the DG scheme performs well, followed by the QBD-RAP scheme, then the uniformisation scheme performs worst.
+However, the DG scheme results in an oscillatory solution which does not represent a CDF, as shown in Figure~\ref{fig: ffq2 oscillation}.
+
+\subsection{Summary}
+Although the initial condition in the DG scheme violated the axioms of probability, for the smooth problem the DG scheme was highly effective, significantly outperforming the positivity preserving QBD-RAP and uniformisation methods.
+Interestingly the oscillations in the initial condition were not present in the first return distribution.
+
+For the discontinuous problem, the DG scheme displayed oscillations (in both the initial condition and first-return distribution) and did not produce a legitimate CDF, but it did produce the lowest errors.
+Of the positivity preserving methods (QBD-RAP and uniformisation scheme), the QBD-RAP scheme performed the best for both problems.
 \FloatBarrier
@@
 In this chapter we have numerically investigated some properties of the QBD-RAP approximation scheme and compared it to existing schemes; the uniformisation scheme of \cite{bo2013} and the discontinuous Galerkin scheme.
-In general, the numerical experiments that the smoother the problem is the better the performance of the DG scheme, and it emphatically outperforms the other two schemes for smooth problems.
-However, for problems with discontinuities the DG approximation can exhibit oscillations, while the QBD-RAP and uniformisation approximations avoid this and, of the latter two, the QBD-RAP scheme often converges faster.
-
-To avoid the problems of oscillations we can sometimes employ a \emph{slope limiter} with the DG scheme, which effectively reduces the scheme to linear in the regions where oscillations are detected.
-In some contexts, we implemented two slope-limited DG schemes, the DG-lim scheme which takes a high order DG scheme and limits the solution as necessary, and the DG-lin-lim, which is a linear approximation on a finer grid designed to use approximately the same computational resources as the other schemes we consider.
+In general, the numerical experiments show that the smoother the problem is, the better the performance of the DG scheme, and it emphatically outperforms the other two schemes for smooth problems.
+However, for problems with discontinuities the DG approximation can exhibit oscillations and result in illegitimate approximations to probability distributions which violate the axioms of probability.
+The QBD-RAP and uniformisation approximations are guaranteed to produce legitimate distributions and, of the two, the QBD-RAP scheme almost always produces lower errors.
+
+To avoid the problems of oscillations we can sometimes employ a \emph{slope limiter} with the DG scheme which reduces the scheme to linear in the regions where oscillations are detected.
+We implemented two slope-limited DG schemes, the DG-lim scheme which takes a high order DG scheme and limits the solution as necessary, and the DG-lin-lim, which is a linear approximation on a finer grid designed to use approximately the same computational resources as the other schemes considered.
 The numerical experiments demonstrate a significant loss of accuracy in the approximation when a DG-lim scheme is used for discontinuous problems.
+We conclude that the DG-lim scheme is not a viable scheme for the discontinuous problems considered here.
 The numerical experiments suggest that the DG-lin-lim scheme can perform well, and is similar to the performance of the QBD-RAP scheme, in the presence of discontinuities.
@@
 
-As a first step in the numerical experiments, in Section~\ref{sec: recon num}, we examined the ability of each scheme to approximate different initial conditions.
+As a first step, in Section~\ref{sec: recon num}, we examined the ability of each scheme to approximate various initial conditions.
 For the DG scheme, this is equivalent to a projection of the initial condition on to a set of basis polynomials, for the uniformisation scheme this is equivalent to projecting the initial condition on to a basis of constant functions, and for the QBD-RAP scheme the approximation of the initial condition is as described in Sections~\ref{sec: initial conditions} and~\ref{sec: closing}.
 Section~\ref{sec: comp} demonstrates that the DG scheme (projection) can result in oscillations and negative regions in the approximation when the initial condition is discontinuous.
-The uniformisation and QBD-RAP schemes avoid this problem, but appear to have higher errors and the QBD-RAP scheme appears to have the largest errors.
+The uniformisation and QBD-RAP schemes avoid this, but appear to have higher errors and the QBD-RAP scheme appears to have the largest errors.
 For discontinuous initial conditions the rates of convergence are comparable for all three schemes.
-However, when the initial condition to be approximated is sufficiently smooth, then the DG approximation is superior.
+When the initial condition to be approximated is sufficiently smooth, then the DG approximation is superior.
 The QBD-RAP scheme performed the poorest for the examples considered.
-
-Next, in Section~\ref{sec: wave num}, we instrumented the performance of approximations for a simple travelling wave model with various initial conditions.
-For this model the solution is given in terms of the initial condition by \(f(x,t) = f(x-t,0)\).
-This model is useful as it allows us to instrument the ability of the schemes to capture the flow of probability without stochastic dynamics, and since the solution is known, so there is no need for simulation.
-With the travelling wave model we demonstrate that, for problems with discontinuities, the DG scheme can display oscillations, while the other schemes (DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes) avoid oscillations.
+%However, once dynamics are introduced into the problem the approximation from the QBD-RAP scheme can be more accurate than that of the uniformisation scheme suggesting that the QBD-RAP scheme is better able to resolve the movement of probability between cells, as demonstrated in Section~\ref{sec: wave num}.
+
+Next, in Section~\ref{sec: wave num}, we investigated the performance of approximations for a simple travelling wave model with various initial conditions.
+The travelling wave model is useful as it allows us to investigate the ability of the schemes to capture the flow of probability without stochastic dynamics and, since the solution is known, there is no need for simulation.
+With the travelling wave model we demonstrate that, for problems with discontinuities, the DG scheme can display oscillations and violate the axioms of probability, while the other schemes (DG-lim, DG-lin-lim, uniformisation and QBD-RAP schemes) avoid oscillations.
 For discontinuous problems the rates of convergence of the QBD-RAP and DG-lin-lim schemes is similar.
@@
 
-We then instrumented the performance of the approximations on a simple fluid queue with two phases.
-First, in Section~\ref{sec:stat}, we looked at the stationary distribution, which is known to be smooth.
+We then investigated the performance of the approximations on a simple fluid queue with two phases.
+First, in Section~\ref{sec:stat}, we looked at the limiting distribution, which is known to be smooth.
 Since the problem is smooth, then the DG scheme was superior as expected.
@@
 
-In Section~\ref{sec: transient approx} we turned our attention to approximating transient distributions for the same model and consider two different initial conditions, a point-mass and an exponential initial condition.
+In Section~\ref{sec: transient approx} we turned our attention to approximating transient distributions for the same model and considered two different initial conditions, a point-mass and an exponential initial condition.
 The discontinuous initial condition results in a discontinuous transient distribution.
 As for the exponential initial condition, this example demonstrates that, even if the initial condition appears `nice', it can still result in non-smooth (i.e.~non-differetiable) solutions.
-The numerical evidence suggests that the DG scheme can display oscillations, while the other schemes do not.
-The DG-lim scheme detects the oscillations and reduces the approximation to linear.
-Of the uniformisation, DG-lin-lim and QBD-RAP schemes, the latter two performs similarly and better than the uniformisation scheme.
+The numerical evidence suggests that the lack of smoothness in the problem with the exponential initial condition is not problematic for the DG scheme, however, the DG scheme exhibits severe oscillations for the problem with the discontinuous initial condition.
+For both initial conditions, the DG-lim scheme detects oscillations and reduces the approximation to linear which renders the scheme not viable for these problems.
+Of the uniformisation, DG-lin-lim and QBD-RAP schemes, the latter two perform similarly, and better than the uniformisation scheme.
 Of the DG-lin-lim and QBD-RAP schemes the QBD-RAP scheme performed better in the presence of a point mass while the DG-lin-lim scheme performed better with the exponential initial condition.
 
-Next, Section~\ref{sec: return approx} looked at hitting times for the same fluid queue with two initial conditions, an exponential initial condition and a point-mass.
-We looked at the hitting time of the fluid level on the points \(\{0\}\) and \(\{1\}\).
-For this problem there is never any in-flow of mass at the boundaries of the interval and so, for a solution to be continuous, the initial condition needs to be chosen carefully, otherwise discontinuities in the transient distribution may result, as is the case for both initial conditions here.
-The numerical results suggest that, due to the discontinuities in the problems, the DG scheme may display oscillations.
+Section~\ref{sec: hit approx} investigated the performance of the schemes for approximating the hitting time distribution for the same fluid queue with two initial conditions, an exponential initial condition and a point-mass.
+For this problem there is never any in-flow of mass at the boundaries of the interval and so, for a solution to be continuous, the initial condition needs to be chosen carefully, otherwise discontinuities in the transient distribution may result, as was the case for both initial conditions considered here.
+The numerical results suggest that, due to discontinuities in the problems, the DG scheme may display oscillations and violate the axioms of probability.
 Since the uniformisation, DG-lin-lim, and QBD-RAP schemes can handle discontinuities, they perform suitably, with the QBD-RAP and DG-lin-lim schemes performing similarly and better than the uniformisation.
@@
 Lastly, we applied the DG, uniformisation and QBD-RAP schemes to two simple fluid-fluid queues in Section~\ref{sec: ffq num}.
-In the first fluid-fluid queue, which appears to have a smooth solution, the DG scheme performs very well.
-Of the two positivity preserving schemes the QBD-RAP scheme performs better than the uniformisation scheme.
-In the second example, which has a discontinuity, the DG scheme produces the lowest errors, but exhibits oscillations in the solution.
+In the first fluid-fluid queue, which appears to have a smooth first return distribution, the DG scheme performs very well (however, the initial condition for the DG scheme displayed oscillations).
+Of the two positivity preserving schemes considered, the QBD-RAP scheme performs better than the uniformisation scheme.
+In the second example, which has a discontinuity in the first return distribution, the DG scheme produces the lowest errors, but exhibits oscillations in the solution.
 The QBD-RAP and uniformisation schemes do not produce oscillations and, of the two, the QBD-RAP scheme performs best.
 
+As we mentioned in the introduction to this chapter, in Sections~\ref{sec: recon num}, \ref{sec: comp}, \ref{sec:stat}, and \ref{sec: ffq num} slope limiting is not used as it is not applicable: in these cases slope limiting is practically equivalent to post-processing the solution by detecting oscillations, then projecting oscillatory regions of the solution onto a basis of linear functions, and making some corrections to the solution if it happens to be negative.
+
+Moreover, when computing approximations to the first-return operator of a fluid-fluid queue in Section~\ref{sec: ffq num}, the QBD-RAP and uniformisation schemes lead to an algorithm that is theoretically justified.
+In contrast, to-date there no theory surrounding the DG approximation to the same operator (let alone any positivity preserving variants of the DG scheme).
+
 In conclusion, when the problem is known to be smooth, the DG scheme is very likely to produce excellent results.
-However, for discontinuous problems, the scheme can show oscillations and infeasible or negative regions of the solution.
+However, for discontinuous problems, the scheme can show oscillations and may lead to illegitimate approximations to distribution functions which violate the axioms of probability.
 The slope limiter overcomes this, but reduces the accuracy of the DG scheme to linear near discontinuities, sometimes severely affecting the quality of the approximation.
-A piecewise-linear DG scheme with a limiter, such as the DG-lin-lim scheme considered here, can produce satisfactory results.
-The uniformisation and QBD-RAP schemes are other alternative approximation schemes which avoid oscillatory solutions and guarantee a non-decreasing approximation of the CDF and have a stochastic interpretation.
+A piecewise-linear DG scheme with a limiter, such as the DG-lin-lim scheme considered here, can produce satisfactory results for some problems.
+The uniformisation and QBD-RAP schemes are other alternative approximation schemes which avoid oscillatory solutions and guarantee a legitimate approximation to distributions of fluid queues.
 Of the uniformisation and QBD-RAP schemes, the latter often produced lower errors and performed similarly to the DG-lin-lim scheme.
-Moreover, the uniformisation and QBD-RAP schemes have interpretations which can aid in their analysis (such as justifying the Riccati equation to approximate \(\mathbb \Psi\) for a fluid-fluid queue).
-The uniformisation and QBD-RAP schemes also result in linear approximations, whereas the DG-lim and DG-lin-lim schemes do not and this can be beneficial in certain contexts.
+Moreover, the uniformisation and QBD-RAP schemes have stochastic interpretations which can aid in their analysis.
+For the application to fluid-fluid queues in this chapter the stochastic interpretation is important as it justifies (and defines) the approximation to the first return operator.
+For the DG scheme there is no such theory (that I know of) and this means that the approximation to the first return operator using the DG scheme is not a rigorously defined object.
+On a similar note, the uniformisation and QBD-RAP schemes result in operators which are linear, whereas the DG-lim and DG-lin-lim schemes do not result in linear operators.
+When approximating the Riccati equation for the first return operator this is necessary as the uniformisation and QBD-RAP schemes result in a matrix-Riccati equation which can be easily solved in software, whereas slope-limited DG schemes do not lead to matrix formulations as they are non-linear.
+In summary, the QBD-RAP and DG-lin-lim schemes perform similarly and guarantee solutions which obey the axioms of probability, however approximating operators of fluid-fluid queues is justified when using the QBD-RAP scheme, it is not justified for the DG-lin-lim (or DG schemes in general).

### 2022-06-19 chapter7/chapter.tex
Ch7 2nd draft

@@
 Given the success of fluid queues it is believed that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success.
-\cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and stationary distribution of a fluid-fluid queue.
+\cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and limiting distribution of a fluid-fluid queue.
 

### 2022-06-19 front/abstract.tex
Ch7 2nd draft

@@
 
-This thesis provides approximations to fluid queue so that we can approximate the operators in \cite{bo2014}.
-It investigates three main approximation schemes; the DG scheme (Chapter~\ref{ch:galerkin}) which is a popular finite-element scheme, the uniformisation scheme of \cite{bo2013} which approximates a fluid queue but a continuous-time Markov chain (specifically, a quasi-birth-and-death-process (QBD)), and the QBD-RAP scheme which is a generalisation of a QBD to allow matrix exponential inter event times.
+This thesis provides approximations to fluid queues so that we can approximate the operators in \cite{bo2014}.
+It investigates three main approximation schemes; the DG scheme (Chapter~\ref{ch:galerkin}) which is a popular finite-element scheme, the uniformisation scheme of \cite{bo2013} which approximates a fluid queue by a continuous-time Markov chain (specifically, a quasi-birth-and-death-process (QBD)), and the QBD-RAP scheme which is a generalisation of a QBD to allow matrix exponential inter event times.
 The QBD-RAP scheme is novel; we describe the construction of the scheme in Chapter~\ref{sec: construction and modelling}, and provide an analysis to show that it is convergent in Chapters~\ref{sec: conv} and \ref{ch: global results}.

### 2022-06-19 intro/chapter.tex
Ch7 2nd draft

@@
 
-More recently, \cite{bo2014} extends fluid queues to so-called \emph{stochastic fluid-fluid queues}.
+% Of particular relevance to this thesis is the matrix-analytic analysis of fluid queues in \cite{bean2005}.
+Sample paths of fluid queues can be broken up into contiguous (random) intervals of time on which either, the phase remains in some subset of phases, \(\varphi(t) \in \calS_+'\) say, and the level process of the fluid queue is non-decreasing, or the phase remains in some subset of phases, \(\varphi(t) \in \calS_-'\) say, and the level process of the fluid queue is non-increasing.
+For \(m\in\{+,-\}\), the distribution of
+% \begin{align}
+% 	&\mathbb P(X(t)\geq x, \varphi(t)=j\in\calS_m', t\leq E^\lambda \mid X(0)=x_0, \varphi(0)=i\in\calS_m') \label{eqn: LLLppL}
+% \end{align}
+% where \(E^\lambda\) is an exponential random variable with rate \(\lambda\) is readily computable as a Phase-type distribution.
+In particular, the probability (\ref{eqn: LLLppL}) is characterised in terms of the \emph{generator} of the phase process, and the rates at which the level process moves in each phase.
+%Moreover, taking the Laplace transform with respect to time of the distribution yields
+% \begin{align*}
+% 	\mathbb P(X(t)\geq x, \varphi(u)\in\calS_+, u\in[0,t), \varphi(t)=j\in\calS_+, t\leq E^\lambda \mid \varphi(0)=i\in\calS_+)
+% \end{align*}
+
+More recently, \cite{bo2014} extend fluid queues to so-called \emph{stochastic fluid-fluid queues}.
 In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\), so \((X(t),\varphi(t))\) determines the rate at which \(\{Y(t)\}\) moves.
@@
 For example, \cite{hydro,bo2014} describe a model for a hydroelectric power generator which can be operated `on design' and `off design'.
-When the system operates `off design' there system is less efficient and there is more wear, but may be used to optimise overall system performance.
+When the system operates `off design' the system is less efficient and there is more wear, but may be used to optimise overall system performance.
 Further, maintenance of the generator takes place periodically to improve its performance and lifespan.
@@
 
-The addition of the second level process \(\{Y(t)\}\) greatly increases the flexibility and modelling potential, however comes at some cost to the ease of their analysis.
+The addition of the second level process \(\{Y(t)\}\) greatly increases the flexibility and modelling potential.
+However, the added flexibility comes at a cost to the in the analysis of the model.
 The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
 For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used.
-To this end, to date, two possible discretisation have been suggested.
+To this end, to date, two possible discretisations have been suggested.
 Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise the operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process.
@@
 Furthermore, for the DG method there is a limited theoretical backing as to why we may approximate the operator-Riccati equation with the matrix-Riccati equation other than it is a sensible thing to do, and it seems to work in practice.
-In the case of the QBD approximation of \cite{bo2013} the approximate matrix-Riccati is more justified as in it can be derived by considering the first-return probabilities for a fluid queue driven by the approximating QBD where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\) \citep{bean2005}.
+In the case of the QBD approximation of \cite{bo2013} the approximate matrix-Riccati equation is more justified as in it can be derived by considering the first-return probabilities for a fluid queue driven by the approximating QBD where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\) \citep{bean2005}.
 
@@
 
-As the QBD-RAP has a stochastic interpretation then the approximations it produces are guaranteed to have non-negative density functions.
+As the QBD-RAP has a stochastic interpretation the approximations it produces are guaranteed to have non-negative density functions.
 Further, for the QBD-RAP (and uniformisation) scheme non-negativity is guaranteed for any initial condition without any further computation or post-processing.
@@
 The rest of Chapter~\ref{ch: global results} then attempt to prove a global converge result.
-The main result being Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (weakly with respect to the spatial and temporal variable) to the distribution of the fluid queue.
+The main result being Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (with respect to the spatial and temporal variable) to the distribution of the fluid queue.
 
@@
 
-To help the reader understand the notations used in the DG scheme, we provide a small toy model example in Appendix~\ref{appendix:example}.
+To help the reader understand the notation used in the DG scheme, we provide a small toy model example in Appendix~\ref{appendix:example}.
 Appendix~\ref{sec:properties} provides a proof that the DG scheme conserves probability.

### 2022-08-05 chapter1/chapter.tex
ch 7/8 comments

@@
 Here we very briefly introduce semigroups and their \emph{infinitesimal generators}.
+We refer to the reader to \cite{ethierkurtz} for more details (see also \cite{kallenberg}, or the more approachable course notes of \cite{shalizi}).
 
 \subsubsection{Operators}
-Let \(L\) be a Banach space (a vector space with a metric and where Cauchy sequences converge to a point within the space).
-An operator is a mapping which takes elements of a vector space and sends them to elements of a vector space (not necessarily the same space).
-That is, let \(V\) and \(W\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\subset V\to W\) is an operator.
-The set \(\mathcal D(\mathbb B)\) is the domain of the operator, and is not necessarily all of \(V\).
+Let \(L\) be a \emph{Banach space}.
+A Banach space is a \emph{complete, normed} vector space.
+Complete means that every Cauchy sequence of points in \(L\) has a limit which also lies in \(L\).
+Normed means that a norm is defined on \(L\).
+Intuitively, a norm is a function which tells us the distance between two points in \(L\).
+
+An \emph{operator} is a mapping which takes elements of a vector space and sends them to elements of a vector space (not necessarily the same space).
+That is, let \(V\) and \(W\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\to W\) is an operator.
+The set \(\mathcal D(\mathbb B)\subseteq V\) is the domain of the operator, and is not necessarily all of \(V\).
 When specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to W\).
+
 Let \(c\in \mathbb R\) and \(v\in V\), \(w\in W\).
 Then \(\mathbb B\) is a \emph{linear} operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\).
-We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms attached to \(V\) and \(W\), respectively.
+We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms defined on \(V\) and \(W\), respectively.
 An operator is said to be bounded if \(||\mathbb B||<M\) for some \(M<\infty, M\in \mathbb R\), for every \(v\in V\).
@@
 \subsubsection{Semigroups}  A family of linear operators is a collection of indexed linear operators \(\{\mathbb B(i)\}_{i\in I}\) where \(I\) is some index set.
-That is, for each \(i\in I\), \(\mathbb B(i)\) is a linear operator.
+For each \(i\in I\), \(\mathbb B(i)\) is a linear operator.
 A family of operators \(\{\mathbb T(t)\}_{t\geq 0}\), \(\mathbb T(t):L\to L\), is said to have the \emph{semigroup} property if \(\mathbb T(t+s)=\mathbb T(t)\mathbb T(s)\).
-If \(\mathbb T(0)=I\), the identity operator, then the family \(\{\mathbb T(t)_{t\geq 0}\) is known as a semigroup.
-If \(\mathbb T(t)f\to f\) as \(t\to 0^+\), then \(\{\mathbb T(t)\}_{t\geq 0}\) is said to be strongly continuous.
+If \(\mathbb T(0)=I\), the identity operator, then the family \(\{\mathbb T(t)_{t\geq 0}\) is known as an operator semigroup.
+If \(f\in L,\, \mathbb T(t)f\to f\) as \(t\to 0^+\), then \(\{\mathbb T(t)\}_{t\geq 0}\) is said to be strongly continuous.
 If \(||\mathbb T(t)||\leq 1\) for all \(t\geq 0\) then \(\{\mathbb T(t)\}_{t\geq 0}\) is known as a contraction semigroup.
@@
 
-% \section{Markov processes and semigroups}
-% Semigroups arise naturally in Markov processes.
+\subsubsection{Markov processes and semigroups}
+Semigroups arise naturally in Markov processes.
 Let \(\{X(t):t\geq 0\}\) be a Markov process with state space \(\mathcal S\).
-The Markov property states that for \(s,t\geq 0\), \(X(t+s)\) is independent of \(X(u),\, 0\leq u<t\) given \(X(t)\).
-That is, \(P(X(t+s)\in\mathcal{E} \mid X(t), X(u), 0\leq u<t)=P(X(t+s)\in\mathcal{E} \mid X(t))\).
-If \(P(X(t+s)\in\mathcal{E} \mid X(t))=P(X(s)\in\mathcal{E} \mid X(0))\), that is the distributions are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.
-Where necessary, we will use the notation \(X_\mu(t)\) to denote the Markov process with initial distribution \(\mu\), i.e.~\(P(X(0)\in\cdot)=\mu(\cdot)\), and \(X_x(t)\) to be the Markov process with initial distribution \(\delta(x)\).
-Markov processes can be characterised by \emph{probability kernels}.
-
-% Let \((U,\mathcal U)\) and \((V,\mathcal V)\) be two measure spaces, then a probability kernel \(\nu\), is a mapping \(\nu:U\times \mathcal V\to \mathbb R_+\) such that \(\nu(u,\mathcal{E})\) is \(\mathcal U\)-measurable in \(u\in U\) for fixed \(\mathcal{E}\in \mathcal V\), and a probability measure in \(\mathcal{E}\in \mathcal V\) for fixed \(u\in U\).
+The Markov property states that, given \(X(t)\), then for \(s,t\geq 0\), \(X(t+s)\) is independent of \(X(u),\, 0\leq u<t\).
+That is, \[P(X(t+s)\in\mathcal{E} \mid X(t), X(u), 0\leq u<t)=P(X(t+s)\in\mathcal{E} \mid X(t)).\]
+If \(P(X(t+s)\in\mathcal{E} \mid X(t))=P(X(s)\in\mathcal{E} \mid X(0))\), are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.
+
+% Markov processes can be characterised by \emph{probability kernels}.
+Let \((U,\mathcal U)\) and \((V,\mathcal V)\) be two measure spaces, then a probability kernel \(\nu\), is a mapping \(\nu:U\times \mathcal V\to \mathbb R_+\) such that \(\nu(u,\mathcal{E})\) is \(\mathcal U\)-measurable in \(u\in U\) for fixed \(\mathcal{E}\in \mathcal V\), and a probability measure in \(\mathcal{E}\in \mathcal V\) for fixed \(u\in U\).
 We say that \(\nu\) is a probability kernel from \(U\) to \(V\).
@@
 
-% Let \(\mu_{t}(\cdot)=P(X(t)\in \cdot)\) for \(t\geq 0\) be a one-parameter family of measures on \((\mathcal S,\mathcal F)\), and \(\nu_{t}(x,\cdot)=P(X(t)\in\cdot \mid X(0)=x)\) for \(t\geq 0\) be a one-parameter family of probability kernels from \(\mathcal S\) to \(\mathcal S\).
+Consider the expectation
+\[\mathbb E[f(X(t))\mid X(0)=x]=\int_{y\in\calS} f(y) P(X(t)\in\wrt y\mid X(0)=x),\]
+where \(f\) is some bounded function.
+Expectation is an operator which is acting on a function \(f\).
+The expectation depends on \(t\) and \(x\) as these define the distribution with which we are taking the expectation, hence, let us write
+\[\mathbb T(t)f(x) = \mathbb E[f(X(t))\mid X(0)=x].\]
+The result of the operator acting on \(f\), \(\mathbb T(t)f(x)\), is another function.
+
+By the Markov property, for some \(s\in[0,t]\), 
+\begin{align*}
+	&\int_{y\in\calS} f(y) P(X(t)\in\wrt y \mid X(0)=x)
+	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}f(y) P(X(t)\in\wrt y\mid X(s)=z) P(X(s)\in\wrt z\mid X(0)=x)
+	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}\mathbb E[f(X(t))\mid X(s)=z] P(X(s)\in\wrt z\mid X(0)=x)
+	\\&=\int_{y\in\calS}\int_{z\in\mathcal S} \mathbb E[f(X(t-s))\mid X(0)=z]P(X(s)\in\wrt z\mid X(0)=x),
+	\\&=\int_{y\in\calS}\int_{z\in\mathcal S} (\mathbb T(t-s)f(z))P(X(s)\in\wrt z\mid X(0)=x),
+	\\&=\mathbb T(s) (\mathbb T(t-s)f)(x),
+\end{align*}
+where the third equality holds by time-homogeneity.
+Hence, \(\{\mathbb T(t)(\cdot)\}_{t\geq 0}\) is an operator semigroup.
+Moreover, since \(f\) is bounded, then \(\mathbb T(t)f(x) = \mathbb E[f(X(t))\mid X(0)=x] \leq F\), which means that \(\{\mathbb T(t)\}\) is a contraction semigroup.
+
+% Let \(\nu_{t}(x,\cdot)=P(X(t)\in\cdot \mid X(0)=x)\) for \(t\geq 0\) be a one-parameter family of probability kernels from \(\mathcal S\) to \(\mathcal S\).
 Then \(\mu_t\) and \(\nu_{t}\) characterise the time-homogeneous Markov process \(X(t)\).
@@
 \begin{thm}[Theorem 4.1.4, \cite{MEinAP}]\label{thm: 4.1.4}
-	Consider the nonvanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}\neq 0,\,d_{jk}\neq 0,\) or \(e_{jk}\neq 0\).
+	Consider the non-vanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}\neq 0,\,d_{jk}\neq 0,\) or \(e_{jk}\neq 0\).
 Among the corresponding eigenvalues \(\lambda_j\), there is a real dominating eigenvalue \(\kappa\), say.
@@
 %By CITE it is known that \(\mathcal{E}\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).
+
+\subsubsection*{Concentrated matrix exponential distributions}
+Recently, the class of \emph{concentrated matrix exponential distributions} (CMEs) has been investigated \citep{ert2006,hstz2016,ert2006,hht2020,mt2021}.
+As the name suggests, concentrated matrix exponential distributions are matrix exponential distributions with a very low coefficient of variation (variance relative to the mean).
+As the order, \(p\), of the representation increases, the variance of concentrated matrix exponential distributions decreases at rate approximately \(\mathcal O(1/p^2)\).
+For comparison, the variance of an Erlang distribution, the most concentrated Phase-type distribution, decreases at rate \(\mathcal O(1/p)\) as the order \(p\) increases \citep{as1987}.
+In this thesis we use CMEs to approximate deterministic events.
+For a given order, concentrated matrix exponential distributions have a much lower coefficient of variation than any phase type of the same order, and therefore a better ability to model determinism.
+
+The class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hht2020} and computationally efficient expressions for the density and moments of CMEs are provided in \citep{hht2020}.
+
+CMEs exist for all orders, however, in our computations we only use CMEs with odd order.
+The justification for considering representations of odd orders only is that the variance of CME distributions of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}.
+Hence, if we construct an approximation with a representation of order \(2p\) we expect it to perform only marginally better than an approximation constructed with representations of order \(2p-1\).
+However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.
+
+From the definition of the class of CME distributions, for a given CME with odd order, \(p\), and representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part \cite{hht2020}.
+
+For CMEs considered here, the vector function \(\bs k(t)=\cfrac{\bs \alpha e^{St}}{\bs \alpha e^{St}\bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
+We leverage this fact to simplify some numerical integration in Chapter~\ref{sec: construction and modelling}.
 
@@
 
+\section{Discontinuous Galerkin}
+The discontinuous Galerkin (DG) method is a general methodology for numerically approximating solutions to differential equations.
+The discontinuous Galerkin method is a finite-element method which approximates differential operators by finite dimensional matrices.
+In this thesis we use the discontinuous Galerkin method to approximate the infinitesimal generator of a fluid queue.
+Here, we give a very brief introduction to discontinuous Galerkin methods in general and refer the reader to Chapter~\ref{ch:galerkin} for details on the application to fluid queues, and to \cite{nodalDGBook} for more details on the method in general.
+
+Intuitively, the DG method works by first partitioning the spatial domain into cells, on each cell a basis of (typically polynomial) functions is defined, then the differential equation is projected onto the basis of polynomials.
+Thus, the differential equation is discretised it two ways, by the partition of space into cells, then by the projection onto a finite dimensional basis.
+Linking the cells together essentially amounts to defining boundary conditions for each cell.
+To form the boundary condition for cells adjacent to the boundary we take the boundary condition defined by the problem we are solving and discretise it.
+For cells in the interior the boundary condition is defined by the solution on the adjacent cell.
+In the context of conservation laws (which is what we solve in this thesis), the boundary conditions on each cell are known as the \emph{flux}, as they describe the rate, with respect to time, that the conserved quantity flows across the boundaries of the cell.
+The flux is discretised and is known as the \emph{numerical flux}.
+
+Typically, polynomial functions are used as the basis for the DG method.
+The main choices to make in defining the scheme are the grid which defines the cells, the order of the polynomial basis, and the choice of numerical flux.
+
+\cite{nodalDGBook} claim that the first discontinuous Galerkin finite element scheme was developed in \cite{reed1973} where they use the method to solve the first order differential equation
+\[\sigma u + \nabla \cdot (a u) = f,\]
+in a 2-dimensional spatial domain, where \(\sigma\) is a constant, \(a\) is a piecewise constant function, and \(u\) is the unknown.
+The literature has since expanded and includes (but is not limited to) theory on the convergence rate of the scheme, the choice of \emph{numerical flux} used (we will define this later), the choice of spatial discretisation used, and applications to a range of other problems including higher-order differential equations and to partial differential equations (PDEs) such as non-linear conservation law.
+The PDE which describes the evolution of the distribution of a fluid queue is a conservation law as probability must be conserved.
+
+In one dimension, linear conservation laws take the form 
+\begin{align}
+	&\cfrac{\partial u( x, t)}{\partial t} - \cfrac{\partial}{\partial x} ( c(t) u( x, t)) = 0,\,  x \in \Omega, \label{eqn: cons law}
+	\\& u( x,t)= g( x,t),\tag{boundary condition}
+	\\& u( x,0)=f( x).\tag{initial condition}
+\end{align}
+where \(\Omega\) is the region of interest. Here \(u\) is the conserved quantity, which we will refer to as \emph{mass}, but it could really be any conserved quantity.
+
+To solve conservation laws with the DG method the spatial domain \(\Omega\) is partitioned into cells (intervals) \(\mathcal D_k = [y_k,y_{k+1}]\), \(k=0,...,K-1\).
+For each cell \(\calD_k\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), compactly supported on \(\calD_{k}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k}\)).
+Typically, we take \(\phi_k^r(x)\) to be a basis of polynomial functions.
+The functions \(\{\phi^r_k\}_{r=1}^{p_k}\) span some function space, \(W_k\) say, and further form a basis for this space.
+It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{p_k}\) as a basis of \emph{Lagrange interpolating polynomials} defined by the \emph{Gauss-Lobatto quadrature points} \citep{nodalDGBook} (see also Section~\ref{eqn: la}).
+For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials.
+On each cell \(\mathcal D_{k}\) we approximate
+\[u(x,t)\approx u_{k}(x,t)=\sum\limits_{r=1}^{p_k}a_{k}^r(t)\phi^r_k(x),\] 
+where \(a_{k}^r(t)\) are yet-to-be-determined time-dependent coefficients.
+The approximation \(u_{k}\) is referred to as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k=0}^{K-1} u_{k}\) on the whole domain \(\Omega\).
+The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\{0,...,K-1\}\).
+
+First, consider multiplying (\ref{eqn: cons law}) by the row vector \(\bs \phi_k(x)=[\phi_k^1(x),...,\phi_k^{p_k}(x)]\) and integrating by parts over \(\calD_k\),
+\begin{align}
+	&\int_{x\in \calD_k}\cfrac{\partial u(x,t)}{\partial t}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) u(x,t)\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x - c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}} = 0.\label{eqn: hadndjj}
+\end{align}
+For a sufficiently large basis of functions, then (\ref{eqn: hadndjj}) is equivalent to the PDE (\ref{eqn: cons law}).
+The term \(- c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}}\) is known as the \emph{flux} as it describes the rate at which mass flows into and out of the cell \(\calD_k\).
+The flux is essentially a boundary condition on the cell \(\calD_k\) and is specified by the value of the solution on the adjoining cells.
+When \(c(t)\) is positive (negative) then mass flows into (out of) cell \(\calD_k\) at the left-hand boundary and out of (into) cell \(\calD_k\) at the right-hand boundary.
+Suppose that \(c(t)\) is positive.
+When \(u\) is continuous, then \(u(y_{k}^-,t)=u(y_k,t),\) where \(y^-\) is the left limit at \(y\), hence the flux into cell \(\calD_k\) in Equation (\ref{eqn: cons law}) is \(c(t)u(y_k^-,t)\bs \phi_k(y_k)\).
+The flux out of cell \(\calD_k\) is \(c(t)u(y_{k+1}^-,t)\bs \phi_k(y_{k+1})\).
+Using this, rewrite (\ref{eqn: hadndjj})
+\begin{align}
+	&\int_{x\in \calD_k}\cfrac{\partial u(x,t)}{\partial t}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) u(x,t)\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x \nonumber
+	\\&\qquad\qquad{}- c(t)[u(y_{k+1}^-,t)\bs \phi_k(y_{k+1}) - u(y_{k}^-,t)\bs \phi_k(y_{k})] = 0.\label{eqn: hadndjjj}
+\end{align}
+We can proceed analogously when \(c(t)\) is negative.
+
+Upon substituting the approximations \(u_k\) and \(u_{k-1}\) into (\ref{eqn: hadndjjj}) 
+\begin{align}
+	&\int_{x\in \calD_k}\cfrac{\partial\bs a_k(t)}{\partial t}\bs \phi_k(x)\tr{}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) \bs a_k(t)\bs\phi_k(x)\tr{}\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x \nonumber
+	\\&\qquad\qquad{}- c(t)[\bs a_k(t)\bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}) - \bs a_{k-1}(t)\bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k})].\nonumber
+	\\&=\cfrac{\wrt \bs a_k(t)}{\wrt t}\bs M_k + c(t) \bs a_k(t)\bs G_k - c(t)[\bs a_k(t)\bs F_{k,k} - \bs a_{k-1}(t)\bs F_{k-1,k}]\nonumber
+	\\&= 0,\label{eqn: hadndj}
+\end{align}
+where 
+\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x,\]
+\[\bs F_{k,k} := \bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}),\quad \bs F_{k-1,k} := \bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k}),\]
+and are known as the \emph{mass}, \emph{stiffness} and \emph{flux} matrices.
+Thus, the approximation has reduced the problem to a system of ordinary differential equations.
+One can determine the coefficients \(\bs a_k(t)\), \(k=0,...,K-1\), from the initial condition \(\bs a_k(0)\), \(k=0,...,K-1\), then integrating over time, perhaps numerically.
+The initial coefficients \(\bs a_k(0)\), \(k=0,...,K-1\), can be found by projecting the initial condition onto the basis of polynomial functions, \(\{\phi_k^r\}_{k,r}\).
+
+One subtlety is the choice of numerical flux.
+We arrived at the numerical flux above using a continuity argument.
+However, the approximate solution \(\sum_{k=0}^{K-1}u_k(x,t)\) is discontinuous at the points \({y_{k}}\).
+It turns out that the choice of numerical flux is not unique and there are range of values we may choose from.
+The numerical flux above is known as the \emph{upwind} flux -- we use it throughout this thesis and do not explore others.
+Other choices which define a discontinuous Galerkin scheme are the basis functions used and the cell geometries.
+We refer the reader to \citep{nodalDGBook} and reference therein for more details.
+
+Typically, DG schemes are used to solve physical problems in 2 or 3 dimensions.
+The cell geometries become more complex in higher dimensions and this complicates matters somewhat, but the underlying principles and intuition is the same.
+In this thesis we solve problems with one spatial dimension only.
+
+Two properties make the DG scheme a natural choice to solve the problem present in this thesis.
+High-order accuracy enabled by the use of basis functions to model the solution on each cell, and the cell-based structure of the method with cells related to each other at cell edges only and only to neighbouring cells.
+Finite difference methods can achieve high-order accuracy, however they are not cell-based, which makes them not a natural choice to solve the problems in this thesis.
+Finite volume methods are cell-based, but do not have the same high-order accuracy as the DG method.
+Further, approximating flux terms with greater accuracy in the finite volume method can introduce dependencies between cells which are not neighbours and this creates difficulties for the application in this thesis.
+
+\subsection{Time-integration schemes}\label{sec: time integration}
+In this thesis we numerically integrate ODEs of the form
+\begin{align}\label{eqn: DG ODE w BCs 3}
+	\cfrac{\wrt}{\wrt t} \boldsymbol a(t)
+	% 
+	= \boldsymbol a(t) \bs Q
+\end{align}
+where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix, given an initial condition \(\bs a(0)\).
+To do so, we employ the strong stability preserving Runge-Kutta (SSPRK) scheme of order 4 with 5 stages \citep{sr2002}.
+SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``...can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
+
+In our context, the SSPRK method with \(s\) stages has the form 
+\begin{align*}
+	&\bs v^{(0)} = \bs a(t),\\
+	%
+	&\bs v^{(\ell)} = \sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q,\, \ell = 1,\dots,s,\\
+	%
+	&\bs a(t+h) =\bs v^{(s)},
+\end{align*}
+where \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are coefficients which define the scheme and \(h\) is the \(t\)-step size.
+To ensure the stability-preserving property the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are chosen to be positive so that \(\bs v^{(s)}\) is a convex combination of forward-Euler steps.
+Moreover, the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are optimised so that the allowable \(t\)-step size is as large as possible.
+The maximum allowable \(t\)-step size is \(h_{RK}\), and we require
+\[h_{RK} \leq \min_{\ell k}\cfrac{\alpha_{\ell k}}{\beta_{\ell k}}h_E,\]
+where \(h_E\) is the maximum allowable \(t\)-step size for the forward-Euler scheme.
+
+The \(t\)-step size needs to be chosen to be sufficiently small so that the time integration is stable. Here, for an order \(p\) DG scheme we choose the time step to be less than 
+\[\max_{i\in\calS}|c_i|\min_{r\in{2,...,p+1}}\left(z_{r}-z_{r-1}\right) \cfrac{\min_{k}{\Delta}_k}{2}\]
+where \(z_{r-1}\leq z_r\), \(z_1=-1\), \(z_{p+1}=1\) and \(z_r, r=2,...,p,\) are the \(p-1\) zeros of the first derivative of \(P_{p}(x)\), the order \(p\) Legendre polynomial.
+This follows advice from \cite{nodalDGBook}.
+
+\subsection{Slope limiters}\label{sec: slope limiting}
+A well-know problem with DG schemes is that, in the presence of steep gradients or discontinuities, spurious oscillations in the approximate solution can occur and positivity is not guaranteed (see \cite{nodalDGBook}~Section~5.6, and \cite{koltai2011}~Section~3.3, for example).
+To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
+Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
+However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
+As a result, the limiter may unnecessarily reduce the accuracy of the scheme in the presence of natural oscillations (see \citep{nodalDGBook},~Example~5.8).
+Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear.
+
+We now describe a simple slope limiter known as the \emph{Generalised MUSCL} scheme \citep{c99} (see also \citep{nodalDGBook},~Section~5.6.1). Define the \emph{minmod} function 
+\begin{align*}
+	m(a_1,a_2,a_3)=\begin{cases}
+		s\min_{i\in\{1,2,3\}}|a_i|, & |s|=1, \\
+		0, &\mbox{otherwise},
+	\end{cases}
+\end{align*}
+where \(s=\frac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
+When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
+In the context of DG the three arguments \(a_1,\,a_2\) and \(a_3\) are the estimates of the average gradient of the approximate solution near a given cell.
+When the signs of \(a_1,\,a_2\) and \(a_3\) differ, this suggests an oscillation in the approximate solution.
+
+Now, define \(\overline u_{k}\) as the average value of the approximate solution on cell \(k\), then the slope limited solution on cell \(k\) is the linear approximant 
+\begin{align*}
+	\Pi_{k}^{lim}(u_k)=\overline u_{k} + (x-\overline y_k)m\left(\Pi^1_{\partial x}u_k,\cfrac{\overline u_{k+1}-\overline u_{k}}{\Delta_k},\cfrac{\overline u_{k}-\overline u_{k-1}}{\Delta_k}\right),
+\end{align*}
+where \(\Pi^1_{\partial x}u_k\) is the slope of the linear projection of \(u_k\) on cell \(k\) and \(\overline y_k=(y_k+y_{k+1})/2\) is the centre of the \(k\)th cell.
+
+In non-oscillatory regions of the solution the slope limiter is unnecessary and reduces accuracy, so we only apply the limiter to cells \(k\) where oscillations are detected.
+To determine which cells need limiting we compute
+\begin{align*}
+	v_k^L &:= \overline{u}_{k} - m(\overline{u}_{k} - {u}_{k}^L,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+	\\v_k^R &:= \overline{u}_{k} + m(\overline{u}_{k} - {u}_{k}^R,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+\end{align*}
+where \({u}_{k}^L\) and \({u}_{k}^R\) are the values of the approximate solution on cell \(k\) evaluated at the left-hand and right-hand edges of the cell.
+We apply the slope limiter to cell \(k\) if \(v_k^L\neq u_k^L\) or \(v_k^R\neq u_k^R\), otherwise the approximation on cell \(k\) is left unaltered.
+
+% Notice that the process of limiting is essentially a post-processing of the approximate solution after applying
+
+% In the context of approximating the first return operator of a fluid-fluid queue there is no direct equivalent of slope limiting, other than, perhaps, to re-compute the solution with constant basis function if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
+
+\subsection{Slope limiters and time integration}\label{subsec: slope lim and int}
+To use slope limiters within the SSPRK scheme we first project the initial condition on to \(W\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
+To find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the (limited) initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation~(\ref{eqn: DG ODE w BCs 3}).
+At each stage of the time-integration we apply the slope limiter, i.e.~the scheme above with a slope limiter is
+\begin{align*}
+	&\bs v^{(0)} = \bs a(t),\\
+	%
+	&\bs v^{(\ell)} = \Pi^{lim}\left(\sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q\right),\, \ell = 1,\dots,s,\\
+	%
+	&\bs a(t+h) =\bs v^{(s)},
+\end{align*} 
+where \(\Pi^{lim}\) is the slope limiter function which determines on which cells the solution needs limiting, and if so, applies the limiter \(\Pi_k^{lim}\).
+
+We refer to this (vanilla) application of a limiter to the DG scheme as the DG-lim scheme.
+
+\paragraph{Consequences of slope limiting} We have already mentioned that the Generalised MUSCL slope limiter reduces the order to linear in regions where oscillatory solutions are detected.
+When slope limiting is used in conjunction with a time-integration scheme the reduction in order may not be local in time and may persist past the time when the limiter first detected oscillations.
+For example, consider a problem with an initial condition which introduces oscillations into the numerical approximation.
+If the slope limiter is applied to the initial condition then, in the region around the oscillations, the approximation to the initial condition will be linear.
+Even if the slope limiter is not applied (or not needed) after this point, the initial condition has still been altered, perhaps significantly, from the original approximation and this can affect the accuracy of transient approximations.
+This is not to say that the slope-limited-regions of the approximation will \emph{always} remain linear.
+When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of generalised slope limiter described.
+Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy of subsequent computations.
+
+\subsection{Another slope limiting scheme}\label{sec: anoth pos pres}
+The advantage of the approach above is that, in areas of the approximation where the solution is smooth, then the high-order accuracy of the method can be retained, while maintaining positivity where necessary.
+However, if the problem at hand is dominated by discontinuities and/or if there are no regions of the solution of interest which are smooth, then there is no real advantage in the approach above as the limiter will reduce the approximate solution to linear in all regions of interest.
+Hence, we may as well just use a linear scheme and save on some computation.
+
+Perhaps a better approach would be to use a linear scheme with a smaller cell width such that the computational work remains approximately the same.
+For example, for a DG scheme with \(2p\) basis functions on a cell, \(\mathcal D_k\), of width \(\Delta\), say, we construct block matrices of dimension \(2p\times 2p\) (i.e.~the mass, stiffness and flux matrices, \(\bs M_k\), \(\bs G_k\), and \(\bs F^{k\ell}\), respectively) and compute the approximate solution using these matrices.
+Alternatively, consider breaking the cell, \(\mathcal D_k\), up into \(p\) sub-cells of width \(\Delta/p\) and using a DG scheme with \(2\) basis functions on each sub-cell (i.e.~a linear basis on each cell).
+On each of the \(p\) sub-cells we construct block matrices of size \(2\times 2\).
+To approximate the solution on the original interval \(\mathcal D_k\) we form the \(p\) approximations on each sub-cell into larger matrices which are of dimension \(2p\times 2p\), and then use in the same way as before.
+Thus, the scheme with \(2p\) basis functions on the cell \(\calD_k\), and the scheme with a 2 basis functions on each of the \(p\) sub-cell have approximately the same computational cost (the latter scheme results in a sparser matrix).
+
+If we know \emph{a priori} that we will apply the slope limiter to cell \(\calD_k\), then intuitively we suspect that the latter scheme may be far more accurate; it will approximate the solution on \(\calD_k\) by a piecewise linear function with \(p\) pieces, whereas the DG scheme with \(2p\) basis functions on cell \(\calD_k\) will, after limiting, represent the solution as a single linear function on \(\calD_k\).
+
+We refer to this scheme as the DG-lin-lim scheme, as it is a linear DG scheme with a limiter.
+
+\subsection{Briefly, on accuracy}
+As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method from \cite[Sections~5.5,~5.8]{nodalDGBook} for a similar kind of problem to the one under study here.
+For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
+Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
+
+% To quote Section~5.7 of \cite{nodalDGBook} SSPRK methods ``can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
+
+% why the MUSCL limiter?
+%   it requires no tuning and guarentees non-negativity
+%   some other positivity-preserving methods require tuning and can still produce negative results if not tuned properly
+%   ^ same goes for filtering
+
+% Consider the partial differential equation 
+% \[\cfrac{\partial f}{\partial t}+\cfrac{\partial c(f)}{\partial x}=0,\, x\in[a,b],\]
+% where \(c\) is some function, and subject to appropriate initial, \(f(x,0)=f_0(x)\), and boundary conditions, \(f(a,t)=g(t),\) assuming \(c\geq 0\).
+The method first supposes that the solution is of the form \(f(x,t)\) translates the above differential equation, which is in \emph{strong form}, into a \emph{weak form} by multiplying by a test function \(\psi\), integrating with respect to \(x\) and applying integration by parts;
+% \begin{align*}
+% 	% &\int_{x=a}^b\cfrac{\partial f(x)}{\partial t}\psi(x)\wrt x+\int_{x=a}^b\cfrac{\partial c(f)}{\partial x}\psi(x)\wrt x
+% 	\int_{x=a}^b\cfrac{\partial f(x)}{\partial t}\psi(x)\wrt x+\int_{x=a}^b c(f)\cfrac{\wrt\psi(x)}{\wrt x}\wrt x - [c(f(x,t))]_{x=a}^b = 0. 
+% \end{align*}
+% The weak form
+
+\section{Sundry mathematical concepts}\label{eqn: la}
+\subsection*{Polynomials}
+In this thesis, we use the discontinuous Galerkin method to approximate the generator, a differential operator, of a fluid queue.
+The discontinuous Galerkin method involves projecting the operator equation onto a basis of functions, typically polynomials.
+A convenient basis with which to work is the interpolating Lagrange polynomials.
+The order \(p-1\) Lagrange polynomials are defined by a set of points, \(\xi_i, i=1,...,p\), (the \(\xi_i\)'s must be distinct), and are given by
+\[\ell_i(r)=\prod_{\substack{j=1\\j\neq i}}^p \cfrac{r-\xi_j}{\xi_i-\xi_j},\, i=1,...,p.\]
+A convenient property of the Lagrange polynomials is 
+\[l_i(\xi_j)=\begin{cases}1 & \mbox{ if } i=j, \\ 0 & \mbox{ otherwise.} \end{cases}\]
+
+Sometimes it is more convenient to work with \emph{orthogonal} polynomials. For example, the Legendre polynomials, which can be defined recursively by 
+\begin{align*}
+	P_0(x)&=1,
+	\\P_1(x)&=x, 
+	\\(n+1)P_{n+1}(x)&=(2n+1)xP_n(x)-nP_{n-1}(x),
+\end{align*}
+and are orthogonal on \([-1,1]\).
+
+The zeros of \((1-x^2)P_n'(x)\) define the \emph{Legendre-Gauss-Lobatto} points, which are used, among other things, for numerically approximating integrals via quadrature.
+
+\subsection*{Numerical integration}
+To numerically approximate integrals in the discontinuous Galerkin schemes in this thesis we can use Gauss-Lobatto quadrature.
+Consider the integral of some function \(f\) over the interval \([-1,1]\).
+Quadrature approximates the integral by evaluating the function on a set of points, \(x_i,\, i=1,...,p\), and computing the weighted sum
+\[\int_{-1}^1f(x)\wrt x \approx \sum_{i=1}^p f(x_i)w_i,\]
+where \(w_i\) are weights. There are various quadrature schemes which one can use.
+
+For the discontinuous Galerkin method, Gauss-Lobatto quadrature is convenient as in both we evaluate the function at the end points of the interval.
+For Gauss-Lobatto quadrature, the nodes, \(x_i, i=1,...,p\) is the zero of \((1-x^2)P_n'(x)\) where \(P_n\) is the \(n\)th Legendre polynomial, the weights are given by
+\begin{align*}
+	w_1&=1,
+	\\w_i&=\cfrac{2}{n(n-1)[P_{n-1}(x_i)]^2}, \, i\neq 1, p,
+	\\w_p&=1.
+\end{align*}
+Gauss-Lobatto quadrature is accurate for polynomials up to degree \(2p-3\). An integral over the interval \([a,b]\) can be approximated by 
+\[\int_{a}^bf(x)\wrt x \approx \cfrac{b-a}{2}\sum_{i=1}^p f\left(\cfrac{b-a}{2}x_i+\cfrac{a+b}{2}\right)w_i.\]
+
+We also use a trapezoidal integration rule at times. Consider the integral \(\int_{a}^bf(x)\wrt x\). A trapezoidal rule with \(p\) evenly spaced grid-points approximates the integral via 
+\[\int_{a}^bf(x)\wrt x \approx \sum_{i=2}^p \cfrac{f(x_{i-1}+f(x_i))}{2}\Delta,\]
+where \(x_i=a+(i-1)\Delta\), \(i=1,...,p\), and \(\Delta = \cfrac{b-a}{p-1}.\)
+
+\subsection*{Measuring the difference between distributions}
+In the numerical investigations in this thesis we evaluate approximation schemes by comparing the resulting approximations they produce with a ground truth.
+For comparing distributions we use \(L^p\) \emph{norms} and the Kolmogorov-Smirnov distance.
+The \(L^p\) norm of a function \(f\) is given by
+\[\left(\int |f(x)|^p\wrt x\right)^{1/p}.\]
+The Kolmogorov-Smirnov distance between two distribution functions, \(F_1\), \(F_2\) is 
+\[\sup_{x}|F_1(x)-F_2(x)|.\]
+
 \section{More on relevant literature \& the context of the thesis} 
@@
 
-\paragraph{Transient analysis of fluid queues} The analysis of fluid queues can be classified broadly into two approaches, matrix-analytic methods (for example, \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}) and differential equations based approaches (such as \cite{anick1982,kk1995,blnos2022}).
+\paragraph{Transient analysis of fluid queues} The analysis of fluid queues can be classified broadly into three approaches, matrix-analytic methods (for example, \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}), differential equations based approaches (such as \cite{anick1982,kk1995,blnos2022}) and analyses based on recurrence relations (for example, \cite{sericola1998,sericola1999,sericola2001}).
 Often, it is the limiting distribution of the fluid queue which is of interest.

### 2022-08-05 chapter2/chapter.tex
ch 7/8 comments

@@
 	\begin{minipage}{0.8\textwidth}
-		\textit{Apart from Sections~\ref{sec: other applications}-\ref{sec: slope limiting}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notation, so that this chapter is consistent with the rest of the thesis.
+		\textit{Apart from Sections~\ref{sec: other applications} and \ref{sec: limiting and linearity}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notation, so that this chapter is consistent with the rest of the thesis.
 I am a co-author of the paper \cite{blnos2022}.
@@
 Discontinuous Galerkin (DG) methods can be used to approximate the solutions to systems of partial differential equations (PDEs).
-For a most readable and practical text on DG methods we refer the reader to \cite{nodalDGBook}.
 Here we describe how one can apply the DG method to approximate fluid queues and fluid-fluid queues.
-In finite-element methods, such as the DG method, the domain of approximation is partitioned into intervals, referred to individually as \textit{cells} and collectively as a \textit{mesh}.
-On each cell we construct a finite-dimensional smooth Sobolev space using piecewise-polynomial basis functions, and then project the partial differential equations onto this space.
+In finite-element methods we project the partial differential equations onto a set of piecewise polynomial functions.
 This projection leads to a new system of equations which represent a \textit{weak form} of the original system of PDEs.
-Next, we approximate the \textit{flux operator} which moves probability from one cell to another, in a manner similar to the underlying principle of a finite-volume approximation.
+Next, we approximate the \textit{flux operator} which moves probability from one cell to another.
 This creates a system of ordinary differential equations which is known as a \emph{semi-discrete system} (discrete in the spatial variable but continuous in time).
@@
 Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{p_k}\) to be the space of polynomials of degree \(p_k-1\).
-It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{p_k}\) as a basis of Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature points, since our approximations inherit nice properties from this \citep{nodalDGBook}.
-However, the constructions presented here are general, and any basis can be used.
-For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials.
+For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials, but any polynomial basis can be used.
 On each cell \(\mathcal D_{k,i}\) we approximate
@@
 We refer to \(u_{k,i}\) as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_{k,i}\) on the whole domain.
-The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\mathbb Z\).
+%The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\mathbb Z\).
 
@@
 Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE.
-For a sufficiently rich set of test functions \(W\) the weak and strong forms of the PDE are equivalent.
 Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE.
@@
 
-For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\), the same as the set of basis functions of our solution space.
+For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\) (the direct sum of \(W^k\)), the same as the set of basis functions of our solution space.
 Proceeding as described above, the weak formulation is
@@
 \section{Mass, stiffness, and flux matrices}
-For \(k\in\mathbb Z\), define local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
+Recall, for \(k\in\mathbb Z\), we defined local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
 \[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x.\]
@@
 \end{align}
-where \(\bs P_{-,+} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{+}}\) and \(\bs P_{-,{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\).
+where \(\bs P_{-,+} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{+}}\), \(\bs P_{-,{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\) and \(p_{ij}^{-1}\) is the probability of instantaneous transition of the phase process from Phase \(i\) to Phase \(j\) upon the process hitting the lower boundary.
 Equation~(\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(\bs T_{{-1},{-1}}\), and that the flux of probability density into the point masses is \(\boldsymbol f_-(0,t)\bs P_{-,{-1}}\bs C_{-1}\).
@@
 \end{align*}
-where \(\bs P_{+,-} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{-}}\) and \(\bs P_{+,K+1} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{K+1}}\). Using the same arguments as above, 
+where \(\bs P_{+,-} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{-}}\), \(\bs P_{+,K+1} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{K+1}}\) and \(p_{ij}^{K+1}\) is the probability of instantaneous transition of the phase process from Phase \(i\) to Phase \(j\) upon the process hitting the lower boundary.
+Using the same arguments as above,
 \begin{align*}
@@
 % APPLICATION TO INTERGRATING OVER TIME
-\section{Time-integration schemes}\label{sec: time integration}
-In this thesis we numerically integrate ODEs of the form
-\begin{align}\label{eqn: DG ODE w BCs 3}
-	\cfrac{\wrt}{\wrt t} \boldsymbol a(t)
-	% 
-	= \boldsymbol a(t) \bs Q
-\end{align}
-where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix, given an initial condition \(\bs a(0)\).
-To do so, we employ the strong stability preserving Runge-Kutta (SSPRK) scheme of order 4 with 5 stages \citep{sr2002}.
-SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``...can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
-
-In our context, the SSPRK method with \(s\) stages has the form 
-\begin{align*}
-	&\bs v^{(0)} = \bs a(t),\\
-	%
-	&\bs v^{(\ell)} = \sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q,\, \ell = 1,\dots,s,\\
-	%
-	&\bs a(t+h) =\bs v^{(s)},
-\end{align*}
-where \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are coefficients which define the scheme and \(h\) is the \(t\)-step size.
-To ensure the stability-preserving property the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are chosen to be positive so that \(\bs v^{(s)}\) is a convex combination of forward-Euler steps.
-Moreover, the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are optimised so that the allowable \(t\)-step size is as large as possible.
-The maximum allowable \(t\)-step size is \(h_{RK}\), and we require
-\[h_{RK} \leq \min_{\ell k}\cfrac{\alpha_{\ell k}}{\beta_{\ell k}}h_E,\]
-where \(h_E\) is the maximum allowable \(t\)-step size for the forward-Euler scheme.
-
-The \(t\)-step size needs to be chosen to be sufficiently small so that the time integration is stable. Here, for an order \(p\) DG scheme we choose the time step to be less than 
-\[\max_{i\in\calS}|c_i|\min_{r\in{2,...,p+1}}\left(z_{r}-z_{r-1}\right) \cfrac{\min_{k}{\Delta}_k}{2}\]
-where \(z_{r-1}\leq z_r\), \(z_1=-1\), \(z_{p+1}=1\) and \(z_r, r=2,...,p,\) are the \(p-1\) zeros of the first derivative of \(P_{p}(x)\), the order \(p\) Legendre polynomial.
-This follows advice from \cite{nodalDGBook}.
-
-\section{Slope limiters}\label{sec: slope limiting}
-A well-know problem with DG schemes is that, in the presence of steep gradients or discontinuities, spurious oscillations in the approximate solution can occur and positivity is not guaranteed (see \cite{nodalDGBook}~Section~5.6, and \cite{koltai2011}~Section~3.3, for example).
-To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
-Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
-However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
-As a result, the limiter may unnecessarily reduce the accuracy of the scheme in the presence of natural oscillations (see \citep{nodalDGBook},~Example~5.8).
-Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear.
-
-We now describe a simple slope limiter known as the \emph{Generalised MUSCL} scheme \citep{c99} (see also \citep{nodalDGBook},~Section~5.6.1). Define the \emph{minmod} function 
-\begin{align*}
-	m(a_1,a_2,a_3)=\begin{cases}
-		s\min_{i\in\{1,2,3\}}|a_i|, & |s|=1, \\
-		0, &\mbox{otherwise},
-	\end{cases}
-\end{align*}
-where \(s=\frac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
-When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
-In the context of DG the three arguments \(a_1,\,a_2\) and \(a_3\) are the estimates of the average gradient of the approximate solution near a given cell.
-When the signs of \(a_1,\,a_2\) and \(a_3\) differ, this suggests an oscillation in the approximate solution.
-
-Now, define \(\overline u_{k}\) as the average value of the approximate solution on cell \(k\), then the slope limited solution on cell \(k\) is the linear approximant 
-\begin{align*}
-	\Pi_{k}^{lim}(u_k)=\overline u_{k} + (x-\overline y_k)m\left(\Pi^1_{\partial x}u_k,\cfrac{\overline u_{k+1}-\overline u_{k}}{\Delta_k},\cfrac{\overline u_{k}-\overline u_{k-1}}{\Delta_k}\right),
-\end{align*}
-where \(\Pi^1_{\partial x}u_k\) is the slope of the linear projection of \(u_k\) on cell \(k\) and \(\overline y_k=(y_k+y_{k+1})/2\) is the centre of the \(k\)th cell.
-
-In non-oscillatory regions of the solution the slope limiter is unnecessary and reduces accuracy, so we only apply the limiter to cells \(k\) where oscillations are detected.
-To determine which cells need limiting we compute
-\begin{align*}
-	v_k^L &:= \overline{u}_{k} - m(\overline{u}_{k} - {u}_{k}^L,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
-	\\v_k^R &:= \overline{u}_{k} + m(\overline{u}_{k} - {u}_{k}^R,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
-\end{align*}
-where \({u}_{k}^L\) and \({u}_{k}^R\) are the values of the approximate solution on cell \(k\) evaluated at the left-hand and right-hand edges of the cell.
-We apply the slope limiter to cell \(k\) if \(v_k^L\neq u_k^L\) or \(v_k^R\neq u_k^R\), otherwise the approximation on cell \(k\) is left unaltered.
-
-% Notice that the process of limiting is essentially a post-processing of the approximate solution after applying
-
-% In the context of approximating the first return operator of a fluid-fluid queue there is no direct equivalent of slope limiting, other than, perhaps, to re-compute the solution with constant basis function if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
-
-\subsection{Slope limiters and time integration}\label{subsec: slope lim and int}
-To use slope limiters within the SSPRK scheme we first project the initial condition on to \(W\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
-To find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the (limited) initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation~(\ref{eqn: DG ODE w BCs 3}).
-At each stage of the time-integration we apply the slope limiter, i.e.~the scheme above with a slope limiter is
-\begin{align*}
-	&\bs v^{(0)} = \bs a(t),\\
-	%
-	&\bs v^{(\ell)} = \Pi^{lim}\left(\sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q\right),\, \ell = 1,\dots,s,\\
-	%
-	&\bs a(t+h) =\bs v^{(s)},
-\end{align*} 
-where \(\Pi^{lim}\) is the slope limiter function which determines on which cells the solution needs limiting, and if so, applies the limiter \(\Pi_k^{lim}\).
-
-We refer to this (vanilla) application of a limiter to the DG scheme as the DG-lim scheme.
-
-\paragraph{Consequences of slope limiting} We have already mentioned that the Generalised MUSCL slope limiter reduces the order to linear in regions where oscillatory solutions are detected.
-When slope limiting is used in conjunction with a time-integration scheme the reduction in order may not be local in time and may persist past the time when the limiter first detected oscillations.
-For example, consider a problem with an initial condition which introduces oscillations into the numerical approximation.
-If the slope limiter is applied to the initial condition then, in the region around the oscillations, the approximation to the initial condition will be linear.
-Even if the slope limiter is not applied (or not needed) after this point, the initial condition has still been altered, perhaps significantly, from the original approximation and this can affect the accuracy of transient approximations.
-This is not to say that the slope-limited-regions of the approximation will \emph{always} remain linear.
-When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of generalised slope limiter described.
-Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy of subsequent computations.
-
-\subsection{Another slope limiting scheme}\label{sec: anoth pos pres}
-The advantage of the approach above is that, in areas of the approximation where the solution is smooth, then the high-order accuracy of the method can be retained, while maintaining positivity where necessary.
-However, if the problem at hand is dominated by discontinuities and/or if there are no regions of the solution of interest which are smooth, then there is no real advantage in this approach as the limiter will reduce the approximate solution to linear in all regions of interest.
-Hence, we may as well just use a linear scheme and save on some computation.
-
-Perhaps a better approach would be to use a linear scheme with a smaller cell width such that the computational work remains approximately the same.
-For example, for a DG scheme with \(2p\) basis functions on a cell, \(\mathcal D_k\), of width \(\Delta\), say, we construct block matrices of dimension \(2p\times 2p\) (i.e.~the mass, stiffness and flux matrices, \(\bs M_k\), \(\bs G_k\), and \(\bs F^{k\ell}\), respectively) and compute the approximate solution using these matrices.
-Alternatively, consider breaking the cell, \(\mathcal D_k\), up into \(p\) sub-cells of width \(\Delta/p\) and using a DG scheme with \(2\) basis functions on each sub-cell (i.e.~a linear basis on each cell).
-On each of the \(p\) sub-cells we construct block matrices of size \(2\times 2\).
-To approximate the solution on the original interval \(\mathcal D_k\) we form the \(p\) approximations on each sub-cell into larger matrices which are of dimension \(2p\times 2p\), and then use in the same way as before.
-Thus, the scheme with \(2p\) basis functions on the cell \(\calD_k\), and the scheme with a 2 basis functions on each of the \(p\) sub-cell have approximately the same computational cost (the latter scheme results in a sparser matrix).
-
-If we know \emph{a priori} that we will apply the slope limiter to cell \(\calD_k\), then intuitively we suspect that the latter scheme may be far more accurate; it will approximate the solution on \(\calD_k\) by a piecewise linear function with \(p\) pieces, whereas the DG scheme with \(2p\) basis functions on cell \(\calD_k\) will, after limiting, represent the solution as a single linear function on \(\calD_k\).
-
-We refer to this scheme as the DG-lin-lim scheme, as it is a linear DG scheme with a limiter.
-
-\subsection{Slope limiters and linear operators}\label{sec: limiting and linearity}
-The process of slope limiting, as described above, essentially amounts to a non-linear post-processing of the approximate solution after applying the operator \(\bs B\).
+\section{Remarks on slope limiters, linear operators and application to fluid-fluid queues}\label{sec: limiting and linearity}
+The process of slope limiting, as described in Section~\ref{sec: slope limiting}, essentially amounts to a non-linear post-processing of the approximate solution after applying the operator \(\bs B\).
 Thus, the slope-limited scheme arrives at an approximation which is the action of a non-linear operator applied to the initial condition.
-The fact that the approximate operator is non-linear makes solving the Riccati equation~\ref{eqn:RiccatiPsi} difficult/impossible, whereas there are efficient techniques to solve matrix-Riccati equations \cite{bot08}.
+The fact that the approximate operator is non-linear makes solving the Riccati Equation~(\ref{eqn:RiccatiPsi}) difficult/impossible, whereas there are efficient techniques to solve matrix-Riccati equations \cite{bot08}.
 Thus, for the application to fluid-fluid queues slope limiting is not a viable option.
@@
 %
-
-\subsection{Briefly, on accuracy}
-As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method for a similar kind of problem to the one under study here.
-For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
-Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
-
-% To quote Section~5.7 of \cite{nodalDGBook} SSPRK methods ``can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
-
-% why the MUSCL limiter?
-%   it requires no tuning and guarentees non-negativity
-%   some other positivity-preserving methods require tuning and can still produce negative results if not tuned properly
-%   ^ same goes for filtering

### 2022-08-05 chapter6/chapter.tex
ch 7/8 comments

@@
 
-Lastly, for the DG scheme I know of no rigorous argument which shows that we can use the DG approximation to the generator of a fluid queue to approximate performance measures of fluid-fluid queues, such as the operator \(\mathbb \Psi\).
-Here, we rely on the intuition that substituting the DG approximation for the true operator \emph{should} give reasonable results, and further, \emph{assume} that the resulting approximations are polynomial.
-For example this means that, other than being the solution to a matrix Riccati equation, the DG approximations to \(\mathbb \Psi\) that we compute in this chapter are not actually rigorously defined.
-
-On the other hand, using the stochastic approximations (QBD-RAP of uniformisation), results in rigorously defined operators as proved in \cite{bgnp2021}.
-
-In summary, the QBD-RAP and uniformisation schemes are the only two schemes which result in rigorously defined approximations, preserve the axioms of probability, and, of the two, the numerical experiments demonstrate that the QBD-RAP methods is superior.
+Although we have been unable to supply a rigorous argument to show that we can use the DG approximation to the generator of a fluid queue to approximate performance measures of fluid-fluid queues, such as the operator \(\mathbb \Psi\), we can rely on the intuition that substituting the DG approximation for the true operator \emph{should} give reasonable results, and further, \emph{assume} that the resulting approximations are polynomial.
+
+On the other hand, using the stochastic approximations (QBD-RAP or uniformisation), results in rigorously defined operators as proved in \cite{bgnp2021}.
+
+In summary, the QBD-RAP and uniformisation schemes are the only two schemes which result in rigorously defined approximations, preserve the axioms of probability, and, of the two, the numerical experiments demonstrate that the QBD-RAP method is superior.
 
@@
 
-The main focus of the numerical experiments is on the error as the number of basis functions per cell is increase.
+The main focus of the numerical experiments is on the error as the number of basis functions per cell is increased.
 In the numerical experiments the number of basis functions used to approximate the solutions is the same on all cells and the number of basis functions on a cell (and hence all cells) is referred to as the \emph{dimension} of the scheme.
-For the discontinuous Galerkin scheme, increasing the dimension of the scheme means increasing the number of polynomial basis functions used to approximate the solution within each cell.
+For the discontinuous Galerkin scheme, the dimension of the scheme means is the number of polynomial basis functions used to approximate the solution within each cell.
 For example, if we use 3 basis functions in the discontinuous Galerkin scheme, we approximate the solution by 3 linearly independent quadratics on each cell and hence the approximation is a quadratic on each cell.
-For the QBD-RAP scheme, increasing the dimension of the scheme means increasing the order of the CME distribution used to construct the scheme.
+For the QBD-RAP scheme, the dimension of the scheme is the order of the CME distribution used to construct the scheme.
 To make a comparable equivalent for the uniformisation scheme, we divide each cell into smaller sub-cells over which we approximate the solution.
@@
 
-We implement the discontinuous Galerkin scheme with and without a slope-limiter (\cite{c99}, \cite{nodalDGBook}~Section~5.6.2, see also Section~\ref{sec: slope limiting}).
-The slope limiter stops oscillations appearing in the approximations and hence also avoids non-probabilistic approximations to PDFs and CDFs.
 We investigate two slope-limited schemes.
@@
 In the \emph{DG-lin-lim} scheme, we follow the construction described in Section~\ref{sec: anoth pos pres}, which is effectively a DG scheme with cells of width \(2\Delta/(p+1)\), for \(p=1,3,5,...,21\), and a basis of dimension \(2\) (linear) on each cell, with an upwind flux, and we apply the Generalised MUSCL limiter.
-Thus, the DG-lin-lim scheme uses \(p+1\), \(p=1,3,5,...,21\), basis functions to represent the solution on each of the original cells \(\calD_k\), and we define the dimension of the scheme to be \(p+1\).
+Thus, the DG-lin-lim scheme uses \(p+1\) basis functions to represent the solution on each of the original cells \(\calD_k\), and we define the dimension of the scheme to be \(p+1\).
 In the presence of discontinuities, slope-limiting forces the approximation to be linear, i.e.~an order~1 polynomial, around the discontinuity.
@@
 However, there are possibly better choices for this post-processing, such as filtering, \citep[Section~5.6.1]{nodalDGBook}.
-One could also view the composition of the DG operator and the limiting operator as a one, but the result is a non-linear operator which is of no use for simplifying the Riccati Equation which characterises \(\mathbb \Psi\).
-Considering there is no theory on the use of the DG scheme to approximate \(\mathbb \Psi\) and hence the resultant approximation is not rigorously-defined, it is not known if the use of a limiter in this context is possible.
+One could also view the composition of the DG operator and the limiting operator as one, but the result is a non-linear operator which is of no use for simplifying the Riccati Equation which characterises \(\mathbb \Psi\).
+Considering there is no theory on the use of the DG scheme to approximate \(\mathbb \Psi\) and hence the resultant approximation is not rigorously-defined, the author does not know if the use of a limiter in this context is possible.
 
@@
 Instead, we fix the time-integration step size for each numerical experiment at a suitably small value to obey a certain stability criterion (a CFL-like condition, \citep[Section~4.8]{nodalDGBook}).
-Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (\cite{sr2002}, \cite{nodalDGBook}~Section~5.7, see also Section~\ref{sec: time integration}), which claims to introduce no more oscillations into the solution as we integrate over time.
+Moreover, we always implement the strong stability preserving Runge-Kutta method of order 4 with 5 stages (\cite{sr2002}, \cite{nodalDGBook}~Section~5.7, see also Section~\ref{sec: time integration}), which is claimed to introduce no more oscillations into the solution as we integrate over time.
 
@@
 Section~\ref{sec:stat} investigates approximations to the limiting distribution of the queue, Section~\ref{sec: transient approx} investigates approximations to the transient distribution of the queue for two initial conditions, and Section~\ref{sec: hit approx} investigates approximations the first hitting time of the fluid level on the set \(\{0,1\}\).
-Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num}, compute approximations to \(\mathbb \Psi\) and ultimately, approximate the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\).
+Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num}, compute approximations to \(\mathbb \Psi\) and ultimately, approximate the distribution of \(\{(X(\zeta_{Y}(\{0\})),\varphi(\zeta_{Y}(\{0\})))\}\) at time \(\zeta_{Y}(\{0\})\) at which \(\{Y(t)\}\) first returns to \(0\).
 
@@
 \subsection{QBD-RAP closing operators}\label{sec: closing vecs numerics}
-The three closing operators we investigate are the following. The unnormalised closing vector, 
+We investigate the three closing operators introduced in \ref{sec: closing}.
+For convenience, we restate the three closing operators from Section~\ref{sec: closing} here.
+The unnormalised closing vector,
 \begin{align}
@@
 
-	Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the non-linear normalised and normalised reconstructions perform very similarly -- this is the case throughout much of this section.
-The difference between the non-linear normalised and normalised reconstructions is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward).
+	Figure~\ref{fig: fun 2 ks error qbdrap closing vecs} also shows that the non-linear normalised and normalised closing operators perform very similarly -- this is the case throughout much of this section.
+The difference between the non-linear normalised and normalised closing operators is how they treat mass in the tail of a matrix exponential distribution (from \(2\Delta\) onward).
 Intuitively, there should be very little mass in the tail of the distribution as we are using concentrated matrix exponential distributions in the reconstruction.
@@
 
-Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised (\ref{eqn:546}) and normalised (\ref{eqn:5462}) closing operators for dimension 1, 3, 5 and 7, as well as the true PDF.
+Some insight is gained by looking at Figure~\ref{fig: pdf reconstructed} where we plot the reconstructed PDFs for the unnormalised (\ref{eqn: alal}) and normalised (\ref{eqn:5462}) closing operators for dimension 1, 3, 5 and 7, as well as the true PDF.
 Observing Figure~\ref{fig: pdf reconstructed} notice that the unnormalised reconstruction fails to capture the density at the left of each of the plots.
@@
 %\exampleFloatBarrier
-Figure~\ref{fig: pdf reconstructed} also show that both closing operators fail to approximate the initial condition well at the right-hand side of the interval.
+Figure~\ref{fig: pdf reconstructed} also shows that both closing operators fail to approximate the initial condition well at the right-hand side of the interval.
 A potential area for further research is to determine if there is a different reconstruction method or another closing operator which could alleviate this issue.
@@
 	Next we consider approximating the initial distribution with density \(-6x^2+6x\) on \([0,1)\).
-Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the KS error against the dimension of the reconstruction for the three closing operators, we once again see that the reconstruction using the unnormalised closing operator performs the worst, while the performance of the two normalised reconstructions is indistinguishable.
+Observing the left-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the KS error against the dimension of the matrix exponential approximation for the three closing operators, we once again see that the unnormalised closing operator performs the worst, while the performance of the two normalised closing operators is indistinguishable.
 However, if we instead look at the right-hand panel of Figure~\ref{fig: fun 6 ks error qbdrap closing vecs} which plots the \(L^2\) error between the reconstructed PDF and the true PDF, then the unnormalised closing operator performs better than the two normalised ones for dimension 5 and above.
@@
 The fact that the unnormalised closing operator outperforms the two normalised ones can be explained by the `folding' of the normalised operators around \(\Delta\).
-In Figure~\ref{fig: pdf reconstructed quadratic} we plot the unnormalised and normalised reconstructions along with the true PDF, \(-6x^2+6x\).
-In Figure~\ref{fig: pdf reconstructed quadratic} at the left of the plots, we observe that the normalised reconstructions over-estimate the density function, whereas the unnormalised reconstruction looks to be doing better.
-The `folding' in the normalised closing operator manifests itself as more mass at the left of these plots compared to the unnormalised reconstructions.
+In Figure~\ref{fig: pdf reconstructed quadratic} we plot the reconstructions obtained using the unnormalised and normalised closing operators, along with the true PDF, \(-6x^2+6x\).
+In Figure~\ref{fig: pdf reconstructed quadratic} at the left of the plots, we observe that the normalised closing operator over-estimate the density function, whereas the unnormalised closing operator looks to be doing better.
+The `folding' in the normalised closing operator manifests itself as more mass at the left of these plots compared to the unnormalised closing operator.
 \end{example}
@@
 Considering what we have learnt so far about the unnormalised operator underestimating in this region, we expect that the unnormalised closing operator will perform relatively poorly in this case.
-Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that this is the case; the normalised reconstructions perform relatively well compared to the unnormalised reconstruction, as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF).
+Indeed, observing Figures~\ref{fig: fun 7 ks error qbdrap closing vecs} we see that this is the case; the normalised closing operators perform relatively well compared to the unnormalised closing operator, as measured by both error metrics (KS statistic and \(L^2\) norm on the PDF).
 Plots of the PDFs (omitted), confirm that this is due to the loss of mass at the left-hand side of the region due to the truncation.
@@
 Table~\ref{table: error rates for QBD-RAP closing} summarises the rates of convergence of the \(L^2\) error on the PDFs for unnormalised and normalised closing operators in the examples above (i.e.~the slopes of the black dotted lines in the error plots).
-In general, the normalised closing operators converge at a rate approximately as fast, or faster, than the unnormalised reconstruction.
+In general, the errors of the approximations obtained using normalised closing operators converge at a rate approximately as fast, or faster, than the approximations obtained using the unnormalised closing operator.
 \begin{table}
@@
 Here, the uniformisation scheme appears to perform the best, while the QBD-RAP scheme performs the worst.
-The rate of convergence of the QBD-RAP scheme appears to be similar to the rate of convergence of the DG scheme.
 
@@
 %\exampleFloatBarrier
-In Figure~\ref{fig: pdf comp fun 1} we plot approximated CDFs from the DG, uniformisation and QBD-RAP schemes alongside the true CDF.
+In Figure~\ref{fig: pdf comp fun 1} we plot approximate CDFs from the DG, uniformisation and QBD-RAP schemes alongside the true CDF.
 The DG scheme displays undesirable features for an approximation to a CDF -- it is not monotonically increasing and is not constrained to the bounds of probability, \([0,1]\).
@@
 	%\exampleFloatBarrier
-	Now consider approximating the initial distribution with PDF \(1(x\leq 0.5)\).
+	Now consider approximating the initial distribution with PDF \(2\times 1(x\leq 0.5)\).
 Figure~\ref{fig: fun 2 comp} plots the KS error (left) and \(L^2\) error between the true and approximate PDFs (right).
@@
 At the other extreme we now consider an initial distribution with density \(-6x^2+6x\) on \([0,1)\).
-In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDFs.
+In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximate PDFs.
 Since the DG method projects the initial condition onto a polynomial basis, then, for a dimension 3 approximation and above, the DG scheme can represent the initial condition exactly.
@@
 
-We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space and discretise the interval \([0,10]\) into \(10\) cells, each of width 1.
+We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space and discretising the interval \([0,10]\) into \(10\) cells, each of width 1.
 We use $10,001$ points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which to evaluate the CDFs to approximate the KS metric.
@@
 \end{figure}
-Figure~\ref{fig: fun 4 wave} plots the KS error between the approximated and true CDFs and \(L^2\) error between the approximated and true PDFs.
+Figure~\ref{fig: fun 4 wave} plots the KS error between the approximate and true CDFs and \(L^2\) error between the approximate and true PDFs.
 Comparing Figure~\ref{fig: fun 4 wave cp} with Figure~\ref{fig: fun 4 wave}, each scheme seems to show a similar rate of decay for these two error metrics as they did for the cell-wise errors.
@@
 	\end{figure}
-	Since the discontinuity at \(t=4,\,x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG scheme will perform well for this example.
+	Since the magnitude of the discontinuity at \(t=4,\,x=4\) is small (much smaller than numerical integration errors in the evaluation of the error metrics), and the distribution is otherwise smooth, we expect that the DG scheme will perform well for this example.
 Figures~\ref{fig: fun 6 wave cp} and~\ref{fig: fun 6 wave} confirms that this is indeed the case.
@@
 This is characteristic of the DG scheme for the smooth problems we investigate throughout this chapter.
-For low order DG schemes there are regions where the approximated PDF is negative, however, as the order of the DG scheme increases, these regions disappear.
+For low order DG schemes there are regions where the approximate PDF is negative, however, as the order of the DG scheme increases, these regions disappear.
 
@@
 
-	Interestingly, this solution is discontinuous at a cell-edge and even though the DG scheme\footnote{All the schemes considered allow discontinuities at cell-egdes.} allows for discontinuities in the PDF at cell edges, it is unable to capture this discontinuity without producing oscillations.
+	Interestingly, this solution is discontinuous at a cell-edge and even though the DG scheme\footnote{All the schemes considered allow discontinuities at a cell-edges.} allows for discontinuities in the PDF at cell edges, it is unable to capture this discontinuity without producing oscillations.
 \end{example}
@@
 Figure~\ref{fig: reflecting transient pm} (right) plots the cell-wise error metric (\ref{eqn: cell errors 2}) for Model~\ref{model: simple} with the point mass initial condition.
-Figure~\ref{fig: reflecting transient pm} (right) suggests that approximating the cell-wise error seems to be a difficult problem.
+Figure~\ref{fig: reflecting transient pm} (right) suggests that approximating the cell-wise probabilities seems to be a difficult problem.
 This is likely caused by the discontinuity at \(x=2\) in Phase \(1\), which lies exactly on a cell boundary.
@@
 Once again, we use simulation and bootstrapping to approximate the true distribution.
-The plot (not shown) of the \(L^1\) error between the CDFs for the model at times \(t=2\) and \(t=2.1\) is relatively similar to the one show and not noteworthy.
+The plot (not shown) of the \(L^1\) error between the CDFs for the model at times \(t=2\) and \(t=2.1\) is relatively similar to the one shown and not noteworthy.
 The plot of the cell-wise error metric at time \(t=2.1\) in Figure~\ref{fig: reflecting transient pm t2.1} is somewhat more interesting.
@@
 To account for Monte-Carlo errors we take 1,000 bootstrap resamples of the original \(5\times 10^{10}\) samples and compute the empirical CDF of the hitting times for each bootstrap sample.
-For each bootstrap sample, we resample \(5\times 10^{10}\) points with replacement from the original \(5\times 10^{10}\) realisations, then compute the error metrics between the empirical and approximated CDFs and, ultimately, estimate the 5th and 95th percentile of the distribution of the errors.
+For each bootstrap sample, we resample \(5\times 10^{10}\) points with replacement from the original \(5\times 10^{10}\) realisations, then compute the error metrics between the empirical and approximate CDFs and, ultimately, estimate the 5th and 95th percentile of the distribution of the errors.
 
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/exp/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the exponential initial condition.
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the exponential initial condition.
 The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/hitting_times_model/hitting_times/point_mass/l1_cdf_error_formatted.pdf}
-	% \caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
+	% \caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
 The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
@@
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
-	\caption{\(L^1\) errors between the simulated and approximated first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
+	\caption{\(L^1\) errors between the simulated and approximate first hitting time CDFs in (\ref{eqn: hit cdf}) for Model~\ref{model: simple} with the point mass initial condition.
 The approximations were obtained via the DG (blue solid line), DG-lim (orange dashed line), uniformisation (green dashed line), QBD-RAP (purple dotted line) and DG-lin-lim (gold dashed line) schemes.
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) in (\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}.
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate CDFs of \(X(\zeta_{Y}(\{0\}))\) in (\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}.
 The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cell_probs_error_formatted.pdf}%
-	\caption{Cell-wise errors between the simulated and approximated probabilities of \(X(\zeta_{Y}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}.
+	\caption{Cell-wise errors between the simulated and approximate probabilities of \(X(\zeta_{Y}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}.
 The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
@@
 
-By modifying slightly Model~\ref{model: ffq}, we can construct a first return distribution which is discontinuous. 
+By modifying Model~\ref{model: ffq} slightly, we can construct a first return distribution which is discontinuous. 
 \begin{model}\label{model: ffq2}
@@
 Moreover, for \(u\in[0,0.25]\), \(X(u)\geq x^*\) and \(r_{00}(X(u))=\theta_2=1>0\), and for \(u\in(0.25,0.5]\), \(X(u)\in(0,x^*)\) and \(r_{00}(X(u))=-\theta_2=-1<0\).
-Hence, for \(u\leq 0.5\) the second fluid level is \(Y(u)=t1(t\leq 0.25)+(0.25-t)1(t>0.25)\) and at time \(t=0.5\), the second fluid level first returns to \(0\), i.e.~\(\zeta_{Y}(\{0\})=0.5\) on this sample path.
+Hence, for \(u\leq 0.5\) the second fluid level is \(Y(u)=u1(u\leq 0.25)+(0.25-u)1(u>0.25)\) and at time \(u=0.5\), the second fluid level first returns to \(0\), i.e.~\(\zeta_{Y}(\{0\})=0.5\) on this sample path.
 The probability of this sample path is \(e^{-(\beta_1+\beta_2)\times 0.5}\).
-Thus, in the distribution of the first return time of the second fluid level there is a point mass at \(t=0.5\) with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\).
+Thus, in the distribution of the first return time of the second fluid level there is a point mass at \(u=0.5\) with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\).
 Furthermore, on the same sample path \(X(\zeta_Y(\{0\}))=X(0.5)=1.2\), and hence there is a point mass with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\) in the distribution of \(X(\zeta_Y(\{0\}))\) also.
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximated CDFs of \(X(\zeta_{Y}(\{0\}))\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}.
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate CDFs of \(X(\zeta_{Y}(\{0\}))\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}.
 The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
@@
 Moreover, the uniformisation and QBD-RAP schemes have stochastic interpretations which can aid in their analysis.
-For the application to fluid-fluid queues in this chapter the stochastic interpretation is important as it justifies (and defines) the approximation to the first return operator.
-For the DG scheme there is no such theory (that I know of) and this means that the approximation to the first return operator using the DG scheme is not a rigorously defined object.
+For the application to fluid-fluid queues in this chapter the stochastic interpretation is important as it justifies (and defines) an approximation to the first return operator.
+For the DG scheme there is no such theory (that the author is aware of) and this means that the approximation to the first return operator using the DG scheme is not a rigorously defined object.
 On a similar note, the uniformisation and QBD-RAP schemes result in operators which are linear, whereas the DG-lim and DG-lin-lim schemes do not result in linear operators.

### 2022-08-05 chapter7/chapter.tex
ch 7/8 comments

@@
 Fluid queues have found a wide variety of applications including risk processes, telecommunications, and environmental modelling, among others.
-Given the success of fluid queues it is believed that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success.
+Given the success of fluid queues it is plausible that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success.
 \cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and limiting distribution of a fluid-fluid queue.
@@
 In Chapter~\ref{ch:galerkin} we introduced a discontinuous Galerkin (DG) scheme for the approximation of the generator of a fluid queue and show how to approximate the operators from \cite{bo2014}.
-High-order DG schemes are known to perform well for smooth problems, but problems can occur in the presence of discontinuities, such as negative probability estimates.
+High-order DG schemes are known to perform well for smooth problems, but problems, such as negative probability estimates, can occur in the presence of discontinuities.
 In some contexts, slope limiting can be used to rectify this, but they result in approximations to operators which are non-linear which complicates the computation of the operators in \cite{bo2014}.
@@
 The DG method is well known, but its application to fluid and fluid-fluid queues has not been well-studied.
-This thesis introduces the DG scheme in the context of fluid and fluid-fluid queues.
 The QBD-RAP scheme is a novel approach which overcomes some known issues of the DG scheme, and converges faster than the uniformisation scheme.

### 2022-08-05 intro/chapter.tex
ch 7/8 comments

@@
 However, the added flexibility comes at a cost to the in the analysis of the model.
+For both fluid-fluid queue and classical fluid queues, performance measures such has the stationary distribution, and the Laplace-transforms of hitting times and return times, can be derived in terms of the infinitesimal generator of the modulating process.
+For fluid queues the modulating process is a CTMC for which the infinitesimal generator is a matrix.
+Thus, expressions for the performance measures of fluid queues are matrix expressions which are readily computable.
+For fluid-fluid queue the modulating processes is a fluid queue for which the infinitesimal generator is a matrix of differential operators.
+Thus, expressions for the performance measures of fluid-fluid queues are matrices of differential operators which, in all but the simplest of cases, are not readily computable.
+
 The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).

### 2022-08-05 chapter1/chapter.tex
intro/ch2 edits

@@
 
-A stochastic process is \emph{stationary} if \(X(t_n),X(t_{n-1}),...,X(t_0)\) and \(X(t_n+t),X(t_{n-1}+t),...,X(t_0+t)\) have the same distribution for any \(t\). That is, 
+A stochastic process is \emph{stationary} if, for any \(n\), \(t_0<t_1<...<t_n\) and any \(t\), \(X(t_n),X(t_{n-1}),...,X(t_0)\) and \(X(t_n+t),X(t_{n-1}+t),...,X(t_0+t)\) have the same distribution.
+That is,
 \begin{align}\label{eqn: stationary}
@@
 \end{align}
-for all \(n\in\{0,1,2,...\}\) and \(i,i_0,...,i_{n-1},j\in\calS,\) and refer to (\ref{eqn: Markov property}) as the Markov property.
+for all \(n\in\{0,1,2,...\}\) and \(i,i_0,...,i_{n-1},j\in\calS,\) and referred to (\ref{eqn: Markov property}) as the \emph{Markov property}.
 The process \(\{X(n)\}_{n\in\{0,1,2,...\}}\) is said to be time-homogeneous if
@@
 \end{align}
-and the elements \(T_{ij}\) are the \emph{transition rates}. A CTMC is said to be \emph{conservative} if \(\bs P(t)\bs 1=\bs 1\) for all \(t\geq 0\) where \(\bs 1\) is a column vector of ones.
+and the elements \(T_{ij}\) are the \emph{transition rates}. A CTMC is said to be \emph{honest} if \(\bs P(t)\bs 1=\bs 1\) for all \(t\geq 0\) where \(\bs 1\) is a column vector of ones.
 
@@
 Normed means that a norm is defined on \(L\).
-Intuitively, a norm is a function which tells us the distance between two points in \(L\).
+Intuitively, a norm is a function which tells us the size of a vector in \(L\), hence we can use a norm to compute a distance between two points \(f,g\in L\) by computing the norm of \(f-g\).
 
@@
 We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms defined on \(V\) and \(W\), respectively.
-An operator is said to be bounded if \(||\mathbb B||<M\) for some \(M<\infty, M\in \mathbb R\), for every \(v\in V\).
+An operator is said to be bounded if \(||\mathbb B||<M\) for some \(M<\infty\).
 
@@
 A family of operators \(\{\mathbb T(t)\}_{t\geq 0}\), \(\mathbb T(t):L\to L\), is said to have the \emph{semigroup} property if \(\mathbb T(t+s)=\mathbb T(t)\mathbb T(s)\).
-If \(\mathbb T(0)=I\), the identity operator, then the family \(\{\mathbb T(t)_{t\geq 0}\) is known as an operator semigroup.
-If \(f\in L,\, \mathbb T(t)f\to f\) as \(t\to 0^+\), then \(\{\mathbb T(t)\}_{t\geq 0}\) is said to be strongly continuous.
+If \(\mathbb T(0)=I\), the identity operator, then the family \(\{\mathbb T(t)_{t\geq 0}\}\) is known as an operator semigroup.
+If, for all \(f\in L,\, \mathbb T(t)f\to f\) as \(t\to 0^+\), then \(\{\mathbb T(t)\}_{t\geq 0}\) is said to be strongly continuous.
 If \(||\mathbb T(t)||\leq 1\) for all \(t\geq 0\) then \(\{\mathbb T(t)\}_{t\geq 0}\) is known as a contraction semigroup.
 
-For example, for a bounded linear operator \(\mathbb B:\mathcal D(\mathbb B)\to \mathcal D(\mathbb B)\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\) where, for \(f\in \mathcal D(\mathbb B)\), \(\mathbb B^kf\) is defined recursively by \(\mathbb B^k f = \mathbb B^{k-1} (\mathbb B f)\).
-It can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), hence \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup.
+For a bounded linear operator \(\mathbb B:\mathcal D(\mathbb B)\to \mathcal D(\mathbb B)\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\) where, for \(f\in \mathcal D(\mathbb B)\), \(\mathbb B^kf\) is defined recursively by \(\mathbb B^k f = \mathbb B^{k-1} (\mathbb B f)\).
+It can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), which implies that \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup.
 As another example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega(\mathbb R) \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\).
-Then \(\mathbb T(t)f(x)=\sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is the Taylor series of \(f\) about the point \(x\) and the operator \(\mathbb T(t)f(x)=f(x-t)\) is the shift operator.
+Then \(\mathbb T(t)f(x)=\sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is the Taylor series of \(f\) about the point \(x\), and we see that the operator \(\mathbb T(t)f(x)=f(x-t)\) is the shift operator.
 The semigroup property follows since
@@
 \[\mathbb E[f(X(t))\mid X(0)=x]=\int_{y\in\calS} f(y) P(X(t)\in\wrt y\mid X(0)=x),\]
-where \(f\) is some bounded function.
-Expectation is an operator which is acting on a function \(f\).
+where \(f<F\) is some bounded function.
+We can think of expectation as an operator which is acting on the function \(f\).
 The expectation depends on \(t\) and \(x\) as these define the distribution with which we are taking the expectation, hence, let us write
@@
 
-By the Markov property, for some \(s\in[0,t]\), 
-\begin{align*}
-	&\int_{y\in\calS} f(y) P(X(t)\in\wrt y \mid X(0)=x)
+By the Markov property, for any \(s\in[0,t]\), 
+\begin{align*}
+	\mathbb T(t)f(x)&=\int_{y\in\calS} f(y) P(X(t)\in\wrt y \mid X(0)=x)
 	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}f(y) P(X(t)\in\wrt y\mid X(s)=z) P(X(s)\in\wrt z\mid X(0)=x)
-	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}\mathbb E[f(X(t))\mid X(s)=z] P(X(s)\in\wrt z\mid X(0)=x)
-	\\&=\int_{y\in\calS}\int_{z\in\mathcal S} \mathbb E[f(X(t-s))\mid X(0)=z]P(X(s)\in\wrt z\mid X(0)=x),
-	\\&=\int_{y\in\calS}\int_{z\in\mathcal S} (\mathbb T(t-s)f(z))P(X(s)\in\wrt z\mid X(0)=x),
+	\\&=\int_{z\in\mathcal S}\mathbb E[f(X(t))\mid X(s)=z] P(X(s)\in\wrt z\mid X(0)=x)
+	\\&=\int_{z\in\mathcal S} \mathbb E[f(X(t-s))\mid X(0)=z]P(X(s)\in\wrt z\mid X(0)=x)
+	\\&=\int_{z\in\mathcal S} (\mathbb T(t-s)f(z))P(X(s)\in\wrt z\mid X(0)=x)
 	\\&=\mathbb T(s) (\mathbb T(t-s)f)(x),
@@
 In Chapters~\ref{sec: conv} and~\ref{ch: global results} we use the Dominated Convergence Theorem to aid our convergence arguments.
-In applied probability we often want to prove convergence of certain probabilistic expressions.
+In applied probability we often want to prove convergence of certain expressions.
 An approach which can simplify matters is to partition the expression on certain events where we have a simpler characterisation, thereby enabling us to prove convergence on each element in the partition.
-The original expression can be written as a sum over the partition.
+The original expression can be written as an integral over the partition.
 To establish the convergence result we initially desired, we can use the convergence of each element of the partition and the Dominated Convergence Theorem.
@@
 \[\int f = \int f \wrt x = \int f \wrt m(x),\]
-where \(m\) denotes Lebesgue measure, to denote the Lebesgue integral. With their notation clarified we now quote their results.
+where \(m\) denotes Lebesgue measure, to denote the Lebesgue integral. 
 \begin{thm}[Dominated Convergence Theorem]
@@
 \end{thm}
-\cite{steinreal} then follow ``Clearly, the [Fubini] theorem is symmetric in \(x\) and \(y\) so that we also may conclude that the slice \(f_x\) is integrable on \(\mathbb R^{d_2}\) for almost every \(x\).
+\cite{steinreal} then state \begin{quotation}``Clearly, the [Fubini] theorem is symmetric in \(x\) and \(y\) so that we also may conclude that the slice \(f_x\) is integrable on \(\mathbb R^{d_2}\) for almost every \(x\).
 Moreover, \(\int_{\mathbb R^{d_2}}f_x(y)\wrt y\) is integrable and
@@
 \[\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y=\displaystyle\int_{R^{d_1}}\left(\int_{R^{d_2}}f(x,y)\wrt y\right)\wrt x = \int_{\mathbb R^d}f.\mbox{''}\]
+\end{quotation}
 It is this last statement which is most powerful. Effectively, if either
@@
 In Chapter~\ref{sec: numerics} we investigate the numerical performance of the uniformisation approximation and show that it is effective.
-However, of the methods considered in Chapter~\ref{sec: numerics}, its performance is the poorest and its rate of convergence is the slowest for the numerical experiments we conduct.
-Though we should not rush to discount the uniformisation approximation.
-Due to the stochastic interpretation of the uniformisation approximation as a QBD, approximations to probabilities are guaranteed to be non-negative, and we can employ probabilistic techniques to analyse the approximation (as \cite{bo2013} did).
-Moreover, its simplicity lends itself more easily to mathematical analysis.
-Further, when the uniformisation scheme is used in the context of approximating a fluid-fluid queue the resultant approximation is itself a fluid queue, and this stochastic interpretation could aid in the analysis of this approximation.
+Of the methods considered in Chapter~\ref{sec: numerics}, the performance of the uniformisation approximation is the poorest and its rate of convergence is the slowest for the numerical experiments we conduct.
+However, due to the stochastic interpretation of the uniformisation approximation as a QBD, approximations to probabilities are guaranteed to be non-negative, and we can employ probabilistic techniques to analyse the approximation (as \cite{bo2013} did).
+Moreover, the simplicity of the uniformisation scheme lends itself more easily to mathematical analysis.
+For example, when the uniformisation scheme is used in the context of approximating a fluid-fluid queue the resultant approximation is itself a fluid queue, and this stochastic interpretation could aid in the analysis of this approximation.
 

### 2022-08-05 front/abstract.tex
intro/ch2 edits

@@
 Fluid queues have found a wide variety of applications including risk processes, telecommunications, and environmental modelling, among others.
-Given the success of fluid queues it is believed that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success.
+Given the success of fluid queues it is plausible that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success.
 \cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and stationary distribution of a fluid-fluid queue.

### 2022-08-05 front/declaration.tex
intro/ch2 edits

@@
 Prior to this boundary conditions were approximated by a small interval in the DG scheme (in this thesis, I generalised the boundary conditions further).
-    \item Added and proved Corollary~4.1.
-    \item Formally deriving, and describing, the approximation to the operator \(\mathbb R\) in Section~5.1. 
-    \item Introduced the concept of operator kernels to the paper (used throughout Section~5.5).
-    \item Wrote the final version of the code (Vikram did the initial version).
-    \item Extended the numerical experiments to consider higher-order bases, smaller cell-widths, elucidating issues around truncation errors, and computing confidence intervals. 
-    \item Added Appendix~A on the properties of the DG approximation to the generator of the fluid queue.
-    \item Added details to Appendix~B. 
-    \item I also made significant contributions to the writing (and rewriting) the manuscript, submission, and responding to reviewers comments. 
+    \item The addition and proof of Corollary~4.1.
+    \item The formal definition and derivation of the approximation to the operator \(\mathbb R\) in Section~5.1. 
+    \item Introduction of the concept of operator kernels to the paper (used throughout Section~5.5).
+    \item The final version of the code (Vikram did the initial version).
+    \item Extension of the numerical experiments to consider higher-order bases, smaller cell-widths, elucidating issues around truncation errors, and computing confidence intervals. 
+    \item Addition of Appendix~A on the properties of the DG approximation to the generator of the fluid queue.
+    \item Additional details in Appendix~B. 
+    \item I also made significant contributions to the writing (and rewriting) the manuscript, submission, and responding to reviewers' comments. 
 \end{itemize}
@@
     \item Section~\ref{sec: ffq intro},
-    \item Chapter~\ref{ch:galerkin}, except for Sections~\ref{sec: other applications}-\ref{sec: slope limiting},
+    \item Chapter~\ref{ch:galerkin}, except for Sections~\ref{sec: other applications} and \ref{sec: limiting and linearity},
     \item Appendix~\ref{appendix:example},

### 2022-08-05 intro/chapter.tex
intro/ch2 edits

@@
 For example, \cite{anick1982} model a \emph{data handling switch} in which there are \(N\) information sources which asynchronously switch from ON to OFF, or OFF to ON.
-The phase process \(\varphi(t)\) models the number of data sources which are ON at time \(t\), and the rate at which data is received when the buffer is on is assumed (without loss of generality) to be 1.
+The phase process \(\varphi(t)\) models the number of data sources which are ON at time \(t\), and the rate at which data is received from a single source when the buffer is on is assumed (without loss of generality) to be 1.
 The switch stores data in a buffer and processes the data at rate \(c\), thus the net rate of change of the buffer at time \(t\) when there are \(r\) information sources which are ON is \(r-c\).
-The level process, \(X(t)\), models the amount of data in the buffer at time \(t\) and the rate at which \(X(t)\) changes which \(\varphi(t)=r\) is \(c_r=r-c\).
+The level process, \(X(t)\), models the amount of data in the buffer at time \(t\) and the rate at which \(X(t)\) changes when \(\varphi(t)=r\) is \(c_r=(r-c)\times 1(X(t)>0)\).
 The model described above could be used to aid in the analysis and design of the data handling switch, for example, by determining the minimum rate \(r\) for which the system is \emph{stable} in steady state, which practically means that the processing power of the switch is large enough that the amount of data in the buffer does not grow without bound.
@@
 However, the added flexibility comes at a cost to the in the analysis of the model.
-For both fluid-fluid queue and classical fluid queues, performance measures such has the stationary distribution, and the Laplace-transforms of hitting times and return times, can be derived in terms of the infinitesimal generator of the modulating process.
+For both fluid-fluid queue and classical fluid queues, performance measures such as the stationary distribution, and the Laplace-transforms of hitting times and return times, can be derived in terms of the infinitesimal generator of the modulating process.
 For fluid queues the modulating process is a CTMC for which the infinitesimal generator is a matrix.
@@
 
-The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
+The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
 For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used.
@@
 In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way that makes clear exactly how the approximations we develop correspond to the theoretical operators.
+Chapter~\ref{ch:math} also introduces the discontinuous Galerkin (DG) method for approximating solutions to differential equations.
+Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:math} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and negativity.
 Chapter~\ref{ch:math} also introduces the existing literature and gives context to this thesis.
 
-Chapter~\ref{ch:galerkin} demonstrates a way that we can use the discontinuous Galerkin (DG) method to approximate certain operators and distributions of fluid and fluid-fluid queues.
-Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:galerkin} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and negativity.
-At the end of Chapter~\ref{ch:galerkin} we also introduce other numerical procedures which we use in our numerical investigations in Chapter~\ref{sec: numerics}.
+Chapter~\ref{ch:galerkin} demonstrates a way that we can use the discontinuous Galerkin method to approximate certain operators and distributions of fluid and fluid-fluid queues.
+The DG method provides an effective tool for approximating performance measures of fluid-fluid queues.
+However, the DG method can exhibit oscillations in the presence of discontinuities and these oscillations can cause approximations to certain probabilities to be negative.
+This motivates the next chapter.
 
-In the following chapter, Chapter~\ref{sec: construction and modelling}, we develop a new QBD-RAP approximation scheme which, due to its interpretation as a stochastic process, ensures all estimates of probabilities will be non-negative.
+In the next chapter, Chapter~\ref{sec: construction and modelling}, we develop a new QBD-RAP approximation scheme which, due to its interpretation as a stochastic process, ensures all estimates of probabilities will be non-negative.
 The chapter takes a stochastic modelling approach to developing the approximation scheme.
 
-Once we have established the new approximation scheme, we then prove a convergence of the approximation scheme in Chapters~\ref{sec: conv}~and~\ref{ch: global results} via certain Laplace transforms with respect to time.
-Chapter~\ref{sec: conv} utilises methods and ideas relating specifically to QBD-RAPs and matrix exponential distributions.
+Once we have established the new approximation scheme, we then prove its convergence in Chapters~\ref{sec: conv}~and~\ref{ch: global results}.
 Ultimately, Chapter~\ref{sec: conv} proves that, on the event that the QBD-RAP has not-yet seen an \emph{orbit restart epoch} (which is much like a change of level, but not always), certain Laplace transforms with respect to time of the QBD-RAP scheme converge to corresponding Laplace transforms with respect to time of the fluid queue.
@@
 Chapter~\ref{ch: global results} then looks to extend the local convergence results of Chapter~\ref{sec: conv} to a global convergence result on the whole domain of approximation.
-Chapter~\ref{ch: global results} uses more traditional Markov process arguments which rely on things such as the strong Markov property, time-homogeneity and the Law of total probability.
-In Chapter~\ref{ch: global results} we first consider the discrete-time process which the process embedded in the QBD-RAP at times when the QBD-RAP sees an orbit restart epoch.
+Chapter~\ref{ch: global results} uses more traditional Markov process arguments which rely on properties such as the strong Markov property, time-homogeneity and the law of total probability.
+In Chapter~\ref{ch: global results} we first consider the discrete-time process which is the process embedded in the QBD-RAP at times when the QBD-RAP sees an orbit restart epoch.
 We prove that the embedded process converges in distribution to a corresponding embedded process of the fluid queue.
-The rest of Chapter~\ref{ch: global results} then attempt to prove a global converge result.
-The main result being Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (with respect to the spatial and temporal variable) to the distribution of the fluid queue.
+The rest of Chapter~\ref{ch: global results} proves a global convergence result.
+The main result is Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (with respect to the spatial and temporal variable) to the distribution of the fluid queue.
 
@@
 The rest of the appendices contain technical results relating to proving the convergence of the QBD-RAP approximation scheme.
-Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operators} introduced as part of the QBD-RAP scheme have the properties we claim they do to prove convergence in Chapter~\ref{sec: conv}.
+Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operators} introduced as part of the QBD-RAP scheme have the properties we claim in proving convergence in Chapter~\ref{sec: conv}.
 Appendix~\ref{app:extend conv} extends some results from Chapter~\ref{sec: conv} to a setting which requires slightly less computation.

### 2022-09-04 appendix_kronecker/appendix.tex
is it finished?? :o

@@
 %!TEX root = ../thesis.tex
-\chapter{Kronecker properties}\label{appendix: kronecker}
-Here we detail some properties of Kronecker sums, products, and exponential (see \citep{MEinAP}, Appendix A.4).
-
-Let 
-\[\bs A = \left[\begin{array}{ccc}a_{11} & \dots & a_{1m}\\\hdots & & \hdots \\ a_{n1}&\dots & a_{nm}\end{array}\right]
-\qquad
-\bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
-be matrices with dimensions \(n\times m\) and \(n'\times m'\), respectively. The operator \(\otimes\) is the Kronecker product of two matrices; 
-\[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
-which is an \(nn'\times mm'\) matrix.
-
-Let \(\bs{C},\bs{D}\)  be matrices with dimensions \(m\times k\) and \(m'\times k'\). A property of the Kronecker Product is 
-\begin{align}
-	\left(\bs A\otimes \bs{B}\right)\left(\bs{C}\otimes \bs{D}\right) &= \bs A\bs{C}\otimes \bs{B}\bs{D}.\label{eqn:mpr}\tag{Mixed Product Rule}
-\end{align}
-\begin{proof}
-	The proof follows from 
-	\begin{align*}
-		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j}\\\vdots\\c_{nj}\bs{D} \end{array}\right] 
-		%
-		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
-		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
-	\end{align*}
-\end{proof}
-
-If \(\bs A\) and \(\bs{B}\) are invertible matrices, then 
-\begin{align}\label{eqn:kron inverse}
-	\left(\bs A\otimes \bs{B}\right)^{-1} = \bs A^{-1}\otimes \bs{B}^{-1}.
-\end{align}
-
-Let \(\bs A\) and \(\bs{B}\) be \(n\times n\) and \(m\times m\) matrices, respectively. The Kronecker sum of \(\bs A\) and \(\bs{B}\) is denoted by \(\oplus\) and defined as 
-\[\bs A\oplus \bs{B} := \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]
-
-The exponential of a square matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}\bs B^n.\]
-
-A property of the Kronecker sum is 
-\begin{align}\label{eqn:kron exp}
-	e^{\bs A\oplus \bs{B}}= e^{\bs A}\otimes e^{\bs{B}}.
-\end{align}
-\begin{proof}
-	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
-	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
-	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\).
-The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\).
-The former follows from
-	\begin{align}
-	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
-	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
-	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
-	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
-	\end{align}
-
-	Therefore 
-	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
-	and the result follows by the mixed product rule. 
-\end{proof}
-
-\begin{lem}\label{lem: lst mpr}
-	Let \(\bs{T}\) and \(\bs{C}\) be \(n\times n\), square matrices with \(\bs{C}\) diagonal and invertible; let \(\bs{S}\) be a \(p\times p\) matrix.
-Further, suppose \(\left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]\) is invertible for \(\lambda>0\).
-Then
-\begin{align}
-	&\int_{t=0}^\infty e^{-\lambda t}  e^{{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t}} \wrt t 
-	%
-	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}.  \label{eqn:lstsimplify}\end{align}
-\end{lem}
-\begin{proof}
-	Computing the integral on the left-hand side and then factorising the result and using the~\ref{eqn:mpr} multiple times gives
-	\begin{align}
-            	\int_{t=0}^\infty e^{-\lambda t} e^{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t} \wrt t\nonumber 
-            	%
-            	&= - \left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]^{-1}
-		%
-		\\&= -  \left[\bs{T}\otimes \bs{I} + \left(\bs{C}\otimes \bs{I}\right)\left(\bs{I}\otimes \bs{S}\right) - \lambda \bs{I}\right]^{-1}\nonumber
-		%
-		\\&= -  \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}\otimes \bs{I}\right)^{-1}\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}\otimes \bs{I}\right)^{-1}\lambda \bs{I}\right)\right]^{-1}.
-\label{eqn: ref this one 12}
-		%
-	\end{align}
-	By Equation~(\ref{eqn:kron inverse}) and since \(\bs{C}\) is invertible, (\ref{eqn: ref this one 12}) is equal to
-	\begin{align}
-		& - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\otimes \bs{I}\right)\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\otimes \bs{I}\right)\lambda \bs{I}\right)\right]^{-1}.
-\label{eqn: ref this one 13}
-		%
-	\end{align}
-	{Using the~\ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
-	\begin{align}
-		&- \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\bs{T}\right)\otimes \bs{I} + \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\lambda \bs{I}\right)\otimes \bs{I}\right)\right]^{-1} \nonumber
-		%
-		\\&= - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right)\right]^{-1} \nonumber
-		%
-		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1} \nonumber
-		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1},\label{eqn: is an integral}
-	\end{align}
-	by definition of the Kronecker sum.
-
-	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore, (\ref{eqn: is an integral}) is 
-	\begin{align*}
-		-\left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1}
-		&= \int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x\right)\oplus \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1}.
-	\end{align*}
-	{Using the rule in Equation~(\ref{eqn:kron exp}) gives }
-	\begin{align*}
-		&\int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)x}\otimes e^{ \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1},
-	\end{align*}
-	which is the result.
-\end{proof}
-
+\chapter{Algebraic manipulations of certain Kronecker products}\label{appendix: kronecker}
 %\section{Matrix exponentials}\label{appendix: matrix exponentials}

### 2022-09-04 chapter1/chapter.tex
is it finished?? :o

@@
 
+\section{Discontinuous Galerkin}
+The discontinuous Galerkin (DG) method is a general methodology for numerically approximating solutions to differential equations.
+The discontinuous Galerkin method is a finite-element method which approximates differential operators by finite dimensional matrices.
+In this thesis we use the discontinuous Galerkin method to approximate the infinitesimal generator of a fluid queue.
+Here, we give a very brief introduction to discontinuous Galerkin methods in general and refer the reader to Chapter~\ref{ch:galerkin} for details on the application to fluid queues, and to \cite{nodalDGBook} for more details on the method in general.
+
+Intuitively, the DG method works by first partitioning the spatial domain into cells, on each cell a basis of (typically polynomial) functions is defined, then the differential equation is projected onto the basis of polynomials.
+Thus, the differential equation is discretised it two ways, by the partition of space into cells, then by the projection onto a finite dimensional basis.
+Linking the cells together essentially amounts to defining boundary conditions for each cell.
+To form the boundary condition for cells adjacent to the boundary we take the boundary condition defined by the problem we are solving and discretise it.
+For cells in the interior the boundary condition is defined by the solution on the adjacent cell.
+In the context of conservation laws (which is what we solve in this thesis), the boundary conditions on each cell are known as the \emph{flux}, as they describe the rate, with respect to time, that the conserved quantity flows across the boundaries of the cell.
+The flux is discretised and is known as the \emph{numerical flux}.
+
+Typically, polynomial functions are used as the basis for the DG method.
+The main choices to make in defining the scheme are the grid which defines the cells, the order of the polynomial basis, and the choice of numerical flux.
+
+\cite{nodalDGBook} claim that the first discontinuous Galerkin finite element scheme was developed in \cite{reed1973} where they use the method to solve the first order differential equation
+\[\sigma u + \nabla \cdot (a u) = f,\]
+in a 2-dimensional spatial domain, where \(\sigma\) is a constant, \(a\) is a piecewise constant function, and \(u\) is the unknown.
+The literature has since expanded and includes (but is not limited to) theory on the convergence rate of the scheme, the choice of \emph{numerical flux} used (we will define this later), the choice of spatial discretisation used, and applications to a range of other problems including higher-order differential equations and to partial differential equations (PDEs) such as non-linear conservation law.
+The PDE which describes the evolution of the distribution of a fluid queue is a conservation law as probability must be conserved.
+
+In one dimension, linear conservation laws take the form 
+\begin{align}
+	&\cfrac{\partial u( x, t)}{\partial t} - \cfrac{\partial}{\partial x} ( c(t) u( x, t)) = 0,\,  x \in \Omega, \label{eqn: cons law}
+	\\& u( x,t)= g( x,t),\tag{boundary condition}
+	\\& u( x,0)=f( x).\tag{initial condition}
+\end{align}
+where \(\Omega\) is the region of interest. Here \(u\) is the conserved quantity, which we will refer to as \emph{mass}, but it could really be any conserved quantity.
+
+To solve conservation laws with the DG method the spatial domain \(\Omega\) is partitioned into cells (intervals) \(\mathcal D_k = [y_k,y_{k+1}]\), \(k=0,...,K-1\).
+For each cell \(\calD_k\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), compactly supported on \(\calD_{k}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k}\)).
+Typically, we take \(\phi_k^r(x)\) to be a basis of polynomial functions.
+The functions \(\{\phi^r_k\}_{r=1}^{p_k}\) span some function space, \(W_k\) say, and further form a basis for this space.
+It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{p_k}\) as a basis of \emph{Lagrange interpolating polynomials} defined by the \emph{Gauss-Lobatto quadrature points} \citep{nodalDGBook} (see also Section~\ref{eqn: la}).
+For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials.
+On each cell \(\mathcal D_{k}\) we approximate
+\[u(x,t)\approx u_{k}(x,t)=\sum\limits_{r=1}^{p_k}a_{k}^r(t)\phi^r_k(x),\] 
+where \(a_{k}^r(t)\) are yet-to-be-determined time-dependent coefficients.
+The approximation \(u_{k}\) is referred to as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k=0}^{K-1} u_{k}\) on the whole domain \(\Omega\).
+The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\{0,...,K-1\}\).
+
+First, consider multiplying (\ref{eqn: cons law}) by the row vector \(\bs \phi_k(x)=[\phi_k^1(x),...,\phi_k^{p_k}(x)]\) and integrating by parts over \(\calD_k\),
+\begin{align}
+	&\int_{x\in \calD_k}\cfrac{\partial u(x,t)}{\partial t}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) u(x,t)\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x - c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}} = 0.\label{eqn: hadndjj}
+\end{align}
+For a sufficiently large basis of functions, then (\ref{eqn: hadndjj}) is equivalent to the PDE (\ref{eqn: cons law}).
+The term \(- c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}}\) is known as the \emph{flux} as it describes the rate at which mass flows into and out of the cell \(\calD_k\).
+The flux is essentially a boundary condition on the cell \(\calD_k\) and is specified by the value of the solution on the adjoining cells.
+When \(c(t)\) is positive (negative) then mass flows into (out of) cell \(\calD_k\) at the left-hand boundary and out of (into) cell \(\calD_k\) at the right-hand boundary.
+Suppose that \(c(t)\) is positive.
+When \(u\) is continuous, then \(u(y_{k}^-,t)=u(y_k,t),\) where \(y^-\) is the left limit at \(y\), hence the flux into cell \(\calD_k\) in Equation (\ref{eqn: cons law}) is \(c(t)u(y_k^-,t)\bs \phi_k(y_k)\).
+The flux out of cell \(\calD_k\) is \(c(t)u(y_{k+1}^-,t)\bs \phi_k(y_{k+1})\).
+Using this, rewrite (\ref{eqn: hadndjj})
+\begin{align}
+	&\int_{x\in \calD_k}\cfrac{\partial u(x,t)}{\partial t}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) u(x,t)\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x \nonumber
+	\\&\qquad\qquad{}- c(t)[u(y_{k+1}^-,t)\bs \phi_k(y_{k+1}) - u(y_{k}^-,t)\bs \phi_k(y_{k})] = 0.\label{eqn: hadndjjj}
+\end{align}
+We can proceed analogously when \(c(t)\) is negative.
+
+Upon substituting the approximations \(u_k\) and \(u_{k-1}\) into (\ref{eqn: hadndjjj}) 
+\begin{align}
+	&\int_{x\in \calD_k}\cfrac{\partial\bs a_k(t)}{\partial t}\bs \phi_k(x)\tr{}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) \bs a_k(t)\bs\phi_k(x)\tr{}\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x \nonumber
+	\\&\qquad\qquad{}- c(t)[\bs a_k(t)\bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}) - \bs a_{k-1}(t)\bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k})].\nonumber
+	\\&=\cfrac{\wrt \bs a_k(t)}{\wrt t}\bs M_k + c(t) \bs a_k(t)\bs G_k - c(t)[\bs a_k(t)\bs F_{k,k} - \bs a_{k-1}(t)\bs F_{k-1,k}]\nonumber
+	\\&= 0,\label{eqn: hadndj}
+\end{align}
+where 
+\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x,\]
+\[\bs F_{k,k} := \bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}),\quad \bs F_{k-1,k} := \bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k}),\]
+and are known as the \emph{mass}, \emph{stiffness} and \emph{flux} matrices.
+Thus, the approximation has reduced the problem to a system of ordinary differential equations.
+One can determine the coefficients \(\bs a_k(t)\), \(k=0,...,K-1\), from the initial condition \(\bs a_k(0)\), \(k=0,...,K-1\), then integrating over time, perhaps numerically.
+The initial coefficients \(\bs a_k(0)\), \(k=0,...,K-1\), can be found by projecting the initial condition onto the basis of polynomial functions, \(\{\phi_k^r\}_{k,r}\).
+
+One subtlety is the choice of numerical flux.
+We arrived at the numerical flux above using a continuity argument.
+However, the approximate solution \(\sum_{k=0}^{K-1}u_k(x,t)\) is discontinuous at the points \({y_{k}}\).
+It turns out that the choice of numerical flux is not unique and there are range of values we may choose from.
+The numerical flux above is known as the \emph{upwind} flux -- we use it throughout this thesis and do not explore others.
+Other choices which define a discontinuous Galerkin scheme are the basis functions used and the cell geometries.
+We refer the reader to \citep{nodalDGBook} and reference therein for more details.
+
+Typically, DG schemes are used to solve physical problems in 2 or 3 dimensions.
+The cell geometries become more complex in higher dimensions and this complicates matters somewhat, but the underlying principles and intuition is the same.
+In this thesis we solve problems with one spatial dimension only.
+
+Two properties make the DG scheme a natural choice to solve the problem present in this thesis.
+High-order accuracy enabled by the use of basis functions to model the solution on each cell, and the cell-based structure of the method with cells related to each other at cell edges only and only to neighbouring cells.
+Finite difference methods can achieve high-order accuracy, however they are not cell-based, which makes them not a natural choice to solve the problems in this thesis.
+Finite volume methods are cell-based, but do not have the same high-order accuracy as the DG method.
+Further, approximating flux terms with greater accuracy in the finite volume method can introduce dependencies between cells which are not neighbours and this creates difficulties for the application in this thesis.
+
+\subsection{Time-integration schemes}\label{sec: time integration}
+In this thesis we numerically integrate ODEs of the form
+\begin{align}\label{eqn: DG ODE w BCs 3}
+	\cfrac{\wrt}{\wrt t} \boldsymbol a(t)
+	% 
+	= \boldsymbol a(t) \bs Q
+\end{align}
+where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix, given an initial condition \(\bs a(0)\).
+To do so, we employ the strong stability preserving Runge-Kutta (SSPRK) scheme of order 4 with 5 stages \citep{sr2002}.
+SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``...can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
+
+In our context, the SSPRK method with \(s\) stages has the form 
+\begin{align*}
+	&\bs v^{(0)} = \bs a(t),\\
+	%
+	&\bs v^{(\ell)} = \sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q,\, \ell = 1,\dots,s,\\
+	%
+	&\bs a(t+h) =\bs v^{(s)},
+\end{align*}
+where \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are coefficients which define the scheme and \(h\) is the \(t\)-step size.
+To ensure the stability-preserving property the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are chosen to be positive so that \(\bs v^{(s)}\) is a convex combination of forward-Euler steps.
+Moreover, the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are optimised so that the allowable \(t\)-step size is as large as possible.
+The maximum allowable \(t\)-step size is \(h_{RK}\), and we require
+\[h_{RK} \leq \min_{\ell k}\cfrac{\alpha_{\ell k}}{\beta_{\ell k}}h_E,\]
+where \(h_E\) is the maximum allowable \(t\)-step size for the forward-Euler scheme.
+
+The \(t\)-step size needs to be chosen to be sufficiently small so that the time integration is stable. Here, for an order \(p\) DG scheme we choose the time step to be less than 
+\[\max_{i\in\calS}|c_i|\min_{r\in{2,...,p+1}}\left(z_{r}-z_{r-1}\right) \cfrac{\min_{k}{\Delta}_k}{2}\]
+where \(z_{r-1}\leq z_r\), \(z_1=-1\), \(z_{p+1}=1\) and \(z_r, r=2,...,p,\) are the \(p-1\) zeros of the first derivative of \(P_{p}(x)\), the order \(p\) Legendre polynomial.
+This follows advice from \cite{nodalDGBook}.
+
+\subsection{Slope limiters}\label{sec: slope limiting}
+A well-know problem with DG schemes is that, in the presence of steep gradients or discontinuities, spurious oscillations in the approximate solution can occur and positivity is not guaranteed (see \cite{nodalDGBook}~Section~5.6, and \cite{koltai2011}~Section~3.3, for example).
+To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
+Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
+However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
+As a result, the limiter may unnecessarily reduce the accuracy of the scheme in the presence of natural oscillations (see \citep{nodalDGBook},~Example~5.8).
+Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear.
+
+We now describe a simple slope limiter known as the \emph{Generalised MUSCL} scheme \citep{c99} (see also \citep{nodalDGBook},~Section~5.6.1). Define the \emph{minmod} function 
+\begin{align*}
+	m(a_1,a_2,a_3)=\begin{cases}
+		s\min_{i\in\{1,2,3\}}|a_i|, & |s|=1, \\
+		0, &\mbox{otherwise},
+	\end{cases}
+\end{align*}
+where \(s=\frac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
+When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
+In the context of DG the three arguments \(a_1,\,a_2\) and \(a_3\) are the estimates of the average gradient of the approximate solution near a given cell.
+When the signs of \(a_1,\,a_2\) and \(a_3\) differ, this suggests an oscillation in the approximate solution.
+
+Now, define \(\overline u_{k}\) as the average value of the approximate solution on cell \(k\), then the slope limited solution on cell \(k\) is the linear approximant 
+\begin{align*}
+	\Pi_{k}^{lim}(u_k)=\overline u_{k} + (x-\overline y_k)m\left(\Pi^1_{\partial x}u_k,\cfrac{\overline u_{k+1}-\overline u_{k}}{\Delta_k},\cfrac{\overline u_{k}-\overline u_{k-1}}{\Delta_k}\right),
+\end{align*}
+where \(\Pi^1_{\partial x}u_k\) is the slope of the linear projection of \(u_k\) on cell \(k\) and \(\overline y_k=(y_k+y_{k+1})/2\) is the centre of the \(k\)th cell.
+
+In non-oscillatory regions of the solution the slope limiter is unnecessary and reduces accuracy, so we only apply the limiter to cells \(k\) where oscillations are detected.
+To determine which cells need limiting we compute
+\begin{align*}
+	v_k^L &:= \overline{u}_{k} - m(\overline{u}_{k} - {u}_{k}^L,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+	\\v_k^R &:= \overline{u}_{k} + m(\overline{u}_{k} - {u}_{k}^R,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+\end{align*}
+where \({u}_{k}^L\) and \({u}_{k}^R\) are the values of the approximate solution on cell \(k\) evaluated at the left-hand and right-hand edges of the cell.
+We apply the slope limiter to cell \(k\) if \(v_k^L\neq u_k^L\) or \(v_k^R\neq u_k^R\), otherwise the approximation on cell \(k\) is left unaltered.
+
+% Notice that the process of limiting is essentially a post-processing of the approximate solution after applying
+
+% In the context of approximating the first return operator of a fluid-fluid queue there is no direct equivalent of slope limiting, other than, perhaps, to re-compute the solution with constant basis function if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
+
+\subsection{Slope limiters and time integration}\label{subsec: slope lim and int}
+To use slope limiters within the SSPRK scheme we first project the initial condition on to \(W\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
+To find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the (limited) initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation~(\ref{eqn: DG ODE w BCs 3}).
+At each stage of the time-integration we apply the slope limiter, i.e.~the scheme above with a slope limiter is
+\begin{align*}
+	&\bs v^{(0)} = \bs a(t),\\
+	%
+	&\bs v^{(\ell)} = \Pi^{lim}\left(\sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q\right),\, \ell = 1,\dots,s,\\
+	%
+	&\bs a(t+h) =\bs v^{(s)},
+\end{align*} 
+where \(\Pi^{lim}\) is the slope limiter function which determines on which cells the solution needs limiting, and if so, applies the limiter \(\Pi_k^{lim}\).
+
+We refer to this (vanilla) application of a limiter to the DG scheme as the DG-lim scheme.
+
+\paragraph{Consequences of slope limiting} We have already mentioned that the Generalised MUSCL slope limiter reduces the order to linear in regions where oscillatory solutions are detected.
+When slope limiting is used in conjunction with a time-integration scheme the reduction in order may not be local in time and may persist past the time when the limiter first detected oscillations.
+For example, consider a problem with an initial condition which introduces oscillations into the numerical approximation.
+If the slope limiter is applied to the initial condition then, in the region around the oscillations, the approximation to the initial condition will be linear.
+Even if the slope limiter is not applied (or not needed) after this point, the initial condition has still been altered, perhaps significantly, from the original approximation and this can affect the accuracy of transient approximations.
+This is not to say that the slope-limited-regions of the approximation will \emph{always} remain linear.
+When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of generalised slope limiter described.
+Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy of subsequent computations.
+
+\subsection{Another slope limiting scheme}\label{sec: anoth pos pres}
+The advantage of the approach above is that, in areas of the approximation where the solution is smooth, then the high-order accuracy of the method can be retained, while maintaining positivity where necessary.
+However, if the problem at hand is dominated by discontinuities and/or if there are no regions of the solution of interest which are smooth, then there is no real advantage in the approach above as the limiter will reduce the approximate solution to linear in all regions of interest.
+Hence, we may as well just use a linear scheme and save on some computation.
+
+Perhaps a better approach would be to use a linear scheme with a smaller cell width such that the computational work remains approximately the same.
+For example, for a DG scheme with \(2p\) basis functions on a cell, \(\mathcal D_k\), of width \(\Delta\), say, we construct block matrices of dimension \(2p\times 2p\) (i.e.~the mass, stiffness and flux matrices, \(\bs M_k\), \(\bs G_k\), and \(\bs F^{k\ell}\), respectively) and compute the approximate solution using these matrices.
+Alternatively, consider breaking the cell, \(\mathcal D_k\), up into \(p\) sub-cells of width \(\Delta/p\) and using a DG scheme with \(2\) basis functions on each sub-cell (i.e.~a linear basis on each cell).
+On each of the \(p\) sub-cells we construct block matrices of size \(2\times 2\).
+To approximate the solution on the original interval \(\mathcal D_k\) we form the \(p\) approximations on each sub-cell into larger matrices which are of dimension \(2p\times 2p\), and then use in the same way as before.
+Thus, the scheme with \(2p\) basis functions on the cell \(\calD_k\), and the scheme with a 2 basis functions on each of the \(p\) sub-cell have approximately the same computational cost (the latter scheme results in a sparser matrix).
+
+If we know \emph{a priori} that we will apply the slope limiter to cell \(\calD_k\), then intuitively we suspect that the latter scheme may be far more accurate; it will approximate the solution on \(\calD_k\) by a piecewise linear function with \(p\) pieces, whereas the DG scheme with \(2p\) basis functions on cell \(\calD_k\) will, after limiting, represent the solution as a single linear function on \(\calD_k\).
+
+We refer to this scheme as the DG-lin-lim scheme, as it is a linear DG scheme with a limiter.
+
+\subsection{Briefly, on accuracy}
+As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method from \cite[Sections~5.5,~5.8]{nodalDGBook} for a similar kind of problem to the one under study here.
+For certain scalar conservation-law problems, under certain regularity conditions on the continuity and boundedness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
+Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
+
+% To quote Section~5.7 of \cite{nodalDGBook} SSPRK methods ``can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
+
+% why the MUSCL limiter?
+%   it requires no tuning and guarentees non-negativity
+%   some other positivity-preserving methods require tuning and can still produce negative results if not tuned properly
+%   ^ same goes for filtering
+
+% Consider the partial differential equation 
+% \[\cfrac{\partial f}{\partial t}+\cfrac{\partial c(f)}{\partial x}=0,\, x\in[a,b],\]
+% where \(c\) is some function, and subject to appropriate initial, \(f(x,0)=f_0(x)\), and boundary conditions, \(f(a,t)=g(t),\) assuming \(c\geq 0\).
+The method first supposes that the solution is of the form \(f(x,t)\) translates the above differential equation, which is in \emph{strong form}, into a \emph{weak form} by multiplying by a test function \(\psi\), integrating with respect to \(x\) and applying integration by parts;
+% \begin{align*}
+% 	% &\int_{x=a}^b\cfrac{\partial f(x)}{\partial t}\psi(x)\wrt x+\int_{x=a}^b\cfrac{\partial c(f)}{\partial x}\psi(x)\wrt x
+% 	\int_{x=a}^b\cfrac{\partial f(x)}{\partial t}\psi(x)\wrt x+\int_{x=a}^b c(f)\cfrac{\wrt\psi(x)}{\wrt x}\wrt x - [c(f(x,t))]_{x=a}^b = 0. 
+% \end{align*}
+% The weak form
+
 \section{Quasi-birth-and-death processes with rational arrival process components}\label{qbd-rap intro}
@@
 
+\section{Kronecker products and related results}
+Here we describe Kronecker products and sums and some of their properties (see \cite{MEinAP}, Appendix A.4, for further details).
+We use the results in this section to manipulate certain matrix expressions in Chapter~\ref{sec: conv}.
+
+Let 
+\[\bs A = \left[\begin{array}{ccc}a_{11} & \dots & a_{1m}\\\hdots & & \hdots \\ a_{n1}&\dots & a_{nm}\end{array}\right]
+\qquad
+\bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
+be matrices with dimensions \(n\times m\) and \(n'\times m'\), respectively. The operator \(\otimes\) is the Kronecker product of two matrices; 
+\[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
+which is an \(nn'\times mm'\) matrix.
+
+Let \(\bs{C},\bs{D}\)  be matrices with dimensions \(m\times k\) and \(m'\times k'\). A property of the Kronecker Product is 
+\begin{align}
+	\left(\bs A\otimes \bs{B}\right)\left(\bs{C}\otimes \bs{D}\right) &= \bs A\bs{C}\otimes \bs{B}\bs{D}.\label{eqn:mpr}\tag{Mixed Product Rule}
+\end{align}
+\begin{proof}
+	The proof follows from 
+	\begin{align*}
+		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j}\\\vdots\\c_{nj}\bs{D} \end{array}\right] 
+		%
+		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
+		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
+	\end{align*}
+\end{proof}
+
+If \(\bs A\) and \(\bs{B}\) are invertible matrices, then 
+\begin{align}\label{eqn:kron inverse}
+	\left(\bs A\otimes \bs{B}\right)^{-1} = \bs A^{-1}\otimes \bs{B}^{-1}.
+\end{align}
+
+Let \(\bs A\) and \(\bs{B}\) be \(n\times n\) and \(m\times m\) matrices, respectively. The Kronecker sum of \(\bs A\) and \(\bs{B}\) is denoted by \(\oplus\) and defined as 
+\[\bs A\oplus \bs{B} := \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]
+
+The exponential of a square matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}\bs B^n.\]
+
+A property of the Kronecker sum is 
+\begin{align}\label{eqn:kron exp}
+	e^{\bs A\oplus \bs{B}}= e^{\bs A}\otimes e^{\bs{B}}.
+\end{align}
+\begin{proof}
+	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
+	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
+	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\).
+The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\).
+The former follows from
+	\begin{align}
+	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
+	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
+	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
+	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
+	\end{align}
+
+	Therefore 
+	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
+	and the result follows by the mixed product rule. 
+\end{proof}
+
+\begin{lem}\label{lem: lst mpr}
+	Let \(\bs{T}\) and \(\bs{C}\) be \(n\times n\), square matrices with \(\bs{C}\) diagonal and invertible; let \(\bs{S}\) be a \(p\times p\) matrix.
+Further, suppose \(\left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]\) is invertible for \(\lambda>0\).
+Then
+\begin{align}
+	&\int_{t=0}^\infty e^{-\lambda t}  e^{{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t}} \wrt t 
+	%
+	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}.  \label{eqn:lstsimplify}\end{align}
+\end{lem}
+\begin{proof}
+	Computing the integral on the left-hand side and then factorising the result and using the~\ref{eqn:mpr} multiple times gives
+	\begin{align}
+            	\int_{t=0}^\infty e^{-\lambda t} e^{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t} \wrt t\nonumber 
+            	%
+            	&= - \left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]^{-1}
+		%
+		\\&= -  \left[\bs{T}\otimes \bs{I} + \left(\bs{C}\otimes \bs{I}\right)\left(\bs{I}\otimes \bs{S}\right) - \lambda \bs{I}\right]^{-1}\nonumber
+		%
+		\\&= -  \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}\otimes \bs{I}\right)^{-1}\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}\otimes \bs{I}\right)^{-1}\lambda \bs{I}\right)\right]^{-1}.
+\label{eqn: ref this one 12}
+		%
+	\end{align}
+	By Equation~(\ref{eqn:kron inverse}) and since \(\bs{C}\) is invertible, (\ref{eqn: ref this one 12}) is equal to
+	\begin{align}
+		& - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\otimes \bs{I}\right)\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\otimes \bs{I}\right)\lambda \bs{I}\right)\right]^{-1}.
+\label{eqn: ref this one 13}
+		%
+	\end{align}
+	{Using the~\ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
+	\begin{align}
+		&- \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\bs{T}\right)\otimes \bs{I} + \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\lambda \bs{I}\right)\otimes \bs{I}\right)\right]^{-1} \nonumber
+		%
+		\\&= - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right)\right]^{-1} \nonumber
+		%
+		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1} \nonumber
+		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1},\label{eqn: is an integral}
+	\end{align}
+	by definition of the Kronecker sum.
+
+	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore, (\ref{eqn: is an integral}) is 
+	\begin{align*}
+		-\left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1}
+		&= \int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x\right)\oplus \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1}.
+	\end{align*}
+	{Using the rule in Equation~(\ref{eqn:kron exp}) gives }
+	\begin{align*}
+		&\int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)x}\otimes e^{ \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1},
+	\end{align*}
+	which is the result.
+\end{proof}
+
 % vector space? Polynomial basis?
@@
 
+Laplace transforms of probability measures with non-negative support and where the Laplace transform variable, \(\lambda\), is real and non-negative can be given a stochastic interpretation.
+Let \(W\) be a random variable with distribution function \(F_W(w)= \mathbb P(W<w)\), then \(\displaystyle\int_{w=0}^\infty e^{-\lambda w} \wrt F_W(w) = \mathbb P(W < E^{\lambda})\).
+That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(W\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
+
 \section{Convergence theorems}
@@
 First, let's define the notion of weak convergence.
-We follow \cite{portmanteaubook}.
-
-Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra.
+We follow \cite{billingsleyconvergence}.
+
+Let \(S\) be a metric space and let \(\mathcal S\) be the Borel \(\sigma\)-algebra generated by the open sub-sets of \(S\).
+A probability measure, \(P\), is a function which maps elements of \(\mathcal S\) (elements of \(\mathcal S\) are sets) to real numbers in the interval \([0,1]\), with \(P(S)=1\).
+Further, \(P\) is \emph{countably-additive}, which means that for any countable collection of \emph{disjoint} sets \(A_1, A_2,...\in\mathcal S\), then
+\[P\left(\bigcup\limits_{n=1}^\infty A_n\right) = \sum\limits_{n=1}^\infty P\left(A_n\right).\]
+Define the notation \(Pf = \int_S f \wrt P\) where \(f:S\to \mathbb R\) is a function.
+Let \(P_1,P_2,...\), be a sequence of probability measures.
+For a given function \(f\), the sequence \(P_1f,P_2f,...\), is simply a sequence of real numbers.
+The sequence of probability measures \(P_1,P_2,....\), is said to \emph{converge weakly} to \(P\) if \(P_nf\to Pf\) for every bounded continuous real function \(f\).
+\cite{billingsleyconvergence} uses the notation \(P_n\Rightarrow P\) to denote this weak convergence.
+
+We need a few more definitions before stating The Portmanteau Theorem.
+A set \(A\in\mathcal S\) is said to be a \(P\)-\emph{continuity set} if \(\partial A\) (the boundary of the set \(A\)) satisfies \(P(\partial A)=0\).
+Define the \emph{limit inferior} and \emph{limit superior} of a sequence of real numbers \(x_1,x_2,...,\) as
+\begin{align*}
+	\liminf_{n\to\infty} x_n &:= \lim_{n\to \infty}	\left(\inf_{m\geq n}x_m\right)
+	\intertext{and}
+	\\\limsup_{n\to\infty} x_n &:= \lim_{n\to \infty}	\left(\sup_{m\geq n}x_m\right),
+\end{align*}
+respectively.
+
+\begin{thm}[Portmanteau Theorem I, Theorem~2.1 of \cite{billingsleyconvergence}]
+	These five conditions are equivalent.
+	\begin{enumerate}
+		\item[(i)] \(P_n\Rightarrow P\). 
+		\item[(ii)] \(P_nf\to Pf\) for all bounded, uniformly continuous \(f\).
+		\item[(iii)] \(\limsup_n P_n F \leq P F\) for all closed [sets] \(F\).
+		\item[(iv)] \(\liminf_n P_n G \leq P G\) for all open [sets] \(G\).
+		\item[(v)] \(P_n A \to P A\) for all \(P\)-continuity sets \(F\).
+	\end{enumerate}
+\end{thm}
+Some authors, such as \cite{portmanteaubook}, also include the following equivalent condition in the Portmanteau Theorem. 
+\begin{thm}[Portmanteau Theorem II, Theorem~13.16 of \cite{portmanteaubook}]\label{thm: Portmanteau}
+	\(\,\)
+	\begin{enumerate}
+		\item[(vi)] \(P_nf \to P f\) for all bounded, Lipschitz continuous \(f\).
+	\end{enumerate}
+\end{thm}
+
+Recall that a stochastic process is a sequence of random variables \(\{X(t)\}_{t\in\mathcal T}\).
+The \emph{finite-dimensional distributions} of \(\{X(t)\}_{t\in\mathcal T}\) are the joint distributions of the random vector \((X(t_1),X(t_2),...,X(t_n))\), where \(t_1,t_2,...,t_n\in \mathcal T\) is a finite collection of times.
+
+Key concepts in establishing weak convergence of stochastic processes are the notions \emph{tight} and \emph{relatively compact}.
+Again, we follow \cite{billingsleyconvergence}.
+A probability measure \(P\) is said to be \emph{tight} if for each \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\).
+A family of probability measures, \(\Pi\), is said to be tight if for every \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\) for every probability measure \(P\) in \(\Pi\).
+The family \(\Pi\) is \emph{relatively compact} if every sequence of elements of \(\Pi\) contains a subsequence which converges weakly.
+
+Let \(P_1,P_2,...,\) and \(P\) be probability measures of stochastic processes. \cite{billingsleyconvergence} provides the following result.
+\begin{thm}\label{thm: aofaa}
+	If \(\{P_n\}\) is relatively compact and the finite-dimensional distributions of \(P_n\) converge weakly to those of \(P\), then \(P_n\Rightarrow_n P\). 
+\end{thm}
+Further, the condition that \(\{P_n\}\) is relatively compact can be replaced by tightness due Prohorov's Theorem \citep{billingsleyconvergence}. 
+\begin{thm}
+	If \(\Pi\) is tight, then it is relatively compact.
+\end{thm}
+
+% Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra.
 Further, denote by \(C_b(E)\) the set of continuous bounded functions on \(E\).
@@
 We say that \(\{\mu_n\}_{n\in\mathbb N}\) converges weakly to \(\mu\), formally \(\mu_n\to \mu\) weakly as \(n\to\infty\), if
-\[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
-We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}.
+% \[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
+% We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}.
 Let \(\mathcal M_{\leq 1}(E):=\{\mu \in \mathcal M_f(E)\mid \mu(E)\leq 1\},\) the set of all sub-probability measures on \((E,\mathcal E)\).
-\begin{thm}[[Part of] The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
-	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
-	\begin{enumerate}
-		\item[(i)] \(\mu_n\to \mu\) weakly as \(n\to \infty\). 
-		\item[(ii)] \(\int f\wrt \mu_n \to \int f\wrt \mu\) for all bounded, Lipschitz continuous \(f\).
-		\\ \(\vdots\) 
-	\end{enumerate}
-\end{thm}
-There are 8 parts to The Portmanteau Theorem in \cite{portmanteaubook}, Theorem~13.16.
+% \begin{thm}[(Part of) The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
+% 	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
+% 	\begin{enumerate}
+% 		\item[(i)] \(\mu_n\to \mu\) weakly as \(n\to \infty\). 
+% 		\item[(ii)] \(\int f\wrt \mu_n \to \int f\wrt \mu\) for all bounded, Lipschitz continuous \(f\).
+% 		\\ \(\vdots\) 
+% 	\end{enumerate}
+% \end{thm}
+% There are 8 parts to The Portmanteau Theorem in \cite{portmanteaubook}, Theorem~13.16.
 Here we only quote to relevant parts; the other 6 parts require us to define other concepts which are not relevant to this thesis, so they are omitted.
@@
 
-\section{Discontinuous Galerkin}
-The discontinuous Galerkin (DG) method is a general methodology for numerically approximating solutions to differential equations.
-The discontinuous Galerkin method is a finite-element method which approximates differential operators by finite dimensional matrices.
-In this thesis we use the discontinuous Galerkin method to approximate the infinitesimal generator of a fluid queue.
-Here, we give a very brief introduction to discontinuous Galerkin methods in general and refer the reader to Chapter~\ref{ch:galerkin} for details on the application to fluid queues, and to \cite{nodalDGBook} for more details on the method in general.
-
-Intuitively, the DG method works by first partitioning the spatial domain into cells, on each cell a basis of (typically polynomial) functions is defined, then the differential equation is projected onto the basis of polynomials.
-Thus, the differential equation is discretised it two ways, by the partition of space into cells, then by the projection onto a finite dimensional basis.
-Linking the cells together essentially amounts to defining boundary conditions for each cell.
-To form the boundary condition for cells adjacent to the boundary we take the boundary condition defined by the problem we are solving and discretise it.
-For cells in the interior the boundary condition is defined by the solution on the adjacent cell.
-In the context of conservation laws (which is what we solve in this thesis), the boundary conditions on each cell are known as the \emph{flux}, as they describe the rate, with respect to time, that the conserved quantity flows across the boundaries of the cell.
-The flux is discretised and is known as the \emph{numerical flux}.
-
-Typically, polynomial functions are used as the basis for the DG method.
-The main choices to make in defining the scheme are the grid which defines the cells, the order of the polynomial basis, and the choice of numerical flux.
-
-\cite{nodalDGBook} claim that the first discontinuous Galerkin finite element scheme was developed in \cite{reed1973} where they use the method to solve the first order differential equation
-\[\sigma u + \nabla \cdot (a u) = f,\]
-in a 2-dimensional spatial domain, where \(\sigma\) is a constant, \(a\) is a piecewise constant function, and \(u\) is the unknown.
-The literature has since expanded and includes (but is not limited to) theory on the convergence rate of the scheme, the choice of \emph{numerical flux} used (we will define this later), the choice of spatial discretisation used, and applications to a range of other problems including higher-order differential equations and to partial differential equations (PDEs) such as non-linear conservation law.
-The PDE which describes the evolution of the distribution of a fluid queue is a conservation law as probability must be conserved.
-
-In one dimension, linear conservation laws take the form 
-\begin{align}
-	&\cfrac{\partial u( x, t)}{\partial t} - \cfrac{\partial}{\partial x} ( c(t) u( x, t)) = 0,\,  x \in \Omega, \label{eqn: cons law}
-	\\& u( x,t)= g( x,t),\tag{boundary condition}
-	\\& u( x,0)=f( x).\tag{initial condition}
-\end{align}
-where \(\Omega\) is the region of interest. Here \(u\) is the conserved quantity, which we will refer to as \emph{mass}, but it could really be any conserved quantity.
-
-To solve conservation laws with the DG method the spatial domain \(\Omega\) is partitioned into cells (intervals) \(\mathcal D_k = [y_k,y_{k+1}]\), \(k=0,...,K-1\).
-For each cell \(\calD_k\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), compactly supported on \(\calD_{k}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k}\)).
-Typically, we take \(\phi_k^r(x)\) to be a basis of polynomial functions.
-The functions \(\{\phi^r_k\}_{r=1}^{p_k}\) span some function space, \(W_k\) say, and further form a basis for this space.
-It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{p_k}\) as a basis of \emph{Lagrange interpolating polynomials} defined by the \emph{Gauss-Lobatto quadrature points} \citep{nodalDGBook} (see also Section~\ref{eqn: la}).
-For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials.
-On each cell \(\mathcal D_{k}\) we approximate
-\[u(x,t)\approx u_{k}(x,t)=\sum\limits_{r=1}^{p_k}a_{k}^r(t)\phi^r_k(x),\] 
-where \(a_{k}^r(t)\) are yet-to-be-determined time-dependent coefficients.
-The approximation \(u_{k}\) is referred to as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k=0}^{K-1} u_{k}\) on the whole domain \(\Omega\).
-The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\{0,...,K-1\}\).
-
-First, consider multiplying (\ref{eqn: cons law}) by the row vector \(\bs \phi_k(x)=[\phi_k^1(x),...,\phi_k^{p_k}(x)]\) and integrating by parts over \(\calD_k\),
-\begin{align}
-	&\int_{x\in \calD_k}\cfrac{\partial u(x,t)}{\partial t}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) u(x,t)\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x - c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}} = 0.\label{eqn: hadndjj}
-\end{align}
-For a sufficiently large basis of functions, then (\ref{eqn: hadndjj}) is equivalent to the PDE (\ref{eqn: cons law}).
-The term \(- c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}}\) is known as the \emph{flux} as it describes the rate at which mass flows into and out of the cell \(\calD_k\).
-The flux is essentially a boundary condition on the cell \(\calD_k\) and is specified by the value of the solution on the adjoining cells.
-When \(c(t)\) is positive (negative) then mass flows into (out of) cell \(\calD_k\) at the left-hand boundary and out of (into) cell \(\calD_k\) at the right-hand boundary.
-Suppose that \(c(t)\) is positive.
-When \(u\) is continuous, then \(u(y_{k}^-,t)=u(y_k,t),\) where \(y^-\) is the left limit at \(y\), hence the flux into cell \(\calD_k\) in Equation (\ref{eqn: cons law}) is \(c(t)u(y_k^-,t)\bs \phi_k(y_k)\).
-The flux out of cell \(\calD_k\) is \(c(t)u(y_{k+1}^-,t)\bs \phi_k(y_{k+1})\).
-Using this, rewrite (\ref{eqn: hadndjj})
-\begin{align}
-	&\int_{x\in \calD_k}\cfrac{\partial u(x,t)}{\partial t}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) u(x,t)\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x \nonumber
-	\\&\qquad\qquad{}- c(t)[u(y_{k+1}^-,t)\bs \phi_k(y_{k+1}) - u(y_{k}^-,t)\bs \phi_k(y_{k})] = 0.\label{eqn: hadndjjj}
-\end{align}
-We can proceed analogously when \(c(t)\) is negative.
-
-Upon substituting the approximations \(u_k\) and \(u_{k-1}\) into (\ref{eqn: hadndjjj}) 
-\begin{align}
-	&\int_{x\in \calD_k}\cfrac{\partial\bs a_k(t)}{\partial t}\bs \phi_k(x)\tr{}\bs\phi_k(x)\wrt x + \int_{x\in \calD_k} c(t) \bs a_k(t)\bs\phi_k(x)\tr{}\cfrac{\partial}{\partial x} \bs \phi_k( x)\wrt x \nonumber
-	\\&\qquad\qquad{}- c(t)[\bs a_k(t)\bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}) - \bs a_{k-1}(t)\bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k})].\nonumber
-	\\&=\cfrac{\wrt \bs a_k(t)}{\wrt t}\bs M_k + c(t) \bs a_k(t)\bs G_k - c(t)[\bs a_k(t)\bs F_{k,k} - \bs a_{k-1}(t)\bs F_{k-1,k}]\nonumber
-	\\&= 0,\label{eqn: hadndj}
-\end{align}
-where 
-\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x,\]
-\[\bs F_{k,k} := \bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}),\quad \bs F_{k-1,k} := \bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k}),\]
-and are known as the \emph{mass}, \emph{stiffness} and \emph{flux} matrices.
-Thus, the approximation has reduced the problem to a system of ordinary differential equations.
-One can determine the coefficients \(\bs a_k(t)\), \(k=0,...,K-1\), from the initial condition \(\bs a_k(0)\), \(k=0,...,K-1\), then integrating over time, perhaps numerically.
-The initial coefficients \(\bs a_k(0)\), \(k=0,...,K-1\), can be found by projecting the initial condition onto the basis of polynomial functions, \(\{\phi_k^r\}_{k,r}\).
-
-One subtlety is the choice of numerical flux.
-We arrived at the numerical flux above using a continuity argument.
-However, the approximate solution \(\sum_{k=0}^{K-1}u_k(x,t)\) is discontinuous at the points \({y_{k}}\).
-It turns out that the choice of numerical flux is not unique and there are range of values we may choose from.
-The numerical flux above is known as the \emph{upwind} flux -- we use it throughout this thesis and do not explore others.
-Other choices which define a discontinuous Galerkin scheme are the basis functions used and the cell geometries.
-We refer the reader to \citep{nodalDGBook} and reference therein for more details.
-
-Typically, DG schemes are used to solve physical problems in 2 or 3 dimensions.
-The cell geometries become more complex in higher dimensions and this complicates matters somewhat, but the underlying principles and intuition is the same.
-In this thesis we solve problems with one spatial dimension only.
-
-Two properties make the DG scheme a natural choice to solve the problem present in this thesis.
-High-order accuracy enabled by the use of basis functions to model the solution on each cell, and the cell-based structure of the method with cells related to each other at cell edges only and only to neighbouring cells.
-Finite difference methods can achieve high-order accuracy, however they are not cell-based, which makes them not a natural choice to solve the problems in this thesis.
-Finite volume methods are cell-based, but do not have the same high-order accuracy as the DG method.
-Further, approximating flux terms with greater accuracy in the finite volume method can introduce dependencies between cells which are not neighbours and this creates difficulties for the application in this thesis.
-
-\subsection{Time-integration schemes}\label{sec: time integration}
-In this thesis we numerically integrate ODEs of the form
-\begin{align}\label{eqn: DG ODE w BCs 3}
-	\cfrac{\wrt}{\wrt t} \boldsymbol a(t)
-	% 
-	= \boldsymbol a(t) \bs Q
-\end{align}
-where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix, given an initial condition \(\bs a(0)\).
-To do so, we employ the strong stability preserving Runge-Kutta (SSPRK) scheme of order 4 with 5 stages \citep{sr2002}.
-SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``...can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
-
-In our context, the SSPRK method with \(s\) stages has the form 
-\begin{align*}
-	&\bs v^{(0)} = \bs a(t),\\
-	%
-	&\bs v^{(\ell)} = \sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q,\, \ell = 1,\dots,s,\\
-	%
-	&\bs a(t+h) =\bs v^{(s)},
-\end{align*}
-where \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are coefficients which define the scheme and \(h\) is the \(t\)-step size.
-To ensure the stability-preserving property the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are chosen to be positive so that \(\bs v^{(s)}\) is a convex combination of forward-Euler steps.
-Moreover, the coefficients \(\alpha_{\ell k}\) and \(\beta_{\ell k}\) are optimised so that the allowable \(t\)-step size is as large as possible.
-The maximum allowable \(t\)-step size is \(h_{RK}\), and we require
-\[h_{RK} \leq \min_{\ell k}\cfrac{\alpha_{\ell k}}{\beta_{\ell k}}h_E,\]
-where \(h_E\) is the maximum allowable \(t\)-step size for the forward-Euler scheme.
-
-The \(t\)-step size needs to be chosen to be sufficiently small so that the time integration is stable. Here, for an order \(p\) DG scheme we choose the time step to be less than 
-\[\max_{i\in\calS}|c_i|\min_{r\in{2,...,p+1}}\left(z_{r}-z_{r-1}\right) \cfrac{\min_{k}{\Delta}_k}{2}\]
-where \(z_{r-1}\leq z_r\), \(z_1=-1\), \(z_{p+1}=1\) and \(z_r, r=2,...,p,\) are the \(p-1\) zeros of the first derivative of \(P_{p}(x)\), the order \(p\) Legendre polynomial.
-This follows advice from \cite{nodalDGBook}.
-
-\subsection{Slope limiters}\label{sec: slope limiting}
-A well-know problem with DG schemes is that, in the presence of steep gradients or discontinuities, spurious oscillations in the approximate solution can occur and positivity is not guaranteed (see \cite{nodalDGBook}~Section~5.6, and \cite{koltai2011}~Section~3.3, for example).
-To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
-Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
-However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
-As a result, the limiter may unnecessarily reduce the accuracy of the scheme in the presence of natural oscillations (see \citep{nodalDGBook},~Example~5.8).
-Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear.
-
-We now describe a simple slope limiter known as the \emph{Generalised MUSCL} scheme \citep{c99} (see also \citep{nodalDGBook},~Section~5.6.1). Define the \emph{minmod} function 
-\begin{align*}
-	m(a_1,a_2,a_3)=\begin{cases}
-		s\min_{i\in\{1,2,3\}}|a_i|, & |s|=1, \\
-		0, &\mbox{otherwise},
-	\end{cases}
-\end{align*}
-where \(s=\frac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
-When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
-In the context of DG the three arguments \(a_1,\,a_2\) and \(a_3\) are the estimates of the average gradient of the approximate solution near a given cell.
-When the signs of \(a_1,\,a_2\) and \(a_3\) differ, this suggests an oscillation in the approximate solution.
-
-Now, define \(\overline u_{k}\) as the average value of the approximate solution on cell \(k\), then the slope limited solution on cell \(k\) is the linear approximant 
-\begin{align*}
-	\Pi_{k}^{lim}(u_k)=\overline u_{k} + (x-\overline y_k)m\left(\Pi^1_{\partial x}u_k,\cfrac{\overline u_{k+1}-\overline u_{k}}{\Delta_k},\cfrac{\overline u_{k}-\overline u_{k-1}}{\Delta_k}\right),
-\end{align*}
-where \(\Pi^1_{\partial x}u_k\) is the slope of the linear projection of \(u_k\) on cell \(k\) and \(\overline y_k=(y_k+y_{k+1})/2\) is the centre of the \(k\)th cell.
-
-In non-oscillatory regions of the solution the slope limiter is unnecessary and reduces accuracy, so we only apply the limiter to cells \(k\) where oscillations are detected.
-To determine which cells need limiting we compute
-\begin{align*}
-	v_k^L &:= \overline{u}_{k} - m(\overline{u}_{k} - {u}_{k}^L,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
-	\\v_k^R &:= \overline{u}_{k} + m(\overline{u}_{k} - {u}_{k}^R,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
-\end{align*}
-where \({u}_{k}^L\) and \({u}_{k}^R\) are the values of the approximate solution on cell \(k\) evaluated at the left-hand and right-hand edges of the cell.
-We apply the slope limiter to cell \(k\) if \(v_k^L\neq u_k^L\) or \(v_k^R\neq u_k^R\), otherwise the approximation on cell \(k\) is left unaltered.
-
-% Notice that the process of limiting is essentially a post-processing of the approximate solution after applying
-
-% In the context of approximating the first return operator of a fluid-fluid queue there is no direct equivalent of slope limiting, other than, perhaps, to re-compute the solution with constant basis function if a higher-order approximation happens to be oscillatory, but this is a post-hoc solution which would essentially require computing two solutions.
-
-\subsection{Slope limiters and time integration}\label{subsec: slope lim and int}
-To use slope limiters within the SSPRK scheme we first project the initial condition on to \(W\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
-To find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the (limited) initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation~(\ref{eqn: DG ODE w BCs 3}).
-At each stage of the time-integration we apply the slope limiter, i.e.~the scheme above with a slope limiter is
-\begin{align*}
-	&\bs v^{(0)} = \bs a(t),\\
-	%
-	&\bs v^{(\ell)} = \Pi^{lim}\left(\sum_{k=0}^{\ell-1} \alpha_{\ell k}\bs v^{(k)} + h\beta_{\ell k}  \bs v^{(k)}\bs Q\right),\, \ell = 1,\dots,s,\\
-	%
-	&\bs a(t+h) =\bs v^{(s)},
-\end{align*} 
-where \(\Pi^{lim}\) is the slope limiter function which determines on which cells the solution needs limiting, and if so, applies the limiter \(\Pi_k^{lim}\).
-
-We refer to this (vanilla) application of a limiter to the DG scheme as the DG-lim scheme.
-
-\paragraph{Consequences of slope limiting} We have already mentioned that the Generalised MUSCL slope limiter reduces the order to linear in regions where oscillatory solutions are detected.
-When slope limiting is used in conjunction with a time-integration scheme the reduction in order may not be local in time and may persist past the time when the limiter first detected oscillations.
-For example, consider a problem with an initial condition which introduces oscillations into the numerical approximation.
-If the slope limiter is applied to the initial condition then, in the region around the oscillations, the approximation to the initial condition will be linear.
-Even if the slope limiter is not applied (or not needed) after this point, the initial condition has still been altered, perhaps significantly, from the original approximation and this can affect the accuracy of transient approximations.
-This is not to say that the slope-limited-regions of the approximation will \emph{always} remain linear.
-When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of generalised slope limiter described.
-Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy of subsequent computations.
-
-\subsection{Another slope limiting scheme}\label{sec: anoth pos pres}
-The advantage of the approach above is that, in areas of the approximation where the solution is smooth, then the high-order accuracy of the method can be retained, while maintaining positivity where necessary.
-However, if the problem at hand is dominated by discontinuities and/or if there are no regions of the solution of interest which are smooth, then there is no real advantage in the approach above as the limiter will reduce the approximate solution to linear in all regions of interest.
-Hence, we may as well just use a linear scheme and save on some computation.
-
-Perhaps a better approach would be to use a linear scheme with a smaller cell width such that the computational work remains approximately the same.
-For example, for a DG scheme with \(2p\) basis functions on a cell, \(\mathcal D_k\), of width \(\Delta\), say, we construct block matrices of dimension \(2p\times 2p\) (i.e.~the mass, stiffness and flux matrices, \(\bs M_k\), \(\bs G_k\), and \(\bs F^{k\ell}\), respectively) and compute the approximate solution using these matrices.
-Alternatively, consider breaking the cell, \(\mathcal D_k\), up into \(p\) sub-cells of width \(\Delta/p\) and using a DG scheme with \(2\) basis functions on each sub-cell (i.e.~a linear basis on each cell).
-On each of the \(p\) sub-cells we construct block matrices of size \(2\times 2\).
-To approximate the solution on the original interval \(\mathcal D_k\) we form the \(p\) approximations on each sub-cell into larger matrices which are of dimension \(2p\times 2p\), and then use in the same way as before.
-Thus, the scheme with \(2p\) basis functions on the cell \(\calD_k\), and the scheme with a 2 basis functions on each of the \(p\) sub-cell have approximately the same computational cost (the latter scheme results in a sparser matrix).
-
-If we know \emph{a priori} that we will apply the slope limiter to cell \(\calD_k\), then intuitively we suspect that the latter scheme may be far more accurate; it will approximate the solution on \(\calD_k\) by a piecewise linear function with \(p\) pieces, whereas the DG scheme with \(2p\) basis functions on cell \(\calD_k\) will, after limiting, represent the solution as a single linear function on \(\calD_k\).
-
-We refer to this scheme as the DG-lin-lim scheme, as it is a linear DG scheme with a limiter.
-
-\subsection{Briefly, on accuracy}
-As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method from \cite[Sections~5.5,~5.8]{nodalDGBook} for a similar kind of problem to the one under study here.
-For certain scalar conservation-law problems, under certain regularity conditions on the continuity and bounded-ness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
-Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
-
-% To quote Section~5.7 of \cite{nodalDGBook} SSPRK methods ``can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''
-
-% why the MUSCL limiter?
-%   it requires no tuning and guarentees non-negativity
-%   some other positivity-preserving methods require tuning and can still produce negative results if not tuned properly
-%   ^ same goes for filtering
-
-% Consider the partial differential equation 
-% \[\cfrac{\partial f}{\partial t}+\cfrac{\partial c(f)}{\partial x}=0,\, x\in[a,b],\]
-% where \(c\) is some function, and subject to appropriate initial, \(f(x,0)=f_0(x)\), and boundary conditions, \(f(a,t)=g(t),\) assuming \(c\geq 0\).
-The method first supposes that the solution is of the form \(f(x,t)\) translates the above differential equation, which is in \emph{strong form}, into a \emph{weak form} by multiplying by a test function \(\psi\), integrating with respect to \(x\) and applying integration by parts;
-% \begin{align*}
-% 	% &\int_{x=a}^b\cfrac{\partial f(x)}{\partial t}\psi(x)\wrt x+\int_{x=a}^b\cfrac{\partial c(f)}{\partial x}\psi(x)\wrt x
-% 	\int_{x=a}^b\cfrac{\partial f(x)}{\partial t}\psi(x)\wrt x+\int_{x=a}^b c(f)\cfrac{\wrt\psi(x)}{\wrt x}\wrt x - [c(f(x,t))]_{x=a}^b = 0. 
-% \end{align*}
-% The weak form
-
 \section{Sundry mathematical concepts}\label{eqn: la}

### 2022-09-04 chapter2/chapter.tex
is it finished?? :o

@@
 This projection leads to a new system of equations which represent a \textit{weak form} of the original system of PDEs.
-Next, we approximate the \textit{flux operator} which moves probability from one cell to another.
+Next, we approximate the \textit{flux operator} which moves mass from one cell to another.
 This creates a system of ordinary differential equations which is known as a \emph{semi-discrete system} (discrete in the spatial variable but continuous in time).
@@
 \section{The partial differential equation}
-We start by introducing the PDE from which we will extract the approximation to the generator \(\mathbb B\).
+We start by introducing the PDE (assuming the PDE exists) from which we will extract the approximation to the generator \(\mathbb B\).
 
-Let $f_i(x,t)$ be the joint density of $(X_t, \varphi_t)$: 
+Let $f_i(x,t)$ be the joint density of an unbounded fluid queue $(\ddot X_t, \varphi_t)$ (assuming a density exists): 
 	% 
 	\begin{align*} 
-	    f_i(x,t) := \cfrac{\partial}{\partial x} 	\mathbb{P}\left(X(t) \leq x, \varphi(t) = i\right),\quad 0<x<b,\, i \in\calS,
+	    f_i(x,t) := \cfrac{\partial}{\partial x} 	\mathbb{P}\left(X(t) \leq x, \varphi(t) = i\right),\quad i \in\calS,
 	\end{align*} 
@@
 \begin{align*}
-\cfrac{\partial}{\partial t} f_i(x,t)& = \sum_{j\in \mathcal{S}}  f_j(x,t)T_{ji} - c_i \cfrac{\partial}{\partial x} f_i(x,t), \quad 0<x<b,\, i\in\mathcal S,
+\cfrac{\partial}{\partial t} f_i(x,t)& = \sum_{j\in \mathcal{S}}  f_j(x,t)T_{ji} - c_i \cfrac{\partial}{\partial x} f_i(x,t), \quad i\in\mathcal S,
 \end{align*}
@@
 \section{Cells, test functions, and weak formulation}
-To begin with, consider an unbounded first fluid level \(\{\ddot X_t,t\geq0\}\), \(\ddot X_t\in(-\infty,\infty)\).
-We will eventually truncate this space so that we have a finite dimensional approximation; however, this requires a discussion on boundary conditions which we save for later.
+To begin with, consider an unbounded fluid queue with fluid level \(\{\ddot X_t,t\geq0\}\), \(\ddot X_t\in(-\infty,\infty)\).
+Later, we will use the approximation to the generator of the fluid queue to approximate performance measures of fluid-fluid queues, but for now, we can consider just the driving fluid queue.
+We will eventually truncate the state space, \(\ddot X_t\in(-\infty,\infty)\), so that we have a finite dimensional approximation; however, this requires a discussion of boundary conditions which we save for later.
 Let \(\mathcal D_{k,i} = [y_k,y_{k+1}),\, k\in\mathbb Z\) for \(i\in\calS_+\cup\calS_0\) and \(\mathcal D_{k,i} = (y_k,y_{k+1}],\, k\in\mathbb Z\) for \(i\in\calS_-\), \(y_k<y_{k+1}\), partition the domain \((-\infty,\infty)\).
@@
 
-For each \(k\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), compactly supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) to form a basis for the space \(W_k\), in which we formulate the approximation.
-Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{p_k}\) to be the space of polynomials of degree \(p_k-1\).
+For each \(k\in \mathbb Z\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) for all \(i\in\calS\), to form a basis for the space \(W_k\), in which we formulate the approximation.
+Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{p_k}\) to be a basis for the space of polynomials of degree \(p_k-1\).
 For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials, but any polynomial basis can be used.
@@
 Let \(\mathcal N_k := \left\{1,\dots,p_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
-\[\boldsymbol \phi_k(x) = (\phi^r_k(x))_{r\in\mathcal N_k}, \quad \boldsymbol a_{k,i}(x) = (a_{k,i}^r(x))_{r\in\mathcal N_k},\,i\in\mathcal S.\]
+\[\boldsymbol \phi_k(x) = (\phi^r_k(x))_{r\in\mathcal N_k}, \quad \boldsymbol a_{k,i}(t) = (a_{k,i}^r(t))_{r\in\mathcal N_k},\,i\in\mathcal S.\]
 %Also define \textit{global} row-vectors 
@@
 for \(j\in\calS\).
-It is common to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero.
+It is convenient to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero.
 Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE.
@@
 
-For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\) (the direct sum of \(W^k\)), the same as the set of basis functions of our solution space.
-Proceeding as described above, the weak formulation is
+For the problem at hand, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\) (the direct sum of \(W^k\)), the same as the set of basis functions of our solution space.
+Proceeding as described above, for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z,\) the weak formulation is
 \begin{align*}
@@
 \end{align*}
-since \(\phi^r_k\) is compactly supported on \(\calD_{k,j}\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\).
+since \(\phi^r_k\) is compactly supported on \(\calD_{k,j}\).
+Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), the orthogonal complement of \(W\).
 Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi^r_k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\).
@@
 \end{align*}
-	which is equivalent to
+	which reduces to
 \begin{align}
@@
 \end{align}
-Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), where the prime, \({}\tr{}\), denotes the transpose, which we now substitute into (\ref{eqn:hash})
+Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), where the prime denotes the transpose, which we now substitute into (\ref{eqn:hash})
 %\begin{align*}
@@
 \section{Boundary conditions}\label{subsec: boundary DG}
-To enable computation, this numerical approximation has to take place on a finite interval, which means we must consider a bounded domain and specify boundary conditions.
+To enable computation, the numerical approximation has to take place on a finite interval, which means we must consider a bounded domain and specify boundary conditions.
 Recall that we wish to impose lower and upper boundaries at \(0\) and \(b\), respectively.
+On the interior of the space, the PDE (\ref{eqn:pde_density_matrix}) (assuming it exists) still describes the evolution of the distribution of the fluid queue.
+Regarding the approximation, the constructions above can still be used to approximate the evolution of the fluid queue on the interior of the space.
+The imposition of boundary conditions amounts to truncating the operator \(\ddot{\bs B}\), adding boundary elements to the system and describing how mass moves between the boundary and the elements adjacent to the boundary.
 %In the numerics chapter, Chapter~\ref{sec: numerics}, we wish to approximate a fluid-fluid queue where the first fluid level, \(\dot X_t\), is bounded below at 0, only: in that case the first step in the approximation scheme is to approximate \(\dot X_t\) by \( X_t\).
@@
 Similarly, the coefficients \( {\boldsymbol{p}}\) are given by 
-	\begin{equation}\vligne{\bs{p}^{-}  & \bs{p}^{0}} = z \vligne{{\bs\xi} & \bs{0}} 
+	\begin{equation}\vligne{\bs{p}^{-}  & \bs{p}^{0}} = Z \vligne{{\bs\xi} & \bs{0}} 
 	\left(-\left[\begin{array}{ll} 
@@
 		\end{array} \right] \right)^{-1},\label{eqn:pisystem1}\end{equation}
-		where \(z\) is a normalising constant. The coefficients \(\bs \pi(y)\) are given by 
+		where \(Z\) is a normalising constant. The coefficients \(\bs \pi(y)\) are given by 
 \begin{align} 
@@
 Thus, for the application to fluid-fluid queues slope limiting is not a viable option.
-We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{Y(t)\}\) first return to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation.
+We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{Y(t)\}\) first returns to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation.
 The application of the limiter/filter is dependent on the initial condition, so the post-processing needs to be done for every initial condition under consideration.

### 2022-09-04 chapter5/chapter.tex
is it finished?? :o

@@
 In this chapter we prove convergence results which build on the main results of Chapter~\ref{sec: conv}.
-We consider the discrete-time embedded processes formed by observing the QBD-RAP at the orbit restart epochs and by observing the fluid queue at the hitting times of the level process at the on points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
-In Corollary~\ref{cor: aln22222212} we prove that the transition probability of the embedded process of the QBD-RAP converge those of the embedded process of the fluid queue.
+We consider the discrete-time embedded processes formed by observing the QBD-RAP at the orbit restart epochs and by observing the fluid queue at the hitting times of the level process on the points \(\{y_\ell\}\), which are the boundaries of the intervals \(\{\calD_\ell\}\).
+In Corollary~\ref{cor: aln22222212} we prove that the transition probabilities of the embedded process of the QBD-RAP converge those of the embedded process of the fluid queue.
 In Corollary~\ref{cor: short cor} we prove that the distribution of the sojourn time of the QBD-RAP in a given level converges to the distribution of the sojourn time of the fluid queue in the corresponding interval.
@@
 
-Consider the embedded process formed by observing the QBD-RAP at the orbit restart epochs. Let \(\{\tau_n^{(p)}\}_{n\geq 0, n \in \mathbb Z}\), \(\tau_0^{(p)}=0\), and
+Consider the embedded process formed by observing the QBD-RAP at the orbit restart epochs. Let \(\tau_0^{(p)}=0\), and
 \[\tau_{n}^{(p)} = \inf\left\{t\geq \tau_{n-1}^{(p)} \mid L^{(p)}(t)\neq L^{(p)}(\tau_{n-1}^{(p)}), \mbox{or \(\{\bs Y(t)\}\) hits a boundary}\right\},\]
@@
 To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed.
-Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the orbit restart epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
+Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the orbit restart epochs.
 The subscript \(\bs \alpha \) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\), unless the process is temporarily absorbed in a sticky boundary, in which case \(\bs A^{(p)}(\tau_n^{(p)})=1\).
@@
 We have the following result on the convergence of the embedded processes \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\), which we also will utilise later to prove a global result.
-\begin{cor}\label{cor: aln222222} For \(\ell_0\in\mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
-	for \(n=0\), then
+\begin{lem}\label{cor: aln222222} For \(\ell_0\in\mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
+	then
 	\begin{align}
@@
 	% \end{align}
-\end{cor}
+\end{lem}
 \begin{proof}
@@
 \end{proof}
-A direct corollary of Corollary~\ref{cor: aln222222} is the convergence of the transition probabilities of the embedded process. 
+A direct corollary of Lemma~\ref{cor: aln222222} is the convergence of the transition probabilities of the embedded process. 
 \begin{cor}\label{cor: aln22222212} For \(\ell_0\in \mathcal K,\) \(x_0\in\mathcal D_{\ell_0,i}\), \(i\in\mathcal S\), 
-		for \(n=0\), then
+		then
 		\begin{align}
@@
 \begin{proof}
-	Since \(\tau_1^{(p)}<\infty\) almost surely, as is \(\tau_1^X\), then taking \(\lambda \to 0\) in Corollary~\ref{cor: aln222222} yields the result. 
+	Since \(\tau_1^{(p)}<\infty\) almost surely, as is \(\tau_1^X\), then taking \(\lambda \to 0\) in Lemma~\ref{cor: aln222222} yields the result. 
 \end{proof}
@@
 Thus, the finite-dimensional distributions of \(\{\Ydp(n)\}\) converge, and if the space \(\mathcal K\times \calS\) is finite, then the sequence of distributions of \(\{\Ydp(n)\}\) is tight.
-Thus, we can establish the convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).
+Thus, by Theorem~\ref{thm: aofaa}, we establish convergence in distribution of \(\{\Ydp(n)\}\) and \(\{\bs X(\tau_n^X)\}\).
 
@@
 Second, considering the Laplace transform with respect to time of the distribution of the QBD-RAP \emph{between} the \(n\)th and \((n+1)\)th orbit restart epoch.
-Third, summing over the number of orbit restart epochs, \(n\), and via the Dominated Convergence Theorem, we claim a convergence of the Laplace transforms with respect to time of the QBD-RAP and fluid queue.
-Lastly, we apply the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} to claim a weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
+Third, summing over the number of orbit restart epochs, \(n\), and using the Dominated Convergence Theorem, we establish a convergence of the Laplace transforms with respect to time of the QBD-RAP and fluid queue.
+Lastly, we apply the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} to establish a weak convergence (in space and time) of the QBD-RAP approximation scheme to the fluid queue.
 
@@
 The idea for the analysis of the QBD-RAP at the \(n\)th orbit restart epoch is to partition the distribution of the QBD-RAP on the phase and level at the times \(\tau_1^{(p)}\) and \(\tau_2^{(p)},...,\tau_{n-1}^{(p)}\) and then use the strong Markov property to write the expression in terms of the distributions of the QBD-RAP at time \(\tau_k^{p}\) given the distribution of the QBD-RAP at time \(\tau_{k-1}^{(p)}\), \(k=1,...,n\).
-The distribution at time \(\tau_1^{(p)}\) depends on the initial condition, so we treat it separately and use Corollary~\ref{cor: aln222} to claim convergence to the fluid queue.
+The distribution at time \(\tau_1^{(p)}\) depends on the initial condition, so we treat it separately and use Corollary~\ref{cor: aln222} to establish convergence to the fluid queue.
 By time-homogeneity, the distributions of the QBD-RAP at times \(\tau_k^{p}\) given the position of the QBD-RAP at time \(\tau_{k-1}^{(p)}\) is equal to the distribution of the QBD-RAP at time \(\tau_1^{(p)}\) given an initial distribution of the QBD-RAP.
-Hence, we may also use Corollary~\ref{cor: aln222} to claim convergence to the fluid queue for each of these distributions.
+Hence, we may also use Corollary~\ref{cor: aln222} to establish convergence to the fluid queue for each of these distributions.
 Putting all the elements together yields the convergence at the \(n\)th orbit restart epoch.
@@
 
-Thus, returning to (\ref{eqn: 161222}), taking the limit as \(p\to\infty\), and using the convergence established in (\ref{eqn: aMMMMM}) and (\ref{eqn: lim 12}), then 
+Thus, returning to (\ref{eqn: 161222}), taking the limit as \(p\to\infty\), and using the convergence established in (\ref{eqn: aMMMMM}) and (\ref{eqn: lim 12}), we see that  
 \begin{align}
@@
 \end{align}
-by the strong Markov property of the fluid queue and the Law of total probability.
+by the strong Markov property of the fluid queue and the Law of Total Probability.
 
@@
 	t\in\mathcal T_0^{(p)}\mid \Ydp(0) = (\ell,j_n))\wrt t \psi(x) \nonumber 
-	 \\&= \int_{x\in\calD_{\ell,j}} \widehat f^{\ell,(p)} (\lambda)(x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x)\wrt x \label{eqn: Jk jJ KK}
+	 \\&= \int_{x\in\calD_{\ell,j}} \widehat f^{\ell,(p)} (\lambda)(x,j;y_{\ell+1(j_n\in\mathcal S_-)},j_n) \psi(x)\wrt x. \label{eqn: Jk jJ KK}
 \end{align}
@@
 
-Hence, we have shown the following result 
+Hence, we have shown
 \begin{lem}\label{lem: LAkAKFnvnb mav h}
@@
 The arguments for an initial phase \(i\in\mathcal S_-\) are analogous.
-For an orbit restart epochs to occur, either;
+For an orbit restart epoch to occur, either;
 	\begin{enumerate}
@@
 	\end{align}
-	Partitioning on the time of the first orbit restart epoch, \(\tau_1\), and the phase and level at time \(\tau_1\), and using the strong Markov property, then (\ref{eqn: alalal}) is equal to 
+	Partitioning on the time of the first orbit restart epoch, \(\tau_1\), and the phase and level at time \(\tau_1\), and using the strong Markov property, (\ref{eqn: alalal}) is equal to 
 	\begin{align}
@@
 
-	Since \(\{\bs Y(t)\in(\ell,\calD_{\ell,j},j),t\in\mathcal T_n\}\subseteq \{t\in\mathcal T_n\}\), then the expression 
+	Since \(\{\bs Y(t)\in(\ell,\calD_{\ell,j},j),t\in\mathcal T_n\}\subseteq \{t\in\mathcal T_n\}\), the expression 
 	\begin{align}
@@
 \subsection{Global convergence}\label{sec: sub global}
-Finally, we combine the convergence result of Lemma~\ref{lem: LAkAKFnvnb mav h} and the domination condition from Corollary~\ref{vcor: cdks d} via the Dominated Convergence Theorem to claim convergence of the Laplace transform of the QBD-RAP given by
+Finally, we combine the convergence result of Lemma~\ref{lem: LAkAKFnvnb mav h} and the domination condition from Corollary~\ref{vcor: cdks d} via the Dominated Convergence Theorem to establish convergence of the Laplace transform of the QBD-RAP given by
 \begin{align}
@@
 However, for \(p\) sufficiently large, there exists a \(p_0<\infty\) and a \(B\) with \(B<1\) such that \(b^{(p)}<B\) for all \(p>p_0\).
-Hence, we can apply the Dominated Convergence Theorem to (\ref{eqn: KLDNVnav}) and claim that
+Hence, we can apply the Dominated Convergence Theorem to (\ref{eqn: KLDNVnav}) and establish
 \begin{align}
@@
 	\end{align}
-	since, to be in level \(\ell\) after starting in level \(\ell_0\), there must be at least \(|\ell_0-\ell|\) orbit restart epochs.
-By Lemma~\ref{lem: another bound 2} then (\ref{eqn:SKAM.
+	observing that, to be in level \(\ell\) after starting in level \(\ell_0\), there must be at least \(|\ell_0-\ell|\) orbit restart epochs.
+By Lemma~\ref{lem: another bound 2} (\ref{eqn:SKAM.
 fqj}) is bounded by \(\left(b^{(p)}\right)^{|\ell-\ell_0|-1}\) for \(|\ell-\ell_0|\geq 2\) and by \(1\) otherwise.
@@
 
-The Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} can now be used to claim that the QBD-RAP approximation scheme converges weakly (in space and time) to the fluid queue.
+Theorem~\ref{thm: ext cont thm}, The Extended Continuity Theorem for Laplace transforms, can now be used to establish weak convergence of the QBD-RAP approximation scheme (in space and time) to the fluid queue.
 
@@
 \begin{proof}
-	Combine the Extended Continuity Theorem for Laplace transforms \cite[Chapter XIII, Theorem 2a]{feller1957} with the convergence of Laplace transforms in Corollary~\ref{cor: lk}. 
+	Combine Theorem~\ref{thm: ext cont thm} with the convergence of Laplace transforms in Corollary~\ref{cor: lk}. 
 \end{proof}
@@
 for almost all \(t\geq 0\).
-At such values of \(t\), since \(\psi\) is arbitrary and bounded, then The Portmanteau Theorem (Theorem~13.16 of \cite{portmanteaubook}) states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
+At such values of \(t\), since \(\psi\) is arbitrary and bounded, then The Portmanteau Theorem \ref{thm: Portmanteau} states that the QBD-RAP approximation scheme converges in distribution to the fluid queue.
 
@@
 
-The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, i.e.~the hypotheses of the theorem depend only on the interval/level \({\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval/level, and not on any other interval/level.
+The arguments which prove Theorem~\ref{thm: a thm!} are independent of all other levels/intervals, that is the hypotheses of the theorem depend only on the interval/level \({\ell_0}\), and the sequence of matrix exponential distributions used to model the behaviour of the fluid queue on this interval/level, and not on any other interval/level.
 Thus, Lemma~\ref{thm: a thm!} holds independently on each interval, as does Corollary~\ref{cor: aln222}.
@@
 	where 
-	\[b^{(p)} = \max\left\{\sup_{\ell_0}b_{\ell_0}^{(p)}, \cfrac{q}{\lambda + q}\right\}.\]
+	\[b^{(p)} = \max\left\{\sup_{\ell_0}b_{\ell_0}^{(p)}, \cfrac{q}{\lambda + q}\right\},\]
+	and \(\sup_{\ell_0}b_{\ell_0}^{(p)}<1\) for \(p\) sufficiently large.
 \end{lem}
@@
 	\end{align*}
+	Now, let \(p_{\ell_0}\) be such that \(b_{\ell_0}^{(p)}<1\) for all \(p\geq p_{\ell_0}\) and let \(p^*=\sup_{\ell_0}p_{\ell_0}\).
+Since \(\mathcal K^\circ\) has finitely many elements, then \(p^*\) is finite and therefore \(\sup_{\ell_0}b_{\ell_0}^{(p)}<1\) for all \(p\geq p^*\).
 \end{proof}

### 2022-09-19 chapter1/chapter.tex
finito

@@
 We may omit the index set and write \(\{X(t)\}\) in place of \(\{X(t)\}_{t\in\mathcal T}\) when it is not explicitly needed, or we may write \(\{X(t)\}_{t\geq t_0}\) to mean \(\{X(t)\}_{t\in[t_0,\infty)}\).
-The \emph{state space} of \(X(t)\) is the set of possible values that \(X(t)\) can take at any time \(t\in\mathcal T\).
+The \emph{state space}, \(\mathcal S\), of \(X(t)\) is the set of possible values that \(X(t)\) can take at any time \(t\in\mathcal T\).
 
@@
 
-A stochastic process is \emph{stationary} if, for any \(n\), \(t_0<t_1<...<t_n\) and any \(t\), \(X(t_n),X(t_{n-1}),...,X(t_0)\) and \(X(t_n+t),X(t_{n-1}+t),...,X(t_0+t)\) have the same distribution.
+A stochastic process is \emph{stationary} if, for any \(n\), \(t_0<t_1<...<t_n\) and any \(t\), then \(X(t_n),X(t_{n-1}),...,X(t_0)\) and \(X(t_n+t),X(t_{n-1}+t),...,X(t_0+t)\) have the same distribution.
 That is,
@@
 Also, as in \cite[Section~1.3]{MEinAP}, we call \(\{X(t)\}_{t\geq 0}\) a \emph{continuous-time Markov chain} (CTMC) if \(\calS\) is countable and for all \(t_{n+1} > t_n > t_{n-1} > ...
-> t_0 > 0\) and \(i,i_0,...,i_{n-1},j\in\calS,\)
+> t_0 \geq 0\) and \(i,i_0,...,i_{n-1},j\in\calS,\)
 \begin{align}\label{eqn: Markov property cmtc}
-	\mathbb P(X(t_{n+1})=j \mid X(n)=i, X(n-1)=i_{n-1},...,X(0)=i_0) = \mathbb P(X(t_{n+1})=j \mid X(t_n)=i),
+	\mathbb P(X(t_{n+1})=j \mid X(t_n)=i, X(t_{n-1})=i_{n-1},...,X(t_0)=i_0) = \mathbb P(X(t_{n+1})=j \mid X(t_n)=i),
 \end{align}
@@
 
-Most generally, we call \(\{X(t)\}_{t\geq 0}\) a \emph{Markov process} if for all \(t_n > t_{n-1} > ... > t_0 > 0,\, t_k\in\mathcal T,\, k=0,...,n+1,\) and \(E_0,...,E_n\subseteq \mathcal S\)
+Most generally, we call \(\{X(t)\}_{t\geq 0}\) a \emph{Markov process} if for all \(t_n > t_{n-1} > ... > t_0 \geq 0,\, t_k\in\mathcal T,\, k=0,...,n,\) and \(E_0,...,E_n\subseteq \mathcal S\)
 \begin{align}\label{eqn: Markov process}
-	\mathbb P(X(t_n)\in E_n,X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
+	\mathbb P(X(t_n)\in E_n,\mid X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
 	= \mathbb P(X(t_n)\in E_n\mid X(t_{n-1})),
 \end{align}
-and is time homogeneous if \(\mathbb P(X(t+s)\in E_n\mid X(s))\) does not depend on \(s\).
+and we say that it is time homogeneous if \(\mathbb P(X(t+s)\in E_n\mid X(s))\) does not depend on \(s\).
 
@@
 
-Let \(c\in \mathbb R\) and \(v\in V\), \(w\in W\).
+Let \(c\in \mathbb R\) (\(\mathbb R\) is the reals), \(v\in V\) and \(w\in W\).
 Then \(\mathbb B\) is a \emph{linear} operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\).
@@
 \[\mathbb E[f(X(t))\mid X(0)=x]=\int_{y\in\calS} f(y) P(X(t)\in\wrt y\mid X(0)=x),\]
-where \(f<F\) is some bounded function.
+where \(f\) is some bounded function, \(|f|<F\).
 We can think of expectation as an operator which is acting on the function \(f\).
@@
 \[\mathbb T(t)f(x) = \mathbb E[f(X(t))\mid X(0)=x].\]
-The result of the operator acting on \(f\), \(\mathbb T(t)f(x)\), is another function.
+The result of the operator acting on \(f\), \(\mathbb T(t)f\), is another function (a function of the initial point \(x\)).
 
@@
 
-The level process is given by 
+Now, back to fluid queues. The level process is given by 
 \[\ddot X(t) = \ddot X(0) + \int_{s=0}^t c_{\varphi(s)}\wrt s.\]
@@
 Moreover, these techniques often lead to quantities which have direct probabilistic interpretations which further aids in their analysis and also applications to other problems \citep{ar2003,dasilva2005,bnp2018} (we also leverage these stochastic interpretations in Chapter~\ref{sec: conv} to derive expressions for certain the Laplace transforms of the fluid queue).
-In some contexts, the actual transient distributions are not required, and Laplace transforms or moments are all that are required.
-In other cases, we can invert the Laplace transforms using known methods (\cite{aw2006}, or in the case of the discontinuities we might prefer \cite{hhat2020}, which uses the same concentrated matrix exponential distribution that we do).
+In some contexts, the actual transient distributions are not required, and Laplace transforms or moments, which can be obtained relatively straightforwardly from the Laplace transforms, are all that are required.
+In other cases, the Laplace transforms may be inverted using known methods (\cite{aw2006}, or in the case of the discontinuities we might prefer \cite{hhat2020}, which uses the same concentrated matrix exponential distribution that we do).
 
@@
 As a result, there are only three types of sample paths to consider.
-Further, \cite{bean2005} similarly argue that the generator \(\bs Q(\lambda)\) can be partitioned into blocks \(\bs Q_{mn}(\lambda) = \left[Q_{ij}(\lambda)\right]_{i\in\calS_m,j\in\calS_n}\), \(m\in\{+,-\},\,n\in\{+,-,0\}\) where
+Further, \cite{bean2005} argue that the generator \(\bs Q(\lambda)\) can be partitioned into blocks \(\bs Q_{mn}(\lambda) = \left[Q_{ij}(\lambda)\right]_{i\in\calS_m,j\in\calS_n}\), \(m\in\{+,-\},\,n\in\{+,-,0\}\) where
 \begin{align*}
@@
 \begin{align}
-	\int_{y=0}^\infty e^{\bs Q_{++}(\lambda)y} \bs \Psi_x(\lambda)\bs Q_{-+}(\lambda)\bs \Psi_X(\lambda) e^{\bs Q_{--}(\lambda)y}\wrt y.
+	\int_{y=0}^\infty e^{\bs Q_{++}(\lambda)y} \bs \Psi_X(\lambda)\bs Q_{-+}(\lambda)\bs \Psi_X(\lambda) e^{\bs Q_{--}(\lambda)y}\wrt y.
 \end{align}
@@
 Then we need to partition \(\mathbb B\) on sets for which \(\{\dot Y(t)\}\) is increasing, decreasing or constant -- these are the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), below.
-We can then derive an operator-Riccati equation for the first return operator for the fluid-fluid queue, \(\mathbb \Psi\), in terms of the partitioned generator \(\mathbb B\), the rates \(r_i(x)\).
-The solution to the operator-Riccati equation gives the distribution of the driving process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to \(0\).
+We can then derive an operator-Riccati equation for the first return operator for the fluid-fluid queue, \(\mathbb \Psi\), in terms of the partitioned generator and the rates \(r_i(x)\).
+The solution to the operator-Riccati equation is an operator which can be used to obtain the distribution of the driving process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to \(0\).
 
@@
 Here, we approximate \(\mathbb B\) by a finite-dimensional matrix which we then partition according to the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\).
-We then substitute the resulting matrices into the operator-Riccati equation and this gives us a matrix-Riccati equation which we can then solve using known methods \cite{bot08}.
+We then substitute the resulting matrices into the operator-Riccati equation and this gives us a matrix-Riccati equation which we can then solve using known methods \citep{bot08}.
 
@@
 The keen observer will notice that (\ref{eqn:jointpi}) assumes a certain differentiability condition on the limiting distribution.
-Indeed, it is not always the case that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) (trivially, consider \(\dot Y(t)=y\) for all \(t\)).
-Conditions on fluid-fluid queues to ensure that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) are not known, and here we just assume that (\ref{eqn:jointpi}) is differentiable.
+Indeed, it is not always the case that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) (trivially, consider \(\dot Y(t)=y_0\) for all \(t\)).
+Conditions on fluid-fluid queues to ensure that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) are not known, and here we just assume that it is.
 
@@
 The generator of a fluid queue is a differential operator and to enable computation of the first-return operator, approximation methods are needed.
-The approximation schemes which we discuss in this thesis are all cell-based methods which discretise the level of the fluid queue into cells.
-Thus, one complexity in approximation is to reconcile the partition according to whether the second fluid is increasing, decreasing or constant and the partition which the approximation method uses.
-As long as the rates of the fluid-fluid queue are not `too-badly' behaved, this is not terribly difficult to do, but it does require some discussion and notation.
+The approximation schemes which we discuss in this thesis are all cell-based methods which discretise the level of the fluid queue into intervals called cells.
+Thus, one complexity in approximation is to reconcile the partition according to whether the second fluid is increasing, decreasing or constant, and the partition which the approximation method uses.
+%As long as the rates of the fluid-fluid queue are not `too-badly' behaved, this is not terribly difficult to do, but it does require some discussion and notation.
 
@@
 We claim that numerical schemes are needed to approximate the analytic operator equations introduced in \cite{bo2014}.
-The schemes we choose to use here work by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \({\calD_k,i}\) then constructing an approximation to \(\overline{\mathbb B}\) on each interval.
+The schemes we choose to use here work by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \({\calD_{k,i}}\) then constructing an approximation to \(\overline{\mathbb B}\) on each interval.
 To help elucidate the connection between the operators \(\{ \overline{\mathbb  V}(t)\}\), \( \overline{\mathbb  B}\) and their approximate counterparts we take a slightly different approach to partitioning these operators than that taken in \cite{bo2014}.
@@
 
-By an appropriate choice of the intervals \(\{\mathcal D_k\},\, k\in\mathcal K\), the partition used in \cite{bo2014} can be recovered.
+By an appropriate choice of the intervals \(\{\mathcal D_{k,i}\}\), the partition used in \cite{bo2014} can be recovered.
 Intuitively, we must ensure that each of the boundaries of \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), align with a boundary of a cell \(\mathcal D_{k,i}\).
-Then, each set \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), can be written as a union of cells, \(\mathcal D_k,\,k\in\mathcal K\), sans a collection of points which have measure zero for all measures in \(\mathcal M_{0,b}\), and this collection of points is inconsequential for the purposes of the approximations presented here.
+Then, each set \(\calF_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), can be written as a union of cells, \(\mathcal D_{k,i},\,k\in\mathcal K\), sans a collection of points which have measure zero for all measures in \(\mathcal M_{0,b}\), and this collection of points is inconsequential for the purposes of the approximations presented here.
 
@@
 \begin{align*}
-	\mu_{k,i}\mathbb B_{ij}^{kk}(\calD_{k,i})&=\int_{x\in\calD_k}v_{k,i}(x)T_{ij}\wrt x,\\
-	\mu_{k,i}\mathbb B_{ii}^{kk}(\calD_{k,i})&=\int_{x\in\calD_k}v_{k,i}(x)T_{ii}\wrt x - c_iv_{k,i}(y_{k+1}^-) 1{(c_i>0)} + c_iv_{k,i}(y_{k}^+) 1{(c_i<0)},
+	\mu_{k,i}\mathbb B_{ij}^{kk}(\calD_{k,i})&=\int_{x\in\calD_{k,i}}v_{k,i}(x)T_{ij}\wrt x,\\
+	\mu_{k,i}\mathbb B_{ii}^{kk}(\calD_{k,i})&=\int_{x\in\calD_{k,i}}v_{k,i}(x)T_{ii}\wrt x - c_iv_{k,i}(y_{k+1}^-) 1{(c_i>0)} + c_iv_{k,i}(y_{k}^+) 1{(c_i<0)},
 \end{align*}
@@
 \begin{align*}
-\mu_{k,i}\mathbb B_{ii}^{k,k+1}(\calD_{k+1})&= c_iv_{k,i}(y_{k+1}^-) 1{(c_i>0)},\mbox{ for all \(i\in\calS\), \(k\in\{0,1,2,...,K-1\}\),}
-\\\mu_{k,i}\mathbb B_{ii}^{k,k-1}(\calD_{k-1})&= -c_iv_{k,i}(y_{k}^+) 1{(c_i<0)}, \mbox{  for all \(i\in\calS\), \(k\in\{1,2,3,...,K\}\)}.
+\mu_{k,i}\mathbb B_{ii}^{k,k+1}(\calD_{k+1,i})&= c_iv_{k,i}(y_{k+1}^-) 1{(c_i>0)},\mbox{ for all \(i\in\calS\), \(k\in\{0,1,2,...,K-1\}\),}
+\\\mu_{k,i}\mathbb B_{ii}^{k,k-1}(\calD_{k-1,i})&= -c_iv_{k,i}(y_{k}^+) 1{(c_i<0)}, \mbox{  for all \(i\in\calS\), \(k\in\{1,2,3,...,K\}\)}.
 \end{align*}
@@
 Intuitively, the DG method works by first partitioning the spatial domain into cells, on each cell a basis of (typically polynomial) functions is defined, then the differential equation is projected onto the basis of polynomials.
-Thus, the differential equation is discretised it two ways, by the partition of space into cells, then by the projection onto a finite dimensional basis.
+Thus, the differential equation is discretised it two ways, by the partition of the state space into cells, then by the projection onto a finite dimensional basis.
 Linking the cells together essentially amounts to defining boundary conditions for each cell.
-To form the boundary condition for cells adjacent to the boundary we take the boundary condition defined by the problem we are solving and discretise it.
-For cells in the interior the boundary condition is defined by the solution on the adjacent cell.
-In the context of conservation laws (which is what we solve in this thesis), the boundary conditions on each cell are known as the \emph{flux}, as they describe the rate, with respect to time, that the conserved quantity flows across the boundaries of the cell.
+To form the boundary condition for cells adjacent to the boundary we take the boundary condition defined by the problem we are solving and discretise it by projecting it onto a set of basis functions.
+For cells in the interior the boundary condition is defined by the solutions on adjacent cells.
+In the context of conservation laws (which is what we solve in this thesis), the boundary conditions on each cell are known as the \emph{flux}, as they describe the rate, with respect to time, that the conserved quantity flows across the boundaries of the cells.
 The flux is discretised and is known as the \emph{numerical flux}.
@@
 	&\cfrac{\partial u( x, t)}{\partial t} - \cfrac{\partial}{\partial x} ( c(t) u( x, t)) = 0,\,  x \in \Omega, \label{eqn: cons law}
-	\\& u( x,t)= g( x,t),\tag{boundary condition}
+	\\& u( x,t)= g( x,t),\, x \in \partial \Omega\tag{boundary condition}
 	\\& u( x,0)=f( x).\tag{initial condition}
 \end{align}
-where \(\Omega\) is the region of interest. Here \(u\) is the conserved quantity, which we will refer to as \emph{mass}, but it could really be any conserved quantity.
+where \(\Omega\) is the region of interest and \(\partial \Omega\) its boundary. Here \(u\) is the conserved quantity, which we will refer to as \emph{mass}, but it could be any conserved quantity.
 
@@
 On each cell \(\mathcal D_{k}\) we approximate
-\[u(x,t)\approx u_{k}(x,t)=\sum\limits_{r=1}^{p_k}a_{k}^r(t)\phi^r_k(x),\] 
+\[u(x,t)\approx u_{k}(x,t)=\sum\limits_{r=1}^{p_k}a_{k}^r(t)\phi^r_k(x),\, x \in\calD_k,\] 
 where \(a_{k}^r(t)\) are yet-to-be-determined time-dependent coefficients.
@@
 
-First, consider multiplying (\ref{eqn: cons law}) by the row vector \(\bs \phi_k(x)=[\phi_k^1(x),...,\phi_k^{p_k}(x)]\) and integrating by parts over \(\calD_k\),
+Now, consider multiplying (\ref{eqn: cons law}) by the row vector \(\bs \phi_k(x)=()\phi_k^1(x),...,\phi_k^{p_k}(x))\) and integrating by parts over \(\calD_k\),
 \begin{align}
@@
 \end{align}
-For a sufficiently large basis of functions, then (\ref{eqn: hadndjj}) is equivalent to the PDE (\ref{eqn: cons law}).
-The term \(- c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}}\) is known as the \emph{flux} as it describes the rate at which mass flows into and out of the cell \(\calD_k\).
+The term \(- c(t)[u(x,t)\bs \phi_k(x)]_{x=y_k}^{y_{k+1}}\) is known as the \emph{flux}, and it describes the rate at which mass flows into and out of the cell \(\calD_k\).
 The flux is essentially a boundary condition on the cell \(\calD_k\) and is specified by the value of the solution on the adjoining cells.
-When \(c(t)\) is positive (negative) then mass flows into (out of) cell \(\calD_k\) at the left-hand boundary and out of (into) cell \(\calD_k\) at the right-hand boundary.
+When \(c(t)\) is positive (negative) then mass flows into (out of) cell \(\calD_k\) at the left-hand boundary, and out of (into) cell \(\calD_k\) at the right-hand boundary.
 Suppose that \(c(t)\) is positive.
@@
 Thus, the approximation has reduced the problem to a system of ordinary differential equations.
-One can determine the coefficients \(\bs a_k(t)\), \(k=0,...,K-1\), from the initial condition \(\bs a_k(0)\), \(k=0,...,K-1\), then integrating over time, perhaps numerically.
+One can determine the coefficients \(\bs a_k(t)\), \(k=0,...,K-1\), from the initial condition \(\bs a_k(0)\), \(k=0,...,K-1\), and integrating over time, perhaps numerically.
 The initial coefficients \(\bs a_k(0)\), \(k=0,...,K-1\), can be found by projecting the initial condition onto the basis of polynomial functions, \(\{\phi_k^r\}_{k,r}\).
@@
 The numerical flux above is known as the \emph{upwind} flux -- we use it throughout this thesis and do not explore others.
-Other choices which define a discontinuous Galerkin scheme are the basis functions used and the cell geometries.
+
+Other choices we have made to define our discontinuous Galerkin scheme are the basis functions used and the cell geometries.
 We refer the reader to \citep{nodalDGBook} and reference therein for more details.
@@
 
-Two properties make the DG scheme a natural choice to solve the problem present in this thesis.
-High-order accuracy enabled by the use of basis functions to model the solution on each cell, and the cell-based structure of the method with cells related to each other at cell edges only and only to neighbouring cells.
-Finite difference methods can achieve high-order accuracy, however they are not cell-based, which makes them not a natural choice to solve the problems in this thesis.
+Two properties which make the DG scheme a natural choice to solve the problem present in this thesis are: high-order accuracy enabled by the use of basis functions to model the solution on each cell, and the cell-based structure of the method with cells related to each other at cell edges only and only to neighbouring cells.
+Finite difference methods can achieve high-order accuracy, however they are not cell-based, which makes them not a natural choice to solve the fluid-fluid queue problems in this thesis.
 Finite volume methods are cell-based, but do not have the same high-order accuracy as the DG method.
-Further, approximating flux terms with greater accuracy in the finite volume method can introduce dependencies between cells which are not neighbours and this creates difficulties for the application in this thesis.
+Further, approximating flux terms with greater accuracy in the finite volume method can introduce dependencies between cells which are not neighbours and this creates difficulties when partitioning the approximate operator which we need to do for our application to fluid-fluid queues.
 
@@
 To rectify this, in some contexts, slope limiting can be used (see \cite{c99}, or \cite{nodalDGBook},~Section~5.6.2 and references therein).
-Slope limiting alters the DG approximation in regions where oscillations are detected to ensure non-oscillatory by reducing the order of the approximation to linear and limiting the gradient of the approximation in these regions.
+Slope limiting alters the DG approximation in regions where oscillations are detected to ensure the approximate solution is non-oscillatory.
+It does so by reducing the order of the approximation to linear and then limiting the gradient of the approximation in these regions.
 However, limiting does not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme.
@@
 Furthermore, slope limiting adds an extra computational cost on top of the approximation scheme and means that the resultant approximation to certain operators are no longer linear.
+This is of consequence in our application to fluid-fluid queues where we approximate an operator-Riccati equation by a matrix-Riccati equation, which is only possible when the approximate operators are matrices.
 
@@
 \end{align*}
-where \(s=\frac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
+where \(s=\cfrac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
 When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
@@
 When no oscillations are detected, the DG scheme can use the full power of the high-order polynomial basis to approximate the solution, and this is one of the advantages of generalised slope limiter described.
-Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy of subsequent computations.
+Instead, we want to point out that once the limiter is applied at one time point, the approximation becomes linear which will affect the accuracy at subsequent time points.
 
@@
 For comparison, the variance of an Erlang distribution, the most concentrated Phase-type distribution, decreases at rate \(\mathcal O(1/p)\) as the order \(p\) increases \citep{as1987}.
-In this thesis we use CMEs to approximate deterministic events.
-For a given order, concentrated matrix exponential distributions have a much lower coefficient of variation than any phase type of the same order, and therefore a better ability to model determinism.
-
-The class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hht2020} and computationally efficient expressions for the density and moments of CMEs are provided in \citep{hht2020}.
+In this thesis we use CMEs to approximate the distribution of deterministic events.
+For a given order, concentrated matrix exponential distributions have a much lower coefficient of variation than any Phase-type of the same order, and therefore a better ability to model determinism.
+
+The class of concentrated matrix exponential distributions (CMEs) is found numerically in \citep{hht2020} and computationally efficient expressions for the density and moments of CMEs are provided in \citep{hht2020}.
 
@@
 
-For CMEs considered here, the vector function \(\bs k(t)=\cfrac{\bs \alpha e^{St}}{\bs \alpha e^{St}\bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
+For CMEs the vector function \(\bs k(t)=\cfrac{\bs \alpha e^{St}}{\bs \alpha e^{St}\bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
 We leverage this fact to simplify some numerical integration in Chapter~\ref{sec: construction and modelling}.
@@
 \section{Kronecker products and related results}
-Here we describe Kronecker products and sums and some of their properties (see \cite{MEinAP}, Appendix A.4, for further details).
+Here we describe Kronecker products and sums and some of their properties (see \cite{MEinAP}, and also Appendix A.4, for further details).
 We use the results in this section to manipulate certain matrix expressions in Chapter~\ref{sec: conv}.
@@
 \section{Laplace Transforms}
-In Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with an object called the \emph{Laplace transform}.
+We have already mentioned \emph{Laplace transforms}, but we have not-yet defined them.
+We define them now.
+In Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with {Laplace transforms} to prove convergence of the QBD-RAP scheme.
 For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transform of \(\mu\) to be
@@
 which gives a probabilistic interpretation of the Laplace transform.
+That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(Z\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
 
@@
 
-Laplace transforms of probability measures with non-negative support and where the Laplace transform variable, \(\lambda\), is real and non-negative can be given a stochastic interpretation.
+%Laplace transforms of probability measures with non-negative support and where the Laplace transform variable, \(\lambda\), is real and non-negative can be given a stochastic interpretation.
 Let \(W\) be a random variable with distribution function \(F_W(w)= \mathbb P(W<w)\), then \(\displaystyle\int_{w=0}^\infty e^{-\lambda w} \wrt F_W(w) = \mathbb P(W < E^{\lambda})\).
@@
 \subsection*{Polynomials}
-In this thesis, we use the discontinuous Galerkin method to approximate the generator, a differential operator, of a fluid queue.
 The discontinuous Galerkin method involves projecting the operator equation onto a basis of functions, typically polynomials.
@@
 In contrast, the work of \cite{bo2014} treats more general models and derives operator-analytic expressions for the joint limiting distribution.
-Therefore, compared to \cite{lnp13}, the methods of this thesis therefore apply to a general class of models and enable to approximation of the joint limiting distribution.
-However, unlike \cite{lnp13}, the analysis of \cite{bo2014} does not consider a level-dependent driving process such as this, but I believe that it would be relatively straightforward to extend their results (and the results of this thesis) to the level-dependent case where the rates \(c_i\) depend on \(X(t)\) in a piecewise-constant way.
+Therefore, compared to \cite{lnp13}, the methods of this thesis apply to a more general class of models and enable to approximation of the joint limiting distribution.
+However, unlike \cite{lnp13}, the analysis of \cite{bo2014} does not consider a level-dependent driving process.
+%but I believe that it would be relatively straightforward to extend their results (and the results of this thesis) to the level-dependent case where the rates \(c_i\) depend on \(X(t)\) in a piecewise-constant way.
 \cite{bo2013b} analyse a model which is simple special case of the fluid-fluid queue.
@@
 They derive matrix-analytic expressions for certain Laplace transforms, with respect to the position of the first fluid \(\{X(t)\}\) relative to its starting position \(X(0)\), of the probability of certain transient events (such as first return and draining/filling times) of \(\{Y(t)\}\).
-In contrast, \cite{bo2013b} together with this thesis, treat a more general model and together, give computational methods for various performance measures of fluid-fluid queues.
-The yet-unpublished work of \cite{bop2020} treats a similar model to that of \cite{bo2013b}, except that the driving process has a regulated lower boundary at \(0\); this significantly complicates the analysis.
+%In contrast, \cite{bo2014} treats a more general model.
+The work of \cite{bop2020} treats a similar model to that of \cite{bo2013b}, except that the driving process has a regulated lower boundary at \(0\); this significantly complicates the analysis.
 They derive matrix-analytic expressions for the first return operator of the second fluid given a specific exponential form of the initial distribution and a certain boundary condition is met.
-Once again, \cite{bo2013b} together with this thesis, treat a more general model and together, give computational methods for various performance measures of fluid-fluid queues.
-
-Returning now to \cite{bo2014} which introduces the fluid-fluid queue in its generality and derives operator-analytic expressions for the first-return operator of the second fluid and the joint limiting distribution.
+%Once again, \cite{bo2013b} together with this thesis, treat a more general model and together, give computational methods for various performance measures of fluid-fluid queues.
+
+Returning now to \cite{bo2014} which treats the most general form of the fluid-fluid queue.
 Their analysis is inspired by the analysis of classical fluid queues in \cite{bean2005} whereby matrix-analytic expressions for the fluid level are derived in terms of the generator of the driving process.
@@
 Often, it is the limiting distribution of the fluid queue which is of interest.
-In the context of the analysis of fluid-fluid queue as in \cite{bo2014}, however, we require information about the transient distribution of the fluid queue.
+In the context of the analysis of fluid-fluid queues as in \cite{bo2014}, however, we require information about the transient distribution of the fluid queue.
 Specifically, on the event that the second fluid is non-decreasing (non-increasing), we need to know the amount of fluid to have flowed into or out of the second level process, the time (or Laplace transform of time) taken to do so, and the position of the first fluid and phase at the time when the in-out process of the second fluid reaches a given height.
@@
 None of the existing literature gives exact expressions which have all the aforementioned properties.
-Even if we consider a simpler model where the rates \(r_i(x)\) are constant on intervals \(\calD_{k,i}\), then the existing results are not satisfactory.
-Perhaps the closest is the work of \cite{bean2009} who compute expressions for the Laplace transform with respect to the time take for the fluid to exit an interval, \(\calD_{k,i}\), say, on the even that fluid exits in some phase \(j\), given it started within the interval at some point \(x_0\) and in some phase \(i\).
+Even if we consider a simpler model where the rates \(r_i(x)\) are constant on intervals \(\calD_{k}\), then the existing results are not satisfactory.
+Perhaps the closest is the work of \cite{bean2009} who compute expressions for the Laplace transform with respect to the time take for the fluid to exit an interval, \(\calD_{k}\), say, on the even that fluid exits in some phase \(j\), given it started within the interval at some point \(x_0\) and in some phase \(i\).
 These expressions could perhaps be used to analyse a fluid-fluid queue where the rates \(r_i(x)=r_k\) for all \(i\in\calS\), and \(x\in\calD_{k,i}\).
@@
 
-An approximate method which is appropriate for the analysis of fluid-fluid queues has been proposed.
-The method of \cite{bo2013}, which we refer to as the uniformisation method as it is derived via uniformisation arguments, approximates a fluid queue by a discrete-time Markov chain.
+Approximation methods which are appropriate for the application to fluid-fluid queues have been proposed.
+The uniformisation method of \cite{bo2013} approximates a fluid queue by a discrete-time Markov chain.
 \cite{bo2013} approximate the fluid queue \(\{(\dot X(t),\varphi(t))\}\) by the quasi-birth-and-death-process \(\{(L(t),\varphi(t))\}\).
-In both processes, \(\{\varphi(t)\}\) is the same so there is no approximation in phase process.
+In both processes, \(\{\varphi(t)\}\) is the same so there is no approximation in the phase process.
 The level process \(\{L(t)\}\) approximates the level process \(\{\dot X(t)\}\).
-Specifically, \cite{bo2013} derive the level process such that the event that \(L(t)=k\) and \(\varphi(t)\) approximates the event \(\dot X(t)\in \calD_{k,i}\) and \(\varphi(t)=i\), where all the levels \(\{\calD_{k,i}\}_{k}\) have the same width, \(\Delta=y_{k+1}-y_k\).
+Specifically, \cite{bo2013} derive the level process such that the event that \(L(t)=k\) and \(\varphi(t)=i\) approximates the event \(\dot X(t)\in \calD_{k,i}\) and \(\varphi(t)=i\), where all the levels \(\{\calD_{k,i}\}_{k}\) have the same width, \(\Delta=y_{k+1}-y_k\).
 By considering arbitrarily small \(\Delta\), they prove that the distribution of the quasi-birth-and-death-process \(\{(L(t),\varphi(t))\}\) can be made arbitrarily close to the distribution of the fluid queue \(\{(\dot X(t),\varphi(t))\}\).
@@
 Of the methods considered in Chapter~\ref{sec: numerics}, the performance of the uniformisation approximation is the poorest and its rate of convergence is the slowest for the numerical experiments we conduct.
-However, due to the stochastic interpretation of the uniformisation approximation as a QBD, approximations to probabilities are guaranteed to be non-negative, and we can employ probabilistic techniques to analyse the approximation (as \cite{bo2013} did).
+%However, due to the stochastic interpretation of the uniformisation approximation as a QBD, approximations to probabilities are guaranteed to be non-negative, and we can employ probabilistic techniques to analyse the approximation (as \cite{bo2013} did).
 Moreover, the simplicity of the uniformisation scheme lends itself more easily to mathematical analysis.
@@
 
-It turns out that the uniformisation approximation of \cite{bo2013} is equivalent to a certain finite-element approximation which is a specific application of the \emph{discontinuous Galerkin (DG) method}.
-We introduce the discontinuous Galerkin method and its application to fluid queues, and fluid-fluid queues, in Chapter~\ref{ch:galerkin}.
+As mentioned earlier, the uniformisation approximation of \cite{bo2013} is equivalent to a certain application of the {discontinuous Galerkin (DG) method}.
 DG approximation schemes have found great success for approximating solutions of partial differential equations, however, to my knowledge, they have not been applied to fluid queues, or fluid-fluid queue to date (except for our paper \cite{blnos2022}, of which I am a co-author and the relevant parts of that work are included in this thesis).
-The DG scheme has the advantage that its rate of convergence is rapid for smooth problems, however, it can produce oscillatory approximations and result in negative probabilities.
-
-In Chapter~\ref{sec: construction and modelling} we attempt to derive a new approximation scheme which converges faster than the uniformisation scheme and still has a stochastic interpretation so that estimates of probability are guaranteed to be positive.
+%The DG scheme has the advantage that its rate of convergence is rapid for smooth problems, however, it can produce oscillatory approximations and result in negative probabilities.
+
+%In Chapter~\ref{sec: construction and modelling} we attempt to derive a new approximation scheme which converges faster than the uniformisation scheme and still has a stochastic interpretation so that estimates of probability are guaranteed to be positive.
 

### 2022-09-19 chapter2/chapter.tex
finito

@@
 	\begin{minipage}{0.8\textwidth}
-		\textit{Apart from Sections~\ref{sec: other applications} and \ref{sec: limiting and linearity}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notation, so that this chapter is consistent with the rest of the thesis.
+		\textit{Apart from Section~\ref{sec: other applications}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notation, so that this chapter is consistent with the rest of the thesis.
 I am a co-author of the paper \cite{blnos2022}.
@@
 \section{The partial differential equation}
-We start by introducing the PDE (assuming the PDE exists) from which we will extract the approximation to the generator \(\mathbb B\).
+We start by recalling the PDE (assuming the PDE exists) from which we will extract the approximation to the generator \(\mathbb B\).
 
@@
 It is convenient to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero.
-Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE.
+%Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE.
 Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE.
-For example, this may allow discontinuities with respect to \(x\) in the solution -- something which is ill-defined for the strong form.
+%For example, this may allow discontinuities with respect to \(x\) in the solution -- something which is ill-defined for the strong form.
 
@@
 It remains to approximate the \textit{flux}, \(f_j(x,t)\) at the cell edges \(y_k,\,k\in\mathbb Z\), so that we may evaluate the terms \([f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}\), \(r=1,...,p_k,\,k\in\mathbb Z\).
-This is the key for DG -- it joins the local approximations on each cell into a global approximation on the whole domain of approximation.
+%This is the key for DG -- it joins the local approximations on each cell into a global approximation on the whole domain of approximation.
 The flux is the instantaneous rate (with respect to time) at which density moves across the boundaries \(y_k,\,k\in\mathbb Z\).
@@
 Indeed, the DG scheme has the potential to answer many mathematical questions for fluid queues.
-However, one reason DG schemes may not have found such widespread use in the context of fluid queues is the difficulties surrounding discontinuous solutions, which we elaborate on in Section~\ref{sec: slope limiting} and demonstrate in Chapter~\ref{sec: numerics} (see also Section~5.6~\cite{nodalDGBook}), which we often want to consider.
+However, one reason DG schemes may not have found such widespread use in the context of fluid queues is the difficulties surrounding discontinuous solutions, which is something we often want to consider.
 Moreover, there are sometimes other techniques which can give exact solutions, for example exact expressions for transient quantities are derived in \cite{rs2003,ar2004,bean2005,bean2009}, and for the limiting distribution are derived in \cite{ajr2005,dasilva2005,lnp13,s2017}.
-
-Nonetheless, the DG scheme can be used to answer a very broad range of questions about fluid queues (such as transient distributions, hitting times on complex geometries, and of course analysis of fluid-fluid queues) which have no exact, readily computable solution.
-Moreover, we will numerically investigate the performance of the DG scheme (and other numerical schemes which we will develop) in Chapter~\ref{sec: numerics} which require numerical integration with respect to time.
+Nonetheless, the DG scheme can be used to answer a very broad range of questions about fluid queues (such as transient distributions, hitting times on complex geometries, and of course analysis of fluid-fluid queues) which have no exact, readily computable solution.% Moreover, we will numerically investigate the performance of the DG scheme (and other numerical schemes which we will develop) in Chapter~\ref{sec: numerics} which require numerical integration with respect to time.
 Hence, we now briefly introduce the numerical time integration scheme which we use in this thesis, as well as some issues pertaining to oscillatory approximations.
@@
 % APPLICATION TO INTERGRATING OVER TIME
-\section{Remarks on slope limiters, linear operators and application to fluid-fluid queues}\label{sec: limiting and linearity}
-The process of slope limiting, as described in Section~\ref{sec: slope limiting}, essentially amounts to a non-linear post-processing of the approximate solution after applying the operator \(\bs B\).
+% \section{Remarks on slope limiters, linear operators and application to fluid-fluid queues}\label{sec: limiting and linearity}
+% The process of slope limiting, as described in Section~\ref{sec: slope limiting}, essentially amounts to a non-linear post-processing of the approximate solution after applying the operator \(\bs B\).
 Thus, the slope-limited scheme arrives at an approximation which is the action of a non-linear operator applied to the initial condition.
 The fact that the approximate operator is non-linear makes solving the Riccati Equation~(\ref{eqn:RiccatiPsi}) difficult/impossible, whereas there are efficient techniques to solve matrix-Riccati equations \cite{bot08}.
-Thus, for the application to fluid-fluid queues slope limiting is not a viable option.
-We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{Y(t)\}\) first returns to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation.
+Thus, for the application to fluid-fluid queues, slope limiting is not a viable option.
+%We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{Y(t)\}\) first returns to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation.
 The application of the limiter/filter is dependent on the initial condition, so the post-processing needs to be done for every initial condition under consideration.
@@
 
-On a related note, the advantage of a linear approximation can also be useful if we want to approximate transient distributions of fluid queues for many initial conditions.
+% On a related note, a linear-operator can also be useful if we want to approximate transient distributions of fluid queues for many initial conditions.
 For example, consider solving (\ref{eqn: DG ODE w BCs 3}) numerically with a forward-Euler scheme to integrate over time.
 To advance the coefficients from time \(t\) to time \(t+h\) we compute
-\begin{align*}
-	&\bs a(t+h) = \bs a(t) + h \bs a(t)\bs Q = \bs a(t)\left[\bs I + h\bs Q\right].
-\end{align*} 
-Thus, the coefficients at time \(t_0=nh\) are given by 
-\begin{align*}
-	&\bs a(nh) = \bs a(0)\left[\bs I + h\bs Q\right]^n,
-\end{align*} 
-which is a linear transformation of the coefficients \(\bs a(0)\).
+% \begin{align*}
+% 	&\bs a(t+h) = \bs a(t) + h \bs a(t)\bs Q = \bs a(t)\left[\bs I + h\bs Q\right].
+% \end{align*} 
+% Thus, the coefficients at time \(t_0=nh\) are given by 
+% \begin{align*}
+% 	&\bs a(nh) = \bs a(0)\left[\bs I + h\bs Q\right]^n,
+% \end{align*} 
+% which is a linear transformation of the coefficients \(\bs a(0)\).
 Thus, if we determine the matrix \(\left[\bs I + h\bs Q\right]^n\), then we may apply it to any initial condition to approximate the solution at time \(t_0=nh\).
@@
 
-Another point to note is that limiters incur a computational cost.
+% Another point to note is that limiters incur a computational cost.
 There is a computational cost in reducing the solution to linear when oscillations are detected, and also in determining whether there are oscillations or not.
@@
 
-One advantage of the QBD-RAP (and uniformisation) scheme we develop in this thesis is that the resulting operator is positivity-preserving and linear.
+% One advantage of the QBD-RAP (and uniformisation) scheme we develop in this thesis is that the resulting operator is positivity-preserving and linear.
 % This raises two points.

### 2022-09-19 chapter6/chapter.tex
finito

@@
 In general, the QBD-RAP scheme performs similarly to the DG-lin-lim scheme, while the uniformisation scheme is a viable approximation method, but produces larger errors than the QBD-RAP and DG-lin-lim scheme, for the examples considered here.
-
-Furthermore, when slope limiting is used, the resultant approximation is at best linear around discontinuities.
+When slope limiting is used, the resultant approximation is at best linear around discontinuities.
 Moreover, there is a computational cost in applying a limiter (or filter), which could be significant, particularly if many initial conditions are to be considered.
@@
 
-The concept of slope-limiting originates in the context of numerically integrating (typically with respect to a \emph{time} variable) a so-called \emph{semi-discrete system}, as we do in Section~\ref{sec: transient approx}.
+%The concept of slope-limiting originates in the context of numerically integrating (typically with respect to a \emph{time} variable) a so-called \emph{semi-discrete system}, as we do in Section~\ref{sec: transient approx}.
 However, approximating the operator \(\mathbb\Psi\) of a fluid-fluid queue amounts to solving an implicit matrix-equation and there is no concept of numerical integration with respect to time.
@@
 
-We obtain approximations to the generator of the fluid queue \(\{(X(t),\varphi(t))\}\) via the DG, QBD-RAP and uniformisation schemes (as discussed at the start of this chapter, we do not apply slope limiting for this calculation).
+We obtain approximations to the generator of the fluid queue \(\{(X(t),\varphi(t))\}\) via the DG, QBD-RAP and uniformisation schemes.
+We do not apply slope limiting in this case as slope limiting results in a non-linear operator which means slope limiting schemes cannot be used to approximate the operator-Riccati equation by a matrix-Riccati equation.
 All the discretisations use a cell width of \(\Delta=0.4\).
@@
 The numerical experiments suggest that the DG-lin-lim scheme can perform well, and is similar to the performance of the QBD-RAP scheme, in the presence of discontinuities.
-For highly discontinuous problems (i.e.~problems with point masses) the QBD-RAP scheme can outperform the DG-lin-lim scheme.
+%For highly discontinuous problems (i.e.~problems with point masses) the QBD-RAP scheme can outperform the DG-lin-lim scheme.
 %For the application of the schemes to fluid-fluid queues, there is no obvious way to apply the concept of a slope limiter.
@@
 The QBD-RAP and uniformisation schemes do not produce oscillations and, of the two, the QBD-RAP scheme performs best.
-
-As we mentioned in the introduction to this chapter, in Sections~\ref{sec: recon num}, \ref{sec: comp}, \ref{sec:stat}, and \ref{sec: ffq num} slope limiting is not used as it is not applicable: in these cases slope limiting is practically equivalent to post-processing the solution by detecting oscillations, then projecting oscillatory regions of the solution onto a basis of linear functions, and making some corrections to the solution if it happens to be negative.
-
 Moreover, when computing approximations to the first-return operator of a fluid-fluid queue in Section~\ref{sec: ffq num}, the QBD-RAP and uniformisation schemes lead to an algorithm that is theoretically justified.
 In contrast, to-date there no theory surrounding the DG approximation to the same operator (let alone any positivity preserving variants of the DG scheme).
+
+%As we mentioned in the introduction to this chapter, in Sections~\ref{sec: recon num}, \ref{sec:stat}, and \ref{sec: ffq num} slope limiting is not used as it is not applicable: in these cases slope limiting is practically equivalent to post-processing the solution by detecting oscillations, then projecting oscillatory regions of the solution onto a basis of linear functions, and making some corrections to the solution if it happens to be negative.
 

### 2022-09-19 chapter7/chapter.tex
finito

@@
 \cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and limiting distribution of a fluid-fluid queue.
-
-Motivated by computation of the theoretical operators for stochastic fluid-fluid queues in \cite{bo2014}, this thesis investigates approximations method for the generator of stochastic fluid queues.
+Motivated by the computation of the theoretical operators for stochastic fluid-fluid queues in \cite{bo2014}, this thesis investigates approximations method for the generator of stochastic fluid queues.
 Cell-based approximation methods are particularly useful as they allow flexible partitioning of the approximate operators, which is required in the analysis of fluid-fluid queues.
@@
 The QBD-RAP scheme performed similarly to the linear DG approximation on a finer grid with a slope limiter (the DG-lin-lim scheme).
-The QBD-RAP scheme performed slightly better on problems with point masses, and the DG-lin-lim scheme performed slightly better on the smoother problems.
-However, the QBD-RAP scheme constructs an approximation to the generator of the fluid queue which is linear and this is advantageous for the application to fluid-fluid queues.
+%The QBD-RAP scheme performed slightly better on problems with point masses, and the DG-lin-lim scheme performed slightly better on the smoother problems.
+However, unlike the slope limited DG-lin-lim scheme, the QBD-RAP scheme constructs an approximation to the generator of the fluid queue which is linear and this is advantageous for the application to fluid-fluid queues.
 The uniformisation scheme performed better than the DG-lim scheme but worse than all the others.
@@
 In summary, this thesis investigated three main computational frameworks for the analysis of fluid queues and fluid-fluid queues, namely, the uniformisation, DG, and QBD-RAP schemes.
-The uniformisation scheme is from \cite{bo2013}.
 The DG method is well known, but its application to fluid and fluid-fluid queues has not been well-studied.

### 2022-09-19 front/abstract.tex
finito

@@
 \label{ch:abstract}
-A fluid-fluid queue is a stochastic fluid queue, where the driving process is a fluid queue itself.
+A Fluid queue is a piecewise-linear stochastic process where the driving process is a continuous-time Markov chain.
 Fluid queues provide a model for a single continuous performance measure of a system in the presence of a random environment.
 Fluid queues have found a wide variety of applications including risk processes, telecommunications, and environmental modelling, among others.
+A \emph{fluid-fluid queue} is a stochastic fluid queue, where the driving process is a fluid queue itself.
 Given the success of fluid queues it is plausible that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success.
@@
 This thesis provides approximations to fluid queues so that we can approximate the operators in \cite{bo2014}.
-It investigates three main approximation schemes; the DG scheme (Chapter~\ref{ch:galerkin}) which is a popular finite-element scheme, the uniformisation scheme of \cite{bo2013} which approximates a fluid queue by a continuous-time Markov chain (specifically, a quasi-birth-and-death-process (QBD)), and the QBD-RAP scheme which is a generalisation of a QBD to allow matrix exponential inter event times.
+It investigates three main approximation schemes; the DG scheme (Chapter~\ref{ch:galerkin}) which is a popular finite-element scheme, the uniformisation scheme of \cite{bo2013} which approximates a fluid queue by a continuous-time Markov chain (specifically, a quasi-birth-and-death-process (QBD)), and the QBD-RAP scheme which is a generalisation of a QBD to allow matrix exponential inter-event times.
 The QBD-RAP scheme is novel; we describe the construction of the scheme in Chapter~\ref{sec: construction and modelling}, and provide an analysis to show that it is convergent in Chapters~\ref{sec: conv} and \ref{ch: global results}.
 We demonstrate the effectiveness of the approximation schemes in Chapter~\ref{sec: numerics}, focussing on problems with discontinuous solutions.
-In general, we find that the DG scheme performs remarkably well for smooth problems, but can produce oscillations and negative probability estimates in the presence of discontinuities, the QBD-RAP approximation performs well in the presence of discontinuities, but does not perform as well as the DG scheme for smooth problems, the uniformisation scheme produces reliable approximations in the presence of discontinuities, but converges slowest for all problems.
+In general, we find that the DG scheme performs remarkably well for smooth problems, but can produce oscillations solutions and negative probability estimates in the presence of discontinuities, the QBD-RAP approximation performs well in the presence of discontinuities, but does not perform as well as the DG scheme for smooth problems, the uniformisation scheme produces reliable approximations in the presence of discontinuities, but its numerical convergence is slowest.

### 2022-09-19 intro/chapter.tex
finito

@@
 For example, \cite{anick1982} model a \emph{data handling switch} in which there are \(N\) information sources which asynchronously switch from ON to OFF, or OFF to ON.
-The phase process \(\varphi(t)\) models the number of data sources which are ON at time \(t\), and the rate at which data is received from a single source when the buffer is on is assumed (without loss of generality) to be 1.
-The switch stores data in a buffer and processes the data at rate \(c\), thus the net rate of change of the buffer at time \(t\) when there are \(r\) information sources which are ON is \(r-c\).
-The level process, \(X(t)\), models the amount of data in the buffer at time \(t\) and the rate at which \(X(t)\) changes when \(\varphi(t)=r\) is \(c_r=(r-c)\times 1(X(t)>0)\).
-The model described above could be used to aid in the analysis and design of the data handling switch, for example, by determining the minimum rate \(r\) for which the system is \emph{stable} in steady state, which practically means that the processing power of the switch is large enough that the amount of data in the buffer does not grow without bound.
+The phase process \(\varphi(t)\) models the number of data sources which are ON at time \(t\), and the rate at which data is received from a single source when the buffer is on is assumed to be 1 (without loss of generality).
+The switch stores data in a buffer and processes the data at rate \(c\), thus the net rate of change of the buffer at time \(t\), when there are \(r\) information sources which are ON, is \(r-c\).
+The level process, \(X(t)\), models the amount of data in the buffer at time \(t\), and the rate at which \(X(t)\) changes when \(\varphi(t)=r\) is \(c_r=(r-c)\times 1(X(t)>0)\).
+The model described above could be used to aid in the analysis and design of the data handling switch, for example, by determining the minimum processing rate \(c\) for which the system is \emph{stable} in steady state.
+In practical terms, the minimum processing rate \(c\) determines the minimum processing power of the switch so that it is sufficient to prevent the amount of data in the buffer growing without bound.
 
-The success of fluid queues largely lies in their mix of flexibility, any (finite) number of phases can be used, and their parsimony and analytic tractability which enables effective transient and stationary analysis \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018,bnp2018}.
+The success of fluid queues largely lies in their mix of flexibility, any (finite) number of phases can be used, and their parsimony and analytic tractability which enables effective transient and stationary analysis (such as the analyses \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018,bnp2018}).
 Fluid queues are relatively well studied.
@@
 In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\), so \((X(t),\varphi(t))\) determines the rate at which \(\{Y(t)\}\) moves.
-The addition of the second level process \(\{Y(t)\}\) greatly increases the modelling potential.
-For example, \cite{hydro,bo2014} describe a model for a hydroelectric power generator which can be operated `on design' and `off design'.
+The addition of the second level process \(\{Y(t)\}\) increases the modelling potential.
+For example, \cite{hydro} describe a fluid queue model for a hydroelectric power generator which can be operated `on design' and `off design'.
 When the system operates `off design' the system is less efficient and there is more wear, but may be used to optimise overall system performance.
@@
 
-The addition of the second level process \(\{Y(t)\}\) greatly increases the flexibility and modelling potential.
-However, the added flexibility comes at a cost to the in the analysis of the model.
-For both fluid-fluid queue and classical fluid queues, performance measures such as the stationary distribution, and the Laplace-transforms of hitting times and return times, can be derived in terms of the infinitesimal generator of the modulating process.
-For fluid queues the modulating process is a CTMC for which the infinitesimal generator is a matrix.
+The addition of the second level process \(\{Y(t)\}\) increases the modelling potential, however, it comes at a cost to the simplicity of the analysis of the model.
+For both fluid-fluid queues and classical fluid queues, performance measures, such as the stationary distribution or the Laplace-transforms of hitting times and return times, can be derived in terms of the infinitesimal generator of the modulating process.
+For fluid queues the modulating process is a CTMC and the infinitesimal generator is a matrix.
 Thus, expressions for the performance measures of fluid queues are matrix expressions which are readily computable.
-For fluid-fluid queue the modulating processes is a fluid queue for which the infinitesimal generator is a matrix of differential operators.
+For fluid-fluid queues the modulating processes is a fluid queue and the infinitesimal generator is a matrix of differential operators.
 Thus, expressions for the performance measures of fluid-fluid queues are matrices of differential operators which, in all but the simplest of cases, are not readily computable.
+This thesis investigates approximations to these differential operator expressions.
 
-The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
+The analysis of fluid-fluid queues in \cite{bo2014} is, in principle, similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the generator of the driving fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
 For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used.
-To this end, to date, two possible discretisations have been suggested.
-Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise the operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process.
+To this end, two discretisations have been suggested.
+Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise the operator, while \cite{bo2013} take a stochastic modelling approach to approximate the fluid queue by a quasi-birth-and-death (QBD) process.
 The QBD approximation of \cite{bo2013} is derived via a uniformisation argument, so we refer to it as the uniformisation scheme throughout this thesis.
 Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations.
-It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the former; the uniformisation scheme can be viewed as the simplest DG scheme where cells are all have equal width, the operator is projected onto a basis of piecewise constant functions, and an upwind flux is used to approximate the flow of mass between cells.
+It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the DG scheme; the uniformisation scheme can be viewed as the simplest DG scheme where cells all have equal width, the operator is projected onto a basis of piecewise constant functions, and an upwind flux is used to approximate the flow of mass between cells.
 This is also equivalent to a finite-volume scheme.
@@
 One drawback of the uniformisation scheme is that convergence can be relatively slow compared to higher-order DG schemes.
-However, in the context of approximating fluid queues, one advantage of the uniformisation scheme and, equivalently, a DG scheme with constant basis functions, is that the uniformisation scheme guarantees probabilities computed from the approximation are positive \citep[Section~3.3]{koltai2011}.
+However, in the context of approximating fluid queues, one advantage of the uniformisation scheme is that the uniformisation scheme guarantees probabilities computed from the approximation are positive \citep[Section~3.3]{koltai2011}.
 One justification for the positivity preserving property of the uniformisation scheme is from its interpretation as a stochastic process.
+
 For higher order DG schemes there is no such interpretation and positivity is not guaranteed \citep[Section~3.3]{koltai2011}.
-Moreover, higher-order DG approximation schemes may produce negative and oscillatory solutions, particularly when discontinuities or steep gradients are present.
+Higher-order DG approximation schemes may produce negative and oscillatory solutions, particularly when discontinuities or steep gradients are present.
 This is problematic for probabilistic problems as we know that probabilities must be positive.
@@
 
-Depending on the context, filtering of the approximate solution to remove oscillations may not necessarily guarantee a strictly non-negative approximation, or may result in severe smearing of the solution at discontinuities or regions with steep gradients \citep[Section~5.6.1]{nodalDGBook}.
+Depending on the context, filtering the approximate solution to remove oscillations may not necessarily guarantee a strictly non-negative approximation, or may result in severe smearing of the solution at discontinuities or regions with steep gradients \citep[Section~5.6.1]{nodalDGBook}.
 Moreover, filtering requires us to make a trade-off between filtering enough of the oscillations away while maintaining sufficient accuracy -- a choice which may not be obvious \emph{a priori}.
 Slope limiting does guarantee positivity but reduces the approximation to linear where oscillations in the approximate solution are detected \citep[Section~5.6.1]{nodalDGBook}.
-Furthermore, limiting and filtering do not distinguish between natural oscillations, which are a fundamental feature of the solution and spurious oscillations which are caused by the approximation scheme, and they may remove both from the approximation.
+Limiting and filtering do not distinguish between natural oscillations, which are a fundamental feature of the solution, and spurious oscillations, which are caused by the approximation scheme, and they may remove both from the approximation.
 This can lead to an unnecessary loss of accuracy in the approximation (see \citep{nodalDGBook},~Example~5.8).
-Other drawbacks of these methods is that they result in non-linear operators, so we cannot use matrix algebra for computations, and they incur an additional computational cost on top of the DG scheme.
 
+Further, slope limiting is a non-linear operator which means it cannot be represented as matrices and this hinders its application to fluid-fluid queues.
+%In addition, filtering and limiting also incur an additional computational cost on top of the DG scheme.
 % In the context of approximating the first return operator of a fluid-fluid queue, the application of a slope limiter is not possible since we need the approximated operators to be linear. 
+In the process of approximating the first return operator of a fluid-fluid queue we approximate an operator-Riccati equation with a matrix-Riccati equation by substituting in matrix approximations to the operators.
+The operators are related to the generator of the fluid queue and approximations to them can be constructed via the DG scheme \citep{blnos2022}.
+We then solve the matrix-Riccati equation via an iterative procedure \citep{bean2005b,blnos2022} and the solution is an approximation to the first-return operator of the fluid-fluid queue.
+Since the operators constructed with limiters are non-linear (not matrices) then we cannot apply them in this case.
+%As an alternative, we could apply a post-processing based on slope limiters, or filtering, but we do not consider this approach in this thesis.
+%Both of these post-processing methods incur a computational cost and need to be applied on a case-by-case basis depending on the initial condition.
 
-In approximating the first return operator of a fluid-fluid queue we first approximate an operator-Riccati equation by a matrix-Riccati equation by substituting in matrix approximations to the infinitesimal generator of the fluid queue which are constructed via the DG scheme \citep{blnos2022}.
-We then solve the matrix-Riccati equation via an iterative procedure \citep{bean2005b,blnos2022}.
-Since the operators constructed with limiters are non-linear (not matrices) then we cannot apply limiters in this case.
-Alternatively, we could apply a post-processing based on slope limiters, or filtering.
-Both of these post-processing methods incur a computational cost and need to be applied on a case-by-case basis depending on the initial condition.
-
-Furthermore, for the DG method there is a limited theoretical backing as to why we may approximate the operator-Riccati equation with the matrix-Riccati equation other than it is a sensible thing to do, and it seems to work in practice.
-In the case of the QBD approximation of \cite{bo2013} the approximate matrix-Riccati equation is more justified as in it can be derived by considering the first-return probabilities for a fluid queue driven by the approximating QBD where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\) \citep{bean2005}.
+The approach described above relies on the intuition that substituting the DG approximation in place of the true operators in the Riccati equation is a reasonable thing to do, however there is limited theory (to the authors' knowledge) on why this approach works.
+In contrast, for the QBD approximation of \cite{bo2013} the approximate matrix-Riccati equation is theoretically justified and in it can be derived by considering the first-return probabilities for a fluid queue driven by the approximating QBD, where the rates of the fluid are determined by a piecewise constant approximation to \(r_i(x)\) \citep{bean2005}.
 
 Hence, for the approximation of fluid queues with application to fluid-fluid queues, on the one hand we can use the DG scheme which can produce high-order approximations when the problem is smooth, but can display spurious oscillations and result in negative probability estimates in the presence of discontinuities.
-While, on the other hand, we can use the uniformisation scheme which constructs in a linear operator to approximate the generator and guarantees non-negative probability estimates, but convergence is relatively slower.
+On the other hand, we can use the uniformisation scheme for the approximation which guarantees non-negative probability estimates and has a sound theoretical justification, but convergence is relatively slower.
 
 Motivated by this, this thesis derives a new approximation to a fluid queue.
-The construction of the approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributions to approximate the sojourn time of the fluid level in a given interval on the event that the phase of the fluid is constant.
-The sojourn time in a given interval on the event that the phase of the fluid is constant is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution, so in this sense, the best approximation to this deterministic sojourn time.
-Thus, it appears that the approximation of \cite{bo2013} is the best-possible Markov chain approximation.
-Recently, there has been much work on a class of concentrated matrix exponential distributions \citep{hhat2020} which are postulated to be the least-variable matrix exponential distributions of a given order.
+The construction of the approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributed random variables to approximate the sojourn time of the fluid level in a given interval on the event that the phase of the fluid is constant.
+On the event that the phase of the fluid is constant, the sojourn time in a given interval is a deterministic event.
+It is known that the Erlang distribution is the least-variable Phase-type distribution of a given order.
+In this sense, in the class of all Phase-type distributions of a given order, an Erlang-distribution is the best-possible approximation to the distribution of the deterministic sojourn time.
+Thus, it appears that the approximation of \cite{bo2013} is the best-possible Markov chain approximation to a fluid queue.
+Recently, \cite{hhat2020} have developed the class of \emph{concentrated matrix exponential distributions} which are postulated to be the least-variable matrix exponential distributions of a given order.
 Matrix exponential distributions generalise Phase-type distributions; they have the same functional form, without the restriction that the distribution has an interpretation in terms of the absorption time of a continuous-time Markov chain.
 A class of stochastic processes, known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \citep{bn2010}, extends QBDs, which have Phase-type inter-event times, to allow matrix-exponentially distributed inter-event times.
-Thus, by using matrix exponential distributions, we attempt to construct a QBD-RAP which better captures the dynamics of the fluid queue than the QBD approximation in \cite{bo2013}, while retaining a stochastic interpretation.
+Thus, by using concentrated matrix exponential distributions, one aim of this thesis is to construct a QBD-RAP to approximate a fluid queue with the idea that it will be more accurate than the QBD approximation in \cite{bo2013}, while still retaining a stochastic interpretation which ensures positivity and aids in the application to fluid-fluid queues.
 
-As the QBD-RAP has a stochastic interpretation the approximations it produces are guaranteed to have non-negative density functions.
-Further, for the QBD-RAP (and uniformisation) scheme non-negativity is guaranteed for any initial condition without any further computation or post-processing.
-For the approximation of fluid-fluid queues, the matrix-Riccati equation we solve to approximate the first-return distribution of a fluid-fluid queue can be derived by considering the first-return probabilities for a RAP-modulated fluid queue driven by the QBD-RAP \citep{p2019,bgnp2021}.
-The stochastic interpretation can also be leveraged to aid in the analysis of the approximation.
-Another attractive property is that the approximations to operators obtained via the QBD-RAP (and uniformisation) scheme are linear which is necessary for the application to approximating the first-return operator of fluid-fluid queues and may also provide other computational and analytical benefits.
+As a QBD-RAP has a stochastic interpretation the resulting approximations are guaranteed to have non-negative density functions and give non-negative estimates of probabilities.
+This non-negativity is guaranteed for any initial condition without any further computation or post-processing.
+Further, for the application to fluid-fluid queues, the matrix-Riccati equation that we solve to approximate the first-return distribution can be derived by considering the first-return probabilities for a RAP-modulated fluid queue driven by the QBD-RAP \citep{p2019,bgnp2021} and hence it has a sound theoretical justification.
+In addition, the stochastic interpretation can be leveraged to aid in the analysis of the approximation.
+%Another attractive property is that the approximations to operators obtained via the QBD-RAP scheme are linear, which is necessary for the application to approximating the first-return operator of fluid-fluid queues and may also provide other computational and analytical benefits.
 
 The structure of this thesis is as follows.
-The next chapter, Chapter~\ref{ch:math}, is dedicated to mathematical preliminaries and introduces the main mathematical objects and tools which we will need.
-In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way that makes clear exactly how the approximations we develop correspond to the theoretical operators.
+The next chapter, Chapter~\ref{ch:math}, is dedicated to mathematical preliminaries and introduces the main mathematical objects and tools that we will need.
+In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way that makes clear exactly how the approximations correspond to the theoretical operators.
 Chapter~\ref{ch:math} also introduces the discontinuous Galerkin (DG) method for approximating solutions to differential equations.
 Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:math} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and negativity.
-Chapter~\ref{ch:math} also introduces the existing literature and gives context to this thesis.
+Chapter~\ref{ch:math} also introduces the existing literature and gives more context to this thesis.
 
-Chapter~\ref{ch:galerkin} demonstrates a way that we can use the discontinuous Galerkin method to approximate certain operators and distributions of fluid and fluid-fluid queues.
+Chapter~\ref{ch:galerkin} demonstrates a way to use the discontinuous Galerkin method to approximate certain operators and distributions of fluid and fluid-fluid queues.
 The DG method provides an effective tool for approximating performance measures of fluid-fluid queues.
-However, the DG method can exhibit oscillations in the presence of discontinuities and these oscillations can cause approximations to certain probabilities to be negative.
+However, it can exhibit oscillations in the presence of discontinuities and these oscillations can cause approximations to certain probabilities to be negative.
 This motivates the next chapter.
 
-In the next chapter, Chapter~\ref{sec: construction and modelling}, we develop a new QBD-RAP approximation scheme which, due to its interpretation as a stochastic process, ensures all estimates of probabilities will be non-negative.
+In Chapter~\ref{sec: construction and modelling}, we develop a new QBD-RAP approximation scheme which, due to its interpretation as a stochastic process, ensures all estimates of probabilities will be non-negative.
 The chapter takes a stochastic modelling approach to developing the approximation scheme.
@@
 Once we have established the new approximation scheme, we then prove its convergence in Chapters~\ref{sec: conv}~and~\ref{ch: global results}.
-Ultimately, Chapter~\ref{sec: conv} proves that, on the event that the QBD-RAP has not-yet seen an \emph{orbit restart epoch} (which is much like a change of level, but not always), certain Laplace transforms with respect to time of the QBD-RAP scheme converge to corresponding Laplace transforms with respect to time of the fluid queue.
+In Chapter~\ref{sec: conv} we consider the QBD-RAP up until the first \emph{orbit restart epoch} (which is the same as a change of level, except in the case that the process hits a boundary and is reflected).
+Ultimately, Chapter~\ref{sec: conv} proves that, on the event that an \emph{orbit restart epoch} has not-yet occurred, certain Laplace transforms with respect to time of the QBD-RAP scheme converge to corresponding Laplace transforms with respect to time of the fluid queue.
 This is a local convergence result as it relates only to convergence to the fluid queue in an interval.
 
-Chapter~\ref{ch: global results} then looks to extend the local convergence results of Chapter~\ref{sec: conv} to a global convergence result on the whole domain of approximation.
+Chapter~\ref{ch: global results} then extends the local convergence results of Chapter~\ref{sec: conv} to a global convergence result on the whole domain of approximation.
 Chapter~\ref{ch: global results} uses more traditional Markov process arguments which rely on properties such as the strong Markov property, time-homogeneity and the law of total probability.
-In Chapter~\ref{ch: global results} we first consider the discrete-time process which is the process embedded in the QBD-RAP at times when the QBD-RAP sees an orbit restart epoch.
+In Chapter~\ref{ch: global results} we first consider the discrete-time process which is the process embedded in the QBD-RAP at orbit restart epochs.
 We prove that the embedded process converges in distribution to a corresponding embedded process of the fluid queue.
 The rest of Chapter~\ref{ch: global results} proves a global convergence result.
-The main result is Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (with respect to the spatial and temporal variable) to the distribution of the fluid queue.
+The main result is Theorem~\ref{thm: big thm} which states that the QBD-RAP approximation scheme converges weakly (with respect to the spatial and temporal variables) to the distribution of the fluid queue.
 
@@
 Given that the QBD-RAP scheme was developed so that it guarantees positivity of the approximation, we largely focus on problems with discontinuities.
+Chapter~\ref{sec: numerics} can be read independently of Chapters~\ref{sec: conv}~and~\ref{ch: global results}, which are somewhat technical and could be skipped on a first-pass if the reader wishes.
 
@@
 The rest of the appendices contain technical results relating to proving the convergence of the QBD-RAP approximation scheme.
-Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operators} introduced as part of the QBD-RAP scheme have the properties we claim in proving convergence in Chapter~\ref{sec: conv}.
+Appendix~\ref{appendix: sec: 2} proves that the \emph{closing operators} introduced as part of the QBD-RAP scheme have the properties we claim they do when we are proving convergence in Chapter~\ref{sec: conv}.
 Appendix~\ref{app:extend conv} extends some results from Chapter~\ref{sec: conv} to a setting which requires slightly less computation.

### 2022-09-19 summary/summary.tex
finito

@@
 \maketitle
-The stochastic fluid-fluid model (SFFM) is a Markov process {\(\{(X_t,Y_t,\varphi_t)\}_{t\geq 0}\),} where {\(\{\varphi_t\}_{t\geq 0}\)} is a continuous-time Markov chain, the first fluid, {\(\{X_t\}_{t\geq 0}\),} is a classical stochastic fluid process driven by {\(\{\varphi_t\}_{t\geq0}\),} and the second fluid, {\(\{Y_t\}_{t\geq0}\),} is driven by the pair {\(\{(X_t,\varphi_t)\}_{t\geq 0}\).} Operator-analytic expressions for the stationary distribution of the SFFM, in terms of the infinitesimal generator of the process {\(\{(X_t,\varphi_t)\}_{t\geq 0}\),} are known \cite{bo2014}.
-However, these operator-analytic expressions do not lend themselves to direct computation.
+The stochastic fluid-fluid model (SFFM) is a Markov process {\(\{(X_t,Y_t,\varphi_t)\}_{t\geq 0}\),} where {\(\{\varphi_t\}_{t\geq 0}\)} is a continuous-time Markov chain, the first fluid, {\(\{X_t\}_{t\geq 0}\),} is a classical stochastic fluid process driven by {\(\{\varphi_t\}_{t\geq0}\),} and the second fluid, {\(\{Y_t\}_{t\geq0}\),} is driven by the pair {\(\{(X_t,\varphi_t)\}_{t\geq 0}\).} Operator-analytic expressions for performance measures of fluid-fluid queues, such as the stationary distribution and first return distribution of the SFFM, have been derived \cite{bo2014}.
+These performance measures are in terms of the infinitesimal generator of the process {\(\{(X_t,\varphi_t)\}_{t\geq 0}\)}, which is a differential operator.
+Thus, the operator-analytic expressions do not lend themselves to direct computation.
 
-In this thesis we investigate approximations to the infinitesimal generator of the process {\(\{(X_t,\varphi_t)\}_{t\geq 0}\)} so that one may approximate the fluid-fluid performance measures derived in \cite{bo2014}.
+In this thesis we investigate approximations to the infinitesimal generator of the process {\(\{(X_t,\varphi_t)\}_{t\geq 0}\)} so that one may approximate the performance measures derived in \cite{bo2014}.
 As a first step we first introduce modified versions of the operators derived in \cite{bo2014} which are partitioned so that we can directly correspond elements of the true and approximated operators.
 
-Perhaps the simplest approximation is to approximate the continuous fluid process by a discrete-time Markov chain.
-This was considered in \cite{bo2013} and results in a quasi-birth-and-death process (QBD).
-Another approach is to use the finite-element method -- in Chapter~2 we apply the discontinuous Galerkin (DG) scheme.
-It turns out that the simplest DG method is equivalent to the Markov chain approximation of \cite{bo2013}.
-In some circumstances the DG method can result in approximations to the cumulative distribution function (CDF) which oscillate and are non-monotonic.
-This is clearly undesirable as we know that CDFs must be monotonic and non-decreasing.
-Due to its interpretation as a stochastic process, the Markov chain approximation of \cite{bo2013} ensures that the resulting approximations to the CDF are monotonic and non-decreasing.
-However, the rate of convergence of the Markov chain approximation of \cite{bo2013} is, in general, much slower than the DG scheme.
+Perhaps the simplest approximation is to approximate the continuous fluid process \(\{(X_t,\varphi_t)\}_t\) by a discrete-time Markov chain.
+This was considered in \cite{bo2013} where they use a quasi-birth-and-death process (QBD) to approximate \(\{(X_t,\varphi_t)\}_t\).
+Another approach is to use a finite-element method to approximate the infinitesimal generator of \(\{(X_t,\varphi_t)\}_t\).
+In Chapter~2 we show how we can use the discontinuous Galerkin (DG) method to derive approximations to the infinitesimal generator of \(\{(X_t,\varphi_t)\}_t\), and then use these to approximate the operator-analytic expressions from \cite{bo2014}.
+When discontinuities or steep gradients are present in the distribution of \(\{(X_t,\varphi_t)\}_t\), the DG method can result in approximations to the cumulative distribution function (CDF) which oscillate and, in the worst cases, are non-monotonic and can take values outside the interval \([0,1]\).
+This is clearly undesirable.
+In contrast, the QBD approximation of \cite{bo2013} does not suffer the same problems due to its interpretation as a stochastic process, which ensures that the resulting approximations to the CDF are well-behaved.
+However, the rate of convergence of the QBD approximation is, in general, much slower than the DG approximation.
 
 Motivated by this, the thesis proceeds to derive a new approximation to a fluid queue.
-The approximation is inspired by the observation that the Markov chain approximation of \cite{bo2013} effectively uses Erlang distributions to model the sojourn time in a given interval on the event that the phase of the fluid is constant.
-The sojourn time in a given interval on the event that the phase of the fluid is constant is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution so, in this sense, the best approximation to this deterministic sojourn time.
-Thus, it appears that the approximation of \cite{bo2013} is the best-possible Markov chain approximation.
-Recently, there has been much work on a class of concentrated matrix exponential distributions \cite{hhat2020} which are postulated to be the least-variable matrix exponential distribution.
+The approximation is inspired by the observation that the QBD approximation of \cite{bo2013} effectively uses an Erlang distribution to model the sojourn time of \(\{(X_t,\varphi_t)\}_t\) in a given interval on the event that the phase of the fluid is constant.
+The aforementioned sojourn time is a deterministic event, and it is known that the Erlang distribution is the least-variable Phase-type distribution so, in this sense, the Erlang distribution is the best approximation to the distribution of this deterministic sojourn time.
+Thus, we argue that the approximation of \cite{bo2013} is the best-possible Markov chain approximation to the fluid queue \(\{(X_t,\varphi_t)\}_t\).
+In light of this, we look to a broader class of models known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \cite{bn2010} to construct our approximation.
+QBDs have Phase-type distributed inter-event times.
+QBD-RAPs extend QBDs to allow \emph{matrix-exponentially} distributed inter-event times.
 Matrix exponential distributions have the same functional form as Phase-type distributions without the restriction that the distribution has an interpretation in terms of the absorption time of a continuous-time Markov chain.
-A class of stochastic processes, known as quasi-birth-and-death-processes with rational-arrival-process components (QBD-RAPs) \cite{bn2010}, extend QBDs, which have Phase-type inter-event times, to allow matrix-exponentially distributed inter-event times.
-Thus, by using matrix exponential distributions, Chapter~3 attempts to construct a QBD-RAP which better captures the dynamics of the fluid queue than the QBD approximation in \cite{bo2013}.
-As the QBD-RAP has a stochastic interpretation then the approximations of to CDFs it produces are guaranteed to be monotonic and non-decreasing.
+Recently, there has been much work on a class of \emph{concentrated matrix exponential distributions} \cite{hhat2020} which are postulated to be the least-variable matrix exponential distribution.
+Thus, by using concentrated matrix exponential distributions, Chapter~3 construct a QBD-RAP which better captures the dynamics of the fluid queue than the QBD approximation in \cite{bo2013}.
+As the QBD-RAP has a stochastic interpretation then the approximations of CDFs it produces are guaranteed to be monotonic, non-decreasing and take values in \([0,1]\).
 
-The thesis then moves on to mathematical analysis of the QBD-RAP process.
+The thesis then moves on to mathematical analysis of the constructed QBD-RAP process.
 Chapter~4 proves a convergence of the QBD-RAP process to the fluid queue up to the time that the fluid queue leaves a given interval.
 Chapter~5 then uses the results of Chapter~4 to prove a global convergence results of the QBD-RAP approximation scheme to the fluid queue.
-Chapter~4 relies more on matrix-exponential specific arguments, while Chapter~5 uses more traditional Markov process arguments, such as partitioning on the position of the process at convenient stopping times, then exploiting the strong Markov property and time homogeneity.
+Chapter~4 uses more matrix-exponential-specific arguments to show its convergence result, while Chapter~5 uses more traditional Markov process arguments, such as partitioning on the position of the process at convenient stopping times, then exploiting the strong Markov property and time homogeneity.
 
-Chapter~6 numerically investigates the performance of the DG, QBD-RAP and Markov chain approximations.
+Chapter~6 numerically investigates the performance of the DG, QBD-RAP and QBD approximations.
 The numerics demonstrate that, if the solution is smooth enough, then the discontinuous Galerkin method is highly-effective and displays rapid convergence to the solution as the number of basis functions is increased.

### 2022-12-30 reviews/response.tex
first iteration of changes for reviewer 2 and response to reviewer 2 comments (to do: reviewer 1 changes and responses)

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{soul}
\usepackage{etoolbox}
\AtBeginEnvironment{quote}{\itshape}

\title{Response to Thesis Reviews}
\author{Angus Lewis}

\begin{document}
\maketitle
The reviewers comments are in plain text and not indented. 
\begin{quote}
    My comments follow are are in indented and are in italics. 
\end{quote}
Text that is \st{struck-out} has already been addressed.

\section{Reviewer 2}

\textbf{2 About the main sections}

Section 2 is of slightly uneven level of detail. Maybe some of the more basic topics are not even necessary. I also suggest re-ordering the subsections from basic to advanced
\begin{quote}
    \begin{itemize}
        \item I'm not sure what the reviewer means by "uneven level of detail" -- sure, some sections are more technical than others. But I think that's just the nature of the topics and the they are presented in the detail required to understand the relevant parts of the thesis.

        \item I don't think removing basic topics is the answer. I'd rather explain more and the reader can skip bits they don't think they need.

        \item I like the order of the introduction. It builds in a logical order: some fundamental mathematical concepts -- just enought to get going, then I introduce the fluid and fluid-fluid queue, then tools for their approximations, then other sundry mathematical concepts used in the thesis. I don't think suggested order would add value.  
    \end{itemize}
\end{quote}

Section 3 presents the discontinuous Galerkin (DG) method. Unfortunately, no explicit error bounds are available.

Section 4 presents the QBDRAP method. It uses concentrated matrix exponential (CME) functions. It is nice to see that certain properties of CME functions (such as identical real part of all eigenvalues) are utilised.

Sections 5 and 6 proves convergence for QBDRAP. These sections are rather long and heavy, but the arguments are mostly familiar from the literature. Everything is properly referenced. Sections 5 and 6 could be merged into a single section. Section 5 is difficult to follow; maybe arranging the calculations in Sections 5.2 and 5.3 into lemmas or steps would result in a better structure.
\begin{quote}
    \begin{itemize}
        \item Yes, these sections are long and heavy. I think it's hard to avoid. We have talked about this a lot. I am happy with the content of these sections, but I know Peter objects to including some of the details which make the presentation more complex. 
        \item Happy to merge 5 and 6. What do you think? I seperated them because they used distinct techniques, but they ultimately combine to prove a single result. 
        \item Item, happy to add more steps and signposts in Sections 5.2/5.3 or convert them to Lemmas -- perhaps structuring as Lemmas is the way to go as it condeses the important points into a single statement and leaves details to a proof, allowing the reader to forget about the details.
    \end{itemize}
\end{quote}

Section 7 presents numerical investigations for a fairly broad selection of examples. Despite the lack of explicit error bounds, we see DG performing well in many cases. Fluid-fluid models are welcome. A general remark: CME distributions only get really concentrated for high orders (dimension). Ideally, I would recommend using at least an order of 30, although I understand there are computational limitations. Either way, it should not be surprising that their performance is not great for orders below 10 or so.
\begin{quote}
    \begin{itemize}
        \item   Not sure what ``fluid-fluid models are welcome'' means.
        \item I went up to order 21 in the numerical investigations. For some of the more trivial examples in this section I could use higher-order approximations. However, for the interesting examples, simulation is required to measure accuracy and this is computationally expensive. I'm already doing \(5\times 10^{10}\) realisations! To me, it seems unlikely that investigation of higher-order examples would reveal much more insight -- my intuition is that the error decay of the QBD-RAP method is exponential in the order of the ME used. 
    \end{itemize}
\end{quote}

\textbf{3 List of minor remarks and typos}

Minor remarks/typos:
\begin{itemize}
\item general: definitions are sometimes in italics, sometimes not; defining formulas are
sometimes denoted by :=, sometimes by =
\begin{quote}
    I'm lazy. Do I need to fix this?
\end{quote}
\item \st{general: in formulas, functions like diag, sign etc. should be in mathrm font}
\item page 7 bottom, stopping time definition
\begin{quote}
    Not sure what's wrong with it...
\end{quote}
\item \st{page 13: 'otherwise it is reflected' - I understand why the word 'reflected' is used here, but usually a reflecting boundary condition means evolving with -ci instead of ci. Here, this is simply an absorbing boundary condition, with transition upon reaching the boundary.}
\item page 15: Why is the Laplace domain variable sometimes \(s\), sometimes \(\lambda\)?
\begin{quote}
    I have change this on page 15. Should I change it throughout?
\end{quote}
\item \st{page 17, bottom formula: extra ) bracket}
\item \st{page 18: 'which is only possibly'  'which is only possible'}
\item page 19: 'Conditions on fluid-fluid queues to ensure that (2.20) will be differentiable with respect to y are not known, and here we just assume that it is.' - this seems like an innocent assumption. Why don't we have this at least for some nice cases? A few more words on this would be welcome.
\begin{quote}
    This is a large can of worms... how much do we want to open it?
\end{quote}
\item \st{page 20, third row from top: some ) brackets are missing}
\item \st{page 20: a point masses to a point mass}
\item \st{page 30, before (2.35): extra ) bracket in ()}
\item page 35: Section 2.5.5. about accuracy is indeed very brief and seems highly non-
rigorous.
\begin{quote}
    Hnnn. Yep, it is. I could tighten this up and quote the result from the reference I cite, but it requires some more preliminaries which are not relevant to the rest of the thesis. Thoughts? 
\end{quote}
\item \st{page 36, (2.39): m + 2 to m2 in the second sum}
\item \st{page 36, bottom line: m + 1 to m1 in the first sum}
\item \st{page 37: the term representation is used before its definition. The correct order
would also help out Cor. 2.6.}
\item \st{page 37, near bottom: possibility defective to possibly defective I recommend a
thorough read-through of Section 2.6.1.}
\item \st{page 39: 'the process may not actually jump at these times' - reword to something
like 'the process does not necessarily jump at these times'}
\item \st{page 42, Section 2.8: is this section even necessary? Laplace-Stieltjes transform has been used heavily earlier.}
\begin{quote}
    I have moved this section to section 2.3.
\end{quote}
\item \st{page 48, near the top: 'Fubini's Theorem us to' to 'Fubini's Theorem allows us to'}
\item \st{page 53: 'The DG method conserves probability' - so any quantity from [0,1] remains
in [0,1] after the approximation, or does it stay completely unchanged? Or does this
mean something else? Maybe a reference to Cor. 3.1 on page 60? In any case, this
should be clarified here.}
\begin{quote}
    I have clarified that it conerves total probability of the system and added a reference to cor 3.1.
\end{quote}
\item \st{page 55: W k or Wk? Both are used, but denote the same space. Also, it would be
nice to recall the definition of Wk here.}
\begin{quote}
    Fixed and recalled.
\end{quote}
\item page 55: orthogonal complement spaces and such could be introduced earlier, in
Section 2.
\begin{quote}
    They are only used here, in this specific place, and only takes up less than 2 lines. Do I need to move it?
\end{quote}
\item \st{page 56: There are different choices for the flux - rather, for the choice of approxi-
mation for the flux, right?}
\item page 60, middle of page: overlong line
\begin{quote}
    Fixed? Check once compiled.
\end{quote}
\item \st{page 60, bottom formula, first line: the 'or' part should be re-formulated (e.g.
something like k = l 6 = 0 or K)}
\item \st{page 70, middle: 'we can show even show' to 'we can even show'}
\item \st{page 70: 'on the event that phi(t) is constant, then X(t) moves deterministically' -
'then' is unnecessary}
\item \st{page 80, near top: 'Other choices are possible' ... 'are other possible choices' -
unnecessary repeat}
\item \st{page 88, top line: 'events epochs' to 'event epochs'}
\item \st{page 102, bottom: use the same function (e.g. g(x)) in both formulas}
\item page 104: (5.10) and (5.11) seem to be a single formula broken into two parts
unnecessarily.
\begin{quote}
    Check once compiled.
\end{quote}
\item \st{page 106, near bottom: 'in terms of the of first return matrices' to 'in terms of the
first return matrices'}
\item \st{page 108, line before (5.25): you could simply write q, r in +, - (later as well)}
\item page 111: maybe Corollary D.3 from Appendix D could be moved to the main text
\begin{quote}
    It could be but I dont think it needs to be moved. Thoughts?
\end{quote}
\item \st{page 151: 'and the last inequality is the Law of total probability' to 'and the last
equality is the Law of total probability'}
\item \st{page 154: 'where the swap of the intergals' to 'where the swap of the integrals'}
\item page 155: 'The probability of 1 is' - this looks a bit strange, maybe use A and B to
denote those events
\begin{quote}
    Check once compiled.
\end{quote}
\item \st{page 161: the bottom formula can fit in a single line}
\item page 162, about pointwise convergence: weak convergence is fine, that should cover
the most relevant performance measures. That said, the subsequent remarks and
Chapter 6.3 are welcome.
\begin{quote}
    Thanks?
\end{quote}
\item page 170, Fig. 1: good choice of plot markers, as the non-linear plot would be
invisible otherwise
\begin{quote}
    Thanks?
\end{quote}
\item \st{page 171: 'as more mass at the left' to 'as more mass is to the left'}
\item \st{page 175, Fig. 7.7: for the L1 error, QBDRAP seems to have a faster convergence
rate than DG. It would be nice to see the plot for higher dimension too. Either
way, I would be careful with conclusions about which method performs better in
this case.}
\begin{quote}
    Reworded and clarified that the QBD-RAP performs worst for these dimension but may be better for higher dimensions (speculative). See comment on higher dims.
\end{quote}
\item page 182, Fig. 7.14: once again, it would be nice to go higher with the dimension
\begin{quote}
    See previous comment on higher dims.
\end{quote}
\item page 184, Fig. 7.15 explanation: 'Interestingly, the error curve for the DG scheme
is not monotonic.' - any ideas why?
\begin{quote}
    added ``Interestingly, the error curve for the DG scheme
    is not monotonic which is likely to be caused by the specific location of oscillations in the approximation for different dimension approximations.''
\end{quote}
\item page 187, Fig. 7.18: nice to see a plot with higher dimension.
\begin{quote}
    See previous comments on higher order.
\end{quote}
\item \st{page 265: Borthwick (n.d.) - why is there no date? I found a 2016 edition.}
\end{itemize}

\begin{thebibliography}{9}
    \bibitem{bn2010}
    Bean, N.~G.~and Nielsen, B.~F. (2010). {`Quasi-birth-and-death processes with rational arrival process components.}' \textit{Stochastic Models}, 26(3), 309-334.
\end{thebibliography}
\end{document}

### 2022-12-30 chapter1/chapter.tex
first iteration of changes for reviewer 2 and response to reviewer 2 comments (to do: reviewer 1 changes and responses)

@@
 for any \(t,t_0,...,t_n\in\mathcal T\) with \(t_0+t,...,t_n+t\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\).
-A random variable, \(\tau\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\tau\in\sigma(X(s), s\leq t),\) where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s), s\leq t\}\).
+A random variable, \(\tau\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\tau\in\sigma(X(s), s\leq t),\) for all \(t\in T\), where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s)\}_{s\leq t}\).
 
@@
 
+\section{Laplace Transforms}
+The Laplace transform is an important tool in many areas of mathematics, and is particularly useful in the analysis of fluid queues and fluid-fluid queues.
+Furthermore, in Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with {Laplace transforms} to prove convergence of the QBD-RAP scheme.
+
+For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transform of \(\mu\) to be
+\[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
+where the \emph{region of convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
+\[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
+When \(\mu\) is the probability measure associated with a random variable, \(Z\), say, then we may write 
+\[\widehat \mu(\lambda)=\mathbb E[e^{-\lambda Z}],\]
+and the region of convergence is at least \([0,\infty)\). 
+Further, letting \(E^\lambda\) be an exponentially distributed random variable with rate \(\lambda\) and noting that \(\mathbb P(E^\lambda >t ) = e^{-\lambda t}\), then 
+\[\widehat \mu(\lambda)=\mathbb P(Z<E^\lambda),\]
+which gives a probabilistic interpretation of the Laplace transform.
+That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(Z\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
+
+A convenient property of the Laplace transform which we utilise is the Convolution Theorem. 
+\begin{thm}[Convolution Theorem]
+	Let \(f,g: [0,\infty) \to \mathbb R\) be integrable functions, then 
+	\[\mathcal L\left(\int_{u=0}^tf(u)g(t-u)\wrt u\right)(\lambda) = \mathcal L\left(f\right)\cdot \mathcal L\left(g\right).\]
+\end{thm}
+The Convolution Theorem states that the Laplace transform of the convolution, given by \(\displaystyle \int_{u=0}^tf(u)g(t-u)\wrt u\), is equal to the product of the Laplace transform of \(f\) and \(g\).
+
+The Laplace transform is unique in the sense that, if \(\mu\) and \(\nu\) are two measures on the Borel sets of \([0,\infty)\) and 
+\[\widehat \mu(\lambda) = \widehat \nu(\lambda)\] 
+for all \(\lambda > a\) with \(a<\infty\), then \(\mu\) and \(\nu\) are the same. In terms of functions, \(f,g: [0,\infty) \to \mathbb R\), if 
+\[\mathcal L(f)(\lambda) = \mathcal L(g)(\lambda),\]
+for all \(\lambda > a\) with \(a<\infty\), and \(f\) and \(g\) are continuous, then \(f(t)=g(t)\) for all \(t\geq 0\). Without knowing \(f\) and \(g\) are continuous, then we can only claim that 
+\(f(t)=g(t)\) for all \(t\geq 0, t\notin \mathcal N,\) where \(\mathcal N\) is a \emph{null set} with respect to Lebesgue measure.
+
 \section{Fluid queues}\label{sec: fqs}
@@
 \begin{align*}
-	\bs C &= \left[\begin{array}{ccc} \bs C_+ && \\ &\bs C_-& \\ && \bs 0\end{array}\right], && \bs C_+ = diag(c_i,i\in\calS_+), && \bs C_- = diag(|c_i|,i\in\calS_-),
+	\bs C &= \left[\begin{array}{ccc} \bs C_+ && \\ &\bs C_-& \\ && \bs 0\end{array}\right], && \bs C_+ = \mathrm{diag}(c_i,i\in\calS_+), && \bs C_- = \mathrm{diag}(|c_i|,i\in\calS_-),
 \end{align*}
@@
 \(
-	\widehat{\bs C} = diag(c_i,i\in\calS),
+	\widehat{\bs C} = \mathrm{diag}(c_i,i\in\calS),
 \)
-where \(diag(a_i,i\in\mathcal I)\) denotes a diagonal matrix with entries \(a_i\) down the diagonal.
+where \(\mathrm{diag}(a_i,i\in\mathcal I)\) denotes a diagonal matrix with entries \(a_i\) down the diagonal.
 
@@
 Here, we consider a mixture of \emph{regulated} and \emph{reflecting} boundary conditions.
-Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(sign(c_i)=sign(c_j)\) or \(sign(c_j)=0\) then the process is absorbed in the boundary, otherwise it is reflected.
+Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(\mathrm{sign}(c_i)=\mathrm{sign}(c_j)\) or \(\mathrm{sign}(c_j)=0\) then the process is absorbed in the boundary, otherwise it immediately leaves the boundary.
 At a lower boundary, if \(j\in\calS_0\cup\calS_-\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\), and the phase process continues to evolve according to the sub-generator
@@
 Consider an unbounded fluid queue \(\{(\ddot X(t), \varphi(t))\}\).
-Let \(\zeta_X(E)\) be the random variable which is the first hitting time of \(\{\ddot X(t)\}\) on the set \(E\) and define the matrix \(\bs \Psi_X(s)\) with elements \(\left[\bs \Psi_X(s)\right]_{ij}\), \(i\in\calS_+\), \(j\in\calS_-\), given by
+Let \(\zeta_X(E)\) be the random variable which is the first hitting time of \(\{\ddot X(t)\}\) on the set \(E\) and define the matrix \(\bs \Psi_X(\lambda)\) with elements \(\left[\bs \Psi_X(\lambda)\right]_{ij}\), \(i\in\calS_+\), \(j\in\calS_-\), given by
 \begin{align}
@@
 \begin{align*} 
-	\ddot{Y}(t) := \ddot Y(0)) + \int_0^t r_{\varphi(s)}({X}(s)) \wrt s.
+	\ddot{Y}(t) := \ddot Y(0) + \int_0^t r_{\varphi(s)}({X}(s)) \wrt s.
 \end{align*} 
@@
 	% 
-for $i \in \mathcal{S} = \{1,...,N\}$. Let $\bs R(x) := \diag(r_i(x))_{i \in \mathcal{S}}$ be the diagonal fluid-rate matrix of functions for $\{\dot Y(t)\}$.
+for $i \in \mathcal{S} = \{1,...,N\}$. Let $\bs R(x) := (r_i(x))_{i \in \mathcal{S}}$ be the diagonal fluid-rate matrix of functions for $\{\dot Y(t)\}$.
 
@@
 
-The problem we have is to solve the operator-Riccati equation -- something which is only possibly in the very simplest of cases.
+The problem we have is to solve the operator-Riccati equation -- something which is only possible in the very simplest of cases.
 This is where the methods in this thesis come in.
@@
 Indeed, it is not always the case that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) (trivially, consider \(\dot Y(t)=y_0\) for all \(t\)).
-Conditions on fluid-fluid queues to ensure that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) are not known, and here we just assume that it is.
+Conditions on fluid-fluid queues to ensure that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) are not known, in here we just assume that it is.
 
@@
 The measures \(\overline{\mu_i}(\cdot)\) represent an initial distribution, \(\overline{\mu_i}(\cdot) = \mathbb P(X(0)\in\cdot, \varphi(0) = i)\).
-Let \(\{\overline{\mathbb V}(t)\}_{t\geq 0},\, \overline{\mathbb V}(t):\mathcal{M}(\mathcal{S} \times [0,b]\mapsto \mathcal{M}(\mathcal{S} \times [0,b]\) be the semigroup describing the evolution of probability for \(\{(X(t),\varphi(t))\}_{t\geq 0}\) structured as a matrix of operators, \(\left[ \overline{\mathbb V}(t)\right]_{ij}= \overline{\mathbb V}_{ij}(t)\) where,
+Let \(\{\overline{\mathbb V}(t)\}_{t\geq 0},\, \overline{\mathbb V}(t):\mathcal{M}(\mathcal{S} \times [0,b])\mapsto \mathcal{M}(\mathcal{S} \times [0,b])\) be the semigroup describing the evolution of probability for \(\{(X(t),\varphi(t))\}_{t\geq 0}\) structured as a matrix of operators, \(\left[ \overline{\mathbb V}(t)\right]_{ij}= \overline{\mathbb V}_{ij}(t)\) where,
 \[\overline\mu_i \overline{ \mathbb V}_{ij}(t)(\mathcal{E}) = \int_{x\in[0,b]} \wrt \overline\mu_i(x)\mathbb P (X(t) \in\mathcal{E},\varphi(t) = j \mid X(0) = x, \varphi(0) = i).\]
@@
 Specifically, the domain of \(\overline{\mathbb B}\) is the set of measures, \(\overline{\bs \mu} = (\overline\mu_i)_{i\in\calS}\), for which each \(\overline\mu_i\) admits an absolutely continuous density on \((0,b)\), and can have a point mass at \(0\) if \(i\in\mathcal S_{-1}\) or a point mass at \(b\) if \(i\in\mathcal S_{K+1}\); call this set of measures \(\mathcal M_{0,b}\).
-The measures \(\overline\mu_i\) cannot have a point masses at 0 if \(i\notin \mathcal S_{-1}\), nor can they have a point masses at \(b\) if \(i\notin\calS_{{K+1}}\), nor can they have point masses in \((0,b)\) for any \(i\in\calS\).
+The measures \(\overline\mu_i\) cannot have a point mass at 0 if \(i\notin \mathcal S_{-1}\), nor can they have a point mass at \(b\) if \(i\notin\calS_{{K+1}}\), nor can they have point masses in \((0,b)\) for any \(i\in\calS\).
 In the sequel we write \(\overline v_i(x)\) as the density of \(\overline\mu_i\), and \(q_{{-1},i}\) and \(q_{{K+1},i}\) as the point masses of \(\overline\mu_i\) at \(0\) and \(b\), respectively (if such point masses exist).
@@
 	% 
-	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} := diag(\mathbb{R}_i^{k},\,i \in \mathcal{S})$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
+	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} := \mathrm{diag}(\mathbb{R}_i^{k},\,i \in \mathcal{S})$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
 	\begin{align*} 
@@
 
-Now, consider multiplying (\ref{eqn: cons law}) by the row vector \(\bs \phi_k(x)=()\phi_k^1(x),...,\phi_k^{p_k}(x))\) and integrating by parts over \(\calD_k\),
+Now, consider multiplying (\ref{eqn: cons law}) by the row vector \(\bs \phi_k(x)=(\phi_k^1(x),...,\phi_k^{p_k}(x))\) and integrating by parts over \(\calD_k\),
 \begin{align}
@@
 \end{align*}
-where \(s=\cfrac{1}{3}\sum\limits_{i=1}^{3}\mbox{sign}(a_i)\).
+where \(s=\cfrac{1}{3}\sum\limits_{i=1}^{3}\mathrm{sign}(a_i)\).
 When all three arguments, \(a_1,\,a_2\) and \(a_3\) have the same sign, the minmod function returns the smallest of the three arguments with the correct sign, otherwise the signs of the three arguments differ and the minmod function returns zero.
@@
 The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) In general, in \cite{MEinAP}, there is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility.
+
+The class of matrix exponential distributions is characterised as the class of probability distributions which have a rational Laplace transform.
+That is, \(\displaystyle \int_{x=0}^\infty e^{-\lambda x}\bs \alpha e^{\bs S x}\bs s \wrt x\) is a ratio of two polynomial functions in \(\lambda\).
+Matrix exponential distributions are an extension of Phase-type distributions, where for the latter, \(\bs S\) must be a sub-generator matrix of a CTMC, \(\bs s = -\bs S\bs e\) where \(\bs e\) is a \(1\times p\) vector of ones, and \(\bs \alpha\) is a discrete probability distribution.
+
+A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix exponential distribution with this representation.
+The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\).
+Representations of matrix-exponential distributions are not unique \citep{MEinAP}.
+A representation is called \emph{minimal} when \(\bs S\) has the smallest possible dimension.
+Throughout this work, we assume that the representation of any matrix exponential distribution is minimal.
+Let \(\bs e_i\) be a vector with a 1 in the \(i\)th position and zeros elsewhere.
+We assume that \(\bs s = -\bs S\bs e\), and that \((\bs e_i,\bs S,\bs s)\) for \(i=1,\dots,p\) are representations of matrix exponential distributions.
+It is always possible to find such a representation \cite[Theorem 4.5.17, Corollary 4.5.18]{MEinAP}.
+As such, we abbreviate our notation \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to \(Z\sim ME(\bs \alpha, \bs S)\).
+Further, given \(\bs s=-\bs S\bs e\) then observe that  \(\displaystyle \int_{x=0}^\infty e^{\bs S x} \bs s\wrt x = (-\bs S)^{-1}\bs s= \bs e\).
+
+For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal{A}\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a (possibly defective) matrix exponential distribution.
+
 The condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\).
@@
 	\begin{align}
-		\psi(x)&=\sum_{j=1}^{m_1} \sum_{k=1}^{p_j} c_{jk} \cfrac{x^{k-1}}{(k-1)!} e^{\mu_jx} + \sum_{j=1}^{m+2}\sum_{k=1}^{q_j} d_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\cos(\sigma_jx) \nonumber 
+		\psi(x)&=\sum_{j=1}^{m_1} \sum_{k=1}^{p_j} c_{jk} \cfrac{x^{k-1}}{(k-1)!} e^{\mu_jx} + \sum_{j=1}^{m_2}\sum_{k=1}^{q_j} d_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\cos(\sigma_jx) \nonumber 
 		\\&\quad{}+ \sum_{j=1}^{m_2}\sum_{k=1}^{q_j}e_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\sin(\sigma_jx), \label{eqn: spec}
@@
 Here \(\mu_j,\, j=1,\dots,m_1\) are the real eigenvalues of \(\bs S\), while \(\eta_j+i\sigma_j, \, \eta_j-i\sigma_j,\, j=1,\dots,m_2\) denote its complex eigenvalues, which come in conjugate pairs.
-Thus, \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m+1}p_j + 2\sum_{j=1}^{m_2}q_j\).
+Thus, \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m_1}p_j + 2\sum_{j=1}^{m_2}q_j\).
 \end{thm}
@@
 We define \(dev(\bs S)\) to be the real dominating eigenvalue of \(\bs S\), that is \(dev(\bs S)=\kappa\) in Theorem~\ref{thm: 4.1.4}.
-
-The class of matrix exponential distributions is characterised as the class of probability distributions which have a rational Laplace transform.
-That is, \(\displaystyle \int_{x=0}^\infty e^{-\lambda x}\bs \alpha e^{\bs S x}\bs s \wrt x\) is a ratio of two polynomial functions in \(\lambda\).
-Matrix exponential distributions are an extension of Phase-type distributions, where for the latter, \(\bs S\) must be a sub-generator matrix of a CTMC, \(\bs s = -\bs S\bs e\) where \(\bs e\) is a \(1\times p\) vector of ones, and \(\bs \alpha\) is a discrete probability distribution.
-
-A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix exponential distribution with this representation.
-The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\).
-Representations of matrix-exponential distributions are not unique \citep{MEinAP}.
-A representation is called \emph{minimal} when \(\bs S\) has the smallest possible dimension.
-Throughout this work, we assume that the representation of any matrix exponential distribution is minimal.
-Let \(\bs e_i\) be a vector with a 1 in the \(i\)th position and zeros elsewhere.
-We assume that \(\bs s = -\bs S\bs e\), and that \((\bs e_i,\bs S,\bs s)\) for \(i=1,\dots,p\) are representations of matrix exponential distributions.
-It is always possible to find such a representation \cite[Theorem 4.5.17, Corollary 4.5.18]{MEinAP}.
-As such, we abbreviate our notation \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to \(Z\sim ME(\bs \alpha, \bs S)\).
-Further, given \(\bs s=-\bs S\bs e\) then observe that  \(\displaystyle \int_{x=0}^\infty e^{\bs S x} \bs s\wrt x = (-\bs S)^{-1}\bs s= \bs e\).
-
-For a given \(p\times p\) matrix \(\bs S\), denote by \(\mathcal{A}\subset \mathbb R^p\) the space of all possible vectors \(\bs a\) such that \((\bs a, \bs S)\) is a valid representation of a (possibility defective) matrix exponential distribution.
 %By CITE it is known that \(\mathcal{E}\) is a closed, bounded, convex, affine subspace of \(\mathbb R^p\).
@@
 
-The class of concentrated matrix exponential distributions (CMEs) is found numerically in \citep{hht2020} and computationally efficient expressions for the density and moments of CMEs are provided in \citep{hht2020}.
+The class of concentrated matrix exponential distributions (CMEs) is found numerically in \citep{hht2020} and computationally efficient expressions for the density and moments of CMEs are provided therein.
 
@@
 
-From the definition of the class of CME distributions, for a given CME with odd order, \(p\), and representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part \cite{hht2020}.
+From the definition of the class of CME distributions, for a given CME with odd order, \(p\), and representation \((\bs\alpha,\bs S)\), the matrix \(\bs S\) has one real eigenvalue, and \(p-1\) complex eigenvalues and all eigenvalues have the same real part \citep{hht2020}.
 
@@
 where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
-The process \(\{\bs A(t)\}\) can jump at event times of \(N\) (the process may not actually jump at these times, but we may still refer to it a `jump' and typically the dynamics change at this point).
+The process \(\{\bs A(t)\}\) can jump at event times of \(N\) (the process does not necessarily jump at these times, but we may still refer to it a `jump' and typically the dynamics change at this point).
 At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\), i.e.~\(\mathbb P(N(t)=n,N(t+dt)=n+1)=\bs A(t) \bs D\bs e\wrt t\).
@@
 % me basis? semigroup? generator?
-\section{Laplace Transforms}
-We have already mentioned \emph{Laplace transforms}, but we have not-yet defined them.
-We define them now.
-In Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with {Laplace transforms} to prove convergence of the QBD-RAP scheme.
-For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transform of \(\mu\) to be
-\[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
-where the \emph{region of convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
-\[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
-When \(\mu\) is the probability measure associated with a random variable, \(Z\), say, then we may write 
-\[\widehat \mu(\lambda)=\mathbb E[e^{-\lambda Z}],\]
-and the region of convergence is at least \([0,\infty)\). 
-Further, letting \(E^\lambda\) be an exponentially distributed random variable with rate \(\lambda\) and noting that \(\mathbb P(E^\lambda >t ) = e^{-\lambda t}\), then 
-\[\widehat \mu(\lambda)=\mathbb P(Z<E^\lambda),\]
-which gives a probabilistic interpretation of the Laplace transform.
-That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(Z\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
-
-A convenient property of the Laplace transform which we utilise is the Convolution Theorem. 
-\begin{thm}[Convolution Theorem]
-	Let \(f,g: [0,\infty) \to \mathbb R\) be integrable functions, then 
-	\[\mathcal L\left(\int_{u=0}^tf(u)g(t-u)\wrt u\right)(\lambda) = \mathcal L\left(f\right)\cdot \mathcal L\left(g\right).\]
-\end{thm}
-The Convolution Theorem states that the Laplace transform of the convolution, given by \(\displaystyle \int_{u=0}^tf(u)g(t-u)\wrt u\), is equal to the product of the Laplace transform of \(f\) and \(g\).
-
-The Laplace transform is unique in the sense that, if \(\mu\) and \(\nu\) are two measures on the Borel sets of \([0,\infty)\) and 
-\[\widehat \mu(\lambda) = \widehat \nu(\lambda)\] 
-for all \(\lambda > a\) with \(a<\infty\), then \(\mu\) and \(\nu\) are the same. In terms of functions, \(f,g: [0,\infty) \to \mathbb R\), if 
-\[\mathcal L(f)(\lambda) = \mathcal L(g)(\lambda),\]
-for all \(\lambda > a\) with \(a<\infty\), and \(f\) and \(g\) are continuous, then \(f(t)=g(t)\) for all \(t\geq 0\). Without knowing \(f\) and \(g\) are continuous, then we can only claim that 
-\(f(t)=g(t)\) for all \(t\geq 0, t\notin \mathcal N,\) where \(\mathcal N\) is a \emph{null set} with respect to Lebesgue measure.
 
@@
 Tonelli's Theorem allows us to exchange the order of integration for any non-negative extended real-valued function, and includes the case where the value of the integral is infinite.
-In contrast, Fubini's Theorem us to exchange the order of integration for any real-valued function, provided that the integral is finite.
+In contrast, Fubini's Theorem allows us to exchange the order of integration for any real-valued function, provided that the integral is finite.
 

### 2022-12-30 chapter2/chapter.tex
first iteration of changes for reviewer 2 and response to reviewer 2 comments (to do: reviewer 1 changes and responses)

@@
 We can solve the semi-discrete system to approximate transient distributions of the fluid queue or, as we do in Section~\ref{sec:DGSFFM}, we can extract and manipulate an approximation to the generator of the fluid queue to approximate performance measures of fluid-fluid queues.
-The DG method conserves probability.
 Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the limiting distribution of a stochastic fluid-fluid queue.
+The DG method is convenient for approximating stochastic processes as it conserves the total probability of the system (as shown in Corollary~\ref{cor: googgy}).
 
@@
 
-For the problem at hand, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\) (the direct sum of \(W^k\)), the same as the set of basis functions of our solution space.
+For the problem at hand, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of \(W_k\), where \(W_k\) is the domain of the approximation on the interval \(\calD_k\).
+Further, we use the same basis functions for the test-function space and the approximation space.
 Proceeding as described above, for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z,\) the weak formulation is
@@
 The flux is the instantaneous rate (with respect to time) at which density moves across the boundaries \(y_k,\,k\in\mathbb Z\).
-There are different choices for the flux, and we refer the reader to \citep{c99,nodalDGBook}, and references therein, for some discussion of the topic.
+There are different choices for the numerical flux, and we refer the reader to \citep{c99,nodalDGBook}, and references therein, for some discussion of the topic.
 Here, we choose the \textit{upwind} scheme, which, as we shall see, closely resembles the flux terms from the generator \(\mathbb B\).
 The approximate flux, also known as the \textit{numerical flux}, is given by
-\[f^*_j(x,t) = sign(c_j)\lim_{\varepsilon\to0^+}u_j(x-\varepsilon c_j,t),\]
+\[f^*_j(x,t) = \mathrm{sign}(c_j)\lim_{\varepsilon\to0^+}u_j(x-\varepsilon c_j,t),\]
 at each \(x=y_k,\,k\in\mathbb Z\). 
@@
 \[\sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(0)\tr{} p_{ij}^{K+1} c_i,\]
-\(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{{K+1}}(t)\) is \(\bs T_{{K+1},{K+1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b)\), \(j\in\calS_-\) and the DG approximation to reflection of density at the upper boundary is \(- \sum_{i\in\calS_+}c_ip_{ij}^{K+1}\bs a_{0,i}(t)\bs F_j^{0,0}\).
+\(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{{K+1}}(t)\) is \(\bs T_{{K+1},{K+1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b)\), \(j\in\calS_-\) and the approximation to reflection of density at the upper boundary is \(- \sum_{i\in\calS_+}c_ip_{ij}^{K+1}\bs a_{0,i}(t)\bs F_j^{0,0}\).
 
@@
 \begin{align*}
-	{{\bs B}}^{k\ell} &= \ddot{{\bs B}}^{k\ell}, \mbox{ for \(k\in\mathcal K^\circ\), \(\ell\in\{k-1,k,k+1\}\), \(k=\ell\neq 0\) or \(k=\ell\neq K\),}
+	{{\bs B}}^{k\ell} &= \ddot{{\bs B}}^{k\ell}, \mbox{ for \(k\in\mathcal K^\circ\), \(\ell\in\{k-1,k,k+1\}\), \(k=\ell,\, k,\ell\notin \{0,K\}\),}
 	\\ \bs B^{00} &= \ddot{\bs B}^{00} - \left[c_ip_{ij}^{-1} 1(c_i<0, c_j>0)\right]_{i\in\calS,j\in\calS}\otimes \bs F_-^{00},
@@
 However, it is convenient to prove properties of the generator.
-\begin{cor}
+\begin{cor} \label{cor: googgy}
 	The approximate generator \( \bs B\) conserves probability. That is, for all \(t\geq 0\),
@@
 Let \(\boldsymbol a_{-1}(t):=\boldsymbol q_{-1}(t)\) and \(\boldsymbol a_{K+1}(t):=\boldsymbol q_{K+1}(t)\), and define basis functions \(\bs\phi_{-1}(x) = \phi_{-1}^1(x) = \delta(x)\) and \(\bs\phi_{K+1}(x) = \phi_{K+1}^1(x) = \delta(x-b)\), where \(\delta\) is the Dirac delta, \(p_{-1} = p_{K+1} = 1\), and \(\mathcal N_{-1} = \mathcal N_{K+1} = \{1\}\).
-Also define \({\bs M}_{-1}=\bs I_{|\calS_{-1}|}\), \({\bs M}_{K+1}=\bs I_{|\calS_{K+1}|}\), the block-diagonal matrix \(\bs M=diag(\bs M_k,k\in\mathcal K)\), and row-vectors
+Also define \({\bs M}_{-1}=\bs I_{|\calS_{-1}|}\), \({\bs M}_{K+1}=\bs I_{|\calS_{K+1}|}\), the block-diagonal matrix \(\bs M=\mathrm{diag}(\bs M_k,k\in\mathcal K)\), and row-vectors
 \[\boldsymbol \phi(x) = (\bs\phi_k(x))_{k\in\mathcal K}, \quad \boldsymbol a_i(t) = (\bs a_{k,i}(t))_{k\in\mathcal K},\,i\in\calS.\]
@@
 Since \(\boldsymbol a_{k,i}\) is arbitrary, we see that we approximate \(\mathbb R_{k,i}\) by \(  \bs R_{k,i} = \bs M_k^r\bs M_k^{-1},\)
-and \(\mathbb R^k\) by \(  \bs R^k = diag(  \bs R_{k,i},{i\in\calS^\bullet_k})\).
+and \(\mathbb R^k\) by \(  \bs R^k = \mathrm{diag}(  \bs R_{k,i},{i\in\calS^\bullet_k})\).
 
@@
 \[\bs\phi^+(x){\bs M}_+^{-1}\bs \Psi(s)\bs\phi^-(y)\tr{}\wrt y = \bs \phi^+(x)\bs M_+ \left[\left[\bs \Psi_{ij}^{k\ell}\right]_{i\in\calS_k^+,j\in\calS_\ell^-}\right]_{k\in\mathcal K^+,\ell\in\mathcal K^-}\bs \phi^-(y)'\wrt y\]%\left[\left[\bs\phi^k(x)\bs {M}_k^{-1}\bs \Psi_{ij}^{k\ell}(s)\bs\phi^\ell(y)\tr{}\wrt y\right]_{i\in\calS_k^+,j\in\calS_\ell^-}\right]_{ k\in\mathcal K^+, \ell\in\mathcal K^-},\]
-where \(\bs\phi^+(x) = (\bs\phi_k(x))_{i\in\calS_k^+,k\in\mathcal K^+}\) and \(\bs\phi^-(y) = (\bs\phi_k(y))_{i\in\calS_k^-,k\in\mathcal K^-}\) are row-vectors, \(\bs \Psi(s)\) is a matrix of constants for a given \(s\) with the same size as \(\bs D^{+-}\), and \({\bs M}_m,\) \(m\in\{+,-,0\}\) is a block diagonal matrix \({\bs M}_m = diag\left(  \bs M_k, {i\in\calS_k^m}, {k\in\mathcal K^m}\right)\), \(m\in\{+,-,0\}\).
+where \(\bs\phi^+(x) = (\bs\phi_k(x))_{i\in\calS_k^+,k\in\mathcal K^+}\) and \(\bs\phi^-(y) = (\bs\phi_k(y))_{i\in\calS_k^-,k\in\mathcal K^-}\) are row-vectors, \(\bs \Psi(s)\) is a matrix of constants for a given \(s\) with the same size as \(\bs D^{+-}\), and \({\bs M}_m,\) \(m\in\{+,-,0\}\) is a block diagonal matrix \({\bs M}_m = \mathrm{diag}\left(  \bs M_k, {i\in\calS_k^m}, {k\in\mathcal K^m}\right)\), \(m\in\{+,-,0\}\).
 Now replace the theoretical kernels in Theorem~\ref{theo:Psi} by their DG approximations to get

### 2022-12-30 chapter5/chapter.tex
first iteration of changes for reviewer 2 and response to reviewer 2 comments (to do: reviewer 1 changes and responses)

@@
 \end{align}
-where the swapping of limits and sums in the first equality is justified as the sums are finite, the swapping limits and products in the first equality is justified as the product is finite and all limits exist, and the last inequality is the Law of total probability.
+where the swapping of limits and sums in the first equality is justified as the sums are finite, the swapping limits and products in the first equality is justified as the product is finite and all limits exist, and the last equality is the Law of total probability.
 
@@
 \end{align}
-where the swap of the intergals and sums is justified by the Fubini-Tonelli Theorem.
+where the swap of the integrals and sums is justified by the Fubini-Tonelli Theorem.
 Now, (\ref{eqn:GHJKLKJHGHJKL}) is an infinite sum of terms appearing in Lemma~\ref{lem: LAkAKFnvnb mav h}.
@@
 	Hence, for sample paths which contribute to the Laplace transform, one of two things must happen, either; 
-	\begin{enumerate}
+	\begin{enumerate}[a]
 		\item the phase remains \(i\) until there is an orbit restart epoch and \(E^\lambda\) does not occur before the orbit restart epoch, or, \label{item: 1}
@@
 		&\leq2\sum_{n=1}^\infty B^{n-1}+1
-		\\&=\cfrac{2}{1-B}+1
-		\\&<\infty,
+		=\cfrac{2}{1-B}+1
+		<\infty,
 	\end{align*}

### 2022-12-30 chapter6/chapter.tex
first iteration of changes for reviewer 2 and response to reviewer 2 comments (to do: reviewer 1 changes and responses)

@@
 In comparison, the reconstruction with the normalised closing operator is much better in this region due to the `folding' around \(\Delta\) in the construction of the closing operator.
-The `folding' in the normalised closing operator manifests itself as more mass at the left of these plots compared to the unnormalised reconstructions.
+The `folding' in the normalised closing operator manifests itself as more mass is to the left of these plots compared to the unnormalised reconstructions.
 % Recall that, as discussed in Section~\ref{}, for the reconstruction we must to allocate the initial condition to a phase with positive rate, or negative rate, so that we can use the appropriate approximation and reconstruction method as discussed in Section~\ref{sec: closing}.
@@
 Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the schemes appear to show the convergent behaviour we expect.
-Here, the uniformisation scheme appears to perform the best, while the QBD-RAP scheme performs the worst.
+Here, for dimensions up to 21, the uniformisation scheme appears to perform the best, while the QBD-RAP scheme performs the worst, however the QBD-RAP approximation appears to be converging at a faster rate and may outperform the DG approximation if a higher dimension is used.
 
@@
 Once again, the DG-lim scheme does not appear to converge.
-Interestingly, the error curve for the DG scheme is not monotonic.
+Interestingly, the error curve for the DG scheme is not monotonic which is likely to be caused by the specific location of oscillations in the approximation for different dimension approximations.
 Furthermore, the DG scheme can result in negative estimates of the cell probabilities in (\ref{eqn: cell errors})-(\ref{eqn: cell probs boundary}) (not shown).

### 2022-12-31 reviews/response.tex
beginning of comments to reviewer 1

@@
 
+\pagebreak
+\section{Reviewer 1}
+
+\textbf{OVERVIEW}
+
+This thesis considers approximations of fluid-fluid queues by approximating the dri-
+ving fluid process by either a QBD or RAP-QBD process. By discretizing the state-
+space of the driving fluid process, one does not obtain a QBD automatically, but by
+replacing the constant drifts with random ones, governed by birth- and death pro-
+cesses that mimic the drift, one can approximate the fluid flow process by a QBD.
+This is the content of Chapter 3, in which part the author is the co-author of a
+published paper.
+\begin{quote}
+    This is not the content of Chapter 3. 
+\end{quote}
+
+Since one is interested in approximating the deterministic hitting times caused by
+the constant drift, one might think that Erlangization is a good idea to improve on
+this. It turns out, however, that Erlangization is already a part of the QBD construc-
+tion, so nothing is gained in this direction.
+
+Then the author, in the second part of the thesis, considers an interesting and far
+from trivial approach by employing RAP-QBD's as replacements for the QBD appro-
+ximations for the sake of using concentrated matrix-exponential distribution as a
+more efficient approximation to the deterministic hitting times.
+
+The second chapter contains some background chosen by the author. It resumes ope-
+rator theory, weak convergence, Laplace transform and many other subjects that are
+readily available in the mathematical literature and might be considered as common
+background. When it comes to reviewing fluid and fluid-fluid models, then these are
+explained with the same level of detail. However, since this is much more advanced
+and more specialized, one might have hoped for a more rigorous introduction to both
+fluid queues and RAPs for the sake of readability. This, however, is not a compulsory
+request from my side.
+\begin{quote}
+    I'd be happy to add more to the background on fluid queus and RAP, but, personally, I don't think there is much more to say which is relevant to the thesis.
+This thesis does not have a great deal to do with the existing analyses of fluid queues.
+Thoughts?
+\end{quote}
+
+The thesis is large and contains a lot of material but lacks order. It is hard to find
+details on notation, for example, \(\phi\) and \(\varphi\). This makes the task of reading it much
+harder, in particular, if one does not read the whole thesis from scratch.
+\begin{quote}
+    Yes. Sorry about that. 
+\end{quote}
+
+I believe this is mainly due to the lack of structure. It would be useful to employ
+a more methodological approach using Definitions, Theorems, etc. and frequently
+pinpoint back to these results/definitions whenever used.
+\begin{quote}
+    IMO, I have used Theorem, Lemma, Corllary environments to help with this. Perhaps I could use Definition environments too, this might help. Thoughts? 
+\end{quote}
+
+Chapter 5 is long and technical, while Chapter 6 is short and more straightforward.
+Also, the ultimate aim of both chapters is the result of Chapter 6, so to me, it would
+make more sense to merge the two chapters.
+\begin{quote}
+    I'm happy either way.
+These two chapters were originally one long chapter but i decided to break them in two because they use distinct technical tools.
+Perhaps I should make them a single chapter with two main sections in the chapter.
+\end{quote}
+
+In conclusion, the thesis is long, highly technical, and somewhat unstructured, ma-
+king it non-accessible to a broader community.
+
+\textbf{ASSESSMENT}
+
+The author is a co-author of Chapter 3, the material of which is original and publis-
+hed. The material in Chapters 4,5 and 6 is original. The material of the thesis has a
+reasonable narrow scope. It considers an important approximation problem using a
+recent approach. The communication is very technical, and lacks some structure for
+a more smooth communication. The contribution is significant but could have been
+broader in its scope.
+\begin{quote}
+    Interesting that he didn't mention Chapter 7... perhaps he does not recognise numerical investigation in the same regard as the technical chapters, haha.
+
+    Yes, the scope of the thesis is narrow. It was one of my main worries about the work. I could have added more, but, as he already pointed out, the thesis is already long and hard to read.
+
+    Specific comments on structure to be addressed below. 
+\end{quote}
+
+\textbf{CHAPTER 1-3}
+
+\st{p. 8 l.end-6: restricted to t = 0.}
+
+p. 12 l.end-6: you have already used matices extensively.
+\begin{quote}
+    I have. I just wanted to clarify notation, and point out that matrices use square brackets. 
+\end{quote}
+
+p. 13 l.end-10: how can the underlying phase-process change state upon hitting a boun-
+dary? Then it should be level dependent?
+\begin{quote}
+    It is level-dependent.
+I have added some commentary: "Some boundary conditions may introduce a dependence of the phase process, \(\{\varphi(t)\}\), on the level process \(\{X(t)\}\).
+Specifically there may be a change of phase upon the event that the level process hits a boundary.
+When the phase process depends on the level process then the fluid queue is \emph{level-dependent}, otherwise it is \emph{level independent}.
+For a level-dependent fluid queue the phase process \(\{\varphi(t)\}\) is not the only random element of the fluid queue."
+\end{quote}
+
+\st{Formula (2.10): parenthesis.}
+
+\st{p.21: B(t), argument t missing at some places.}
+\begin{quote}
+    Actually, there shouldn't be a (t)
+\end{quote}
+
+\st{p.40: second last line, a D missing.}
+
+(2.42): one does announce a proof to a formula/equation. Why do you proved it all? Most other items around this section are left unproven.
+\begin{quote}
+    I don't think there is a need for the proof. Happy to omit it. Same for the MPR on the previous page?
+\end{quote}
+
+Section 2.4 reproduces some fluid-fluid theory. The section could have benefitted
+from including, at least some, proofs of the ascertions to gain a working knowled-
+ge.
+\begin{quote}
+    happy to add proofs, but they are relatively technical and not absolutely necessary for the thesis IMO. They could help the reader intuit the fluid queue and its operators. Thoughts?
+\end{quote}
+
+Chapter 3 is taken from a paper, reproduced as such and in a good shape.
+\begin{quote}
+    I disagree, but okay. IMO this is one of the hardest-to-read sections!
+\end{quote}
+
+\pagebreak
 \section{Reviewer 2}

### 2022-12-31 chapter1/chapter.tex
beginning of comments to reviewer 1

@@
 \(\mathbb P(X(t+s)=j \mid X(s)=i)\) does not depend on \(s\). For a time-homogeneous CTMC, the \emph{transition function} is the matrix function 
-\[\bs P(t) = [\mathbb P(X(t)=j \mid X(t)=i)]_{i,j\in\calS}.\]
+\[\bs P(t) = [\mathbb P(X(t)=j \mid X(0)=i)]_{i,j\in\calS}.\]
 The Chapman-Kolmogorov equations state that 
@@
 \begin{align} 
-	&\nonumber \bs T = [T_{ij}]_{i,j\in\calS} = \cfrac{\wrt}{\wrt t}P(t),
+	&\nonumber \bs T = [T_{ij}]_{i,j\in\calS} = \left. \cfrac{\wrt}{\wrt t}P(t)\right|_{t=0},
 \end{align}
@@
 Sample paths of \(\{\ddot X (t)\}\) are continuous and piecewise linear, with \(\cfrac{\wrt }{\wrt t} \ddot X(t) = c_\varphi(t)\), when \(\ddot X(t)\) is differentiable.
-Given sample paths of \(\{\varphi(t)\}\), then \(\{\ddot X(t)\}\) is deterministic, and in this sense, \(\{\varphi(t)\}\) is the only stochastic element of the fluid queue.
+Given sample paths of \(\{\varphi(t)\}\), then \(\{\ddot X(t)\}\) is deterministic, and in this sense, \(\{\varphi(t)\}\) is the only stochastic element of the unbounded fluid queue.
 
@@
 Here, we consider a mixture of \emph{regulated} and \emph{reflecting} boundary conditions.
+Some boundary conditions may introduce a dependence of the phase process, \(\{\varphi(t)\}\), on the level process \(\{X(t)\}\).
+Specifically there may be a change of phase upon the event that the level process hits a boundary.
+When the phase process depends on the level process then the fluid queue is \emph{level-dependent}, otherwise it is \emph{level independent}.
+For a level-dependent fluid queue the phase process \(\{\varphi(t)\}\) is not the only random element of the fluid queue.
+
 Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(\mathrm{sign}(c_i)=\mathrm{sign}(c_j)\) or \(\mathrm{sign}(c_j)=0\) then the process is absorbed in the boundary, otherwise it immediately leaves the boundary.
@@
 \begin{align}
-	\left[\bs \Psi_X(\lambda)\right]_{ij} = \mathbb E\left[ e^{-\zeta_X(z)\lambda} 1(\zeta_X(z)<\infty, \varphi(\zeta_X(z)=j)\mid \ddot X(0)=z, \varphi(0)=i\right].
+	\left[\bs \Psi_X(\lambda)\right]_{ij} = \mathbb E\left[ e^{-\zeta_X(z)\lambda} 1(\zeta_X(z)<\infty, \varphi(\zeta_X(z)=j))\mid \ddot X(0)=z, \varphi(0)=i\right].
 \end{align}
@@
 Now define \(\mathbb B=\left.\cfrac{\wrt}{\wrt t}\mathbb V(t)\right|_{t=0}\) as the infinitesimal generator of \(\{\mathbb V(t)\}\), resulting in the tridiagonal matrix of operators 
-\[\mathbb B(t) = \left[
+\[\mathbb B = \left[
 	\begin{array}{ccccc}
-		\mathbb B^{{-1},{-1}}(t)&\mathbb B^{{-1},0}(t)& & & \\
-		\mathbb B^{0,{-1}}(t)&\mathbb B^{0,0}(t)&\mathbb B^{0,1}(t)&& \\
-		&\mathbb B^{1,0}(t)&\mathbb B^{1,1}(t)&\ddots& \\
+		\mathbb B^{{-1},{-1}}&\mathbb B^{{-1},0}& & & \\
+		\mathbb B^{0,{-1}}&\mathbb B^{0,0}&\mathbb B^{0,1}&& \\
+		&\mathbb B^{1,0}&\mathbb B^{1,1}&\ddots& \\
 		& & \ddots & \ddots & \mathbb B^{K,{K+1}}\\
@@
 %\end{align*}
-where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\).\footnote{We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
+where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\).\footnote{We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
 The operators denoted with an overline play a minor role in the introductory sections of this thesis, but do not appear again.
@@
 \end{align}
-\begin{proof}
-	The proof follows from 
-	\begin{align*}
-		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j}\\\vdots\\c_{nj}\bs{D} \end{array}\right] 
-		%
-		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
-		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
-	\end{align*}
-\end{proof}
+
+% \begin{proof}
+% 	The proof follows from 
+% 	\begin{align*}
+% 		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j} \bs D \\\vdots\\c_{nj}\bs{D} \end{array}\right] 
+% 		%
+% 		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
+% 		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
+% 	\end{align*}
+% \end{proof}
 
@@
 \end{align}
-\begin{proof}
-	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
-	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
-	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\).
+% \begin{proof}
+% 	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
+% 	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
+% 	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\).
 The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\).
 The former follows from
-	\begin{align}
-	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
-	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
-	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
-	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
-	\end{align}
-
-	Therefore 
-	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
-	and the result follows by the mixed product rule. 
-\end{proof}
+% 	\begin{align}
+% 	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
+% 	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
+% 	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
+% 	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
+% 	\end{align}
+
+% 	Therefore 
+% 	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
+% 	and the result follows by the mixed product rule. 
+% \end{proof}
 

### 2023-01-05 reviews/response.tex
.

@@
 \documentclass[a4paper]{article}
-\usepackage{amsmath}
+\usepackage{amsmath,amssymb}
 \usepackage{soul}
@@
 \begin{quote}
-    My comments follow are are in indented and are in italics. 
+    My comments follow are in indented and in \emph{italics}. 
 \end{quote}
@@
 
-\textbf{OVERVIEW}
+\noindent\textbf{OVERVIEW}
 
-This thesis considers approximations of fluid-fluid queues by approximating the dri-
-ving fluid process by either a QBD or RAP-QBD process. By discretizing the state-
-space of the driving fluid process, one does not obtain a QBD automatically, but by
-replacing the constant drifts with random ones, governed by birth- and death pro-
-cesses that mimic the drift, one can approximate the fluid flow process by a QBD.
+This thesis considers approximations of fluid-fluid queues by approximating the driving fluid process by either a QBD or RAP-QBD process.
+By discretizing the state space of the driving fluid process, one does not obtain a QBD automatically, but by
+replacing the constant drifts with random ones, governed by birth and death processes that mimic the drift, one can approximate the fluid flow process by a QBD.
 This is the content of Chapter 3, in which part the author is the co-author of a
@@
 the constant drift, one might think that Erlangization is a good idea to improve on
-this. It turns out, however, that Erlangization is already a part of the QBD construc-
-tion, so nothing is gained in this direction.
+this. It turns out, however, that Erlangization is already a part of the QBD construction, so nothing is gained in this direction.
 
 Then the author, in the second part of the thesis, considers an interesting and far
-from trivial approach by employing RAP-QBD's as replacements for the QBD appro-
-ximations for the sake of using concentrated matrix-exponential distribution as a
+from trivial approach by employing RAP-QBD's as replacements for the QBD approximations for the sake of using concentrated matrix-exponential distribution as a
 more efficient approximation to the deterministic hitting times.
 
-The second chapter contains some background chosen by the author. It resumes ope-
-rator theory, weak convergence, Laplace transform and many other subjects that are
+The second chapter contains some background chosen by the author. It resumes operator theory, weak convergence, Laplace transform and many other subjects that are
 readily available in the mathematical literature and might be considered as common
@@
 
-In conclusion, the thesis is long, highly technical, and somewhat unstructured, ma-
-king it non-accessible to a broader community.
+In conclusion, the thesis is long, highly technical, and somewhat unstructured, making it non-accessible to a broader community.
 
-\textbf{ASSESSMENT}
+\noindent\textbf{ASSESSMENT}
 
-The author is a co-author of Chapter 3, the material of which is original and publis-
-hed. The material in Chapters 4,5 and 6 is original. The material of the thesis has a
+The author is a co-author of Chapter 3, the material of which is original and published. The material in Chapters 4,5 and 6 is original. The material of the thesis has a
 reasonable narrow scope. It considers an important approximation problem using a
@@
 
-\textbf{CHAPTER 1-3}
+\noindent\textbf{CHAPTER 1-3}
 
@@
 
-p. 12 l.end-6: you have already used matices extensively.
+p. 12 l.end-6: you have already used matrices extensively.
 \begin{quote}
@@
 
-p. 13 l.end-10: how can the underlying phase-process change state upon hitting a boun-
-dary? Then it should be level dependent?
+p. 13 l.end-10: how can the underlying phase-process change state upon hitting a boundary? Then it should be level dependent?
 \begin{quote}
@@
 Section 2.4 reproduces some fluid-fluid theory. The section could have benefitted
-from including, at least some, proofs of the ascertions to gain a working knowled-
-ge.
+from including, at least some, proofs of the assertions to gain a working knowledge.
 \begin{quote}
@@
 
+%%%
+\noindent\textbf{CHAPTER 4}
+
+p.
+68 l.
+end-11: Highly confusing use of \(\varphi\) and \(\phi\).
+It is probably misprinted here, but also later, it is never clear which one is which.
+Here definitions to point back to would be useful (or at least a page number).
+\begin{quote}
+    Fixed typo on p. 68. However, I still need to clarify this notation throughout. 
+\end{quote}
+
+\st{p. 69: You are in a new chapter, and suddenly T appears again (back from Sec. 2.3.). Refer back or set up the basic notation again.}
+\begin{quote}
+    Defined \(T\) and \(c_i\) again. 
+\end{quote}
+
+p.
+69: formula (4.2), why is the lower Erlang structure transposed as compared to the upper?
+Is this necessary?
+I suppose it is due to running the phases in reverse order.
+Perhaps commenting on this might be a good idea.
+Also, why are the rates \(\Delta/2\), implying 4 times longer holding time?
+Should they not be \(2\Delta\) for a \(\Delta/2\) grid?
+\begin{quote}
+    -Added ``Intuitively, the transposition of the Erlang sub-blocks, \(\left[\begin{array}{cc} -2\Delta & 2\Delta \\ 0 & -2\Delta \end{array}\right]\), in negative phases relative to positive phases is due to ``reversing'' the progression of the jump process associated with the Erlang variables in negative phases relative to positive phase.''
+
+    -Corrected rates \(\Delta/2\) to \(2\Delta\).  
+\end{quote}
+
+p. 70 l. 14: \(\phi\) again. What does “phase-process” mean here? 
+\begin{quote}
+    Clarified: ``In this chapter we choose to use an equivalent, but slightly different, representation of the QBD-RAP, we write \(\{(L(t),\boldsymbol A(t),\phi(t))\}_{t\geq 0}\), where we refer to \(\{\phi(t)\}\) as the \emph{phase process} of the QBD-RAP.
+Such a representation does not always exist for a QBD-RAP.
+However, for the specific QBD-RAP introduced in this thesis such a representation does exist, by construction.''
+\end{quote}
+
+\st{p. 72 l. 10: seems that the symbol Ri(u) is used before introduced just below. I believe
+it should say Zi instead and remove u from the interval.}
+\begin{quote}
+    Changed
+\end{quote}
+
+p. 73 l. end-9: It this the best trade-off between small \(\varepsilon\) and small variance (power)? Does
+the choice matter? Comment on the choice of \(\varepsilon\).
+\begin{quote}
+    I guess the choice is relatively arbitrary at this point. This was a choice used by the CME group. Given the coarseness of Chebyshev's inequality, I do not think the choice is important.
+\end{quote}
+
+p.
+75 l.
+5: representations for matrix-exponential distributions are both denoted by a triple and a pair (with or with exit-rate vector).
+Since you always choose a representation with s = -Se you might as well use pair unless otherwise needed.
+\begin{quote}
+    It is a pair? I assume this was misread by the reviewer? 
+\end{quote}
+
+\st{p. 84 l. end-1: something wrong in (4.18)? Perhaps wrongly placed closing parenthesis.}
+
+p.
+84 l.
+end-1: where does this result come from?
+Earlier, you introduced this type of result only for the phase-type case, but I do not see where it comes from.
+You must refer back or justify why it should hold.
+\begin{quote}
+    I'm not entirely sure what the reviewer means by this.
+I realised that the ODE description of the QBD-RAP evolution was missing so I added it to the preliminaries, and I added a reference for it too.
+Do they want me to clarify/derive the equation?
+\end{quote}
+
+p.
+85: Why is A(t)D e = 1.
+Since the orbits and phases are independent, it is rather trivial what you are calculating.
+In fact, this could apply to the whole section up to Th.
+4.1.
+Please check if this is not the case.
+\begin{quote}
+    \begin{itemize}
+        \item I am not calculating anything. I am describing and summarising the dynamics of the stochastic process which we just introduced.
+        \item Sure, it's trivial if you're the person who invented and wrote the book on RAPs! For mortals, it is less trivial. 
+    \end{itemize}
+\end{quote}
+
+p. 86: Theorem 4.1. A lot of confusion regarding what is what (\(\varphi, \phi\), etc.). The first formula in the proof, where does that come from? Refer back.
+\begin{quote}
+    Yep, I need to go through the whole thesis and clarify this...
+\end{quote}
+
+p. 88: Second half page, why these forms of the matrices? Refer back or deduce.
+\begin{quote}
+    These matrices can be deduced from the description of the process above (which is apparently ``trivial'')
+\end{quote}
+
+p. 89: There can't be a phase change exactly at the time upon hitting. Also, the notation \(p_{ij}^{-1}\) is terrible (or dangerous!). What does it mean? It should be explained. Again \(ij\)
+the probability seems to be independent of the rest (at best), and the calculations on page 90 seem, therefore, trivial. Do I miss something?
+\begin{quote}
+    \begin{itemize}
+        \item Yes, there can be a change upon hitting due to the boundary conditions we have specified. 
+        \item Okay, I will update notation from \(p_{ij}^{-1}\) to something else. 
+        \item You do not miss something. The calculations are rudimentary (not trivial, but why get rid of them? 
+    \end{itemize}
+\end{quote}
+
+p. 91: matrices are put up without explanation.
+\begin{quote}
+    How should I explain these?
+\end{quote}
+
+\st{p. 92: k(x) suddenly appears again. Write its formula or refer to (4.5) since it has not been used since.}
+
+p. 93: Again, justify the matrix forms. How do we know that we can actually find such an augmentation which is still a RAP? This would be “trivial” for the QBD case, but for RAP is more delicate.
+\begin{quote}
+    \begin{itemize}
+        \item TO DO: justify matrices. 
+        \item I think it is still relatively straight forward to ensure that it is a QBD-RAP. I could elaborate? 
+    \end{itemize}
+\end{quote}
+
+p. 94 l. end-10: \(t + du\) is not an interval.
+\begin{quote}
+    I believe I have referred to is as an infinitesimal interval. Is that clear enough?
+\end{quote}
+
+p. 93: Section 4.7 is quite messy, seems experimental, and lacks a conclusion or result. Again, introducing some definitions and theorems would help to understand what you would like to communicate.
+\begin{quote}
+    I believe I explain it.
+The purpose is to introduce the concept of closing operators so that we can get approximations of density functions within intervals.
+I accept that it is messy and experimental.
+I will add some definition environments, or something to help clear it up.
+\end{quote}
+
+p. 97: the whole point of all the previous work, as far as I understood, was 4.8. This section is rather inconclusive.
+\begin{quote}
+    This is a ``how to'' approximate operators of fluid-fluid queues. Not sure what he wants me to add here? Perhaps the title is misleading?
+\end{quote}
+
+\st{p. 98: Section 4.9 the shortest section I have ever seen. May incorporate it into 4.8.}
+
+\noindent\textbf{CHAPTER 5}
+
+One may wonder whether it is really necessary to apply such a long and tedious proof of convergence.
+Essentially, since it is before \(\tau_1\), it is the approximation of the time to the change of level that is approximated.
+This is ME distributed, which is dense, so only using this fact should be enough.
+Maybe the proof could be carried out for a general case of dense approximants and thereby gain transparency.
+\begin{quote}
+    \begin{itemize}
+        \item Indeed, the proof is long and tedious, and there might be a shorter proof. I wondered this myself. To me, this is the most obvious proof. 
+        \item I'm not sure that the full extent of the result is appreciated in the comment above.
+The approximant is ME distributed, but the true solution is, in general, not ME distributed.
+It's not clear, to me, how denseness, in and of itself, would help the proof.
+To me, all denseness allows us to say is ``there exists a sequence of ME's which converge to the distribution of interest'' --it does not tell us what the MEs are.
+    \end{itemize}
+\end{quote}
+
+Much of the chapter continues the style by not structuring the exposition into notation, definition and theorems.
+For example in the announcement of Lemma 5.4, the \(R_{v,2}^{(p)}\), \(\varepsilon^{(p)}\) and \(G\) are not defined or referred back to.
+They do not appear in Section 5.4.
+above, and I could not find them browsing backwards either.
+\begin{quote}
+    They are defined in the preliminaries to the chapter. I have now recalled them in Lemma 5.4, for convenience. 
+\end{quote}
+
+It is very, very hard to read this section, and I shall not go through it in more detail until it has been restructured. Here I pinpoint some details.
+\begin{quote}
+    sorry about that 
+\end{quote}
+
+p. 104: (5.10), what is the meaning of this expression? The same applies to the subsequent formulas. The reader is left wondering why you introduce these formulas and what they mean.
+\begin{quote}
+    Fair enough.
+However, I do say in this section that we form a partition of \(f^{\ell_0,(p)}(t)(x,j;x_0,i)\), and the \(f_{m,r,s}^{\ell_0,(p)}\) the reviewer refers to is an element of this partition.
+\end{quote}
+
+p. 105: comment on how you obtain (5.11). Probabilistically it is clear if it would be a QBD, but less so if it is a RAP-QBD. So refer to the appropriate result you use. 
+\begin{quote}
+    I have now cited the characterisation of QBD-RAPs from Bean and Nielsen.
+\end{quote}
+
+p.
+109: the formulas for \(Q_{++}(\lambda)\) have nice interpretations in terms of \(T -\lambda I\) in the QBD setting.
+Worth mentioning and providing intuition.
+Also, be more specific about their whereabouts in da Silva!
+\begin{quote}
+    \begin{itemize}
+        \item I'm not sure what interpretation they are referring to. Do you have any idea?
+        \item Not sure if I mean to cite da Silva here, but I'm not sure if it is the best reference for these equations, so I have removed it as the expressions in it do not concern time variables.  
+    \end{itemize}
+\end{quote}
+
+p. 110: how do you deduce (5.31). There is a missing link from (5.26) which could be elaborated on.
+\begin{quote}
+    I should elaborate on this. Need help devising a plan of attack. Maybe an inductive argument might be easiest? 
+\end{quote}
+
+p.
+128: top, did I miss that you introduced the multivariate O-function somewhere previously?
+And concerning the error term r4(n), is that actually introduced here in this Lemma?
+Then you might clarify this by writing “...where r4(n) is an error term satisfying...”
+\begin{quote}
+    \begin{itemize}
+        \item I don't think I have introduced big-O notation. I shall do that. It's not multivariate. 
+        \item It is introduced in this Lemma --all the \(r_\ell(n)\) terms are introduced in Lemmas. I wonder why this \(r_4\) was singled out.
+    \end{itemize}
+\end{quote}
+
+\noindent\textbf{CHAPTER 6}
+
+The first part of the chapter shows convergence of the discrete-time approximated level process to the embedded level process of the fluid queue.
+The second part is dedicated to the global convergence of the continuous-time processes.
+This can be achieved by the use of the results of Chapter 5, dominated convergence and tedious calculations!
+This chapter is easier to read since it has more structure.
+
+As you comment yourself at the end of the chapter, ideally would have liked to have pointwise convergence, but what you prove is weak convergence.
+
+(6.23): could you elaborate/comment on the insertion of the exponential r.v.? It seems you are integrating out over the tail, and a comment would be in place. Is there a “=” missing?
+\begin{quote}
+    \begin{itemize}
+        \item elaborated.
+``The exponential random variable in the last line of (6.23) appears due to the interpretation of the Laplace transform as probability that \(\tau_n^{(p)}\) occurs before and exponential random variable, \(E^\lambda\).''
+        \item no, I don't think an = is missing.
+    \end{itemize}
+\end{quote}
+
+Several places: you carefully verify conditions in Tonelli’s theorem for how to swap summation and integration.
+I think, alternatively, the Beppo-Levi theorem stating that you can always swap the order whenever the integrant is non-negative would apply.
+\begin{quote}
+    From memory, I justified the swap by the non-negativeness of the summand/integrand where possible.
+I did not assume \(\psi\) was positive, so I had to use Tonelli's theorem many times.
+Also, my Googling tells me Beppo-Levi is like (the same as?) the monotone convergence theorem, but I do not assume any monotonic behaviour, so it's not immediate how I might use it.
+\end{quote}
+
+\noindent\textbf{CHAPTER 7}
+
+This chapter considers some numerical experiments.
+A general problem in this chap-ter is the quality of the plots up to 7.5, particularly the ones concerning the KS-error and L2-errors.
+It is hard to distinguish the different lines in the plots.
+From 7.5 on-wards, it is better though not full publishing quality.
+\begin{quote}
+    Yeah, maybe I do need to remake the plots up to 7.5. Perhaps a solid orange line would be better. 
+\end{quote}
+
+It remains a bit of mystery to me what is happening under the hood.
+How did you choose the parametrisation of the approximating QBD-RAP?
+How do they look like?
+Is 21 really enough?
+Only comparative graphs are presented, but no examples of representations of the approximations.
+\begin{quote}
+    \begin{itemize}
+        \item How did I choose the QBD-RAP? This was the entire purpose of Chapter 4!!! 
+        \item Is 21 enough? Perhaps not.
+        \item I have a couple of examples of approximated PDFs in there -- I could add more. For example, Fig 7.21, 7.18, 7.27, 7.29, 7.31, 7.35. 
+    \end{itemize}
+\end{quote}
+
+In these static cases, you might be able to assess the size of the error and employ a Richardson extrapolation on the dimension (e.g.
+10 and 20) to obtain a much better approximation.
+The same can probably be said about the other schemes.
+\begin{quote}
+    Is this beyond the scope of the thesis? 
+\end{quote}
+
+You conclude overall that the QBD-RAP scheme performs better than the uniformisation scheme; however, in my perception through the examples, and also being very much example dependent, my view is a bit muddier.
+Also, everything seems to be a trade-off between convergence vs.
+quality.
+Of course, converging to a useless approximation might not be considered “convergence” at all.
+\begin{quote}
+    I do not understand this criticism.
+The QBD-RAP beats uniformisation almost all the time.
+I do cherry-pick the examples because I am trying to argue that the QBD-RAP method works well for those``typically hard'' examples.
+What do you mean by convergence and quality?
+It's converging to the true solution!
+What is useless about that?
+\end{quote}
+
+p.177, mid: you mention the problem of integration errors.
+Those are, of course, present also in all other cases.
+How did you control these errors?
+Did you consider employing Richardson extrapolation to reduce the errors?
+\begin{quote}
+    I did not specifically control for these errors.
+I did use the same integration scheme and step size for all methods, so these errors should be somewhat consistent between the methods.
+I did not use Richardson.
+\end{quote}
+
+\st{Example 7.9: what do you mean by a PDF 1(x < 1). Uniform dist. over (0, 1)?}
+
+What about CPU times for the different experiments? Was the limitation of 21 dimensions dictated by execution times?
+\begin{quote}
+    I did not measure CPU times.
+
+    I could do higher than dimension 21 for most examples, except Example 7.13 which is limited by simulations.
+\end{quote}
+
+While Erlangization/Uniformisation is straightforward and hence automatic to implement, I believe this is not the case for the concentrated ME. Please comment on this.
+\begin{quote}
+    They are relatively similar to implement. What sort of commentary would this require?
+\end{quote}
+
 \pagebreak
@@
 
-\textbf{2 About the main sections}
+\noindent\textbf{2 About the main sections}
 
@@
     \begin{itemize}
-        \item I'm not sure what the reviewer means by "uneven level of detail" -- sure, some sections are more technical than others.
+        \item I'm not sure what the reviewer means by "uneven level of detail" -sure, some sections are more technical than others.
 But I think that's just the nature of the topics and the they are presented in the detail required to understand the relevant parts of the thesis.
@@
         \item I like the order of the introduction.
-It builds in a logical order: some fundamental mathematical concepts -- just enought to get going, then I introduce the fluid and fluid-fluid queue, then tools for their approximations, then other sundry mathematical concepts used in the thesis.
+It builds in a logical order: some fundamental mathematical concepts -just enought to get going, then I introduce the fluid and fluid-fluid queue, then tools for their approximations, then other sundry mathematical concepts used in the thesis.
 I don't think suggested order would add value.
@@
 I am happy with the content of these sections, but I know Peter objects to including some of the details which make the presentation more complex.
-        \item Happy to merge 5 and 6. What do you think? I seperated them because they used distinct techniques, but they ultimately combine to prove a single result. 
-        \item Item, happy to add more steps and signposts in Sections 5.2/5.3 or convert them to Lemmas -- perhaps structuring as Lemmas is the way to go as it condeses the important points into a single statement and leaves details to a proof, allowing the reader to forget about the details.
+        \item Happy to merge 5 and 6. What do you think? I separated them because they used distinct techniques, but they ultimately combine to prove a single result. 
+        \item Happy to add more steps and signposts in Sections 5.2/5.3 or convert them to Lemmas -perhaps structuring as Lemmas is the way to go as it condenses the important points into a single statement and leaves details to a proof, allowing the reader to forget about the details.
     \end{itemize}
@@
 I'm already doing \(5\times 10^{10}\) realisations!
-To me, it seems unlikely that investigation of higher-order examples would reveal much more insight -- my intuition is that the error decay of the QBD-RAP method is exponential in the order of the ME used.
+To me, it seems unlikely that investigation of higher-order examples would reveal much more insight -my intuition is that the error decay of the QBD-RAP method is exponential in the order of the ME used.
     \end{itemize}
@@
 
-\textbf{3 List of minor remarks and typos}
+\noindent\textbf{3 List of minor remarks and typos}
 
@@
 \end{quote}
-\item \st{page 13: 'otherwise it is reflected' - I understand why the word 'reflected' is used here, but usually a reflecting boundary condition means evolving with -ci instead of ci.
+\item \st{page 13: 'otherwise it is reflected' I understand why the word 'reflected' is used here, but usually a reflecting boundary condition means evolving with -ci instead of ci.
 Here, this is simply an absorbing boundary condition, with transition upon reaching the boundary.}
@@
 \item \st{page 18: 'which is only possibly'  'which is only possible'}
-\item page 19: 'Conditions on fluid-fluid queues to ensure that (2.20) will be differentiable with respect to y are not known, and here we just assume that it is.' - this seems like an innocent assumption.
+\item page 19: 'Conditions on fluid-fluid queues to ensure that (2.20) will be differentiable with respect to y are not known, and here we just assume that it is.' this seems like an innocent assumption.
 Why don't we have this at least for some nice cases?
@@
 \item \st{page 30, before (2.35): extra ) bracket in ()}
-\item page 35: Section 2.5.5. about accuracy is indeed very brief and seems highly non-
-rigorous.
+\item page 35: Section 2.5.5. about accuracy is indeed very brief and seems highly nonrigorous.
 \begin{quote}
@@
 thorough read-through of Section 2.6.1.}
-\item \st{page 39: 'the process may not actually jump at these times' - reword to something
+\item \st{page 39: 'the process may not actually jump at these times' reword to something
 like 'the process does not necessarily jump at these times'}
@@
 \item \st{page 48, near the top: 'Fubini's Theorem us to' to 'Fubini's Theorem allows us to'}
-\item \st{page 53: 'The DG method conserves probability' - so any quantity from [0,1] remains
+\item \st{page 53: 'The DG method conserves probability' so any quantity from [0,1] remains
 in [0,1] after the approximation, or does it stay completely unchanged? Or does this
@@
 \begin{quote}
-    I have clarified that it conerves total probability of the system and added a reference to cor 3.1.
+    I have clarified that it conserves total probability of the system and added a reference to cor 3.1.
 \end{quote}
@@
 \end{quote}
-\item \st{page 56: There are different choices for the flux - rather, for the choice of approxi-
-mation for the flux, right?}
-\item page 60, middle of page: overlong line
-\begin{quote}
-    Fixed? Check once compiled.
-\end{quote}
+\item \st{page 56: There are different choices for the flux rather, for the choice of approximation for the flux, right?}
+\item \st{page 60, middle of page: overlong line}
 \item \st{page 60, bottom formula, first line: the 'or' part should be re-formulated (e.g.
@@
 \item \st{page 70, middle: 'we can show even show' to 'we can even show'}
-\item \st{page 70: 'on the event that phi(t) is constant, then X(t) moves deterministically' -
-'then' is unnecessary}
-\item \st{page 80, near top: 'Other choices are possible' ... 'are other possible choices' -
-unnecessary repeat}
+\item \st{page 70: 'on the event that phi(t) is constant, then X(t) moves deterministically' 'then' is unnecessary}
+\item \st{page 80, near top: 'Other choices are possible' ... 'are other possible choices' unnecessary repeat}
 \item \st{page 88, top line: 'events epochs' to 'event epochs'}
-\item \st{page 102, bottom: use the same function (e.g. g(x)) in both formulas}
-\item page 104: (5.10) and (5.11) seem to be a single formula broken into two parts
-unnecessarily.
-\begin{quote}
-    Check once compiled.
-\end{quote}
+\item \st{page 102, bottom: use the same function (e.g. g(x)) in both formulae}
+\item \st{page 104: (5.10) and (5.11) seem to be a single formula broken into two parts
+unnecessarily.}
 \item \st{page 106, near bottom: 'in terms of the of first return matrices' to 'in terms of the
 first return matrices'}
-\item \st{page 108, line before (5.25): you could simply write q, r in +, - (later as well)}
+\item \st{page 108, line before (5.25): you could simply write q, r in +, (later as well)}
 \item page 111: maybe Corollary D.3 from Appendix D could be moved to the main text
 \begin{quote}
-    It could be but I dont think it needs to be moved. Thoughts?
+    It could be but I don't think it needs to be moved. Thoughts?
 \end{quote}
@@
 \item \st{page 154: 'where the swap of the intergals' to 'where the swap of the integrals'}
-\item page 155: 'The probability of 1 is' - this looks a bit strange, maybe use A and B to
-denote those events
-\begin{quote}
-    Check once compiled.
-\end{quote}
+\item \st{page 155: 'The probability of 1 is' this looks a bit strange, maybe use A and B to
+denote those events}
 \item \st{page 161: the bottom formula can fit in a single line}
@@
 \item page 184, Fig. 7.15 explanation: 'Interestingly, the error curve for the DG scheme
-is not monotonic.' - any ideas why?
+is not monotonic.' any ideas why?
 \begin{quote}
@@
 \end{quote}
-\item \st{page 265: Borthwick (n.d.) - why is there no date? I found a 2016 edition.}
+\item \st{page 265: Borthwick (n.d.) why is there no date? I found a 2016 edition.}
 \end{itemize}
 
-\begin{thebibliography}{9}
-    \bibitem{bn2010}
-    Bean, N.~G.~and Nielsen, B.~F. (2010). {`Quasi-birth-and-death processes with rational arrival process components.}' \textit{Stochastic Models}, 26(3), 309-334.
-\end{thebibliography}
 \end{document}

### 2023-01-22 appendix_dg_toy/appendix.tex
post-reviews, pre supervisor feedback

@@
 
-Consider a process \(\{( \dot Y(t),X(t),\varphi(t))\}_{t\geq 0}\) with two phases, \(\varphi(t)\in\calS=\{1,2\}\) and generator matrix \(\bs T\).
-Let \(b = 1.8\), and the cell edges be \(y_0=0,\,y_1=1,\,y_2=1.8\).
+Consider a process \(\{( \dot W(t),X(t),\varphi(t))\}_{t\geq 0}\) with two phases, \(\varphi(t)\in\calS=\{1,2\}\) and generator matrix \(\bs T\).
+Let \(b = 1.8\), and the cell edges be \(w_0=0,\,w_1=1,\,w_2=1.8\).
 We choose a basis of Lagrange polynomials of order 1 to define our approximation space.
@@
 \end{tikzpicture}
-\caption{A mesh with nodes \(y_0=0\), \(y_1=1\) and \(y_2=1.8\). There are two basis functions on each cell. Boundaries are located at \(y_0=0\) and \(y_2=1.8\).\label{fig: mesh}}
+\caption{A mesh with nodes \(w_0=0\), \(w_1=1\) and \(w_2=1.8\). There are two basis functions on each cell. Boundaries are located at \(w_0=0\) and \(w_2=1.8\).\label{fig: mesh}}
 \end{figure}
@@
 
-Next, given a value of \(y\), we solve the system \eqref{eqn:pisystem1}-\eqref{eqn:pisystem2} to find \( {\bs p}= {\bs p}^-\) and \( {\bs \pi}(y)\).
+Next, given a value of \(w\), we solve the system \eqref{eqn:pisystem1}-\eqref{eqn:pisystem2} to find \( {\bs p}= {\bs p}^-\) and \( {\bs \pi}(w)\).
 
@@
 \[ {\bs p}^- = \left[\begin{array}{c|cc|cc|c} {p}_{-1,2} &  {p}_{0,2}^1 &  {p}_{0,2}^2 &  {p}_{1,1}^1 &  {p}_{1,1}^2 &  {p}_{K+1,1} \end{array}\right],\]
-where \( {p}_{-1,2}\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t)=0,X({t}) =0, \varphi_{t} = 2\right)\) and \( {p}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t)=0, X(t) =1.8, \varphi({t}) = 1\right)\).
-For \(x\in\calD_{0,2}\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t)=0, X(t) \in \wrt x, \varphi({t}) = 2\right)\), is constructed as \( {p}_{0,2}^1\phi_0^1(x) +  {p}_{0,2}^2\phi_0^2(x)\).
-For \(x\in\calD_{1,1}\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t)=0, X(t) \in \wrt x, \varphi({t}) = 1\right)\), is constructed as \( {p}_{1,1}^1\phi_1^1(x) +  {p}_{1,1}^2\phi_1^2(x).\)
+where \( {p}_{-1,2}\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t)=0,X({t}) =0, \varphi_{t} = 2\right)\) and \( {p}_{K+1,1}\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t)=0, X(t) =1.8, \varphi({t}) = 1\right)\).
+For \(x\in\calD_{0,2}\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left( \dot W(t)=0, X(t) \in \wrt x, \varphi({t}) = 2\right)\), is constructed as \( {p}_{0,2}^1\phi_0^1(x) +  {p}_{0,2}^2\phi_0^2(x)\).
+For \(x\in\calD_{1,1}\), an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t)=0, X(t) \in \wrt x, \varphi({t}) = 1\right)\), is constructed as \( {p}_{1,1}^1\phi_1^1(x) +  {p}_{1,1}^2\phi_1^2(x).\)
 
-Similarly, for \( {\bs\pi}^-(y)\), we have 
-\[ {\bs \pi}^-(y) = \left[\begin{array}{c|cc|cc|c} {\pi}_{-1,2}(y) &  {\pi}_{0,2}^1(y) &  {\pi}_{0,2}^2(y) &  {\pi}_{1,1}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{K+1,1}(y) \end{array}\right],\]
-where \( {\pi}_{-1,2}(y)\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(  Y(t)\in\wrt y, X(t) =0, \varphi({t}) = 2\right)\) and \( {\pi}_{K+1,1}(y)\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t)\in \wrt y, X(t) =1.8, \varphi({t}) = 1\right)\).
-For \(x\in\calD_{0,2}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y, X(t) \in \wrt x, \varphi({t}) = 2\right)\), is constructed as \( {\pi}_{0,2}^1(y)\phi_0^1(x) +  {\pi}_{0,2}^2(y)\phi_0^2(x)\).
-For \(x\in\calD_{1,1}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t) \in \wrt y, X(t) \in \wrt x, \varphi({t}) = 1\right)\), is constructed as \( {\pi}_{1,1}^1(y)\phi_1^1(x) +  {\pi}_{1,1}^2(y)\phi_1^2(x).\)
+Similarly, for \( {\bs\pi}^-(w)\), we have 
+\[ {\bs \pi}^-(w) = \left[\begin{array}{c|cc|cc|c} {\pi}_{-1,2}(w) &  {\pi}_{0,2}^1(w) &  {\pi}_{0,2}^2(w) &  {\pi}_{1,1}^1(w) &  {\pi}_{1,1}^2(w) &  {\pi}_{K+1,1}(w) \end{array}\right],\]
+where \( {\pi}_{-1,2}(w)\) is an approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left(  \dot W(t)\in\wrt w, X(t) =0, \varphi({t}) = 2\right)\) and \( {\pi}_{K+1,1}(w)\) is an approximation to the artificial point mass \( \lim\limits_{t\to\infty}\mathbb{P}\left( \dot W(t)\in \wrt w, X(t) =1.8, \varphi({t}) = 1\right)\).
+For \(x\in\calD_{0,2}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t) \in \wrt w, X(t) \in \wrt x, \varphi({t}) = 2\right)\), is constructed as \( {\pi}_{0,2}^1(w)\phi_0^1(x) +  {\pi}_{0,2}^2(w)\phi_0^2(x)\).
+For \(x\in\calD_{1,1}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left( \dot W(t) \in \wrt w, X(t) \in \wrt x, \varphi({t}) = 1\right)\), is constructed as \( {\pi}_{1,1}^1(w)\phi_1^1(x) +  {\pi}_{1,1}^2(w)\phi_1^2(x).\)
 
-For \( {\bs\pi}^+(y)\), we have 
-\[ {\bs \pi}^+(y) = \left[\begin{array}{cc|cc}  {\pi}_{0,1}^1(y) &  {\pi}_{0,1}^2(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{1,2}^2(y) \end{array}\right].\]
-For \(x\in\calD_{0,1}\) an approximation to the joint density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y, X_t \in \wrt x, \varphi({t}) = 1\right)\) is constructed as \( {\pi}_{0,1}^1(y)\phi_0^1(x) +  {\pi}_{0,1}^2(y)\phi_0^2(x)\).
-For \(x\in\calD_{1,2}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y,  X(t) \in \wrt x, \varphi({t}) = 2\right)\) is constructed as \( {\pi}_{1,2}^1(y)\phi_1^1(x) +  {\pi}_{1,2}^2(y)\phi_1^2(x).\)
+For \( {\bs\pi}^+(w)\), we have 
+\[ {\bs \pi}^+(w) = \left[\begin{array}{cc|cc}  {\pi}_{0,1}^1(w) &  {\pi}_{0,1}^2(w) &  {\pi}_{1,2}^1(w) &  {\pi}_{1,2}^2(w) \end{array}\right].\]
+For \(x\in\calD_{0,1}\) an approximation to the joint density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t) \in \wrt w, X_t \in \wrt x, \varphi({t}) = 1\right)\) is constructed as \( {\pi}_{0,1}^1(w)\phi_0^1(x) +  {\pi}_{0,1}^2(w)\phi_0^2(x)\).
+For \(x\in\calD_{1,2}\) an approximation to the density of \( \lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t) \in \wrt w,  X(t) \in \wrt x, \varphi({t}) = 2\right)\) is constructed as \( {\pi}_{1,2}^1(w)\phi_1^1(x) +  {\pi}_{1,2}^2(w)\phi_1^2(x).\)
 
@@
 \begin{align*}
-	\lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y,  X(t) \in \wrt x, \varphi({t}) = i\right) &\approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {\pi}_{i,k}^r(y)\phi_k^r(x)\wrt x \wrt y ,\intertext{ for \(x\in(0,1.8),\, y>0, and\)}
-	\lim\limits_{t\to\infty}\mathbb{P}\left( Y(t)=0, X(t) \in \wrt x, \varphi({t}) = i\right) & \approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {p}_{k,i}^r\phi_k^r(x)\wrt x ,\quad x\in(0,1.8),\\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(Y(t) \in \wrt y,X(t) =0 ,\varphi({t}) = i\right) & \approx{\pi}_{-1,i}(y)\wrt y,\quad y>0, \\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t) =0, X(t) =0,\varphi({t}) = i\right) & \approx {p}_{-1,i},\quad \\ 
-	 \lim\limits_{t\to\infty}\mathbb{P}\left(  Y(t) \in \wrt y, X(t) = 1.8, \varphi({t}) = i\right) & \approx {\pi}_{K+1,i}(y)\wrt y,\quad y>0, \\
-	 \lim\limits_{t\to\infty}\mathbb{P}\left( Y(t) = 0, X(t) = 1.8, \varphi({t}) = i\right) & \approx {p}_{K+1,i}.\quad
+	\lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t) \in \wrt w,  X(t) \in \wrt x, \varphi({t}) = i\right) &\approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {\pi}_{i,k}^r(w)\phi_k^r(x)\wrt x \wrt w ,\intertext{ for \(x\in(0,1.8),\, w>0, and\)}
+	\lim\limits_{t\to\infty}\mathbb{P}\left( \dot W(t)=0, X(t) \in \wrt x, \varphi({t}) = i\right) & \approx \displaystyle \sum_{r\in\{1,2\},k\in\{0,1\}}  {p}_{k,i}^r\phi_k^r(x)\wrt x ,\quad x\in(0,1.8),\\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(\dot W(t) \in \wrt w,X(t) =0 ,\varphi({t}) = i\right) & \approx{\pi}_{-1,i}(w)\wrt w,\quad w>0, \\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left( \dot W(t) =0, X(t) =0,\varphi({t}) = i\right) & \approx {p}_{-1,i},\quad \\ 
+	 \lim\limits_{t\to\infty}\mathbb{P}\left(  \dot W(t) \in \wrt w, X(t) = 1.8, \varphi({t}) = i\right) & \approx {\pi}_{K+1,i}(w)\wrt w,\quad w>0, \\
+	 \lim\limits_{t\to\infty}\mathbb{P}\left( \dot W(t) = 0, X(t) = 1.8, \varphi({t}) = i\right) & \approx {p}_{K+1,i}.\quad
 \end{align*}
-%\[\left[\begin{array}{c|cc|cc|cc|cc|c} {\pi}_{2,{-1}}(y) &  {\pi}_{1,1}^1(y) &  {\pi}_{1,2}^1(y) &  {\pi}_{2,1}^1(y) &  {\pi}_{2,2}^1(y) &  {\pi}_{1,1}^2(y) &  {\pi}_{1,2}^2(y) &  {\pi}_{2,1}^2(y) &  {\pi}_{2,2}^2(y) &  {\pi}_{1,{K+1}}(y) \end{array}\right]\]
+%\[\left[\begin{array}{c|cc|cc|cc|cc|c} {\pi}_{2,{-1}}(w) &  {\pi}_{1,1}^1(w) &  {\pi}_{1,2}^1(w) &  {\pi}_{2,1}^1(w) &  {\pi}_{2,2}^1(w) &  {\pi}_{1,1}^2(w) &  {\pi}_{1,2}^2(w) &  {\pi}_{2,1}^2(w) &  {\pi}_{2,2}^2(w) &  {\pi}_{1,{K+1}}(w) \end{array}\right]\]

### 2023-01-22 appendix_kronecker/appendix.tex
post-reviews, pre supervisor feedback

@@
 \end{align}
-Denote by \(\widehat{\bs H}_{11}(\lambda):=\displaystyle \int_{t=0}^\infty e^{-\lambda t}\bs H_{11}(t)\wrt t\) and by \(\widehat{\bs H}_{12}(\lambda):=\displaystyle \int_{t=0}^\infty e^{-\lambda t}\bs H_{12}(t)\wrt t\), the Laplace transforms of \(\bs H_{11}(t)\) and \(\bs H_{12}(t)\), respectively.
+Denote by \(\widehat{\bs H}_{11}(\lambda)=\displaystyle \int_{t=0}^\infty e^{-\lambda t}\bs H_{11}(t)\wrt t\) and by \(\widehat{\bs H}_{12}(\lambda)=\displaystyle \int_{t=0}^\infty e^{-\lambda t}\bs H_{12}(t)\wrt t\), the Laplace transforms of \(\bs H_{11}(t)\) and \(\bs H_{12}(t)\), respectively.
 Using Lemma~\ref{lem: sfkjgn} we can show the following result.

### 2023-01-22 appendix_math_background/appendix.tex
.
post-reviews, pre supervisor feedback

@@
 %!TEX root = ../thesis.tex
-\chapter{More mathematical background}\label{app: math background}
-\section{Time-integration schemes}\label{sec: time integration}
-% SSPRK
-% CFL
+\chapter{Mathematical background}\label{app: math background}
+\section{Laplace Transforms}
+The Laplace transform is an important tool in many areas of mathematics, and is particularly useful in the analysis of fluid queues and fluid-fluid queues.
+Furthermore, in Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with {Laplace transforms} to prove convergence of the QBD-RAP scheme.
+
+For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transform of \(\mu\) to be
+\[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
+where the \emph{region of convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
+\[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
+When \(\mu\) is the probability measure associated with a random variable, \(Z\), say, then we may write 
+\[\widehat \mu(\lambda)=\mathbb E[e^{-\lambda Z}],\]
+and the region of convergence is at least \([0,\infty)\). 
+Further, letting \(E^\lambda\) be an exponentially distributed random variable with rate \(\lambda\) and noting that \(\mathbb P(E^\lambda >t ) = e^{-\lambda t}\), then 
+\[\widehat \mu(\lambda)=\mathbb P(Z<E^\lambda),\]
+which gives a probabilistic interpretation of the Laplace transform.
+That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(Z\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
+
+A convenient property of the Laplace transform which we utilise is the Convolution Theorem. 
+\begin{thm}[Convolution Theorem]
+	Let \(f,g: [0,\infty) \to \mathbb R\) be integrable functions, then 
+	\[\mathcal L\left(\int_{u=0}^tf(u)g(t-u)\wrt u\right)(\lambda) = \mathcal L\left(f\right)\cdot \mathcal L\left(g\right).\]
+\end{thm}
+The Convolution Theorem states that the Laplace transform of the convolution, given by \(\displaystyle \int_{u=0}^tf(u)g(t-u)\wrt u\), is equal to the product of the Laplace transform of \(f\) and \(g\).
+
+The Laplace transform is unique in the sense that, if \(\mu\) and \(\nu\) are two measures on the Borel sets of \([0,\infty)\) and 
+\[\widehat \mu(\lambda) = \widehat \nu(\lambda)\] 
+for all \(\lambda > a\) with \(a<\infty\), then \(\mu\) and \(\nu\) are the same. In terms of functions, \(f,g: [0,\infty) \to \mathbb R\), if 
+\[\mathcal L(f)(\lambda) = \mathcal L(g)(\lambda),\]
+for all \(\lambda > a\) with \(a<\infty\), and \(f\) and \(g\) are continuous, then \(f(t)=g(t)\) for all \(t\geq 0\). Without knowing \(f\) and \(g\) are continuous, then we can only claim that 
+\(f(t)=g(t)\) for all \(t\geq 0, t\notin \mathcal N,\) where \(\mathcal N\) is a \emph{null set} with respect to Lebesgue measure.
+
+\section{Kronecker products and related results}\label{sec:ksdfkkakaaaaaa}
+Here we describe Kronecker products and sums and some of their properties (see \cite{MEinAP}, and also Appendix A.4, for further details).
+We use the results in this section to manipulate certain matrix expressions in Chapter~\ref{sec: conv}.
+
+Let 
+\[\bs A = \left[\begin{array}{ccc}a_{11} & \dots & a_{1m}\\\hdots & & \hdots \\ a_{n1}&\dots & a_{nm}\end{array}\right]
+\qquad
+\bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
+be matrices with dimensions \(n\times m\) and \(n'\times m'\), respectively. The operator \(\otimes\) is the Kronecker product of two matrices; 
+\[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
+which is an \(nn'\times mm'\) matrix.
+
+Let \(\bs{C},\bs{D}\)  be matrices with dimensions \(m\times k\) and \(m'\times k'\). A property of the Kronecker Product is 
+\begin{align}
+	\left(\bs A\otimes \bs{B}\right)\left(\bs{C}\otimes \bs{D}\right) &= \bs A\bs{C}\otimes \bs{B}\bs{D}.\label{eqn:mpr}\tag{Mixed Product Rule}
+\end{align}
+
+% \begin{proof}
+% 	The proof follows from 
+% 	\begin{align*}
+% 		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j} \bs D \\\vdots\\c_{nj}\bs{D} \end{array}\right] 
+% 		%
+% 		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
+% 		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
+% 	\end{align*}
+% \end{proof}
+
+If \(\bs A\) and \(\bs{B}\) are invertible matrices, then 
+\begin{align}\label{eqn:kron inverse}
+	\left(\bs A\otimes \bs{B}\right)^{-1} = \bs A^{-1}\otimes \bs{B}^{-1}.
+\end{align}
+
+Let \(\bs A\) and \(\bs{B}\) be \(n\times n\) and \(m\times m\) matrices, respectively. The Kronecker sum of \(\bs A\) and \(\bs{B}\) is denoted by \(\oplus\) and defined as 
+\[\bs A\oplus \bs{B} = \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]
+
+The exponential of a square matrix \(\bs B\) is \[e^{\bs B} = \sum_{n=0}^\infty \cfrac{1}{n!}\bs B^n.\]
+
+A property of the Kronecker sum is 
+\begin{align}\label{eqn:kron exp}
+	e^{\bs A\oplus \bs{B}}= e^{\bs A}\otimes e^{\bs{B}}.
+\end{align}
+% \begin{proof}
+% 	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
+% 	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
+% 	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\).
+The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\).
+The former follows from
+% 	\begin{align}
+% 	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
+% 	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
+% 	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
+% 	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
+% 	\end{align}
+
+% 	Therefore 
+% 	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
+% 	and the result follows by the mixed product rule. 
+% \end{proof}
+
+\begin{lem}\label{lem: lst mpr}
+	Let \(\bs{T}\) and \(\bs{C}\) be \(n\times n\), square matrices with \(\bs{C}\) diagonal and invertible; let \(\bs{S}\) be a \(p\times p\) matrix.
+Further, suppose \(\left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]\) is invertible for \(\lambda>0\).
+Then
+\begin{align}
+	&\int_{t=0}^\infty e^{-\lambda t}  e^{{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t}} \wrt t 
+	%
+	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}.  \label{eqn:lstsimplify}\end{align}
+\end{lem}
+\begin{proof}
+	Computing the integral on the left-hand side and then factorising the result and using the~\ref{eqn:mpr} multiple times gives
+	\begin{align}
+            	\int_{t=0}^\infty e^{-\lambda t} e^{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t} \wrt t\nonumber 
+            	%
+            	&= - \left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]^{-1}
+		%
+		\\&= -  \left[\bs{T}\otimes \bs{I} + \left(\bs{C}\otimes \bs{I}\right)\left(\bs{I}\otimes \bs{S}\right) - \lambda \bs{I}\right]^{-1}\nonumber
+		%
+		\\&= -  \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}\otimes \bs{I}\right)^{-1}\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}\otimes \bs{I}\right)^{-1}\lambda \bs{I}\right)\right]^{-1}.
+\label{eqn: ref this one 12}
+		%
+	\end{align}
+	By Equation~(\ref{eqn:kron inverse}) and since \(\bs{C}\) is invertible, (\ref{eqn: ref this one 12}) is equal to
+	\begin{align}
+		& - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\otimes \bs{I}\right)\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\otimes \bs{I}\right)\lambda \bs{I}\right)\right]^{-1}.
+\label{eqn: ref this one 13}
+		%
+	\end{align}
+	{Using the~\ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
+	\begin{align}
+		&- \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\bs{T}\right)\otimes \bs{I} + \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\lambda \bs{I}\right)\otimes \bs{I}\right)\right]^{-1} \nonumber
+		%
+		\\&= - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right)\right]^{-1} \nonumber
+		%
+		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1} \nonumber
+		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1},\label{eqn: is an integral}
+	\end{align}
+	by definition of the Kronecker sum.
+
+	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore, (\ref{eqn: is an integral}) is 
+	\begin{align*}
+		-\left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1}
+		&= \int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x\right)\oplus \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1}.
+	\end{align*}
+	{Using the rule in Equation~(\ref{eqn:kron exp}) gives }
+	\begin{align*}
+		&\int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)x}\otimes e^{ \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1},
+	\end{align*}
+	which is the result.
+\end{proof}
+
+% vector space? Polynomial basis?
+% me basis? semigroup? generator?
+
+%Laplace transforms of probability measures with non-negative support and where the Laplace transform variable, \(\lambda\), is real and non-negative can be given a stochastic interpretation.
+Let \(W\) be a random variable with distribution function \(F_W(w)= \mathbb P(W<w)\), then \(\displaystyle\int_{w=0}^\infty e^{-\lambda w} \wrt F_W(w) = \mathbb P(W < E^{\lambda})\).
+That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(W\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
+
+\section{Convergence theorems}
+We use the following convergence theorems to help us prove that the QBD-RAP scheme converges weakly to the distribution of the fluid queue.
+The first result we state, the Portmanteau Theorem, is a sweeping statement about convergence of measures.
+First, let's define the notion of weak convergence.
+We follow \cite{billingsleyconvergence}.
+
+Let \(S\) be a metric space and let \(\mathcal S\) be the Borel \(\sigma\)-algebra generated by the open sub-sets of \(S\).
+A probability measure, \(P\), is a function which maps elements of \(\mathcal S\) (elements of \(\mathcal S\) are sets) to real numbers in the interval \([0,1]\), with \(P(S)=1\).
+Further, \(P\) is \emph{countably-additive}, which means that for any countable collection of \emph{disjoint} sets \(A_1, A_2,...\in\mathcal S\), then
+\[P\left(\bigcup\limits_{n=1}^\infty A_n\right) = \sum\limits_{n=1}^\infty P\left(A_n\right).\]
+Define the notation \(Pf = \int_S f \wrt P\) where \(f:S\to \mathbb R\) is a function.
+Let \(P_1,P_2,...\), be a sequence of probability measures.
+For a given function \(f\), the sequence \(P_1f,P_2f,...\), is simply a sequence of real numbers.
+The sequence of probability measures \(P_1,P_2,....\), is said to \emph{converge weakly} to \(P\) if \(P_nf\to Pf\) for every bounded continuous real function \(f\).
+\cite{billingsleyconvergence} uses the notation \(P_n\Rightarrow P\) to denote this weak convergence.
+
+We need a few more definitions before stating The Portmanteau Theorem.
+A set \(A\in\mathcal S\) is said to be a \(P\)-\emph{continuity set} if \(\partial A\) (the boundary of the set \(A\)) satisfies \(P(\partial A)=0\).
+Define the \emph{limit inferior} and \emph{limit superior} of a sequence of real numbers \(x_1,x_2,...,\) as
+\begin{align*}
+	\liminf_{n\to\infty} x_n &= \lim_{n\to \infty}	\left(\inf_{m\geq n}x_m\right)
+	\intertext{and}
+	\\\limsup_{n\to\infty} x_n &= \lim_{n\to \infty}	\left(\sup_{m\geq n}x_m\right),
+\end{align*}
+respectively.
+
+\begin{thm}[Portmanteau Theorem I, Theorem~2.1 of \cite{billingsleyconvergence}]
+	These five conditions are equivalent.
+	\begin{enumerate}
+		\item[(i)] \(P_n\Rightarrow P\). 
+		\item[(ii)] \(P_nf\to Pf\) for all bounded, uniformly continuous \(f\).
+		\item[(iii)] \(\limsup_n P_n F \leq P F\) for all closed [sets] \(F\).
+		\item[(iv)] \(\liminf_n P_n G \leq P G\) for all open [sets] \(G\).
+		\item[(v)] \(P_n A \to P A\) for all \(P\)-continuity sets \(F\).
+	\end{enumerate}
+\end{thm}
+Some authors, such as \cite{portmanteaubook}, also include the following equivalent condition in the Portmanteau Theorem. 
+\begin{thm}[Portmanteau Theorem II, Theorem~13.16 of \cite{portmanteaubook}]\label{thm: Portmanteau}
+	\(\,\)
+	\begin{enumerate}
+		\item[(vi)] \(P_nf \to P f\) for all bounded, Lipschitz continuous \(f\).
+	\end{enumerate}
+\end{thm}
+
+Recall that a stochastic process is a sequence of random variables \(\{X(t)\}_{t\in\mathcal T}\).
+The \emph{finite-dimensional distributions} of \(\{X(t)\}_{t\in\mathcal T}\) are the joint distributions of the random vector \((X(t_1),X(t_2),...,X(t_n))\), where \(t_1,t_2,...,t_n\in \mathcal T\) is a finite collection of times.
+
+Key concepts in establishing weak convergence of stochastic processes are the notions \emph{tight} and \emph{relatively compact}.
+Again, we follow \cite{billingsleyconvergence}.
+A probability measure \(P\) is said to be \emph{tight} if for each \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\).
+A family of probability measures, \(\Pi\), is said to be tight if for every \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\) for every probability measure \(P\) in \(\Pi\).
+The family \(\Pi\) is \emph{relatively compact} if every sequence of elements of \(\Pi\) contains a subsequence which converges weakly.
+
+Let \(P_1,P_2,...,\) and \(P\) be probability measures of stochastic processes. \cite{billingsleyconvergence} provides the following result.
+\begin{thm}\label{thm: aofaa}
+	If \(\{P_n\}\) is relatively compact and the finite-dimensional distributions of \(P_n\) converge weakly to those of \(P\), then \(P_n\Rightarrow_n P\). 
+\end{thm}
+Further, the condition that \(\{P_n\}\) is relatively compact can be replaced by tightness due Prohorov's Theorem \citep{billingsleyconvergence}. 
+\begin{thm}
+	If \(\Pi\) is tight, then it is relatively compact.
+\end{thm}
+
+% Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra.
+Further, denote by \(C_b(E)\) the set of continuous bounded functions on \(E\).
+Let \(\mu,\,\mu_1,\,\mu_2,...\in \mathcal M_f(E)\).
+We say that \(\{\mu_n\}_{n\in\mathbb N}\) converges weakly to \(\mu\), formally \(\mu_n\to \mu\) weakly as \(n\to\infty\), if
+% \[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
+% We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}.
+Let \(\mathcal M_{\leq 1}(E)=\{\mu \in \mathcal M_f(E)\mid \mu(E)\leq 1\},\) the set of all sub-probability measures on \((E,\mathcal E)\).
+% \begin{thm}[(Part of) The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
+% 	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
+% 	\begin{enumerate}
+% 		\item[(i)] \(\mu_n\to \mu\) weakly as \(n\to \infty\). 
+% 		\item[(ii)] \(\int f\wrt \mu_n \to \int f\wrt \mu\) for all bounded, Lipschitz continuous \(f\).
+% 		\\ \(\vdots\) 
+% 	\end{enumerate}
+% \end{thm}
+% There are 8 parts to The Portmanteau Theorem in \cite{portmanteaubook}, Theorem~13.16.
+Here we only quote to relevant parts; the other 6 parts require us to define other concepts which are not relevant to this thesis, so they are omitted.
+See \cite{portmanteaubook}, Theorem~13.16 for details.
+
+Another tool we can use to show convergence of measures is to show that the Laplace transforms converge, as stated in the following theorem.
+\begin{thm}[Extended Continuity Theorem, \cite{feller1957}, Theorem 2a]\label{thm: ext cont thm}
+	For \(p=1,2,...,\) let \(U_p\) be a measure with Laplace transform \(\zeta_p\).
+If \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda > a\geq 0\), then \(\zeta\) is the Laplace transform of a measure \(U\) and \(U_p\to U\) [weakly].
+
+	Conversely, if \(U_p\to U\)[weakly] and the sequence \(\{\zeta_p(a)\}\) is bounded, then \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda >a\). 
+\end{thm}
+
+In Chapters~\ref{sec: conv} and~\ref{ch: global results} we use the Dominated Convergence Theorem to aid our convergence arguments.
+In applied probability we often want to prove convergence of certain expressions.
+An approach which can simplify matters is to partition the expression on certain events where we have a simpler characterisation, thereby enabling us to prove convergence on each element in the partition.
+The original expression can be written as an integral over the partition.
+To establish the convergence result we initially desired, we can use the convergence of each element of the partition and the Dominated Convergence Theorem.
+I have taken the following from Theorem~1.13~\cite{steinreal}.
+\cite{steinreal} use the notation
+\[\int f = \int f \wrt x = \int f \wrt m(x),\]
+where \(m\) denotes Lebesgue measure, to denote the Lebesgue integral. 
+\begin{thm}[Dominated Convergence Theorem]
+	Suppose \(\{f_n\}\) is a sequence of measurable functions such that \(f_n(x)\to f(x)\) almost everywhere with respect to \(x\), as \(n\) tends to infinity.
+If \(|f_n(x)|\leq g(x)\), where \(g\) is integrable, then
+	\[\int|f_n-f|\to 0 \mbox{ as } n \to \infty,\]
+	and consequently 
+	\[\int f_n\to\int f \mbox{ as } n\to \infty.\]
+\end{thm}
+
+Also in Chapters~\ref{sec: conv} and~\ref{ch: global results} we want to manipulate infinite sums or integrals and rearrange the order of summation or integration.
+However, things can go awry when we swap the order of integration/summation if we are not careful.
+The next few results give some conditions under which we have equality under before and after swapping the order of summation/integration.
+Once again, we follow \cite{steinreal} quoting their Theorem~2.13.
+If \(f\) is a function in \(\mathbb R^{d}=\mathbb R^{d_1}\times \mathbb R^{d_2}\), the \emph{slice} of \(f\) corresponding to \(y\in\mathbb R^{d_2}\) is the function \(f^y\) of the \(x\in\mathbb R^{d_1}\) variable, given by
+\[f^y(x)=f(x,y).\]
+Similarly, the slice of \(f\) for a fixed \(x\in\mathbb R^{d_1}\) is \(f_x(y)=f(x,y)\). 
+\begin{thm}[Fubini's Theorem]\label{Fubini}
+	Suppose \(f(x,y)\) is integrable on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
+	\begin{itemize}
+		\item The slice \(f^y\) is integrable on \(\mathbb R^{d_1}\).
+		\item The function defined by \(\int_{R^{d_1}} f^y(x)\wrt x\) is integrable on \(\mathbb R^{d_2}\). 
+	\end{itemize}
+	Moreover:
+	\begin{itemize}
+		\item[(iii)] \(\displaystyle\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y = \int_{\mathbb R^d}f.\)
+	\end{itemize}
+\end{thm}
+\cite{steinreal} then state \begin{quotation}``Clearly, the [Fubini] theorem is symmetric in \(x\) and \(y\) so that we also may conclude that the slice \(f_x\) is integrable on \(\mathbb R^{d_2}\) for almost every \(x\).
+Moreover, \(\int_{\mathbb R^{d_2}}f_x(y)\wrt y\) is integrable and
+\[\displaystyle\int_{R^{d_1}}\left(\int_{R^{d_2}}f(x,y)\wrt y\right)\wrt x = \int_{\mathbb R^d}f.\]
+In particular, Fubini's theorem states that the integral of \(f\) on \(\mathbb R^d\) can be computed by iterating lower-dimensional integrals, and that the iterations can be taken in any order
+\[\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y=\displaystyle\int_{R^{d_1}}\left(\int_{R^{d_2}}f(x,y)\wrt y\right)\wrt x = \int_{\mathbb R^d}f.\mbox{''}\]
+\end{quotation}
+It is this last statement which is most powerful. Effectively, if either
+\[\int_{R^{d_2}}\left(\int_{R^{d_1}}|f(x,y)|\wrt x\right)\wrt y<\infty,\]
+or
+\[\int_{R^{d_1}}\left(\int_{R^{d_2}}|f(x,y)|\wrt y\right)\wrt x<\infty,\]
+then we can swap the order of integration.
+
+A closely related theorem which is often used alongside Fubini's Theorem is Tonelli's Theorem.
+Define the \emph{extended Lebesgue integral} of an extended valued (it can take the value \(+\infty\)) non-negative function \(f\) by
+\[\int f(x)\wrt x = \sup_{g}\int g(x)\wrt x.\]
+\begin{thm}[Tonelli's Theorem]\label{Tonelli}
+	Suppose \(f(x,y)\) is a non-negative measurable function on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
+	\begin{itemize}
+		\item The slice \(f^y\) is integrable on \(\mathbb R^{d_1}\).
+		\item The function defined by \(\int_{R^{d_1}} f^y(x)\wrt x\) is integrable on \(\mathbb R^{d_2}\). 
+	\end{itemize}
+	Moreover:
+	\begin{itemize}
+		\item[(iii)] \(\displaystyle\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y = \int_{\mathbb R^d}f(x,y)\wrt x\wrt y \mbox{ in the extended sense}.\)
+	\end{itemize}
+\end{thm}
+
+Once again, we note that the theorem is symmetric in \(x\) and \(y\), so we can establish that we may swap the order of integration provided that \(f\) is non-negative.
+Tonelli's Theorem allows us to exchange the order of integration for any non-negative extended real-valued function, and includes the case where the value of the integral is infinite.
+In contrast, Fubini's Theorem allows us to exchange the order of integration for any real-valued function, provided that the integral is finite.
+
+Collectively, we refer to Theorems~\ref{Fubini} and~\ref{Tonelli} together as the Fubini-Tonelli Theorem, but they are otherwise known collectively as just Fubini's Theorem.
+Often, they are used in conjunction.
+Since \(|f|\) is a non-negative function, then we may use Tonelli's Theorem and compute (or bound) the integral \(\int |f|\) via computing an iterated integral.
+If this is found to be finite, then Fubini's Theorem applies so \(f\) is integrable, and we may evaluate \(\int f\) via an iterated integral.
+
+In the context of probability, the function \(f\) which we are integrating is often positive, so Tonelli's Theorem is all that is required to justify a swap of iterated integrals.
+
+\section{Sundry mathematical concepts}\label{eqn: la}
+\subsection*{Polynomials}
+The discontinuous Galerkin method involves projecting the operator equation onto a basis of functions, typically polynomials.
+A convenient basis with which to work is the interpolating Lagrange polynomials.
+The order \(p-1\) Lagrange polynomials are defined by a set of points, \(\xi_i, i=1,...,p\), (the \(\xi_i\)'s must be distinct), and are given by
+\[\ell_i(r)=\prod_{\substack{j=1\\j\neq i}}^p \cfrac{r-\xi_j}{\xi_i-\xi_j},\, i=1,...,p.\]
+A convenient property of the Lagrange polynomials is 
+\[l_i(\xi_j)=\begin{cases}1 & \mbox{ if } i=j, \\ 0 & \mbox{ otherwise.} \end{cases}\]
+
+Sometimes it is more convenient to work with \emph{orthogonal} polynomials.
+Two functions \(f\) and \(g\) are orthogonal on a set \(\calD\) if \(\int_{x\in\calS}f(x)g(x)\wrt x = 0\).
+Let \(U\) be a vector space of functions (for example, the space of polynomials of order \(p-1\)).
+The orthogonal complement of \(U\), denoted \(U^\perp\), is the set of functions \(f\) such that \(f\) is orthogonal to every \(g\in U\), \(\int_{x\in\calD} f(x)g(x)\wrt x\).
+Any function \(h(x)\) can be decomposed into \(h(x) = h^{U}(x)+h^\perp(x)\) where \(h^{U}\in U\) and \(h^\perp \in U^\perp\), the orthogonal complement of \(U\).
+
+The Legendre polynomials are a set of orthogonal polynomial which are defined recursively by 
+\begin{align*}
+	P_0(x)&=1,
+	\\P_1(x)&=x, 
+	\\(n+1)P_{n+1}(x)&=(2n+1)xP_n(x)-nP_{n-1}(x),
+\end{align*}
+and are orthogonal on \([-1,1]\).
+
+The zeros of \((1-x^2)P_n'(x)\) define the \emph{Legendre-Gauss-Lobatto} points, which are used, among other things, for numerically approximating integrals via quadrature.
+
+\subsection*{Numerical integration}
+To numerically approximate integrals in the discontinuous Galerkin schemes in this thesis we can use Gauss-Lobatto quadrature.
+Consider the integral of some function \(f\) over the interval \([-1,1]\).
+Quadrature approximates the integral by evaluating the function on a set of points, \(x_i,\, i=1,...,p\), and computing the weighted sum
+\[\int_{-1}^1f(x)\wrt x \approx \sum_{i=1}^p f(x_i)w_i,\]
+where \(w_i\) are weights. There are various quadrature schemes which one can use.
+
+For the discontinuous Galerkin method, Gauss-Lobatto quadrature is convenient as in both we evaluate the function at the end points of the interval.
+For Gauss-Lobatto quadrature, the nodes, \(x_i, i=1,...,p\) is the zero of \((1-x^2)P_n'(x)\) where \(P_n\) is the \(n\)th Legendre polynomial, the weights are given by
+\begin{align*}
+	w_1&=1,
+	\\w_i&=\cfrac{2}{n(n-1)[P_{n-1}(x_i)]^2}, \, i\neq 1, p,
+	\\w_p&=1.
+\end{align*}
+Gauss-Lobatto quadrature is accurate for polynomials up to degree \(2p-3\). An integral over the interval \([a,b]\) can be approximated by 
+\[\int_{a}^bf(x)\wrt x \approx \cfrac{b-a}{2}\sum_{i=1}^p f\left(\cfrac{b-a}{2}x_i+\cfrac{a+b}{2}\right)w_i.\]
+
+We also use a trapezoidal integration rule at times. Consider the integral \(\int_{a}^bf(x)\wrt x\). A trapezoidal rule with \(p\) evenly spaced grid-points approximates the integral via 
+\[\int_{a}^bf(x)\wrt x \approx \sum_{i=2}^p \cfrac{f(x_{i-1}+f(x_i))}{2}\Delta,\]
+where \(x_i=a+(i-1)\Delta\), \(i=1,...,p\), and \(\Delta = \cfrac{b-a}{p-1}.\)
+
+\subsection*{Measuring the difference between distributions}
+In the numerical investigations in this thesis we evaluate approximation schemes by comparing the resulting approximations they produce with a ground truth.
+For comparing distributions we use \(L^p\) \emph{norms} and the Kolmogorov-Smirnov distance.
+The \(L^p\) norm of a function \(f\) is given by
+\[\left(\int |f(x)|^p\wrt x\right)^{1/p}.\]
+The Kolmogorov-Smirnov distance between two distribution functions, \(F_1\), \(F_2\) is 
+\[\sup_{x}|F_1(x)-F_2(x)|.\]

### 2023-01-22 appendix_properties_of_B/appendix.tex
finito
first iteration of changes for reviewer 2 and response to reviewer 2 comments (to do: reviewer 1 changes and responses)
post-reviews, pre supervisor feedback

@@
 
-Recall that the coefficients \({\boldsymbol a}_{k,i}(t)\) can be used to construct an approximate solution to a differential equation at time \(t\) as \(u_{k,i}(x,t)={\boldsymbol a}_{k,i}(t)\boldsymbol \phi_k(x)\tr{}.\) For \(i\in\calS,\,k\in\mathcal K,\,r\in\{1,...,p_k\}\), define \(\alpha_{k,i}^r(t) := a_{k,i}^r(t)\displaystyle\int_{x=y_k}^{y_{k+1}}\phi^r_k(x)\wrt x,\) and row-vectors \(\boldsymbol \alpha_{k,i}(t)=(\alpha_{k,i}^r(t))_{r\in\{1,...,p_k\}}\).
+Recall that the coefficients \({\boldsymbol a}_{k,i}(t)\) can be used to construct an approximate solution to a differential equation at time \(t\) as \(u_{k,i}(x,t)={\boldsymbol a}_{k,i}(t)\boldsymbol \phi_k(x)\tr{}.\) For \(i\in\calS,\,k\in\mathcal K,\,r\in\{1,...,p_k\}\), define \(\alpha_{k,i}^r(t) = a_{k,i}^r(t)\displaystyle\int_{x=y_k}^{y_{k+1}}\phi^r_k(x)\wrt x,\) and row-vectors \(\boldsymbol \alpha_{k,i}(t)=(\alpha_{k,i}^r(t))_{r\in\{1,...,p_k\}}\).
 Motivated by the fact that we may be interested in approximations of the probabilities \(\mathbb P(X(t)\in\calD_{k,i},\varphi(t)=i)\) rather than the function \(u_{k,i}\) itself, we can pose the problem equivalently in terms of the integrals
@@
 \begin{align*}
-\bs P_k &= \diag\left(\int_{x=y_k}^{y_{k+1}}\phi^r_k(x)\wrt x\right)_{ r\in\{1,...,p_k\}},\, k\in\mathcal K^\circ,
+\bs P_k &= \mathrm{diag}\left(\int_{x=y_k}^{y_{k+1}}\phi^r_k(x)\wrt x\right)_{ r\in\{1,...,p_k\}},\, k\in\mathcal K^\circ,
 \\\bs P &=\left[\begin{array}{ccc}
@@
 \begin{align*}
- {  {\mathfrak{\bs B}}}^{{-1}0} &:= \bs T_{{-1}+}\otimes\left( \boldsymbol\phi_0(0)\bs M_0^{-1}\bs P_0\right),
+ {  {\mathfrak{\bs B}}}^{{-1}0} &= \bs T_{{-1}+}\otimes\left( \boldsymbol\phi_0(0)\bs M_0^{-1}\bs P_0\right),
  %
- \\ {  {\mathfrak{\bs B}}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1} 1{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}}\otimes \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{},
+ \\ {  {\mathfrak{\bs B}}}^{0{-1}} &=-\left[c_i\widecheck p_{ij} 1{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}}\otimes \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{},
  %
- \\ {  {\mathfrak{\bs B}}}^{{K+1} K} &:= \bs T_{{K+1}-}\otimes \left(\boldsymbol\phi_K(b)\bs M_K^{-1}\bs P_K\right)
+ \\ {  {\mathfrak{\bs B}}}^{{K+1} K} &= \bs T_{{K+1}-}\otimes \left(\boldsymbol\phi_K(b)\bs M_K^{-1}\bs P_K\right)
  %
- \\ {  {\mathfrak{\bs B}}}^{K {K+1}} &:= \left[c_ip_{ij}^{K+1} 1{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes  \bs P_K^{-1} \boldsymbol\phi_K(b)\tr{},
+ \\ {  {\mathfrak{\bs B}}}^{K {K+1}} &= \left[c_i\widehat p_{ij}  1{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes  \bs P_K^{-1} \boldsymbol\phi_K(b)\tr{},
  %
- \\\mathfrak{\bs B}^{kk}_{ij} &:= \begin{cases}
+ \\\mathfrak{\bs B}^{kk}_{ij} &= \begin{cases}
     	T_{ii}\bs I_{p_k} + c_i\bs P_k^{-1}(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1}\bs P_k & i = j, \\
@@
  %
- \\\mathfrak{\bs B}^{00}_{ij} &:= \begin{cases}
+ \\\mathfrak{\bs B}^{00}_{ij} &= \begin{cases}
 	T_{ii}\bs I_{p_0} + c_i\bs P_0^{-1}(\bs F_i^{00}+\bs G_0)\bs M_0^{-1}\bs P_0 & i = j, \\
-T_{ij}\bs I_{p_0} - c_ip_{ij}^{-1} 1(c_i<0)\bs F_i^{00}\bs M_0^{-1}\bs P_0 & i \neq j,
+T_{ij}\bs I_{p_0} - c_i\widecheck p_{ij} 1(c_i<0)\bs F_i^{00}\bs M_0^{-1}\bs P_0 & i \neq j,
 \end{cases}
 %
-\\\mathfrak{\bs B}^{KK}_{ij} &:= \begin{cases}
+\\\mathfrak{\bs B}^{KK}_{ij} &= \begin{cases}
 	T_{ii}\bs I_{p_K} + c_i\bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K & i = j, \\
-T_{ij}\bs I_{p_K} + c_ip_{ij}^{K+1} 1(c_i>0)\bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K & i \neq j,
+T_{ij}\bs I_{p_K} + c_i\widehat p_{ij} 1(c_i>0)\bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K & i \neq j,
 \end{cases}
 %
- \\\mathfrak{\bs B}^{k,k+1}_{ij} &:= \begin{cases}
+ \\\mathfrak{\bs B}^{k,k+1}_{ij} &= \begin{cases}
 	c_i\bs P_k^{-1}\bs F_i^{k,k+1}\bs M_{k+1}^{-1}\bs P_{k+1} & i = j, \\
@@
  %
- \\\mathfrak{\bs B}^{k-1,k}_{ij} &:= \begin{cases}
+ \\\mathfrak{\bs B}^{k-1,k}_{ij} &= \begin{cases}
     	c_i\bs P_k^{-1}\bs F_i^{k,k-1}\bs M_{k-1}^{-1}\bs P_{k-1} & i = j, \\
@@
 		&\sum_{j\in\calS}\mathfrak{\bs B}^{KK}_{ij}\bs 1 + \mathfrak{\bs B}^{K{K+1}}_{ij}\bs 1
-		\\&= \sum_{j\in\calS}T_{ij}\bs I_{p_K}\bs 1-\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 -\sum_{j\in\calS_{{K+1} }}c_ip_{ij}^{K+1} \bs P_K^{-1}\boldsymbol \phi_K(0)\tr{}\bs 1
+		\\&= \sum_{j\in\calS}T_{ij}\bs I_{p_K}\bs 1-\sum_{j\in\calS_-} c_i\widehat p_{ij}  \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 -\sum_{j\in\calS_{{K+1} }}c_i\widehat p_{ij} \bs P_K^{-1}\boldsymbol \phi_K(0)\tr{}\bs 1
 		\\&\quad{} + c_i \bs P_K^{-1}(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1}\bs P_K\boldsymbol 1 
-		\\&= -\sum_{j\in\calS_-} c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 
-		\\&{}\quad-\sum_{j\in\calS_{{K+1}}}c_ip_{ij}^{K+1} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\bs 1+ c_i \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\boldsymbol 1
+		\\&= -\sum_{j\in\calS_-} c_i\widehat p_{ij} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K \bs 1 
+		\\&{}\quad-\sum_{j\in\calS_{{K+1}}}c_i\widehat p_{ij} \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\bs 1+ c_i \bs P_K^{-1}\bs F_i^{KK}\bs M_K^{-1}\bs P_K\boldsymbol 1
 		\\&\quad{}+ c_i \bs P_K^{-1}\bs G_K\bs M_K^{-1}\bs P_K\boldsymbol 1
@@
 		\sum_{j\in\calS}\mathfrak{\bs B}^{00}_{ij}\bs 1 + \mathfrak{\bs B}^{0{-1}}_{ij}\bs 1
-		&= \sum_{j\in\calS}T_{ij}\bs I_{p_0}\bs 1-\sum_{j\in\calS_+} c_ip_{ij}^{-1} \bs P_0^{-1}\bs F_i^{00}\bs M_0^{-1}\bs P_0 \bs 1 -\sum_{j\in\calS_{{-1} }}c_ip_{ij}^{-1} \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{}\bs 1
+		&= \sum_{j\in\calS}T_{ij}\bs I_{p_0}\bs 1-\sum_{j\in\calS_+} c_i\widecheck p_{ij} \bs P_0^{-1}\bs F_i^{00}\bs M_0^{-1}\bs P_0 \bs 1 -\sum_{j\in\calS_{{-1} }}c_i\widecheck p_{ij} \bs P_0^{-1}\boldsymbol \phi_0(0)\tr{}\bs 1
 		\\&\quad{} + c_i \bs P_0^{-1}(\bs F_i^{0,0}+\bs G_0)\bs M_0^{-1}\bs P_0\boldsymbol 1 

### 2023-01-22 chapter1/chapter.tex
.
post-reviews, pre supervisor feedback

@@
 for any \(t,t_0,...,t_n\in\mathcal T\) with \(t_0+t,...,t_n+t\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\).
-A random variable, \(\tau\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\tau\in\sigma(X(s), s\leq t),\) for all \(t\in T\), where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s)\}_{s\leq t}\).
-
-As in \cite[Chapter~1.2]{MEinAP}, we call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable and
+
+As in \cite[Section~2.1]{MEinAP}, a random variable, \(\tau\) taking values in \(\mathcal T\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\{\tau\leq t\}\in\sigma(X(s), s\leq t),\) for all \(t\in T\), where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s)\}_{s\leq t}\).
+
+As in \cite[Section~1.2]{MEinAP}, we call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable and
 \begin{align}\label{eqn: Markov property}
@@
 An \emph{operator} is a mapping which takes elements of a vector space and sends them to elements of a vector space (not necessarily the same space).
-That is, let \(V\) and \(W\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\to W\) is an operator.
+That is, let \(V\) and \(U\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\to U\) is an operator.
 The set \(\mathcal D(\mathbb B)\subseteq V\) is the domain of the operator, and is not necessarily all of \(V\).
-When specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to W\).
-
-Let \(c\in \mathbb R\) (\(\mathbb R\) is the reals), \(v\in V\) and \(w\in W\).
+When specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to U\).
+
+Let \(c\in \mathbb R\) (\(\mathbb R\) is the reals), \(v\in V\) and \(w\in U\).
 Then \(\mathbb B\) is a \emph{linear} operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\).
-We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms defined on \(V\) and \(W\), respectively.
+We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_U}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_U\) are the norms defined on \(V\) and \(U\), respectively.
 An operator is said to be bounded if \(||\mathbb B||<M\) for some \(M<\infty\).
@@
 
-For a bounded linear operator \(\mathbb B:\mathcal D(\mathbb B)\to \mathcal D(\mathbb B)\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\) where, for \(f\in \mathcal D(\mathbb B)\), \(\mathbb B^kf\) is defined recursively by \(\mathbb B^k f = \mathbb B^{k-1} (\mathbb B f)\).
+For a bounded linear operator \(\mathbb B:\mathcal D(\mathbb B)\to \mathcal D(\mathbb B)\), define the operator exponential as \(e^{\mathbb Bt}=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\) where, for \(f\in \mathcal D(\mathbb B)\), \(\mathbb B^kf\) is defined recursively by \(\mathbb B^k f = \mathbb B^{k-1} (\mathbb B f)\).
 It can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), which implies that \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup.
@@
 The Markov property states that, given \(X(t)\), then for \(s,t\geq 0\), \(X(t+s)\) is independent of \(X(u),\, 0\leq u<t\).
-That is, \[P(X(t+s)\in\mathcal{E} \mid X(t), X(u), 0\leq u<t)=P(X(t+s)\in\mathcal{E} \mid X(t)).\]
-If \(P(X(t+s)\in\mathcal{E} \mid X(t))=P(X(s)\in\mathcal{E} \mid X(0))\), are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.
+That is, \[\mathbb P(X(t+s)\in\mathcal{E} \mid X(t), X(u), 0\leq u<t)=\mathbb P(X(t+s)\in\mathcal{E} \mid X(t)).\]
+If \(\mathbb P(X(t+s)\in\mathcal{E} \mid X(t))=\mathbb P(X(s)\in\mathcal{E} \mid X(0))\), are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.
 
@@
 Consider the expectation
-\[\mathbb E[f(X(t))\mid X(0)=x]=\int_{y\in\calS} f(y) P(X(t)\in\wrt y\mid X(0)=x),\]
+\[\mathbb E[f(X(t))\mid X(0)=x]=\int_{y\in\calS} f(y) \mathbb P(X(t)\in\wrt y\mid X(0)=x),\]
 where \(f\) is some bounded function, \(|f|<F\).
@@
 \begin{align*}
-	\mathbb T(t)f(x)&=\int_{y\in\calS} f(y) P(X(t)\in\wrt y \mid X(0)=x)
-	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}f(y) P(X(t)\in\wrt y\mid X(s)=z) P(X(s)\in\wrt z\mid X(0)=x)
-	\\&=\int_{z\in\mathcal S}\mathbb E[f(X(t))\mid X(s)=z] P(X(s)\in\wrt z\mid X(0)=x)
-	\\&=\int_{z\in\mathcal S} \mathbb E[f(X(t-s))\mid X(0)=z]P(X(s)\in\wrt z\mid X(0)=x)
-	\\&=\int_{z\in\mathcal S} (\mathbb T(t-s)f(z))P(X(s)\in\wrt z\mid X(0)=x)
+	\mathbb T(t)f(x)&=\int_{y\in\calS} f(y) \mathbb P(X(t)\in\wrt y \mid X(0)=x)
+	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}f(y) \mathbb P(X(t)\in\wrt y\mid X(s)=z) \mathbb P(X(s)\in\wrt z\mid X(0)=x)
+	\\&=\int_{z\in\mathcal S}\mathbb E[f(X(t))\mid X(s)=z] \mathbb P(X(s)\in\wrt z\mid X(0)=x)
+	\\&=\int_{z\in\mathcal S} \mathbb E[f(X(t-s))\mid X(0)=z]\mathbb P(X(s)\in\wrt z\mid X(0)=x)
+	\\&=\int_{z\in\mathcal S} (\mathbb T(t-s)f(z))\mathbb P(X(s)\in\wrt z\mid X(0)=x)
 	\\&=\mathbb T(s) (\mathbb T(t-s)f)(x),
@@
 % Under certain conditions, a semigroup, (and hence also a Feller process), can be characterised by its generator, \(B\) defined by;
-% \[Bf:=\lim_{t\to 0^+} \frac{1}{t}(K(t)-I)f,\]
+% \[Bf=\lim_{t\to 0^+} \frac{1}{t}(K(t)-I)f,\]
 % on the set of functions \(f\) for which this limit exists.
@@
 
-\section{Laplace Transforms}
-The Laplace transform is an important tool in many areas of mathematics, and is particularly useful in the analysis of fluid queues and fluid-fluid queues.
-Furthermore, in Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with {Laplace transforms} to prove convergence of the QBD-RAP scheme.
-
-For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transform of \(\mu\) to be
-\[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
-where the \emph{region of convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
-\[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
-When \(\mu\) is the probability measure associated with a random variable, \(Z\), say, then we may write 
-\[\widehat \mu(\lambda)=\mathbb E[e^{-\lambda Z}],\]
-and the region of convergence is at least \([0,\infty)\). 
-Further, letting \(E^\lambda\) be an exponentially distributed random variable with rate \(\lambda\) and noting that \(\mathbb P(E^\lambda >t ) = e^{-\lambda t}\), then 
-\[\widehat \mu(\lambda)=\mathbb P(Z<E^\lambda),\]
-which gives a probabilistic interpretation of the Laplace transform.
-That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(Z\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
-
-A convenient property of the Laplace transform which we utilise is the Convolution Theorem. 
-\begin{thm}[Convolution Theorem]
-	Let \(f,g: [0,\infty) \to \mathbb R\) be integrable functions, then 
-	\[\mathcal L\left(\int_{u=0}^tf(u)g(t-u)\wrt u\right)(\lambda) = \mathcal L\left(f\right)\cdot \mathcal L\left(g\right).\]
-\end{thm}
-The Convolution Theorem states that the Laplace transform of the convolution, given by \(\displaystyle \int_{u=0}^tf(u)g(t-u)\wrt u\), is equal to the product of the Laplace transform of \(f\) and \(g\).
-
-The Laplace transform is unique in the sense that, if \(\mu\) and \(\nu\) are two measures on the Borel sets of \([0,\infty)\) and 
-\[\widehat \mu(\lambda) = \widehat \nu(\lambda)\] 
-for all \(\lambda > a\) with \(a<\infty\), then \(\mu\) and \(\nu\) are the same. In terms of functions, \(f,g: [0,\infty) \to \mathbb R\), if 
-\[\mathcal L(f)(\lambda) = \mathcal L(g)(\lambda),\]
-for all \(\lambda > a\) with \(a<\infty\), and \(f\) and \(g\) are continuous, then \(f(t)=g(t)\) for all \(t\geq 0\). Without knowing \(f\) and \(g\) are continuous, then we can only claim that 
-\(f(t)=g(t)\) for all \(t\geq 0, t\notin \mathcal N,\) where \(\mathcal N\) is a \emph{null set} with respect to Lebesgue measure.
-
 \section{Fluid queues}\label{sec: fqs}
-An unbounded fluid queue is a two-dimensional stochastic process which we denote by \(\{\ddot{\bs X}(t)\} = \{(\ddot X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{\ddot X(t)\}_{t\geq0}\) is known as the level process or buffer.
-The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\), and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\).
-We assume that \(\bs T\) is {conservative} and time-homogeneous.
-Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\) which determine the rate at which \(\{X(t)\}\) moves.
-
-Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{i\in\calS\mid c_i<0\}\), \(\calS_0 = \{i\in\calS\mid c_i=0\}\), \(\mathcal S_{-1}=\{i\in\calS\mid c_i\leq0\},\, \mathcal S_{K+1}=\{i\in\calS\mid c_i\geq0\}\)\footnote{The notation \(\mathcal S_{-1}\) and \(\mathcal S_{K+1}\) will make sense later when we introduce the partition into cells for the approximation schemes into \(K\) cells, plus a lower and upper boundary which we represent with indices \(-1\) and \(K+1\) respectively.}.
-We assume, without loss of generality, that the generator \(\bs T\) is partitioned into sub-matrices
-\[\bs T = \left[\begin{array}{ccc}\bs T_{++} & \bs T_{+-} & \bs T_{+0} \\ \bs T_{-+} & \bs T_{--} & \bs T_{-0} \\ \bs T_{0+} & \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
-where \(\bs T_{mn} = [T_{ij}]_{i\in\mathcal S_m, j\in\mathcal S_n}\), \(m,n\in\{+,-,0\}\).
-
-{Let's clarify some notation.
+An unbounded fluid queue is a two-dimensional stochastic process which we denote by \(\{\ddot{\bs X}(t)\} = \{(\ddot X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process of the fluid queue, and \(\{\ddot X(t)\}_{t\geq0}\) is known as the level process or buffer.
+The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\), and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\)\footnote{Let's clarify some notation.
 We use the notation \(\bs u = (u_h)_{h\in \mathcal H}\) to denote a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some index set.
@@
 Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\).
-The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).}
+The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).}.
+We assume that \(\bs T\) is {conservative} and time-homogeneous.
+Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\) which determine the rate at which \(\{X(t)\}\) moves.
+
+Partition the state space \(\calS\) into \(\calS_+ = \{i\in\calS\mid c_i>0\}\), \(\calS_- = \{i\in\calS\mid c_i<0\}\), \(\calS_0 = \{i\in\calS\mid c_i=0\}\), \(\mathcal S_{-1}=\{i\in\calS\mid c_i\leq0\},\, \mathcal S_{K+1}=\{i\in\calS\mid c_i\geq0\}\)\footnote{The notation \(\mathcal S_{-1}\) and \(\mathcal S_{K+1}\) will make sense later when we introduce the partition into cells for the approximation schemes into \(K\) cells, plus a lower and upper boundary which we represent with indices \(-1\) and \(K+1\) respectively.}.
+We assume, without loss of generality, that the generator \(\bs T\) is partitioned into sub-matrices
+\[\bs T = \left[\begin{array}{ccc}\bs T_{++} & \bs T_{+-} & \bs T_{+0} \\ \bs T_{-+} & \bs T_{--} & \bs T_{-0} \\ \bs T_{0+} & \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
+where \(\bs T_{mn} = [T_{ij}]_{i\in\mathcal S_m, j\in\mathcal S_n}\), \(m,n\in\{+,-,0\}\).
 
@@
 
-Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(\mathrm{sign}(c_i)=\mathrm{sign}(c_j)\) or \(\mathrm{sign}(c_j)=0\) then the process is absorbed in the boundary, otherwise it immediately leaves the boundary.
-At a lower boundary, if \(j\in\calS_0\cup\calS_-\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\), and the phase process continues to evolve according to the sub-generator
+Upon hitting the lower (upper) boundary we suppose that, with probability \(\widecheck p_{ij},\,i,j\in\mathcal S\) (\(\widehat p_{ij},\,i,j\in\mathcal S\)), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(\mathrm{sign}(c_i)=\mathrm{sign}(c_j)\) or \(\mathrm{sign}(c_j)=0\) then the process is absorbed in the boundary, otherwise it immediately leaves the boundary.
+At the lower boundary, if \(j\in\calS_0\cup\calS_-\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\), and the phase process continues to evolve according to the sub-generator
 \[\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
 until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_+\), at which time \(\{X(t)\}\) leaves the boundary.
-Similarly, at an upper boundary if \(j\in\calS_0\cup\calS_+\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\) and the phase process continues to evolve according to the sub-generator
+Similarly, at the upper boundary if \(j\in\calS_0\cup\calS_+\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\) and the phase process continues to evolve according to the sub-generator
 \[\left[\begin{array}{cc} \bs T_{++} & \bs T_{+0} \\ \bs T_{0+} & \bs T_{00}  \end{array}\right],\]
@@
 \(\left[\bs \Psi_X(\lambda)\right]_{ij}\) is the Laplace-Stieltjes transform of the time taken for the fluid queue to first return to level \(z\) and do so in phase \(j\in\calS_-\), given it started at level \(z\) in phase \(i\in\calS_+\).
-Define the in-out fluid level by $\beta_X(t) := \int_0^t \left| c_{\varphi(z)} \right|  \wrt z$, which is the total amount of fluid to flow in to or out of the buffer \(\{\ddot X(t)\}\) by time \(t\), and also define $\eta_X(y) := \inf \{t > 0: \beta_X(t) = y\}$ as the first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\).
+Define the in-out fluid level by $\beta_X(t) = \int_0^t \left| c_{\varphi(z)} \right|  \wrt z$, which is the total amount of fluid to flow in to or out of the buffer \(\{\ddot X(t)\}\) by time \(t\), and also define $\eta_X(y) = \inf \{t > 0: \beta_X(t) = y\}$ as the first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\).
 Further, let \(\bs H(\lambda,y)\) be the matrix with elements \(h_{ij}(\lambda,y)\), \(i,j\in\calS_+\cup\calS_-\), given by
@@
 \begin{align}
-	\bs H^{++}(\lambda,y)&= \left[h_{ij}^{++}(\lambda,y)\right]_{i\in \mathcal S_+,j\in\mathcal S_{+}\cup\calS_{+0}} := e^{\bs{Q}_{++}(\lambda)y}\vligne{\bs C_+^{-1} & \bs Q_{+0}(\lambda)},  \label{eqn: lst 1a}
-	\\\bs H^{--}(\lambda,y) &= \left[h_{ij}^{--}(\lambda,y)\right]_{i\in\mathcal S_-,j\in\mathcal S_{-}\cup\calS_{-0}}:= e^{\bs{Q}_{--}(\lambda)y}\vligne{\bs C_-^{-1} & \bs Q_{-0}(\lambda)},
-	\\\bs H^{+-}(\lambda,y)  &= \left[h_{ij}^{+-}(\lambda,y)\right]_{i\in\mathcal S_+,\,j\in\mathcal S_-}:= e^{\bs{Q}_{++}(\lambda)y}\bs{Q}_{+-}(\lambda), 
-	\\\bs H^{-+}(\lambda,y)&= \left[h_{ij}^{-+}(\lambda,y)\right]_{i\in\mathcal S_-,\, j\in\mathcal S_+} := e^{\bs{Q}_{--}(\lambda)y}\bs{Q}_{-+}(\lambda), \label{eqn: lst 4a}
+	\bs H^{++}(\lambda,y)&= \left[h_{ij}^{++}(\lambda,y)\right]_{i\in \mathcal S_+,j\in\mathcal S_{+}\cup\calS_{+0}} = e^{\bs{Q}_{++}(\lambda)y}\vligne{\bs C_+^{-1} & \bs Q_{+0}(\lambda)},  \label{eqn: lst 1a}
+	\\\bs H^{--}(\lambda,y) &= \left[h_{ij}^{--}(\lambda,y)\right]_{i\in\mathcal S_-,j\in\mathcal S_{-}\cup\calS_{-0}}= e^{\bs{Q}_{--}(\lambda)y}\vligne{\bs C_-^{-1} & \bs Q_{-0}(\lambda)},
+	\\\bs H^{+-}(\lambda,y)  &= \left[h_{ij}^{+-}(\lambda,y)\right]_{i\in\mathcal S_+,\,j\in\mathcal S_-}= e^{\bs{Q}_{++}(\lambda)y}\bs{Q}_{+-}(\lambda), 
+	\\\bs H^{-+}(\lambda,y)&= \left[h_{ij}^{-+}(\lambda,y)\right]_{i\in\mathcal S_-,\, j\in\mathcal S_+} = e^{\bs{Q}_{--}(\lambda)y}\bs{Q}_{-+}(\lambda), \label{eqn: lst 4a}
 \end{align}
@@
 The function \(h_{ij}^{++}(\lambda,y)\) (\(h_{ij}^{--}(\lambda,y)\)) is the Laplace transform (with respect to time) of the time taken for the fluid level to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), given the phase was initially \(i\in\mathcal S_+\) (\(i\in\mathcal S_-\)).
-The function \(h_{ij}^{+-}(\lambda,y)\) (\(h_{ij}^{-+}(\lambda,y)\)) is the Laplace transform (with respect to time) of the time taken for the fluid level, \(\{Y(t)\}\) to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), after which time the phase instantaneously changes to \(j\in\mathcal S_-\) (\(\mathcal S_+\)), given the phase was initially \(i\in\mathcal S_+\) (\(\mathcal S_-\)) \citep{bean2005}.
+The function \(h_{ij}^{+-}(\lambda,y)\) (\(h_{ij}^{-+}(\lambda,y)\)) is the Laplace transform (with respect to time) of the time taken for the fluid level to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_+\cup\calS_{+0}\) (\(\mathcal S_-\cup\calS_{-0}\)), after which time the phase instantaneously changes to \(j\in\mathcal S_-\) (\(\mathcal S_+\)), given the phase was initially \(i\in\mathcal S_+\) (\(\mathcal S_-\)) \citep{bean2005}.
 
@@
 
-A stochastic fluid-fluid queue \citep{bo2013} is a Markov process with three elements, $\{( \ddot{Y}(t),{X}(t), \varphi(t))\}_{t \geq 0},$ where $\{(X(t),\varphi(t))\}_{t\geq 0}$ is a classical fluid queue\footnote{We assume \(X(t)\) is doubly-bounded as this results in a finite-dimensional approximation, but this is not necessary} and $\ddot{Y}(t)$ is the second fluid, which varies at rate $r_{\varphi(t)}(\dot{X}(t))$:
-% 
-\begin{align*} 
-	\ddot{Y}(t) := \ddot Y(0) + \int_0^t r_{\varphi(s)}({X}(s)) \wrt s.
+A stochastic fluid-fluid queue \citep{bo2013} is a Markov process with three elements, $\{( \ddot{W}(t),{X}(t), \varphi(t))\}_{t \geq 0},$ where $\{(X(t),\varphi(t))\}_{t\geq 0}$ is a classical fluid queue\footnote{We assume \(X(t)\) is doubly-bounded as this results in a finite-dimensional approximation, but this is not necessary} and $\ddot{W}(t)$ is the second fluid, which varies at rate $r_{\varphi(t)}(\dot{X}(t))$:
+% 
+\begin{align*} 
+	\ddot{W}(t) = \ddot W(0) + \int_0^t r_{\varphi(s)}({X}(s)) \wrt s.
 \end{align*} 
@@
 Regulated boundaries may also be included for the second fluid level.
-In this thesis we consider the second fluid to have a regulated boundary at \(y=0\) and unbounded above.
-To distinguish between unbounded and bounded processes, we use the notation \(\ddot Y(t)\) to denote the unbounded process and \(\dot Y(t)\) to denote the second fluid level process with a regulated lower boundary at 0.
+To distinguish between unbounded and bounded processes, we use the notation \(\ddot W(t)\) to denote the unbounded process and \(\dot W(t)\) to denote the second fluid level process with a regulated lower boundary at 0.
 
@@
 
-In the following, we assume that $\dot Y(t) \in [0,\infty)$ and that there is a boundary at level $0$ for both the first and second fluid levels and the first fluid level is also bounded above at \(b\).
+In the following, we assume that $\dot W(t) \in [0,\infty)$ and that there is a boundary at level $0$ for both the first and second fluid levels and the first fluid level is also bounded above at \(b\).
 Thus, the rates of change of the fluid levels can be summarised as:
@@
 	\begin{align*} 
-		\frac{\wrt}{\wrt t}  X(t) & := \max\{0, c_i\} \quad \mbox{if }  X(t) = 0 \mbox{ and } \varphi(t) = i, \\
-		\frac{\wrt}{\wrt t}  X(t) & := \min\{0, c_i\} \quad \mbox{if }  X(t) = b \mbox{ and } \varphi(t) = i, \\
-          	\frac{\wrt}{\wrt t} \dot Y(t) & := \max\{0, r_i(x)\} \quad \mbox{if } \dot Y(t) = 0,\, X(t) = x \mbox{ and } \varphi(t) = i, 	
-	\end{align*} 
-	% 
-for $i \in \mathcal{S} = \{1,...,N\}$. Let $\bs R(x) := (r_i(x))_{i \in \mathcal{S}}$ be the diagonal fluid-rate matrix of functions for $\{\dot Y(t)\}$.
-
-For the remainder of this section, we summarise the findings of~\cite{bo2014} on the joint limiting distribution of $\{( \dot Y(t), X(t), \varphi(t))\}_{t \geq 0}$.
-The derivation of the limiting distribution relies on obtaining the operator \(\mathbb \Psi\) which gives the distribution of the process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to the level 0, given \(\dot Y(0)=0\).
+		\frac{\wrt}{\wrt t}  X(t) & = \max\{0, c_i\} \quad \mbox{if }  X(t) = 0 \mbox{ and } \varphi(t) = i, \\
+		\frac{\wrt}{\wrt t}  X(t) & = \min\{0, c_i\} \quad \mbox{if }  X(t) = b \mbox{ and } \varphi(t) = i, \\
+          	\frac{\wrt}{\wrt t} \dot W(t) & = \max\{0, r_i(x)\} \quad \mbox{if } \dot W(t) = 0,\, X(t) = x \mbox{ and } \varphi(t) = i, 	
+	\end{align*} 
+	% 
+for $i \in \mathcal{S} = \{1,...,N\}$. Let $\bs R(x) =\diag(r_i(x))_{i \in \mathcal{S}}$ be the diagonal fluid-rate matrix of functions for $\{\dot W(t)\}$.
+
+For the remainder of this section, we summarise the findings of~\cite{bo2014} on the joint limiting distribution of $\{( \dot W(t), X(t), \varphi(t))\}_{t \geq 0}$.
+The derivation of the limiting distribution relies on obtaining the operator \(\mathbb \Psi\) which gives the distribution of the process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot W(t)\}\) first returns to the level 0, given \(\dot W(0)=0\).
 
@@
 First, we need the infinitesimal generator of the driving process, \(\{(X(t),\varphi(t))\}\), which we denote by \(\mathbb B\).
-Then we need to partition \(\mathbb B\) on sets for which \(\{\dot Y(t)\}\) is increasing, decreasing or constant -- these are the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), below.
+Then we need to partition \(\mathbb B\) on sets for which \(\{\dot W(t)\}\) is increasing, decreasing or constant -- these are the sets \(\mathcal F_i^m,\, i\in\calS,\,m\in\{+,-,0\}\), below.
 We can then derive an operator-Riccati equation for the first return operator for the fluid-fluid queue, \(\mathbb \Psi\), in terms of the partitioned generator and the rates \(r_i(x)\).
-The solution to the operator-Riccati equation is an operator which can be used to obtain the distribution of the driving process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot Y(t)\}\) first returns to \(0\).
+The solution to the operator-Riccati equation is an operator which can be used to obtain the distribution of the driving process \(\{(X(t),\varphi(t))\}\) at the time when \(\{\dot W(t)\}\) first returns to \(0\).
 
@@
 
-For each Markovian state $i \in \mathcal{S}$, we partition the state space of \( X(t)\), \([0,b]\), according to the rates of change $r_i(\cdot)$ for the second fluid $\{\dot Y(t)\}$: $[0,b] := \mathcal{F}^{+}_i \cup \mathcal{F}^{-}_i \cup \mathcal{F}^{0}_i,$
+For each Markovian state $i \in \mathcal{S}$, we partition the state space of \( X(t)\), \([0,b]\), according to the rates of change $r_i(\cdot)$ for the second fluid $\{\dot W(t)\}$: $[0,b] = \mathcal{F}^{+}_i \cup \mathcal{F}^{-}_i \cup \mathcal{F}^{0}_i,$
 where 
@@
 %				\label{eqn:Fi1}
-			\mathcal{F}^{+}_i & := \{u \in [0,b] : r_i(u) > 0\},  \; \nonumber
+			\mathcal{F}^{+}_i & = \{u \in [0,b] : r_i(u) > 0\},  \; \nonumber
 %				\label{eqn:Fi2}
-			\\\mathcal{F}^{-}_i &:=  \{u \in [0,b]:  r_i(u) < 0\}, \; \nonumber
+			\\\mathcal{F}^{-}_i &=  \{u \in [0,b]:  r_i(u) < 0\}, \; \nonumber
 %				\label{eqn:Fi3}
-			\\\mathcal{F}^{0}_i &:= \{u \in [0,b]: r_i(u) = 0\}.\label{eqn:fil}
+			\\\mathcal{F}^{0}_i &= \{u \in [0,b]: r_i(u) = 0\}.\label{eqn:fil}
 		\end{align} 
@@
 
-We assume that the process $\{(\dot Y(t), X(t), \varphi(t))\}_{t\geq 0}$ is positive recurrent, in order to guarantee the existence of the joint limiting distribution. Define limiting operators 
+We assume that the process $\{(\dot W(t), X(t), \varphi(t))\}_{t\geq 0}$ is positive recurrent, in order to guarantee the existence of the joint limiting distribution. Define limiting operators 
 \begin{align} 
 		\label{eqn:jointpi} 
-		\bbpi_i(y)(\mathcal{E}) & := \lim_{t \rightarrow \infty} \frac{\partial}{\partial y} \mathbb{P}\left( \dot Y(t) \leq y,  X(t) \in \mathcal{E}, \varphi(t) = i\right), \quad y>0,\\
+		\bbpi_i(w)(\mathcal{E}) & =  \frac{\partial}{\partial y} \lim_{t \rightarrow \infty} \mathbb{P}\left( \dot W(t) \leq w,  X(t) \in \mathcal{E}, \varphi(t) = i\right), \quad w>0,\\
 % 
 		\label{eqn:jointmass}
-		{\mathbb p}_i(\mathcal{E}) & := \lim_{t \rightarrow \infty}  \mathbb{P}( \dot Y(t) = 0,  X(t) \in \mathcal{E}, \varphi(t) = i]),
+		{\mathbb p}_i(\mathcal{E}) & = \lim_{t \rightarrow \infty}  \mathbb{P}( \dot W(t) = 0,  X(t) \in \mathcal{E}, \varphi(t) = i]),
 \end{align} 
 where $\mathcal{E} \subset [0,b]$.
-The keen observer will notice that (\ref{eqn:jointpi}) assumes a certain differentiability condition on the limiting distribution.
-Indeed, it is not always the case that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) (trivially, consider \(\dot Y(t)=y_0\) for all \(t\)).
-Conditions on fluid-fluid queues to ensure that (\ref{eqn:jointpi}) will be differentiable with respect to \(y\) are not known, in here we just assume that it is.
-
-Now, let ${{\bbpi}}(y) = ( \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint limiting density operators and ${\mathbb{p}} = ( {\mathbb p}_i)_{i \in \mathcal{S}}$ be a vector containing the joint limiting mass operators.
-The determination of ${{\bbpi}}(y)$ involves two important matrices of operators, $ {\mathbb D}$ and $ {\mathbb \Psi}$ which we now introduce.
+Notice that (\ref{eqn:jointpi}) assumes a certain differentiability condition on the limiting distribution.
+Indeed, it is not always the case that (\ref{eqn:jointpi}) will be differentiable with respect to \(w\) (trivially, consider \(\dot W(t)=w_0\) for all \(t\)).
+To the author's knowledge, conditions on fluid-fluid queues to ensure that (\ref{eqn:jointpi}) will be differentiable with respect to \(w\) are not known, and here we just assume that it is.
+For further reading, further theoretical developments and some examples of fluid-fluid queues are analysed in \cite{boz2022}.
+
+Now, let ${{\bbpi}}(w) = ( \bbpi_i(w))_{i \in \mathcal{S}}$ be a vector containing the joint limiting density operators and ${\mathbb{p}} = ( {\mathbb p}_i)_{i \in \mathcal{S}}$ be a vector containing the joint limiting mass operators.
+The determination of ${{\bbpi}}(w)$ involves two important matrices of operators, $ {\mathbb D}$ and $ {\mathbb \Psi}$ which we now introduce.
 %The operator \( {\mathbb B}\) is the \textit{infinitesimal generator} of the process \(\{(X(t),\varphi(t))\}\).
 %Intuitively, for a set $\mathcal{E} \subset \mathcal{F}$ and a measure vector $\boldsymbol{\mu} = (\mu_i)_{i \in \mathcal{S}}$, $\bs{\mu}V(t)(\mathcal{E})$, where \(V(t)\) is a semigroup with infinitesimal generator \(B\), gives the conditional probability of $X(t) \in \mathcal{E}$,
-% The operator \( {\mathbb \Psi}\) is such that ${\boldsymbol{\mu}}^{+} {\mathbb \Psi} (\mathcal{E})$ gives the conditional probability of $\{\dot Y(t)\}$ returning to level zero and doing so when $X(t) \in \mathcal{E}$, given that the initial distribution of \((X(0),\varphi(0))\) in the set \(\bigcup_{i\in\calS}(\mathcal F_i^+,i)\), is ${\boldsymbol{\mu}}^{+}$.
-
-\subsection{The infinitesimal generator, \(\mathbb B\), of the driving process}
+% The operator \( {\mathbb \Psi}\) is such that ${\boldsymbol{\mu}}^{+} {\mathbb \Psi} (\mathcal{E})$ gives the conditional probability of $\{\dot W(t)\}$ returning to level zero and doing so when $X(t) \in \mathcal{E}$, given that the initial distribution of \((X(0),\varphi(0))\) in the set \(\bigcup_{i\in\calS}(\mathcal F_i^+,i)\), is ${\boldsymbol{\mu}}^{+}$.
+
+\subsection{The infinitesimal generator, \(\mathbb B\), of the driving process}\label{subsec: operatorsoffq}
 Given the discussion above, if we are to replicate the arguments of \cite{bean2005} to derive the Laplace-Stieltjes transform of the first return operator of a fluid-fluid queue, we first need an expression for the generator of the driving process.
@@
 Intuitively, the operator \( \overline{\mathbb V}(t)\) maps an initial measure \(\overline{\boldsymbol \mu}\) on \((X(0),\varphi(0))\) to the measure \(\mathbb P(X(t)\in \mathcal{E}, \varphi(t)=j)=:\overline\mu_j(t)(\mathcal{E})\).
-The matrix of operators \( \overline{\mathbb B}:=[\overline{\mathbb B}_{ij}]_{i,j\in\mathcal S}\) is the \textit{infinitesimal generator} of the semigroup \(\{ \overline{\mathbb V}(t)\}\) defined by 
+The matrix of operators \( \overline{\mathbb B}=[\overline{\mathbb B}_{ij}]_{i,j\in\mathcal S}\) is the \textit{infinitesimal generator} of the semigroup \(\{ \overline{\mathbb V}(t)\}\) defined by 
 \[ \overline{\mathbb B} =\left.\cfrac{\wrt}{\wrt t} \overline{\mathbb V}(t)\right|_{t=0},\]
@@
 To use the operators \(\{ \overline{\mathbb  V}(t)\}\) and \( \overline{\mathbb  B}\) to analyse the fluid-fluid queue, \cite{bo2014} explicitly track when \((X(t),\varphi(t))\in(\mathcal F_i^m,i)\) for \(i\in\mathcal S,\, m \in \{+,-,0\}\) by partitioning the operators \( \overline{\mathbb  V}(t)\) and \( \overline{\mathbb  B}\) into \( \overline{\mathbb  V}_{ij}^{m n}\) and \( \overline{\mathbb  B}_{ij}^{m n}\), for \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\), where
-\[\left.\overline\mu_i\right|_{\calF_i^m}  \overline{\mathbb  V}_{ij}^{m n}(t)(\mathcal{E}) := \int_{x\in[0,b]} \left.
+\[\left.\overline\mu_i\right|_{\calF_i^m}  \overline{\mathbb  V}_{ij}^{m n}(t)(\mathcal{E}) = \int_{x\in[0,b]} \left.
 \wrt \overline\mu_i\right|_{\calF_i^m}(x)\mathbb P (X(t) \in\mathcal{E}\cap \calF_j^n,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
@@
 
-Let us first partition the space \([0,b]\) into \(\calD_{-1,i}=\calD_{-1} = \{0\}\), \(i\in\calS_{-1}\), \(\calD_{{K+1},i}=\calD_{{K+1}}=\{b\}\), \(i\in\calS_{K+1}\), and non-trivial intervals \(\calD_{k,i}=[y_k,y_{k+1})\setminus \{0\},\) \(i\in\calS_+\cup\calS_0\) and \(\calD_{k,i}=(y_k,y_{k+1}]\setminus \{b\}\), \(i\in\calS_-\), with \(y_0=0,\,y_{K+1}=b,\, y_k<y_{k+1},\,k\in\mathcal K^\circ:=\{0,1,2,...,K\}\) and define \(\mathcal K = \{-1,K+1\}\cup\mathcal K^\circ\).
+Let us first partition the space \([0,b]\) into \(\calD_{-1,i}=\calD_{-1} = \{0\}\), \(i\in\calS_{-1}\), \(\calD_{{K+1},i}=\calD_{{K+1}}=\{b\}\), \(i\in\calS_{K+1}\), and non-trivial intervals \(\calD_{k,i}=[y_k,y_{k+1})\setminus \{0\},\) \(i\in\calS_+\cup\calS_0\) and \(\calD_{k,i}=(y_k,y_{k+1}]\setminus \{b\}\), \(i\in\calS_-\), with \(y_0=0,\,y_{K+1}=b,\, y_k<y_{k+1},\,k\in\mathcal K^\circ=\{0,1,2,...,K\}\) and define \(\mathcal K = \{-1,K+1\}\cup\mathcal K^\circ\).
 For $\bs{\mu} \in \mathcal{M}_{0,b}(\mathcal{S} \times [0,b])$ we write \(\bs{\mu} = (\mu_{k,i})_{i \in \mathcal{S},k\in\mathcal K},\) where $\mu_{k,i}(\cdot) = \mu_i(\cdot \cap \calD_{k,i}),\,k\in\mathcal K.$ We denote by \(v_{k,i}(x),\,x>0\), the density associated with the measure, \(\mu_{k,i},\,k\in \mathcal K^\circ\).
 For \( i,j\in\calS,\,k,\ell\in\mathcal K\) define the operators
-\[\mu_{k,i} \mathbb V_{ij}^{k \ell}(t)(\mathcal{E}) := \int_{x\in\calD_{k,i}} \wrt \mu_{k,i}(x)\mathbb P (X(t) \in\mathcal{E}\cap \calD_{\ell,i},\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
-and the matrices of operators \(\mathbb V^{k\ell}(t) := \vligne{\mathbb V_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\) and write 
+\[\mu_{k,i} \mathbb V_{ij}^{k \ell}(t)(\mathcal{E}) = \int_{x\in\calD_{k,i}} \wrt \mu_{k,i}(x)\mathbb P (X(t) \in\mathcal{E}\cap \calD_{\ell,i},\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
+and the matrices of operators \(\mathbb V^{k\ell}(t) = \vligne{\mathbb V_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\) and write 
 \[\mathbb V(t) = \left[
@@
 %\end{align*}
-where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\).\footnote{We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
+where the blocks \(\mathbb B^{k\ell}= \vligne{\mathbb B_{ij}^{k\ell}}_{i,j\in\mathcal S},\,k,\ell\in\mathcal K\).\footnote{We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}).
 The operators denoted with an overline play a minor role in the introductory sections of this thesis, but do not appear again.
@@
 
-Let \(\calS_k^+=\{i\in\calS\mid r_i(x)>0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^0=\{i\in\calS\mid r_i(x)=0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^-=\{i\in\calS\mid r_i(x)<0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^\bullet=\{i\in\calS\mid r_i(x)\neq0,\,\forall x \in\calD_{k,i}\}\) for \(k\in\mathcal K^\circ\) and \(\calS_{k}^+=\{i\in\calS_{k}\mid r_i(y_{k})>0\}\), \(\calS_{k}^-=\{i\in\calS_{k}\mid r_i(y_{k})<0\}\), \(\calS_{k}^0=\{i\in\calS_{k}\mid r_i(y_{k})=0\}\), \(\calS_{k}^\bullet=\{i\in\calS_{k}\mid r_i(y_{k})\neq 0\}\) for \(k\in\{-1,K+1\}\).
+Let \(\calS_k^+=\{i\in\calS\mid r_i(x)>0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^0=\{i\in\calS\mid r_i(x)=0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^-=\{i\in\calS\mid r_i(x)<0,\,\forall x \in\calD_{k,i}\}\), \(\calS_k^\bullet=\{i\in\calS\mid r_i(x)\neq0,\,\forall x \in\calD_{k,i}\}\) for \(k\in\mathcal K^\circ\).
 For later reference, we need the following constructions.
@@
 When \(v_k\) is differentiable we can write \(\mu_{k,i}\mathbb B_{ij}^{k\ell}(\mathcal{E})\) in kernel form as \[\displaystyle\int_{x\in\calD_{k,i},y\in\mathcal{E}}\wrt \mu_{k,i}(x)\mathbb B_{ij}^{k\ell}(x,dy).\] It is known that
-\begin{align}\label{eqn: lfj22}\mu_{k,i}\mathbb B_{ij}^{kk}(\wrt y):=\int_{x\in\calD_{k,i}}\wrt\mu_{k,i}(x)\mathbb B_{ij}^{kk}(x,\wrt y)=\begin{cases}
+\begin{align}\label{eqn: lfj22}\mu_{k,i}\mathbb B_{ij}^{kk}(\wrt y)=\int_{x\in\calD_{k,i}}\wrt\mu_{k,i}(x)\mathbb B_{ij}^{kk}(x,\wrt y)=\begin{cases}
 v_{k,i}(y)T_{ij}\wrt y, & i\neq j,\\
@@
 	\begin{align*}
-	\mu_{0,i}\mathbb B_{ij}^{0,{-1}} &= -c_ip^{{-1}}_{ij}v_{0,i}(0^+),\mbox{ if }c_i< 0,\,c_j\leq 0.
-	\end{align*}
-	In \(\calD_{0,i}\), if \(c_i<0\) and \(c_j>0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{-1}\) to phase \(j\), hence
-	\begin{align*}
-	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_{0,i})&=\int_{x\in\calD_{0,i}}v_{0,i}(x)T_{ij}\wrt x - c_iv_{0,i}(0^+)p^{-1}_{ij}.
+	\mu_{0,i}\mathbb B_{ij}^{0,{-1}} &= -c_i\widecheck p_{ij}v_{0,i}(0^+),\mbox{ if }c_i< 0,\,c_j\leq 0.
+	\end{align*}
+	In \(\calD_{0,i}\), if \(c_i<0\) and \(c_j>0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(\widecheck{p}_{ij}\) to phase \(j\), hence
+	\begin{align*}
+	\mu_{0,i}\mathbb B_{ij}^{0,0}(\calD_{0,i})&=\int_{x\in\calD_{0,i}}v_{0,i}(x)T_{ij}\wrt x - c_iv_{0,i}(0^+)\widecheck p_{ij}.
 	\end{align*}
@@
 	\begin{align*}
-	\mu_{K,i}\mathbb B_{ij}^{K,{K+1}} &= c_ip^{{K+1}}_{ij}v_{K,i}(0^+),\mbox{ if }c_i> 0,\,c_j\geq 0.
-	\end{align*}
-	In \(\calD_{K,i}\), if \(c_i>0\) and \(c_j<0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(p_{ij}^{K+1}\) to phase \(j\),
-	\begin{align*}
-	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_{K,i})&=\int_{x\in\calD_{K,i}}v_{K,i}(x)T_{ij}\wrt x + c_iv_{K,i}(b^-)p^{K+1}_{ij},\,c_i>0,\,c_j<0.
+	\mu_{K,i}\mathbb B_{ij}^{K,{K+1}} &= c_i\widehat p_{ij}v_{K,i}(0^+),\mbox{ if }c_i> 0,\,c_j\geq 0.
+	\end{align*}
+	In \(\calD_{K,i}\), if \(c_i>0\) and \(c_j<0\) the phase changes from \(i\) to \(j\) either from a stochastic jump at rate \(T_{ij}\), or from \(\{X(t)\}\) hitting the boundary in phase \(i\) and transitioning with probability \(\widehat p_{ij}\) to phase \(j\),
+	\begin{align*}
+	\mu_{K,i}\mathbb B_{ij}^{K,K}(\calD_{K,i})&=\int_{x\in\calD_{K,i}}v_{K,i}(x)T_{ij}\wrt x + c_iv_{K,i}(b^-)\widehat p_{ij},\,c_i>0,\,c_j<0.
 	\end{align*}
@@
 
-\subsection{The infinitesimal generator, \(\mathbb D\), of an in-out process}
-Let $\beta(t) := \int_0^t \left| r_{\varphi(z)}(X(z)) \right|  \wrt z$ be the total unregulated amount of fluid that has flowed into or out of the second buffer during $[0,t]$, and let $\eta(y) := \inf \{t > 0: \beta(t) = y\}$ be the first time this accumulated in-out amount hits level $y$.
-Note that at the stopping time \(\eta(y)\) it must be that \((X({\eta(y)}),\varphi({\eta(y)}))\in(\calF_i^m,i)\) for some \(i\in\calS\) and \(m\in\{+,-\}\), i.e.~\(m\neq0\).
-We define the operators $\mathbb{U}_{ij}^{k\ell}(y,s): \mathcal{M}_{0,b}(\mathcal{D}_{k,i}\cap\calF_i^m) \mapsto \mathcal{M}_{0,b} (\calD_{\ell,j}\cap\mathcal{F}_j^n)$, for $k\in\mathcal K_i^+\cup\mathcal K_i^-$,  $\ell\in\mathcal K_j^+\cup\mathcal K_j^-$, and $i \in \mathcal{S}_k^\bullet,\,j \in \mathcal{S}_k^\bullet$, by
-% 
-	\begin{align*} 
-		&\mu_{k,i}\mathbb{U}_{ij}^{k\ell} (y,s) (\mathcal{E}) 
-		\\&:= \int_{x \in \calD_{k,i}} \wrt  \mu_{k,i}(x) \mathbb{E}\left[e^{-s\eta(y)}{1}\left\{X({\eta(y)}) \in \mathcal{E}\cap\calD_{\ell,j},\;\varphi({\eta(y)}) = j\right\} \mid \varphi(0) = i, X(0) = x\right].
+\subsection{The infinitesimal generator, \(\mathbb D\), of an in-out process}\label{subsec: afjakje}
+Let $\beta(t) = \int_0^t \left| r_{\varphi(z)}(X(z)) \right|  \wrt z$ be the total unregulated amount of fluid that has flowed into or out of the second buffer during $[0,t]$, and let $\eta(w) = \inf \{t > 0: \beta(t) = w\}$ be the first time this accumulated in-out amount hits level $w$.
+Note that at the stopping time \(\eta(w)\) it must be that \((X({\eta(w)}),\varphi({\eta(w)}))\in(\calF_i^m,i)\) for some \(i\in\calS\) and \(m\in\{+,-\}\), i.e.~\(m\neq0\).
+We define the operators $\mathbb{U}_{ij}^{k\ell}(w,s): \mathcal{M}_{0,b}(\mathcal{D}_{k,i}\cap\calF_i^m) \mapsto \mathcal{M}_{0,b} (\calD_{\ell,j}\cap\mathcal{F}_j^n)$, for $k\in\mathcal K^+\cup\mathcal K^-$,  $\ell\in\mathcal K^+\cup\mathcal K^-$, and $i \in \mathcal{S}_k^\bullet,\,j \in \mathcal{S}_\ell^\bullet$, by
+% 
+	\begin{align*} 
+		&\mu_{k,i}\mathbb{U}_{ij}^{k\ell} (w,s) (\mathcal{E}) 
+		\\&= \int_{x \in \calD_{k,i}} \wrt  \mu_{k,i}(x) \mathbb{E}\left[e^{-s\eta(w)}{1}\left\{X({\eta(w)}) \in \mathcal{E}\cap\calD_{\ell,j},\;\varphi({\eta(w)}) = j\right\} \mid \varphi(0) = i, X(0) = x\right].
 	\end{align*} 
@@
 \begin{align*}
-%\mathbb{U}^{k \ell}(y,s)&:=[\mathbb U_{ij}^{k\ell}(y,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}, \, k,\ell\in \mathcal K^+\cup\mathcal K^-,
-\mathbb U(y,s) &:= \left[[\mathbb U_{ij}^{k\ell}(y,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}\right]_{k,\ell\in\mathcal K^+\cup\mathcal K^-}.\end{align*}
+%\mathbb{U}^{k \ell}(w,s)&=[\mathbb U_{ij}^{k\ell}(w,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}, \, k,\ell\in \mathcal K^+\cup\mathcal K^-,
+\mathbb U(w,s) &= \left[[\mathbb U_{ij}^{k\ell}(w,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}\right]_{k,\ell\in\mathcal K^+\cup\mathcal K^-}.\end{align*}
 %\[\mathbb U(t) = \left[
@@
 %	\begin{align*} 
-%		\mathbb{U}(y,s) = \left[\begin{array}{llllll} 
+%		\mathbb{U}(w,s) = \left[\begin{array}{llllll} 
 %		\ddots & & & &~\reflectbox{$\ddots$}\\
-%            		& \mathbb{U}^{k-1,k-1}(y,s) & \mathbb{U}^{k-1,k}(y,s) & \mathbb{U}^{k-1,k+1}(y,s) \\
-%			& \mathbb{U}^{k,k-1}(y,s) & \mathbb{U}^{k,k}(y,s) & \mathbb{U}^{k,k+1}(y,s) \\
-%			& \mathbb{U}^{k+1,k-1}(y,s) & \mathbb{U}^{k+1,k}(y,s) & \mathbb{U}^{k+1,k+1}(y,s) \\	
+%            		& \mathbb{U}^{k-1,k-1}(w,s) & \mathbb{U}^{k-1,k}(w,s) & \mathbb{U}^{k-1,k+1}(w,s) \\
+%			& \mathbb{U}^{k,k-1}(w,s) & \mathbb{U}^{k,k}(w,s) & \mathbb{U}^{k,k+1}(w,s) \\
+%			& \mathbb{U}^{k+1,k-1}(w,s) & \mathbb{U}^{k+1,k}(w,s) & \mathbb{U}^{k+1,k+1}(w,s) \\	
 %			\reflectbox{$\ddots$}& & & & \ddots		
@@
 	% 
-%	where $\mathbb{U}^{k \ell}:=\mathbb U_{ij}^{k\ell}$ is an $|\mathcal{S}| \times |\mathcal{S}|$ matrix of operators for $y > 0$, $s \in \mathbb{C}$, and Re$(s) \geq 0$. 
-The matrix of operators \(\mathbb D(s)\) is the infinitesimal generator of the semigroup \(\{\mathbb U(y,s)\}_{y\geq 0}\) defined by 
-\[\mathbb D(s) = \cfrac{\wrt}{\wrt y}\mathbb U(y,s)|_{y=0},\]
+%	where $\mathbb{U}^{k \ell}=\mathbb U_{ij}^{k\ell}$ is an $|\mathcal{S}| \times |\mathcal{S}|$ matrix of operators for $w > 0$, $s \in \mathbb{C}$, and Re$(s) \geq 0$. 
+The matrix of operators \(\mathbb D(s)\) is the infinitesimal generator of the semigroup \(\{\mathbb U(w,s)\}_{w\geq 0}\) defined by 
+\[\mathbb D(s) = \cfrac{\wrt}{\wrt w}\mathbb U(w,s)|_{w=0},\]
 whenever this limit exists.
@@
 \begin{lem}\label{lemma: D(s)}
-	For $y \geq 0$, $s \in \mathbb{C}$ with \textit{Re}$(s) \geq 0$, $i,j \in \mathcal{S}$, $k\in \mathcal K_i^+\cup\mathcal K_i^-$, $\ell\in \mathcal K_j^+\cup\mathcal K_j^-$,
+	For $w \geq 0$, $s \in \mathbb{C}$ with \textit{Re}$(s) \geq 0$, $i,j \in \mathcal{S}$, $k\in \mathcal K_i^+\cup\mathcal K_i^-$, $\ell\in \mathcal K_j^+\cup\mathcal K_j^-$,
 	% 
@@
 	% 
-	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} := \mathrm{diag}(\mathbb{R}_i^{k},\,i \in \mathcal{S})$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
-	\begin{align*} 
-		{\mu}_{k,i}\mathbb{R}_{k,i}(\mathcal{E}) := \int_{x \in \mathcal{E} \cap \mathcal{D}_{k,i}} \frac{1}{|r_i(x)|}\wrt  \mu_{k,i}(x),\quad k\in \mathcal K_i^+\cup\mathcal K_i^-.
+	where $\mathbb I$ is the identity operator, and $\mathbb{R}^{k} = \mathrm{diag}(\mathbb{R}_i^{k},\,i \in \mathcal{S})$ is a diagonal matrix of operators $\mathbb{R}_i^{k}$ given by 
+	\begin{align*} 
+		{\mu}_{k,i}\mathbb{R}_{i}^k(\mathcal{E}) = \int_{x \in \mathcal{E} \cap \mathcal{D}_{k,i}} \frac{1}{|r_i(x)|}\wrt  \mu_{k,i}(x),\quad k\in \mathcal K_i^+\cup\mathcal K_i^-.
 	\end{align*} 
@@
 	\begin{align*}
-            	%\mathbb D^{k \ell} &:= \left[\mathbb D_{ij}^{k \ell}\right]_{i,j\in\mathcal S},\quad k,\ell\in\{{-1},1,2,...\},
-		\mathbb D^{m n} &:= \left[\left[\mathbb D_{ij}^{k \ell}\right]_{i\in\calS_k^m,j\in\calS_k^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n}.
-		%\\ \mathbb D_{ij}^{m n} &:= \left[\mathbb D_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\quad i,j\in\calS,\,m,n\in\{+,-,0\}.
+            	%\mathbb D^{k \ell} &= \left[\mathbb D_{ij}^{k \ell}\right]_{i,j\in\mathcal S},\quad k,\ell\in\{{-1},1,2,...\},
+		\mathbb D^{m n} &= \left[\left[\mathbb D_{ij}^{k \ell}\right]_{i\in\calS_k^m,j\in\calS_k^n}\right]_{k\in\mathcal K^m,\ell\in\mathcal K^n}.
+		%\\ \mathbb D_{ij}^{m n} &= \left[\mathbb D_{ij}^{k \ell}\right]_{k\in\mathcal K_i^m,\ell\in\mathcal K_j^n},\quad i,j\in\calS,\,m,n\in\{+,-,0\}.
         \end{align*}
@@
 \subsection{The first-return operator, $\mathbb\Psi(s)$}\label{sec: intro Psi}
-We denote by $\mathbb \Psi(s)$ the matrix of operators with the same dimensions as \(\mathbb D^{+-}\), recording the Laplace-Stieltjes transforms of the time for $\{\dot Y(t)\}$ to return, for the first time, to the initial level of zero as introduced in \cite{bo2014} but constructed with respect to the finer partition \(\{\calD_{k,i}\}\).
-Define the stopping time $\zeta_Y(E):= \inf \{t > 0: \dot Y(t) \in E\}$ to be the first time $\{\dot Y(t)\}$ hits the set $E$, then each component $\mathbb \Psi_{ij}^{k\ell}(s): \mathcal{M}_{0,b}(\mathcal D_{k,i}) \mapsto \mathcal{M}_{0,b}(\mathcal D_{\ell,j}), i,j \in \mathcal{S},\,k\in\mathcal K_i^+$ and $\ell \in \mathcal K_j^-$, is given by
+We denote by $\mathbb \Psi(s)$ the matrix of operators with the same dimensions as \(\mathbb D^{+-}\), recording the Laplace-Stieltjes transforms of the time for $\{\dot W(t)\}$ to return, for the first time, to the initial level of zero as introduced in \cite{bo2014} but constructed with respect to the finer partition \(\{\calD_{k,i}\}\).
+Define the stopping time $\zeta_W(E)= \inf \{t > 0: \dot W(t) \in E\}$ to be the first time $\{\dot W(t)\}$ hits the set $E$, then each component $\mathbb \Psi_{ij}^{k\ell}(s): \mathcal{M}_{0,b}(\mathcal D_{k,i}) \mapsto \mathcal{M}_{0,b}(\mathcal D_{\ell,j}), i,j \in \mathcal{S},\,k\in\mathcal K_i^+$ and $\ell \in \mathcal K_j^-$, is given by
 % 
@@
 	&\mu_{k,i}\mathbb\Psi_{ij}^{k\ell}(s) (\mathcal{E}) 
-	\\& := \int_{x \in \calD_{k,i}} \wrt \mu_{k,i}(x)
-	% 
-	 \mathbb{E}\big[e^{-s\zeta_Y(\{0\})}1\left(X({\zeta_Y(\{0\})}) \in \mathcal{E}\cap\calD_{\ell,j},\; \varphi({\zeta_Y(\{0\})}) = j \right) \mid X(0) = x, 
-	 \\&\qquad{} \dot Y(0) = 0, \varphi(0) = i\big].
+	\\& = \int_{x \in \calD_{k,i}} \wrt \mu_{k,i}(x)
+	% 
+	 \mathbb{E}\big[e^{-s\zeta_W(\{0\})}1\left(X({\zeta_W(\{0\})}) \in \mathcal{E}\cap\calD_{\ell,j},\; \varphi({\zeta_W(\{0\})}) = j \right) \mid X(0) = x, 
+	 \\&\qquad{} \dot W(0) = 0, \varphi(0) = i\big].
 	 % 
@@
 
-Let $\mathbb\Psi := \mathbb\Psi(0)$.
-We define $\zeta_Y^n(\{0\}) := \inf\{t \geq \zeta_Y^{n - 1}(\{0\}): \dot Y(t) = 0\}$, for $n \geq  2$, to be the sequence of hitting times to level $0$ of $\dot Y(t)$, with $\zeta_Y^1(\{0\}): = \zeta_Y(\{0\})$.
-Consider the discrete-time Markov process $\{X({\zeta_Y^n(\{0\})}), \varphi(\zeta_Y^n(\{0\}))\}_{n \geq 1}$, and for $i \in \mathcal{S},\,k\in\mathcal K_i^-$ define the measures $\bbxi_{k,i}$ given by
-	%
-	\begin{align*}
-		\bbxi_{k,i}(\mathcal{E}) := \lim_{n \rightarrow \infty} \mathbb{P}\left(X(\zeta_Y^n(\{0\})) \in \mathcal{E}\cap\calD_{k,i}, \varphi(\zeta_Y^n(\{0\})) = i\right).
-	\end{align*} 
-	By \cite{bo2014}, the vector of measures $\boldsymbol{\bbxi} := (\bbxi_{k,i})_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}$ satisfies the set of equations 
+Let $\mathbb\Psi = \mathbb\Psi(0)$.
+We define $\zeta_W^n(\{0\}) = \inf\{t \geq \zeta_W^{n - 1}(\{0\}): \dot W(t) = 0\}$, for $n \geq  2$, to be the sequence of hitting times to level $0$ of $\dot W(t)$, with $\zeta_W^1(\{0\}): = \zeta_W(\{0\})$.
+Consider the discrete-time Markov process $\{X({\zeta_W^n(\{0\})}), \varphi(\zeta_W^n(\{0\}))\}_{n \geq 1}$, and for $i \in \mathcal{S},\,k\in\mathcal K_i^-$ define the measures $\bbxi_{k,i}$ given by
+	%
+	\begin{align*}
+		\bbxi_{k,i}(\mathcal{E}) = \lim_{n \rightarrow \infty} \mathbb{P}\left(X(\zeta_W^n(\{0\})) \in \mathcal{E}\cap\calD_{k,i}, \varphi(\zeta_W^n(\{0\})) = i\right).
+	\end{align*} 
+	By \cite{bo2014}, the vector of measures $\boldsymbol{\bbxi} = (\bbxi_{k,i})_{i \in \mathcal{S}_k^-,k\in\mathcal K^-}$ satisfies the set of equations 
  % 
@@
 
-We reproduce Theorem 2 of \cite{bo2014} below, which gives the joint limiting distribution of $\{(\dot Y(t),X(t),  \varphi(t))\}$.
-Recall that the joint limiting density operator ${\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{(\dot Y(t), X(t), \varphi(t))\}$ and the joint limiting mass operator ${\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$ are defined by~\eqref{eqn:jointpi} and \eqref{eqn:jointmass}, respectively.
-%Following out notational convention, define an equivalent joint limiting density operator $\boldsymbol{\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{X(t), Y(t), \varphi(t)\}$ and an equivalent joint limiting mass operator $\boldsymbol{\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$, but partitioned accprsing .
+We reproduce Theorem 2 of \cite{bo2014} below, which gives the joint limiting distribution of $\{(\dot W(t),X(t),  \varphi(t))\}$.
+Recall that the joint limiting density operator ${\bbpi}(w) = (\bbpi_i(w))_{i \in \mathcal{S}}$ for $\{(\dot W(t), X(t), \varphi(t))\}$ and the joint limiting mass operator ${\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$ are defined by~\eqref{eqn:jointpi} and \eqref{eqn:jointmass}, respectively.
+%Following out notational convention, define an equivalent joint limiting density operator $\boldsymbol{\bbpi}(y) = (\bbpi_i(y))_{i \in \mathcal{S}}$ for $\{X(t), W(t), \varphi(t)\}$ and an equivalent joint limiting mass operator $\boldsymbol{\mathbb p} = (\mathbb p_i)_{i \in \mathcal{S}}$, but partitioned accprsing .
 We can partition \(\bbpi\) so that 
@@
 	\begin{align*} 
-		\bbpi(y) &= \vligne{\bbpi^{+}(y) & \bbpi^{-}(y) & \bbpi^{0}(y)} 
-		\\&= \vligne{\left(\bbpi_{k,i}(y)\right)_{i \in \mathcal{S}_k^+,k\in\mathcal K^+} & \left(\bbpi_{k,i}(y)\right)_{i \in \mathcal{S}_k^-,k\in\mathcal K^-} & \left(\bbpi_{k,i}(y)\right)_{i \in \mathcal{S}_k^0,k\in\mathcal K^0}},
+		\bbpi(w) &= \vligne{\bbpi^{+}(w) & \bbpi^{-}(w) & \bbpi^{0}(w)} 
+		\\&= \vligne{\left(\bbpi_{k,i}(w)\right)_{i \in \mathcal{S}_k^+,k\in\mathcal K^+} & \left(\bbpi_{k,i}(w)\right)_{i \in \mathcal{S}_k^-,k\in\mathcal K^-} & \left(\bbpi_{k,i}(w)\right)_{i \in \mathcal{S}_k^0,k\in\mathcal K^0}},
 	\end{align*} 
@@
 	\begin{align*} 
-		\bbpi_{k,i}(y)(\mathcal{E}) = \bbpi_i(y)(\mathcal{E}\cap \mathcal D_{k,i}).
+		\bbpi_{k,i}(w)(\mathcal{E}) = \bbpi_i(w)(\mathcal{E}\cap \mathcal D_{k,i}).
 	\end{align*} 
@@
 	\label{theo:density} 
-The operator ${\bbpi}^{m}(y)$, for $m \in \{+,-,0\}$ and $y > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the set of equations:
-% 	 
-	\begin{align} 
-	&\label{11} \; \bbpi^{0}(y) = \vligne{\bbpi^{+}(y) & \bbpi^{-}(y)}\left[\begin{array}{l} \mathbb{B}^{+0} \\ \mathbb{B}^{-0} \end{array} \right]\left(-\mathbb{B}^{00}\right)^{-1}, \\
-	% 
-	&  \vligne{\bbpi^{+}(y) & \bbpi^{-}(y)} = \vligne{\mathbb{p}^{-} & \mathbb{p}^{0}}\left[\begin{array}{l} \mathbb{B}^{-+} \\ \mathbb{B}^{0+} \end{array} \right]\vligne{e^{\mathbb{K}y} & e^{\mathbb{K}y}\mathbb\Psi}\left[\begin{array}{cc} \mathbb{R}^{+} & 0 \\ 0 & \mathbb{R}^{-}\end{array}\right], \\
+The operator ${\bbpi}^{m}(w)$, for $m \in \{+,-,0\}$ and $w > 0$, and the probability mass $\mathbb{p}^{m}$, for $m \in \{-,0\}$, satisfy the set of equations:
+% 	 
+	\begin{align} 
+	&\label{11} \; \bbpi^{0}(w) = \vligne{\bbpi^{+}(w) & \bbpi^{-}(w)}\left[\begin{array}{l} \mathbb{B}^{+0} \\ \mathbb{B}^{-0} \end{array} \right]\left(-\mathbb{B}^{00}\right)^{-1}, \\
+	% 
+	&  \vligne{\bbpi^{+}(w) & \bbpi^{-}(w)} = \vligne{\mathbb{p}^{-} & \mathbb{p}^{0}}\left[\begin{array}{l} \mathbb{B}^{-+} \\ \mathbb{B}^{0+} \end{array} \right]\vligne{e^{\mathbb{K}w} & e^{\mathbb{K}w}\mathbb\Psi}\left[\begin{array}{cc} \mathbb{R}^{+} & 0 \\ 0 & \mathbb{R}^{-}\end{array}\right], \\
 	% 
@@
 	%  
-	& \sum_{m \in \{+,-,0\}}\sum_{i \in \mathcal{S}} \int_{y = 0}^{\infty} \bbpi_i^{m}(y)(\mathcal{F}^{m}_i)\wrt y + \sum_{m \in \{-,0\}} \sum_{i \in \mathcal{S}}\mathbb p^{m}_i(\mathcal{F}^{m}_i) = 1, \label{14}
-	\end{align}
-	% 
-	where $\mathbb{K} := \mathbb{D}^{++}(0) + \mathbb\Psi\mathbb{D}^{(-+)}(0)$ and $z$ is a normalising constant. 
+	& \sum_{m \in \{+,-,0\}}\sum_{i \in \mathcal{S}} \int_{w = 0}^{\infty} \bbpi_i^{m}(w)(\mathcal{F}^{m}_i)\wrt w + \sum_{m \in \{-,0\}} \sum_{i \in \mathcal{S}}\mathbb p^{m}_i(\mathcal{F}^{m}_i) = 1, \label{14}
+	\end{align}
+	% 
+	where $\mathbb{K} = \mathbb{D}^{++}(0) + \mathbb\Psi\mathbb{D}^{(-+)}(0)$ and $z$ is a normalising constant. 
 \end{theo}
@@
 
-\section{Discontinuous Galerkin}
+\section{Discontinuous Galerkin}\label{sec: dgdgdgdg}
 The discontinuous Galerkin (DG) method is a general methodology for numerically approximating solutions to differential equations.
@@
 Typically, we take \(\phi_k^r(x)\) to be a basis of polynomial functions.
-The functions \(\{\phi^r_k\}_{r=1}^{p_k}\) span some function space, \(W_k\) say, and further form a basis for this space.
+The functions \(\{\phi^r_k\}_{r=1}^{p_k}\) span some function space, \(U_k\) say, and further form a basis for this space.
 It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{p_k}\) as a basis of \emph{Lagrange interpolating polynomials} defined by the \emph{Gauss-Lobatto quadrature points} \citep{nodalDGBook} (see also Section~\ref{eqn: la}).
@@
 The approximation \(u_{k}\) is referred to as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k=0}^{K-1} u_{k}\) on the whole domain \(\Omega\).
-The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\{0,...,K-1\}\).
+The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} U_k\), the direct sum of the spaces \(U_k\), \(k\in\{0,...,K-1\}\).
 
@@
 	\\&\qquad\qquad{}- c(t)[\bs a_k(t)\bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}) - \bs a_{k-1}(t)\bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k})].\nonumber
-	\\&=\cfrac{\wrt \bs a_k(t)}{\wrt t}\bs M_k + c(t) \bs a_k(t)\bs G_k - c(t)[\bs a_k(t)\bs F_{k,k} - \bs a_{k-1}(t)\bs F_{k-1,k}]\nonumber
+	\\&=\cfrac{\wrt \bs a_k(t)}{\wrt t}\bs M_k + c(t) \bs a_k(t)\bs G_k - c(t)[\bs a_k(t)\bs F^{k,k} - \bs a_{k-1}(t)\bs F^{k-1,k}]\nonumber
 	\\&= 0,\label{eqn: hadndj}
@@
 where 
-\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x,\]
-\[\bs F_{k,k} := \bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}),\quad \bs F_{k-1,k} := \bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k}),\]
+\[\bs M_k = \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k = \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x,\]
+\[\bs F^{k,k} = \bs\phi_k(y_{k+1}^-)\tr{}\bs \phi_k(y_{k+1}),\quad \bs F^{k-1,k} = \bs\phi_{k-1}(y_{k}^-)\tr{}\bs \phi_k(y_{k}),\]
 and are known as the \emph{mass}, \emph{stiffness} and \emph{flux} matrices.
@@
 \begin{align*}
-	v_k^L &:= \overline{u}_{k} - m(\overline{u}_{k} - {u}_{k}^L,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
-	\\v_k^R &:= \overline{u}_{k} + m(\overline{u}_{k} - {u}_{k}^R,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+	v_k^L &= \overline{u}_{k} - m(\overline{u}_{k} - {u}_{k}^L,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
+	\\v_k^R &= \overline{u}_{k} + m(\overline{u}_{k} - {u}_{k}^R,\overline{u}_{k}-\overline{u}_{k-1},\overline{u}_{k+1}-\overline{u}_{k}),
 \end{align*}
@@
 \subsection{Slope limiters and time integration}\label{subsec: slope lim and int}
-To use slope limiters within the SSPRK scheme we first project the initial condition on to \(W\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
+To use slope limiters within the SSPRK scheme we first project the initial condition on to \(U\), the set of polynomial basis functions, and apply the slope limiter to the projection to determine the initial coefficients \(\bs a(0)\).
 To find the coefficients for the approximate solution to the transient distribution at time \(t_0\) we take the (limited) initial condition \(\bs a(0)\) and evolve it forward in time until \(t_0\) via numerical integration of the differential equation~(\ref{eqn: DG ODE w BCs 3}).
@@
 \subsection{Briefly, on accuracy}
-As a rough guide on accuracy we now paraphrase some known results on the accuracy of the DG method from \cite[Sections~5.5,~5.8]{nodalDGBook} for a similar kind of problem to the one under study here.
-For certain scalar conservation-law problems, under certain regularity conditions on the continuity and boundedness of the solution and its derivatives, on the grid used, on the flux function, and on the numerical flux, then one can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\) is used and a second-order SSPRK method is used to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
-Applying this result to the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, we might expect accuracy of order \((\Delta/p)^2\) under the same regularity conditions.
+The accuracy of the DG scheme for smooth problems is well-known. For example, in \cite[Section~5.5]{nodalDGBook} there is the following result. 
+\begin{thm}
+	Consider the one-dimensional scalar conservation law 
+	\begin{align*}
+		\cfrac{\partial u}{\partial t}+\cfrac{\partial f}{\partial t}=g,
+	\end{align*}
+	with \(x \in\Omega \).
+
+	Assume that the flux, \(f\in \mathrm{C}^3\), and the exact solution \(u\) is sufficiently smooth with bounded derivatives.
+Let \(u_h\) be a piecewise polynomial semidiscrete solution of the discontinuous Galerkin approximation to the one-dimensional scalar conservation law; then
+	\[|| u(t)-u_h(t)||_{\Omega,h}\leq C(t)h^{N+1},\]
+	provided a regular grid of \(h=\max h^k\) [and an upwind flux] is used.
+The constant \(C\) depends on \(u,\, N\) and time \(t\), but not on \(h\).
+[Here, \(h\) is the maximum width of the cells used, \(N\) is the order of the polynomial basis and \(||\cdot||_{\Omega,h}\) denotes a global `broken' norm \(||u||^2_{\Omega,h}=\sum_{k=1}^K\int_{\calD_k} u^2 \wrt x\) where \(\calD_k\) are the non-overlapping elements, or cells.]
+\end{thm}
+
+% A similar result is established for one-dimensional hyperbolic linear systems in \cite[Section~4.2]{nodalDGBook}. Consider 
+% \[\cfrac{\partial \bs u}{\partial t} + \bs A \cfrac{\partial \bs u}{\partial x}=0,\]
+% where \(\bs A\) is diagonalizable and we assume that the appropriate boundary conditions are available (see \cite[Section~4.2]{nodalDGBook}).
+
+Thus, in the numerical experiments in Chapter~\ref{sec: numerics}, for smooth problems, we can expect accuracy proportional to \(\Delta^{p}\) for the DG approximation of the solution, provided the DG scheme uses a polynomial basis with \(p\) functions on each cell, a regular grid of \(\Delta = \max \Delta_k\), and a second-order SSPRK method to integrate over time \citep[Sections~5.5,~5.8, and references therein]{nodalDGBook}.
+However, for the atypical applications of the DG approximation in Chapter~\ref{sec: numerics}, such as determining the first-return operator of the fluid-fluid queue, the author does know of any rigorous bounds on the accuracy of the approximation.
+Although, it seems reasonable to expect to see similar behaviour in the approximation errors in these cases, provided the solution is sufficiently smooth.
+
+Recall that for the DG-lin-lim scheme in Section~\ref{sec: anoth pos pres}, instead of using \(p\) basis functions on each cell, we instead use linear functions on smaller cells with width \(\Delta/(p/2)\).
+Thus, for the DG-lin-lim scheme we can expect accuracy of order \((\Delta/(p/2))^2\) (the exponent is 2 as the scheme uses linear functions for the approximation on each cell, and the base is \(\Delta/(p/2)\) as that is the reduced grid size) \citep[Sections~5.5,~5.8]{nodalDGBook}.
 
@@
 QBD-RAP processes are built from matrix-exponentially distributed inter-event times.
-We now introduce the class of matrix exponential distributions and recount some important properties before introducing the QBD-RAP.
+We now introduce the class of matrix exponential distributions (see \cite{MEinAP} for a more detailed exposition) and recount some important properties before introducing the QBD-RAP (see \cite{bn2010} for more on QBD-RAPs).
 
 \subsection{Matrix exponential distributions}
-Here we recount some facts about matrix exponential distributions.
-See \cite{MEinAP} for a more detailed exposition.
-A random variable, \(Z\), is said to have a matrix exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
+A random variable, \(Z\), is said to have a matrix exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} = \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential.
 The density function of \(Z\) is given by \(f_Z(x) = \bs\alpha e^{\bs S x}\bs s\).
-The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) In general, in \cite{MEinAP}, there is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility.
+The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) In general there is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility.
 
@@
 A \emph{representation} of a matrix exponential distribution is a triplet \((\bs \alpha, \bs S, \bs s)\), and we write \(Z\sim ME(\bs \alpha, \bs S, \bs s)\) to denote that \(Z\) has a matrix exponential distribution with this representation.
-The order of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the matrix exponential distribution is said to be of order \(p\).
-Representations of matrix-exponential distributions are not unique \citep{MEinAP}.
+The \emph{order} of the representation is the dimension of the square matrix \(\bs S\), i.e.~if \(\bs S\) is \(p\times p\), then the representation is said to be of order \(p\).
+Representations of matrix exponential distributions are not unique \citep{MEinAP}.
 A representation is called \emph{minimal} when \(\bs S\) has the smallest possible dimension.
 Throughout this work, we assume that the representation of any matrix exponential distribution is minimal.
+
 Let \(\bs e_i\) be a vector with a 1 in the \(i\)th position and zeros elsewhere.
-We assume that \(\bs s = -\bs S\bs e\), and that \((\bs e_i,\bs S,\bs s)\) for \(i=1,\dots,p\) are representations of matrix exponential distributions.
+Throughout this thesis, unless otherwise stated, we also assume that the closing vectors of matrix exponential distributions are given by \(\bs s = -\bs S\bs e\), and that \((\bs e_i,\bs S,\bs s)\) for \(i=1,\dots,p\) are representations of matrix exponential distributions.
 It is always possible to find such a representation \cite[Theorem 4.5.17, Corollary 4.5.18]{MEinAP}.
@@
 \begin{cor}[Corollary 4.1.5, \cite{MEinAP}]
-If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix exponential distribution, then \(\bs S\) has a real dominating eigenvalues.
+If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix exponential distribution, then \(\bs S\) has a real dominating eigenvalue.
 \end{cor}
@@
 The class of concentrated matrix exponential distributions (CMEs) is found numerically in \citep{hht2020} and computationally efficient expressions for the density and moments of CMEs are provided therein.
-
-CMEs exist for all orders, however, in our computations we only use CMEs with odd order.
-The justification for considering representations of odd orders only is that the variance of CME distributions of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}.
-Hence, if we construct an approximation with a representation of order \(2p\) we expect it to perform only marginally better than an approximation constructed with representations of order \(2p-1\).
+We refer the reader to \citep{hht2020} for the specific parameters of CMEs.
+
+CMEs exist for all orders, however, in our computations we use CMEs with odd orders only.
+The justification for considering representations of odd orders only is that the variance of CME distributions of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to model deterministic events \citep{hht2020}.
+Hence, when we construct our approximations (in Chapter~\ref{sec: construction and modelling}) using a representation of order \(2p\) we expect it to perform only marginally better than an approximation constructed with a representation of order \(2p-1\).
 However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.
@@
 
-For CMEs the vector function \(\bs k(t)=\cfrac{\bs \alpha e^{St}}{\bs \alpha e^{St}\bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
+For CMEs the vector function \(\bs k(t)=\cfrac{\bs \alpha e^{\bs St}}{\bs \alpha e^{\bs St}\bs e}\) is periodic with period \(\rho = 2\pi/\omega\) where \(\omega=min_i(|\Im(\lambda_i)|)\), \(\lambda_i\) are the eigenvalues of \(\bs{S}\) and \(\Im(z)\) is the imaginary component of a complex number \(z\).
 We leverage this fact to simplify some numerical integration in Chapter~\ref{sec: construction and modelling}.
@@
 %
-%RAPs are an extension of Markovian arrival processes (MAPs) to include matrix-exponential inter-arrival times.
+%RAPs are an extension of Markovian arrival processes (MAPs) to include matrix exponential inter-arrival times.
 For MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain.
@@
 \subsection{QBD-RAPs}\label{sec: qbd-rap}
-% As RAPs are an extension of Markovian arrival processes to include matrix-exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix-exponential times between level changes.
-
-To define a QBD-RAP we first need a Batch (Marked) rational-arrival-process (RAP).
-Let \(\mathscr K\subset \mathbb Z\) be a set of \emph{marks}.
-Let \(N\) be a point process, and \(Y_0=0<Y_1<Y_2\cdots\) be event times of \(N\).
-Let \(\{N(t)\}\) be the counting process associated with \(N\) such that \(N(t)\) returns the number of events by time \(t\).
-Associated with the \(n\)th event is a mark \(M_n\).
+% As RAPs are an extension of Markovian arrival processes to include matrix exponential inter-arrival times, QBD-RAPs are extensions of QBDs to include matrix exponential times between level changes.
+
+To define a QBD-RAP we first define a Marked (Batch) rational-arrival-process (BRAP), and to define the BRAP, we first define the rational-arrival-process (RAP).
+As in \citep{ab1999} (see also Section~10.5 of \cite{MEinAP}) let \((\Omega, \mathcal F, \mathbb P)\) be a given probability space, let \(\mathscr N\) be the set of all counting measures on \((0,\infty)\), let \(\mathscr B(\mathscr N)\) be the Borel \(\sigma\)-algebra of \(\mathscr N\) and let \(\mathscr M(\mathscr N)\) denote the set of finite signed measures on \(\mathscr N\).
+Consider a point process \(N:\Omega\to\mathscr N\), and denote by \(Y_0=0<Y_1<Y_2\cdots\) the event times of \(N\).
+Let \(\{N(t)\}\) be the counting process associated with \(N\) such that \(N(t)\) returns the number of events by time \(t\), and let \(\theta_t:\mathscr N \to\mathscr N\) be the shift operator defined by \[\theta_t N=\{N(t+s)\}_{s\geq 0}=\{N(s)\}_{s\geq t}.\]
+\begin{defn}[\cite{ab1999}]
+	We call a point process \(N\) a rational arrival process if \(\mathbb P(N(0,\infty)=\infty)=1\) and there exists a finite dimensional subspace \(V\) of \(\mathscr M(\mathscr N)\) such that for any \(t\), \(\mathbb P(\theta_tN\in\cdot\mid \mathcal F_t)\) has a version \(\mu(t,\cdot)\) with \(\mu(t,\omega)\in V\) for all \(\omega\in\Omega\).
+\end{defn}
+
+Now, to define the Batch RAP, let \(\mathscr K\subset \mathbb Z\) be a set of \emph{marks}.
+Suppose that associated with the \(n\)th event  of \(N\) is a mark \(M_n\).
 For \(i\in\mathscr K\), let \(N_i\) be simple point processes associated with events with marks of type \(i\) only, and let \(\{N_i(t)\}_{t\geq 0}\) be the associated counting processes of events of mark \(i\).
-Denote by \(f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n)\) the joint density, probability mass function of the first \(n\) inter-arrival times, \(Y_1,Y_2-Y_1,\dots,Y_n-Y_{n-1}\), and the associated marks \(M_n\).
-From Bean and Nielsen \cite[Theorem 1]{bn2010} we have the following.
-\begin{thm}
+\begin{defn}[\cite{bn2010}]
+	We call a marked point process, taking marks in a countable set \(\mathscr K\), a batch rational arrival process (BRAP), if for all \(i\in\mathscr K\), \[\mathbb P(N_i(0,\infty)=\infty)=1\] with \(N_i(0,\infty)\) being the total number of marks of type \(i\) on the positive half-line, and there exists a finite dimensional subspace \(V\) of \(\mathscr M(\mathscr N)\) such that, for any \(t\), \(\mathbb P(\theta_t N\in \cdot \mid \mathcal F_t)\) has a version \(\mu(t,\cdot)\) with \(\mu(t,\omega)\in V\) for all \(\omega \in \Omega\).
+\end{defn}
+
+Denote by \(f_{N,n}(y_1,m_1,y_2,m_2,\dots,y_n,m_n)\) the joint density, probability mass function of the first \(n\) inter-arrival times, \(Y_1,Y_2-Y_1,\dots,Y_n-Y_{n-1}\), and the associated marks \(M_1,...,M_n\).
+From Theorem~1~of~\cite{bn2010} we have the following characterisation of a Marked RAP (analogous to the characterisation of RAPs in \cite{ab1999}).
+\begin{thm}\label{thm:qbdrapchar}
 A process \(N\) is a Marked RAP if there exist matrices \(\bs S\), \(\bs D_i,\,i\in\mathscr K\), and a row vector \(\bs \alpha\) such that \(dev(\bs S)<0\), \(dev(\bs S+\bs D)=0\), \((\bs S+\bs D)\bs e = 0\), \(\bs D = \sum_{i\in\mathscr K} \bs D_i\), and
@@
 \[\bs A(t) = \cfrac{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}}{\bs A(Y_{N(t)}^-)e^{\bs S(t-Y_{N(t)})}\bs e},\]
-where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\).
+where \(\bs A(Y_{N(t)}^-) = \lim_{u\to 0^+}\bs A(Y_{N(t)}-u)\). Equivalently, \(\{\bs A(t)\}\) evolves deterministically according to differential equation
+\[\cfrac{\wrt}{\wrt t} \bs A(t) = \bs A(t)\bs S - \bs A(t)\bs S\bs e\cdot \bs A(t),\] 
+between events.
+
 The process \(\{\bs A(t)\}\) can jump at event times of \(N\) (the process does not necessarily jump at these times, but we may still refer to it a `jump' and typically the dynamics change at this point).
 At time \(t\) the intensity with which \(\{\bs A(t)\}\) has a jump is \(\bs A(t) \bs D\bs e\), i.e.~\(\mathbb P(N(t)=n,N(t+dt)=n+1)=\bs A(t) \bs D\bs e\wrt t\).
-Upon an event the event is associated with mark \(i\) with probability \(\bs A(t) \bs D_i\bs e/\bs A(t) \bs D\bs e\).
+Upon an event at some time \(t\), the event is associated with mark \(i\) with probability \(\bs A(t) \bs D_i\bs e/\bs A(t) \bs D\bs e\).
 Upon on an event at time \(t\) with mark \(i\), the new position of the orbit is \(\bs A(t) = \bs A(t^-) \bs D_i/\bs A(t^-) \bs D_i\bs e\).
-It is important to note that the jumps of the orbit process are \emph{linear} transformations of the orbit process immediately before the time of the jump.
-
-Marked RAPs are an extension of Marked Markovian arrival processes to include matrix-exponential inter-arrival times.
+Thus, jumps of the orbit process are \emph{linear} transformations of the orbit process immediately at the time immediately before the jump.
+
+A key property of Marked RAPs is that the distribution of the process, conditional on the orbit process up to and including time \(t\), can be expressed in terms of \(\bs A(t)\) and measures \(v_1,...,v_p\in V\) which form a basis for the space \(V\);
+\begin{align}\label{eqn: rapproperty}
+	\mu(t,\cdot)=\mathbb P(\theta_tN\in \cdot\mid \mathcal F_t)=\bs A(t)\bs v(\cdot),
+\end{align}
+where \(\bs v=(v_1,...,v_p)'\) form a basis for \(V\).
+
+Marked RAPs are an extension of Marked Markovian arrival processes to include matrix exponential inter-arrival times.
 For Marked MAPs, the vector \(\bs A(t)\) is a vector of posterior probabilities of a continuous-time Markov chain.
 
-Intuitively, \(\bs A(t)\) encodes all the information about the event times of the Marked RAP and associated marks up to time \(t\) that is needed to determine the future behaviour of the point process.
+Intuitively, \(\bs A(t)\) encodes all the information about the event times of the Marked RAP, and the associated marks up to time \(t\), that is needed to determine the future behaviour of the point process.
 Let \(\mathcal F_{t}\) be the \(\sigma\)-algebra generated by \(N(u), u\in[0,t]\).
@@
 Now consider a Marked RAP, \(N\sim BRAP(\bs \alpha, \bs S, \bs D_i,\,i\in\{-1,0,+1\}).\) The process \(\{(L(t),\bs A(t))\}_{t\geq0}\) formed by letting \(L(t) = N_{+1}(t) - N_{-1}(t)\) is a QBD-RAP.
-QBDs are QBD-RAPs where the inter-event times are of Phase-type.
-
-\section{Kronecker products and related results}
-Here we describe Kronecker products and sums and some of their properties (see \cite{MEinAP}, and also Appendix A.4, for further details).
-We use the results in this section to manipulate certain matrix expressions in Chapter~\ref{sec: conv}.
-
-Let 
-\[\bs A = \left[\begin{array}{ccc}a_{11} & \dots & a_{1m}\\\hdots & & \hdots \\ a_{n1}&\dots & a_{nm}\end{array}\right]
-\qquad
-\bs{B} = \left[\begin{array}{ccc}b_{11} & \dots & b_{1m'}\\\hdots & & \hdots \\ b_{n'1}&\dots & b_{n'm'}\end{array}\right]\]
-be matrices with dimensions \(n\times m\) and \(n'\times m'\), respectively. The operator \(\otimes\) is the Kronecker product of two matrices; 
-\[\bs A\otimes \bs{B} = \left[\begin{array}{ccc}a_{11}\bs{B} & \dots & a_{1m}\bs{B}\\\hdots & & \hdots \\ a_{n1}\bs{B}&\dots & a_{nm}\bs{B}\end{array}\right],\]
-which is an \(nn'\times mm'\) matrix.
-
-Let \(\bs{C},\bs{D}\)  be matrices with dimensions \(m\times k\) and \(m'\times k'\). A property of the Kronecker Product is 
-\begin{align}
-	\left(\bs A\otimes \bs{B}\right)\left(\bs{C}\otimes \bs{D}\right) &= \bs A\bs{C}\otimes \bs{B}\bs{D}.\label{eqn:mpr}\tag{Mixed Product Rule}
-\end{align}
-
-% \begin{proof}
-% 	The proof follows from 
-% 	\begin{align*}
-% 		\left[\begin{array}{cccc}a_{i1}\bs{B} & a_{i2}\bs{B}&\dots&a_{in}\bs{B}\end{array}\right]\left[\begin{array}{c}c_{1j}\bs{D}\\c_{2j} \bs D \\\vdots\\c_{nj}\bs{D} \end{array}\right] 
-% 		%
-% 		&= \left(\sum_\ell a_{i\ell}c_{\ell j}\right) \bs{B}\bs{D}
-% 		\\&= \left(\bs A\bs{C}\right)_{ij}\bs{B}\bs{D}.
-% 	\end{align*}
-% \end{proof}
-
-If \(\bs A\) and \(\bs{B}\) are invertible matrices, then 
-\begin{align}\label{eqn:kron inverse}
-	\left(\bs A\otimes \bs{B}\right)^{-1} = \bs A^{-1}\otimes \bs{B}^{-1}.
-\end{align}
-
-Let \(\bs A\) and \(\bs{B}\) be \(n\times n\) and \(m\times m\) matrices, respectively. The Kronecker sum of \(\bs A\) and \(\bs{B}\) is denoted by \(\oplus\) and defined as 
-\[\bs A\oplus \bs{B} := \bs A\otimes \bs{I}_{m} + \bs{I}_{n}\otimes \bs{B}.\]
-
-The exponential of a square matrix \(\bs B\) is \[e^{\bs B} := \sum_{n=0}^\infty \cfrac{1}{n!}\bs B^n.\]
-
-A property of the Kronecker sum is 
-\begin{align}\label{eqn:kron exp}
-	e^{\bs A\oplus \bs{B}}= e^{\bs A}\otimes e^{\bs{B}}.
-\end{align}
-% \begin{proof}
-% 	First, the matrices \(\bs A\otimes \bs{I}_m\) and \(\bs{I}_n\otimes \bs{B}\) commute; from the mixed product rule their product is \(\bs A\otimes \bs{B}\). Hence, 
-% 	\[e^{\bs A\oplus \bs{B}} = e^{\bs A\otimes \bs{I}_m}e^{\bs{I}_n\otimes \bs{B}}.\]
-% 	We now show that \(e^{\bs A\otimes \bs{I}_m} = e^{\bs A}\otimes \bs{I}_m\) and \(e^{\bs{I}_n\otimes \bs{B}}=\bs{I}_n\otimes e^{\bs{B}}\).
-The latter follows from the fact that \(\bs{I}_n\otimes \bs{B}\) is a block diagonal matrix with blocks \(\bs{B}\), hence its exponential is also block diagonal with blocks equal to the exponential of \(\bs{B}\).
-The former follows from
-% 	\begin{align}
-% 	e^{\bs A\otimes \bs{I}_m} &= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A\otimes \bs{I}_m\right)^n \nonumber
-% 	\\&= \sum_{n=0}^\infty \cfrac{1}{n!}\left(\bs A^n\otimes \bs{I}_m\right) \nonumber
-% 	\\&=\left(\sum_{n=0}^\infty \cfrac{1}{n!}\bs A^n\otimes \bs{I}_m\right) \nonumber
-% 	\\&=e^{\bs A}\otimes \bs{I}_m.\label{eqn:09ksdjgah}
-% 	\end{align}
-
-% 	Therefore 
-% 	\[e^{\bs A\oplus \bs{B}} = \left(e^{\bs A}\otimes \bs{I}_m\right)\left(\bs{I}_n\otimes e^{\bs{B}}\right),\]
-% 	and the result follows by the mixed product rule. 
-% \end{proof}
-
-\begin{lem}\label{lem: lst mpr}
-	Let \(\bs{T}\) and \(\bs{C}\) be \(n\times n\), square matrices with \(\bs{C}\) diagonal and invertible; let \(\bs{S}\) be a \(p\times p\) matrix.
-Further, suppose \(\left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]\) is invertible for \(\lambda>0\).
-Then
-\begin{align}
-	&\int_{t=0}^\infty e^{-\lambda t}  e^{{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t}} \wrt t 
-	%
-	=   \int_{x=0}^\infty e^{{\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x}}\otimes e^{\bs{S}x} \wrt x \left(\bs{C}\otimes \bs{I}\right)^{-1}.  \label{eqn:lstsimplify}\end{align}
-\end{lem}
-\begin{proof}
-	Computing the integral on the left-hand side and then factorising the result and using the~\ref{eqn:mpr} multiple times gives
-	\begin{align}
-            	\int_{t=0}^\infty e^{-\lambda t} e^{\left(\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S}\right)t} \wrt t\nonumber 
-            	%
-            	&= - \left[\bs{T}\otimes \bs{I} + \bs{C}\otimes \bs{S} - \lambda \bs{I}\right]^{-1}
-		%
-		\\&= -  \left[\bs{T}\otimes \bs{I} + \left(\bs{C}\otimes \bs{I}\right)\left(\bs{I}\otimes \bs{S}\right) - \lambda \bs{I}\right]^{-1}\nonumber
-		%
-		\\&= -  \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}\otimes \bs{I}\right)^{-1}\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}\otimes \bs{I}\right)^{-1}\lambda \bs{I}\right)\right]^{-1}.
-\label{eqn: ref this one 12}
-		%
-	\end{align}
-	By Equation~(\ref{eqn:kron inverse}) and since \(\bs{C}\) is invertible, (\ref{eqn: ref this one 12}) is equal to
-	\begin{align}
-		& - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\otimes \bs{I}\right)\left(\bs{T}\otimes \bs{I} \right)+ \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\otimes \bs{I}\right)\lambda \bs{I}\right)\right]^{-1}.
-\label{eqn: ref this one 13}
-		%
-	\end{align}
-	{Using the~\ref{eqn:mpr} and algebraic manipulation, (\ref{eqn: ref this one 13}) is equal to }
-	\begin{align}
-		&- \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\bs{T}\right)\otimes \bs{I} + \bs{I}\otimes \bs{S} - \left(\bs{C}^{-1}\lambda \bs{I}\right)\otimes \bs{I}\right)\right]^{-1} \nonumber
-		%
-		\\&= - \left[\left(\bs{C}\otimes \bs{I}\right)\left(\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right)\right]^{-1} \nonumber
-		%
-		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\otimes \bs{I} + \bs{I}\otimes \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1} \nonumber
-		\\&= - \left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1},\label{eqn: is an integral}
-	\end{align}
-	by definition of the Kronecker sum.
-
-	Now, for an invertible matrix \(\bs A\) we can write \(-\bs A^{-1} = \displaystyle\int_{x=0}^\infty e^{\bs Ax}\wrt x\). Therefore, (\ref{eqn: is an integral}) is 
-	\begin{align*}
-		-\left[\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)\oplus \bs{S}\right]^{-1}\left(\bs{C}\otimes \bs{I}\right)^{-1}
-		&= \int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)x\right)\oplus \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1}.
-	\end{align*}
-	{Using the rule in Equation~(\ref{eqn:kron exp}) gives }
-	\begin{align*}
-		&\int_{x=0}^\infty e^{\left(\bs{C}^{-1}\left(\bs{T}-\lambda \bs{I}\right)\right)x}\otimes e^{ \bs{S}x}\wrt x\left(\bs{C}\otimes \bs{I}\right)^{-1},
-	\end{align*}
-	which is the result.
-\end{proof}
-
-% vector space? Polynomial basis?
-% me basis? semigroup? generator?
-
-%Laplace transforms of probability measures with non-negative support and where the Laplace transform variable, \(\lambda\), is real and non-negative can be given a stochastic interpretation.
-Let \(W\) be a random variable with distribution function \(F_W(w)= \mathbb P(W<w)\), then \(\displaystyle\int_{w=0}^\infty e^{-\lambda w} \wrt F_W(w) = \mathbb P(W < E^{\lambda})\).
-That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(W\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.
-
-\section{Convergence theorems}
-We use the following convergence theorems to help us prove that the QBD-RAP scheme converges weakly to the distribution of the fluid queue.
-The first result we state, the Portmanteau Theorem, is a sweeping statement about convergence of measures.
-First, let's define the notion of weak convergence.
-We follow \cite{billingsleyconvergence}.
-
-Let \(S\) be a metric space and let \(\mathcal S\) be the Borel \(\sigma\)-algebra generated by the open sub-sets of \(S\).
-A probability measure, \(P\), is a function which maps elements of \(\mathcal S\) (elements of \(\mathcal S\) are sets) to real numbers in the interval \([0,1]\), with \(P(S)=1\).
-Further, \(P\) is \emph{countably-additive}, which means that for any countable collection of \emph{disjoint} sets \(A_1, A_2,...\in\mathcal S\), then
-\[P\left(\bigcup\limits_{n=1}^\infty A_n\right) = \sum\limits_{n=1}^\infty P\left(A_n\right).\]
-Define the notation \(Pf = \int_S f \wrt P\) where \(f:S\to \mathbb R\) is a function.
-Let \(P_1,P_2,...\), be a sequence of probability measures.
-For a given function \(f\), the sequence \(P_1f,P_2f,...\), is simply a sequence of real numbers.
-The sequence of probability measures \(P_1,P_2,....\), is said to \emph{converge weakly} to \(P\) if \(P_nf\to Pf\) for every bounded continuous real function \(f\).
-\cite{billingsleyconvergence} uses the notation \(P_n\Rightarrow P\) to denote this weak convergence.
-
-We need a few more definitions before stating The Portmanteau Theorem.
-A set \(A\in\mathcal S\) is said to be a \(P\)-\emph{continuity set} if \(\partial A\) (the boundary of the set \(A\)) satisfies \(P(\partial A)=0\).
-Define the \emph{limit inferior} and \emph{limit superior} of a sequence of real numbers \(x_1,x_2,...,\) as
-\begin{align*}
-	\liminf_{n\to\infty} x_n &:= \lim_{n\to \infty}	\left(\inf_{m\geq n}x_m\right)
-	\intertext{and}
-	\\\limsup_{n\to\infty} x_n &:= \lim_{n\to \infty}	\left(\sup_{m\geq n}x_m\right),
-\end{align*}
-respectively.
-
-\begin{thm}[Portmanteau Theorem I, Theorem~2.1 of \cite{billingsleyconvergence}]
-	These five conditions are equivalent.
-	\begin{enumerate}
-		\item[(i)] \(P_n\Rightarrow P\). 
-		\item[(ii)] \(P_nf\to Pf\) for all bounded, uniformly continuous \(f\).
-		\item[(iii)] \(\limsup_n P_n F \leq P F\) for all closed [sets] \(F\).
-		\item[(iv)] \(\liminf_n P_n G \leq P G\) for all open [sets] \(G\).
-		\item[(v)] \(P_n A \to P A\) for all \(P\)-continuity sets \(F\).
-	\end{enumerate}
-\end{thm}
-Some authors, such as \cite{portmanteaubook}, also include the following equivalent condition in the Portmanteau Theorem. 
-\begin{thm}[Portmanteau Theorem II, Theorem~13.16 of \cite{portmanteaubook}]\label{thm: Portmanteau}
-	\(\,\)
-	\begin{enumerate}
-		\item[(vi)] \(P_nf \to P f\) for all bounded, Lipschitz continuous \(f\).
-	\end{enumerate}
-\end{thm}
-
-Recall that a stochastic process is a sequence of random variables \(\{X(t)\}_{t\in\mathcal T}\).
-The \emph{finite-dimensional distributions} of \(\{X(t)\}_{t\in\mathcal T}\) are the joint distributions of the random vector \((X(t_1),X(t_2),...,X(t_n))\), where \(t_1,t_2,...,t_n\in \mathcal T\) is a finite collection of times.
-
-Key concepts in establishing weak convergence of stochastic processes are the notions \emph{tight} and \emph{relatively compact}.
-Again, we follow \cite{billingsleyconvergence}.
-A probability measure \(P\) is said to be \emph{tight} if for each \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\).
-A family of probability measures, \(\Pi\), is said to be tight if for every \(\varepsilon\) there exists a compact set \(K\) such that \(P(K)>1-\varepsilon\) for every probability measure \(P\) in \(\Pi\).
-The family \(\Pi\) is \emph{relatively compact} if every sequence of elements of \(\Pi\) contains a subsequence which converges weakly.
-
-Let \(P_1,P_2,...,\) and \(P\) be probability measures of stochastic processes. \cite{billingsleyconvergence} provides the following result.
-\begin{thm}\label{thm: aofaa}
-	If \(\{P_n\}\) is relatively compact and the finite-dimensional distributions of \(P_n\) converge weakly to those of \(P\), then \(P_n\Rightarrow_n P\). 
-\end{thm}
-Further, the condition that \(\{P_n\}\) is relatively compact can be replaced by tightness due Prohorov's Theorem \citep{billingsleyconvergence}. 
-\begin{thm}
-	If \(\Pi\) is tight, then it is relatively compact.
-\end{thm}
-
-% Denote by \(\mathcal M_f(E)\) the set of all finite measures on \((E,\mathcal E)\), where \(E\) is a non-empty set and \(\mathcal E\) is a \(\sigma\)-algebra.
-Further, denote by \(C_b(E)\) the set of continuous bounded functions on \(E\).
-Let \(\mu,\,\mu_1,\,\mu_2,...\in \mathcal M_f(E)\).
-We say that \(\{\mu_n\}_{n\in\mathbb N}\) converges weakly to \(\mu\), formally \(\mu_n\to \mu\) weakly as \(n\to\infty\), if
-% \[\int f \wrt \mu_n \to \int f\wrt \mu,\mbox{ for all }f \in C_b(E).\]
-% We use part of The Portmanteau Theorem in Chapter~\ref{ch: global results}.
-Let \(\mathcal M_{\leq 1}(E):=\{\mu \in \mathcal M_f(E)\mid \mu(E)\leq 1\},\) the set of all sub-probability measures on \((E,\mathcal E)\).
-% \begin{thm}[(Part of) The Portmanteau Theorem, Theorem~13.16 of \cite{portmanteaubook}]
-% 	Let \(E\) be a metric space and let \(\mu,\, \mu_1,\,\mu_2,... \in \mathcal M_{\leq 1}(E)\). The following are equivalent.
-% 	\begin{enumerate}
-% 		\item[(i)] \(\mu_n\to \mu\) weakly as \(n\to \infty\). 
-% 		\item[(ii)] \(\int f\wrt \mu_n \to \int f\wrt \mu\) for all bounded, Lipschitz continuous \(f\).
-% 		\\ \(\vdots\) 
-% 	\end{enumerate}
-% \end{thm}
-% There are 8 parts to The Portmanteau Theorem in \cite{portmanteaubook}, Theorem~13.16.
-Here we only quote to relevant parts; the other 6 parts require us to define other concepts which are not relevant to this thesis, so they are omitted.
-See \cite{portmanteaubook}, Theorem~13.16 for details.
-
-Another tool we can use to show convergence of measures is to show that the Laplace transforms converge, as stated in the following theorem.
-\begin{thm}[Extended Continuity Theorem, \cite{feller1957}, Theorem 2a]\label{thm: ext cont thm}
-	For \(p=1,2,...,\) let \(U_p\) be a measure with Laplace transform \(\zeta_p\).
-If \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda > a\geq 0\), then \(\zeta\) is the Laplace transform of a measure \(U\) and \(U_p\to U\) [weakly].
-
-	Conversely, if \(U_p\to U\)[weakly] and the sequence \(\{\zeta_p(a)\}\) is bounded, then \(\zeta_p(\lambda)\to\zeta(\lambda)\) for \(\lambda >a\). 
-\end{thm}
-
-In Chapters~\ref{sec: conv} and~\ref{ch: global results} we use the Dominated Convergence Theorem to aid our convergence arguments.
-In applied probability we often want to prove convergence of certain expressions.
-An approach which can simplify matters is to partition the expression on certain events where we have a simpler characterisation, thereby enabling us to prove convergence on each element in the partition.
-The original expression can be written as an integral over the partition.
-To establish the convergence result we initially desired, we can use the convergence of each element of the partition and the Dominated Convergence Theorem.
-I have taken the following from Theorem~1.13~\cite{steinreal}.
-\cite{steinreal} use the notation
-\[\int f = \int f \wrt x = \int f \wrt m(x),\]
-where \(m\) denotes Lebesgue measure, to denote the Lebesgue integral. 
-\begin{thm}[Dominated Convergence Theorem]
-	Suppose \(\{f_n\}\) is a sequence of measurable functions such that \(f_n(x)\to f(x)\) almost everywhere with respect to \(x\), as \(n\) tends to infinity.
-If \(|f_n(x)|\leq g(x)\), where \(g\) is integrable, then
-	\[\int|f_n-f|\to 0 \mbox{ as } n \to \infty,\]
-	and consequently 
-	\[\int f_n\to\int f \mbox{ as } n\to \infty.\]
-\end{thm}
-
-Also in Chapters~\ref{sec: conv} and~\ref{ch: global results} we want to manipulate infinite sums or integrals and rearrange the order of summation or integration.
-However, things can go awry when we swap the order of integration/summation if we are not careful.
-The next few results give some conditions under which we have equality under before and after swapping the order of summation/integration.
-Once again, we follow \cite{steinreal} quoting their Theorem~2.13.
-If \(f\) is a function in \(\mathbb R^{d}=\mathbb R^{d_1}\times \mathbb R^{d_2}\), the \emph{slice} of \(f\) corresponding to \(y\in\mathbb R^{d_2}\) is the function \(f^y\) of the \(x\in\mathbb R^{d_1}\) variable, given by
-\[f^y(x)=f(x,y).\]
-Similarly, the slice of \(f\) for a fixed \(x\in\mathbb R^{d_1}\) is \(f_x(y)=f(x,y)\). 
-\begin{thm}[Fubini's Theorem]\label{Fubini}
-	Suppose \(f(x,y)\) is integrable on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
-	\begin{itemize}
-		\item The slice \(f^y\) is integrable on \(\mathbb R^{d_1}\).
-		\item The function defined by \(\int_{R^{d_1}} f^y(x)\wrt x\) is integrable on \(\mathbb R^{d_2}\). 
-	\end{itemize}
-	Moreover:
-	\begin{itemize}
-		\item[(iii)] \(\displaystyle\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y = \int_{\mathbb R^d}f.\)
-	\end{itemize}
-\end{thm}
-\cite{steinreal} then state \begin{quotation}``Clearly, the [Fubini] theorem is symmetric in \(x\) and \(y\) so that we also may conclude that the slice \(f_x\) is integrable on \(\mathbb R^{d_2}\) for almost every \(x\).
-Moreover, \(\int_{\mathbb R^{d_2}}f_x(y)\wrt y\) is integrable and
-\[\displaystyle\int_{R^{d_1}}\left(\int_{R^{d_2}}f(x,y)\wrt y\right)\wrt x = \int_{\mathbb R^d}f.\]
-In particular, Fubini's theorem states that the integral of \(f\) on \(\mathbb R^d\) can be computed by iterating lower-dimensional integrals, and that the iterations can be taken in any order
-\[\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y=\displaystyle\int_{R^{d_1}}\left(\int_{R^{d_2}}f(x,y)\wrt y\right)\wrt x = \int_{\mathbb R^d}f.\mbox{''}\]
-\end{quotation}
-It is this last statement which is most powerful. Effectively, if either
-\[\int_{R^{d_2}}\left(\int_{R^{d_1}}|f(x,y)|\wrt x\right)\wrt y<\infty,\]
-or
-\[\int_{R^{d_1}}\left(\int_{R^{d_2}}|f(x,y)|\wrt y\right)\wrt x<\infty,\]
-then we can swap the order of integration.
-
-A closely related theorem which is often used alongside Fubini's Theorem is Tonelli's Theorem.
-Define the \emph{extended Lebesgue integral} of an extended valued (it can take the value \(+\infty\)) non-negative function \(f\) by
-\[\int f(x)\wrt x = \sup_{g}\int g(x)\wrt x.\]
-\begin{thm}[Tonelli's Theorem]\label{Tonelli}
-	Suppose \(f(x,y)\) is a non-negative measurable function on \(\mathbb R^{d_1}\times \mathbb R^{d_2}\). Then for almost every \(y\in\mathbb R^{d_2}\):
-	\begin{itemize}
-		\item The slice \(f^y\) is integrable on \(\mathbb R^{d_1}\).
-		\item The function defined by \(\int_{R^{d_1}} f^y(x)\wrt x\) is integrable on \(\mathbb R^{d_2}\). 
-	\end{itemize}
-	Moreover:
-	\begin{itemize}
-		\item[(iii)] \(\displaystyle\int_{R^{d_2}}\left(\int_{R^{d_1}}f(x,y)\wrt x\right)\wrt y = \int_{\mathbb R^d}f(x,y)\wrt x\wrt y \mbox{ in the extended sense}.\)
-	\end{itemize}
-\end{thm}
-
-Once again, we note that the theorem is symmetric in \(x\) and \(y\), so we can establish that we may swap the order of integration provided that \(f\) is non-negative.
-Tonelli's Theorem allows us to exchange the order of integration for any non-negative extended real-valued function, and includes the case where the value of the integral is infinite.
-In contrast, Fubini's Theorem allows us to exchange the order of integration for any real-valued function, provided that the integral is finite.
-
-Collectively, we refer to Theorems~\ref{Fubini} and~\ref{Tonelli} together as the Fubini-Tonelli Theorem, but they are otherwise known collectively as just Fubini's Theorem.
-Often, they are used in conjunction.
-Since \(|f|\) is a non-negative function, then we may use Tonelli's Theorem and compute (or bound) the integral \(\int |f|\) via computing an iterated integral.
-If this is found to be finite, then Fubini's Theorem applies so \(f\) is integrable, and we may evaluate \(\int f\) via an iterated integral.
-
-In the context of probability, the function \(f\) which we are integrating is often positive, so Tonelli's Theorem is all that is required to justify a swap of iterated integrals.
-
-\section{Sundry mathematical concepts}\label{eqn: la}
-\subsection*{Polynomials}
-The discontinuous Galerkin method involves projecting the operator equation onto a basis of functions, typically polynomials.
-A convenient basis with which to work is the interpolating Lagrange polynomials.
-The order \(p-1\) Lagrange polynomials are defined by a set of points, \(\xi_i, i=1,...,p\), (the \(\xi_i\)'s must be distinct), and are given by
-\[\ell_i(r)=\prod_{\substack{j=1\\j\neq i}}^p \cfrac{r-\xi_j}{\xi_i-\xi_j},\, i=1,...,p.\]
-A convenient property of the Lagrange polynomials is 
-\[l_i(\xi_j)=\begin{cases}1 & \mbox{ if } i=j, \\ 0 & \mbox{ otherwise.} \end{cases}\]
-
-Sometimes it is more convenient to work with \emph{orthogonal} polynomials. For example, the Legendre polynomials, which can be defined recursively by 
-\begin{align*}
-	P_0(x)&=1,
-	\\P_1(x)&=x, 
-	\\(n+1)P_{n+1}(x)&=(2n+1)xP_n(x)-nP_{n-1}(x),
-\end{align*}
-and are orthogonal on \([-1,1]\).
-
-The zeros of \((1-x^2)P_n'(x)\) define the \emph{Legendre-Gauss-Lobatto} points, which are used, among other things, for numerically approximating integrals via quadrature.
-
-\subsection*{Numerical integration}
-To numerically approximate integrals in the discontinuous Galerkin schemes in this thesis we can use Gauss-Lobatto quadrature.
-Consider the integral of some function \(f\) over the interval \([-1,1]\).
-Quadrature approximates the integral by evaluating the function on a set of points, \(x_i,\, i=1,...,p\), and computing the weighted sum
-\[\int_{-1}^1f(x)\wrt x \approx \sum_{i=1}^p f(x_i)w_i,\]
-where \(w_i\) are weights. There are various quadrature schemes which one can use.
-
-For the discontinuous Galerkin method, Gauss-Lobatto quadrature is convenient as in both we evaluate the function at the end points of the interval.
-For Gauss-Lobatto quadrature, the nodes, \(x_i, i=1,...,p\) is the zero of \((1-x^2)P_n'(x)\) where \(P_n\) is the \(n\)th Legendre polynomial, the weights are given by
-\begin{align*}
-	w_1&=1,
-	\\w_i&=\cfrac{2}{n(n-1)[P_{n-1}(x_i)]^2}, \, i\neq 1, p,
-	\\w_p&=1.
-\end{align*}
-Gauss-Lobatto quadrature is accurate for polynomials up to degree \(2p-3\). An integral over the interval \([a,b]\) can be approximated by 
-\[\int_{a}^bf(x)\wrt x \approx \cfrac{b-a}{2}\sum_{i=1}^p f\left(\cfrac{b-a}{2}x_i+\cfrac{a+b}{2}\right)w_i.\]
-
-We also use a trapezoidal integration rule at times. Consider the integral \(\int_{a}^bf(x)\wrt x\). A trapezoidal rule with \(p\) evenly spaced grid-points approximates the integral via 
-\[\int_{a}^bf(x)\wrt x \approx \sum_{i=2}^p \cfrac{f(x_{i-1}+f(x_i))}{2}\Delta,\]
-where \(x_i=a+(i-1)\Delta\), \(i=1,...,p\), and \(\Delta = \cfrac{b-a}{p-1}.\)
-
-\subsection*{Measuring the difference between distributions}
-In the numerical investigations in this thesis we evaluate approximation schemes by comparing the resulting approximations they produce with a ground truth.
-For comparing distributions we use \(L^p\) \emph{norms} and the Kolmogorov-Smirnov distance.
-The \(L^p\) norm of a function \(f\) is given by
-\[\left(\int |f(x)|^p\wrt x\right)^{1/p}.\]
-The Kolmogorov-Smirnov distance between two distribution functions, \(F_1\), \(F_2\) is 
-\[\sup_{x}|F_1(x)-F_2(x)|.\]
+A QBD is a QBD-RAP where the inter-event times are Phase-type.
 
@@
 
-\paragraph{Fluid-fluid queues and related models} Related to the analysis of fluid-fluid queues are the works of \cite{mz2012,lnp13,bo2013b} and \cite{bop2020}.
+\paragraph{Fluid-fluid queues and related models} Related to the analysis of fluid-fluid queues are the works of \cite{mz2012,lnp13,bo2013b} and \cite{boz2022}.
 \cite{mz2012} analyse related discrete-time multidimensional Markov additive processes and derive operator-analytic expressions for the limiting distribution.
@@
 \cite{bo2013b} analyse a model which is simple special case of the fluid-fluid queue.
-They consider a model where the second level process \(\{\dot Y(t)\}\) has constant rates which do not depend on \(\ddot X(t)\) (but do depend on \(\varphi(t)\)), while \(\{(\ddot X(t),\varphi(t))\}\) is an \emph{unbounded} fluid queue.
-They derive matrix-analytic expressions for certain Laplace transforms, with respect to the position of the first fluid \(\{X(t)\}\) relative to its starting position \(X(0)\), of the probability of certain transient events (such as first return and draining/filling times) of \(\{Y(t)\}\).
+They consider a model where the second level process \(\{\dot W(t)\}\) has constant rates which do not depend on \(\ddot X(t)\) (but do depend on \(\varphi(t)\)), while \(\{(\ddot X(t),\varphi(t))\}\) is an \emph{unbounded} fluid queue.
+They derive matrix-analytic expressions for certain Laplace transforms, with respect to the position of the first fluid \(\{X(t)\}\) relative to its starting position \(X(0)\), of the probability of certain transient events (such as first return and draining/filling times) of \(\{\dot W(t)\}\).
 %In contrast, \cite{bo2014} treats a more general model.
-The work of \cite{bop2020} treats a similar model to that of \cite{bo2013b}, except that the driving process has a regulated lower boundary at \(0\); this significantly complicates the analysis.
+The work of \cite{boz2022} treats a similar model to that of \cite{bo2013b}, except that the driving process has a regulated lower boundary at \(0\); this significantly complicates the analysis.
 They derive matrix-analytic expressions for the first return operator of the second fluid given a specific exponential form of the initial distribution and a certain boundary condition is met.

### 2023-01-22 chapter2/chapter.tex
post-reviews, pre supervisor feedback

@@
 	\begin{align*} 
-	    f_i(x,t) := \cfrac{\partial}{\partial x} 	\mathbb{P}\left(X(t) \leq x, \varphi(t) = i\right),\quad i \in\calS,
+	    f_i(x,t) = \cfrac{\partial}{\partial x} 	\mathbb{P}\left(X(t) \leq x, \varphi(t) = i\right),\quad i \in\calS,
 	\end{align*} 
@@
 
-For each \(k\in \mathbb Z\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) for all \(i\in\calS\), to form a basis for the space \(W_k\), in which we formulate the approximation.
+For each \(k\in \mathbb Z\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) for all \(i\in\calS\), to form a basis for the space \(U_k\), in which we formulate the approximation.
 Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{p_k}\) to be a basis for the space of polynomials of degree \(p_k-1\).
@@
 
-Let \(\mathcal N_k := \left\{1,\dots,p_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
+Let \(\mathcal N_k = \left\{1,\dots,p_k\right\},\, k \in \mathbb Z\). For \(k\in\mathbb Z,\) define \textit{local} row-vectors 
 \[\boldsymbol \phi_k(x) = (\phi^r_k(x))_{r\in\mathcal N_k}, \quad \boldsymbol a_{k,i}(t) = (a_{k,i}^r(t))_{r\in\mathcal N_k},\,i\in\mathcal S.\]
@@
 The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation~(\ref{eqn:pde_density_matrix}).
-In general, to construct the weak-form we need a set of test functions, say \(W\).
-Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to get
+In general, to construct the weak-form we need a set of test functions, say \(U\).
+Now, take the strong form of the PDE, multiply it by some test function \(\psi\in U\), integrate with respect to \(x\), and apply integration by parts to get
 \begin{align}
@@
 for \(j\in\calS\).
-It is convenient to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero.
-%Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE.
+It is convenient to choose \(U\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero.
+%Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in U\) gives the weak-formulation of the PDE.
 Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE.
@@
 
-For the problem at hand, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of \(W_k\), where \(W_k\) is the domain of the approximation on the interval \(\calD_k\).
+For the problem at hand, we take the set of test functions to be \(U = \bigoplus\limits_{k\in\mathbb Z} U_k\), the direct sum of \(U_k\), where \(U_k\) is the domain of the approximation on the interval \(\calD_k\).
 Further, we use the same basis functions for the test-function space and the approximation space.
@@
 since \(\phi^r_k\) is compactly supported on \(\calD_{k,j}\).
-Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), the orthogonal complement of \(W\).
-Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi^r_k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\).
-Also, note that \(\cfrac{\wrt}{\wrt x}\phi^r_k(x)\in W\).
+Now, as any function \(g(x)\) can be decomposed as \(g(x) = g^{U}(x)+g^\perp(x)\) where \(g^{U}\in U\) and \(g^\perp \in U^\perp\), the orthogonal complement of \(U\).
+Since \(g^\perp\) is orthogonal to \(U\), \(\displaystyle\int_{x}g^\perp(x)\phi^r_k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\).
+Also, note that \(\cfrac{\wrt}{\wrt x}\phi^r_k(x)\in U\).
 Using this, we can write
 \begin{align*}
-	\begin{multlined}[t]\int_{x=y_{k}}^{y_{k+1}}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi^r_k(x)\wrt x 
-	= \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi^r_k(x)\wrt x  
+	\begin{multlined}[t]\int_{x=y_{k}}^{y_{k+1}}\cfrac{\partial}{\partial t} \left(f_j^U(x,t)+f_j^\perp(x,t)\right)\phi^r_k(x)\wrt x 
+	= \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \left(f_i^U(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi^r_k(x)\wrt x  
 	%
 	\\ {}
-	+  \int_{x=y_k}^{y_{k+1}} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
+	+  \int_{x=y_k}^{y_{k+1}} \left(f_j^U(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
 \end{align*}
@@
 \begin{align}
-	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi^r_k(x)\wrt x  
+	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j^U(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} f_i^U(x,t)T_{ij}\phi^r_k(x)\wrt x  
 	%
-	\\{} + \int_{x=y_k}^{y_{k+1}} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}. \end{multlined} \label{eqn:hash}
+	\\{} + \int_{x=y_k}^{y_{k+1}} f_j^U(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}. \end{multlined} \label{eqn:hash}
 \end{align}
-Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), where the prime denotes the transpose, which we now substitute into (\ref{eqn:hash})
+Now, \(f_j^U(x,t)\in U\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t)=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), where the prime denotes the transpose, which we now substitute into (\ref{eqn:hash})
 %\begin{align*}
@@
 
-\section{Mass, stiffness, and flux matrices}
+\section{Mass, stiffness, and flux matrices}\label{sec:llaaksksbnnn}
 Recall, for \(k\in\mathbb Z\), we defined local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
-\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x.\]
+\[\bs M_k = \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k = \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x.\]
 %and global mass and stiffness matrices by
 %\begin{align*}
-%	M &:= \left[\begin{array}{ccc} M_1 & & \\ & \ddots & \\ & & M_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\boldsymbol \phi(x)\wrt x,
-%	\\G &:= \left[\begin{array}{ccc} G_1 & & \\ & \ddots & \\ & & G_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\cfrac{\wrt}{\wrt x}\boldsymbol \phi(x)\wrt x,
+%	M &= \left[\begin{array}{ccc} M_1 & & \\ & \ddots & \\ & & M_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\boldsymbol \phi(x)\wrt x,
+%	\\G &= \left[\begin{array}{ccc} G_1 & & \\ & \ddots & \\ & & G_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\cfrac{\wrt}{\wrt x}\boldsymbol \phi(x)\wrt x,
 %\end{align*}
@@
 \end{align*}
-where, for \(j\in \calS\) with \(c_j>0\), we define \(\bs F_+^{k,k}=\bs F_j^{k,k} := -\boldsymbol \phi_k(y_{k+1}^-)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z\) and \(\bs F_j^{k-1,k} := \boldsymbol \phi_{k-1}(y_{k}^-)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
+where, for \(j\in \calS\) with \(c_j>0\), we define \(\bs F_+^{k,k}=\bs F_j^{k,k} = -\boldsymbol \phi_k(y_{k+1}^-)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z\) and \(\bs F_j^{k-1,k} = \boldsymbol \phi_{k-1}(y_{k}^-)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
 
@@
 \end{align*}
-where, for \(j\in \calS\) with \(c_j<0\), we define \(\bs F_j^{k+1,k} := -\boldsymbol \phi_{k+1}(y_{k+1}^+)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z,\) and \(\bs F_-^{k,k}=\bs F_j^{k,k} := \boldsymbol \phi_{k}(y_{k}^+)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
+where, for \(j\in \calS\) with \(c_j<0\), we define \(\bs F_j^{k+1,k} = -\boldsymbol \phi_{k+1}(y_{k+1}^+)\tr{}\boldsymbol\phi_k(y_{k+1}),\,k\in\mathbb Z,\) and \(\bs F_-^{k,k}=\bs F_j^{k,k} = \boldsymbol \phi_{k}(y_{k}^+)\tr{}\boldsymbol\phi_k(y_{k}),\,k\in\mathbb Z\).
 
@@
 Let $[0,b]$ be the domain of the approximation, where $b < \infty$.
-We partition the space $[0,b]$ into \(\mathcal D_{-1}=\{0\},\) \(\mathcal D_{K+1}=\{b\},\) and \(K\) non-trivial intervals, \(\calD_{k,i}=[y_k,y_{k+1})\setminus \{0\},\,i\in\calS_+\cup\calS_0,\calD_{k,i}=(y_k,y_{k+1}]\setminus \{0\},\,i\in\calS_-,\, y_k<y_{k+1},\, k\in\mathcal K^\circ\), \(y_0=0,\,y_{K+1}=b\) and define \(\Delta_k := y_{k+1}-y_k\).
+We partition the space $[0,b]$ into \(\mathcal D_{-1}=\{0\},\) \(\mathcal D_{K+1}=\{b\},\) and \(K\) non-trivial intervals, \(\calD_{k,i}=[y_k,y_{k+1})\setminus \{0\},\,i\in\calS_+\cup\calS_0,\calD_{k,i}=(y_k,y_{k+1}]\setminus \{0\},\,i\in\calS_-,\, y_k<y_{k+1},\, k\in\mathcal K^\circ\), \(y_0=0,\,y_{K+1}=b\) and define \(\Delta_k = y_{k+1}-y_k\).
 
@@
 \end{align}
-where \(\bs P_{-,+} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{+}}\), \(\bs P_{-,{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\) and \(p_{ij}^{-1}\) is the probability of instantaneous transition of the phase process from Phase \(i\) to Phase \(j\) upon the process hitting the lower boundary.
+where \(\bs P_{-,+} = \left[\widecheck p_{ij}\right]_{i\in\calS_-,j\in\calS_{+}}\), \(\bs P_{-,{-1}} = \left[\widecheck p_{ij}\right]_{i\in\calS_-,j\in\calS_{-1}}\) and \(\widecheck p_{ij}\) is the probability of instantaneous transition of the phase process from Phase \(i\) to Phase \(j\) upon the process hitting the lower boundary.
 Equation~(\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(\bs T_{{-1},{-1}}\), and that the flux of probability density into the point masses is \(\boldsymbol f_-(0,t)\bs P_{-,{-1}}\bs C_{-1}\).
@@
 \begin{equation*}\label{eqn:DGBC1}
-\cfrac{\wrt}{\wrt t} q_{{-1},j}(t) = \sum_{i\in\calS_{-1}} q_{{-1},i}(t) T_{i j} - \sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0^+)\tr{}p_{ij}^{-1} c_i.
+\cfrac{\wrt}{\wrt t} q_{{-1},j}(t) = \sum_{i\in\calS_{-1}} q_{{-1},i}(t) T_{i j} - \sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0^+)\tr{}\widecheck p_{ij} c_i.
 \end{equation*}
@@
 	&-c_j[f_j(x,t)\boldsymbol\phi_0(x)]_{x=0}^{x=y_{0}} 
-	\\& =  -c_jf_j(y_0^-,t)\boldsymbol\phi_0(y_0)+c_jf_j(0^+,t)\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_if_i(0^+,t)p_{ij}^{-1}\bs \phi_0(0)
-	\\& \approx -c_j f_j^*(y_0,t)\boldsymbol\phi_0(y_0)+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs \phi_0(0)\tr{}\bs \phi_0(0)
-	\\& = c_j\boldsymbol a_{0,j}(t)\bs F_j^{0,0}+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs F_j^{0,0},
+	\\& =  -c_jf_j(y_0^-,t)\boldsymbol\phi_0(y_0)+c_jf_j(0^+,t)\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_if_i(0^+,t)\widecheck p_{ij}\bs \phi_0(0)
+	\\& \approx -c_j f_j^*(y_0,t)\boldsymbol\phi_0(y_0)+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_i\widecheck p_{ij}\bs a_{0,i}(t)\bs \phi_0(0)\tr{}\bs \phi_0(0)
+	\\& = c_j\boldsymbol a_{0,j}(t)\bs F_j^{0,0}+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_i\widecheck p_{ij}\bs a_{0,i}(t)\bs F_j^{0,0},
 \end{align*}
@@
 
-Thus, the DG approximation of the flux into point masses \(q_{{-1},j}(t)\) is \[-\sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0)\tr{} p_{ij}^{-1} c_i,\,j\in\calS_-,\] the rate of transition of point mass within \(\bs q_{{-1}}(t)\) is \(\bs T_{{-1},{-1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0),\,j\in\calS_+\), and the DG approximation to density reflected at the lower boundary is \(\sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs F_j^{0,0}\).
+Thus, the DG approximation of the flux into point masses \(q_{{-1},j}(t)\) is \[-\sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0)\tr{} p_{ij}^{-1} c_i,\,j\in\calS_-,\] the rate of transition of point mass within \(\bs q_{{-1}}(t)\) is \(\bs T_{{-1},{-1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0),\,j\in\calS_+\), and the DG approximation to density reflected at the lower boundary is \(\sum_{i\in\calS_-}c_i\widecheck p_{ij}\bs a_{0,i}(t)\bs F_j^{0,0}\).
 
@@
 \end{align*}
-where \(\bs P_{+,-} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{-}}\), \(\bs P_{+,K+1} = \left[p_{ij}^{K+1}\right]_{i\in\calS_+,j\in\calS_{K+1}}\) and \(p_{ij}^{K+1}\) is the probability of instantaneous transition of the phase process from Phase \(i\) to Phase \(j\) upon the process hitting the lower boundary.
+where \(\bs P_{+,-} = \left[\widehat p_{ij}\right]_{i\in\calS_+,j\in\calS_{-}}\), \(\bs P_{+,K+1} = \left[\widehat p_{ij} \right]_{i\in\calS_+,j\in\calS_{K+1}}\) and \(\widehat p_{ij} \) is the probability of instantaneous transition of the phase process from Phase \(i\) to Phase \(j\) upon the process hitting the lower boundary.
 Using the same arguments as above,
 \begin{align*}
-\cfrac{\wrt}{\wrt t} q_{{K+1},j}(t) &= \sum_{i\in\calS_{K+1}}q_{{K+1},i}(t) T_{ij} + \sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(b)\tr{}p_{ij}^{K+1} c_i,
-\\-c_j[f_j(x,t)\boldsymbol\phi_K(x)]_{x=y_K}^{x=b} & \approx c_j\boldsymbol a_{K,j}(t)F_j^{K,K}+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b) 
-\\&\quad{} - \sum_{i\in\calS_+}c_ip_{ij}^{K+1}\bs a_{0,i}(t)\bs F_j^{0,0}
+\cfrac{\wrt}{\wrt t} q_{{K+1},j}(t) &= \sum_{i\in\calS_{K+1}}q_{{K+1},i}(t) T_{ij} + \sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(b)\tr{}\widehat p_{ij} c_i,
+\\-c_j[f_j(x,t)\boldsymbol\phi_K(x)]_{x=y_K}^{x=b} & \approx c_j\boldsymbol a_{K,j}(t)\bs F_j^{K,K}+\sum_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b) 
+\\&\quad{} - \sum_{i\in\calS_+}c_i\widehat p_{ij} \bs a_{0,i}(t)\bs F_j^{0,0}
 \end{align*}
@@
 Thus, the DG approximation of the flux into the point mass \(q_{{K+1},j}(t)\) is 
-\[\sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(0)\tr{} p_{ij}^{K+1} c_i,\]
-\(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{{K+1}}(t)\) is \(\bs T_{{K+1},{K+1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b)\), \(j\in\calS_-\) and the approximation to reflection of density at the upper boundary is \(- \sum_{i\in\calS_+}c_ip_{ij}^{K+1}\bs a_{0,i}(t)\bs F_j^{0,0}\).
+\[\sum_{i\in\calS_+}\boldsymbol a_{K,i}(t)\boldsymbol \phi_K(0)\tr{} \widehat p_{ij} c_i,\]
+\(j\in\calS_+\), the rate of transition of point mass within \(\bs q_{{K+1}}(t)\) is \(\bs T_{{K+1},{K+1}}\), the DG approximation of the transition of point mass to density is \(\sum\limits_{i\in\calS_{K+1}}q_{{K+1},i}(t)T_{ij}\boldsymbol\phi_K(b)\), \(j\in\calS_-\) and the approximation to reflection of density at the upper boundary is \(- \sum_{i\in\calS_+}c_i\widehat p_{ij} \bs a_{0,i}(t)\bs F_j^{0,0}\).
 
@@
 \[{{\bs B}} = \left[\begin{array}{llllll}
-	\bs T_{{-1},{-1}}& {{\bs B}}^{{-1}0} & & & & \\
-	{{\bs B}}^{0{-1}} & {{\bs B}}^{00} & {{\bs B}}^{01} & & & \\
-	& {{\bs B}}^{10} & {{\bs B}}^{11} & {{\bs B}}^{12} & & \\
+	\bs T_{{-1},{-1}}& {{\bs B}}^{{-1},0} & & & & \\
+	{{\bs B}}^{0,{-1}} & {{\bs B}}^{0,0} & {{\bs B}}^{0,1} & & & \\
+	& {{\bs B}}^{1,0} & {{\bs B}}^{1,1} & {{\bs B}}^{1,2} & & \\
 	& & \ddots & \ddots & \ddots & \\
@@
 \begin{align*}
-	{{\bs B}}^{k\ell} &= \ddot{{\bs B}}^{k\ell}, \mbox{ for \(k\in\mathcal K^\circ\), \(\ell\in\{k-1,k,k+1\}\), \(k=\ell,\, k,\ell\notin \{0,K\}\),}
-	\\ \bs B^{00} &= \ddot{\bs B}^{00} - \left[c_ip_{ij}^{-1} 1(c_i<0, c_j>0)\right]_{i\in\calS,j\in\calS}\otimes \bs F_-^{00},
-	\\ \bs B^{KK} &= \ddot{\bs B}^{KK} + \left[c_ip_{ij}^{-1} 1(c_i>0, c_j<0)\right]_{i\in\calS,j\in\calS}\otimes \bs F_+^{00}, 
-	\\ {{\bs B}}^{{-1}0} &:= \bs T_{{-1}+}\otimes \boldsymbol\phi^0(0), 
-	\\ {{\bs B}}^{0{-1}} &:=-\left[c_ip_{ij}^{-1} 1{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}} \otimes \boldsymbol \phi^0(0)\tr{}, 
-	\\ {{\bs B}}^{{K+1} K} &:= \bs T_{{K+1}-}\otimes \boldsymbol\phi^K(b),
-	\\ {{\bs B}}^{K,{K+1}} &:= \left[c_ip_{ij}^{K+1} 1{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes \boldsymbol \phi^K(b)\tr{},
+	{{\bs B}}^{k,\ell} &= \ddot{{\bs B}}^{k,\ell}, \mbox{ for \(k\in\mathcal K^\circ\), \(\ell\in\{k-1,k,k+1\}\), \(k=\ell,\, k,\ell\notin \{0,K\}\),}
+	\\ \bs B^{0,0} &= \ddot{\bs B}^{0,0} - \left[c_i\widecheck p_{ij}1(c_i<0, c_j>0)\right]_{i\in\calS,j\in\calS}\otimes \bs F_-^{0,0},
+	\\ \bs B^{K,K} &= \ddot{\bs B}^{K,K} + \left[c_i\widecheck p_{ij} 1(c_i>0, c_j<0)\right]_{i\in\calS,j\in\calS}\otimes \bs F_+^{0,0}, 
+	\\ {{\bs B}}^{{-1},0} &= \bs T_{{-1},+}\otimes \boldsymbol\phi^0(0), 
+	\\ {{\bs B}}^{0,{-1}} &=-\left[c_i \widecheck p_{ij}  1{(c_i<0)}\right]_{i\in\calS,j\in\calS_{{-1}}} \otimes \boldsymbol \phi^0(0)\tr{}, 
+	\\ {{\bs B}}^{{K+1}, K} &= \bs T_{{K+1},-}\otimes \boldsymbol\phi^K(b),
+	\\ {{\bs B}}^{K,{K+1}} &= \left[c_i\widehat p_{ij}  1{(c_i>0)}\right]_{i\in\calS,j\in\calS_{K+1}} \otimes \boldsymbol \phi^K(b)\tr{},
 \end{align*} 
@@
 
-For future reference, we also define the matrices \({{\bs B}}^{k\ell}_{ij}\) for \(k\in\{2,\dots,K-1\}\), \(\ell\in\{k-1,k\},\) \(i,j\in\calS\), by
+For future reference, we also define the matrices \({{\bs B}}^{k,\ell}_{ij}\) for \(k\in\{2,\dots,K-1\}\), \(\ell\in\{k-1,k\},\) \(i,j\in\calS\), by
 \begin{align*}
-	{{\bs B}}^{kk}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{kk}+\bs G_k)\bs M_k^{-1} & i=j,\\T_{ij} \bs I_{p_k}& i\neq j,\end{cases}
-\\	{{\bs B}}^{k\ell}_{ij} &= \begin{cases}c_i\bs F_i^{k\ell}\bs M_\ell^{-1} & i=j,\\\bs 0 & i\neq j,\end{cases}\quad \ell \in \{k-1,k\}
+	{{\bs B}}^{k,k}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{k,k}+\bs G_k)\bs M_k^{-1} & i=j,\\T_{ij} \bs I_{p_k}& i\neq j,\end{cases}
+\\	{{\bs B}}^{k,\ell}_{ij} &= \begin{cases}c_i\bs F_i^{k,\ell}\bs M_\ell^{-1} & i=j,\\\bs 0 & i\neq j,\end{cases}\quad \ell \in \{k-1,k\}
 \end{align*}
@@
 \begin{align*}
-	{{\bs B}}^{00}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{00}+\bs G_0)\bs M_0^{-1} & i=j,\\T_{ij}\bs I_{p_k} - 1(c_i<0,c_j>0)c_ip_{ij}^{-1} \bs F_i^{00}\bs M_0^{-1} & i\neq j,\end{cases}
-\\	{{\bs B}}^{KK}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{KK}+\bs G_K)\bs M_K^{-1} & i=j,\\T_{ij}\bs I_{p_K} + 1(c_i>0,c_j<0)c_ip_{ij}^{K+1} \bs F_i^{KK}\bs M_K^{-1} & i\neq j.\end{cases}
+	{{\bs B}}^{0,0}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{0,0}+\bs G_0)\bs M_0^{-1} & i=j,\\T_{ij}\bs I_{p_k} - 1(c_i<0,c_j>0)c_i\widecheck p_{ij} \bs F_i^{0,0}\bs M_0^{-1} & i\neq j,\end{cases}
+\\	{{\bs B}}^{K,K}_{ij} &= \begin{cases}T_{ij}\bs I_{p_k} + c_i(\bs F_i^{K,K}+\bs G_K)\bs M_K^{-1} & i=j,\\T_{ij}\bs I_{p_K} + 1(c_i>0,c_j<0)c_i\widehat p_{ij} \bs F_i^{K,K}\bs M_K^{-1} & i\neq j.\end{cases}
 \end{align*}
@@
 
-Let \(\boldsymbol a_{-1}(t):=\boldsymbol q_{-1}(t)\) and \(\boldsymbol a_{K+1}(t):=\boldsymbol q_{K+1}(t)\), and define basis functions \(\bs\phi_{-1}(x) = \phi_{-1}^1(x) = \delta(x)\) and \(\bs\phi_{K+1}(x) = \phi_{K+1}^1(x) = \delta(x-b)\), where \(\delta\) is the Dirac delta, \(p_{-1} = p_{K+1} = 1\), and \(\mathcal N_{-1} = \mathcal N_{K+1} = \{1\}\).
+Let \(\boldsymbol a_{-1}(t)=\boldsymbol q_{-1}(t)\) and \(\boldsymbol a_{K+1}(t)=\boldsymbol q_{K+1}(t)\), and define basis functions \(\bs\phi_{-1}(x) = \phi_{-1}^1(x) = \delta(x)\) and \(\bs\phi_{K+1}(x) = \phi_{K+1}^1(x) = \delta(x-b)\), where \(\delta\) is the Dirac delta, \(p_{-1} = p_{K+1} = 1\), and \(\mathcal N_{-1} = \mathcal N_{K+1} = \{1\}\).
 Also define \({\bs M}_{-1}=\bs I_{|\calS_{-1}|}\), \({\bs M}_{K+1}=\bs I_{|\calS_{K+1}|}\), the block-diagonal matrix \(\bs M=\mathrm{diag}(\bs M_k,k\in\mathcal K)\), and row-vectors
@@
 
-To pose the approximation \(\bs B\) in kernel form let \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in W,\,i\in\calS\) be the initial density of the process, and \(\boldsymbol \phi(x)\bs b_i\tr{}\in W,\,i\in\calS\) be a test function.
+To pose the approximation \(\bs B\) in kernel form let \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in U,\,i\in\calS\) be the initial density of the process, and \(\boldsymbol \phi(x)\bs b_i\tr{}\in U,\,i\in\calS\) be a test function.
 Observe that, from our DG construction earlier and the definition of \({\bs M}\),
@@
 Essentially, the operator \(\mathbb R^k\) takes an initial measure \(\boldsymbol \mu_k\) and multiplies each element by \(1/|r_i(x)|\) on cells \(\calD_k\) where \(r_i(x)\neq 0\).
-In the context of DG the initial distribution is given by \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in W,\,i\in\calS\).
+In the context of DG the initial distribution is given by \(\boldsymbol a_i \boldsymbol \phi(x)\tr{}\in U,\,i\in\calS\).
 Thus, for \(k\in\mathcal K\) such that \(r_i(x)\neq0\) on \(\calD_k\), we have
 \[\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}\mathbb R^k_i = \cfrac{\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}}{|r_i(x)|}.\]
-Decompose the right-hand side into a component which lies in \(W\) and another orthogonal to \(W\): 
-\[\cfrac{\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}}{|r_i(x)|} = \bs \rho_{k,i} \bs \phi_k(x)\tr{} + g_i^\perp(x),\] where \(\bs \rho_{k,i} \bs\phi_k(x)\tr{}\in W\), \(g_i^\perp \in W^\perp\).
+Decompose the right-hand side into a component which lies in \(U\) and another orthogonal to \(U\): 
+\[\cfrac{\boldsymbol a_{k,i} \boldsymbol \phi_k(x)\tr{}}{|r_i(x)|} = \bs \rho_{k,i} \bs \phi_k(x)\tr{} + g_i^\perp(x),\] where \(\bs \rho_{k,i} \bs\phi_k(x)\tr{}\in U\), \(g_i^\perp \in U^\perp\).
 Now, multiply by test functions \(\{\phi^r_k(x)\}_{r=1}^{p_k}\) and integrate over \([0,b]\):
@@
 \end{align*}
-since \(g_i(x)^\perp\in W^\perp\). Define the matrix \(\bs M_k^r := \displaystyle\int_{x\in[0,b]} \cfrac{ \boldsymbol \phi_k(x)\tr{}\bs \phi_k(x)}{|r_i(x)|}\wrt x\), then 
+since \(g_i(x)^\perp\in U^\perp\). Define the matrix \(\bs M_k^r = \displaystyle\int_{x\in[0,b]} \cfrac{ \boldsymbol \phi_k(x)\tr{}\bs \phi_k(x)}{|r_i(x)|}\wrt x\), then 
 \(
@@
 	%
-	\( \bbxi_{k,i} \) & \(i\in\calS_k^-,\,k\in\mathcal K^-\)  & \(\bs \xi_{k,i} := (\xi_{k,i}^r)_{r\in\mathcal N_k}\) & 
+	\( \bbxi_{k,i} \) & \(i\in\calS_k^-,\,k\in\mathcal K^-\)  & \(\bs \xi_{k,i} = (\xi_{k,i}^r)_{r\in\mathcal N_k}\) & 
 	\(%\begin{array}{c}
 	\bbxi_{k,i}(\wrt x)\approx  \bs{\xi}_{k,i} \bs \phi^k(x)\tr{}\wrt x,
-	 %\\\bbxi_i^{-1}(\{0\}) %:=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
+	 %\\\bbxi_i^{-1}(\{0\}) %=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
 	%\approx \bs\xi_{i}^{-1},
-	%\\\bbxi_i^{K+1}(\{b\}) %:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]
+	%\\\bbxi_i^{K+1}(\{b\}) %= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]
 	%\approx \bs\xi_{i}^{K+1}.%\end{array}
@@
 %
-	\(\mathbb p_{k,i}\) & \(\begin{array}{c}i\in\calS_k^-\cup\calS_k^0,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_m\end{array}\) & \(\bs p_{k,i} := (p_{k,i}^r)_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
-	\mathbb p_{k,i}(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+	\(\mathbb p_{k,i}\) & \(\begin{array}{c}i\in\calS_k^-\cup\calS_k^0,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_m\end{array}\) & \(\bs p_{k,i} = (p_{k,i}^r)_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
+	\mathbb p_{k,i}(\wrt x) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
 	\approx\bs{p}_{k,i} \bs\phi_k(x)\tr{}\wrt x
-	%\\ \mathbb p_{i}^{-1}(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
-	%\approx \bs p_{i}^{-1},\\ \mathbb p_{i}^{K+1}(\{b\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = b, \varphi_{t} = i\right]
+	%\\ \mathbb p_{i}^{-1}(\{0\}) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t=0, X_{t} = 0, \varphi_{t} = i\right]
+	%\approx \bs p_{i}^{-1},\\ \mathbb p_{i}^{K+1}(\{b\}) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t=0, X_{t} = b, \varphi_{t} = i\right]
 	%\approx \bs p_{i}^{K+1}.% \end{array}
@@
 %
-	 \(\bbpi_{k,i}(y)\)  & \(\begin{array}{c}i\in\calS,\\k\in\mathcal K\end{array}\) & \(\bs \pi_{k,i}(y) := (\pi_{k,i}^r(y))_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
-	 \bbpi_{k,i}(y)(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+	 \(\bbpi_{k,i}(y)\)  & \(\begin{array}{c}i\in\calS,\\k\in\mathcal K\end{array}\) & \(\bs \pi_{k,i}(y) = (\pi_{k,i}^r(y))_{r\in\mathcal N_k}\) & \(%\begin{array}{c}
+	 \bbpi_{k,i}(y)(\wrt x) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
 	\approx\bs{\pi}_{k,i}(y) \bs \phi_k(x)\tr{}\wrt x
-	%\\ \bbpi_{i}^{-1}(y)(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
-	%\approx \bs \pi_{i}^{-1}(y),\\ \bbpi_{i}^{K+1}(y)(\{b\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = b, \varphi_{t} = i\right]
+	%\\ \bbpi_{i}^{-1}(y)(\{0\}) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t=0, X_{t} = 0, \varphi_{t} = i\right]
+	%\approx \bs \pi_{i}^{-1}(y),\\ \bbpi_{i}^{K+1}(y)(\{b\}) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t=0, X_{t} = b, \varphi_{t} = i\right]
 	%\approx \bs \pi_{i}^{K+1}(y). \end{array}
@@
 %	%
-%	\( \bbxi_{k,i}(\wrt x) \) & \(\xi_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\mathcal K_i^-\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(%:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\\
+%	\( \bbxi_{k,i}(\wrt x) \) & \(\xi_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\mathcal K_i^-\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(%= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} \in \wrt x, \varphi_{\theta_n} = i\right]\\
 %	 \displaystyle\sum\limits_{r=1}^{N_k} {\xi}_{k,i}^r \phi^r_k(x)\wrt x,\,x\in\calD_k.\)\\\hline
 %
-%	&\(\xi_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\mathcal K_i^-\cap \{{-1}\cup{K+1}\}.\end{array}\) & \(\begin{array}{l}\bbxi_{k,i}(\{0\}) %:=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
+%	&\(\xi_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\mathcal K_i^-\cap \{{-1}\cup{K+1}\}.\end{array}\) & \(\begin{array}{l}\bbxi_{k,i}(\{0\}) %=\lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = 0, \varphi_{\theta_n} = i\right]
 %	\approx \xi_{i,r}^{-1},
-%	\\\bbxi_{k,i}(\{b\}) %:= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]
+%	\\\bbxi_{k,i}(\{b\}) %= \lim\limits_{n\to\infty}\mathbb{P}\left[X_{\theta_n} = b, \varphi_{\theta_n} = i\right]
 %	\approx \xi_{i,r}^{K+1}.\end{array}\)\\\hline 
 %
-%	&\(p_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(\mathbb p_{k,i}(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+%	&\(p_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(\mathbb p_{k,i}(\wrt x) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
 %	\approx\displaystyle\sum\limits_{r=1}^{N_k} {p}_{k,i}^r \phi^r_k(x)\wrt x,\,x\in\calD_k.\)\\\hline
 %
-%	&\(p_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\}.\end{array}\) & \(\begin{array}{l} \mathbb p_{k,i}(\{0\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]
-%	\approx p_{i,r}^{-1},\\ \mathbb p_{k,i}(\{b\}) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = b, \varphi_{t} = i\right]
+%	&\(p_{k,i}^r\) & \(\begin{array}{l}i\in\calS,\,r=1,\\k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\}.\end{array}\) & \(\begin{array}{l} \mathbb p_{k,i}(\{0\}) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t=0, X_{t} = 0, \varphi_{t} = i\right]
+%	\approx p_{i,r}^{-1},\\ \mathbb p_{k,i}(\{b\}) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t=0, X_{t} = b, \varphi_{t} = i\right]
 %	\approx p_{i,r}^{K+1}. \end{array}\) \\\hline
 %
-%	&\(\pi_{k,i}^r(y)\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(\bbpi_{k,i}(y)(\wrt x) %:= \lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
+%	&\(\pi_{k,i}^r(y)\) & \(\begin{array}{l}i\in\calS,\,r\in\{1,...,N_k\},\\k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\setminus \{{-1}\cup{K+1}\}.\end{array}\) & \(\bbpi_{k,i}(y)(\wrt x) %= \lim\limits_{t\to\infty}\mathbb{P}\left[W_t \in \wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\\
 %	\approx\displaystyle\sum\limits_{r=1}^{N_k} {\pi}_{k,i}^r(y) \phi^r_k(x)\wrt x,\,y>0,\, x\in\calD_k.\)\\\hline
@@
 	%
-	{\bs{\xi}}_k &:= ( \bs\xi_{k,i})_{i\in\calS_k^-}, \quad  {k\in\mathcal K_i^-},
+	{\bs{\xi}}_k &= ( \bs\xi_{k,i})_{i\in\calS_k^-}, \quad  {k\in\mathcal K_i^-},
 	%
-	\\ {\bs{\xi}}& := ( \bs \xi_k)_{k\in\mathcal K^-},
+	\\ {\bs{\xi}}& = ( \bs \xi_k)_{k\in\mathcal K^-},
 	% 
@@
 	%
-	\\\bs{p}^{m}_k &:= (  \bs p_{k,i})_{i\in\calS_k^m}, \quad  k\in{\mathcal K^m},\, m\in\{-,0\},
+	\\\bs{p}^{m}_k &= (  \bs p_{k,i})_{i\in\calS_k^m}, \quad  k\in{\mathcal K^m},\, m\in\{-,0\},
 	%
-	\\ {\bs{p}^m} &:= (  \bs p^{m}_k)_{k\in\mathcal K^m},\quad m\in\{-,0\},
+	\\ {\bs{p}^m} &= (  \bs p^{m}_k)_{k\in\mathcal K^m},\quad m\in\{-,0\},
 	%
-	\\ {\bs{p}} &:= (  \bs p^m)_{m\in\{-,0\}},
+	\\ {\bs{p}} &= (  \bs p^m)_{m\in\{-,0\}},
 	%
@@
 	%
-	\\{\bs{\pi}}^{k}_m(y) &:= (  \bs \pi_{k,i}(y))_{i\in\calS_k^m},\quad  k\in\mathcal K,\,m\in\{+,-,0\},
+	\\{\bs{\pi}}^{k}_m(y) &= (  \bs \pi_{k,i}(y))_{i\in\calS_k^m},\quad  k\in\mathcal K,\,m\in\{+,-,0\},
 	%
-	\\{\bs{\pi}}^m(y) &:= (\bs \pi^{k}_m(y))_{k\in\mathcal K^m},\quad  m\in\{+,-,0\},
+	\\{\bs{\pi}}^m(y) &= (\bs \pi^{k}_m(y))_{k\in\mathcal K^m},\quad  m\in\{+,-,0\},
 	%
-	\\ {\bs{\pi}}(y) &:= (  \bs \pi^m(y))_{m\in\{+,-,0\}}.
+	\\ {\bs{\pi}}(y) &= (  \bs \pi^m(y))_{m\in\{+,-,0\}}.
  \end{align*}
@@
 	% 
-	where $ {\bs K} :=  {\bs D}^{++}(0) + \bs \Psi(0) {\bs D}^{(-+)}(0)$, and $z$ is a normalising constant.
+	where $ {\bs K} =  {\bs D}^{++}(0) + \bs \Psi(0) {\bs D}^{(-+)}(0)$, and $z$ is a normalising constant.
 
@@
 %\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}}p_{k,i}^r \phi^r_k(x),\quad x \in \calD_k,\]
-%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
-For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\},\) define point masses \(p_{k,i}^r\) such that \(p_{i,r}^{-1}\) and \(p_{i,r}^{K+1}\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t=0, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t = 0, X_{t} = b, \varphi_{t} = i\right]\), respectively.
+%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[W_t = 0, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\},\) define point masses \(p_{k,i}^r\) such that \(p_{i,r}^{-1}\) and \(p_{i,r}^{K+1}\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[W_t=0, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[W_t = 0, X_{t} = b, \varphi_{t} = i\right]\), respectively.
 In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\), let \( {\bs{p}}_{k,i} = (  p_{k,i}^r)_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m\), let \( {\bs{p}}^k = (  p_{k,i})_{\{i\in\calS\mid r_i(x)\leq 0,x\in\calD_k\}}\) and let \( {\bs{p}} = (  p^k)_{k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{-,0\}}\mathcal K_i^m}\).
@@
 %%	
-%%	For \(i\in\calS_{-1}, m\in\{-,0\}\) define \(  p^m_{{-1},i}\) as the approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, Y_t=0, \varphi_{t} = i\right]\).
+%%	For \(i\in\calS_{-1}, m\in\{-,0\}\) define \(  p^m_{{-1},i}\) as the approximation to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, W_t=0, \varphi_{t} = i\right]\).
 %%	
@@
 %%	Now define \( {\bs p}^m_{{-1}},\, {\bs p}^m_{\circ},\, {\bs p}^m_{{K+1}}\) as the approximation to \(\bs p\) at the left-hand boundary, interior, and on the right-hand boundary of \([0,b]\), respectively.
-%%	\[ {\bs p}^m := \vligne{ {\bs p}^m_{-1}(y) &  {\bs p}^m_\circ(y) &  {\bs p}^m_{K+1}(y)},\quad m\in\{-,0\},\]
+%%	\[ {\bs p}^m = \vligne{ {\bs p}^m_{-1}(y) &  {\bs p}^m_\circ(y) &  {\bs p}^m_{K+1}(y)},\quad m\in\{-,0\},\]
 %%where Specifically,  
@@
 %\[\displaystyle\sum\limits_{r\in\{1,...,N_k\}}\pi_{k,i}^r(y) \phi^r_k(x),\quad x \in \calD_k,\]
-%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in\wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
-For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\}\) define point masses \(\pi_{k,i}^r(y)\) such that \(\pi_{i,r}^{-1}(y)\) and \(\pi_{i,r}^{K+1}(y)\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t\in\wrt y, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[Y_t \in\wrt y, X_{t} = b, \varphi_{t} = i\right]\), respectively.
+%is an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[W_t \in\wrt y, X_{t} \in \wrt x, \varphi_{t} = i\right]\) on \(x\in\calD_k\).
+For \(i,k,r\) such that \(i\in\calS,\,r=1,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\cap \{{-1}\cup{K+1}\}\) define point masses \(\pi_{k,i}^r(y)\) such that \(\pi_{i,r}^{-1}(y)\) and \(\pi_{i,r}^{K+1}(y)\) approximate the point masses \(\lim\limits_{t\to\infty}\mathbb{P}\left[W_t\in\wrt y, X_{t} = 0, \varphi_{t} = i\right]\) and \(\lim\limits_{t\to\infty}\mathbb{P}\left[W_t \in\wrt y, X_{t} = b, \varphi_{t} = i\right]\), respectively.
 In vector form, for \(i,k\) such that \(i\in\calS,\,k\in\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\), let \( {\bs{\pi}}_{k,i}(y) = (  \pi_{k,i}^r(y))_{r\in\{1,...,N_k\}}\), for \(k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m\), let \( {\bs{\pi}}^k(y) = (  \pi_{k,i}(y))_{\{i\in\calS\mid r_i(x)\leq 0,x\in\calD_k\}}\) and let \( {\bs{\pi}}(y) = (  \pi^k(y))_{k\in\bigcup\limits_{i\in\calS}\bigcup\limits_{m\in\{+,-,0\}}\mathcal K_i^m}\).
@@
 %	% 
-%	where $ {K} :=  {D}^{++}(0) + \Psi(0) {D}^{(-+)}(0)$, and $z$ is a normalising constant.
+%	where $ {K} =  {D}^{++}(0) + \Psi(0) {D}^{(-+)}(0)$, and $z$ is a normalising constant.
 %	%
-%%Also \( {\bs p}^-_{{K+1}}=(  p^-_{{K+1},i})_{\{i\in\calS_{K+1}\mid r_i(b)<0\}}\) and \( {\bs p}^0_{{K+1}}=(  p^0_{{K+1},i})_{\{i\in\calS_{K+1}\mid r_i(b)=0\}}\) where \(  p^m_{{K+1},i},\,m\in\{-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =b, Y_t=0, \varphi_{t} = i\right]\).
+%%Also \( {\bs p}^-_{{K+1}}=(  p^-_{{K+1},i})_{\{i\in\calS_{K+1}\mid r_i(b)<0\}}\) and \( {\bs p}^0_{{K+1}}=(  p^0_{{K+1},i})_{\{i\in\calS_{K+1}\mid r_i(b)=0\}}\) where \(  p^m_{{K+1},i},\,m\in\{-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =b, W_t=0, \varphi_{t} = i\right]\).
 On the interior, we have
 %%\( {\bs p}^m_{\circ}=( {\bs p}_{k,i})_{i\in\calS,k\in\mathcal K_i^m}\) where \( {\bs p}_{k,i}=( {p}_{k,i}^r)_{r=1,...,N_k}\).
-On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, Y_t=0, \varphi_{t} = i\right]\) is given by \( {\bs p}_{k,i}\bs \phi^k(x)\tr{}\).
+On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, W_t=0, \varphi_{t} = i\right]\) is given by \( {\bs p}_{k,i}\bs \phi^k(x)\tr{}\).
 %%
@@
 %%%
-%%Specifically, \( {\bs \pi}^+_{{-1}}(y)=(  \pi^+_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)>0\}}\), \( {\bs \pi}^-_{{-1}}(y)=(  \pi^-_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)<0\}}\) and \( {\bs \pi}^0_{{-1}}(y)=(  \pi^0_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)=0\}}\) where \(  \pi^m_{{-1},i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, Y_t=\wrt y, \varphi_{t} = i\right]\).
+%%Specifically, \( {\bs \pi}^+_{{-1}}(y)=(  \pi^+_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)>0\}}\), \( {\bs \pi}^-_{{-1}}(y)=(  \pi^-_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)<0\}}\) and \( {\bs \pi}^0_{{-1}}(y)=(  \pi^0_{{-1},i}(y))_{\{i\in\calS_{-1}\mid r_i(0)=0\}}\) where \(  \pi^m_{{-1},i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =0, W_t=\wrt y, \varphi_{t} = i\right]\).
 %%%
-%%Also \( {\bs \pi}^+_{{K+1}}(y)=(  \pi^+_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)>0\}}\), \( {\bs \pi}^-_{{K+1}}(y)=(  \pi^-_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)<0\}}\) and \( {\bs \pi}^0_{{K+1}}(y)=(  \pi^0_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)=0\}}\) where \(  \pi^m_{{K+1},i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =b, Y_t\in\wrt y, \varphi_{t} = i\right]\).
+%%Also \( {\bs \pi}^+_{{K+1}}(y)=(  \pi^+_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)>0\}}\), \( {\bs \pi}^-_{{K+1}}(y)=(  \pi^-_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)<0\}}\) and \( {\bs \pi}^0_{{K+1}}(y)=(  \pi^0_{{K+1},i}(y))_{\{i\in\calS_{K+1}\mid r_i(b)=0\}}\) where \(  \pi^m_{{K+1},i}(y),\,m\in\{+,-,0\}\) are approximations to \( \lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =b, W_t\in\wrt y, \varphi_{t} = i\right]\).
 %%%
@@
 %%\( {\bs \pi}^m_{\circ}(y)=( {\bs \pi}_{k,i}(y))_{i\in\calS,k\in\mathcal K_i^m}\) where \( {\bs \pi}_{k,i}(y)=( {\pi}_{k,i}^r(y))_{r=1,...,N_k}\).
-On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, Y_t\in\wrt y, \varphi_{t} = i\right]\) is given by \( {\bs \pi}_{k,i}(y)\bs \phi^k(x)\tr{}\).
+On \(\mathcal D_k\) an approximation to the density of \(\lim\limits_{t\to\infty}\mathbb{P}\left[X_{t} =\wrt x, W_t\in\wrt y, \varphi_{t} = i\right]\) is given by \( {\bs \pi}_{k,i}(y)\bs \phi^k(x)\tr{}\).
 
@@
 Thus, for the application to fluid-fluid queues, slope limiting is not a viable option.
-%We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{Y(t)\}\) first returns to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation.
+%We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{W(t)\}\) first returns to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation.
 The application of the limiter/filter is dependent on the initial condition, so the post-processing needs to be done for every initial condition under consideration.

### 2023-01-22 chapter5/chapter.tex
.
post-reviews, pre supervisor feedback

@@
 To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed.
-Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the orbit restart epochs.
+Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\rapphase(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the level and phase of the discrete-time process embedded at the orbit restart epochs.
 The subscript \(\bs \alpha \) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\), unless the process is temporarily absorbed in a sticky boundary, in which case \(\bs A^{(p)}(\tau_n^{(p)})=1\).
@@
 % \begin{align}
-% 	&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1(j\in\calS_-)+1(j\in\calS_+), \varphi(\tau_1^{(p)}) = j
+% 	&\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1(j\in\calS_-)+1(j\in\calS_+), \rapphase(\tau_1^{(p)}) = j
 % 			 \mid \bs Y^{(p)}(0) = (\ell_0 ,\bs\alpha, i)).\nonumber
 % 	%  \intertext{for \(j\in\calS_+\)}
-% 	%  &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \varphi(\tau_1^{(p)}) = j
+% 	%  &\mathbb P(L^{(p)}(\tau_1^{(p)}) = \ell_0-1, \rapphase(\tau_1^{(p)}) = j
 % 	% 		 \mid \bs Y^{(p)}(0) = \bs y_0^{(p)}) \nonumber
@@
 To simplify notation, we may drop the superscript \(p\) where it is not explicitly needed.
-Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\varphi^{(p)}(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the value of the level and phase of the discrete-time process embedded at the change-of-level epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
+Further, let \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z} = \{(L^{(p)}(\tau_n^{(p)}),\rapphase^{(p)}(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\) be the value of the level and phase of the discrete-time process embedded at the change-of-level epochs, \(\{\tau_n^{(p)}\}_{n\geq 0}\).
 The subscript \(\bs \alpha^{(p)}\) refers to the fact that \(\bs A^{(p)}(\tau_n^{(p)})=\bs \alpha^{(p)}\) for \(n\geq 1\).
@@
 by the time homogenous property of the QBD-RAP, the memoryless property of the exponential distribution, and the convolution theorem of Laplace transforms.
+The exponential random variable in the last line of (\ref{eqn: l whkrqvlkjbrd}) appears due to the interpretation of the Laplace transform as probability that \(\tau_n^{(p)}\) occurs before and exponential random variable, \(E^\lambda\).
 The swap of integrals and sums is justified by the Fubini-Tonelli Theorem.
@@
 	Hence, for sample paths which contribute to the Laplace transform, one of two things must happen, either; 
-	\begin{enumerate}[a]
+	\begin{enumerate}[a)]
 		\item the phase remains \(i\) until there is an orbit restart epoch and \(E^\lambda\) does not occur before the orbit restart epoch, or, \label{item: 1}
@@
 
-Let \(\mathcal R(L(t),\bs A(t),\varphi(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\varphi(t)}(x)\), \(x\in\calD_{L(t),\varphi(t)}\), then
-\begin{align}
-	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )  \mid \nonumber
+Let \(\mathcal R(L(t),\bs A(t),\rapphase(t))\) be the random variable with density function \(\bs A(t)\bs v_{L(t),\rapphase(t)}(x)\), \(x\in\calD_{L(t),\rapphase(t)}\), then
+\begin{align}
+	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L(t),\bs A(t),\rapphase(t)),\rapphase(t) )  \mid \nonumber
 	\bs Y (0)=\bs y_0 \Bigg] \wrt t 
-	% \\&=\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\int_{x\in\calD_{L(t),\varphi(t)}}\bs A(t)\bs v_{L(t),\varphi(t)}(x)\psi(x,\varphi(t)),\varphi(t) )\wrt x  \mid \nonumber
+	% \\&=\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\int_{x\in\calD_{L(t),\rapphase(t)}}\bs A(t)\bs v_{L(t),\rapphase(t)}(x)\psi(x,\rapphase(t)),\rapphase(t) )\wrt x  \mid \nonumber
 	% \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t 
@@
 	Let \(\psi: \mathbb R\times \mathcal S \to \mathbb R\) be an arbitrary, bounded function with \(|\psi(\cdot,\cdot)|\leq F\).
-%Further, let \(\mathcal L(X(t)) = \sum_{k\in\mathcal K}k1(X(t)\in\mathcal D_k)\) and \(\widetilde{\bs X}(t) = (\mathcal L(X(t)),X(t),\varphi(t)))\).
+%Further, let \(\mathcal L(X(t)) = \sum_{k\in\mathcal K}k1(X(t)\in\mathcal D_k)\) and \(\widetilde{\bs X}(t) = (\mathcal L(X(t)),X(t),\rapphase(t)))\).
 	For each \(i\in\mathcal S\), \(\ell_0\in\mathcal K\), \(x_0\in\mathcal D_{\ell_0,i}\),
 	\begin{align}
-		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )  \mid \nonumber
+		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi(\mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )  \mid \nonumber
 		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber
@@
 	\begin{align}
-		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid 
+		&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )   \mid 
 		  \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
-		\\&= \int_{t=0}^\infty e^{-\lambda t}\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )  1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
+		\\&= \int_{t=0}^\infty e^{-\lambda t}\sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )  1({L}^{(p)}(t)=\ell,\rapphase^{(p)}(t)=j))  \mid \nonumber
 		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t \nonumber 
-		\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[ \psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid 
+		\\&= \sum_{\ell\in\mathcal K}\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[ \psi( \mathcal R(\ell,\bs A^{(p)}(t),j),j )1({L}^{(p)}(t)=\ell,\rapphase^{(p)}(t)=j))  \mid 
 		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t, \label{eqnL AKJF cjk ajhJK}
@@
 	 \begin{align}
-	 	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j), j )\wrt x 1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
+	 	&\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j), j )\wrt x 1({L}^{(p)}(t)=\ell,\rapphase^{(p)}(t)=j))  \mid \nonumber
 		 \bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t
@@
 	\begin{align}
-		&\Bigg|\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),\varphi(t) )  1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
+		&\Bigg|\sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[\psi( \mathcal R(\ell,\bs A^{(p)}(t),j),\rapphase^{(p)}(t) )  1({L}^{(p)}(t)=\ell,\rapphase^{(p)}(t)=j))  \mid \nonumber
 		\bs Y^{(p)}(0)=\bs y_0^{(p)} \Bigg] \wrt t\Bigg| \nonumber 
 		%
-		\\&\leq F \sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[1({L}^{(p)}(t)=\ell,\varphi(t)=j))  \mid \nonumber
+		\\&\leq F \sum_{j\in\mathcal S}\int_{t=0}^\infty e^{-\lambda t}\mathbb E\Bigg[1({L}^{(p)}(t)=\ell,\rapphase^{(p)}(t)=j))  \mid \nonumber
 		\bs Y^{(p)}(0)=\bs y_0^{(p)}\Bigg] \wrt t \nonumber 
@@
 	\begin{align}
-			&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] \nonumber 
+			&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] \nonumber 
 			\to \mathbb E\left[\psi({\bs X}(t))\mid \bs X(0)=(x_0, i) \right] \nonumber
@@
 is a continuous function of \(t\) (it is a Feller semigroup), and, for \(p<\infty\), 
-\[\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \]
+\[\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \]
  is continuous in \(t\), the limit 
- \[\lim\limits_{p\to\infty}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) ) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \] need not be continuous in \(t\).
+ \[\lim\limits_{p\to\infty}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) ) \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \] need not be continuous in \(t\).
 Nonetheless, we can claim that
 \begin{align}
-		&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )    \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \nonumber
+		&\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )    \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right]  \nonumber
 		\to \mathbb E\left[\psi({\bs X}(t) )\mid \bs X(0)=(x_0, i) \right] \nonumber
@@
 A sufficient condition to upgrade the convergence from weak to point-wise (in the variable \(t\)) is to show that for \(t\geq 0\) 
-\[\sup_{p}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )  \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \leq M(t)<\infty\]
-and the sequence \(\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \) is eventually equicontinuous in \(t\).
+\[\sup_{p}\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )  \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \leq M(t)<\infty\]
+and the sequence \(\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right]  \) is eventually equicontinuous in \(t\).
 That is, for every \(\varepsilon>0\) there exists a \(\delta(t,\varepsilon)>0\) and a \(p_0(t,\varepsilon)\) such that \(|t-u|<\delta(t,\varepsilon)\) implies that
 \begin{align}
-	&\nonumber \Bigg|\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
-	\\&{}-\mathbb E\left[\psi( \mathcal R(L^{(p)}(u),\bs A^{(p)}(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| \nonumber 
+	&\nonumber \Bigg|\mathbb E\left[\psi( \mathcal R(L^{(p)}(t),\bs A^{(p)}(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
+	\\&{}-\mathbb E\left[\psi( \mathcal R(L^{(p)}(u),\bs A^{(p)}(u),\rapphase^{(p)}(u)),\rapphase^{(p)}(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| \nonumber 
 	<\varepsilon
@@
 % 	\begin{align}
-% 		&\Bigg|\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\varphi(t)),\varphi(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
-% 		\\&{}-\mathbb E\left[\psi( \mathcal R(L(u),\bs A(u),\varphi(u)),\varphi(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| < Lc_{max}|t-u| \nonumber
+% 		&\Bigg|\mathbb E\left[\psi( \mathcal R(L(t),\bs A(t),\rapphase^{(p)}(t)),\rapphase^{(p)}(t) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)} \right] 
+% 		\\&{}-\mathbb E\left[\psi( \mathcal R(L(u),\bs A(u),\rapphase^{(p)}(u)),\rapphase^{(p)}(u) )   \mid \bs Y^{(p)}(0)=\bs y_0^{(p)}\right] \Bigg| < Lc_{max}|t-u| \nonumber
 % 	\end{align}
@@
 % 	\begin{align*}
-% 		&\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\varphi(t+u)),\varphi(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
-% 		%
-% 		\\&=\mathbb E\left[\int_{x\in\calD_{L(t+u),\varphi(t+u)}}\bs A(t+u)\bs v_{L(t+u),\varphi(t+u)}\psi(x,\varphi(t+u) )\wrt x   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
-% 		%
-% 		\\&=\int_{L(t+u),\varphi(t+u)}\bs a e^{(c_i\bs S + T_{ii}\bs I)u}\bs v_{\ell,i}(x)\psi(x,i)\wrt x 
-% 		%
-% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
-% 		%
-% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
-% 		%
-% 		\\&\quad{}+\int_{L(t+u),\varphi(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}\bs v_{\ell+1,i}(x)\psi(x,i)\wrt x 
+% 		&\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\rapphase^{(p)}(t+u)),\rapphase^{(p)}(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \rapphase^{(p)}(t)=i \right]
+% 		%
+% 		\\&=\mathbb E\left[\int_{x\in\calD_{L(t+u),\rapphase^{(p)}(t+u)}}\bs A(t+u)\bs v_{L(t+u),\rapphase^{(p)}(t+u)}\psi(x,\rapphase^{(p)}(t+u) )\wrt x   \mid L(t)=\ell,\bs A(t)=\bs a, \rapphase^{(p)}(t)=i \right]
+% 		%
+% 		\\&=\int_{L(t+u),\rapphase^{(p)}(t+u)}\bs a e^{(c_i\bs S + T_{ii}\bs I)u}\bs v_{\ell,i}(x)\psi(x,i)\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_-}\int_{L(t+u),\rapphase^{(p)}(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
+% 		%
+% 		\\&\quad{}+\sum_{j\in\calS_+}\int_{L(t+u),\rapphase^{(p)}(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}T_{ij}\bs D e^{(c_j\bs S + T_{jj}\bs I)(u-v)}\bs v_{\ell,j}(x)\psi(x,j)\wrt v\wrt x 
+% 		%
+% 		\\&\quad{}+\int_{L(t+u),\rapphase^{(p)}(t+u)} \int_{v=0}^u\bs a e^{(c_i\bs S + T_{ii}\bs I)v}c_i\bs s\bs \alpha e^{(c_i\bs S + T_{ii}\bs I)(u-v)}\bs v_{\ell+1,i}(x)\psi(x,i)\wrt x 
 % 		%
@@
 % 	\begin{align*}
-% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\varphi(t+u)),\varphi(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \varphi(t)=i \right]
+% 		&\lim_{u\to 0^+}u^{-1}\Bigg(\mathbb E\left[\psi( \mathcal R(L(t+u),\bs A(t+u),\rapphase^{(p)}(t+u)),\rapphase^{(p)}(t+u) )   \mid L(t)=\ell,\bs A(t)=\bs a, \rapphase^{(p)}(t)=i \right]
 % 		\\&{}-\int_{x\in\calD_{\ell,i}}\bs a\bs v_{\ell,i}(x)\psi(x,i)\wrt x \Bigg)

### 2023-01-22 chapter6/chapter.tex
.
post-reviews, pre supervisor feedback

@@
 As we will see (in Section~\ref{sec: transient approx}), even if a fluid-queue is initialised with a smooth initial density, the boundary dynamics may induce transient discontinuities or non-smooth behaviour into the problem.
-A very specific set of conditions on the initial density and point masses must hold for the transient distribution to remain continuous as it evolves over time \citep{bo2014,bop2020}.
+A very specific set of conditions on the initial density and point masses must hold for the transient distribution to remain continuous as it evolves over time \citep{bo2014,boz2022}.
 Further, it is possible that discontinuities are present in the performance measures of fluid-fluid queues (see, for example, Section~\ref{sec: ffq num}).
@@
 Section~\ref{sec:stat} investigates approximations to the limiting distribution of the queue, Section~\ref{sec: transient approx} investigates approximations to the transient distribution of the queue for two initial conditions, and Section~\ref{sec: hit approx} investigates approximations the first hitting time of the fluid level on the set \(\{0,1\}\).
-Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num}, compute approximations to \(\mathbb \Psi\) and ultimately, approximate the distribution of \(\{(X(\zeta_{Y}(\{0\})),\varphi(\zeta_{Y}(\{0\})))\}\) at time \(\zeta_{Y}(\{0\})\) at which \(\{Y(t)\}\) first returns to \(0\).
+Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num}, compute approximations to \(\mathbb \Psi\) and ultimately, approximate the distribution of \(\{(X(\zeta_{W}(\{0\})),\varphi(\zeta_{W}(\{0\})))\}\) at time \(\zeta_{W}(\{0\})\) at which \(\{\dot W(t)\}\) first returns to \(0\).
+
+\paragraph{On the application of the QBD-RAP approximation to approximating operators of fluid-fluid queues.}
+A few remarks about applying the QBD-RAP approximation to approximate fluid queues and operators of fluid-fluid queues from \cite{bo2014} are in order.
+
+Despite the construction of the QBD-RAP in Chapter~\ref{sec: construction and modelling} and the proof of convergence to the fluid queue spanning Chapters~\ref{sec: conv} and \ref{ch: global results} being relatively technical, the actual computations involved are relatively simple once we have the generator of the QBD-RAP.
+Construction of the generator matrix of the QBD-RAP, \(\bs B\), is a matter of obtaining the parameters \(\bs\alpha,\,\bs S\) and \(\bs s\) of a concentrated matrix exponential distribution from \cite{hht2020}, computing the matrix \(\bs D\) as in the last part of Section~\ref{sec:kkkrjrjrjjrj}, then assembling the parameters into the tridiagonal form of the generator as in (\ref{eqn: BBBB})-(\ref{eqn: BBBB3}) and including boundary conditions as in (\ref{eqn: BBBBbndry})-(\ref{eqn: BBBBbndry2}).
+The initial vector of the QBD-RAP, \(\bs a=(\bs a_{k,i})_{i\in\calS,k\in\mathcal K}\), can be constructed as discussed in Section~\ref{sec: initial conditions}, perhaps using numerical integration depending on the choice of initial condition.
+After constructing the generator and initial condition, approximations to the fluid queue can be obtained relatively easily.
+For example, the expected orbit positions in each phase and level, \(\mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=i)]\), are found by computing
+\[(\mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=i)])_{i\in\calS, \ell \in\mathcal K}=\bs a \exp(\bs Bt),\]
+which, in this chapter, is computed numerically via the SSPRK integration scheme in Section~\ref{sec: time integration}.
+The approximation to the transient density is obtained from \(\mathbb E[\bs A(t)1(L(t)=\ell, \phi(t)=i)]\) as discussed in Section~\ref{sec: closing}.
+
+To approximate operators of fluid-fluid queues with the QBD-RAP  we need to approximate the operator \(\mathbb R\), defined in Lemma~\ref{lemma: D(s)}.
+Recall from Section~\ref{sec: approx r}, in the context of the DG scheme we could approximate \(\mathbb R\) as a projection.
+In the context of the QBD-RAP scheme it is less clear how one can approximate \(\mathbb R\), in general.
+If the rate functions \(r_i(x)\) are piecewise constant with \(r_i(x)=r_i^k\) on cell \(\calD_{k,i}\), then a suitable approximation for \(\mathbb R^k_i\) is \(1/|r_i^k|\bs I_{p}\), a diagonal matrix with \(1/|r_i^k|\) down the diagonal -- this is the case in the numerical experiments in Chapter~\ref{sec: numerics}.
+For more general cases we could consider approximating \(r_i(x)\) by a piecewise constant function, then apply the method above.
+
+Next, the approximation \(\bs \Psi\) is computed by solving the matrix Riccati equation 
+\begin{align}\label{eqn:RiccatiPsi2222aa}
+    \bs D^{+-}
++ \bs \Psi   \bs D^{-+}\bs \Psi
++   \bs D^{++}\bs \Psi
++ \bs \Psi  \bs D^{--}
+= \bs 0,
+\end{align}
+where 
+\[\bs D^{mn} = \bs{R}^m (\bs{B}^{mn} + \bs{B}^{m0}(\bs B^{00})^{-1}\bs{B}^{0n}), \, m,n\in\{+,-\},\]
+and \(\bs{B}^{mn},\, m,n\in\{+,-,0\}\) are matrices of parameters extracted from the QBD-RAP which approximate \(\mathbb B^{mn}\).
+For example, \(\bs B^{+-}\) is obtained by extracting all the elements of \(\bs B\) which correspond to starting in phase \(i\) in level \(k\) and ending in phase \(j\) in level \(\ell\) and have \(r_i(x)>0\) for \(x\in\calD_{k,i}\) and \(r_j(x)<0\) for \(x\in\calD_{\ell,j}\).
+Given the stochastic interpretation of the QBD-RAP scheme, and assuming the rate functions \(r_i(x)\) are piecewise constant with \(r_i(x)=r_i^k\) on each cell \(\calD_{k,i}\), the QBD-RAP approximation of the fluid-fluid queue forms a RAP-modulated fluid process \citep{p2019,bgnp2021}.
+Hence, the approximation \(\bs \Psi\) has a stochastic interpretation in terms of the first-return operator of the RAP-modulated fluid process.
+This is important.
+Equation (\ref{eqn:RiccatiPsi2222aa}) determines, but does not define \(\bs \Psi\).
+When we use the DG scheme to approximate \(\mathbb \Psi\) we still solve (\ref{eqn:RiccatiPsi2222aa}) (with the DG approximation to the generator \(\mathbb B\)), however, this does not define what the solution means in terms of the fluid-fluid queue (other than it is the solution to (\ref{eqn:RiccatiPsi2222aa})).
+In comparison, there is a rigorous definition of \(\bs \Psi\) when the QBD-RAP scheme is used in the approximation \citep{p2019,bgnp2021}.
+
+By Proposition~4.3 of \cite{p2019}, for the QBD-RAP scheme the approximation \(\bs \Psi\) has the property \(\bs \Psi \bs 1=\bs 1\) as long as the event that level process of the RAP-modulated fluid returns to \(0\) in finite time occurs with probability 1.
+
+Third, note that unlike the positivity preserving DG schemes which use slope limiting which are not linear operators in the presence of oscillations, the operator obtained via the QBD-RAP approximation scheme is linear.
+
+{Lastly, a comment on expected accuracy of the QBD-RAP scheme.} For \(p\) odd, \cite{hht2020} estimate that the variance of the class of CMEs decreases at order \(1/p^{2.14}\) asymptotically.
+Hence, we might expect the rate of convergence of the QBD-RAP scheme to be approximately order \(1/p^2\) for certain problems.
+This is in line with the positivity preserving DG scheme mentioned in~\ref{sec: anoth pos pres}.
+%The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}.
+Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform only marginally better than an approximation constructed with representations of order \(2p-1\).
+However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.
 
@@
 Section~\ref{sec:stat} investigates the ability of the schemes to approximate the limiting distribution of the model, Section~\ref{sec: transient approx} investigates the ability of the schemes to approximate the transient distribution of the model for two initial conditions, and Section~\ref{sec: return approx} investigates the ability of the schemes to approximate the first hitting time of the fluid level on the boundary of the interval \([0,1]\).
-Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num} and approximate the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{Y(t)\}\) first returns to \(0\).
+Lastly, we apply the approximation methods to a fluid-fluid queue in Section~\ref{sec: ffq num} and approximate the distribution of \(\{(X(t),\varphi(t))\}\) at the time at which \(\{W(t)\}\) first returns to \(0\).
 
@@
 		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun2/l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: another example} for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves of the non-linear normalised (orange) and normalised closing vectors are almost coincident.
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: another example} for the three closing vectors considered; unnormalised (green solid line), non-linear normalised (blue dashed line) and normalised (orange dash-dotted line).
+The error curves of the non-linear normalised and normalised closing vectors are almost coincident.
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}%To estimate the asymptotic rate of convergence, we fit a linear model of \(\log_{10}(error) = \beta_0 + \beta_1\log_{10}(Dim)\) to the the last 6 data points on the curves above, which corresponds to a rate of convergence \(error \propto N^{-\beta_1}\).
@@
 		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: al} for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves of the error curves of the non-linear normalised (orange) and normalised closing vectors are almost coincident.
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) of Example~\ref{ex: al} for the three closing vectors considered; unnormalised (green solid line), non-linear normalised (blue dashed line) and normalised (orange dash-dotted line).
+The error curves of the error curves of the non-linear normalised and normalised closing vectors are almost coincident.
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
@@
 	\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun4/pdfs_formatted.pdf}
-	\caption{Approximations to the PDF \(f(x)=1,\, x\in[0,1)\) (orange dotted line) for Example~\ref{ex: al}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
+	\caption{Approximations to the PDF \(f(x)=1,\, x\in[0,1)\) (blue dashed line) for Example~\ref{ex: al}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (green solid line) and normalised closing operator (orange dash-dotted line).}
 	\label{fig: pdf reconstructed}
@@
 		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error for the PDFs (right) of Example~\ref{ex: aasmm}, for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves for the non-linear normalised (orange) and normalised closing vectors are almost coincident.
+		\caption{KS error (left) and \(L^2\) error for the PDFs (right) of Example~\ref{ex: aasmm}, for the three closing vectors considered; unnormalised (green solid line), non-linear normalised (blue dashed line) and normalised (orange dash-dotted line).
+The error curves for the non-linear normalised and normalised closing vectors are almost coincident.
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
@@
 		\includegraphics[width=\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun6/pdfs_formatted.pdf}
-		\caption{Approximations to the PDF \(f(x)=-6x^2+6x\) (orange dotted line) for Example~\ref{ex: aasmm}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (blue solid line) and normalised closing operator (green dash-dotted line).}
+		\caption{Approximations to the PDF \(f(x)=-6x^2+6x\) (blue dashed line) for Example~\ref{ex: aasmm}, constructed using matrix exponential distributions of various dimensions, and using the unnormalised closing operator (green solid line) and normalised closing operator (orange dash-dotted line).}
 		\label{fig: pdf reconstructed quadratic}
@@
 		\includegraphics[width=0.5\textwidth,trim={0.5cm 0.8cm 0.2cm 1.25cm},clip]{chapter6/figs/qbdrap_closing_vec/fun7/l2_pdf_error_formatted.pdf}
-		\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: kjfk} for the three closing vectors considered; unnormalised (blue solid line), non-linear normalised (orange dashed line) and normalised (green dash-dotted line).
-The error curves for the non-linear normalised (orange) and normalised closing vectors are almost coincident.
+		\caption{KS error (left) and \(L^2\) error of the PDF (right) for Example~\ref{ex: kjfk} for the three closing vectors considered; unnormalised (green solid line), non-linear normalised (blue dashed line) and normalised (orange dash-dotted line).
+The error curves for the non-linear normalised and normalised closing vectors are almost coincident.
 The black dotted lines are linear least-squares fits to the last 8 data points and the slopes of the least square lines are written next to the last point.}
@@
 	\end{figure}
-First consider the initial condition with PDF \(1(x<1)\).
+First consider the initial condition with PDF \(1(x<1)\), a uniform distribution over \((0,1)\).
 The level of the fluid queue is uniformly distributed over the first cell.
@@
 	\begin{align}
-		\mu([0,x)) := \mathbb P(X(0)\leq x, \varphi(0)=1)=\cfrac{\Phi((x-2.5)/0.5)-\Phi(-2.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}, \label{eqn: kdjfksdf}
+		\mu([0,x)) = \mathbb P(X(0)\leq x, \varphi(0)=1)=\cfrac{\Phi((x-2.5)/0.5)-\Phi(-2.5/0.5)}{\Phi(7.5/0.5)-\Phi(-2.5/0.5)}, \label{eqn: kdjfksdf}
 	\end{align}
@@
 Recall, in the analysis of fluid-fluid queues in Section~\ref{sec: ffq intro}, we partition sample paths of the second fluid into periods where it is either increasing, decreasing, or constant.
-The position of \((\varphi(t),X(t))\) determines the rate at which the second fluid moves, hence the hitting times on the boundaries of the sets \(\mathcal F_i^+, \mathcal F_i^-,\) and \(\mathcal F_i^0\) determine the periods of time when the second fluid is either increasing, decreasing, or constant.
+The position of \((X(t),\varphi(t))\) determines the rate at which the second fluid moves, hence the hitting times on the boundaries of the sets \(\mathcal F_i^+, \mathcal F_i^-,\) and \(\mathcal F_i^0\) determine the periods of time when the second fluid is either increasing, decreasing, or constant.
 
@@
 \begin{model}\label{model: ffq}
-	Consider a stochastic fluid-fluid queue $\{(Y(t),X(t),\varphi(t))\}_{t\geq0},$ where $\{X(t)\}$ and $\{Y(t)\}$ represent the workloads in Buffers~1 and~2 at time $t \geq 0$, respectively, both driven by the phase $\{\varphi(t)\},$ which is a Markov chain on the state space $\mathcal{S} = \{11,10,01,00\}$.
-Both $\{X(t)\}$ and $\{Y(t)\}$ have a regulated boundary at 0.
+	Consider a stochastic fluid-fluid queue $\{(\dot W(t),X(t),\varphi(t))\}_{t\geq0},$ where $\{X(t)\}$ and $\{\dot W(t)\}$ represent the workloads in Buffers~1 and~2 at time $t \geq 0$, respectively, both driven by the phase $\{\varphi(t)\},$ which is a Markov chain on the state space $\mathcal{S} = \{11,10,01,00\}$.
+Both $\{X(t)\}$ and $\{\dot W(t)\}$ have a regulated boundary at 0.
 Here, the state $11$ indicates inputs to both buffers being \textnormal{\textsc{on}}, the state $00$ indicates both being \textnormal{\textsc{off}}, the state $10$ is when only the first input is \textnormal{\textsc{on}}, and the state $01$ is when only the second is \textnormal{\textsc{on}}.
@@
 	% 
-	and the net rates of change for $Y(t)$, denoted $r_i$, are as follows  
+	and the net rates of change for $\dot W(t)$, denoted $r_i$, are as follows  
 	%  
@@
 	(r_{11},r_{10},r_{01},r_{00})  & = \left\{ \begin{array}{lrrrll}  
-	% (\lambda_2 -\kappa, & \;\;\;\;0, & \lambda_2 - \kappa, &  \;\;\;\;0) & \text{if } X_t = 0,  Y_t = 0,\\
+	% (\lambda_2 -\kappa, & \;\;\;\;0, & \lambda_2 - \kappa, &  \;\;\;\;0) & \text{if } X_t = 0,  W_t = 0,\\
 	%  \vspace*{-0.3cm} \\
@@
 												\vspace*{-0.3cm} \\
-	%	(\lambda_2 -\theta_2,  &  0,  & \lambda_2 - \theta_2,  & \;\;\;\; 0) & \text{if } X_t \in (0,x^*), Y_t = 0,\\
+	%	(\lambda_2 -\theta_2,  &  0,  & \lambda_2 - \theta_2,  & \;\;\;\; 0) & \text{if } X_t \in (0,x^*), W_t = 0,\\
 	%											  \vspace*{-0.3cm} \\
@@
 		\end{align} 
-	We consider the initial distribution which is a point mass at \(Y(0)=0,\, X(0)=5,\, \varphi(0)=01\). 
+	We consider the initial distribution which is a point mass at \(W(0)=0,\, X(0)=5,\, \varphi(0)=01\). 
 \end{model}
@@
 Here, we assume that it is given by the solution to (\ref{eqn:RiccatiPsi2}) and is a polynomial approximation.
-%\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{Y}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}
+%\footnote{I believe that the resulting operator is a projection operator which, given an initial distribution, projects the distribution of \(X(\zeta_{W}(\{0\}))\) on to a set of polynomial basis functions, much like the DG method itself.}
 
@@
 \begin{align}\label{eqn: first return Y 1}
-	\mathbb P(X(\zeta_{Y}(\{0\}))\leq x, \varphi(\zeta_{Y}(\{0\}))=i\mid \bs X(0)\sim \mu),
-\end{align}
-where we recall \(\zeta_{Y}(E) = \inf\{t>0\mid Y(t)\in E\}\) is the first hitting time of \(\{Y(t)\}\) on the set \(E\).
-For Model~\ref{model: ffq}, it is only possible for the process \(Y(t)\) to return to \(0\) at time \(t\) when \((X(t),\varphi(t))\in[0,1.6)\times \{10,00\}\), so we evaluate the approximations over this region only.
+	\mathbb P(X(\zeta_{W}(\{0\}))\leq x, \varphi(\zeta_{W}(\{0\}))=i\mid \bs X(0)\sim \mu),
+\end{align}
+where we recall \(\zeta_{W}(E) = \inf\{t>0\mid \dot W(t)\in E\}\) is the first hitting time of \(\{\dot W(t)\}\) on the set \(E\).
+For Model~\ref{model: ffq}, it is only possible for the process \(\dot W(t)\) to return to \(0\) at time \(t\) when \((X(t),\varphi(t))\in[0,1.6)\times \{10,00\}\), so we evaluate the approximations over this region only.
 We use a grid of 10,001 points on which we evaluate the approximations of the CDF in each phase.
 
-For a `ground truth' comparison, we simulated \(5\times 10^{10}\) realisations of the fluid-fluid queue and recorded the value of \(\{X(t)\}\) and \(\{\varphi(t)\}\) at the time of first return of the second fluid, \(\{Y(t)\}\).
+For a `ground truth' comparison, we simulated \(5\times 10^{10}\) realisations of the fluid-fluid queue and recorded the value of \(\{X(t)\}\) and \(\{\varphi(t)\}\) at the time of first return of the second fluid, \(\{\dot W(t)\}\).
 The empirical approximation of (\ref{eqn: first return Y 1}) was then constructed, and error metrics for the difference between the empirical CDF and the approximations were computed.
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate CDFs of \(X(\zeta_{Y}(\{0\}))\) in (\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}.
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate CDFs of \(X(\zeta_{W}(\{0\}))\) in (\ref{eqn: first return Y 1}) for Model~\ref{model: ffq}.
 The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/cts/l1_cell_probs_error_formatted.pdf}%
-	\caption{Cell-wise errors between the simulated and approximate probabilities of \(X(\zeta_{Y}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}.
+	\caption{Cell-wise errors between the simulated and approximate probabilities of \(X(\zeta_{W}(\{0\}))\) residing on each cell \(\mathcal D_k\) or at the boundary for Model~\ref{model: ffq}.
 The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
@@
 	\end{align}
-	with an initial distribution which is a point-mass at \(Y(0)=0,\, X(0)=2, \varphi(0)=00\). 
+	with an initial distribution which is a point-mass at \(W(0)=0,\, X(0)=2, \varphi(0)=00\). 
 \end{model}
@@
 Moreover, for \(u\in[0,0.25]\), \(X(u)\geq x^*\) and \(r_{00}(X(u))=\theta_2=1>0\), and for \(u\in(0.25,0.5]\), \(X(u)\in(0,x^*)\) and \(r_{00}(X(u))=-\theta_2=-1<0\).
-Hence, for \(u\leq 0.5\) the second fluid level is \(Y(u)=u1(u\leq 0.25)+(0.25-u)1(u>0.25)\) and at time \(u=0.5\), the second fluid level first returns to \(0\), i.e.~\(\zeta_{Y}(\{0\})=0.5\) on this sample path.
+Hence, for \(u\leq 0.5\) the second fluid level is \(W(u)=u1(u\leq 0.25)+(0.25-u)1(u>0.25)\) and at time \(u=0.5\), the second fluid level first returns to \(0\), i.e.~\(\zeta_{W}(\{0\})=0.5\) on this sample path.
 The probability of this sample path is \(e^{-(\beta_1+\beta_2)\times 0.5}\).
 Thus, in the distribution of the first return time of the second fluid level there is a point mass at \(u=0.5\) with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\).
-Furthermore, on the same sample path \(X(\zeta_Y(\{0\}))=X(0.5)=1.2\), and hence there is a point mass with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\) in the distribution of \(X(\zeta_Y(\{0\}))\) also.
+Furthermore, on the same sample path \(X(\zeta_W(\{0\}))=X(0.5)=1.2\), and hence there is a point mass with magnitude \(e^{-(\beta_1+\beta_2)\times 0.5}\) in the distribution of \(X(\zeta_W(\{0\}))\) also.
 
@@
 	\includegraphics[width=0.5\textwidth,trim={0.75cm 0.8cm 0.25cm 1.25cm},clip]{chapter6/figs/ffq/discts/l1_cdf_error_formatted.pdf}
-	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate CDFs of \(X(\zeta_{Y}(\{0\}))\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}.
+	\caption{KS (left) and \(L^1\) (right) errors between the simulated and approximate CDFs of \(X(\zeta_{W}(\{0\}))\) (Equation~\ref{eqn: first return Y 1}) for Model~\ref{model: ffq2}.
 The approximations were obtained via the DG (blue solid line), uniformisation (green dashed line) and QBD-RAP (purple dotted line) schemes.
@@
 	\includegraphics[width=0.8\textwidth,trim={0cm 1.25cm 0cm 1.25cm},clip]{chapter6/figs/ffq/discts/phase_4_cdf.pdf}%
-	\caption{Approximations of the CDF of the distribution of \(X\) at the time \(\zeta_Y(\{0\})\) in Phase~1 for Model~\ref{model: ffq2}.
+	\caption{Approximations of the CDF of the distribution of \(X\) at the time \(\zeta_W(\{0\})\) in Phase~1 for Model~\ref{model: ffq2}.
 The blue line was obtained from the dimension-21 DG scheme, the purple dotted line from the dimension-21 QBD-RAP scheme, and the gold dotted line is the empirical CDF obtained via simulation.

### 2023-01-22 front/notation.tex
post-reviews, pre supervisor feedback

\chapter*{Notation}
\subsection*{General notation}
\begin{longtable}{p{0.45\textwidth}p{0.55\textwidth}}
  \(\bs u = (u_h)_{h\in \mathcal H}\)
      & Denotes a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some index set. \\
  \(\bs u = (\bs u_h)_{h\in\mathcal H}\) 
      & A row-vector defined by a collection of row-vectors \(\bs u_h\). The notation \(\bs u_m=(u_h)_{h\in\mathcal H_m}\) refers to the vector containing the subset of elements corresponding to \(\mathcal H_m\subseteq \mathcal H\). When the index set is empty, the resulting vector \(\bs u_m\) is a vector of dimension 0. In cases when there are two indices, we order the elements of the vector according to the first index, then the second; i.e.~\(\bs u = (u_{g}^h)_{g\in\mathcal G,h\in\mathcal H} = ((u_g^h)_{g\in\mathcal G})_{h\in\mathcal H}\). Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\). \\
  \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) 
      & Denotes a matrix defined by its elements, or sub-blocks, \(u_{gh}\). \\
  \(\diag(v_n, n\in N)\) 
      & A matrix with diagonal elements \(\{v_n, n\in N\}\), and all off-diagonal elements are zero. \\
  \(\{\cdot\}_{t\in I}\) e.g.~(\(\{X(t)\}_{t\geq}\))
      & Curly braces denote a sequence and the subscript denote the index set for the sequence. The subscript may be omitted in it is not necessary. \\
  \(x^-\) 
      & The left limit at \(x\). \\ 
  \(x^+\) 
      & The right limit at \(x\). \\ 
  \({}'\) 
      & The transpose of a vector or matrix (e.g.~\(\bs v'\) denotes the transpose of \(\bs v\)). \\ 
  \(\otimes, \oplus\) 
      & The Kronecker product and sum, e.g.~\(\bs A\otimes \bs B\) the Kronecker product of matrices \(A\) and \(B\). See Appendix~\ref{sec:ksdfkkakaaaaaa}. \\ 
\end{longtable}

\subsection*{Glossary of Symbols}
\begin{longtable}{p{0.45\textwidth}p{0.55\textwidth}}
  \(1(\cdot)\) 
      & The indicator function (equal to 1 when the argument is true, otherwise equal to 0) \\ 
   \(\bs\alpha^{(p)}\) 
      & See entry for \(\{Z^{(p)}\}\). \\
  \(\beta(t)= \int_0^t \left| r_{\varphi(z)}(X(z)) \right|  \wrt z \)
      & The total unregulated amount of fluid that has flowed into or out of the second buffer during $[0,t]$. See Section~\ref{subsec: afjakje}. \\
  \(\beta_X(t)\)
      & The in-out fluid level (equal to \(\int_0^t \left| c_{\varphi(z)} \right|  \wrt z\)). See Section~\ref{sec: transient ffq intro}. \\
  $\eta(w) = \inf \{t > 0: \beta(t) = w\}$ 
      & The first time the accumulated in-out amount for the second fluid hits level $w$. See Section~\ref{subsec: afjakje}. \\ 
  \(\eta_X(y)\) 
      & The first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\). See Section~\ref{sec: transient ffq intro}. \\
  \(\varepsilon^{(p)}\) 
      & A ``small'' error term which we get to choose so long as it tends to zero as \(p\to\infty\). This is set to \(\var\left(Z^{(p)}\right)^{1/3}\) in the convergence arguments. It arises in the application of Chebyshev's inequality. \\ 
  \(\Gamma_m,\, \Sigma_m\) 
      & The time of the \(m\)th down-up and up-down transition of the QBD-RAP, respectively. See Section~\ref{sec: qbd dists}. \\ 
  \(\hGamma_m,\, \hSigma_m\) 
      & The time of the \(m\)th down-up and up-down transition of the fluid queue, respectively. See Section~\ref{sec: no change}. \\ 
  \(\mu^{\ell_0}(t)(\cdot,j; x_0,i)\) 
      & The joint density/mass function of the fluid queue restricted to level \(\ell_0\). See Section~\ref{sec: no change}. \\
  \(\mu^{\ell_0}(t)(\cdot,j; x_0,i)\) 
      & The joint density/mass function of the fluid queue restricted to level \(\ell_0\) partitioned on the event that there are \(m\) down-up or up-down transitions, and \(i\in\calS_r\), \(j\in\calS_s\). See Section~\ref{sec: no change}. \\
  \(\widehat \mu^{\ell_0}(\lambda)(\cdot,j; x_0,i)\) 
      & The Laplace transform with respect to time of \(\mu^{\ell_0}(t)(\cdot,j; x_0,i)\). \\ 
  \(\widehat \mu^{\ell_0}(\lambda)(\cdot,j; x_0,i)\) 
      & The Laplace transform with respect to time of \(\mu^{\ell_0}(t)(\cdot,j; x_0,i)\). \\  
  \(\bbpi_i(y)(\mathcal{E})\) 
      & The stationary density operator of a fluid-fluid queue. See Equation (\ref{eqn:jointpi}). \\
  \(\{\phi(t)\}_{t\geq 0}\)   
      & The phase variable of the QBD-RAP approximation to a fluid queue. \\
  \(\phi_k^r\) 
      & The \(r\)th basis function of the DG approximation scheme on cell \(k\) (i.e.~the \(r\)th basis function of \(W_k\)). See Section~\ref{sec: dgdgdgdg}. \\ 
  \(\bs \phi_k\) 
      & A row vector of the basis functions on cell \(k\). See Section~\ref{sec: dgdgdgdg}. \\ 
  \(\{\varphi(t)\}_{t\geq0}\) 
      & Phase process of the fluid process. \\ 
  \(\{\varphi^*(t)\}_{t\geq0}\) 
      & Phase process of the augmented state space fluid process. See Section~\ref{sec: zero states} \\ 
  \(\mathbb \Psi(s)\) 
      & The first-return operator of the second fluid level which maps initial distributions of the driving fluid queue, to the distribution of the fluid queue at the time that the second fluid level first returns to its initial level. See Section~\ref{sec: intro Psi}. \\
  \(\psi(\cdot)\) 
      & A test function. \\ 
  \(\bs \Psi_X(\lambda)=\left[\bs \Psi_X(\lambda)\right]_{i\in\calS_+,j\in\calS_-}\)
      & The Laplace transform of the first return time to a fluid queue. See Section~\ref{sec: transient ffq intro}. \\
  \(\tau_1^{(p)}\) 
      & The first orbit restart epoch of the approximating QBD-RAP. See Section~\ref{sec: qbd dists}. \\
  \(\tau_n^{(p)}\) 
      & The \(n\)th orbit restart epoch of the QBD-RAP. See Section~\ref{sec: embedded}. \\ 
  \(\tau_1^X\)
      & The minimum of the first time the level variable of the fluid queue hits the cell edges \(y_0,...,y_{K+1}\), or the fluid queue leaves the boundary. See Section~\ref{sec: qbd dists}. \\ 
  \(\tau_{n}^X\).  
      & The \(n\)th time the level process of the fluid queue hits \(y_0,...,y_{K+1}\), or leaves the boundary. See Section~\ref{sec: embedded}. \\ 
  \(\zeta_W(E)= \inf \{t > 0: \dot W(t) \in E\}\) 
      & The stopping time which is the first time that the second fluid level hits the set \(E\).\\
  \(\zeta_X(E)\) \\
      & The random variable which is the first hitting time of \(\{\ddot X(t)\}\) on a set \(E\). See Section~\ref{sec: transient ffq intro}. \\
  \(\mathcal A\) 
      & The space of initial vectors \(\bs a\in\mathcal A\) such that \(\bs ae^{\bs Sx}\bs s\) is a valid probability density function. See Section~\ref{qbd-rap intro}. \\ 
  \(\{\bs A(t)\}_{t\geq 0}\) 
      & The orbit process of a QBD-RAP (or RAP). See Section~\ref{qbd-rap intro}. \\ 
  \(\mathbb B\) 
      & The generator of a fluid queue. See Section~\ref{subsec: operatorsoffq}. \\
  \(\ddot{\bs B}\) 
      & An approximation to the generator of an unbounded fluid queue. \\ 
  \({\bs B}\) 
      & An approximation to the generator of a bounded fluid queue. \\ 
  \(b^{(p)}\) 
      & A bound defined in Lemma~\ref{lem: another bound}. The dependence on \(p\) may be omitted. \\ 
  \(\bs C_m=\diag(|c_i|, i\in\calS_m),\) \(m\in\{+,-,0\}\)
      & A diagonal matrix of the absolute value of rates of a fluid queue for a subset of phases. \\
  \(\bs C=\left[\begin{array}{ccc}\bs C_+ && \\ &\bs C_-& \\ &&\bs C_0\end{array}\right]\)
      & A diagonal matrix of the absolute value of rates of a fluid queue. \\
  \(\widehat{\bs C}_m=\diag(c_i, i\in\calS_m),\) \( m\in\{+,-,0\}\)
      & A diagonal matrix of fluid queue rates for a subset of phases. \\
  \(\widehat{\bs C}=\left[\begin{array}{ccc}\widehat{\bs C}_+ && \\ &\widehat{\bs C}_-& \\ &&\widehat{\bs C}_0\end{array}\right]\)
      & A diagonal matrix of fluid queue rates. \\
  \(c_i\) 
      & The rate of change (with respect to time) of the level variable of a fluid queue when the phase variable is \(i\). \\ 
  \(\bs D\) 
      & The jump matrix of the QBD-RAP approximation. See the end of Section~\ref{sec:kkkrjrjrjjrj}. \\
  \(\mathbb D(s)\)  
      & The generator of \(\mathbb U(y,s)\). \\ 
  \(\mathcal D_{-1} = {0}\)
      & Convenient notation for the approximation cell at the lower boundary of a bounded fluid queue. \\
  \(\mathcal D_{K+1} = {b}\)
      & Convenient notation for the approximation cell at the upper boundary of a bounded fluid queue. \\
  \(\mathcal D_{k,i},\, i\in\calS\) 
      & The ``cells'' of the approximation schemes. Equal to \([y_k,y_{k+1})\) for \(i\in\calS_+\cup\calS_0\) or \((y_k,y_{k+1}]\) for \(i\in\calS_-\). \\
  \(\mathcal D_k\) 
      & The \(k\)th cell of the approximation scheme, \(\mathcal D_k=[y_k,y_{k+1}]\). \\
  \(\Delta_k\)
      & The width of approximation cell \(k\). The subscript \(k\) may be dropped when \(\Delta_k\) is equal for all cells. \\ 
  \(\bs e\) 
      & A vector of all ones. \\
  \(\bs e_i\) 
      & A vector of all zeros except there is a 1 in the \(i\)th position. \\ 
  \(E^\lambda\) 
      & An exponential random variable with rate \(\lambda>0\). \\
  \(\mathcal F_i^+={u\in[0,b]: r_i(u)>0}\)
      & The set of values \(u\in[0,b]\) such that the second fluid rate for phase \(i\in\calS\) is positive. See Equation (\ref{eqn:fil}). \\
  \(\mathcal F_i^-={u\in[0,b]: r_i(u)<0}\)
      & The set of values \(u\in[0,b]\) such that the second fluid rate for phase \(i\in\calS\) is negative. See Equation (\ref{eqn:fil}). \\
  \(\mathcal F_i^0={u\in[0,b]: r_i(u)>0}\)
      & The set of values \(u\in[0,b]\) such that the second fluid rate for phase \(i\in\calS\) is zero. See Equation (\ref{eqn:fil}). \\
  \(\bs F_{k,\ell}\) 
      & The flux matrix for the flow of mass from cell \(k\) to cell \(l\in\{k-1,k+1\}\), for the DG scheme. See Sections~\ref{sec: dgdgdgdg} and \ref{sec:llaaksksbnnn}. \\ 
  \(f_i(x,t)\) 
      & The density of a fluid queue at time \(t\) evaluated at \(\bs X(t)=(X(t),\varphi(t))=(x,i)\). \\ 
  \(\bs f(x,t)=(f_i(x,t))_{i\in\calS}\) 
      & A vector of density functions of a fluid queue. \\ 
  \(f^{\ell_0,(p)}(t)(x,j;x_0,i)\) 
      & The QBD-RAP approximation to the joint density/mass function of the fluid queue restricted to level \(\ell_0\). See Section~\ref{sec: qbd dists}. \\
  \(f^{\ell_0,(p)}_{m,r,s}(t)(x,j;x_0,i)\) 
      & The QBD-RAP approximation to the joint density/mass function of the fluid queue restricted to level \(\ell_0\) partitioned on the event that there are \(m\) down-up or up-down transitions, and \(i\in\calS_r\), \(j\in\calS_s\). See Section~\ref{sec: qbd dists}. \\ 
  \(\widehat f^{\ell_0,(p)}(\lambda)(x,j;x_0,i)\) 
      & The Laplace transform with respect to time of \(f^{\ell_0,(p)}(t)(x,j;x_0,i)\). \\ 
  \(\widehat f^{\ell_0,(p)}_{m,r,s}(\lambda)(x,j;x_0,i)\) 
      & The Laplace transform with respect to time of \(f^{\ell_0,(p)}_{m,r,s}(t)(x,j;x_0,i)\). \\ 
  \(g(x),\, g_1,g_2,...\) 
      & Arbitrary functions satisfying Assumptions~\ref{asu: g}. \\ 
  \(G\) 
      & A bound on \(g,g_1,g_2,....<G\). See Assumptions~\ref{asu: g}. \\
  \(\widehat G\) 
      & A bound on \(\int_{x=0}^\infty g(x) \wrt x\leq \widehat G\), \(\int_{x=0}^\infty g_k(x) \wrt x\leq \widehat G, \, k=1,2,....\) See Assumptions~\ref{asu: g}. \\ 
  \(G_{\bs v}, \, \widetilde G_{\bs v}\) 
      & Bounds on closing operators. See Properties~\ref{properties: some props}. \\
  \(\bs G_k\) 
      & The stiffness matrix for cell \(k\) of the DG scheme. See Sections~\ref{sec: dgdgdgdg} and \ref{sec:llaaksksbnnn}. \\ 
  \(\bs H(\lambda, y)=\left[h_{ij}(\lambda,y)\right]_{i,j\in\calS_+\cup\calS_-}\)
      & The Laplace-Stieltjes transform of the time taken for \(y\) amount of fluid to flow in or out of the fluid queue and to be in phase \(j\) at this time, given the initial level of the fluid queue was \(0\) and the initial phase was \(i\). See Section~\ref{sec: transient ffq intro} and \ref{sec: lst on no change}. \\
  \(\bs H^{mm}(\lambda, y)=\left[h^{mm}_{ij}(\lambda,y)\right]_{i,\in\calS_m,j\in\calS_m},\) \( m\in\{+,-\}\) 
      & The Laplace transform (with respect to time) of the time taken for the fluid level to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_m\cup\calS_{m0}\), given the phase was initially \(i\in\mathcal S_m\). See Section~\ref{sec: transient ffq intro} and \ref{sec: lst on no change}. \\
  \(\bs H^{mn}(\lambda, y)=\left[h^{mn}_{ij}(\lambda,y)\right]_{i,\in\calS_m\cup\in\calS_n},\) \( m,n\in \{+,-\}, m\neq n\)
      & The Laplace transform (with respect to time) of the time taken for the fluid level to shift by an amount \(y\) whilst remaining in phases in \(\mathcal S_m\cup\calS_{m0}\), after which time the phase instantaneously changes to \(j\in\mathcal S_n\), given the phase was initially \(i\in\mathcal S_m\). See Section~\ref{sec: transient ffq intro} and \ref{sec: lst on no change}. \\ 
  \(\bs k(t)=\bs \alpha e^{\bs St}/(\bs \alpha e^{\bs St}\bs e)\)
      & A row-vector valued function related to the orbit process of the QBD-RAP. \\
  \(\mathcal K^\circ=\{0,1,...,K\}\)
      & The index set for the interior cells/intervals for the approximation schemes. \\
  \(\mathcal K=\{-1,K+1\}\cup\{0,1,...,K\}\)
      & The index set for the cells/intervals for the approximation schemes. \\
  \(\mathcal K^m_i = \{k\in\mathcal K\mid  l(\calD_{k,i}\cap\calF_i^m)\) where \(l\) is Lebesgue measure.
      & The set of indices of the cells which are subsets of \(\calF_i^m\). Then \(\bigcup\limits_{k\in\mathcal K_i^m} \calD_{k,i}\) and \(\mathcal F_i^m\) are equal up to a set of \(\mathcal M_{0,b}\)-measure 0. \\
  \(\mathcal K^m = \bigcup\limits_{i\in\calS}\mathcal K_i^m\), \(m\in\{+,-,0\}\). 
      & \\ 
  \(\{\bs L(t)\}_{t\geq 0}\) 
      & The level process of a QBD-RAP. See Section~\ref{qbd-rap intro}. \\ 
  \(L\) 
      & The Lipschitz constant of \(g, g_1,g_2,...\). See Assumptions~\ref{asu: g}. \\ 
  \(\bs M_k\) 
      & The mass matrix for cell \(k\) of the DG scheme. See Sections~\ref{sec: dgdgdgdg} and \ref{sec:llaaksksbnnn}. \\ 
  \(ME(\bs\alpha, \bs S, \bs s)\) (or \(ME(\bs\alpha, \bs S)\))
      & A matrix exponential distribution with initial vector \(\bs \alpha\), generator \(\bs S\), and closing vector \(\bs s\). The closing vector may be dropped from the notation when it is assumed that \(\bs s=-\bs S\bs e\). See Section~\ref{qbd-rap intro}. \\
  \({\mathbb p}_i(\mathcal{E})\) 
      & The stationary point-mass operator of a fluid-fluid queue. See Equation (\ref{eqn:jointmass}). \\
  \(\widecheck p_{ij},\,i,j\in\calS,\,k\in\{-1,K+1\}\) 
      & The probability of a phase change of a fluid queue from phase \(i\) to phase \(j\) which occurs instantaneously upon the fluid queue hitting the lower boundary at \(0\). \\
  \(\widehat p_{ij},\,i,j\in\calS,\,k\in\{-1,K+1\}\) 
      & The probability of a phase change of a fluid queue from phase \(i\) to phase \(j\) which occurs instantaneously upon the fluid queue hitting the upper boundary. \\
  \(\widecheck{\bs P}_{-+} = [\widecheck p_{ij}]_{i\in\calS_-,j\in\calS_+}\) 
      & A matrix describing the probability of transitioning from a negative phase to a positive phase upon the first fluid level hitting the lower boundary. \\
  \(\widecheck{\bs P}_{-0} = [\widecheck p_{ij}]_{i\in\calS_-,j\in\calS_0}\)  
      & A matrix describing the probability of transitioning from a negative phase to a phase with rate zero upon the first fluid level hitting the lower boundary. \\
  \(\widecheck{\bs P}_{--} = [\widecheck p_{ij}]_{i\in\calS_-,j\in\calS_-}\)
      & A matrix describing the probability of transitioning from a negative phase to a negative phase upon the first fluid level hitting the lower boundary. \\ 
  \(\widehat{\bs P}_{+-} = [\widehat p_{ij}]_{i\in\calS_+,j\in\calS_{-}}\) 
      & A matrix describing the probability of transitioning from a positive phase to a negative phase upon the first fluid level hitting the upper boundary. \\
  \(\widehat{\bs P}_{+0} = [\widehat p_{ij}]_{i\in\calS_+,j\in\calS_0}\) 
      & A matrix describing the probability of transitioning from a positive phase to a phase with rate zero upon the first fluid level hitting the upper boundary. \\ 
  \(\widehat{\bs P}_{++}^{K+1} = [\widehat p_{ij}]_{i\in\calS_+,j\in\calS_+}\) 
      & A matrix describing the probability of transitioning from a positive phase to a positive phase upon the first fluid level hitting the upper boundary. \\
  \(\bs Q(\lambda),\, m\in\{+,-\}\)  
      & The generator of \(\bs H(\lambda,y)\). See Section~\ref{sec: transient ffq intro} and \ref{sec: lst on no change}. \\ 
  \(\bs R(x)=\diag(r_i(x))\)
      & The diagonal matrix function of rates of the second level process of a fluid-fluid queue. \\
  \(\mathbb R\) 
      & The operator which maps measures to the same measures divided by the rate of the second fluid level. See Section~\ref{subsec: afjakje}. \\
  \(r_i(x)\) 
      & The rate of change of the second fluid level of a fluid-fluid queue when the phase is \(i\) and the first level is \(x\). \\
  \(R_i(u)\) 
      & The residual time of \(Z_i\) given \(Z_i>u\). \\ 
  \( R_{\bs v,1}^{(p)}, \, R_{\bs v,2}^{(p)}\) 
      & Bounds on closing operators which tend to \(0\) as \(p\to \infty\). See Properties~\ref{properties: some props}. \\
  \(r_1(n)^{(p)}\) 
      & An error term defined in Corollary~\ref{cor: lh and rh}. The dependence on \((p)\) may be suppressed. \\
  \(r_2(n)^{(p)}\) 
      & An error term defined in Lemma~\ref{lemma:bound}. The dependence on \((p)\) may be suppressed. \\
  \(r_3(u+v)\) 
      & An error term defined in Corollary~\ref{cor: cond bnd 2}. The dependence on \((p)\) may be suppressed. \\ 
  \(r_4(n)\) 
      & An error term defined in Lemma~\ref{lem: lst convergence}. The dependence on \((p)\) may be suppressed. \\ 
  \(r_5(n)\) 
      & An error term defined in Lemma~\ref{lem: lst convergence}. The dependence on \((p)\) may be suppressed. \\ 
  \(r_6^M\) 
      & An error term defined in Lemma~\ref{lem: gkjljklgagjklagsjlk}. (There is no dependence on \(p\).) \\ 
  \(\bs S^{(p)}\) 
      & See entry for \(\{Z^{(p)}\}\). \\
  \(\calS=\{1,...,N\}\) 
      & The state space of \(\{\varphi(t)\}\) and \(\{\phi(t)\}\). \\
  \(\bs S_i,\, \bs s_i\) 
      & \(|c_i|\bs S\) and \(|c_i|\bs s\). \\ 
  \(\calS_+=\{i\in\calS \mid c_i>0\}\) 
      & The set of phases of a fluid queue for which the rate of change of the fluid level is positive. \\
  \(\calS_-=\{i\in\calS \mid c_i<0\}\) 
      & The set of phases of a fluid queue for which the rate of change of the fluid level is negative. \\
  \(\calS_0=\{i\in\calS \mid c_i=0\}\) 
      & The set of phases of a fluid queue for which the rate of change of the fluid level is equal to 0. \\
  \(\bs s^{(p)}\) 
      & See entry for \(\{Z^{(p)}\}\). \\
  \(\calS_{m0}\) 
      & A copy of \(\calS_0\) which is only accessible via \(\calS_m\), \(m\in{+,-}\). \\ 
  \(\calS_k^+=\{i\in\calS\mid r_i(x)>0,\,\forall x \in\calD_{k,i}\}\) 
      & Phase of a fluid queue where the rates of the second fluid are positive on the sets \(\mathcal D_{k,i},\, i\in\calS\). \\
  \(\calS_k^0=\{i\in\calS\mid r_i(x)=0,\,\forall x \in\calD_{k,i}\}\)
      & Phase of a fluid queue where the rates of the second fluid are positive on the sets \(\mathcal D_{k,i},\, i\in\calS\).\\
  \(\calS_k^-=\{i\in\calS\mid r_i(x)<0,\,\forall x \in\calD_{k,i}\}\)
      & Phase of a fluid queue where the rates of the second fluid are positive on the sets \(\mathcal D_{k,i},\, i\in\calS\).\\
  \(\calS_k^\bullet=\{i\in\calS\mid r_i(x)\neq0,\,\forall x \in\calD_{k,i}\}\). 
      & Phase of a fluid queue where the rates of the second fluid are non-zero on the sets \(\mathcal D_{k,i},\, i\in\calS\).\\ 
  \(\bs T=[T_{ij}]\) 
      & The generator of a CTMC which is the driving process of the fluid queue. \\
  \(U_k\) 
      & The space of function in which we formulate the DG approximation for cell \(k\) (the set of polynomials of order \(p_k\)). \\ 
  \(U=\bigoplus_{k=0,...,K} U_k\) 
      & The space of function in which we formulate the DG approximation (also used as the set of test functions). \\
  $\mathbb{U}_{ij}^{k\ell}(y,s)$ 
      & The Laplace transform of the operator which maps the initial distribution of a fluid queue to the distribution of the fluid queue at the time that the in-out fluid if the second fluid level, \(\beta(t)\), reaches \(y\), given \(W(0)=0\) and the driving fluid process was initially at \(\varphi(0)=i\) and \(X(0)=x\). Defined for indices $k\in\mathcal K^+\cup\mathcal K^-$,  $\ell\in\mathcal K^+\cup\mathcal K^-$, and $i \in \mathcal{S}_k^\bullet,\,j \in \mathcal{S}_\ell^\bullet$. See Section~\ref{subsec: afjakje}. \\
  \(\mathbb U(y,s) = \left[[\mathbb U_{ij}^{k\ell}(y,s)]_{i\in\calS_k^\bullet,j\in\calS_\ell^\bullet}\right]_{k,\ell\in\mathcal K^+\cup\mathcal K^-}\) 
      & The operators \(\mathbb{U}_{ij}^{k\ell}(y,s)\) assembled into a matrix. See Section~\ref{subsec: afjakje}.\\
  \(\mathbb V(t)\) 
      & The semigroup which maps the initial distribution of a fluid queue to the distribution at time \(t\). (The transition operator). See Section~\ref{subsec: operatorsoffq}. \\
  \(\var(\cdot)\) 
      & The variance operator. \\ 
  \(\bs v(x)\)
      & A closing operator. See Section~\ref{sec: closing}. \\
  \(\{\bs v^{(p)}\}_p \)
      & A sequence of closing operators. See Properties~\ref{properties: some props}. \\
  \(\{\dot W(t)\}_{t\geq 0}\) 
      & The second fluid level of a fluid-fluid queue which is bounded below at 0.  \\
  \(\{\ddot W(t)\}_{t\geq 0}\) 
      & The second fluid level of a fluid-fluid queue which is unbounded.  \\
  \(\{X(t)\}_{t\geq 0}\)   
      & The fluid level of a fluid queue which is bounded below at \(0\) and above at \(b>0\).  \\
  \(\{\dot X(t)\}_{t\geq 0}\)   
      & The fluid level of a fluid queue which is bounded below at \(0\) and unbounded above.  \\
  \(\{\ddot X(t)\}_{t\geq 0}\)   
      & The fluid level of a fluid queue which is unbounded.  \\
  \(\{\bs X(t)\}_{t\geq 0}\) & The fluid process \(\bs X(t)=(X(t),\varphi(t))\). \\
  \(\{\bs Y(t)\}_{t\geq0}=\{(L(t), \bs A(t), \phi(t))\}_{t\geq 0}\) 
      & The QBD-RAP approximation to a fluid queue. \\ 
  \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z}\)
      & The process formed by observing the QBD-RAP at the orbit restart epochs \(\{\Ydp(n)\}_{n\geq 0, n \in \mathbb Z}=\{(L^{(p)}(\tau_n^{(p)}),\varphi(\tau_n^{(p)}))\}_{n\geq 0, n \in \mathbb Z}\). \\
  \(Z\) 
      & A ME distributed random variable with parameters \((\bs \alpha, \bs S)\). \\ 
  \(Z_i\) 
      & A ME distributed random variable with parameters \((\bs \alpha, |c_i|\bs S)\). \\ 
  \(\{Z^{(p)}\}_{p}\) 
      & A sequence of ME random variables with \(\var\left(Z^{(p)}\right)\to 0\) as \(p\to \infty\), where each 
      \(Z^{(p)}\) has parameters \(\bs \alpha^{(p)},\, \bs S^{(p)},\, \bs s^{(p)}\). \\
\end{longtable}

### 2023-01-22 intro/chapter.tex
post-reviews, pre supervisor feedback

@@
 More recently, \cite{bo2014} extend fluid queues to so-called \emph{stochastic fluid-fluid queues}.
-In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\), so \((X(t),\varphi(t))\) determines the rate at which \(\{Y(t)\}\) moves.
-The addition of the second level process \(\{Y(t)\}\) increases the modelling potential.
+In a fluid-fluid queue there is a second level process, \(\{W(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\), so \((X(t),\varphi(t))\) determines the rate at which \(\{W(t)\}\) moves.
+The addition of the second level process \(\{W(t)\}\) increases the modelling potential.
 For example, \cite{hydro} describe a fluid queue model for a hydroelectric power generator which can be operated `on design' and `off design'.
@@
 The level variable \(\{X(t)\}\) take values in \([0,1]\) and represents the deterioration level of the system with 0 being perfect working order and 1 meaning the system needs replacement.
-The second level process \(\{Y(t)\}\) could be used to model the revenue of the power generator, which, logically, should depend on the state of the system, \(\{\varphi(t)\}\).
+The second level process \(\{W(t)\}\) could be used to model the revenue of the power generator, which, logically, should depend on the state of the system, \(\{\varphi(t)\}\).
 Moreover, the revenue could also depend on the deterioration level \(\{X(t)\}\), as, for example, when the system is brand new it may operate more efficiently, and when it is near broken it may be less efficient.
@@
 % the level X (t) at time t, we would like to model some other continuous performance measure
-% Y (t), such as the net profit at time t for example. To illustrate this, consider a hydro-power
+% W (t), such as the net profit at time t for example. To illustrate this, consider a hydro-power
 % generator, which can be operated “on design” and “off design”. The latter causes increased
@@
 % restarts itself from level 0 [12].
-% Now, in order to describe the continuous revenue process {Y (t), t ≥ 0} associated with
+% Now, in order to describe the continuous revenue process {W (t), t ≥ 0} associated with
 % the process {(φ(t), X (t)), t ≥ 0}, it may be plausible to use a SFM driven by {φ(t), t ≥ 0}.
@@
 % depend on the phase φ(t). However, for practical reasons, we may additionally require that the
-% revenue Y (t) at time t depends on the operating level X (t). For example, if the generator is
+% revenue W (t) at time t depends on the operating level X (t). For example, if the generator is
 % newer, it may operate more efficiently, produce more energy and require less-costly maintenance.
-% Hence, the rates at which the revenue level Y (t) is increasing/decreasing may depend both on
+% Hence, the rates at which the revenue level W (t) is increasing/decreasing may depend both on
 % the phase i ∈ S and the operating level X (t) = x at time t, and so to reflect this, we denote
 % these rates by ri (x). Summarizing this, we propose a new class of fluid models, denoted by
-% {(φ(t), X (t), Y (t)), t ≥ 0}. We refer to this class as the stochastic fluid–fluid model, since it is
-% a stochastic fluid model with level Y (t) driven by {(φ(t), X (t)), t ≥ 0}, which is also a stochastic
+% {(φ(t), X (t), W (t)), t ≥ 0}. We refer to this class as the stochastic fluid–fluid model, since it is
+% a stochastic fluid model with level W (t) driven by {(φ(t), X (t)), t ≥ 0}, which is also a stochastic
 % fluid model with level X (t). The buffers collecting fluid from the process {X (t), t ≥ 0} and
-% {Y (t), t ≥ 0} are referred to as X and Y , respectively.
+% {W (t), t ≥ 0} are referred to as X and W , respectively.
 
-The addition of the second level process \(\{Y(t)\}\) increases the modelling potential, however, it comes at a cost to the simplicity of the analysis of the model.
+The addition of the second level process \(\{W(t)\}\) increases the modelling potential, however, it comes at a cost to the simplicity of the analysis of the model.
 For both fluid-fluid queues and classical fluid queues, performance measures, such as the stationary distribution or the Laplace-transforms of hitting times and return times, can be derived in terms of the infinitesimal generator of the modulating process.
@@
 
-The analysis of fluid-fluid queues in \cite{bo2014} is, in principle, similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the generator of the driving fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
+The analysis of fluid-fluid queues in \cite{bo2014} is, in principle, similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{W(t)\}_{t\geq0}\) in terms of the generator of the driving fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\).
 For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used.
@@
 In particular, we go into detail describing operators arising in the analysis of fluid-fluid queues and introducing them in such a way that makes clear exactly how the approximations correspond to the theoretical operators.
-Chapter~\ref{ch:math} also introduces the discontinuous Galerkin (DG) method for approximating solutions to differential equations.
+%Chapter~\ref{ch:math} also introduces the discontinuous Galerkin (DG) method for approximating solutions to differential equations.
 Due to issues with oscillations in the approximations produced by the DG scheme, at the end of Chapter~\ref{ch:math} we also describe \emph{slope limiting} -- a method which can be used to prevent oscillations and negativity.
@@
 Chapter~\ref{sec: numerics} can be read independently of Chapters~\ref{sec: conv}~and~\ref{ch: global results}, which are somewhat technical and could be skipped on a first-pass if the reader wishes.
+Although the construction of the QBD-RAP and the proof that it converges is somewhat technical, the actual computations involved in Chapter~\ref{sec: numerics} are relatively straightforward once we have the generator of the QBD-RAP and knowledge of how the level process of the QBD-RAP corresponds to the level process of the fluid queue.
 
@@
 
+Some of the more general mathematical background has been reserved for Appendix~\ref{app: math background}.
 To help the reader understand the notation used in the DG scheme, we provide a small toy model example in Appendix~\ref{appendix:example}.

### 2023-01-22 summary/summary.tex
post-reviews, pre supervisor feedback

@@
 \maketitle
-The stochastic fluid-fluid model (SFFM) is a Markov process {\(\{(X_t,Y_t,\varphi_t)\}_{t\geq 0}\),} where {\(\{\varphi_t\}_{t\geq 0}\)} is a continuous-time Markov chain, the first fluid, {\(\{X_t\}_{t\geq 0}\),} is a classical stochastic fluid process driven by {\(\{\varphi_t\}_{t\geq0}\),} and the second fluid, {\(\{Y_t\}_{t\geq0}\),} is driven by the pair {\(\{(X_t,\varphi_t)\}_{t\geq 0}\).} Operator-analytic expressions for performance measures of fluid-fluid queues, such as the stationary distribution and first return distribution of the SFFM, have been derived \cite{bo2014}.
+The stochastic fluid-fluid model (SFFM) is a Markov process {\(\{(X_t,W_t,\varphi_t)\}_{t\geq 0}\),} where {\(\{\varphi_t\}_{t\geq 0}\)} is a continuous-time Markov chain, the first fluid, {\(\{X_t\}_{t\geq 0}\),} is a classical stochastic fluid process driven by {\(\{\varphi_t\}_{t\geq0}\),} and the second fluid, {\(\{W_t\}_{t\geq0}\),} is driven by the pair {\(\{(X_t,\varphi_t)\}_{t\geq 0}\).} Operator-analytic expressions for performance measures of fluid-fluid queues, such as the stationary distribution and first return distribution of the SFFM, have been derived \cite{bo2014}.
 These performance measures are in terms of the infinitesimal generator of the process {\(\{(X_t,\varphi_t)\}_{t\geq 0}\)}, which is a differential operator.

### 2023-01-22 reviews/response.tex
ignore reviews

@@
 \AtBeginEnvironment{quote}{\itshape}
-
+\DeclareFontFamily{U}{mathx}{}
+\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
+\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
+\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
 \title{Response to Thesis Reviews}
@@
 published paper.
-\begin{quote}
-    This is not the content of Chapter 3. 
-\end{quote}
 
@@
 \begin{quote}
-    I'd be happy to add more to the background on fluid queus and RAP, but, personally, I don't think there is much more to say which is relevant to the thesis.
-This thesis does not have a great deal to do with the existing analyses of fluid queues.
-Thoughts?
+    Upon discussion with my supervisors, it was deemed that the introduction to fluid queues is sufficient for the purposes of this thesis.
+
+    Regarding RAPs (and Marked RAPs) I have added a formal definition of both (by their characterisation in terms of finite-dimensional basis of measures) and appropriate references.
+I have also added further characterisation of the ODE associated with these processes and also a property of the characterisation of the conditional distributions of the process.
+
+    I thank the reviewer for this welcome feedback. 
 \end{quote}
@@
 \begin{quote}
-    Yes. Sorry about that. 
+    To help clarify, we have added a glossary of the persistent notation used throughout the thesis.
+
+    The notation \(\phi\) and \(\varphi\) has been clarified, and I have now made it explicit when we are referring to \(\phi\) or \(\varphi\). 
 \end{quote}
@@
 \begin{quote}
-    IMO, I have used Theorem, Lemma, Corllary environments to help with this. Perhaps I could use Definition environments too, this might help. Thoughts? 
+    [IMO, I have used Theorem, Lemma, Corllary environments to help with this. Perhaps I could use Definition environments too, this might help. Thoughts?]
 \end{quote}
@@
 \begin{quote}
-    I'm happy either way.
-These two chapters were originally one long chapter but i decided to break them in two because they use distinct technical tools.
-Perhaps I should make them a single chapter with two main sections in the chapter.
+    I originally had Chapters 5 and 6 combined into one long chapter, however, I (after discussion with my supervisors), decided to break the chapter into two to help readability.
+The techniques in Chapter 5 emphasise matrix-exponential and QBD-RAP specific techniques/properties.
+In contrast, the techniques/properties in Chapter 6 are more typical such as the application of the law of total probability and (strong) Markov property.
+I have decided to leave this as two distinct chapters.
+Also, Chapter 5 relies on Appendix~D and is related to the results in Appendix~E, whereas Chapter 6 only depends on Chapter 5, via Lemma~5.13 and Corollary~5.15 only and is otherwise independent of the appendices.
 \end{quote}
@@
 broader in its scope.
-\begin{quote}
-    Interesting that he didn't mention Chapter 7... perhaps he does not recognise numerical investigation in the same regard as the technical chapters, haha.
-
-    Yes, the scope of the thesis is narrow. It was one of my main worries about the work. I could have added more, but, as he already pointed out, the thesis is already long and hard to read.
-
-    Specific comments on structure to be addressed below. 
-\end{quote}
 
@@
 
-p. 12 l.end-6: you have already used matrices extensively.
+\st{p. 12 l.end-6: you have already used matrices extensively.}
 \begin{quote}
-    I have. I just wanted to clarify notation, and point out that matrices use square brackets. 
+    I have moved this section to a footnote and clarified matrix notation in the Glossary. 
 \end{quote}
@@
 \begin{quote}
-    Actually, there shouldn't be a (t)
+    Actually, there shouldn't be a (t) at all. This has been corrected.
 \end{quote}
@@
 
-(2.42): one does announce a proof to a formula/equation. Why do you proved it all? Most other items around this section are left unproven.
+\st{(2.42): one does announce a proof to a formula/equation. Why do you proved it all? Most other items around this section are left unproven.}
 \begin{quote}
-    I don't think there is a need for the proof. Happy to omit it. Same for the MPR on the previous page?
+    I have omitted the proofs and also moved this section to the mathematical background appendix.
 \end{quote}
@@
 \begin{quote}
-    happy to add proofs, but they are relatively technical and not absolutely necessary for the thesis IMO. They could help the reader intuit the fluid queue and its operators. Thoughts?
+    Section~2.4 summarises results of other authors, who also provide proof of these results.
+The proofs of these results are long, technical, and, in our opinion, do not add a great deal of insight into the results or process, so they are omitted.
 \end{quote}
@@
 Chapter 3 is taken from a paper, reproduced as such and in a good shape.
-\begin{quote}
-    I disagree, but okay. IMO this is one of the hardest-to-read sections!
-\end{quote}
 
@@
 \begin{quote}
-    Fixed typo on p. 68. However, I still need to clarify this notation throughout. 
+    Fixed typo on p. 68.
+
+    I have clarified the notation \(\varphi\) and \(\phi\) and made references to these variables explicit throughout the thesis. A Glossary has been added to help clarify notation too. 
 \end{quote}
@@
 \begin{quote}
-    Defined \(T\) and \(c_i\) again. 
+    I have defined \(T\) and \(c_i\) again. 
 \end{quote}
@@
 \begin{quote}
-    -Added ``Intuitively, the transposition of the Erlang sub-blocks, \(\left[\begin{array}{cc} -2\Delta & 2\Delta \\ 0 & -2\Delta \end{array}\right]\), in negative phases relative to positive phases is due to ``reversing'' the progression of the jump process associated with the Erlang variables in negative phases relative to positive phase.''
+    -I have added the comment ``Intuitively, the transposition of the Erlang sub-blocks, \(\left[\begin{array}{cc} -2\Delta & 2\Delta \\ 0 & -2\Delta \end{array}\right]\), in negative phases relative to positive phases is due to ``reversing'' the progression of the jump process associated with the Erlang variables in negative phases relative to positive phase.''
 
@@
 \begin{quote}
-    Clarified: ``In this chapter we choose to use an equivalent, but slightly different, representation of the QBD-RAP, we write \(\{(L(t),\boldsymbol A(t),\phi(t))\}_{t\geq 0}\), where we refer to \(\{\phi(t)\}\) as the \emph{phase process} of the QBD-RAP.
+    I have now clarified: ``In this chapter we choose to use an equivalent, but slightly different, representation of the QBD-RAP, we write \(\{(L(t),\boldsymbol A(t),\phi(t))\}_{t\geq 0}\), where we refer to \(\{\phi(t)\}\) as the \emph{phase process} of the QBD-RAP.
 Such a representation does not always exist for a QBD-RAP.
 However, for the specific QBD-RAP introduced in this thesis such a representation does exist, by construction.''
+
+    I have also clarified the notation throughout. 
 \end{quote}
 
-\st{p. 72 l. 10: seems that the symbol Ri(u) is used before introduced just below. I believe
-it should say Zi instead and remove u from the interval.}
+p. 72 l. 10: seems that the symbol\(R_i(u)\) is used before introduced just below. I believe
+it should say \(Z_i\) instead and remove \(u\) from the interval.
 \begin{quote}
-    Changed
+    This has been corrected. 
 \end{quote}
@@
 \begin{quote}
-    I guess the choice is relatively arbitrary at this point. This was a choice used by the CME group. Given the coarseness of Chebyshev's inequality, I do not think the choice is important.
+    I have added the comment ``Choose \(\varepsilon = \mbox{Var}(Z)^{1/3}\) (similar to the choice of \(\varepsilon\) in the proof of Theorem~4 of Horvath, Horvath \& Telek (2020)).
+Since we use Chebyshev's inequality to obtain the bound, which is general and likely to be far from sharp, we expect that the choice of \(\varepsilon\) is not important.
+All we require is that when \(\mbox{Var}(Z)\) is small, then \(\mbox{Var}(Z)/\varepsilon^2\) is also small.
+''
 \end{quote}
@@
 \begin{quote}
-    It is a pair? I assume this was misread by the reviewer? 
+    I have used a pair unless otherwise needed. 
 \end{quote}
@@
 \begin{quote}
-    I'm not entirely sure what the reviewer means by this.
-I realised that the ODE description of the QBD-RAP evolution was missing so I added it to the preliminaries, and I added a reference for it too.
-Do they want me to clarify/derive the equation?
+    This comment could refer to two possible things.
+1) Justification of the ODE which describes the movement of the orbit.
+Or, 2) Justification of the parameters of the ODE.
+I have addressed them both in the following way.
+
+    1) I have added a reference for the ODE and also added it to the preliminaries.
+
+    2) I have referenced back to the discussion about the choice of \(\boldsymbol S_i\) to model changes of level and the choice of \(-T_{ii}\) to model changes of phase. 
 \end{quote}
@@
     \begin{itemize}
-        \item I am not calculating anything. I am describing and summarising the dynamics of the stochastic process which we just introduced.
-        \item Sure, it's trivial if you're the person who invented and wrote the book on RAPs! For mortals, it is less trivial. 
+        \item I have clarified that \(\boldsymbol A(t)\boldsymbol D\boldsymbol e=1\) as \(\boldsymbol D\boldsymbol e=1\) by assumption and \(\boldsymbol A(t)\boldsymbol e=1\) is a property of QBD-RAPs.
+        \item This is the case. The calculations are rudimentary. However, I believe them to be insightful, so I have left them as it. 
     \end{itemize}
@@
 \begin{quote}
-    Yep, I need to go through the whole thesis and clarify this...
+    I have clarified what \(\varphi\) and \(\phi\) refer to in text. See also the other responses regarding \(\varphi\) and \(\phi\). 
 \end{quote}
@@
 \begin{quote}
-    These matrices can be deduced from the description of the process above (which is apparently ``trivial'')
+    I have removed the matrices which appeared earlier as they were a duplicate of those on p 88.
+
+    I have now added a description of the generator. \textbf{Peter, Nigel,} please can you read the last 4 paragraphs of Section 4.4 and verify that it is accurate and sufficient.
 \end{quote}
@@
     \begin{itemize}
-        \item Yes, there can be a change upon hitting due to the boundary conditions we have specified. 
-        \item Okay, I will update notation from \(p_{ij}^{-1}\) to something else. 
-        \item You do not miss something. The calculations are rudimentary (not trivial, but why get rid of them? 
+        \item There can be a change upon hitting due to the boundary conditions we have specified. 
+        \item I have changed \(p_{ij}^{-1}\) to \(\widecheck p_{ij}\) and clarified notation with a glossary. 
+        \item You do not miss something.
+The calculations are rudimentary, however they describe and exhibit that we can model the phase dynamics of the fluid queue exactly with the phase of the QBD-RAP.
     \end{itemize}
@@
 \begin{quote}
-    How should I explain these?
+    I have added comments on how the parameters of the QBD-RAP relate to its generator as I describe the process dynamics.
+
+    \textbf{Nigel}, could you please take a look at Section 4.4 and check that it makes sense. Also ready page 84, after the displayed matrices.
 \end{quote}
@@
     \begin{itemize}
-        \item TO DO: justify matrices. 
-        \item I think it is still relatively straight forward to ensure that it is a QBD-RAP. I could elaborate? 
+        \item Similar to my comment about his feedback about matrices on p.~91 -- not sure what more he wants here.
+        \item I have added the following.
+``Since \(\boldsymbol a_{\ell,i}(x)\in\mathcal A\) for all \(x\geq 0\) and \(\mu_k\) is a probability measure, then \(\boldsymbol a_\ell^r=\int_{\mathcal D_{\ell,i}}\boldsymbol a_{\ell,i}(x)d \mu_k\in \mathcal A\) is a convex combination of vectors in \(\mathcal A\) and, since \(\mathcal A\) is convex, so \(\boldsymbol a_\ell^r\in\mathcal A\).
+Therefore, the orbit after the transition out of the ephemeral set \(\mathcal S_{0}^{*,k}\) is in \(\mathcal A\) and therefore the QBD-RAP with the additional ephemeral states is a valid QBD-RAP.''
     \end{itemize}
@@
 \begin{quote}
-    I believe I have referred to is as an infinitesimal interval. Is that clear enough?
+    Fixed. I replaced \(t+du\) by the infinitesimal interval \([t+u,t+u+du]\). 
 \end{quote}
@@
 \begin{quote}
-    I believe I explain it.
-The purpose is to introduce the concept of closing operators so that we can get approximations of density functions within intervals.
-I accept that it is messy and experimental.
-I will add some definition environments, or something to help clear it up.
+   It is experimental.
+I have changed the name of the section to clarify its intent; the section title is now \textit{Recovering intra-level approximations to the distribution of \(X(t)\) using \(\boldsymbol A(t)\) -- some intuitive ideas}.
+
+   I have added some comments at the start to clarify the purpose of this section.
+
+   I have also added a concluding paragraph: ``In this section we have established some intuitive ideas about how to extract an approximation to the density of \(X(t)\) from the orbit \(\boldsymbol A(t)\) using the idea of a closing operator, \(\boldsymbol v(x)\).
+As we shall see, given certain properties of \({\boldsymbol v}(x)\) we can prove that the approximation scheme converges, and ensures positivity.
+In the cases above, all the closing operators (4.25)-(4.26) lead to an approximation which converges and, due to their interpretation as probability densities, ensure positivity.
+Numerical experiments are used to investigate the closing operator from this section in Chapter~7.''
 \end{quote}
@@
 \begin{quote}
-    This is a ``how to'' approximate operators of fluid-fluid queues. Not sure what he wants me to add here? Perhaps the title is misleading?
+    I have moved this section to the preliminaries for Chapter~7 -- as we discussed, it fits better there.
 \end{quote}
@@
 \begin{quote}
-    \begin{itemize}
-        \item Indeed, the proof is long and tedious, and there might be a shorter proof. I wondered this myself. To me, this is the most obvious proof. 
-        \item I'm not sure that the full extent of the result is appreciated in the comment above.
-The approximant is ME distributed, but the true solution is, in general, not ME distributed.
-It's not clear, to me, how denseness, in and of itself, would help the proof.
-To me, all denseness allows us to say is ``there exists a sequence of ME's which converge to the distribution of interest'' --it does not tell us what the MEs are.
-    \end{itemize}
+    The QBD-RAP approximation to the fluid queue approximates the distribution of the fluid queue at all times \(t\), including the time up to and including \(\tau_1\).
+The approximation to the distribution of the \(X(t)\) within a given band \(\mathcal D_k\) is obtained using the value of the orbit process of the QBD-RAP, \(\boldsymbol A(t)\) (the band which \(X(t)\) is in, and the phase are approximated by \(L(t)\) and \(\phi(t)\), respectively).
+So, in this sense, it is significantly more than just \(\tau_1\) which is approximated -- it is the entire transient distribution of the fluid queue.
+Further, while \(\tau_1\) is matrix exponentially distributed, the distributions of interest (the transient distributions and hitting times of fluid queues) are, in general, not matrix exponentially distributed.
+Indeed, MEs are dense, which implies that we can approximate the distribution of interest arbitrary well with an ME.
+However, the fact that MEs are dense is not constructive and gives now way to find such an ME.
+In contrast, we show that by choosing any matrix exponential distribution with sufficiently small variance, the distribution of the fluid queue can be approximated arbitrarily closely.
+We use the class of concentrated matrix exponential distributions for our numerical experiments.
+
+    I would greatly appreciate a simpler, more general and more transparent proof.
+The proof shown here is, to me, the most natural, and also useful for our purpose as it suggests that all we need to do choose an ME with small variance, and we should obtain a decent approximation.
 \end{quote}
@@
 \begin{quote}
-    They are defined in the preliminaries to the chapter. I have now recalled them in Lemma 5.4, for convenience. 
+    I have now referenced them and all similar terms throughout that chapter.
+
+    I have also added a glossary of terms to help the reader navigate the notation, and the glossary has a reference to where that notation is defined. 
 \end{quote}
@@
 It is very, very hard to read this section, and I shall not go through it in more detail until it has been restructured. Here I pinpoint some details.
-\begin{quote}
-    sorry about that 
-\end{quote}
 
@@
 \begin{quote}
-    Fair enough.
-However, I do say in this section that we form a partition of \(f^{\ell_0,(p)}(t)(x,j;x_0,i)\), and the \(f_{m,r,s}^{\ell_0,(p)}\) the reviewer refers to is an element of this partition.
+    I have added a sentence before defining \(f_{m,r,s}^{\ell_0,(p)}\) which states why we are interested in them and I have added a sentence after defining \(f_{m,+,+}^{\ell_0,(p)}\) which describes its meaning.
 \end{quote}
@@
     \begin{itemize}
-        \item I'm not sure what interpretation they are referring to. Do you have any idea?
-        \item Not sure if I mean to cite da Silva here, but I'm not sure if it is the best reference for these equations, so I have removed it as the expressions in it do not concern time variables.  
+        \item I have added interpretations to this section. \textbf{Nigel}, could you please take a look at the first 3 paragraphs of Section~5.3.
+        \item I believe that this is a typo or accidental reference as Da Silva does not consider time-dependent analysis. I have removed the reference. 
     \end{itemize}
@@
 \begin{quote}
-    I should elaborate on this. Need help devising a plan of attack. Maybe an inductive argument might be easiest? 
+    Yes, a link is missing which is not obvious. I have now elaborated.
+
+    \textbf{Nigel,} please read up to Equation (5.35). 
 \end{quote}
@@
     \begin{itemize}
-        \item I don't think I have introduced big-O notation. I shall do that. It's not multivariate. 
-        \item It is introduced in this Lemma --all the \(r_\ell(n)\) terms are introduced in Lemmas. I wonder why this \(r_4\) was singled out.
+        \item I have now mentioned that this is big O notation (it is not multivariate). 
+        \item It is introduced in this Lemma -- all the \(r_\ell(n)\) terms are introduced in Lemmas. I have now clarified this for all \(r_\ell(n)\) terms.
     \end{itemize}
@@
 ``The exponential random variable in the last line of (6.23) appears due to the interpretation of the Laplace transform as probability that \(\tau_n^{(p)}\) occurs before and exponential random variable, \(E^\lambda\).''
-        \item no, I don't think an = is missing.
+        \item I don't think an = is missing.
     \end{itemize}
@@
 
-Several places: you carefully verify conditions in Tonelli’s theorem for how to swap summation and integration.
+Several places: you carefully verify conditions in Tonelli's theorem for how to swap summation and integration.
 I think, alternatively, the Beppo-Levi theorem stating that you can always swap the order whenever the integrant is non-negative would apply.
 \begin{quote}
-    From memory, I justified the swap by the non-negativeness of the summand/integrand where possible.
-I did not assume \(\psi\) was positive, so I had to use Tonelli's theorem many times.
-Also, my Googling tells me Beppo-Levi is like (the same as?) the monotone convergence theorem, but I do not assume any monotonic behaviour, so it's not immediate how I might use it.
+    The Beppo-Levi Theorem applies to sequences of increasing, positive functions.
+Here, we do not, in general assume any monotonic properties of integrands, nor do we assume positivity.
+For example, we do not assume \(\psi\) in (6.23) is positive, and it is not known, in general, if the integrands in (6.32) are monotonic (in the sequence index \(p\).
+It may be possible to assume \(\psi\) is positive.
+In fact Wikipedia tells me that to show convergence in distribution (of a sequence of r.v.~\(X_p\) to \(X\)) using a positive \(\psi\) requires us to show
+    \[\liminf E[\psi(X_p)]\geq E[\psi(X)].\]
+
+    I guess we could show \[\lim E[\psi(X_p)]= E[\psi(X)]\] for positive \(\psi\) which implies \[\liminf E[\psi(X_p)]\geq E[\psi(X)]\] and that would be simpler....
+
+    Where possible, I have used the positiveness of an integrand to justify a swap of integrals via Tonelli's Theorem. 
 \end{quote}
@@
 \begin{quote}
-    Yeah, maybe I do need to remake the plots up to 7.5. Perhaps a solid orange line would be better. 
+    I have changed the markers and colours for plots 7.1-7.7 (and the captions to match) to make the plots clearer. 
 \end{quote}
@@
     \begin{itemize}
-        \item How did I choose the QBD-RAP? This was the entire purpose of Chapter 4!!! 
-        \item Is 21 enough? Perhaps not.
-        \item I have a couple of examples of approximated PDFs in there -- I could add more. For example, Fig 7.21, 7.18, 7.27, 7.29, 7.31, 7.35. 
+        \item To demystify the computations two comments have been added, one in the introduction and one in Chapter~7.
+They both mention that once we have the parameters of the QBD-RAP and have constructed the generator, then we just use the usual formulae to approximate the fluid queue or fluid-fluid queue.
+        \item Dimension 21 appears to be sufficient to give a reasonable indication about which methods perform better than others and to give some indication of what the convergence characteristics of the schemes might be.
+For the problems which use simulation to obtain a ground truth we are limited by computational resources -- we are already doing \(5\times 10^10\) simulations and even with this many the variance of the approximations is still significant relative to the errors (for example in Figure~7.23).
+For smaller problems like the travelling wave problems or hitting times problem, we do have enough resources to use higher-dimension schemes, but for problems like the transient distribution in Section~7.5 it takes at least a couple of days to run on my machine.
+        \item Representative examples of the approximations are shown in Figures~7.3, 7.5, 7.8, 7.14, 7.18, 7.21, 7.27, 7.29, 7.31, and 7.35.
+Perhaps, \textbf{Nigel}, could you run your eye over these figures to see if they are sufficient?
     \end{itemize}
@@
 \begin{quote}
-    Is this beyond the scope of the thesis? 
+    I'm not sure which `static cases' he refers to -- perhaps the stationary distribution and first return distribution?
+I am also unfamiliar with the Richardson extrapolation -- it appears to be a way to obtain better approximations given two coarse approximations which are constructed with different `step sizes'.
+An interesting idea for further research, but beyond the scope of the thesis.
 \end{quote}
@@
 \begin{quote}
-    I do not understand this criticism.
-The QBD-RAP beats uniformisation almost all the time.
-I do cherry-pick the examples because I am trying to argue that the QBD-RAP method works well for those``typically hard'' examples.
-What do you mean by convergence and quality?
-It's converging to the true solution!
-What is useless about that?
+    I disagree.
+For all examples except the trivial problem of approximating the initial condition, the QBD-RAP outperforms the Uniformisation method, often significantly.
+Indeed, the relative performance will be example dependent, but the QBD-RAP still outperforms the Uniformisation method in every example -- from smooth problems like the stationary distribution to hard problems like the hitting times problem.
+
+    In all examples we have compared the approximations to a `ground truth'.
+Often the ground truth is known analytically, but sometimes it has to be simulated.
+We ensure that the measurements of performance are relatively accurate by doing many \(5\times 10^10\) simulations so that the variance of the estimates is sufficiently small.
 \end{quote}
@@
 \begin{quote}
-    I did not specifically control for these errors.
-I did use the same integration scheme and step size for all methods, so these errors should be somewhat consistent between the methods.
-I did not use Richardson.
+    I did not consider using Richardson extrapolation.
+
+    To control for numerical errors I used the same SSPRK time integration scheme for all examples, keeping the time-step constant and sufficiently small to ensure stability.
+Double-precision floating point number were used.
 \end{quote}
@@
 \begin{quote}
-    They are relatively similar to implement. What sort of commentary would this require?
+    See my response to the computations being a mystery above.
+\textbf{Nigel}, please take a look at the first four paragraphs of the second subsection of Section~7.1 (titled \emph{On the application of the QBD-RAP...}).
 \end{quote}
@@
     \begin{itemize}
-        \item I'm not sure what the reviewer means by "uneven level of detail" -sure, some sections are more technical than others.
-But I think that's just the nature of the topics and the they are presented in the detail required to understand the relevant parts of the thesis.
-
-        \item I don't think removing basic topics is the answer. I'd rather explain more and the reader can skip bits they don't think they need.
-
-        \item I like the order of the introduction.
-It builds in a logical order: some fundamental mathematical concepts -just enought to get going, then I introduce the fluid and fluid-fluid queue, then tools for their approximations, then other sundry mathematical concepts used in the thesis.
-I don't think suggested order would add value.
+        \item I have moved some of the more general mathematical preliminaries to the appendix and left the more essential sections. This should now be a more even level of detail.
     \end{itemize}
@@
     \begin{itemize}
-        \item Yes, these sections are long and heavy.
-I think it's hard to avoid.
-We have talked about this a lot.
-I am happy with the content of these sections, but I know Peter objects to including some of the details which make the presentation more complex.
-        \item Happy to merge 5 and 6. What do you think? I separated them because they used distinct techniques, but they ultimately combine to prove a single result. 
-        \item Happy to add more steps and signposts in Sections 5.2/5.3 or convert them to Lemmas -perhaps structuring as Lemmas is the way to go as it condenses the important points into a single statement and leaves details to a proof, allowing the reader to forget about the details.
+        \item I have reorganised and rewritten Sections 5.1, 5.2 and 5.3 into lemmas and corollaries which should help readability.
+It also makes it more clear what is a definition and what is a result.
+        \item I separated Chapters 5 and 6 to help readability and because the proof techniques use in each chapter accentuate different elements of their arguments.
     \end{itemize}
@@
     \begin{itemize}
-        \item   Not sure what ``fluid-fluid models are welcome'' means.
+        \item See also my comment to Reviewer 1 about using approximations up to order 21.
+
         \item I went up to order 21 in the numerical investigations.
@@
 \noindent\textbf{3 List of minor remarks and typos}
+% [IM UP TO HERE]
 
@@
 \begin{quote}
-    I'm lazy. Do I need to fix this?
+    Changed := to =. 
 \end{quote}
-\item \st{general: in formulas, functions like diag, sign etc. should be in mathrm font}
+\item {general: in formulas, functions like diag, sign etc. should be in mathrm font}
+\begin{quote}
+    Fixed
+\end{quote}
 \item page 7 bottom, stopping time definition
 \begin{quote}
-    Not sure what's wrong with it...
+    Fixed. 
 \end{quote}
-\item \st{page 13: 'otherwise it is reflected' I understand why the word 'reflected' is used here, but usually a reflecting boundary condition means evolving with -ci instead of ci.
+\item {page 13: 'otherwise it is reflected' I understand why the word 'reflected' is used here, but usually a reflecting boundary condition means evolving with -ci instead of ci.
 Here, this is simply an absorbing boundary condition, with transition upon reaching the boundary.}
+\begin{quote}
+    Good point. I have now clarified.
+\end{quote}
 \item page 15: Why is the Laplace domain variable sometimes \(s\), sometimes \(\lambda\)?
 \begin{quote}
-    I have change this on page 15. Should I change it throughout?
+   It should be \(\lambda\). Fixed.
 \end{quote}
-\item \st{page 17, bottom formula: extra ) bracket}
-\item \st{page 18: 'which is only possibly'  'which is only possible'}
+\item page 17, bottom formula: extra ) bracket
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 18: 'which is only possibly'  'which is only possible'}
+\begin{quote}
+    Fixed. 
+\end{quote}
 \item page 19: 'Conditions on fluid-fluid queues to ensure that (2.20) will be differentiable with respect to y are not known, and here we just assume that it is.' this seems like an innocent assumption.
@@
 \begin{quote}
-    This is a large can of worms... how much do we want to open it?
+    I added ``For further reading, further theoretical developments and some examples of fluid-fluid queues are analysed in \cite{boz2022}.''
 \end{quote}
-\item \st{page 20, third row from top: some ) brackets are missing}
-\item \st{page 20: a point masses to a point mass}
-\item \st{page 30, before (2.35): extra ) bracket in ()}
+\item {page 20, third row from top: some ) brackets are missing}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 20: a point masses to a point mass}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 30, before (2.35): extra ) bracket in ()}
+\begin{quote}
+    Fixed. 
+\end{quote}
 \item page 35: Section 2.5.5. about accuracy is indeed very brief and seems highly nonrigorous.
 \begin{quote}
-    Hnnn. Yep, it is. I could tighten this up and quote the result from the reference I cite, but it requires some more preliminaries which are not relevant to the rest of the thesis. Thoughts? 
+   I have rewritten this section completely. Please take a quick look.  
 \end{quote}
-\item \st{page 36, (2.39): m + 2 to m2 in the second sum}
-\item \st{page 36, bottom line: m + 1 to m1 in the first sum}
-\item \st{page 37: the term representation is used before its definition. The correct order
+\item {page 36, (2.39): m + 2 to m2 in the second sum}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 36, bottom line: m + 1 to m1 in the first sum}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 37: the term representation is used before its definition. The correct order
 would also help out Cor. 2.6.}
-\item \st{page 37, near bottom: possibility defective to possibly defective I recommend a
+\begin{quote}
+    I have reordered this entire section. 
+\end{quote}
+\item {page 37, near bottom: possibility defective to possibly defective I recommend a
 thorough read-through of Section 2.6.1.}
-\item \st{page 39: 'the process may not actually jump at these times' reword to something
+\begin{quote}
+    Fixed \& read-through,
+\end{quote}
+\item {page 39: 'the process may not actually jump at these times' reword to something
 like 'the process does not necessarily jump at these times'}
-\item \st{page 42, Section 2.8: is this section even necessary? Laplace-Stieltjes transform has been used heavily earlier.}
 \begin{quote}
-    I have moved this section to section 2.3.
+    Fixed. 
 \end{quote}
-\item \st{page 48, near the top: 'Fubini's Theorem us to' to 'Fubini's Theorem allows us to'}
-\item \st{page 53: 'The DG method conserves probability' so any quantity from [0,1] remains
+\item {page 42, Section 2.8: is this section even necessary? Laplace-Stieltjes transform has been used heavily earlier.}
+\begin{quote}
+    I have moved this section to an appendix. 
+\end{quote}
+\item {page 48, near the top: 'Fubini's Theorem us to' to 'Fubini's Theorem allows us to'}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 53: 'The DG method conserves probability' so any quantity from [0,1] remains
 in [0,1] after the approximation, or does it stay completely unchanged? Or does this
@@
 \end{quote}
-\item \st{page 55: W k or Wk? Both are used, but denote the same space. Also, it would be
+\item {page 55: W k or Wk? Both are used, but denote the same space. Also, it would be
 nice to recall the definition of Wk here.}
@@
 \begin{quote}
-    They are only used here, in this specific place, and only takes up less than 2 lines. Do I need to move it?
+    Added to the polynomials section in the appendix. 
 \end{quote}
-\item \st{page 56: There are different choices for the flux rather, for the choice of approximation for the flux, right?}
-\item \st{page 60, middle of page: overlong line}
-\item \st{page 60, bottom formula, first line: the 'or' part should be re-formulated (e.g.
+\item {page 56: There are different choices for the flux rather, for the choice of approximation for the flux, right?}
+\begin{quote}
+    Fixed grammar.
+\end{quote}
+\item {page 60, middle of page: overlong line}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 60, bottom formula, first line: the 'or' part should be re-formulated (e.g.
 something like k = l 6 = 0 or K)}
-\item \st{page 70, middle: 'we can show even show' to 'we can even show'}
-\item \st{page 70: 'on the event that phi(t) is constant, then X(t) moves deterministically' 'then' is unnecessary}
-\item \st{page 80, near top: 'Other choices are possible' ... 'are other possible choices' unnecessary repeat}
-\item \st{page 88, top line: 'events epochs' to 'event epochs'}
-\item \st{page 102, bottom: use the same function (e.g. g(x)) in both formulae}
-\item \st{page 104: (5.10) and (5.11) seem to be a single formula broken into two parts
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 70, middle: 'we can show even show' to 'we can even show'}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 70: 'on the event that phi(t) is constant, then X(t) moves deterministically' 'then' is unnecessary}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 80, near top: 'Other choices are possible' ... 'are other possible choices' unnecessary repeat}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 88, top line: 'events epochs' to 'event epochs'}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 102, bottom: use the same function (e.g. g(x)) in both formulae}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 104: (5.10) and (5.11) seem to be a single formula broken into two parts
 unnecessarily.}
-\item \st{page 106, near bottom: 'in terms of the of first return matrices' to 'in terms of the
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 106, near bottom: 'in terms of the of first return matrices' to 'in terms of the
 first return matrices'}
-\item \st{page 108, line before (5.25): you could simply write q, r in +, (later as well)}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 108, line before (5.25): you could simply write q, r in +, (later as well)}
+\begin{quote}
+    Fixed. 
+\end{quote}
 \item page 111: maybe Corollary D.3 from Appendix D could be moved to the main text
 \begin{quote}
-    It could be but I don't think it needs to be moved. Thoughts?
+    The Corollary and proof are not insightful and are basically just algebraic manipulations, so I have left them in the appendix.
 \end{quote}
-\item \st{page 151: 'and the last inequality is the Law of total probability' to 'and the last
+\item {page 151: 'and the last inequality is the Law of total probability' to 'and the last
 equality is the Law of total probability'}
-\item \st{page 154: 'where the swap of the intergals' to 'where the swap of the integrals'}
-\item \st{page 155: 'The probability of 1 is' this looks a bit strange, maybe use A and B to
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 154: 'where the swap of the intergals' to 'where the swap of the integrals'}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 155: 'The probability of 1 is' this looks a bit strange, maybe use A and B to
 denote those events}
-\item \st{page 161: the bottom formula can fit in a single line}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 161: the bottom formula can fit in a single line}
+\begin{quote}
+    Fixed. 
+\end{quote}
 \item page 162, about pointwise convergence: weak convergence is fine, that should cover
@@
 \begin{quote}
-    Thanks?
+    Excellent.
 \end{quote}
@@
 \begin{quote}
-    Thanks?
+    I updated these plots addressing Reviewer 1's comments.
 \end{quote}
-\item \st{page 171: 'as more mass at the left' to 'as more mass is to the left'}
-\item \st{page 175, Fig. 7.7: for the L1 error, QBDRAP seems to have a faster convergence
+\item {page 171: 'as more mass at the left' to 'as more mass is to the left'}
+\begin{quote}
+    Fixed. 
+\end{quote}
+\item {page 175, Fig. 7.7: for the L1 error, QBDRAP seems to have a faster convergence
 rate than DG. It would be nice to see the plot for higher dimension too. Either
@@
 \begin{quote}
-    Reworded and clarified that the QBD-RAP performs worst for these dimension but may be better for higher dimensions (speculative). See comment on higher dims.
+    Reworded and clarified that the QBD-RAP performs worst for these dimension but may be better for higher dimensions (speculative). See previous comments on higher dim approximation schemes.
 \end{quote}
@@
 \begin{quote}
-    added ``Interestingly, the error curve for the DG scheme
-    is not monotonic which is likely to be caused by the specific location of oscillations in the approximation for different dimension approximations.''
+    added ``Interestingly, the error curve for the DG scheme is not monotonic which is likely to be caused by the specific location of oscillations in the approximation for different dimension approximations.''
 \end{quote}
@@
 \end{quote}
-\item \st{page 265: Borthwick (n.d.) why is there no date? I found a 2016 edition.}
+\item {page 265: Borthwick (n.d.) why is there no date? I found a 2016 edition.}
+\begin{quote}
+    Fixed. 
+\end{quote}
 \end{itemize}
 
+% -----
+
+% Things to go back and do 
+% \begin{itemize}
+%     \item CPU times
+%     \item Re-read 5.1, 5.2, 5.3
+%     \item Reread 2.5.5
+% \end{itemize}
+
+% -----
+
 \end{document}

### 2023-01-22 reviews/response.tex
.

--- 2022-12-31 reviews/response.tex
@@
 \begin{document}
-\maketitle
-The reviewers comments are in plain text and not indented. 
-\begin{quote}
-    My comments follow are are in indented and are in italics. 
-\end{quote}
-Text that is \st{struck-out} has already been addressed.
-
-\pagebreak
-\section{Reviewer 1}
-
-\textbf{OVERVIEW}
-
-This thesis considers approximations of fluid-fluid queues by approximating the dri-
-ving fluid process by either a QBD or RAP-QBD process. By discretizing the state-
-space of the driving fluid process, one does not obtain a QBD automatically, but by
-replacing the constant drifts with random ones, governed by birth- and death pro-
-cesses that mimic the drift, one can approximate the fluid flow process by a QBD.
-This is the content of Chapter 3, in which part the author is the co-author of a
-published paper.
-\begin{quote}
-    This is not the content of Chapter 3. 
-\end{quote}
-
-Since one is interested in approximating the deterministic hitting times caused by
-the constant drift, one might think that Erlangization is a good idea to improve on
-this. It turns out, however, that Erlangization is already a part of the QBD construc-
-tion, so nothing is gained in this direction.
-
-Then the author, in the second part of the thesis, considers an interesting and far
-from trivial approach by employing RAP-QBD's as replacements for the QBD appro-
-ximations for the sake of using concentrated matrix-exponential distribution as a
-more efficient approximation to the deterministic hitting times.
-
-The second chapter contains some background chosen by the author. It resumes ope-
-rator theory, weak convergence, Laplace transform and many other subjects that are
-readily available in the mathematical literature and might be considered as common
-background. When it comes to reviewing fluid and fluid-fluid models, then these are
-explained with the same level of detail. However, since this is much more advanced
-and more specialized, one might have hoped for a more rigorous introduction to both
-fluid queues and RAPs for the sake of readability. This, however, is not a compulsory
-request from my side.
-\begin{quote}
-    I'd be happy to add more to the background on fluid queus and RAP, but, personally, I don't think there is much more to say which is relevant to the thesis.
-This thesis does not have a great deal to do with the existing analyses of fluid queues.
-Thoughts?
-\end{quote}
-
-The thesis is large and contains a lot of material but lacks order. It is hard to find
-details on notation, for example, \(\phi\) and \(\varphi\). This makes the task of reading it much
-harder, in particular, if one does not read the whole thesis from scratch.
-\begin{quote}
-    Yes. Sorry about that. 
-\end{quote}
-
-I believe this is mainly due to the lack of structure. It would be useful to employ
-a more methodological approach using Definitions, Theorems, etc. and frequently
-pinpoint back to these results/definitions whenever used.
-\begin{quote}
-    IMO, I have used Theorem, Lemma, Corllary environments to help with this. Perhaps I could use Definition environments too, this might help. Thoughts? 
-\end{quote}
-
-Chapter 5 is long and technical, while Chapter 6 is short and more straightforward.
-Also, the ultimate aim of both chapters is the result of Chapter 6, so to me, it would
-make more sense to merge the two chapters.
-\begin{quote}
-    I'm happy either way.
-These two chapters were originally one long chapter but i decided to break them in two because they use distinct technical tools.
-Perhaps I should make them a single chapter with two main sections in the chapter.
-\end{quote}
-
-In conclusion, the thesis is long, highly technical, and somewhat unstructured, ma-
-king it non-accessible to a broader community.
-
-\textbf{ASSESSMENT}
-
-The author is a co-author of Chapter 3, the material of which is original and publis-
-hed. The material in Chapters 4,5 and 6 is original. The material of the thesis has a
-reasonable narrow scope. It considers an important approximation problem using a
-recent approach. The communication is very technical, and lacks some structure for
-a more smooth communication. The contribution is significant but could have been
-broader in its scope.
-\begin{quote}
-    Interesting that he didn't mention Chapter 7... perhaps he does not recognise numerical investigation in the same regard as the technical chapters, haha.
-
-    Yes, the scope of the thesis is narrow. It was one of my main worries about the work. I could have added more, but, as he already pointed out, the thesis is already long and hard to read.
-
-    Specific comments on structure to be addressed below. 
-\end{quote}
-
-\textbf{CHAPTER 1-3}
-
-\st{p. 8 l.end-6: restricted to t = 0.}
-
-p. 12 l.end-6: you have already used matices extensively.
-\begin{quote}
-    I have. I just wanted to clarify notation, and point out that matrices use square brackets. 
-\end{quote}
-
-p. 13 l.end-10: how can the underlying phase-process change state upon hitting a boun-
-dary? Then it should be level dependent?
-\begin{quote}
-    It is level-dependent.
-I have added some commentary: "Some boundary conditions may introduce a dependence of the phase process, \(\{\varphi(t)\}\), on the level process \(\{X(t)\}\).
-Specifically there may be a change of phase upon the event that the level process hits a boundary.
-When the phase process depends on the level process then the fluid queue is \emph{level-dependent}, otherwise it is \emph{level independent}.
-For a level-dependent fluid queue the phase process \(\{\varphi(t)\}\) is not the only random element of the fluid queue."
-\end{quote}
-
-\st{Formula (2.10): parenthesis.}
-
-\st{p.21: B(t), argument t missing at some places.}
-\begin{quote}
-    Actually, there shouldn't be a (t)
-\end{quote}
-
-\st{p.40: second last line, a D missing.}
-
-(2.42): one does announce a proof to a formula/equation. Why do you proved it all? Most other items around this section are left unproven.
-\begin{quote}
-    I don't think there is a need for the proof. Happy to omit it. Same for the MPR on the previous page?
-\end{quote}
-
-Section 2.4 reproduces some fluid-fluid theory. The section could have benefitted
-from including, at least some, proofs of the ascertions to gain a working knowled-
-ge.
-\begin{quote}
-    happy to add proofs, but they are relatively technical and not absolutely necessary for the thesis IMO. They could help the reader intuit the fluid queue and its operators. Thoughts?
-\end{quote}
-
-Chapter 3 is taken from a paper, reproduced as such and in a good shape.
-\begin{quote}
-    I disagree, but okay. IMO this is one of the hardest-to-read sections!
-\end{quote}
-
-\pagebreak
-\section{Reviewer 2}
-
-\textbf{2 About the main sections}
-
-Section 2 is of slightly uneven level of detail. Maybe some of the more basic topics are not even necessary. I also suggest re-ordering the subsections from basic to advanced
-\begin{quote}
-    \begin{itemize}
-        \item I'm not sure what the reviewer means by "uneven level of detail" -- sure, some sections are more technical than others.
-But I think that's just the nature of the topics and the they are presented in the detail required to understand the relevant parts of the thesis.
-
-        \item I don't think removing basic topics is the answer. I'd rather explain more and the reader can skip bits they don't think they need.
-
-        \item I like the order of the introduction.
-It builds in a logical order: some fundamental mathematical concepts -- just enought to get going, then I introduce the fluid and fluid-fluid queue, then tools for their approximations, then other sundry mathematical concepts used in the thesis.
-I don't think suggested order would add value.
-    \end{itemize}
-\end{quote}
-
-Section 3 presents the discontinuous Galerkin (DG) method. Unfortunately, no explicit error bounds are available.
-
-Section 4 presents the QBDRAP method.
-It uses concentrated matrix exponential (CME) functions.
-It is nice to see that certain properties of CME functions (such as identical real part of all eigenvalues) are utilised.
-
-Sections 5 and 6 proves convergence for QBDRAP.
-These sections are rather long and heavy, but the arguments are mostly familiar from the literature.
-Everything is properly referenced.
-Sections 5 and 6 could be merged into a single section.
-Section 5 is difficult to follow; maybe arranging the calculations in Sections 5.2 and 5.3 into lemmas or steps would result in a better structure.
-\begin{quote}
-    \begin{itemize}
-        \item Yes, these sections are long and heavy.
-I think it's hard to avoid.
-We have talked about this a lot.
-I am happy with the content of these sections, but I know Peter objects to including some of the details which make the presentation more complex.
-        \item Happy to merge 5 and 6. What do you think? I seperated them because they used distinct techniques, but they ultimately combine to prove a single result. 
-        \item Item, happy to add more steps and signposts in Sections 5.2/5.3 or convert them to Lemmas -- perhaps structuring as Lemmas is the way to go as it condeses the important points into a single statement and leaves details to a proof, allowing the reader to forget about the details.
-    \end{itemize}
-\end{quote}
-
-Section 7 presents numerical investigations for a fairly broad selection of examples.
-Despite the lack of explicit error bounds, we see DG performing well in many cases.
-Fluid-fluid models are welcome.
-A general remark: CME distributions only get really concentrated for high orders (dimension).
-Ideally, I would recommend using at least an order of 30, although I understand there are computational limitations.
-Either way, it should not be surprising that their performance is not great for orders below 10 or so.
-\begin{quote}
-    \begin{itemize}
-        \item   Not sure what ``fluid-fluid models are welcome'' means.
-        \item I went up to order 21 in the numerical investigations.
-For some of the more trivial examples in this section I could use higher-order approximations.
-However, for the interesting examples, simulation is required to measure accuracy and this is computationally expensive.
-I'm already doing \(5\times 10^{10}\) realisations!
-To me, it seems unlikely that investigation of higher-order examples would reveal much more insight -- my intuition is that the error decay of the QBD-RAP method is exponential in the order of the ME used.
-    \end{itemize}
-\end{quote}
-
-\textbf{3 List of minor remarks and typos}
-
-Minor remarks/typos:
-\begin{itemize}
-\item general: definitions are sometimes in italics, sometimes not; defining formulas are
-sometimes denoted by :=, sometimes by =
-\begin{quote}
-    I'm lazy. Do I need to fix this?
-\end{quote}
-\item \st{general: in formulas, functions like diag, sign etc. should be in mathrm font}
-\item page 7 bottom, stopping time definition
-\begin{quote}
-    Not sure what's wrong with it...
-\end{quote}
-\item \st{page 13: 'otherwise it is reflected' - I understand why the word 'reflected' is used here, but usually a reflecting boundary condition means evolving with -ci instead of ci.
-Here, this is simply an absorbing boundary condition, with transition upon reaching the boundary.}
-\item page 15: Why is the Laplace domain variable sometimes \(s\), sometimes \(\lambda\)?
-\begin{quote}
-    I have change this on page 15. Should I change it throughout?
-\end{quote}
-\item \st{page 17, bottom formula: extra ) bracket}
-\item \st{page 18: 'which is only possibly'  'which is only possible'}
-\item page 19: 'Conditions on fluid-fluid queues to ensure that (2.20) will be differentiable with respect to y are not known, and here we just assume that it is.' - this seems like an innocent assumption.
-Why don't we have this at least for some nice cases?
-A few more words on this would be welcome.
-\begin{quote}
-    This is a large can of worms... how much do we want to open it?
-\end{quote}
-\item \st{page 20, third row from top: some ) brackets are missing}
-\item \st{page 20: a point masses to a point mass}
-\item \st{page 30, before (2.35): extra ) bracket in ()}
-\item page 35: Section 2.5.5. about accuracy is indeed very brief and seems highly non-
-rigorous.
-\begin{quote}
-    Hnnn. Yep, it is. I could tighten this up and quote the result from the reference I cite, but it requires some more preliminaries which are not relevant to the rest of the thesis. Thoughts? 
-\end{quote}
-\item \st{page 36, (2.39): m + 2 to m2 in the second sum}
-\item \st{page 36, bottom line: m + 1 to m1 in the first sum}
-\item \st{page 37: the term representation is used before its definition. The correct order
-would also help out Cor. 2.6.}
-\item \st{page 37, near bottom: possibility defective to possibly defective I recommend a
-thorough read-through of Section 2.6.1.}
-\item \st{page 39: 'the process may not actually jump at these times' - reword to something
-like 'the process does not necessarily jump at these times'}
-\item \st{page 42, Section 2.8: is this section even necessary? Laplace-Stieltjes transform has been used heavily earlier.}
-\begin{quote}
-    I have moved this section to section 2.3.
-\end{quote}
-\item \st{page 48, near the top: 'Fubini's Theorem us to' to 'Fubini's Theorem allows us to'}
-\item \st{page 53: 'The DG method conserves probability' - so any quantity from [0,1] remains
-in [0,1] after the approximation, or does it stay completely unchanged? Or does this
-mean something else? Maybe a reference to Cor. 3.1 on page 60? In any case, this
-should be clarified here.}
-\begin{quote}
-    I have clarified that it conerves total probability of the system and added a reference to cor 3.1.
-\end{quote}
-\item \st{page 55: W k or Wk? Both are used, but denote the same space. Also, it would be
-nice to recall the definition of Wk here.}
-\begin{quote}
-    Fixed and recalled.
-\end{quote}
-\item page 55: orthogonal complement spaces and such could be introduced earlier, in
-Section 2.
-\begin{quote}
-    They are only used here, in this specific place, and only takes up less than 2 lines. Do I need to move it?
-\end{quote}
-\item \st{page 56: There are different choices for the flux - rather, for the choice of approxi-
-mation for the flux, right?}
-\item page 60, middle of page: overlong line
-\begin{quote}
-    Fixed? Check once compiled.
-\end{quote}
-\item \st{page 60, bottom formula, first line: the 'or' part should be re-formulated (e.g.
-something like k = l 6 = 0 or K)}
-\item \st{page 70, middle: 'we can show even show' to 'we can even show'}
-\item \st{page 70: 'on the event that phi(t) is constant, then X(t) moves deterministically' -
-'then' is unnecessary}
-\item \st{page 80, near top: 'Other choices are possible' ... 'are other possible choices' -
-unnecessary repeat}
-\item \st{page 88, top line: 'events epochs' to 'event epochs'}
-\item \st{page 102, bottom: use the same function (e.g. g(x)) in both formulas}
-\item page 104: (5.10) and (5.11) seem to be a single formula broken into two parts
-unnecessarily.
-\begin{quote}
-    Check once compiled.
-\end{quote}
-\item \st{page 106, near bottom: 'in terms of the of first return matrices' to 'in terms of the
-first return matrices'}
-\item \st{page 108, line before (5.25): you could simply write q, r in +, - (later as well)}
-\item page 111: maybe Corollary D.3 from Appendix D could be moved to the main text
-\begin{quote}
-    It could be but I dont think it needs to be moved. Thoughts?
-\end{quote}
-\item \st{page 151: 'and the last inequality is the Law of total probability' to 'and the last
-equality is the Law of total probability'}
-\item \st{page 154: 'where the swap of the intergals' to 'where the swap of the integrals'}
-\item page 155: 'The probability of 1 is' - this looks a bit strange, maybe use A and B to
-denote those events
-\begin{quote}
-    Check once compiled.
-\end{quote}
-\item \st{page 161: the bottom formula can fit in a single line}
-\item page 162, about pointwise convergence: weak convergence is fine, that should cover
-the most relevant performance measures. That said, the subsequent remarks and
-Chapter 6.3 are welcome.
-\begin{quote}
-    Thanks?
-\end{quote}
-\item page 170, Fig. 1: good choice of plot markers, as the non-linear plot would be
-invisible otherwise
-\begin{quote}
-    Thanks?
-\end{quote}
-\item \st{page 171: 'as more mass at the left' to 'as more mass is to the left'}
-\item \st{page 175, Fig. 7.7: for the L1 error, QBDRAP seems to have a faster convergence
-rate than DG. It would be nice to see the plot for higher dimension too. Either
-way, I would be careful with conclusions about which method performs better in
-this case.}
-\begin{quote}
-    Reworded and clarified that the QBD-RAP performs worst for these dimension but may be better for higher dimensions (speculative). See comment on higher dims.
-\end{quote}
-\item page 182, Fig. 7.14: once again, it would be nice to go higher with the dimension
-\begin{quote}
-    See previous comment on higher dims.
-\end{quote}
-\item page 184, Fig. 7.15 explanation: 'Interestingly, the error curve for the DG scheme
-is not monotonic.' - any ideas why?
-\begin{quote}
-    added ``Interestingly, the error curve for the DG scheme
-    is not monotonic which is likely to be caused by the specific location of oscillations in the approximation for different dimension approximations.''
-\end{quote}
-\item page 187, Fig. 7.18: nice to see a plot with higher dimension.
-\begin{quote}
-    See previous comments on higher order.
-\end{quote}
-\item \st{page 265: Borthwick (n.d.) - why is there no date? I found a 2016 edition.}
-\end{itemize}
-
-\begin{thebibliography}{9}
-    \bibitem{bn2010}
-    Bean, N.~G.~and Nielsen, B.~F. (2010). {`Quasi-birth-and-death processes with rational arrival process components.}' \textit{Stochastic Models}, 26(3), 309-334.
-\end{thebibliography}
+ 1
 \end{document}
